### 2008-10-17 lda-paper.tex
2008-10-17 10:32:37 abez

\documentclass{report}
\usepackage{graphicx,amsmath,comment,fullpage}
\newcommand{\igH}[1]{\includegraphics[height=.9\textheight]{#1}}
\newcommand{\igW}[1]{\includegraphics[width=.9\textwidth]{#1}}
\newcommand{\igWh}[1]{\includegraphics[width=.7\textwidth]{#1}}
\newcommand{\igWhalf}[1]{\includegraphics[width=.45\textwidth]{#1}}
\newcommand{\lda}{Latent Dirichlet Allocation}

\author{Abram Hindle\\
ahindle@cs.uwaterloo.ca\\
Software Architecture Group (SWAG)\\
Supervised by\\
Micheal Godfrey and Ric Holt
}
\title{Lab 0006: Latent Dirichlet Allocation Applied to Revisions}

\usepackage{rotating}   
\begin{document}

\maketitle
\begin{abstract}

  In this lab report we seek to explore how topics shift over time in
  the source control repository of a project. We analyze commit
  comments, and even source code per revision per each time window and
  extract the topics prevelant in that window. We expect that topic
  sets will change over time and the change in topic sets indicate a
  change in the focus of development. We suppose this could help us to
  segment the revisions on the time line to discover parts of an
  iteration. For topic extraction we use \lda (LDA).

\end{abstract}

\section{Introduction}

%?

In this lab report we seek to explore how topics shift over time in
the source control repository of a project. We analyze commit
comments, and even source code per revision per each time window and
extract the topics prevelant in that window. We expect that topic sets
will change over time and the change in topic sets indicate a change
in the focus of development. We suppose this could help us to segment
the revisions on the time line to discover parts of an iteration. For
topic extraction we use \lda (LDA).

\section{Methodology}

\subsection{Datasets and Tools}

Extractors:
\begin{itemize}
\item CVSSuck - CVS Suck mirrors RCS files from a CVS repository. 
\item softChange - extract CVS facts to a PostgreSQL database.
\item bt2csv - Convert BitKeeper repositories to facts in CSV
  databases.
\end{itemize}

Analysis Tools:
\begin{itemize}
\item Hiraldo-Grok - An OCaml based spin off of Grok used for
  answering queries.
\item Gnuplot - graph plotting package
\end{itemize}

\section{First Impression}

% 2007-10-12

Looking at 30 day non-overlapping windows of revisions for MySQL 3.23
(20 topics, top 10 words per topic) we found there were common words
across topic clusters such as diffs, annotate and history. There were
notable transitional topics such as in the first window the word
bitkeeper appears (probably when they adopted bitkeeper) yet in the
following windows there were no mentions of bitkeeper. ``RENAME'' also
only appeared once in the first window.

Often words are shared across topics, see talbe \ref{tab:portability}
for tracking topics about portability.

It is apparent that we have to compare these topics some how by
itemset or similarity.

\begin{table}
\centering
\begin{tabular}{|ccc|l|}
\hline
2000 &  Jul &  31 &    chmod \\
2000 &  Sep &  29 &    fixes benchmark logging windows \\
2000 &  Nov &  28 &    typo fix insert\_multi\_value \\
2001 &  Jan &  27 &    FIxes Innobase Cleanups auto-union \\
2001 &  Mar &  28 &    2 topics bugfix, logging , TEMPORARY,  \\
\hline
2001 &  Jul &  26 &    Update Allow TABLES LOCK [a] \\

2001 &  Aug &  25 &    Tables row version [a] \\
\hline
2001 &  Sep &  24 &    Update checksum merge \\
2001 &  Oct &  24 &    fixed fix \\
2001 &  Dec &  23 &    HPUX SCO fix \\
\hline
2002 &  Feb &  21 &    net buffer length  max\_allowed\_packet [b] \\
2002 &  Mar &  23 &    small buf fix [b]  \\
\hline
2002 &  May &  22 &    [popular] fix SCO OSF1 table\_name \\
2002 &  Nov &  18 &    HPUX11 compiler HP \\
\hline
2003 &  Feb &  16 &    Linux errno  [c] \\
2003 &  Mar &  18 &    alarm bookmark bug [c] \\
\hline
2003 &  Sep &  14 &    Auto logging merge windows distribution fix 64-bit 4.0 Cleanup \\
\hline
\end{tabular}
\caption{Tracking topics associated with the word portability, note some continuous blocks}
\label{tab:portability}
\end{table}

%LDA Skeleton
%1. [ ] Abstract 
\begin{abstract}
%2. [ ] Introduction
\section{Introduction}

%3. [ ] Previous Work [0/2]
\section{Previous Work}
%   1. [ ] Intro to LDA
\subsection{Intro to LDA}
%   2. [ ] LDA Work [0/6]
\subsection{LDA Work}
%      1. [ ] LDA itself
%      2. [ ] Maletic stuff
%      3. [ ] WCRE: Scott/Jim
%      4. [ ] WCRE: Bug LDA stuff
%      5. [ ] MSR: Challenge work
%      6. [ ] FCA work
%4. [ ] Methodology [0/9]
\section{Methodology}
%   1. [ ] Extract Repos
\subsection{Extract Repos}
%   2. [ ] Extract Documents
\subsection{Extract Documents}
%   3. [ ] Windowed Sets
\subsection{Windowed Sets}
%   4. [ ] Cluster Windowed Set w/ LDA
\subsection{Cluster Windowed Set w/ LDA}
%   5. [ ] Extract Clusters
\subsection{Extract Clusters}
%   6. [ ] Cluster Similarity [0/1]
\subsection{Cluster Similarity}
%      1. [ ] need the tool
%             grab from the cluster analysis
%   7. [ ] Continuous Cluster Analysis
\subsection{Continuous Cluster Analysis}
%   8. [ ] Topic Smear
\subsection{Topic Smear}
%   9. [ ] Visualization [0/1]
\subsection{Visualization}
%       1. [ ] need the visualizer
%5. [ ] Tools [0/5]
\section{Tools}
%   1. [ ] LDA
\subsection{LDA}
%   2. [ ] Extractor
\subsection{Extractor}
%   3.   Similarity
\subsection{Similarity}
%   4.   Smearing
\subsection{Smearing}
%   5.   Visualizer
\subsection{Visualizer}
%6.   Results [0/4]
\section{Results}
%   1. Interesting smears 
%   2.   postgresql
\subsection{Postgresql}
%   3.   mysql
\subsection{Mysql}
%   4.   firebird
\subsection{Firebird}
%   5.   what's that other db?
\subsection{ what's that other db?}
%7.   Validation [0/1]
\subsection{Validation}
%   1. well i'm stumped
%   2. look at it?
%   3. investigate those revisions and check the proportion etc
%   4. exploratory
%   5.   work it out
%8.   Validity Threats [0/5]
\section{Validity Threats}
%   1.   validation
%   2.   LDA is questionable
%   3.   blackbox problem
%   4.   not in the token
%   5.   multiple X token
%9.    Future Work
\section{ Future Work}
%10.   Conclusions
\section{Conclusions}
%11.   Start file file:/home/abez/projects/lda-paper/lda-paper.tex

\end{document}

### 2008-10-23 lda-paper.tex
more lda paper
2008-10-22 23:25:30 abez

@@
 topic extraction we use \lda (LDA).
+
+%
+
+In the presence of fine grained changes and detailed changelogs
+
+Developers topics the focus of develop
+
+many changelogs how to sort by change type
+
+interleaved changes relate similar changes
+
+What happened in the previous iteration? What were the developers
+working on? What was the developer in the next cubicle working on?
+What were the common topics and issues dealt with in the current
+release of the software? What were the topics of the previous release?
+
+We propose to extend common concept analysis techniques by executing
+them over windows of documents over time. We hope by looking at topic
+clusters over time we can I identify continuous effort on a select
+group of topics.
+
+Topic analysis of fine-grained revisions is important as often
+revisions belonging to different topics are interleaved or even mixed
+inside of commits.
 
@@
 \begin{abstract}
+\end{abstract}
 %2. [ ] Introduction

### 2008-10-24 lda-paper.tex
2008-10-24 10:44:28 abez

@@
 revisions belonging to different topics are interleaved or even mixed
-inside of commits.
+inside of commits. Using powerful topic analysis like Latent Semantic
+Indexing (LSI) or Latent Dirichlet Allocation (LDA) one can seperate
+these mixed signals into coherent topics. In this case a topic is a
+set of tokens that are related to each other.
+
+\section{Previous Work}
+%   1. [ ] Intro to LDA
+\subsection{Intro to LDA}
+
+\subsection{LDA Work}
+%      1. [ ] LDA itself
+%      2. [ ] Maletic stuff
+%      3. [ ] WCRE: Scott/Jim
+%      4. [ ] WCRE: Bug LDA stuff
+%      5. [ ] MSR: Challenge work
+%      6. [ ] FCA work
+%4. [ ] Methodology [0/9]
 
 \section{Methodology}
+
+\subsection{Extract Repos}
+%   2. [ ] Extract Documents
+\subsection{Extract Documents}
+%   3. [ ] Windowed Sets
+\subsection{Windowed Sets}
+%   4. [ ] Cluster Windowed Set w/ LDA
+\subsection{Cluster Windowed Set w/ LDA}
+%   5. [ ] Extract Clusters
+\subsection{Extract Clusters}
+%   6. [ ] Cluster Similarity [0/1]
+\subsection{Cluster Similarity}
+%      1. [ ] need the tool
+%             grab from the cluster analysis
+%   7. [ ] Continuous Cluster Analysis
+\subsection{Continuous Cluster Analysis}
+%   8. [ ] Topic Smear
+\subsection{Topic Smear}
+%   9. [ ] Visualization [0/1]
+\subsection{Visualization}
+%       1. [ ] need the visualizer
 
@@
 
+%   1. [ ] LDA
+
+%   2. [ ] Extractor
+
+%   3.   Similarity
+
+%   4.   Smearing
+
+%   5.   Visualizer
+
 \section{First Impression}
@@
 
-Often words are shared across topics, see talbe \ref{tab:portability}
+Often words are shared across topics, see table \ref{tab:portability}
 for tracking topics about portability.
@@
 %1. [ ] Abstract 
-\begin{abstract}
-\end{abstract}
 %2. [ ] Introduction
-\section{Introduction}
 
 %3. [ ] Previous Work [0/2]
-\section{Previous Work}
-%   1. [ ] Intro to LDA
-\subsection{Intro to LDA}
 %   2. [ ] LDA Work [0/6]
-\subsection{LDA Work}
-%      1. [ ] LDA itself
-%      2. [ ] Maletic stuff
-%      3. [ ] WCRE: Scott/Jim
-%      4. [ ] WCRE: Bug LDA stuff
-%      5. [ ] MSR: Challenge work
-%      6. [ ] FCA work
-%4. [ ] Methodology [0/9]
-\section{Methodology}
-%   1. [ ] Extract Repos
-\subsection{Extract Repos}
-%   2. [ ] Extract Documents
-\subsection{Extract Documents}
-%   3. [ ] Windowed Sets
-\subsection{Windowed Sets}
-%   4. [ ] Cluster Windowed Set w/ LDA
-\subsection{Cluster Windowed Set w/ LDA}
-%   5. [ ] Extract Clusters
-\subsection{Extract Clusters}
-%   6. [ ] Cluster Similarity [0/1]
-\subsection{Cluster Similarity}
-%      1. [ ] need the tool
-%             grab from the cluster analysis
-%   7. [ ] Continuous Cluster Analysis
-\subsection{Continuous Cluster Analysis}
-%   8. [ ] Topic Smear
-\subsection{Topic Smear}
-%   9. [ ] Visualization [0/1]
-\subsection{Visualization}
-%       1. [ ] need the visualizer
-%5. [ ] Tools [0/5]
-\section{Tools}
-%   1. [ ] LDA
-\subsection{LDA}
-%   2. [ ] Extractor
-\subsection{Extractor}
-%   3.   Similarity
-\subsection{Similarity}
-%   4.   Smearing
-\subsection{Smearing}
-%   5.   Visualizer
-\subsection{Visualizer}
-%6.   Results [0/4]
 \section{Results}

### 2008-10-24 lda-paper.tex
previous work being worked on!

@@
 
-In this lab report we seek to explore how topics shift over time in
-the source control repository of a project. We analyze commit
-comments, and even source code per revision per each time window and
-extract the topics prevelant in that window. We expect that topic sets
-will change over time and the change in topic sets indicate a change
-in the focus of development. We suppose this could help us to segment
-the revisions on the time line to discover parts of an iteration. For
-topic extraction we use \lda (LDA).
-
 %
 
-In the presence of fine grained changes and detailed changelogs
+% In the presence of fine grained changes and detailed changelogs
 
-Developers topics the focus of develop
+% Developers topics the focus of develop
 
-many changelogs how to sort by change type
+% many changelogs how to sort by change type
 
-interleaved changes relate similar changes
+% interleaved changes relate similar changes
 
@@
 revisions belonging to different topics are interleaved or even mixed
-inside of commits. Using powerful topic analysis like Latent Semantic
-Indexing (LSI) or Latent Dirichlet Allocation (LDA) one can seperate
-these mixed signals into coherent topics. In this case a topic is a
-set of tokens that are related to each other.
+inside of commits.  Sometimes a single revision to a file contains
+multiple bug fixes.  Using powerful topic analysis like Latent
+Semantic Indexing (LSI) or Latent Dirichlet Allocation (LDA) one can
+seperate these mixed signals into coherent topics. In this case a
+topic is a set of tokens that are related to each other.
+
+% We suppose this could help us to segment the revisions on the time
+% line to discover parts of an iteration.
+
+Topic analysis is useful because it can help us partition a project's
+time-line into sub-iterations. By breaking apart an iteration we might
+be able to extract the underlying software development process from
+these artifacts.
+
+In this paper we seek to explore how topics shift over time in
+the source control repository of a project. We analyze commit
+comments, and even source code per revision per each time window and
+extract the topics in that time window. We expect that topic sets
+will change over time and the change in topic sets indicate a change
+in the focus of development.
 
 \section{Previous Work}
-%   1. [ ] Intro to LDA
+
 \subsection{Intro to LDA}
 
+XXX explain LDA \cite{944937}
+
 \subsection{LDA Work}
+
+Lukins, Kraft and Etzkorn~\cite{lukins2008} use LDA to help query for
+finding related documents during bug localization tasks. They used LDA
+to build a topic model and then query against it using sample
+documents, the query would return relevant documentation.
+
+LSI is related to LDA and has been used to identify topics in software
+artifacts~\cite{1421013,1374321,10.1109/ICPC.2007.13,10.1109/ICPC.2006.17}.
+
+Semantic Clustering \cite{1698774,1566153}
+
+Linstead et al has proposed an author source code model using
+LDA.\cite{10.1109/MSR.2007.20,NIPS2007637,1321709}
+
+Grant et al. have used an alternative technique to LSI and LDA, they
+used Independant Component Analysis to seperate topic signals from
+software code.
+
+Our technique differs in the sense we apply LDA temporally and we
+apply to change comments.
+
 %      1. [ ] LDA itself
@@
 
+\bibliographystyle{alpha}
+\bibliography{paper1}
+
 \end{document}

### 2008-10-25 lda-paper.tex
2008-10-24 21:59:21 abez

@@
-\documentclass{report}
+\documentclass[times, 10pt,twocolumn]{article} 
+\usepackage{times}
+
 \usepackage{graphicx,amsmath,comment,fullpage}
@@
 
-\author{Abram Hindle\\
-ahindle@cs.uwaterloo.ca\\
+\author{Abram Hindle and Michael Godfrey and Ric Holt \\
+\{ahindle,migod,holt\}@cs.uwaterloo.ca\\
 Software Architecture Group (SWAG)\\
-Supervised by\\
-Micheal Godfrey and Ric Holt
+University of Waterloo\\
 }
-\title{Lab 0006: Latent Dirichlet Allocation Applied to Revisions}
+\title{Temporal Developer Concept Witchery: Latent Dirichlet Allocation Applied to Revisions}
 
@@
 
-\subsection{Intro to LDA}
+%\subsection{Intro to LDA}
 
-XXX explain LDA \cite{944937}
+%XXX explain LDA \cite{944937}
 
-\subsection{LDA Work}
+Latent Dirichlet Alloation (LDA)~\cite{944937} is a competitor to
+latent semantic indexing (LSI), probabilistic latent semantic indexing
+(pLSI) and semantic clustering. It is used for document modelling,
+document clustering and collaborative filtering. What LDA attempts to
+do is model a documents as mixture model of topics.  The common use of
+LDA in software engineering
+literature~\cite{lukins2008,10.1109/MSR.2007.20,NIPS2007637,1321709}
+is to extract topic clusters from documents such as methods, bug
+reports, source code and files.
+
+Our blackbox view of LDA and LSI is that they extract topic clusters
+from the documents that are independant from one another. LDA doesn't
+name these clusters but the clusters consist of words whose
+relationship is often obvious to someone familar with the
+corpus. Technically we could swap LDA for LSI. Our data is posed as
+documents with word distributions (word counts per documents) and LDA
+extracts distinct topics (clusters of words) from the documents.
+
+\subsection{LDA, LSI and Semantic Clustering}
+
+Linstead et al has proposed an author source code model using
+LDA~\cite{10.1109/MSR.2007.20,NIPS2007637,1321709}. They used an
+author-topic (AT) model~\cite{1036902} to infer the association or
+expertise of a particular developer with a particular code topic
+extracted from the source control system.
 
@@
 to build a topic model and then query against it using sample
-documents, the query would return relevant documentation.
+documents, the query would the proportion of topics that were
+associated with the query and thus the related documents.
 
 LSI is related to LDA and has been used to identify topics in software
-artifacts~\cite{1421013,1374321,10.1109/ICPC.2007.13,10.1109/ICPC.2006.17}.
+artifacts for forumal concept analysis and concept
+location~\cite{1421013,1374321,10.1109/ICPC.2007.13,10.1109/ICPC.2006.17}.
+This is when concepts are automatically extracted from source code or
+documents are related to part of the source code.  Concept location is
+how high level concepts relate to low level entities like source code,
+for instance when fixing a bug how do the concepts in the bug report
+map back to the source code?  Semantic Clustering has also been used
+for similar reason~\cite{1698774,1566153}.
 
-Semantic Clustering \cite{1698774,1566153}
-
-Linstead et al has proposed an author source code model using
-LDA.\cite{10.1109/MSR.2007.20,NIPS2007637,1321709}
-
-Grant et al. have used an alternative technique to LSI and LDA, they
-used Independant Component Analysis to seperate topic signals from
-software code.
+Grant et al.~\cite{scottcordy} have used an alternative technique to
+LSI and LDA, they used Independant Component Analysis to seperate
+topic signals from software code.
 
@@
 
-\section{First Impression}
+\section{First Impressions}
 
-% 2007-10-12
+In our first exploratory pass we wanted to see
 
@@
 notable transitional topics such as in the first window the word
-bitkeeper appears (probably when they adopted bitkeeper) yet in the
-following windows there were no mentions of bitkeeper. ``RENAME'' also
-only appeared once in the first window.
+``bitkeeper'' appears (probably when they adopted Bitkeeper for their
+source control system) yet in the following windows there were no
+mentions of Bitkeeper. ``RENAME'' also only appeared once in the first
+window and never again.
 
-Often words are shared across topics, see table \ref{tab:portability}
-for tracking topics about portability.
+A subsampling of the notable topic words is displayed in Table
+\ref{tab:portability}. We tracked one topic cluster which seemed
+related to portability.
 
-It is apparent that we have to compare these topics some how by
-itemset or similarity.
+%Often words are shared across topics, see table \ref{tab:portability}
+%for tracking topics about portability.
 
-\begin{table}
+After viewing these topic is was obvious that we would have to combine
+try to track the evolution of topics over the windows or rank topics
+by similarity.
+
+%It is apparent that we have to compare these topics some how by
+%itemset or similarity.
+
+\begin{table*}
 \centering
@@
 2000 &  Nov &  28 &    typo fix insert\_multi\_value \\
-2001 &  Jan &  27 &    FIxes Innobase Cleanups auto-union \\
+2001 &  Jan &  27 &    fixes Innobase Cleanups auto-union \\
 2001 &  Mar &  28 &    2 topics bugfix, logging , TEMPORARY,  \\
 \hline
-2001 &  Jul &  26 &    Update Allow TABLES LOCK [a] \\
+2001 &  Jul &  26 &    update Allow TABLES LOCK [a] \\
 
-2001 &  Aug &  25 &    Tables row version [a] \\
+2001 &  Aug &  25 &    tables row version [a] \\
 \hline
-2001 &  Sep &  24 &    Update checksum merge \\
+2001 &  Sep &  24 &    update checksum merge \\
 2001 &  Oct &  24 &    fixed fix \\
@@
 \label{tab:portability}
-\end{table}
+\end{table*}
 
@@
 \bibliographystyle{alpha}
-\bibliography{paper1}
+\bibliography{lda-paper}
 

### 2008-10-26 lda-paper.tex
research & paper change!

@@
 
-Extractors:
-\begin{itemize}
-\item CVSSuck - CVS Suck mirrors RCS files from a CVS repository. 
-\item softChange - extract CVS facts to a PostgreSQL database.
-\item bt2csv - Convert BitKeeper repositories to facts in CSV
-  databases.
-\end{itemize}
+To extract data from CVS and BitKeeper we used softChange to extract
+CVS repositories and provide a schema for our revision data. We used
+bt2csv to extract BitKeeper revisions from BitKeeper web repositories.
 
-Analysis Tools:
-\begin{itemize}
-\item Hiraldo-Grok - An OCaml based spin off of Grok used for
-  answering queries.
-\item Gnuplot - graph plotting package
-\end{itemize}
+% Extractors:
+% \begin{itemize}
+% \item CVSSuck - CVS Suck mirrors RCS files from a CVS repository. 
+% \item softChange - extract CVS facts to a PostgreSQL database.
+% \item bt2csv - Convert BitKeeper repositories to facts in CSV
+%   databases.
+% \end{itemize}
+
+Our analysis tools consisted of: Hiraldo-Grok, an OCaml based spin off
+of Grok used for answering queries; Gnuplot, a graph plotting package,
+lda-c a LDA package by Biel
 

### 2008-11-06 lda-paper.tex
little update

--- 2008-10-24 lda-paper.tex
@@
 
+In this lab report we seek to explore how topics shift over time in
+the source control repository of a project. We analyze commit
+comments, and even source code per revision per each time window and
+extract the topics prevelant in that window. We expect that topic sets
+will change over time and the change in topic sets indicate a change
+in the focus of development. We suppose this could help us to segment
+the revisions on the time line to discover parts of an iteration. For
+topic extraction we use \lda (LDA).
+
 %
 
-% In the presence of fine grained changes and detailed changelogs
+In the presence of fine grained changes and detailed changelogs
 
-% Developers topics the focus of develop
+Developers topics the focus of develop
 
-% many changelogs how to sort by change type
+many changelogs how to sort by change type
 
-% interleaved changes relate similar changes
+interleaved changes relate similar changes
 
@@
 revisions belonging to different topics are interleaved or even mixed
-inside of commits.  Sometimes a single revision to a file contains
-multiple bug fixes.  Using powerful topic analysis like Latent
-Semantic Indexing (LSI) or Latent Dirichlet Allocation (LDA) one can
-seperate these mixed signals into coherent topics. In this case a
-topic is a set of tokens that are related to each other.
-
-% We suppose this could help us to segment the revisions on the time
-% line to discover parts of an iteration.
-
-Topic analysis is useful because it can help us partition a project's
-time-line into sub-iterations. By breaking apart an iteration we might
-be able to extract the underlying software development process from
-these artifacts.
-
-In this paper we seek to explore how topics shift over time in
-the source control repository of a project. We analyze commit
-comments, and even source code per revision per each time window and
-extract the topics in that time window. We expect that topic sets
-will change over time and the change in topic sets indicate a change
-in the focus of development.
+inside of commits. Using powerful topic analysis like Latent Semantic
+Indexing (LSI) or Latent Dirichlet Allocation (LDA) one can seperate
+these mixed signals into coherent topics. In this case a topic is a
+set of tokens that are related to each other.
 
 \section{Previous Work}
-
+%   1. [ ] Intro to LDA
 \subsection{Intro to LDA}
 
-XXX explain LDA \cite{944937}
-
 \subsection{LDA Work}
-
-Lukins, Kraft and Etzkorn~\cite{lukins2008} use LDA to help query for
-finding related documents during bug localization tasks. They used LDA
-to build a topic model and then query against it using sample
-documents, the query would return relevant documentation.
-
-LSI is related to LDA and has been used to identify topics in software
-artifacts~\cite{1421013,1374321,10.1109/ICPC.2007.13,10.1109/ICPC.2006.17}.
-
-Semantic Clustering \cite{1698774,1566153}
-
-Linstead et al has proposed an author source code model using
-LDA.\cite{10.1109/MSR.2007.20,NIPS2007637,1321709}
-
-Grant et al. have used an alternative technique to LSI and LDA, they
-used Independant Component Analysis to seperate topic signals from
-software code.
-
-Our technique differs in the sense we apply LDA temporally and we
-apply to change comments.
-
 %      1. [ ] LDA itself
@@
 
-\bibliographystyle{alpha}
-\bibliography{paper1}
-
 \end{document}

### 2009-02-17 lda-paper.tex

--- 2008-10-25 lda-paper.tex
@@
 \subsection{Extract Repos}
+
+We extracted the repositories with software such as rsync, CVSSuck,
+\texttt{softChange} and bt2csv.  \texttt{softChange} provided a
+general schema for storing revisions. CVSSuck and rsync mirrored CVS
+repositories while bt2csv mirrored web accessible bitkeeper
+repositories.
+
 %   2. [ ] Extract Documents
@@
 %   3. [ ] Windowed Sets
+
+The document we are primarily using are the changelog comments, these
+are the comments added when a revision to the source code is
+committed, it could be per file or per commit of changes to multiple
+files, this depends on the source control system used.
+
+Documents are converted into word count distributions. The
+distributions include all of the possible words in the system.
+
 \subsection{Windowed Sets}
+
+We then create windows which are sets of documents grouped by time
+window. Time windows can overlap but don't have to. Windowing by time
+allows for many levels of granularity. We usually used one month as
+the length of our time windows.
+
 %   4. [ ] Cluster Windowed Set w/ LDA
@@
 %   5. [ ] Extract Clusters
+
+We take our topic clusterer and cluster the documents by their topics
+and get the $n$ unique topic clusters out. In this paper we used LDA
+but LSI could be used in a similar way.
+
+This is a very slow step as LDA takes a long time to run.
+
 \subsection{Extract Clusters}
+
 %   6. [ ] Cluster Similarity [0/1]
@@
 %   7. [ ] Continuous Cluster Analysis
+
+Once we get our clusters, we analyze them and look for repeating
+topics. We do this by taking the top 10 most common words and
+comparing them. Given a fraction such as 8/10 we joined topics where
+8/10 of the top ten topic words were shared.
+
+Clusters in the same period which are this similar are grouped
+together. Anything that matches either of those clusters 8/10 will
+match.
+
 \subsection{Continuous Cluster Analysis}
+
+We extend the cluster matching across contiguous months.q
+
 %   8. [ ] Topic Smear

### 2009-02-17 presentation/presentation.out.tex

\documentclass[titlepage,usenames,a4,landscape,semhelv]{seminar}
\newcommand{\presenter}{Abram Hindle}
\newcommand{\conference}{}
\newcommand{\gettitle}{Temporal Developer Concept Witchery: Latent Dirichlet Allocation Applied to Revisions}

\newcommand{\gettitleproper}{\gettitle}
\newcommand{\names}{Abram Hindle, Micheal W. Godfrey, Richard C. Holt}
\author{
\names \\ 
{\small Software Architecture Group }\\
\small David R. Cheriton School of Computer Science\\
\small University of Waterloo\\
\small Canada\\
ahindle@cs.uwaterloo.ca
}

\include{header}

\newcommand{\imageslide}[2]{
\newslide
\includegraphics[width=0.9\textwidth]{#2}
}

\newcommand{\figcaption}[1]{
\begin{center}
#1\\
\end{center}
}

\newcommand{\ig}[1]{\includegraphics{#1}}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
\pagestyle{fancy} %bars..
\begin{slide}

\begin{center}
{\bf \LARGE \gettitleproper }

{\names }

{\small Software Architecture Group }\\[-.5em]
{\small David R. Cheriton School of Computer Science}\\[-.5em]
{\small University of Waterloo}\\[-.5em]
{\small Canada}\\[-.5em]
{\small http://swag.uwaterloo.ca/}\\
\{ahindle,migod,holt\}@cs.uwaterloo.ca

\end{center}

\sslide{Introduction}
\begin{itemize}
\item Intro

\newslide

\begin{table*}
\centering
\begin{tabular}{|ccc|l|}

%\hline
%2000 &	Jul &	31 &		chmod \\
%2000 &	Sep &	29 &		fixes benchmark logging windows \\
%2000 &	Nov &	28 &		typo fix insert\_multi\_value \\
%2001 &	Jan &	27 &		fixes Innobase Cleanups auto-union \\
%2001 &	Mar &	28 &		2 topics bugfix, logging , TEMPORARY,	\\
\hline
2001 &	Jul &	26 &		update Allow TABLES LOCK [a] \\

2001 &	Aug &	25 &		tables row version [a] \\
\hline
%2001 &	Sep &	24 &		update checksum merge \\
%2001 &	Oct &	24 &		fixed fix \\
%2001 &	Dec &	23 &		HPUX SCO fix \\
\hline
2002 &	Feb &	21 &		net buffer length	max\_allowed\_packet [b] \\
2002 &	Mar &	23 &		small buf fix [b]	\\
%\hline
%2002 &	May &	22 &		[popular] fix SCO OSF1 table\_name \\
%2002 &	Nov &	18 &		HPUX11 compiler HP \\
\hline
\hline
2003 &	Feb &	16 &		Linux errno	[c] \\
2003 &	Mar &	18 &		alarm bookmark bug [c] \\
%\hline
%2003 &	Sep &	14 &		Auto logging merge windows distribution fix 64-bit 4.0 Cleanup \\
\hline
\end{tabular}
\caption{Continuous blocks found while tracking topics associated with the word portability in MySQL 3.23}
\label{tab:portability}
\end{table*}

\newslide

% \begin{figure}
%	 \centering
%	 \includegraphics[width=1.5\textwidth]{lda}
%	 \caption{Example of topics extracked from MySQL 3.23}
%	 \label{fig:mysql323}
% \end{figure}

\igbigslidecap{lda}{Example of topics extracted from MySQL 3.23}

\end{itemize}
\end{slide}

\end{document}

### 2009-02-17 presentation/presentation.out.tex
transitive closure diagram

@@
 \begin{itemize}
-\item Intro
+\item Stakeholders want to know what the focus of the last iteration was
+\item Evidence of work done at Retrospective meetings
+\item Difference between developer opinion and developer commits
+
+\end{itemize}
+\sslide{Topic Clustering }
+\begin{itemize}
+\item Need to track continuous topics across 
+\item Similarity between topics
+\item Clusters of the transitive closure of topics with X\% similarity
+	\begin{itemize}
+	\item Fill flood along similarity, make subsets of everyone who X\% similar to any of their neighbors, make that a cluster
+
+\newslide
+
+%\igbigslidecap{transitiveclosure}{Clustering topics (black dots) by Transitive Closure based on similarity (arcs)}
+
+\begin{center}
+\includegraphics[width=\textwidth]{transitiveclosure}
+Clustering topics (black dots) by Transitive Closure based on similarity (arcs)
+\end{center}
 
@@
 \end{itemize}
+\end{itemize}
 \end{slide}

### 2009-02-17 presentation/presentation.tex
transitive closure diagram

--- 2009-02-17 presentation/presentation.out.tex
@@
 
-\sslide{Introduction}
-\begin{itemize}
-\item Intro
+=Introduction
+* Stakeholders want to know what the focus of the last iteration was
+* Evidence of work done at Retrospective meetings
+* Difference between developer opinion and developer commits
+
+=Topic Clustering 
+* Need to track continuous topics across 
+* Similarity between topics
+* Clusters of the transitive closure of topics with X\% similarity
+** Fill flood along similarity, make subsets of everyone who X\% similar to any of their neighbors, make that a cluster
+
+\newslide
+
+%\igbigslidecap{transitiveclosure}{Clustering topics (black dots) by Transitive Closure based on similarity (arcs)}
+
+\begin{center}
+\includegraphics[width=\textwidth]{transitiveclosure}
+Clustering topics (black dots) by Transitive Closure based on similarity (arcs)
+\end{center}
 
@@
 
-\end{itemize}
+=
 \end{slide}

### 2009-02-18 presentation/presentation.out.tex
commit to topics drawing fucking with sizing and everything

@@
 
-\begin{center}
-\includegraphics[width=\textwidth]{transitiveclosure}
-Clustering topics (black dots) by Transitive Closure based on similarity (arcs)
-\end{center}
+% \begin{center}
+% \includegraphics[width=\textwidth]{transitiveclosure}
+% Clustering topics (black dots) by Transitive Closure based on similarity (arcs)
+% \end{center}
+
+%\igninetycap{transitiveclosure}{Clustering topics (black dots) by Transitive Closure based on similarity (arcs)}
+\igonecap{transitiveclosure}{Clustering topics (black dots) by Transitive Closure based on similarity (arcs)}
 
@@
 
-\begin{table*}
+%\begin{table*}
+\begin{specquotef}
 \centering
@@
 \hline
-\end{tabular}
-\caption{Continuous blocks found while tracking topics associated with the word portability in MySQL 3.23}
+\end{tabular} \\
+%\caption{Continuous blocks found while tracking topics associated with the word portability in MySQL 3.23}
+{Continuous blocks found while tracking topics associated with the word portability in MySQL 3.23}
 \label{tab:portability}
-\end{table*}
+%\end{table*}
+\end{specquotef}
 
@@
 
-\igbigslidecap{lda}{Example of topics extracted from MySQL 3.23}
+%\igbigslidecap{lda}{Example of topics extracted from MySQL 3.23}
+%\igonecap{lda}{Example of topics extracted from MySQL 3.23}
+%\igbigcapb{lda}{Example of topics extracted from MySQL 3.23}
+%\igbigthirtycapb{lda}{Example of topics extracted from MySQL 3.23}
+\igbigthirtyfivecapb{lda}{Example of topics extracted from MySQL 3.23}
 

### 2009-02-18 presentation/presentation.tex
commit to topics drawing fucking with sizing and everything

@@
 
+=Dictionary
+* Message - in this a CVS Commit
+* Word Distribution - count of words occuring in a message
+* Topic - a word distribution that is common amongst messages
+* Trend - a topic that re-occurs
+
 =Topic Clustering 

### 2009-02-18 presentation/presentation.out.tex
plots added

@@
-\documentclass[titlepage,usenames,a4,landscape,semhelv]{seminar}
+\documentclass[titlepage,usenames,a4,landscape,semhelv,16pt]{seminar}
 \newcommand{\presenter}{Abram Hindle}
@@
 \end{itemize}
+\sslide{Dictionary}
+\begin{itemize}
+\item Message - in this a CVS Commit
+\item Word Distribution - count of words occuring in a message
+\item Topic - a word distribution that is common amongst messages
+\item Trend - a topic that re-occurs
+
+\newslide
+\igbigcapb{commit-to-topics}{How commits are aggregated into Topics}
+
+\end{itemize}
+\sslide{Developer Trends}
+\begin{itemize}
+\item Trends
+	\begin{itemize}
+	\item Continuous or repeating topics
+	\item Occur over multiple time windows
+
+\end{itemize}
+\end{itemize}
+\sslide{Topic Similarity}
+\begin{itemize}
+\item Given 2 topics
+	\begin{itemize}
+	\item if their top 10 words 
+		\begin{itemize}
+		\item share X words (7 or more)
+			\begin{itemize}
+			\item they are similar
+
+		\end{itemize}
+	\end{itemize}
+\end{itemize}
+\end{itemize}
+\sslide{LDA}
+\begin{itemize}
+\item Latent Dirichlet Allocation
+\item Like Latent Semantic Indexing (LSI)
+\item Like ICA/PCA
+\item Black box view:
+	\begin{itemize}
+	\item Input word distributions of messages, 
+	\item Output topics consisting of a word distribution and related messages
+
+\end{itemize}
+\end{itemize}
 \sslide{Topic Clustering }
@@
 \igonecap{transitiveclosure}{Clustering topics (black dots) by Transitive Closure based on similarity (arcs)}
+
+\end{itemize}
+\end{itemize}
+\sslide{Case Study: Portability}
+\begin{itemize}
+\item MySQL 3.23 
+	\begin{itemize}
+	\item Discussions relating to portability
+	\item Some topics smeared across multiple months
 
@@
 \end{itemize}
+\sslide{Observations}
+\begin{itemize}
+\item Topics could be named
+	\begin{itemize}
+	\item Certain words indicate a focus on
+		\begin{itemize}
+		\item Maintenance
+		\item Bug Fixing
+		\item Portability
+		\item Potential 'ilities
+
+	\end{itemize}
+\end{itemize}
+\end{itemize}
+\sslide{Observations}
+\begin{itemize}
+\item Powerlaw distribution of repeating topic counts
+	\begin{itemize}
+	\item Most occured once
+	\item A few were repeated 
+		\begin{itemize}
+		\item less than 10\% of topics reoccur
+
+	\end{itemize}
+\end{itemize}
+\end{itemize}
+\sslide{Unfinished}
+\begin{itemize}
+\item Total Topics versus Monthly Topics
+
+\end{itemize}
+\sslide{Future work}
+\begin{itemize}
+\item Extension, associate Trends or Topics with purpoes by using wordnet and Taxonomies of software quality
+
+\end{itemize}
 \end{slide}

### 2009-02-18 presentation/presentation.tex
plots added

@@
 
+\newslide
+\igbigcapb{commit-to-topics}{How commits are aggregated into Topics}
+
+=Developer Trends
+* Trends
+** Continuous or repeating topics
+** Occur over multiple time windows
+
+=Topic Similarity
+* Given 2 topics
+** if their top 10 words 
+*** share X words (7 or more)
+****  they are similar
+
+=LDA
+* Latent Dirichlet Allocation
+* Like Latent Semantic Indexing (LSI)
+* Like ICA/PCA
+* Black box view:
+** Input word distributions of messages, 
+** Output topics consisting of a word distribution and related messages
+
 =Topic Clustering 
@@
 \igonecap{transitiveclosure}{Clustering topics (black dots) by Transitive Closure based on similarity (arcs)}
+
+=Case Study: Portability
+* MySQL 3.23 
+** Discussions relating to portability
+** Some topics smeared across multiple months
 
@@
 
+\igbigthirtyfivecapb{fixed-time-smear-plot-cropped}{}
+\igbigthirtyfivecapb{fixed-time-smear-plot-scaled}{}
+
+\igbigthirtyfivecapb{histogram-cropped}{}
+\igbigthirtyfivecapb{histogram-cropped-scaled}{}
+
+%class-smear-plot-crop
+
+\igbigthirtyfivecapb{class-smear-plot-crop-scaled}{}         
+\igbigthirtyfivecapb{class-smear-plot-crop-scaled-zoomed}{}
+
+=Observations
+* Topics could be named
+** Certain words indicate a focus on
+*** Maintenance
+*** Bug Fixing
+*** Portability
+*** Potential 'ilities
+
+=Observations
+* Powerlaw distribution of repeating topic counts
+** Most occured once
+** A few were repeated 
+*** less than 10\% of topics reoccur
+
+=Unfinished
+* Total Topics versus Monthly Topics
+
+=Future work
+* Extension, associate Trends or Topics with purpoes by using wordnet and Taxonomies of software quality
+
 =

### 2009-02-18 presentation/presentation.out.tex
Renamed
more presentation changes

@@
-\documentclass[titlepage,usenames,a4,landscape,semhelv,16pt]{seminar}
+\documentclass[titlepage,usenames,a4,landscape,semhelv]{seminar}
 \newcommand{\presenter}{Abram Hindle}
@@
 
-\newslide
-\igbigcapb{commit-to-topics}{How commits are aggregated into Topics}
-
 \end{itemize}
@@
 
-\end{itemize}
-\end{itemize}
-\sslide{Topic Similarity}
-\begin{itemize}
-\item Given 2 topics
-	\begin{itemize}
-	\item if their top 10 words 
-		\begin{itemize}
-		\item share X words (7 or more)
-			\begin{itemize}
-			\item they are similar
-
-		\end{itemize}
-	\end{itemize}
 \end{itemize}
@@
 
-\newslide
+%\newslide
+
+\igonecap{transitiveclosure}{Clustering topics (black dots) by Transitive Closure based on similarity (arcs)}
+
+\end{itemize}
+\end{itemize}
+\sslide{Topic Similarity}
+\begin{itemize}
+\item Given 2 topics
+	\begin{itemize}
+	\item if their top 10 words 
+		\begin{itemize}
+		\item share X words (7 or more)
+			\begin{itemize}
+			\item they are similar
 
@@
 %\igninetycap{transitiveclosure}{Clustering topics (black dots) by Transitive Closure based on similarity (arcs)}
-\igonecap{transitiveclosure}{Clustering topics (black dots) by Transitive Closure based on similarity (arcs)}
 
+\newslide
+
+%\igbigcapb{commit-to-topics}{How commits are aggregated into Topics}
+\begin{specquotei}
+\begin{center}
+\noindent
+\includegraphics[width=1.25\textwidth]{commit-to-topics} \\
+{How commits are aggregated into Topics}
+\end{center}
+\end{specquotei}
+
+		\end{itemize}
+	\end{itemize}
 \end{itemize}
@@
 	\item Discussions relating to portability
-	\item Some topics smeared across multiple months
+	\item Some topics trends (topics smeared across multiple months)
 
@@
 
-\newslide
-
 % \begin{figure}
@@
 %\igbigthirtycapb{lda}{Example of topics extracted from MySQL 3.23}
-\igbigthirtyfivecapb{lda}{Example of topics extracted from MySQL 3.23}
+
+%\newslide
+
+\igbigthirtyfivecapb{lda}{Example of topics extracted from MySQL 3.23. This is the kind of plot we eventually want to produce: named topics and topic trends}
+
+%\newslide
+
+\end{itemize}
+\end{itemize}
+\sslide{Case Study: MaxDB 7.500}
+\begin{itemize}
+\item Autogenerated plots from MaxDB 7.500
+\item Multiple Presentations
+	\begin{itemize}
+	\item Condensed view
+	\item Trend Histogram 
+	\item Trend View
+	\item Trend Timeline
+
+\igbigcap{fixed-time-smear-plot-cropped}{Slice of Topics per Month for MaxDB 7.500}
+
+%\newslide
+
+\igonecap{fixed-time-smear-plot-scaled}{Entire View of Topics per Month for MaxDB 7.500}
+
+%\newslide
+
+%\igninetycap{histogram-cropped}{}
+
+%\newslide
+
+\igHeigthyfivecap{histogram-cropped-scaled}{Top part of trend histogram, ordered by topic occurance}
+
+\newslide
+
+\igonecap{class-smear-plot-crop-scaled}{Trend Time Line: Trends plotted over time of MaxDB 7.500}
+
+\newslide
+
+%\igbigcapb{class-smear-plot-crop-scaled-zoomed}{Zoomed in view of the repeating topics plotted}
+\begin{specquoteh}
+\begin{center}
+\includegraphics[width=1.25\textwidth]{class-smear-plot-crop-scaled-zoomed} \\
+Zoomed in view of the trend time line
+\end{center}
+\end{specquoteh}
+
+%\igbigcapb{class-smear-plot-crop-scaled-zoomed}{Zoomed in view of the repeating topics plotted}
 
@@
 \item Total Topics versus Monthly Topics
+\item Clean up plots
 
@@
 \begin{itemize}
-\item Extension, associate Trends or Topics with purpoes by using wordnet and Taxonomies of software quality
+\item Cluster Naming
+\item Associate trends with software taxonomies
+	\begin{itemize}
+	\item Extension, associate Trends or Topics with purpoes by using wordnet and Taxonomies of software quality as suggested by Ernst et al.
+\end{itemize}
+\item Validation
 
+\end{itemize}
+\sslide{Conclusions}
+\begin{itemize}
+\item Contributions
+	\begin{itemize}
+	\item Auto extract trends
+		\begin{itemize}
+		\item Present trends
+\end{itemize}
+	\item Apply topic analysis on timewindows rather than the entire project
+
+\end{itemize}
 \end{itemize}

### 2009-02-18 presentation/presentation.tex
more presentation changes

--- 2009-02-18 presentation/presentation.tex
@@
 
+=Developer Trends
+* Trends
+** Continuous or repeating topics
+** Occur over multiple time windows
+
+=LDA
+* Latent Dirichlet Allocation
+* Like Latent Semantic Indexing (LSI)
+* Like ICA/PCA
+* Black box view:
+** Input word distributions of messages, 
+** Output topics consisting of a word distribution and related messages
+
 =Topic Clustering 
@@
 
-\newslide
+%\newslide
+
+\igonecap{transitiveclosure}{Clustering topics (black dots) by Transitive Closure based on similarity (arcs)}
+
+=Topic Similarity
+* Given 2 topics
+** if their top 10 words 
+*** share X words (7 or more)
+****  they are similar
 
@@
 %\igninetycap{transitiveclosure}{Clustering topics (black dots) by Transitive Closure based on similarity (arcs)}
-\igonecap{transitiveclosure}{Clustering topics (black dots) by Transitive Closure based on similarity (arcs)}
+
+\newslide
+
+%\igbigcapb{commit-to-topics}{How commits are aggregated into Topics}
+\begin{specquotei}
+\begin{center}
+\noindent
+\includegraphics[width=1.25\textwidth]{commit-to-topics} \\
+{How commits are aggregated into Topics}
+\end{center}
+\end{specquotei}
+
+=Case Study: Portability
+* MySQL 3.23 
+** Discussions relating to portability
+** Some topics trends (topics smeared across multiple months)
 
@@
 %\igbigthirtycapb{lda}{Example of topics extracted from MySQL 3.23}
-\igbigthirtyfivecapb{lda}{Example of topics extracted from MySQL 3.23}
+
+%\newslide
+
+\igbigthirtyfivecapb{lda}{Example of topics extracted from MySQL 3.23. This is the kind of plot we eventually want to produce: named topics and topic trends}
+
+%\newslide
+
+=Case Study: MaxDB 7.500
+* Autogenerated plots from MaxDB 7.500
+* Multiple Presentations
+** Condensed view
+** Trend Histogram 
+** Trend View
+** Trend Timeline
+
+\igbigcap{fixed-time-smear-plot-cropped}{Slice of Topics per Month for MaxDB 7.500}
+
+%\newslide
+
+\igonecap{fixed-time-smear-plot-scaled}{Entire View of Topics per Month for MaxDB 7.500}
+
+%\newslide
+
+%\igninetycap{histogram-cropped}{}
+
+%\newslide
+
+\igHeigthyfivecap{histogram-cropped-scaled}{Top part of trend histogram, ordered by topic occurance}
+
+\newslide
+
+\igonecap{class-smear-plot-crop-scaled}{Trend Time Line: Trends plotted over time of MaxDB 7.500}
+
+\newslide
+
+%\igbigcapb{class-smear-plot-crop-scaled-zoomed}{Zoomed in view of the repeating topics plotted}
+\begin{specquoteh}
+\begin{center}
+\includegraphics[width=1.25\textwidth]{class-smear-plot-crop-scaled-zoomed} \\
+Zoomed in view of the trend time line
+\end{center}
+\end{specquoteh}
+
+%\igbigcapb{class-smear-plot-crop-scaled-zoomed}{Zoomed in view of the repeating topics plotted}
+
+=Observations
+* Topics could be named
+** Certain words indicate a focus on
+*** Maintenance
+*** Bug Fixing
+*** Portability
+*** Potential 'ilities
+
+=Observations
+* Powerlaw distribution of repeating topic counts
+** Most occured once
+** A few were repeated 
+*** less than 10\% of topics reoccur
+
+=Unfinished
+* Total Topics versus Monthly Topics
+* Clean up plots
+
+=Future work
+* Cluster Naming
+* Associate trends with software taxonomies
+** Extension, associate Trends or Topics with purpoes by using wordnet and Taxonomies of software quality as suggested by Ernst et al.
+* Validation
+
+=Conclusions
+* Contributions
+** Auto extract trends
+*** Present trends
+** Apply topic analysis on timewindows rather than the entire project
 

### 2009-03-13 lda-paper.tex
abstract

@@
 }
-\title{Temporal Developer Concept Witchery: Latent Dirichlet Allocation Applied to Revisions}
+\title{Windowed Developer Topic Analysis}
 
@@
 \begin{abstract}
+
+  At different times, developers focus on different topics and tasks
+  during development.  These different topics and times of focus might
+  be obvious to developers familar with the project, but other
+  stakeholders, like managers, might be less aware of the topics of
+  development. Imagine the scenario where developers had to deal with
+  issues orthogonal to the features suggested, at the end of the
+  iteration a manager would wonder where all the time went, the
+  developers might have just blocked all sides issues in with features
+  they were initially working on, they might not have noticed their
+  focus had changed.  Currently it is difficult for stakeholders to
+  audit what actually occurred and when certain features or topics
+  were focussed on. We propose and provide a visualization of the
+  changing development topics over time. Using
 

### 2009-03-13 lda-paper.tex
more lda paper changes abstract

@@
   were focussed on. We propose and provide a visualization of the
-  changing development topics over time. Using
+  changing development topics over time. Our approach takes a topic
+  analysis tool like Latent Dirichlet Allocation (LDA) or Latent
+  Semantic Indexing (LSI) and instead of applying the tool to the
+  entire corpus we apply to windows of the corpus. This allows us to
+  see the unique topics of development occuring during a particular
+  time period and then see if these individual topics ever reappear.
 

### 2009-03-14 lda-paper.tex
more changes to intro
lda paper

abram

@@
 
-  In this lab report we seek to explore how topics shift over time in
-  the source control repository of a project. We analyze commit
-  comments, and even source code per revision per each time window and
-  extract the topics prevelant in that window. We expect that topic
-  sets will change over time and the change in topic sets indicate a
-  change in the focus of development. We suppose this could help us to
-  segment the revisions on the time line to discover parts of an
-  iteration. For topic extraction we use \lda (LDA).
+%   In this lab report we seek to explore how topics shift over time in
+%   the source control repository of a project. We analyze commit
+%   comments, and even source code per revision per each time window and
+%   extract the topics prevelant in that window. We expect that topic
+%   sets will change over time and the change in topic sets indicate a
+%   change in the focus of development. We suppose this could help us to
+%   segment the revisions on the time line to discover parts of an
+%   iteration. For topic extraction we use \lda (LDA).
 
@@
 release of the software? What were the topics of the previous release?
+Given the requirements agreed to at the start of the iteration did our
+developers work on them? What else did they work on. Were there issues
+which were orthogonal to the
+
+Topic analysis, with respect to Software Control Systems (SCSs), is
+valuable as it has many uses for the stakeholders involved.
+
+Example scenarios where the usage of this technique would prove useful
+are numerous. Imagine a development team has agreed to implement some
+features at the start of an iteration. By the end of the iteration
+only some of the agreed upon features are implemented. The developers
+swear they were focusing solely on the features agreed upon. Topic
+analysis could help by teasing out the distinct topics or focusses of
+the developers. Perhaps the developers had to deal with issues
+orthogonal to the features they were working on, but were probably
+necessary to complete such features. Developers might have grouped
+those orthogonal changes into one of the features and might not
 
@@
 
-\section{Previous Work}
+\section{Background}
+
+Some terms we will be using for the rest of the paper include:
+message, word distribution, topic and trend. Messages are block of
+text written by developers, in this paper messages will be the CVS
+commit comments. Word distributions are the summary of a message by
+word count, effectively the word distribution will be over all
+messages looked at but since most words will not appear in a message,
+a word distribution is effectively a word count divided by the message
+size. A topic is a unique word distribution, but effectively a set of
+words unique to an independant behaviour in the text corpus, in this
+paper we often summarize topics by their top 10 more frequent words.
+A trend is a one or more similar topics which reoccur over
+time. Trends are important because they are sets of topics which
+reoccur over the development of a project.
 

### 2009-03-16 lda-paper.tex
more changes

@@
 Topic analysis, with respect to Software Control Systems (SCSs), is
-valuable as it has many uses for the stakeholders involved.
-
-Example scenarios where the usage of this technique would prove useful
-are numerous. Imagine a development team has agreed to implement some
+valuable as it has many uses for the stakeholders involved.  Example
+scenarios where the usage of these techniques would prove useful are
+numerous. Imagine a development team has agreed to implement some
 features at the start of an iteration. By the end of the iteration
@@
 necessary to complete such features. Developers might have grouped
-those orthogonal changes into one of the features and might not
+those orthogonal changes into one of the features they were working
+on, thus when asked about how their time was spent the orthogonal
+changes would not be covered. This is what we hope that windowed topic
+analysis can uncover. If windowed topic analysis can discover
+important but overlooked topics, it can help both developers and
+managers.
 
@@
 \section{Validity Threats}
+
+In this study we are explicitly trusting that the programmers annotate
+their changes with relevant information. We rely on the descriptions
+they provide. If the language of checkin comments was automated
+
 %   1.   validation

### 2009-03-16 lda-paper.tex
lda paper

@@
 important but overlooked topics, it can help both developers and
-managers.
+managers. Developer can reflect and see a good summary of what worked
+on, and managers can see the focus of development while still
+remaining relatively high level.
 
@@
 them over windows of documents over time. We hope by looking at topic
-clusters over time we can I identify continuous effort on a select
+clusters over time we can identify continuous effort on a select
 group of topics.
@@
 inside of commits.  Sometimes a single revision to a file contains
-multiple bug fixes.  Using powerful topic analysis like Latent
-Semantic Indexing (LSI) or Latent Dirichlet Allocation (LDA) one can
-seperate these mixed signals into coherent topics. In this case a
-topic is a set of tokens that are related to each other.
+multiple bug fixes.  Using powerful topic analysis techniques like
+Latent Semantic Indexing (LSI) or Latent Dirichlet Allocation (LDA)
+one can seperate these mixed signals into coherent topics. In this
+case a topic is a set of tokens that are related to each other.
 
@@
 
-Topic analysis is useful because it can help us partition a project's
-time-line into sub-iterations. By breaking apart an iteration we might
-be able to extract the underlying software development process from
-these artifacts.
+Another use of topic analysis is to help us partition a project's
+time-line into sub-iterations of developer focus. By breaking apart an
+iteration we might be able to extract the underlying software
+development process from these artifacts or at the very least what
+maintenance was occurring in the repository at that time.
 
-In this paper we seek to explore how topics shift over time in
-the source control repository of a project. We analyze commit
-comments, and even source code per revision per each time window and
-extract the topics in that time window. We expect that topic sets
-will change over time and the change in topic sets indicate a change
-in the focus of development.
+In this paper we seek to explore how topics shift over time in the
+source control repository of a project. We analyze commit comments per
+revision in each time window and we extract the topics in that time
+window. We expect that topics in each window will change over time and
+the similarities and differences in these sets of topics will indicate
+both developer focus and changes in developer focus.
 
@@
 Latent Dirichlet Alloation (LDA)~\cite{944937} is a competitor to
-latent semantic indexing (LSI), probabilistic latent semantic indexing
+Latent Semantic Indexing (LSI), probabilistic Latent Semantic Indexing
 (pLSI) and semantic clustering. It is used for document modelling,
@@
 
-\section{Methodology}
-
-\subsection{Extract Repos}
-
-We extracted the repositories with software such as rsync, CVSSuck,
-\texttt{softChange} and bt2csv.  \texttt{softChange} provided a
-general schema for storing revisions. CVSSuck and rsync mirrored CVS
-repositories while bt2csv mirrored web accessible bitkeeper
-repositories.
-
-%   2. [ ] Extract Documents
-\subsection{Extract Documents}
-%   3. [ ] Windowed Sets
-
-The document we are primarily using are the changelog comments, these
-are the comments added when a revision to the source code is
-committed, it could be per file or per commit of changes to multiple
-files, this depends on the source control system used.
-
-Documents are converted into word count distributions. The
-distributions include all of the possible words in the system.
-
-\subsection{Windowed Sets}
-
-We then create windows which are sets of documents grouped by time
-window. Time windows can overlap but don't have to. Windowing by time
-allows for many levels of granularity. We usually used one month as
-the length of our time windows.
-
-%   4. [ ] Cluster Windowed Set w/ LDA
-\subsection{Cluster Windowed Set w/ LDA}
-%   5. [ ] Extract Clusters
-
-We take our topic clusterer and cluster the documents by their topics
-and get the $n$ unique topic clusters out. In this paper we used LDA
-but LSI could be used in a similar way.
-
-This is a very slow step as LDA takes a long time to run.
-
-\subsection{Extract Clusters}
-
-%   6. [ ] Cluster Similarity [0/1]
-\subsection{Cluster Similarity}
-%      1. [ ] need the tool
-%             grab from the cluster analysis
-%   7. [ ] Continuous Cluster Analysis
-
-Once we get our clusters, we analyze them and look for repeating
-topics. We do this by taking the top 10 most common words and
-comparing them. Given a fraction such as 8/10 we joined topics where
-8/10 of the top ten topic words were shared.
-
-Clusters in the same period which are this similar are grouped
-together. Anything that matches either of those clusters 8/10 will
-match.
-
-\subsection{Continuous Cluster Analysis}
-
-We extend the cluster matching across contiguous months.q
-
-%   8. [ ] Topic Smear
-\subsection{Topic Smear}
-%   9. [ ] Visualization [0/1]
-\subsection{Visualization}
-%       1. [ ] need the visualizer
-
-\subsection{Datasets and Tools}
-
-To extract data from CVS and BitKeeper we used softChange to extract
-CVS repositories and provide a schema for our revision data. We used
-bt2csv to extract BitKeeper revisions from BitKeeper web repositories.
-
-% Extractors:
-% \begin{itemize}
-% \item CVSSuck - CVS Suck mirrors RCS files from a CVS repository. 
-% \item softChange - extract CVS facts to a PostgreSQL database.
-% \item bt2csv - Convert BitKeeper repositories to facts in CSV
-%   databases.
-% \end{itemize}
-
-Our analysis tools consisted of: Hiraldo-Grok, an OCaml based spin off
-of Grok used for answering queries; Gnuplot, a graph plotting package,
-lda-c a LDA package by Biel
-
-%   1. [ ] LDA
-
-%   2. [ ] Extractor
-
-%   3.   Similarity
-
-%   4.   Smearing
-
-%   5.   Visualizer
-
 \section{First Impressions}
@@
 \ref{tab:portability}. We tracked one topic cluster which seemed
-related to portability.
+related to portability. We picked out terms which seem to summarize
+those topics.
 
@@
 
+\section{Methodology}
+
+\subsection{Extract Repos}
+
+We extracted the repositories with software such as rsync, CVSSuck,
+\texttt{softChange} and bt2csv.  \texttt{softChange} provided a
+general schema for storing revisions. CVSSuck and rsync mirrored CVS
+repositories while bt2csv mirrored web accessible bitkeeper
+repositories.
+
+%   2. [ ] Extract Documents
+\subsection{Extract Documents}
+%   3. [ ] Windowed Sets
+
+The document we are analyzing are the changelog comments, these are
+the comments added when a revision to the source code is committed.
+Documents are converted into word count distributions. These
+distributions are can be normalized by by the size of the
+message. After all messages are processes the distributions are
+extended to include all words in all distributions (each distribution
+is extended with word bins of count 0).
+
+\subsection{Windowed Sets}
+
+Given all messages, we group the messages into windows. We could use
+overlapping windows, but in this paper we use non-overlapping windows
+of a set time unit.  Windowing by time allows for many levels of
+granularity. We usually used one month as the length of our time
+windows, while we could use different granularities for the study we
+think that a month is a sizable unit of development (4 weeks of
+development).
+
+\subsection{Apply Topic Analysis to each Window}
+
+Once we have our data windowed, we apply our Topic Analysis tool to
+each window and extract the top $N$ topic clusters, we used $20$
+topics in our case study. Our Topic Analysis tool is an implementation
+of LDA, but we could've used other similar tools like LSI.
+
+This is a very slow step as LDA takes a long time to run, and takes
+even longer since we run it once per window.
+
+%\subsection{Extract Clusters}
+
+%   6. [ ] Cluster Similarity [0/1]
+\subsection{Cluster Similarity}
+%      1. [ ] need the tool
+%             grab from the cluster analysis
+%   7. [ ] Continuous Cluster Analysis
+
+Once we have our topics extracted for each window , we analyze them
+and look for repeating topics. We do this by taking the top 10 most
+common words and then comparing them.
+
+We then abstract these topics by comparing them to each other.  Each
+topic is compared to each other topic in the system, given a certain
+threshold of top 10 similarity, like 8/10 words. Then given this topic
+similarity matrix, we find the transitive closure of similar topics,
+that is we fill flood along the similarity arcs until we have
+partitioned the topics into clusters of similar topics.  These
+clusters of topics are called trends. A trend is probably interesting
+and valuable if there is more than 1 topic in the trend.
+
+This technique has weaknesses in that nodes a few neighbors away in
+similarity might not share any similar words.  We do this because we
+want to be able to color or highlight topics which are related to each
+other and track them throughout time.
+
+Once we have determined our similarity clusters we can choose to
+analyze and to plot the topics.
+
+\begin{comment}
+Topic Clustering
+ · Need to track continuous topics across
+ · Similarity between topics
+ · Clusters of the transitive closure of topics with X%
+   similarity
+    ­ Fill flood along similarity, make subsets of
+      everyone who X% similar to any of their neighbors,
+      make that a cluster
+\end{comment}
+
+\subsection{Visualization}
+%       1. [ ] need the visualizer
+
+We have come up with multiple ways to visualize these topics and
+trends.  For all of these plots if we trends which have continuous
+segments we plot those segments as joined across time.  We have a
+compressed stacked trend view, which displays trends arbitrarily
+across the y axis but are placed where they occur on the time-based x
+axis. Another plot we have is the trend-time plot, a each trend gets
+it own distinct-y value , while the x-axis is time; these topics are
+then each plotted on their own line across time as they occur. The
+last plot we have is the histogram plot, where we plot each trend on
+its own line but stack up the segments of the trend, much like a plot
+of counts or a histogram.
+
+The first visualization is a compressed view, we try to sink the
+larger trends to the bottom, once the larger trends are stacked, we
+fill in the gaps with the smaller trends in the same window and then
+stack the smaller trends on top. Although there is a chance that gaps
+will be left due to nonoptimal stacking in practice there are lot of
+small trends (90\% of all trends contain 1 topic) which fill in these
+gaps quickly. The trend segments that share the same color are the
+same trend, colors are randomly chosen for each trend, so color
+similarity is meaningless, only exact color matching matters.
+The compressed view is nice because it shows all of the
+information and the repeating continuous trends are easy to pick out,
+unfortunately repeating discontinuous trends are harder to spot.
+
+The second visualization is the trend-time plot, it displays repeating
+trends more clearly by dedicating a horizonal line for trend segments
+dedicated to one trend. This means that if a trend contains
+discontinuous segments the segments appear on the same line. One down
+side to this view is the least common trends need to be pruned
+otherwise it'd take too many horizontal lines to plot them.
+
+The 3rd visualization is the histogram plot which is a lot like the
+time-trend plot except the trends are plotted together by stacking to
+their row so time information is lost and the trends are ordered by
+the number of topics in the trend. The histogram plot shows the count
+of instances of a trend and thus indicates which trends occur the
+most. Unfortunately due to the large number of topics given allotted
+space it is often best to crop off the trends with only one topic,
+otherwise the tail is long.
+
+\subsection{Datasets and Tools}
+
+To extract data from CVS and BitKeeper we used softChange to extract
+CVS repositories and provide a schema for our revision data. We used
+bt2csv to extract BitKeeper revisions from BitKeeper web repositories.
+
+We extracted the repositories of: MySQL, PostGreSQL, MaxDB and
+Firebird.
+
+% Extractors:
+% \begin{itemize}
+% \item CVSSuck - CVS Suck mirrors RCS files from a CVS repository. 
+% \item softChange - extract CVS facts to a PostgreSQL database.
+% \item bt2csv - Convert BitKeeper repositories to facts in CSV
+%   databases.
+% \end{itemize}
+
+Our analysis tools consisted of: Hiraldo-Grok, an OCaml based spin off
+of Grok used for answering queries; Gnuplot, a graph plotting package;
+lda-c, a LDA package implemented by Biel; and our topic plotter,
+implemented in Haskell.
+
+%   1. [ ] LDA
+
+%   2. [ ] Extractor
+
+%   3.   Similarity
+
+%   4.   Smearing
+
+%   5.   Visualizer
+
 %LDA Skeleton
@@
 %   2. [ ] LDA Work [0/6]
+
 \section{Results}
@@
 %7.   Validation [0/1]
+
+\subsection{Compare to topics over entire development}
+
 \subsection{Validation}
+
 %   1. well i'm stumped
@@
 they provide. If the language of checkin comments was automated
+
+We truncated the topics to top 10 tokens, this summary of a topic
+might not have been as useful as determining the actual topic distance
+between two word topic distributions.
 

### 2009-03-16 lda-paper.tex
changed the lda paper more, added much more stuff very draft like including new figures ripped from presentation

@@
 \documentclass[times, 10pt,twocolumn]{article} 
-\usepackage{times}
+\usepackage{latex8}
 
@@
 }
+
 \title{Windowed Developer Topic Analysis}
@@
 \begin{document}
+
+\newcommand{\affaddr}[1]{#1}
+\newcommand{\aemail}[1]{#1}
+%\numberofauthors{5}
+\author{
+%\alignauthor
+Abram Hindle\\
+%\affaddr{David Cheriton School of Computer Science}\\
+\affaddr{University of Waterloo}\\
+\affaddr{Waterloo, Ontario}\\
+\affaddr{Canada}\\
+\aemail{ahindle@cs.uwaterloo.ca}
+%\alignauthor
+\and
+Michael W. Godfrey\\
+%\affaddr{David Cheriton School of Computer Science}\\
+\affaddr{University of Waterloo}\\
+\affaddr{Waterloo, Ontario}\\
+\affaddr{Canada}\\
+\aemail{migod@cs.uwaterloo.ca}
+%\alignauthor
+\and
+Richard C. Holt\\
+%\affaddr{David Cheriton School of Computer Science}\\
+\affaddr{University of Waterloo}\\
+\affaddr{Waterloo, Ontario}\\
+\affaddr{Canada}\\
+\aemail{holt@cs.uwaterloo.ca}
+%\alignauthor
+%\and
+%Gregorio Robles\\
+%\affaddr{GSyc}\\
+%\affaddr{Universidad Rey Juan Carlos}\\
+%\affaddr{Madrid}\\
+%\affaddr{Spain}\\
+%\aemail{grex@gsyc.escet.urjc.es}
+}
 
@@
 
+\begin{figure*}
+  \centering
+  \includegraphics[width=0.9\textwidth]{lda}
+  \caption{Example of topics extracted from MySQL 3.23. This is the kind of plot we eventually want to produce: named topics and topic trends}
+  \label{fig:lda}
+\end{figure*}
+
 \section{Methodology}
+\begin{figure*}
+  \centering
+  \includegraphics[width=0.9\textwidth]{commit-to-topics} 
+  \caption{How commits are analyzed and aggregated into Topics and Trends}
+  \label{fig:commits}
+\end{figure*}
 
@@
 
+\begin{figure}
+  \centering
+  \includegraphics[width=0.45\textwidth]{transitiveclosure}
+  \caption{Clusting topics by the transitive closure of topic similarity. Nodes are topics and arcs imply similarity}
+\label{fig:closure}
+\end{figure}
+
 \subsection{Visualization}
@@
 otherwise the tail is long.
+
+\begin{figure*}
+  \centering
+  \includegraphics[width=1.0\textwidth]{fixed-time-smear-plot-scaled}
+  \caption{Compact Trend Plot Topics per Month for MaxDB 7.500}
+  \label{fig:topicsmear}
+\end{figure*}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=0.45\textwidth]{fixed-time-smear-plot-cropped}
+  \caption{Zoomed in Slice Compact Trend Plot of Topics per Month for MaxDB 7.500}
+  \label{fig:zoomedsmear}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=0.45\textwidth]{class-smear-plot-crop-scaled}
+  \caption{Trend Time Line: Trends plotted over time of MaxDB 7.500}         
+  \label{fig:trendtimeline}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=0.45\textwidth]{histogram-cropped-scaled}
+  \caption{Top part of trend histogram, ordered by topic occurance for MaxDB 7.500}       
+  \label{fig:histogram}
+\end{figure}
 
@@
 %   4.   firebird
+
+XXX need results!
+
 \subsection{Firebird}
 %   5.   what's that other db?
-\subsection{ what's that other db?}
-%7.   Validation [0/1]
+
+XXX needs results
+
+\subsection{MaxDB 7.500}
 
 \subsection{Compare to topics over entire development}
+\begin{figure}
+  \centering
+  \includegraphics[width=0.45\textwidth]{maxdb7500-everything-by-month}
+  \caption{MaxDB 7.500 Topics Analyzed with Static Topic Analysis, 20
+    topics over the entire development history of MaxDB 7.500. Each
+    box indicates the number of documents matching that topic in that
+    month.}
+  \label{fig:statictopics}
+\end{figure}
 
@@
 %9.    Future Work
-\section{ Future Work}
+
 %10.   Conclusions
+
 \section{Conclusions}
@@
 
-\bibliographystyle{alpha}
+\subsection{ Future Work}
+
+One avenue of future work we think is quite important is automatic
+cluster labelling. Given a word distribution we should be able to
+figure out a word or term which best describes that distribution
+automatically.
+Another avenue we wish to investigate the is value of applying
+static community generated software taxonomies to the naming or
+grouping of clusters
+
+\bibliographystyle{latex8}
 \bibliography{lda-paper}
 
+%\bibliographystyle{abbrv}
+
+\section{Appendix}
+
+\begin{itemize}
+\item We are missing a comparison of our technique to the classical technique
+\end{itemize}
+
 \end{document}

### 2009-03-19 lda-paper.tex
More changes to stuff

@@
 \subsection{Compare to topics over entire development}
+
+The previous work which leverage LSI or LDA usually extracted a
+certain number of topics and tracked them over time. This meant a
+static set of topics was tracked across the whole development history
+of a project. Thus we really had 20 trends.
+
+We attempted the same kind of analysis, as shown in
+Figure~\ref{fig:statictopics} we extracted 20 topics and then plotted
+the number of messages per month which related to that topic. What we
+found was that often one topic would dominate the results, while
+messages related to other topics didn't appear that often.
+
+This technique seems useful if the 20 extracted topics are relevant
+and useful, that said one weakness of this approach is that documents
+for a topic might only occur in one time window, yet we're tracking
+messages that could be related to it over the entire lifetime of the
+project.
+
 \begin{figure}

### 2009-03-20 lda-paper.tex
More lda paper

@@
 \section{Results}
+
+We applied our tools and visualizations to the repositories of
+multiple database systems.
+
 %   1. Interesting smears 
@@
 \subsection{MaxDB 7.500}
+
+The plots we produced of MaxDB 7.500 were interesting, there was a
+period where no real development occured and thus there were no topics
+whatsoever. We have two periods to evaluate for MaxDB 7.500, the first
+period from June 2004 to jan 2005, and the second period from June
+2005 to June 2006.
+
+The largest common trend has references to build system files like
+SYSDD.punix and MONITOR.punix etc. Other tokens mentioned are Sutcheck
+v1.09, the prefix SUT stands for Storable Unit Type. Sutcheck would
+also automate perforce check-ins.
+
+The second largest common trend seems to be a side effect of an
+automated checkin which is annotated as ``implicit checkin''. These
+were checkins that were produced when importing changes from an
+external Perforce repository.
+
+The third most common trend, seen on Figure \ref{fig:trend}, seemed to
+include tokens related to operating system support, such as Linux and
+Windows, as well as architecture support, AMD64 and Opteron. The word
+problem was common among all of those trends. This trend seemed
+related to the smaller fourth largest trend which occurred nearby it
+that had AMD64 and Windows tokens.
 
@@
 their changes with relevant information. We rely on the descriptions
-they provide. If the language of checkin comments was automated
+they provide. If the language of checkin comments was automated we
+would only be analyzing that.
 
@@
 might not have been as useful as determining the actual topic distance
-between two word topic distributions.
+between two word topic distributions. By truncating tokens we could be
+losing valuable data.
+
+Each project seemed to have their own kind of stop words, we did not
+remove them, but perhaps dropping some of these stop words would aid
+the clarity of such topic clusters, alternatively it might cause our
+topic similarity plots to be fundamentally different. Perhaps our
+methods of suggesting that topics are similar is not appropriate.
+
+The number of commits per month is inconsistent some months have a lot
+of changes where as some months have almost none.
 

### 2009-03-27 lda-paper.tex
described firebird

@@
 %   3.   mysql
-\subsection{Mysql}
-%   4.   firebird
-
-XXX need results!
 
@@
 
-XXX needs results
+%XXX needs results
+
+We tracked Firebird from August 2000 to January 2006.
+Our first attempt with Firebird resulted in a set of trends that were
+dominated by one large trend containing tokens which were mostly stop
+words such as ``the to a for of is it'' and references to one author
+\texttt{carlosga05}.
+
+The second largest trend related to branching and possibly merging, it
+contained language such as ``was added on branch'' and contained
+references to tags and branches.
+
+The third largest trend contained reference to the changelog, the
+build system, architecture and platform support. It specifically
+mentioned tokens such as ``Solaris port x86''.
+
+Other trends included topics regarding AIX PPC compilation, updating
+the build process, internationalization and UTF8, Darwin build support
+and bug fixing.
+
+Topics that were not trends that were interesting were external bug
+fixes submitted to the project, the developers would write thank yous
+in their changelog comments: ``thanks Bill Lam''. Other interesting
+topics include tokens and topics such as: ``compiler workarounds'', ``nightly updates'', packets and MSVC++,
 
@@
 The plots we produced of MaxDB 7.500 were interesting, there was a
-period where no real development occured and thus there were no topics
+period where no real development occured and thus there were no topics or trends
 whatsoever. We have two periods to evaluate for MaxDB 7.500, the first

### 2009-03-29 lda-paper.tex
minor changes
postgresql

@@
 
-We extracted the repositories of: MySQL, PostGreSQL, MaxDB and
+We extracted the repositories of: PostGreSQL, MaxDB and
 Firebird.
@@
 %   3.   mysql
+
+PostgreSQL did not have many trends with two or more topics, but it
+did have one large mega-trend which stretched across most of its
+existence. Much of the mega-trend consisted of stop words but it did
+include topics which had words such as patch, fix, update.
+
+The second largest topic was interesting because it dealt changes from
+1998, from developers using Central Europe Time (CEST), the two
+notable tokens were 1998 and CEST. These changes seem to be patches
+submitted by one particular developer who used that timezone.
+
+The third largest trend seemed to be about copying and moving files
+between branches as it referenced the old branch. This is a side
+effect of the programmers manually recording such a merge across
+branches because CVS cannot.
 

### 2009-03-31 lda-paper.tex
modified the lda paper some more?

@@
 
+Once stop words were removed the some of hte purposes became more
+clear. The mega trend disappeared, the largest trend became bug fixes
+submitted by Dal Zotto which is a short trend.
+
+% Large Trends is not the point!!!
+% That's what normal LDA gives us
+
+But large trends are not the point, by investigating windowed trends
+we want more local information about trends being worked on.
+
 \subsection{Firebird}

### 2009-04-01 lda-paper.tex
SOme conclusions

@@
 both developer focus and changes in developer focus.
+
+Our contributions include:
+\begin{itemize}
+\item Demonstrating the value of windowed topic analysis
+\item Multiple visualizations of trends over time
+\end{itemize}
 
@@
 
+Even with our liberal topic similarity metrics which produced both
+long and short and trends we showed that there are only a few trends
+in a repository which re-appear. Since so few trends re-appear and so
+many trends appear only once this suggests that global topic analysis
+might be ignoring local unique topics.
+
 \subsection{ Future Work}

### 2009-04-03 lda-paper.tex
ric edits

--- 2009-03-31 lda-paper.tex
@@
   At different times, developers focus on different topics and tasks
-  during development.  These different topics and times of focus might
+  during development.  These different topics and tasks might
   be obvious to developers familar with the project, but other
   stakeholders, like managers, might be less aware of the topics of
-  development. Imagine the scenario where developers had to deal with
+  development. Imagine the scenario where developers unexpectedly had to deal with
   issues orthogonal to the features suggested, at the end of the
-  iteration a manager would wonder where all the time went, the
-  developers might have just blocked all sides issues in with features
-  they were initially working on, they might not have noticed their
-  focus had changed.  Currently it is difficult for stakeholders to
+  iteration a manager would wonder where all the effort went to.
+  Currently it is difficult for stakeholders to
   audit what actually occurred and when certain features or topics
-  were focussed on. We propose and provide a visualization of the
-  changing development topics over time. Our approach takes a topic
+  were focussed on. We propose a visualization of the
+  changing development topics over time. Our approach uses a topic
   analysis tool like Latent Dirichlet Allocation (LDA) or Latent
-  Semantic Indexing (LSI) and instead of applying the tool to the
-  entire corpus we apply to windows of the corpus. This allows us to
-  see the unique topics of development occuring during a particular
-  time period and then see if these individual topics ever reappear.
+  Semantic Indexing (LSI). Instead of applying the tool to the
+  entire corpus we apply it to windows (periods) of the corpus. This allows us to
+  see the evolving stream of topics of development occuring
+  over time.
 
@@
 
-What happened in the previous iteration? What were the developers
+What happened in the previous maintenance or development iteration? What were the developers
 working on? What was the developer in the next cubicle working on?
@@
 Given the requirements agreed to at the start of the iteration did our
-developers work on them? What else did they work on. Were there issues
-which were orthogonal to the
+developers work on them? What else did they work on.
 
-Topic analysis, with respect to Software Control Systems (SCSs), is
-valuable as it has many uses for the stakeholders involved.  Example
-scenarios where the usage of these techniques would prove useful are
-numerous. Imagine a development team has agreed to implement some
-features at the start of an iteration. By the end of the iteration
-only some of the agreed upon features are implemented. The developers
-swear they were focusing solely on the features agreed upon. Topic
-analysis could help by teasing out the distinct topics or focusses of
-the developers. Perhaps the developers had to deal with issues
-orthogonal to the features they were working on, but were probably
-necessary to complete such features. Developers might have grouped
-those orthogonal changes into one of the features they were working
-on, thus when asked about how their time was spent the orthogonal
-changes would not be covered. This is what we hope that windowed topic
-analysis can uncover. If windowed topic analysis can discover
-important but overlooked topics, it can help both developers and
-managers. Developer can reflect and see a good summary of what worked
-on, and managers can see the focus of development while still
-remaining relatively high level.
+Topic analysis, with respect to Software Control Systems (SCSs) like
+CVS, Subversion, Git etc., is potentially valuable as it has many uses
+for various stakeholders and uses.
+Imagine a
+development team has agreed to implement some features at the start of
+an iteration. By the end of the iteration only some of the agreed upon
+features are implemented. The developers swear they were focusing
+solely on the features agreed upon. Topic analysis could help by
+teasing out the distinct topics or focusses of the developers. Perhaps
+the developers had to deal with issues.
+Developers might have grouped those orthogonal changes into
+one of the features they were working on, thus when asked about how
+their time was spent the orthogonal changes would not be covered. This
+is what we hope that windowed topic analysis can uncover. If windowed
+topic analysis can discover important but overlooked topics, it can
+help both developers and managers. Developer can see a
+summary of what worked on, and managers can development.
 
-We propose to extend common concept analysis techniques by executing
+Topic analysis of revisions in SCSs is important as often revisions
+belonging to different topics are interleaved or even mixed inside of
+commits.  Sometimes a single revision to a file contains multiple bug
+fixes.  Using powerful topic analysis techniques like Latent Semantic
+Indexing (LSI) or Latent Dirichlet Allocation (LDA) one can seperate
+these mixed comments into topics. These topics describe seperate ideas
+that are interleaved throughtout the development. They indicate groups
+of comments which are related to each other by their language, often
+the content of the comments are related.  In this case a topic is a
+set of tokens that are related to each other.
+
+% We suppose this could help us to segment the revisions on the time
+% line to discover parts of an iteration.
+
+One use of topic analysis is to help us partition a project's
+time-line into periods  of developer focus. By breaking apart an
+iteration we may be able to recognize the underlying software
+development process from these revision or at the very least what
+maintenance was occurring in the repository at a particular time.
+
+We propose to extend the use of common topic analysis techniques by executing
 them over windows of documents over time. We hope by looking at topic
@@
 
-Topic analysis of fine-grained revisions is important as often
-revisions belonging to different topics are interleaved or even mixed
-inside of commits.  Sometimes a single revision to a file contains
-multiple bug fixes.  Using powerful topic analysis techniques like
-Latent Semantic Indexing (LSI) or Latent Dirichlet Allocation (LDA)
-one can seperate these mixed signals into coherent topics. In this
-case a topic is a set of tokens that are related to each other.
-
-% We suppose this could help us to segment the revisions on the time
-% line to discover parts of an iteration.
-
-Another use of topic analysis is to help us partition a project's
-time-line into sub-iterations of developer focus. By breaking apart an
-iteration we might be able to extract the underlying software
-development process from these artifacts or at the very least what
-maintenance was occurring in the repository at that time.
-
-In this paper we seek to explore how topics shift over time in the
+In this paper we explore how topics shift over time in the
 source control repository of a project. We analyze commit comments per
@@
 
+Our contributions include:
+\begin{itemize}
+\item Demonstrating the value of windowed topic analysis
+\item Multiple visualizations of trends over time
+\end{itemize}
+
 \section{Background}
@@
 Some terms we will be using for the rest of the paper include:
-message, word distribution, topic and trend. Messages are block of
-text written by developers, in this paper messages will be the CVS
-commit comments. Word distributions are the summary of a message by
-word count, effectively the word distribution will be over all
-messages looked at but since most words will not appear in a message,
-a word distribution is effectively a word count divided by the message
+message, word distribution, topic and trend. \emph{Messages} are
+blocks of text written by developers. In this paper messages will be
+the CVS commit comments made when the user commits changes to files in
+a repository. Word distributions are the summary of a message by word
+count. The word distribution will be over all messages
+looked at but since most words will not appear in a message, a word
+distribution is effectively a word count divided by the message
 size. A topic is a unique word distribution, but effectively a set of
@@
 
-Our blackbox view of LDA and LSI is that they extract topic clusters
-from the documents that are independant from one another. LDA doesn't
+Each of these, LDA and LSI, extract topic clusters
+from the documents that are independent from one another. LDA doesn't
 name these clusters but the clusters consist of words whose
@@
 
+Even with our liberal topic similarity metrics which produced both
+long and short trends, we showed that there are only a few trends in a
+repository which re-appear. Since so few trends re-appear and so many
+trends appear only once this suggests that global topic analysis might
+be ignoring locally unique topics.
+
+The utility of global topic analysis is questionable if the value of
+information decreases as it becomes older. Perhaps older trends will
+dominate the topic analysis. Windowed localized topic analysis show
+what are the unique topics.
+
+We presented multiple visualizations which focused on different
+aspects of the data: temporality of trends, trend size, and a
+condensed overall trend view. The condensed trend view shows much more
+information than the views that global analysis could show.
+
 \subsection{ Future Work}
@@
 
-\section{Appendix}
+% \section{Appendix}
 
-\begin{itemize}
-\item We are missing a comparison of our technique to the classical technique
-\end{itemize}
+% \begin{itemize}
+% \item We are missing a comparison of our technique to the classical technique
+% \end{itemize}
 

### 2009-04-03 lda-paper.tex
abstract edit

@@
 
-  At different times, developers focus on different topics and tasks
-  during development.  These different topics and tasks might
-  be obvious to developers familar with the project, but other
-  stakeholders, like managers, might be less aware of the topics of
-  development. Imagine the scenario where developers unexpectedly had to deal with
-  issues orthogonal to the features suggested, at the end of the
-  iteration a manager would wonder where all the effort went to.
-  Currently it is difficult for stakeholders to
-  audit what actually occurred and when certain features or topics
-  were focussed on. We propose a visualization of the
-  changing development topics over time. Our approach uses a topic
-  analysis tool like Latent Dirichlet Allocation (LDA) or Latent
-  Semantic Indexing (LSI). Instead of applying the tool to the
-  entire corpus we apply it to windows (periods) of the corpus. This allows us to
-  see the evolving stream of topics of development occuring
-  over time.
+%   At different times, developers focus on different topics and tasks
+%   during development.  These different topics and tasks might
+%   be obvious to developers familar with the project, but other
+%   stakeholders, like managers, might be less aware of the topics of
+%   development. Imagine the scenario where developers unexpectedly had to deal with
+%   issues orthogonal to the features suggested, at the end of the
+%   iteration a manager would wonder where all the effort went to.
+%   Currently it is difficult for stakeholders to
+%   audit what actually occurred and when certain features or topics
+%   were focussed on. We propose a visualization of the
+%   changing development topics over time. Our approach uses a topic
+%   analysis tool like Latent Dirichlet Allocation (LDA) or Latent
+%   Semantic Indexing (LSI). Instead of applying the tool to the
+%   entire corpus we apply it to windows (periods) of the corpus. This allows us to
+%   see the evolving stream of topics of development occuring
+%   over time.
+
+  During different periods, developers focus on different topics and
+  different tasks during development. Stake-holders, such as managers,
+  not directly involved with the development might have less awareness
+  of the project and the topics and tasks that the developers are
+  focussed on.  Imagine the scenario where a manager is trying to
+  discover where all the developers' effort went after features have
+  not been implemented because the developers unexpectedly had to deal
+  with issues orthogonal to these features.  We propose to help
+  stake-holder understand the topics of development by extracting
+  topics from the descriptions of changes that programmers provide
+  when they make changes.  We leverage topic analysis tools such as
+  Latent Dirichlet Allocation (LDA) or Latent Semantic Indexing (LSI)
+  to analyze periods, such as months, within a project's
+  lifetime. Previous work applied LDA and LSI to
+  to a project's entire lifetime.  These tools associate
+  descriptions with independent topics extracted from these
+  descriptions.  We propose a visualization of changing development
+  topics over time that allows us to see the evolving stream of topics
+  of development occurring over time. We also show that windowed topic
+  analysis offers benefits compare to entire lifetime topic analysis
+  because of the independence of topics between windows.
 
@@
 CVS, Subversion, Git etc., is potentially valuable as it has many uses
-for various stakeholders and uses.
+for various stake-holders and uses.
 Imagine a
@@
 fixes.  Using powerful topic analysis techniques like Latent Semantic
-Indexing (LSI) or Latent Dirichlet Allocation (LDA) one can seperate
-these mixed comments into topics. These topics describe seperate ideas
-that are interleaved throughtout the development. They indicate groups
+Indexing (LSI) or Latent Dirichlet Allocation (LDA) one can separate
+these mixed comments into topics. These topics describe separate ideas
+that are interleaved throughout the development. They indicate groups
 of comments which are related to each other by their language, often
@@
 size. A topic is a unique word distribution, but effectively a set of
-words unique to an independant behaviour in the text corpus, in this
+words unique to an independent behavior in the text corpus, in this
 paper we often summarize topics by their top 10 more frequent words.
@@
 
-Latent Dirichlet Alloation (LDA)~\cite{944937} is a competitor to
+Latent Dirichlet Allocation (LDA)~\cite{944937} is a competitor to
 Latent Semantic Indexing (LSI), probabilistic Latent Semantic Indexing
@@
 name these clusters but the clusters consist of words whose
-relationship is often obvious to someone familar with the
-corpus. Technically we could swap LDA for LSI. Our data is posed as
+relationship is often obvious to someone familiar with the
+corpus. We could swap LDA for LSI as produce similar results. Our data is posed as
 documents with word distributions (word counts per documents) and LDA
@@
 LSI is related to LDA and has been used to identify topics in software
-artifacts for forumal concept analysis and concept
+artifacts for formal concept analysis and concept
 location~\cite{1421013,1374321,10.1109/ICPC.2007.13,10.1109/ICPC.2006.17}.
@@
 Grant et al.~\cite{scottcordy} have used an alternative technique to
-LSI and LDA, they used Independant Component Analysis to seperate
+LSI and LDA, they used Independent Component Analysis to separate
 topic signals from software code.
@@
 stack the smaller trends on top. Although there is a chance that gaps
-will be left due to nonoptimal stacking in practice there are lot of
+will be left due to non-optimal stacking in practice there are lot of
 small trends (90\% of all trends contain 1 topic) which fill in these
@@
 The second visualization is the trend-time plot, it displays repeating
-trends more clearly by dedicating a horizonal line for trend segments
+trends more clearly by dedicating a horizontal line for trend segments
 dedicated to one trend. This means that if a trend contains
@@
 
-Once stop words were removed the some of hte purposes became more
-clear. The mega trend disappeared, the largest trend became bug fixes
+Once stop words were removed the some of the purposes became more
+clear. The mega- trend disappeared, the largest trend became bug fixes
 submitted by Dal Zotto which is a short trend.
@@
 The plots we produced of MaxDB 7.500 were interesting, there was a
-period where no real development occured and thus there were no topics or trends
+period where no real development occurred and thus there were no topics or trends
 whatsoever. We have two periods to evaluate for MaxDB 7.500, the first

### 2009-04-04 lda-paper.tex
abstract

@@
 
+%   During different periods, developers focus on different topics and
+%   different tasks during development. Stake-holders, such as managers,
+%   not directly involved with the development might have less awareness
+%   of the project and the topics and tasks that the developers are
+%   focussed on.  Imagine the scenario where a manager is trying to
+%   discover where all the developers' effort went after features have
+%   not been implemented because the developers unexpectedly had to deal
+%   with issues orthogonal to these features.  We propose to help
+%   stake-holder understand the topics of development by extracting
+%   topics from the descriptions of changes that programmers provide
+%   when they make changes.  We leverage topic analysis tools such as
+%   Latent Dirichlet Allocation (LDA) or Latent Semantic Indexing (LSI)
+%   to analyze periods, such as months, within a project's
+%   lifetime. Previous work applied LDA and LSI to
+%   to a project's entire lifetime.  These tools associate
+%   descriptions with independent topics extracted from these
+%   descriptions.  We propose a visualization of changing development
+%   topics over time that allows us to see the evolving stream of topics
+%   of development occurring over time. We also show that windowed topic
+%   analysis offers benefits compare to entire lifetime topic analysis
+%   because of the independence of topics between windows.
+
 During different periods, developers focus on different topics and
-  different tasks during development. Stake-holders, such as managers,
-  not directly involved with the development might have less awareness
-  of the project and the topics and tasks that the developers are
-  focussed on.  Imagine the scenario where a manager is trying to
-  discover where all the developers' effort went after features have
-  not been implemented because the developers unexpectedly had to deal
-  with issues orthogonal to these features.  We propose to help
-  stake-holder understand the topics of development by extracting
-  topics from the descriptions of changes that programmers provide
-  when they make changes.  We leverage topic analysis tools such as
-  Latent Dirichlet Allocation (LDA) or Latent Semantic Indexing (LSI)
-  to analyze periods, such as months, within a project's
-  lifetime. Previous work applied LDA and LSI to
-  to a project's entire lifetime.  These tools associate
-  descriptions with independent topics extracted from these
-  descriptions.  We propose a visualization of changing development
-  topics over time that allows us to see the evolving stream of topics
-  of development occurring over time. We also show that windowed topic
-  analysis offers benefits compare to entire lifetime topic analysis
-  because of the independence of topics between windows.
+different tasks during development. Stake-holders, such as managers, who are
+not directly involved with the project's development might have less awareness of the
+project and the topics and tasks that the developers have focussed on.
+Imagine the scenario where a manager is trying to discover where all the
+developers' effort went after features have not been implemented because the
+developers unexpectedly had to deal with issues orthogonal to these
+features.  We propose to help stakeholders understand the topics of
+development by extracting topics from the descriptions of changes that
+programmers provide when they make commits.  We leverage topic analysis
+tools such as Latent Dirichlet Allocation (LDA) or Latent Semantic Indexing
+(LSI) to analyze periods, such as months, within a project's lifetime,
+whereas previous work applied LDA and LSI to to a project's entire lifetime.  These
+tools associate descriptions with independent topics extracted from these
+descriptions.  
+We propose a visualization of changing development topics
+over time that allows us to see the evolving stream of topics of development
+occurring over time. We demonstrate that windowed topic analysis offers
+benefits over entire lifetime topic analysis because of the
+independence of topics between windows.
 

### 2009-04-05 lda-paper.tex
my latest changes to the paper
Some changes to the paper about figures and a change to the commit to topics figure (more labels)

@@
 
-We propose to extend the use of common topic analysis techniques by executing
-them over windows of documents over time. We hope by looking at topic
-clusters over time we can identify continuous effort on a select
-group of topics.
+We propose to extend the use of common topic analysis techniques by
+executing them over windows of documents over time. We hope by looking
+at topic clusters over time we can identify continuous effort on a
+select group of topics. Figure \ref{fig:lda} is an example of the kind
+of visualization we want to automatically generate. We want to see
+topics which are unique to a certain time period, displayed across the
+time axis. Those topics that reoccur, or occur over a larger period
+are plotted continuously. In our example we have summarized the topics
+with a single word.
 
@@
   \includegraphics[width=0.9\textwidth]{lda}
-  \caption{Example of topics extracted from MySQL 3.23. This is the kind of plot we eventually want to produce: named topics and topic trends}
+  \caption{Example of topics extracted from MySQL 3.23.
+This is the kind of plot we eventually want to produce: named topics and topic trends.
+The horizontal axis is time in months and the vertical axis used to stack topics that occur at the same time.
+Longer topics are topics which reoccur in adjacent window.}
   \label{fig:lda}
@@
   \includegraphics[width=0.9\textwidth]{commit-to-topics} 
-  \caption{How commits are analyzed and aggregated into Topics and Trends}
+  \caption{How commits are analyzed and aggregated into Topics and Trends.
+Commits are first extracted, then abstracted into word counts or word distributions.
+These word distributions are then given to topic clusterer like LDA which finds independant word distributions (topics) that these documents are related too and clusters or associates these documents with these topics.}
   \label{fig:commits}
@@
   \includegraphics[width=0.45\textwidth]{transitiveclosure}
-  \caption{Clusting topics by the transitive closure of topic similarity. Nodes are topics and arcs imply similarity}
+  \caption{Clusting topics by the transitive closure (connectedness or
+    a fill flood along arcs) of topic similarity. Nodes are topics and
+    arcs imply similarity. Similarity could be a measure such as
+    topics share 8 out of 10 top ten words.}
 \label{fig:closure}
@@
   \includegraphics[width=1.0\textwidth]{fixed-time-smear-plot-scaled}
-  \caption{Compact Trend Plot Topics per Month for MaxDB 7.500}
+  \caption{Compact Trend Plot Topics per Month for MaxDB 7.500. The
+    X-axis is time in months, the Y axis is used to stack topics
+    occurring at the same time. Trends which are continuous are
+    plotted as continuous blocks. The top 10 words in the topics are
+    joined and embedded in box representing the topics.}
   \label{fig:topicsmear}
@@
   \includegraphics[width=0.45\textwidth]{fixed-time-smear-plot-cropped}
-  \caption{Zoomed in Slice Compact Trend Plot of Topics per Month for MaxDB 7.500}
+  \caption{Zoomed in Slice Compact Trend Plot of Topics per Month for MaxDB 7.500. The topic text is visible in each topic box. Trends are plotted across time continuously.}
   \label{fig:zoomedsmear}
@@
   \includegraphics[width=0.45\textwidth]{class-smear-plot-crop-scaled}
-  \caption{Trend Time Line: Trends plotted over time of MaxDB 7.500}         
+  \caption{Trend Time Line: Trends plotted over time of MaxDB 7.500.
+Time in months are plotted along the X-Axis, each row on the Y-axis is associated with a trend ranked by size in descending order}
   \label{fig:trendtimeline}
@@
   \includegraphics[width=0.45\textwidth]{histogram-cropped-scaled}
-  \caption{Top part of trend histogram, ordered by topic occurance for MaxDB 7.500}       
+  \caption{Top part of trend histogram, ordered by topic occurance for MaxDB 7.500.
+X axis determines the number of continuous months of a trend.
+Trends are ranked  by the number of topics that a trend contains in descending order.}
   \label{fig:histogram}
@@
   \caption{MaxDB 7.500 Topics Analyzed with Static Topic Analysis, 20
-    topics over the entire development history of MaxDB 7.500. Each
-    box indicates the number of documents matching that topic in that
-    month.}
+    topics over the entire development history of MaxDB 7.500. The
+    color of the topic indicates the number of documents matching that
+    topic in that month relative to the number of documents (white is
+    most, black is least).}
   \label{fig:statictopics}
@@
 
-\subsection{Validation}
+%\subsection{Validation}
 

### 2009-04-06 lda-paper.tex
more lda paper

@@
 executing them over windows of documents over time. We hope by looking
-at topic clusters over time we can identify continuous effort on a
-select group of topics. Figure \ref{fig:lda} is an example of the kind
-of visualization we want to automatically generate. We want to see
-topics which are unique to a certain time period, displayed across the
-time axis. Those topics that reoccur, or occur over a larger period
-are plotted continuously. In our example we have summarized the topics
-with a single word.
+at topic clusters during periods or windows over time we can identify
+continuous effort on a select group of topics as well reflect topics
+that were local to that period. Figure \ref{fig:lda} is an example of
+the kind of visualization we want to automatically generate. We want
+to see topics which are unique to a certain time period, displayed
+across the time axis. Those topics that reoccur, or occur over a
+larger period are plotted continuously. In our example we have
+summarized the topics with a single word.
 
@@
 \end{figure*}
+
+Our methodology is to first extract the commits from repositories such
+as CVS or Bitkeeper. Then we extract the revisions and commits,
+grouping the revisions into commits if necessary. We extract the
+commit log message of each commit, count the occurence of each word in
+the message, remove stop words, and then produce word distributions
+for each message.
 

### 2009-04-07 lda-paper.tex
some minor intro edits

@@
 
-Topic analysis of revisions in SCSs is important as often revisions
-belonging to different topics are interleaved or even mixed inside of
-commits.  Sometimes a single revision to a file contains multiple bug
-fixes.  Using powerful topic analysis techniques like Latent Semantic
-Indexing (LSI) or Latent Dirichlet Allocation (LDA) one can separate
-these mixed comments into topics. These topics describe separate ideas
-that are interleaved throughout the development. They indicate groups
-of comments which are related to each other by their language, often
-the content of the comments are related.  In this case a topic is a
-set of tokens that are related to each other.
+Commits in SCSs are not necessarily consistent, they can have multiple
+purposes, thus we hope that multipurposes manifest themselves as
+topics interleaved within the commits.  Sometimes a single revision to
+a file contains multiple bug fixes.  Others have demonstrated that by
+using powerful topic analysis techniques like Latent Semantic Indexing
+(LSI) or Latent Dirichlet Allocation (LDA) one can separate topics
+from these mixed comments. These topics often describe separate ideas
+that are interleaved throughout the development. A topic represents a
+groups of commit comments which are related to each other by their
+content.  In this case a topic is a set of tokens that are related to
+each other.
 
@@
 time-line into periods of developer focus. By breaking apart an
-iteration we may be able to recognize the underlying software
-development process from these revision or at the very least what
-maintenance was occurring in the repository at a particular time.
+iteration into sets of topics and trends (topics which reoccur), we
+may be able to recognize the underlying software development process
+from these commits. Alternatively we can track down what maintenance was
+occurring in the repository at a particular time.
 

### 2009-04-07 lda-paper.tex
Latest changes

@@
 \begin{itemize}
-\item Demonstrating the value of windowed topic analysis
-\item Multiple visualizations of trends over time
+\item Demonstrating the value of windowed topic analysis using trends
+\item Multiple visualizations topics and trends over time
 \end{itemize}
@@
 Some terms we will be using for the rest of the paper include:
-message, word distribution, topic and trend. \emph{Messages} are
+message, word distribution, topics and trends. \emph{Messages} are
 blocks of text written by developers. In this paper messages will be
 the CVS commit comments made when the user commits changes to files in
-a repository. Word distributions are the summary of a message by word
-count. The word distribution will be over all messages
-looked at but since most words will not appear in a message, a word
+a repository. \emph{Word distributions} are the summary of a message
+by word count. The word distribution will be over all messages looked
+at but since most words will not appear in a message, a word
 distribution is effectively a word count divided by the message
-size. A topic is a unique word distribution, but effectively a set of
-words unique to an independent behavior in the text corpus, in this
-paper we often summarize topics by their top 10 more frequent words.
-A trend is a one or more similar topics which reoccur over
-time. Trends are important because they are sets of topics which
-reoccur over the development of a project.
+size. A \emph{topic} is essentially a set of words which form a word
+distribution that is unique and independant within the set of
+documents in our total corpus. One could think of a topic as the
+center distribution of a group of messages. In this paper we often
+summarize topics by their top 10 most frequent words in their word
+distribution.  A \emph{trend} is a one or more similar topics which
+reoccur over time. Trends are important because they are sets of
+topics which reoccur over the development of a project.
 
@@
 
-Latent Dirichlet Allocation (LDA)~\cite{944937} is a competitor to
-Latent Semantic Indexing (LSI), probabilistic Latent Semantic Indexing
-(pLSI) and semantic clustering. It is used for document modelling,
+In terms of clustering and finding topic distributions, Latent
+Dirichlet Allocation (LDA)~\cite{944937} competes with Latent Semantic
+Indexing (LSI), probabilistic Latent Semantic Indexing (pLSI) and
+semantic clustering. These tools is used for document modelling,
 document clustering and collaborative filtering. What LDA attempts to
-do is model a documents as mixture model of topics.  The common use of
-LDA in software engineering
+do is encode documents as mixture model of topics.  LDA is used in
+software engineering
 literature~\cite{lukins2008,10.1109/MSR.2007.20,NIPS2007637,1321709}
@@
 
-Each of these, LDA and LSI, extract topic clusters
-from the documents that are independent from one another. LDA doesn't
-name these clusters but the clusters consist of words whose
-relationship is often obvious to someone familiar with the
-corpus. We could swap LDA for LSI as produce similar results. Our data is posed as
-documents with word distributions (word counts per documents) and LDA
-extracts distinct topics (clusters of words) from the documents.
+Each of these, LDA, LSI and semantic clustering, extract topic
+clusters from the documents that are independent from one another. LDA
+doesn't name these clusters but the clusters consist of words whose
+relationship is often obvious to someone familiar with the corpus. We
+could swap LDA for LSI or semantic clustering and produce similar
+results. Our data is posed as documents with word distributions (word
+counts per documents) and LDA extracts distinct topics (clusters of
+words) from the documents.
 
@@
 
-Linstead et al has proposed an author source code model using
-LDA~\cite{10.1109/MSR.2007.20,NIPS2007637,1321709}. They used an
-author-topic (AT) model~\cite{1036902} to infer the association or
-expertise of a particular developer with a particular code topic
-extracted from the source control system.
+In order to infer or associate the expertise of an author with topics
+extracted from SCS, Linstead et al proposed an author-sourcecode model
+using LDA~\cite{10.1109/MSR.2007.20,NIPS2007637,1321709}. This model
+is essentially an extension of the author-topic (AT)
+model~\cite{1036902}, which associated authors and topics extracted
+with LDA.
 
-Lukins, Kraft and Etzkorn~\cite{lukins2008} use LDA to help query for
-finding related documents during bug localization tasks. They used LDA
-to build a topic model and then query against it using sample
-documents, the query would the proportion of topics that were
-associated with the query and thus the related documents.
+Lukins, Kraft and Etzkorn~\cite{lukins2008} use LDA to help bug
+localization by querying for related documents to a bug's topic. They
+used LDA to build a topic model and then queried against it using
+sample documents. These queries would indicate the proportion of
+topics that were associated with the query and thus possible the
+related documents.
 
@@
 location~\cite{1421013,1374321,10.1109/ICPC.2007.13,10.1109/ICPC.2006.17}.
-This is when concepts are automatically extracted from source code or
-documents are related to part of the source code.  Concept location is
-how high level concepts relate to low level entities like source code,
-for instance when fixing a bug how do the concepts in the bug report
-map back to the source code?  Semantic Clustering has also been used
-for similar reason~\cite{1698774,1566153}.
+This is where concepts are automatically extracted from source code or
+documents related to the source code.  Concept location is how high
+level concepts relate to low level entities like source code, for
+instance when fixing a bug how do the concepts in the bug report map
+back to the source code?  Semantic Clustering has also been used for
+similar reasons~\cite{1698774,1566153} as it is similar to LSI.
 

### 2009-04-08 lda-paper.tex
More lda paper changes from ric's suggestions

@@
 location~\cite{1421013,1374321,10.1109/ICPC.2007.13,10.1109/ICPC.2006.17}.
-This is where concepts are automatically extracted from source code or
-documents related to the source code.  Concept location is how high
-level concepts relate to low level entities like source code, for
-instance when fixing a bug how do the concepts in the bug report map
-back to the source code?  Semantic Clustering has also been used for
-similar reasons~\cite{1698774,1566153} as it is similar to LSI.
+Concept analysis aims to automatically extract concepts from source
+code or documents related to the source code.  Concept location is how
+high level concepts relate to low level entities like source code, for
+instance when fixing a bug how and where do the concepts in the bug
+report map back to the source code?  Semantic Clustering has also been
+used for similar reasons~\cite{1698774,1566153} as it is similar to
+LSI.
 
-Grant et al.~\cite{scottcordy} have used an alternative technique to
-LSI and LDA, they used Independent Component Analysis to separate
-topic signals from software code.
+Grant et al.~\cite{scottcordy} have used an alternative technique,
+rather than LSI and LDA, called Independent Component Analysis to
+separate topic signals from software code.
 
-Our technique differs in the sense we apply LDA temporally and we
-apply to change comments.
+Our technique differ from the previous techniques because we apply LDA
+locally to month-long windows of commit comments, where as other
+approaches apply LDA once to the entire project.
 
@@
 
-In our first exploratory pass we wanted to see
-
-Looking at 30 day non-overlapping windows of revisions for MySQL 3.23
-(20 topics, top 10 words per topic) we found there were common words
-across topic clusters such as diffs, annotate and history. There were
-notable transitional topics such as in the first window the word
-``bitkeeper'' appears (probably when they adopted Bitkeeper for their
-source control system) yet in the following windows there were no
-mentions of Bitkeeper. ``RENAME'' also only appeared once in the first
-window and never again.
+In our first exploratory pass we wanted to see if LDA could provide
+interesting topics extracted from a real system. We took the
+repository of MySQL 3.23, extracted the commits, grouped them by 30
+day non-overlapping windows then applied LDA to each of these
+windows. We asked LDA to make 20 topics per window and we would pay
+attention to the top 10 most frequent words in that topic.  We found
+there were common words across topic clusters, such as diffs, annotate
+and history, these words probably should be viewed as stop
+words. There were notable transitional topics such as in the first
+window the word ``bitkeeper'' appears (probably MySQL they adopted
+Bitkeeper for their source control system) yet in the following
+windows there were no mentions of Bitkeeper. ``RENAME'' also only
+appeared once in the first window and never again. Per each topic we
+tried to name the purpose of the topic, we found that with little
+familarity with the code base that naming the topic was not difficult.
 
 A subsampling of the notable topic words is displayed in Table
-\ref{tab:portability}. We tracked one topic cluster which seemed
-related to portability. We picked out terms which seem to summarize
-those topics.
-
-%Often words are shared across topics, see table \ref{tab:portability}
-%for tracking topics about portability.
-
-After viewing these topic is was obvious that we would have to combine
-try to track the evolution of topics over the windows or rank topics
-by similarity.
+\ref{tab:portability}.  To name the topics we picked out terms which
+seem to summarize those topics.  After viewing these topic is was
+obvious that we would have to combine try to track the evolution of
+topics over the windows or rank topics by similarity, thus if they
+were similar enough we joined them into a trend. Figure \ref{fig:lda}
+displays a sample plot of the extracted topics in table
+\ref{tab:portability}.
 
@@
 
+% \begin{table*}
+% \centering
+% \begin{tabular}{|ccc|l|}
+% \hline
+% 2000 &  Jul &  31 &    chmod \\
+% 2000 &  Sep &  29 &    fixes benchmark logging windows \\
+% 2000 &  Nov &  28 &    typo fix insert\_multi\_value \\
+% 2001 &  Jan &  27 &    fixes Innobase Cleanups auto-union \\
+% 2001 &  Mar &  28 &    2 topics bugfix, logging , TEMPORARY,  \\
+% \hline
+% 2001 &  Jul &  26 &    update Allow TABLES LOCK [a] \\
+
+% 2001 &  Aug &  25 &    tables row version [a] \\
+% \hline
+% 2001 &  Sep &  24 &    update checksum merge \\
+% 2001 &  Oct &  24 &    fixed fix \\
+% 2001 &  Dec &  23 &    HPUX SCO fix \\
+% \hline
+% 2002 &  Feb &  21 &    net buffer length  max\_allowed\_packet [b] \\
+% 2002 &  Mar &  23 &    small buf fix [b]  \\
+% \hline
+% 2002 &  May &  22 &    [popular] fix SCO OSF1 table\_name \\
+% 2002 &  Nov &  18 &    HPUX11 compiler HP \\
+% \hline
+% 2003 &  Feb &  16 &    Linux errno  [c] \\
+% 2003 &  Mar &  18 &    alarm bookmark bug [c] \\
+% \hline
+% 2003 &  Sep &  14 &    Auto logging merge windows distribution fix 64-bit 4.0 Cleanup \\
+% \hline
+% \end{tabular}
+% \caption{Tracking topics associated with the word portability, note some continuous blocks}
+% \label{tab:portability}
+% \end{table*}
+
 \begin{table*}
 \centering
-\begin{tabular}{|ccc|l|}
+\begin{tabular}{|cc|l|}
 \hline
-2000 &  Jul &  31 &    chmod \\
-2000 &  Sep &  29 &    fixes benchmark logging windows \\
-2000 &  Nov &  28 &    typo fix insert\_multi\_value \\
-2001 &  Jan &  27 &    fixes Innobase Cleanups auto-union \\
-2001 &  Mar &  28 &    2 topics bugfix, logging , TEMPORARY,  \\
+2000 &  Jul &      chmod \\
+2000 &  Sep &      fixes benchmark logging Win32 \\
+2000 &  Nov &      fixes insert\_multi\_value \\
+2001 &  Jan &      fixes Innobase Cleanups auto-union \\
+2001 &  Mar &      bugfix logging  TEMPORARY  \\
 \hline         
-2001 &  Jul &  26 &    update Allow TABLES LOCK [a] \\
+2001 &  Jul &      TABLES update Allow LOCK \\
 
-2001 &  Aug &  25 &    tables row version [a] \\
+2001 &  Aug &      TABLES row version \\
 \hline         
-2001 &  Sep &  24 &    update checksum merge \\
-2001 &  Oct &  24 &    fixed fix \\
-2001 &  Dec &  23 &    HPUX SCO fix \\
+2001 &  Sep &      update checksum merge \\
+2001 &  Oct &      fixed fix \\
+2001 &  Dec &      HPUX SCO fix \\
 \hline         
-2002 &  Feb &  21 &    net buffer length  max\_allowed\_packet [b] \\
-2002 &  Mar &  23 &    small buf fix [b]  \\
+2002 &  Feb &      net buffer length  max\_allowed\_packet \\
+2002 &  Mar &      small buf fix  \\
 \hline         
-2002 &  May &  22 &    [popular] fix SCO OSF1 table\_name \\
-2002 &  Nov &  18 &    HPUX11 compiler HP \\
+2002 &  May &      [popular] fix SCO OSF1 table\_name \\
+2002 &  Nov &      HPUX11 compiler HP \\
 \hline         
-2003 &  Feb &  16 &    Linux errno  [c] \\
-2003 &  Mar &  18 &    alarm bookmark bug [c] \\
+2003 &  Feb &      Linux errno   \\
+2003 &  Mar &      alarm bookmark bug \\
 \hline         
-2003 &  Sep &  14 &    Auto logging merge windows distribution fix 64-bit 4.0 Cleanup \\
+2003 &  Sep &      Auto logging merge windows distribution fix 64-bit 4.0 Cleanup \\
 \hline
 \end{tabular}
-\caption{Tracking topics associated with the word portability, note some continuous blocks}
+\caption{Sub-sampled topics from MySQL 3.23, some with continuous topics, these tokens were pulled from LDA topic. Each token is a summary of one LDA generated topic from MySQL 3.23 commit logs.}
 \label{tab:portability}
@@
 The horizontal axis is time in months and the vertical axis used to stack topics that occur at the same time.
-Longer topics are topics which reoccur in adjacent window.}
+Longer topics are topics which reoccur in adjacent window.
+Colors are arbitrary.}
   \label{fig:lda}

### 2009-04-08 lda-paper.tex
sent to mike and ric

@@
 %   during development.  These different topics and tasks might
-%   be obvious to developers familar with the project, but other
-%   stakeholders, like managers, might be less aware of the topics of
+%   be obvious to developers familiar with the project, but other
+%   stake-holders, like managers, might be less aware of the topics of
 %   development. Imagine the scenario where developers unexpectedly had to deal with
@@
 %   iteration a manager would wonder where all the effort went to.
-%   Currently it is difficult for stakeholders to
+%   Currently it is difficult for stake-holders to
 %   audit what actually occurred and when certain features or topics
@@
 %   entire corpus we apply it to windows (periods) of the corpus. This allows us to
-%   see the evolving stream of topics of development occuring
+%   see the evolving stream of topics of development occurring
 %   over time.
@@
 developers unexpectedly had to deal with issues orthogonal to these
-features.  We propose to help stakeholders understand the topics of
+features.  We propose to help stake-holders understand the topics of
 development by extracting topics from the descriptions of changes that
@@
 %   comments, and even source code per revision per each time window and
-%   extract the topics prevelant in that window. We expect that topic
+%   extract the topics prevalent in that window. We expect that topic
 %   sets will change over time and the change in topic sets indicate a
@@
 
-% In the presence of fine grained changes and detailed changelogs
+% In the presence of fine grained changes and detailed change-logs
 
@@
 
-% many changelogs how to sort by change type
+% many change-logs how to sort by change type
 
@@
 Commits in SCSs are not necessarily consistent, they can have multiple
-purposes, thus we hope that multipurposes manifest themselves as
+purposes, thus we hope that these multiple purposes manifest themselves as
 topics interleaved within the commits.  Sometimes a single revision to
@@
 size. A \emph{topic} is essentially a set of words which form a word
-distribution that is unique and independant within the set of
+distribution that is unique and independent within the set of
 documents in our total corpus. One could think of a topic as the
@@
 In order to infer or associate the expertise of an author with topics
-extracted from SCS, Linstead et al proposed an author-sourcecode model
+extracted from SCS, Linstead et al proposed an author-source-code model
 using LDA~\cite{10.1109/MSR.2007.20,NIPS2007637,1321709}. This model
@@
 tried to name the purpose of the topic, we found that with little
-familarity with the code base that naming the topic was not difficult.
+familiarity with the code base that naming the topic was not difficult.
 
-A subsampling of the notable topic words is displayed in Table
+A sub-sampling of the notable topic words is displayed in Table
 \ref{tab:portability}.  To name the topics we picked out terms which
@@
 %It is apparent that we have to compare these topics some how by
-%itemset or similarity.
+%item-set or similarity.
 
@@
 Commits are first extracted, then abstracted into word counts or word distributions.
-These word distributions are then given to topic clusterer like LDA which finds independant word distributions (topics) that these documents are related too and clusters or associates these documents with these topics.}
+These word distributions are then given to topic clusterer like LDA which finds independent word distributions (topics) that these documents are related too and clusters or associates these documents with these topics.}
   \label{fig:commits}
@@
 grouping the revisions into commits if necessary. We extract the
-commit log message of each commit, count the occurence of each word in
+commit log message of each commit, count the occurrence of each word in
 the message, remove stop words, and then produce word distributions
@@
 that is we fill flood along the similarity arcs until we have
-partitioned the topics into clusters of similar topics.  These
-clusters of topics are called trends. A trend is probably interesting
-and valuable if there is more than 1 topic in the trend.
+partitioned the topics into clusters of similar topics. Figure
+\ref{fig:closure} provides an example of clustering topics by the
+topic similarity. These clusters of topics are called trends. A trend
+is probably interesting and valuable if there is more than 1 topic in
+the trend.
 
@@
   \includegraphics[width=0.45\textwidth]{transitiveclosure}
-  \caption{Clusting topics by the transitive closure (connectedness or
-    a fill flood along arcs) of topic similarity. Nodes are topics and
+  \caption{Topic similarity demonstrated by clustering topics by the transitive closure (connectedness)
+    of topic similarity. Nodes are topics and
     arcs imply similarity. Similarity could be a measure such as
@@
 segments we plot those segments as joined across time.  We have a
-compressed stacked trend view, which displays trends arbitrarily
+compressed stacked trend view, shown in Figure \ref{fig:topicsmear}
+and Figure \ref{fig:zoomedsmear}, which displays trends arbitrarily
 across the y axis but are placed where they occur on the time-based x
-axis. Another plot we have is the trend-time plot, a each trend gets
-it own distinct-y value , while the x-axis is time; these topics are
-then each plotted on their own line across time as they occur. The
-last plot we have is the histogram plot, where we plot each trend on
+axis. Another plot we have is the trend-time line plot, shown in
+Figure \ref{fig:timeline}, each trend gets it own distinct-y value ,
+while the x-axis is time; these topics are then each plotted on their
+own line across time as they occur. Our final plot is the histogram
+plot, shown in Figure \ref{fig:histogram} where we plot each trend on
 its own line but stack up the segments of the trend, much like a plot
-of counts or a histogram.
+of counts or a histogram .
 
-The first visualization is a compressed view, we try to sink the
+The first visualization is a compressed view (Figure \ref{fig:topicsmear}
+and Figure \ref{fig:zoomedsmear}), we try to sink the
 larger trends to the bottom, once the larger trends are stacked, we
@@
 
-The second visualization is the trend-time plot, it displays repeating
-trends more clearly by dedicating a horizontal line for trend segments
-dedicated to one trend. This means that if a trend contains
-discontinuous segments the segments appear on the same line. One down
-side to this view is the least common trends need to be pruned
-otherwise it'd take too many horizontal lines to plot them.
+The second visualization is the trend-time line plot (Figure
+\ref{fig:timeline}), it displays repeating trends more clearly by
+dedicating a horizontal line for trend segments dedicated to one
+trend. This means that if a trend contains discontinuous segments the
+segments appear on the same line. One down side to this view is the
+least common trends need to be pruned otherwise it'd take too many
+horizontal lines to plot them.
 
-The 3rd visualization is the histogram plot which is a lot like the
-time-trend plot except the trends are plotted together by stacking to
-their row so time information is lost and the trends are ordered by
-the number of topics in the trend. The histogram plot shows the count
-of instances of a trend and thus indicates which trends occur the
-most. Unfortunately due to the large number of topics given allotted
-space it is often best to crop off the trends with only one topic,
-otherwise the tail is long.
+The third visualization is the histogram plot (Figure
+\ref{fig:histogram}) which is a lot like the time-trend plot except
+the trends are plotted together by stacking to their row so time
+information is lost and the trends are ordered by the number of topics
+in the trend. The histogram plot shows the count of instances of a
+trend and thus indicates which trends occur the most. Unfortunately
+due to the large number of topics given allotted space it is often
+best to crop off the trends with only one topic, otherwise the tail is
+long.
 
@@
   \includegraphics[width=0.45\textwidth]{class-smear-plot-crop-scaled}
-  \caption{Trend Time Line: Trends plotted over time of MaxDB 7.500.
+  \caption{Trend-time Line: Trends plotted over time of MaxDB 7.500.
 Time in months are plotted along the X-Axis, each row on the Y-axis is associated with a trend ranked by size in descending order}
@@
 %   2.   postgresql
+
+%\subsection{MySQL}
+%XXX DO MYSQL?
+
 \subsection{Postgresql}
@@
 
+%XXX Redo?
+
 PostgreSQL did not have many trends with two or more topics, but it
-did have one large mega-trend which stretched across most of its
-existence. Much of the mega-trend consisted of stop words but it did
+did have one large trend which stretched across most of its
+existence. Much of the large trend consisted of stop words but it did
 include topics which had words such as patch, fix, update.
@@
 Once stop words were removed the some of the purposes became more
-clear. The mega- trend disappeared, the largest trend became bug fixes
+clear. The large trend disappeared, the largest trend became bug fixes
 submitted by Dal Zotto which is a short trend.
@@
 
-%XXX needs results
+%XXX REDO?
 
@@
 \subsection{MaxDB 7.500}
+
+%XXX REDO?
 

### 2009-04-08 lda-paper.tex
uhu

@@
 
-\title{Windowed Developer Topic Analysis}
+%\title{Windowed Developer Topic Analysis}
+\title{What's hot and what's not:\\Windowing developer topic analysis}
 
@@
 
-During different periods, developers focus on different topics and
-different tasks during development. Stake-holders, such as managers, who are
-not directly involved with the project's development might have less awareness of the
-project and the topics and tasks that the developers have focussed on.
-Imagine the scenario where a manager is trying to discover where all the
-developers' effort went after features have not been implemented because the
-developers unexpectedly had to deal with issues orthogonal to these
-features.  We propose to help stake-holders understand the topics of
-development by extracting topics from the descriptions of changes that
-programmers provide when they make commits.  We leverage topic analysis
-tools such as Latent Dirichlet Allocation (LDA) or Latent Semantic Indexing
-(LSI) to analyze periods, such as months, within a project's lifetime,
-whereas previous work applied LDA and LSI to to a project's entire lifetime.  These
-tools associate descriptions with independent topics extracted from these
-descriptions.  
-We propose a visualization of changing development topics
-over time that allows us to see the evolving stream of topics of development
-occurring over time. We demonstrate that windowed topic analysis offers
-benefits over entire lifetime topic analysis because of the
-independence of topics between windows.
+% During different periods, developers focus on different topics and
+% different tasks during development. Stake-holders, such as managers, who are
+% not directly involved with the project's development might have less awareness of the
+% project and the topics and tasks that the developers have focussed on.
+% Imagine the scenario where a manager is trying to discover where all the
+% developers' effort went after features have not been implemented because the
+% developers unexpectedly had to deal with issues orthogonal to these
+% features.  We propose to help stake-holders understand the topics of
+% development by extracting topics from the descriptions of changes that
+% programmers provide when they make commits.  We leverage topic analysis
+% tools such as Latent Dirichlet Allocation (LDA) or Latent Semantic Indexing
+% (LSI) to analyze periods, such as months, within a project's lifetime,
+% whereas previous work applied LDA and LSI to to a project's entire lifetime.  These
+% tools associate descriptions with independent topics extracted from these
+% descriptions.  
+% We propose a visualization of changing development topics
+% over time that allows us to see the evolving stream of topics of development
+% occurring over time. We demonstrate that windowed topic analysis offers
+% benefits over entire lifetime topic analysis because of the
+% independence of topics between windows.
+
+As development on a software project progresses, developers will naturally
+shift their focus between different topics and tasks many times.  Managers
+and newcomer developers often seek simple ways of understanding what tasks
+have recently been worked on and how much effort has gone into each; for
+example, a manager might wonder what unexpected tasks occupied their team's
+attention during a period when they were supposed to have been implementing
+a set of new features.  Tools such as Latent Dirichlet Allocation (LDA) and
+Latent Semantic Indexing (LSI) can be used to analyze check-in comments
+over the entire lifetime of a project, and previous work on developer topic
+analysis has leveraged these tools to associate commit log comments with
+independent topics extracted from these commit log comments.  In this work, we use
+LDA and LSI to analyze periods, such as months, within a project's lifetime
+to create a time-windowed model of changing development topics.  We propose
+a visualization of this model that allows us to explore the evolving stream
+of topics of development occurring over time, and we demonstrate that
+windowed topic analysis offers strong advantages over entire lifetime topic
+analysis because of the independence of topics between windows.
 
@@
 CVS, Subversion, Git etc., is potentially valuable as it has many uses
-for various stake-holders and uses.
+for various stakeholders and uses.
 Imagine a
@@
 
-\begin{table*}
+\begin{table}
 \centering
@@
 2001 &  Sep &      update checksum merge \\
-2001 &  Oct &      fixed fix \\
-2001 &  Dec &      HPUX SCO fix \\
-\hline         
-2002 &  Feb &      net buffer length  max\_allowed\_packet \\
-2002 &  Mar &      small buf fix  \\
-\hline         
-2002 &  May &      [popular] fix SCO OSF1 table\_name \\
-2002 &  Nov &      HPUX11 compiler HP \\
-\hline         
-2003 &  Feb &      Linux errno   \\
-2003 &  Mar &      alarm bookmark bug \\
-\hline         
-2003 &  Sep &      Auto logging merge windows distribution fix 64-bit 4.0 Cleanup \\
+% 2001 &  Oct &      fixed fix \\
+% 2001 &  Dec &      HPUX SCO fix \\
+% \hline         
+% 2002 &  Feb &      net buffer length  max\_allowed\_packet \\
+% 2002 &  Mar &      small buf fix  \\
+% \hline         
+% 2002 &  May &      [popular] fix SCO OSF1 table\_name \\
+% 2002 &  Nov &      HPUX11 compiler HP \\
+% \hline         
+% 2003 &  Feb &      Linux errno   \\
+% 2003 &  Mar &      alarm bookmark bug \\
+% \hline         
+% 2003 &  Sep &      Auto logging merge windows distribution fix 64-bit 4.0 Cleanup \\
 \hline
@@
 \label{tab:portability}
-\end{table*}
+\end{table}
 

### 2009-04-09 lda-paper.tex
big migod edits

--- 2009-04-08 lda-paper.tex
@@
 working on? What was the developer in the next cubicle working on?
-What were the common topics and issues dealt with in the current
+Which common topics and issues were dealt with in the current
 release of the software? What were the topics of the previous release?
 Given the requirements agreed to at the start of the iteration did our
-developers work on them? What else did they work on.
+developers work on them? What else did they work on?
 
-Topic analysis, with respect to Software Control Systems (SCSs) like
-CVS, Subversion, Git etc., is potentially valuable as it has many uses
-for various stake-holders and uses.
-Imagine a
-development team has agreed to implement some features at the start of
-an iteration. By the end of the iteration only some of the agreed upon
-features are implemented. The developers swear they were focusing
-solely on the features agreed upon. Topic analysis could help by
-teasing out the distinct topics or focusses of the developers. Perhaps
-the developers had to deal with issues.
-Developers might have grouped those orthogonal changes into
-one of the features they were working on, thus when asked about how
-their time was spent the orthogonal changes would not be covered. This
-is what we hope that windowed topic analysis can uncover. If windowed
-topic analysis can discover important but overlooked topics, it can
-help both developers and managers. Developer can see a
-summary of what worked on, and managers can development.
+Topic analysis, with respect to Software Control Systems (SCSs) such
+as CVS, Subversion, Git etc., is potentially valuable as it has many
+uses for various stake-holders and uses.  Imagine a development team
+has agreed to implement some features at the start of an iteration. By
+the end of the iteration only some of the agreed upon features are
+implemented. The developers insist they were focusing solely on the
+features agreed upon, but what actually happened? Topic analysis could
+help by teasing out the distinct topics or focusses of the
+developers. Perhaps the developers had to deal with issues.
+Developers might have grouped those orthogonal changes into one of the
+features they were working on, thus when asked about how their time
+was spent the orthogonal changes would not be covered. This is the kind of underlying story that we
+hope that windowed topic analysis can uncover. If windowed topic
+analysis can discover important but overlooked topics, it can help
+both developers and managers. Developer can see a summary of what
+worked on, and managers can development.
 
@@
 purposes, thus we hope that these multiple purposes manifest themselves as
-topics interleaved within the commits.  Sometimes a single revision to
-a file contains multiple bug fixes.  Others have demonstrated that by
-using powerful topic analysis techniques like Latent Semantic Indexing
-(LSI) or Latent Dirichlet Allocation (LDA) one can separate topics
-from these mixed comments. These topics often describe separate ideas
+topics interleaved within the commits.   Others have demonstrated that by
+using powerful topic analysis techniques, such as Latent Semantic Indexing
+(LSI) and Latent Dirichlet Allocation (LDA), can aid in separating topics
+from these mixed comments. These topics often describe distinct ideas
 that are interleaved throughout the development. A topic represents a
-groups of commit comments which are related to each other by their
-content.  In this case a topic is a set of tokens that are related to
-each other.
+groups of commit comments that are related to each other by their
+content.  In this case a topic is a set of tokens.
 
@@
 time-line into periods of developer focus. By breaking apart an
-iteration into sets of topics and trends (topics which reoccur), we
-may be able to recognize the underlying software development process
-from these commits. Alternatively we can track down what maintenance was
-occurring in the repository at a particular time.
+iteration into sets of topics and trends (topics which recur), we may
+be able to recognize the underlying software development process from
+these commits. Alternatively, we can determine what particular
+maintenance task was being performed are a given time.
 
@@
 the kind of visualization we want to automatically generate. We want
-to see topics which are unique to a certain time period, displayed
-across the time axis. Those topics that reoccur, or occur over a
+to see topics that are unique to a certain time period, displayed
+across the time axis. Those topics that recur, or occur over a
 larger period are plotted continuously. In our example we have
-summarized the topics with a single word.
+summarized each topic with a single word.
 
-In this paper we explore how topics shift over time in the
-source control repository of a project. We analyze commit comments per
-revision in each time window and we extract the topics in that time
-window. We expect that topics in each window will change over time and
-the similarities and differences in these sets of topics will indicate
-both developer focus and changes in developer focus.
+In this paper we explore how topics shift over time in the source
+control repository of a project, using several open source database
+systems as examples. We analyze commit comments per revision in each
+time window and we extract the topics in that time window. We expect
+that topics in each window will change over time and the similarities
+and differences in these sets of topics will indicate both developer
+focus and changes in developer focus.
 
@@
 \item Multiple visualizations topics and trends over time
+\item An exploratory case study using these techniques on several open source database systems.
 \end{itemize}
@@
 
-Some terms we will be using for the rest of the paper include:
-message, word distribution, topics and trends. \emph{Messages} are
-blocks of text written by developers. In this paper messages will be
-the CVS commit comments made when the user commits changes to files in
-a repository. \emph{Word distributions} are the summary of a message
-by word count. The word distribution will be over all messages looked
-at but since most words will not appear in a message, a word
-distribution is effectively a word count divided by the message
-size. A \emph{topic} is essentially a set of words which form a word
-distribution that is unique and independent within the set of
-documents in our total corpus. One could think of a topic as the
+We now define some terms that we will use in the rest of the paper: A
+\emph{message} is a block of text written by developers. In this
+paper, messages will be the CVS commit comments made when the user
+commits changes to files in a repository. A \emph{word distribution}
+is the summary of a message by word count. The word distribution will
+be over all messages looked at but since most words will not appear in
+a message, a word distribution is effectively a word count divided by
+the message size. A \emph{topic} is essentially a set of words that
+form a word distribution that is unique and independent within the set
+of documents in our total corpus. One could think of a topic as the
 center distribution of a group of messages. In this paper we often
@@
 distribution.  A \emph{trend} is a one or more similar topics which
-reoccur over time. Trends are important because they are sets of
-topics which reoccur over the development of a project.
+recur over time.  Trends are particularly important, as they indicate
+long-lived and recurring topics that may provide key insights into the
+development of the project.
 
@@
 %XXX explain LDA \cite{944937}
-
+%XXX What has pLSI in it?
 In terms of clustering and finding topic distributions, Latent
 Dirichlet Allocation (LDA)~\cite{944937} competes with Latent Semantic
-Indexing (LSI), probabilistic Latent Semantic Indexing (pLSI) and
-semantic clustering. These tools is used for document modelling,
+Indexing (LSI)~\cite{1421013,1374321,10.1109/ICPC.2007.13,10.1109/ICPC.2006.17}, probabilistic Latent Semantic Indexing (pLSI) and
+semantic clustering~\cite{1698774,1566153}. These tools are used for document modelling,
 document clustering and collaborative filtering. What LDA attempts to
@@
 
-Each of these, LDA, LSI and semantic clustering, extract topic
+LDA, LSI and semantic clustering extract topic
 clusters from the documents that are independent from one another. LDA
-doesn't name these clusters but the clusters consist of words whose
+does not name these clusters but the clusters consist of words whose
 relationship is often obvious to someone familiar with the corpus. We
@@
 In order to infer or associate the expertise of an author with topics
-extracted from SCS, Linstead et al proposed an author-source-code model
+extracted from SCS, Linstead et al. proposed an author-source-code model
 using LDA~\cite{10.1109/MSR.2007.20,NIPS2007637,1321709}. This model
@@
 Lukins, Kraft and Etzkorn~\cite{lukins2008} use LDA to help bug
-localization by querying for related documents to a bug's topic. They
+localization by querying for documents that are related to a bug's topic. They
 used LDA to build a topic model and then queried against it using
@@
 Concept analysis aims to automatically extract concepts from source
-code or documents related to the source code.  Concept location is how
-high level concepts relate to low level entities like source code, for
-instance when fixing a bug how and where do the concepts in the bug
+code or documents that are related to the source code.  Concept location concerns how
+high level concepts relate to low level entities such as source code, for
+example, when fixing a bug how and where do the concepts in the bug
 report map back to the source code?  Semantic Clustering has also been
-used for similar reasons~\cite{1698774,1566153} as it is similar to
+used for similar purposes~\cite{1698774,1566153} as it is similar to
 LSI.
@@
 Grant et al.~\cite{scottcordy} have used an alternative technique,
-rather than LSI and LDA, called Independent Component Analysis to
+ called Independent Component Analysis to
 separate topic signals from software code.
 
-Our technique differ from the previous techniques because we apply LDA
-locally to month-long windows of commit comments, where as other
-approaches apply LDA once to the entire project.
+Our technique differs from the previous techniques because we apply
+LDA locally to month-long windows of commit comments, where as other
+approaches apply LDA once to the entire project. This allows us to
+examine the time-based windowing of topics over its development
+history.
 
@@
 
-\section{First Impressions}
+\section{Preliminary Case Study}
 
@@
 repository of MySQL 3.23, extracted the commits, grouped them by 30
-day non-overlapping windows then applied LDA to each of these
-windows. We asked LDA to make 20 topics per window and we would pay
-attention to the top 10 most frequent words in that topic.  We found
-there were common words across topic clusters, such as diffs, annotate
-and history, these words probably should be viewed as stop
-words. There were notable transitional topics such as in the first
-window the word ``bitkeeper'' appears (probably MySQL they adopted
-Bitkeeper for their source control system) yet in the following
+day non-overlapping windows, and then applied LDA to each of these
+windows. We asked LDA to make 20 topics per window and we then
+carefully examined the top 10 most frequent words in that topic.  We
+found there were common words across topic clusters, such as
+``diffs'', ``annotate'' and ``history'', these words probably should
+be viewed as stop words. There were notable transitional topics such
+as in the first window the word ``bitkeeper'' appears because MySQL
+adopted Bitkeeper for their source control system yet in the following
 windows there were no mentions of Bitkeeper. ``RENAME'' also only
 appeared once in the first window and never again. Per each topic we
-tried to name the purpose of the topic, we found that with little
-familiarity with the code base that naming the topic was not difficult.
+tried to identify the purpose of the topic; to our surprise we found
+that even with onlyy a little familiarity with the code base that
+naming the topic was straightforward.
 
 A sub-sampling of the notable topic words is displayed in Table
-\ref{tab:portability}.  To name the topics we picked out terms which
-seem to summarize those topics.  After viewing these topic is was
+\ref{tab:portability}.  To name the topics we selected terms that
+seemed to best summarize those topics.  After viewing these topics, it was
 obvious that we would have to combine try to track the evolution of
@@
 The horizontal axis is time in months and the vertical axis used to stack topics that occur at the same time.
-Longer topics are topics which reoccur in adjacent window.
+Longer topics are topics which recur in adjacent window.
 Colors are arbitrary.}
@@
 
-Our methodology is to first extract the commits from repositories such
+Our methodology is to first extract the commits from SCS repositories such
 as CVS or Bitkeeper. Then we extract the revisions and commits,
@@
 
-\subsection{Extract Repos}
+\subsection{Extract Repositories}
 
-We extracted the repositories with software such as rsync, CVSSuck,
-\texttt{softChange} and bt2csv.  \texttt{softChange} provided a
-general schema for storing revisions. CVSSuck and rsync mirrored CVS
-repositories while bt2csv mirrored web accessible bitkeeper
-repositories.
+We mirrored the repositories and their revisions using software such
+as rsync, CVSSuck, \texttt{softChange} and bt2csv.
+\texttt{softChange} provided a general schema for storing
+revisions. CVSSuck and rsync mirrored CVS repositories while bt2csv
+mirrored web accessible bitkeeper repositories.
 
@@
 
-The document we are analyzing are the changelog comments, these are
-the comments added when a revision to the source code is committed.
+The document we are analyzing are the changelog comments i.e., the comments that are added  when a revision to the source code is committed.
 Documents are converted into word count distributions. These
-distributions are can be normalized by by the size of the
-message. After all messages are processes the distributions are
+distributions, which are then normalized by by the size of the
+message. After all messages are processed the distributions are
 extended to include all words in all distributions (each distribution
@@
 overlapping windows, but in this paper we use non-overlapping windows
-of a set time unit.  Windowing by time allows for many levels of
-granularity. We usually used one month as the length of our time
-windows, while we could use different granularities for the study we
-think that a month is a sizable unit of development (4 weeks of
-development).
+of a set time unit because it simplifies analysis.  Windowing by time
+allows for many levels of granularity. We usually used one month as
+the length of our time windows, while we could use different
+granularities for the study we think that a month is a sizable unit of
+development, 4 weeks of development, which we consider to be fine
+enough to show meaningful shifts but coarse enough not to show too
+much detail.
 
@@
 topics in our case study. Our Topic Analysis tool is an implementation
-of LDA, but we could've used other similar tools like LSI.
+of LDA, but we could have used other similar tools like LSI, but
+previous showed more promising topic analysis with LDA.
 
 This is a very slow step as LDA takes a long time to run, and takes
-even longer since we run it once per window.
+even longer since we run it once per window. We note that this step is
+a slow one, as executing even one instance of LDA involves significant
+computation, and we perform LDA once per window.
 
@@
 
-Once we have our topics extracted for each window , we analyze them
-and look for repeating topics. We do this by taking the top 10 most
-common words and then comparing them.
+Once we have our topics extracted for each window, we analyze them and
+look for repeating topics. We find repeating topics by comparing the
+10 most common words of each topic.
 
@@
 topic is compared to each other topic in the system, given a certain
-threshold of top 10 similarity, like 8/10 words. Then given this topic
-similarity matrix, we find the transitive closure of similar topics,
-that is we fill flood along the similarity arcs until we have
+threshold of top 10 similarity, such as 8/10 words. Then given this topic
+similarity matrix, we find the transitive closure of similar topics;
+that is, we fill flood along the similarity arcs until we have
 partitioned the topics into clusters of similar topics. Figure
@@
 topic similarity. These clusters of topics are called trends. A trend
-is probably interesting and valuable if there is more than 1 topic in
+is probably interesting and valuable if there is more than one topic in
 the trend.
 
-This technique has weaknesses in that nodes a few neighbors away in
+This technique has weaknesses in that nodes that are  a few neighbors away in
 similarity might not share any similar words.  We do this because we
-want to be able to color or highlight topics which are related to each
+want to be able to color or highlight topics that are related to each
 other and track them throughout time.
@@
 
-We have come up with multiple ways to visualize these topics and
-trends.  For all of these plots if we trends which have continuous
-segments we plot those segments as joined across time.  We have a
+We have devised several ways to visualize these topics and
+trends.  For all of these plots if we find trends that have continuous
+segments, then we plot those segments as joined across time.  We have a
 compressed stacked trend view, shown in Figure \ref{fig:topicsmear}
-and Figure \ref{fig:zoomedsmear}, which displays trends arbitrarily
-across the y axis but are placed where they occur on the time-based x
-axis. Another plot we have is the trend-time line plot, shown in
-Figure \ref{fig:timeline}, each trend gets it own distinct-y value ,
+and Figure \ref{fig:zoomedsmear}, 
+which displays trends as they occur
+on the time-based x-axis (placement along the y-axis is arbitrary).
+ Another plot we have is the trend-time line plot, shown in
+Figure \ref{fig:trendtimeline}, each trend gets it own distinct y-value ,
 while the x-axis is time; these topics are then each plotted on their
@@
 its own line but stack up the segments of the trend, much like a plot
-of counts or a histogram .
+of counts or a histogram.
 
-The first visualization is a compressed view (Figure \ref{fig:topicsmear}
-and Figure \ref{fig:zoomedsmear}), we try to sink the
-larger trends to the bottom, once the larger trends are stacked, we
-fill in the gaps with the smaller trends in the same window and then
-stack the smaller trends on top. Although there is a chance that gaps
-will be left due to non-optimal stacking in practice there are lot of
-small trends (90\% of all trends contain 1 topic) which fill in these
-gaps quickly. The trend segments that share the same color are the
-same trend, colors are randomly chosen for each trend, so color
-similarity is meaningless, only exact color matching matters.
-The compressed view is nice because it shows all of the
+The first visualization is a compressed view
+(Figure\ref{fig:topicsmear} and Figure \ref{fig:zoomedsmear}) where we
+try to sink the larger trends to the bottom.  Once the larger trends
+are stacked we fill in the gaps with the smaller trends in the same
+window, and then stack the smaller trends on top.  Although there is a
+chance that gaps will be left due to non-optimal stacking in practice
+there are lot of small trends (90\% of all trends contain 1 topic)
+which fill in these gaps quickly.  Different instances of the same
+trend share the same colour; apart from that, the color of trends is
+randomly chosen and color similarity is not meaningful.
+The compressed view is convienant because it shows all of the
 information and the repeating continuous trends are easy to pick out,
-unfortunately repeating discontinuous trends are harder to spot.
+although repeating discontinuous trends are harder to spot.
 
 The second visualization is the trend-time line plot (Figure
-\ref{fig:timeline}), it displays repeating trends more clearly by
+\ref{fig:trendtimeline}), it displays repeating trends more clearly by
 dedicating a horizontal line for trend segments dedicated to one
-trend. This means that if a trend contains discontinuous segments the
-segments appear on the same line. One down side to this view is the
-least common trends need to be pruned otherwise it'd take too many
-horizontal lines to plot them.
+trend. This means that if a trend contains discontinuous segments then the
+segments appear on the same line. 
+However, the least common trends need to be pruned away or the view will be cluttered with a lot of noise.
 
 The third visualization is the histogram plot (Figure
-\ref{fig:histogram}) which is a lot like the time-trend plot except
-the trends are plotted together by stacking to their row so time
-information is lost and the trends are ordered by the number of topics
-in the trend. The histogram plot shows the count of instances of a
-trend and thus indicates which trends occur the most. Unfortunately
-due to the large number of topics given allotted space it is often
-best to crop off the trends with only one topic, otherwise the tail is
-long.
+\ref{fig:histogram}) which superficially resembles the time-trend
+plot.  However, in this view the trends are plotted together by
+stacking to their row so time information is lost, and the trends are
+ordered by the number of topics in the trend.  The histogram plot
+shows the count of instances of a trend and thus indicates which
+trends occur the most. Unfortunately due to the large number of topics
+given allotted space it is often best to crop off the trends with only
+one topic, otherwise the tail is long.
 
+%XXX Describe something useful like a trend
 \begin{figure*}
@@
     X-axis is time in months, the Y axis is used to stack topics
-    occurring at the same time. Trends which are continuous are
+    occurring at the same time. Trends that are continuous are
     plotted as continuous blocks. The top 10 words in the topics are
@@
 
-Our analysis tools consisted of: Hiraldo-Grok, an OCaml based spin off
-of Grok used for answering queries; Gnuplot, a graph plotting package;
-lda-c, a LDA package implemented by Biel; and our topic plotter,
-implemented in Haskell.
+Our analysis tools consisted of: Hiraldo-Grok, an OCaml-based variant
+of the Grok query language; Gnuplot, a graph plotting package; lda-c,
+a LDA package implemented by Biel; and our topic plotter, implemented
+in Haskell.
 
@@
 
-We applied our tools and visualizations to the repositories of
-multiple database systems.
+We applied our tools and visualizations to the repositories of several
+open source database systems: PostgreSQL, MaxDB and Firebird.
 
@@
 
-PostgreSQL did not have many trends with two or more topics, but it
-did have one large trend which stretched across most of its
-existence. Much of the large trend consisted of stop words but it did
-include topics which had words such as patch, fix, update.
+When we examined PostgreSQL, we did not find many trends with two or more topics; however, we did find one large trend which stretched across most of its existence.
+ Much of the large trend consisted of stop words but it did
+include topics which had non-stop words such as \emph{patch}, \emph{fix}, \emph{update}.
 
-The second largest topic was interesting because it dealt changes from
-1998, from developers using Central Europe Time (CEST), the two
-notable tokens were 1998 and CEST. These changes seem to be patches
-submitted by one particular developer who used that timezone.
+The second largest topic was interesting because it concerned changes from
+1998, from developers using Central Europe Time (CEST), where the two
+notable tokens were 1998 and CEST. These changes appear to be patches
+submitted by one particular developer who lived in that timezone.
 
-The third largest trend seemed to be about copying and moving files
-between branches as it referenced the old branch. This is a side
-effect of the programmers manually recording such a merge across
-branches because CVS cannot.
+The third largest trend we found appeared to concerned with copying
+and moving files between branches as it referenced the old branch.
+branches.  We inferred this trend based on the fact that the old
+version identifier was referenced, and we know this this is common
+practice when performing a branch merge, since CVS cannot record such
+information directly.
 
@@
 clear. The large trend disappeared, the largest trend became bug fixes
-submitted by Dal Zotto which is a short trend.
+submitted by a developer named Dal Zotto which is a short trend.
 
+%XXX REDO POSTEGRESQL?
 % Large Trends is not the point!!!
@@
 
-But large trends are not the point, by investigating windowed trends
-we want more local information about trends being worked on.
+%But large trends are not the point, by investigating windowed trends
+%we want more local information about trends being worked on.
 
@@
 Our first attempt with Firebird resulted in a set of trends that were
-dominated by one large trend containing tokens which were mostly stop
+dominated by one large trend containing tokens that were mostly stop
 words such as ``the to a for of is it'' and references to one author
@@
 
-The second largest trend related to branching and possibly merging, it
+The second largest trend related to branching and possibly merging; it
 contained language such as ``was added on branch'' and contained
@@
 
-The third largest trend contained reference to the changelog, the
+The third largest trend contained references to the changelog, the
 build system, architecture and platform support. It specifically
@@
 
-Topics that were not trends that were interesting were external bug
-fixes submitted to the project, the developers would write thank yous
-in their changelog comments: ``thanks Bill Lam''. Other interesting
-topics include tokens and topics such as: ``compiler workarounds'', ``nightly updates'', packets and MSVC++,
+Topics that were not trends but appeared to be  interesting were mostly external bug
+fixes submitted to the project.
+ In these cases, the developers would express thanks in their changelog comments, such as ``Thanks, Bill Lam''.  Other easily discernible topics included 
+ tokens and topics such as: ``compiler workarounds'', ``nightly updates'', packets and MSVC++,
 
@@
 
-The plots we produced of MaxDB 7.500 were interesting, there was a
+%XXX Migod todo
+% I think that you should enlarge this discussion of MaxDB (at the
+% expense of the other systems) and make careful references to how the
+% diagrams give you extra info about the evolution of the system (in
+% terms of topics).
+
+% We are trying to argue that (a) the approach is more informative that
+% the naive static approach and (b) the diagrams are useful.
+
+The plots we produced of MaxDB 7.500 were 
+unlike those of the other systems, as 
+ there was a
 period where no real development occurred and thus there were no topics or trends
@@
 v1.09, the prefix SUT stands for Storable Unit Type. Sutcheck would
-also automate perforce check-ins.
+also automate check-ins using a Perforce SCS tool.
 
 The second largest common trend seems to be a side effect of an
-automated checkin which is annotated as ``implicit checkin''. These
-were checkins that were produced when importing changes from an
+automated check-in which is annotated as ``implicit check-in''. These
+were check-ins that were produced when importing changes from an
 external Perforce repository.
@@
 
-The previous work which leverage LSI or LDA usually extracted a
-certain number of topics and tracked them over time. This meant a
-static set of topics was tracked across the whole development history
-of a project. Thus we really had 20 trends.
+%XXX Migod notes This section is perhaps the most important of the
+% whole paper.  You need expand it.  Maybe you should move the first
+% two paras of the conclusions in here too.
+
+% The thing that is most disappointing is that the argument is entire qualitative.  Can you give examples to make this sound less hypothetical?
+
+% Can you add numbers in here to make the argument stronger?
+
+Previous work on topic analysis that employed LSI and LDA typically
+extracted a certain number of topics and tracked them over time. This
+means that a single, static set of topics was tracked across the whole
+development history of a project.
 
 We attempted the same kind of analysis, as shown in
-Figure~\ref{fig:statictopics} we extracted 20 topics and then plotted
-the number of messages per month which related to that topic. What we
-found was that often one topic would dominate the results, while
-messages related to other topics didn't appear that often.
-
-This technique seems useful if the 20 extracted topics are relevant
-and useful, that said one weakness of this approach is that documents
-for a topic might only occur in one time window, yet we're tracking
-messages that could be related to it over the entire lifetime of the
-project.
+Figure~\ref{fig:statictopics}. We extracted 20 topics and then plotted
+the number of messages per month that were related to that topic. We
+found was that one topic would often dominate the results, while
+messages related to other topics did not appear very often.
+% This technique seems useful if the 20 extracted topics are relevant
+% and useful, that said one weakness of this approach is that documents
+% for a topic might only occur in one time window, yet we are tracking
+% messages that could be related to it over the entire lifetime of the
+% project.
+This approach seems reasonable if most of the extracted topics are of
+broad interest during most of the development process.  However, it
+may be that some topics are of strong interest but only briefly; in
+such a case, a window-based model gives a much stronger indication of
+the fleeting importance of such topics, and can help to put such a
+topic into its proper context.
 
@@
 their changes with relevant information. We rely on the descriptions
-they provide. If the language of checkin comments was automated we
+they provide. If the language of check-in comments was automated we
 would only be analyzing that.
@@
 remove them, but perhaps dropping some of these stop words would aid
-the clarity of such topic clusters, alternatively it might cause our
-topic similarity plots to be fundamentally different. Perhaps our
-methods of suggesting that topics are similar is not appropriate.
+the clarity of such topic cluster; alternatively it might cause our
+topic similarity plots to be fundamentally different.
 
-The number of commits per month is inconsistent some months have a lot
-of changes where as some months have almost none.
-
+The number of
+commits per month is inconsistent as some months have a lot of changes
+while others months have almost none.
 %   1.   validation
@@
 
-Even with our liberal topic similarity metrics which produced both
+%XXX migod comments:
+
+% The first two paras here are good stuff, but I think it belongs in
+% section 5 at the end, as a summary of the utility of the approach.
+% But can you add more meat to it in terms of numbers?
+
+% In the conclusions you can simply restate what we did and what we
+% learned without doing the arguing here.
+
+% The third para is a good one for the conclusions, tho there need to
+% be more detail.
+
+Even with our liberal topic similarity metrics that produced both
 long and short trends, we showed that there are only a few trends in a
-repository which re-appear. Since so few trends re-appear and so many
+repository which recur. Since so few trends recur and so many
 trends appear only once this suggests that global topic analysis might
@@
 
-We presented multiple visualizations which focused on different
+We presented multiple visualizations that focused on different
 aspects of the data: temporality of trends, trend size, and a
@@
 cluster labelling. Given a word distribution we should be able to
-figure out a word or term which best describes that distribution
-automatically.
+automatically select a word or term that best describes that distribution.
 Another avenue we wish to investigate the is value of applying
-static community generated software taxonomies to the naming or
+static community-generated software taxonomies to the naming or
 grouping of clusters

### 2009-04-09 lda-paper.tex
Some latest edits

@@
 
-The third most common trend, seen on Figure \ref{fig:trend}, seemed to
+The third most common trend, seen on Figure \ref{fig:trendtimeline}, seemed to
 include tokens related to operating system support, such as Linux and

### 2009-04-09 lda-paper.tex
pLSI

@@
 as CVS, Subversion, Git etc., is potentially valuable as it has many
-uses for various stake-holders and uses.  Imagine a development team
+uses for various stakeholders and uses.  Imagine a development team
 has agreed to implement some features at the start of an iteration. By
@@
 
-%XXX explain LDA \cite{944937}
-%XXX What has pLSI in it?
 In terms of clustering and finding topic distributions, Latent
 Dirichlet Allocation (LDA)~\cite{944937} competes with Latent Semantic
-Indexing (LSI)~\cite{1421013,1374321,10.1109/ICPC.2007.13,10.1109/ICPC.2006.17}, probabilistic Latent Semantic Indexing (pLSI) and
-semantic clustering~\cite{1698774,1566153}. These tools are used for document modelling,
-document clustering and collaborative filtering. What LDA attempts to
-do is encode documents as mixture model of topics.  LDA is used in
-software engineering
+Indexing
+(LSI)~\cite{1421013,1374321,10.1109/ICPC.2007.13,10.1109/ICPC.2006.17},
+probabilistic Latent Semantic Indexing (pLSI)~\cite{944937} and
+semantic clustering~\cite{1698774,1566153}. These tools are used for
+document modelling, document clustering and collaborative
+filtering. What LDA attempts to do is encode documents as mixture
+model of topics.  LDA is used in software engineering
 literature~\cite{lukins2008,10.1109/MSR.2007.20,NIPS2007637,1321709}
@@
   \caption{Compact Trend Plot Topics per Month for MaxDB 7.500. The
-    X-axis is time in months, the Y axis is used to stack topics
+    x-axis is time in months, the Y axis is used to stack topics
     occurring at the same time. Trends that are continuous are
@@
   \caption{Trend-time Line: Trends plotted over time of MaxDB 7.500.
-Time in months are plotted along the X-Axis, each row on the Y-axis is associated with a trend ranked by size in descending order}
+Time in months are plotted along the x-Axis, each row on the y-axis is associated with a trend ranked by size in descending order}
   \label{fig:trendtimeline}
@@
 of the Grok query language; Gnuplot, a graph plotting package; lda-c,
-a LDA package implemented by Biel; and our topic plotter, implemented
+a LDA package implemented by Blei et al. \cite{944937}; and our topic plotter, implemented
 in Haskell.
@@
 
-% The thing that is most disappointing is that the argument is entire qualitative.  Can you give examples to make this sound less hypothetical?
+% The thing that is most disappointing is that the argument is entire
+% qualitative.  Can you give examples to make this sound less
+% hypothetical?
 

### 2009-04-09 lda-paper.tex
Some edits and some more maxdb75 stuff Need more maxdb75 stuff

@@
 
+%XXX Edit
+% Was reading lda.report
+Bug tracker URLs were common during some months. For instance in the
+last month of MaxDB 7.500 development every topic contained 1 unique
+Bug tracker URL, this pattern did not occur in the previou month. We
+investigated the actual revisions and we did find there was difference
+in behaviour between and that the last month of development commits
+were mostly referencing the bug tracker. If the topics of one month
+were 20 or more unique bug tickets being addressed, the global LDA
+analysis would probably miss this, yet these bug tickets were the
+focus of development for that month and not necessarily globally
+relevant.
+
 \subsection{Compare to topics over entire development}

### 2009-04-09 lda-paper.tex
some edits but more data to look at

@@
 found there were common words across topic clusters, such as
-``diffs'', ``annotate'' and ``history'', these words probably should
+\emph{diffs}, \emph{annotate} and \emph{history}, these words probably should
 be viewed as stop words. There were notable transitional topics such
-as in the first window the word ``bitkeeper'' appears because MySQL
+as in the first window the word \emph{bitkeeper} appears because MySQL
 adopted Bitkeeper for their source control system yet in the following
-windows there were no mentions of Bitkeeper. ``RENAME'' also only
+windows there were no mentions of Bitkeeper. \emph{RENAME} also only
 appeared once in the first window and never again. Per each topic we
@@
 dominated by one large trend containing tokens that were mostly stop
-words such as ``the to a for of is it'' and references to one author
-\texttt{carlosga05}.
+words such as references to one author
+\texttt{carlosga05} and the words: \emph{the, to, a, for, of, is, it}.
 
@@
 build system, architecture and platform support. It specifically
-mentioned tokens such as ``Solaris port x86''.
+mentioned tokens such as \emph{Solaris}, \emph{port} and \emph{x86}.
 
@@
  In these cases, the developers would express thanks in their changelog comments, such as ``Thanks, Bill Lam''.  Other easily discernible topics included 
- tokens and topics such as: ``compiler workarounds'', ``nightly updates'', packets and MSVC++,
+ tokens and topics such as: \emph{compiler workarounds}, \emph{nightly updates}, \emph{packets} and \emph{MSVC++}.
 
@@
 % Was reading lda.report
-Bug tracker URLs were common during some months. For instance in the
-last month of MaxDB 7.500 development every topic contained 1 unique
-Bug tracker URL, this pattern did not occur in the previou month. We
-investigated the actual revisions and we did find there was difference
-in behaviour between and that the last month of development commits
-were mostly referencing the bug tracker. If the topics of one month
-were 20 or more unique bug tickets being addressed, the global LDA
-analysis would probably miss this, yet these bug tickets were the
-focus of development for that month and not necessarily globally
-relevant.
+Bug tracker URLs dominated unique topics during some months. For
+instance in the last month of MaxDB 7.500 development every topic
+contained 1 unique Bug tracker URL, this pattern did not occur in the
+previou month. We investigated the actual revisions and we did find
+there was difference in behaviour between and that the last month of
+development commits were mostly referencing the bug tracker. If the
+topics of one month were 20 or more unique bug tickets being
+addressed, the global LDA analysis would probably miss this, yet these
+bug tickets were the focus of development for that month and not
+necessarily globally relevant.
+
+We noticed that Perforce and Changelist were in just about every
+topic, so we added them to the stop words list. Longer trends were
+shortened and in some cases the total number of topics found per month
+was less. Evaluating different similarity thresholds showed that by
+removing common words one reduces the general similarity of topics,
+that said the larger topics still existed. Thus if you add more
+relevant stop words one should tune your topic similarity parameters
+to handle this change.
 

### 2009-04-09 lda-paper.tex
submitted to migod and ric

@@
 
+The query optimizer was a topic which recurred multiple times during
+MaxDB's development. In our plots, topics which mention optimizer
+occur four times yet in the large static plot (Figure
+\ref{fig:statictopics}) it is not in any of the topics. A query
+optimizer is rather important to an DBMS, but as we've shown it just
+doesn't pop up as a topic on its own. We tried to remove words to see
+if we could get optimizer topic out, after removing stop words and
+then two of the most common words the global analysis finally had a
+topic with optimizer in its topic ten words. This shows that optimizer
+was important but it had been obscured, while it would have been
+noticed using a more locally relevant approach that we propose.
+
 We noticed that Perforce and Changelist were in just about every
@@
 topic into its proper context.
+
+XXX Numbers and more argument
+
+Even with our liberal topic similarity metrics that produced both
+long and short trends, we showed that there are only a few trends in a
+repository which recur. Since so few trends recur and so many
+trends appear only once this suggests that global topic analysis might
+be ignoring locally unique topics.
+
+The utility of global topic analysis is questionable if the value of
+information decreases as it becomes older. Perhaps older trends will
+dominate the topic analysis. Windowed localized topic analysis show
+what are the unique topics.
 
@@
 
-Even with our liberal topic similarity metrics that produced both
-long and short trends, we showed that there are only a few trends in a
-repository which recur. Since so few trends recur and so many
-trends appear only once this suggests that global topic analysis might
-be ignoring locally unique topics.
+% Even with our liberal topic similarity metrics that produced both
+% long and short trends, we showed that there are only a few trends in a
+% repository which recur. Since so few trends recur and so many
+% trends appear only once this suggests that global topic analysis might
+% be ignoring locally unique topics.
 
-The utility of global topic analysis is questionable if the value of
-information decreases as it becomes older. Perhaps older trends will
-dominate the topic analysis. Windowed localized topic analysis show
-what are the unique topics.
+% The utility of global topic analysis is questionable if the value of
+% information decreases as it becomes older. Perhaps older trends will
+% dominate the topic analysis. Windowed localized topic analysis show
+% what are the unique topics.
+
+XXX summarize our position
+
+XXX Add more detail here
 

### 2009-04-09 dist.sh
missing files

@@ -1,0 +1,9 @@
+#make clean paper1.pdf
+#FILE=indentation.paper1.`date +"%Y%m%d.%H%M%S.%h.%d.%Y"`.pdf
+#cp paper1.pdf $FILE
+#vex $FILE
+FILEO=lda-paper.pdf
+make clean $FILEO 
+FILE=lda-paper.`date +"%Y%m%d.%H%M%S.%h.%d.%Y"`.pdf
+cp $FILEO $FILE
+vex $FILE

### 2009-04-10 lda-paper.tex
in the middle of hacking some numbers

@@
 
-The plots we produced of MaxDB 7.500 were 
-unlike those of the other systems, as 
- there was a
-period where no real development occurred and thus there were no topics or trends
-whatsoever. We have two periods to evaluate for MaxDB 7.500, the first
-period from June 2004 to jan 2005, and the second period from June
-2005 to June 2006.
+The plots we produced of MaxDB 7.500 were unlike those of the other
+systems, as there was a period where no real development occurred and
+thus there were no topics or trends whatsoever. We have two periods to
+evaluate for MaxDB 7.500, the first period from June 2004 to jan 2005,
+and the second period from June 2005 to June 2006.
 
 The largest common trend has references to build system files like
-SYSDD.punix and MONITOR.punix etc. Other tokens mentioned are Sutcheck
-v1.09, the prefix SUT stands for Storable Unit Type. Sutcheck would
-also automate check-ins using a Perforce SCS tool.
+SYSDD.punix and MONITOR.punix.  This trend is partially displayed at
+the top of the zoomed in compressed view plot (Figure
+\ref{fig:zoomedsmear}) and at the top of the histogram plot (Figure
+\ref{fig:histogram}).  Other tokens mentioned are Sutcheck v1.09, the
+prefix SUT stands for Storable Unit Type. Sutcheck would also automate
+check-ins using a Perforce SCS tool which was exported to CVS.
 
 The second largest common trend seems to be a side effect of an
-automated check-in which is annotated as ``implicit check-in''. These
-were check-ins that were produced when importing changes from an
-external Perforce repository.
+automated check-in which is annotated as ``implicit check-in'' (see
+the bottom of Figure \ref{fig:zoomedsmear}). These were check-ins that
+were produced when importing changes from an external Perforce
+repository.
 
-The third most common trend, seen on Figure \ref{fig:trendtimeline}, seemed to
-include tokens related to operating system support, such as Linux and
-Windows, as well as architecture support, AMD64 and Opteron. The word
-problem was common among all of those trends. This trend seemed
-related to the smaller fourth largest trend which occurred nearby it
-that had AMD64 and Windows tokens.
+The third most common trend, seen on Figure \ref{fig:trendtimeline},
+seemed to include tokens related to operating system support, such as
+Linux and Windows, as well as architecture support, AMD64 and
+Opteron. The word \emph{problem} was common among all of these
+trends. This trend seemed related to the smaller fourth largest trend
+which occurred nearby which referenced tokens \emph{AMD64} and
+\emph{Windows}, this shows that topics can overlap but still not match
+via our similarity measure.
 
@@
 
-The query optimizer was a topic which recurred multiple times during
+The query optimizer was a topic which recurred during
 MaxDB's development. In our plots, topics which mention optimizer
@@
 to handle this change.
+
+UNICODE
+
+DEC/OSF1
+
+ODBC
+
+crash
+
+% PHP yet Perl is mentioned is less but it part
+
+SQLCLI
+%V75
+
+SYSDD.cpnix
+backup
+
+select
+make
+memory
+view
+
+debug
+
+% V75 is mentioned many times maxdb version guh
 
@@
 topic into its proper context.
+
+If we approach the difference via common tokens we can see that common
+token tend to dominate both results, at least in the case of MaxDB
+7.500 our local topic approach netted us topics which contained these
+relatively important and common words: \emph{UNICODE},
+\emph{DEC/OSF1}, \emph{ODBC}, \emph{crash}, \emph{SQLCLI},
+\emph{SYSDD.cpnix}, \emph{backup}, \emph{select}, \emph{make},
+\emph{memory}, \emph{view}, and finally \emph{debug}. None of these
+tokens were part of the static topic analysis topics but they were
+part of 494 commits (5\% of the entire system) to MaxDB 7.500.
+
+% Revisions 26+1+103+19+20+48+35+50+77+49+48+18
+% Revisions 494 
+% Topics 1+1+19+1+5+10+7+10+11+9+9+7
+% Topics 90
 

### 2009-04-10 data/countrevs.sh

@@ -1,0 +1,5 @@
+FILE="UNICODE DEC/OSF1 ODBC crash SQLCLI SYSDD.cpnix kacup select make memory view debug"
+for file in $FILE
+do
+	fgrep -w -i -- "$file" query1.mrs2
+done

### 2009-04-10 data/counttopics.sh

@@ -1,0 +1,5 @@
+FILE="UNICODE DEC/OSF1 ODBC crash SQLCLI SYSDD.cpnix kacup select make memory view debug"
+for file in $FILE
+do
+	fgrep -w -i -- "$file" maxdb7500-default-stopwords.report
+done

### 2009-04-10 lda-paper.tex
Added more content more edits, submitted to migod and holt

@@
 tried to identify the purpose of the topic; to our surprise we found
-that even with onlyy a little familiarity with the code base that
+that even with only a little familiarity with the code base that
 naming the topic was straightforward.
@@
 which fill in these gaps quickly.  Different instances of the same
-trend share the same colour; apart from that, the color of trends is
+trend share the same color; apart from that, the color of trends is
 randomly chosen and color similarity is not meaningful.
-The compressed view is convienant because it shows all of the
+The compressed view is convenient because it shows all of the
 information and the repeating continuous trends are easy to pick out,
@@
   \includegraphics[width=0.45\textwidth]{histogram-cropped-scaled}
-  \caption{Top part of trend histogram, ordered by topic occurance for MaxDB 7.500.
+  \caption{Top part of trend histogram, ordered by topic occurrence for MaxDB 7.500.
 X axis determines the number of continuous months of a trend.
@@
 
-We extracted the repositories of: PostGreSQL, MaxDB and
+We extracted the repositories of: PosgreSQL, MaxDB and
 Firebird.
@@
 
-\subsection{Postgresql}
+\subsection{PostgreSQL}
 %   3.   mysql
@@
 thus there were no topics or trends whatsoever. We have two periods to
-evaluate for MaxDB 7.500, the first period from June 2004 to jan 2005,
+evaluate for MaxDB 7.500, the first period from June 2004 to January 2005,
 and the second period from June 2005 to June 2006.
@@
 \ref{fig:statictopics}) it is not in any of the topics. A query
-optimizer is rather important to an DBMS, but as we've shown it just
-doesn't pop up as a topic on its own. We tried to remove words to see
+optimizer is rather important to an DBMS, but as we have shown it just
+does not pop up as a topic on its own. We tried to remove words to see
 if we could get optimizer topic out, after removing stop words and
@@
 noticed using a more locally relevant approach that we propose.
+
+We noticed that PHP was in a third of the commits that Perl was in
+(not the same commits) in as many commits as Perl yet PHP did not
+appear in the global static topic analysis topics while Perl did. Our
+local topic analysis mentioned PHP in 5 different topics, yet
+mentioned Perl in 4 different topics. Perhaps the Perl mentions were
+not as independent topics as the PHP topics were.
 
@@
 
-UNICODE
-
-DEC/OSF1
-
-ODBC
-
-crash
+% UNICODE
+
+% DEC/OSF1
+
+% ODBC
+
+% crash
 
@@
 
-SQLCLI
-%V75
-
-SYSDD.cpnix
-backup
-
-select
-make
-memory
-view
-
-debug
+% SQLCLI
+% %V75
+
+% SYSDD.cpnix
+% backup
+
+% select
+% make
+% memory
+% view
+
+% debug
 
@@
 
+% Revisions 26+1+103+19+20+48+35+50+77+49+48+18
+% Revisions 494 
+% Topics 1+1+19+1+5+10+7+10+11+9+9+7
+% Topics 90
+
 If we approach the difference via common tokens we can see that common
@@
 \emph{memory}, \emph{view}, and finally \emph{debug}. None of these
-tokens were part of the static topic analysis topics but they were
-part of 494 commits (5\% of the entire system) to MaxDB 7.500.
-
-% Revisions 26+1+103+19+20+48+35+50+77+49+48+18
-% Revisions 494 
-% Topics 1+1+19+1+5+10+7+10+11+9+9+7
-% Topics 90
-
-XXX Numbers and more argument
+tokens were part of the static topic analysis topics, but there were
+part of 566 commits (6\% of the entire system) to MaxDB 7.500 which
+contained these. These tokens were part of 87 out of 500 of our
+locally generated topics (ignoring similar topics).
+
+%XXX Numbers and more argument
 
@@
 
-XXX summarize our position
-
-XXX Add more detail here
-
-We presented multiple visualizations that focused on different
-aspects of the data: temporality of trends, trend size, and a
-condensed overall trend view. The condensed trend view shows much more
-information than the views that global analysis could show.
+We demonstrated the application of windowed topic analysis, topic
+analysis applied to commit messages during periods of
+development. This approach differed from previous work, as previous
+work would apply topic analysis globally and then analyze related
+documents. Instead we show that many topics that exist locally are
+relevant and interesting yet would often not be included in global
+topic analysis. We handle the case of recurring topics by using a
+topic similarity measure which looks for topics which recur and mutate
+throughout the development of the software project.
+
+The value of windowed topic analysis was demonstrated on a case study
+of three database systems with a particular focus on MaxDB 7.500. We
+showed that static topic analysis missed important database concepts
+such as ODBC.
+
+%XXX summarize our position
+
+%XXX Add more detail here
+
+We presented multiple visualizations that focused on different aspects
+of the data: temporality of trends, trend size, and a condensed
+overall trend view. The condensed trend view shows much more
+information than the views that global analysis could show, it
+indicates how focused a period is by the total number of topics as
+well shows topics by similarity so one can track trends across time.
+Our histogram trend-view immediately hi-lights and measures the size of
+trends while our time-line trend view shows how a topic reoccurs over
+time. These visualizations help us get an over feeling about the
+common and unique topics that developers focus on. If implemented
+properly a user can easily zoom and query a topic.
 

### 2009-04-10 data/countrevs.sh

@@ -1,5 +1,5 @@
 FILE="UNICODE DEC/OSF1 ODBC crash SQLCLI SYSDD.cpnix kacup select make memory view debug"
-for file in $FILE
-do
+(for file in $FILE
+ do
 	fgrep -w -i -- "$file" query1.mrs2
-done
+ done) | sort -u |  wc -l

### 2009-04-10 data/counttopics.sh

@@ -1,5 +1,5 @@
 FILE="UNICODE DEC/OSF1 ODBC crash SQLCLI SYSDD.cpnix kacup select make memory view debug"
-for file in $FILE
-do
+(for file in $FILE
+ do
 	fgrep -w -i -- "$file" maxdb7500-default-stopwords.report
-done
+ done ) | sort -u | wc -l

### 2009-04-10 lda-paper.tex
More edits and changes cleaned up descriptions

@@
 
-Our methodology is to first extract the commits from SCS repositories such
-as CVS or Bitkeeper. Then we extract the revisions and commits,
+Our methodology is to first extract the commits from SCS repositories
+such as CVS or Bitkeeper. Then we extract the revisions and commits,
 grouping the revisions into commits if necessary. We extract the
-commit log message of each commit, count the occurrence of each word in
-the message, remove stop words, and then produce word distributions
+commit log message of each commit, count the occurrence of each word
+in the message, remove stop words, and then produce word distributions
 for each message.
@@
 revisions. CVSSuck and rsync mirrored CVS repositories while bt2csv
-mirrored web accessible bitkeeper repositories.
+mirrored web accessible BitKeeper repositories.
 
@@
 
-The document we are analyzing are the changelog comments i.e., the comments that are added  when a revision to the source code is committed.
+The documents we are analyzing are the commit log comments i.e., the
+comments that are added when revisions to the project are committed.
 Documents are converted into word count distributions. These
-distributions, which are then normalized by by the size of the
-message. After all messages are processed the distributions are
-extended to include all words in all distributions (each distribution
-is extended with word bins of count 0).
+distributions are then normalized by the size of the message. After
+all messages are processed the distributions are extended to include
+all words within all of the  distributions.
 
@@
 granularities for the study we think that a month is a sizable unit of
-development, 4 weeks of development, which we consider to be fine
-enough to show meaningful shifts but coarse enough not to show too
-much detail.
+development, which we consider to be fine enough to show meaningful
+shifts in focus, but coarse enough not to show too much detail.
 
@@
 each window and extract the top $N$ topic clusters, we used $20$
-topics in our case study. Our Topic Analysis tool is an implementation
-of LDA, but we could have used other similar tools like LSI, but
-previous showed more promising topic analysis with LDA.
-
-This is a very slow step as LDA takes a long time to run, and takes
-even longer since we run it once per window. We note that this step is
-a slow one, as executing even one instance of LDA involves significant
-computation, and we perform LDA once per window.
+topics in our case studies. Our Topic Analysis tool is an
+implementation of LDA, but we could have used other similar tools like
+LSI, but previous work showed more promising topic analysis results
+with LDA.  We note that this step is a slow one, as executing even one
+instance of LDA involves significant computation, and we perform LDA
+once per window.
 
@@
 topic is compared to each other topic in the system, given a certain
-threshold of top 10 similarity, such as 8/10 words. Then given this topic
-similarity matrix, we find the transitive closure of similar topics;
-that is, we fill flood along the similarity arcs until we have
+threshold of top 10 similarity, such as 8 out of 10 words. Then given
+this topic similarity matrix, we find the transitive closure of
+similar topics; that is, if we model topics are nodes and arcs as
+similarity, we fill flood along the similarity arcs until we have
 partitioned the topics into clusters of similar topics. Figure
@@
 topic similarity. These clusters of topics are called trends. A trend
-is probably interesting and valuable if there is more than one topic in
-the trend.
-
-This technique has weaknesses in that nodes that are  a few neighbors away in
-similarity might not share any similar words.  We do this because we
-want to be able to color or highlight topics that are related to each
-other and track them throughout time.
+is probably interesting and valuable if there is more than one topic
+in the trend.
+
+This technique has weaknesses in that nodes that are a few neighbors
+away in similarity might not share any similar words.  We do this
+because we want to be able to color or highlight topics that are
+related to each other and track them throughout time.
 
@@
 
-We have devised several ways to visualize these topics and
-trends.  For all of these plots if we find trends that have continuous
-segments, then we plot those segments as joined across time.  We have a
-compressed stacked trend view, shown in Figure \ref{fig:topicsmear}
-and Figure \ref{fig:zoomedsmear}, 
-which displays trends as they occur
+We have devised several ways to visualize these topics and trends.
+For all of these plots if we find trends that have continuous
+segments, then we plot those segments as joined across time.  We have
+a compressed stacked trend view, shown in Figure \ref{fig:topicsmear}
+and Figure \ref{fig:zoomedsmear}, which displays trends as they occur
 on the time-based x-axis (placement along the y-axis is arbitrary).
- Another plot we have is the trend-time line plot, shown in
-Figure \ref{fig:trendtimeline}, each trend gets it own distinct y-value ,
+Another plot we have is the trend-time line plot, shown in Figure
+\ref{fig:trendtimeline}, each trend gets it own distinct y-value ,
 while the x-axis is time; these topics are then each plotted on their
@@
 plot, shown in Figure \ref{fig:histogram} where we plot each trend on
-its own line but stack up the segments of the trend, much like a plot
-of counts or a histogram.
+its own line but stack up the segments of the trend, much like a
+histogram.
 
@@
 trend share the same color; apart from that, the color of trends is
-randomly chosen and color similarity is not meaningful.
-The compressed view is convenient because it shows all of the
-information and the repeating continuous trends are easy to pick out,
-although repeating discontinuous trends are harder to spot.
+randomly chosen and color similarity is not meaningful.  The
+compressed view is convenient because it shows all of the information
+and the repeating continuous trends are easy to pick out, although
+repeating discontinuous trends are harder to spot.
 
@@
 \ref{fig:trendtimeline}), it displays repeating trends more clearly by
-dedicating a horizontal line for trend segments dedicated to one
-trend. This means that if a trend contains discontinuous segments then the
-segments appear on the same line. 
-However, the least common trends need to be pruned away or the view will be cluttered with a lot of noise.
+dedicating a horizontal line for trend segments belonging to one
+trend. This means that if a trend contains discontinuous segments then
+the segments appear on the same line.  However, the least common
+trends need to be pruned away or the view will be very long.
 
@@
 plot.  However, in this view the trends are plotted together by
-stacking to their row so time information is lost, and the trends are
-ordered by the number of topics in the trend.  The histogram plot
-shows the count of instances of a trend and thus indicates which
-trends occur the most. Unfortunately due to the large number of topics
-given allotted space it is often best to crop off the trends with only
-one topic, otherwise the tail is long.
+stacking to the left of their row, thus time information is lost.  The
+trends are ordered by the number of topics in the trend.  The
+histogram plot shows the count of instances of a trend and thus
+indicates which trends occur the most. Unfortunately due to the large
+number of topics given allotted space it is often best to crop off the
+trends with only one topic, otherwise the tail is long.
 
@@
   \includegraphics[width=1.0\textwidth]{fixed-time-smear-plot-scaled}
-  \caption{Compact Trend Plot Topics per Month for MaxDB 7.500. The
-    x-axis is time in months, the Y axis is used to stack topics
-    occurring at the same time. Trends that are continuous are
+  \caption{Compact trend plot shows topics per month for MaxDB
+    7.500. The x-axis is time in months, the Y axis is used to stack
+    topics occurring at the same time. Trends that are continuous are
     plotted as continuous blocks. The top 10 words in the topics are
@@
   \includegraphics[width=0.45\textwidth]{fixed-time-smear-plot-cropped}
-  \caption{Zoomed in Slice Compact Trend Plot of Topics per Month for MaxDB 7.500. The topic text is visible in each topic box. Trends are plotted across time continuously.}
+  \caption{A zoomed in slice of compact trend plot of topics per month of MaxDB 7.500. The topic text is visible in each topic box. Trends are plotted across time continuously.}
   \label{fig:zoomedsmear}
@@
   \includegraphics[width=0.45\textwidth]{class-smear-plot-crop-scaled}
-  \caption{Trend-time Line: Trends plotted over time of MaxDB 7.500.
+  \caption{Trend-time Line: Trends plotted per month of MaxDB 7.500.
 Time in months are plotted along the x-Axis, each row on the y-axis is associated with a trend ranked by size in descending order}
@@
   \includegraphics[width=0.45\textwidth]{histogram-cropped-scaled}
-  \caption{Top part of trend histogram, ordered by topic occurrence for MaxDB 7.500.
-X axis determines the number of continuous months of a trend.
-Trends are ranked  by the number of topics that a trend contains in descending order.}
+  \caption{The top part of a trend histogram of MaxDB 7.500, ordered
+    by topic occurrence. X axis determines the number of continuous
+    months of a trend. Trends are ranked by the number of topics that
+    a trend contains in descending order.}
   \label{fig:histogram}
@@
 of the Grok query language; Gnuplot, a graph plotting package; lda-c,
-a LDA package implemented by Blei et al. \cite{944937}; and our topic plotter, implemented
-in Haskell.
+a LDA package implemented by Blei et al. \cite{944937}; and our topic
+plotter, implemented in Haskell.
 
@@
 
-When we examined PostgreSQL, we did not find many trends with two or more topics; however, we did find one large trend which stretched across most of its existence.
- Much of the large trend consisted of stop words but it did
-include topics which had non-stop words such as \emph{patch}, \emph{fix}, \emph{update}.
-
-The second largest topic was interesting because it concerned changes from
-1998, from developers using Central Europe Time (CEST), where the two
-notable tokens were 1998 and CEST. These changes appear to be patches
-submitted by one particular developer who lived in that timezone.
-
-The third largest trend we found appeared to concerned with copying
-and moving files between branches as it referenced the old branch.
-branches.  We inferred this trend based on the fact that the old
-version identifier was referenced, and we know this this is common
-practice when performing a branch merge, since CVS cannot record such
-information directly.
-
-Once stop words were removed the some of the purposes became more
-clear. The large trend disappeared, the largest trend became bug fixes
-submitted by a developer named Dal Zotto which is a short trend.
+When we examined PostgreSQL, we did not find many trends with two or
+more topics with a similarity of $7/10$. These trends were not very
+large, lasting on 3 months at most. The first, second largest trends
+were dedicated to the work of two external developers: Dal Zotto, Dan
+McGuirk. The fifth largest trend related to work by PostgreSQL developer D'Arcy. Other
+topics of the larger trends were changes to the TODO list in HTML, and
+time string formating topics relating to timezones.
+
+If we kept the stop words we found that the large trend consisted
+mostly of stop words and non-stop words such as \emph{patch},
+\emph{fix}, \emph{update}.  By decreasing the similarity constraint to
+$1/2$ similarity the largest most the largest most common trend, which
+stretched across development, contained these same words
+(\emph{patch}, \emph{fix}, \emph{update}). The second most common
+topic mentions Dal Zotto, while the third most common topic mentions
+the \texttt{[PATCHES]} mailinglist and the names of some patch
+contributors.  Other repeating topics refer to portability with Win32,
+Central Europe Time (CEST) from email headers, issues with ALTER TABLE
+and CVS branch merging (because CVS doesn't record merges explicitly).
+
+% When we examined PostgreSQL, we did not find many trends with two or
+% more topics; however, we did find one large trend which stretched
+% across most of its existence.  Much of the large trend consisted of
+% stop words but it did include topics which had non-stop words such as
+% \emph{patch}, \emph{fix}, \emph{update}.
+
+% The second largest topic was interesting because it concerned changes from
+% 1998, from developers using Central Europe Time (CEST), where the two
+% notable tokens were 1998 and CEST. These changes appear to be patches
+% submitted by one particular developer who lived in that timezone.
+
+% The third largest trend we found appeared to concerned with copying
+% and moving files between branches as it referenced the old branch.
+% branches.  We inferred this trend based on the fact that the old
+% version identifier was referenced, and we know this this is common
+% practice when performing a branch merge, since CVS cannot record such
+% information directly.
+
+% Once stop words were removed the some of the purposes became more
+% clear. The large trend disappeared, the largest trend became bug fixes
+% submitted by a developer named Dal Zotto which is a short trend.
 
@@
 
-We tracked Firebird from August 2000 to January 2006.
-Our first attempt with Firebird resulted in a set of trends that were
-dominated by one large trend containing tokens that were mostly stop
-words such as references to one author
-\texttt{carlosga05} and the words: \emph{the, to, a, for, of, is, it}.
-
-The second largest trend related to branching and possibly merging; it
-contained language such as ``was added on branch'' and contained
-references to tags and branches.
-
-The third largest trend contained references to the changelog, the
-build system, architecture and platform support. It specifically
-mentioned tokens such as \emph{Solaris}, \emph{port} and \emph{x86}.
-
-Other trends included topics regarding AIX PPC compilation, updating
-the build process, internationalization and UTF8, Darwin build support
-and bug fixing.
-
-Topics that were not trends but appeared to be  interesting were mostly external bug
-fixes submitted to the project.
- In these cases, the developers would express thanks in their changelog comments, such as ``Thanks, Bill Lam''.  Other easily discernible topics included 
- tokens and topics such as: \emph{compiler workarounds}, \emph{nightly updates}, \emph{packets} and \emph{MSVC++}.
+We tracked Firebird from August 2000 to January 2006. We found that with a
+similarity of $7/10$ that Firebird had far more continuous and
+recurring trends than PosgreSQL.  The first large trend was segmented
+across time but explicitly references one author \texttt{carlosga05}
+and words like \emph{added}, \emph{fixed} and \emph{updated}.
+
+The second largest trend was during the month of March 2001 relating
+to incremental building and the porting of developer Konstantin's
+Solaris port of the Firebird build files. The third largest trend was
+about JDBC, which is how Firebird and Java communicate.  Other trends
+included topics regarding AIX PPC compilation, updating the build
+process, internationalization and UTF8, Darwin build support and bug
+fixing.
+
+Topics that were not trends but appeared to be interesting were mostly
+external bug fixes submitted to the project.  In these cases, the
+developers would express gratitude in their commit log comments, such
+as ``Thanks, Bill Lam''.  Other easily discernible topics included
+tokens and phrases such as: \emph{compiler workarounds}, \emph{nightly
+  updates}, \emph{packets} and \emph{MSVC++}.
+
+% We tracked Firebird from August 2000 to January 2006.
+% Our first attempt with Firebird resulted in a set of trends that were
+% dominated by one large trend containing tokens that were mostly stop
+% words such as references to one author
+% \texttt{carlosga05} and the words: \emph{the, to, a, for, of, is, it}.
+
+% The second largest trend related to branching and possibly merging; it
+% contained language such as ``was added on branch'' and contained
+% references to tags and branches.
+
+% The third largest trend contained references to the changelog, the
+% build system, architecture and platform support. It specifically
+% mentioned tokens such as \emph{Solaris}, \emph{port} and \emph{x86}.
+
+% Other trends included topics regarding AIX PPC compilation, updating
+% the build process, internationalization and UTF8, Darwin build support
+% and bug fixing.
+
+% Topics that were not trends but appeared to be  interesting were mostly external bug
+% fixes submitted to the project.
+%  In these cases, the developers would express thanks in their commit log comments, such as ``Thanks, Bill Lam''.  Other easily discernible topics included 
+%  tokens and topics such as: \emph{compiler workarounds}, \emph{nightly updates}, \emph{packets} and \emph{MSVC++}.
 
@@
 systems, as there was a period where no real development occurred and
-thus there were no topics or trends whatsoever. We have two periods to
-evaluate for MaxDB 7.500, the first period from June 2004 to January 2005,
-and the second period from June 2005 to June 2006.
+thus there were no topics or trends whatsoever. Using a topic
+similarity of $7/10$ we evaluated of MaxDB 7.500, the first period
+from June 2004 to January 2005, and the second period from June 2005
+to June 2006.
 
 The largest common trend has references to build system files like
-SYSDD.punix and MONITOR.punix.  This trend is partially displayed at
-the top of the zoomed in compressed view plot (Figure
-\ref{fig:zoomedsmear}) and at the top of the histogram plot (Figure
-\ref{fig:histogram}).  Other tokens mentioned are Sutcheck v1.09, the
-prefix SUT stands for Storable Unit Type. Sutcheck would also automate
-check-ins using a Perforce SCS tool which was exported to CVS.
+\texttt{SYSDD.punix} and \texttt{MONITOR.punix}.  This trend is
+partially displayed at the top of the zoomed in compressed view plot
+(Figure \ref{fig:zoomedsmear}) and at the top of the histogram plot
+(Figure \ref{fig:histogram}).  Other tokens mentioned are
+\emph{Sutcheck v1.09}, the prefix SUT stands for Storable Unit
+Type. Sutcheck would also automate check-ins using a Perforce SCS tool
+which was exporting check-ins to CVS.
 
@@
 trends. This trend seemed related to the smaller fourth largest trend
-which occurred nearby which referenced tokens \emph{AMD64} and
-\emph{Windows}, this shows that topics can overlap but still not match
-via our similarity measure.
+that had tokens \emph{AMD64} and \emph{Windows}. This shows that
+topics can overlap but still not match via our similarity measure.
 
@@
 contained 1 unique Bug tracker URL, this pattern did not occur in the
-previou month. We investigated the actual revisions and we did find
-there was difference in behaviour between and that the last month of
+previous month. We investigated the actual revisions and we did find
+there was difference in behavior between and that the last month of
 development commits were mostly referencing the bug tracker. If the
@@
 
-We noticed that PHP was in a third of the commits that Perl was in
-(not the same commits) in as many commits as Perl yet PHP did not
-appear in the global static topic analysis topics while Perl did. Our
-local topic analysis mentioned PHP in 5 different topics, yet
-mentioned Perl in 4 different topics. Perhaps the Perl mentions were
-not as independent topics as the PHP topics were.
+We noticed that commits which mentioned \emph{PHP} occured two thirds
+less frequently than commits which mentioned \emph{Perl}, but
+\emph{Perl} related topics appeared in the global static topics for
+MaxDB while \emph{PHP} related topics did not.  Our local topic
+analysis mentioned \emph{PHP} in 5 different topics, yet only
+mentioned \emph{Perl} in 4 different topics. Perhaps the Perl mentions
+were not as independent topics as the PHP topics were.
 
@@
 relevant stop words one should tune your topic similarity parameters
-to handle this change.
+to handle these changes.
 
@@
 
-If we approach the difference via common tokens we can see that common
-token tend to dominate both results, at least in the case of MaxDB
-7.500 our local topic approach netted us topics which contained these
-relatively important and common words: \emph{UNICODE},
-\emph{DEC/OSF1}, \emph{ODBC}, \emph{crash}, \emph{SQLCLI},
-\emph{SYSDD.cpnix}, \emph{backup}, \emph{select}, \emph{make},
-\emph{memory}, \emph{view}, and finally \emph{debug}. None of these
-tokens were part of the static topic analysis topics, but there were
-part of 566 commits (6\% of the entire system) to MaxDB 7.500 which
-contained these. These tokens were part of 87 out of 500 of our
-locally generated topics (ignoring similar topics).
+If we approach the difference of static global topic analysis and
+windowed topic analysis via common tokens we can see that common token
+tend to dominate both results, at least in the case of MaxDB 7.500 our
+local topic approach netted us topics which contained these relatively
+important and common words: \emph{UNICODE}, \emph{DEC/OSF1},
+\emph{ODBC}, \emph{crash}, \emph{SQLCLI}, \emph{SYSDD.cpnix},
+\emph{backup}, \emph{select}, \emph{make}, \emph{memory}, \emph{view},
+and finally \emph{debug}. None of these tokens were part of the static
+topic analysis topics, but they were part of 566 commits (6\% of the
+entire system) to MaxDB 7.500. These tokens were part of 87 out of 500
+(25 months, 20 topics per month) of our locally generated topics.
 
@@
 remove them, but perhaps dropping some of these stop words would aid
-the clarity of such topic cluster; alternatively it might cause our
-topic similarity plots to be fundamentally different.
-
-The number of
-commits per month is inconsistent as some months have a lot of changes
-while others months have almost none.
+the clarity of topics produced through topic analysis; alternatively
+it might cause our topic similarity plots to be fundamentally
+different.
+
+The number of commits per month is inconsistent as some months have a
+lot of changes while others months have almost none.
 %   1.   validation
@@
 
-We demonstrated the application of windowed topic analysis, topic
-analysis applied to commit messages during periods of
-development. This approach differed from previous work, as previous
-work would apply topic analysis globally and then analyze related
-documents. Instead we show that many topics that exist locally are
-relevant and interesting yet would often not be included in global
-topic analysis. We handle the case of recurring topics by using a
-topic similarity measure which looks for topics which recur and mutate
-throughout the development of the software project.
+We proposed and demonstrated the application of windowed topic
+analysis, that is, topic analysis applied to commit messages during
+periods of development. This approach differed from previous work that
+would apply topic analysis globally to a project. We showed that many
+topics that exist locally are relevant and interesting yet would often
+not be included in global topic analysis. We identify recurring topics
+with a topic similarity measure which looks for topics which recur and
+mutate throughout the development of the software project.
 
@@
 of three database systems with a particular focus on MaxDB 7.500. We
-showed that static topic analysis missed important database concepts
-such as ODBC.
+showed that static topic analysis missed important topics like ODBC
+changes.
 
@@
 We presented multiple visualizations that focused on different aspects
-of the data: temporality of trends, trend size, and a condensed
-overall trend view. The condensed trend view shows much more
+of the data: temporality of trends, trend size, and a compressed
+overall trend view. The compressed trend view shows much more
 information than the views that global analysis could show, it

### 2009-04-10 lda-paper.tex
some ric edits plus new month plot

@@
 
-As development on a software project progresses, developers will naturally
-shift their focus between different topics and tasks many times.  Managers
-and newcomer developers often seek simple ways of understanding what tasks
-have recently been worked on and how much effort has gone into each; for
-example, a manager might wonder what unexpected tasks occupied their team's
-attention during a period when they were supposed to have been implementing
-a set of new features.  Tools such as Latent Dirichlet Allocation (LDA) and
-Latent Semantic Indexing (LSI) can be used to analyze check-in comments
-over the entire lifetime of a project, and previous work on developer topic
-analysis has leveraged these tools to associate commit log comments with
-independent topics extracted from these commit log comments.  In this work, we use
-LDA and LSI to analyze periods, such as months, within a project's lifetime
-to create a time-windowed model of changing development topics.  We propose
-a visualization of this model that allows us to explore the evolving stream
-of topics of development occurring over time, and we demonstrate that
-windowed topic analysis offers strong advantages over entire lifetime topic
-analysis because of the independence of topics between windows.
+  As development on a software project progresses, developers will
+  naturally shift their focus between different topics and tasks many
+  times.  Managers and newcomer developers often seek ways of
+  understanding what tasks have recently been worked on and how much
+  effort has gone into each; for example, a manager might wonder what
+  unexpected tasks occupied their team's attention during a period
+  when they were supposed to have been implementing a set of new
+  features.  Tools such as Latent Dirichlet Allocation (LDA) and
+  Latent Semantic Indexing (LSI) can be used to analyze check-in
+  comments over the entire lifetime of a project, and previous work on
+  developer topic analysis has leveraged these tools to associate
+  commit log comments with independent topics extracted from these
+  commit log comments.  In this work, we use LDA and LSI to analyze
+  periods, such as months, within a project's lifetime to create a
+  time-windowed model of changing development topics.  We propose a
+  visualization of this model that allows us to explore the evolving
+  stream of topics of development occurring over time.  We demonstrate
+  that windowed topic analysis offers advantages over topic analysis
+  applied to a projects lifetime because many topics are quite local.
 
@@
 uses for various stakeholders and uses.  Imagine a development team
-has agreed to implement some features at the start of an iteration. By
-the end of the iteration only some of the agreed upon features are
-implemented. The developers insist they were focusing solely on the
-features agreed upon, but what actually happened? Topic analysis could
-help by teasing out the distinct topics or focusses of the
-developers. Perhaps the developers had to deal with issues.
-Developers might have grouped those orthogonal changes into one of the
+has agreed to implement some features at the start of a development
+iteration. By the end of the iteration only some of the agreed upon
+features are implemented. The developers insist they were focusing
+ on the features agreed upon, but what actually happened? Topic
+analysis could help by teasing out the distinct topics or focusses of
+the developers. Perhaps the developers had to deal with other pressing issues.
+Developers might have grouped those orthogonal issues into the
 features they were working on, thus when asked about how their time
-was spent the orthogonal changes would not be covered. This is the kind of underlying story that we
-hope that windowed topic analysis can uncover. If windowed topic
-analysis can discover important but overlooked topics, it can help
-both developers and managers. Developer can see a summary of what
-worked on, and managers can development.
-
-Commits in SCSs are not necessarily consistent, they can have multiple
-purposes, thus we hope that these multiple purposes manifest themselves as
-topics interleaved within the commits.   Others have demonstrated that by
-using powerful topic analysis techniques, such as Latent Semantic Indexing
-(LSI) and Latent Dirichlet Allocation (LDA), can aid in separating topics
-from these mixed comments. These topics often describe distinct ideas
-that are interleaved throughout the development. A topic represents a
+was spent the orthogonal changes would not be covered. This is the
+kind of underlying story that we hope that windowed topic analysis can
+uncover. If windowed topic analysis can discover important but
+overlooked topics, it can help both developers and managers. Developers
+can see a summary of what worked on, and managers can development.
+
+Commits in SCSs are not necessarily consistent and they can have
+multiple purposes. Thus we hope that our research will help uncover
+these multiple purposes that manifest themselves as topics interleaved
+within the commits.  Others have demonstrated that by using 
+topic analysis techniques, such as Latent Semantic Indexing (LSI) and
+Latent Dirichlet Allocation (LDA), can aid in separating topics from
+these mixed comments. These topics often describe distinct ideas that
+are interleaved throughout the development. A topic represents a
 groups of commit comments that are related to each other by their
@@
 these commits. Alternatively, we can determine what particular
-maintenance task was being performed are a given time.
-
-We propose to extend the use of common topic analysis techniques by
-executing them over windows of documents over time. We hope by looking
+maintenance task was being performed are at a given time.
+
+We propose to extend the use of common existing topic analysis techniques by
+applying them to windows of documents over time. We hope by looking
 at topic clusters during periods or windows over time we can identify
-continuous effort on a select group of topics as well reflect topics
-that were local to that period. Figure \ref{fig:lda} is an example of
+ongoing areas of focus on a select group of topics as well reflect more intermittent topics. Figure \ref{fig:lda} illustrates
 the kind of visualization we want to automatically generate. We want
@@
 larger period are plotted continuously. In our example we have
-summarized each topic with a single word.
+titled each topic with a single word.
 
@@
 time window and we extract the topics in that time window. We expect
-that topics in each window will change over time and the similarities
-and differences in these sets of topics will indicate both developer
-focus and changes in developer focus.
+that topics will change over time and the similarities
+and differences in these topics will indicate developer
+focus and changes in developer focus as it changes over time.
 
@@
 commits changes to files in a repository. A \emph{word distribution}
-is the summary of a message by word count. The word distribution will
-be over all messages looked at but since most words will not appear in
-a message, a word distribution is effectively a word count divided by
-the message size. A \emph{topic} is essentially a set of words that
-form a word distribution that is unique and independent within the set
-of documents in our total corpus. One could think of a topic as the
-center distribution of a group of messages. In this paper we often
-summarize topics by their top 10 most frequent words in their word
-distribution.  A \emph{trend} is a one or more similar topics which
-recur over time.  Trends are particularly important, as they indicate
-long-lived and recurring topics that may provide key insights into the
-development of the project.
+is the summary of a message by word count. Each word distribution will
+be over the words in all messages. However, most words will not appear
+in a message, a word distribution is effectively a word count divided
+by the message size. A \emph{topic} is a word distribution. It is
+essentially a set of words that form a word distribution that is
+unique and independent within the set of documents in our total
+corpus. One could think of a topic as a distribution of the centroid
+of a group of messages. In this paper we often summarize topics by the
+top 10 most frequent words of their word distribution.  A \emph{trend}
+is a one or more similar topics which recur over time.  Trends are
+particularly important, as they indicate long-lived and recurring
+topics that may provide key insights into the development of the
+project.
 
@@
 document modelling, document clustering and collaborative
-filtering. What LDA attempts to do is encode documents as mixture
-model of topics.  LDA is used in software engineering
+filtering. LDA attempts to encode documents as mixture
+model, a combination, of topics.  LDA is used in software engineering
 literature~\cite{lukins2008,10.1109/MSR.2007.20,NIPS2007637,1321709}
-is to extract topic clusters from documents such as methods, bug
+is to extract topics from documents such as methods, bug
 reports, source code and files.
@@
 code or documents that are related to the source code.  Concept location concerns how
-high level concepts relate to low level entities such as source code, for
+high level concepts relate to low level entities such as source code. For
 example, when fixing a bug how and where do the concepts in the bug
@@
 Our technique differs from the previous techniques because we apply
-LDA locally to month-long windows of commit comments, where as other
+LDA locally to month-long windows of commit comments, where-as other
 approaches apply LDA once to the entire project. This allows us to
@@
 windows. We asked LDA to make 20 topics per window and we then
-carefully examined the top 10 most frequent words in that topic.  We
+ examined the top 10 most frequent words in that topic.  We
 found there were common words across topic clusters, such as
@@
 
-A sub-sampling of the notable topic words is displayed in Table
-\ref{tab:portability}.  To name the topics we selected terms that
-seemed to best summarize those topics.  After viewing these topics, it was
-obvious that we would have to combine try to track the evolution of
-topics over the windows or rank topics by similarity, thus if they
-were similar enough we joined them into a trend. Figure \ref{fig:lda}
-displays a sample plot of the extracted topics in table
-\ref{tab:portability}.
+A sampling of the notable topic words is displayed in Table
+\ref{tab:portability}, we chose topics that we felt confident we could
+name.  To name each topic we selected a term that seemed to best
+summarized that topic.  After extracting these topics, we attempted to
+track the evolution of topics by visualizing the topics and joining
+similar topics into trends.  Figure \ref{fig:lda} displays a sample
+plot of the extracted topics in Table \ref{tab:portability}.
 
@@
 \end{tabular}
-\caption{Sub-sampled topics from MySQL 3.23, some with continuous topics, these tokens were pulled from LDA topic. Each token is a summary of one LDA generated topic from MySQL 3.23 commit logs.}
+\caption{Sampled topics from MySQL 3.23, some with continuous topics. These tokens were pulled from LDA topic. Each token is a summary of one LDA generated topic from MySQL 3.23 commit logs.}
 \label{tab:portability}
@@
   \includegraphics[width=0.9\textwidth]{lda}
-  \caption{Example of topics extracted from MySQL 3.23.
-This is the kind of plot we eventually want to produce: named topics and topic trends.
-The horizontal axis is time in months and the vertical axis used to stack topics that occur at the same time.
-Longer topics are topics which recur in adjacent window.
-Colors are arbitrary.}
+  \caption{Example of topics extracted from MySQL 3.23. This is the
+    kind of plot we eventually want to illustrate: named topics and
+    topic trends. The horizontal axis is time by month examined. The
+    vertical axis is used to stack topics that occur at the same
+    time. Longer topics are topics which recur in adjacent
+    window. Colors are arbitrary.}
   \label{fig:lda}
@@
 Commits are first extracted, then abstracted into word counts or word distributions.
-These word distributions are then given to topic clusterer like LDA which finds independent word distributions (topics) that these documents are related too and clusters or associates these documents with these topics.}
+These word distributions are then given a topic analysis tool like LDA.
+LDA finds independent word distributions (topics) that these documents are related to.}
   \label{fig:commits}
@@
 
-\subsection{Extract Repositories}
+\subsection{Extraction of  Repositories and Documents}
 
@@
 mirrored web accessible BitKeeper repositories.
-
-%   2. [ ] Extract Documents
-\subsection{Extract Documents}
-%   3. [ ] Windowed Sets
 
@@
 allows for many levels of granularity.  We usually used one month as
-the length of our time windows, while we could use different
-granularities for the study we think that a month is a sizable unit of
-development, which we consider to be fine enough to show meaningful
-shifts in focus, but coarse enough not to show too much detail.
+the length of our time windows. While we could use different window
+lengths for the study we think that a month is a sizable unit of
+development, which is large enough to show shifts in focus, but coarse
+enough not to show too much detail.
 
@@
 Once we have our data windowed, we apply our Topic Analysis tool to
-each window and extract the top $N$ topic clusters, we used $20$
+each window and extract the top $N$ topics. We used $20$
 topics in our case studies. Our Topic Analysis tool is an
@@
 %   6. [ ] Cluster Similarity [0/1]
-\subsection{Cluster Similarity}
+\subsection{Topic Similarity}
 %      1. [ ] need the tool
@@
 Once we have our topics extracted for each window, we analyze them and
-look for repeating topics. We find repeating topics by comparing the
-10 most common words of each topic.
-
-We then abstract these topics by comparing them to each other.  Each
-topic is compared to each other topic in the system, given a certain
-threshold of top 10 similarity, such as 8 out of 10 words. Then given
-this topic similarity matrix, we find the transitive closure of
-similar topics; that is, if we model topics are nodes and arcs as
-similarity, we fill flood along the similarity arcs until we have
-partitioned the topics into clusters of similar topics. Figure
-\ref{fig:closure} provides an example of clustering topics by the
+look for topics that recur across windows. Our input for topic
+similarity is the top 10 most common words of each topic.
+
+We then cluster these topics into trends by comparing them to each
+other.  Each topic is compared to each other topic in the system,
+given a certain threshold of top 10 similarity, such as 8 out of 10
+words. Then given this topic similarity matrix, we find the transitive
+closure of similar topics; that is, if we model topics are nodes and
+arcs as similarity, we fill flood along the similarity arcs until we
+have partitioned the topics into clusters of similar topics. Figure
+\ref{fig:closure} illustrates clustering of topics by the
 topic similarity. These clusters of topics are called trends. A trend
-is probably interesting and valuable if there is more than one topic
-in the trend.
+is probably interesting if it contains more than one topic.
 
@@
 
-We have devised several ways to visualize these topics and trends.
-For all of these plots if we find trends that have continuous
-segments, then we plot those segments as joined across time.  We have
-a compressed stacked trend view, shown in Figure \ref{fig:topicsmear}
-and Figure \ref{fig:zoomedsmear}, which displays trends as they occur
-on the time-based x-axis (placement along the y-axis is arbitrary).
-Another plot we have is the trend-time line plot, shown in Figure
-\ref{fig:trendtimeline}, each trend gets it own distinct y-value ,
-while the x-axis is time; these topics are then each plotted on their
-own line across time as they occur. Our final plot is the histogram
-plot, shown in Figure \ref{fig:histogram} where we plot each trend on
-its own line but stack up the segments of the trend, much like a
-histogram.
-
-The first visualization is a compressed view
-(Figure\ref{fig:topicsmear} and Figure \ref{fig:zoomedsmear}) where we
-try to sink the larger trends to the bottom.  Once the larger trends
-are stacked we fill in the gaps with the smaller trends in the same
+We have devised several techniques for visualizing these topics and
+trends.  For all of these techniques if we find trends that have
+continuous segments, then we plot those segments as joined
+horizontally across time. One technique is the \emph{compact-trend-view},
+shown in Figure \ref{fig:topicsmear} and Figure \ref{fig:zoomedsmear},
+which displays trends as they occur on the time-based x-axis
+(placement along the y-axis is arbitrary).  Another technique is
+the \emph{trend-time-line}, shown in Figure \ref{fig:trendtimeline},
+each trend gets it own distinct y-value , while the x-axis is time;
+these topics are then each plotted on their own line across time as
+they occur. Our final technique is the \emph{trend-histogram}, shown in Figure
+\ref{fig:histogram} where we plot each trend on its own line but stack
+up the segments of the trend, much like a histogram.
+
+The \emph{compact-trend-view} (Figure \ref{fig:topicsmear}) tries to
+sink the larger trends to the bottom.  Once the larger trends are
+stacked we fill in the gaps with the smaller trends in the same
 window, and then stack the smaller trends on top.  Although there is a
@@
 trend share the same color; apart from that, the color of trends is
-randomly chosen and color similarity is not meaningful.  The
-compressed view is convenient because it shows all of the information
-and the repeating continuous trends are easy to pick out, although
-repeating discontinuous trends are harder to spot.
-
-The second visualization is the trend-time line plot (Figure
-\ref{fig:trendtimeline}), it displays repeating trends more clearly by
+randomly chosen and color similarity is not meaningful, this is easier
+to see in the zoomed view of MaxDB 7.500's compact-trend-view shown in
+Figure \ref{fig:zoomedsmear}.  The compact-trend-view is convenient
+because it shows all of the information and the repeating continuous
+trends are easy to pick out, although repeating discontinuous trends
+are harder to spot.
+
+The \emph{trend-time-line} (Figure
+\ref{fig:trendtimeline}) displays repeating trends more clearly by
 dedicating a horizontal line for trend segments belonging to one
@@
 
-The third visualization is the histogram plot (Figure
-\ref{fig:histogram}) which superficially resembles the time-trend
-plot.  However, in this view the trends are plotted together by
-stacking to the left of their row, thus time information is lost.  The
-trends are ordered by the number of topics in the trend.  The
-histogram plot shows the count of instances of a trend and thus
-indicates which trends occur the most. Unfortunately due to the large
-number of topics given allotted space it is often best to crop off the
-trends with only one topic, otherwise the tail is long.
+The {trend-histogram} (Figure \ref{fig:histogram}) superficially
+resembles the trend-time-line.  However, in this view the trends are
+plotted together by stacking to the left of their row, thus time
+information is lost.  The trends are ordered by the number of topics
+in the trend.  The trend-histogram shows the count of instances of a
+trend and thus indicates which trends occur the most. Unfortunately
+due to the large number of topics given allotted space it is often
+best to crop off the trends with only one topic, otherwise the tail is
+long.
 
@@
   \includegraphics[width=1.0\textwidth]{fixed-time-smear-plot-scaled}
-  \caption{Compact trend plot shows topics per month for MaxDB
+  \caption{Compact-trend-view shows topics per month for MaxDB
     7.500. The x-axis is time in months, the Y axis is used to stack
@@
   \includegraphics[width=0.45\textwidth]{fixed-time-smear-plot-cropped}
-  \caption{A zoomed in slice of compact trend plot of topics per month of MaxDB 7.500. The topic text is visible in each topic box. Trends are plotted across time continuously.}
+  \caption{A zoomed in slice of compact-trend-view of topics per month of MaxDB 7.500. The topic text is visible in each topic box. Trends are plotted across time continuously.}
   \label{fig:zoomedsmear}
@@
   \includegraphics[width=0.45\textwidth]{class-smear-plot-crop-scaled}
-  \caption{Trend-time Line: Trends plotted per month of MaxDB 7.500.
-Time in months are plotted along the x-Axis, each row on the y-axis is associated with a trend ranked by size in descending order}
+  \caption{Trend-time-line: Trends plotted per month of MaxDB 7.500.
+Time in months are plotted along the x-axis, each row on the y-axis is associated with a trend ranked by size in descending order}
   \label{fig:trendtimeline}
@@
   \includegraphics[width=0.45\textwidth]{histogram-cropped-scaled}
-  \caption{The top part of a trend histogram of MaxDB 7.500, ordered
+  \caption{The top part of a trend-histogram of MaxDB 7.500, ordered
     by topic occurrence. X axis determines the number of continuous
@@
 \texttt{SYSDD.punix} and \texttt{MONITOR.punix}.  This trend is
-partially displayed at the top of the zoomed in compressed view plot
-(Figure \ref{fig:zoomedsmear}) and at the top of the histogram plot
+partially displayed at the top of the zoomed in compact-trend-view
+(Figure \ref{fig:zoomedsmear}) and at the top of the trend-histogram
 (Figure \ref{fig:histogram}).  Other tokens mentioned are
@@
 We noticed that Perforce and Changelist were in just about every
-topic, so we added them to the stop words list. Longer trends were
-shortened and in some cases the total number of topics found per month
-was less. Evaluating different similarity thresholds showed that by
-removing common words one reduces the general similarity of topics,
-that said the larger topics still existed. Thus if you add more
-relevant stop words one should tune your topic similarity parameters
-to handle these changes.
+topic, so we added them to the stop words list. As a result, longer
+trends were shortened and in some cases the total number of topics
+found per month was less.  Evaluating different similarity thresholds
+showed that by removing common words one reduces the general
+similarity of topics. That said the larger topics still existed. Thus
+if more relevant stop words are added one should tune your topic
+similarity parameters to handle these changes.
 
@@
 
-\subsection{Compare to topics over entire development}
+\subsection{Compare with topics over the entire development}
 
@@
 
-We attempted the same kind of analysis, as shown in
-Figure~\ref{fig:statictopics}. We extracted 20 topics and then plotted
+We carried out the same kind of analysis, as shown in
+Figure~\ref{fig:statictopics}. We extracted 20 topics  and plotted
 the number of messages per month that were related to that topic. We
-found was that one topic would often dominate the results, while
+found was that one topic would often dominate the results, as shown , while
 messages related to other topics did not appear very often.
@@
 topic analysis topics, but they were part of 566 commits (6\% of the
-entire system) to MaxDB 7.500. These tokens were part of 87 out of 500
-(25 months, 20 topics per month) of our locally generated topics.
+entire system) to MaxDB 7.500. These tokens were part of 87 out of 520
+(26 months, 20 topics per month) of our locally generated topics.
 
@@
   \centering
-  \includegraphics[width=0.45\textwidth]{maxdb7500-everything-by-month}
+  %\includegraphics[width=0.45\textwidth]{maxdb7500-everything-by-month}
+  \includegraphics[width=0.45\textwidth]{month}
   \caption{MaxDB 7.500 Topics Analyzed with Static Topic Analysis, 20
-    topics over the entire development history of MaxDB 7.500. The
-    color of the topic indicates the number of documents matching that
+    topics over the entire development history of MaxDB 7.500 (26 months). The
+    shade of the topic indicates the number of documents matching that
     topic in that month relative to the number of documents (white is
@@
 We presented multiple visualizations that focused on different aspects
-of the data: temporality of trends, trend size, and a compressed
-overall trend view. The compressed trend view shows much more
+of the data: temporality of trends, trend size, and a
+compact-trend-view. The compact-trend-view shows much more
 information than the views that global analysis could show, it
@@
 well shows topics by similarity so one can track trends across time.
-Our histogram trend-view immediately hi-lights and measures the size of
-trends while our time-line trend view shows how a topic reoccurs over
-time. These visualizations help us get an over feeling about the
+Our trend-histogram immediately hi-lights and measures the size
+of trends while our trend-time-line view shows how a topic reoccurs
+over time. These visualizations help us get an over feeling about the
 common and unique topics that developers focus on. If implemented

### 2009-04-10 lda-paper.tex
global topics

@@
 \title{What's hot and what's not:\\Windowing developer topic analysis}
+
+\newcommand{\shrinkit}{\vspace*{-.3em}}
 
@@
 
+\shrinkit
 \section{Introduction}
+\shrinkit
 
@@
 
+\shrinkit
 \section{Background}
+\shrinkit
 
@@
 
+\shrinkit
 \subsection{LDA, LSI and Semantic Clustering}
+\shrinkit
 
@@
 
+\shrinkit
 \section{Preliminary Case Study}
+\shrinkit
 
@@
 be viewed as stop words. There were notable transitional topics such
-as in the first window the word \emph{bitkeeper} appears because MySQL
+as in the first window the word \emph{BitKeeper} appears because MySQL
 adopted Bitkeeper for their source control system yet in the following
@@
 
+\shrinkit
 \section{Methodology}
+\shrinkit
 \begin{figure*}
@@
 
+\shrinkit
 \subsection{Extraction of  Repositories and Documents}
+\shrinkit
 
@@
 
+\shrinkit
 \subsection{Windowed Sets}
+\shrinkit
 
@@
 
+\shrinkit
 \subsection{Apply Topic Analysis to each Window}
+\shrinkit
 
@@
 %   6. [ ] Cluster Similarity [0/1]
+\shrinkit
 \subsection{Topic Similarity}
+\shrinkit
 %      1. [ ] need the tool
@@
 
+\shrinkit
 \subsection{Visualization}
+\shrinkit
 %       1. [ ] need the visualizer
@@
 
+\shrinkit
 \subsection{Datasets and Tools}
+\shrinkit
 
@@
 
-We extracted the repositories of: PosgreSQL, MaxDB and
+We extracted the repositories of: PostgreSQL, MaxDB and
 Firebird.
@@
 
+\shrinkit
 \section{Results}
+\shrinkit
 
@@
 %XXX DO MYSQL?
-
+\shrinkit
 \subsection{PostgreSQL}
+\shrinkit
 %   3.   mysql
@@
 
+\shrinkit
 \subsection{Firebird}
+\shrinkit
 %   5.   what's that other db?
@@
 
+\shrinkit
 \subsection{MaxDB 7.500}
+\shrinkit
 
@@
 systems, as there was a period where no real development occurred and
-thus there were no topics or trends whatsoever. Using a topic
-similarity of $7/10$ we evaluated of MaxDB 7.500, the first period
-from June 2004 to January 2005, and the second period from June 2005
-to June 2006.
+thus there were no topics or trends whatsoever (see the gap in Figure
+\ref{fig:topicsmear}). Using a topic similarity of $7/10$ we evaluated
+of MaxDB 7.500, the first period from June 2004 to January 2005, and
+the second period from June 2005 to June 2006.
 
@@
 (Figure \ref{fig:histogram}).  Other tokens mentioned are
-\emph{Sutcheck v1.09}, the prefix SUT stands for Storable Unit
-Type. Sutcheck would also automate check-ins using a Perforce SCS tool
+\emph{Sutcheck v1.09} (the prefix SUT stands for Storable Unit
+Type), a tool that would also automate check-ins using a Perforce SCS tool
 which was exporting check-ins to CVS.
@@
 
-The query optimizer was a topic which recurred during
-MaxDB's development. In our plots, topics which mention optimizer
-occur four times yet in the large static plot (Figure
-\ref{fig:statictopics}) it is not in any of the topics. A query
-optimizer is rather important to an DBMS, but as we have shown it just
-does not pop up as a topic on its own. We tried to remove words to see
-if we could get optimizer topic out, after removing stop words and
-then two of the most common words the global analysis finally had a
-topic with optimizer in its topic ten words. This shows that optimizer
-was important but it had been obscured, while it would have been
-noticed using a more locally relevant approach that we propose.
-
-We noticed that commits which mentioned \emph{PHP} occured two thirds
+The query optimizer was a topic which recurred during MaxDB's
+development. In our plots, topics which mention optimizer occur four
+times yet in the global-trend-view (Figure \ref{fig:statictopics}) it
+is not in any of the topics. A query optimizer is rather important to
+an DBMS, but as we have shown it just does not pop up as a topic on
+its own. We tried to remove words to see if we could get an topic
+optimizer topic. After removing stop words and then two of the most
+common words the global analysis finally had a topic with optimizer in
+its topic ten words. This shows that optimizer was important but it
+had been obscured, while it would have been noticed using a more
+locally relevant approach that we propose.
+
+We noticed that commits which mentioned \emph{PHP} occurred two thirds
 less frequently than commits which mentioned \emph{Perl}, but
@@
 
+\shrinkit
 \subsection{Compare with topics over the entire development}
+\shrinkit
 
@@
 means that a single, static set of topics was tracked across the whole
-development history of a project.
+development history of a project. \emph{Global topic analysis} is
+topic analysis applied to entire lifetime of a project.
 
 We carried out the same kind of analysis, as shown in
-Figure~\ref{fig:statictopics}. We extracted 20 topics  and plotted
-the number of messages per month that were related to that topic. We
-found was that one topic would often dominate the results, as shown , while
-messages related to other topics did not appear very often.
+Figure~\ref{fig:statictopics}. We extracted 20 topics and plotted the
+number of messages per month that were related to that topic. We found
+was that one topic would often dominate the results, as shown in the
+third row of Figure \ref{fig:statictopics}, while messages related to
+other topics did not appear very often.
 % This technique seems useful if the 20 extracted topics are relevant
@@
 
-If we approach the difference of static global topic analysis and
+If we approach the difference of global topic analysis and
 windowed topic analysis via common tokens we can see that common token
@@
 \emph{backup}, \emph{select}, \emph{make}, \emph{memory}, \emph{view},
-and finally \emph{debug}. None of these tokens were part of the static
+and finally \emph{debug}. None of these tokens were part of the global 
 topic analysis topics, but they were part of 566 commits (6\% of the
@@
   %\includegraphics[width=0.45\textwidth]{maxdb7500-everything-by-month}
-  \includegraphics[width=0.45\textwidth]{month}
-  \caption{MaxDB 7.500 Topics Analyzed with Static Topic Analysis, 20
-    topics over the entire development history of MaxDB 7.500 (26 months). The
-    shade of the topic indicates the number of documents matching that
-    topic in that month relative to the number of documents (white is
-    most, black is least).}
+  \includegraphics[width=0.46\textwidth]{month}
+  \caption{MaxDB 7.500 topics analyzed with global topic analysis, 20
+    topics (each row) and their document counts plotted over the
+    entire development history of MaxDB 7.500 (26 months). The shade
+    of the topic indicates the number of documents matching that topic
+    in that month relative to the number of documents (white is most,
+    black is least).}
   \label{fig:statictopics}
@@
 %8.   Validity Threats [0/5]
+\shrinkit
 \section{Validity Threats}
+\shrinkit
 
@@
 
+\shrinkit
 \section{Conclusions}
+\shrinkit
 %11.   Start file file:/home/abez/projects/lda-paper/lda-paper.tex
@@
 We proposed and demonstrated the application of windowed topic
-analysis, that is, topic analysis applied to commit messages during
-periods of development. This approach differed from previous work that
+analysis, that is, which is topic analysis applied to commit messages during
+periods of development. This approach differs from previous work that
 would apply topic analysis globally to a project. We showed that many
 topics that exist locally are relevant and interesting yet would often
-not be included in global topic analysis. We identify recurring topics
+not be detected via global topic analysis. We identify recurring topics
 with a topic similarity measure which looks for topics which recur and
-mutate throughout the development of the software project.
+mutate repeatedly throughout the development of the software project.
 
@@
 of three database systems with a particular focus on MaxDB 7.500. We
-showed that static topic analysis missed important topics like ODBC
+showed that global topic analysis missed important topics like ODBC
 changes.
@@
 
-We presented multiple visualizations that focused on different aspects
-of the data: temporality of trends, trend size, and a
-compact-trend-view. The compact-trend-view shows much more
+We presented several visualization techniques that focused on
+different aspects of the data: temporality of trends, trend size, and
+a compact-trend-view. The compact-trend-view shows much more
 information than the views that global analysis could show, it
@@
 well shows topics by similarity so one can track trends across time.
-Our trend-histogram immediately hi-lights and measures the size
-of trends while our trend-time-line view shows how a topic reoccurs
-over time. These visualizations help us get an over feeling about the
+Our trend-histogram immediately hi-lights and measures the size of
+trends while our trend-time-line view shows how a topic reoccurs over
+time. These visualizations help us get an over feeling about the
 common and unique topics that developers focus on. If implemented
@@
 
+\shrinkit
 \subsection{ Future Work}
-
-One avenue of future work we think is quite important is automatic
-cluster labelling. Given a word distribution we should be able to
-automatically select a word or term that best describes that distribution.
-Another avenue we wish to investigate the is value of applying
-static community-generated software taxonomies to the naming or
-grouping of clusters
+\shrinkit
+
+One avenue of future work is automatic topic labelling. Given a word
+distribution we should be able to automatically select a word or term
+that describes that distribution. Potential external sources of topic
+names include: software engineering taxonomies, ontologies and
+standards.
 

### 2009-04-11 lda-paper.tex
Getting closer to a 10 pager + migod edits

@@
 
-Topic analysis, with respect to Software Control Systems (SCSs) such
-as CVS, Subversion, Git etc., is potentially valuable as it has many
-uses for various stakeholders and uses.  Imagine a development team
-has agreed to implement some features at the start of a development
-iteration. By the end of the iteration only some of the agreed upon
-features are implemented. The developers insist they were focusing
- on the features agreed upon, but what actually happened? Topic
-analysis could help by teasing out the distinct topics or focusses of
-the developers. Perhaps the developers had to deal with other pressing issues.
-Developers might have grouped those orthogonal issues into the
-features they were working on, thus when asked about how their time
-was spent the orthogonal changes would not be covered. This is the
-kind of underlying story that we hope that windowed topic analysis can
-uncover. If windowed topic analysis can discover important but
-overlooked topics, it can help both developers and managers. Developers
-can see a summary of what worked on, and managers can development.
-
-Commits in SCSs are not necessarily consistent and they can have
+% XXX Come back here, clear up and cut 
+% Topic analysis, with respect to
+% Software Control Systems (SCSs) such as CVS, Subversion, Git etc.,
+% is potentially valuable as it has many uses for various
+% stakeholders.  Imagine a development team has agreed to implement
+% some features at the start of a development iteration.  By the end
+% of the iteration only some of the agreed upon features are
+% implemented.  The developers insist they were focusing on the
+% features agreed upon, but what actually happened?  Topic analysis
+% could help by teasing out the distinct topics or focusses of the
+% developers.  Perhaps the developers had to deal with other pressing
+% issues.  Developers might have grouped those orthogonal issues into
+% the features they were working on, thus when asked about how their
+% time was spent the orthogonal changes would not be covered.  This is
+% the kind of underlying story that we hope that windowed topic
+% analysis can uncover. If windowed topic analysis can discover
+% important but overlooked topics, it can help both developers and
+% managers.  Developers and Managers can see a summary of what worked
+% on and the topics of development.
+
+Imagine that a development team have agreed upon a set of features to
+implement for their next interation, but these features were not
+finished by the end of the iteration. The developers insist they were
+focussed on the features, yet their manager and is not sure what
+happened. In a scenario like this, topic analysis can help tease out
+the topics of development that the developers were focussed on. Topic
+analysis could highlight the orthogonal issues that developers were
+spending time on.  Topic analysis extracts groups of words which from
+documents (commit log messages) which can be used to relate
+documents. These extracted topics often correlate with actual
+development topics that the developers discussed. By applying topic
+analysis locally to an iteration, both the developers and managers
+might be able to see what were the topics of development.
+
+Commits in SCSs can have
 multiple purposes. Thus we hope that our research will help uncover
 these multiple purposes that manifest themselves as topics interleaved
-within the commits.  Others have demonstrated that by using 
+within the commits.  Others have demonstrated that 
 topic analysis techniques, such as Latent Semantic Indexing (LSI) and
@@
 
-One use of topic analysis is to help us partition a project's
+Topic analysis can aid in partitioning a project's
 time-line into periods of developer focus. By breaking apart an
-iteration into sets of topics and trends (topics which recur), we may
+iteration into sets of topics and trends (topics that recur), we may
 be able to recognize the underlying software development process from
 these commits. Alternatively, we can determine what particular
-maintenance task was being performed are at a given time.
-
-We propose to extend the use of common existing topic analysis techniques by
+maintenance task was being performed at a given time.
+
+We propose to extend existing topic analysis techniques by
 applying them to windows of documents over time. We hope by looking
@@
 top 10 most frequent words of their word distribution.  A \emph{trend}
-is a one or more similar topics which recur over time.  Trends are
+is a one or more similar topics that recur over time.  Trends are
 particularly important, as they indicate long-lived and recurring
@@
 Indexing
-(LSI)~\cite{1421013,1374321,10.1109/ICPC.2007.13,10.1109/ICPC.2006.17},
+%(LSI)~\cite{1421013,1374321,10.1109/ICPC.2007.13,10.1109/ICPC.2006.17},
+(LSI)~\cite{1374321,10.1109/ICPC.2007.13},
 probabilistic Latent Semantic Indexing (pLSI)~\cite{944937} and
@@
 document modelling, document clustering and collaborative
-filtering. LDA attempts to encode documents as mixture
-model, a combination, of topics.  LDA is used in software engineering
-literature~\cite{lukins2008,10.1109/MSR.2007.20,NIPS2007637,1321709}
-is to extract topics from documents such as methods, bug
+filtering. LDA attempts to encode documents as a mixture
+model, a combination of topics.  LDA is used in software engineering
+literature~\cite{lukins2008,10.1109/MSR.2007.20}%,NIPS2007637,1321709}
+ to extract topics from documents such as methods, bug
 reports, source code and files.
@@
 relationship is often obvious to someone familiar with the corpus. We
-could swap LDA for LSI or semantic clustering and produce similar
+could swap LDA for LSI or semantic clustering and likely produce similar
 results. Our data is posed as documents with word distributions (word
@@
 extracted from SCS, Linstead et al. proposed an author-source-code model
-using LDA~\cite{10.1109/MSR.2007.20,NIPS2007637,1321709}. This model
+using LDA~\cite{10.1109/MSR.2007.20}%,NIPS2007637,1321709}. 
+This model
 is essentially an extension of the author-topic (AT)
@@
 artifacts for formal concept analysis and concept
-location~\cite{1421013,1374321,10.1109/ICPC.2007.13,10.1109/ICPC.2006.17}.
+%location~\cite{1421013,1374321,10.1109/ICPC.2007.13,10.1109/ICPC.2006.17}.
+location~\cite{1374321,10.1109/ICPC.2007.13}
 Concept analysis aims to automatically extract concepts from source
 code or documents that are related to the source code.  Concept location concerns how
-high level concepts relate to low level entities such as source code. For
+high-level concepts relate to low level entities such as source code. For
 example, when fixing a bug how and where do the concepts in the bug
@@
 Our technique differs from the previous techniques because we apply
-LDA locally to month-long windows of commit comments, where-as other
+LDA locally to month-long windows of commit comments, whereas other
 approaches apply LDA once to the entire project. This allows us to
@@
 
-\begin{figure*}
+\begin{figure*}[t]
   \centering
@@
 \shrinkit
-\begin{figure*}
+\begin{figure*}[t]
   \centering
@@
   \label{fig:commits}
+\end{figure*}
+
+\begin{figure*}
+  \centering
+  \includegraphics[width=1.0\textwidth]{fixed-time-smear-plot-scaled}
+  \caption{Compact-trend-view shows topics per month for MaxDB
+    7.500. The x-axis is time in months, the Y axis is used to stack
+    topics occurring at the same time. Trends that are continuous are
+    plotted as continuous blocks. The top 10 words in the topics are
+    joined and embedded in box representing the topics.}
+  \label{fig:topicsmear}
 \end{figure*}
@@
 there are lot of small trends (90\% of all trends contain 1 topic)
-which fill in these gaps quickly.  Different instances of the same
+that fill in these gaps quickly.  Different instances of the same
 trend share the same color; apart from that, the color of trends is
-randomly chosen and color similarity is not meaningful, this is easier
-to see in the zoomed view of MaxDB 7.500's compact-trend-view shown in
-Figure \ref{fig:zoomedsmear}.  The compact-trend-view is convenient
-because it shows all of the information and the repeating continuous
-trends are easy to pick out, although repeating discontinuous trends
-are harder to spot.
+randomly chosen and color similarity is not meaningful.  The
+compact-trend-view is convenient because it shows all of the topic
+information, as shown by the zoomed-in view of MaxDB 7.500's
+compact-trend-view in Figure \ref{fig:zoomedsmear}. This view makes
+repeating continuous trends easy to pick out, although discontinuous
+trends are harder to spot.
 
@@
 
-The {trend-histogram} (Figure \ref{fig:histogram}) superficially
+The \emph{trend-histogram} (Figure \ref{fig:histogram}) superficially
 resembles the trend-time-line.  However, in this view the trends are
@@
 long.
-
-%XXX Describe something useful like a trend
-\begin{figure*}
-  \centering
-  \includegraphics[width=1.0\textwidth]{fixed-time-smear-plot-scaled}
-  \caption{Compact-trend-view shows topics per month for MaxDB
-    7.500. The x-axis is time in months, the Y axis is used to stack
-    topics occurring at the same time. Trends that are continuous are
-    plotted as continuous blocks. The top 10 words in the topics are
-    joined and embedded in box representing the topics.}
-  \label{fig:topicsmear}
-\end{figure*}
 
@@
 
-\shrinkit
-\subsection{Datasets and Tools}
-\shrinkit
-
-To extract data from CVS and BitKeeper we used softChange to extract
-CVS repositories and provide a schema for our revision data. We used
-bt2csv to extract BitKeeper revisions from BitKeeper web repositories.
-
-We extracted the repositories of: PostgreSQL, MaxDB and
-Firebird.
+%\shrinkit
+%\subsection{Datasets and Tools}
+%\shrinkit
+
+%   1. [ ] LDA
+
+%   2. [ ] Extractor
+
+%   3.   Similarity
+
+%   4.   Smearing
+
+%   5.   Visualizer
+
+%LDA Skeleton
+%1. [ ] Abstract 
+%2. [ ] Introduction
+
+%3. [ ] Previous Work [0/2]
+%   2. [ ] LDA Work [0/6]
+
+\shrinkit
+\section{Results}
+\shrinkit
+
+We applied our methodology to multiple database systems that we
+extracted.  To analyze these extracted repositories we used:
+Hiraldo-Grok, an OCaml-based variant of the Grok query language;
+Gnuplot, a graph plotting package; lda-c, a LDA package implemented by
+Blei et al.~\cite{944937}; and our topic plotter, implemented in
+Haskell.
+
+We applied our tools and visualizations to the repositories of several
+open source database systems: PostgreSQL, MaxDB and Firebird.
 
@@
 
-Our analysis tools consisted of: Hiraldo-Grok, an OCaml-based variant
-of the Grok query language; Gnuplot, a graph plotting package; lda-c,
-a LDA package implemented by Blei et al. \cite{944937}; and our topic
-plotter, implemented in Haskell.
-
-%   1. [ ] LDA
-
-%   2. [ ] Extractor
-
-%   3.   Similarity
-
-%   4.   Smearing
-
-%   5.   Visualizer
-
-%LDA Skeleton
-%1. [ ] Abstract 
-%2. [ ] Introduction
-
-%3. [ ] Previous Work [0/2]
-%   2. [ ] LDA Work [0/6]
-
-\shrinkit
-\section{Results}
-\shrinkit
-
-We applied our tools and visualizations to the repositories of several
-open source database systems: PostgreSQL, MaxDB and Firebird.
-
 %   1. Interesting smears 
@@
 When we examined PostgreSQL, we did not find many trends with two or
-more topics with a similarity of $7/10$. These trends were not very
-large, lasting on 3 months at most. The first, second largest trends
-were dedicated to the work of two external developers: Dal Zotto, Dan
-McGuirk. The fifth largest trend related to work by PostgreSQL developer D'Arcy. Other
-topics of the larger trends were changes to the TODO list in HTML, and
-time string formating topics relating to timezones.
+more topics with a similarity of $7/10$. Those trends that we did find
+were not very large, lasting on 3 months at most. The first and second
+largest trends were dedicated to the work of two external developers:
+Dal Zotto, Dan McGuirk. The fifth largest trend related to work by
+PostgreSQL developer D'Arcy. Other topics of the larger trends were
+changes to the TODO list in HTML, and time string formating topics
+relating to timezones.
 
@@
 stretched across development, contained these same words
-(\emph{patch}, \emph{fix}, \emph{update}). The second most common
-topic mentions Dal Zotto, while the third most common topic mentions
-the \texttt{[PATCHES]} mailinglist and the names of some patch
+(\emph{patch}, \emph{fix}, \emph{update}). The second largest trend
+mentions Dal Zotto, while the third largest trend mentions the
+\texttt{[PATCHES]} mailinglist and the names of some patch
 contributors.  Other repeating topics refer to portability with Win32,
 Central Europe Time (CEST) from email headers, issues with ALTER TABLE
-and CVS branch merging (because CVS doesn't record merges explicitly).
+and CVS branch merging (because CVS does not record merges explicitly).
 
@@
 previous month. We investigated the actual revisions and we did find
-there was difference in behavior between and that the last month of
+there was difference in behavior between, and that the last month of
 development commits were mostly referencing the bug tracker. If the
@@
 
-The query optimizer was a topic which recurred during MaxDB's
-development. In our plots, topics which mention optimizer occur four
-times yet in the global-trend-view (Figure \ref{fig:statictopics}) it
+The query optimizer was a topic that recurred during MaxDB's
+development. In our plots, topics that mention \emph{optimizer} occur four
+times, yet in the global-trend-view (Figure \ref{fig:statictopics}) it
 is not in any of the topics. A query optimizer is rather important to
 an DBMS, but as we have shown it just does not pop up as a topic on
-its own. We tried to remove words to see if we could get an topic
-optimizer topic. After removing stop words and then two of the most
-common words the global analysis finally had a topic with optimizer in
+its own. We tried to remove words to see if we could get an 
+\emph{optimizer} topic. After removing stop words and then two of the most
+common words, the global analysis finally had a topic with optimizer in
 its topic ten words. This shows that optimizer was important but it
@@
 
-We noticed that commits which mentioned \emph{PHP} occurred two thirds
-less frequently than commits which mentioned \emph{Perl}, but
-\emph{Perl} related topics appeared in the global static topics for
-MaxDB while \emph{PHP} related topics did not.  Our local topic
+We noticed that commits that mentioned \emph{PHP} occurred two thirds
+less frequently than commits that mentioned \emph{Perl}, but
+\emph{Perl}-related topics appeared in the global static topics for
+MaxDB while \emph{PHP}-related topics did not.  Our local topic
 analysis mentioned \emph{PHP} in 5 different topics, yet only
-mentioned \emph{Perl} in 4 different topics. Perhaps the Perl mentions
-were not as independent topics as the PHP topics were.
+mentioned \emph{Perl} in 4 different windoed topics and 1 global
+topic. Perhaps this is because there was a cluster of \emph{Perl}
+mentions during one month while \emph{PHP} was mentions were more
+spread out.
 
@@
 showed that by removing common words one reduces the general
-similarity of topics. That said the larger topics still existed. Thus
-if more relevant stop words are added one should tune your topic
-similarity parameters to handle these changes.
+similarity of topics. That said, the larger topics still clearly
+apparent. Thus if more relevant stop words are added one should tune
+the topic similarity parameters to handle these changes.
 
@@
 Previous work on topic analysis that employed LSI and LDA typically
-extracted a certain number of topics and tracked them over time. This
+extracted a specified number of topics and then tracked them over time. This
 means that a single, static set of topics was tracked across the whole
@@
 
-We carried out the same kind of analysis, as shown in
-Figure~\ref{fig:statictopics}. We extracted 20 topics and plotted the
+We carried out the same kind of analysis, in order to compare the
+results against our time-windowed approach.  To produce
+figure~\ref{fig:statictopics}, we extracted 20 topics and plotted the
 number of messages per month that were related to that topic. We found
@@
 
-If we approach the difference of global topic analysis and
-windowed topic analysis via common tokens we can see that common token
-tend to dominate both results, at least in the case of MaxDB 7.500 our
-local topic approach netted us topics which contained these relatively
+If we approach the difference of global topic analysis and windowed
+topic analysis via common tokens we can see that common token tend to
+dominate both results, at least in the case of MaxDB 7.500 our local
+topic approach netted us topics which contained these relatively
 important and common words: \emph{UNICODE}, \emph{DEC/OSF1},
@@
 \emph{backup}, \emph{select}, \emph{make}, \emph{memory}, \emph{view},
-and finally \emph{debug}. None of these tokens were part of the global 
-topic analysis topics, but they were part of 566 commits (6\% of the
-entire system) to MaxDB 7.500. These tokens were part of 87 out of 520
-(26 months, 20 topics per month) of our locally generated topics.
+and finally \emph{debug}. \emph{ODBC} is quite important because it
+often determines how databases communicate with software. None of
+these tokens were part of the global topic analysis topics, but they
+were part of 566 commits (6\% of the entire system) to MaxDB
+7.500. These tokens were part of 87 out of 520 (26 months, 20 topics
+per month) of our locally generated topics.
 
@@
 
-Even with our liberal topic similarity metrics that produced both
-long and short trends, we showed that there are only a few trends in a
-repository which recur. Since so few trends recur and so many
-trends appear only once this suggests that global topic analysis might
-be ignoring locally unique topics.
+Even with our liberal topic similarity metrics that produced both long
+and short trends, we showed that there are only a few trends in a
+repository that recur. Since so few trends recur and so many trends
+appear only once this suggests that global topic analysis might be
+ignoring locally unique topics.
 
@@
 information decreases as it becomes older. Perhaps older trends will
-dominate the topic analysis. Windowed localized topic analysis show
-what are the unique topics.
-
-\begin{figure}
+dominate the topic analysis. Windowed localized topic analysis shows
+what are the unique topics, yet seems to give a more nuanced view of
+the timeliness of the important topics.
+
+\begin{figure}[t]
   \centering
   %\includegraphics[width=0.45\textwidth]{maxdb7500-everything-by-month}
-  \includegraphics[width=0.46\textwidth]{month}
+  \includegraphics[width=0.45\textwidth]{month}
   \caption{MaxDB 7.500 topics analyzed with global topic analysis, 20
@@
 they provide. If the language of check-in comments was automated we
-would only be analyzing that.
-
-We truncated the topics to top 10 tokens, this summary of a topic
+would  be analyzing only that.
+
+We truncated the topics to top 10 tokens, so this summary of a topic
 might not have been as useful as determining the actual topic distance
@@
 
-Each project seemed to have their own kind of stop words, we did not
+Each project seemed to have their own set of stop words, and we did not
 remove them, but perhaps dropping some of these stop words would aid
@@
 We proposed and demonstrated the application of windowed topic
-analysis, that is, which is topic analysis applied to commit messages during
+analysis, that is, topic analysis applied to commit messages during
 periods of development. This approach differs from previous work that
-would apply topic analysis globally to a project. We showed that many
-topics that exist locally are relevant and interesting yet would often
-not be detected via global topic analysis. We identify recurring topics
-with a topic similarity measure which looks for topics which recur and
-mutate repeatedly throughout the development of the software project.
+would applied topic analysis globally to the entire history of a
+project without regard to time.. We showed that many topics that exist
+locally are relevant and interesting yet would often not be detected
+via global topic analysis. We identify recurring topics with a topic
+similarity measure which looks for topics which recur and mutate
+repeatedly throughout the development of the software project.
 
@@
 of three database systems with a particular focus on MaxDB 7.500. We
-showed that global topic analysis missed important topics like ODBC
-changes.
+showed that global topic analysis missed important topics like such as
+ODBC changes.
 
@@
 
+\shrinkit
 \bibliographystyle{latex8}

### 2009-04-11 lda-paper.tex
minor change
Figures in Order..
spell checked

@@
 Imagine that a development team have agreed upon a set of features to
-implement for their next interation, but these features were not
+implement for their next iteration, but these features were not
 finished by the end of the iteration. The developers insist they were
@@
 
+Our methodology is to first extract the commits from SCS repositories
+such as CVS or Bitkeeper. Then we extract the revisions and commits,
+grouping the revisions into commits if necessary. We extract the
+commit log message of each commit, count the occurrence of each word
+in the message, remove stop words, and then produce word distributions
+for each message.
+
+\shrinkit
+\subsection{Extraction of  Repositories and Documents}
+\shrinkit
+
+We mirrored the repositories and their revisions using software such
+as rsync, CVSSuck, \texttt{softChange} and bt2csv.
+\texttt{softChange} provided a general schema for storing
+revisions. CVSSuck and rsync mirrored CVS repositories while bt2csv
+mirrored web accessible BitKeeper repositories.
+
+The documents we are analyzing are the commit log comments i.e., the
+comments that are added when revisions to the project are committed.
+Documents are converted into word count distributions. These
+distributions are then normalized by the size of the message. After
+all messages are processed the distributions are extended to include
+all words within all of the  distributions.
+
+\shrinkit
+\subsection{Windowed Sets}
+\shrinkit
+
+Given all messages, we group the messages into windows. We could use
+overlapping windows, but in this paper we use non-overlapping windows
+of a set time unit because it simplifies analysis.  Windowing by time
+allows for many levels of granularity.  We usually used one month as
+the length of our time windows. While we could use different window
+lengths for the study we think that a month is a sizable unit of
+development, which is large enough to show shifts in focus, but coarse
+enough not to show too much detail.
+
+\shrinkit
+\subsection{Apply Topic Analysis to each Window}
+\shrinkit
+
+Once we have our data windowed, we apply our Topic Analysis tool to
+each window and extract the top $N$ topics. We used $20$
+topics in our case studies. Our Topic Analysis tool is an
+implementation of LDA, but we could have used other similar tools like
+LSI, but previous work showed more promising topic analysis results
+with LDA.  We note that this step is a slow one, as executing even one
+instance of LDA involves significant computation, and we perform LDA
+once per window.
+
+%\subsection{Extract Clusters}
+
+%   6. [ ] Cluster Similarity [0/1]
+\shrinkit
+\subsection{Topic Similarity}
+\shrinkit
+%      1. [ ] need the tool
+%             grab from the cluster analysis
+%   7. [ ] Continuous Cluster Analysis
+
+Once we have our topics extracted for each window, we analyze them and
+look for topics that recur across windows. Our input for topic
+similarity is the top 10 most common words of each topic.
+
+We then cluster these topics into trends by comparing them to each
+other.  Each topic is compared to each other topic in the system,
+given a certain threshold of top 10 similarity, such as 8 out of 10
+words. Then given this topic similarity matrix, we find the transitive
+closure of similar topics; that is, if we model topics are nodes and
+arcs as similarity, we fill flood along the similarity arcs until we
+have partitioned the topics into clusters of similar topics. Figure
+\ref{fig:closure} illustrates clustering of topics by the
+topic similarity. These clusters of topics are called trends. A trend
+is probably interesting if it contains more than one topic.
+
+This technique has weaknesses in that nodes that are a few neighbors
+away in similarity might not share any similar words.  We do this
+because we want to be able to color or highlight topics that are
+related to each other and track them throughout time.
+
+Once we have determined our similarity clusters we can choose to
+analyze and to plot the topics.
+
+\begin{comment}
+Topic Clustering
+ · Need to track continuous topics across
+ · Similarity between topics
+ · Clusters of the transitive closure of topics with X%
+   similarity
+    ­ Fill flood along similarity, make subsets of
+      everyone who X% similar to any of their neighbors,
+      make that a cluster
+\end{comment}
+
+\begin{figure}[t]
+  \centering
+  \includegraphics[width=0.45\textwidth]{transitiveclosure}
+  \caption{Topic similarity demonstrated by clustering topics by the transitive closure (connectedness)
+    of topic similarity. Nodes are topics and
+    arcs imply similarity. Similarity could be a measure such as
+    topics share 8 out of 10 top ten words.}
+\label{fig:closure}
+\end{figure}
+
 \begin{figure*}
@@
 \end{figure*}
-
-Our methodology is to first extract the commits from SCS repositories
-such as CVS or Bitkeeper. Then we extract the revisions and commits,
-grouping the revisions into commits if necessary. We extract the
-commit log message of each commit, count the occurrence of each word
-in the message, remove stop words, and then produce word distributions
-for each message.
-
-\shrinkit
-\subsection{Extraction of  Repositories and Documents}
-\shrinkit
-
-We mirrored the repositories and their revisions using software such
-as rsync, CVSSuck, \texttt{softChange} and bt2csv.
-\texttt{softChange} provided a general schema for storing
-revisions. CVSSuck and rsync mirrored CVS repositories while bt2csv
-mirrored web accessible BitKeeper repositories.
-
-The documents we are analyzing are the commit log comments i.e., the
-comments that are added when revisions to the project are committed.
-Documents are converted into word count distributions. These
-distributions are then normalized by the size of the message. After
-all messages are processed the distributions are extended to include
-all words within all of the  distributions.
-
-\shrinkit
-\subsection{Windowed Sets}
-\shrinkit
-
-Given all messages, we group the messages into windows. We could use
-overlapping windows, but in this paper we use non-overlapping windows
-of a set time unit because it simplifies analysis.  Windowing by time
-allows for many levels of granularity.  We usually used one month as
-the length of our time windows. While we could use different window
-lengths for the study we think that a month is a sizable unit of
-development, which is large enough to show shifts in focus, but coarse
-enough not to show too much detail.
-
-\shrinkit
-\subsection{Apply Topic Analysis to each Window}
-\shrinkit
-
-Once we have our data windowed, we apply our Topic Analysis tool to
-each window and extract the top $N$ topics. We used $20$
-topics in our case studies. Our Topic Analysis tool is an
-implementation of LDA, but we could have used other similar tools like
-LSI, but previous work showed more promising topic analysis results
-with LDA.  We note that this step is a slow one, as executing even one
-instance of LDA involves significant computation, and we perform LDA
-once per window.
-
-%\subsection{Extract Clusters}
-
-%   6. [ ] Cluster Similarity [0/1]
-\shrinkit
-\subsection{Topic Similarity}
-\shrinkit
-%      1. [ ] need the tool
-%             grab from the cluster analysis
-%   7. [ ] Continuous Cluster Analysis
-
-Once we have our topics extracted for each window, we analyze them and
-look for topics that recur across windows. Our input for topic
-similarity is the top 10 most common words of each topic.
-
-We then cluster these topics into trends by comparing them to each
-other.  Each topic is compared to each other topic in the system,
-given a certain threshold of top 10 similarity, such as 8 out of 10
-words. Then given this topic similarity matrix, we find the transitive
-closure of similar topics; that is, if we model topics are nodes and
-arcs as similarity, we fill flood along the similarity arcs until we
-have partitioned the topics into clusters of similar topics. Figure
-\ref{fig:closure} illustrates clustering of topics by the
-topic similarity. These clusters of topics are called trends. A trend
-is probably interesting if it contains more than one topic.
-
-This technique has weaknesses in that nodes that are a few neighbors
-away in similarity might not share any similar words.  We do this
-because we want to be able to color or highlight topics that are
-related to each other and track them throughout time.
-
-Once we have determined our similarity clusters we can choose to
-analyze and to plot the topics.
-
-\begin{comment}
-Topic Clustering
- · Need to track continuous topics across
- · Similarity between topics
- · Clusters of the transitive closure of topics with X%
-   similarity
-    ­ Fill flood along similarity, make subsets of
-      everyone who X% similar to any of their neighbors,
-      make that a cluster
-\end{comment}
-
-\begin{figure}
-  \centering
-  \includegraphics[width=0.45\textwidth]{transitiveclosure}
-  \caption{Topic similarity demonstrated by clustering topics by the transitive closure (connectedness)
-    of topic similarity. Nodes are topics and
-    arcs imply similarity. Similarity could be a measure such as
-    topics share 8 out of 10 top ten words.}
-\label{fig:closure}
-\end{figure}
 
@@
 
-\begin{figure}
+\begin{figure}[t]
   \centering
@@
 similarity of $7/10$ that Firebird had far more continuous and
-recurring trends than PosgreSQL.  The first large trend was segmented
+recurring trends than PostgreSQL.  The first large trend was segmented
 across time but explicitly references one author \texttt{carlosga05}
@@
 analysis mentioned \emph{PHP} in 5 different topics, yet only
-mentioned \emph{Perl} in 4 different windoed topics and 1 global
+mentioned \emph{Perl} in 4 different windowed topics and 1 global
 topic. Perhaps this is because there was a cluster of \emph{Perl}
@@
 
-We noticed that Perforce and Changelist were in just about every
-topic, so we added them to the stop words list. As a result, longer
-trends were shortened and in some cases the total number of topics
-found per month was less.  Evaluating different similarity thresholds
-showed that by removing common words one reduces the general
-similarity of topics. That said, the larger topics still clearly
-apparent. Thus if more relevant stop words are added one should tune
-the topic similarity parameters to handle these changes.
+We noticed that \emph{Perforce} and \emph{Changelist} were in just
+about every topic, so we added them to the stop words list. As a
+result, longer trends were shortened and in some cases the total
+number of topics found per month was less.  Evaluating different
+similarity thresholds showed that by removing common words one reduces
+the general similarity of topics. That said, the larger topics still
+clearly apparent. Thus if more relevant stop words are added one
+should tune the topic similarity parameters to handle these changes.
 

### 2009-04-11 lda-paper.tex
editted the intro

@@
   Latent Semantic Indexing (LSI) can be used to analyze check-in
-  comments over the entire lifetime of a project, and previous work on
+  comments over the entire lifetime of a project. Previous work on
   developer topic analysis has leveraged these tools to associate
@@
 
-What happened in the previous maintenance or development iteration? What were the developers
-working on? What was the developer in the next cubicle working on?
-Which common topics and issues were dealt with in the current
-release of the software? What were the topics of the previous release?
-Given the requirements agreed to at the start of the iteration did our
-developers work on them? What else did they work on?
+What happened in the previous maintenance or development iteration?
+What were the developers working on? What was the developer in the
+next cubicle working on?  Which common topics and issues were dealt
+with in the current release of the software? What were the topics of
+the previous release?  Given the requirements agreed to at the start
+of the iteration did our developers work on them? What else did they
+work on?
 
@@
 
-Imagine that a development team have agreed upon a set of features to
-implement for their next iteration, but these features were not
-finished by the end of the iteration. The developers insist they were
-focussed on the features, yet their manager and is not sure what
+Imagine that a development team had agreed upon a set of features to
+implement for a development iteration, but these features were not
+finished by the end of the iteration. The developers insist they
+focussed on those features, yet their manager is not sure what
 happened. In a scenario like this, topic analysis can help tease out
@@
 analysis could highlight the orthogonal issues that developers were
-spending time on.  Topic analysis extracts groups of words which from
-documents (commit log messages) which can be used to relate
-documents. These extracted topics often correlate with actual
-development topics that the developers discussed. By applying topic
-analysis locally to an iteration, both the developers and managers
-might be able to see what were the topics of development.
+spending time on.  Topic analysis extracts groups of words from
+documents (commit log messages) that are word distributions, topics,
+which characterize clusters of documents. These extracted topics often
+correlate with actual development topics that the developers
+discussed. By applying topic analysis locally to an iteration, both
+the developers and managers might be able to see what were the topics
+of development during that iteration. %XXX RM during that iteration
 
@@
 are interleaved throughout the development. A topic represents a
-groups of commit comments that are related to each other by their
+groups of commit log comments that are related to each other by their
 content.  In this case a topic is a set of tokens.
@@
 We propose to extend existing topic analysis techniques by
-applying them to windows of documents over time. We hope by looking
+applying them to windows of documents over time. We hope that by
+looking %XX X ADDED
 at topic clusters during periods or windows over time we can identify
@@
 In this paper we explore how topics shift over time in the source
-control repository of a project, using several open source database
-systems as examples. We analyze commit comments per revision in each
-time window and we extract the topics in that time window. We expect
-that topics will change over time and the similarities
-and differences in these topics will indicate developer
-focus and changes in developer focus as it changes over time.
+control system (SCS) of a project, using several open source database
+systems as examples. We analyze commit log comments in each
+time window and we extract the topics of that time window. We expect
+that topics will change over time and the similarities and differences
+in these topics will indicate developer focus and changes in developer
+focus as it changes over time.
 
@@
 \begin{itemize}
-\item Demonstrating the value of windowed topic analysis using trends
-\item Multiple visualizations topics and trends over time
+\item Demonstrating the value of windowed topic analysis using trends.
+\item Multiple visualizations of topics and trends over time.
 \item An exploratory case study using these techniques on several open source database systems.

### 2009-04-11 lda-paper.tex
Edits and figure clear up

@@
 spending time on.  Topic analysis extracts groups of words from
-documents (commit log messages) that are word distributions, topics,
+documents (commit log comments) that are word distributions, topics,
 which characterize clusters of documents. These extracted topics often
@@
 \emph{message} is a block of text written by developers. In this
-paper, messages will be the CVS commit comments made when the user
+paper, messages will be the CVS commit log comments made when the user
 commits changes to files in a repository. A \emph{word distribution}
@@
 in a message, a word distribution is effectively a word count divided
-by the message size. A \emph{topic} is a word distribution. It is
-essentially a set of words that form a word distribution that is
-unique and independent within the set of documents in our total
-corpus. One could think of a topic as a distribution of the centroid
-of a group of messages. In this paper we often summarize topics by the
-top 10 most frequent words of their word distribution.  A \emph{trend}
-is a one or more similar topics that recur over time.  Trends are
-particularly important, as they indicate long-lived and recurring
-topics that may provide key insights into the development of the
-project.
+by the message size. A \emph{topic} is a word distribution, a
+set of words that form a word distribution that is unique and
+independent within the set of documents in our total corpus. One could
+think of a topic as a distribution of the centroid of a group of
+messages. In this paper we often summarize topics by the top 10 most
+frequent words of their word distribution.  A \emph{trend} is a one or
+more similar topics that recur over time.  Trends are particularly
+important, as they indicate long-lived and recurring topics that may
+provide key insights into the development of the project.
 
@@
 sample documents. These queries would indicate the proportion of
-topics that were associated with the query and thus possible the
+topics that were associated with the query and the
 related documents.
@@
 high-level concepts relate to low level entities such as source code. For
-example, when fixing a bug how and where do the concepts in the bug
+example, when fixing a bug, how and where do the concepts in the bug
 report map back to the source code?  Semantic Clustering has also been
@@
 Our technique differs from the previous techniques because we apply
-LDA locally to month-long windows of commit comments, whereas other
+LDA locally to month-long windows of commit log comments, whereas other
 approaches apply LDA once to the entire project. This allows us to
@@
 windows. We asked LDA to make 20 topics per window and we then
- examined the top 10 most frequent words in that topic.  We
-found there were common words across topic clusters, such as
-\emph{diffs}, \emph{annotate} and \emph{history}, these words probably should
-be viewed as stop words. There were notable transitional topics such
-as in the first window the word \emph{BitKeeper} appears because MySQL
+examined the top 10 most frequent words in that topic.  We found there
+were common words across topic clusters, such as \emph{diffs},
+\emph{annotate} and \emph{history}; these words probably should be
+viewed as stop words. There were notable transitional topics such as
+in the first window the word \emph{BitKeeper} appears because MySQL
 adopted Bitkeeper for their source control system yet in the following
-windows there were no mentions of Bitkeeper. \emph{RENAME} also only
+windows there were no mentions of BitKeeper. \emph{RENAME} also only
 appeared once in the first window and never again. Per each topic we
@@
 track the evolution of topics by visualizing the topics and joining
-similar topics into trends.  Figure \ref{fig:lda} displays a sample
-plot of the extracted topics in Table \ref{tab:portability}.
+similar topics into trends.  Figure \ref{fig:lda} displays a manually
+created sample plot of the extracted topics in Table
+\ref{tab:portability}.
 
@@
 \end{tabular}
-\caption{Sampled topics from MySQL 3.23, some with continuous topics. These tokens were pulled from LDA topic. Each token is a summary of one LDA generated topic from MySQL 3.23 commit logs.}
+\caption{Sampled topics from MySQL 3.23, some with continuous topics. These tokens were pulled from LDA topic. Each token is a summary of one LDA generated topic from MySQL 3.23 commit comments.}
 \label{tab:portability}
@@
     time. Longer topics are topics which recur in adjacent
-    window. Colors are arbitrary.}
+    windows. Colors are arbitrary.}
   \label{fig:lda}
@@
 grouping the revisions into commits if necessary. We extract the
-commit log message of each commit, count the occurrence of each word
+commit log comment of each commit, count the occurrence of each word
 in the message, remove stop words, and then produce word distributions
@@
 all messages are processed the distributions are extended to include
-all words within all of the  distributions.
+all words from all of the  distributions.
 
@@
 have partitioned the topics into clusters of similar topics. Figure
-\ref{fig:closure} illustrates clustering of topics by the
+\ref{fig:closure} illustrates clustering of topics by 
 topic similarity. These clusters of topics are called trends. A trend
@@
 This technique has weaknesses in that nodes that are a few neighbors
-away in similarity might not share any similar words.  We do this
+away in similarity might not share any similar words.  We use this measure
 because we want to be able to color or highlight topics that are
@@
   \centering
-  \includegraphics[width=1.0\textwidth]{fixed-time-smear-plot-scaled}
+%  \includegraphics[width=1.0\textwidth]{fixed-time-smear-plot-scaled}
+  \includegraphics[width=1.0\textwidth]{time-smear-plot}%fixed-time-smear-plot-scaled}
   \caption{Compact-trend-view shows topics per month for MaxDB

### 2009-04-11 lda-paper.tex
submitted once

@@
 %\title{Windowed Developer Topic Analysis}
-\title{What's hot and what's not:\\Windowing developer topic analysis}
+\title{What's hot and what's not:\\Windowed developer topic analysis}
 
@@
 the \emph{trend-time-line}, shown in Figure \ref{fig:trendtimeline},
-each trend gets it own distinct y-value , while the x-axis is time;
+each trend gets it own distinct y-value, while the x-axis is time;
 these topics are then each plotted on their own line across time as
@@
 trend and thus indicates which trends occur the most. Unfortunately
-due to the large number of topics given allotted space it is often
+due to the large number of topics, given the allotted space, it is often
 best to crop off the trends with only one topic, otherwise the tail is
@@
   \caption{The top part of a trend-histogram of MaxDB 7.500, ordered
-    by topic occurrence. X axis determines the number of continuous
+    by topic occurrence. X-axis determines the number of continuous
     months of a trend. Trends are ranked by the number of topics that
@@
 Gnuplot, a graph plotting package; lda-c, a LDA package implemented by
-Blei et al.~\cite{944937}; and our topic plotter, implemented in
+Blei et al.~\cite{944937}; and our trend plotter, implemented in
 Haskell.
@@
 When we examined PostgreSQL, we did not find many trends with two or
-more topics with a similarity of $7/10$. Those trends that we did find
-were not very large, lasting on 3 months at most. The first and second
+more topics, using a similarity of $7/10$. Those trends that we did find
+were not very large, lasting only 3 months at most. The first and second
 largest trends were dedicated to the work of two external developers:
@@
 PostgreSQL developer D'Arcy. Other topics of the larger trends were
-changes to the TODO list in HTML, and time string formating topics
+changes to the TODO, and time string formating topics
 relating to timezones.
@@
 \emph{fix}, \emph{update}.  By decreasing the similarity constraint to
-$1/2$ similarity the largest most the largest most common trend, which
-stretched across development, contained these same words
+$1/2$ similarity the largest most common trend, which
+stretched across the entire development, contained these same words
 (\emph{patch}, \emph{fix}, \emph{update}). The second largest trend
@@
 Central Europe Time (CEST) from email headers, issues with ALTER TABLE
-and CVS branch merging (because CVS does not record merges explicitly).
+and CVS branch merging, because CVS does not record merges explicitly.
 

### 2009-04-11 lda-paper.tex
what

@@
 
-The second largest trend was during the month of March 2001 relating
-to incremental building and the porting of developer Konstantin's
-Solaris port of the Firebird build files. The third largest trend was
-about JDBC, which is how Firebird and Java communicate.  Other trends
-included topics regarding AIX PPC compilation, updating the build
-process, internationalization and UTF8, Darwin build support and bug
-fixing.
+The second largest trend was during the month of March 2001. It was
+related to incremental building and the porting of developer
+Konstantin's Solaris port of the Firebird build files. The third
+largest trend was about JDBC, which is how Firebird and Java
+communicate.  Other trends included topics regarding AIX PPC
+compilation, updating the build process, internationalization and
+UTF8, Darwin build support and bug fixing.
 

### 2009-04-11 lda-paper.tex
removed a bunch of comments
main edits and spelling

@@
 
-%   At different times, developers focus on different topics and tasks
-%   during development.  These different topics and tasks might
-%   be obvious to developers familiar with the project, but other
-%   stake-holders, like managers, might be less aware of the topics of
-%   development. Imagine the scenario where developers unexpectedly had to deal with
-%   issues orthogonal to the features suggested, at the end of the
-%   iteration a manager would wonder where all the effort went to.
-%   Currently it is difficult for stake-holders to
-%   audit what actually occurred and when certain features or topics
-%   were focussed on. We propose a visualization of the
-%   changing development topics over time. Our approach uses a topic
-%   analysis tool like Latent Dirichlet Allocation (LDA) or Latent
-%   Semantic Indexing (LSI). Instead of applying the tool to the
-%   entire corpus we apply it to windows (periods) of the corpus. This allows us to
-%   see the evolving stream of topics of development occurring
-%   over time.
-
-%   During different periods, developers focus on different topics and
-%   different tasks during development. Stake-holders, such as managers,
-%   not directly involved with the development might have less awareness
-%   of the project and the topics and tasks that the developers are
-%   focussed on.  Imagine the scenario where a manager is trying to
-%   discover where all the developers' effort went after features have
-%   not been implemented because the developers unexpectedly had to deal
-%   with issues orthogonal to these features.  We propose to help
-%   stake-holder understand the topics of development by extracting
-%   topics from the descriptions of changes that programmers provide
-%   when they make changes.  We leverage topic analysis tools such as
-%   Latent Dirichlet Allocation (LDA) or Latent Semantic Indexing (LSI)
-%   to analyze periods, such as months, within a project's
-%   lifetime. Previous work applied LDA and LSI to
-%   to a project's entire lifetime.  These tools associate
-%   descriptions with independent topics extracted from these
-%   descriptions.  We propose a visualization of changing development
-%   topics over time that allows us to see the evolving stream of topics
-%   of development occurring over time. We also show that windowed topic
-%   analysis offers benefits compare to entire lifetime topic analysis
-%   because of the independence of topics between windows.
-
-% During different periods, developers focus on different topics and
-% different tasks during development. Stake-holders, such as managers, who are
-% not directly involved with the project's development might have less awareness of the
-% project and the topics and tasks that the developers have focussed on.
-% Imagine the scenario where a manager is trying to discover where all the
-% developers' effort went after features have not been implemented because the
-% developers unexpectedly had to deal with issues orthogonal to these
-% features.  We propose to help stake-holders understand the topics of
-% development by extracting topics from the descriptions of changes that
-% programmers provide when they make commits.  We leverage topic analysis
-% tools such as Latent Dirichlet Allocation (LDA) or Latent Semantic Indexing
-% (LSI) to analyze periods, such as months, within a project's lifetime,
-% whereas previous work applied LDA and LSI to to a project's entire lifetime.  These
-% tools associate descriptions with independent topics extracted from these
-% descriptions.  
-% We propose a visualization of changing development topics
-% over time that allows us to see the evolving stream of topics of development
-% occurring over time. We demonstrate that windowed topic analysis offers
-% benefits over entire lifetime topic analysis because of the
-% independence of topics between windows.
-
   As development on a software project progresses, developers will
@@
 
-%   In this lab report we seek to explore how topics shift over time in
-%   the source control repository of a project. We analyze commit
-%   comments, and even source code per revision per each time window and
-%   extract the topics prevalent in that window. We expect that topic
-%   sets will change over time and the change in topic sets indicate a
-%   change in the focus of development. We suppose this could help us to
-%   segment the revisions on the time line to discover parts of an
-%   iteration. For topic extraction we use \lda (LDA).
-
 \end{abstract}
@@
 %?
-
-%
-
-% In the presence of fine grained changes and detailed change-logs
-
-% Developers topics the focus of develop
-
-% many change-logs how to sort by change type
-
-% interleaved changes relate similar changes
-
 What happened in the previous maintenance or development iteration?
@@
 work on?
-
-% XXX Come back here, clear up and cut 
-% Topic analysis, with respect to
-% Software Control Systems (SCSs) such as CVS, Subversion, Git etc.,
-% is potentially valuable as it has many uses for various
-% stakeholders.  Imagine a development team has agreed to implement
-% some features at the start of a development iteration.  By the end
-% of the iteration only some of the agreed upon features are
-% implemented.  The developers insist they were focusing on the
-% features agreed upon, but what actually happened?  Topic analysis
-% could help by teasing out the distinct topics or focusses of the
-% developers.  Perhaps the developers had to deal with other pressing
-% issues.  Developers might have grouped those orthogonal issues into
-% the features they were working on, thus when asked about how their
-% time was spent the orthogonal changes would not be covered.  This is
-% the kind of underlying story that we hope that windowed topic
-% analysis can uncover. If windowed topic analysis can discover
-% important but overlooked topics, it can help both developers and
-% managers.  Developers and Managers can see a summary of what worked
-% on and the topics of development.
 
@@
 \ref{tab:portability}.
-
-%It is apparent that we have to compare these topics some how by
-%item-set or similarity.
-
-% \begin{table*}
-% \centering
-% \begin{tabular}{|ccc|l|}
-% \hline
-% 2000 &  Jul &  31 &    chmod \\
-% 2000 &  Sep &  29 &    fixes benchmark logging windows \\
-% 2000 &  Nov &  28 &    typo fix insert\_multi\_value \\
-% 2001 &  Jan &  27 &    fixes Innobase Cleanups auto-union \\
-% 2001 &  Mar &  28 &    2 topics bugfix, logging , TEMPORARY,  \\
-% \hline
-% 2001 &  Jul &  26 &    update Allow TABLES LOCK [a] \\
-
-% 2001 &  Aug &  25 &    tables row version [a] \\
-% \hline
-% 2001 &  Sep &  24 &    update checksum merge \\
-% 2001 &  Oct &  24 &    fixed fix \\
-% 2001 &  Dec &  23 &    HPUX SCO fix \\
-% \hline
-% 2002 &  Feb &  21 &    net buffer length  max\_allowed\_packet [b] \\
-% 2002 &  Mar &  23 &    small buf fix [b]  \\
-% \hline
-% 2002 &  May &  22 &    [popular] fix SCO OSF1 table\_name \\
-% 2002 &  Nov &  18 &    HPUX11 compiler HP \\
-% \hline
-% 2003 &  Feb &  16 &    Linux errno  [c] \\
-% 2003 &  Mar &  18 &    alarm bookmark bug [c] \\
-% \hline
-% 2003 &  Sep &  14 &    Auto logging merge windows distribution fix 64-bit 4.0 Cleanup \\
-% \hline
-% \end{tabular}
-% \caption{Tracking topics associated with the word portability, note some continuous blocks}
-% \label{tab:portability}
-% \end{table*}
 
@@
 
-%\shrinkit
-%\subsection{Datasets and Tools}
-%\shrinkit
-
-%   1. [ ] LDA
-
-%   2. [ ] Extractor
-
-%   3.   Similarity
-
-%   4.   Smearing
-
-%   5.   Visualizer
-
-%LDA Skeleton
-%1. [ ] Abstract 
-%2. [ ] Introduction
-
-%3. [ ] Previous Work [0/2]
-%   2. [ ] LDA Work [0/6]
-
 \shrinkit
@@
 
-% Extractors:
-% \begin{itemize}
-% \item CVSSuck - CVS Suck mirrors RCS files from a CVS repository. 
-% \item softChange - extract CVS facts to a PostgreSQL database.
-% \item bt2csv - Convert BitKeeper repositories to facts in CSV
-%   databases.
-% \end{itemize}
-
-%   1. Interesting smears 
-%   2.   postgresql
-
 %\subsection{MySQL}
@@
 %   3.   mysql
-
-%XXX Redo?
 
@@
 
-% When we examined PostgreSQL, we did not find many trends with two or
-% more topics; however, we did find one large trend which stretched
-% across most of its existence.  Much of the large trend consisted of
-% stop words but it did include topics which had non-stop words such as
-% \emph{patch}, \emph{fix}, \emph{update}.
-
-% The second largest topic was interesting because it concerned changes from
-% 1998, from developers using Central Europe Time (CEST), where the two
-% notable tokens were 1998 and CEST. These changes appear to be patches
-% submitted by one particular developer who lived in that timezone.
-
-% The third largest trend we found appeared to concerned with copying
-% and moving files between branches as it referenced the old branch.
-% branches.  We inferred this trend based on the fact that the old
-% version identifier was referenced, and we know this this is common
-% practice when performing a branch merge, since CVS cannot record such
-% information directly.
-
-% Once stop words were removed the some of the purposes became more
-% clear. The large trend disappeared, the largest trend became bug fixes
-% submitted by a developer named Dal Zotto which is a short trend.
-
-%XXX REDO POSTEGRESQL?
-% Large Trends is not the point!!!
-% That's what normal LDA gives us
-
-%But large trends are not the point, by investigating windowed trends
-%we want more local information about trends being worked on.
-
 \shrinkit
@@
 %   5.   what's that other db?
-
-%XXX REDO?
 
@@
 
-% We tracked Firebird from August 2000 to January 2006.
-% Our first attempt with Firebird resulted in a set of trends that were
-% dominated by one large trend containing tokens that were mostly stop
-% words such as references to one author
-% \texttt{carlosga05} and the words: \emph{the, to, a, for, of, is, it}.
-
-% The second largest trend related to branching and possibly merging; it
-% contained language such as ``was added on branch'' and contained
-% references to tags and branches.
-
-% The third largest trend contained references to the changelog, the
-% build system, architecture and platform support. It specifically
-% mentioned tokens such as \emph{Solaris}, \emph{port} and \emph{x86}.
-
-% Other trends included topics regarding AIX PPC compilation, updating
-% the build process, internationalization and UTF8, Darwin build support
-% and bug fixing.
-
-% Topics that were not trends but appeared to be  interesting were mostly external bug
-% fixes submitted to the project.
-%  In these cases, the developers would express thanks in their commit log comments, such as ``Thanks, Bill Lam''.  Other easily discernible topics included 
-%  tokens and topics such as: \emph{compiler workarounds}, \emph{nightly updates}, \emph{packets} and \emph{MSVC++}.
-
 \shrinkit
@@
 
-%XXX REDO?
-
-%XXX Migod todo
+% Migod todo
 % I think that you should enlarge this discussion of MaxDB (at the
@@
 \ref{fig:topicsmear}). Using a topic similarity of $7/10$ we evaluated
-of MaxDB 7.500, the first period from June 2004 to January 2005, and
-the second period from June 2005 to June 2006.
+MaxDB 7.500. MaxDB 7.500's first period was from June 2004 to January
+2005, and its second period was from June 2005 to June 2006.
 
@@
 (Figure \ref{fig:histogram}).  Other tokens mentioned are
-\emph{Sutcheck v1.09} (the prefix SUT stands for Storable Unit
-Type), a tool that would also automate check-ins using a Perforce SCS tool
-which was exporting check-ins to CVS.
+\emph{Sutcheck v1.09} (the prefix SUT stands for Storable Unit Type),
+that is a tool that would also automate check-ins using a Perforce SCS
+tool which was exporting check-ins to CVS.
 
@@
 seemed to include tokens related to operating system support, such as
-Linux and Windows, as well as architecture support, AMD64 and
-Opteron. The word \emph{problem} was common among all of these
+\emph{Linux} and \emph{Windows}, as well as architecture support, \emph{AMD64} and
+\emph{Opteron}. The word \emph{problem} was common among all of these
 trends. This trend seemed related to the smaller fourth largest trend
@@
 previous month. We investigated the actual revisions and we did find
-there was difference in behavior between, and that the last month of
+there was a difference in behavior between, and that the last month of
 development commits were mostly referencing the bug tracker. If the
-topics of one month were 20 or more unique bug tickets being
-addressed, the global LDA analysis would probably miss this, yet these
-bug tickets were the focus of development for that month and not
+topics of one month were about unique bug tickets being
+addressed, the global topic analysis would probably miss this, yet these
+bug tickets were the focus of development for that month but not
 necessarily globally relevant.
@@
 its topic ten words. This shows that optimizer was important but it
-had been obscured, while it would have been noticed using a more
+had been obscured, while it would have been noticed using the more
 locally relevant approach that we propose.
@@
 topic. Perhaps this is because there was a cluster of \emph{Perl}
-mentions during one month while \emph{PHP} was mentions were more
+mentions during one month while the \emph{PHP} mentions were more
 spread out.
@@
 similarity thresholds showed that by removing common words one reduces
-the general similarity of topics. That said, the larger topics still
+the general similarity of topics. That said, the larger topics were still
 clearly apparent. Thus if more relevant stop words are added one
@@
 % UNICODE
-
 % DEC/OSF1
-
 % ODBC
-
 % crash
-
 % PHP yet Perl is mentioned is less but it part
-
 % SQLCLI
 % %V75
-
 % SYSDD.cpnix
 % backup
-
 % select
@@
 % view
-
 % debug
-
 % V75 is mentioned many times maxdb version guh
@@
 
-%XXX Migod notes This section is perhaps the most important of the
+% Migod notes This section is perhaps the most important of the
 % whole paper.  You need expand it.  Maybe you should move the first
@@
 Previous work on topic analysis that employed LSI and LDA typically
-extracted a specified number of topics and then tracked them over time. This
-means that a single, static set of topics was tracked across the whole
-development history of a project. \emph{Global topic analysis} is
-topic analysis applied to entire lifetime of a project.
+extracted a specified number of topics and then tracked them over
+time. This means that a single, static set of topics was tracked
+across the whole development history of a project. \emph{Global topic
+  analysis} is topic analysis applied against the entire lifetime of a
+project.
 
@@
 If we approach the difference of global topic analysis and windowed
-topic analysis via common tokens we can see that common token tend to
+topic analysis via common tokens we can see that common tokens tend to
 dominate both results, at least in the case of MaxDB 7.500 our local
@@
 The number of commits per month is inconsistent as some months have a
-lot of changes while others months have almost none.
+lot of changes while other months have almost none.
 %   1.   validation
@@
 
-% Even with our liberal topic similarity metrics that produced both
-% long and short trends, we showed that there are only a few trends in a
-% repository which recur. Since so few trends recur and so many
-% trends appear only once this suggests that global topic analysis might
-% be ignoring locally unique topics.
-
-% The utility of global topic analysis is questionable if the value of
-% information decreases as it becomes older. Perhaps older trends will
-% dominate the topic analysis. Windowed localized topic analysis show
-% what are the unique topics.
-
 We proposed and demonstrated the application of windowed topic
@@
 periods of development. This approach differs from previous work that
-would applied topic analysis globally to the entire history of a
-project without regard to time.. We showed that many topics that exist
+applied topic analysis globally to the entire history of a
+project without regard to time. We showed that many topics that exist
 locally are relevant and interesting yet would often not be detected
@@
 of three database systems with a particular focus on MaxDB 7.500. We
-showed that global topic analysis missed important topics like such as
-ODBC changes.
-
-%XXX summarize our position
-
-%XXX Add more detail here
+showed that global topic analysis missed important topics such as
+\emph{ODBC}.
 
@@
 information than the views that global analysis could show, it
-indicates how focused a period is by the total number of topics as
-well shows topics by similarity so one can track trends across time.
-Our trend-histogram immediately hi-lights and measures the size of
-trends while our trend-time-line view shows how a topic reoccurs over
-time. These visualizations help us get an over feeling about the
-common and unique topics that developers focus on. If implemented
-properly a user can easily zoom and query a topic.
+indicates how focused a period is by the total number of topics, as
+well, it shows topics by similarity so one can track trends across
+time.  Our trend-histogram immediately highlights and measures the
+size of trends while our trend-time-line view shows how a topic
+reoccurs over time. These visualizations help us get an overall
+feeling about the common and unique topics that developers focus
+on. If implemented properly a user could easily zoom in and query a
+summary of a topic or trend.
 

### 2009-04-11 lda-paper.tex
latest edits

@@
 shown in Figure \ref{fig:topicsmear} and Figure \ref{fig:zoomedsmear},
-which displays trends as they occur on the time-based x-axis
+that displays trends as they occur on the time-based x-axis
 (placement along the y-axis is arbitrary).  Another technique is
@@
 \emph{Sutcheck v1.09} (the prefix SUT stands for Storable Unit Type),
-that is a tool that would also automate check-ins using a Perforce SCS
-tool which was exporting check-ins to CVS.
+\emph{Sutcheck} is a tool that would also automate check-ins using a Perforce SCS
+tool, which was exporting check-ins to CVS.
 
 The second largest common trend seems to be a side effect of an
-automated check-in which is annotated as ``implicit check-in'' (see
+automated check-in that is annotated as ``implicit check-in'' (see
 the bottom of Figure \ref{fig:zoomedsmear}). These were check-ins that
@@
 trends. This trend seemed related to the smaller fourth largest trend
-that had tokens \emph{AMD64} and \emph{Windows}. This shows that
+that had tokens \emph{AMD64} and \emph{Windows}. This example shows that
 topics can overlap but still not match via our similarity measure.
@@
 contained 1 unique Bug tracker URL, this pattern did not occur in the
-previous month. We investigated the actual revisions and we did find
+previous month. We investigated the revisions and we did find
 there was a difference in behavior between, and that the last month of
@@
 about every topic, so we added them to the stop words list. As a
-result, longer trends were shortened and in some cases the total
-number of topics found per month was less.  Evaluating different
+result, longer trends were shortened and sometimes the total
+number of topics found per month were less.  Evaluating different
 similarity thresholds showed that by removing common words one reduces
@@
 
-We carried out the same kind of analysis, in order to compare the
+We carried out the same kind of analysis,          to compare the
 results against our time-windowed approach.  To produce
@@
 third row of Figure \ref{fig:statictopics}, while messages related to
-other topics did not appear very often.
+other topics did not appear often.
 % This technique seems useful if the 20 extracted topics are relevant
@@
 topic analysis via common tokens we can see that common tokens tend to
-dominate both results, at least in the case of MaxDB 7.500 our local
-topic approach netted us topics which contained these relatively
-important and common words: \emph{UNICODE}, \emph{DEC/OSF1},
-\emph{ODBC}, \emph{crash}, \emph{SQLCLI}, \emph{SYSDD.cpnix},
-\emph{backup}, \emph{select}, \emph{make}, \emph{memory}, \emph{view},
-and finally \emph{debug}. \emph{ODBC} is quite important because it
+dominate both results. For MaxDB 7.500, our local topic approach
+netted us topics that contained these relatively important and common
+words, which did not occur in the topics produced by global topic
+analysis: \emph{UNICODE}, \emph{DEC/OSF1}, \emph{ODBC}, \emph{crash},
+\emph{SQLCLI}, \emph{SYSDD.cpnix}, \emph{backup}, \emph{select},
+\emph{make}, \emph{memory}, \emph{view}, and finally
+\emph{debug}. \emph{ODBC} is an important topic token, because it
 often determines how databases communicate with software. None of
@@
 
-The number of commits per month is inconsistent as some months have a
-lot of changes while other months have almost none.
+The number of commits per month is inconsistent as some months have
+many changes while other months have almost none.
 %   1.   validation
@@
 via global topic analysis. We identify recurring topics with a topic
-similarity measure which looks for topics which recur and mutate
+similarity measure that looks for topics which recur and mutate
 repeatedly throughout the development of the software project.
@@
 size of trends while our trend-time-line view shows how a topic
-reoccurs over time. These visualizations help us get an overall
+reoccurs over time. These visualizations help us get an general
 feeling about the common and unique topics that developers focus
-on. If implemented properly a user could easily zoom in and query a
+on during development. If implemented properly a user could easily zoom in and query a
 summary of a topic or trend.

### 2009-04-11 lda-paper.tex
some more edits but it puts it at 11 pages ;_;

@@
   commit log comments with independent topics extracted from these
-  commit log comments.  In this work, we use LDA and LSI to analyze
+  commit log comments.  In this work, we use LDA  to analyze
   periods, such as months, within a project's lifetime to create a
-  time-windowed model of changing development topics.  We propose a
-  visualization of this model that allows us to explore the evolving
+  time-windowed model of changing development topics.  We propose 
+  visualizations of this model that allows us to explore the evolving
   stream of topics of development occurring over time.  We demonstrate
@@
 
-Commits in SCSs can have
-multiple purposes. Thus we hope that our research will help uncover
-these multiple purposes that manifest themselves as topics interleaved
-within the commits.  Others have demonstrated that 
-topic analysis techniques, such as Latent Semantic Indexing (LSI) and
-Latent Dirichlet Allocation (LDA), can aid in separating topics from
-these mixed comments. These topics often describe distinct ideas that
-are interleaved throughout the development. A topic represents a
-groups of commit log comments that are related to each other by their
-content.  In this case a topic is a set of tokens.
+%XXX RM 
+Commits in source control systems (SCSs) can have multiple
+purposes. Thus we hope that our research will help uncover these
+multiple purposes that manifest themselves as topics interleaved
+within the commits.  Others have demonstrated that topic analysis
+techniques, such as Latent Semantic Indexing (LSI) and Latent
+Dirichlet Allocation (LDA), can aid in separating topics from these
+mixed comments. These topics often describe distinct ideas that are
+interleaved throughout the development. A topic represents both a word
+distribution and a group of commit log comments that are related to
+each other by their content.  In this case a topic is a set of tokens.
 
@@
 
-We propose to extend existing topic analysis techniques by
-applying them to windows of documents over time. We hope that by
+We propose to extend existing topic analysis techniques by applying
+them to windows of documents over time. We hope that by
 looking %XX X ADDED
 at topic clusters during periods or windows over time we can identify
-ongoing areas of focus on a select group of topics as well reflect more intermittent topics. Figure \ref{fig:lda} illustrates
-the kind of visualization we want to automatically generate. We want
-to see topics that are unique to a certain time period, displayed
-across the time axis. Those topics that recur, or occur over a
-larger period are plotted continuously. In our example we have
-titled each topic with a single word.
+ongoing areas of focus on a select group of topics as well reflect
+more intermittent topics. Figure \ref{fig:lda} illustrates the kind of
+visualization we want to automatically generate. We want to see topics
+that are unique to a certain time period, displayed across the time
+axis. Those topics that recur, or occur over a larger period are
+plotted continuously. In our example we have titled each topic with a
+single word drawn from itself.
 
-In this paper we explore how topics shift over time in the source
-control system (SCS) of a project, using several open source database
-systems as examples. We analyze commit log comments in each
-time window and we extract the topics of that time window. We expect
-that topics will change over time and the similarities and differences
-in these topics will indicate developer focus and changes in developer
-focus as it changes over time.
+In this paper we explore how topics shift over time in the SCS of a
+project, using several open source database systems as examples. We
+analyze commit log comments in each time window and we extract the
+topics of that time window. We expect that topics will change over
+time and the similarities and differences in these topics will
+indicate developer focus and changes in developer focus as it changes
+over time.
 
@@
 \emph{message} is a block of text written by developers. In this
-paper, messages will be the CVS commit log comments made when the user
-commits changes to files in a repository. A \emph{word distribution}
-is the summary of a message by word count. Each word distribution will
-be over the words in all messages. However, most words will not appear
-in a message, a word distribution is effectively a word count divided
-by the message size. A \emph{topic} is a word distribution, a
-set of words that form a word distribution that is unique and
-independent within the set of documents in our total corpus. One could
-think of a topic as a distribution of the centroid of a group of
-messages. In this paper we often summarize topics by the top 10 most
-frequent words of their word distribution.  A \emph{trend} is a one or
-more similar topics that recur over time.  Trends are particularly
-important, as they indicate long-lived and recurring topics that may
-provide key insights into the development of the project.
+paper, messages will be the CVS and BitKeeper commit log comments made
+when the user commits changes to files in a repository. A \emph{word
+  distribution} is the summary of a message by word count. Each word
+distribution will be over the words in all messages. However, most
+words will not appear in a message, a word distribution is effectively
+a word count divided by the message size. A \emph{topic} is a word
+distribution, a set of words that form a word distribution that is
+unique and independent within the set of documents in our total
+corpus. One could think of a topic as a distribution of the centroid
+of a group of messages. In this paper we often summarize topics by the
+top 10 most frequent words of their word distribution.  A \emph{trend}
+is a one or more similar topics that recur over time.  Trends are
+particularly important, as they indicate long-lived and recurring
+topics that may provide key insights into the development of the
+project.
 
@@
 Grant et al.~\cite{scottcordy} have used an alternative technique,
- called Independent Component Analysis to
-separate topic signals from software code.
+called Independent Component Analysis to separate topic signals from
+software code.
 
@@
 name.  To name each topic we selected a term that seemed to best
-summarized that topic.  After extracting these topics, we attempted to
+summarize that topic.  After extracting these topics, we attempted to
 track the evolution of topics by visualizing the topics and joining
@@
 
-Our methodology is to first extract the commits from SCS repositories
-such as CVS or Bitkeeper. Then we extract the revisions and commits,
-grouping the revisions into commits if necessary. We extract the
-commit log comment of each commit, count the occurrence of each word
-in the message, remove stop words, and then produce word distributions
-for each message.
+%XXX Smaller?
+
+Our methodology is to first extract the commit log comments from a
+project's SCS repository. We filter out stop words and produce word
+distributions from these messages. These distributions are bucketted
+into windows, and then each window is subject to topic analysis and
+then we analyze and visualize the results.
 
@@
 as rsync, CVSSuck, \texttt{softChange} and bt2csv.
-\texttt{softChange} provided a general schema for storing
-revisions. CVSSuck and rsync mirrored CVS repositories while bt2csv
-mirrored web accessible BitKeeper repositories.
+\texttt{softChange} provided a general schema for storing revisions
+and grouped CVS revisions into commits. CVSSuck and rsync mirrored CVS
+repositories while bt2csv mirrored web accessible BitKeeper
+repositories.
+
+%XXX Smaller?
 
@@
 comments that are added when revisions to the project are committed.
-Documents are converted into word count distributions. These
-distributions are then normalized by the size of the message. After
-all messages are processed the distributions are extended to include
-all words from all of the  distributions.
+Per each commit log comment, we count the occurrence of each word in
+the message, remove stop words, and then produce word distributions
+for each message.  These distributions are then normalized by the size
+of the message. After all messages are processed the distributions are
+extended to include all words from all of the distributions.
 
@@
 of a set time unit because it simplifies analysis.  Windowing by time
-allows for many levels of granularity.  We usually used one month as
-the length of our time windows. While we could use different window
-lengths for the study we think that a month is a sizable unit of
+allows for many levels of granularity.  We used one month as the
+length of our time windows. While we could use different window
+lengths for this study we think that a month is a sizable unit of
 development, which is large enough to show shifts in focus, but coarse
@@
 Once we have our data windowed, we apply our Topic Analysis tool to
-each window and extract the top $N$ topics. We used $20$
-topics in our case studies. Our Topic Analysis tool is an
-implementation of LDA, but we could have used other similar tools like
-LSI, but previous work showed more promising topic analysis results
-with LDA.  We note that this step is a slow one, as executing even one
-instance of LDA involves significant computation, and we perform LDA
-once per window.
+each window and extract the top $N$ topics. We used $20$ topics in our
+case studies. Our Topic Analysis tool is an implementation of LDA, but
+we could have used other similar tools like LSI, but previous work
+showed more promising topic analysis results with LDA.  We note that
+this step is a slow one, as executing even one instance of LDA
+involves significant computation, and we perform LDA once per window.
 

### 2009-04-11 lda-paper.tex
lame

@@
 Once we have our topics extracted for each window, we analyze them and
-look for topics that recur across windows. Our input for topic
-similarity is the top 10 most common words of each topic.
+look for topics that recur across windows.  We then cluster these
+topics into trends by comparing them to each other using topic similarity.
 
-We then cluster these topics into trends by comparing them to each
-other.  Each topic is compared to each other topic in the system,
-given a certain threshold of top 10 similarity, such as 8 out of 10
-words. Then given this topic similarity matrix, we find the transitive
-closure of similar topics; that is, if we model topics are nodes and
-arcs as similarity, we fill flood along the similarity arcs until we
-have partitioned the topics into clusters of similar topics. Figure
-\ref{fig:closure} illustrates clustering of topics by 
-topic similarity. These clusters of topics are called trends. A trend
-is probably interesting if it contains more than one topic.
+Our input for topic similarity is the top 10 most common words of each
+topic.  Each topic is compared to each other topic in the system,
+given a certain threshold of top 10 word similarity, such as 8 out of
+top 10 matching words. Then given this topic similarity matrix, we
+find the transitive closure of similar topics; that is, if we model
+topics are nodes and arcs as similarity, we fill flood along the
+similarity arcs until we have partitioned the topics into clusters of
+similar topics. Figure \ref{fig:closure} illustrates clustering of
+topics by topic similarity. These clusters of topics are called
+trends. A trend is probably interesting if it contains more than one
+topic.
 

### 2009-04-12 lda-paper.tex
One good edit pass

@@
 \shrinkit
+
 \begin{figure*}[t]
@@
 This technique has weaknesses in that nodes that are a few neighbors
-away in similarity might not share any similar words.  We use this measure
-because we want to be able to color or highlight topics that are
-related to each other and track them throughout time.
+away in similarity might not share any similar words.  We use this
+measure because we want to be able to color or highlight topics that
+are related to each other and track them throughout time.
 
@@
 window, and then stack the smaller trends on top.  Although there is a
-chance that gaps will be left due to non-optimal stacking in practice
+chance that gaps will be left due to non-optimal stacking, in practice
 there are lot of small trends (90\% of all trends contain 1 topic)
@@
 \emph{fix}, \emph{update}.  By decreasing the similarity constraint to
-$1/2$ similarity the largest most common trend, which
-stretched across the entire development, contained these same words
-(\emph{patch}, \emph{fix}, \emph{update}). The second largest trend
-mentions Dal Zotto, while the third largest trend mentions the
-\texttt{[PATCHES]} mailinglist and the names of some patch
-contributors.  Other repeating topics refer to portability with Win32,
-Central Europe Time (CEST) from email headers, issues with ALTER TABLE
-and CVS branch merging, because CVS does not record merges explicitly.
+$1/2$ similarity the largest most common trend, which stretched across
+the entire development, contained these same words (\emph{patch},
+\emph{fix}, \emph{update}). The second largest trend mentions Dal
+Zotto, while the third largest trend mentions the \texttt{[PATCHES]}
+mailinglist and the names of some patc contributors.  Other repeating
+topics refer to portability with Win32, Central Europe Time (CEST)
+from email headers, issues with ALTER TABLE and CVS branch merging
+(CVS does not record merges explicitly).
 
@@
 contained 1 unique Bug tracker URL, this pattern did not occur in the
-previous month. We investigated the revisions and we did find
-there was a difference in behavior between, and that the last month of
-development commits were mostly referencing the bug tracker. If the
-topics of one month were about unique bug tickets being
-addressed, the global topic analysis would probably miss this, yet these
-bug tickets were the focus of development for that month but not
+previous month. We investigated the revisions and we found that
+developers were referencing the bug tracker more during the last
+month.  If the topics of one month were about unique bug tickets being
+addressed, the global topic analysis would probably miss this, yet
+these bug tickets were the focus of development for that month but not
 necessarily globally relevant.
@@
 The query optimizer was a topic that recurred during MaxDB's
-development. In our plots, topics that mention \emph{optimizer} occur four
-times, yet in the global-trend-view (Figure \ref{fig:statictopics}) it
-is not in any of the topics. A query optimizer is rather important to
-an DBMS, but as we have shown it just does not pop up as a topic on
-its own. We tried to remove words to see if we could get an 
-\emph{optimizer} topic. After removing stop words and then two of the most
-common words, the global analysis finally had a topic with optimizer in
-its topic ten words. This shows that optimizer was important but it
-had been obscured, while it would have been noticed using the more
-locally relevant approach that we propose.
+development. In our plots, topics that mention \emph{optimizer} occur
+four times, yet in the global-trend-view (Figure
+\ref{fig:statictopics}) it is not in any of the topics. A query
+optimizer is an important component of an DBMS, but as we have shown
+it does not appear as a topic on its own. We tried to remove words to
+see if we could get an \emph{optimizer} topic. After removing stop
+words and then two of the most common words, the global analysis
+finally had a topic with optimizer in its topic ten words. This shows
+that optimizer was important but it had been obscured, but would have
+been noticed using the more local topic analysis.
 
@@
 analysis mentioned \emph{PHP} in 5 different topics, yet only
-mentioned \emph{Perl} in 4 different windowed topics and 1 global
+mentioned \emph{Perl} in 4 different topics and 1 global
 topic. Perhaps this is because there was a cluster of \emph{Perl}
@@
 
-We noticed that \emph{Perforce} and \emph{Changelist} were in just
-about every topic, so we added them to the stop words list. As a
-result, longer trends were shortened and sometimes the total
-number of topics found per month were less.  Evaluating different
-similarity thresholds showed that by removing common words one reduces
-the general similarity of topics. That said, the larger topics were still
+Just about every topic included the words \emph{Perforce} and
+\emph{Changelist}, so we added them to the stop words list. As a
+result, longer trends were shortened and sometimes the total number of
+topics found per month were reduced.  Evaluating different similarity
+thresholds showed that by removing common words one reduces the
+general similarity of topics. That said, the larger topics were still
 clearly apparent. Thus if more relevant stop words are added one
@@
 time. This means that a single, static set of topics was tracked
-across the whole development history of a project. \emph{Global topic
-  analysis} is topic analysis applied against the entire lifetime of a
-project.
+across the whole development history of a project, we refer to this as \emph{global topic
+  analysis}.% is topic analysis applied against the entire lifetime of a
+%project.
 
-We carried out the same kind of analysis,          to compare the
-results against our time-windowed approach.  To produce
+We carried otu \emph{global topic analysis} on MaxDB 7.500 and
+compared this against our windowed topic analysis.  To produce
 figure~\ref{fig:statictopics}, we extracted 20 topics and plotted the
@@
 may be that some topics are of strong interest but only briefly; in
-such a case, a window-based model gives a much stronger indication of
-the fleeting importance of such topics, and can help to put such a
-topic into its proper context.
+such a case, a windowed topic analysis gives a much stronger
+indication of the fleeting importance of such topics, and can help to
+put such a topic into its proper context.
 
@@
 
-We truncated the topics to top 10 tokens, so this summary of a topic
-might not have been as useful as determining the actual topic distance
-between two word topic distributions. By truncating tokens we could be
-losing valuable data.
+We compared topics using the top 10 tokens, this approach could be
+throwing data away and might not be as useful as determining the
+actual distance between two word topic distributions.
 
-Each project seemed to have their own set of stop words, and we did not
-remove them, but perhaps dropping some of these stop words would aid
-the clarity of topics produced through topic analysis; alternatively
-it might cause our topic similarity plots to be fundamentally
-different.
+% We truncated the topics to the top 10 tokens, so this summary of a topic
+% might not have been as useful as determining the actual topic distance
+% between two word topic distributions. By truncating tokens we could be
+% losing valuable data.
+
+Adding and removing stop words produced different results. Our choice
+of stopwords could be biased, and could affect the results.
+%, and our bias for or against such stop words could affect
+%the results.
+
+% Each project seemed to have their own set of stop words, and we did not
+% remove them, but perhaps dropping some of these stop words would aid
+% the clarity of topics produced through topic analysis; alternatively
+% it might cause our topic similarity plots to be fundamentally
+% different.
 
@@
 size of trends while our trend-time-line view shows how a topic
-reoccurs over time. These visualizations help us get an general
-feeling about the common and unique topics that developers focus
-on during development. If implemented properly a user could easily zoom in and query a
-summary of a topic or trend.
+reoccurs over time. These visualizations help us get a general feeling
+about the common and unique topics that developers focus on during
+development. If implemented interactively a user could easily zoom in
+and query for a summary of a topic or trend.
 
@@
 \subsection{ Future Work}
+\shrinkit
 \shrinkit
@@
 \bibliographystyle{latex8}
+
+\shrinkit
 \bibliography{lda-paper}

### 2009-04-12 lda-paper.tex
some spellcheck

@@
 project's SCS repository. We filter out stop words and produce word
-distributions from these messages. These distributions are bucketted
+distributions from these messages. These distributions are bucketed
 into windows, and then each window is subject to topic analysis and
@@
 Zotto, while the third largest trend mentions the \texttt{[PATCHES]}
-mailinglist and the names of some patc contributors.  Other repeating
+mailing-list and the names of some patch contributors.  Other repeating
 topics refer to portability with Win32, Central Europe Time (CEST)

### 2009-04-12 lda-paper.tex
minor spell checks

@@
 
-We carried otu \emph{global topic analysis} on MaxDB 7.500 and
+We carried out \emph{global topic analysis} on MaxDB 7.500 and
 compared this against our windowed topic analysis.  To produce
@@
 Adding and removing stop words produced different results. Our choice
-of stopwords could be biased, and could affect the results.
+of stop words could be biased, and could affect the results.
 %, and our bias for or against such stop words could affect

### 2009-04-12 lda-paper.tex
Minor edits

@@
 top 10 most frequent words of their word distribution.  A \emph{trend}
-is a one or more similar topics that recur over time.  Trends are
+is one or more similar topics that recur over time.  Trends are
 particularly important, as they indicate long-lived and recurring
@@
 Our technique differs from the previous techniques because we apply
-LDA locally to month-long windows of commit log comments, whereas other
-approaches apply LDA once to the entire project. This allows us to
-examine the time-based windowing of topics over its development
-history.
+LDA locally to month-long windows of commit log comments, whereas
+other approaches apply LDA once to the entire project. Windowed topic
+analysis allows us to examine the time-based windowing of topics over
+its development history.
 
@@
 words and then two of the most common words, the global analysis
-finally had a topic with optimizer in its topic ten words. This shows
+finally had a topic with optimizer in its topic ten words. Our analysis shows
 that optimizer was important but it had been obscured, but would have
@@
 analysis mentioned \emph{PHP} in 5 different topics, yet only
-mentioned \emph{Perl} in 4 different topics and 1 global
+mentioned \emph{Perl} in four different topics and one global
 topic. Perhaps this is because there was a cluster of \emph{Perl}
@@
 about the common and unique topics that developers focus on during
-development. If implemented interactively a user could easily zoom in
+development. If implemented interactively an user could easily zoom in
 and query for a summary of a topic or trend.

### 2009-04-12 lda-paper.tex
Removed some indefinites

@@
 dedicating a horizontal line for trend segments belonging to one
-trend. This means that if a trend contains discontinuous segments then
+trend. Therefore if a trend contains discontinuous segments then
 the segments appear on the same line.  However, the least common
@@
 
+% Previous work on topic analysis that employed LSI and LDA typically
+% extracted a specified number of topics and then tracked their
+% relationship to documents over time. This means that a single, static
+% set of topics was tracked across the whole development history of a
+% project, we refer to this as \emph{global topic
+%   analysis}.% is topic analysis applied against the entire lifetime of a
+% %project.
+
 Previous work on topic analysis that employed LSI and LDA typically
-extracted a specified number of topics and then tracked them over
-time. This means that a single, static set of topics was tracked
-across the whole development history of a project, we refer to this as \emph{global topic
-  analysis}.% is topic analysis applied against the entire lifetime of a
+extracted a specified number of topics from the entire development
+history of a project and then tracked their relationships to documents
+over time, we call this \emph{global topic analysis}.
+
+% This means that a single, static
+%set of topics was tracked across the whole development history of a
+%project, we refer to this as \emph{global topic
+%  analysis}.% is topic analysis applied against the entire lifetime of a
 %project.

### 2009-04-12 lda-paper.tex
et al stuff

@@
 In order to infer or associate the expertise of an author with topics
-extracted from SCS, Linstead et al. proposed an author-source-code model
+extracted from SCS, Linstead et al.\  proposed an author-source-code model
 using LDA~\cite{10.1109/MSR.2007.20}%,NIPS2007637,1321709}. 

### 2009-06-15 lda-paper.tex

@@
 \end{abstract}
+
+\section{Real Introduction}
+
+We care about the topics of development. We want to know given a
+certain period of time, what were the popular topics being worked on
+during that period. We also want to know if any of those topics
+occurred again, whether before or after our current period.
+
+These topics could help stakeholders determine what their coworkers
+were working on, what were common issues, what requirements were being
+worked on, etc.
+
+Topic analysis extracts groups of words from documents (commit log
+comments) that are word distributions, topics, which characterize
+clusters of documents. These extracted topics often correlate with
+actual development topics that the developers discussed.
+
+Usually topic analysis is applied to the entire history of a project,
+where as we suggest that more topics are probably local, if a
+programmer is working on a story card, they might never deal with that
+story card ever again. Thus that topic might be quite relevant to a
+small window of time but never relevant ever again.  By applying topic
+analysis locally to an iteration, both the developers and managers
+might be able to see what were the topics of development during that
+iteration, rather than the whole project.
+
+Topic analysis often allows for a single document, such as a commit
+message, to be related to multiple documents. This maps well to
+commits which often have more than one purpose.
+
+Commits in source control systems (SCSs) can have multiple
+purposes. Thus we hope that our research will help uncover these
+multiple purposes that manifest themselves as topics interleaved
+within the commits.  Others have demonstrated that topic analysis
+techniques, such as Latent Semantic Indexing (LSI) and Latent
+Dirichlet Allocation (LDA), can aid in separating topics from these
+mixed comments. These topics often describe distinct ideas that are
+interleaved throughout the development. A topic represents both a word
+distribution and a group of commit log comments that are related to
+each other by their content.  In this case a topic is a set of tokens.
 

### 2009-06-16 lda-paper.tex

@@
 
-\section{Real Introduction}
+\section{Introduction}
 
@@
 
+Topic analysis often allows for a single document, such as a commit
+message, to be related to multiple topics. This maps well to
+commits which often have more than one purpose.  A topic represents
+both a word distribution and a group of commit log comments that are
+related to each other by their content.  In this case a topic is a set
+of tokens.
+
 Usually topic analysis is applied to the entire history of a project,
@@
 
-Topic analysis often allows for a single document, such as a commit
-message, to be related to multiple documents. This maps well to
-commits which often have more than one purpose.
+Figure \ref{fig:lda} illustrates the kind of visualization we want to
+automatically generate. We want to see topics that are unique to a
+certain time period, displayed across the time axis. Those topics that
+recur, or occur over a larger period are plotted continuously. In our
+example we have titled each topic with a single word drawn from
+itself.
 
-Commits in source control systems (SCSs) can have multiple
-purposes. Thus we hope that our research will help uncover these
-multiple purposes that manifest themselves as topics interleaved
-within the commits.  Others have demonstrated that topic analysis
-techniques, such as Latent Semantic Indexing (LSI) and Latent
-Dirichlet Allocation (LDA), can aid in separating topics from these
-mixed comments. These topics often describe distinct ideas that are
-interleaved throughout the development. A topic represents both a word
-distribution and a group of commit log comments that are related to
-each other by their content.  In this case a topic is a set of tokens.
-
-\shrinkit
-\section{Introduction}
-\shrinkit
-
-%?
-What happened in the previous maintenance or development iteration?
-What were the developers working on? What was the developer in the
-next cubicle working on?  Which common topics and issues were dealt
-with in the current release of the software? What were the topics of
-the previous release?  Given the requirements agreed to at the start
-of the iteration did our developers work on them? What else did they
-work on?
-
-Imagine that a development team had agreed upon a set of features to
-implement for a development iteration, but these features were not
-finished by the end of the iteration. The developers insist they
-focussed on those features, yet their manager is not sure what
-happened. In a scenario like this, topic analysis can help tease out
-the topics of development that the developers were focussed on. Topic
-analysis could highlight the orthogonal issues that developers were
-spending time on.  Topic analysis extracts groups of words from
-documents (commit log comments) that are word distributions, topics,
-which characterize clusters of documents. These extracted topics often
-correlate with actual development topics that the developers
-discussed. By applying topic analysis locally to an iteration, both
-the developers and managers might be able to see what were the topics
-of development during that iteration. %XXX RM during that iteration
-
-%XXX RM 
-Commits in source control systems (SCSs) can have multiple
-purposes. Thus we hope that our research will help uncover these
-multiple purposes that manifest themselves as topics interleaved
-within the commits.  Others have demonstrated that topic analysis
-techniques, such as Latent Semantic Indexing (LSI) and Latent
-Dirichlet Allocation (LDA), can aid in separating topics from these
-mixed comments. These topics often describe distinct ideas that are
-interleaved throughout the development. A topic represents both a word
-distribution and a group of commit log comments that are related to
-each other by their content.  In this case a topic is a set of tokens.
-
-% We suppose this could help us to segment the revisions on the time
-% line to discover parts of an iteration.
-
-Topic analysis can aid in partitioning a project's
-time-line into periods of developer focus. By breaking apart an
-iteration into sets of topics and trends (topics that recur), we may
-be able to recognize the underlying software development process from
-these commits. Alternatively, we can determine what particular
-maintenance task was being performed at a given time.
-
-We propose to extend existing topic analysis techniques by applying
-them to windows of documents over time. We hope that by
-looking %XX X ADDED
-at topic clusters during periods or windows over time we can identify
-ongoing areas of focus on a select group of topics as well reflect
-more intermittent topics. Figure \ref{fig:lda} illustrates the kind of
-visualization we want to automatically generate. We want to see topics
-that are unique to a certain time period, displayed across the time
-axis. Those topics that recur, or occur over a larger period are
-plotted continuously. In our example we have titled each topic with a
-single word drawn from itself.
-
-In this paper we explore how topics shift over time in the SCS of a
-project, using several open source database systems as examples. We
-analyze commit log comments in each time window and we extract the
-topics of that time window. We expect that topics will change over
-time and the similarities and differences in these topics will
-indicate developer focus and changes in developer focus as it changes
-over time.
+In this paper we explore how topics shift over time in the source
+control system (SCS) of a project, using several open source database
+systems as examples. We analyze commit log comments in each time
+window and we extract the topics of that time window. We expect that
+topics will change over time and the similarities and differences in
+these topics will indicate developer focus and changes in developer
+focus as it changes over time.
 
@@
 \begin{itemize}
-\item Demonstrating the value of windowed topic analysis using trends.
-\item Multiple visualizations of topics and trends over time.
+\item Demonstrating that most topics are local in scope, thus would be
+  missed by a global topic analysis.
+\item Multiple visualizations of topics and trends over time to aid
+  communication and analysis of topics and trends.
 \item An exploratory case study using these techniques on several open source database systems.
@@
 
+%%%   
+%%%   
+%%%   \shrinkit
+%%%   \section{Introduction}
+%%%   \shrinkit
+%%%   
+%%%   %?
+%%%   What happened in the previous maintenance or development iteration?
+%%%   What were the developers working on? What was the developer in the
+%%%   next cubicle working on?  Which common topics and issues were dealt
+%%%   with in the current release of the software? What were the topics of
+%%%   the previous release?  Given the requirements agreed to at the start
+%%%   of the iteration did our developers work on them? What else did they
+%%%   work on?
+%%%   
+%%%   
+%%%   Imagine that a development team had agreed upon a set of features to
+%%%   implement for a development iteration, but these features were not
+%%%   finished by the end of the iteration. The developers insist they
+%%%   focussed on those features, yet their manager is not sure what
+%%%   happened. In a scenario like this, topic analysis can help tease out
+%%%   the topics of development that the developers were focussed on. Topic
+%%%   analysis could highlight the orthogonal issues that developers were
+%%%   spending time on.  Topic analysis extracts groups of words from
+%%%   documents (commit log comments) that are word distributions, topics,
+%%%   which characterize clusters of documents. These extracted topics often
+%%%   correlate with actual development topics that the developers
+%%%   discussed. By applying topic analysis locally to an iteration, both
+%%%   the developers and managers might be able to see what were the topics
+%%%   of development during that iteration. %XXX RM during that iteration
+%%%   
+%%%   
+%%%   %XXX RM 
+%%%   Commits in source control systems (SCSs) can have multiple
+%%%   purposes. Thus we hope that our research will help uncover these
+%%%   multiple purposes that manifest themselves as topics interleaved
+%%%   within the commits.  Others have demonstrated that topic analysis
+%%%   techniques, such as Latent Semantic Indexing (LSI) and Latent
+%%%   Dirichlet Allocation (LDA), can aid in separating topics from these
+%%%   mixed comments. These topics often describe distinct ideas that are
+%%%   interleaved throughout the development. A topic represents both a word
+%%%   distribution and a group of commit log comments that are related to
+%%%   each other by their content.  In this case a topic is a set of tokens.
+%%%   
+%%%   
+%%%   % We suppose this could help us to segment the revisions on the time
+%%%   % line to discover parts of an iteration.
+%%%   
+%%%   Topic analysis can aid in partitioning a project's
+%%%   time-line into periods of developer focus. By breaking apart an
+%%%   iteration into sets of topics and trends (topics that recur), we may
+%%%   be able to recognize the underlying software development process from
+%%%   these commits. Alternatively, we can determine what particular
+%%%   maintenance task was being performed at a given time.
+%%%   
+%%%   We propose to extend existing topic analysis techniques by applying
+%%%   them to windows of documents over time. We hope that by
+%%%   looking %XX X ADDED
+%%%   at topic clusters during periods or windows over time we can identify
+%%%   ongoing areas of focus on a select group of topics as well reflect
+%%%   more intermittent topics. Figure \ref{fig:lda} illustrates the kind of
+%%%   visualization we want to automatically generate. We want to see topics
+%%%   that are unique to a certain time period, displayed across the time
+%%%   axis. Those topics that recur, or occur over a larger period are
+%%%   plotted continuously. In our example we have titled each topic with a
+%%%   single word drawn from itself.
+%%%   
+%%%   In this paper we explore how topics shift over time in the SCS of a
+%%%   project, using several open source database systems as examples. We
+%%%   analyze commit log comments in each time window and we extract the
+%%%   topics of that time window. We expect that topics will change over
+%%%   time and the similarities and differences in these topics will
+%%%   indicate developer focus and changes in developer focus as it changes
+%%%   over time.
+%%%   
+%%%   Our contributions include:
+%%%   \begin{itemize}
+%%%   \item Demonstrating the value of windowed topic analysis using trends.
+%%%   \item Multiple visualizations of topics and trends over time.
+%%%   \item An exploratory case study using these techniques on several open source database systems.
+%%%   \end{itemize}
+%%%   
 \shrinkit

### 2009-06-16 lda-paper.tex
latest changes - intro and stuff

@@
 these topics will indicate developer focus and changes in developer
-focus as it changes over time.
+focus as it changes over time. We will show that most topics are
+locally relevant and thus a windowed analysis can allow for both
+different kinds of global and local analysis than global topic
+analysis would.
 
@@
 \begin{itemize}
-\item Demonstrating that most topics are local in scope, thus would be
-  missed by a global topic analysis.
+\item Demonstrating that windowed topic analysis most topics are local
+  in scope, thus would be missed by a global topic analysis.
 \item Multiple visualizations of topics and trends over time to aid
   communication and analysis of topics and trends.
-\item An exploratory case study using these techniques on several open source database systems.
+\item An exploratory case study using these techniques on several open
+  source database systems.
 \end{itemize}
@@
 extracted from SCS, Linstead et al.\  proposed an author-source-code model
-using LDA~\cite{10.1109/MSR.2007.20}%,NIPS2007637,1321709}. 
+using LDA~\cite{10.1109/MSR.2007.20}.%,NIPS2007637,1321709}. 
 This model
@@
 %location~\cite{1421013,1374321,10.1109/ICPC.2007.13,10.1109/ICPC.2006.17}.
-location~\cite{1374321,10.1109/ICPC.2007.13}
+location~\cite{1374321,10.1109/ICPC.2007.13}.
 Concept analysis aims to automatically extract concepts from source
@@
 into windows, and then each window is subject to topic analysis and
-then we analyze and visualize the results.
+then we analyze and visualize the results. Figure \ref{fig:commits}
+depicts the general process for processing and analyzing the commit
+messages.
 
@@
 involves significant computation, and we perform LDA once per window.
+Figure \ref{fig:commits} shows how LDA is applied to a set of
+messages, and how the topics are extracted and related to the
+messages.
 
@@
 given a certain threshold of top 10 word similarity, such as 8 out of
-top 10 matching words. Then given this topic similarity matrix, we
-find the transitive closure of similar topics; that is, if we model
-topics are nodes and arcs as similarity, we fill flood along the
+top 10 matching words. We then apply the transitive closure on similar
+topics to this topic similarity matrix; this is similar to modelling
+topics as nodes and similarity as arcs, then fill flooding along the
 similarity arcs until we have partitioned the topics into clusters of
@@
     plotted as continuous blocks. The top 10 words in the topics are
-    joined and embedded in box representing the topics.}
+    joined and embedded in box representing the topics. The gap in the
+    middle is a period where no development took place.}
   \label{fig:topicsmear}
@@
 trend and thus indicates which trends occur the most. Unfortunately
-due to the large number of topics, given the allotted space, it is often
-best to crop off the trends with only one topic, otherwise the tail is
-long.
+due to the large number of topics (approximately $N$ topics multiplied
+by $M$ periods), given the allotted space, it is often best to crop
+off the trends with only one topic ($1\%$ to $10\%$ of the total
+topics), otherwise the tail is long.
 
@@
 We tracked Firebird from August 2000 to January 2006. We found that with a
-similarity of $7/10$ that Firebird had far more continuous and
+similarity of $7/10$ Firebird had far more continuous and
 recurring trends than PostgreSQL.  The first large trend was segmented
@@
 The plots we produced of MaxDB 7.500 were unlike those of the other
-systems, as there was a period where no real development occurred and
+systems, as there was a period where no development occurred and
 thus there were no topics or trends whatsoever (see the gap in Figure
@@
 compared this against our windowed topic analysis.  To produce
-figure~\ref{fig:statictopics}, we extracted 20 topics and plotted the
-number of messages per month that were related to that topic. We found
-was that one topic would often dominate the results, as shown in the
-third row of Figure \ref{fig:statictopics}, while messages related to
-other topics did not appear often.
+Figure~\ref{fig:statictopics}, we extracted 20 topics and plotted the
+number of messages per month that were related to that topic.  One
+topic would often dominate the results, as shown in the third row of
+Figure \ref{fig:statictopics}, while the other topics did not appear
+as often.
+
 % This technique seems useful if the 20 extracted topics are relevant
@@
 
-The value of windowed topic analysis was demonstrated on a case study
-of three database systems with a particular focus on MaxDB 7.500. We
-showed that global topic analysis missed important topics such as
-\emph{ODBC}.
+Windowed topic analysis demonstrated its value to hilight local topics
+and indentify global trends. This was shown in our case study of MaxDB
+7.500. Global topic analysis missed important topics such as
+\emph{ODBC} while windowed topic analysis identified them.
 

### 2009-06-17 lda-paper.tex
more edits and latex8 updates

@@
 \usepackage{latex8}
-
-\usepackage{graphicx,amsmath,comment,fullpage}
+\usepackage{times}
+\usepackage{graphicx,amsmath,comment}
 \newcommand{\igH}[1]{\includegraphics[height=.9\textheight]{#1}}
@@
 \newcommand{\lda}{Latent Dirichlet Allocation}
+\pagestyle{empty}
 
@@
 \maketitle
+\thispagestyle{empty}
+
 \begin{abstract}
@@
 
-\section{Introduction}
-
-We care about the topics of development. We want to know given a
-certain period of time, what were the popular topics being worked on
+\Section{Introduction}
+
+We want to know about the topics of development that programmers
+encountered while developing a software project. We want to know given
+a certain period of time, what were the popular topics being worked on
 during that period. We also want to know if any of those topics
-occurred again, whether before or after our current period.
-
-These topics could help stakeholders determine what their coworkers
-were working on, what were common issues, what requirements were being
+occurred again, whether before or after our current period.  These
+topics could help stakeholders determine what their coworkers were
+working on, what were common issues, what requirements were being
 worked on, etc.
@@
 Topic analysis extracts groups of words from documents (commit log
-comments) that are word distributions, topics, which characterize
-clusters of documents. These extracted topics often correlate with
-actual development topics that the developers discussed.
+comments) that are word distributions, topics, that are combined to
+characterize clusters of documents. These extracted topics often
+correlate with actual development topics that the developers discussed
+during the development of the software project.
 
 Topic analysis often allows for a single document, such as a commit
-message, to be related to multiple topics. This maps well to
-commits which often have more than one purpose.  A topic represents
-both a word distribution and a group of commit log comments that are
-related to each other by their content.  In this case a topic is a set
-of tokens.
+message, to be related to multiple topics. This maps well to commits
+to source code, which often have more than one purpose.  A topic
+represents both a word distribution and a group of commit log comments
+that are related to each other by their content.  In this case a topic
+is a set of tokens extracted from commit messages found within a
+projects source control system (SCS).
 
@@
 recur, or occur over a larger period are plotted continuously. In our
-example we have titled each topic with a single word drawn from
-itself.
+example we have titled each topic with a single word drawn from its word
+distribution.
 
@@
 focus as it changes over time. We will show that most topics are
-locally relevant and thus a windowed analysis can allow for both
-different kinds of global and local analysis than global topic
-analysis would.
+locally relevant and thus a windowed topic analysis (local topic
+analysis of windows of messages) can allow for both different kinds of
+global and local analysis than global topic analysis would.
 
@@
 \begin{itemize}
-\item Demonstrating that windowed topic analysis most topics are local
+\item We demonstrate that windowed topic analysis most topics are local
   in scope, thus would be missed by a global topic analysis.
-\item Multiple visualizations of topics and trends over time to aid
+\item We present multiple visualizations of topics and trends over time to aid
   communication and analysis of topics and trends.
-\item An exploratory case study using these techniques on several open
-  source database systems.
+\item We provide an exploratory case study using these techniques on
+  several open source database systems.
 \end{itemize}
@@
 %%%   \shrinkit
-%%%   \section{Introduction}
+%%%   \Section{Introduction}
 %%%   \shrinkit
@@
 \shrinkit
-\section{Background}
+\Section{Background}
 \shrinkit
@@
 
-%\subsection{Intro to LDA}
+%\SubSection{Intro to LDA}
 
@@
 \shrinkit
-\subsection{LDA, LSI and Semantic Clustering}
+\SubSection{LDA, LSI and Semantic Clustering}
 \shrinkit
@@
 \shrinkit
-\section{Preliminary Case Study}
+\Section{Preliminary Case Study}
 \shrinkit
@@
 were common words across topic clusters, such as \emph{diffs},
-\emph{annotate} and \emph{history}; these words probably should be
-viewed as stop words. There were notable transitional topics such as
-in the first window the word \emph{BitKeeper} appears because MySQL
-adopted Bitkeeper for their source control system yet in the following
-windows there were no mentions of BitKeeper. \emph{RENAME} also only
-appeared once in the first window and never again. Per each topic we
-tried to identify the purpose of the topic; to our surprise we found
-that even with only a little familiarity with the code base that
-naming the topic was straightforward.
+\emph{annotate} and \emph{history}; since these words occur so often
+in the MySQL 3.23 messages, we probably should be viewed as stop
+words. There were notable transitional topics such as in the first
+window the word \emph{BitKeeper} appears because MySQL adopted
+Bitkeeper for their source control system yet in the following windows
+there were no mentions of BitKeeper. \emph{RENAME} also only appeared
+once in the first window and never again. Per each topic we tried to
+identify the purpose of the topic; to our surprise we found that even
+with only a little familiarity with the code base that naming the
+topic was straightforward.
 
@@
 \shrinkit
-\section{Methodology}
+\Section{Methodology}
 \shrinkit
@@
 \shrinkit
-\subsection{Extraction of  Repositories and Documents}
+\SubSection{Extraction of  Repositories and Documents}
 \shrinkit
@@
 \shrinkit
-\subsection{Windowed Sets}
+\SubSection{Windowed Sets}
 \shrinkit
@@
 \shrinkit
-\subsection{Apply Topic Analysis to each Window}
+\SubSection{Apply Topic Analysis to each Window}
 \shrinkit
@@
 
-%\subsection{Extract Clusters}
+%\SubSection{Extract Clusters}
 
@@
 \shrinkit
-\subsection{Topic Similarity}
+\SubSection{Topic Similarity}
 \shrinkit
@@
 \shrinkit
-\subsection{Visualization}
+\SubSection{Visualization}
 \shrinkit
@@
 \shrinkit
-\section{Results}
+\Section{Results}
 \shrinkit
@@
 
-%\subsection{MySQL}
+%\SubSection{MySQL}
 %XXX DO MYSQL?
 \shrinkit
-\subsection{PostgreSQL}
+\SubSection{PostgreSQL}
 \shrinkit
@@
 \shrinkit
-\subsection{Firebird}
+\SubSection{Firebird}
 \shrinkit
@@
 \shrinkit
-\subsection{MaxDB 7.500}
+\SubSection{MaxDB 7.500}
 \shrinkit
@@
 words and then two of the most common words, the global analysis
-finally had a topic with optimizer in its topic ten words. Our analysis shows
-that optimizer was important but it had been obscured, but would have
-been noticed using the more local topic analysis.
+finally had a topic with optimizer in its topic ten words. Our
+analysis shows that optimizer was important but it had been obscured
+by the global topic analysis, which used the entire history of
+messages, but would have been noticed using the more local topic
+analysis such as our windowed topic analysis, which used a window of
+messages.
 
@@
 \shrinkit
-\subsection{Compare with topics over the entire development}
+\SubSection{Compare with topics over the entire development}
 \shrinkit
@@
 
-%\subsection{Validation}
+%\SubSection{Validation}
 
@@
 \shrinkit
-\section{Validity Threats}
+\Section{Validity Threats}
 \shrinkit
@@
 \shrinkit
-\section{Conclusions}
+\Section{Conclusions}
 \shrinkit
@@
 \shrinkit
-\subsection{ Future Work}
+\SubSection{ Future Work}
 \shrinkit
@@
 
-% \section{Appendix}
+% \Section{Appendix}
 

### 2009-06-18 lda-paper.tex
- [X] Add an LDA example (Dogs/cats - but with developers)     - [X] Make it clear what we mean when we do "local topic analysis"           "Our analysis shows that optimizer was important but it had           been obscured, but would have been noticed using the more           local topic analysis." What does local topic mean? A smaller           window?           - In the introduction we try to clarify this and in MaxDB 7.500     - [X] Why were certain values chosen? • Sometimes some choices are not so clear/well explained: one month as          the length of time windows, 20 topics, 10 words

@@
 %\alignauthor
-Abram Hindle\\
+Abram Hindle and Michael W. Godfrey and Richard C. Holt\\
 %\affaddr{David Cheriton School of Computer Science}\\
@@
 \affaddr{Canada}\\
-\aemail{ahindle@cs.uwaterloo.ca}
+\aemail{\{ahindle,migod,holt\}@uwaterloo.ca}
 %\alignauthor
-\and
-Michael W. Godfrey\\
-%\affaddr{David Cheriton School of Computer Science}\\
-\affaddr{University of Waterloo}\\
-\affaddr{Waterloo, Ontario}\\
-\affaddr{Canada}\\
-\aemail{migod@cs.uwaterloo.ca}
-%\alignauthor
-\and
-Richard C. Holt\\
-%\affaddr{David Cheriton School of Computer Science}\\
-\affaddr{University of Waterloo}\\
-\affaddr{Waterloo, Ontario}\\
-\affaddr{Canada}\\
-\aemail{holt@cs.uwaterloo.ca}
+% \and
+% Michael W. Godfrey\\
+% %\affaddr{David Cheriton School of Computer Science}\\
+% \affaddr{University of Waterloo}\\
+% \affaddr{Waterloo, Ontario}\\
+% \affaddr{Canada}\\
+% \aemail{migod@cs.uwaterloo.ca}
+% %\alignauthor
+% \and
+% Richard C. Holt\\
+% %\affaddr{David Cheriton School of Computer Science}\\
+% \affaddr{University of Waterloo}\\
+% \affaddr{Waterloo, Ontario}\\
+% \affaddr{Canada}\\
+% ;\aemail{holt@cs.uwaterloo.ca}
 %\alignauthor
@@
   features.  Tools such as Latent Dirichlet Allocation (LDA) and
-  Latent Semantic Indexing (LSI) can be used to analyze check-in
+  Latent Semantic Indexing (LSI) can be used to analyze commit log
   comments over the entire lifetime of a project. Previous work on
@@
 
+What happened in the previous maintenance or development iteration?
+What were the developers working on? What was the developer in the
+next cubicle working on? Which common topics and issues were dealt
+with in the current release of the software? What were the topics of
+the previous release? Given the requirements agreed to at the start
+of the iteration did our developers work on them?  What else did they
+work on?
+
 We want to know about the topics of development that programmers
@@
 working on, what were common issues, what requirements were being
-worked on, etc.
+worked on, etc. In order to extract and evaluate these topics we plan
+to rely on topic analysis.
 
@@
 
+% An example use of LDA
+% * take a bunch of commit messages
+% ** occurred during a period with bug fixing, adding a GUI and documentation changes
+% ** LDA would find that bug fixing topic was associated to the words bug, fix, failure, ticket
+% *** GUI - widget, panel, button, menu, window
+% *** Documentation - chapter, section, figure and words shared between GUI and bug fixing
+% * a particular commit would be associated with some of these topics 
+% ** a commit about addressing a fix to a bug in a ticket would be heavily associated with the bug topic
+% ** a commit that deals with documenting the GUI associated with GUI and Documentation
+% ** a commit that was independant of all three could've induced a seperate topic altogether
+
+Imagine that we are analyzing one month of development in a project
+where the main focus was on bug fixing, adding a new GUI and
+documentation updates relating to the system and the GUI. To use LDA
+we'd get the word distributions of each commit messages, pass them to
+LDA and tell LDA to find some topics, in this case we said 4. The 4
+topics it discovered would probably be bug fixing, GUI, documentation
+and other. These topics are essentially word distributions, for
+example the bug fixing topic would include words like \emph{bug},
+\emph{fix}, \emph{failure} and \emph{ticket}. The GUI topic would
+include words like \emph{widget}, \emph{panel}, \emph{window}, and
+\emph{menu}. The documentation topic would include words like
+\emph{section}, \emph{figure}, and \emph{chapter}. The other topic
+would include words found in topics generally independant of the other
+topics. If we analyzed a commit of a bug fix it'd probably contain
+words like \emph{ticket} and thus heavily associate with the bug fix
+topic. A commit that was about user documentation regarding the GUI
+would be associated with the documentation topic and the GUI topic
+because it shared words from both topics. Where as a topic dealing
+neither bugs, GUIs or documentation might be associated with the other
+topic. This example demonstrates how LDA can associate 1 document (a
+commit message) with more than one topic, how topics are related to
+documents via their word distributions. LDA also would discover these
+independant topics in an unsupervised manner, all the end user has to
+do is name the topic based upon the most frequent words in that topic,
+which is a word distribution.
+
 %      1. [ ] LDA itself
@@
 windows. We asked LDA to make 20 topics per window and we then
-examined the top 10 most frequent words in that topic.  We found there
+examined the top 10 most frequent words in that topic.  We chose one
+month because it was smaller than the time between minor releases but
+large enough for there to be many commits to analyze. We chose 20
+topics because past experimentation showed that few topics might
+aggregate multiple unique topics while anymore topics seemed to dilute
+the results and create some indistinct topics. 10 most frequent words
+were chosen as sometimes 1 common word dominated topics so we wanted
+to have enough words to distingush topics.
+
+We found there
 were common words across topic clusters, such as \emph{diffs},
@@
 \end{tabular}
-\caption{Sampled topics from MySQL 3.23, some with continuous topics. These tokens were pulled from LDA topic. Each token is a summary of one LDA generated topic from MySQL 3.23 commit comments.}
+\caption{Sampled topics from MySQL 3.23, some with continuous topics.
+These tokens were pulled from the top 10 most common words found in LDA extracted topics.
+Each token is a summary of one LDA generated topic from MySQL 3.23 commit comments.}
 \label{tab:portability}
@@
 development, which is large enough to show shifts in focus, but coarse
-enough not to show too much detail.
+enough not to show too much detail. Choosing a month as the window
+provided us with enough documents to analyze. One could overlap the
+windows, which would increase the likelihood of trends, but for this
+study we lacked the space and were more concerned if topics ever
+repeated, overlapping might skew that result.
 
@@
 each window and extract the top $N$ topics. We used $20$ topics in our
-case studies. Our Topic Analysis tool is an implementation of LDA, but
+case studies because we experimented before and that $20$ was about
+all we can handle before too many topics seemed too indistinct from
+each other to differeniate.
+
+ Our Topic Analysis tool is an implementation of LDA, but
 we could have used other similar tools like LSI, but previous work
@@
 given a certain threshold of top 10 word similarity, such as 8 out of
-top 10 matching words. We then apply the transitive closure on similar
-topics to this topic similarity matrix; this is similar to modelling
-topics as nodes and similarity as arcs, then fill flooding along the
-similarity arcs until we have partitioned the topics into clusters of
-similar topics. Figure \ref{fig:closure} illustrates clustering of
-topics by topic similarity. These clusters of topics are called
-trends. A trend is probably interesting if it contains more than one
-topic.
+top 10 matching words.  10 words were chosen because often people care
+about top 10 comparison and it allowed for some common words to exist
+but not cause topics to match so closely.  We then apply the
+transitive closure on similar topics to this topic similarity matrix;
+this is similar to modelling topics as nodes and similarity as arcs,
+then fill flooding along the similarity arcs until we have partitioned
+the topics into clusters of similar topics. Figure \ref{fig:closure}
+illustrates clustering of topics by topic similarity. These clusters
+of topics are called trends. A trend is probably interesting if it
+contains more than one topic.
 
@@
 When we examined PostgreSQL, we did not find many trends with two or
-more topics, using a similarity of $7/10$. Those trends that we did find
+more topics, using a similarity of $7/10$. $7/10$ was chosen because
+it preserved indepedant topics but seemed to be a borderline threshold
+value where more serious trends started to appear.
+
+Those trends that we did find
 were not very large, lasting only 3 months at most. The first and second
@@
 
-One avenue of future work is automatic topic labelling. Given a word
+We want to explore parameters more across multiple projects, what
+motivates the need for more or less topics, can we find a good
+estimate of a good choice of number of topics to evaluate. We want to
+investigate the effects of overlapping windows on this analysis, do
+they make the trends too long? Do they highlight some trends
+unnecessarily or are they more because they are time agnostic.
+
+Another avenue of future work is automatic topic labelling. Given a word
 distribution we should be able to automatically select a word or term

### 2009-06-19 lda-paper.tex
- [X] Address how we found the purpose of commits You say that you tries to identify the purpose of the topic. How did you do it? Was it by looking at the commit messages, or just by reading the list of words? Did you check your assumptions?

@@
 
-We found there
-were common words across topic clusters, such as \emph{diffs},
-\emph{annotate} and \emph{history}; since these words occur so often
-in the MySQL 3.23 messages, we probably should be viewed as stop
-words. There were notable transitional topics such as in the first
-window the word \emph{BitKeeper} appears because MySQL adopted
-Bitkeeper for their source control system yet in the following windows
-there were no mentions of BitKeeper. \emph{RENAME} also only appeared
-once in the first window and never again. Per each topic we tried to
-identify the purpose of the topic; to our surprise we found that even
-with only a little familiarity with the code base that naming the
-topic was straightforward.
+We found there were common words across topic clusters, such as
+\emph{diffs}, \emph{annotate} and \emph{history}; since these words
+occur so often in the MySQL 3.23 messages, we probably should be
+viewed as stop words. There were notable transitional topics such as
+in the first window the word \emph{BitKeeper} appears because MySQL
+adopted Bitkeeper for their source control system yet in the following
+windows there were no mentions of BitKeeper. \emph{RENAME} also only
+appeared once in the first window and never again. Per each topic we
+tried to identify the purpose of the topic; to our surprise we found
+that even with only a little familiarity with the code base that
+naming the topic was straightforward. To find the purpose of a commit
+we looked at the most frequent words in the word distribution of the
+topic and tried to summarize the topic, then we looked into the
+commits related to that topic to investigate if we were correct; since
+the commit messages and the word distribution share the same words the
+purpose extracted from commit comments was usually accurate.
 

### 2009-06-19 lda-paper.tex
- [X] Address this comment: You say that you remove stop words, but in Figure 4 I noticed words like "to" or "on". These are stop words.
- [X] Specify that if in 1 CASE we removed these words we got different results: You state that words like "diffs, annotate, history" should be treated as stop words. Why? Is it because they do not denote a domain concept? What if the domain actually include these words, are they still not relevant?  Dealt with in the case study

@@
 \emph{diffs}, \emph{annotate} and \emph{history}; since these words
-occur so often in the MySQL 3.23 messages, we probably should be
-viewed as stop words. There were notable transitional topics such as
-in the first window the word \emph{BitKeeper} appears because MySQL
-adopted Bitkeeper for their source control system yet in the following
-windows there were no mentions of BitKeeper. \emph{RENAME} also only
-appeared once in the first window and never again. Per each topic we
-tried to identify the purpose of the topic; to our surprise we found
-that even with only a little familiarity with the code base that
-naming the topic was straightforward. To find the purpose of a commit
-we looked at the most frequent words in the word distribution of the
-topic and tried to summarize the topic, then we looked into the
-commits related to that topic to investigate if we were correct; since
-the commit messages and the word distribution share the same words the
-purpose extracted from commit comments was usually accurate.
+occur so often in the MySQL 3.23 messages. These words probably should
+be viewed as stop words since occur across so many topics, even though
+they are related to version control usage.  There were notable
+transitional topics such as in the first window the word
+\emph{BitKeeper} appears because MySQL adopted Bitkeeper for their
+source control system yet in the following windows there were no
+mentions of BitKeeper. \emph{RENAME} also only appeared once in the
+first window and never again. Per each topic we tried to identify the
+purpose of the topic; to our surprise we found that even with only a
+little familiarity with the code base that naming the topic was
+straightforward. To find the purpose of a commit we looked at the most
+frequent words in the word distribution of the topic and tried to
+summarize the topic, then we looked into the commits related to that
+topic to investigate if we were correct; since the commit messages and
+the word distribution share the same words the purpose extracted from
+commit comments was usually accurate.
 
@@
     plotted as continuous blocks. The top 10 words in the topics are
-    joined and embedded in box representing the topics. The gap in the
+    joined and embedded in box representing the topics. No stop words were removed. The gap in the
     middle is a period where no development took place.}

### 2009-06-19 lda-paper.tex
more lda - paper changes    - [X] Add references?  • Some references are lacking; for example in “We mirrored the repositories and their revisions using software such as rsync, CVSSuck, softChange and bt2csv” (page 3) - [X]  Add more refernces "Others have demonstrated ..." -> some references would help - Count - [X] General case study metrics • Maybe some information regarding the number of analyzed commit per case studies could be useful to understand a little bit better the experiemnts.  - updated per each case study

@@
 
-Usually topic analysis is applied to the entire history of a project,
+Usually topic analysis is applied to the entire history of a project~\cite{1321709,lukins2008,10.1109/ICPC.2007.13},
 where as we suggest that more topics are probably local, if a
@@
 
-Grant et al.~\cite{scottcordy} have used an alternative technique,
-called Independent Component Analysis to separate topic signals from
+Grant et al.~\cite{scottcordy} and have used an alternative technique,
+called Independent Component Analysis~\cite{delac} to separate topic signals from
 software code.
@@
 We mirrored the repositories and their revisions using software such
-as rsync, CVSSuck, \texttt{softChange} and bt2csv.
+as rsync~\cite{rsync}, CVSsuck~\cite{cvssuck},
+\texttt{softChange}~\cite{dmgseke2004}, and bt2csv~\cite{methodology}.
 \texttt{softChange} provided a general schema for storing revisions
-and grouped CVS revisions into commits. CVSSuck and rsync mirrored CVS
+and grouped CVS revisions into commits. CVSsuck and rsync mirrored CVS
 repositories while bt2csv mirrored web accessible BitKeeper
@@
 \ref{fig:histogram} where we plot each trend on its own line but stack
-up the segments of the trend, much like a histogram.
+up the segments of the trend, much like a histogram. Each topic has its top 10 words 
+listed in descending order of frequency from top to bottom, this text is embedded inside the topic's box.
+A trend has all of its topic text embedded side by side within the same trend box.
 
@@
 We applied our tools and visualizations to the repositories of several
-open source database systems: PostgreSQL, MaxDB and Firebird.
+open source database systems: PostgreSQL, MaxDB and Firebird. The
+total number of commits analyzed was over $66 000$.
 
@@
 
-When we examined PostgreSQL, we did not find many trends with two or
-more topics, using a similarity of $7/10$. $7/10$ was chosen because
-it preserved indepedant topics but seemed to be a borderline threshold
+When we examined PostgreSQL's history from 1996 to 2004, which include
+over $20 000$ commits. We did not find many trends with two or more
+topics, using a similarity of $7/10$. $7/10$ was chosen because it
+preserved indepedant topics but seemed to be a borderline threshold
 value where more serious trends started to appear.
@@
 
-We tracked Firebird from August 2000 to January 2006. We found that with a
+We tracked Firebird from August 2000 to January 2006, we extracted comments from $38 000$ commits. We found that with a
 similarity of $7/10$ Firebird had far more continuous and
@@
 The plots we produced of MaxDB 7.500 were unlike those of the other
-systems, as there was a period where no development occurred and
-thus there were no topics or trends whatsoever (see the gap in Figure
+systems, as there was a period where no development occurred and thus
+there were no topics or trends whatsoever (see the gap in Figure
 \ref{fig:topicsmear}). Using a topic similarity of $7/10$ we evaluated
 MaxDB 7.500. MaxDB 7.500's first period was from June 2004 to January
-2005, and its second period was from June 2005 to June 2006.
+2005, and its second period was from June 2005 to June 2006. There
+were a total of $8 600$ commits analyzed.
 

### 2009-06-19 lda-paper.tex
hilighted what changed

@@
 \pagestyle{empty}
+
+\usepackage[usenames]{color}
+\definecolor{changecolor1}{rgb}{0.5,0.3,0.08} 
+\definecolor{changecolor2}{rgb}{0.3,0.08,0.45} 
+\definecolor{changecolor3}{rgb}{0.3,0.45,0.45} 
+\definecolor{changecolor4}{rgb}{0.0,0.3,0.0} 
+\newcommand{\hl}[1]{{\color{changecolor1} #1 }}
+\newcommand{\hla}[1]{{\color{changecolor2} #1 }}
+\newcommand{\hlc}[1]{{\color{changecolor3} #1 }}
+\newcommand{\hld}[1]{{\color{changecolor4} #1 }}
 
@@
   features.  Tools such as Latent Dirichlet Allocation (LDA) and
+  \hla{
     Latent Semantic Indexing (LSI) can be used to analyze commit log
+  }
   comments over the entire lifetime of a project. Previous work on
@@
 
+\hla{
 What happened in the previous maintenance or development iteration?
@@
 work on?
-
+}
+
+\hla{
 We want to know about the topics of development that programmers
@@
 to rely on topic analysis.
+}
 
@@
 Topic analysis often allows for a single document, such as a commit
+\hla{
 message, to be related to multiple topics. This maps well to commits
@@
 projects source control system (SCS).
-
+}
+
+\hla{
 Usually topic analysis is applied to the entire history of a project~\cite{1321709,lukins2008,10.1109/ICPC.2007.13},
+}
 where as we suggest that more topics are probably local, if a
@@
 focus as it changes over time. We will show that most topics are
+\hla{
 locally relevant and thus a windowed topic analysis (local topic
@@
 global and local analysis than global topic analysis would.
-
+}
+
+\hla{
 Our contributions include:
@@
 \end{itemize}
+}
 
@@
 Grant et al.~\cite{scottcordy} and have used an alternative technique,
+\hla{
 called Independent Component Analysis~\cite{delac} to separate topic signals from
+}
 software code.
@@
 
+\hla{
 Imagine that we are analyzing one month of development in a project
@@
 which is a word distribution.
+}
 
@@
 windows. We asked LDA to make 20 topics per window and we then
+\hla{
 examined the top 10 most frequent words in that topic.  We chose one
@@
 commit comments was usually accurate.
+}
 
@@
 \end{tabular}
-\caption{Sampled topics from MySQL 3.23, some with continuous topics.
+\caption{
+\hla{
+Sampled topics from MySQL 3.23, some with continuous topics.
 These tokens were pulled from the top 10 most common words found in LDA extracted topics.
-Each token is a summary of one LDA generated topic from MySQL 3.23 commit comments.}
+Each token is a summary of one LDA generated topic from MySQL 3.23 commit comments.
+}
+}
 \label{tab:portability}
@@
 
+\hla{
 We mirrored the repositories and their revisions using software such
@@
 repositories.
+}
 
@@
 development, which is large enough to show shifts in focus, but coarse
+\hla{
 enough not to show too much detail. Choosing a month as the window
@@
 repeated, overlapping might skew that result.
+}
 
@@
 each window and extract the top $N$ topics. We used $20$ topics in our
+\hla{
 case studies because we experimented before and that $20$ was about
@@
  Our Topic Analysis tool is an implementation of LDA, but
+}
 we could have used other similar tools like LSI, but previous work
@@
 given a certain threshold of top 10 word similarity, such as 8 out of
+\hla{
 top 10 matching words.  10 words were chosen because often people care
@@
 contains more than one topic.
+}
 
@@
 \ref{fig:histogram} where we plot each trend on its own line but stack
+\hla{
 up the segments of the trend, much like a histogram. Each topic has its top 10 words 
@@
 A trend has all of its topic text embedded side by side within the same trend box.
+}
 
@@
 We applied our tools and visualizations to the repositories of several
+\hla{
 open source database systems: PostgreSQL, MaxDB and Firebird. The
 total number of commits analyzed was over $66 000$.
+}
 
@@
 
+\hla{
 We tracked Firebird from August 2000 to January 2006, we extracted comments from $38 000$ commits. We found that with a
+}
 similarity of $7/10$ Firebird had far more continuous and
@@
 The plots we produced of MaxDB 7.500 were unlike those of the other
+\hla{
 systems, as there was a period where no development occurred and thus
@@
 were a total of $8 600$ commits analyzed.
+}
 
@@
 words and then two of the most common words, the global analysis
+\hla{
 finally had a topic with optimizer in its topic ten words. Our
@@
 messages.
+}
 
@@
 
+\hla{
 We want to explore parameters more across multiple projects, what
@@
 unnecessarily or are they more because they are time agnostic.
+}
 

### 2009-06-22 lda-paper.tex
more edits

@@
 
-\hla{
-What happened in the previous maintenance or development iteration?
-What were the developers working on? What was the developer in the
-next cubicle working on? Which common topics and issues were dealt
-with in the current release of the software? What were the topics of
-the previous release? Given the requirements agreed to at the start
-of the iteration did our developers work on them?  What else did they
-work on?
-}
+\hla{ The questions that we wish to address deal with the past and
+  topics of development that occurred in the past.  What happened in
+  the previous iteration of a project's development?  What were the
+  developers working on? What was the developer in the next cubicle
+  working on? Which common topics and issues were dealt with in the
+  current release of the software? What were the topics of the
+  previous release? Given the requirements agreed to at the start of
+  the iteration did our developers work on them?  What else did they
+  work on?  }
 
@@
 during the development of the software project. 
-
 Topic analysis often allows for a single document, such as a commit
@@
 recur, or occur over a larger period are plotted continuously. In our
-example we have titled each topic with a single word drawn from its word
-distribution.
+example we have titled each topic with a single word drawn from its
+word distribution.  Given a visualization, such as this, topic
+analysis can aid in partitioning a project's time-line into periods of
+developer focus. By breaking apart an iteration into sets of topics
+and trends (topics that recur), we may be able to recognize the
+underlying software development process from these
+commits. Alternatively, we can determine what particular maintenance
+task was being performed at a given time.
 

### 2009-06-22 lda-paper.tex
- Motivate why we bother to visualize - [X] Motivate why the visualizations are important - [X] A. state why we visualize, B. state the goal of each visualization In section 4.5 you enumerate several visualizations, but it is again unclear what the goal is. Why do you need these?  What do you intend to detect?  - Clean up intro - [X] intro needs focus on what we did - [X] Sharpen Intro - [X] Make the goal and the achievements more explicit

--- 2009-06-19 lda-paper.tex
@@
 
-\author{Abram Hindle and Michael Godfrey and Ric Holt \\
+\author{Abram Hindle, Michael Godfrey and Ric Holt \\
 \{ahindle,migod,holt\}@cs.uwaterloo.ca\\
@@
 
+\hlb{
+We want to see topics that are unique to a
+certain time period, displayed across the time axis.
 Figure \ref{fig:lda} illustrates the kind of visualization we want to
-automatically generate. We want to see topics that are unique to a
-certain time period, displayed across the time axis. Those topics that
+automatically generate. } Those topics that
 recur, or occur over a larger period are plotted continuously. In our
-example we have titled each topic with a single word drawn from its word
-distribution.
+example we have titled each topic with a single word drawn from its
+word distribution.  
+\hdb{
+Given a visualization, such as Figure \ref{fig:lda}, topic
+analysis can aid in partitioning a project's time-line into periods of
+developer focus.
+}
+By breaking apart an iteration into sets of topics
+and trends (topics that recur), we may be able to recognize the
+underlying software development process and maintenance tasks from these
+commits.
 
@@
 %       1. [ ] need the visualizer
+\hdb{ Visualization is an integral part of our topics analysis which
+  allows us to quickly explore the topics and trends of a project.
+  Visualization provides us with a framework that allows us to
+  visually answer many questions about: the spread of topics, how many
+  topics were independant per period, what the repeating trends were,
+  if the trends dominated or did local topics dominate, how continuous
+  trends were, and what the topic text in topic, trend, or period was.
+  Since we have multiple questions to answer we have multiple
+  visualization, described within this section, that help answer
+  different questions.  }
 
@@
 
-The \emph{compact-trend-view} (Figure \ref{fig:topicsmear}) tries to
-sink the larger trends to the bottom.  Once the larger trends are
-stacked we fill in the gaps with the smaller trends in the same
-window, and then stack the smaller trends on top.  Although there is a
-chance that gaps will be left due to non-optimal stacking, in practice
-there are lot of small trends (90\% of all trends contain 1 topic)
-that fill in these gaps quickly.  Different instances of the same
-trend share the same color; apart from that, the color of trends is
-randomly chosen and color similarity is not meaningful.  The
-compact-trend-view is convenient because it shows all of the topic
-information, as shown by the zoomed-in view of MaxDB 7.500's
-compact-trend-view in Figure \ref{fig:zoomedsmear}. This view makes
-repeating continuous trends easy to pick out, although discontinuous
-trends are harder to spot.
-
-The \emph{trend-time-line} (Figure
-\ref{fig:trendtimeline}) displays repeating trends more clearly by
+\hlb{ The \emph{compact-trend-view} (Figure \ref{fig:topicsmear})
+  attempts to show all topics and trends at once across time in a
+  compact view that could fit on one page. It tries to give a compact
+  summary of the topics and trends across time.  } The
+\emph{compact-trend-view} (Figure \ref{fig:topicsmear}) tries to sink
+the larger trends to the bottom.  Once the larger trends are stacked
+we fill in the gaps with the smaller trends in the same window, and
+then stack the smaller trends on top.  Although there is a chance that
+gaps will be left due to non-optimal stacking, in practice there are
+lot of small trends (90\% of all trends contain 1 topic) that fill in
+these gaps quickly.  Different instances of the same trend share the
+same color; apart from that, the color of trends is randomly chosen
+and color similarity is not meaningful.  The compact-trend-view is
+convenient because it shows all of the topic information, as shown by
+the zoomed-in view of MaxDB 7.500's compact-trend-view in Figure
+\ref{fig:zoomedsmear}. The \emph{compact-trend-view} makes repeating
+continuous trends easy to pick out, although discontinuous trends are
+harder to spot, and provides a general summary of the topics within a project.
+
+\hdb{
+The \emph{trend-time-line} (Figure \ref{fig:trendtimeline})
+attempts to show a summary of trends seperated from single topics. This view shows how a trend persists across time, which aids time-wise analysis of trends.
+}
+The \emph{trend-time-line} displays repeating trends more clearly by
 dedicating a horizontal line for trend segments belonging to one
@@
 trends need to be pruned away or the view will be very long. 
-
-The \emph{trend-histogram} (Figure \ref{fig:histogram}) superficially
-resembles the trend-time-line.  However, in this view the trends are
-plotted together by stacking to the left of their row, thus time
-information is lost.  The trends are ordered by the number of topics
-in the trend.  The trend-histogram shows the count of instances of a
-trend and thus indicates which trends occur the most. Unfortunately
-due to the large number of topics (approximately $N$ topics multiplied
-by $M$ periods), given the allotted space, it is often best to crop
-off the trends with only one topic ($1\%$ to $10\%$ of the total
-topics), otherwise the tail is long.
+Thus the \emph{trend-time-line} view is used to analyze trends across time.
+
+\hdb{ The \emph{trend-histogram} (Figure \ref{fig:histogram}) attempts
+  to show a count of how often a trend reoccurs and how many topics
+  are related to a trend. It is meant to show the distribution of
+  trends by their size in topics and time-span.  } The
+\emph{trend-histogram} superficially resembles the trend-time-line.
+However, in this view the trends are plotted together by stacking to
+the left of their row, thus time information is lost.  The trends are
+ordered by the number of topics in the trend.  The trend-histogram
+shows the count of instances of a trend and thus indicates which
+trends occur the most. Unfortunately due to the large number of topics
+(approximately $N$ topics multiplied by $M$ periods), given the
+allotted space, it is often best to crop off the trends with only one
+topic ($1\%$ to $10\%$ of the total topics), otherwise the tail is
+long. \hdb{ The \emph{trend-histogram} summarizes the distribution of
+trends ordered by size. }
+
+\hdb{ All of these visualization combine to enable an analysis of
+  trends and topics. Some views like the \emph{compact-trend-view}
+  enable an analysis of local topics while the \emph{trend-histogram}
+  and \emph{trend-time-line} focus more on trends. We used these
+  visualizations to analyze the database systems that we discuss in
+  the following results section.  }
 

### 2009-06-23 lda-paper.tex
some ric changes to the LDA paper

@@
 
-  As development on a software project progresses, developers will
-  naturally shift their focus between different topics and tasks many
+  As development on a software project progresses, developers shift their focus between different topics and tasks many
   times.  Managers and newcomer developers often seek ways of
@@
   commit log comments with independent topics extracted from these
-  commit log comments.  In this work, we use LDA  to analyze
+  commit log comments.  In this paper, we use LDA  to analyze
   periods, such as months, within a project's lifetime to create a
@@
   that windowed topic analysis offers advantages over topic analysis
-  applied to a projects lifetime because many topics are quite local.
+  applied to a project's lifetime because many topics are quite local.
 
@@
 \hla{ The questions that we wish to address deal with the past and
-  topics of development that occurred in the past.  What happened in
+  topics development that occurred in the past.  What happened in
   the previous iteration of a project's development?  What were the
@@
 \hla{
-We want to know about the topics of development that programmers
+We want to know about the topics of that programmers
 encountered while developing a software project. We want to know given
@@
 
-Topic analysis extracts groups of words from documents (commit log
-comments) that are word distributions, topics, that are combined to
-characterize clusters of documents. These extracted topics often
+\hdb{
+Topic analysis extracts independant word distributions (topics) from
+documents (commit log comments).
+Ideally these extracted topics }
 correlate with actual development topics that the developers discussed
@@
 Topic analysis often allows for a single document, such as a commit
-\hla{
-message, to be related to multiple topics. This maps well to commits
+\hlb{
+message, to be related to multiple topics. Documents represented as a mixture of topics maps well to commits
+}
+\hla{
 to source code, which often have more than one purpose.  A topic
@@
 \hla{
-Usually topic analysis is applied to the entire history of a project~\cite{1321709,lukins2008,10.1109/ICPC.2007.13},
-}
-where as we suggest that more topics are probably local, if a
-programmer is working on a story card, they might never deal with that
+In the past, topic analysis has been applied to the entire history of a project~\cite{1321709,lukins2008,10.1109/ICPC.2007.13},
+}
+\hlb{
+where-as this paper suggests that most topics are local. For example,  
+programmer may be working on a story card,} they might never deal with that
 story card ever again. Thus that topic might be quite relevant to a
 small window of time but never relevant ever again.  By applying topic
-analysis locally to an iteration, both the developers and managers
+analysis locally to a window or an iteration, both the developers and managers
 might be able to see what were the topics of development during that
@@
 recur, or occur over a larger period are plotted continuously. In our
-example we have titled each topic with a single word drawn from its
+example we have titled each topic with a word drawn from its
 word distribution.  

### 2009-06-23 lda-paper.tex
more ric edits

@@
 focus as it changes over time. We will show that most topics are
-\hla{
-locally relevant and thus a windowed topic analysis (local topic
-analysis of windows of messages) can allow for both different kinds of
-global and local analysis than global topic analysis would.
+\hlb{
+locally relevant, i.e., relevant to a single window of time.  
+Our windowing approach supports both local (single window) analysis
+and global (entire history) analysis.
 }
@@
 \item We demonstrate that windowed topic analysis most topics are local
-  in scope, thus would be missed by a global topic analysis.
-\item We present multiple visualizations of topics and trends over time to aid
-  communication and analysis of topics and trends.
+  in scope, which would be missed by a global topic analysis.
+\item We present a number of visualizations of topics and their trends over time to aid
+  communication and analysis of these topics and trends.
 \item We provide an exploratory case study using these techniques on
@@
 distribution will be over the words in all messages. However, most
-words will not appear in a message, a word distribution is effectively
+words will not appear in each message. A word distribution is effectively
 a word count divided by the message size. A \emph{topic} is a word
-distribution, a set of words that form a word distribution that is
+distribution, i.e. a set of words that form a word distribution that is
 unique and independent within the set of documents in our total
@@
 model, a combination of topics.  LDA is used in software engineering
-literature~\cite{lukins2008,10.1109/MSR.2007.20}%,NIPS2007637,1321709}
- to extract topics from documents such as methods, bug
+literature~\cite{lukins2008,10.1109/MSR.2007.20}%,NIPS2007637,1321709}.
+LDA extracts topics from documents such as methods, bug
 reports, source code and files.
@@
 does not name these clusters but the clusters consist of words whose
-relationship is often obvious to someone familiar with the corpus. We
+relationship is often obvious to someone familiar with the corpus. For our purposes, we
 could swap LDA for LSI or semantic clustering and likely produce similar
@@
 documentation updates relating to the system and the GUI. To use LDA
-we'd get the word distributions of each commit messages, pass them to
-LDA and tell LDA to find some topics, in this case we said 4. The 4
-topics it discovered would probably be bug fixing, GUI, documentation
+we would get the word distributions of each commit messages, pass them to
+LDA and tell LDA to find some topics, in this case we
+\hlb{
+ in this case we chose to concentrate on four topics.
+} The four
+topics it discovered would ideally be bug fixing, GUI, documentation
 and other. These topics are essentially word distributions, for
@@
 would include words found in topics generally independant of the other
-topics. If we analyzed a commit of a bug fix it'd probably contain
+topics. If we analyzed a commit of a bug fix it would probably contain
 words like \emph{ticket} and thus heavily associate with the bug fix
@@
 neither bugs, GUIs or documentation might be associated with the other
-topic. This example demonstrates how LDA can associate 1 document (a
-commit message) with more than one topic, how topics are related to
+topic. This example demonstrates how LDA can associate one document (a
+commit message) with more than one topic, and how topics are related to
 documents via their word distributions. LDA also would discover these
-independant topics in an unsupervised manner, all the end user has to
+independant topics in an unsupervised manner. All the end user has to
 do is name the topic based upon the most frequent words in that topic,
@@
 large enough for there to be many commits to analyze. We chose 20
-topics because past experimentation showed that few topics might
-aggregate multiple unique topics while anymore topics seemed to dilute
-the results and create some indistinct topics. 10 most frequent words
-were chosen as sometimes 1 common word dominated topics so we wanted
-to have enough words to distingush topics.
+topics because past experimentation showed that fewer topics might
+aggregate multiple unique topics while any more topics seemed to dilute
+the results and create indistinct topics. 
+\hlb{
+  We chose the top ten most frequent words because sometimes a couple
+  of common words dominated topics so we wanted to have enough words
+  to distinguish topics.
+}
 
@@
 mentions of BitKeeper. \emph{RENAME} also only appeared once in the
-first window and never again. Per each topic we tried to identify the
+first window and never again. For each topic we tried to name the
 purpose of the topic; to our surprise we found that even with only a
@@
 similar topics into trends.  Figure \ref{fig:lda} displays a manually
-created sample plot of the extracted topics in Table
+created plot of the extracted topics in Table
 \ref{tab:portability}.
@@
 
-The documents we are analyzing are the commit log comments i.e., the
+The documents we are analyzing are the commit log comments, i.e. the
 comments that are added when revisions to the project are committed.
-Per each commit log comment, we count the occurrence of each word in
+For each commit log comment, we count the occurrence of each word in
 the message, remove stop words, and then produce word distributions
 for each message.  These distributions are then normalized by the size
-of the message. After all messages are processed the distributions are
+of the message, each count was divided by the total number of
+tokens. After all messages are processed the distributions are
 extended to include all words from all of the distributions.
@@
 overlapping windows, but in this paper we use non-overlapping windows
-of a set time unit because it simplifies analysis.  Windowing by time
+of a set time unit because it simplifies analysis.  
+One could overlap the
+windows, which would increase the likelihood of trends, but for this
+study we lacked the space and were more concerned if topics ever
+repeated, overlapping might skew that result.
+Windowing by time
 allows for many levels of granularity.  We used one month as the
@@
 enough not to show too much detail. Choosing a month as the window
-provided us with enough documents to analyze. One could overlap the
-windows, which would increase the likelihood of trends, but for this
-study we lacked the space and were more concerned if topics ever
-repeated, overlapping might skew that result.
+provided us with enough documents to analyze. 
 }

### 2009-06-24 lda-paper.tex
another change

@@
 \newcommand{\hla}[1]{{\color{changecolor2} #1 }}
-\newcommand{\hlc}[1]{{\color{changecolor3} #1 }}
-\newcommand{\hld}[1]{{\color{changecolor4} #1 }}
+\newcommand{\hlb}[1]{{\color{changecolor3} #1 }}
+\newcommand{\hlc}[1]{{\color{changecolor4} #1 }}
 
@@
 
-\hdb{
+\hla{
 Topic analysis extracts independant word distributions (topics) from
@@
 word distribution.  
-\hdb{
+\hla{
 Given a visualization, such as Figure \ref{fig:lda}, topic
@@
 %       1. [ ] need the visualizer
-\hdb{ Visualization is an integral part of our topics analysis which
+\hla{ Visualization is an integral part of our topics analysis which
   allows us to quickly explore the topics and trends of a project.
@@
 
-\hdb{
+\hla{
 The \emph{trend-time-line} (Figure \ref{fig:trendtimeline})
@@
 
-\hdb{ The \emph{trend-histogram} (Figure \ref{fig:histogram}) attempts
+\hla{ The \emph{trend-histogram} (Figure \ref{fig:histogram}) attempts
   to show a count of how often a trend reoccurs and how many topics
@@
 topic ($1\%$ to $10\%$ of the total topics), otherwise the tail is
-long. \hdb{ The \emph{trend-histogram} summarizes the distribution of
+long. \hla{ The \emph{trend-histogram} summarizes the distribution of
 trends ordered by size. }
 
-\hdb{ All of these visualization combine to enable an analysis of
+\hla{ All of these visualization combine to enable an analysis of
   trends and topics. Some views like the \emph{compact-trend-view}

### 2009-06-25 lda-paper.tex
ric edits

@@
 
-\hla{
+\hlb{
 Topic analysis extracts independant word distributions (topics) from
@@
 word distribution.  
-\hla{
+\hlb{
 Given a visualization, such as Figure \ref{fig:lda}, topic
@@
   \includegraphics[width=0.9\textwidth]{commit-to-topics} 
-  \caption{How commits are analyzed and aggregated into Topics and Trends.
+  \caption{How commits are analyzed and aggregated into topics and trends.
 Commits are first extracted, then abstracted into word counts or word distributions.
-These word distributions are then given a topic analysis tool like LDA.
-LDA finds independent word distributions (topics) that these documents are related to.}
+These word distributions are then given to topic analysis tool like LDA.
+LDA finds independent word distributions (topics) that these documents are related to \hlb{ (the numbers indicate similarity between documents and topics)}.}
   \label{fig:commits}
@@
 
- Our Topic Analysis tool is an implementation of LDA, but
-}
-we could have used other similar tools like LSI, but previous work
+ Our Topic Analysis tool is an implementation of LDA.
+}
+We could have used other similar tools like LSI, but previous work
 showed more promising topic analysis results with LDA.  We note that
@@
 illustrates clustering of topics by topic similarity. These clusters
-of topics are called trends. A trend is probably interesting if it
-contains more than one topic.
+of topics are called trends. 
+Trends are interesting because they that topic during development has occured over more thanone period.
 }
@@
 
-Once we have determined our similarity clusters we can choose to
-analyze and to plot the topics.
+Once we have determined our similarity clusters we are ready analyze
+and to plot the topics.
 
@@
     joined and embedded in box representing the topics. No stop words were removed. The gap in the
-    middle is a period where no development took place.}
+    middle is a period where no development took place (see Section \ref{sec:maxdb}).}
   \label{fig:topicsmear}
@@
 %       1. [ ] need the visualizer
-\hla{ Visualization is an integral part of our topics analysis which
+\hlb{ Visualization is an integral part of our topics analysis which
   allows us to quickly explore the topics and trends of a project.
@@
 
-\hla{
+\hlb{
 The \emph{trend-time-line} (Figure \ref{fig:trendtimeline})
@@
 
-\hla{ The \emph{trend-histogram} (Figure \ref{fig:histogram}) attempts
+\hlb{ The \emph{trend-histogram} (Figure \ref{fig:histogram}) attempts
   to show a count of how often a trend reoccurs and how many topics
@@
 shows the count of instances of a trend and thus indicates which
-trends occur the most. Unfortunately due to the large number of topics
+trends occur the most. Due to the large number of topics
 (approximately $N$ topics multiplied by $M$ periods), given the
@@
 topic ($1\%$ to $10\%$ of the total topics), otherwise the tail is
-long. \hla{ The \emph{trend-histogram} summarizes the distribution of
+long. \hlb{ The \emph{trend-histogram} summarizes the distribution of
 trends ordered by size. }
 
-\hla{ All of these visualization combine to enable an analysis of
+\hlb{ All of these visualization combine to enable an analysis of
   trends and topics. Some views like the \emph{compact-trend-view}
@@
   \includegraphics[width=0.45\textwidth]{fixed-time-smear-plot-cropped}
-  \caption{A zoomed in slice of compact-trend-view of topics per month of MaxDB 7.500. The topic text is visible in each topic box. Trends are plotted across time continuously.}
+  \caption{A zoomed-in slice of compact-trend-view of topics per month of MaxDB 7.500. The topic text is visible in each topic box. Trends are plotted across time continuously.}
   \label{fig:zoomedsmear}
@@
 
-We applied our tools and visualizations to the repositories of several
+We applied our tools and visualizations to the repositories of these
 \hla{
@@
 
-When we examined PostgreSQL's history from 1996 to 2004, which include
+When we examined PostgreSQL's history from 1996 to 2004, which includes
 over $20 000$ commits. We did not find many trends with two or more
 topics, using a similarity of $7/10$. $7/10$ was chosen because it
-preserved indepedant topics but seemed to be a borderline threshold
+preserve indepedent topics but seemed to be a threshold
 value where more serious trends started to appear.
@@
 were not very large, lasting only 3 months at most. The first and second
-largest trends were dedicated to the work of two external developers:
+largest trends directly referenced two external developers:
 Dal Zotto, Dan McGuirk. The fifth largest trend related to work by
 PostgreSQL developer D'Arcy. Other topics of the larger trends were
-changes to the TODO, and time string formating topics
+changes to the ``to do'' list, and time string formating topics
 relating to timezones.
@@
 \emph{fix}, \emph{update}.  By decreasing the similarity constraint to
-$1/2$ similarity the largest most common trend, which stretched across
+$1/2$ the largest most common trend, which stretched across
 the entire development, contained these same words (\emph{patch},
@@
 similarity of $7/10$ Firebird had far more continuous and
-recurring trends than PostgreSQL.  The first large trend was segmented
+recurring trends than PostgreSQL.  The first large trend was discontinuous
 across time but explicitly references one author \texttt{carlosga05}
@@
 \shrinkit
+
+\label{sec:maxdb}
 
@@
 instance in the last month of MaxDB 7.500 development every topic
-contained 1 unique Bug tracker URL, this pattern did not occur in the
+contained one unique Bug tracker URL. This pattern did not occur in the
 previous month. We investigated the revisions and we found that
@@
 four times, yet in the global-trend-view (Figure
-\ref{fig:statictopics}) it is not in any of the topics. A query
+\ref{fig:statictopics}, explained in Section \ref{ref:compare}) it is not in any of the topics. A query
 optimizer is an important component of an DBMS, but as we have shown
@@
 
+\label{sec:compare}
+
 % Migod notes This section is perhaps the most important of the
@@
 
-Windowed topic analysis demonstrated its value to hilight local topics
+Windowed topic analysis demonstrated its ability to hilight local topics
 and indentify global trends. This was shown in our case study of MaxDB
@@
 different aspects of the data: temporality of trends, trend size, and
-a compact-trend-view. The compact-trend-view shows much more
-information than the views that global analysis could show, it
-indicates how focused a period is by the total number of topics, as
+a compact-trend-view. The compact-trend-view shows more
+information than the views that global analysis could show, and it
+indicates how focused a period is by the total number of topics. As
 well, it shows topics by similarity so one can track trends across
-time.  Our trend-histogram immediately highlights and measures the
+time.  Our trend-histogram highlights and measures the
 size of trends while our trend-time-line view shows how a topic
-reoccurs over time. These visualizations help us get a general feeling
-about the common and unique topics that developers focus on during
+reoccurs over time. These visualizations help us  understand the common topics that developers focus on during
 development. If implemented interactively an user could easily zoom in
 and query for a summary of a topic or trend.
+In summary, our work on windowed topic analysis shows the potential
+for automatically determining key topics and trends across software
+development projects .
 
@@
 
-\hla{
-We want to explore parameters more across multiple projects, what
-motivates the need for more or less topics, can we find a good
-estimate of a good choice of number of topics to evaluate. We want to
-investigate the effects of overlapping windows on this analysis, do
-they make the trends too long? Do they highlight some trends
-unnecessarily or are they more because they are time agnostic.
-}
+\hla{ We want to explore parameters more across multiple projects,
+  what motivates the need for more or less topics.  Parameters
+  include: window overlapping, window overlap size, window size,
+  number of topics, etc.  We want to investigate the effects of
+  overlapping windows on this analysis. Do they make the trends too
+  long? Do they highlight some trends unnecessarily or are they more
+  because they are time agnostic.  }
 

### 2009-06-25 lda-paper.tex
minor touchups and     - [X] Fix the Last Diagram - more contrast?

@@
 \definecolor{changecolor4}{rgb}{0.0,0.3,0.0} 
-\newcommand{\hl}[1]{{\color{changecolor1} #1 }}
-\newcommand{\hla}[1]{{\color{changecolor2} #1 }}
-\newcommand{\hlb}[1]{{\color{changecolor3} #1 }}
-\newcommand{\hlc}[1]{{\color{changecolor4} #1 }}
+% \newcommand{\hl}[1]{{\color{changecolor1} #1 }}
+% \newcommand{\hla}[1]{{\color{changecolor2} #1 }}
+% \newcommand{\hlb}[1]{{\color{changecolor3} #1 }}
+% \newcommand{\hlc}[1]{{\color{changecolor4} #1 }}
+
+\newcommand{\hl}[1]{#1}
+\newcommand{\hla}[1]{#1}
+\newcommand{\hlb}[1]{#1}
+\newcommand{\hlc}[1]{#1}
 
@@
 %\alignauthor
-Abram Hindle and Michael W. Godfrey and Richard C. Holt\\
+Abram Hindle, Michael W. Godfrey and Richard C. Holt\\
 %\affaddr{David Cheriton School of Computer Science}\\
@@
   the previous iteration of a project's development?  What were the
-  developers working on? What was the developer in the next cubicle
-  working on? Which common topics and issues were dealt with in the
-  current release of the software? What were the topics of the
+  developers working on? What were the topics of the
   previous release? Given the requirements agreed to at the start of
@@
 four times, yet in the global-trend-view (Figure
-\ref{fig:statictopics}, explained in Section \ref{ref:compare}) it is not in any of the topics. A query
+\ref{fig:statictopics}, explained in Section \ref{sec:compare}) it is not in any of the topics. A query
 optimizer is an important component of an DBMS, but as we have shown

### 2009-06-25 lda-paper.tex
some minor changes

--- 2009-06-25 lda-paper.tex
@@
 
+% \newcommand{\hl}[1]{#1}
+% \newcommand{\hla}[1]{#1}
+% \newcommand{\hlb}[1]{#1}
+% \newcommand{\hlc}[1]{#1}
+
 \author{Abram Hindle, Michael Godfrey and Ric Holt \\
@@
 \hlb{
-Topic analysis extracts independant word distributions (topics) from
+Topic analysis extracts independent word distributions (topics) from
 documents (commit log comments).
@@
 \emph{section}, \emph{figure}, and \emph{chapter}. The other topic
-would include words found in topics generally independant of the other
+would include words found in topics generally independent of the other
 topics. If we analyzed a commit of a bug fix it would probably contain
@@
 documents via their word distributions. LDA also would discover these
-independant topics in an unsupervised manner. All the end user has to
+independent topics in an unsupervised manner. All the end user has to
 do is name the topic based upon the most frequent words in that topic,
@@
 all we can handle before too many topics seemed too indistinct from
-each other to differeniate.
+each other to differentiate.
 
@@
 of topics are called trends. 
-Trends are interesting because they that topic during development has occured over more thanone period.
+Trends are interesting because they that topic during development has occurred over more than one period.
 }
@@
   visually answer many questions about: the spread of topics, how many
-  topics were independant per period, what the repeating trends were,
+  topics were independent per period, what the repeating trends were,
   if the trends dominated or did local topics dominate, how continuous
@@
 The \emph{trend-time-line} (Figure \ref{fig:trendtimeline})
-attempts to show a summary of trends seperated from single topics. This view shows how a trend persists across time, which aids time-wise analysis of trends.
+attempts to show a summary of trends separated from single topics. This view shows how a trend persists across time, which aids time-wise analysis of trends.
 }
@@
 topics, using a similarity of $7/10$. $7/10$ was chosen because it
-preserve indepedent topics but seemed to be a threshold
+preserve independent topics but seemed to be a threshold
 value where more serious trends started to appear.
@@
 changes to the ``to do'' list, and time string formating topics
-relating to timezones.
+relating to time-zones.
 
@@
 
-Windowed topic analysis demonstrated its ability to hilight local topics
-and indentify global trends. This was shown in our case study of MaxDB
+Windowed topic analysis demonstrated its ability to hi-light local topics
+and identify global trends. This was shown in our case study of MaxDB
 7.500. Global topic analysis missed important topics such as

### 2009-06-25 lda-paper.tex
slight edits

@@
 
-  As development on a software project progresses, developers shift their focus between different topics and tasks many
-  times.  Managers and newcomer developers often seek ways of
-  understanding what tasks have recently been worked on and how much
-  effort has gone into each; for example, a manager might wonder what
-  unexpected tasks occupied their team's attention during a period
-  when they were supposed to have been implementing a set of new
-  features.  Tools such as Latent Dirichlet Allocation (LDA) and
-  \hla{
-    Latent Semantic Indexing (LSI) can be used to analyze commit log
-  }
-  comments over the entire lifetime of a project. Previous work on
-  developer topic analysis has leveraged these tools to associate
-  commit log comments with independent topics extracted from these
-  commit log comments.  In this paper, we use LDA  to analyze
-  periods, such as months, within a project's lifetime to create a
-  time-windowed model of changing development topics.  We propose 
+  As development on a software project progresses, developers shift
+  their focus between different topics and tasks many times. Managers
+  and newcomer developers often seek ways of understanding what tasks
+  have recently been worked on and how much effort has gone into each;
+  for example, a manager might wonder what unexpected tasks occupied
+  their team's attention during a period when they were supposed to
+  have been implementing a set of new features.  Tools such as Latent
+  Dirichlet Allocation (LDA) and Latent Semantic Indexing (LSI) can be
+  used to extract a set of independent topics from a corpus of commit
+  log comments.  Previous work in the area has created a single set of
+  topics by analyzing comments from the entire lifetime of the
+  project.  In this paper, we propose windowing the topic analysis to
+  give a more nuanced view of the system's evolution. By using a
+  defined time-window of, for example, one month, we can track which
+  topics come and go over time, and which ones recur. We propose
   visualizations of this model that allows us to explore the evolving
@@
   that windowed topic analysis offers advantages over topic analysis
-  applied to a project's lifetime because many topics are quite local.
+  applied to a project's lifetime because many topics are quite
+  local.
+
+%   As development on a software project progresses, developers shift
+%   their focus between different topics and tasks many times.  Managers
+%   and newcomer developers often seek ways of understanding what tasks
+%   have recently been worked on and how much effort has gone into each;
+%   for example, a manager might wonder what unexpected tasks occupied
+%   their team's attention during a period when they were supposed to
+%   have been implementing a set of new features.  Tools such as Latent
+%   Dirichlet Allocation (LDA) and \hla{ Latent Semantic Indexing (LSI)
+%     can be used to analyze commit log } comments over the entire
+%   lifetime of a project. Previous work on developer topic analysis has
+%   leveraged these tools to associate commit log comments with
+%   independent topics extracted from these commit log comments.  In
+%   this paper, we use LDA to analyze periods, such as months, within a
+%   project's lifetime to create a time-windowed model of changing
+%   development topics.  We propose visualizations of this model that
+%   allows us to explore the evolving stream of topics of development
+%   occurring over time.  We demonstrate that windowed topic analysis
+%   offers advantages over topic analysis applied to a project's
+%   lifetime because many topics are quite local.
 

### 2009-06-26 lda-paper.tex
minor lda-paper changes

@@
 %   have been implementing a set of new features.  Tools such as Latent
-%   Dirichlet Allocation (LDA) and \hla{ Latent Semantic Indexing (LSI)
-%     can be used to analyze commit log } comments over the entire
+%   Dirichlet Allocation (LDA) and  Latent Semantic Indexing (LSI)
+%     can be used to analyze commit log  comments over the entire
 %   lifetime of a project. Previous work on developer topic analysis has
@@
 
+\shrinkit
 \Section{Introduction}
-
-\hla{ The questions that we wish to address deal with the past and
+\shrinkit
+
+  The questions that we wish to address deal with the past and
   topics development that occurred in the past.  What happened in
@@
   the iteration did our developers work on them?  What else did they
-  work on?  }
-
-\hla{
+  work on?
+
 We want to know about the topics of that programmers
@@
 to rely on topic analysis.
-}
-
-\hlb{
+
 Topic analysis extracts independent word distributions (topics) from
 documents (commit log comments).
-Ideally these extracted topics }
+Ideally these extracted topics 
 correlate with actual development topics that the developers discussed
@@
 Topic analysis often allows for a single document, such as a commit
-\hlb{
 message, to be related to multiple topics. Documents represented as a mixture of topics maps well to commits
-}
-\hla{
 to source code, which often have more than one purpose.  A topic
@@
 projects source control system (SCS).
-}
-
-\hla{
+
 In the past, topic analysis has been applied to the entire history of a project~\cite{1321709,lukins2008,10.1109/ICPC.2007.13},
-}
-\hlb{
 where-as this paper suggests that most topics are local. For example,  
-programmer may be working on a story card,} they might never deal with that
+programmer may be working on a story card, they might never deal with that
 story card ever again. Thus that topic might be quite relevant to a
@@
 
-\hlb{
 We want to see topics that are unique to a
@@
 Figure \ref{fig:lda} illustrates the kind of visualization we want to
-automatically generate. } Those topics that
+automatically generate. Those topics that
 recur, or occur over a larger period are plotted continuously. In our
@@
 word distribution.  
-\hlb{
 Given a visualization, such as Figure \ref{fig:lda}, topic
@@
 developer focus.
-}
 By breaking apart an iteration into sets of topics
@@
 focus as it changes over time. We will show that most topics are
-\hlb{
 locally relevant, i.e., relevant to a single window of time.  
@@
 and global (entire history) analysis.
-}
-
-\hla{
+
 Our contributions include:
@@
 \end{itemize}
-}
 
@@
 Grant et al.~\cite{scottcordy} and have used an alternative technique,
-\hla{
-called Independent Component Analysis~\cite{delac} to separate topic signals from
-}
+called Independent Component Analysis~\cite{delacshort} to separate topic signals from
 software code.
@@
 
-\hla{
 Imagine that we are analyzing one month of development in a project
@@
 LDA and tell LDA to find some topics, in this case we
-\hlb{
 in this case we chose to concentrate on four topics.
-} The four
+ The four
 topics it discovered would ideally be bug fixing, GUI, documentation
@@
 which is a word distribution.
-}
 
@@
 windows. We asked LDA to make 20 topics per window and we then
-\hla{
 examined the top 10 most frequent words in that topic.  We chose one
@@
 the results and create indistinct topics. 
-\hlb{
   We chose the top ten most frequent words because sometimes a couple
@@
   to distinguish topics.
-}
 
@@
 commit comments was usually accurate.
-}
 
@@
 \caption{
-\hla{
 Sampled topics from MySQL 3.23, some with continuous topics.
@@
 Each token is a summary of one LDA generated topic from MySQL 3.23 commit comments.
-}
 }
@@
 These word distributions are then given to topic analysis tool like LDA.
-LDA finds independent word distributions (topics) that these documents are related to \hlb{ (the numbers indicate similarity between documents and topics)}.}
+LDA finds independent word distributions (topics) that these documents are related to (the numbers indicate similarity between documents and topics).}
   \label{fig:commits}
@@
 
-\hla{
 We mirrored the repositories and their revisions using software such
 as rsync~\cite{rsync}, CVSsuck~\cite{cvssuck},
-\texttt{softChange}~\cite{dmgseke2004}, and bt2csv~\cite{methodology}.
+\texttt{softChange}~\cite{dmgseke2004short}, and bt2csv~\cite{methodology}.
 \texttt{softChange} provided a general schema for storing revisions
@@
 repositories.
-}
 
@@
 Windowing by time
-allows for many levels of granularity.  We used one month as the
+allows for many levels of granularity.  
+We used one month as the
 length of our time windows. While we could use different window
@@
 development, which is large enough to show shifts in focus, but coarse
-\hla{
 enough not to show too much detail. Choosing a month as the window
 provided us with enough documents to analyze.
-}
 
@@
 each window and extract the top $N$ topics. We used $20$ topics in our
-\hla{
 case studies because we experimented before and that $20$ was about
@@
  Our Topic Analysis tool is an implementation of LDA.
-}
 We could have used other similar tools like LSI, but previous work
@@
 given a certain threshold of top 10 word similarity, such as 8 out of
-\hla{
 top 10 matching words.  10 words were chosen because often people care
@@
 Trends are interesting because they that topic during development has occurred over more than one period.
-}
 
@@
   \centering
-  \includegraphics[width=0.45\textwidth]{transitiveclosure}
+  \includegraphics[width=0.42\textwidth]{transitiveclosure}
   \caption{Topic similarity demonstrated by clustering topics by the transitive closure (connectedness)
@@
 %       1. [ ] need the visualizer
-\hlb{ Visualization is an integral part of our topics analysis which
+  Visualization is an integral part of our topics analysis which
   allows us to quickly explore the topics and trends of a project.
@@
   visualization, described within this section, that help answer
-  different questions.  }
+  different questions.
 
@@
 \ref{fig:histogram} where we plot each trend on its own line but stack
-\hla{
 up the segments of the trend, much like a histogram. Each topic has its top 10 words 
@@
 A trend has all of its topic text embedded side by side within the same trend box.
-}
-
-\hlb{ The \emph{compact-trend-view} (Figure \ref{fig:topicsmear})
+
+ The \emph{compact-trend-view} (Figure \ref{fig:topicsmear})
   attempts to show all topics and trends at once across time in a
   compact view that could fit on one page. It tries to give a compact
-  summary of the topics and trends across time.  } The
+  summary of the topics and trends across time.   The
 \emph{compact-trend-view} (Figure \ref{fig:topicsmear}) tries to sink
@@
 
-\hlb{
 The \emph{trend-time-line} (Figure \ref{fig:trendtimeline})
 attempts to show a summary of trends separated from single topics. This view shows how a trend persists across time, which aids time-wise analysis of trends.
-}
 The \emph{trend-time-line} displays repeating trends more clearly by
@@
 
-\hlb{ The \emph{trend-histogram} (Figure \ref{fig:histogram}) attempts
+ The \emph{trend-histogram} (Figure \ref{fig:histogram}) attempts
   to show a count of how often a trend reoccurs and how many topics
   are related to a trend. It is meant to show the distribution of
-  trends by their size in topics and time-span.  } The
+  trends by their size in topics and time-span.  The
 \emph{trend-histogram} superficially resembles the trend-time-line.
@@
 topic ($1\%$ to $10\%$ of the total topics), otherwise the tail is
-long. \hlb{ The \emph{trend-histogram} summarizes the distribution of
-trends ordered by size. }
-
-\hlb{ All of these visualization combine to enable an analysis of
+long. The \emph{trend-histogram} summarizes the distribution of
+trends ordered by size.
+
+ All of these visualization combine to enable an analysis of
   trends and topics. Some views like the \emph{compact-trend-view}
@@
   visualizations to analyze the database systems that we discuss in
-  the following results section.  }
+  the following results section.
 
@@
   \centering
-  \includegraphics[width=0.45\textwidth]{fixed-time-smear-plot-cropped}
+  \includegraphics[width=0.42\textwidth]{fixed-time-smear-plot-cropped}
   \caption{A zoomed-in slice of compact-trend-view of topics per month of MaxDB 7.500. The topic text is visible in each topic box. Trends are plotted across time continuously.}
@@
   \centering
-  \includegraphics[width=0.45\textwidth]{histogram-cropped-scaled}
+  \includegraphics[width=0.42\textwidth]{histogram-cropped-scaled}
   \caption{The top part of a trend-histogram of MaxDB 7.500, ordered
@@
 We applied our tools and visualizations to the repositories of these
-\hla{
 open source database systems: PostgreSQL, MaxDB and Firebird. The
 total number of commits analyzed was over $66 000$.
-}
 
@@
 
-If we kept the stop words we found that the large trend consisted
+If we kept the stop words, we found that the large trend consisted
 mostly of stop words and non-stop words such as \emph{patch},
@@
 
-\hla{
 We tracked Firebird from August 2000 to January 2006, we extracted comments from $38 000$ commits. We found that with a
-}
 similarity of $7/10$ Firebird had far more continuous and
@@
 The plots we produced of MaxDB 7.500 were unlike those of the other
-\hla{
 systems, as there was a period where no development occurred and thus
@@
 were a total of $8 600$ commits analyzed.
-}
 
@@
 words and then two of the most common words, the global analysis
-\hla{
 finally had a topic with optimizer in its topic ten words. Our
@@
 messages.
-}
 
@@
 
+% In this study we are explicitly trusting that the programmers annotate
+% their changes with relevant information. We rely on the descriptions
+% they provide. If the language of check-in comments was automated we
+% would  be analyzing only that.
+% We compared topics using the top 10 tokens, this approach could be
+% throwing data away and might not be as useful as determining the
+% actual distance between two word topic distributions.
+% Adding and removing stop words produced different results. Our choice
+% of stop words could be biased, and could affect the results.
+% The number of commits per month is inconsistent as some months have
+% many changes while other months have almost none.
+
 \shrinkit
@@
 
-\hla{ We want to explore parameters more across multiple projects,
-  what motivates the need for more or less topics.  Parameters
-  include: window overlapping, window overlap size, window size,
-  number of topics, etc.  We want to investigate the effects of
-  overlapping windows on this analysis. Do they make the trends too
-  long? Do they highlight some trends unnecessarily or are they more
-  because they are time agnostic.  }
-
+We wish to extend this study by further exploring parameter
+  choices and their effects in terms of window overlap size and number
+ of topics.
+%  We want to explore parameters more across multiple projects,
+%   what motivates the need for more or less topics.  Parameters
+%   include: window overlapping, window overlap size, window size,
+%   number of topics, etc.  We want to investigate the effects of
+%   overlapping windows on this analysis. Do they make the trends too
+%   long? Do they highlight some trends unnecessarily or are they more
+%   because they are time agnostic. 
 Another avenue of future work is automatic topic labelling. Given a word
@@
 that describes that distribution. Potential external sources of topic
-names include: software engineering taxonomies, ontologies and
+names include software engineering taxonomies, ontologies and
 standards.

### 2009-06-26 lda-paper.tex
migod edits

@@
 
+%XXX Drop down to 200
 \begin{abstract}
@@
   their team's attention during a period when they were supposed to
-  have been implementing a set of new features.  Tools such as Latent
-  Dirichlet Allocation (LDA) and Latent Semantic Indexing (LSI) can be
-  used to extract a set of independent topics from a corpus of commit
-  log comments.  Previous work in the area has created a single set of
-  topics by analyzing comments from the entire lifetime of the
-  project.  In this paper, we propose windowing the topic analysis to
-  give a more nuanced view of the system's evolution. By using a
-  defined time-window of, for example, one month, we can track which
-  topics come and go over time, and which ones recur. We propose
+  have been implementing new features.  
+%  have been implementing a set of new features.  
+%
+  Tools such as Latent Dirichlet Allocation (LDA) and Latent Semantic
+%  Indexing (LSI) can be used to extract a set of independent topics
+  Indexing (LSI) extract a set of independent topics
+  from a corpus of commit-log comments.  Previous work in the area has
+  created a single set of topics by analyzing comments from the entire
+  lifetime of the project.  In this paper, we propose windowing the
+  topic analysis to give a more nuanced view of the system's
+  evolution. By using a defined time-window of, for example, one
+  month, we can track which topics come and go over time, and which
+  ones recur.
+%
+%
+%   Tools such as Latent
+%   Dirichlet Allocation (LDA) and Latent Semantic Indexing (LSI) can be
+%   used to extract a set of independent topics from a corpus of commit
+%   log comments.  Previous work in the area has created a single set of
+%   topics by analyzing comments from the entire lifetime of the
+%   project.  In this paper, we propose windowing the topic analysis to
+%   give a more nuanced view of the system's evolution. By using a
+%   defined time-window of, for example, one month, we can track which
+%   topics come and go over time, and which ones recur.
+%  
+  We propose
   visualizations of this model that allows us to explore the evolving
@@
 
-  The questions that we wish to address deal with the past and
-  topics development that occurred in the past.  What happened in
-  the previous iteration of a project's development?  What were the
-  developers working on? What were the topics of the
-  previous release? Given the requirements agreed to at the start of
-  the iteration did our developers work on them?  What else did they
-  work on?
-
-We want to know about the topics of that programmers
-encountered while developing a software project. We want to know given
-a certain period of time, what were the popular topics being worked on
-during that period. We also want to know if any of those topics
-occurred again, whether before or after our current period.  These
-topics could help stakeholders determine what their coworkers were
-working on, what were common issues, what requirements were being
-worked on, etc. In order to extract and evaluate these topics we plan
-to rely on topic analysis.
-
-Topic analysis extracts independent word distributions (topics) from
-documents (commit log comments).
+Software managers know that --- during any given development cycle ---
+there is usually a difference between the set of tasks that developers
+are supposed to be working on and the tasks they are actually working
+on.  To better understand the realities of development, it would be
+helpful to be able to answer questions such as: What topics dominated
+the previous release cycle? Given the requirements agreed to at the
+start of the iteration how much work did our developers spend on them?
+What other tasks did they work on?  Which topics seem to recur over
+and over between cycles?  How long do different topics persist?  Are
+there latent cause-effect relationships between different topics?
+
+%   The questions that we wish to address deal with the past and
+%   topics development that occurred in the past.  What happened in
+%   the previous iteration of a project's development?  What were the
+%   developers working on? What were the topics of the
+%   previous release? Given the requirements agreed to at the start of
+%   the iteration did our developers work on them?  What else did they
+%   work on?
+
+% We want to know about the topics of that programmers
+% encountered while developing a software project. We want to know given
+% a certain period of time, what were the popular topics being worked on
+% during that period. We also want to know if any of those topics
+% occurred again, whether before or after our current period.  These
+% topics could help stakeholders determine what their coworkers were
+% working on, what were common issues, what requirements were being
+% worked on, etc. In order to extract and evaluate these topics we plan
+% to rely on topic analysis.
+
+Topic analysis 
+uses tools such as Latent Dirichlet Allocation (LDA) and Latent Semantic Indexing (LSI) to extract
+independent word distributions (topics) from
+documents (commit log comments)~\cite{1374321,10.1109/ICPC.2007.13,lukins2008,10.1109/MSR.2007.20}.
 Ideally these extracted topics 
@@
 
-In the past, topic analysis has been applied to the entire history of a project~\cite{1321709,lukins2008,10.1109/ICPC.2007.13},
-where-as this paper suggests that most topics are local. For example,  
-programmer may be working on a story card, they might never deal with that
-story card ever again. Thus that topic might be quite relevant to a
-small window of time but never relevant ever again.  By applying topic
-analysis locally to a window or an iteration, both the developers and managers
-might be able to see what were the topics of development during that
-iteration, rather than the whole project.
-
-We want to see topics that are unique to a
-certain time period, displayed across the time axis.
-Figure \ref{fig:lda} illustrates the kind of visualization we want to
-automatically generate. Those topics that
+Previous work on topic analysis has been applied to the entire history of a 
+project~\cite{1321709,lukins2008,10.1109/ICPC.2007.13}
+to produce a single set of topics. In this paper we explore the idea
+of time-windowed topic analysis; that is, we pick a period of, say,
+one month and perform topic analysis within each window.  In addition
+to being able to study just which topics were actually worked on and
+when, we can also analyze the historical timeline of topics for
+patterns and trends.  For example, we might wish to categorize each
+topic as being local to a single time window, spanning multiple
+consecutive windows, or recurring periodically.  Alternatively, we
+might explore if any topics often recur together, if one particular
+topic usually precedes another, if unplanned topics outnumber planned
+topics, if particular topics recur near the end of development
+iterations, etc.  whereas this paper suggests that most topics are
+local.
+
+% In the past, topic analysis has been applied to the entire history of a project~\cite{1321709,lukins2008,10.1109/ICPC.2007.13},
+% whereas this paper suggests that most topics are local. For example,  
+% programmer may be working on a story card, they might never deal with that
+% story card ever again. Thus that topic might be quite relevant to a
+% small window of time but never relevant ever again.  By applying topic
+% analysis locally to a window or an iteration, both the developers and managers
+% might be able to see what were the topics of development during that
+% iteration, rather than the whole project.
+
+% We want to see topics that are unique to a
+% certain time period, displayed across the time axis.
+% Figure \ref{fig:lda} illustrates the kind of visualization we want to
+% automatically generate. 
+As part of this work, we wish to explore approaches to automatically
+generating visualizations of the topic trends and development timelines, so
+that managers may be able to more easily grasp the topic-related
+activities of the developers.  Figure \ref{fig:lda} gives an example.
+Those topics that
 recur, or occur over a larger period are plotted continuously. In our
@@
 We now define some terms that we will use in the rest of the paper: A
-\emph{message} is a block of text written by developers. In this
+\emph{message} is a block of text written by a developer. In this
 paper, messages will be the CVS and BitKeeper commit log comments made
@@
 a word count divided by the message size. A \emph{topic} is a word
-distribution, i.e. a set of words that form a word distribution that is
+distribution, i.e., a set of words that form a word distribution that is
 unique and independent within the set of documents in our total
@@
 filtering. LDA attempts to encode documents as a mixture
-model, a combination of topics.  LDA is used in software engineering
+model, a combination of topics.  LDA has been used in software engineering
 literature~\cite{lukins2008,10.1109/MSR.2007.20}%,NIPS2007637,1321709}.
-LDA extracts topics from documents such as methods, bug
+to extract topics from documents such as methods, bug
 reports, source code and files.
@@
 LDA, LSI and semantic clustering extract topic
-clusters from the documents that are independent from one another. LDA
-does not name these clusters but the clusters consist of words whose
+clusters from the documents that are independent from one another. 
+The clusters are not named by the process, and consist of
+%LDA
+%does not name these clusters but the clusters consist of
+words whose
 relationship is often obvious to someone familiar with the corpus. For our purposes, we
@@
 
-Our technique differs from the previous techniques because we apply
+Our technique differs from the previous approaches to topic analysis because we apply
 LDA locally to month-long windows of commit log comments, whereas
@@
 Imagine that we are analyzing one month of development in a project
-where the main focus was on bug fixing, adding a new GUI and
+where there were three major tasks: bug fixing, adding a new GUI, and
 documentation updates relating to the system and the GUI. To use LDA
-we would get the word distributions of each commit messages, pass them to
+we would get the word distributions of each commit messages, 
+\begin{comment}
+pass them to
 LDA and tell LDA to find some topics, in this case we
@@
 which is a word distribution.
+\end{comment}
+% (please preserve the italics you put in ... I had to edit this in plain text so they got lost)
+and have the LDA tool extract a set of topics from the data.  Ideally,
+the LDA analysis would return four topics, corresponding to bug
+fixing, GUI implementation, documentation, and ``other''.  Each topic
+essentially corresponds to a word distribution.  For example, the bug
+fixing topic might include words such as \emph{bug}, \emph{fix}, \emph{failure}, and
+\emph{ticket}, while the GUI topic might include \emph{widget}, \emph{panel}, \emph{window}, and
+\emph{menu}, and the documentation topic might include \emph{section}, \emph{figure}, and
+\emph{chapter}.  The last topic would include words that were not commonly
+found in bug reports, GUI fixing, or documentation commits; that is,
+the LDA analysis would put into this last category the words that were
+independent of the other topics it found.
+
+LDA is an unsupervised technique.  The user provides a corpus of
+documents and specifies various parameters, and then LDA automatically
+generates a set of likely topics based on word frequencies.  It is
+then up to the user to decide if the topics are meaningful by looking
+at the most frequent words in each and deciding what each discovered
+topic should be named.
+
+In addition to finding the topics, LDA can also associate the
+documents (i.e., commit messages) in its corpus to one or more of the
+topics it discovers.  For example, a commit of a simple bug fix would
+probably contain a word such as \emph{ticket}, and thus heavily associate
+with the bug fix topic.  A commit that concerned user documentation of
+the GUI would likely be associated with both the documentation and GUI
+topics if it shared words from both topics.  And a commit that did not
+concern bugs, GUIs, or documentation would likely be related to the other topic.
 
@@
 
+%   We decided [did we?] to treat these as stop words, since they occur
+%   across many of the topics and since they relate to the mechanisms of
+%   version control rather than the semantic content of the commit.
+
 We found there were common words across topic clusters, such as
-\emph{diffs}, \emph{annotate} and \emph{history}; since these words
-occur so often in the MySQL 3.23 messages. These words probably should
-be viewed as stop words since occur across so many topics, even though
-they are related to version control usage.  There were notable
-transitional topics such as in the first window the word
-\emph{BitKeeper} appears because MySQL adopted Bitkeeper for their
-source control system yet in the following windows there were no
-mentions of BitKeeper. \emph{RENAME} also only appeared once in the
-first window and never again. For each topic we tried to name the
-purpose of the topic; to our surprise we found that even with only a
+\emph{diffs}, \emph{annotate} and \emph{history}.
+Words like these could be treated as stop words, since they occur
+across many of the topics and since they relate to the mechanisms of
+version control rather than the semantic content of the commit.
+% since these words
+%occur so often in the MySQL 3.23 messages. These words probably should
+%be viewed as stop words since occur across so many topics, even though
+%they are related to version control usage.  
+There were topics which appeared only during some transitional period and never again
+such as \emph{RENAME} and \emph{BitKeeper}. \emph{BitKeeper} which appears when MySQL switched to  BitKeeper for their SCS.
+% There were notable
+% transitional topics such as in the first window the word
+% \emph{BitKeeper} appears because MySQL adopted Bitkeeper for their
+% source control system yet in the following windows there were no
+% mentions of BitKeeper. 
+%\emph{RENAME} also only appeared once in the
+%first window and never again.
+For each topic we tried to find a word to describe its purpose.
+To our surprise we found that even with only a
 little familiarity with the code base that naming the topic was
@@
   \includegraphics[width=0.9\textwidth]{lda}
-  \caption{Example of topics extracted from MySQL 3.23. This is the
-    kind of plot we eventually want to illustrate: named topics and
-    topic trends. The horizontal axis is time by month examined. The
+  \caption{Example of topics extracted from MySQL 3.23. 
+%     This is the
+%     kind of plot we eventually want to illustrate: named topics and
+%     topic trends. 
+    The horizontal axis is time by month examined. The
     vertical axis is used to stack topics that occur at the same
@@
   \includegraphics[width=0.9\textwidth]{commit-to-topics} 
-  \caption{How commits are analyzed and aggregated into topics and trends.
-Commits are first extracted, then abstracted into word counts or word distributions.
-These word distributions are then given to topic analysis tool like LDA.
+  \caption{How commits are analyzed and aggregated into topics and trends: commits are first extracted, then abstracted into word counts or word distributions which are then given to topic analysis tool like LDA.
 LDA finds independent word distributions (topics) that these documents are related to (the numbers indicate similarity between documents and topics).}
@@
 
-The documents we are analyzing are the commit log comments, i.e. the
+The documents we are analyzing are the commit log comments, i.e., the
 comments that are added when revisions to the project are committed.
@@
 Once we have our data windowed, we apply our Topic Analysis tool to
-each window and extract the top $N$ topics. We used $20$ topics in our
-case studies because we experimented before and that $20$ was about
-all we can handle before too many topics seemed too indistinct from
-each other to differentiate.
-
- Our Topic Analysis tool is an implementation of LDA.
-We could have used other similar tools like LSI, but previous work
-showed more promising topic analysis results with LDA.  We note that
-this step is a slow one, as executing even one instance of LDA
-involves significant computation, and we perform LDA once per window.
-Figure \ref{fig:commits} shows how LDA is applied to a set of
-messages, and how the topics are extracted and related to the
-messages.
+each window and extract the top $N$ topics.  We decided to use a
+threshold of $20$ topics based on our previous experience with LDA; we
+found that using more than $20$ makes the boundaries between topics less
+distinct.
+
+% We used $20$ topics in our
+% case studies because we experimented before and that $20$ was about
+% all we can handle before too many topics seemed too indistinct from
+% each other to differentiate.
+
+Our Topic Analysis tool is based around a third-party implementation
+of LDA.  While we could have used LSI or similar techniques as the
+engine, our preliminary studies found LDA to provide more promising
+results for topic analysis.  We note that this step is a slow one, as
+executing even one instance of LDA involves significant computation,
+and we perform LDA once per window.  Figure \ref{fig:commits} shows
+how LDA is applied to a set of messages, and how the topics are
+extracted and related to the messages.
 
@@
 illustrates clustering of topics by topic similarity. These clusters
-of topics are called trends. 
+of topics are called \emph{trends}. 
 Trends are interesting because they that topic during development has occurred over more than one period.
 
-This technique has weaknesses in that nodes that are a few neighbors
+This approach does have a weakness, in that nodes that are a few neighbors
 away in similarity might not share any similar words.  We use this
@@
 
-Once we have determined our similarity clusters we are ready analyze
+Once we have determined our similarity clusters we are ready to analyze
 and to plot the topics.
@@
     of topic similarity. Nodes are topics and
-    arcs imply similarity. Similarity could be a measure such as
-    topics share 8 out of 10 top ten words.}
+    arcs imply 
+some notion of similarity, e.g.,  topics share 8 out of 10 top ten words.}
 \label{fig:closure}
@@
   allows us to quickly explore the topics and trends of a project.
-  Visualization provides us with a framework that allows us to
-  visually answer many questions about: the spread of topics, how many
-  topics were independent per period, what the repeating trends were,
-  if the trends dominated or did local topics dominate, how continuous
-  trends were, and what the topic text in topic, trend, or period was.
-  Since we have multiple questions to answer we have multiple
-  visualization, described within this section, that help answer
-  different questions.
+  Visualizing the results is desirable as it allows us to explore the
+  data from different points of view to address different questions:
+  How are the topics spread over time?  How many independent topics
+  per period were there?  What repeating trends were there?  Did the
+  trends dominate, or did local topics dominate?  How broken up were
+  the trends?  
+  Since we have a variety of questions to answer we suggest multiple
+  visualizations, which we now describe.
+
+%   Visualization provides us with a framework that allows us to
+%   visually answer many questions about: the spread of topics, how many
+%   topics were independent per period, what the repeating trends were,
+%   if the trends dominated or did local topics dominate, how continuous
+%   trends were, and what the topic text in topic, trend, or period was.
+%   Since we have multiple questions to answer we have multiple
+%   visualization, described within this section, that help answer
+%   different questions.
 
@@
 (placement along the y-axis is arbitrary).  Another technique is
-the \emph{trend-time-line}, shown in Figure \ref{fig:trendtimeline},
+the \emph{trend-time-line}, shown in Figure \ref{fig:trendtimeline}, where
 each trend gets it own distinct y-value, while the x-axis is time;
@@
 listed in descending order of frequency from top to bottom, this text is embedded inside the topic's box.
-A trend has all of its topic text embedded side by side within the same trend box.
+A trend has all of its topic text embedded side by side within the same trend box, which at this resolution requires the reader to zoom in electronically.
 
@@
   attempts to show all topics and trends at once across time in a
-  compact view that could fit on one page. It tries to give a compact
-  summary of the topics and trends across time.   The
+  compact view that could fit on one page.   The
 \emph{compact-trend-view} (Figure \ref{fig:topicsmear}) tries to sink
@@
 
-We applied our tools and visualizations to the repositories of these
-open source database systems: PostgreSQL, MaxDB and Firebird. The
-total number of commits analyzed was over $66 000$.
+We applied our tools and visualizations to the repositories of three
+open source database systems: PostgreSQL, MaxDB, and Firebird. The
+total number of commits analyzed was over $66,000$.
 
@@
 When we examined PostgreSQL's history from 1996 to 2004, which includes
-over $20 000$ commits. We did not find many trends with two or more
+over $20,000$ commits. We did not find many trends with two or more
 topics, using a similarity of $7/10$. $7/10$ was chosen because it
@@
 
-We tracked Firebird from August 2000 to January 2006, we extracted comments from $38 000$ commits. We found that with a
+We tracked Firebird from August 2000 to January 2006, we extracted comments from $38,000$ commits. We found that with a
 similarity of $7/10$ Firebird had far more continuous and
@@
 across time but explicitly references one author \texttt{carlosga05}
-and words like \emph{added}, \emph{fixed} and \emph{updated}.
+and words like \emph{added}, \emph{fixed}, and \emph{updated}.
 
@@
 words and then two of the most common words, the global analysis
-finally had a topic with optimizer in its topic ten words. Our
+finally found a topic with optimizer in its topic ten words. Our
 analysis shows that optimizer was important but it had been obscured

### 2009-06-27 lda-paper.tex
some minor editting

@@
 Lukins, Kraft and Etzkorn~\cite{lukins2008} use LDA to help bug
-localization by querying for documents that are related to a bug's topic. They
-used LDA to build a topic model and then queried against it using
-sample documents. These queries would indicate the proportion of
-topics that were associated with the query and the
-related documents.
+localization by querying for documents that are related to a bug's
+topic. They used LDA to build a topic model and then searched for
+similar documents by querying with sample documents.
+%These queries would indicate the proportion of
+%topics that were associated with the query and the
+%related documents.
 
@@
 code or documents that are related to the source code.  Concept location concerns how
-high-level concepts relate to low level entities such as source code. For
-example, when fixing a bug, how and where do the concepts in the bug
-report map back to the source code?  Semantic Clustering has also been
+high-level concepts, such as bugs, relate to low level entities such as source code. 
+%For
+%example, when fixing a bug, how and where do the concepts in the bug
+%report map back to the source code?  
+Semantic Clustering has also been
 used for similar purposes~\cite{1698774,1566153} as it is similar to
@@
 addressed, the global topic analysis would probably miss this, yet
-these bug tickets were the focus of development for that month but not
-necessarily globally relevant.
+these bug tickets were the focus of development for that month.
+% but not
+% necessarily globally relevant.
 

### 2009-06-29 lda-paper.tex
Omg now we have 10 pages. Jesus christ

@@
 independent word distributions (topics) from
-documents (commit log comments)~\cite{1374321,10.1109/ICPC.2007.13,lukins2008,10.1109/MSR.2007.20}.
+documents (commit log comments)~\cite{1374321-short,10.1109/ICPC.2007.13-short,lukins2008,10.1109/MSR.2007.20-short}.
 Ideally these extracted topics 
@@
 Previous work on topic analysis has been applied to the entire history of a 
-project~\cite{1321709,lukins2008,10.1109/ICPC.2007.13}
+project~\cite{1321709-short,lukins2008,10.1109/ICPC.2007.13-short}
 to produce a single set of topics. In this paper we explore the idea
@@
 
+%SHORTEN
 We now define some terms that we will use in the rest of the paper: A
@@
 %(LSI)~\cite{1421013,1374321,10.1109/ICPC.2007.13,10.1109/ICPC.2006.17},
-(LSI)~\cite{1374321,10.1109/ICPC.2007.13},
+(LSI)~\cite{1374321-short,10.1109/ICPC.2007.13-short},
 probabilistic Latent Semantic Indexing (pLSI)~\cite{944937} and
-semantic clustering~\cite{1698774,1566153}. These tools are used for
+semantic clustering~\cite{1698774-short,1566153-short}. These tools are used for
 document modelling, document clustering and collaborative
@@
 model, a combination of topics.  LDA has been used in software engineering
-literature~\cite{lukins2008,10.1109/MSR.2007.20}%,NIPS2007637,1321709}.
+literature~\cite{lukins2008,10.1109/MSR.2007.20-short}%,NIPS2007637,1321709}.
 to extract topics from documents such as methods, bug
@@
 extracted from SCS, Linstead et al.\  proposed an author-source-code model
-using LDA~\cite{10.1109/MSR.2007.20}.%,NIPS2007637,1321709}. 
-  This model
-is essentially an extension of the author-topic (AT)
-model~\cite{1036902}, which associated authors and topics extracted
-with LDA.
+using LDA~\cite{10.1109/MSR.2007.20-short}.%,NIPS2007637,1321709}. 
+This model associates authors and topics via their associated documents.
+
+%   This model
+% is essentially an extension of the author-topic (AT)
+% model~\cite{1036902}, which associated authors and topics extracted
+% with LDA.
 
@@
 %location~\cite{1421013,1374321,10.1109/ICPC.2007.13,10.1109/ICPC.2006.17}.
-location~\cite{1374321,10.1109/ICPC.2007.13}.
+location~\cite{1374321-short,10.1109/ICPC.2007.13-short}.
 Concept analysis aims to automatically extract concepts from source
@@
 Semantic Clustering has also been
-used for similar purposes~\cite{1698774,1566153} as it is similar to
+used for similar purposes~\cite{1698774-short,1566153-short} as it is similar to
 LSI.
@@
 Grant et al.~\cite{scottcordy} and have used an alternative technique,
-called Independent Component Analysis~\cite{delacshort} to separate topic signals from
+called Independent Component Analysis~\cite{delac-short} to separate topic signals from
 software code.
@@
 the results and create indistinct topics. 
-  We chose the top ten most frequent words because sometimes a couple
-  of common words dominated topics so we wanted to have enough words
-  to distinguish topics.
+%  We chose the top 10 most frequent words because sometimes a couple
+%  of common words dominated topics so we wanted to have enough words
+%  to distinguish topics.
+We chose the top 10 words because we wanted to be able to distinguish
+topics even if some common words dominated the topics.
 
@@
 the word distribution share the same words the purpose extracted from
-commit comments was usually accurate.
-
+top 10 words was usually accurate.
+
+%SHORTEN could shorten 1 line
 A sampling of the notable topic words is displayed in Table
@@
 as rsync~\cite{rsync}, CVSsuck~\cite{cvssuck},
-\texttt{softChange}~\cite{dmgseke2004short}, and bt2csv~\cite{methodology}.
+\texttt{softChange}~\cite{dmgseke2004short}, and bt2csv~\cite{methodology-short}.
 \texttt{softChange} provided a general schema for storing revisions
@@
 
-%XXX Smaller?
-
+%XXX maybe fix
 The documents we are analyzing are the commit log comments, i.e., the
 comments that are added when revisions to the project are committed.
-For each commit log comment, we count the occurrence of each word in
-the message, remove stop words, and then produce word distributions
-for each message.  These distributions are then normalized by the size
-of the message, each count was divided by the total number of
-tokens. After all messages are processed the distributions are
-extended to include all words from all of the distributions.
+For each commit log comment we produce word distributions by counting
+the occurrence of each word in the message, removing stop words and
+normalizing the distribution by the size of message, in tokens.
+% (dividing counts by the number of tokens).  
+After
+all messages are processed the distributions are extended to include
+all words from all of the distributions.
+
+%, and then produce word distributions
+%for each message.  
+%These distributions are then normalized by the size
+%of the message, each count was divided by the total number of
+%tokens.
 
@@
 of a set time unit because it simplifies analysis.  
-One could overlap the
-windows, which would increase the likelihood of trends, but for this
+Overlapping windows would 
+%One could overlap the
+%windows, which would 
+increase the likelihood of trends, 
+but for this
 study we lacked the space and were more concerned if topics ever
@@
 topic.  Each topic is compared to each other topic in the system,
-given a certain threshold of top 10 word similarity, such as 8 out of
+given a certain threshold of similarity, such as 8 out of
 top 10 matching words.  10 words were chosen because often people care
 about top 10 comparison and it allowed for some common words to exist
-but not cause topics to match so closely.  We then apply the
-transitive closure on similar topics to this topic similarity matrix;
+%but not cause topics to match so closely.  
+yet not produce too many trends.
+We then apply the
+transitive closure on similar topics to our set of topic similarities;
 this is similar to modelling topics as nodes and similarity as arcs,
@@
 of topics are called \emph{trends}. 
-Trends are interesting because they that topic during development has occurred over more than one period.
+Trends indicate that a topic has occurred over more than one period during development.
 
@@
 
-\begin{figure}[t]
+\begin{figure}
   \centering
@@
     arcs imply 
-some notion of similarity, e.g.,  topics share 8 out of 10 top ten words.}
+some notion of similarity, e.g.,  topics share 8 out of  top 10 words.}
 \label{fig:closure}
@@
   \caption{Compact-trend-view shows topics per month for MaxDB
-    7.500. The x-axis is time in months, the Y axis is used to stack
+    7.500. The x-axis is time in months, the y-axis is used to stack
     topics occurring at the same time. Trends that are continuous are
@@
   allows us to quickly explore the topics and trends of a project.
-  Visualizing the results is desirable as it allows us to explore the
+  Visualizing the results %is desirable as it 
+  allows us to explore the
   data from different points of view to address different questions:
   How are the topics spread over time?  How many independent topics
-  per period were there?  What repeating trends were there?  Did the
+  per period were there?  What were the repeating trends?  Did the
   trends dominate, or did local topics dominate?  How broken up were
   the trends?  
-  Since we have a variety of questions to answer we suggest multiple
-  visualizations, which we now describe.
+  To address these multiple questions we will describe and employ multiple visualizations.
+%  Since we have a variety of questions to answer we suggest multiple
+%  visualizations, which we now describe.
 
@@
 up the segments of the trend, much like a histogram. Each topic has its top 10 words 
-listed in descending order of frequency from top to bottom, this text is embedded inside the topic's box.
-A trend has all of its topic text embedded side by side within the same trend box, which at this resolution requires the reader to zoom in electronically.
+listed in descending order of frequency from top to bottom, this text is embedded inside the topic's box, which at this resolution requires the reader to zoom in electronically.
+A trend has all of its topic text embedded side by side within the same trend box.
+
+\begin{figure}[t]
+  \centering
+  \includegraphics[width=0.42\textwidth]{fixed-time-smear-plot-cropped}
+  \caption{A zoomed-in slice of compact-trend-view of topics per month of MaxDB 7.500. The topic text is visible in each topic box. Trends are plotted across time continuously.}
+  \label{fig:zoomedsmear}
+\end{figure}
 
@@
 gaps will be left due to non-optimal stacking, in practice there are
-lot of small trends (90\% of all trends contain 1 topic) that fill in
+many small trends (90\% of all trends contain 1 topic) that fill in
 these gaps quickly.  Different instances of the same trend share the
 same color; apart from that, the color of trends is randomly chosen
-and color similarity is not meaningful.  The compact-trend-view is
-convenient because it shows all of the topic information, as shown by
-the zoomed-in view of MaxDB 7.500's compact-trend-view in Figure
-\ref{fig:zoomedsmear}. The \emph{compact-trend-view} makes repeating
+and color similarity is not meaningful.  
+%The compact-trend-view is
+%convenient because it shows all of the topic information, as shown by
+%the zoomed-in view of MaxDB 7.500's compact-trend-view in Figure
+%\ref{fig:zoomedsmear}. 
+The \emph{compact-trend-view} makes repeating
 continuous trends easy to pick out, although discontinuous trends are
@@
   \centering
-  \includegraphics[width=0.42\textwidth]{fixed-time-smear-plot-cropped}
-  \caption{A zoomed-in slice of compact-trend-view of topics per month of MaxDB 7.500. The topic text is visible in each topic box. Trends are plotted across time continuously.}
-  \label{fig:zoomedsmear}
-\end{figure}
-
-\begin{figure}
-  \centering
   \includegraphics[width=0.45\textwidth]{class-smear-plot-crop-scaled}
@@
 words and then two of the most common words, the global analysis
-finally found a topic with optimizer in its topic ten words. Our
+finally found a topic with optimizer in its top 10 words. Our
 analysis shows that optimizer was important but it had been obscured

### 2009-06-29 lda-paper.tex
made it fit to 10 pages!

@@
 comments that are added when revisions to the project are committed.
+% For each commit log comment, we count the occurrence of each word in
+% the message, remove stop words, and then produce word distributions
+% for each message.  These distributions are then normalized by the size
+% of the message, each count was divided by the total number of
+% tokens. After all messages are processed the distributions are
+% extended to include all words from all of the distributions.
+% For each commit log comment, we count the occurrence of each word in
+% the message, remove stop words, and then produce word distributions
+% for each message, which are normalized (divided by) by the size
+% of the message.
+% After all messages are processed the distributions are
+% extended to include all words from all of the distributions.
+% 
+%
 For each commit log comment we produce word distributions by counting
 the occurrence of each word in the message, removing stop words and
-normalizing the distribution by the size of message, in tokens.
-% (dividing counts by the number of tokens).  
+then normalizing the distribution by the size of message, in tokens.
+% (dividing word counts by the number of tokens).  
 After

### 2009-06-29 lda-paper.tex
migod's edits

@@
 %  Indexing (LSI) can be used to extract a set of independent topics
-  Indexing (LSI) extract a set of independent topics
+  Indexing (LSI) can be used to extract a set of independent topics
   from a corpus of commit-log comments.  Previous work in the area has
@@
 consecutive windows, or recurring periodically.  Alternatively, we
-might explore if any topics often recur together, if one particular
+might explore evolutionary patterns, such as if topics often recur together, if one particular
 topic usually precedes another, if unplanned topics outnumber planned
 topics, if particular topics recur near the end of development
-iterations, etc.  whereas this paper suggests that most topics are
-local.
+iterations, etc. % whereas this paper suggests that most topics are
+%local.
 
@@
 recur, or occur over a larger period are plotted continuously. In our
-example we have titled each topic with a word drawn from its
+example we have titled each topic with a word chosen from its
 word distribution.  
@@
 \begin{itemize}
-\item We demonstrate that windowed topic analysis most topics are local
-  in scope, which would be missed by a global topic analysis.
+%\item We demonstrate that windowed topic analysis most topics are local
+%  in scope, which would be missed by a global topic analysis.
+
+\item  We introduce windowed topic analysis, and demonstrate its utility
+  compared to global topic analysis.
+
 \item We present a number of visualizations of topics and their trends over time to aid
@@
 
-Lukins, Kraft and Etzkorn~\cite{lukins2008} use LDA to help bug
+Lukins et al.~\cite{lukins2008} use LDA to help bug
 localization by querying for documents that are related to a bug's
@@
 topics if it shared words from both topics.  And a commit that did not
-concern bugs, GUIs, or documentation would likely be related to the other topic.
+concern bugs, GUIs, or documentation would likely be related to the ``other'' topic.
 
@@
 each window and extract the top $N$ topics.  We decided to use a
-threshold of $20$ topics based on our previous experience with LDA; we
+threshold of $N=20$ topics based on our previous experience with LDA; we
 found that using more than $20$ makes the boundaries between topics less
@@
   per period were there?  What were the repeating trends?  Did the
-  trends dominate, or did local topics dominate?  How broken up were
+  trends dominate, or did local topics dominate?  How contiguous were
   the trends?  
-  To address these multiple questions we will describe and employ multiple visualizations.
+  To address these questions we will describe and employ multiple visualizations.
 %  Since we have a variety of questions to answer we suggest multiple
@@
 
-Adding and removing stop words produced different results. Our choice
-of stop words could be biased, and could affect the results.
+Our choice of the number of topics and adding and removing stop words
+produced different results. Our choice of stop words could be biased,
+and could affect the results.
 %, and our bias for or against such stop words could affect

### 2009-07-01 lda-paper.tex
Greying some images and describing it

@@
     topics occurring at the same time. Trends that are continuous are
-    plotted as continuous blocks. The top 10 words in the topics are
+    plotted as continuous blocks. Trends with more than one topic are colored the same unique color, while topics that do not reoccur are colored grey.
+    The top 10 words in the topics are
     joined and embedded in box representing the topics. No stop words were removed. The gap in the
@@
 these gaps quickly.  Different instances of the same trend share the
-same color; apart from that, the color of trends is randomly chosen
-and color similarity is not meaningful.  
+same color; apart from that, the color of a trend is randomly chosen
+while the color of a unique topic is grey. Color similarity is not meaningful.  
 %The compact-trend-view is
@@
 standards.
-
+\newline\noindent{\small\textbf{Acknowledgements:} Thanks to Tudor Girba and Neil Ernst.}
+%for valuable feedback
 \shrinkit

### 2009-07-01 lda-paper.tex
Some more edits

@@
 independent word distributions (topics) from
-documents (commit log comments)~\cite{1374321-short,10.1109/ICPC.2007.13-short,lukins2008,10.1109/MSR.2007.20-short}.
+documents (commit log comments)~\cite{1374321-short,10.1109/ICPC.2007.13-short,lukins2008-short,10.1109/MSR.2007.20-short}.
 Ideally these extracted topics 
@@
 Previous work on topic analysis has been applied to the entire history of a 
-project~\cite{1321709-short,lukins2008,10.1109/ICPC.2007.13-short}
+project~\cite{1321709-short,lukins2008-short,10.1109/ICPC.2007.13-short}
 to produce a single set of topics. In this paper we explore the idea
@@
 model, a combination of topics.  LDA has been used in software engineering
-literature~\cite{lukins2008,10.1109/MSR.2007.20-short}%,NIPS2007637,1321709}.
+literature~\cite{lukins2008-short,10.1109/MSR.2007.20-short}%,NIPS2007637,1321709}.
 to extract topics from documents such as methods, bug
@@
 
-Lukins et al.~\cite{lukins2008} use LDA to help bug
+Lukins et al.~\cite{lukins2008-short} use LDA to help bug
 localization by querying for documents that are related to a bug's
@@
 same color; apart from that, the color of a trend is randomly chosen
-while the color of a unique topic is grey. Color similarity is not meaningful.  
+while topics that do not reoccur are colored grey.
+%the color of a unique topic is grey. 
+Color similarity is not meaningful.  
 %The compact-trend-view is
@@
 \newline\noindent{\small\textbf{Acknowledgements:} Thanks to Tudor Girba and Neil Ernst.}
-%for valuable feedback
+%for their valuable feedback.}
 \shrinkit

### 2009-07-01 lda-paper.tex
ack change
Thank yous
Final edits!

@@
 What other tasks did they work on?  Which topics seem to recur over
-and over between cycles?  How long do different topics persist?  Are
+and over between releases?  How long do different topics persist?  Are
 there latent cause-effect relationships between different topics?
@@
 model, a combination of topics.  LDA has been used in software engineering
-literature~\cite{lukins2008-short,10.1109/MSR.2007.20-short}%,NIPS2007637,1321709}.
-to extract topics from documents such as methods, bug
+literature~\cite{lukins2008-short,10.1109/MSR.2007.20-short} to %,NIPS2007637,1321709}.
+extract topics from documents such as methods, bug
 reports, source code and files.
@@
 extracted from SCS, Linstead et al.\  proposed an author-source-code model
-using LDA~\cite{10.1109/MSR.2007.20-short}.%,NIPS2007637,1321709}. 
-This model associates authors and topics via their associated documents.
+using LDA~\cite{10.1109/MSR.2007.20-short}. This %,NIPS2007637,1321709}. 
+model associates authors and topics via their associated documents.
 
@@
 
-Grant et al.~\cite{scottcordy} and have used an alternative technique,
+Grant et al.~\cite{scottcordy-short} and have used an alternative technique,
 called Independent Component Analysis~\cite{delac-short} to separate topic signals from
@@
 documentation updates relating to the system and the GUI. To use LDA
-we would get the word distributions of each commit messages, 
+we would get the word distributions of each commit message, 
 \begin{comment}
@@
 %they are related to version control usage.  
-There were topics which appeared only during some transitional period and never again
+There were topics which appeared only during some transitional period and never again,
 such as \emph{RENAME} and \emph{BitKeeper}. \emph{BitKeeper} which appears when MySQL switched to  BitKeeper for their SCS.
@@
 the word distribution share the same words the purpose extracted from
-top 10 words was usually accurate.
+the top 10 words was usually accurate.
 
@@
   \includegraphics[width=0.9\textwidth]{commit-to-topics} 
-  \caption{How commits are analyzed and aggregated into topics and trends: commits are first extracted, then abstracted into word counts or word distributions which are then given to topic analysis tool like LDA.
+  \caption{How commits are analyzed and aggregated into topics and trends: commits are first extracted, then abstracted into word counts or word distributions which are then given to a topic analysis tool like LDA.
 LDA finds independent word distributions (topics) that these documents are related to (the numbers indicate similarity between documents and topics).}
@@
 the occurrence of each word in the message, removing stop words and
-then normalizing the distribution by the size of message, in tokens.
+then normalizing the distribution by the size of the message, in tokens.
 % (dividing word counts by the number of tokens).  
@@
 top 10 matching words.  10 words were chosen because often people care
-about top 10 comparison and it allowed for some common words to exist
+about top 10 comparisons and it allowed for some common words to exist
 %but not cause topics to match so closely.  
@@
 %       1. [ ] need the visualizer
-  Visualization is an integral part of our topics analysis which
+  Visualization is an integral part of our topic analysis which
   allows us to quickly explore the topics and trends of a project.
@@
 the \emph{trend-time-line}, shown in Figure \ref{fig:trendtimeline}, where
-each trend gets it own distinct y-value, while the x-axis is time;
+each trend gets its own distinct y-value, while the x-axis is time;
 these topics are then each plotted on their own line across time as
@@
 allotted space, it is often best to crop off the trends with only one
-topic ($1\%$ to $10\%$ of the total topics), otherwise the tail is
+topic ($90\%$ to $99\%$ of the total topics), otherwise the tail is
 long. The \emph{trend-histogram} summarizes the distribution of
@@
   \caption{Trend-time-line: Trends plotted per month of MaxDB 7.500.
-Time in months are plotted along the x-axis, each row on the y-axis is associated with a trend ranked by size in descending order}
+Time in months are plotted along the x-axis, each row on the y-axis is associated with a trend ranked by size in descending order.}
   \label{fig:trendtimeline}
@@
 
-When we examined PostgreSQL's history from 1996 to 2004, which includes
+We examined PostgreSQL's history from 1996 to 2004, which includes
 over $20,000$ commits. We did not find many trends with two or more
@@
 \emph{fix}, \emph{update}.  By decreasing the similarity constraint to
-$1/2$ the largest most common trend, which stretched across
+$1/2$, the largest most common trend, which stretched across
 the entire development, contained these same words (\emph{patch},
@@
 topics refer to portability with Win32, Central Europe Time (CEST)
-from email headers, issues with ALTER TABLE and CVS branch merging
+from email headers, issues with ALTER TABLE, and CVS branch merging
 (CVS does not record merges explicitly).
@@
 2005, and its second period was from June 2005 to June 2006. There
-were a total of $8 600$ commits analyzed.
+were a total of $8,600$ commits analyzed.
 
@@
 \ref{fig:statictopics}, explained in Section \ref{sec:compare}) it is not in any of the topics. A query
-optimizer is an important component of an DBMS, but as we have shown
+optimizer is an important component of a DBMS, but as we have shown
 it does not appear as a topic on its own. We tried to remove words to
@@
 messages, but would have been noticed using the more local topic
-analysis such as our windowed topic analysis, which used a window of
+analysis such as our windowed topic analysis, which used a smaller window of
 messages.
@@
 reoccurs over time. These visualizations help us  understand the common topics that developers focus on during
-development. If implemented interactively an user could easily zoom in
+development. If implemented interactively, a user could easily zoom in
 and query for a summary of a topic or trend.
@@
 standards.
-\newline\noindent{\small\textbf{Acknowledgements:} Thanks to Tudor Girba and Neil Ernst.}
+%\newline\noindent{\small\textbf{Acknowledgements:} We'd like to thank Tudor Girba and Neil Ernst.}
+%\newline\noindent{\small\textbf{Acknowledgements:} We thank Tudor Girba for his comments.}
+\newline\noindent{\small\textbf{Acknowledgements:} We would like to thank Tudor Girba and Neil Ernst for their feedback and ideas.}
+%\newline\noindent{\small\textbf{Acknowledgements:} We thank Tudor Girba and Neil Ernst.}
+%We'd like to thank Tudor Girba and Neil Ernst.}
 %for their valuable feedback.}
@@
 \bibliographystyle{latex8}
-
+\shrinkit
 \shrinkit

### 2009-08-13 lda-experiment.tex
lda experiment start

\documentclass{article}

\begin{document}

\section{Word lists}

Make a word list (a file) of related words and a label. Used wordnet
similarity (unfiltered) and concepts from \cite{5072519}.

\end{document}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: t
%%% End:

@INPROCEEDINGS{5072519,
title={Towards an Ontology for Software Product Quality Attributes},
author={Kayed, A. and Hirzalla, N. and Samhan, A.A. and Alfayoumi, M.},
booktitle={Internet and Web Applications and Services, 2009. ICIW '09. Fourth International Conference on},
year={2009},
month={May},
volume={},
number={},
pages={200-204},
abstract={Recently, quality assurance concept has been developed increasingly to be included in many of our life existing fields; financial, industrial, trading, computing, etc. Software quality product attributes (SWQAs) have been created as a matter of applying the QA concept on the results of Web or desktop application development process, to fit the products with the organizational and global market standards and goals, and to provide it with a competitive advantage value. Web application or software product quality is composed of many attributes such as portability, usability, reliability, modularity. During the recent years, many researchers discussed and presented software attributes in their works which showed that till now there is a lack of consensus on the semantic of many of concepts and terminologies used in this field. Our work is focusing on studying software product quality attributes concepts and terminologies. We conduct several experiments to extract the main concepts for SWQAs. The results show that there is a number of concepts that are frequently used to describe these attributes. Summarizing and formalizing the semantic of the attributes into these concepts presents a common understanding and agreement on the semantic of SWQPAs which can be used by software engineers, researchers, practitioners, and stakeholders.},
keywords={quality assurance, software portability, software quality, software reliability, standardsWeb application development process, competitive advantage value, desktop application development process, global market standards, modularity, organizational market standards, portability, quality assurance, reliability, software product quality attributes, usability},
doi={10.1109/ICIW.2009.36},
ISSN={}, }

### 2009-08-13 src/WordnetOcaml-0.1/related.sh

@@ -1,0 +1,2 @@
+export WNSEARCHDIR=/usr/share/wordnet/
+./related-words $*

### 2009-08-13 src/WordnetOcaml-0.1/run.sh

@@ -1,0 +1,3 @@
+export WNSEARCHDIR=/usr/share/wordnet/
+#make demo
+./demo $*

### 2009-08-13 src/exp1/exp1-gen.sh

@@ -1,0 +1,19 @@
+./related.sh maintainability maintainenance				> wordlist.maintainability
+./related.sh efficiency efficient > wordlist.efficiency
+./related.sh usability  usable                             > wordlist.usability             
+./related.sh performance                            > wordlist.performance
+./related.sh portability portable                            > wordlist.portability           
+./related.sh security                                > wordlist.security              
+./related.sh reliability reliable                            > wordlist.reliability           
+./related.sh understandability  understand                     > wordlist.understandability     
+./related.sh testability  testable test                           > wordlist.testability           
+./related.sh integrity                               > wordlist.integrity             
+./related.sh verifiability verification verify                          > wordlist.verifiability         
+./related.sh correctness    correct                         > wordlist.correctness           
+./related.sh traceability   trace   tracing                      > wordlist.traceability          
+./related.sh flexibility    flexibile                         > wordlist.flexibility           
+./related.sh modifiability   modifiable                        > wordlist.modifiability         
+./related.sh accuracy      accurate                          > wordlist.accuracy              
+./related.sh modularity      module                        > wordlist.modularity            
+./related.sh interoperability   interoperate                     > wordlist.interoperability      
+./related.sh availability      available                      > wordlist.availability          

### 2009-08-13 src/exp1/related.sh

@@ -1,0 +1,4 @@
+#!/bin/bash
+pushd ../WordnetOcaml-0.1 > /dev/null
+./related.sh $*
+popd > /dev/null

### 2009-08-15 lda-experiment.tex
More expiremental stuff

@@
 similarity (unfiltered) and concepts from \cite{5072519}.
+
+\subsection{Observations}
+
+We ran it and we found that wordnet for correctness was too
+generic. We ran it on MaxDB 7.500 with the words: integrity, security,
+interoperability, testability, maintainability, traceability,
+accuracy, modifiability, understandability, availability, modularity,
+usability, correctness, performance, verifiability, efficiency,
+portability, flexibility, and reliability.
+
+We had 500 topics extracted.
 

### 2009-08-15 src/exp2/names.py

@@ -1,0 +1,84 @@
+class Taxonomy():
+    """ class to store lists of various terms of interest. Each element/term in the list will be queried once."""
+    #TODO account for misspelinges 
+
+    usability_spell = ['usbility', 'useability',]
+    usability_syn = ['usability', 'serviceability', 'serviceableness', 'usableness', 'useableness'] #synonyms
+    usability_hyper = ['utility', 'usefulness'] #hypernyms
+    usability_deriv = ['serviceable', 'usable', 'useable'] # derived forms
+    usability_mccall = ['flexibility', 'interface'] #mccall defns
+    usability_meronym = ['Learnability', 'Understandability', 'Operability'] #as defined in iso9126
+    usability_user = ['screen', 'user', 'friendly', 'convention', 'human', 'default', \
+                        'click', 'guidelines', 'dialog', 'ugly', 'icons', 'ui', 'focus', \
+                        'feature', 'standard', 'convention', 'configure', 'menu', 'accessibility', 'gui'] #from mailing list
+    usability =  usability_mccall + usability_user + usability_syn + usability_hyper + usability_deriv + usability_meronym #+ usability_spell
+    
+    functionality_spell = []
+    functionality_mccall = ['accuracy', 'correctness']
+    functionality_user = ['vulnerability', 'secure', 'accurate', 'vulnerability', 'vulnerable', 'trustworthy', 'malicious', \
+                            'policy', '"buffer overflow"', 'secured', 'certificate', 'exploit', 'compliant' ]
+    functionality_syn = ['functionality']
+    functionality_hyper = ['practicality']
+    functionality_deriv = ['functional']
+    functionality_meronym = ['Suitability', 'Interoperability', 'Accuracy', 'Compliance', 'Security'] #as defined in iso9126
+    functionality = functionality_mccall + functionality_user + functionality_syn + functionality_hyper + functionality_deriv + functionality_meronym #+ _spell
+       
+    reliability_spell = []
+    reliability_syn = ['dependability', 'dependableness', 'reliability', 'reliableness']
+    reliability_hyper = ['responsibility', 'responsibleness']
+    reliability_mccall = ['integrity', 'resilience']
+    reliability_deriv = ['dependable', 'reliable']
+    reliability_user = ['failure', 'error', 'redundancy', 'fails', 'bug', 'crash', 'stable', 'stability']
[…]

### 2009-08-15 src/hiraldo_grok/hs/ATime.hs

@@ -1,0 +1,144 @@
+module ATime where
+import Time
+import System.Time
+
+-- Todo remove time, set it day year month
+
+--data atime = AMonth { year :: Int, 
+--                      month :: Time.Month } |
+--             AYear { year :: Int } |
+--             ADay { time :: CalendarTime }
+--                  deriving (Eq, Ord, Read, Show)
+data ATime = 
+    AMonth { year :: Int
+           , month:: Time.Month 
+           } |
+    AYear { year :: Int } |
+    ADay  { year :: Int, 
+            month :: Time.Month, 
+            day :: Int }
+          deriving (Ord, Read, Show)
+
+fromPOSIXTime time = let nt = (toUTCTime (System.Time.TOD time 0))
+                         day = (Time.ctDay nt)
+                         month = (Time.ctMonth nt)
+                         year = (Time.ctYear nt)
+                     in ADay { year = year , month = month , day = day }
+fromPOSIXTimeToDay = fromPOSIXTime 
+fromPOSIXTimeToMonth = toAMonth . fromPOSIXTime
+fromPOSIXTimeToYear = toAYear . fromPOSIXTime
+
+toAMonth x@(AMonth { year = year , month = month }) = x
+toAMonth (AYear { year = year }) = AMonth { year = year , month = Time.January }
+toAMonth (ADay { year = year, month = month, day = day }) = AMonth { year = year, month = month }
+
+ymdToCalendarTime year month day = ct
+    where ct = CalendarTime {     ctYear  = year
+                            ,     ctMonth = month
+                            ,     ctDay = day 
+                            ,     ctHour = 0
+                            ,     ctMin = 0
+                            ,     ctSec = 0
+                            ,     ctPicosec = 0
+                            ,     ctWDay = Sunday -- lol
+                            ,     ctYDay = 0 -- lol 
+                            ,     ctTZName = "" -- lol
+                            ,     ctTZ = 0 -- lol
+                            ,     ctIsDST = False -- lol
+                            }
+
+toADay (AMonth { year = year , month = month }) = ADay { year = year, month = month, day = 1 }
+toADay (AYear { year = year }) = ADay { year = year, month = January, day = 1 }
+toADay day@(ADay { year = year, month = month, day = day'}) = day
+
+toAYear year@(AYear _) = year
[…]

### 2009-08-15 src/hiraldo_grok/hs/EqClass.hs

@@ -1,0 +1,75 @@
+module EqClass where
+import Data.Graph
+import Data.Map
+import Data.Set
+import Array
+
+--
+-- ** Labnotes <2008-11-12 Wed 00:10> Global versus local comparison?     :smearer:
+--    I think there's some problem with the algorithm, there might be a difference between local
+--    comparison and global. Maybe pre-coloring would be a better idea, where topics are
+--    globally compared to each, equivilence classes found and then maybe it'll be easier
+--    to find smears
+--  So we're here to find equivilance classes where if you are reachable (like a fill flood)
+--  you are color'd the same
+
+-- notes: reachable means itself and itschildren and their children
+
+-- getEqClasses:: (a -> a -> Bool) -> [a] -> [[a]]
+-- getEqClasses simf elms =
+reachableFromRoot simf elms =
+    let len = length elms
+        simftuple (_,x) (_,y) = simf x y                                
+        enumelms = zip [1..len] elms
+        simlist = Prelude.map (\(i,elm) -> (i,elm,(Prelude.filter (simftuple (i,elm)) enumelms))) enumelms
+        -- simlist = [ (elm,sims) | elm <- enumelms, sims <- (filter (simftuple elm) enumelms) ] -- could be smarter, do we care?
+        edgelist = concatMap (\(i,_,ls) -> 
+                                  Prelude.map (\(j,_) -> (i,j)) ls) simlist
+        graph = Data.Graph.buildG (1,len) edgelist
+    in
+      reachable graph 1
+
+
+getEqClasses:: Ord a => (a -> a -> Bool) -> [a] -> (Int,(Data.Map.Map a Int))
+getEqClasses simf elms =
+    let len = length elms
+        simftuple (_,x) (_,y) = simf x y                                
+        enumelms = zip [1..len] elms
+        elmarray = array (1,len) enumelms
+        simlist = Prelude.map (\(i,elm) -> (i,elm,(Prelude.filter (simftuple (i,elm)) enumelms))) enumelms
+        -- simlist = [ (elm,sims) | elm <- enumelms, sims <- (filter (simftuple elm) enumelms) ] -- could be smarter, do we care?
+        halflist = concatMap (\(i,_,ls) -> 
+                                  Prelude.map (\(j,_) -> (i,j)) ls) simlist
+        edgelist = halflist ++ (Prelude.map (\(i,j) -> (j,i)) halflist)
+        graph = Data.Graph.buildG (1,len) edgelist
[…]

### 2009-08-15 src/hiraldo_grok/hs/Lda_layout.hs

@@ -1,0 +1,467 @@
+module Lda_layout where
+import Array
+import Lda_parser
+import System.Random
+import Data.List
+import Data.Map (Map)
+import qualified Data.Map as Map
+
+intToDouble:: Int -> Double
+intToDouble x =    fromInteger  $ toInteger x
+
+
+
+tinyTopicFontSize = 10.0
+bigTopicFontSize = 30.0
+defaultSplotchWidth = 150.0
+defaultSplotchHeight = 150.0
+
+type Color = (Double,Double,Double)
+
+data Point = Point Double Double deriving (Ord, Show, Eq)
+data Bounds = Bounds Double Double deriving (Ord, Show, Eq)
+data Region = Region Point Bounds deriving (Ord, Show, Eq)
+-- The bounds of a label Include the space for the text
+data DrawObj = DrawString String Double Point Bounds |
+               Label DrawObj Point Bounds | -- DrawString
+               Smear (Double,Double,Double) DrawObj [DrawObj] Region | -- DrawString Label and Topic
+               DTopic [DrawObj] Region |
+               Box (Double, Double, Double) Region |
+               Translate DrawObj Point
+
+drawObjWidth (DrawString _ _ _ (Bounds w _)) = w
+drawObjWidth (Label _ _ (Bounds w _)) = w
+drawObjWidth (Smear _ _ _ (Region _ (Bounds w _))) = w
+drawObjWidth (DTopic _ (Region _ (Bounds w _))) = w
+drawObjWidth (Box _ (Region _ (Bounds w _))) = w
+drawObjWidth (Translate x _) = drawObjWidth x
+
+drawObjRegion (DrawString _ _ pts bds) = Region pts bds
+drawObjRegion (Label _ pts bds ) = Region pts bds
+drawObjRegion (Smear _ _ _ r) = r
+drawObjRegion (DTopic _ r) = r
+drawObjRegion (Box _ r) = r
+drawObjRegion (Translate v pt) = Region pt bds
+    where (Region _ bds) = drawObjRegion v
+
+-- makes an assumption of 0,0 existing
+enclosingRegion objs =
+    let l = map drawObjRegion objs in
+    foldl combineRegions (Region (Point 0 0) (Bounds 0 0)) l
+
+combineRegions (Region (Point x1 y1) (Bounds w1 h1)) (Region (Point x2 y2) (Bounds w2 h2)) =
+    let nx = min x1 x2
[…]

### 2009-08-15 src/hiraldo_grok/hs/Lda_parser.hs

@@ -1,0 +1,562 @@
+module Lda_parser where
+import System.IO
+import System.Locale
+import System.Time
+import Data.List
+import Time
+import Data.Map (Map)
+import qualified Data.Map as Map
+import Char
+import Monad
+import Array
+import EqClass
+
+data Window = EmptyWindow | TopicWindow String [[String]]
+
+showsTopicWindow (TopicWindow time topics) = "TopicWindow [" ++ time ++ "] [" 
+                                                            ++ (concat (map unwords topics))
+                                                            ++ "]"
+
+showsTopicWindow (EmptyWindow) = "EmptyWindow"
+
+instance Show Window where 
+    show = show . showsTopicWindow
+
+
+
+data Topic = Topic [String]
+
+showsTopic (Topic x) = "Topic "++(show x)
+instance Show Topic where 
+    show = show . showsTopic
+
+
+topicCompare (Topic topic1) (Topic topic2) = compare topic1 topic2
+-- I don't like this, also, I think topics should be sets.
+topicEq (Topic topic1) (Topic topic2) = topic1 == topic2
+instance Eq Topic where
+    (==) = topicEq
+instance Ord Topic where
+    compare = topicCompare
+
+
+topicWords (Topic words) = words
+
+isEmptyTopic EmptyWindow = True
+isEmptyTopic _  = False                 
+
+addTopic (TopicWindow topicTime wordwords) topicNumber words = 
+    if (timeToYearInt (strToTime topicTime) < 1981) 
+    then error (show (topicTime,topicNumber,words))
+    else (TopicWindow topicTime (words:wordwords))
+
+makeTopicWindow time number words  = 
+    if (timeToYearInt (strToTime time) < 1981) 
+    then error (show (time,number,(TopicWindow time [words])))
+    else (TopicWindow time [words])
+              
+
+emptyTopicWindow = EmptyWindow
+--getTopics (TopicWindow time words)
+getTopics (TopicWindow time words) = words
+getTime (TopicWindow time words) = time
+
+getIDTopics (TopicWindow time words) = map (\x -> (time, x)) words
+
+-- Start w/ EmptyWindow and lines
+lineParser EmptyWindow [] = [] 
+lineParser x [] = [x]
+lineParser current (x:xs) = 
+    case words x of
+      []  -> current : (lineParser emptyTopicWindow xs)
[…]

### 2009-08-15 src/hiraldo_grok/hs/Lda_relator.hs

@@ -1,0 +1,149 @@
+import Lda_parser
+import List
+import Data.List
+import Directory
+import Char
+import Monad
+
+data WordList = WordList String [String]
+wordListName (WordList name _) = name
+wordListWords (WordList _ words) = words
+
+-- ranks by number of hits
+topicWordListSimilarity (Topic topicWords) (WordList name words) =
+    length $ List.intersect topicWords words
+
+
+
+-- Ok left to do 
+-- it should work?
+-- it has to load the wordlist and iterate per each topic seeing if it is similar
+-- then we'll make it print a report
+
+sim = 7 / 10
+
+getWordListFileNames () = do
+  d <- getCurrentDirectory
+  l <- getDirectoryContents d
+  return (filter (isPrefixOf "wordlist.") l)
+
+getWordLists () = do
+  fileNames <- getWordListFileNames ()
+  wordList <- mapM parseWordList fileNames
+  return wordList
+
+processWord word = 
+    map (\c -> if (c == '_') then ' ' else toLower c) word
+
+parseWordList file = do
+  contents <- readFile file
+  let lines = Data.List.lines contents
+  let words = map processWord lines 
+  return (WordList (head words) (nub words))
+
+isTopicSimilarToWordList topic wordList =  (topicWordListSimilarity topic wordList) > 0
+
+topicRelations sim l wl =
+  let plda = processLDA sim l
+      -- topicmap is a Data.Map that maps from a topic to a color
+      (tp,maxcolor,topicmap) = plda
+      -- labels are basically the names of the months and years.. (great)..
+      -- smears are Topic Placements 
+      (TopicPlot labels smears) = tp
+      topics = concatMap placementGetTopics smears        
+      topicRelations = map (\topic -> 
+                                let relatedWordLists = filter (isTopicSimilarToWordList topic) wl
+                                in
+                                (topic, relatedWordLists) 
+                                
+                           ) topics
+  in
+    topicRelations
+
+presentTopicRelations tr =
+    map (\((Topic topicWords), wls) ->
+            concat [ "Topic: " , 
+                     (intercalate ", " $ map wordListName wls) , 
+                     " [ " , 
[…]

### 2009-08-15 src/hiraldo_grok/hs/Lda_report.hs

@@ -1,0 +1,491 @@
+module Lda_report where
+import System.IO
+import Data.List
+import Char
+import Monad
+import Array
+import Graphics.Rendering.Cairo
+import qualified Graphics.Rendering.Cairo.Matrix as M
+import Lda_parser
+import Lda_layout
+import EqClass
+import Data.Map (Map)
+import qualified Data.Map as Map
+import Control.Concurrent
+import Control.Parallel
+import Control.Parallel.Strategies
+
+textBounds (TextExtents _ _ w h _ _) = (w, h)
+
+replaceChar l c r = map (\x -> if (x == c) then r else x) l
+
+drawText size str x y = do
+  lineWidth <- getLineWidth
+  setFontSize  size
+  stroke
+  moveTo x y
+  textPath str
+  setSourceRGBA 0 0 0 1 -- text
+  fillPreserve
+  setSourceRGBA 0 0 0 0.5 -- text outline
+  setLineWidth 0.5 -- linewidth
+  stroke
+  return ()
+
+  
+
+drawTopics :: String -> [String] -> Double -> Double -> Render ()
+drawTopics text topics x y = do
+  save
+
+  lineWidth <- getLineWidth
+  setFontSize  bigTopicFontSize
+  bigextents <- fontExtents             
+  setFontSize  tinyTopicFontSize
+  tinyextents <- fontExtents             
+
+  let bigfontHeight = fontExtentsHeight bigextents
+  let tinyfontHeight = fontExtentsHeight tinyextents
+
+  (TextExtents xb yb w h _ _) <- textExtents text
+  setFontSize tinyTopicFontSize
+  textExtentsMonads <- mapM textExtents topics
+  let extents = map textBounds textExtentsMonads
+  let (topicw,topich) = if (extents == []) 
+                        then (0,0) 
+                        else foldlOnFirst (\(w,h) (mw,mh) -> ((max mw w),(max mh h))) extents
+  let tlen = length topics
+
+
+  rectangle (x + xb - lineWidth)
+                  (y + yb - lineWidth)
+                  (w + topicw + 5 * lineWidth)
+                  ((max h ((fromInteger (toInteger tlen)) * topich)) + 2 * lineWidth)
+  stroke
+  moveTo x y
+  setFontSize  bigTopicFontSize
+  textPath text
+  setSourceRGBA 0 0 0 1 -- text
+  fillPreserve
+  setSourceRGBA 0 0 0 0.5 -- text outline
+  setLineWidth 3.0
[…]

### 2009-08-15 src/hiraldo_grok/hs/Lda_report_driver.hs

@@ -1,0 +1,3 @@
+import Lda_report
+
+main = Lda_report.lda_report_main

### 2009-08-15 src/hiraldo_grok/hs/M2.hs

@@ -1,0 +1,4 @@
+module M2(tinyTopicFontSize,bigTopicFontSize,v) where
+tinyTopicFontSize = 100.0
+bigTopicFontSize = 400.0
+v x = tinyTopicFontSize * x

### 2009-08-15 src/hiraldo_grok/hs/M3.hs

@@ -1,0 +1,4 @@
+module M3 where
+import M2 hiding (tinyTopicFontSize)
+tinyTopicFontSize = 30.0
+main = putStr (show (v 10))

### 2009-08-15 src/hiraldo_grok/hs/OurTime.hs

@@ -1,0 +1,38 @@
+module OurTime where
+import System.Locale
+import System.Time
+import Time
+import Char
+import Data.List
+
+strToInt:: String -> Integer
+strToInt str =
+    let len = length str in
+    read (if (last str == '.') 
+          then take (len - 1) str
+          else str)
+strToTime str =
+    toUTCTime (TOD x 0) where 
+                   x = strToInt str
+        
+timeToMonthStr time =
+    formatCalendarTime System.Locale.defaultTimeLocale "%Y %b" time
+
+isFloatStr [] = True
+isFloatStr (x:xs) = if ((Char.isDigit x) || x == '.') 
+                            then isFloatStr xs
+                            else False
+isIntStr [] = True
+isIntStr (x:xs) = if (Char.isDigit x) 
+                            then isFloatStr xs
+                            else False
+
+epochToTimeString = timeToMonthStr . strToTime 
+
+timeTest = do
+  contents <- readFile "timetest.txt"
+  let lines = Data.List.lines contents in
+      return (map epochToTimeString lines)
+
+
+    

### 2009-08-15 src/hiraldo_grok/hs/RandTest.hs

@@ -1,0 +1,6 @@
+import System.Random
+lol = [1..]
+z = [(i,j,k) | x <- lol, i <- (3*(x-1) + 1), j <- (i + 1), k <- (j + 1)]
+
+rs = randoms (mkStdGen 4)
+main = do print $ take 10 (rs :: [Double])

### 2009-08-15 src/hiraldo_grok/hs/Ranked.hs

@@ -1,0 +1,151 @@
+module Ranked where
+import System.IO
+import System.Locale
+import System.Time
+import Time
+import Data.List
+import Data.Map (Map)
+import qualified Data.Map as Map
+import Time
+import Char
+import Monad
+import Array
+import ATime
+import Lda_layout
+import Lda_report
+
+read_ocaml_float :: String -> Double
+read_ocaml_float str =
+    let len = length str in
+    read (if (last str == '.') 
+          then take (len - 1) str
+          else str)
+
+
+parse_ranked_docs filename = do
+  contents <- readFile filename
+  let lines = Data.List.lines contents in
+    return (map (\line ->
+                      case words line of
+                        []    -> error "Bad line, no elements"
+                        x:y:_ -> (((read x)::Integer) , (floor (read_ocaml_float y)))) lines)
+
+-- Stupid Ord
+-- (Int, Integer) is TOPIC, TIME
+-- output is (bucketlabel, [(Topic Time)])
+-- This is ignoring topics
+-- bucketize_by :: (Ord a) => ((Integer, Integer) -> a) -> [(Integer, Integer)] -> [ ( a,  [(Integer, Integer)]) ]
+bucketize_by f l = buckets
+    where m = foldl (\m x -> 
+                       let key = (f x) 
+                       in Map.insert key (x : (Map.findWithDefault [] key m)) m)
+                   Map.empty l
+          buckets = Map.toAscList m
+
+
+bucket_key_day (topic,ptime) = toUTCTime (System.Time.TOD ptime 0)
+
+bucket_key (topic,ptime) = ATime.fromPOSIXTime ptime
+
+bucketize_by_month l = bucketize_by month_key l
+    where month_key = ATime.toAMonth . bucket_key 
+
+bucketize_by_year l = bucketize_by year_key l
+    where year_key = ATime.toAYear . bucket_key 
+
+bucketize_by_day l = bucketize_by day_key l
+    where day_key = bucket_key
+
+data PlotSettings = 
+    PlotSettings { pWidth :: Double
+                 , pHeight :: Double
+                 }
+
+
+-- addGaps takes 2 lists and shoves the missing elements of (in list y) into list x
+-- Assumptions:
+--  First list is increasing succ x > x
+--  Second list is increasing succ y > y
+--  x maps to y 1 to 1 but y can be larger
+--  y is a range of x (domain?)
[…]

### 2009-08-15 src/hiraldo_grok/hs/Ranked_driver.hs

@@ -1,0 +1,4 @@
+import Ranked
+
+main =  main_ranked ()
+

### 2009-08-15 src/hiraldo_grok/hs/getline.hs

@@ -1,0 +1,18 @@
+import System.IO
+
+slurpFile file = 
+    do fd <- openFile file ReadMode
+       let reader = do eof <- hIsEOF fd
+                       if eof
+                         then return []
+                         else (do x <- hGetLine fd
+                                  xs <- reader
+                                  return (x:xs))
+       do x <- reader 
+          hClose fd;
+          return x
+
+testSlurpFile = do lines <- slurpFile "bigfile" 
+                   putStrLn (lines!!100)
+
+main = testSlurpFile

### 2009-08-15 src/hiraldo_grok/hs/iotest.hs

@@ -1,0 +1,14 @@
+import Data.Char
+
+run =  getLine 
+       >>= \x -> 
+           let c = x!!2 in
+           putStrLn (if isLower c then "LOWER" else "UPPER")
+
+run2 = do x <- getLine
+          let c = x !! 2 in
+              putStrLn (if isLower c then "LOWER" else "UPPER")
+
+main = run2
+       
+

### 2009-08-15 src/hiraldo_grok/hs/mandelbrot.hs

@@ -1,0 +1,108 @@
+{-# LANGUAGE BangPatterns #-}
+{-# OPTIONS -O2 -funbox-strict-fields -fvia-C -optc-O2 -fexcess-precision #-}
+--
+-- The Computer Language Benchmarks Game
+-- http://shootout.alioth.debian.org/
+--
+-- Contributed by Spencer Janssen, Trevor McCort, Christophe Poucet and Don Stewart
+-- Parallelised by Tim Newsham
+--
+-- Must be compiled with the -fexcess-precision flag as a pragma. GHC
+-- currently doesn't recognise the -fexcess-precision flag on the command
+-- line (!).
+--
+-- The following flags are suggested when compiling:
+--
+--      ghc -optc-march=pentium4 -optc-mfpmath=sse -optc-msse2 -threaded --make
+--
+-- Run with -N6 on a quad core (more capabilities to hide latency)
+--
+--      $ time ./A 6400 +RTS -N6
+--
+
+import System
+import System.IO
+import Foreign
+import Foreign.Marshal.Array
+import Control.Concurrent
+import Control.Concurrent.MVar
+import Control.Concurrent.Chan
+import Control.Monad
+
+main = do
+    -- width in pixels
+    w <- getArgs >>= readIO . head
+        -- width in bytes
+    let n      = w `div` 8
+        -- width of a pixel in the complex plane
+        m  = 2 / fromIntegral w
+        coords = [T 1 0 y (fromIntegral y * m - 1) | y <- [0..w-1]]
+    q <- newChan
+    replies <- replicateM w newEmptyMVar
+    mapM_ (writeChan q) $ zip coords replies
+    replicateM_ 4 . forkIO $ worker q w m n
+
+    putStrLn ("P4\n"++show w++" "++show w)
+    mapM_ (takeMVar >=> \b -> hPutBuf stdout b n) replies
+
+-- Worker computes one line of the image and sends it to the master
+-- q - work queue
+-- w - width in pixels
+-- m - width of a pixel in the complex plane
+-- n - width in bytes
+worker q w m n = forever (do
+    (coord, reply) <- readChan q
+    p <- mallocArray0 n
+    unfold (next_x w m n) p coord
+    putMVar reply p)
+
+-- f - takes coordinates and returns Nothing if done
+--     or the next byte of the bitmap otherwise.
+-- ptr - buffer to write to
[…]

### 2009-08-15 src/hiraldo_grok/hs/permute.hs

@@ -1,0 +1,61 @@
+-- 
+--             DO WHAT THE FUCK YOU WANT TO PUBLIC LICENSE
+--                     Version 2, December 2004
+-- 
+--  Copyright (C) 2009 Abram Hindle
+--  Everyone is permitted to copy and distribute verbatim or modified
+--  copies of this license document, and changing it is allowed as long
+--  as the name is changed.
+-- 
+--             DO WHAT THE FUCK YOU WANT TO PUBLIC LICENSE
+--    TERMS AND CONDITIONS FOR COPYING, DISTRIBUTION AND MODIFICATION
+-- 
+--   0. You just DO WHAT THE FUCK YOU WANT TO. 
+-- 
+-- This program is free software. It comes without any warranty, to
+-- the extent permitted by applicable law. You can redistribute it
+-- and/or modify it under the terms of the Do What The Fuck You Want
+-- To Public License, Version 2, as published by Sam Hocevar. See
+-- http://sam.zoy.org/wtfpl/COPYING for more details. */
+
+module Permute where 
+import System.Random
+import qualified Data.Map as Map
+
+-- Taken from
+-- http://en.wikipedia.org/wiki/Fisher-Yates_shuffle
+
+--     public static void shuffle (int[] array) 
+--     {
+--         Random rng = new Random();   // i.e., java.util.Random.
+--         int n = array.length;        // The number of items left to shuffle (loop invariant).
+--         while (n > 1) 
+--         {
+--             int k = rng.nextInt(n);  // 0 <= k < n.
+--             n--;                     // n is now the last pertinent index;
+--             int temp = array[n];     // swap array[n] with array[k] (does nothing if k == n).
+--             array[n] = array[k];
+--             array[k] = temp;
+--         }
+--     }
+
+
+-- Given a seed, permute the list l 
+permuteList seed l = map snd $ loop len dm rl
+    where len = length l 
+          rl   = randoms (mkStdGen seed)
+          indexed_list = zip [0..] l
+          dm = Map.fromAscList indexed_list
+          -- this is basically a fold that decrements n each time
+          -- and pops a new random number off the random list 
+          loop n dm rl = 
+              if (n > 1) 
+              then let r' = mod (head rl) n  -- danger
+                       (Just x) = Map.lookup (n - 1) dm
+                       (Just y) = Map.lookup r' dm
+                       dm' = Map.insert (n - 1) y (Map.insert r' x dm) in
[…]

### 2009-08-15 src/hiraldo_grok/hs/regex-dna.hs

@@ -1,0 +1,44 @@
+-- The Computer Language Benchmarks Game
+-- http://shootout.alioth.debian.org/
+-- Contributed by: Sergei Matusevich 2007
+-- Parallelised by Tim Newsham using Haskell's parallel strategies
+
+import Control.Parallel.Strategies
+import List
+import Text.Regex.PCRE          -- requires haskell-regex-pcre-builtin
+import qualified Data.ByteString.Char8 as B
+
+variants = [
+  "agggtaaa|tttaccct",
+  "[cgt]gggtaaa|tttaccc[acg]",
+  "a[act]ggtaaa|tttacc[agt]t",
+  "ag[act]gtaaa|tttac[agt]ct",
+  "agg[act]taaa|ttta[agt]cct",
+  "aggg[acg]aaa|ttt[cgt]ccct",
+  "agggt[cgt]aa|tt[acg]accct",
+  "agggta[cgt]a|t[acg]taccct",
+  "agggtaa[cgt]|[acg]ttaccct" ]
+
+main = do
+  file <- B.getContents
+  let [s1,s2,s3] = map (B.concat . tail) $ groupBy notHeader $ B.split '\n' file
+      showVars r = r ++ ' ' : show ((s2 =~ r :: Int) + (s3 =~ r :: Int))
+  mapM_ putStrLn $ parMap rnf showVars variants
+  putChar '\n'
+  print (B.length file)
+  print (B.length s1 + B.length s2 + B.length s3)
+  print (B.length s1 + B.length s3 + length (B.unpack s2 >>= substCh))
+  where notHeader _ s = B.null s || B.head s /= '>'
+        substCh 'B' = "(c|g|t)"
+        substCh 'D' = "(a|g|t)"
+        substCh 'H' = "(a|c|t)"
+        substCh 'K' = "(g|t)"
+        substCh 'M' = "(a|c)"
+        substCh 'N' = "(a|c|g|t)"
+        substCh 'R' = "(a|g)"
+        substCh 'S' = "(c|g)"
+        substCh 'V' = "(a|c|g)"
+        substCh 'W' = "(a|t)"
+        substCh 'Y' = "(c|t)"
+        substCh etc = [etc]
+

### 2009-08-15 src/hiraldo_grok/hs/smearer.hs

@@ -1,0 +1,18 @@
+data Window time word = EmptyWindow | TopicWindow time [[word]]
+
+data Topic words = Topic words
+
+isEmptyTopic EmptyWindow = True
+isEmptyTopic _  = False                 
+
+addTopic (TopicWindow topicTime wordwords) topicNumber words = 
+    (TopicWindow topicTime (words:wordwords))
+
+makeTopicWindow time number words  = (TopicWindow time [words])
+
+emptyTopicWindow = EmptyWindow
+getTopics (TopicWindow time words)
+getTopics (TopicWindow time words) = words
+getTime (TopicWindow time words) = words
+
+getIDTopics (TopicWindow time words) = map (\x -> (time, x)) words

### 2009-08-15 src/hiraldo_grok/hs/spectral-norm.hs

@@ -1,0 +1,121 @@
+{-# LANGUAGE BangPatterns, MagicHash #-}
+{-# OPTIONS -fvia-C -fexcess-precision #-}
+--
+-- The Computer Language Benchmarks Game
+-- http://shootout.alioth.debian.org/
+--
+-- Translation from Clean by Don Stewart
+-- Modified by Ryan Trinkle: 1) change from divInt# to uncheckedIShiftRA#
+--                           2) changed -optc-O to -optc-O3
+--                           3) added -optc-ffast-math
+-- Parallelised by Don Stewart
+--
+-- Should be compiled with:
+--      -O2 -threaded --make
+--      -optc-O2 -optc-march=native -optc-mfpmath=sse -optc-msse2 -optc-ffast-math
+--
+
+import System
+import Foreign.Marshal.Array
+import Foreign
+import Text.Printf
+import Control.Monad
+import Data.ByteString.Internal
+
+import GHC.Base
+import GHC.Float
+import GHC.Int
+
+import Control.Concurrent
+import Control.Exception
+
+type Reals = Ptr Double
+
+main = do
+    n <- getArgs >>= readIO . head
+
+    u <- mallocArray n :: IO Reals
+    forM_ [0..n-1] $ \i -> pokeElemOff u i 1
+    v <- mallocArray n :: IO Reals
+    forM_ [0..n-1] $ \i -> pokeElemOff v i 0
+
+    powerMethod 10 n u v
+    printf "%.9f\n" (eigenvalue n u v 0 0 0)
+
+------------------------------------------------------------------------
+
+eigenvalue :: Int -> Reals -> Reals -> Int -> Double -> Double -> Double
+eigenvalue !n !u !v !i !vBv !vv
+    | i < n     = eigenvalue n u v (i+1) (vBv + ui * vi) (vv + vi * vi)
+    | otherwise = sqrt $! vBv / vv
+    where
+       ui = inlinePerformIO (peekElemOff u i)
+       vi = inlinePerformIO (peekElemOff v i)
+
+------------------------------------------------------------------------
+
+powerMethod :: Int -> Int -> Reals -> Reals -> IO ()
+powerMethod i n u v = do
+    -- roll our own synchronisation
+    children <- newMVar []
+    replicateM_ i $ do
+        me <- newEmptyMVar
+        cs <- takeMVar children
+        putMVar children (me : cs)
[…]

### 2009-08-15 src/hiraldo_grok/hs/test.hs

@@ -1,0 +1,18 @@
+import System.IO
+import Monad
+import List
+
+data Huh = Lol | Hy
+
+test1 Lol = Hy
+test1 Hy =  Lol
+test2 Lol = "Lol"
+test2 Hy = "Hy"
+
+p1 = sequence_ (map 
+                     (\x -> putStrLn (test2 (test1 x)))
+                     [ Hy, Lol, Hy, Lol ])
+  
+p2 = sequence_ (map (\x -> putStrLn (show x))  (intersect ["a","b","c"] ["b","c","c","d"]))
+
+main = p2

### 2009-08-15 src/hiraldo_grok/hs/testtime.hs

@@ -1,0 +1,14 @@
+module TestTime where
+import System.Time
+import Time
+import System.Locale
+
+main = 
+  let ti = toUTCTime (TOD 1234567890 0)
+      cs = Time.calendarTimeToString ti
+      cs2 = Time.formatCalendarTime System.Locale.defaultTimeLocale "%Y %M %D" ti
+  in do
+    print cs;
+    print cs2;
+    print (show (Time.ctMonth ti));
+    return ()

### 2009-08-15 lda-experiment.tex
more expirement and some more write up and reporting.

--- 2009-08-13 lda-experiment.tex
@@
 
-\section{Word lists}
+% \section{Word lists}
 
-Make a word list (a file) of related words and a label. Used wordnet
-similarity (unfiltered) and concepts from \cite{5072519}.
+% Make a word list (a file) of related words and a label. Used wordnet
+% similarity (unfiltered) and concepts from \cite{5072519}.
+
+% \subsection{Observations}
+
+% We ran it and we found that wordnet for correctness was too
+% generic. We ran it on MaxDB 7.500 with the words: integrity, security,
+% interoperability, testability, maintainability, traceability,
+% accuracy, modifiability, understandability, availability, modularity,
+% usability, correctness, performance, verifiability, efficiency,
+% portability, flexibility, and reliability.
+
+% We had 500 topics extracted.
+
+\section{Exp1: Word Lists on MaxDb 7.500 commit log comments}
+
+We extracted the commit log comments from MaxDB 7.500 and per month extracted 20 topics using LDA.
+There are 25 months, thus a total of topic 500 possible topics, but there were a couple of months with no commits.
+
+Make a word list (a file) of related words and a label. We used
+wordnet similarity (unfiltered) on concepts extracted from
+\cite{5072519}.  The concepts used included: integrity, security,
+interoperability, testability, maintainability, traceability,
+accuracy, modifiability, understandability, availability, modularity,
+usability, correctness, performance, verifiability, efficiency,
+portability, flexibility, and reliability.
+
+We determined if a word list was similar to a topic if they shared 1
+or more words in common.
+
+\subsection{Observations}
+
+We ran it and we found that wordnet generated word lists for
+\emph{correctness} were too large andgeneric. Results are listed below
+in the results section
+
+\subsection{Future attempts:}
+
+\begin{itemize}
+\item  Make better manual word lists
+\item  Take that ISO standard that Neil used
+\item  Take that ISO Architecture standard that ric mentioned
+\item  Different similarity measure
+\item  Utilize wordnet distance
+\item  Utilize LDA to query its own model with single wordlists to get related documents
+\end{itemize}
+
+\subsection{Results}
+
+Here's the output of what I got:
+\begin{verbatim}
+Total Topics: 500
+Total Named Topics: 281
+Total UnNamed Topics: 219
+Total Empty Topics: 80
+Total UnNamed Non-empty Topics: 139
+
+Wordlists associated with n topics:
+interoperability 1
+accuracy 3
+usability 4
+modularity 5
+understandability 6
+efficiency 8
+security 8
+integrity 9
+performance 58
+verifiability 78
+traceability 95
+testability 121
+correctness 182
+
+Now wordlists that were correlated with each other
+correctness, interoperability 1
+efficiency, performance 1
+efficiency, security 1
+efficiency, testability 1
+efficiency, verifiability 1
+integrity, traceability 1
+integrity, verifiability 1
+interoperability, interoperability 1
+modularity, performance 1
+performance, security 1
+performance, usability 1
+traceability, understandability 1
+traceability, usability 1
+understandability, verifiability 1
+efficiency, traceability 2
+security, testability 2
+testability, usability 2
+accuracy, accuracy 3
+accuracy, correctness 3
+accuracy, testability 3
+accuracy, understandability 3
+integrity, testability 3
+correctness, integrity 4
+security, verifiability 4
+usability, usability 4
+correctness, efficiency 5
+correctness, modularity 5
+correctness, security 5
+correctness, understandability 5
+modularity, modularity 5
+security, traceability 5
+testability, understandability 5
+understandability, understandability 6
+efficiency, efficiency 8
+security, security 8
+integrity, integrity 9
+performance, traceability 10
+performance, verifiability 12
+performance, testability 17
+correctness, performance 24
+testability, verifiability 26
+testability, traceability 32
+performance, performance 58
+correctness, testability 62
+correctness, verifiability 76
+traceability, verifiability 76
+verifiability, verifiability 78
+correctness, traceability 86
+traceability, traceability 95
+testability, testability 121
+correctness, correctness 182
+\end{verbatim}
+
+\section{Exp2: Word Lists extracted from ISO9126 Taxonomy}
+
+We extracted the commit log comments from MaxDB 7.500 and per month
+extracted 20 topics using LDA. There are 25 months, thus a total of
+topic 500 possible topics, but there were a couple of months with no
+commits.
+
+Make a word list (a file) of related words and a label.  These words
+were extracted from the ISO9126 Taxonomy.  The main concepts used are:
+functionality, reliability, maintainability, usability, efficiency,
+and portability.
+
+We determined if a word list was similar to a topic if they shared 1
+or more words in common.
+
+\subsection{Observations}
+
+There are many more unnamed topics than before. Only reliability shows up very often.
+
+The words that reliability finds are:
+\begin{verbatim}
+error 279
+bug 25
+fails 2
+crash 14
+failure 1
+\end{verbatim}
+
+Usability words
+\begin{verbatim}
+ui 16
+user 8
+default 4
+gui 3
+feature 2
+\end{verbatim}
+
+Portability words
+\begin{verbatim}
+documentation 5
+migration 1
+\end{verbatim}
+
+Effeciency words
+\begin{verbatim}
+optimize 18
+optimization 4
+fast 3
+slow 1
+\end{verbatim}
+
+And ``dependency'' for maintability
+
+\subsection{Results}
+
+\begin{verbatim}
+Total Topics: 500
+Total Named Topics: 125
+Total UnNamed Topics: 375
+Total Empty Topics: 80
+Total UnNamed Non-empty Topics: 295
+
+maintainability 1
+efficiency 4
+portability 5
+usability 13
+reliability 107
+
+efficiency, reliability 1
+maintainability, maintainability 1
+maintainability, reliability 1
+portability, reliability 1
+reliability, usability 2
+efficiency, efficiency 4
+portability, portability 5
+usability, usability 13
+reliability, reliability 107
+\end{verbatim}
+
+\section{Exp3: Word Lists extracted from ISO9126 Taxonomy combined with Wordnet}
+
+We extended Experiment 2 to include the wordnet words.
+
+\subsection{Observations}
+
+Many more results that before, most topics are named now.
+
+For the efficiency concept:
+\begin{verbatim}
+add 109
+fix 102
+base 24
+optimize 18
+fixing 14
+correction 12
+command 7
+process 5
+frame 5
+string 4
+optimization 4
+fast 3
+end 3
+count 3
+upgrade 2
+speed 2
+solve 2
+record 2
+avoid 2
+slow 1
+shorten 1
+scratch 1
+resolve 1
+quick 1
+loading 1
+frequency 1
+estimate 1
+corrections 1
+control 1
+constant 1
+administration 1
+\end{verbatim}
+
+For functionality:
+\begin{verbatim}
+fix 102
+update 43
+window 22
+tack 9
+key 9
+support 8
+memory 8
+difference 8
+mark 7
+system 6
+ease 6
+lease 5
+grate 5
+frame 5
+ring 4
+strategy 3
+label 3
+source 2
+power 2
+gain 2
+buffer 2
+buff 2
+bar 2
+tool 1
+machine 1
+grant 1
+extract 1
+complex 1
+catch 1
+\end{verbatim}
+
+For maintainability
+\begin{verbatim}
+file 140
+log 19
+difference 8
+mark 7
+ease 6
+point 5
+document 5
+environment 4
+order 3
+improvement 3
+end 3
+action 3
+verify 2
+record 2
+power 2
+post 2
+place 2
+mode 2
+union 1
+tract 1
+tension 1
+sleep 1
+representation 1
+processing 1
+ownership 1
+need 1
+increase 1
+enlargement 1
+dependency 1
+control 1
+condition 1
+\end{verbatim}
+
+For portability:
+\begin{verbatim}
+os 17
+support 8
+form 8
+program 6
+function 6
+documentation 5
+transaction 3
+action 3
+upgrade 2
+routine 1
+migration 1
+integration 1
+\end{verbatim}
+
+For reliability
+\begin{verbatim}
+error 279
+check 196
+change 67
+move 60
+bug 25
+correct 24
+age 20
+set 18
+make 18
+ram 17
+out 17
+ace 15
+crash 14
+lock 10
+sing 9
+run 9
+list 9
+mat 8
+form 8
+case 8
+fail 7
+vent 6
+stall 6
+din 6
+point 5
+grate 5
+get 5
+format 5
+cover 5
+work 4
+ting 4
+ring 4
+fault 4
+default 4
+adjust 4
+turn 3
+rap 3
+pass 3
+order 3
+let 3
+gel 3
+end 3
+adapt 3
+action 3
+static 2
+start 2
+resume 2
+refresh 2
+power 2
+part 2
+modify 2
+ill 2
+flow 2
+find 2
+fails 2
+drop 2
+cut 2
+yield 1
+worker 1
+value 1
+union 1
+stance 1
+sink 1
+scratch 1
+representation 1
+report 1
+regard 1
+prim 1
+pop 1
+ping 1
+ownership 1
+nod 1
+instance 1
+header 1
+grow 1
+go 1
+failure 1
+exit 1
+enlargement 1
+eld 1
+dress 1
+develop 1
+dependency 1
+cycle 1
+corrections 1
+condition 1
+close 1
+catch 1
+calm 1
+administration 1
+\end{verbatim}
+
+For usability
+\begin{verbatim}
+error 279
+check 196
+test 125
+fix 102
+window 22
+monitor 19
+set 18
+ram 17
+ui 16
+view 15
+port 15
+back 15
+elect 12
+trace 11
+user 8
+use 8
+man 8
+compiler 8
+case 8
+miss 7
+name 6
+module 6
+function 6
+longer 5
+etch 5
+cover 5
+diagnose 4
+default 4
+gui 3
+end 3
+aries 3
+application 3
+wrapper 2
+wrap 2
+wall 2
+top 2
+stop 2
+possibility 2
+match 2
+liver 2
+libra 2
+feature 2
+cut 2
+cap 2
+bar 2
+worker 1
+ward 1
+sort 1
+show 1
+self 1
+resolve 1
+relative 1
+relation 1
+purpose 1
+protection 1
+present 1
+patch 1
+owner 1
+machine 1
+killer 1
+go 1
+gem 1
+extract 1
+equal 1
+block 1
+binary 1
+\end{verbatim}
+
+\subsection{Results}
+\begin{verbatim}
+Total Topics: 500
+Total Named Topics: 328
+Total UnNamed Topics: 172
+Total Empty Topics: 80
+Total UnNamed Non-empty Topics: 92
+
+maintainability 44
+portability 45
+functionality 94
+efficiency 200
+reliability 225
+usability 265
+
+maintainability, portability 5
+functionality, maintainability 15
+portability, reliability 29
+efficiency, maintainability 31
+maintainability, reliability 34
+maintainability, usability 35
+efficiency, portability 38
+functionality, portability 39
+portability, usability 40
+maintainability, maintainability 44
+portability, portability 45
+functionality, reliability 58
+efficiency, functionality 73
+functionality, usability 85
+functionality, functionality 94
+efficiency, reliability 137
+efficiency, usability 150
+reliability, usability 190
+efficiency, efficiency 200
+reliability, reliability 225
+usability, usability 265
+\end{verbatim}
 

### 2009-08-15 src/exp2/quickcount.sh

@@ -1,0 +1,1 @@
+cat $* | xargs -i echo echo -n \"{} \" \; fgrep -i \"{}\" lda.report \| wc -l | sh | sort -n -r -k 2

### 2009-08-15 src/exp2/quickreport.sh

@@ -1,0 +1,1 @@
+for file in wordlist.*; do echo $file; sh quickcount.sh $file; done | egrep -v ' 0$'

### 2009-08-15 src/exp3/run.sh

@@ -1,0 +1,14 @@
+#!/bin/bash
+DIR=`pwd`
+mkdir wordlists
+cd wordlists
+python ../names.py
+FILES=`echo wordlist.*`
+WLDIR=`pwd`
+cd ../../WordnetOcaml-0.1/
+for file in $FILES
+do
+    echo $file
+    cat $WLDIR/$file > $DIR/$file
+    cat $WLDIR/$file | xargs sh related.sh >> $DIR/$file
+done

### 2009-09-10 presentation-icsm/mkpres-out.sh
presentation compilation and output

@@ -1,0 +1,4 @@
+rm -rf pres-out
+mkdir pres-out
+awk '{print "convert -composite  white.png " $1 " pres-out/" (sprintf("%003d",FNR)) ".png"}' < order  | sh -x
+

### 2009-09-10 presentation-icsm/present.sh

@@ -1,0 +1,1 @@
+find pres-out -iname '*.png' -or -iname '*.jpg' | sort -n | fgrep -v notused | XARGS feh  -b white.png -F -Z  -a 0

### 2009-09-10 presentation-icsm/show.sh

@@ -1,0 +1,1 @@
+find -iname '*.png' -or -iname '*.jpg' | sort -n | fgrep -v notused | XARGS feh -Z 

### 2009-12-09 icpc/icpc2010.tex
first commit of icpc paper skeleton
Signed-off-by: Neil Ernst <neil@neilernst.net>

\documentclass[times, 10pt,twocolumn]{article} 
\usepackage{latex8}
\usepackage{times}

\begin{document}

\title{On the (automated) topic naming of software maintenance activities}

\author{Abram Hindle\\
University of Waterloo\\ ahindle@uwaterloo.ca\\
% For a paper whose authors are all at the same institution, 
% omit the following lines up until the closing ``}''.
% Additional authors and addresses can be added with ``\and'', 
% just like the second author.
\and
Neil Ernst\\
University of Toronto\\ 
nernst@cs.toronto.edu\\

\and
(illustrious supervisors)\\
}

\maketitle
\thispagestyle{empty}

\begin{abstract}
	The field of automated repository mining in software engineering has been a fruitful one. However, while many techniques use automated machine learners to classify their data, there has not been any investigation into the latent topics these classifiers represent. In particular, we would like to know whether such groups can be generalized across software products. We focus on software quality as a potential generalization, since there is some shared belief that these qualities apply broadly across software products. We wanted to name topics based on a static ontology of non-functional requirements and software qualities. Our overarching goal is to use individual topic models to create a list of terms that can be shared across software projects. In essence, this would be adding some form of context to general-purpose ontologies such as Wordnet. In this paper, we discuss how to use domain knowledge about software development in order to better annotate topics of development that are extracted from the change log messages in configuration management systems such as CVS.
\end{abstract}

\Section{Introduction}
In this paper we are going to show how one can `name' topic models generated with LDA. We also show that these topics can be generalized across products.

\section{Methodology}
- Extract topics from changes
- Annotate topics of multiple projects
-- mysql
-- maxdb
- Try simple word matching
- Try simple word match of wordnet words (related words)
- Try machine learning
- Make a simple wordnet?
- Try it

\section{Observations and evaluation}

\section{Related work}
I think somewhere we should make clear why supervised
(Cleland-Huang) methods are not useful (too time-consuming, domain
specific, etc).

\section{Conclusion}
``Can a domain specific wordnet for
   software development provide better accuracy for labelling and other
   lexical related tasks than wordnet and other machine learning
   techniques''.

\bibliographystyle{latex8}
\bibliography{icpc}

\end{document}

### 2010-01-11 src/validate/check_validity.py
minor change

@@ -3,7 +3,7 @@
 def load_period(period): 
     """ load the xml file with the data """
     period_file = period + '.xml'
-    data_dir = "/Users/nernst/Documents/papers/current-papers/abram/naming-paper/data/maxdb-tagged/"
+    data_dir = "../../data/maxdb-tagged/"
     
     import lxml.etree
     doc = lxml.etree.parse(data_dir + period_file)

### 2010-01-26 icpc/icpc2010.tex
reformat for ICPC / IEEE

@@
-\documentclass[times, 10pt,twocolumn]{article} 
-\usepackage{latex8}
-\usepackage{times}
+%!TEX TS-program = xelatex
+%!TEX encoding = UTF-8 Unicode
+\documentclass[10pt, conference, compsocconf]{IEEEtran}
+\usepackage{fontspec}
+\usepackage{xunicode}
+\usepackage{xltxtra}
+\usepackage{booktabs}
+\defaultfontfeatures{Mapping=tex-text}
+\setmainfont{Times New Roman}
+%\usepackage{amsthm}
+\usepackage{color}
+\newcommand{\ppe}[2]{$#1 \xrightarrow{++} #2$}
+\newcommand{\pe}[2]{$#1 \xrightarrow{+} #2$}
+\newcommand{\nne}[2]{$#1 \xrightarrow{--} #2$}
+\newcommand{\nege}[2]{$#1 \xrightarrow{-} #2$}
+\newcommand{\ande}[2]{$#1 \xrightarrow{AND} #2$}
+\newcommand{\ore}[2]{$#1 \xrightarrow{OR} #2$}
+\newcommand{\inode}[2]{$#1 \xrightarrow{I} #2$}
+\newcommand{\cnode}[2]{$#1 \xrightarrow{C} #2$}
+\newcommand{\ab}[1]{}
 
+\newtheorem{mydef}{definition}
+\usepackage{graphicx}
+\newcommand{\slbl}[1]{\textbf{\textsf{#1}}}
+\usepackage{multirow}%,multicolumn}
+\usepackage{slashbox}
+\usepackage{wrapfig}
+\usepackage{booktabs, colortbl}
+\usepackage{amsmath,amssymb}
+\usepackage{algpseudocode}
+%\usepackage[pdfstartview=FitH,xetex,bookmarksopenlevel=3,bookmarks=true]{hyperref} %pdfpagescrop={92 112 523 778},
+\usepackage{algorithm}%,algorithmic-fix}
 \begin{document}
 
-\title{On the (automated) topic naming of software maintenance activities}
+\author{
+\IEEEauthorblockN{Neil A. Ernst, John Mylopoulos}
+\IEEEauthorblockA{Department of Computer Science\\ University of Toronto\\ jm,nernst@cs.toronto.edu}
+\and
+\IEEEauthorblockN{Abram Hindle}
+\IEEEauthorblockA{University of Waterloo\\ ahindle@uwaterloo.ca\\}
+}
 
-\author{Abram Hindle\\
-University of Waterloo\\ ahindle@uwaterloo.ca\\
-% For a paper whose authors are all at the same institution, 
-% omit the following lines up until the closing ``}''.
-% Additional authors and addresses can be added with ``\and'', 
-% just like the second author.
-\and
-Neil Ernst\\
-University of Toronto\\ 
-nernst@cs.toronto.edu\\
-
-\and
-(illustrious supervisors)\\
-}
+\title{What's in a name? On the (automated) topic naming of software maintenance activities}
 
@@
 \begin{abstract}
-	The field of automated repository mining in software engineering has been a fruitful one.
-However, while many techniques use automated machine learners to classify their data, there has not been any investigation into the latent topics these classifiers represent.
-In particular, we would like to know whether such groups can be generalized across software products.
-We focus on software quality as a potential generalization, since there is some shared belief that these qualities apply broadly across software products.
-We wanted to name topics based on a static ontology of non-functional requirements and software qualities.
-Our overarching goal is to use individual topic models to create a list of terms that can be shared across software projects.
-In essence, this would be adding some form of context to general-purpose ontologies such as Wordnet.
-In this paper, we discuss how to use domain knowledge about software development in order to better annotate topics of development that are extracted from the change log messages in configuration management systems such as CVS.
+  The field of automated repository mining in software engineering has
+  been a fruitful one. However, while techniques such as topic modeling and concept location have been used, there has not been any
+  investigation into the latent topics these classifiers represent. In
+  particular, we would like to know whether such groups can be
+  generalized across software products. We focus on software quality
+  as a potential generalization, since there is some shared belief
+  that these qualities apply broadly across software products. We
+  wanted to name topics based on a static ontology of non-functional
+  requirements and software qualities. Our overarching goal is to use
+  individual topic models to create a list of terms that can be shared
+  across software projects. In essence, this would be adding some form
+  of context to general-purpose ontologies such as Wordnet. In this
+  paper, we discuss how to use domain knowledge about software
+  development in order to better annotate topics of development that
+  are extracted from the change log messages in configuration
+  management systems such as CVS.
 \end{abstract}
@@
 \Section{Introduction}
-In this paper we are going to show how one can `name' topic models generated with LDA. We also show that these topics can be generalized across products.
+In this paper we are going to show how one can `name' topic models generated with a topic modeling technique, Latent Dirichlet Allocation. We also show that these topics can be generalized
+across products.
+
+Traditionally topic modeling has been used to characterize the subject of single software projects~\cite{}. This paper introduces \emph{named topic models}, topic models
+
+Typically what is distinct is the software's particular functionality -- whether it offers email composition, prepares tax returns, and so on.
+What is common, however, is the non-functional, or quality requirements, that the software focuses on.
+
+\section{Background}
+
+\subsection{Definitions}
+A \emph{message} is a block of text written by a developer. In this
+paper, messages will be the CVS and BitKeeper commit log comments made
+when the user commits changes to files in a repository. A \emph{word
+  distribution} is the summary of a message by word count. Each word
+distribution will be over the words in all messages. However, most
+words will not appear in each message. A word distribution is effectively
+a word count divided by the message size. A \emph{topic} is a word
+distribution, i.e., a set of words that form a word distribution that is
+unique and independent within the set of documents in our total
+corpus. One could think of a topic as a distribution of the centroid
+of a group of messages. In this paper we often summarize topics by the
+top 10 most frequent words of their word distribution.  A \emph{trend}
+is one or more similar topics that recur over time.  Trends are
+particularly important, as they indicate long-lived and recurring
+topics that may provide key insights into the development of the
+project.
+
+\emph{ROC} values reflect a score, similar to school letter-grades, reflecting how well a particular learner performed for the given data.
+A ROC result of 0.5 would be equivalent to a random learner (randomly classifying data).
+ROC maps to the more familiar concepts of precision/sensitivity and recall/specificity: it plots the true positive rate (sensitivity) vs.
+the false positive rate (1 - specificity).
+A perfect learner has a value of 1.0, reflecting perfect recall and precision.
+
+\subsection{Topic models}
+Topic analysis uses tools such as Latent Dirichlet Allocation (LDA) and Latent Semantic Indexing (LSI) to extract
+independent word distributions (topics) from
+documents (commit log comments)~\cite{marcus04wcre,Poshyvanyk2007,lukins08wcre,Linstead2007}.
+Ideally these extracted topics 
+correlate with actual development topics that the developers discussed
+during the development of the software project. 
+Topic analysis often allows for a single document, such as a commit
+message, to be related to multiple topics. Documents represented as a mixture of topics maps well to commits
+to source code, which often have more than one purpose.  A topic
+represents both a word distribution and a group of commit log comments
+that are related to each other by their content.  In this case a topic
+is a set of tokens extracted from commit messages found within a
+projects source control system (SCS).
+
+One issue that arises with use of unsupervised techniques is how to label the topics.
+While the topic models themselves are generated automatically, what to make of them is less clear.
+For example, in both our previous work, and in \cite{Baldi2008}, topics are named manually: human experts read the highest-frequency members of a topic and assign a keyword accordingly.
+E.g., for the word list ``listener change remove add fire", Baldi et al.
+assign the keyword `event-handling'.
+The labels are reasonable enough, but still require some expert in the field to determine them.
+Our technique is automated, because we match keywords from Wordnet to words in the topic model.
+
+\subsection{Supervised learning}
 
 \section{Methodology}
+We leveraged the data we gathered in \cite{Hindle09ICSM} for this work. To gather the data, we first extract the commit log comments from a
+project's SCS repository. We filter out stop words and produce word distributions from these messages. These distributions are bucketed
+into windows, and then each window is subject to topic analysis and then we analyze and visualize the results. Figure \ref{fig:commits}
+depicts the general process for processing and analyzing the commit
+messages.
+
+Find two projects with open, accessible source code in similar fields.
+We selected MySQL and MaxDB, open-source, near-commercial quality database software.
+We used source code commit data (the message attached to the commit).
+A typical message might look like ``".
+From that dataset, we created an XML file which grouped commits into temporal periods, each approximately X weeks wide.
+Given a text file full of commits, indexed by date and time, we then ran Latent Dirichlet Allocation ~\cite{Blei2003} on the data.
+LDA works by creating topic models -- word frequency distributions -- for each period.
+A topic model is a statistical summary of the important subjects for that dataset.
+E.g., in Period 1, the most important topic might have been ``security", with ``sql'' second, down on to less relevant topics.
+The idea is similar to other IR approaches like TF/IDF (term frequency/inverse document frequency), which find uncommon words that aren't found in other wordlists.
+These approaches are unsupervised in that they rely solely on the source data with no human intervention.
+
+Once we have a list of topics for each period, we generated some word lists to assess to what extent our topics fit word lists.
+Intuitively, we would like to `name' our topic models for each period.
+In this way, we can see how certain words -- subjects -- occur more or less often with each period.
+We use several different word lists.
+One we tried was 9126, the ISO quality model.
+If words in the quality model were also found in a period, we automatically assigned that period to that subject (e.g., security, performance, reliability).
+
+To check our outcomes, we manually annotated each period with the same quality model.
+In other words, we looked at each period, and tried to assess what the LDA analysis output told us was the topic for that period.
+
+We then compared our automated analysis using signifier matching (greps) to the manual tagging to get an error rate.
+
+To compare our LDA approach, we used a suite of supervised classifiers from WEKA (e.g., support vector machines and Bayesian nets).
+These learners, described elsewhere, may also be used to find topics.
+We did a 19/20 analysis (using 1 period to train the learner on the tags, and the other 19 to attempt to learn) and got the following error rates.
+
+We present a ROC curve in Fig.
+\ref{}.
+The chart shows each learner measured in terms of precision and recall.
+In general, we prefer techniques which are closer to the upper-left corner (maximizing precision and maximizing recall).
+
 - Extract topics from changes
@@
 - Try it
+\subsection{Project characteristics}
+MaxDB started in the late 1970s as a research project, and was later acquired by SAP.
+As of version 7.5, released April 2007, the project has 940 thousand lines of C source code\footnote{generated using David A.
+Wheeler's `SLOCCount'.}.
+There are approximately ...
+developers working on it.
+MySQL 3.23, released in early 2001, contains 320 thousand lines of C and C++ source code.
+The MySQL project started in 1994.
 
@@
 \section{Related work}
-I think somewhere we should make clear why supervised
-(Cleland-Huang) methods are not useful (too time-consuming, domain
-specific, etc).
 
+One issue is presenting these results to non-experts.
+ConcernLines~\cite{treude09cl}, for instance, show topic occurrence using color intensity.
+We maintain that these results require more substantive manipulation before being reported to, for example, management.
+Like any large-scale data analysis project, it requires skill to derive the implications.
+Hooking the backend results directly to a reporting tool is unlikely to produce useful results.
+
+Part of our effort with this project is to understand the qualitative and intentional aspects of requirements in software evolution, a notion we first discussed in \cite{ernst07icsm}.
+That idea is derived from, in part, work on narratives of software systems shown in academic work like \cite{anton01}, or more general-purpose works like \cite{waldo93}.
+
+Cleland-Huang and her colleagues published work on mining requirements documents for non-functional requirements (quality requirements) in \cite{Cleland-Huang2006}.
+One approach they tried was similar to this one, with keywords mined from NFR catalogues found in~\cite{chung99}.
+They managed recall of 80\% with precision of 57\% for the Security NFR, but could not find a reliable source of keywords for other NFRs.
+Instead, they developed a supervised classifier by using human experts to identify an NFR training set.
+There are several reasons we did not follow this route.
+One, we believe we have a more comprehensive set of terms due to the taxonomy we chose.
+Secondly, we wanted to compare across projects.
+Their technique was not compared across different projects and the applicability of the training set to different corpora is unclear.
+A common taxonomy allows us to make inter-project comparison (subject to the assumption that all projects conceive of these terms in the same way).
+%Thirdly, while the objective of Cleland-Huang's study was to identify new NFRs (for system development) our study assumes these NFRs are latent in the textual documents of the project.
+Finally, the source text we use is less structured than their requirements documents.
+
+% There has been an explosion of interest in mining OSS repositories.
+Godfrey and Tu \cite{godfrey00} were among the first to use this data to assess software evolution.
+They discussed how well the growth of Linux was predicted by laws that Lehman and colleagues \cite{lehman85} proposed.
+More recently, Herraiz et al.
+\cite{herraiz07icsm} predicted OSS growth using time series analysis.
+Like us, they refer to release windows as a useful unit of analysis.
+
+Mens et al.
+\cite{mens08icsm} conducted an empirical study of Eclipse, the OSS code editor, to verify the claims of Lehman~\cite{lehman97sms}.
+They concerned themselves with source code only, and found Law Seven, ``Declining Quality'', to be too difficult to assess: ``[we lacked an] appropriate measurement of the evolving quality of the system as perceived by the users \cite[p.
+388]{mens08icsm}''.
+This paper examines the notions of quality in terms of a consistent ontology, as Mens et al.
+call for in their conclusions.
+
+	 Massey~\cite{massey02icse} and Scacchi (\cite{scacchi02,scacchi05b}) looked at the topic of requirements in open-source software.
+Their work discusses the source of the requirements and how they are used in the development process.
+German~\cite{german03gnome} looked at GNOME specifically, and listed several sources for requirements: leader interest, mimicry, brainstorming, and prototypes.
+None of this work  addressed quality requirements in OSS, nor did it examine requirements trends.
+
+% Hindle et al.
+\cite{Hindle2007} examined release patterns in OSS.
+They showed that there is a difference between projects regarding maintenance techniques.
+This supports our result that software qualities are not discussed with the same frequency across projects.
+
+% Most of this related research examined project source code.
+Software mining of other project corpora is less common.
+Similarities exist with approaches that begin with structured taxonomies, as with the Hismo software ontology \cite{girba06}.
+Rigby and Hassan \cite{rigby07msr} used a general purpose taxonomy to classify developer email according to temperament.
 \section{Conclusion}
@@
 
-\bibliographystyle{latex8}
+\bibliographystyle{IEEETran}
 \bibliography{icpc}

### 2010-01-26 src/WordnetOcaml-0.1/related.sh

@@ -1,2 +1,2 @@
 export WNSEARCHDIR=/usr/share/wordnet/
-./related-words $*
+./related-words.ml $*

### 2010-01-26 icpc/icpc2010.tex
working with related work and background

@@
 \author{
-\IEEEauthorblockN{Neil A. Ernst, John Mylopoulos}
+\IEEEauthorblockN{Abram Hindle}
+\IEEEauthorblockA{Department of Computer Science\\University of Waterloo\\ ahindle@uwaterloo.ca\\}
+\and
+\IEEEauthorblockN{Neil A. Ernst}
 \IEEEauthorblockA{Department of Computer Science\\ University of Toronto\\ jm,nernst@cs.toronto.edu}
-\and
-\IEEEauthorblockN{Abram Hindle}
-\IEEEauthorblockA{University of Waterloo\\ ahindle@uwaterloo.ca\\}
+
 }
@@
 \begin{abstract}
-  The field of automated repository mining in software engineering has
-  been a fruitful one. However, while techniques such as topic modeling and concept location have been used, there has not been any
-  investigation into the latent topics these classifiers represent. In
-  particular, we would like to know whether such groups can be
+  In automated mining of software repositories, most approaches, such as topic modeling and concept location, focus on applying standard machine learning algorithms to corpora.
+This approach does not make any use of domain-specific information to better understand results.
+While too much specificity can produce non-generalizable results, too little produces broad learners that don't provide much useful detail.%There has not been any   investigation into the latent topics these classifiers represent.
+In this paper, we investigate the latent topics classifier produce, and examine  whether such topics can be
   generalized across software products. We focus on software quality
@@
   wanted to name topics based on a static ontology of non-functional
-  requirements and software qualities. Our overarching goal is to use
-  individual topic models to create a list of terms that can be shared
-  across software projects. In essence, this would be adding some form
+  requirements and software qualities. %Our overarching goal is to use   individual topic models to create a list of terms that can be shared   across software projects. 
+In essence, this would be adding some form
   of context to general-purpose ontologies such as Wordnet. In this
@@
   are extracted from the change log messages in configuration
-  management systems such as CVS.
+  management systems such as CVS. Since our taxonomy is global, our results allow us to compare the relative importance of particular qualities between projects.
 \end{abstract}
 
-\Section{Introduction}
-In this paper we are going to show how one can `name' topic models generated with a topic modeling technique, Latent Dirichlet Allocation. We also show that these topics can be generalized
-across products.
+\section{Introduction}
+In this paper we `name' topic models generated with a topic modeling technique, Latent Dirichlet Allocation. We also show that these topics can be generalized across products.
 
-Traditionally topic modeling has been used to characterize the subject of single software projects~\cite{}. This paper introduces \emph{named topic models}, topic models
+Traditionally topic modeling has required manual annotation to derive domain-relevant information.
+This paper introduces \emph{named topic models}, topic models which are automatically named using a cross-project taxonomy.
+We start with an unsupervised topic learning algorithm, then manually annotate the topics using the ontology.
+From the annotated topics, we apply supervised classifiers to find labels for the remaining topics.
 
@@
 What is common, however, is the non-functional, or quality requirements, that the software focuses on.
+Our approach makes use of software quality taxonomies in order to compare different projects on similar concepts.
+For example, how is the Security quality treated over time by product A versus product B?
+Does product A care about it more as it matures?
+Does the quality occur more often in software commits?
+
+We first introduce some important terminology for our work.
+We then describe our methodology, including our datasets, then highlight our results.
+We conclude with a look at related work and possible improvements.
 
 \section{Background}
-
+We provide a brief overview of software repository mining and information retrieval.
 \subsection{Definitions}
@@
 \subsection{Topic models}
-Topic analysis uses tools such as Latent Dirichlet Allocation (LDA) and Latent Semantic Indexing (LSI) to extract
+Topic analysis uses tools such as Latent Dirichlet Allocation (LDA)~\cite{Blei2003} and Latent Semantic Indexing (LSI) to extract
 independent word distributions (topics) from
-documents (commit log comments)~\cite{marcus04wcre,Poshyvanyk2007,lukins08wcre,Linstead2007}.
+	documents (commit log comments). %~\cite{marcus04wcre,Poshyvanyk2007,lukins08wcre,Linstead2007}
 Ideally these extracted topics 
@@
 during the development of the software project.
-Topic analysis often allows for a single document, such as a commit
-message, to be related to multiple topics. Documents represented as a mixture of topics maps well to commits
-to source code, which often have more than one purpose.  A topic
+For example, our technique might identify a collection of words: `list next has iterator add', which we then identify (manually) as concerning the topics of collections and the Iterator pattern.
+In topic analysis a single document, such as a commit message, can be related to multiple topics.
+Representing documents as a mixture of topics maps well to commits to source code repositories, which often have more than one purpose.
+A topic
 represents both a word distribution and a group of commit log comments
-that are related to each other by their content.  In this case a topic
+that are related to each other by their content.  In this paper a topic
 is a set of tokens extracted from commit messages found within a
-projects source control system (SCS).
+project's source control system (SCS).
+
+In this paper we use LDA to identify topics.
+The input to the algorithm is a set of documents, with contents treated as a bag of words, and a parameter to set the number of topics to generate.
+The algorithm returns a word-document matrix relating the occurrence frequency of a
 
@@
 While the topic models themselves are generated automatically, what to make of them is less clear.
-For example, in both our previous work, and in \cite{Baldi2008}, topics are named manually: human experts read the highest-frequency members of a topic and assign a keyword accordingly.
+For example, in our previous work~\cite{Hindle09ICSM}, and in \cite{Baldi2008}, topics are named manually: human experts read the highest-frequency members of a topic and assign a keyword accordingly.
 E.g., for the word list ``listener change remove add fire", Baldi et al.
@@
 \subsection{Supervised learning}
+While unsupervised techniques (LSI and LDA are both unsupervised) are appealing in their lack of human intervention, supervised learners have the advantage of domain knowledge.
+This typically results in improved results.
+In supervised learning, the data is divided into slices.
+One slice is manually annotated by the domain expert
 
@@
 project's SCS repository. We filter out stop words and produce word distributions from these messages. These distributions are bucketed
-into windows, and then each window is subject to topic analysis and then we analyze and visualize the results. Figure \ref{fig:commits}
-depicts the general process for processing and analyzing the commit
-messages.
+into 30-day non-overlapping windows. Each window is subject to LDA and then we analyze and visualize the results. %Figure \ref{fig:commits}
+depicts the general process for processing and analyzing the commit messages.
 
@@
 
-- Extract topics from changes
-- Annotate topics of multiple projects
--- mysql
--- maxdb
-- Try simple word matching
-- Try simple word match of wordnet words (related words)
-- Try machine learning
-- Make a simple wordnet?
-- Try it
+% - Extract topics from changes
+% - Annotate topics of multiple projects
+% -- mysql
+% -- maxdb
+% - Try simple word matching
+% - Try simple word match of wordnet words (related words)
+% - Try machine learning
+% - Make a simple wordnet?
+% - Try it
+
 \subsection{Project characteristics}
@@
 \section{Related work}
-
-One issue is presenting these results to non-experts.
-ConcernLines~\cite{treude09cl}, for instance, show topic occurrence using color intensity.
-We maintain that these results require more substantive manipulation before being reported to, for example, management.
-Like any large-scale data analysis project, it requires skill to derive the implications.
-Hooking the backend results directly to a reporting tool is unlikely to produce useful results.
-
 Part of our effort with this project is to understand the qualitative and intentional aspects of requirements in software evolution, a notion we first discussed in \cite{ernst07icsm}.
 That idea is derived from, in part, work on narratives of software systems shown in academic work like \cite{anton01}, or more general-purpose works like \cite{waldo93}.
+
+The idea of extracting higher-level `concerns' (also known as `concepts', `aspects' or `requirements') has been attacked in two ways.
 
@@
 A common taxonomy allows us to make inter-project comparison (subject to the assumption that all projects conceive of these terms in the same way).
-%Thirdly, while the objective of Cleland-Huang's study was to identify new NFRs (for system development) our study assumes these NFRs are latent in the textual documents of the project.
+Thirdly, while the objective of Cleland-Huang's study was to identify new NFRs (for system development) our study assumes these NFRs are latent in the textual documents of the project.
 Finally, the source text we use is less structured than their requirements documents.
+
+The other approach is to start with code repositories, and try to extract concerns from there.
+In \cite{marcus04wcre}, the authors describe their use of Latent Semantic Indexing to identify commonly occurring concerns for software maintenance.
+Some results were interesting, but their precision was quite low.
+(such as
+ConcernLines~\cite{treude09cl} shows tag occurrence using color intensity.
+They mined change request tags (such as `milestone 3') and used these to make evolutionary analyses of a single product.
+The presence of a well-maintained set of tags is obviously essential to the success of this technique.
+%We maintain that these results require more substantive manipulation before being reported to, for example, management.
+Like any large-scale data analysis project, it requires skill to derive the implications.
+Hooking the backend results directly to a reporting tool is unlikely to produce useful results.
 
@@
 
-% Hindle et al.
+Hindle et al.
 \cite{Hindle2007} examined release patterns in OSS.
@@
 
-% Most of this related research examined project source code.
+Most of this related research examined project source code.
 Software mining of other project corpora is less common.
@@
 Rigby and Hassan \cite{rigby07msr} used a general purpose taxonomy to classify developer email according to temperament.
+
 \section{Conclusion}
-``Can a domain specific wordnet for
-   software development provide better accuracy for labelling and other
-   lexical related tasks than wordnet and other machine learning
-   techniques''.
+% ``Can a domain specific wordnet for
+%    software development provide better accuracy for labelling and other
+%    lexical related tasks than wordnet and other machine learning
+%    techniques''.
 
-\bibliographystyle{IEEETran}
+\bibliographystyle{IEEEtran}
 \bibliography{icpc}

### 2010-01-27 icpc/icpc2010.tex
updated methodology somewhat

@@
 In this paper we use LDA to identify topics.
+LDA works by creating topic models for each period.
+A topic model is a statistical summary of the important subjects for that dataset.
 The input to the algorithm is a set of documents, with contents treated as a bag of words, and a parameter to set the number of topics to generate.
 The algorithm returns a word-document matrix relating the occurrence frequency of a
+ The idea is similar to other IR approaches like TF/IDF (term frequency/inverse document frequency), which find uncommon words that aren't found in other wordlists.
+These approaches are unsupervised in that they rely solely on the source data with no human intervention.
 
@@
 \section{Methodology}
-We leveraged the data we gathered in \cite{Hindle09ICSM} for this work. To gather the data, we first extract the commit log comments from a
-project's SCS repository. We filter out stop words and produce word distributions from these messages. These distributions are bucketed
-into 30-day non-overlapping windows. Each window is subject to LDA and then we analyze and visualize the results. %Figure \ref{fig:commits}
-depicts the general process for processing and analyzing the commit messages.
+ % To gather the data, we first extract the commit log comments from a
+% project's SCS repository. We filter out stop words and produce word distributions from these messages. These distributions are bucketed
+% into 30-day non-overlapping windows. Each window is subject to LDA analysis. %Figure \ref{fig:commits}
+% depicts the general process for processing and analyzing the commit messages.
 
-Find two projects with open, accessible source code in similar fields.
+We wanted to compare two projects with open, accessible source code in similar functional areas.
+This allows us to partially control for differences in functional requirements.
 We selected MySQL and MaxDB, open-source, near-commercial quality database software.
-We used source code commit data (the message attached to the commit).
-A typical message might look like ``".
-From that dataset, we created an XML file which grouped commits into temporal periods, each approximately X weeks wide.
-Given a text file full of commits, indexed by date and time, we then ran Latent Dirichlet Allocation ~\cite{Blei2003} on the data.
-LDA works by creating topic models -- word frequency distributions -- for each period.
-A topic model is a statistical summary of the important subjects for that dataset.
+MaxDB started in the late 1970s as a research project, and was later acquired by SAP.
+As of version 7.5, released April 2007, the project has 940 thousand lines of C source code\footnote{generated using David A.
+Wheeler's `SLOCCount'.}.
+There are approximately ...
+developers working on it.
+MySQL 3.23, released in early 2001, contains 320 thousand lines of C and C++ source code.
+The MySQL project started in 1994.
+
+From each project, we used source code commit data (the message attached to the commit).
+We leveraged the data we gathered in \cite{Hindle09ICSM} for this work.
+A typical commit message is: ``history annotate diffs bug fixed (if mysql_real_connect() failed there were two pointers to malloc'ed strings, with memory corruption on free(), of course) ".
+We extracted these messages and indexed them by creation time.
+
+From that dataset, we created an XML file which grouped commits into temporal periods, each approximately 4 weeks wide.
+Given a text file full of commits, indexed by date and time, we then ran an LDA modeler on each period, to generate a list of topics per period.
 E.g., in Period 1, the most important topic might have been ``security", with ``sql'' second, down on to less relevant topics.
-The idea is similar to other IR approaches like TF/IDF (term frequency/inverse document frequency), which find uncommon words that aren't found in other wordlists.
-These approaches are unsupervised in that they rely solely on the source data with no human intervention.
+We arbitrarily set the number of topics to generate to 20, based on past experiences.
 
-Once we have a list of topics for each period, we generated some word lists to assess to what extent our topics fit word lists.
+	Once we have a list of topics for each period, we compared them with lists of keywords to assess to what extent our topics fit word lists.
 Intuitively, we would like to `name' our topic models for each period.
 In this way, we can see how certain words -- subjects -- occur more or less often with each period.
-We use several different word lists.
-One we tried was 9126, the ISO quality model.
-If words in the quality model were also found in a period, we automatically assigned that period to that subject (e.g., security, performance, reliability).
+We used several different word lists.
+One we tried was 9126, the ISO quality model~\cite{iso9126}.
+It describes six high-level quality requirements (listed in Table \ref{}).
+There is some debate about the significance and importance of the terms in this model.
+However, it is ``an international standard and thus provides an internationally accepted terminology for software quality \cite[p.
+58]{Bøegh2008},'' which is sufficient for the purposes of this research.
+%exp3
+
+	We then extended our word lists with related words using Wordnet.
+
+	 If words in the quality model were also found in a period, we automatically assigned that period to that concept (e.g., security, performance, reliability).
 
@@
 % - Try it
-
-\subsection{Project characteristics}
-MaxDB started in the late 1970s as a research project, and was later acquired by SAP.
-As of version 7.5, released April 2007, the project has 940 thousand lines of C source code\footnote{generated using David A.
-Wheeler's `SLOCCount'.}.
-There are approximately ...
-developers working on it.
-MySQL 3.23, released in early 2001, contains 320 thousand lines of C and C++ source code.
-The MySQL project started in 1994.
 

### 2010-01-27 icpc/icpc2010.tex
added some sections, worked on method.

@@
 \usepackage{algpseudocode}
-%\usepackage[pdfstartview=FitH,xetex,bookmarksopenlevel=3,bookmarks=true]{hyperref} %pdfpagescrop={92 112 523 778},
+\usepackage[pdfstartview=FitH,xetex,bookmarksopenlevel=3,bookmarks=true]{hyperref} %pdfpagescrop={92 112 523 778},
 \usepackage{algorithm}%,algorithmic-fix}
@@
 \IEEEauthorblockA{Department of Computer Science\\ University of Toronto\\ jm,nernst@cs.toronto.edu}
-
 }
@@
 \section{Introduction}
-In this paper we `name' topic models generated with a topic modeling technique, Latent Dirichlet Allocation. We also show that these topics can be generalized across products.
+\begin{quote}
+	\emph{What's in a name? that which we call a rose\\
+	By any other name would smell as sweet;}
+	\begin{flushright} -- Romeo and Juliet, II:ii\end{flushright}
+\end{quote}
+
+Unlike the poetic licence Shakespeare uses, we show that names are indeed important.
+In this paper we `name' topic models generated with a topic modeling technique, Latent Dirichlet Allocation~\cite{Blei2003}.
+We use these topic models to train supervised learners to classify larger datasets.
+We also show that these topics can be generalized across products.
 
 Traditionally topic modeling has required manual annotation to derive domain-relevant information.
-This paper introduces \emph{named topic models}, topic models which are automatically named using a cross-project taxonomy.
+This paper introduces \emph{named topic models}, topic models which are named using a cross-project taxonomy.
 We start with an unsupervised topic learning algorithm, then manually annotate the topics using the ontology.
-From the annotated topics, we apply supervised classifiers to find labels for the remaining topics.
+From the annotated topics, we apply automated classifiers to find labels for the remaining topics.
 
@@
 Does the quality occur more often in software commits?
+
+Our contributions include:
+\begin{itemize}
+\item  We introduce named topic analysis, and demonstrate its usefulness
+  compared to other approaches.
+\item We show that these topics can be learned and used to classify other datasets.
+\item We present a number of visualizations of named topics and their trends over time to aid
+  communication and analysis.
+\item We use an exploratory case study of
+  several open source database systems to show how named topics can be compared between projects.
+\end{itemize}
 
@@
 
+\subsection{Generating the data}
 From each project, we used source code commit data (the message attached to the commit).
 We leveraged the data we gathered in \cite{Hindle09ICSM} for this work.
-A typical commit message is: ``history annotate diffs bug fixed (if mysql_real_connect() failed there were two pointers to malloc'ed strings, with memory corruption on free(), of course) ".
+A typical commit message is: ``history annotate diffs bug fixed (if mysql\_real\_connect() failed there were two pointers to malloc'ed strings, with memory corruption on free(), of course) ".
 We extracted these messages and indexed them by creation time.
+We apply stop-word removal to eliminate common English words like `the', `at', and so on.
 
-From that dataset, we created an XML file which grouped commits into temporal periods, each approximately 4 weeks wide.
+From that dataset, we created an XML file which grouped commits into temporal periods, each 30 days wide.
+This size of period is smaller than the time between minor releases but large enough for there to be sufficient commits to analyze.
 Given a text file full of commits, indexed by date and time, we then ran an LDA modeler on each period, to generate a list of topics per period.
 E.g., in Period 1, the most important topic might have been ``security", with ``sql'' second, down on to less relevant topics.
-We arbitrarily set the number of topics to generate to 20, based on past experiences.
+We arbitrarily set the number of topics to generate to 20, because past experimentation showed that fewer topics might aggregate multiple unique topics while any more topics seemed to dilute the results and create indistinct topics.
 
-	Once we have a list of topics for each period, we compared them with lists of keywords to assess to what extent our topics fit word lists.
-Intuitively, we would like to `name' our topic models for each period.
-In this way, we can see how certain words -- subjects -- occur more or less often with each period.
-We used several different word lists.
-One we tried was 9126, the ISO quality model~\cite{iso9126}.
-It describes six high-level quality requirements (listed in Table \ref{}).
+\subsection{Associating names}
+% We would like to `name' our topic models for each period. In this way, we can see how certain words -- subjects -- occur more or less often with each period.
+Once we had topics for each period, we compared them with lists of keywords to assess to what extent our topics mapped into word lists.
+Recall that topics are lists of words and associated frequencies.
+We performed simple string matching between the topics and our lists, `naming' a topic if it contained that word or words.
+We used several different word lists for comparison.
+
+Our first list, \textsf{exp1}, was generated using the ontology described in Kayed et al.~\cite{5072519}.
+That paper constructs an ontology for software quality measurement using eighty source documents, including research papers and international standards.
+The word list we used:
+\begin{quotation}
+\small \noindent 
+integrity, security,
+interoperability, testability, maintainability, traceability,
+accuracy, modifiability, understandability, availability, modularity,
+usability, correctness, performance, verifiability, efficiency,
+portability, flexibility, reliability.
+\end{quotation}
+
+Our second list, we relied on the ISO quality model~\cite{iso9126}.
+It describes six high-level quality requirements (listed in Table \ref{tbl:wnsig}).
 There is some debate about the significance and importance of the terms in this model.
@@
 58]{Bøegh2008},'' which is sufficient for the purposes of this research.
-%exp3
+%exp2
 
-	We then extended our word lists with related words using Wordnet.
+\begin{table*}
+	\caption{Qualities and associated signifiers – Wordnet version (\textsf{exp2})}
+	\centering
+	\label{tbl:wnsig}
+\begin{tabular}{c|p{9cm}}
+\toprule
+\textbf{Quality} & \textbf{Signifiers} \\
+\midrule 
+\emph{Maintainability} &
+testability changeability analyzability stability maintain maintainable modularity modifiability understandability\\ \hline
+\emph{Functionality} &
+security compliance accuracy interoperability suitability functional practicality functionality\\ \hline
+\emph{Portability} &
+conformance adaptability replaceability installability portable movableness movability portability\\ \hline
+\emph{Efficiency} &
+“resource behaviour” “time behaviour” efficient efficiency\\ \hline
+\emph{Usability} &
+operability understandability learnability useable usable serviceable usefulness utility useableness usableness serviceableness serviceability usability\\ \hline
+\emph{Reliability} &
+“fault tolerance” recoverability maturity reliable dependable responsibleness responsibility reliableness reliability dependableness dependability\\ 
+\bottomrule
+\end{tabular}
+\end{table*}
 
-	 If words in the quality model were also found in a period, we automatically assigned that period to that concept (e.g., security, performance, reliability).
+We then extended our word lists with related words using Wordnet. %exp3
 
+\subsection{Supervised learning}
 To check our outcomes, we manually annotated each period with the same quality model.
@@
 
-We present a ROC curve in Fig.
+\section{Observations and evaluation}
+
+\subsection{Word list similarity}
+We named topics by matching words from a list to the words in that topic model. We determined if a word list was similar to a topic if they shared one or more words in common.
+
+\subsection{Classifying topics}
+We present a ROC chart in Fig.
 \ref{}.
@@
 
-% - Extract topics from changes
-% - Annotate topics of multiple projects
-% -- mysql
-% -- maxdb
-% - Try simple word matching
-% - Try simple word match of wordnet words (related words)
-% - Try machine learning
-% - Make a simple wordnet?
-% - Try it
+\subsection{Comparing topic use between projects}
 
-\section{Observations and evaluation}
+\subsection{Topic change over time}
+
+\subsection{Threats to validity}
+\textbf{Construct validity} – Focus on code commit messages, rather than mail or bug trackers. Choice of taxonomy
+
+\textbf{Internal validity} -- inter-rater reliability.
+
+\textbf{External validity} – Our data originated from open-source projects in the database domain.
+Of these, the open-source nature of the project seems most problematic for external validity.
+However, both projects are developed by commercial entities.
+Our technique is dependent on development practices including meaningful commit messages, which might not be true in other domains.
 
@@
 Some results were interesting, but their precision was quite low.
-(such as
 ConcernLines~\cite{treude09cl} shows tag occurrence using color intensity.
@@
 
+In \cite{Mei2007}, the authors use context information to automatically name topics.
+They describe probabilistic labeling, using the frequency distribution of words in a topic to create a meaningful phrase.
+They do not use external domain-specific information as we do.
+
 Massey~\cite{massey02icse} and Scacchi (\cite{scacchi02,scacchi05b}) looked at the topic of requirements in open-source software.
@@
 
-\section{Conclusion}
-% ``Can a domain specific wordnet for
-%    software development provide better accuracy for labelling and other
-%    lexical related tasks than wordnet and other machine learning
-%    techniques''.
+\section{Conclusions and future work}
+In this paper we showed that a domain specific wordnet for software development provides better accuracy for labelling topic models than unsupervised approaches.
+
+This paper has not considered the interpretation of attitudes towards qualities (e.g., do developers feel this helps security).
 

### 2010-01-27 icsm09/lda-lab.tex
moving older files into icsm09 directory

\documentclass{report}
\usepackage{graphicx,amsmath,comment,fullpage}
\newcommand{\igH}[1]{\includegraphics[height=.9\textheight]{#1}}
\newcommand{\igW}[1]{\includegraphics[width=.9\textwidth]{#1}}
\newcommand{\igWh}[1]{\includegraphics[width=.7\textwidth]{#1}}
\newcommand{\igWhalf}[1]{\includegraphics[width=.45\textwidth]{#1}}
\newcommand{\lda}{Latent Dirichlet Allocation}

\author{Abram Hindle\\
ahindle@cs.uwaterloo.ca\\
Software Architecture Group (SWAG)\\
Supervised by\\
Micheal Godfrey and Ric Holt
}
\title{Lab 0006: Latent Dirichlet Allocation Applied to Revisions}

\usepackage{rotating}   
\begin{document}

\maketitle
\begin{abstract}

  In this lab report we seek to explore how topics shift over time in
  the source control repository of a project. We analyze commit
  comments, and even source code per revision per each time window and
  extract the topics prevelant in that window. We expect that topic
  sets will change over time and the change in topic sets indicate a
  change in the focus of development. We suppose this could help us to
  segment the revisions on the time line to discover parts of an
  iteration. For topic extraction we use \lda (LDA).

\end{abstract}

\section{Introduction}

%?

In this lab report we seek to explore how topics shift over time in
the source control repository of a project. We analyze commit
comments, and even source code per revision per each time window and
extract the topics prevelant in that window. We expect that topic sets
will change over time and the change in topic sets indicate a change
in the focus of development. We suppose this could help us to segment
the revisions on the time line to discover parts of an iteration. For
topic extraction we use \lda (LDA).

\section{Methodology}

\subsection{Datasets and Tools}

Extractors:
\begin{itemize}
\item CVSSuck - CVS Suck mirrors RCS files from a CVS repository. 
\item softChange - extract CVS facts to a PostgreSQL database.
\item bt2csv - Convert BitKeeper repositories to facts in CSV
  databases.
\end{itemize}

Analysis Tools:
\begin{itemize}
\item Hiraldo-Grok - An OCaml based spin off of Grok used for
  answering queries.
\item Gnuplot - graph plotting package
\end{itemize}

\section{First Impression}

% 2007-10-12

Looking at 30 day non-overlapping windows of revisions for MySQL 3.23
(20 topics, top 10 words per topic) we found there were common words
across topic clusters such as diffs, annotate and history. There were
notable transitional topics such as in the first window the word
bitkeeper appears (probably when they adopted bitkeeper) yet in the
following windows there were no mentions of bitkeeper. ``RENAME'' also
only appeared once in the first window.

Often words are shared across topics, see talbe \ref{tab:portability}
for tracking topics about portability.

It is apparent that we have to compare these topics some how by
itemset or similarity.

\begin{table}
\centering
\begin{tabular}{|ccc|l|}
\hline
2000 &  Jul &  31 &    chmod \\
2000 &  Sep &  29 &    fixes benchmark logging windows \\
2000 &  Nov &  28 &    typo fix insert\_multi\_value \\
2001 &  Jan &  27 &    FIxes Innobase Cleanups auto-union \\
2001 &  Mar &  28 &    2 topics bugfix, logging , TEMPORARY,  \\
\hline
2001 &  Jul &  26 &    Update Allow TABLES LOCK [a] \\

2001 &  Aug &  25 &    Tables row version [a] \\
\hline
2001 &  Sep &  24 &    Update checksum merge \\
2001 &  Oct &  24 &    fixed fix \\
2001 &  Dec &  23 &    HPUX SCO fix \\
\hline
2002 &  Feb &  21 &    net buffer length  max\_allowed\_packet [b] \\
2002 &  Mar &  23 &    small buf fix [b]  \\
\hline
2002 &  May &  22 &    [popular] fix SCO OSF1 table\_name \\
2002 &  Nov &  18 &    HPUX11 compiler HP \\
\hline
2003 &  Feb &  16 &    Linux errno  [c] \\
2003 &  Mar &  18 &    alarm bookmark bug [c] \\
\hline
2003 &  Sep &  14 &    Auto logging merge windows distribution fix 64-bit 4.0 Cleanup \\
\hline
\end{tabular}
\caption{Tracking topics associated with the word portability, note some continuous blocks}
\label{tab:portability}
\end{table}

\end{document}

### 2010-01-27 icpc/icpc2010.tex
misc changes to text
updated with some results and exp2 defn expanded

@@
 MaxDB started in the late 1970s as a research project, and was later acquired by SAP.
-As of version 7.5, released April 2007, the project has 940 thousand lines of C source code\footnote{generated using David A.
+As of version 7.500, released April 2007, the project has 940 thousand lines of C source code\footnote{generated using David A.
 Wheeler's `SLOCCount'.}.
@@
 
-\subsection{Associating names}
+\subsection{Associating labels}
 % We would like to `name' our topic models for each period. In this way, we can see how certain words -- subjects -- occur more or less often with each period.
-Once we had topics for each period, we compared them with lists of keywords to assess to what extent our topics mapped into word lists.
+Once we had topics for each period, we tried to associate them with a label from a list of keywords and related terms.
 Recall that topics are lists of words and associated frequencies.
-We performed simple string matching between the topics and our lists, `naming' a topic if it contained that word or words.
+We performed simple string matching between these topics and our lists, `naming' a topic if it contained that word or words.
 We used several different word lists for comparison.
@@
 That paper constructs an ontology for software quality measurement using eighty source documents, including research papers and international standards.
-The word list we used:
+The labels we used:
 \begin{quotation}
-\small \noindent 
+\small \noindent \textsf{
 integrity, security,
@@
 usability, correctness, performance, verifiability, efficiency,
-portability, flexibility, reliability.
+portability, flexibility, reliability.}
 \end{quotation}
 
-Our second list, we relied on the ISO quality model~\cite{iso9126}.
+Our second list relied on the ISO quality model~\cite{iso9126}.
 It describes six high-level quality requirements (listed in Table \ref{tbl:wnsig}).
@@
 58]{Bøegh2008},'' which is sufficient for the purposes of this research.
+However, these terms may not capture all words associated with the labels.
+For example, the term ‘user friendly’ is one most would agree is relevant to discussion of usability, but is not in the standard.
+Therefore we took the words from the taxonomy and expanded them.
+
+To construct our expanded sets, we also used Wordnet~\cite{Fellbaum1998}, an English-language ‘lexical database’ that contains semantic relations between words, including meronymy and synonymy.
+We then added Boehm’s 1976 software quality model \cite{Boehm+:1976:ICSE}, and classified his eleven ‘ilities’ into their respective ISO9126 qualities.
+We did the same for the quality model produced by McCall et al.
+\cite{mccall1977}.
+Finally, we analyzed two mailing lists from the KDE project to enhance the specificity of the sets.
+We selected KDE-Usability, which focuses on usability discussions for KDE as a whole; and KDE-Konqueror, a list about a long-lived web browser project.
+For each high-level quality in ISO9126, we first searched for our existing labels; we then randomly sampled twenty-five mail messages that were relevant to that quality, and selected co-occurring terms relevant to that quality.
+For example, we add the term ‘performance’ to the synonyms for \emph{Efficiency}, since this term occurs in most KDE mail messages that discuss efficiency.
 %exp2
 
-\begin{table*}
+For the third -- \textsf{exp3} -- list of quality labels, we extended the list from \textsf{exp2} using unfiltered Wordnet similarity matches.
+
+\begin{table*}[h]
 	\caption{Qualities and associated signifiers – Wordnet version (\textsf{exp2})}
@@
 \toprule
-\textbf{Quality} & \textbf{Signifiers} \\
+\textbf{Label} & \textbf{Related terms} \\
 \midrule
 \emph{Maintainability} &
-testability changeability analyzability stability maintain maintainable modularity modifiability understandability\\ \hline
+testability changeability analyzability stability maintain maintainable modularity modifiability understandability + interdependent dependency encapsulation decentralized modular\\ \hline
 \emph{Functionality} &
-security compliance accuracy interoperability suitability functional practicality functionality\\ \hline
+security compliance accuracy interoperability suitability functional practicality functionality + compliant exploit certificate secured “buffer overflow” policy malicious trustworthy vulnerable vulnerability accurate secure vulnerability correctness accuracy\\ \hline
 \emph{Portability} &
-conformance adaptability replaceability installability portable movableness movability portability\\ \hline
+conformance adaptability replaceability installability portable movableness movability portability + specification migration standardized l10n localization i18n internationalization documentation interoperability transferability\\ \hline
 \emph{Efficiency} &
-“resource behaviour” “time behaviour” efficient efficiency\\ \hline
+“resource behaviour” “time behaviour” efficient efficiency + performance profiled optimize sluggish factor penalty slower faster slow fast optimization\\ \hline
 \emph{Usability} &
-operability understandability learnability useable usable serviceable usefulness utility useableness usableness serviceableness serviceability usability\\ \hline
+operability understandability learnability useable usable serviceable usefulness utility useableness usableness serviceableness serviceability usability + gui accessibility menu configure convention standard feature focus ui mouse icons ugly dialog guidelines click default human convention friendly user screen interface flexibility\\ \hline
 \emph{Reliability} &
-“fault tolerance” recoverability maturity reliable dependable responsibleness responsibility reliableness reliability dependableness dependability\\ 
+“fault tolerance” recoverability maturity reliable dependable responsibleness responsibility reliableness reliability dependableness dependability + resilience integrity stability stable crash bug fails redundancy error failure\\
 \bottomrule
@@
 \end{table*}
-
-We then extended our word lists with related words using Wordnet. %exp3
 
@@
 \subsection{Word list similarity}
-We named topics by matching words from a list to the words in that topic model. We determined if a word list was similar to a topic if they shared one or more words in common.
+In general, this approach did not work out well.
+The related words for \emph{correctness}, for example, tended to be too lengthy and non-specific.
+Table \ref{tbl:wordlist} lists results.
+A \emph{named topic} is a topic with a matching label.
+There are \{periods X 20\} topics per project.
+All experiments were run on MaxDB data.
+
+\begin{table*}[h]
+	\caption{Results for automatic topic labeling}
+	\centering
+	\label{tbl:wordlist}
+\begin{tabular}{c|c|c|c}
+\toprule
+Measure &		 \textsf{exp1} &	 \textsf{exp2} &	 \textsf{exp3} \\
+\midrule
+Topics &			500		 &			500  	 & 		500  \\
+Named topics &		281      &			125      &		328  \\
+Unnamed topics &	139      &			295      &      92   \\
+\bottomrule
+\end{tabular}
+\end{table*}
+
+For \textsf{exp1}, our best performing labels (the labels matched with the most topics) were correctness (182 topics) and testability (121).
+We did not get good results for usability or accuracy, which were associated with fewer than ten topics.
 

### 2010-01-29 icpc/icpc2010.tex
methodology section finished

@@
 \usepackage{color}
-\newcommand{\ppe}[2]{$#1 \xrightarrow{++} #2$}
-\newcommand{\pe}[2]{$#1 \xrightarrow{+} #2$}
-\newcommand{\nne}[2]{$#1 \xrightarrow{--} #2$}
-\newcommand{\nege}[2]{$#1 \xrightarrow{-} #2$}
-\newcommand{\ande}[2]{$#1 \xrightarrow{AND} #2$}
-\newcommand{\ore}[2]{$#1 \xrightarrow{OR} #2$}
-\newcommand{\inode}[2]{$#1 \xrightarrow{I} #2$}
-\newcommand{\cnode}[2]{$#1 \xrightarrow{C} #2$}
-\newcommand{\ab}[1]{}
 
@@
 \usepackage{algorithm}%,algorithmic-fix}
+\usepackage{url}
 \begin{document}
@@
 \IEEEauthorblockN{Neil A. Ernst}
-\IEEEauthorblockA{Department of Computer Science\\ University of Toronto\\ jm,nernst@cs.toronto.edu}
+\IEEEauthorblockA{Department of Computer Science\\ University of Toronto\\ nernst@cs.toronto.edu}
 }
@@
 usability, correctness, performance, verifiability, efficiency,
-portability, flexibility, reliability.}
+portability, flexibility, reliability.
+}
 \end{quotation}
@@
 For the third -- \textsf{exp3} -- list of quality labels, we extended the list from \textsf{exp2} using unfiltered Wordnet similarity matches.
+Similarity in Wordnet means siblings in a hypernym tree.
+We do not include these words here for space considerations (but see the Appendix for our data repository).
+It isn't clear the words associated with our labels are specific enough, however: for example, the label \emph{maintainability} is associated with words \emph{ease} and \emph{ownership}.
 
@@
 \subsection{Supervised learning}
-To check our outcomes, we manually annotated each period with the same quality model.
-In other words, we looked at each period, and tried to assess what the LDA analysis output told us was the topic for that period.
+To check our outcomes, we manually annotated each topic in each period with the same quality labels as \textsf{exp2}.
+We looked at each period's topics, and assessed what the data -- consisting of the frequency-weighted word lists and messages -- suggested was the topic for that period.
+For example, for the MaxDB topic consisting of a message ``exit() only used in non NPTL LINUX Versions'', we tagged that topic \emph{portability}.
+This provided us with manually classified topics for the second phase of our analysis, supervised learning (classification).
+%inter-rater reliability
 
-We then compared our automated analysis using signifier matching (greps) to the manual tagging to get an error rate.
+We first compared our previous analysis using label matching to our manual classifications to get an error rate for that process (described below).
 
-To compare our LDA approach, we used a suite of supervised classifiers from WEKA (e.g., support vector machines and Bayesian nets).
-These learners, described elsewhere, may also be used to find topics.
-We did a 19/20 analysis (using 1 period to train the learner on the tags, and the other 19 to attempt to learn) and got the following error rates.
+For supervised learning, we used a suite of supervised classifiers from WEKA~\cite{weka09}.
+WEKA contains a suite of machine learning tools (e.g., support vector machines and Bayesian nets).
+We also used the multi-labeling add-on for WEKA, Mulan\footnote{\url{http://mlkd.csd.auth.gr/multilabel.html}}.
+Traditional classifiers map our topics to a single class, whereas Mulan allows for a mixture of classes per topic (much like our manual classification found overlaps).
+
+To assess the performance of the supervised learners, we did a 20-fold cross-over analysis (using 1 period to train the learner on the tags, and the other 19 to attempt to learn, changing the training set each time).
+We report these results below.
+
+Finally, using this data, we evaluated two research questions: 1) Do label frequencies change over time?
+That is, is a certain quality of more interest at one point in the lifecycle than some other?
+2) Do the different projects differ in their relative topic interest?
+That is, is a particular quality more important to one project than the other projects?
 
@@
 We did not get good results for usability or accuracy, which were associated with fewer than ten topics.
+We also looked for correlation between our labels: Excluding double matches (self-correlation), our highest co-occurring terms were verifiability and traceability, and testability and correctness (76 and 62 matches, respectively).
+
+For \textsf{exp2}, there are many more unnamed topics. Only reliability produces a lot of matches, mostly with the `error' word. Co-occurrence results were poor.
+
+For \textsf{exp3}, we had many more named topics.
+As we mentioned, the wordlists are quite broad, so there are likely to be false-positives.
+See the following sections for our error analysis.
+We found a high of 265 topics for usability, with a low of 44 topics for maintainability.
+Common co-occurrences were reliability and usability, efficiency and reliability, and efficiency and usability (200, 190, and 150 topics in common, respectively).
+
+\subsubsection{Analysing the unsupervised labeling}
+Based on the labels, and our manual topic labeling, we compared the results of the unsupervised word matching approach.
+For each quality we tried to assess whether the manual tag matched the unsupervised label assigned.
+Table \ref{tbl:maxdb-unsup-results} shows our results for MaxDB.
 
@@
 
+It is difficult to map abstract qualities to particular messages.
+If the message is about a particular bug fix, do we map that to reliability?
+Maintainability?
+Is it a performance bug or a usability bug?
+Is a `quality' discussion about more than just corrective maintenance, in other words?
+In this study we take the position that any type of maintenance has to do with software quality.
+
+\section{Acknowledgements and appendix}
+Our data and scripts are available at \url{http://softwareprocess.es/icpc2010}
+
 \bibliographystyle{IEEEtran}

### 2010-01-29 src/validate/check_validity.py
Revised the validity checks for the automated labeling so that a nicer CSV file is printed with the stats.

@@ -10,10 +10,7 @@ def make_count():
 
 # As far as I can tell this script is meant to see if
 # the annotations match or don't match portability efficiency reliability etc..
-
-
-
-def load_period(period, data_dir = "/Users/nernst/Documents/papers/current-papers/abram/naming-paper/data/maxdb-tagged/"): 
+def load_period(period, data_dir = "../../data/maxdb-tagged/"): 
     """ load the xml file with the data """
     period_file = period + '.xml'
     
@@ -102,14 +99,14 @@ def index_matching_wordlists( thresh_el, annotations, quality_map ):
     #     return "no match"
 
 def info_report_string(key, tp, fp, fn, tn):
-    s1 = key + "\t"+ "\t".join(["tp","fp","fn","tn"])+"\n"
-    s2 = key + "\t"+ "\t".join([str(x) for x in [tp,fp,fn,tn]])+"\n"
+    #s1 = key + ","+ ",".join(["tp","fp","fn","tn"])+"\n"
+    #s2 = key + ","+ ",".join([str(x) for x in [tp,fp,fn,tn]])+"\n"
     im = info_theory.info_measures(tp,tn,fp,fn)
     ik = im.keys()
     ik.sort()
-    strhead = key + "\t" + "\t".join(ik)
-    strb = key + "\t" + "\t".join([str(y) for y in [im[x] for x in ik]])
-    return "\n".join([s1,s2,strhead,strb]) + "\n"
+    #strhead = key + "," + ",".join(ik)
+    strb = key + "," + ",".join([str(y) for y in [im[x] for x in ik]])
+    return "\n" + strb #+ "\n"#join([s1,s2,strhead,strb]) + "\n"
 
 
 
@@ -129,6 +126,7 @@ if __name__ == '__main__':
         quality_map = load_wordlists( wld )
         
         for exp in exps:
+            #print exp
             ddir = exp[1]
             periods = exp[2]
             report = exp[0]
@@ -143,8 +141,8 @@ if __name__ == '__main__':
             
             for period in periods:
                 pstr = "Period #" + period
-                print pstr
-                rfile.write(pstr+'\n')
+                #print pstr
+                #rfile.write(pstr+'\n')
                 topics = load_period(period, data_dir = ddir)
                 for t in topics:
                     annotations, thresh_el = create_element_list(t)
@@ -183,8 +181,8 @@ if __name__ == '__main__':
                     ress = "Yes"
                     if not res:
                         ress = "No"
-                    print res
-                    rfile.write(ress+'\n')
[…]

### 2010-01-29 src/validate/get_exps.py

@@ -13,8 +13,16 @@ def get_exps():
                     "1091155753",  "1101523753",  "1122259753",  "1135219753",  "1145587753", 
                     "1093747753",  "1104115753",  "1124851753",  "1137811753",  "1148179753", 
                     "1096339753",  "1106707753",  "1130035753",  "1140403753",  "1150771753"]
-    mysql_periods = ["965099404" ,"967691404" ,"970283404" ,"972875404" ,"975467404" ,"978059404" ,"980651404" ,"983243404" ,"985835404" ,"988427404" ,"991019404" ,"993611404" ,"996203404" ,"998795404" ,"1001387404" ,"1003979404" ,"1006571404" ,"1009163404" ,"1011755404" ,"1014347404" ,"1016939404" ,"1019531404" ,"1022123404" ,"1024715404" ,"1027307404" ,"1029899404" ,"1032491404" ,"1035083404" ,"1037675404" ,"1040267404" ,"1042859404" ,"1045451404" ,"1048043404" ,"1050635404" ,"1053227404" ,"1055819404" ,"1058411404" ,"1061003404" ,"1063595404" ,"1066187404" ,"1068779404" ,"1071371404" ,"1073963404" ,"1076555404" ,"1079147404" ,"1081739404" ,"1084331404" ,"1086923404" ,"1089515404" ,"1092107404"]
-    maxdb_dir = "data/maxdb-tagged/"
-    mysql_dir = "data/mysql-tagged/"
+    mysql_periods = ["965099404" ,"967691404" ,"970283404" ,"972875404" ,"975467404" ,"978059404" ,
+                    "980651404" ,"983243404" ,"985835404" ,"988427404" ,"991019404" ,"993611404" ,
+                    "996203404" ,"998795404" ,"1001387404" ,"1003979404" ,"1006571404" ,"1009163404" ,
+                    "1011755404" ,"1014347404" ,"1016939404" ,"1019531404" ,"1022123404" ,"1024715404" ,
+                    "1027307404" ,"1029899404" ,"1032491404" ,"1035083404" ,"1037675404" ,"1040267404" ,
+                    "1042859404" ,"1045451404" ,"1048043404" ,"1050635404" ,"1053227404" ,"1055819404" ,
+                    "1058411404" ,"1061003404" ,"1063595404" ,"1066187404" ,"1068779404" ,"1071371404" ,
[…]

### 2010-01-29 icpc/icpc2010.tex
Added table with automated labeling results.

@@
 \begin{table*}[h]
-	\caption{Results for automatic topic labeling}
+	\caption{Automatic topic labeling}
 	\centering
@@
 Table \ref{tbl:maxdb-unsup-results} shows our results for MaxDB.
+In general results are poor.
+Using the F-Measure, the weighted average of precision and recall, where 1 is perfect, our best results are 0.6, a few at 0.4, and most around 0.2.
+Similar results using the Matthew's correlation coefficient (used to measure efficacy where classes are of different sizes), and the receiver operating characteristic area under curve.
+
+\begin{table*}[h]
+	\caption{Results for automatic topic labeling. f1 = f-measure, mcc = Matthew's correlation coeff., roca = ROC area}
+	\centering
+	\label{tbl:maxdb-unsup-results}
+\begin{tabular}{c|c|c|c|c|c|c}
+\toprule
+Experiment & label & f1 & mcc & precision & recall & roca  \\ 
+\midrule
+MaxDB exp2 & portability & 0.228 & 0.182 & 0.520 & 0.146 & 0.553 \\ 
+ & efficiency & 0.217 & 0.125 & 0.237 & 0.200 & 0.558\\ 
+ & reliability & 0.380 & 0.340 & 0.246 & 0.829 & 0.765 \\ 
+ & functionality & 0.095 & 0.083 & 0.250 & 0.059 & 0.521 \\ 
+ & maintainability & 0.092 & 0.123 & 0.571 & 0.050 & 0.520 \\ 
+ & usability & 0.175 & 0.138 & 0.113 & 0.389 & 0.620 \\ 
+ & total & 0.236 & 0.127 & 0.248 & 0.225 & 0.561 \\ 
+\midrule
+mysql exp2 & portability & 0.138 & 0.211 & 1.000 & 0.074 & 0.537 \\ 
+ & efficiency & 0.345 & 0.327 & 0.476 & 0.270 & 0.625 \\ 
+ & reliability & 0.425 & 0.287 & 0.348 & 0.545 & 0.669 \\ 
+ & functionality & 0.025 & 0.006 & 0.571 & 0.013 & 0.501 \\ 
+ & maintainability & 0.000 & 0.000 & 0.000 & 0.000 & 0.500 \\ 
+ & usability & 0.175 & 0.135 & 0.200 & 0.156 & 0.560 \\ 
+ & total & 0.167 & 0.095 & 0.403 & 0.105 & 0.527 \\ 
+\midrule
+maxdb exp3 & portability & 0.472 & 0.286 & 0.402 & 0.573 & 0.660 \\ 
+ & efficiency & 0.223 & 0.068 & 0.130 & 0.778 & 0.549 \\ 
+ & reliability & 0.257 & 0.196 & 0.149 & 0.927 & 0.652 \\ 
+ & functionality & 0.236 & 0.187 & 0.138 & 0.824 & 0.665 \\ 
+ & maintainability & 0.338 & 0.112 & 0.266 & 0.463 & 0.566 \\ 
+ & usability & 0.108 & 0.094 & 0.057 & 0.944 & 0.595 \\ 
+ & total & 0.258 & 0.093 & 0.160 & 0.671 & 0.568 \\ 
+\midrule
+mysql exp3 & portability & 0.413 & 0.170 & 0.564 & 0.325 & 0.574 \\ 
+ & efficiency & 0.158 & 0.105 & 0.089 & 0.703 & 0.608 \\ 
+ & reliability & 0.388 & 0.240 & 0.260 & 0.758 & 0.660 \\ 
+ & functionality & 0.652 & 0.240 & 0.655 & 0.649 & 0.620 \\ 
+ & maintainability & 0.203 & 0.007 & 0.240 & 0.175 & 0.503 \\ 
+ & usability & 0.105 & 0.013 & 0.057 & 0.688 & 0.513 \\ 
+ & total & 0.362 & 0.076 & 0.284 & 0.499 & 0.544 \\
+\bottomrule
+\end{tabular}
+\end{table*}
 

### 2010-01-29 lda-experiment.tex
misc changes to text
Added table with automated labeling results.

@@
 We ran it and we found that wordnet generated word lists for
-\emph{correctness} were too large andgeneric. Results are listed below
+\emph{correctness} were too large and generic. Results are listed below
 in the results section
@@
 traceability, verifiability 76
-verifiability, verifiability 78
-correctness, traceability 86
-traceability, traceability 95
-testability, testability 121
-correctness, correctness 182
+
 \end{verbatim}

### 2010-02-09 icpc/icpc2010.tex
updated refs and related work with Mockus 2000

@@
 
+Mockus and Votta studied a large-scale industrial change-tracking system in ~\cite{Mockus00}.
+They also leveraged Wordnet, but only for word roots.
+They felt the synonyms would be non-specific and cause errors.
+A nice contribution was access to system developers, with whom they could validate their labels.
+Since we try to bridge different organizations, these interviews are infeasible (particularly in the distributed world of open-source software).
+
 The other approach is to start with code repositories, and try to extract concerns from there.

### 2010-02-24 icpc/icpc2010.tex
minor icp2010 changes to get it to compile
just figuring out what to do

@@
 \usepackage{booktabs, colortbl}
-\usepackage{amsmath,amssymb}
-\usepackage{algpseudocode}
+\usepackage{amsmath}
+%\usepackage{amsmath,amssymb}
+%\usepackage{algpseudocode}
 \usepackage[pdfstartview=FitH,xetex,bookmarksopenlevel=3,bookmarks=true]{hyperref} %pdfpagescrop={92 112 523 778},
-\usepackage{algorithm}%,algorithmic-fix}
+%\usepackage{algorithm}%,algorithmic-fix}
 \usepackage{url}
@@
 \begin{abstract}
-  In automated mining of software repositories, most approaches, such as topic modeling and concept location, focus on applying standard machine learning algorithms to corpora.
+  In the automated mining of software repositories, most approaches, such as topic modeling and concept location, focus on applying standard machine learning algorithms to corpora.
 This approach does not make any use of domain-specific information to better understand results.
 While too much specificity can produce non-generalizable results, too little produces broad learners that don't provide much useful detail.%There has not been any   investigation into the latent topics these classifiers represent.
-In this paper, we investigate the latent topics classifier produce, and examine  whether such topics can be
-  generalized across software products. We focus on software quality
+In this paper, we investigate what topics the latent topics classifier produce, and examine  whether such topics can be
+  generalized across software products. We focus on nonfunctional requirements related to software quality
   as a potential generalization, since there is some shared belief
@@
 Unlike the poetic licence Shakespeare uses, we show that names are indeed important.
-In this paper we `name' topic models generated with a topic modeling technique, Latent Dirichlet Allocation~\cite{Blei2003}.
+In this paper we `name' and label topic models generated by a topic modeling technique, Latent Dirichlet Allocation~\cite{Blei2003}.
 We use these topic models to train supervised learners to classify larger datasets.
@@
 
+% Clean this up
 Traditionally topic modeling has required manual annotation to derive domain-relevant information.
@@
   compared to other approaches.
+% unsure about this
 \item We show that these topics can be learned and used to classify other datasets.
@@
 \section{Background}
+
 We provide a brief overview of software repository mining and information retrieval.
+
 \subsection{Definitions}
@@
 
-\emph{ROC} values reflect a score, similar to school letter-grades, reflecting how well a particular learner performed for the given data.
+% area under the ROC 
+% check
+\emph{Area of ROC Curve} is the area under the Reciever Operating Characteric curve, we will refer to this as \emph{ROC}.
+ROC values reflect a score, similar to school letter-grades, reflecting how well a particular learner performed for the given data.
 A ROC result of 0.5 would be equivalent to a random learner (randomly classifying data).
@@
 
+%Cite mockus et al
 To construct our expanded sets, we also used Wordnet~\cite{Fellbaum1998}, an English-language ‘lexical database’ that contains semantic relations between words, including meronymy and synonymy.
@@
 
-To assess the performance of the supervised learners, we did a 20-fold cross-over analysis (using 1 period to train the learner on the tags, and the other 19 to attempt to learn, changing the training set each time).
+To assess the performance of the supervised learners, we did a 10-fold cross-over analysis (using 1 period to train the learner on the tags, and the other 9 to attempt to learn, changing the training set each time).
 We report these results below.
 
+% enumerate?
 Finally, using this data, we evaluated two research questions: 1) Do label frequencies change over time?
@@
 \subsection{Word list similarity}
+
+% word list similarity
+
 In general, this approach did not work out well.

### 2010-02-27 src/validate/output/maxdb.mulan.out.transpose.tex
Mulan bug fixes and reporting!

\begin{tabular}{l|rrrrrrrrrrrrrrrrrr}

  & example-Accuracy & example-Fmeasure & example-HammingLoss & example-Precision & example-Recall & example-SubsetAccuracy & label-macro-AUC & label-macro-F1 & label-macro-Precision & label-macro-Recall & label-micro-AUC & label-micro-F1 & label-micro-Precision & label-micro-Recall & rank-AvgPrecision & rank-Coverage & rank-One-error & rank-Ranking Loss \\

\hline \\

BR & 0.49 & 0.55 & 0.19 & 0.50 & 0.61 & 0.40 & NaN & 0.32 & 0.29 & 0.38 & 0.76 & 0.39 & 0.34 & 0.47 & 0.54 & 1.40 & 0.81 & 0.41 \\

CLR & 0.47 & 0.50 & 0.14 & 0.49 & 0.50 & 0.43 & NaN & 0.19 & 0.31 & 0.15 & 0.66 & 0.27 & 0.42 & 0.21 & 0.57 & 1.23 & 0.79 & 0.35 \\

HOMER & 0.51 & 0.53 & 0.14 & 0.53 & 0.53 & 0.45 & NaN & 0.23 & 0.32 & 0.20 & 0.62 & 0.34 & 0.49 & 0.26 & 0.54 & 1.43 & 0.81 & 0.41 \\

IncludeLabels & 0.44 & 0.45 & 0.14 & 0.45 & 0.45 & 0.43 & NaN & 0.11 & 0.21 & 0.08 & 0.55 & 0.13 & 0.33 & 0.09 & 0.40 & 1.84 & 0.90 & 0.58 \\

LP & 0.42 & 0.44 & 0.17 & 0.45 & 0.43 & 0.38 & NaN & 0.19 & 0.23 & 0.17 & 0.62 & 0.25 & 0.30 & 0.22 & 0.40 & 1.93 & 0.89 & 0.60 \\

MLStacking & 0.47 & 0.48 & 0.13 & 0.49 & 0.47 & 0.43 & NaN & 0.12 & 0.18 & 0.10 & 0.68 & 0.21 & 0.49 & 0.14 & 0.60 & 1.16 & 0.76 & 0.32 \\

MLkNN & 0.43 & 0.43 & 0.13 & 0.43 & 0.43 & 0.43 & NaN & 0.00 & 0.00 & 0.00 & 0.65 & 0.00 & 0.00 & 0.00 & 0.58 & 1.20 & 0.79 & 0.33 \\

RAkEL & 0.48 & 0.50 & 0.13 & 0.51 & 0.50 & 0.43 & NaN & 0.26 & 0.40 & 0.23 & 0.68 & 0.34 & 0.49 & 0.27 & 0.51 & 1.54 & 0.79 & 0.46 \\

\end{tabular}

### 2010-02-27 src/validate/output/mysql.mulan.out.tex
Mulan bug fixes and reporting!

\begin{tabular}{l|rrrrrrrr}

  & BR & CLR & HOMER & IncludeLabels & LP & MLStacking & MLkNN & RAkEL \\

\hline \\

example-Accuracy & 0.45 & 0.40 & 0.50 & 0.40 & 0.40 & 0.38 & 0.33 & 0.48 \\

example-Fmeasure & 0.56 & 0.48 & 0.59 & 0.47 & 0.49 & 0.46 & 0.40 & 0.55 \\

example-HammingLoss & 0.24 & 0.19 & 0.18 & 0.19 & 0.23 & 0.19 & 0.20 & 0.17 \\

example-Precision & 0.50 & 0.51 & 0.60 & 0.52 & 0.52 & 0.50 & 0.44 & 0.60 \\

example-Recall & 0.65 & 0.46 & 0.58 & 0.44 & 0.47 & 0.43 & 0.37 & 0.52 \\

example-SubsetAccuracy & 0.19 & 0.24 & 0.31 & 0.24 & 0.23 & 0.23 & 0.20 & 0.31 \\

label-macro-AUC & 0.74 & 0.66 & 0.64 & 0.60 & 0.56 & 0.58 & 0.59 & 0.67 \\

label-macro-F1 & 0.46 & 0.30 & 0.43 & 0.28 & 0.32 & 0.27 & 0.22 & 0.39 \\

label-macro-Precision & 0.41 & 0.41 & 0.49 & 0.39 & 0.37 & 0.34 & 0.34 & 0.52 \\

label-macro-Recall & 0.56 & 0.26 & 0.40 & 0.24 & 0.31 & 0.23 & 0.19 & 0.33 \\

label-micro-AUC & 0.81 & 0.81 & 0.77 & 0.77 & 0.66 & 0.80 & 0.80 & 0.79 \\

label-micro-F1 & 0.59 & 0.53 & 0.61 & 0.51 & 0.50 & 0.51 & 0.45 & 0.59 \\

label-micro-Precision & 0.51 & 0.69 & 0.66 & 0.71 & 0.53 & 0.70 & 0.69 & 0.73 \\

label-micro-Recall & 0.68 & 0.43 & 0.56 & 0.41 & 0.48 & 0.41 & 0.34 & 0.49 \\

rank-AvgPrecision & 0.76 & 0.77 & 0.69 & 0.72 & 0.53 & 0.74 & 0.76 & 0.73 \\

rank-Coverage & 1.51 & 1.47 & 1.94 & 1.72 & 2.99 & 1.57 & 1.54 & 2.00 \\

rank-One-error & 0.40 & 0.39 & 0.47 & 0.44 & 0.63 & 0.43 & 0.40 & 0.36 \\

rank-Ranking Loss & 0.19 & 0.18 & 0.27 & 0.22 & 0.52 & 0.20 & 0.19 & 0.28 \\

\end{tabular}

### 2010-02-27 src/validate/mulan-parser.pl

@@ -2,21 +2,40 @@
 use strict;
 use MulanParser;
 use Fatal qw(open close);
+use LatexTable;
+use Data::Dumper;
+
+MulanParser::test();
 
 my $file = $ARGV[0];
 {
     my $fd;
-    open( $fd, $file );
+    open( $fd, $file ) or die "$file not found!";
     my @lines = <$fd>;
+    close($fd);
     my $h = MulanParser::mulan_parser( @lines );
-    my @learners = sort keys %$h;
+    my @learners = sort keys %{$h};
+    # warn "Learners: ".join(" , ",@learners);
     my @rows = sort keys %{$h->{$learners[0]}};
-    my $lt = LatexTable->create_table( rownames=>\@rows, colnames=> \@learners);
-    foreach my $learn (@learners) {
-        foreach my $metric (@rows) {
-            $lt->set_value($learn, $metric, $h->{$learn}->{$metric});
+    my $lt = LatexTable::create_table( rownames=>\@rows, 
+                                       colnames=> \@learners
+                                     );
+    print Dumper($lt);
+
+    my @arows = ();
+    foreach my $metric (@rows) {
+        my @nrow = ();
+        foreach my $learn (@learners) {
+            my $v = $h->{$learn}->{$metric};
+            push @nrow, $v;
+            #warn "$learn $metric $v";
+            #$lt->set_value($learn, $metric, $v);
         }
+        push @arows, \@nrow;
     }
+    $lt->rows(\@arows);
+    
+    print Dumper($lt);
     my $ltt = $lt->layout();
     my $lttt = $lt->transpose()->layout();
     my $fd;

### 2010-03-01 icpc/icpc2010.tex
intro para 1 what's in a name

@@
 
-Unlike the poetic licence Shakespeare uses, we show that names are indeed important.
+Unlike the poetic license Shakespeare uses, we show that names are indeed important.
+The rose has its own tangible properties which allow us to sense it.
+Unfortunately when we are dealing with data abstracted from program source code, we are already dealing with abstract entities that are not tangible.
+
 In this paper we `name' and label topic models generated by a topic modeling technique, Latent Dirichlet Allocation~\cite{Blei2003}.

### 2010-03-01 icpc/icpc2010.tex
intro editting

@@
 Unfortunately when we are dealing with data abstracted from program source code, we are already dealing with abstract entities that are not tangible.
+
+During the development of a software project developers touch on many issues integral to the purpose and specification of the software project.
+These issues are often referred to as topics.
+These topics can sometimes be automatically extracted from the changelog comments that developers create while committing revisions to the project's source control repository.
+Our previous work dealt with topic-trends, topics that repeat overtime.
+It seemed that these trends were often nonfunctional requirements.
+
+Topics in this paper and the previous paper are word bags or word distributions.
+These word distributions are found via Latent Dirichlet Allocation~\cite{Blei2003}, which find independent word distributions shared among documents (changelog comments).
+The unfortunate aspect of topics that are word distributions is that they lack the tangibility of a rose.
+They do not self idea themselves due to their tangible properties.
+Topics have to be identified by interpretting the prevelant words in the word distribution and by inspecting related documents.
+This is tiresome when one has to handle more than one hundred different topics.
+It would be nice if we could glance at a topic and have some idea what it was about.
+This is the power of labelling and naming.
+
+If a topic has a name or a label we immediately understand and register what it is. In our case the name is the bit of information that was missing from our topics to make them useful and complete.
+
+In this paper we attempt to name or label development topics that were extracted from the history of a development project. In part
 

### 2010-03-02 icpc/icpc2010.tex
introduction

@@
 This is tiresome when one has to handle more than one hundred different topics.
-It would be nice if we could glance at a topic and have some idea what it was about.
+It would be nice if we could, at a glance at , have some idea what the topic is about.
 This is the power of labelling and naming.
 
-If a topic has a name or a label we immediately understand and register what it is. In our case the name is the bit of information that was missing from our topics to make them useful and complete.
+If a topic has a name or a label we immediately understand and register what it is.
+In our case the name is the bit of information, the summary, that was missing from our topics to make them useful and complete.
+Although we are convinced that accurate naming is not completely feasible, so we have focussed on labelling topics by non-functional requirements.
 
-In this paper we attempt to name or label development topics that were extracted from the history of a development project. In part
-
-In this paper we `name' and label topic models generated by a topic modeling technique, Latent Dirichlet Allocation~\cite{Blei2003}.
-We use these topic models to train supervised learners to classify larger datasets.
-We also show that these topics can be generalized across products.
+%In this paper we \textit{name} and \textit{label} topics generated by %a topic modeling technique, Latent Dirichlet %Allocation~\cite{Blei2003}.
+We use these topic models to train %supervised learners to classify larger datasets.
+We also show that %these topics can be generalized across products.
 
 % Clean this up
-Traditionally topic modeling has required manual annotation to derive domain-relevant information.
-This paper introduces \emph{named topic models}, topic models which are named using a cross-project taxonomy.
-We start with an unsupervised topic learning algorithm, then manually annotate the topics using the ontology.
-From the annotated topics, we apply automated classifiers to find labels for the remaining topics.
+Traditionally topic extraction has required manual annotation to derive domain-relevant information.
+This paper attempts to implement \emph{labelled topic extraction}, topic extracted are given labels relating to a cross-project taxonomy.
+This is an ontological approach where we attempt to relate networks of words to labels and then search for these terms in our topic.
+We also compare this approach to machine leaners.
+%We start with an unsupervised topic learning algorithm, then manually %annotate the topics using the ontology.
+From the annotated topics, we %apply automated classifiers to find labels for the remaining topics.
 
+% XXX Not sure about this
+\begin{comment}
 Typically what is distinct is the software's particular functionality -- whether it offers email composition, prepares tax returns, and so on.
@@
 Does the quality occur more often in software commits?
+\end{comment}
 
@@
 \begin{itemize}
-\item  We introduce named topic analysis, and demonstrate its usefulness
+\item  We introduce labelled topic extraction, and demonstrate its usefulness
   compared to other approaches.

### 2010-03-02 icpc/icpc2010.tex
latest changes intro and background

@@
 
+\usepackage{comment}
 \newtheorem{mydef}{definition}
@@
 We provide a brief overview of software repository mining and information retrieval.
+This work is related to the mining software repositories (MSR)~\cite{msr} research community as it deals expressly with analyzing a project's source control repository, and the messages associated with revisions therein.
 
 \subsection{Definitions}
+We will use the following terminology in the rest of the paper.
 A \emph{message} is a block of text written by a developer. In this
@@
 topics that may provide key insights into the development of the
-project.
+project. A \emph{label} is part of a title we attach to a topic, whether manually or automatically.
 
@@
 % check
-\emph{Area of ROC Curve} is the area under the Reciever Operating Characteric curve, we will refer to this as \emph{ROC}.
-ROC values reflect a score, similar to school letter-grades, reflecting how well a particular learner performed for the given data.
+\emph{Area of ROC Curve} is the area under the Reciever Operating Characteric curve, we will refer to this as \emph{ROC} although it is sometimes referred to as \emph{AUC}.
+ROC values reflect a score, similar to school letter-grades (A is 0.9, C is 0.6), reflecting how well a particular learner performed for the given data.
 A ROC result of 0.5 would be equivalent to a random learner (randomly classifying data).
@@
 the false positive rate (1 - specificity).
-A perfect learner has a value of 1.0, reflecting perfect recall and precision.
+A perfect learner has a ROC value of 1.0, reflecting perfect recall and precision.
 
-\subsection{Topic models}
+\subsection{Topic and Concept Extraction}
 Topic analysis uses tools such as Latent Dirichlet Allocation (LDA)~\cite{Blei2003} and Latent Semantic Indexing (LSI) to extract
 independent word distributions (topics) from
-	documents (commit log comments). %~\cite{marcus04wcre,Poshyvanyk2007,lukins08wcre,Linstead2007}
+	documents (commit log comments).
+Many researchers~\cite{marcus04wcre,Poshyvanyk2007,lukins08wcre,Linstead2007} have applied tools like LSI and LDA for mining repositories, in particular analyzing source code, bugs or developer comments.
+\begin{comments}
 Ideally these extracted topics 
@@
 during the development of the software project.
-For example, our technique might identify a collection of words: `list next has iterator add', which we then identify (manually) as concerning the topics of collections and the Iterator pattern.
+For example, our technique might identify a collection of words: ``list next has iterator add'', which we then identify (manually) as concerning the topics of collections and the Iterator pattern.
+\end{comments}
+
+Typically a topic analysis tool like LDA will try to find $N$ indepdent word distributions found from within the word distributions of all the messages.
+
 In topic analysis a single document, such as a commit message, can be related to multiple topics.

### 2010-03-02 icpc/icpc2010.tex
methodology change

@@
 \subsection{Topic and Concept Extraction}
-Topic analysis uses tools such as Latent Dirichlet Allocation (LDA)~\cite{Blei2003} and Latent Semantic Indexing (LSI) to extract
+Topic extraction, sometimes called concept extraction, uses tools such as Latent Dirichlet Allocation (LDA)~\cite{Blei2003} and Latent Semantic Indexing (LSI) to extract
 independent word distributions (topics) from
@@
 
-Typically a topic analysis tool like LDA will try to find $N$ indepdent word distributions found from within the word distributions of all the messages.
+Typically a topic analysis tool like LDA will try to find $N$ independent word distributions found within the word distributions of all the messages.
+Linear combinations of these $N$ word distributions are then meant to be able to recreate the word distributions of all of the underlying messages.
+These $N$ word distributions effectively form topics: cross cutting collections of words relevant to 1 or more documents.
+Our problem is that these topics are not easy to interpret and we feel that automatic labelling or naming of these topics would be helpful with respect to interpreting what a topic is about.
+LDA's topics extracted in an unsupervised manner, LDA relies solely on the source data with no human intervention.
 
 In topic analysis a single document, such as a commit message, can be related to multiple topics.
-Representing documents as a mixture of topics maps well to commits to source code repositories, which often have more than one purpose.
-A topic
-represents both a word distribution and a group of commit log comments
-that are related to each other by their content.  In this paper a topic
-is a set of tokens extracted from commit messages found within a
-project's source control system (SCS).
+Representing documents as a mixture of topics maps well to commits to source code repositories, which often have more than one purpose~\cite{hindle09icsm}.
+A topic represents both a word distribution and a group of commit log comments that are related to each other by their content.
+In this paper a topic is a set of tokens extracted from commit messages found within a project's source control system (SCS).
 
+\begin{comment}
 In this paper we use LDA to identify topics.
@@
 These approaches are unsupervised in that they rely solely on the source data with no human intervention.
+\end{comment}
 
@@
 One slice is manually annotated by the domain expert
+In this paper, we employ the WEKA and Mulan machine learning frameworks in order to test supervised learning.
 
@@
 
-We wanted to compare two projects with open, accessible source code in similar functional areas.
+% XXX how many do we have
+We wanted to compare two mature software projects, that had open repositories, within the same domain.
 This allows us to partially control for differences in functional requirements.

### 2010-03-02 src/validate/output/maxdb.mulan.out.tex

\begin{tabular}{l|rrrrrrrr}

  & BR & CLR & HOMER & IncludeLabels & LP & MLStacking & MLkNN & RAkEL \\

\hline \\

example-Accuracy & 0.49 & 0.47 & 0.51 & 0.44 & 0.42 & 0.47 & 0.43 & 0.48 \\

example-Fmeasure & 0.55 & 0.50 & 0.53 & 0.45 & 0.44 & 0.48 & 0.43 & 0.50 \\

example-HammingLoss & 0.19 & 0.14 & 0.14 & 0.14 & 0.17 & 0.13 & 0.13 & 0.13 \\

example-Precision & 0.50 & 0.49 & 0.53 & 0.45 & 0.45 & 0.49 & 0.43 & 0.51 \\

example-Recall & 0.61 & 0.50 & 0.53 & 0.45 & 0.43 & 0.47 & 0.43 & 0.50 \\

example-SubsetAccuracy & 0.40 & 0.43 & 0.45 & 0.43 & 0.38 & 0.43 & 0.43 & 0.43 \\

label-macro-AUC & NaN & NaN & NaN & NaN & NaN & NaN & NaN & NaN \\

label-macro-F1 & 0.32 & 0.19 & 0.23 & 0.11 & 0.19 & 0.12 & 0.00 & 0.26 \\

label-macro-Precision & 0.29 & 0.31 & 0.32 & 0.21 & 0.23 & 0.18 & 0.00 & 0.40 \\

label-macro-Recall & 0.38 & 0.15 & 0.20 & 0.08 & 0.17 & 0.10 & 0.00 & 0.23 \\

label-micro-AUC & 0.76 & 0.66 & 0.62 & 0.55 & 0.62 & 0.68 & 0.65 & 0.68 \\

label-micro-F1 & 0.39 & 0.27 & 0.34 & 0.13 & 0.25 & 0.21 & 0.00 & 0.34 \\

label-micro-Precision & 0.34 & 0.42 & 0.49 & 0.33 & 0.30 & 0.49 & 0.00 & 0.49 \\

label-micro-Recall & 0.47 & 0.21 & 0.26 & 0.09 & 0.22 & 0.14 & 0.00 & 0.27 \\

rank-AvgPrecision & 0.54 & 0.57 & 0.54 & 0.40 & 0.40 & 0.60 & 0.58 & 0.51 \\

rank-Coverage & 1.40 & 1.23 & 1.43 & 1.84 & 1.93 & 1.16 & 1.20 & 1.54 \\

rank-One-error & 0.81 & 0.79 & 0.81 & 0.90 & 0.89 & 0.76 & 0.79 & 0.79 \\

rank-Ranking Loss & 0.41 & 0.35 & 0.41 & 0.58 & 0.60 & 0.32 & 0.33 & 0.46 \\

\end{tabular}

### 2010-03-02 src/validate/output/mysql.mulan.out.transpose.tex

\begin{tabular}{l|rrrrrrrrrrrrrrrrrr}

  & example-Accuracy & example-Fmeasure & example-HammingLoss & example-Precision & example-Recall & example-SubsetAccuracy & label-macro-AUC & label-macro-F1 & label-macro-Precision & label-macro-Recall & label-micro-AUC & label-micro-F1 & label-micro-Precision & label-micro-Recall & rank-AvgPrecision & rank-Coverage & rank-One-error & rank-Ranking Loss \\

\hline \\

BR & 0.45 & 0.56 & 0.24 & 0.50 & 0.65 & 0.19 & 0.74 & 0.46 & 0.41 & 0.56 & 0.81 & 0.59 & 0.51 & 0.68 & 0.76 & 1.51 & 0.40 & 0.19 \\

CLR & 0.40 & 0.48 & 0.19 & 0.51 & 0.46 & 0.24 & 0.66 & 0.30 & 0.41 & 0.26 & 0.81 & 0.53 & 0.69 & 0.43 & 0.77 & 1.47 & 0.39 & 0.18 \\

HOMER & 0.50 & 0.59 & 0.18 & 0.60 & 0.58 & 0.31 & 0.64 & 0.43 & 0.49 & 0.40 & 0.77 & 0.61 & 0.66 & 0.56 & 0.69 & 1.94 & 0.47 & 0.27 \\

IncludeLabels & 0.40 & 0.47 & 0.19 & 0.52 & 0.44 & 0.24 & 0.60 & 0.28 & 0.39 & 0.24 & 0.77 & 0.51 & 0.71 & 0.41 & 0.72 & 1.72 & 0.44 & 0.22 \\

LP & 0.40 & 0.49 & 0.23 & 0.52 & 0.47 & 0.23 & 0.56 & 0.32 & 0.37 & 0.31 & 0.66 & 0.50 & 0.53 & 0.48 & 0.53 & 2.99 & 0.63 & 0.52 \\

MLStacking & 0.38 & 0.46 & 0.19 & 0.50 & 0.43 & 0.23 & 0.58 & 0.27 & 0.34 & 0.23 & 0.80 & 0.51 & 0.70 & 0.41 & 0.74 & 1.57 & 0.43 & 0.20 \\

MLkNN & 0.33 & 0.40 & 0.20 & 0.44 & 0.37 & 0.20 & 0.59 & 0.22 & 0.34 & 0.19 & 0.80 & 0.45 & 0.69 & 0.34 & 0.76 & 1.54 & 0.40 & 0.19 \\

RAkEL & 0.48 & 0.55 & 0.17 & 0.60 & 0.52 & 0.31 & 0.67 & 0.39 & 0.52 & 0.33 & 0.79 & 0.59 & 0.73 & 0.49 & 0.73 & 2.00 & 0.36 & 0.28 \\

\end{tabular}

### 2010-03-03 icpc/icpc2010.tex
minimal edit
generating the data

@@
 For example, in our previous work~\cite{Hindle09ICSM}, and in \cite{Baldi2008}, topics are named manually: human experts read the highest-frequency members of a topic and assign a keyword accordingly.
-E.g., for the word list ``listener change remove add fire", Baldi et al.
-assign the keyword `event-handling'.
+E.g., for the word list ``listener change remove add fire'', Baldi et al.
+assign the keyword \emph{event-handling}.
 The labels are reasonable enough, but still require some expert in the field to determine them.
@@
 \section{Methodology}
+
+\begin{comment}
 % To gather the data, we first extract the commit log comments from a
@@
 % depicts the general process for processing and analyzing the commit messages.
+\end{comment}
 
 % XXX how many do we have
-We wanted to compare two mature software projects, that had open repositories, within the same domain.
+We wanted to compare two mature software projects, that had open repositories, within the same application domain.
 This allows us to partially control for differences in functional requirements.
@@
 MaxDB started in the late 1970s as a research project, and was later acquired by SAP.
-As of version 7.500, released April 2007, the project has 940 thousand lines of C source code\footnote{generated using David A.
-Wheeler's `SLOCCount'.}.
-There are approximately ...
-developers working on it.
-MySQL 3.23, released in early 2001, contains 320 thousand lines of C and C++ source code.
-The MySQL project started in 1994.
+As of version 7.500, released April 2007, the project has 940 thousand lines of C source code~\footnote{generated using David A.
+Wheeler's \emph{SLOCCount}.}.
+%How many developers?
+%There are approximately ... developers working on it. 
+%XXX version #?
+The MySQL project started in 1994 but MySQL 3.23 was released in early 2001. MySQL contains 320 thousand lines of C and C++ source code.
 
 \subsection{Generating the data}
-From each project, we used source code commit data (the message attached to the commit).
-We leveraged the data we gathered in \cite{Hindle09ICSM} for this work.
-A typical commit message is: ``history annotate diffs bug fixed (if mysql\_real\_connect() failed there were two pointers to malloc'ed strings, with memory corruption on free(), of course) ".
+From each project, we used source control commit comments, these are the messages that programmers write when they commit revisions to a source control repository.
+We leveraged the data that we gathered in \cite{Hindle09ICSM} for this work.
+An example of a typical commit message is: ``history annotate diffs bug fixed (if mysql\_real\_connect() failed there were two pointers to malloc'ed strings, with memory corruption on free(), of course)''.
 We extracted these messages and indexed them by creation time.
-We apply stop-word removal to eliminate common English words like `the', `at', and so on.
+We summarized each message as a word distribution but removed stop-words such as   common English words like \emph{the} and \emph{at}.
 
-From that dataset, we created an XML file which grouped commits into temporal periods, each 30 days wide.
+From that dataset, we created an XML file which windowed commits into
+monthly windows.
 This size of period is smaller than the time between minor releases but large enough for there to be sufficient commits to analyze.
-Given a text file full of commits, indexed by date and time, we then ran an LDA modeler on each period, to generate a list of topics per period.
-E.g., in Period 1, the most important topic might have been ``security", with ``sql'' second, down on to less relevant topics.
-We arbitrarily set the number of topics to generate to 20, because past experimentation showed that fewer topics might aggregate multiple unique topics while any more topics seemed to dilute the results and create indistinct topics.
+We applied an Blei's LDA implementation against the word distributions of these commits, and generated list of topics per period.
+%A topic extracted in the first period Period 1, the most important %topic might have been ``security", with ``sql'' second, down on to %less relevant topics. 
+We arbitrarily set the number of topics to generate to $20$, because past experimentation showed that fewer topics might aggregate multiple unique topics while any more topics seemed to dilute the results and create indistinct topics.
+As well, more than $20$ topics quickly became infeasible for inspection and it was difficult to discern the difference in topics.
 
 \subsection{Associating labels}
+\begin{comment}
 % We would like to `name' our topic models for each period. In this way, we can see how certain words -- subjects -- occur more or less often with each period.
+\end{comment}
+
+Topics are word distributions, essentially ranked lists of words, thus they can be burdensome to interpret and hard to distinguish.
 Once we had topics for each period, we tried to associate them with a label from a list of keywords and related terms.

### 2010-03-03 icpc/icpc2010.tex
acm stuff

@@
 %!TEX encoding = UTF-8 Unicode
-\documentclass[10pt, conference, compsocconf]{IEEEtran}
+%\documentclass[10pt, conference, compsocconf]{IEEEtran}
+\documentclass{acm_proc_article-sp}
 \usepackage{fontspec}
@@
 Many researchers~\cite{marcus04wcre,Poshyvanyk2007,lukins08wcre,Linstead2007} have applied tools like LSI and LDA for mining repositories, in particular analyzing source code, bugs or developer comments.
-\begin{comments}
+\begin{comment}
 Ideally these extracted topics 
@@
 For example, our technique might identify a collection of words: ``list next has iterator add'', which we then identify (manually) as concerning the topics of collections and the Iterator pattern.
-\end{comments}
+\end{comment}
 
@@
 In topic analysis a single document, such as a commit message, can be related to multiple topics.
-Representing documents as a mixture of topics maps well to commits to source code repositories, which often have more than one purpose~\cite{hindle09icsm}.
+Representing documents as a mixture of topics maps well to commits to source code repositories, which often have more than one purpose~\cite{Hindle09ICSM}.
 A topic represents both a word distribution and a group of commit log comments that are related to each other by their content.
@@
 We extracted these messages and indexed them by creation time.
-We summarized each message as a word distribution but removed stop-words such as   common English words like \emph{the} and \emph{at}.
+We summarized each emessage as a word distribution but removed stop-words such as   common English words like \emph{the} and \emph{at}.
 
@@
 
-Topics are word distributions, essentially ranked lists of words, thus they can be burdensome to interpret and hard to distinguish.
+Topics are word distributions, essentially lists of words ranked by frequency, thus they can be burdensome to interpret and hard to distinguish.
 Once we had topics for each period, we tried to associate them with a label from a list of keywords and related terms.
-Recall that topics are lists of words and associated frequencies.
 We performed simple string matching between these topics and our lists, `naming' a topic if it contained that word or words.
@@
 
-Our second list relied on the ISO quality model~\cite{iso9126}.
+Our second list relied on the ISO quality model (ISO9126)~\cite{iso9126}.
 It describes six high-level quality requirements (listed in Table \ref{tbl:wnsig}).

### 2010-03-03 icpc/icpc2010.tex
more edits acm style
more edits

@@
 \begin{document}
-
+\numberofauthors{5}
 \author{
-\IEEEauthorblockN{Abram Hindle}
-\IEEEauthorblockA{Department of Computer Science\\University of Waterloo\\ ahindle@uwaterloo.ca\\}
+\alignauthor 
+Abram Hindle\\
+   \affaddr{David Cheriton School of Computer Science}\\
+   \affaddr{University of Waterloo}\\
+   \affaddr{Waterloo, Ontario, CANADA}\\
+   \email{ahindle@uwaterloo.ca}
+\alignauthor  Neil A. Ernst\\
+   \affaddr{Department of Computer Science}\\
+   \affaddr{University of Toronto}\\
+   \affaddr{Toronto, Ontario, CANADA}\\
+   \email{nernst@cs.toronto.edu}
 \and
-\IEEEauthorblockN{Neil A. Ernst}
-\IEEEauthorblockA{Department of Computer Science\\ University of Toronto\\ nernst@cs.toronto.edu}
-}
+\alignauthor Michael W. Godfrey\\
+   \affaddr{David Cheriton School of Computer Science}\\
+   \affaddr{University of Waterloo}\\
+   \affaddr{Waterloo, Ontario, CANADA}\\
+   \email{migod@uwaterloo.ca}
+\alignauthor Richard C. Holt \\
+   \affaddr{David Cheriton School of Computer Science}\\
+   \affaddr{University of Waterloo}\\
+   \affaddr{Waterloo, Ontario, CANADA}\\
+   \email{holt@uwaterloo.ca}
+\alignauthor Neil's Supervisor\\
+   \affaddr{Department of Computer Science}\\
+   \affaddr{University of Toronto}\\
+   \affaddr{Toronto, Ontario, CANADA}\\
+   \email{@cs.toronto.edu}}
 
@@
 
-Topics are word distributions, essentially lists of words ranked by frequency, thus they can be burdensome to interpret and hard to distinguish.
+Topics are word distributions: essentially lists of words ranked by frequency, thus they can be burdensome to interpret and hard to distinguish and understand.
 Once we had topics for each period, we tried to associate them with a label from a list of keywords and related terms.
@@
 
-Our second list relied on the ISO quality model (ISO9126)~\cite{iso9126}.
+Our second list, \textsf{exp2}, relied on the ISO quality model (ISO9126)~\cite{iso9126}.
 It describes six high-level quality requirements (listed in Table \ref{tbl:wnsig}).
@@
 %Cite mockus et al
-To construct our expanded sets, we also used Wordnet~\cite{Fellbaum1998}, an English-language ‘lexical database’ that contains semantic relations between words, including meronymy and synonymy.
+
+%This is confusing/confused
+%XXX redo this 
+To construct our expanded wordlists, we used Wordnet~\cite{Fellbaum1998}, an English-language ``lexical database'' that contains semantic relations between words, including meronymy and synonymy.
 We then added Boehm’s 1976 software quality model \cite{Boehm+:1976:ICSE}, and classified his eleven ‘ilities’ into their respective ISO9126 qualities.
@@
 \subsection{Supervised learning}
-To check our outcomes, we manually annotated each topic in each period with the same quality labels as \textsf{exp2}.
+
+In order to validate how effective these wordbag approaches to topic labelling would be we had to make a data set to test against.
+For MySQL and MaxDB, we manually annotated each extracted topic in each period with the same quality labels as \textsf{exp2} (software qualities).
 We looked at each period's topics, and assessed what the data -- consisting of the frequency-weighted word lists and messages -- suggested was the topic for that period.
+We were able to pinpoint the appropriate label using auxillary information as well, such as the actual revisions and files that were related to this topic.
 For example, for the MaxDB topic consisting of a message ``exit() only used in non NPTL LINUX Versions'', we tagged that topic \emph{portability}.
-This provided us with manually classified topics for the second phase of our analysis, supervised learning (classification).
+We compared against this dataset but we also used it for our supervised learning based topic classification.
 %inter-rater reliability
@@
 For supervised learning, we used a suite of supervised classifiers from WEKA~\cite{weka09}.
-WEKA contains a suite of machine learning tools (e.g., support vector machines and Bayesian nets).
+WEKA contains a suite of machine learning tools such as support vector machines and Bayes nets.
 We also used the multi-labeling add-on for WEKA, Mulan\footnote{\url{http://mlkd.csd.auth.gr/multilabel.html}}.
-Traditional classifiers map our topics to a single class, whereas Mulan allows for a mixture of classes per topic (much like our manual classification found overlaps).
+Traditional classifiers map our topics to a single class, whereas Mulan allows for a mixture of classes per topic, much like our manual classification found overlaps.
 
-To assess the performance of the supervised learners, we did a 10-fold cross-over analysis (using 1 period to train the learner on the tags, and the other 9 to attempt to learn, changing the training set each time).
-We report these results below.
+To assess the performance of the supervised learners, we did a 10-fold cross-validation.
+This is when a set is partitioned into 10 partitions and then each partition is used once as a test set and 9 other times as part of the training set of 9 paritions.
+We haved reported these results below.
 
 % enumerate?
-Finally, using this data, we evaluated two research questions: 1) Do label frequencies change over time?
-That is, is a certain quality of more interest at one point in the lifecycle than some other?
-2) Do the different projects differ in their relative topic interest?
-That is, is a particular quality more important to one project than the other projects?
+Finally, using this data, we evaluated two research questions:
+\begin{enumerate}
+\item Do label frequencies change over time? That is, is a certain quality of more interest at one point in the life-cycle than some other? 
+\item  Do the different projects differ in their relative topic interest? That is, is a particular quality more important to one project than the other projects?  
+\end{enumerate}
 
@@
 
+%XXX this table should have mysql
 \begin{table*}[h]
@@
 
-For \textsf{exp1}, our best performing labels (the labels matched with the most topics) were correctness (182 topics) and testability (121).
+%This is the counts
+
+For \textsf{exp1}, our best performing labels (the labels matched with the most topics) were \emph{correctness} (182 topics) and \emph{testability} (121).
 We did not get good results for usability or accuracy, which were associated with fewer than ten topics.
@@
 \begin{table*}[h]
-	\caption{Results for automatic topic labeling. f1 = f-measure, mcc = Matthew's correlation coeff., roca = ROC area}
+	\caption{Results for automatic topic labeling. f1 = f-measure, mcc = Matthew's correlation coeff., ROC = Area under ROC curve}
 	\centering
@@
 \toprule
-Experiment & label & f1 & mcc & precision & recall & roca  \\ 
+Experiment & label & f1 & mcc & precision & recall & ROC  \\ 
 \midrule
@@
 \subsection{Classifying topics}
+%XXX oh god more more more wtf
 We present a ROC chart in Fig.
@@
 
+%XXX do we have time here
+
 \subsection{Topic change over time}
 
+%XXX I don't know!
+
 \subsection{Threats to validity}
-\textbf{Construct validity} – Focus on code commit messages, rather than mail or bug trackers. Choice of taxonomy
+\textbf{Construct validity} -- Focus on code commit messages, rather than mail or bug trackers. Choice of taxonomy
 
@@
 
-\textbf{External validity} – Our data originated from open-source projects in the database domain.
+\textbf{External validity} -- Our data originated from open-source projects in the database domain.
 Of these, the open-source nature of the project seems most problematic for external validity.
@@
 \section{Related work}
+
 Part of our effort with this project is to understand the qualitative and intentional aspects of requirements in software evolution, a notion we first discussed in \cite{ernst07icsm}.
@@
 \section{Conclusions and future work}
-In this paper we showed that a domain specific wordnet for software development provides better accuracy for labelling topic models than unsupervised approaches.
+
+% No we didn't!
+% In this paper we showed that a domain specific wordnet for software % development provides better accuracy for labelling topic models than % unsupervised approaches.
+%
+
+% XXX WRITE THIS
 

### 2010-03-04 icpc/icpc2010.tex
added weka results

@@
 \subsection{Classifying topics}
-%XXX oh god more more more wtf
-We present a ROC chart in Fig.
-\ref{}.
+
+%XXX Describe table
+
+We present a ROC chart in Figure \ref{fig:best-learn-per-tag}.
 The chart shows each learner measured in terms of precision and recall.
 In general, we prefer techniques which are closer to the upper-left corner (maximizing precision and maximizing recall).
+
+% Cribbed from: 
+% src/validate/latex-out/best-learner-per-project-roc-table.tx
+\begin{table*}
+\begin{tabular}{ccc|rrrr}
+Word & Project & Learner & ROC & ZeroR Acc. & ZeroR. Diff & Fmeasure\\
+\hline
+portability &  mysql &  NaiveBayesMultinomial &  0.74 &  58.53 &  11.43 &  0.70  \\ 
+efficiency &  mysql &  NaiveBayes &  0.67 &  93.69 &  -7.68 &  0.88 \\ 
+reliability &  mysql &  NaiveBayes &  0.73 &  83.11 &  -12.80 &  0.74 \\ 
+functionality &  mysql &  DMNBtext &  0.81 &  54.44 &  21.67 &  0.76 \\ 
+maintainability &  mysql &  DMNBtext &  0.78 &  76.62 &  3.41 &  0.75 \\ 
+usability &  mysql &  NaiveBayes &  0.75 &  94.54 &  -5.80 &  0.90  \\ 
+ \hline 
+portability &  maxdb &  NaiveBayes &  0.84 &  77.12 &  2.06 &  0.80 \\ 
+efficiency &  maxdb &  NaiveBayes &  0.62 &  88.43 &  -11.31 &  0.79 \\ 
+reliability &  maxdb &  DMNBtext &  0.84 &  89.46 &  3.86 &  0.92  \\ 
+functionality &  maxdb &  NaiveBayes &  0.67 &  91.26 &  -6.94 &  0.86  \\ 
+maintainability &  maxdb &  NaiveBayes &  0.70 &  79.43 &  -9.25 &  0.72  \\ 
+usability &  maxdb &  NaiveBayes &  0.56 &  95.37 &  -4.37 &  0.91  \\ 
+ \hline 
+\end{tabular}
+\caption{Per label, per project, the best learner for that label}
+\label{fig:best-learner-per-tag}
+\end{table*}
+
+\subsection{Classifying multiple topic tagging}
+
+%XXX put mulan stuff here
 

### 2010-03-04 icpc/icpc2010.tex
mulan results
Not sure

@@
 \usepackage{color}
+
+\newcommand{\XXX}[1]{\textcolor{red}{{\it \textbf{[XXX: #1]}}}}
 
@@
 \subsubsection{Analysing the unsupervised labeling}
+
 Based on the labels, and our manual topic labeling, we compared the results of the unsupervised word matching approach.
 For each quality we tried to assess whether the manual tag matched the unsupervised label assigned.
-Table \ref{tbl:maxdb-unsup-results} shows our results for MaxDB.
+Table \ref{tbl:maxdb-unsup-results} shows our results for MaxDB and MySQL.
 In general results are poor.
@@
 
+%XXX Describe learners
+
 We present a ROC chart in Figure \ref{fig:best-learn-per-tag}.
@@
 
+\begin{table*}
+\begin{tabular}{l|rrrrrrrr}
+
+  & BR & CLR & HOMER & IncludeLabels & LP & MLStacking & MLkNN & RAkEL \\
+
+\hline \\
+
+example-Accuracy & 0.45 & 0.40 & 0.50 & 0.40 & 0.40 & 0.38 & 0.33 & 0.48 \\
+
+example-Fmeasure & 0.56 & 0.48 & 0.59 & 0.47 & 0.49 & 0.46 & 0.40 & 0.55 \\
+
+example-HammingLoss & 0.24 & 0.19 & 0.18 & 0.19 & 0.23 & 0.19 & 0.20 & 0.17 \\
+
+example-Precision & 0.50 & 0.51 & 0.60 & 0.52 & 0.52 & 0.50 & 0.44 & 0.60 \\
+
+example-Recall & 0.65 & 0.46 & 0.58 & 0.44 & 0.47 & 0.43 & 0.37 & 0.52 \\
+
+example-SubsetAccuracy & 0.19 & 0.24 & 0.31 & 0.24 & 0.23 & 0.23 & 0.20 & 0.31 \\
+
+label-macro-AUC & 0.74 & 0.66 & 0.64 & 0.60 & 0.56 & 0.58 & 0.59 & 0.67 \\
+
+label-macro-F1 & 0.46 & 0.30 & 0.43 & 0.28 & 0.32 & 0.27 & 0.22 & 0.39 \\
+
+label-macro-Precision & 0.41 & 0.41 & 0.49 & 0.39 & 0.37 & 0.34 & 0.34 & 0.52 \\
+
+label-macro-Recall & 0.56 & 0.26 & 0.40 & 0.24 & 0.31 & 0.23 & 0.19 & 0.33 \\
+
+label-micro-AUC & 0.81 & 0.81 & 0.77 & 0.77 & 0.66 & 0.80 & 0.80 & 0.79 \\
+
+label-micro-F1 & 0.59 & 0.53 & 0.61 & 0.51 & 0.50 & 0.51 & 0.45 & 0.59 \\
+
+label-micro-Precision & 0.51 & 0.69 & 0.66 & 0.71 & 0.53 & 0.70 & 0.69 & 0.73 \\
+
+label-micro-Recall & 0.68 & 0.43 & 0.56 & 0.41 & 0.48 & 0.41 & 0.34 & 0.49 \\
+
+rank-AvgPrecision & 0.76 & 0.77 & 0.69 & 0.72 & 0.53 & 0.74 & 0.76 & 0.73 \\
+
+rank-Coverage & 1.51 & 1.47 & 1.94 & 1.72 & 2.99 & 1.57 & 1.54 & 2.00 \\
+
+rank-One-error & 0.40 & 0.39 & 0.47 & 0.44 & 0.63 & 0.43 & 0.40 & 0.36 \\
+
+rank-Ranking Loss & 0.19 & 0.18 & 0.27 & 0.22 & 0.52 & 0.20 & 0.19 & 0.28 \\
+
+\end{tabular}
+\caption{Mysql Mulan results per Mulan learner}
+\end{table*}
+
+\begin{table*}
+% Cribbed from:
+% src/validate/output/maxdb.mulan.out.tex
+\begin{tabular}{l|rrrrrrrr}
+
+  & BR & CLR & HOMER & IncludeLabels & LP & MLStacking & MLkNN & RAkEL \\
+
+\hline \\
+
+example-Accuracy & 0.49 & 0.47 & 0.51 & 0.44 & 0.42 & 0.47 & 0.43 & 0.48 \\
+
+example-Fmeasure & 0.55 & 0.50 & 0.53 & 0.45 & 0.44 & 0.48 & 0.43 & 0.50 \\
+
+example-HammingLoss & 0.19 & 0.14 & 0.14 & 0.14 & 0.17 & 0.13 & 0.13 & 0.13 \\
+
+example-Precision & 0.50 & 0.49 & 0.53 & 0.45 & 0.45 & 0.49 & 0.43 & 0.51 \\
+
+example-Recall & 0.61 & 0.50 & 0.53 & 0.45 & 0.43 & 0.47 & 0.43 & 0.50 \\
+
+example-SubsetAccuracy & 0.40 & 0.43 & 0.45 & 0.43 & 0.38 & 0.43 & 0.43 & 0.43 \\
+
+label-macro-AUC & NaN & NaN & NaN & NaN & NaN & NaN & NaN & NaN \\
+
+label-macro-F1 & 0.32 & 0.19 & 0.23 & 0.11 & 0.19 & 0.12 & 0.00 & 0.26 \\
+
+label-macro-Precision & 0.29 & 0.31 & 0.32 & 0.21 & 0.23 & 0.18 & 0.00 & 0.40 \\
+
+label-macro-Recall & 0.38 & 0.15 & 0.20 & 0.08 & 0.17 & 0.10 & 0.00 & 0.23 \\
+
+label-micro-AUC & 0.76 & 0.66 & 0.62 & 0.55 & 0.62 & 0.68 & 0.65 & 0.68 \\
+
+label-micro-F1 & 0.39 & 0.27 & 0.34 & 0.13 & 0.25 & 0.21 & 0.00 & 0.34 \\
+
+label-micro-Precision & 0.34 & 0.42 & 0.49 & 0.33 & 0.30 & 0.49 & 0.00 & 0.49 \\
+
+label-micro-Recall & 0.47 & 0.21 & 0.26 & 0.09 & 0.22 & 0.14 & 0.00 & 0.27 \\
+
+rank-AvgPrecision & 0.54 & 0.57 & 0.54 & 0.40 & 0.40 & 0.60 & 0.58 & 0.51 \\
+
+rank-Coverage & 1.40 & 1.23 & 1.43 & 1.84 & 1.93 & 1.16 & 1.20 & 1.54 \\
+
+rank-One-error & 0.81 & 0.79 & 0.81 & 0.90 & 0.89 & 0.76 & 0.79 & 0.79 \\
+
+rank-Ranking Loss & 0.41 & 0.35 & 0.41 & 0.58 & 0.60 & 0.32 & 0.33 & 0.46 \\
+
+\end{tabular}
+
+\caption{Maxdb Mulan results per Mulan learner}
+
+\end{table*}
 %XXX put mulan stuff here
+
+\subsection{Summary of techniques}
+
+%XXX Compare the results of Exp2 and Exp3 
+%    versus the mulan results and weka results
 

### 2010-03-04 icpc/icpc2010.tex
XXX

@@
 
-% XXX Not sure about this
+\XXX{Not sure about this}
 \begin{comment}
@@
 
-% XXX how many do we have
+\XXX{how many do we have}
 We wanted to compare two mature software projects, that had open repositories, within the same application domain.
@@
 %There are approximately ... developers working on it. 
-%XXX version #?
+\{XXX version #?}
 The MySQL project started in 1994 but MySQL 3.23 was released in early 2001. MySQL contains 320 thousand lines of C and C++ source code.
@@
 %This is confusing/confused
-%XXX redo this 
+\XXX{redo this}
 To construct our expanded wordlists, we used Wordnet~\cite{Fellbaum1998}, an English-language ``lexical database'' that contains semantic relations between words, including meronymy and synonymy.
@@
 
-%XXX this table should have mysql
+\XXX{this table should have mysql}
 \begin{table*}[h]
@@
 
-%XXX Describe table
+\XXX{Describe table}
 
-%XXX Describe learners
+\XXX{Describe learners}
 
@@
 \end{table*}
-%XXX put mulan stuff here
+\XXX{put mulan stuff here}
 
@@
 
-%XXX Compare the results of Exp2 and Exp3 
+\XXX{Compare the results of Exp2 and Exp3}
 %    versus the mulan results and weka results
@@
 
-%XXX do we have time here
+\XXX{do we have time here}
 
@@
 
-%XXX I don't know!
+\XXX{I don't know!}
 
@@
 
-% XXX WRITE THIS
+\XXX{WRITE THIS}
 

### 2010-03-04 icpc/icpc2010.tex
latest changes about results

@@
 %There are approximately ... developers working on it. 
-\{XXX version #?}
+\XXX{version number?}
 The MySQL project started in 1994 but MySQL 3.23 was released in early 2001. MySQL contains 320 thousand lines of C and C++ source code.
@@
 
-We present a ROC chart in Figure \ref{fig:best-learn-per-tag}.
-The chart shows each learner measured in terms of precision and recall.
-In general, we prefer techniques which are closer to the upper-left corner (maximizing precision and maximizing recall).
+We took our annotated dataset and applied supervised learner to it.
+Because our dataset was of word counts we expected Bayesian techniques, often used in spam filtering, to perform well.
+Table \ref{tab:best-learn-per-tag} shows the performance of the best performing learner per label.
+We considered the best learner for a label to be the one which had the highest ROC value for that label.
+Table \ref{tab:best-learn-per-tag} shows the ROC value, the accuracy of a ZeroR learner ( a learner which just chooses the largest category all of the time), the difference in accuracy between the best learner and ZeroR, and the F-measure for that particular learner.
+The ZeroR difference is often negative, but in particular for labels which are not as common this can be expected because any miscategorization will hurt accuracy, this is why the Fmeasure and the ROC values are useful as they can better present performance on labels which are not applicable to the majority of samples.
+
+Table \ref{tab:best-learn-per-tag} shows that MaxDB and MySQL have quite different results, the ROC values for reliability and functionality seem swapped between projects.
+It should be noted that for both project Bayesian techniques did the best out of a wide variety of machine learners tested.
+Discriminative Multinomial Naive Bayes (DMNBtext), Naive Bayes (NaiveBayes) and Multinomial Naive Bayes (NaiveBayesMultinomal) are all based on Bayes theorem and all assume, naively, that the features are independent.
+The features we used are word counts per message.
+One beneficial aspect of this result is that it suggests we can have very fast training times since Naive Bayes can be calculated in $O(N)$ for $N$ features.
+%http://nlp.stanford.edu/IR-book/html/htmledition/naive-bayes-text-classification-1.html
+
+It seems like the smaller the label the harder it is to get accurate results. Although these results seem better than our previous word bag results of \textsf{exp2} and \textsf{exp3}.
 
@@
 \caption{Per label, per project, the best learner for that label}
-\label{fig:best-learner-per-tag}
+\label{tab:best-learner-per-tag}
 \end{table*}
 
-\subsection{Classifying multiple topic tagging}
+\XXX{What are some words that these classifiers liked?}
+
+\XXX{Summarize the numeric results, is it good/bad}
+
+The limitation of the approach we took here is that we only checked independently for each label.
+Labels could be correlated with each other, as well as we did not evaluate how well the learners performed together.
+
+\subsection{Applying multiple labels to topics}
+
+We applied the Mulan library for Multi-Label learning to our dataset because we feel that topics can have more like one label, much like how a revision can have more than one topic.
+Multi-label learning is more than just classifying entities with more than one label.
+It also includes methods for determining the performance of such techniques.
+Consider that the problem has changed, instead of looking at the precision and recall of applying one label, we have to rank multiple labels at once.
+We have to check if the full subset of labels was applied and then how much of that subset was applied.
+
+Another aspect of multi-label learning is the micro versus macro measurements.
+Macro measurements are aggregated at a class or label level.
+Micro measurements are aggregated at a decision level.
+So a macro-ROC measurment is the average ROC over the ROC values for all labels, where a micro-ROC is the average ROC over all examples that were classified.
+Unfortunately for MaxDB the macro-ROC values are undefined because of poor performance of one of the labels.
+
+Binary Relevance, CLR and HOMER did \XXX{finish this}
 
@@
 
+\subsection{Visualization}
+
+\XXX{We need to have some of the visualizations here}
+
 \subsection{Comparing topic use between projects}
@@
 \XXX{I don't know!}
+
+\subsection{Results Discussion}
+
+\XXX{Discuss how these are CHEAP techniques and although somewhat inaccurate they don't require a tonne of work}
 
@@
 
+\XXX{Boost/COmbine wordbag w/ bayes}
+
 \section{Acknowledgements and appendix}

### 2010-03-04 icpc/icpc2010.tex
Huge figures!

@@
 
-Binary Relevance, CLR and HOMER did \XXX{finish this}
+\XXX{Maybe cite these} Binary Relevance (BR), Calibrated Label Ranking (CLR) and Hierarchy Of Multilabel classifiERs (HOMER), performed the best.
+HOMER and BR acts as a hierarchy of learners, BR is flat while HOMER tries to build a deeper hierarchy in order to build a more accurate learner.
+We feels these classifiers performed because because they have the best micro and macro ROC scores
 
+\XXX{finish this}
+\XXX{Clean up tables with labels}
 \begin{table*}
@@
 
-label-macro-AUC & 0.74 & 0.66 & 0.64 & 0.60 & 0.56 & 0.58 & 0.59 & 0.67 \\
+label-macro-ROC & 0.74 & 0.66 & 0.64 & 0.60 & 0.56 & 0.58 & 0.59 & 0.67 \\
 
@@
 
-label-micro-AUC & 0.81 & 0.81 & 0.77 & 0.77 & 0.66 & 0.80 & 0.80 & 0.79 \\
+label-micro-ROC & 0.81 & 0.81 & 0.77 & 0.77 & 0.66 & 0.80 & 0.80 & 0.79 \\
 
@@
 \XXX{We need to have some of the visualizations here}
+
+\begin{figure*}
+  \centering
+  \includegraphics[height=.9\textheight]{small-maxdb-time-smear-plot-Equal}
+  \caption{MaxDB: Labelled topics over time}
+  \label{fig:maxdb}
+\end{figure*}
+
+\begin{figure*}
+  \centering
+  \includegraphics[width=.99\textwidth]{small-mysql-time-smear-plot-Equal}
+  \caption{MySQL: Labelled topics over time}
+  \label{fig:mysql}
+\end{figure*}
 

### 2010-03-05 icpc/icpc2010.tex
fix neil's supervisor and some random latex error at l. 192
visualization description

@@
    \email{holt@uwaterloo.ca}
-\alignauthor Neil's Supervisor\\
+\alignauthor John Mylopoulos \\
    \affaddr{Department of Computer Science}\\
@@
    \affaddr{Toronto, Ontario, CANADA}\\
-   \email{@cs.toronto.edu}}
+   \email{jm@cs.toronto.edu}}
 
@@
 
-\XXX{We need to have some of the visualizations here}
+We have created two visualizations of the manually labelled data.
+A problem we faced while visualizing was how to display tag overlaps.
+Our solution, while not optimal, was to assign a separate colour to each distinct set of annotations.
+Figures \ref{fig:maxdb} and \ref{fig:mysql} show extracted topics grouped by annotations.
+Annotations or labels that occur in adjacent windows are joined together as trends.
+For MaxDB (Figure \ref{fig:maxdb}) the largest trends were related to maintainability and portability.
+MaxDB supports numerous platforms and portability was a constant issue facing the project's develop.
+MySQL was different, maybe topics overlapped, so there were many different subsets.
+Functionality trends were prevalent throughout its history (the largest reddish streak across the top).
+Two combination  tags of ``functionality portability'' (orange) and ``maintainability portability'' (teal) streaked across most of the history MySQL.
+Since this was MySQL 3.23 and was a stable branch we expect that issues dealings with portability and maintenance would be the primary concerns of this branch: developers probably wanted it to work with current systems and thus had to update it.
 
@@
   \includegraphics[height=.9\textheight]{small-maxdb-time-smear-plot-Equal}
-  \caption{MaxDB: Labelled topics over time}
+  \caption{MaxDB 7.500: Labelled topics over time, continuous topics are trends that occur across time}
   \label{fig:maxdb}
@@
   \includegraphics[width=.99\textwidth]{small-mysql-time-smear-plot-Equal}
-  \caption{MySQL: Labelled topics over time}
+  \caption{MySQL 3.23: Labelled topics over time, continuous topics are trends that occur across time}
   \label{fig:mysql}

### 2010-03-05 icpc/icpc2010.tex
results

@@
 \begin{table*}
+\centering
 \begin{tabular}{ccc|rrrr}
-Word & Project & Learner & ROC & ZeroR Acc. & ZeroR. Diff & Fmeasure\\
+Label & Project & Learner & ROC & ZeroR Acc. & ZeroR. Diff & Fmeasure\\
 \hline
@@
 usability &  mysql &  NaiveBayes &  0.75 &  94.54 &  -5.80 &  0.90  \\ 
+%total & mysql & NaibeBayes & 0.75 & 
  \hline 
@@
 \begin{table*}
+\centering
 \begin{tabular}{l|rrrrrrrr}
@@
 % src/validate/output/maxdb.mulan.out.tex
+\centering
 \begin{tabular}{l|rrrrrrrr}
@@
 
-\XXX{Compare the results of Exp2 and Exp3}
+%\XXX{Compare the results of Exp2 and Exp3}
 %    versus the mulan results and weka results
+
+Very rarely did exp2 and exp3 ever perform as well as the machine learners.
+For MaxDB reliability was slightly better detected using the static word list of \textsf{exp2}.
+In general, the machine learners and \textsf{exp3} did better than \textsf{exp2} for both MaxDB and MySQL.
+In both MySQL and MaxDB usability was better served by \textsf{exp2}.
+
+We found that the multilabel learners of BR, CLR and HOMER did not do as well for Macro-ROC and Micro-F1 as NaiveBayes and other NaiveBayes derived learners did.
+This suggests that by sticking together multiple NaiveBayes leaners we could probably label topics effectively, but it'd require a separate NaiveBayes learner per label.
+
+\begin{comment}
+MySQL Macro ROC 0.746 Micro F1 0.78 Mulan BR ROC 0.74 / 0.81 F1 0.46/0.59/0.56 
+MaxDB Macro ROC 0.7   Micro F1 0.83 Mulan BR ROC NaN  / 0.76 F1 0.32/0.39/0.55
+
+\end{comment}
+
+\begin{comment}
+           exp2 exp3 learn  B 2nd
+maxdb port 0.55 0.66 0.74   L 3
+maxdb eff  0.55 0.64 0.67   L 3
+maxdb rel  0.76 0.65 0.73   2 2
+maxdb fun  0.52 0.66 0.81   L 3
+maxdb main 0.52 0.56 0.78   L 3
+maxdb use  0.62 0.59 0.75   L 2
+
+mysql port 0.53 0.57 0.85   L 3
+mysql eff  0.62 0.60 0.62   2 2 
+mysql rel  0.66 0.66 0.84   L 3
+mysql fun  0.5  0.62 0.67   L 3
+mysql main 0.5  0.5  0.70   L 3
+mysql use  0.56 0.51 0.56   L 2
+\end{comment}
 
@@
 
-\XXX{do we have time here}
+\XXX{Not sure there's enough here to do that!}
+
+We observed that MySQL had more topic subsets than MaxDB.
 
@@
 
-\XXX{Discuss how these are CHEAP techniques and although somewhat inaccurate they don't require a tonne of work}
+%\XXX{Discuss how these are CHEAP techniques and although somewhat %inaccurate they don't require a tonne of work}
+
+With ROC values ranging from $0.6$ to $0.8$ we can see there is promise in these attempts.
+\textsf{exp2} and \textsf{exp3} both indicate that static information can be used to help label topics without any training whatsoever.
+If the techniques used in \textsf{exp2} and \textsf{exp3} were combined with the supervised techniques we could probably reduce the training effort.
+Both Naive Bayesian learners and the wordlist approaches were computationally efficient.
+Thus these results are promising because the results are accurate enough to be useful and cheap enough to execute to be feasible as an automated or semi-automated method of labelling topics by their software qualities.
 

### 2010-03-05 icpc/icpc2010.tex
todo and ric and migod edits

@@
 
-\title{What's in a name? On the (automated) topic naming of software maintenance activities}
+\title{What's in a name? On the automated topic naming of software maintenance activities}
 
@@
 
+%There has not been any   investigation into the latent topics these %classifiers represent.
+
 \begin{abstract}
-  In the automated mining of software repositories, most approaches, such as topic modeling and concept location, focus on applying standard machine learning algorithms to corpora.
-This approach does not make any use of domain-specific information to better understand results.
-While too much specificity can produce non-generalizable results, too little produces broad learners that don't provide much useful detail.%There has not been any   investigation into the latent topics these classifiers represent.
-In this paper, we investigate what topics the latent topics classifier produce, and examine  whether such topics can be
-  generalized across software products. We focus on nonfunctional requirements related to software quality
-  as a potential generalization, since there is some shared belief
-  that these qualities apply broadly across software products. We
-  wanted to name topics based on a static ontology of non-functional
-  requirements and software qualities. %Our overarching goal is to use   individual topic models to create a list of terms that can be shared   across software projects. 
-In essence, this would be adding some form
-  of context to general-purpose ontologies such as Wordnet. In this
-  paper, we discuss how to use domain knowledge about software
-  development in order to better annotate topics of development that
-  are extracted from the change log messages in configuration
-  management systems such as CVS. Since our taxonomy is global, our results allow us to compare the relative importance of particular qualities between projects.
+
+  Within the field of mining software repositories, many approaches,   such as topic modeling and concept location, rely on the automated   application of machine learning algorithms to corpora.
+This kind of   approach typically makes no use of domain-specific knowledge to   improve the results or put them into a context that a developer can   relate to.
+While too much specificity can produce non-generalizable   results, too little produces broad learners that do not provide much   immediately useful detail.
+In this paper, we investigate what   topics the latent topics extractors produces, and examine whether   such topics can be generalized across software products.
+We focus   on non-functional requirements related to software quality as a   potential generalization, since there is some shared belief that   these qualities apply broadly across many software systems and their   development artifacts.
+In our approach, we name topics based on a   static ontology of non-functional requirements and software   qualities; in essence, this adds context to general-purpose   ontologies such as Wordnet.
+In this paper, we discuss how we use   this software development domain knowledge to better annotate topics   of development that are extracted from the change log messages in   configuration management systems such as CVS.
+Since our taxonomy is   global in scope, our results allow us to compare the relative   importance of particular qualities across different projects.
+With   in the topics extracted from MySQL and MaxDB we found that   non-functional requirements were common and often were trends across   time.
+Our automated techniques produce acceptable results without a   lot of loss in computational performance.
+
 \end{abstract}
+%\XXX{abstract too abstract}
 
@@
 
-During the development of a software project developers touch on many issues integral to the purpose and specification of the software project.
+During the development of a software project developers touch on many issues integral to the purpose of the software project.
 These issues are often referred to as topics.
 These topics can sometimes be automatically extracted from the changelog comments that developers create while committing revisions to the project's source control repository.
-Our previous work dealt with topic-trends, topics that repeat overtime.
-It seemed that these trends were often nonfunctional requirements.
+Our previous work dealt with topic-trends~\cite{Hindle09ICSM}, topics that repeat overtime.
+It seemed that these trends were often non-functional requirements.
 
 Topics in this paper and the previous paper are word bags or word distributions.
-These word distributions are found via Latent Dirichlet Allocation~\cite{Blei2003}, which find independent word distributions shared among documents (changelog comments).
+These word distributions are found via Latent Dirichlet Allocation~\cite{Blei2003}, which finds independent word distributions shared among documents (changelog comments).
 The unfortunate aspect of topics that are word distributions is that they lack the tangibility of a rose.
-They do not self idea themselves due to their tangible properties.
-Topics have to be identified by interpretting the prevelant words in the word distribution and by inspecting related documents.
-This is tiresome when one has to handle more than one hundred different topics.
-It would be nice if we could, at a glance at , have some idea what the topic is about.
+They do not self identify themselves with their tangible properties.
+Topics have to be identified by interpreting the prevalent words in the word distribution and by inspecting related documents.
+This is impractical when one has to handle more than one hundred different topics.
+It would be nice if we could have automatic assistance to determine what the topic is about.
 This is the power of labelling and naming.
 
-If a topic has a name or a label we immediately understand and register what it is.
+If a topic has a name or a label we immediately have a good idea of  what it is.
 In our case the name is the bit of information, the summary, that was missing from our topics to make them useful and complete.
@@
 % Clean this up
-Traditionally topic extraction has required manual annotation to derive domain-relevant information.
-This paper attempts to implement \emph{labelled topic extraction}, topic extracted are given labels relating to a cross-project taxonomy.
+Traditionally topic extraction has required manual annotation to derive domain-relevant labelling information.
+This paper attempts to implement \emph{labelled topic extraction}, in which topic are extracted and given labels relating to a cross-project taxonomy.
 This is an ontological approach where we attempt to relate networks of words to labels and then search for these terms in our topic.
@@
 
-\XXX{Not sure about this}
+%\XXX{Not sure about this}
 \begin{comment}
@@
 % unsure about this
-\item We show that these topics can be learned and used to classify other datasets.
-\item We present a number of visualizations of named topics and their trends over time to aid
-  communication and analysis.
+\item We show that these labels with their topics can be learned and used to classify other datasets.
+\item We present visualizations of named topics and their trends over time to aid communication and analysis.
 \item We use an exploratory case study of
@@
 One slice is manually annotated by the domain expert
-In this paper, we employ the WEKA and Mulan machine learning frameworks in order to test supervised learning.
+In this paper, we employ the WEKA~\cite{weka09} and Mulan machine learning frameworks in order to test supervised learning.
 
@@
 
-\XXX{how many do we have}
+%\XXX{how many do we have}
 We wanted to compare two mature software projects, that had open repositories, within the same application domain.
@@
 %There are approximately ... developers working on it. 
-\XXX{version number?}
 The MySQL project started in 1994 but MySQL 3.23 was released in early 2001. MySQL contains 320 thousand lines of C and C++ source code.
@@
 %This is confusing/confused
-\XXX{redo this}
+\XXX{exp2}
 To construct our expanded wordlists, we used Wordnet~\cite{Fellbaum1998}, an English-language ``lexical database'' that contains semantic relations between words, including meronymy and synonymy.
@@
 
-In general, this approach did not work out well.
+In general, this approach did not work out well as common labels dominated the less common labels.
 The related words for \emph{correctness}, for example, tended to be too lengthy and non-specific.
@@
 There are \{periods X 20\} topics per project.
-All experiments were run on MaxDB data.
+All experiments were run on MaxDB and MySQL data.
 
@@
 
-\XXX{Describe table}
-
-\XXX{Describe learners}
-
 We took our annotated dataset and applied supervised learner to it.
 Because our dataset was of word counts we expected Bayesian techniques, often used in spam filtering, to perform well.
+We also tried other learners that WEKA~\cite{weka09} provides: rule learners, tree learners, vector space learners, and support vector machines.
 Table \ref{tab:best-learn-per-tag} shows the performance of the best performing learner per label.
@@
 HOMER and BR acts as a hierarchy of learners, BR is flat while HOMER tries to build a deeper hierarchy in order to build a more accurate learner.
-We feels these classifiers performed because because they have the best micro and macro ROC scores
+We feels these classifiers performed better because they have the best micro and macro ROC scores than the other classifiers.
 
@@
 \end{table*}
-\XXX{put mulan stuff here}
 
 \subsection{Summary of techniques}
-
-%\XXX{Compare the results of Exp2 and Exp3}
-%    versus the mulan results and weka results
 
@@
 
-\subsection{Comparing topic use between projects}
+\subsection{Comparing MaxDB and MySQL}
 
@@
 
-\subsection{Topic change over time}
+%\subsection{Topic change over time}
 
-\XXX{I don't know!}
+%\XXX{I don't know!}
 
-\subsection{Results Discussion}
+\subsection{Effectiveness}
 
@@
 
-This paper has not considered the interpretation of attitudes towards qualities (e.g., do developers feel this helps security).
-
+We want to investigate developer attitudes related to these labels, when we label a topic did was the developer expressing positive or negative qualities about that label.
 It is difficult to map abstract qualities to particular messages.
@@
 \section{Acknowledgements and appendix}
-Our data and scripts are available at \url{http://softwareprocess.es/icpc2010}
+Our data and scripts are available at \url{http://softwareprocess.es/nomen/}
 

### 2010-03-05 icpc/icpc2010.tex
bibliography stuff
This is expiremental

@@
 \XXX{this table should have mysql}
-\begin{table*}[h]
+\begin{table}
 	\caption{Automatic topic labeling}
@@
 \end{tabular}
-\end{table*}
+\end{table}
 
@@
 
-\XXX{WRITE THIS}
+* Demonstrated that domain specific knowledge is valuable
+* Demonstrated the feasibility of static approaches and supervised approaches
+* Showed that non functional requirement pervade much many of the topics that are extracted
+
+In conclusion we have demonstrated that static but domain specific knowledge can improve unsupervised labelling of extracted topics.
+Our \textsf{exp2} experiment had fewer terms but performed mostly just as well as \textsf{exp3} which used more general terms.
 
@@
 
-\bibliographystyle{IEEEtran}
+\bibliographystyle{abbrv}
+%\bibliographystyle{IEEEtran}
 \bibliography{icpc}

### 2010-03-05 src/validate/topics2arff.py

@@ -76,6 +76,47 @@ def to_arff( name, xperiods ): #name, inputdir, periods ):
         data.append(",".join(sgen) + "\n")
     return header + data
 
+def to_arff_data( xperiod, nwords, nannos, l, annotoindex, wordtoindex ):
+    topics = A.flatten([ A.get_topics(x) for x in xperiods ])
+    data = []
+    for t in topics:
+        tannotations, twords = V.create_element_list( t )
+        l = [ 0 for x in range(nwords) ] + [ "0" for x in range(nannos) ]
+        for tanno in tannotations:
+            l[annotoindex[tanno]] = "1"
+        for tword in twords:
+            l[wordtoindex[tword]] += 1
+        sgen = ( str(x) for x in l )
+        data.append(",".join(sgen) + "\n")
+    return data
+
+def to_arff2( name, xperiod1, xperiod2 ): #name, inputdir, periods ):
+    xperiods1 = []
+    xperiods1.append(xperiod1)
+    xperiods1.append(xperiod2)
+    xperiods = A.flatten(xperiods1)
+    topics = A.flatten([ A.get_topics(x) for x in xperiods ])
+    # summarize the words and order by most common
+    word_counts = word_count_of_topics( topics )
+    indextoword, wordtoindex = A.map_words_to_index( word_counts )
+    nwords = len( word_counts.keys() )
+    words = [indextoword[i] for i in range(nwords)]
+    # summarize the annotations and order by most common
+    anno_hist = A.summarize_annotations_of_topics( topics )
+    indextoanno, annotoindex = A.map_words_to_index( anno_hist, cnt = nwords )
+    nannos = len( anno_hist.keys() )
+    # note the + nwords part
+    annos = [indextoanno[i + nwords] for i in range(nannos)]
+    # total # of columns
+    ncolumns = nannos + nwords
+    allcolumns = words + annos
+    # header is a list of strings
+    # too much arff right here.. oh well
+    header = arff_header( name , ["W_"+arff_escape(w)+" INTEGER" for w in words] + ["A_"+arff_escape(a)+" {0,1}" for a in annos] )
+    data1 = to_arff_data( xperiod1, nwords, nannos, l, annotoindex, wordtoindex )
+    data2 = to_arff_data( xperiod2, nwords, nannos, l, annotoindex, wordtoindex )
+    return ((header + data1), (header+data2))
+
 def save_arff( filename, lines):
[…]

### 2010-03-05 src/validate/topics2arff.py
some conclusions and failure to get what I wanted :(

@@ -76,46 +76,44 @@ def to_arff( name, xperiods ): #name, inputdir, periods ):
         data.append(",".join(sgen) + "\n")
     return header + data
 
-def to_arff_data( xperiod, nwords, nannos, l, annotoindex, wordtoindex ):
-    topics = A.flatten([ A.get_topics(x) for x in xperiods ])
-    data = []
-    for t in topics:
-        tannotations, twords = V.create_element_list( t )
-        l = [ 0 for x in range(nwords) ] + [ "0" for x in range(nannos) ]
-        for tanno in tannotations:
-            l[annotoindex[tanno]] = "1"
-        for tword in twords:
-            l[wordtoindex[tword]] += 1
-        sgen = ( str(x) for x in l )
-        data.append(",".join(sgen) + "\n")
-    return data
-
-def to_arff2( name, xperiod1, xperiod2 ): #name, inputdir, periods ):
-    xperiods1 = []
-    xperiods1.append(xperiod1)
-    xperiods1.append(xperiod2)
-    xperiods = A.flatten(xperiods1)
-    topics = A.flatten([ A.get_topics(x) for x in xperiods ])
-    # summarize the words and order by most common
-    word_counts = word_count_of_topics( topics )
-    indextoword, wordtoindex = A.map_words_to_index( word_counts )
-    nwords = len( word_counts.keys() )
-    words = [indextoword[i] for i in range(nwords)]
-    # summarize the annotations and order by most common
-    anno_hist = A.summarize_annotations_of_topics( topics )
-    indextoanno, annotoindex = A.map_words_to_index( anno_hist, cnt = nwords )
-    nannos = len( anno_hist.keys() )
-    # note the + nwords part
-    annos = [indextoanno[i + nwords] for i in range(nannos)]
-    # total # of columns
-    ncolumns = nannos + nwords
-    allcolumns = words + annos
-    # header is a list of strings
-    # too much arff right here.. oh well
-    header = arff_header( name , ["W_"+arff_escape(w)+" INTEGER" for w in words] + ["A_"+arff_escape(a)+" {0,1}" for a in annos] )
-    data1 = to_arff_data( xperiod1, nwords, nannos, l, annotoindex, wordtoindex )
-    data2 = to_arff_data( xperiod2, nwords, nannos, l, annotoindex, wordtoindex )
-    return ((header + data1), (header+data2))
[…]

### 2010-03-05 icpc/icpc2010.tex
some conclusions and failure to get what I wanted :(
more conclusions

@@
 
-* Demonstrated that domain specific knowledge is valuable
-* Demonstrated the feasibility of static approaches and supervised approaches
-* Showed that non functional requirement pervade much many of the topics that are extracted
+%* Demonstrated that domain specific knowledge is valuable
+%* Demonstrated the feasibility of static approaches and supervised %approaches
+%* Showed that non functional requirement pervade much many of the %topics that are extracted
 
@@
 Our \textsf{exp2} experiment had fewer terms but performed mostly just as well as \textsf{exp3} which used more general terms.
+We then showed that with some supervision, and by using efficient machine learners like Naive Bayesian classifiers, we could improve the accuracy of automatic labelling event further.
 
+Our manual inspection and annotation of the topics extracted from MySQL and MaxDB revealed that many of the extracted dealt with non-functional requirements and the topics related to non-functional requirements often occurred across the entire history of a project.
+In the case of MaxDB and MySQL portability was a constant maintenance concern was prevalent through the entire lifetime of the projects.
+
+Thus we have shown that non-functional requirements are often trending topics, that non-functional requirements are quite common in developer topics and that there are efficient methods of semi-automating and automating topic labelling.
+
+There are many avenues of further investigation.
 We want to investigate developer attitudes related to these labels, when we label a topic did was the developer expressing positive or negative qualities about that label.
@@
 
-\XXX{Boost/COmbine wordbag w/ bayes}
+% \XXX{Boost/COmbine wordbag w/ bayes}
 

### 2010-03-05 icpc/icpc2010.tex
mulan citation

@@
 One slice is manually annotated by the domain expert
-In this paper, we employ the WEKA~\cite{weka09} and Mulan machine learning frameworks in order to test supervised learning.
+In this paper, we employ the WEKA~\cite{weka09} and Mulan~\cite{mulan} machine learning frameworks in order to test supervised learning.
 
@@
 WEKA contains a suite of machine learning tools such as support vector machines and Bayes nets.
-We also used the multi-labeling add-on for WEKA, Mulan\footnote{\url{http://mlkd.csd.auth.gr/multilabel.html}}.
+We also used the multi-labeling add-on for WEKA, Mulan~\cite{mulan}\footnote{\url{http://mlkd.csd.auth.gr/multilabel.html}}.
 Traditional classifiers map our topics to a single class, whereas Mulan allows for a mixture of classes per topic, much like our manual classification found overlaps.
@@
 
-We applied the Mulan library for Multi-Label learning to our dataset because we feel that topics can have more like one label, much like how a revision can have more than one topic.
+We applied the Mulan~\cite{mulan} library for Multi-Label learning to our dataset because we feel that topics can have more like one label, much like how a revision can have more than one topic.
 Multi-label learning is more than just classifying entities with more than one label.
@@
 
-\XXX{Maybe cite these} Binary Relevance (BR), Calibrated Label Ranking (CLR) and Hierarchy Of Multilabel classifiERs (HOMER), performed the best.
-HOMER and BR acts as a hierarchy of learners, BR is flat while HOMER tries to build a deeper hierarchy in order to build a more accurate learner.
+ Binary Relevance (BR), Calibrated Label Ranking (CLR) and Hierarchy Of Multilabel classifiERs (HOMER), performed the best.
+HOMER and BR acts as a hierarchy of learners, BR is flat while HOMER tries to build a deeper hierarchy in order to build a more accurate learner~\cite{mulan}.
 We feels these classifiers performed better because they have the best micro and macro ROC scores than the other classifiers.

### 2010-03-05 icpc/icpc2010.tex
updates and edits to conclusions

@@
 In conclusion we have demonstrated that static but domain specific knowledge can improve unsupervised labelling of extracted topics.
-Our \textsf{exp2} experiment had fewer terms but performed mostly just as well as \textsf{exp3} which used more general terms.
-We then showed that with some supervision, and by using efficient machine learners like Naive Bayesian classifiers, we could improve the accuracy of automatic labelling event further.
+Our \textsf{exp2} experiment used small accurate word bags to label topics but performed just as well as \textsf{exp3}, which used many more general terms from WordNet.
+We then showed that with some supervision, and by using efficient machine learners based on Naive Bayesian classifiers, we could improve the accuracy of automatic labelling topics even further.
 
-Our manual inspection and annotation of the topics extracted from MySQL and MaxDB revealed that many of the extracted dealt with non-functional requirements and the topics related to non-functional requirements often occurred across the entire history of a project.
-In the case of MaxDB and MySQL portability was a constant maintenance concern was prevalent through the entire lifetime of the projects.
+Our manual inspection and annotation of the topics extracted from MySQL and MaxDB revealed that many of the extracted dealt with non-functional requirements and these topics were spread across the entire history of a project.
+In the cases of MaxDB and MySQL, portability was a constant maintenance concern and was prevalent throughout the entire lifetime of the projects.
 
-Thus we have shown that non-functional requirements are often trending topics, that non-functional requirements are quite common in developer topics and that there are efficient methods of semi-automating and automating topic labelling.
+Thus we have shown that non-functional requirements are often trending topics, that non-functional requirements are quite common in developer topics, and that there are efficient methods of semi-automating and automating topic labelling.
 
 There are many avenues of further investigation.
-We want to investigate developer attitudes related to these labels, when we label a topic did was the developer expressing positive or negative qualities about that label.
+We want to investigate developer attitudes related to these labels, when we label a topic, was the developer expressing positive or negative qualities about that label.
 It is difficult to map abstract qualities to particular messages.
-If the message is about a particular bug fix, do we map that to reliability?
-Maintainability?
+If the message is about a particular bug fix, do we map that to reliability or maintainability?
 Is it a performance bug or a usability bug?
-Is a `quality' discussion about more than just corrective maintenance, in other words?
+We ask is a ``quality'' discussion about more than just corrective maintenance?
 In this study we take the position that any type of maintenance has to do with software quality.

### 2010-03-05 icpc/icpc2010.tex
some of my edits and migod's first para rewrite

@@
 While too much specificity can produce non-generalizable   results, too little produces broad learners that do not provide much   immediately useful detail.
-In this paper, we investigate what   topics the latent topics extractors produces, and examine whether   such topics can be generalized across software products.
+In this paper, we investigate what   topics the latent topic extractors produce, and examine whether   such topics can be generalized across software products.
 We focus   on non-functional requirements related to software quality as a   potential generalization, since there is some shared belief that   these qualities apply broadly across many software systems and their   development artifacts.
@@
 With   in the topics extracted from MySQL and MaxDB we found that   non-functional requirements were common and often were trends across   time.
-Our automated techniques produce acceptable results without a   lot of loss in computational performance.
+Our automated techniques can label topics without much computational overhead.
 
@@
 \end{quote}
+
+Few would argue that software development is a bed of roses.
+To most of us, the important properties of a rose are sensual in nature, and concern appearance, odour, and feel.
+While horticulturalists and botanists may see more semantic depth than this, one rose is often just as pleasing to us as another.
+Software development artifacts, on the other hand, are abstract and intangible.
+They are often unnamed and ephemeral --- even more so than roses --- and maybe partially derived from other unnamed and ephemeral artifacts.
+Yet to have meaningful discussions about how software development is progressing, we must be able to devise appropriate labels for our work topics and be able to categorize the development artifacts as belonging to one or more of these topics.
+
+%Original is here:
+
+%Unlike the poetic license Shakespeare uses, we show that names are
+%indeed important. The rose has its own tangible properties which
+%allow us to sense it. Unfortunately when we are dealing with data
+%abstracted from program source code, we are already dealing with
+%abstract entities that are not tangible.
 
@@
 However, these terms may not capture all words associated with the labels.
-For example, the term ‘user friendly’ is one most would agree is relevant to discussion of usability, but is not in the standard.
+For example, the term ``user friendly'' is one most would agree is relevant to discussion of usability, but is not in the standard.
 Therefore we took the words from the taxonomy and expanded them.
@@
 
-%This is confusing/confused
-\XXX{exp2}
 To construct our expanded wordlists, we used Wordnet~\cite{Fellbaum1998}, an English-language ``lexical database'' that contains semantic relations between words, including meronymy and synonymy.
@@
 For each high-level quality in ISO9126, we first searched for our existing labels; we then randomly sampled twenty-five mail messages that were relevant to that quality, and selected co-occurring terms relevant to that quality.
-For example, we add the term ‘performance’ to the synonyms for \emph{Efficiency}, since this term occurs in most KDE mail messages that discuss efficiency.
+For example, we add the term ``performance'' to the synonyms for \emph{efficiency}, since this term occurs in most KDE mail messages that discuss efficiency.
 %exp2
@@
 
-\XXX{Not sure there's enough here to do that!}
+%\XXX{Not sure there's enough here to do that!}
 
 We observed that MySQL had more topic subsets than MaxDB.
+
+* we had more MySQL recorded
+* both shared functionality maintainability and portability
 

### 2010-03-05 icpc/icpc2010.tex
Add MSR reference

@@
 
-Unlike the poetic license Shakespeare uses, we show that names are indeed important.
-The rose has its own tangible properties which allow us to sense it.
-Unfortunately when we are dealing with data abstracted from program source code, we are already dealing with abstract entities that are not tangible.
+As development of a software project progresses, developers must concern themselves with a wide variety of issues --- or ``topics'' --- that relate to the problem domain, the processes and tools used, the development artifacts themselves, etc.
+The set of development topics for a given project can sometimes be extracted automatically by analyzing artifacts within software repositories, such as changelog comments that developers create when committing revisions to the project's source control system.
+Our previous work dealt with topic trends, which are topics that recur over time~\cite{Hindle09ICSM}.
+We observed that topic trends were often non-functional requirements.
 
-During the development of a software project developers touch on many issues integral to the purpose of the software project.
-These issues are often referred to as topics.
-These topics can sometimes be automatically extracted from the changelog comments that developers create while committing revisions to the project's source control repository.
-Our previous work dealt with topic-trends~\cite{Hindle09ICSM}, topics that repeat overtime.
-It seemed that these trends were often non-functional requirements.
+%During the development of a software project developers touch on many %issues integral to the purpose of the software project.
+These issues %are often referred to as topics.
+These topics can sometimes be %automatically extracted from the changelog comments that developers %create while committing revisions to the project's source control %repository.
+Our previous work dealt with %topic-trends~\cite{Hindle09ICSM}, topics that repeat overtime.
+It %seemed that these trends were often non-functional requirements.
 
@@
 HOMER and BR acts as a hierarchy of learners, BR is flat while HOMER tries to build a deeper hierarchy in order to build a more accurate learner~\cite{mulan}.
-We feels these classifiers performed better because they have the best micro and macro ROC scores than the other classifiers.
+We feel that these classifiers performed better because they have the best micro and macro ROC scores than the other classifiers.
 
-\XXX{finish this}
-\XXX{Clean up tables with labels}
 \begin{table*}
@@
 We observed that MySQL had more topic subsets than MaxDB.
-
-* we had more MySQL recorded
-* both shared functionality maintainability and portability
+MySQL 3.23 also had more topics and a longer recorded history than MaxDB 7.500.
+We tagged more MySQL topics with annotations than MaxDB topics yet both shared similarities.
+In terms of non-functional requirements both projects had long running trends that focused on functionality, maintainability, and portability, yet MaxDB had more of a focus on efficiency and reliability.
+We feel that MaxDB differed from MySQL because MaxDB was being actively developed where was by half way through our history of MySQL 3.23, other versions of MySQL were being actively developed: MySQL versions 4.0 and 4.1 and eventually 5.0 and 5.1.
+MySQL 3.23 was being maintained rather than actively developed where as MaxDB 7.500 was being actively developed into MaxDB 7.6, and then maintained thereafter.
 
@@
 
+% Not sure how do handle this
 The idea of extracting higher-level `concerns' (also known as `concepts', `aspects' or `requirements') has been attacked in two ways.

### 2010-03-05 src/validate/grepwords.pl
Generate tag clouds

@@ -1,0 +1,5 @@
+while(<>) {
+	chomp;
+	my @matches = ($_ =~ /W_([^ ]+)/);
+	print $_,$/ foreach @matches;
+}

### 2010-03-05 src/validate/mkclouds.sh

@@ -1,0 +1,9 @@
+for project in `cat projects`
+do
+	cat output/$project.*J48.txt | perl grepwords.pl | perl wordcount.pl     > output/$project-cloud.html
+	cat output/$project.*J48.txt | perl grepwords.pl | perl wordcounttex.pl  > output/$project-cloud.tex
+        cd output
+        pdflatex $project-cloud.tex
+        pdfcrop  $project-cloud.pdf
+        cd ..
+done

### 2010-03-05 src/validate/wordcounttex.pl

@@ -1,0 +1,44 @@
+use strict;
+use List::Util qw(shuffle);
+my %hash;
+my $slash = '\\';
+my @sizes = qw(
+	tiny
+	scriptsize
+	footnotesize
+	small
+	normalsize
+	large
+	Large
+	LARGE
+	huge
+	Huge
+);
+my $nsizes = @sizes;
+while(<>) {
+    chomp;
+    $hash{$_}++;
+}
+
+my @keys = keys %hash;
+my @keys = sort { $hash{$a} <=> $hash{$b} } @keys;
+my $max = $hash{$keys[$#keys]};
+@keys = reverse @keys;
+print <<EOF;
+${slash}documentclass[times]{article}
+${slash}usepackage{fullpage}
+${slash}pagestyle{empty}
+${slash}begin{document}
+EOF
+print "${slash}noindent ";
+foreach (@keys) {
+    print "{ ${slash}";
+    print $sizes[int(($nsizes - 1) * ($hash{$_} / $max))];
+    my $a = $_;
+    $a =~ s/_/\\_/g;
+    $a =~ s/$//g;
+    print " $a } $/";
+}
+print <<EOF;
+${slash}end{document}
+EOF

### 2010-03-05 icpc/icpc2010.tex
more migod edits

@@
 \end{abstract}
-%\XXX{abstract too abstract}
 
@@
 The unfortunate aspect of topics that are word distributions is that they lack the tangibility of a rose.
-They do not self identify themselves with their tangible properties.
+They do not self identify with their tangible properties.
 Topics have to be identified by interpreting the prevalent words in the word distribution and by inspecting related documents.
@@
 
-If a topic has a name or a label we immediately have a good idea of  what it is.
-In our case the name is the bit of information, the summary, that was missing from our topics to make them useful and complete.
-Although we are convinced that accurate naming is not completely feasible, so we have focussed on labelling topics by non-functional requirements.
+%If a topic has a name or a label we immediately have a good idea of  %what it is.
+In our case the name is the bit of information, the %summary, that was missing from our topics to make them useful and %complete.
+Although we are convinced that accurate naming is not %completely feasible, so we have focussed on labelling topics by %non-functional requirements.
+
+A topic --- that is, a word distribution in our world --- needs a suggestive name to succinctly convey its intent, and make it easy to use in discussions and analyses about the development of the system.
+Our previous experience leads us to believe that automated topic naming is currently infeasible in the general case, so we have decided to focus on topic labels from the sub-domain of non-functional requirements related to software quality.
 
@@
 The algorithm returns a word-document matrix relating the occurrence frequency of a
- The idea is similar to other IR approaches like TF/IDF (term frequency/inverse document frequency), which find uncommon words that aren't found in other wordlists.
+ The idea is similar to other IR approaches like TF/IDF (term frequency/inverse document frequency), which find uncommon words that are not found in other wordlists.
 These approaches are unsupervised in that they rely solely on the source data with no human intervention.
@@
 
-%\XXX{how many do we have}
-We wanted to compare two mature software projects, that had open repositories, within the same application domain.
-This allows us to partially control for differences in functional requirements.
-We selected MySQL and MaxDB, open-source, near-commercial quality database software.
+To evaluate our approach, we sought candidate systems that were
+mature projects and had openly accessible artifact repositories.
+We also decided to select systems from the same application
+domain, as we felt the functional requirements would probably be
+broadly similar.
+%We wanted to compare two mature software projects, that had open %repositories, within the same application domain. 
+% This allows us to partially control for differences in functional %requirements. 
+We selected MySQL and MaxDB as they were open-source, partially-commercial database systems.
 MaxDB started in the late 1970s as a research project, and was later acquired by SAP.
@@
 
-Our second list, \textsf{exp2}, relied on the ISO quality model (ISO9126)~\cite{iso9126}.
-It describes six high-level quality requirements (listed in Table \ref{tbl:wnsig}).
-There is some debate about the significance and importance of the terms in this model.
-However, it is ``an international standard and thus provides an internationally accepted terminology for software quality \cite[p.
-58]{Bøegh2008},'' which is sufficient for the purposes of this research.
-However, these terms may not capture all words associated with the labels.
-For example, the term ``user friendly'' is one most would agree is relevant to discussion of usability, but is not in the standard.
+Our second list, \textsf{exp2}, relied on the ISO quality model (ISO9126)~\cite{iso9126}. ISO9126 describes six high-level quality requirements (listed in Table \ref{tbl:wnsig}). 
+%There is some debate about the significance and importance of the %terms in this model. 
+%However, it is ``an international standard and thus provides an %internationally accepted terminology for software quality \cite[p.
+%58]{Bøegh2008},'' which is sufficient for the purposes of this %research.
+ISO9126 is ``an international standard and thus provides an internationally accepted terminology for software quality \cite[p. 58]{Bøegh2008},'' that is sufficient for the purposes of this research.
+However, the terms extracted from ISO9126 may not capture all words associated with the labels.
+For example, the term ``redundancy'' is one most would agree is relevant to discussion of reliability, but is not in the standard.
 Therefore we took the words from the taxonomy and expanded them.
@@
 We do not include these words here for space considerations (but see the Appendix for our data repository).
-It isn't clear the words associated with our labels are specific enough, however: for example, the label \emph{maintainability} is associated with words \emph{ease} and \emph{ownership}.
+It is not clear the words associated with our labels are specific enough, however: for example, the label \emph{maintainability} is associated with words \emph{ease} and \emph{ownership}.
 
@@
 
-\XXX{this table should have mysql}
+%\XXX{this table should have mysql}
 \begin{table}
@@
 
-It seems like the smaller the label the harder it is to get accurate results. Although these results seem better than our previous word bag results of \textsf{exp2} and \textsf{exp3}.
+It seems like the smaller the label the harder it is to get accurate results.
+Although these results seem better than our previous word bag results of \textsf{exp2} and \textsf{exp3} because the ROC values are sufficiently higher in most cases other than MaxDB reliability and MySQL efficiency.
+
+The limitation of the approach we took here is that we only checked independently for each label.
+Labels could be correlated with each other, as well as we did not evaluate how well the learners performed together.
 
@@
 \end{table*}
-
-\XXX{What are some words that these classifiers liked?}
-
-\XXX{Summarize the numeric results, is it good/bad}
-
-The limitation of the approach we took here is that we only checked independently for each label.
-Labels could be correlated with each other, as well as we did not evaluate how well the learners performed together.
 
@@
 We found that the multilabel learners of BR, CLR and HOMER did not do as well for Macro-ROC and Micro-F1 as NaiveBayes and other NaiveBayes derived learners did.
-This suggests that by sticking together multiple NaiveBayes leaners we could probably label topics effectively, but it'd require a separate NaiveBayes learner per label.
+This suggests that by sticking together multiple NaiveBayes leaners we could probably label topics effectively, but it would require a separate NaiveBayes learner per label.
 
@@
 
-%\subsection{Topic change over time}
+MySQL and MaxDB's machine learners did dispatch off of some shared words.
+Words that were used to classify topics that were shared between MySQL and MaxDB included: bug, code, compiler, database, HP UX, delete, memory, missing, problems, removed, add, added, changed, problem, and test.
+Potentially adding these words to the word bags of \textsf{exp2} and \textsf{exp3} could improve performance while still being domain specific.
 
-%\XXX{I don't know!}
+\begin{comment}
+ shared-report.sh ; cat output/shared-j48-words
+
+      2 bug
+      2 code
+      2 compiler
+      2 database
+      2 Delete_
+      2 HP_UX
+      2 memory
+      2 missing
+      2 monty_mysql_com
+      2 monty_mysql_fi
+      2 problems
+      2 Removed
+      2 work__home_bk_mysql
+      4 4_0
+      4 Add
+      4 added
+      4 Changed
+      4 problem
+      6 Added
+      6 test
+
+\end{comment}
 
@@
 
-%\XXX{Discuss how these are CHEAP techniques and although somewhat %inaccurate they don't require a tonne of work}
+%{Discuss how these are CHEAP techniques and although somewhat %inaccurate they don't require a tonne of work}
 
@@
 
-% \XXX{Boost/COmbine wordbag w/ bayes}
-
 \section{Acknowledgements and appendix}

### 2010-03-05 src/validate/shared-report.sh
shared report just tells us what j48 words are shared

@@ -1,0 +1,3 @@
+cat output/mysql.*J48.txt | perl grepwords.pl | sort > output/mysql-j48-words
+cat output/maxdb.*J48.txt | perl grepwords.pl | sort > output/maxdb-j48-words
+fgrep -if output/maxdb-j48-words output/mysql-j48-words | sort | uniq -c | sort -n > output/shared-j48-words

### 2010-03-06 icpc/icpc2010.tex
reformatted that mulan table
shared report just tells us what j48 words are shared
some table changes

@@
 
-\begin{table*}[h]
+\begin{table*}
 	\caption{Qualities and associated signifiers – Wordnet version (\textsf{exp2})}
@@
 \begin{tabular}{ccc|rrrr}
+\toprule
 Label & Project & Learner & ROC & ZeroR Acc. & ZeroR. Diff & Fmeasure\\
+\midrule
+portability &  MySQL &  NaiveBayesMultinomial &  0.74 &  58.53 &  11.43 &  0.70  \\ 
+efficiency &  MySQL &  NaiveBayes &  0.67 &  93.69 &  -7.68 &  0.88 \\ 
+reliability &  MySQL &  NaiveBayes &  0.73 &  83.11 &  -12.80 &  0.74 \\ 
+functionality &  MySQL &  DMNBtext &  0.81 &  54.44 &  21.67 &  0.76 \\ 
+maintainability &  MySQL &  DMNBtext &  0.78 &  76.62 &  3.41 &  0.75 \\ 
+usability &  MySQL &  NaiveBayes &  0.75 &  94.54 &  -5.80 &  0.90  \\ 
+%total & MySQL & NaibeBayes & 0.75 & 
  \hline 
-portability &  mysql &  NaiveBayesMultinomial &  0.74 &  58.53 &  11.43 &  0.70  \\ 
-efficiency &  mysql &  NaiveBayes &  0.67 &  93.69 &  -7.68 &  0.88 \\ 
-reliability &  mysql &  NaiveBayes &  0.73 &  83.11 &  -12.80 &  0.74 \\ 
-functionality &  mysql &  DMNBtext &  0.81 &  54.44 &  21.67 &  0.76 \\ 
-maintainability &  mysql &  DMNBtext &  0.78 &  76.62 &  3.41 &  0.75 \\ 
-usability &  mysql &  NaiveBayes &  0.75 &  94.54 &  -5.80 &  0.90  \\ 
-%total & mysql & NaibeBayes & 0.75 & 
- \hline 
-portability &  maxdb &  NaiveBayes &  0.84 &  77.12 &  2.06 &  0.80 \\ 
-efficiency &  maxdb &  NaiveBayes &  0.62 &  88.43 &  -11.31 &  0.79 \\ 
-reliability &  maxdb &  DMNBtext &  0.84 &  89.46 &  3.86 &  0.92  \\ 
-functionality &  maxdb &  NaiveBayes &  0.67 &  91.26 &  -6.94 &  0.86  \\ 
-maintainability &  maxdb &  NaiveBayes &  0.70 &  79.43 &  -9.25 &  0.72  \\ 
-usability &  maxdb &  NaiveBayes &  0.56 &  95.37 &  -4.37 &  0.91  \\ 
+portability &  MaxDB &  NaiveBayes &  0.84 &  77.12 &  2.06 &  0.80 \\ 
+efficiency &  MaxDB &  NaiveBayes &  0.62 &  88.43 &  -11.31 &  0.79 \\ 
+reliability &  MaxDB &  DMNBtext &  0.84 &  89.46 &  3.86 &  0.92  \\ 
+functionality &  MaxDB &  NaiveBayes &  0.67 &  91.26 &  -6.94 &  0.86  \\ 
+maintainability &  MaxDB &  NaiveBayes &  0.70 &  79.43 &  -9.25 &  0.72  \\ 
+usability &  MaxDB &  NaiveBayes &  0.56 &  95.37 &  -4.37 &  0.91  \\ 
  \hline 
@@
 We feel that these classifiers performed better because they have the best micro and macro ROC scores than the other classifiers.
+
+\begin{comment}
 
@@
 
+\end{comment}
+
+\begin{table*}
+\centering
+\begin{tabular}{l|ccc|ccc}
+\toprule
+  & MySQL & MySQL & MySQL &                         MaxDB & MaxDB & MaxDB  \\ 
+Performance Metric  & BR & CLR & HOMER &                              BR & CLR & HOMER \\
+
+\midrule
+%\hline \\
+
+example-Accuracy & 0.45 & 0.40 & 0.50 &           0.49  &  0.47  &  0.51 \\
+
+example-Fmeasure & 0.56 & 0.48 & 0.59 &           0.55  &  0.50  &  0.53 \\
+
+example-HammingLoss & 0.24 & 0.19 & 0.18 &        0.19  &  0.14  &  0.14 \\
+
+example-Precision & 0.50 & 0.51 & 0.60 &          0.50  &  0.49  &  0.53 \\
+
+example-Recall & 0.65 & 0.46 & 0.58 &             0.61  &  0.50  &  0.53 \\
+
+example-SubsetAccuracy & 0.19 & 0.24 & 0.31 &     0.40  &  0.43  &  0.45 \\
+
+label-macro-ROC & 0.74 & 0.66 & 0.64 &            NaN  &  NaN  &  NaN    \\
+
+label-macro-F1 & 0.46 & 0.30 & 0.43 &             0.32  &  0.19  &  0.23 \\
+
+label-macro-Precision & 0.41 & 0.41 & 0.49 &      0.29  &  0.31  &  0.32 \\
+
+label-macro-Recall & 0.56 & 0.26 & 0.40 &         0.38  &  0.15  &  0.20 \\
+
+label-micro-ROC & 0.81 & 0.81 & 0.77 &            0.76  &  0.66  &  0.62 \\
+
+label-micro-F1 & 0.59 & 0.53 & 0.61 &             0.39  &  0.27  &  0.34 \\
+
+label-micro-Precision & 0.51 & 0.69 & 0.66 &      0.34  &  0.42  &  0.49 \\
+
+label-micro-Recall & 0.68 & 0.43 & 0.56 &         0.47  &  0.21  &  0.26 \\
+
+rank-AvgPrecision & 0.76 & 0.77 & 0.69 &          0.54  &  0.57  &  0.54 \\
+
+rank-Coverage & 1.51 & 1.47 & 1.94 &              1.40  &  1.23  &  1.43 \\
+
+rank-One-error & 0.40 & 0.39 & 0.47 &             0.81  &  0.79  &  0.81 \\
+
+rank-Ranking Loss & 0.19 & 0.18 & 0.27 &          0.41  &  0.35  &  0.41 \\ 
+\hline
+\end{tabular}
+\caption{MySQL and MaxDB Mulan results per Multi-Label learner}
+\end{table*}
+
 \subsection{Summary of techniques}
@@
 
-\begin{figure*}
+\begin{figure*}[h]
   \centering

### 2010-03-06 icpc/icpc2010.tex
some table related stuff

@@
 \begin{table*}
-	\caption{Qualities and associated signifiers – Wordnet version (\textsf{exp2})}
 	\centering
-	\label{tbl:wnsig}
 \begin{tabular}{c|p{9cm}}
@@
 \end{tabular}
+	\caption{Qualities and associated signifiers – Wordnet version (\textsf{exp2})}
+	\label{tbl:wnsig}
+
 \end{table*}
@@
 \begin{table}
-	\caption{Automatic topic labeling}
 	\centering
-	\label{tbl:wordlist}
 \begin{tabular}{c|c|c|c}
@@
 \end{tabular}
+	\caption{Automatic topic labeling}
+	\label{tbl:wordlist}
+
 \end{table}
@@
 \begin{table*}[h]
-	\caption{Results for automatic topic labeling. f1 = f-measure, mcc = Matthew's correlation coeff., ROC = Area under ROC curve}
 	\centering
-	\label{tbl:maxdb-unsup-results}
+
 \begin{tabular}{c|c|c|c|c|c|c}
 \toprule
-Experiment & label & f1 & mcc & precision & recall & ROC  \\ 
+Experiment & Label & F1 & MCC & Precision & Recall & ROC  \\ 
 \midrule
@@
 \midrule
-mysql exp2 & portability & 0.138 & 0.211 & 1.000 & 0.074 & 0.537 \\ 
+MySQL exp2 & portability & 0.138 & 0.211 & 1.000 & 0.074 & 0.537 \\ 
  & efficiency & 0.345 & 0.327 & 0.476 & 0.270 & 0.625 \\ 
@@
 \midrule
-maxdb exp3 & portability & 0.472 & 0.286 & 0.402 & 0.573 & 0.660 \\ 
+MaxDB exp3 & portability & 0.472 & 0.286 & 0.402 & 0.573 & 0.660 \\ 
  & efficiency & 0.223 & 0.068 & 0.130 & 0.778 & 0.549 \\ 
@@
 \midrule
-mysql exp3 & portability & 0.413 & 0.170 & 0.564 & 0.325 & 0.574 \\ 
+MySQL exp3 & portability & 0.413 & 0.170 & 0.564 & 0.325 & 0.574 \\ 
  & efficiency & 0.158 & 0.105 & 0.089 & 0.703 & 0.608 \\ 
@@
 \end{tabular}
+	\caption{Results for automatic topic labeling. F1 = f-measure, MCC = Matthew's correlation coeff., ROC = Area under ROC curve}
+	\label{tbl:maxdb-unsup-results}
 \end{table*}
@@
 
-example-Accuracy & 0.45 & 0.40 & 0.50 & 0.40 & 0.40 & 0.38 & 0.33 & 0.48 \\
+example Accuracy & 0.45 & 0.40 & 0.50 & 0.40 & 0.40 & 0.38 & 0.33 & 0.48 \\
 
-example-Fmeasure & 0.56 & 0.48 & 0.59 & 0.47 & 0.49 & 0.46 & 0.40 & 0.55 \\
+example F1 & 0.56 & 0.48 & 0.59 & 0.47 & 0.49 & 0.46 & 0.40 & 0.55 \\
 
-example-HammingLoss & 0.24 & 0.19 & 0.18 & 0.19 & 0.23 & 0.19 & 0.20 & 0.17 \\
+example Hamming loss & 0.24 & 0.19 & 0.18 & 0.19 & 0.23 & 0.19 & 0.20 & 0.17 \\
 
-example-Precision & 0.50 & 0.51 & 0.60 & 0.52 & 0.52 & 0.50 & 0.44 & 0.60 \\
+example Precision & 0.50 & 0.51 & 0.60 & 0.52 & 0.52 & 0.50 & 0.44 & 0.60 \\
 
-example-Recall & 0.65 & 0.46 & 0.58 & 0.44 & 0.47 & 0.43 & 0.37 & 0.52 \\
+example Recall & 0.65 & 0.46 & 0.58 & 0.44 & 0.47 & 0.43 & 0.37 & 0.52 \\
 
-example-SubsetAccuracy & 0.19 & 0.24 & 0.31 & 0.24 & 0.23 & 0.23 & 0.20 & 0.31 \\
+example Subset Accuracy & 0.19 & 0.24 & 0.31 & 0.24 & 0.23 & 0.23 & 0.20 & 0.31 \\
 
-label-macro-ROC & 0.74 & 0.66 & 0.64 & 0.60 & 0.56 & 0.58 & 0.59 & 0.67 \\
+label macro-ROC & 0.74 & 0.66 & 0.64 & 0.60 & 0.56 & 0.58 & 0.59 & 0.67 \\
 
-label-macro-F1 & 0.46 & 0.30 & 0.43 & 0.28 & 0.32 & 0.27 & 0.22 & 0.39 \\
+label macro-F1 & 0.46 & 0.30 & 0.43 & 0.28 & 0.32 & 0.27 & 0.22 & 0.39 \\
 
-label-macro-Precision & 0.41 & 0.41 & 0.49 & 0.39 & 0.37 & 0.34 & 0.34 & 0.52 \\
+label macro-Precision & 0.41 & 0.41 & 0.49 & 0.39 & 0.37 & 0.34 & 0.34 & 0.52 \\
 
-label-macro-Recall & 0.56 & 0.26 & 0.40 & 0.24 & 0.31 & 0.23 & 0.19 & 0.33 \\
+label macro-Recall & 0.56 & 0.26 & 0.40 & 0.24 & 0.31 & 0.23 & 0.19 & 0.33 \\
 
-label-micro-ROC & 0.81 & 0.81 & 0.77 & 0.77 & 0.66 & 0.80 & 0.80 & 0.79 \\
+label micro-ROC & 0.81 & 0.81 & 0.77 & 0.77 & 0.66 & 0.80 & 0.80 & 0.79 \\
 
-label-micro-F1 & 0.59 & 0.53 & 0.61 & 0.51 & 0.50 & 0.51 & 0.45 & 0.59 \\
+label micro-F1 & 0.59 & 0.53 & 0.61 & 0.51 & 0.50 & 0.51 & 0.45 & 0.59 \\
 
-label-micro-Precision & 0.51 & 0.69 & 0.66 & 0.71 & 0.53 & 0.70 & 0.69 & 0.73 \\
+label micro-Precision & 0.51 & 0.69 & 0.66 & 0.71 & 0.53 & 0.70 & 0.69 & 0.73 \\
 
-label-micro-Recall & 0.68 & 0.43 & 0.56 & 0.41 & 0.48 & 0.41 & 0.34 & 0.49 \\
+label micro-Recall & 0.68 & 0.43 & 0.56 & 0.41 & 0.48 & 0.41 & 0.34 & 0.49 \\
 
-rank-AvgPrecision & 0.76 & 0.77 & 0.69 & 0.72 & 0.53 & 0.74 & 0.76 & 0.73 \\
+rank Avg. Precision & 0.76 & 0.77 & 0.69 & 0.72 & 0.53 & 0.74 & 0.76 & 0.73 \\
 
-rank-Coverage & 1.51 & 1.47 & 1.94 & 1.72 & 2.99 & 1.57 & 1.54 & 2.00 \\
+rank Coverage & 1.51 & 1.47 & 1.94 & 1.72 & 2.99 & 1.57 & 1.54 & 2.00 \\
 
-rank-One-error & 0.40 & 0.39 & 0.47 & 0.44 & 0.63 & 0.43 & 0.40 & 0.36 \\
+rank One-error & 0.40 & 0.39 & 0.47 & 0.44 & 0.63 & 0.43 & 0.40 & 0.36 \\
 
-rank-Ranking Loss & 0.19 & 0.18 & 0.27 & 0.22 & 0.52 & 0.20 & 0.19 & 0.28 \\
+rank Ranking Loss & 0.19 & 0.18 & 0.27 & 0.22 & 0.52 & 0.20 & 0.19 & 0.28 \\
 
@@
 
-example-Accuracy & 0.45 & 0.40 & 0.50 &           0.49  &  0.47  &  0.51 \\
+example Accuracy & 0.45 & 0.40 & 0.50 &           0.49  &  0.47  &  0.51 \\
 
-example-Fmeasure & 0.56 & 0.48 & 0.59 &           0.55  &  0.50  &  0.53 \\
+example F1 & 0.56 & 0.48 & 0.59 &           0.55  &  0.50  &  0.53 \\
 
-example-HammingLoss & 0.24 & 0.19 & 0.18 &        0.19  &  0.14  &  0.14 \\
+example Hamming Loss & 0.24 & 0.19 & 0.18 &        0.19  &  0.14  &  0.14 \\
 
-example-Precision & 0.50 & 0.51 & 0.60 &          0.50  &  0.49  &  0.53 \\
+example Precision & 0.50 & 0.51 & 0.60 &          0.50  &  0.49  &  0.53 \\
 
-example-Recall & 0.65 & 0.46 & 0.58 &             0.61  &  0.50  &  0.53 \\
+example Recall & 0.65 & 0.46 & 0.58 &             0.61  &  0.50  &  0.53 \\
 
-example-SubsetAccuracy & 0.19 & 0.24 & 0.31 &     0.40  &  0.43  &  0.45 \\
+example Subset Accuracy & 0.19 & 0.24 & 0.31 &     0.40  &  0.43  &  0.45 \\ 
+      \hline                                                                      
+label macro-ROC & 0.74 & 0.66 & 0.64 &            NaN  &  NaN  &  NaN    \\
 
-label-macro-ROC & 0.74 & 0.66 & 0.64 &            NaN  &  NaN  &  NaN    \\
+label macro-F1 & 0.46 & 0.30 & 0.43 &             0.32  &  0.19  &  0.23 \\
 
-label-macro-F1 & 0.46 & 0.30 & 0.43 &             0.32  &  0.19  &  0.23 \\
+label macro-Precision & 0.41 & 0.41 & 0.49 &      0.29  &  0.31  &  0.32 \\
 
-label-macro-Precision & 0.41 & 0.41 & 0.49 &      0.29  &  0.31  &  0.32 \\
+label macro-Recall & 0.56 & 0.26 & 0.40 &         0.38  &  0.15  &  0.20 \\
 
-label-macro-Recall & 0.56 & 0.26 & 0.40 &         0.38  &  0.15  &  0.20 \\
+label micro-ROC & 0.81 & 0.81 & 0.77 &            0.76  &  0.66  &  0.62 \\
 
-label-micro-ROC & 0.81 & 0.81 & 0.77 &            0.76  &  0.66  &  0.62 \\
+label micro-F1 & 0.59 & 0.53 & 0.61 &             0.39  &  0.27  &  0.34 \\
 
-label-micro-F1 & 0.59 & 0.53 & 0.61 &             0.39  &  0.27  &  0.34 \\
+label micro-Precision & 0.51 & 0.69 & 0.66 &      0.34  &  0.42  &  0.49 \\
 
-label-micro-Precision & 0.51 & 0.69 & 0.66 &      0.34  &  0.42  &  0.49 \\
+label micro-Recall & 0.68 & 0.43 & 0.56 &         0.47  &  0.21  &  0.26 \\ 
+      \hline                                                                      
+rank Avg. Precision & 0.76 & 0.77 & 0.69 &          0.54  &  0.57  &  0.54 \\
 
-label-micro-Recall & 0.68 & 0.43 & 0.56 &         0.47  &  0.21  &  0.26 \\
+rank Coverage & 1.51 & 1.47 & 1.94 &              1.40  &  1.23  &  1.43 \\
 
-rank-AvgPrecision & 0.76 & 0.77 & 0.69 &          0.54  &  0.57  &  0.54 \\
+rank One-error & 0.40 & 0.39 & 0.47 &             0.81  &  0.79  &  0.81 \\
 
-rank-Coverage & 1.51 & 1.47 & 1.94 &              1.40  &  1.23  &  1.43 \\
-
-rank-One-error & 0.40 & 0.39 & 0.47 &             0.81  &  0.79  &  0.81 \\
-
-rank-Ranking Loss & 0.19 & 0.18 & 0.27 &          0.41  &  0.35  &  0.41 \\ 
+rank Ranking Loss & 0.19 & 0.18 & 0.27 &          0.41  &  0.35  &  0.41 \\ 
 \hline
@@
 \caption{MySQL and MaxDB Mulan results per Multi-Label learner}
+\label{tab:mulan}
 \end{table*}

### 2010-03-06 icpc/icpc2010.tex
Tables are referenced
neil's refsq pub

--- 2010-03-06 icpc/icpc2010.tex
@@
 
-Part of our effort with this project is to understand the qualitative and intentional aspects of requirements in software evolution, a notion we first discussed in \cite{ernst07icsm}.
+% Part of our effort with this project is to understand the qualitative and intentional aspects of requirements in software evolution, a notion we first discussed in \cite{ernst07icsm}.
 That idea is derived from, in part, work on narratives of software systems shown in academic work like \cite{anton01}, or more general-purpose works like \cite{waldo93}.
@@
 They do not use external domain-specific information as we do.
+In \cite{ernst10refsq}, we describe an application similar to this to identify change in quality requirements in GNOME software projects; our approach is solely text-matching, however, and does not leverage machine learning strategies.
 
-Massey~\cite{massey02icse} and Scacchi (\cite{scacchi02,scacchi05b}) looked at the topic of requirements in open-source software.
+% Massey~\cite{massey02icse} and Scacchi (\cite{scacchi02,scacchi05b}) looked at the topic of requirements in open-source software.
 Their work discusses the source of the requirements and how they are used in the development process.
@@
 
-Hindle et al.
+% Hindle et al.
 \cite{Hindle2007} examined release patterns in OSS.
@@
 
-Most of this related research examined project source code.
+% Most of this related research examined project source code.
 Software mining of other project corpora is less common.

### 2010-03-06 icpc/icpc2010.tex

--- 2010-03-06 icpc/icpc2010.tex
@@
 
+We have presented the results of Mulan's multi-label learners in Table \ref{tab:mulan}.
 Binary Relevance (BR), Calibrated Label Ranking (CLR) and Hierarchy Of Multilabel classifiERs (HOMER), performed the best.

### 2010-03-06 icpc/icpc2010.tex
related work edits

--- 2010-03-06 icpc/icpc2010.tex
@@
 Part of our effort with this project is to understand the qualitative and intentional aspects of requirements in software evolution, a notion we first discussed in \cite{ernst07icsm}.
-That idea is derived from, in part, work on narratives of software systems shown in academic work like \cite{anton01}, or more general-purpose works like \cite{waldo93}.
+That idea is derived from, in part, work on narratives of software systems shown in academic work from Ant\'{o}n et al.~\cite{anton01}, or more general-purpose work from Waldo~\cite{waldo93}.
 
@@
 
-Cleland-Huang and her colleagues published work on mining requirements documents for non-functional requirements (quality requirements) in \cite{Cleland-Huang2006}.
-One approach they tried was similar to this one, with keywords mined from NFR catalogues found in~\cite{chung99}.
+Cleland-Huang and her colleagues published work on mining requirements documents for non-functional requirements (NFR) (quality requirements)~\cite{Cleland-Huang2006}.
+One approach they tried was similar to this one, with keywords mined from NFR catalogues found in their previous work~\cite{chung99}.
 They managed recall of 80\% with precision of 57\% for the Security NFR, but could not find a reliable source of keywords for other NFRs.
@@
 
-Mockus and Votta studied a large-scale industrial change-tracking system in ~\cite{Mockus00}.
+In the same vein, Mockus and Votta~\cote{Mockus00} studied a large-scale industrial change-tracking system.
 They also leveraged Wordnet, but only for word roots.
@@
 The other approach is to start with code repositories, and try to extract concerns from there.
-In \cite{marcus04wcre}, the authors describe their use of Latent Semantic Indexing to identify commonly occurring concerns for software maintenance.
+Marcus et al.~\cite{marcus04wcre} describe their use of Latent Semantic Indexing to identify commonly occurring concerns for software maintenance.
 Some results were interesting, but their precision was quite low.
@@
 The presence of a well-maintained set of tags is obviously essential to the success of this technique.
+\begin{comment}
 %We maintain that these results require more substantive manipulation before being reported to, for example, management.
@@
 Like us, they refer to release windows as a useful unit of analysis.
+\end{comment}
 
-Mens et al.
-\cite{mens08icsm} conducted an empirical study of Eclipse, the OSS code editor, to verify the claims of Lehman~\cite{lehman97sms}.
+Mens et al.~\cite{mens08icsm} conducted an empirical study of Eclipse, the open source software (OSS) source code editor, to verify the claims of Lehman~\cite{lehman97sms}.
 They concerned themselves with source code only, and found Law Seven, ``Declining Quality'', to be too difficult to assess: ``[we lacked an] appropriate measurement of the evolving quality of the system as perceived by the users \cite[p.
@@
 
-In \cite{Mei2007}, the authors use context information to automatically name topics.
+Mei et al.~\cite{Mei2007} use context information to automatically name topics.
 They describe probabilistic labeling, using the frequency distribution of words in a topic to create a meaningful phrase.
@@
 
-Massey~\cite{massey02icse} and Scacchi (\cite{scacchi02,scacchi05b}) looked at the topic of requirements in open-source software.
+Massey~\cite{massey02icse} and Scacchi~\cite{scacchi02,scacchi05b} looked at the topic of requirements in open-source software.
 Their work discusses the source of the requirements and how they are used in the development process.
@@
 Software mining of other project corpora is less common.
-Similarities exist with approaches that begin with structured taxonomies, as with the Hismo software ontology \cite{girba06}.
-Rigby and Hassan \cite{rigby07msr} used a general purpose taxonomy to classify developer email according to temperament.
+Similarities exist with approaches that begin with structured taxonomies, as with the Hismo software ontology~\cite{girba06}.
+Rigby and Hassan~\cite{rigby07msr} used a general purpose taxonomy to classify developer email according to temperament.
 

### 2010-03-06 icpc/icpc2010.tex

@@
 
+% Part of our effort with this project is to understand the qualitative and intentional aspects of requirements in software evolution, a notion we first discussed in \cite{ernst07icsm}.
+That idea is derived from, in part, work on narratives of software systems shown in academic work like \cite{anton01}, or more general-purpose works like \cite{waldo93}.
+
 % Not sure how do handle this
@@
 They do not use external domain-specific information as we do.
+In \cite{ernst10refsq}, we describe an application similar to this to identify change in quality requirements in GNOME software projects; our approach is solely text-matching, however, and does not leverage machine learning strategies.
 
+% comment out or not
 Massey~\cite{massey02icse} and Scacchi~\cite{scacchi02,scacchi05b} looked at the topic of requirements in open-source software.

### 2010-03-06 icpc/icpc2010.tex
oops cote -> cite
I tried to mine our email messages for notes and this is what I found

@@
 \end{tabular}
-	\caption{Automatic topic labeling}
+	\caption{Automatic topic labeling for MaxDB 7.500}
 	\label{tbl:wordlist}
@@
 
+\begin{comment}
+A quick mind dump before I forget:
+* portability was very popular
+* there were many instances of concurrency
+* There were often topics of crap commits with no comment
+* might need more than the top words to compare
+* Things like tests were noticable more by file type
+* Configuration management is not part of the software quality ontology?
+* I didn't tag none I tagged unknown, so that's some inconsistency between us
+* There were a lot of copyright changes in mysql
+* mysql had optimizer module which did not mean optimization
+* I tried to assign reliability/bug for reliability bugs
+  and Functionality/Correctness for fixes
+* Some security changes came later
+* Mysql 3.23 transitioned from a project being developped to
+  a project being maintained, more frequent backport bugfixes
+\end{comment}
+
 \subsection{Effectiveness}
@@
 
-In the same vein, Mockus and Votta~\cote{Mockus00} studied a large-scale industrial change-tracking system.
+In the same vein, Mockus and Votta~\cite{Mockus00} studied a large-scale industrial change-tracking system.
 They also leveraged Wordnet, but only for word roots.

### 2010-03-06 icpc/icpc2010.tex
observations

@@
 
+\subsection{Annotation Observations}
+
+We also found there were quite a few topics that were not non-functional requirements but were often related to them.
+For instance concurrency was mentioned often in the commit logs and was related to correctness and reliability because concurrency was troublesome.
+Configuration management and source control related changes appeared often and sometimes had topics dedicated to configuration management, these kinds of changes are slightly related to maintainability.
+A non-functional change that was not quality related was licensing and copyright, many changes were simply to do with updating copyrights or ensuring copyright or licenses headers were applied to files.
+
+We noticed that sometimes the names of modules would conflict with words related to other non-functional requirements.
+For instance optimizers are very common modules in database systems, both MySQL and MaxDB have optimizer modules.
+In MySQL the optimizer is mentioned but often the change is deals with correctness or another quality, yet the name of the module has the potential to confuse automated approaches.
+This is why we feel that perhaps a project specific word-bags are needed in order to avoid automated mistakes due to the names of entities and modules with the software project.
+
 \begin{comment}

### 2010-03-06 icpc/icpc2010.tex
spell check

@@
 We focus   on non-functional requirements related to software quality as a   potential generalization, since there is some shared belief that   these qualities apply broadly across many software systems and their   development artifacts.
-In our approach, we name topics based on a   static ontology of non-functional requirements and software   qualities; in essence, this adds context to general-purpose   ontologies such as Wordnet.
+In our approach, we name topics based on a   static ontology of non-functional requirements and software   qualities; in essence, this adds context to general-purpose   ontologies such as WordNet.
 In this paper, we discuss how we use   this software development domain knowledge to better annotate topics   of development that are extracted from the change log messages in   configuration management systems such as CVS.
@@
 As development of a software project progresses, developers must concern themselves with a wide variety of issues --- or ``topics'' --- that relate to the problem domain, the processes and tools used, the development artifacts themselves, etc.
-The set of development topics for a given project can sometimes be extracted automatically by analyzing artifacts within software repositories, such as changelog comments that developers create when committing revisions to the project's source control system.
+The set of development topics for a given project can sometimes be extracted automatically by analyzing artifacts within software repositories, such as change-log comments that developers create when committing revisions to the project's source control system.
 Our previous work dealt with topic trends, which are topics that recur over time~\cite{Hindle09ICSM}.
@@
 Topics in this paper and the previous paper are word bags or word distributions.
-These word distributions are found via Latent Dirichlet Allocation~\cite{Blei2003}, which finds independent word distributions shared among documents (changelog comments).
+These word distributions are found via Latent Dirichlet Allocation~\cite{Blei2003}, which finds independent word distributions shared among documents (change-log comments).
 The unfortunate aspect of topics that are word distributions is that they lack the tangibility of a rose.
@@
 This is an ontological approach where we attempt to relate networks of words to labels and then search for these terms in our topic.
-We also compare this approach to machine leaners.
+We also compare this approach to machine learners.
 %We start with an unsupervised topic learning algorithm, then manually %annotate the topics using the ontology.
@@
 % unsure about this
-\item We show that these labels with their topics can be learned and used to classify other datasets.
+\item We show that these labels with their topics can be learnt and used to classify other data-sets.
 \item We present visualizations of named topics and their trends over time to aid communication and analysis.
@@
 We first introduce some important terminology for our work.
-We then describe our methodology, including our datasets, then highlight our results.
+We then describe our methodology, including our data-sets, then highlight our results.
 We conclude with a look at related work and possible improvements.
@@
 % check
-\emph{Area of ROC Curve} is the area under the Reciever Operating Characteric curve, we will refer to this as \emph{ROC} although it is sometimes referred to as \emph{AUC}.
+\emph{Area of ROC Curve} is the area under the Receiver Operating Characteristic curve, we will refer to this as \emph{ROC} although it is sometimes referred to as \emph{AUC}.
 ROC values reflect a score, similar to school letter-grades (A is 0.9, C is 0.6), reflecting how well a particular learner performed for the given data.
@@
 LDA works by creating topic models for each period.
-A topic model is a statistical summary of the important subjects for that dataset.
+A topic model is a statistical summary of the important subjects for that data-set.
 The input to the algorithm is a set of documents, with contents treated as a bag of words, and a parameter to set the number of topics to generate.
 The algorithm returns a word-document matrix relating the occurrence frequency of a
- The idea is similar to other IR approaches like TF/IDF (term frequency/inverse document frequency), which find uncommon words that are not found in other wordlists.
+ The idea is similar to other IR approaches like TF/IDF (term frequency/inverse document frequency), which find uncommon words that are not found in other word-lists.
 These approaches are unsupervised in that they rely solely on the source data with no human intervention.
@@
 The labels are reasonable enough, but still require some expert in the field to determine them.
-Our technique is automated, because we match keywords from Wordnet to words in the topic model.
+Our technique is automated, because we match keywords from WordNet to words in the topic model.
 
@@
 We extracted these messages and indexed them by creation time.
-We summarized each emessage as a word distribution but removed stop-words such as   common English words like \emph{the} and \emph{at}.
+We summarized each message as a word distribution but removed stop-words such as   common English words like \emph{the} and \emph{at}.
 
-From that dataset, we created an XML file which windowed commits into
+From that data-set, we created an XML file which windowed commits into
 monthly windows.
@@
 
-To construct our expanded wordlists, we used Wordnet~\cite{Fellbaum1998}, an English-language ``lexical database'' that contains semantic relations between words, including meronymy and synonymy.
+To construct our expanded word-lists, we used WordNet~\cite{Fellbaum1998}, an English-language ``lexical database'' that contains semantic relations between words, including meronymy and synonymy.
 We then added Boehm’s 1976 software quality model \cite{Boehm+:1976:ICSE}, and classified his eleven ‘ilities’ into their respective ISO9126 qualities.
@@
 
-For the third -- \textsf{exp3} -- list of quality labels, we extended the list from \textsf{exp2} using unfiltered Wordnet similarity matches.
-Similarity in Wordnet means siblings in a hypernym tree.
+For the third -- \textsf{exp3} -- list of quality labels, we extended the list from \textsf{exp2} using unfiltered WordNet similarity matches.
+Similarity in WordNet means siblings in a hypernym tree.
 We do not include these words here for space considerations (but see the Appendix for our data repository).
@@
 \end{tabular}
-	\caption{Qualities and associated signifiers – Wordnet version (\textsf{exp2})}
+	\caption{Qualities and associated signifiers – WordNet version (\textsf{exp2})}
 	\label{tbl:wnsig}
@@
 
-In order to validate how effective these wordbag approaches to topic labelling would be we had to make a data set to test against.
+In order to validate how effective these word-bag approaches to topic labelling would be we had to make a data set to test against.
 For MySQL and MaxDB, we manually annotated each extracted topic in each period with the same quality labels as \textsf{exp2} (software qualities).
 We looked at each period's topics, and assessed what the data -- consisting of the frequency-weighted word lists and messages -- suggested was the topic for that period.
-We were able to pinpoint the appropriate label using auxillary information as well, such as the actual revisions and files that were related to this topic.
+We were able to pinpoint the appropriate label using auxiliary information as well, such as the actual revisions and files that were related to this topic.
 For example, for the MaxDB topic consisting of a message ``exit() only used in non NPTL LINUX Versions'', we tagged that topic \emph{portability}.
-We compared against this dataset but we also used it for our supervised learning based topic classification.
+We compared against this data-set but we also used it for our supervised learning based topic classification.
 %inter-rater reliability
@@
 WEKA contains a suite of machine learning tools such as support vector machines and Bayes nets.
-We also used the multi-labeling add-on for WEKA, Mulan~\cite{mulan}\footnote{\url{http://mlkd.csd.auth.gr/multilabel.html}}.
+We also used the multi-labelling add-on for WEKA, Mulan~\cite{mulan}\footnote{\url{http://mlkd.csd.auth.gr/multilabel.html}}.
 Traditional classifiers map our topics to a single class, whereas Mulan allows for a mixture of classes per topic, much like our manual classification found overlaps.
@@
 To assess the performance of the supervised learners, we did a 10-fold cross-validation.
-This is when a set is partitioned into 10 partitions and then each partition is used once as a test set and 9 other times as part of the training set of 9 paritions.
-We haved reported these results below.
+This is when a set is partitioned into 10 partitions and then each partition is used once as a test set and 9 other times as part of the training set of 9 partitions.
+We have reported these results below.
 
@@
 \end{tabular}
-	\caption{Automatic topic labeling for MaxDB 7.500}
+	\caption{Automatic topic labelling for MaxDB 7.500}
 	\label{tbl:wordlist}
@@
 For \textsf{exp3}, we had many more named topics.
-As we mentioned, the wordlists are quite broad, so there are likely to be false-positives.
+As we mentioned, the word-lists are quite broad, so there are likely to be false-positives.
 See the following sections for our error analysis.
@@
 
-\subsubsection{Analysing the unsupervised labeling}
+\subsubsection{Analyzing the unsupervised labelling}
 
-Based on the labels, and our manual topic labeling, we compared the results of the unsupervised word matching approach.
+Based on the labels, and our manual topic labelling, we compared the results of the unsupervised word matching approach.
 For each quality we tried to assess whether the manual tag matched the unsupervised label assigned.
@@
 \end{tabular}
-	\caption{Results for automatic topic labeling. F1 = f-measure, MCC = Matthew's correlation coeff., ROC = Area under ROC curve}
+	\caption{Results for automatic topic labelling. F1 = f-measure, MCC = Matthew's correlation coeff., ROC = Area under ROC curve}
 	\label{tbl:maxdb-unsup-results}
@@
 Table \ref{tab:best-learn-per-tag} shows the ROC value, the accuracy of a ZeroR learner ( a learner which just chooses the largest category all of the time), the difference in accuracy between the best learner and ZeroR, and the F-measure for that particular learner.
-The ZeroR difference is often negative, but in particular for labels which are not as common this can be expected because any miscategorization will hurt accuracy, this is why the Fmeasure and the ROC values are useful as they can better present performance on labels which are not applicable to the majority of samples.
+The ZeroR difference is often negative, but in particular for labels which are not as common this can be expected because any miscategorization will hurt accuracy, this is why the F1 (F-measure) and the ROC values are useful as they can better present performance on labels which are not applicable to the majority of samples.
 
@@
 \toprule
-Label & Project & Learner & ROC & ZeroR Acc. & ZeroR. Diff & Fmeasure\\
+Label & Project & Learner & ROC & ZeroR Acc. & ZeroR. Diff & F1\\
 \midrule
@@
 
-We applied the Mulan~\cite{mulan} library for Multi-Label learning to our dataset because we feel that topics can have more like one label, much like how a revision can have more than one topic.
+We applied the Mulan~\cite{mulan} library for Multi-Label learning to our data-set because we feel that topics can have more like one label, much like how a revision can have more than one topic.
 Multi-label learning is more than just classifying entities with more than one label.
@@
 Micro measurements are aggregated at a decision level.
-So a macro-ROC measurment is the average ROC over the ROC values for all labels, where a micro-ROC is the average ROC over all examples that were classified.
+So a macro-ROC measurement is the average ROC over the ROC values for all labels, where a micro-ROC is the average ROC over all examples that were classified.
 Unfortunately for MaxDB the macro-ROC values are undefined because of poor performance of one of the labels.
@@
 
-We found that the multilabel learners of BR, CLR and HOMER did not do as well for Macro-ROC and Micro-F1 as NaiveBayes and other NaiveBayes derived learners did.
-This suggests that by sticking together multiple NaiveBayes leaners we could probably label topics effectively, but it would require a separate NaiveBayes learner per label.
+We found that the multi-label learners of BR, CLR and HOMER did not do as well for Macro-ROC and Micro-F1 as NaiveBayes and other NaiveBayes derived learners did.
+This suggests that by sticking together multiple NaiveBayes learners we could probably label topics effectively, but it would require a separate NaiveBayes learner per label.
 
@@
 * Some security changes came later
-* Mysql 3.23 transitioned from a project being developped to
-  a project being maintained, more frequent backport bugfixes
+* Mysql 3.23 transitioned from a project being developed to
+  a project being maintained, more frequent back-port bug-fixes
 \end{comment}
@@
 If the techniques used in \textsf{exp2} and \textsf{exp3} were combined with the supervised techniques we could probably reduce the training effort.
-Both Naive Bayesian learners and the wordlist approaches were computationally efficient.
+Both Naive Bayesian learners and the word-list approaches were computationally efficient.
 Thus these results are promising because the results are accurate enough to be useful and cheap enough to execute to be feasible as an automated or semi-automated method of labelling topics by their software qualities.
@@
 However, both projects are developed by commercial entities.
-Our technique is dependent on development practices including meaningful commit messages, which might not be true in other domains.
+Our technique is dependent on development practises including meaningful commit messages, which might not be true in other domains.
 
@@
 In the same vein, Mockus and Votta~\cite{Mockus00} studied a large-scale industrial change-tracking system.
-They also leveraged Wordnet, but only for word roots.
+They also leveraged WordNet, but only for word roots.
 They felt the synonyms would be non-specific and cause errors.
@@
 Some results were interesting, but their precision was quite low.
-ConcernLines~\cite{treude09cl} shows tag occurrence using color intensity.
+ConcernLines~\cite{treude09cl} shows tag occurrence using colour intensity.
 They mined change request tags (such as `milestone 3') and used these to make evolutionary analyses of a single product.
@@
 Mei et al.~\cite{Mei2007} use context information to automatically name topics.
-They describe probabilistic labeling, using the frequency distribution of words in a topic to create a meaningful phrase.
+They describe probabilistic labelling, using the frequency distribution of words in a topic to create a meaningful phrase.
 They do not use external domain-specific information as we do.

### 2010-03-06 icpc/icpc2010.tex
possible change to intro

@@
 
-Few would argue that software development is a bed of roses.
+% Few would argue that software development is a bed of roses.
 To most of us, the important properties of a rose are sensual in nature, and concern appearance, odour, and feel.
@@
 Yet to have meaningful discussions about how software development is progressing, we must be able to devise appropriate labels for our work topics and be able to categorize the development artifacts as belonging to one or more of these topics.
-
 %Original is here:
@@
 
-As development of a software project progresses, developers must concern themselves with a wide variety of issues --- or ``topics'' --- that relate to the problem domain, the processes and tools used, the development artifacts themselves, etc.
+To have meaningful discussions about how software development is progressing, we must devise appropriate labels for our work topics, in contrast to the quote above.
+We can then categorize development artifacts as belonging to one or more of these topics.
+
+Topics arise from the wide variety of issues which occurs during a software project's lifecycle.
+These topics can relate to, among others, the problem domain, the processes and tools used, or the development artifacts themselves.
 The set of development topics for a given project can sometimes be extracted automatically by analyzing artifacts within software repositories, such as change-log comments that developers create when committing revisions to the project's source control system.
@@
 
-Part of our effort with this project is to understand the qualitative and intentional aspects of requirements in software evolution, a notion we first discussed in \cite{ernst07icsm}.
+% Part of our effort with this project is to understand the qualitative and intentional aspects of requirements in software evolution, a notion we first discussed in \cite{ernst07icsm}.
 That idea is derived from, in part, work on narratives of software systems shown in academic work from Ant\'{o}n et al.~\cite{anton01}, or more general-purpose work from Waldo~\cite{waldo93}.
@@
 % comment out or not
-Massey~\cite{massey02icse} and Scacchi~\cite{scacchi02,scacchi05b} looked at the topic of requirements in open-source software.
+% Massey~\cite{massey02icse} and Scacchi~\cite{scacchi02,scacchi05b} looked at the topic of requirements in open-source software.
 Their work discusses the source of the requirements and how they are used in the development process.

### 2010-03-06 icpc/icpc2010.tex
half-reversion

--- 2010-03-06 icpc/icpc2010.tex
@@
 
+%To have meaningful discussions about how software development is progressing, we must devise appropriate labels for our work topics, in contrast to the quote above.
+We can then categorize development artifacts as belonging to one or more of these topics.
+
+% This correlates issues with topics
 As development of a software project progresses, developers must concern themselves with a wide variety of issues --- or ``topics'' --- that relate to the problem domain, the processes and tools used, the development artifacts themselves, etc.
+The set of development topics for a given project can sometimes be extracted automatically by analyzing artifacts within software repositories, such as change-log comments that developers create when committing revisions to the project's source control system.
+Our previous work dealt with topic trends, which are topics that recur over time~\cite{Hindle09ICSM}.
+We observed that topic trends were often non-functional requirements.
+
+%Topics arise from the wide variety of issues which occurs during a software project's lifecycle.
+These topics can relate to, among others, the problem domain, the processes and tools used, or the development artifacts themselves.
 The set of development topics for a given project can sometimes be extracted automatically by analyzing artifacts within software repositories, such as change-log comments that developers create when committing revisions to the project's source control system.

### 2010-03-06 icpc/icpc2010.tex
background

@@
 Since our taxonomy is   global in scope, our results allow us to compare the relative   importance of particular qualities across different projects.
-With   in the topics extracted from MySQL and MaxDB we found that   non-functional requirements were common and often were trends across   time.
+Within the topics extracted from MySQL and MaxDB we found that   non-functional requirements were common and often were trends across   time.
 Our automated techniques can label topics without much computational overhead.
@@
 % Clean this up
-Traditionally topic extraction has required manual annotation to derive domain-relevant labelling information.
+Traditionally topic extraction has required manual annotation to derive domain-relevant labels.
 This paper attempts to implement \emph{labelled topic extraction}, in which topic are extracted and given labels relating to a cross-project taxonomy.
-This is an ontological approach where we attempt to relate networks of words to labels and then search for these terms in our topic.
+This is an ontological approach where we attempt to relate networks of words to labels and then search for these terms within our topics.
 We also compare this approach to machine learners.
@@
 paper, messages will be the CVS and BitKeeper commit log comments made
-when the user commits changes to files in a repository. A \emph{word
+when a developer commits changes to files in a repository. A \emph{word
   distribution} is the summary of a message by word count. Each word
@@
 \emph{Area of ROC Curve} is the area under the Receiver Operating Characteristic curve, we will refer to this as \emph{ROC} although it is sometimes referred to as \emph{AUC}.
-ROC values reflect a score, similar to school letter-grades (A is 0.9, C is 0.6), reflecting how well a particular learner performed for the given data.
+ROC values reflect a score, similar to school letter-grades (an \emph{A} is $0.9$, a \emph{C} is $0.6$), reflecting how well a particular learner performed for the given data.
 A ROC result of 0.5 would be equivalent to a random learner (randomly classifying data).

### 2010-03-06 icpc/icpc2010.tex
topic extraction section edit

@@
 	documents (commit log comments).
-Many researchers~\cite{marcus04wcre,Poshyvanyk2007,lukins08wcre,Linstead2007} have applied tools like LSI and LDA for mining repositories, in particular analyzing source code, bugs or developer comments.
+Many researchers~\cite{marcus04wcre,Poshyvanyk2007,lukins08wcre,Linstead2007} have applied tools like LSI and LDA for mining software repositories, in particular analyzing source code, bugs or developer comments.
 \begin{comment}
@@
 
-One issue that arises with use of unsupervised techniques is how to label the topics.
+One issue that arises with the use of unsupervised techniques is how to label the topics.
 While the topic models themselves are generated automatically, what to make of them is less clear.
 For example, in our previous work~\cite{Hindle09ICSM}, and in \cite{Baldi2008}, topics are named manually: human experts read the highest-frequency members of a topic and assign a keyword accordingly.
-E.g., for the word list ``listener change remove add fire'', Baldi et al.
+For example, given the word list ``listener change remove add fire'', Baldi et al.
 assign the keyword \emph{event-handling}.

### 2010-03-06 icpc/icpc2010.tex
Supervised learning

@@
 While unsupervised techniques (LSI and LDA are both unsupervised) are appealing in their lack of human intervention, supervised learners have the advantage of domain knowledge.
-This typically results in improved results.
+This typically produces improved results.
 In supervised learning, the data is divided into slices.
 One slice is manually annotated by the domain expert
-In this paper, we employ the WEKA~\cite{weka09} and Mulan~\cite{mulan} machine learning frameworks in order to test supervised learning.
+In this paper, we employ the WEKA~\cite{weka09} and Mulan~\cite{mulan} machine learning frameworks in order to test various supervised learners on our annotated data-set.
 

### 2010-03-06 icpc/icpc2010.tex
methodology edit
Associating Labels

@@
 To evaluate our approach, we sought candidate systems that were
-mature projects and had openly accessible artifact repositories.
+mature projects and had openly accessible source control repositories.
 We also decided to select systems from the same application
@@
 Once we had topics for each period, we tried to associate them with a label from a list of keywords and related terms.
-We performed simple string matching between these topics and our lists, `naming' a topic if it contained that word or words.
+We performed simple string matching between these topics and our lists, `naming' a topic if it contained that a word from our word list.
 We used several different word lists for comparison.
 
-Our first list, \textsf{exp1}, was generated using the ontology described in Kayed et al.~\cite{5072519}.
+Our first word list set, \textsf{exp1}, was generated using the ontology described in Kayed et al.~\cite{5072519}.
 That paper constructs an ontology for software quality measurement using eighty source documents, including research papers and international standards.
@@
 
-Our second list, \textsf{exp2}, relied on the ISO quality model (ISO9126)~\cite{iso9126}. ISO9126 describes six high-level quality requirements (listed in Table \ref{tbl:wnsig}). 
+Our second word list set, \textsf{exp2}, relied on the ISO quality model (ISO9126)~\cite{iso9126}. ISO9126 describes six high-level quality requirements (listed in Table \ref{tbl:wnsig}). 
 %There is some debate about the significance and importance of the %terms in this model. 

### 2010-03-06 icpc/icpc2010.tex
supervised learning

@@
 
+\label{sec:suplearn}
+
 In order to validate how effective these word-bag approaches to topic labelling would be we had to make a data set to test against.
-For MySQL and MaxDB, we manually annotated each extracted topic in each period with the same quality labels as \textsf{exp2} (software qualities).
+For MySQL 3.23 and MaxDB 7.500, we manually annotated each extracted topic in each period with the same quality labels as \textsf{exp2} (software qualities).
 We looked at each period's topics, and assessed what the data -- consisting of the frequency-weighted word lists and messages -- suggested was the topic for that period.
@@
 
-We first compared our previous analysis using label matching to our manual classifications to get an error rate for that process (described below).
+We first compared our previous analysis using label matching to our manual classifications to get an error rate for that process described below in Section \ref{sec:unsuplabelling}.
 
@@
 We also used the multi-labelling add-on for WEKA, Mulan~\cite{mulan}\footnote{\url{http://mlkd.csd.auth.gr/multilabel.html}}.
-Traditional classifiers map our topics to a single class, whereas Mulan allows for a mixture of classes per topic, much like our manual classification found overlaps.
+Traditional classifiers map our topics to a single class, whereas Mulan allows for a mixture of classes per topic, which maps to what we observed while manually labelling topics.
 
@@
 This is when a set is partitioned into 10 partitions and then each partition is used once as a test set and 9 other times as part of the training set of 9 partitions.
-We have reported these results below.
+We have reported these results below in Section \ref{sec:suplabelling}.
 
-% enumerate?
 Finally, using this data, we evaluated two research questions:
@@
 
-\subsubsection{Analyzing the unsupervised labelling}
+\subsection{Analysis the unsupervised labelling}
+
+\label{sec:unsuplabelling}
 
@@
 
-\subsection{Classifying topics}
+%\subsection{Classifying topics}
+\subsection{Analysis of the supervised labelling}
+\label{sec:suplabelling}
 

### 2010-03-06 icpc/icpc2010.tex
unsupervised

@@
 A \emph{named   topic} is a topic with a matching label.
-There are \{periods X 20\} topics per project.
-All experiments were run on MaxDB and MySQL data.
+There are \{periods X 20\} topics per project as we told LDA to extract 20 topics per period.
+All experiments were run on MaxDB 7.500 and MySQL 3.23 data.
 
@@
 
-For \textsf{exp2}, there are many more unnamed topics. Only reliability produces a lot of matches, mostly with the `error' word. Co-occurrence results were poor.
+For \textsf{exp2}, there are many more unnamed topics. Only reliability produces a lot of matches, mostly with the word `error'. Co-occurrence results were poor.
 
@@
 
-\subsection{Analysis the unsupervised labelling}
+\subsection{Analysis of the unsupervised labelling}
 
@@
 Using the F-Measure, the weighted average of precision and recall, where 1 is perfect, our best results are 0.6, a few at 0.4, and most around 0.2.
-Similar results using the Matthew's correlation coefficient (used to measure efficacy where classes are of different sizes), and the receiver operating characteristic area under curve.
+We achieved similar results using the Matthew's correlation coefficient (used to measure efficacy where classes are of different sizes) and ROC.
+
+Based on these results we find that reliability and usability worked well for MaxDB in \textsf{exp2} and better in \textsf{exp3}.
+MySQL reasonable results within \textsf{exp2} were for reliability and efficiency.
+MySQL's results for efficiency did not improve in \textsf{exp3} but other qualities such as functionality did improve.
+If a \emph{C} grade performance has a ROC value of $0.6$ then most of these tests scored \emph{C} or less, but our results were still better than random chance.
 
@@
 
-We took our annotated dataset and applied supervised learner to it.
+We took our annotated dataset and applied supervised learners to it.
 Because our dataset was of word counts we expected Bayesian techniques, often used in spam filtering, to perform well.

### 2010-03-06 icpc/icpc2010.tex
supervised labelling
changes to intro and refs

--- 2010-03-06 icpc/icpc2010.tex
@@
 Since our taxonomy is   global in scope, our results allow us to compare the relative   importance of particular qualities across different projects.
-Within the topics extracted from MySQL and MaxDB we found that   non-functional requirements were common and often were trends across   time.
+With   in the topics extracted from MySQL and MaxDB we found that   non-functional requirements were common and often were trends across   time.
 Our automated techniques can label topics without much computational overhead.
@@
 
-%To have meaningful discussions about how software development is progressing, we must devise appropriate labels for our work topics, in contrast to the quote above.
+% To have meaningful discussions about how software development is progressing, we must devise appropriate labels for our work topics, in contrast to the quote above.
 We can then categorize development artifacts as belonging to one or more of these topics.
 
-% This correlates issues with topics
-As development of a software project progresses, developers must concern themselves with a wide variety of issues --- or ``topics'' --- that relate to the problem domain, the processes and tools used, the development artifacts themselves, etc.
-The set of development topics for a given project can sometimes be extracted automatically by analyzing artifacts within software repositories, such as change-log comments that developers create when committing revisions to the project's source control system.
-Our previous work dealt with topic trends, which are topics that recur over time~\cite{Hindle09ICSM}.
-We observed that topic trends were often non-functional requirements.
-
-%Topics arise from the wide variety of issues which occurs during a software project's lifecycle.
+Topics arise from the wide variety of issues which occur during a software project's lifecycle.
 These topics can relate to, among others, the problem domain, the processes and tools used, or the development artifacts themselves.
@@
 A topic --- that is, a word distribution in our world --- needs a suggestive name to succinctly convey its intent, and make it easy to use in discussions and analyses about the development of the system.
-Our previous experience leads us to believe that automated topic naming is currently infeasible in the general case, so we have decided to focus on topic labels from the sub-domain of non-functional requirements related to software quality.
+Our previous experience leads us to believe that automated topic naming is currently infeasible in the general case.
+We have therefore decided to focus on topic labels from the sub-domain of non-functional requirements related to software quality.
 
@@
 % Clean this up
-Traditionally topic extraction has required manual annotation to derive domain-relevant labels.
+Traditionally topic extraction has required manual annotation to derive domain-relevant labelling information.
 This paper attempts to implement \emph{labelled topic extraction}, in which topic are extracted and given labels relating to a cross-project taxonomy.
-This is an ontological approach where we attempt to relate networks of words to labels and then search for these terms within our topics.
+This is an ontological approach where we attempt to relate networks of words to labels and then search for these terms in our topic.
 We also compare this approach to machine learners.
@@
 paper, messages will be the CVS and BitKeeper commit log comments made
-when a developer commits changes to files in a repository. A \emph{word
+when the user commits changes to files in a repository. A \emph{word
   distribution} is the summary of a message by word count. Each word
@@
 of a group of messages. In this paper we often summarize topics by the
-top 10 most frequent words of their word distribution.  A \emph{trend}
+top ten most frequent words of their word distribution.  A \emph{trend}
 is one or more similar topics that recur over time.  Trends are
@@
 % check
-\emph{Area of ROC Curve} is the area under the Receiver Operating Characteristic curve, we will refer to this as \emph{ROC} although it is sometimes referred to as \emph{AUC}.
-ROC values reflect a score, similar to school letter-grades (an \emph{A} is $0.9$, a \emph{C} is $0.6$), reflecting how well a particular learner performed for the given data.
+\emph{Area of ROC Curve} is the area under the Receiver Operating Characteristic curve, sometimes referred to as \emph{AUC}.
+ROC values reflect a score, similar to school letter-grades (A is 0.9, C is 0.6), reflecting how well a particular learner performed for the given data.
 A ROC result of 0.5 would be equivalent to a random learner (randomly classifying data).
@@
 	documents (commit log comments).
-Many researchers~\cite{marcus04wcre,Poshyvanyk2007,lukins08wcre,Linstead2007} have applied tools like LSI and LDA for mining software repositories, in particular analyzing source code, bugs or developer comments.
+Many researchers~\cite{marcus04wcre,Poshyvanyk2007,lukins08wcre,Linstead2007} have applied tools like LSI and LDA for mining repositories, in particular analyzing source code, bugs or developer comments.
 \begin{comment}
@@
 Linear combinations of these $N$ word distributions are then meant to be able to recreate the word distributions of all of the underlying messages.
-These $N$ word distributions effectively form topics: cross cutting collections of words relevant to 1 or more documents.
-Our problem is that these topics are not easy to interpret and we feel that automatic labelling or naming of these topics would be helpful with respect to interpreting what a topic is about.
-LDA's topics extracted in an unsupervised manner, LDA relies solely on the source data with no human intervention.
+These $N$ word distributions effectively form topics: cross cutting collections of words relevant to one or more documents.
+Our problem is that these topics are not easy to interpret, as the underlying pattern is not clear.
+We feel that automatic labelling or naming of these topics would be helpful with respect to interpreting the subject of a topic.
+LDA extracts topics in an unsupervised manner; the algorithm relies solely on the source data with no human intervention.
 
 In topic analysis a single document, such as a commit message, can be related to multiple topics.
-Representing documents as a mixture of topics maps well to commits to source code repositories, which often have more than one purpose~\cite{Hindle09ICSM}.
+Representing documents as a mixture of topics maps well to source code repository commits, which often have more than one purpose~\cite{Hindle09ICSM}.
 A topic represents both a word distribution and a group of commit log comments that are related to each other by their content.
@@
 
-One issue that arises with the use of unsupervised techniques is how to label the topics.
+One issue that arises with use of unsupervised techniques is how to label the topics.
 While the topic models themselves are generated automatically, what to make of them is less clear.
-For example, in our previous work~\cite{Hindle09ICSM}, and in \cite{Baldi2008}, topics are named manually: human experts read the highest-frequency members of a topic and assign a keyword accordingly.
-For example, given the word list ``listener change remove add fire'', Baldi et al.
+For example, in our previous work~\cite{Hindle09ICSM}, as well as in \cite{Baldi2008}, topics are named manually: human experts read the highest-frequency members of a topic and assign a keyword accordingly.
+E.g., for the word list \emph{``listener change remove add fire''}, Baldi et al.
 assign the keyword \emph{event-handling}.
-The labels are reasonable enough, but still require some expert in the field to determine them.
-Our technique is automated, because we match keywords from WordNet to words in the topic model.
+The labels are reasonable enough, but still require an expert in the field to determine them.
+Our technique is automated, because we match keywords from WordNet~\cite{Fellbaum1998} to words in the topic model.
 
 \subsection{Supervised learning}
-While unsupervised techniques (LSI and LDA are both unsupervised) are appealing in their lack of human intervention, supervised learners have the advantage of domain knowledge.
-This typically produces improved results.
+While unsupervised techniques (LSI and LDA are both unsupervised) are appealing in their lack of human intervention, and thus lower effort, supervised learners have the advantage of domain knowledge which typically means improved results.
 In supervised learning, the data is divided into slices.
-One slice is manually annotated by the domain expert
-In this paper, we employ the WEKA~\cite{weka09} and Mulan~\cite{mulan} machine learning frameworks in order to test various supervised learners on our annotated data-set.
+One slice is manually annotated by the domain expert, and the classifications he/she determines are applied to the remaining slices.
+In this paper, we employ the WEKA~\cite{weka09} and Mulan~\cite{mulan} machine learning frameworks in order to test supervised learning.
 
@@
 To evaluate our approach, we sought candidate systems that were
-mature projects and had openly accessible source control repositories.
+mature projects and had openly accessible artifact repositories.
 We also decided to select systems from the same application
@@
 \subsection{Generating the data}
-From each project, we used source control commit comments, these are the messages that programmers write when they commit revisions to a source control repository.
+From each project, we used source control commit comments, the messages that programmers write when they commit revisions to a source control repository.
 We leveraged the data that we gathered in \cite{Hindle09ICSM} for this work.
-An example of a typical commit message is: ``history annotate diffs bug fixed (if mysql\_real\_connect() failed there were two pointers to malloc'ed strings, with memory corruption on free(), of course)''.
+An example of a typical commit message is: \textit{``history annotate diffs bug fixed (if mysql\_real\_connect() failed there were two pointers to malloc'ed strings, with memory corruption on free(), of course)''}.
 We extracted these messages and indexed them by creation time.
-We summarized each message as a word distribution but removed stop-words such as   common English words like \emph{the} and \emph{at}.
+We summarized each message as a word distribution but removed stop-words such as common English words \emph{the} and \emph{at}.
 
-From that data-set, we created an XML file which windowed commits into
+From that data-set, we created an XML file which separated commits into
 monthly windows.
 This size of period is smaller than the time between minor releases but large enough for there to be sufficient commits to analyze.
-We applied an Blei's LDA implementation against the word distributions of these commits, and generated list of topics per period.
+We applied Blei's LDA implementation~\cite{Blei2003} against the word distributions of these commits, and generated list of topics per period.
 %A topic extracted in the first period Period 1, the most important %topic might have been ``security", with ``sql'' second, down on to %less relevant topics. 
@@
 
-Topics are word distributions: essentially lists of words ranked by frequency, thus they can be burdensome to interpret and hard to distinguish and understand.
+Topics are word distributions: essentially lists of words ranked by frequency, which can be burdensome to interpret and hard to distinguish and understand.
 Once we had topics for each period, we tried to associate them with a label from a list of keywords and related terms.
-We performed simple string matching between these topics and our lists, `naming' a topic if it contained that a word from our word list.
+We performed simple string matching between these topics and our lists, `naming' a topic if it contained that word or words.
 We used several different word lists for comparison.
 
-Our first word list set, \textsf{exp1}, was generated using the ontology described in Kayed et al.~\cite{5072519}.
+Our first list, \textsf{exp1}, was generated using the ontology described in Kayed et al.~\cite{5072519}.
 That paper constructs an ontology for software quality measurement using eighty source documents, including research papers and international standards.
@@
 
-Our second word list set, \textsf{exp2}, relied on the ISO quality model (ISO9126)~\cite{iso9126}. ISO9126 describes six high-level quality requirements (listed in Table \ref{tbl:wnsig}). 
+Our second list, \textsf{exp2}, relied on the ISO quality model (ISO9126)~\cite{iso9126}. ISO9126 describes six high-level quality requirements (listed in Table \ref{tbl:wnsig}). 
 %There is some debate about the significance and importance of the %terms in this model. 
@@
 They do not use external domain-specific information as we do.
-In \cite{ernst10refsq}, we describe an application similar to this to identify change in quality requirements in GNOME software projects; our approach is solely text-matching, however, and does not leverage machine learning strategies.
+In \cite{ernst10refsq}, we describe our earlier project, similar to this, to identify change in quality requirements in GNOME software projects; our approach is solely text-matching, however, and does not leverage machine learning strategies.
 
@@
 
-In conclusion we have demonstrated that static but domain specific knowledge can improve unsupervised labelling of extracted topics.
+We have demonstrated that static but domain specific knowledge can improve unsupervised labelling of extracted topics.
 Our \textsf{exp2} experiment used small accurate word bags to label topics but performed just as well as \textsf{exp3}, which used many more general terms from WordNet.

### 2010-03-06 icpc/icpc2010.tex
supervised learning

--- 2010-03-06 icpc/icpc2010.tex
@@
 Table \ref{tab:best-learn-per-tag} shows that MaxDB and MySQL have quite different results, the ROC values for reliability and functionality seem swapped between projects.
-It should be noted that for both project Bayesian techniques did the best out of a wide variety of machine learners tested.
+It should be noted that for both projects Bayesian techniques did the best out of a wide variety of machine learners tested.
 Discriminative Multinomial Naive Bayes (DMNBtext), Naive Bayes (NaiveBayes) and Multinomial Naive Bayes (NaiveBayesMultinomal) are all based on Bayes theorem and all assume, naively, that the features are independent.
 The features we used are word counts per message.
-One beneficial aspect of this result is that it suggests we can have very fast training times since Naive Bayes can be calculated in $O(N)$ for $N$ features.
+One beneficial aspect of this result is that it suggests we can have very fast training and classifying times since Naive Bayes can be calculated in $O(N)$ for $N$ features.
 %http://nlp.stanford.edu/IR-book/html/htmledition/naive-bayes-text-classification-1.html

### 2010-03-06 icpc/icpc2010.tex
multi label evaluation

@@
 The features we used are word counts per message.
-One beneficial aspect of this result is that it suggests we can have very fast training and classifying times since Naive Bayes can be calculated in $O(N)$ for $N$ features.
+One beneficial aspect of this result is that it suggests we can have very fast training and classifying  since Naive Bayes can be calculated in $O(N)$ for $N$ features.
 %http://nlp.stanford.edu/IR-book/html/htmledition/naive-bayes-text-classification-1.html
@@
 HOMER and BR acts as a hierarchy of learners, BR is flat while HOMER tries to build a deeper hierarchy in order to build a more accurate learner~\cite{mulan}.
-We feel that these classifiers performed better because they have the best micro and macro ROC scores than the other classifiers.
+We feel that these classifiers performed better than other multilabel classifiers because they have the best micro and macro ROC scores than the other classifiers although their results seem comparable to the naive Bayesian learners we used in Section \ref{sec:suplabelling}.
 

### 2010-03-06 icpc/icpc2010.tex
summary of techniques

@@
 
-Very rarely did exp2 and exp3 ever perform as well as the machine learners.
+Very rarely did \textsf{exp2} and \textsf{exp3} ever perform as well as the machine learners.
 For MaxDB reliability was slightly better detected using the static word list of \textsf{exp2}.
 In general, the machine learners and \textsf{exp3} did better than \textsf{exp2} for both MaxDB and MySQL.
-In both MySQL and MaxDB usability was better served by \textsf{exp2}.
+For both MySQL and MaxDB usability was better served by \textsf{exp2}.
 

### 2010-03-06 icpc/icpc2010.tex
visualization

@@
 For MaxDB (Figure \ref{fig:maxdb}) the largest trends were related to maintainability and portability.
-MaxDB supports numerous platforms and portability was a constant issue facing the project's develop.
-MySQL was different, maybe topics overlapped, so there were many different subsets.
+MaxDB supports numerous platforms and portability was a constant issue facing the project's development.
+MySQL was different, many topics overlapped, so there were many different subsets.
 Functionality trends were prevalent throughout its history (the largest reddish streak across the top).

### 2010-03-06 icpc/icpc2010.tex
visualization II

@@
   \includegraphics[height=.9\textheight]{small-maxdb-time-smear-plot-Equal}
-  \caption{MaxDB 7.500: Labelled topics over time, continuous topics are trends that occur across time}
+  \caption{MaxDB 7.500: Labelled topics related to non-functional software qualities plotted over time, continuous topics are trends that occur across time}
   \label{fig:maxdb}
@@
   \includegraphics[width=.99\textwidth]{small-mysql-time-smear-plot-Equal}
-  \caption{MySQL 3.23: Labelled topics over time, continuous topics are trends that occur across time}
+  \caption{MySQL 3.23: Labelled topics related to non-functional software qualities plotted over time, continuous topics are trends that occur across time}
   \label{fig:mysql}

### 2010-03-06 icpc/icpc2010.tex
comparing maxdb and mysql

@@
 In terms of non-functional requirements both projects had long running trends that focused on functionality, maintainability, and portability, yet MaxDB had more of a focus on efficiency and reliability.
-We feel that MaxDB differed from MySQL because MaxDB was being actively developed where was by half way through our history of MySQL 3.23, other versions of MySQL were being actively developed: MySQL versions 4.0 and 4.1 and eventually 5.0 and 5.1.
+We feel that MaxDB differed from MySQL because MaxDB was being actively developed where was by half way through our history of MySQL 3.23, other versions of MySQL were being actively developed: MySQL versions 4.0, 4.1, and eventually 5.0 and 5.1.
 MySQL 3.23 was being maintained rather than actively developed where as MaxDB 7.500 was being actively developed into MaxDB 7.6, and then maintained thereafter.
 
-MySQL and MaxDB's machine learners did dispatch off of some shared words.
+MySQL and MaxDB's machine learners did make decisions based off some shared words.
 Words that were used to classify topics that were shared between MySQL and MaxDB included: bug, code, compiler, database, HP UX, delete, memory, missing, problems, removed, add, added, changed, problem, and test.
-Potentially adding these words to the word bags of \textsf{exp2} and \textsf{exp3} could improve performance while still being domain specific.
+Adding these words to the word bags of \textsf{exp2} and \textsf{exp3} could improve performance while ensuring they were only domain specific.
 

### 2010-03-06 icpc/icpc2010.tex
annotation observations

@@
 
-We also found there were quite a few topics that were not non-functional requirements but were often related to them.
+We found many topics that were not non-functional requirements but were often related to them.
 For instance concurrency was mentioned often in the commit logs and was related to correctness and reliability because concurrency was troublesome.
-Configuration management and source control related changes appeared often and sometimes had topics dedicated to configuration management, these kinds of changes are slightly related to maintainability.
+Configuration management and source control related changes appeared often and sometimes there were topics dedicated to configuration management, these kinds of changes are slightly related to maintainability.
 A non-functional change that was not quality related was licensing and copyright, many changes were simply to do with updating copyrights or ensuring copyright or licenses headers were applied to files.
@@
 For instance optimizers are very common modules in database systems, both MySQL and MaxDB have optimizer modules.
-In MySQL the optimizer is mentioned but often the change is deals with correctness or another quality, yet the name of the module has the potential to confuse automated approaches.
+In MySQL the optimizer is mentioned but often the change is deals with correctness or another quality, yet the name of the module could indicate the change was always about efficiency.
 This is why we feel that perhaps a project specific word-bags are needed in order to avoid automated mistakes due to the names of entities and modules with the software project.

### 2010-03-06 icpc/icpc2010.tex
effectiveness
qualties
threats to validity

@@
 
-\subsection{Annotation Observations}
+\subsection{Annotation observations}
 
@@
 \subsection{Threats to validity}
+
+\begin{comment}
 \textbf{Construct validity} -- Focus on code commit messages, rather than mail or bug trackers. Choice of taxonomy
@@
 Our technique is dependent on development practises including meaningful commit messages, which might not be true in other domains.
+\end{comment}
+
+Our study suffers from multiple threats to validity.
+\emph{Construct validity} issues include that used only commit messages rather than mail or bug tracker messages, as well we chose our taxonomy and the data to study.
+\emph{Internal validity} issues are to do with inter-rater reliability.
+\emph{External validity} issues are that our data originated from OSS database projects and thus might not be applicable to commercially developed software, as well our analysis techniques relies on a project's commitment to meaningful commit messages.
 
@@
 
+\begin{comment}
 % Part of our effort with this project is to understand the qualitative and intentional aspects of requirements in software evolution, a notion we first discussed in \cite{ernst07icsm}.
@@
 % Not sure how do handle this
+\end{comment}
 The idea of extracting higher-level `concerns' (also known as `concepts', `aspects' or `requirements') has been attacked in two ways.
@@
 The presence of a well-maintained set of tags is obviously essential to the success of this technique.
+
 \begin{comment}
@@
 
+\begin{comment}
 % comment out or not
@@
 This supports our result that software qualities are not discussed with the same frequency across projects.
+\end{comment}
 
@@
 \section{Conclusions and future work}
-
-% No we didn't!
-% In this paper we showed that a domain specific wordnet for software % development provides better accuracy for labelling topic models than % unsupervised approaches.
-%
-
-%* Demonstrated that domain specific knowledge is valuable
-%* Demonstrated the feasibility of static approaches and supervised %approaches
-%* Showed that non functional requirement pervade much many of the %topics that are extracted
 

### 2010-03-06 icpc/icpc2010.tex

@@
 
-%To have meaningful discussions about how software development is progressing, we must devise appropriate labels for our work topics, in contrast to the quote above.
+% To have meaningful discussions about how software development is progressing, we must devise appropriate labels for our work topics, in contrast to the quote above.
 We can then categorize development artifacts as belonging to one or more of these topics.
 
-% This correlates issues with topics
-As development of a software project progresses, developers must concern themselves with a wide variety of issues --- or ``topics'' --- that relate to the problem domain, the processes and tools used, the development artifacts themselves, etc.
+Topics arise from the wide variety of issues which occur during a software project's lifecycle.
+These topics can relate to, among others, the problem domain, the processes and tools used, or the development artifacts themselves.
 The set of development topics for a given project can sometimes be extracted automatically by analyzing artifacts within software repositories, such as change-log comments that developers create when committing revisions to the project's source control system.
@@
 
-%Topics arise from the wide variety of issues which occurs during a software project's lifecycle.
-These topics can relate to, among others, the problem domain, the processes and tools used, or the development artifacts themselves.
+%As development of a software project progresses, developers must concern themselves with a wide variety of issues --- or ``topics'' --- that relate to the problem domain, the processes and tools used, the development artifacts themselves, etc.
 The set of development topics for a given project can sometimes be extracted automatically by analyzing artifacts within software repositories, such as change-log comments that developers create when committing revisions to the project's source control system.
@@
 % check
-\emph{Area of ROC Curve} is the area under the Receiver Operating Characteristic curve, we will refer to this as \emph{ROC} although it is sometimes referred to as \emph{AUC}.
-ROC values reflect a score, similar to school letter-grades (an \emph{A} is $0.9$, a \emph{C} is $0.6$), reflecting how well a particular learner performed for the given data.
+\emph{Area of ROC Curve} is the area under the Receiver Operating Characteristic (ROC) curve, sometimes referred to as \emph{AUC}.
+ROC values reflect a score, similar to school letter-grades (A is 0.9, C is 0.6), reflecting how well a particular learner performed for the given data.
 A ROC result of 0.5 would be equivalent to a random learner (randomly classifying data).
@@
 
-Topics are word distributions: essentially lists of words ranked by frequency, thus they can be burdensome to interpret and hard to distinguish and understand.
+Topics are word distributions: essentially lists of words ranked by frequency, which can be burdensome to interpret and hard to distinguish and understand.
 Once we had topics for each period, we tried to associate them with a label from a list of keywords and related terms.
-We performed simple string matching between these topics and our lists, `naming' a topic if it contained that a word from our word list.
+We performed simple string matching between these topics and our lists, `naming' a topic if it contained that word or words.
 We used several different word lists for comparison.
@@
 
-In conclusion we have demonstrated that static but domain specific knowledge can improve unsupervised labelling of extracted topics.
+In conclusion, we have demonstrated that static but domain specific knowledge can improve unsupervised labelling of extracted topics.
 Our \textsf{exp2} experiment used small accurate word bags to label topics but performed just as well as \textsf{exp3}, which used many more general terms from WordNet.

### 2010-03-06 icpc/icpc2010.tex
removed merge comments in intro

@@
 These topics can relate to, among others, the problem domain, the processes and tools used, or the development artifacts themselves.
-The set of development topics for a given project can sometimes be extracted automatically by analyzing artifacts within software repositories, such as change-log comments that developers create when committing revisions to the project's source control system.
-Our previous work dealt with topic trends, which are topics that recur over time~\cite{Hindle09ICSM}.
-We observed that topic trends were often non-functional requirements.
-
-%As development of a software project progresses, developers must concern themselves with a wide variety of issues --- or ``topics'' --- that relate to the problem domain, the processes and tools used, the development artifacts themselves, etc.
 The set of development topics for a given project can sometimes be extracted automatically by analyzing artifacts within software repositories, such as change-log comments that developers create when committing revisions to the project's source control system.
@@
 While the topic models themselves are generated automatically, what to make of them is less clear.
-For example, in our previous work~\cite{Hindle09ICSM}, as well as in \cite{Baldi2008}, topics are named manually: human experts read the highest-frequency members of a topic and assign a keyword accordingly.
+For example, in our previous work~\cite{Hindle09ICSM}, as well as in Baldi et al.~\cite{Baldi2008}, topics are named manually: human experts read the highest-frequency members of a topic and assign a keyword accordingly.
 E.g., for the word list \emph{``listener change remove add fire''}, Baldi et al.
@@
 For example, the term ``redundancy'' is one most would agree is relevant to discussion of reliability, but is not in the standard.
-Therefore we took the words from the taxonomy and expanded them.
+We therefore took the words from the taxonomy and expanded them.
 
@@
 
-To construct our expanded word-lists, we used WordNet~\cite{Fellbaum1998}, an English-language ``lexical database'' that contains semantic relations between words, including meronymy and synonymy.
+To construct these expanded word-lists, we used WordNet~\cite{Fellbaum1998}, an English-language ``lexical database'' that contains semantic relations between words, including meronymy and synonymy.
 We then added Boehm’s 1976 software quality model \cite{Boehm+:1976:ICSE}, and classified his eleven ‘ilities’ into their respective ISO9126 qualities.
@@
 For example, for the MaxDB topic consisting of a message ``exit() only used in non NPTL LINUX Versions'', we tagged that topic \emph{portability}.
-We compared against this data-set but we also used it for our supervised learning based topic classification.
+We compared against this data-set, but we also used it for our supervised learning based topic classification.
 %inter-rater reliability
@@
 
-Very rarely did \textsf{exp2} and \textsf{exp3} ever perform as well as the machine learners.
+Very rarely did exp2 and exp3 ever perform as well as the machine learners.
 For MaxDB reliability was slightly better detected using the static word list of \textsf{exp2}.
 In general, the machine learners and \textsf{exp3} did better than \textsf{exp2} for both MaxDB and MySQL.
-For both MySQL and MaxDB usability was better served by \textsf{exp2}.
+In both MySQL and MaxDB usability was better served by \textsf{exp2}.
 
@@
 For MaxDB (Figure \ref{fig:maxdb}) the largest trends were related to maintainability and portability.
-MaxDB supports numerous platforms and portability was a constant issue facing the project's development.
-MySQL was different, many topics overlapped, so there were many different subsets.
+MaxDB supports numerous platforms and portability was a constant issue facing the project's develop.
+MySQL was different, maybe topics overlapped, so there were many different subsets.
 Functionality trends were prevalent throughout its history (the largest reddish streak across the top).
@@
   \includegraphics[height=.9\textheight]{small-maxdb-time-smear-plot-Equal}
-  \caption{MaxDB 7.500: Labelled topics related to non-functional software qualities plotted over time, continuous topics are trends that occur across time}
+  \caption{MaxDB 7.500: Labelled topics over time, continuous topics are trends that occur across time}
   \label{fig:maxdb}
@@
   \includegraphics[width=.99\textwidth]{small-mysql-time-smear-plot-Equal}
-  \caption{MySQL 3.23: Labelled topics related to non-functional software qualities plotted over time, continuous topics are trends that occur across time}
+  \caption{MySQL 3.23: Labelled topics over time, continuous topics are trends that occur across time}
   \label{fig:mysql}
@@
 In terms of non-functional requirements both projects had long running trends that focused on functionality, maintainability, and portability, yet MaxDB had more of a focus on efficiency and reliability.
-We feel that MaxDB differed from MySQL because MaxDB was being actively developed where was by half way through our history of MySQL 3.23, other versions of MySQL were being actively developed: MySQL versions 4.0, 4.1, and eventually 5.0 and 5.1.
+We feel that MaxDB differed from MySQL because MaxDB was being actively developed where was by half way through our history of MySQL 3.23, other versions of MySQL were being actively developed: MySQL versions 4.0 and 4.1 and eventually 5.0 and 5.1.
 MySQL 3.23 was being maintained rather than actively developed where as MaxDB 7.500 was being actively developed into MaxDB 7.6, and then maintained thereafter.
 
-MySQL and MaxDB's machine learners did make decisions based off some shared words.
+MySQL and MaxDB's machine learners did dispatch off of some shared words.
 Words that were used to classify topics that were shared between MySQL and MaxDB included: bug, code, compiler, database, HP UX, delete, memory, missing, problems, removed, add, added, changed, problem, and test.
-Adding these words to the word bags of \textsf{exp2} and \textsf{exp3} could improve performance while ensuring they were only domain specific.
+Potentially adding these words to the word bags of \textsf{exp2} and \textsf{exp3} could improve performance while still being domain specific.
 
@@
 
-\subsection{Annotation observations}
+\subsection{Annotation Observations}
 
-We found many topics that were not non-functional requirements but were often related to them.
+We also found there were quite a few topics that were not non-functional requirements but were often related to them.
 For instance concurrency was mentioned often in the commit logs and was related to correctness and reliability because concurrency was troublesome.
-Configuration management and source control related changes appeared often and sometimes there were topics dedicated to configuration management, these kinds of changes are slightly related to maintainability.
+Configuration management and source control related changes appeared often and sometimes had topics dedicated to configuration management, these kinds of changes are slightly related to maintainability.
 A non-functional change that was not quality related was licensing and copyright, many changes were simply to do with updating copyrights or ensuring copyright or licenses headers were applied to files.
@@
 For instance optimizers are very common modules in database systems, both MySQL and MaxDB have optimizer modules.
-In MySQL the optimizer is mentioned but often the change is deals with correctness or another quality, yet the name of the module could indicate the change was always about efficiency.
+In MySQL the optimizer is mentioned but often the change is deals with correctness or another quality, yet the name of the module has the potential to confuse automated approaches.
 This is why we feel that perhaps a project specific word-bags are needed in order to avoid automated mistakes due to the names of entities and modules with the software project.
@@
 \subsection{Threats to validity}
-
-\begin{comment}
 \textbf{Construct validity} -- Focus on code commit messages, rather than mail or bug trackers. Choice of taxonomy
@@
 Our technique is dependent on development practises including meaningful commit messages, which might not be true in other domains.
-\end{comment}
-
-Our study suffers from multiple threats to validity.
-\emph{Construct validity} issues include that used only commit messages rather than mail or bug tracker messages, as well we chose our taxonomy and the data to study.
-\emph{Internal validity} issues are to do with inter-rater reliability.
-\emph{External validity} issues are that our data originated from OSS database projects and thus might not be applicable to commercially developed software, as well our analysis techniques relies on a project's commitment to meaningful commit messages.
 
@@
 
-\begin{comment}
 % Part of our effort with this project is to understand the qualitative and intentional aspects of requirements in software evolution, a notion we first discussed in \cite{ernst07icsm}.
@@
 % Not sure how do handle this
-\end{comment}
 The idea of extracting higher-level `concerns' (also known as `concepts', `aspects' or `requirements') has been attacked in two ways.
@@
 The presence of a well-maintained set of tags is obviously essential to the success of this technique.
-
 \begin{comment}
@@
 
-\begin{comment}
 % comment out or not
@@
 This supports our result that software qualities are not discussed with the same frequency across projects.
-\end{comment}
 
@@
 
-In conclusion, we have demonstrated that static but domain specific knowledge can improve unsupervised labelling of extracted topics.
+% No we didn't!
+% In this paper we showed that a domain specific wordnet for software % development provides better accuracy for labelling topic models than % unsupervised approaches.
+%
+
+%* Demonstrated that domain specific knowledge is valuable
+%* Demonstrated the feasibility of static approaches and supervised %approaches
+%* Showed that non functional requirement pervade much many of the %topics that are extracted
+
+We have demonstrated that static but domain specific knowledge can improve unsupervised labelling of extracted topics.
 Our \textsf{exp2} experiment used small accurate word bags to label topics but performed just as well as \textsf{exp3}, which used many more general terms from WordNet.

### 2010-03-06 icpc/icpc2010.tex
spellcheck

@@
 
-Topics arise from the wide variety of issues which occur during a software project's lifecycle.
+Topics arise from the wide variety of issues which occur during a software project's life-cycle.
 These topics can relate to, among others, the problem domain, the processes and tools used, or the development artifacts themselves.
@@
 
-We took our annotated dataset and applied supervised learners to it.
-Because our dataset was of word counts we expected Bayesian techniques, often used in spam filtering, to perform well.
+We took our annotated data-set and applied supervised learners to it.
+Because our data-set was of word counts we expected Bayesian techniques, often used in spam filtering, to perform well.
 We also tried other learners that WEKA~\cite{weka09} provides: rule learners, tree learners, vector space learners, and support vector machines.
@@
 We have presented the results of Mulan's multi-label learners in Table \ref{tab:mulan}.
-Binary Relevance (BR), Calibrated Label Ranking (CLR) and Hierarchy Of Multilabel classifiERs (HOMER), performed the best.
+Binary Relevance (BR), Calibrated Label Ranking (CLR) and Hierarchy Of Multi-label classifiERs (HOMER), performed the best.
 HOMER and BR acts as a hierarchy of learners, BR is flat while HOMER tries to build a deeper hierarchy in order to build a more accurate learner~\cite{mulan}.
-We feel that these classifiers performed better than other multilabel classifiers because they have the best micro and macro ROC scores than the other classifiers although their results seem comparable to the naive Bayesian learners we used in Section \ref{sec:suplabelling}.
+We feel that these classifiers performed better than other multi-label classifiers because they have the best micro and macro ROC scores than the other classifiers although their results seem comparable to the naive Bayesian learners we used in Section \ref{sec:suplabelling}.
 

### 2010-03-06 icpc/icpc2010.tex
wording fixes
answered the research questions

@@
 
+\paragraph{Do label frequencies change over time?} Yes, MySQL's label
+frequencies decreased as it got older. Usability and reliability
+labels became more and more infrequent as it matured. Maintainability topics
+became more prevalent as MaxDB matured.
+
+\paragraph{Do the different projects differ in their relative topic
+  interest?} Yes they do, MySQL 3.23 had more proportionally more
+functionality labelled topics while MaxDB had proportionally more
+efficiency and portability related topics.
+
 \begin{comment}

### 2010-03-06 icpc/icpc2010.tex
still 10 pages and we answered those questions!
wording fixes

--- 2010-03-06 icpc/icpc2010.tex
@@
 We achieved similar results using the Matthew's correlation coefficient (used to measure efficacy where classes are of different sizes) and ROC.
-
-Based on these results we find that reliability and usability worked well for MaxDB in \textsf{exp2} and better in \textsf{exp3}.
-MySQL reasonable results within \textsf{exp2} were for reliability and efficiency.
-MySQL's results for efficiency did not improve in \textsf{exp3} but other qualities such as functionality did improve.
-If a \emph{C} grade performance has a ROC value of $0.6$ then most of these tests scored \emph{C} or less, but our results were still better than random chance.
-
 \begin{table*}[h]
@@
 
+Based on these results we find that reliability and usability worked well for MaxDB in \textsf{exp2} and better in \textsf{exp3}.
+MySQL had reasonable results within \textsf{exp2} for reliability and efficiency.
+MySQL's results for efficiency did not improve in \textsf{exp3} but other qualities such as functionality did improve.
+If a \emph{C} grade performance has a ROC value of $0.6$ then most of these tests scored \emph{C} or less, but our results were still better than random chance.
+
 %\subsection{Classifying topics}
@@
 We considered the best learner for a label to be the one which had the highest ROC value for that label.
-Table \ref{tab:best-learn-per-tag} shows the ROC value, the accuracy of a ZeroR learner ( a learner which just chooses the largest category all of the time), the difference in accuracy between the best learner and ZeroR, and the F-measure for that particular learner.
-The ZeroR difference is often negative, but in particular for labels which are not as common this can be expected because any miscategorization will hurt accuracy, this is why the F1 (F-measure) and the ROC values are useful as they can better present performance on labels which are not applicable to the majority of samples.
-
-Table \ref{tab:best-learn-per-tag} shows that MaxDB and MySQL have quite different results, the ROC values for reliability and functionality seem swapped between projects.
+Table \ref{tab:best-learn-per-tag} uses the ZeroR learner as a baseline, since it naively chooses the largest category all of the time.
+The ZeroR difference is often negative.
+For labels which are not as common, this can be expected because any miscategorization will hurt accuracy.
+This is why the F1 (F-measure) and the ROC values are useful, as they can better present performance on labels which are not applicable to the majority of samples.
+
+Table \ref{tab:best-learn-per-tag} shows that MaxDB and MySQL have quite different results, as the ROC values for reliability and functionality seem swapped between projects.
 It should be noted that for both projects Bayesian techniques did the best out of a wide variety of machine learners tested.
-Discriminative Multinomial Naive Bayes (DMNBtext), Naive Bayes (NaiveBayes) and Multinomial Naive Bayes (NaiveBayesMultinomal) are all based on Bayes theorem and all assume, naively, that the features are independent.
+Discriminative Multinomial Naive Bayes (DMNBtext), Naive Bayes (NaiveBayes) and Multinomial Naive Bayes (NaiveBayesMultinomal) are all based on Bayes's theorem and all assume, naively, that the features are independent.
 The features we used are word counts per message.
@@
 
-It seems like the smaller the label the harder it is to get accurate results.
-Although these results seem better than our previous word bag results of \textsf{exp2} and \textsf{exp3} because the ROC values are sufficiently higher in most cases other than MaxDB reliability and MySQL efficiency.
-
-The limitation of the approach we took here is that we only checked independently for each label.
-Labels could be correlated with each other, as well as we did not evaluate how well the learners performed together.
+The smaller the label the harder it is to get accurate results.
+Nevertheless, these results are better than our previous word bag results of \textsf{exp2} and \textsf{exp3}, because the ROC values are sufficiently higher in most cases (other than MaxDB reliability and MySQL efficiency).
+
+The limitation of the approach we took here is that we assume labels are independent; however, labels could be correlated with each other.
+We also did not evaluate how well the learners performed together.
 
@@
 \end{tabular}
-\caption{Per label, per project, the best learner for that label}
-\label{tab:best-learner-per-tag}
+\caption{Per label, per project, the best learner for that label.
+ROC value rates learner performance, compared with the ZeroR learner ( a learner which just chooses the largest category all of the time).
+F1 is the F-measure for that particular learner.}
+\label{tab:best-learn-per-tag}
 \end{table*}
@@
 
-We applied the Mulan~\cite{mulan} library for Multi-Label learning to our data-set because we feel that topics can have more like one label, much like how a revision can have more than one topic.
+We applied the Mulan~\cite{mulan} library for Multi-Label learning to our data-set because intuitively, topics can have more than one label, much like how a particular source code revision can have more than one topic.
 Multi-label learning is more than just classifying entities with more than one label.
 It also includes methods for determining the performance of such techniques.
-Consider that the problem has changed, instead of looking at the precision and recall of applying one label, we have to rank multiple labels at once.
-We have to check if the full subset of labels was applied and then how much of that subset was applied.
-
-Another aspect of multi-label learning is the micro versus macro measurements.
+The problem framed in the learners above has changed; instead of looking at the precision and recall of applying one label, we rank multiple labels at once.
+We must check if the full subset of labels was applied, and then how much of that subset was applied.
+
+Another aspect of multi-label learning are micro versus macro measurements.
 Macro measurements are aggregated at a class or label level.
@@
 So a macro-ROC measurement is the average ROC over the ROC values for all labels, where a micro-ROC is the average ROC over all examples that were classified.
-Unfortunately for MaxDB the macro-ROC values are undefined because of poor performance of one of the labels.
+Unfortunately for MaxDB, the macro-ROC values are undefined because of poor performance of one of the labels.
 
@@
 Binary Relevance (BR), Calibrated Label Ranking (CLR) and Hierarchy Of Multi-label classifiERs (HOMER), performed the best.
-HOMER and BR acts as a hierarchy of learners, BR is flat while HOMER tries to build a deeper hierarchy in order to build a more accurate learner~\cite{mulan}.
-We feel that these classifiers performed better than other multi-label classifiers because they have the best micro and macro ROC scores than the other classifiers although their results seem comparable to the naive Bayesian learners we used in Section \ref{sec:suplabelling}.
+HOMER and BR act as a hierarchy of learners: BR is flat, while HOMER tries to build a deeper hierarchy to build a more accurate learner~\cite{mulan}.
+These classifiers performed better than other multi-label classifiers.
+They have the best micro and macro ROC scores, although their results seem comparable to the naive Bayesian learners we used in Section \ref{sec:suplabelling}.
 
@@
 
-Very rarely did \textsf{exp2} and \textsf{exp3} ever perform as well as the machine learners.
-For MaxDB reliability was slightly better detected using the static word list of \textsf{exp2}.
+Very rarely did \textsf{exp2} and \textsf{exp3} (naive word matching) ever perform as well as the machine learners.
+For MaxDB, reliability was slightly better detected using the static word list of \textsf{exp2}.
 In general, the machine learners and \textsf{exp3} did better than \textsf{exp2} for both MaxDB and MySQL.
 For both MySQL and MaxDB usability was better served by \textsf{exp2}.
+Usability was a very infrequent label, however.
 

### 2010-03-06 icpc/icpc2010.tex
Section reference

@@
 We achieved similar results using the Matthew's correlation coefficient (used to measure efficacy where classes are of different sizes) and ROC.
-\begin{table*}[h]
+
+\begin{table*}%[h]
 	\centering

### 2010-03-06 icpc/icpc2010.tex
De-font specc'd

--- 2010-03-06 icpc/icpc2010.tex
@@
 \documentclass{acm_proc_article-sp}
-\usepackage{fontspec}
-\usepackage{xunicode}
-\usepackage{xltxtra}
+%\usepackage{fontspec}
+%\usepackage{xunicode}
+%\usepackage{xltxtra}
 \usepackage{booktabs}
-\defaultfontfeatures{Mapping=tex-text}
-\setmainfont{Times New Roman}
+%\defaultfontfeatures{Mapping=tex-text}
+%\setmainfont{Times New Roman}
 %\usepackage{amsthm}
@@
 %\usepackage{algpseudocode}
-\usepackage[pdfstartview=FitH,xetex,bookmarksopenlevel=3,bookmarks=true]{hyperref} %pdfpagescrop={92 112 523 778},
+%\usepackage[pdfstartview=FitH,xetex,bookmarksopenlevel=3,bookmarks=true]{hyperref}
+% pdfpagescrop={92 112 523 778},
+\usepackage[pdfstartview=FitH,bookmarksopenlevel=3,bookmarks=true]{hyperref} %pdfpagescrop={92 112 523 778},
 %\usepackage{algorithm}%,algorithmic-fix}
@@
 
-Finally, using this data, we evaluated two research questions:
+Finally, using this data, we evaluated two research questions (see
+Section \ref{sec:comparison}):
 \begin{enumerate}
@@
 
-\begin{table*}[h]
+\begin{table*}%[h]
 	\centering
@@
 
+\label{sec:comparison}
 %\XXX{Not sure there's enough here to do that!}

### 2010-03-06 icpc/icpc2010.tex
updates to wording

--- 2010-03-06 icpc/icpc2010.tex
@@
 \defaultfontfeatures{Mapping=tex-text}
-\setmainfont{Times New Roman}
+% \setmainfont{Times New Roman}
 %\usepackage{amsthm}
@@
 We achieved similar results using the Matthew's correlation coefficient (used to measure efficacy where classes are of different sizes) and ROC.
-\begin{table*}[h]
+
+\begin{table*}[ht]
 	\centering
@@
 
-\begin{table*}
+\begin{table*}[h]
 \centering
@@
 MaxDB supports numerous platforms and portability was a constant issue facing the project's development.
-MySQL was different, many topics overlapped, so there were many different subsets.
+MySQL was different, with many topics overlapping, so there were many different subsets.
 Functionality trends were prevalent throughout its history (the largest reddish streak across the top).
-Two combination  tags of ``functionality portability'' (orange) and ``maintainability portability'' (teal) streaked across most of the history MySQL.
-Since this was MySQL 3.23 and was a stable branch we expect that issues dealings with portability and maintenance would be the primary concerns of this branch: developers probably wanted it to work with current systems and thus had to update it.
+Two combination  tags of ``functionality portability'' (orange) and ``maintainability portability'' (teal) streaked across most of the history of MySQL.
+Since this was MySQL 3.23, a stable branch, we expect that issues dealing with portability and maintenance would be the primary concerns of this branch: developers probably wanted it to work with current systems and thus had to update it.
 
@@
 In terms of non-functional requirements both projects had long running trends that focused on functionality, maintainability, and portability, yet MaxDB had more of a focus on efficiency and reliability.
-We feel that MaxDB differed from MySQL because MaxDB was being actively developed where was by half way through our history of MySQL 3.23, other versions of MySQL were being actively developed: MySQL versions 4.0, 4.1, and eventually 5.0 and 5.1.
-MySQL 3.23 was being maintained rather than actively developed where as MaxDB 7.500 was being actively developed into MaxDB 7.6, and then maintained thereafter.
+MaxDB differed from MySQL since MaxDB was being actively developed, whereas halfway through our history of MySQL 3.23, other versions of MySQL were being actively developed: MySQL versions 4.0, 4.1, and eventually 5.0 and 5.1.
+In other words, MySQL 3.23 was being maintained rather than actively developed, whereas MaxDB 7.500 was being actively developed into MaxDB 7.6, and then maintained thereafter.
 
 MySQL and MaxDB's machine learners did make decisions based off some shared words.
-Words that were used to classify topics that were shared between MySQL and MaxDB included: bug, code, compiler, database, HP UX, delete, memory, missing, problems, removed, add, added, changed, problem, and test.
+Words that were used to classify topics that were shared between MySQL and MaxDB included: \textsf{bug, code, compiler, database, HP UX, delete, memory, missing, problems, removed, add, added, changed, problem, and test}.
 Adding these words to the word bags of \textsf{exp2} and \textsf{exp3} could improve performance while ensuring they were only domain specific.
+
+With respect to the questions raised in section \ref{sec:suplearn}:
+\textbf{Do label frequencies change over time?} -- Yes, MySQL's label
+frequencies decreased as it got older. Usability and reliability
+labels became more and more infrequent as it matured. Maintainability topics
+became more prevalent as MaxDB matured.
+
+\textbf{Do the different projects differ in their relative topic
+  interest?} -- Yes. MySQL 3.23 had proportionally more
+functionality labelled topics, while MaxDB had proportionally more
+efficiency and portability related topics.
 
@@
 
-We found many topics that were not non-functional requirements but were often related to them.
-For instance concurrency was mentioned often in the commit logs and was related to correctness and reliability because concurrency was troublesome.
-Configuration management and source control related changes appeared often and sometimes there were topics dedicated to configuration management, these kinds of changes are slightly related to maintainability.
-A non-functional change that was not quality related was licensing and copyright, many changes were simply to do with updating copyrights or ensuring copyright or licenses headers were applied to files.
-
-We noticed that sometimes the names of modules would conflict with words related to other non-functional requirements.
-For instance optimizers are very common modules in database systems, both MySQL and MaxDB have optimizer modules.
-In MySQL the optimizer is mentioned but often the change is deals with correctness or another quality, yet the name of the module could indicate the change was always about efficiency.
-This is why we feel that perhaps a project specific word-bags are needed in order to avoid automated mistakes due to the names of entities and modules with the software project.
+We found many topics that were not non-functional requirements (NFRs) but were often related to them.
+For instance, concurrency was mentioned often in the commit logs and was related to correctness and reliability, because concurrency was troublesome.
+Configuration management and source control related changes appeared often and sometimes there were topics dedicated to configuration management.
+These kinds of changes are slightly related to maintainability.
+A non-functional change that was not quality-related was licensing and copyright; many changes were simply to do with updating copyrights or ensuring copyright or licenses headers were applied to files.
+
+We noticed that occasionally the names of modules would conflict with words related to other non-functional requirements.
+For instance, optimizers are very common modules in database systems: both MySQL and MaxDB have optimizer modules.
+In MySQL the optimizer is mentioned but often the change deals with correctness or another quality.
+Despite this difference, the name of the module could fool our learners into believing the change was always about efficiency.
+Perhaps a project specific word-bag is needed in order to avoid automated mistakes due to the names of entities and modules with the software project.
 
@@
 Both Naive Bayesian learners and the word-list approaches were computationally efficient.
-Thus these results are promising because the results are accurate enough to be useful and cheap enough to execute to be feasible as an automated or semi-automated method of labelling topics by their software qualities.
+These results are promising, because the results are accurate enough to be useful, while still cheap enough to execute to be feasible as an automated or semi-automated method of labelling topics by their software qualities.
 
@@
 Our study suffers from multiple threats to validity.
-\emph{Construct validity} issues include that used only commit messages rather than mail or bug tracker messages, as well we chose our taxonomy and the data to study.
+\emph{Construct validity} issues include that we used only commit messages rather than mail or bug tracker messages.
+We also chose our taxonomy and the data to study.
 \emph{Internal validity} issues are to do with inter-rater reliability.
-\emph{External validity} issues are that our data originated from OSS database projects and thus might not be applicable to commercially developed software, as well our analysis techniques relies on a project's commitment to meaningful commit messages.
+\emph{External validity} issues are that our data originated from OSS database projects and thus might not be applicable to commercially developed software.
+Furthermore, our analysis techniques rely on a project's use of meaningful commit messages.
 
@@
 
-In conclusion, we have demonstrated that static but domain specific knowledge can improve unsupervised labelling of extracted topics.
+We demonstrated that static but domain-specific knowledge can improve unsupervised labelling of extracted topics.
 Our \textsf{exp2} experiment used small accurate word bags to label topics but performed just as well as \textsf{exp3}, which used many more general terms from WordNet.
@@
 
-Our manual inspection and annotation of the topics extracted from MySQL and MaxDB revealed that many of the extracted dealt with non-functional requirements and these topics were spread across the entire history of a project.
+Our manual inspection and annotation of the topics extracted from MySQL and MaxDB revealed that many of the extracted topics dealt with non-functional requirements, and these topics were spread across the entire history of a project.
 In the cases of MaxDB and MySQL, portability was a constant maintenance concern and was prevalent throughout the entire lifetime of the projects.
 
-Thus we have shown that non-functional requirements are often trending topics, that non-functional requirements are quite common in developer topics, and that there are efficient methods of semi-automating and automating topic labelling.
+We showed that non-functional requirements are often trending topics, that non-functional requirements are quite common in developer topics, and that there are efficient methods of semi-automating and automating topic labelling.
 
 There are many avenues of further investigation.
-We want to investigate developer attitudes related to these labels, when we label a topic, was the developer expressing positive or negative qualities about that label.
+We want to investigate developer attitudes related to these labels: i.e., when we label a topic, was the developer expressing positive or negative qualities about that label?
 It is difficult to map abstract qualities to particular messages.
-If the message is about a particular bug fix, do we map that to reliability or maintainability?
-Is it a performance bug or a usability bug?
-We ask is a ``quality'' discussion about more than just corrective maintenance?
-In this study we take the position that any type of maintenance has to do with software quality.
+Is a ``quality'' discussion about more than just corrective maintenance?
+%In this study we take the position that any type of maintenance has to do with software quality.
 

### 2010-03-06 icpc/icpc2010.tex
Desperate attempt to make it all work

--- 2010-03-06 icpc/icpc2010.tex
@@
 
+\newcommand{\shrink}{}
+
 \section{Introduction}
@@
 
+\shrink
 \section{Background}
@@
 
+\shrink
 \subsection{Definitions}
@@
 
+\shrink
 \subsection{Topic and Concept Extraction}
@@
 
+\shrink
 \subsection{Supervised learning}
@@
 
+\shrink 
 \section{Methodology}
@@
 
+\shrink
 \subsection{Generating the data}
@@
 
+\shrink
 \subsection{Associating labels}
@@
 
+\shrink
 \section{Observations and evaluation}
 
+\shrink
 \subsection{Word list similarity}
@@
 
-In general, this approach did not work out well as common labels dominated the less common labels.
-The related words for \emph{correctness}, for example, tended to be too lengthy and non-specific.
-Table \ref{tbl:wordlist} lists results.
+In general, this approach did not work out well as common labels
+dominated the less common labels. The related words for
+\emph{correctness}, for example, tended to be too lengthy and
+non-specific. 
+%Table \ref{tbl:wordlist} lists results. 
 A \emph{named   topic} is a topic with a matching label.
@@
 
+\begin{comment}
 %\XXX{this table should have mysql}
@@
 \end{table}
-
+\end{comment}
 %This is the counts
@@
 
+\shrink
 \subsection{Analysis of the unsupervised labelling}
@@
 %\subsection{Classifying topics}
+\shrink
 \subsection{Analysis of the supervised labelling}
@@
 
+\shrink
 \subsection{Applying multiple labels to topics}
@@
 
+\shrink
 \subsection{Summary of techniques}
@@
 
+\shrink
 \subsection{Visualization}
@@
   \centering
-  \includegraphics[height=.9\textheight]{small-maxdb-time-smear-plot-Equal}
+  \includegraphics[height=.55\textheight]{small-maxdb-time-smear-plot-Equal}
   \caption{MaxDB 7.500: Labelled topics related to non-functional software qualities plotted over time, continuous topics are trends that occur across time}
@@
 
-\begin{figure*}[h]
+\begin{figure*}%[h]
   \centering
-  \includegraphics[width=.99\textwidth]{small-mysql-time-smear-plot-Equal}
+  \includegraphics[width=.97\textwidth]{small-mysql-time-smear-plot-Equal}
   \caption{MySQL 3.23: Labelled topics related to non-functional software qualities plotted over time, continuous topics are trends that occur across time}
@@
 
+\shrink
 \subsection{Comparing MaxDB and MySQL}
@@
 
-\paragraph{Do label frequencies change over time?} Yes, MySQL's label
+\emph{Do label frequencies change over time?} Yes, MySQL's label
 frequencies decreased as it got older. Usability and reliability
@@
 
-\paragraph{Do the different projects differ in their relative topic
+\emph{Do the different projects differ in their relative topic
   interest?} Yes they do, MySQL 3.23 had more proportionally more
@@
 
+\shrink
 \subsection{Annotation observations}
@@
 
+\shrink
 \subsection{Effectiveness}
@@
 
+\shrink
 \subsection{Threats to validity}
@@
 
+\shrink
 \section{Related work}
@@
 
+\shrink
 \section{Conclusions and future work}
@@
 
-\section{Acknowledgements and appendix}
+%\section{Acknowledgements and appendix}
 Our data and scripts are available at \url{http://softwareprocess.es/nomen/}

### 2010-03-06 icpc/icpc2010.tex
slash o for Boegh
added the tables bank

--- 2010-03-06 icpc/icpc2010.tex
@@
 \documentclass{acm_proc_article-sp}
-\usepackage{fontspec}
-\usepackage{xunicode}
-\usepackage{xltxtra}
+%\usepackage{fontspec}
+%\usepackage{xunicode}
+%\usepackage{xltxtra}
 \usepackage{booktabs}
-\defaultfontfeatures{Mapping=tex-text}
-% \setmainfont{Times New Roman}
+%\defaultfontfeatures{Mapping=tex-text}
+%\setmainfont{Times New Roman}
 %\usepackage{amsthm}
@@
 %However, it is ``an international standard and thus provides an %internationally accepted terminology for software quality \cite[p.
-%58]{Bøegh2008},'' which is sufficient for the purposes of this %research.
-ISO9126 is ``an international standard and thus provides an internationally accepted terminology for software quality \cite[p. 58]{Bøegh2008},'' that is sufficient for the purposes of this research.
+%58]{Boegh2008},'' which is sufficient for the purposes of this %research.
+ISO9126 is ``an international standard and thus provides an internationally accepted terminology for software quality \cite[p. 58]{Boegh2008},'' that is sufficient for the purposes of this research.
 However, the terms extracted from ISO9126 may not capture all words associated with the labels.
@@
 
-In general, this approach did not work out well as common labels dominated the less common labels.
-The related words for \emph{correctness}, for example, tended to be too lengthy and non-specific.
+In general, this approach did not work out well as common labels
+dominated the less common labels. The related words for
+\emph{correctness}, for example, tended to be too lengthy and
+non-specific. 
 Table \ref{tbl:wordlist} lists results. 
@@
 
+%\begin{comment}
 %\XXX{this table should have mysql}
@@
 \end{table}
-
+%\end{comment}
 %This is the counts
@@
 
-example Hamming loss & 0.24 & 0.19 & 0.18 & 0.19 & 0.23 & 0.19 & 0.20 & 0.17 \\
+% example Hamming loss & 0.24 & 0.19 & 0.18 & 0.19 & 0.23 & 0.19 & 0.20 & 0.17 \\
 
@@
 
-rank Avg. Precision & 0.76 & 0.77 & 0.69 & 0.72 & 0.53 & 0.74 & 0.76 & 0.73 \\
-
-rank Coverage & 1.51 & 1.47 & 1.94 & 1.72 & 2.99 & 1.57 & 1.54 & 2.00 \\
-
-rank One-error & 0.40 & 0.39 & 0.47 & 0.44 & 0.63 & 0.43 & 0.40 & 0.36 \\
-
-rank Ranking Loss & 0.19 & 0.18 & 0.27 & 0.22 & 0.52 & 0.20 & 0.19 & 0.28 \\
+% rank Avg. Precision & 0.76 & 0.77 & 0.69 & 0.72 & 0.53 & 0.74 & 0.76 & 0.73 \\
+
+% rank Coverage & 1.51 & 1.47 & 1.94 & 1.72 & 2.99 & 1.57 & 1.54 & 2.00 \\
+
+% rank One-error & 0.40 & 0.39 & 0.47 & 0.44 & 0.63 & 0.43 & 0.40 & 0.36 \\
+
+% rank Ranking Loss & 0.19 & 0.18 & 0.27 & 0.22 & 0.52 & 0.20 & 0.19 & 0.28 \\
 
@@
 
-example Accuracy & 0.45 & 0.40 & 0.50 &           0.49  &  0.47  &  0.51 \\
-
-example F1 & 0.56 & 0.48 & 0.59 &           0.55  &  0.50  &  0.53 \\
-
-example Hamming Loss & 0.24 & 0.19 & 0.18 &        0.19  &  0.14  &  0.14 \\
-
-example Precision & 0.50 & 0.51 & 0.60 &          0.50  &  0.49  &  0.53 \\
-
-example Recall & 0.65 & 0.46 & 0.58 &             0.61  &  0.50  &  0.53 \\
+% example Accuracy & 0.45 & 0.40 & 0.50 &           0.49  &  0.47  &  0.51 \\
+
+% example F1 & 0.56 & 0.48 & 0.59 &           0.55  &  0.50  &  0.53 \\
+
+% example Hamming Loss & 0.24 & 0.19 & 0.18 &        0.19  &  0.14  &  0.14 \\
+
+% example Precision & 0.50 & 0.51 & 0.60 &          0.50  &  0.49  &  0.53 \\
+
+% example Recall & 0.65 & 0.46 & 0.58 &             0.61  &  0.50  &  0.53 \\
 

### 2010-03-06 src/validate/summarize_weka.pl
an attempt to fix it

@@ -16,7 +16,7 @@ my @learners = slurp('weka-classifiers');
 
 my $ZEROR = "weka.classifiers.rules.ZeroR";
 my $latex_dir = "latex-out";
-my ($DONE,$NOTFOUND,$STRATIFIED,$CORRECT,$INCORRECT,$WEIGHTED) = (1..100);
+my ($DONE,$NOTFOUND,$STRATIFIED,$CORRECT,$INCORRECT,$CONFUSION,$WEIGHTED) = (1..100);
 
 my $dir="";
 
@@ -163,6 +163,46 @@ my @measures = qw(percent roc recall precision fmeasure);
 
 
 
+sub precision {
+    my ($tp,$fp) = @_;
+    if ($tp + $fp == 0) { return 0 }
+    return $tp / (1.0 * $tp + $fp)
+}
+sub recall { 
+    my ($tp,$fn) = @_;
+    if ($tp + $fn == 0){ return 0 }
+    return $tp / (1.0 * $tp + $fn)
+}
+sub true_negative_rate { 
+    my ($tn,$fp) = @_;
+    return $tn / (1.0 * $tn + $fp)
+}
+sub specificity { 
+    my ($tn,$fp) = @_;
+    return true_negative_rate($tn,$fp)
+}
+sub false_positive_rate {
+    my ($fp,$tn) = @_;
+    return $fp / (1.0*($fp + $tn))
+}
+sub accuracy { 
+    my ($tp,$tn,$fp,$fn) = @_;
+    return ($tp + $tn)/ (1.0 * $tp + $tn + $fp + $fn);
+}
+sub true_positive_rate {
+    my ($tp,$fn) = @_;
+    return $tp /(1.0 * ( $tp + $fn))
+}
+sub sensitivity { 
+    my ($tp,$fn) = @_;
+    return true_positive_rate($tp,$fn)
+}
+sub fmeasure{
+    my ($precision,$recall) = @_;
+    if ($precision + $recall == 0){ return 0 }
+    return 2.0 * ($precision * $recall) / (1.0 * ($precision + $recall))
+}
+
 
 
 
@@ -187,7 +227,7 @@ sub get_correct_incorrect {
     my $correct_percent = 0;
     my $incorrect_count = 0;
     my $incorrect_percent = 0;
-    my ($tp,$fp,$precision,$recall,$fmeasure, $roc) = (0,0,0,0,0,0);
[…]

### 2010-03-06 icpc/icpc2010.tex
ok it is 10 pages again!
-
updated fmeasures

@@
 
-\begin{table*}[ht]
+\begin{table*}%[ht]
 	\centering
@@
 % Cribbed from: 
-% src/validate/latex-out/best-learner-per-project-roc-table.tx
+% src/validate/latex-out/best-learner-per-project-roc-table.txt
 \begin{table*}
@@
 \midrule
-portability &  MySQL &  NaiveBayesMultinomial &  0.74 &  58.53 &  11.43 &  0.70  \\ 
-efficiency &  MySQL &  NaiveBayes &  0.67 &  93.69 &  -7.68 &  0.88 \\ 
-reliability &  MySQL &  NaiveBayes &  0.73 &  83.11 &  -12.80 &  0.74 \\ 
-functionality &  MySQL &  DMNBtext &  0.81 &  54.44 &  21.67 &  0.76 \\ 
-maintainability &  MySQL &  DMNBtext &  0.78 &  76.62 &  3.41 &  0.75 \\ 
-usability &  MySQL &  NaiveBayes &  0.75 &  94.54 &  -5.80 &  0.90  \\ 
-%total & MySQL & NaibeBayes & 0.75 & 
+
+portability &  MySQL &  NaiveBayesMultinomial &  0.74 &  58.53 &  11.43 &  0.62 \\
+efficiency &  MySQL &  NaiveBayes &  0.67 &  93.69 &  -7.68 &  0.23 \\ 
+reliability &  MySQL &  NaiveBayes &  0.73 &  83.11 &  -12.80 &  0.41 \\ 
+functionality &  MySQL &  DMNBtext &  0.81 &  54.44 &  21.67 &  0.77  \\ 
+maintainability &  MySQL &  DMNBtext &  0.78 &  76.62 &  3.41 &  0.32  \\ 
+usability &  MySQL &  NaiveBayes &  0.75 &  94.54 &  -5.80 &  0.21  \\ 
  \hline 
-portability &  MaxDB &  NaiveBayes &  0.84 &  77.12 &  2.06 &  0.80 \\ 
-efficiency &  MaxDB &  NaiveBayes &  0.62 &  88.43 &  -11.31 &  0.79 \\ 
-reliability &  MaxDB &  DMNBtext &  0.84 &  89.46 &  3.86 &  0.92  \\ 
-functionality &  MaxDB &  NaiveBayes &  0.67 &  91.26 &  -6.94 &  0.86  \\ 
-maintainability &  MaxDB &  NaiveBayes &  0.70 &  79.43 &  -9.25 &  0.72  \\ 
-usability &  MaxDB &  NaiveBayes &  0.56 &  95.37 &  -4.37 &  0.91  \\ 
+portability &  MaxDB &  NaiveBayes &  0.84 &  77.12 &  2.06 &  0.61 \\ 
+efficiency &  MaxDB &  NaiveBayes &  0.62 &  88.43 &  -11.31 &  0.25  \\ 
+reliability &  MaxDB &  DMNBtext &  0.84 &  89.46 &  3.86 &  0.57  \\ 
+functionality &  MaxDB &  NaiveBayes &  0.67 &  91.26 &  -6.94 &  0.31  \\ 
+maintainability &  MaxDB &  NaiveBayes &  0.70 &  79.43 &  -9.25 &  0.42  \\ 
+usability &  MaxDB &  NaiveBayes &  0.56 &  95.37 &  -4.37 &  0.00  \\ 
  \hline
+
+% this was before I fixed an fmeasure bug
+% 
+% portability &  MySQL &  NaiveBayesMultinomial &  0.74 &  58.53 &  11.43 &  0.70  \\ 
+% efficiency &  MySQL &  NaiveBayes &  0.67 &  93.69 &  -7.68 &  0.88 \\ 
+% reliability &  MySQL &  NaiveBayes &  0.73 &  83.11 &  -12.80 &  0.74 \\ 
+% functionality &  MySQL &  DMNBtext &  0.81 &  54.44 &  21.67 &  0.76 \\ 
+% maintainability &  MySQL &  DMNBtext &  0.78 &  76.62 &  3.41 &  0.75 \\ 
+% usability &  MySQL &  NaiveBayes &  0.75 &  94.54 &  -5.80 &  0.90  \\ 
+% %total & MySQL & NaibeBayes & 0.75 & 
+%  \hline 
+% portability &  MaxDB &  NaiveBayes &  0.84 &  77.12 &  2.06 &  0.80 \\ 
+% efficiency &  MaxDB &  NaiveBayes &  0.62 &  88.43 &  -11.31 &  0.79 \\ 
+% reliability &  MaxDB &  DMNBtext &  0.84 &  89.46 &  3.86 &  0.92  \\ 
+% functionality &  MaxDB &  NaiveBayes &  0.67 &  91.26 &  -6.94 &  0.86  \\ 
+% maintainability &  MaxDB &  NaiveBayes &  0.70 &  79.43 &  -9.25 &  0.72  \\ 
+% usability &  MaxDB &  NaiveBayes &  0.56 &  95.37 &  -4.37 &  0.91  \\
+
 \end{tabular}
@@
 
-\begin{table*}[h]
+\begin{table*}%[h]
 \centering
@@
 With respect to the questions raised in section \ref{sec:suplearn}:
+
 \textbf{Do label frequencies change over time?} -- Yes, MySQL's label

### 2010-03-06 icpc/icpc2010.tex
work on abstract

--- 2010-03-06 icpc/icpc2010.tex
@@
 \newcommand{\XXX}[1]{\textcolor{red}{{\it \textbf{[XXX: #1]}}}}
+\hyphenation{Max-DB}
 
@@
 \begin{abstract}
-
-  Within the field of mining software repositories, many approaches,   such as topic modeling and concept location, rely on the automated   application of machine learning algorithms to corpora.
+  Within the field of mining software repositories, approaches such as topic modeling and concept location rely on the automated   application of machine learning algorithms to corpora.
+%What is a 'topic'?
+
+% Other approaches
+
 This kind of approach typically makes no use of domain-specific knowledge to improve the results or put them into a context that a developer can relate to.
 While too much specificity can produce non-generalizable results, too little produces broad learners that do not provide much immediately useful detail.
+
  In this paper, we investigate what topics the latent topic extractors produce, and examine whether such topics can be generalized across software products.
@@
 became more prevalent as MaxDB matured.
-
 \textbf{Do the different projects differ in their relative topic
@@
 \end{comment}
-The idea of extracting higher-level `concerns' (also known as `concepts', `aspects' or `requirements') has been attacked in two ways.
+The idea of extracting higher-level `concerns' (also known as `concepts', `aspects' or `requirements') has been approached in two ways.
 

### 2010-03-06 icpc/icpc2010.tex
hi-res version..
easterbrook format for abs

@@
 \begin{abstract}
-  Within the field of mining software repositories, approaches such as topic modeling and concept location rely on the automated   application of machine learning algorithms to corpora.
-%What is a 'topic'?
-
+To have meaningful discussions about how software development is progressing, we must be able to devise appropriate labels for work topics.
+We can then categorize development artifacts as belonging to one or more of these topics.
 % Other approaches
-
-This kind of approach typically makes no use of domain-specific knowledge to improve the results or put them into a context that a developer can relate to.
+In order to do this, approaches such as topic modeling and concept location rely on the automated application of machine learning algorithms to development artifacts. 
+%limitations
+However, these approaches neither use domain-specific knowledge to improve results, nor contextualize those results for developers.
 While too much specificity can produce non-generalizable results, too little produces broad learners that do not provide much immediately useful detail.
-
- In this paper, we investigate what topics the latent topic extractors produce, and examine whether such topics can be generalized across software products.
+%introduce idea
+This paper implements \emph{labelled topic extraction}, in which topics are extracted and given labels relating to a cross-project taxonomy.
 We focus on non-functional requirements related to software quality as a potential generalization, since there is some shared belief that these qualities apply broadly across many software systems and their development artifacts.
 In our approach, we name topics based on a static ontology of non-functional requirements and software qualities; in essence, this adds context to general-purpose ontologies such as WordNet.
+%elaborate
 In this paper, we discuss how we use this software development domain knowledge to better annotate topics of development that are extracted from the change log messages in configuration management systems such as CVS.
 Since our taxonomy is global in scope, our results allow us to compare the relative importance of particular qualities across different projects.
+%sum up
 Within the topics extracted from MySQL and MaxDB we found that non-functional requirements were common and often were trends across time.
-Our automated techniques can label topics without much computational overhead.
+Our automated techniques can label topics with little computational overhead.
 
@@
 Traditionally topic extraction has required manual annotation to derive domain-relevant labels.
-This paper attempts to implement \emph{labelled topic extraction}, in which topic are extracted and given labels relating to a cross-project taxonomy.
+This paper implements \emph{labelled topic extraction}, in which topics are extracted and given labels relating to a cross-project taxonomy.
 This is an ontological approach where we attempt to relate networks of words to labels and then search for these terms within our topics.
@@
 %58]{Boegh2008},'' which is sufficient for the purposes of this %research.
-ISO9126 is ``an international standard and thus provides an internationally accepted terminology for software quality \cite[p. 58]{Boegh2008},'' that is sufficient for the purposes of this research.
+ISO9126 is ``an international standard and thus provides an internationally accepted terminology for software quality \cite[p. 58]{Bøegh2008},'' that is sufficient for the purposes of this research.
 However, the terms extracted from ISO9126 may not capture all words associated with the labels.

### 2010-03-06 icpc/icpc2010.tex
finsihed abs

@@
 We focus on non-functional requirements related to software quality as a potential generalization, since there is some shared belief that these qualities apply broadly across many software systems and their development artifacts.
-In our approach, we name topics based on a static ontology of non-functional requirements and software qualities; in essence, this adds context to general-purpose ontologies such as WordNet.
+%In our approach, we name topics based on a static ontology of non-functional requirements and software qualities; in essence, this adds context to general-purpose ontologies such as WordNet.  
 %elaborate
-In this paper, we discuss how we use this software development domain knowledge to better annotate topics of development that are extracted from the change log messages in configuration management systems such as CVS.
-Since our taxonomy is global in scope, our results allow us to compare the relative importance of particular qualities across different projects.
+%We discuss how we use this software development domain knowledge to better annotate topics of development that are extracted from the change log messages in configuration management systems such as CVS.
+We evaluated our approach with an experimental study on two large-scale database projects, MySQL and MaxDB.
+Our results were generalizable across the two projects, showing that non-functional requirements were commonly discussed, and we identified topic trends over time.
+%Our automated techniques can label topics with little computational overhead.  
 %sum up
-Within the topics extracted from MySQL and MaxDB we found that non-functional requirements were common and often were trends across time.
-Our automated techniques can label topics with little computational overhead.
-
+% Since our taxonomy is global in scope, our results allow us to compare the relative importance of particular qualities across different projects.
+Our labelled topic extraction technique allowed us to devise appropriate, context-sensitive labels across these two projects, providing insight into software development activities.
 \end{abstract}
@@
   \centering
+  % hi-res version
+  % \includegraphics[height=.55\textheight]{maxdb-time-smear-plot-Equal}
   \includegraphics[height=.55\textheight]{small-maxdb-time-smear-plot-Equal}
@@
   \centering
+  % hi-res version
+  % \includegraphics[width=.97\textwidth]{mysql-time-smear-plot-Equal}
   \includegraphics[width=.97\textwidth]{small-mysql-time-smear-plot-Equal}

### 2010-03-06 icpc/icpc2010.tex
a mild edit!

--- 2010-03-06 icpc/icpc2010.tex
@@
 \newcommand{\XXX}[1]{\textcolor{red}{{\it \textbf{[XXX: #1]}}}}
-\hyphenation{Max-DB}
 
@@
 \begin{abstract}
-  Within the field of mining software repositories, approaches such as topic modeling and concept location rely on the automated   application of machine learning algorithms to corpora.
-%What is a 'topic'?
-
-% Other approaches
-
+
+  Within the field of mining software repositories, many approaches,   such as topic modeling and concept location, rely on the automated   application of machine learning algorithms to corpora.
 This kind of   approach typically makes no use of domain-specific knowledge to   improve the results or put them into a context that a developer can   relate to.
 While too much specificity can produce non-generalizable   results, too little produces broad learners that do not provide much   immediately useful detail.
-
 In this paper, we investigate what   topics the latent topic extractors produce, and examine whether   such topics can be generalized across software products.
@@
 Traditionally topic extraction has required manual annotation to derive domain-relevant labels.
-This paper attempts to implement \emph{labelled topic extraction}, in which topic are extracted and given labels relating to a cross-project taxonomy.
+This paper attempts to implement \emph{labelled topic extraction}, in which topics are extracted and given labels relating to a cross-project taxonomy.
 This is an ontological approach where we attempt to relate networks of words to labels and then search for these terms within our topics.
-We also compare this approach to machine learners.
+We also compare this approach to supervised machine learning.
 %We start with an unsupervised topic learning algorithm, then manually %annotate the topics using the ontology.
@@
 \end{comment}
-The idea of extracting higher-level `concerns' (also known as `concepts', `aspects' or `requirements') has been approached in two ways.
+The idea of extracting higher-level `concerns' (also known as `concepts', `aspects' or `requirements') has been attacked in two ways.
 

### 2010-03-06 icpc/icpc2010.tex
my attempt at the abstract

--- 2010-03-06 icpc/icpc2010.tex
@@
 %\documentclass[10pt, conference, compsocconf]{IEEEtran}
-\documentclass{acm_proc_article-sp}
+\documentclass[letterpaper]{acm_proc_article-sp}
 %\usepackage{fontspec}
@@
 
-\begin{abstract}
-To have meaningful discussions about how software development is progressing, we must be able to devise appropriate labels for work topics.
+%  Within the field of mining software repositories, many approaches,   such as topic modeling and concept location, rely on the automated   application of machine learning algorithms to corpora.  
+%This kind of   approach typically makes no use of domain-specific knowledge to   improve the results or put them into a context that a developer can   relate to.  
+%While too much specificity can produce non-generalizable   results, too little produces broad learners that do not provide much   immediately useful detail.  
+%In this paper, we investigate what   topics the latent topic extractors produce, and examine whether   such topics can be generalized across software products.  
+%We focus   on non-functional requirements related to software quality as a   potential generalization, since there is some shared belief that   these qualities apply broadly across many software systems and their   development artifacts.
+%In our approach, we name topics based on a   static ontology of non-functional requirements and software   qualities; in essence, this adds context to general-purpose   ontologies such as WordNet.  
+%In this paper, we discuss how we use   this software development domain knowledge to better annotate topics   of development that are extracted from the change log messages in   configuration management systems such as CVS.
+% Since our taxonomy is   global in scope, our results allow us to compare the relative   importance of particular qualities across different projects. 
+%Within the topics extracted from MySQL and MaxDB we found that   non-functional requirements were common and often were trends across   time. 
+%Our automated techniques can label topics without much computational overhead.
+
+\begin{comment}
+ Within the field of mining software repositories, 
+to have meaningful discussions about how software development is progressing, we must be able to devise appropriate labels for work topics.
 We can then categorize development artifacts as belonging to one or more of these topics.
@@
 %introduce idea
-This paper implements \emph{labelled topic extraction}, in which topics are extracted and given labels relating to a cross-project taxonomy.
+This paper implements \emph{labelled topic extraction}, in which topics are extracted from commit comments and given labels relating to a cross-project taxonomy.
 We focus on non-functional requirements related to software quality as a potential generalization, since there is some shared belief that these qualities apply broadly across many software systems and their development artifacts.
@@
 We evaluated our approach with an experimental study on two large-scale database projects, MySQL and MaxDB.
+We extracted topics from their version control commit logs comments.
 Our results were generalizable across the two projects, showing that non-functional requirements were commonly discussed, and we identified topic trends over time.
@@
 Our labelled topic extraction technique allowed us to devise appropriate, context-sensitive labels across these two projects, providing insight into software development activities.
+\end{comment}
+
+\begin{abstract}
+To have a meaningful discussion about how software development is progressing, we must be able to devise appropriate labels for development topics.
+We can then categorize development artifacts as belonging to one or more of these topics.
+In order to do this, approaches such as topic modeling and concept location rely on the automated application of machine learning algorithms to development artifacts. 
+However, these approaches neither use domain-specific knowledge to improve results, nor contextualize those results for developers.
+These extracted topics are often difficult to distinguish and interpret.
+While too much specificity can produce non-generalizable results, too little produces broad learners that do not provide much immediately useful detail.
+This paper implements \emph{labelled topic extraction}, in which topics are extracted from commit comments and given labels relating to a cross-project taxonomy.
+We focus on non-functional requirements related to software quality as a potential generalization, since there is some shared belief that these qualities apply broadly across many software systems and their development artifacts.
+We evaluated our approach with an experimental study on two large-scale database projects, MySQL and MaxDB.
+We extracted topics using Latent Dirichlet Allocation (LDA) from the commit log comments of their version control systems (CVS and BitKeeper).
+Our results were generalizable across the two projects, showing that non-functional requirements were commonly discussed, and we identified topic trends over time.
+Our labelled topic extraction technique allowed us to devise appropriate, context-sensitive labels across these two projects, providing insight into software development activities.
 \end{abstract}
+
+%\vspace*{4em} % push intro down
 
@@
 They are often unnamed and ephemeral --- even more so than roses --- and maybe partially derived from other unnamed and ephemeral artifacts.
-Yet to have meaningful discussions about how software development is progressing, we must be able to devise appropriate labels for our work topics and be able to categorize the development artifacts as belonging to one or more of these topics.
+Yet to have meaningful discussions about how software development is progressing, we must be able to devise appropriate labels for our development topics and be able to categorize the development artifacts as belonging to one or more of these topics.
 
@@
 % Clean this up
+
 Traditionally topic extraction has required manual annotation to derive domain-relevant labels.
@@
 We also compare this approach to machine learners.
+
 %We start with an unsupervised topic learning algorithm, then manually %annotate the topics using the ontology.
@@
 %58]{Boegh2008},'' which is sufficient for the purposes of this %research.
-ISO9126 is ``an international standard and thus provides an internationally accepted terminology for software quality \cite[p. 58]{Bøegh2008},'' that is sufficient for the purposes of this research.
+% NOTE I HAD TO CHANGE THE O HERE IN THE BIBTEX
+ISO9126 is ``an international standard and thus provides an internationally accepted terminology for software quality \cite[p. 58]{Boegh2008},'' that is sufficient for the purposes of this research.
 However, the terms extracted from ISO9126 may not capture all words associated with the labels.
@@
   % \includegraphics[height=.55\textheight]{maxdb-time-smear-plot-Equal}
-  \includegraphics[height=.55\textheight]{small-maxdb-time-smear-plot-Equal}
+  \includegraphics[height=.57\textheight]{small-maxdb-time-smear-plot-Equal}
   \caption{MaxDB 7.500: Labelled topics related to non-functional software qualities plotted over time, continuous topics are trends that occur across time}

### 2010-03-06 icpc/icpc2010.tex
submitted for safety

@@
 \begin{abstract}
-To have a meaningful discussion about how software development is progressing, we must be able to devise appropriate labels for development topics.
+%  Within the field of mining software repositories, many approaches,   such as topic modeling and concept location, rely on the automated   application of machine learning algorithms to corpora.  
+To have a meaningful discussion about the progress of software development, we must be able to devise appropriate labels for development topics.
 We can then categorize development artifacts as belonging to one or more of these topics.

### 2010-03-06 icpc/icpc2010.tex
better abstract

@@
 \begin{abstract}
-%  Within the field of mining software repositories, many approaches,   such as topic modeling and concept location, rely on the automated   application of machine learning algorithms to corpora.  
-To have a meaningful discussion about the progress of software development, we must be able to devise appropriate labels for development topics.
-We can then categorize development artifacts as belonging to one or more of these topics.
-In order to do this, approaches such as topic modeling and concept location rely on the automated application of machine learning algorithms to development artifacts. 
+ Within the field of mining software repositories, many approaches,   such as topic modeling and concept location, rely on the automated   application of machine learning algorithms to corpora.  
+Unfortunately the output of these tools is often difficult to distinguish and interpret as they are often so abstract.
+Thus to have a meaningful discussion about the topics of software development, we must be able to devise appropriate labels for extracted topics. 
 However, these approaches neither use domain-specific knowledge to improve results, nor contextualize those results for developers.
-These extracted topics are often difficult to distinguish and interpret.
 While too much specificity can produce non-generalizable results, too little produces broad learners that do not provide much immediately useful detail.

### 2010-03-06 icpc/icpc2010.tex
spell checked

@@
 
-\caption{Maxdb Mulan results per Mulan learner}
+\caption{MaxDB Mulan results per Mulan learner}
 
@@
            exp2 exp3 learn  B 2nd
-maxdb port 0.55 0.66 0.74   L 3
-maxdb eff  0.55 0.64 0.67   L 3
-maxdb rel  0.76 0.65 0.73   2 2
-maxdb fun  0.52 0.66 0.81   L 3
-maxdb main 0.52 0.56 0.78   L 3
-maxdb use  0.62 0.59 0.75   L 2
+MaxDB port 0.55 0.66 0.74   L 3
+MaxDB eff  0.55 0.64 0.67   L 3
+MaxDB rel  0.76 0.65 0.73   2 2
+MaxDB fun  0.52 0.66 0.81   L 3
+MaxDB main 0.52 0.56 0.78   L 3
+MaxDB use  0.62 0.59 0.75   L 2
 

### 2010-03-06 icpc/icpc2010.tex
intro comment removal -- go look it up!

@@
 
-%Original is here:
-
-%Unlike the poetic license Shakespeare uses, we show that names are
-%indeed important. The rose has its own tangible properties which
-%allow us to sense it. Unfortunately when we are dealing with data
-%abstracted from program source code, we are already dealing with
-%abstract entities that are not tangible.
-
-% To have meaningful discussions about how software development is progressing, we must devise appropriate labels for our work topics, in contrast to the quote above.
-We can then categorize development artifacts as belonging to one or more of these topics.
-
 Topics arise from the wide variety of issues which occur during a software project's life-cycle.
@@
 We observed that topic trends were often non-functional requirements.
-
-%As development of a software project progresses, developers must concern themselves with a wide variety of issues --- or ``topics'' --- that relate to the problem domain, the processes and tools used, the development artifacts themselves, etc.
-The set of development topics for a given project can sometimes be extracted automatically by analyzing artifacts within software repositories, such as change-log comments that developers create when committing revisions to the project's source control system.
-Our previous work dealt with topic trends, which are topics that recur over time~\cite{Hindle09ICSM}.
-We observed that topic trends were often non-functional requirements.
-
-%During the development of a software project developers touch on many %issues integral to the purpose of the software project.
-These issues %are often referred to as topics.
-These topics can sometimes be %automatically extracted from the changelog comments that developers %create while committing revisions to the project's source control %repository.
-Our previous work dealt with %topic-trends~\cite{Hindle09ICSM}, topics that repeat overtime.
-It %seemed that these trends were often non-functional requirements.
 
@@
 
-%If a topic has a name or a label we immediately have a good idea of  %what it is.
-In our case the name is the bit of information, the %summary, that was missing from our topics to make them useful and %complete.
-Although we are convinced that accurate naming is not %completely feasible, so we have focussed on labelling topics by %non-functional requirements.
-
 A topic --- that is, a word distribution in our world --- needs a suggestive name to succinctly convey its intent, and make it easy to use in discussions and analyses about the development of the system.
@@
 We have therefore decided to focus on topic labels from the sub-domain of non-functional requirements related to software quality.
-
-%In this paper we \textit{name} and \textit{label} topics generated by %a topic modeling technique, Latent Dirichlet %Allocation~\cite{Blei2003}.
-We use these topic models to train %supervised learners to classify larger datasets.
-We also show that %these topics can be generalized across products.
-
-% Clean this up
 
@@
 
-%We start with an unsupervised topic learning algorithm, then manually %annotate the topics using the ontology.
-From the annotated topics, we %apply automated classifiers to find labels for the remaining topics.
-
-%\XXX{Not sure about this}
-\begin{comment}
-Typically what is distinct is the software's particular functionality -- whether it offers email composition, prepares tax returns, and so on.
-What is common, however, is the non-functional, or quality requirements, that the software focuses on.
-Our approach makes use of software quality taxonomies in order to compare different projects on similar concepts.
-For example, how is the Security quality treated over time by product A versus product B?
-Does product A care about it more as it matures?
-Does the quality occur more often in software commits?
-\end{comment}
-
 Our contributions include:
@@
   compared to other approaches.
-% unsure about this
+
 \item We show that these labels with their topics can be learnt and used to classify other data-sets.

### 2010-03-06 icpc/icpc2010.tex
another edit!

@@
 % check
-\emph{Area of ROC Curve} is the area under the Receiver Operating Characteristic (ROC) curve, sometimes referred to as \emph{AUC}.
+\emph{Area of ROC Curve} is the area under the Receiver Operating Characteristic (\emph{ROC}) curve, sometimes referred to as \emph{AUC}.
 ROC values reflect a score, similar to school letter-grades (A is 0.9, C is 0.6), reflecting how well a particular learner performed for the given data.
 A ROC result of 0.5 would be equivalent to a random learner (randomly classifying data).
-ROC maps to the more familiar concepts of precision/sensitivity and recall/specificity: it plots the true positive rate (sensitivity) vs.
-the false positive rate (1 - specificity).
+ROC maps to the more familiar concepts of precision/sensitivity and recall/specificity: it plots the true positive rate (sensitivity) versus the false positive rate (1 - specificity).
 A perfect learner has a ROC value of 1.0, reflecting perfect recall and precision.

### 2010-03-06 icpc/icpc2010.tex
more edits

@@
 	documents (commit log comments).
-Many researchers~\cite{marcus04wcre,Poshyvanyk2007,lukins08wcre,Linstead2007} have applied tools like LSI and LDA for mining software repositories, in particular analyzing source code, bugs or developer comments.
+Many researchers~\cite{marcus04wcre,Poshyvanyk2007,lukins08wcre,Linstead2007} have applied tools like LSI and LDA to mining software repositories, in particular analyzing source code, bugs or developer comments.
 \begin{comment}
@@
 \subsection{Generating the data}
-From each project, we used source control commit comments, the messages that programmers write when they commit revisions to a source control repository.
+For each project, we used source control commit comments, the messages that programmers write when they commit revisions to a source control repository.
 We leveraged the data that we gathered in \cite{Hindle09ICSM} for this work.
@@
 We extracted these messages and indexed them by creation time.
-We summarized each message as a word distribution but removed stop-words such as common English words \emph{the} and \emph{at}.
+We summarized each message as a word distribution but removed stop-words such as common English words like \emph{the} and \emph{at}.
 

### 2010-03-06 icpc/icpc2010.tex
more edits!

@@
 This size of period is smaller than the time between minor releases but large enough for there to be sufficient commits to analyze.
-We applied Blei's LDA implementation~\cite{Blei2003} against the word distributions of these commits, and generated list of topics per period.
+We applied Blei's LDA implementation~\cite{Blei2003} against the word distributions of these commits, and generated lists of topics per period.
 %A topic extracted in the first period Period 1, the most important %topic might have been ``security", with ``sql'' second, down on to %less relevant topics. 
@@
 Finally, we analyzed two mailing lists from the KDE project to enhance the specificity of the sets.
-We selected KDE-Usability, which focuses on usability discussions for KDE as a whole; and KDE-Konqueror, a list about a long-lived web browser project.
+We selected KDE-Usability, which focuses on usability discussions for KDE as a whole; and KDE-Konqueror, a mailing list about a long-lived web browser project.
 For each high-level quality in ISO9126, we first searched for our existing labels; we then randomly sampled twenty-five mail messages that were relevant to that quality, and selected co-occurring terms relevant to that quality.
@@
 MySQL's results for efficiency did not improve in \textsf{exp3} but other qualities such as functionality did improve.
-If a \emph{C} grade performance has a ROC value of $0.6$ then most of these tests scored \emph{C} or less, but our results were still better than random chance.
+If a \emph{C} grade performance has a ROC value of $0.6$ then most of these tests scored a grade of \emph{C} or less, but our results were still better than random chance.
 
@@
 We found that the multi-label learners of BR, CLR and HOMER did not do as well for Macro-ROC and Micro-F1 as NaiveBayes and other NaiveBayes derived learners did.
-This suggests that by sticking together multiple NaiveBayes learners we could probably label topics effectively, but it would require a separate NaiveBayes learner per label.
+This suggests that by sticking together multiple NaiveBayes learners we could probably label sets of topics effectively, but it would require a separate NaiveBayes learner per label.
 
@@
   % hi-res version
-  % \includegraphics[height=.55\textheight]{maxdb-time-smear-plot-Equal}
-  \includegraphics[height=.57\textheight]{small-maxdb-time-smear-plot-Equal}
+  \includegraphics[height=.55\textheight]{maxdb-time-smear-plot-Equal}
+  % \includegraphics[height=.57\textheight]{small-maxdb-time-smear-plot-Equal}
   \caption{MaxDB 7.500: Labelled topics related to non-functional software qualities plotted over time, continuous topics are trends that occur across time}
@@
   % hi-res version
-  % \includegraphics[width=.97\textwidth]{mysql-time-smear-plot-Equal}
-  \includegraphics[width=.97\textwidth]{small-mysql-time-smear-plot-Equal}
+  \includegraphics[width=.97\textwidth]{mysql-time-smear-plot-Equal}
+  %\includegraphics[width=.97\textwidth]{small-mysql-time-smear-plot-Equal}
   \caption{MySQL 3.23: Labelled topics related to non-functional software qualities plotted over time, continuous topics are trends that occur across time}

### 2010-03-06 icpc/icpc2010.tex
edits!

@@
 These kinds of changes are slightly related to maintainability.
-A non-functional change that was not quality-related was licensing and copyright; many changes were simply to do with updating copyrights or ensuring copyright or licenses headers were applied to files.
+A non-functional change that was not quality-related was licensing and copyright; many changes were simply to do with updating copyrights or ensuring copyright or license headers were applied to files.
 
@@
 Despite this difference, the name of the module could fool our learners into believing the change was always about efficiency.
-Perhaps a project specific word-bag is needed in order to avoid automated mistakes due to the names of entities and modules with the software project.
+Perhaps a project specific word-bag is needed in order to avoid automated mistakes due to the names of entities and modules of a software project.
 
@@
 They do not use external domain-specific information as we do.
-In \cite{ernst10refsq}, we describe our earlier project, similar to this, to identify change in quality requirements in GNOME software projects; our approach is solely text-matching, however, and does not leverage machine learning strategies.
+In \cite{ernst10refsq}, we describe our earlier project, similar to this, to identify change in quality requirements in GNOME software projects; our approach is solely text-matching, however, it does not leverage machine learning strategies.
 
@@
 
-%\section{Acknowledgements and appendix}
+\section{Appendix}
 Our data and scripts are available at \url{http://softwareprocess.es/nomen/}

### 2010-08-09 icse/icse2011.tex
ICSE 2011 file and figures.
Signed-off-by: Neil Ernst <neil@neilernst.net>

--- 2010-03-06 icpc/icpc2010.tex
@@
-%!TEX TS-program = xelatex
-%!TEX encoding = UTF-8 Unicode
-%\documentclass[10pt, conference, compsocconf]{IEEEtran}
-\documentclass[letterpaper]{acm_proc_article-sp}
-%\usepackage{fontspec}
-%\usepackage{xunicode}
-%\usepackage{xltxtra}
+ \documentclass[]{sig-alternate}
  \usepackage{booktabs}
-%\defaultfontfeatures{Mapping=tex-text}
-%\setmainfont{Times New Roman}
-%\usepackage{amsthm}
+% \usepackage[colorinlistoftodos]{todonotes}
  \usepackage{color}
@@
 \usepackage{amsmath}
-%\usepackage{amsmath,amssymb}
-%\usepackage{algpseudocode}
-%\usepackage[pdfstartview=FitH,xetex,bookmarksopenlevel=3,bookmarks=true]{hyperref}
-% pdfpagescrop={92 112 523 778},
 \usepackage[pdfstartview=FitH,bookmarksopenlevel=3,bookmarks=true]{hyperref} %pdfpagescrop={92 112 523 778},
-%\usepackage{algorithm}%,algorithmic-fix}
+
 \usepackage{url}
@@
    \affaddr{Toronto, Ontario, CANADA}\\
-   \email{jm@cs.toronto.edu}}
-
-\title{What's in a name? On the automated topic naming of software maintenance activities}
+   \email{jm@cs.toronto.edu}
+}
+
+\title{What's in a name? Automated topic naming of software maintenance activities}
 
@@
 
-%There has not been any   investigation into the latent topics these %classifiers represent.
-
+\begin{abstract}
 %  Within the field of mining software repositories, many approaches,   such as topic modeling and concept location, rely on the automated   application of machine learning algorithms to corpora.  
-%This kind of   approach typically makes no use of domain-specific knowledge to   improve the results or put them into a context that a developer can   relate to.  
-%While too much specificity can produce non-generalizable   results, too little produces broad learners that do not provide much   immediately useful detail.  
-%In this paper, we investigate what   topics the latent topic extractors produce, and examine whether   such topics can be generalized across software products.  
-%We focus   on non-functional requirements related to software quality as a   potential generalization, since there is some shared belief that   these qualities apply broadly across many software systems and their   development artifacts.
-%In our approach, we name topics based on a   static ontology of non-functional requirements and software   qualities; in essence, this adds context to general-purpose   ontologies such as WordNet.  
-%In this paper, we discuss how we use   this software development domain knowledge to better annotate topics   of development that are extracted from the change log messages in   configuration management systems such as CVS.
-% Since our taxonomy is   global in scope, our results allow us to compare the relative   importance of particular qualities across different projects. 
-%Within the topics extracted from MySQL and MaxDB we found that   non-functional requirements were common and often were trends across   time. 
-%Our automated techniques can label topics without much computational overhead.
-
-\begin{comment}
- Within the field of mining software repositories, 
-to have meaningful discussions about how software development is progressing, we must be able to devise appropriate labels for work topics.
-We can then categorize development artifacts as belonging to one or more of these topics.
-% Other approaches
-In order to do this, approaches such as topic modeling and concept location rely on the automated application of machine learning algorithms to development artifacts. 
-%limitations
-However, these approaches neither use domain-specific knowledge to improve results, nor contextualize those results for developers.
+% Unfortunately the output of these tools is often difficult to distinguish and interpret as they are often so abstract.
+% Thus to have a meaningful discussion about the topics of software development, we must be able to devise appropriate labels for extracted topics. 
+% However, these approaches neither use domain-specific knowledge to improve results, nor contextualize those results for developers.
 While too much specificity can produce non-generalizable results, too little produces broad learners that do not provide much immediately useful detail.
-%introduce idea
-This paper implements \emph{labelled topic extraction}, in which topics are extracted from commit comments and given labels relating to a cross-project taxonomy.
+% This paper implements \emph{labelled topic extraction}, in which topics are extracted from commit comments and given labels relating to a cross-project taxonomy.
 We focus on non-functional requirements related to software quality as a potential generalization, since there is some shared belief that these qualities apply broadly across many software systems and their development artifacts.
-%In our approach, we name topics based on a static ontology of non-functional requirements and software qualities; in essence, this adds context to general-purpose ontologies such as WordNet.  
-%elaborate
-%We discuss how we use this software development domain knowledge to better annotate topics of development that are extracted from the change log messages in configuration management systems such as CVS.
-We evaluated our approach with an experimental study on two large-scale database projects, MySQL and MaxDB.
-We extracted topics from their version control commit logs comments.
-Our results were generalizable across the two projects, showing that non-functional requirements were commonly discussed, and we identified topic trends over time.
-%Our automated techniques can label topics with little computational overhead.  
-%sum up
-% Since our taxonomy is global in scope, our results allow us to compare the relative importance of particular qualities across different projects.
-Our labelled topic extraction technique allowed us to devise appropriate, context-sensitive labels across these two projects, providing insight into software development activities.
-\end{comment}
-
-\begin{abstract}
- Within the field of mining software repositories, many approaches,   such as topic modeling and concept location, rely on the automated   application of machine learning algorithms to corpora.  
-Unfortunately the output of these tools is often difficult to distinguish and interpret as they are often so abstract.
-Thus to have a meaningful discussion about the topics of software development, we must be able to devise appropriate labels for extracted topics. 
-However, these approaches neither use domain-specific knowledge to improve results, nor contextualize those results for developers.
-While too much specificity can produce non-generalizable results, too little produces broad learners that do not provide much immediately useful detail.
-This paper implements \emph{labelled topic extraction}, in which topics are extracted from commit comments and given labels relating to a cross-project taxonomy.
-We focus on non-functional requirements related to software quality as a potential generalization, since there is some shared belief that these qualities apply broadly across many software systems and their development artifacts.
-We evaluated our approach with an experimental study on two large-scale database projects, MySQL and MaxDB.
+% We evaluated our approach with an experimental study on two large-scale database projects, MySQL and MaxDB.
 We extracted topics using Latent Dirichlet Allocation (LDA) from the commit log comments of their version control systems (CVS and BitKeeper).
 Our results were generalizable across the two projects, showing that non-functional requirements were commonly discussed, and we identified topic trends over time.
-Our labelled topic extraction technique allowed us to devise appropriate, context-sensitive labels across these two projects, providing insight into software development activities.
+% Our labelled topic extraction technique allowed us to devise appropriate, context-sensitive labels across these two projects, providing insight into software development activities.
 \end{abstract}
 
-%\vspace*{4em} % push intro down
-
 \newcommand{\shrink}{}
@@
 \section{Introduction}
-\begin{quote}
-	\emph{What's in a name? that which we call a rose\\
-	By any other name would smell as sweet;}
-	\begin{flushright} -- Romeo and Juliet, II:ii\end{flushright}
-\end{quote}
-
-Few would argue that software development is a bed of roses.
-To most of us, the important properties of a rose are sensual in nature, and concern appearance, odour, and feel.
-While horticulturalists and botanists may see more semantic depth than this, one rose is often just as pleasing to us as another.
-Software development artifacts, on the other hand, are abstract and intangible.
-They are often unnamed and ephemeral --- even more so than roses --- and maybe partially derived from other unnamed and ephemeral artifacts.
-Yet to have meaningful discussions about how software development is progressing, we must be able to devise appropriate labels for our development topics and be able to categorize the development artifacts as belonging to one or more of these topics.
-
-Topics arise from the wide variety of issues which occur during a software project's life-cycle.
-These topics can relate to, among others, the problem domain, the processes and tools used, or the development artifacts themselves.
-The set of development topics for a given project can sometimes be extracted automatically by analyzing artifacts within software repositories, such as change-log comments that developers create when committing revisions to the project's source control system.
-Our previous work dealt with topic trends, which are topics that recur over time~\cite{Hindle09ICSM}.
-We observed that topic trends were often non-functional requirements.
-
-Topics in this paper and the previous paper are word bags or word distributions.
-These word distributions are found via Latent Dirichlet Allocation~\cite{Blei2003}, which finds independent word distributions shared among documents (change-log comments).
-The unfortunate aspect of topics that are word distributions is that they lack the tangibility of a rose.
-They do not self identify with their tangible properties.
-Topics have to be identified by interpreting the prevalent words in the word distribution and by inspecting related documents.
+The act of developing software generates many artifacts, including source code, bug reports, commit messages, and test cases.
+It does not take long for a software project to grow to a point where keeping track of these artifacts exceeds easy comprehension.
+For these projects, developers interact with abstractions of these artifacts, rather than the artifact itself.
+Such abstractions include project dashboards~\cite{}, tags (Peggy/Christoph), and complexity metrics (\cite{mccabe}).
+
+Developers do more than merely observe these abstractions, however \cite{}.
+They must also use them to assess progress and work with colleagues.
+To have meaningful discussions about how software development is progressing, we must be able to devise appropriate labels for our development artifacts.
+Such labels can be used to compare progress or understand how well a project is meeting requirements.
+These labels then become the topics for development conversations (such as the stream of commentary on a bug report).
+One way to identify topics is to use the artifacts themselves to derive them.
+
+Deriving such topics is difficult.
+Topics have to be identified by interpreting the prevalent words in a word distribution and by inspecting related documents.
 This is impractical when one has to handle more than one hundred different topics.
-It would be nice if we could have automatic assistance to determine what the topic is about.
-This is the power of labelling and naming.
-
-A topic --- that is, a word distribution in our world --- needs a suggestive name to succinctly convey its intent, and make it easy to use in discussions and analyses about the development of the system.
-Our previous experience leads us to believe that automated topic naming is currently infeasible in the general case.
-We have therefore decided to focus on topic labels from the sub-domain of non-functional requirements related to software quality.
-
-Traditionally topic extraction has required manual annotation to derive domain-relevant labels.
-This paper implements \emph{labelled topic extraction}, in which topics are extracted and given labels relating to a cross-project taxonomy.
-This is an ontological approach where we attempt to relate networks of words to labels and then search for these terms within our topics.
-We also compare this approach to machine learners.
-
-Our contributions include:
-\begin{itemize}
-\item  We introduce labelled topic extraction, and demonstrate its usefulness
-  compared to other approaches.
-
-\item We show that these labels with their topics can be learnt and used to classify other data-sets.
-\item We present visualizations of named topics and their trends over time to aid communication and analysis.
-\item We use an exploratory case study of
-  several open source database systems to show how named topics can be compared between projects.
-\end{itemize}
-
-We first introduce some important terminology for our work.
-We then describe our methodology, including our data-sets, then highlight our results.
-We conclude with a look at related work and possible improvements.
-
-\shrink
-\section{Background}
-
-We provide a brief overview of software repository mining and information retrieval.
-This work is related to the mining software repositories (MSR)~\cite{msr} research community as it deals expressly with analyzing a project's source control repository, and the messages associated with revisions therein.
-
-\shrink
-\subsection{Definitions}
-We will use the following terminology in the rest of the paper.
-A \emph{message} is a block of text written by a developer. In this
-paper, messages will be the CVS and BitKeeper commit log comments made
-when a developer commits changes to files in a repository. A \emph{word
-  distribution} is the summary of a message by word count. Each word
-distribution will be over the words in all messages. However, most
-words will not appear in each message. A word distribution is effectively
-a word count divided by the message size. A \emph{topic} is a word
-distribution, i.e., a set of words that form a word distribution that is
-unique and independent within the set of documents in our total
-corpus. One could think of a topic as a distribution of the centroid
-of a group of messages. In this paper we often summarize topics by the
-top ten most frequent words of their word distribution.  A \emph{trend}
-is one or more similar topics that recur over time.  Trends are
-particularly important, as they indicate long-lived and recurring
-topics that may provide key insights into the development of the
-project. A \emph{label} is part of a title we attach to a topic, whether manually or automatically.
-
-% area under the ROC 
-% check
-\emph{Area of ROC Curve} is the area under the Receiver Operating Characteristic (\emph{ROC}) curve, sometimes referred to as \emph{AUC}.
+To scale to meaningfully sized projects, we require automatic assistance to determine the topic label.
+
+In our previous work we dealt with topic trends, which are topics that recur over time~\cite{Hindle09ICSM}.
+We observed that topic trends were often non-functional requirements (NFRs).
+NFRs have the property of being cross-domain and widely applicable.
+In this sense, they are useful abstractions for developer conversations about different software projects.
+
+In this paper, we describe \textbf{Nomen}, which is a tool for \emph{labelled topic extraction}. \textbf{Nomen} addresses two gaps in the topic mining literature:
+\begin{enumerate}
+  \item Topic mining of software has been limited to one project at a time.
+This is because traditional topic mining techniques are specific to a particular dataset.
+\textbf{Nomen} allows for comparisons \textit{between} projects.
+  \item Topic modeling creates bags of words that require interpretation by the user to assign meaning. Like (1), this means
+that it is difficult to discuss results independent of the project context. Our technique automatically assigns labels agnostic of project.
+\end{enumerate}
+
+This paper begins by introducing labelled topic extraction, demonstrating its usefulness compared to other approaches.
+We show that these labels with their topics can be learned and used to classify other data-sets.
+We then present visualizations of named topics and their trends over time to aid communication and analysis.
+We use an exploratory case study of two open source database systems to show how named topics can be compared between projects.
+
+% \section{Background}
+
+% We provide a brief overview of software repository mining and information retrieval.
+% This work is related to the mining software repositories (MSR)~\cite{msr} research community as it deals expressly with analyzing a project's source control repository, and the messages associated with revisions therein.
+
+% \subsection{Definitions}
+% We will use the following terminology in the rest of the paper.
+% A \emph{message} is a block of text written by a developer. In this
+% paper, messages will be the CVS and BitKeeper commit log comments made
+% when a developer commits changes to files in a repository. A \emph{word
+%   distribution} is the summary of a message by word count. Each word
+% distribution will be over the words in all messages. However, most
+% words will not appear in each message. A word distribution is effectively
+% a word count divided by the message size. A \emph{topic} is a word
+% distribution, i.e., a set of words that form a word distribution that is
+% unique and independent within the set of documents in our total
+% corpus. One could think of a topic as a distribution of the centroid
+% of a group of messages. In this paper we often summarize topics by the
+% top ten most frequent words of their word distribution.  A \emph{trend}
+% is one or more similar topics that recur over time.  Trends are
+% particularly important, as they indicate long-lived and recurring
+% topics that may provide key insights into the development of the
+% project. A \emph{label} is part of a title we attach to a topic, whether manually or automatically.
+
+% % area under the ROC 
+% % check
+% \emph{Area of ROC Curve} is the area under the Receiver Operating Characteristic (\emph{ROC}) curve, sometimes referred to as \emph{AUC}.
 ROC values reflect a score, similar to school letter-grades (A is 0.9, C is 0.6), reflecting how well a particular learner performed for the given data.
@@
 
-\shrink
-\subsection{Topic and Concept Extraction}
-Topic extraction, sometimes called concept extraction, uses tools such as Latent Dirichlet Allocation (LDA)~\cite{Blei2003} and Latent Semantic Indexing (LSI) to extract
-independent word distributions (topics) from
-	documents (commit log comments).
+% \shrink
+% \subsection{Topic and Concept Extraction}
+% Topic extraction, sometimes called concept extraction, uses tools such as Latent Dirichlet Allocation (LDA)~\cite{Blei2003} and Latent Semantic Indexing (LSI) to extract
+% independent word distributions (topics) from
+% 	documents (commit log comments).
 Many researchers~\cite{marcus04wcre,Poshyvanyk2007,lukins08wcre,Linstead2007} have applied tools like LSI and LDA to mining software repositories, in particular analyzing source code, bugs or developer comments.
-\begin{comment}
-Ideally these extracted topics 
-correlate with actual development topics that the developers discussed
-during the development of the software project.
+% \begin{comment}
+% Ideally these extracted topics 
+% correlate with actual development topics that the developers discussed
+% during the development of the software project.
 For example, our technique might identify a collection of words: ``list next has iterator add'', which we then identify (manually) as concerning the topics of collections and the Iterator pattern.
-\end{comment}
-
-Typically a topic analysis tool like LDA will try to find $N$ independent word distributions found within the word distributions of all the messages.
+% \end{comment}
+
+% Typically a topic analysis tool like LDA will try to find $N$ independent word distributions found within the word distributions of all the messages.
 Linear combinations of these $N$ word distributions are then meant to be able to recreate the word distributions of all of the underlying messages.
@@
 
-In topic analysis a single document, such as a commit message, can be related to multiple topics.
+% In topic analysis a single document, such as a commit message, can be related to multiple topics.
 Representing documents as a mixture of topics maps well to source code repository commits, which often have more than one purpose~\cite{Hindle09ICSM}.
@@
 
-\begin{comment}
-In this paper we use LDA to identify topics.
-LDA works by creating topic models for each period.
-A topic model is a statistical summary of the important subjects for that data-set.
-The input to the algorithm is a set of documents, with contents treated as a bag of words, and a parameter to set the number of topics to generate.
-The algorithm returns a word-document matrix relating the occurrence frequency of a
- The idea is similar to other IR approaches like TF/IDF (term frequency/inverse document frequency), which find uncommon words that are not found in other word-lists.
-These approaches are unsupervised in that they rely solely on the source data with no human intervention.
-\end{comment}
-
-One issue that arises with use of unsupervised techniques is how to label the topics.
+% One issue that arises with use of unsupervised techniques is how to label the topics.
 While the topic models themselves are generated automatically, what to make of them is less clear.
@@
 
-\shrink
-\subsection{Supervised learning}
-While unsupervised techniques (LSI and LDA are both unsupervised) are appealing in their lack of human intervention, and thus lower effort, supervised learners have the advantage of domain knowledge which typically means improved results.
+% \shrink
+% \subsection{Supervised learning}
+% While unsupervised techniques (LSI and LDA are both unsupervised) are appealing in their lack of human intervention, and thus lower effort, supervised learners have the advantage of domain knowledge which typically means improved results.
 In supervised learning, the data is divided into slices.
@@
 
-\shrink 
 \section{Methodology}
 
-\begin{comment}
-% To gather the data, we first extract the commit log comments from a
-% project's SCS repository. We filter out stop words and produce word distributions from these messages. These distributions are bucketed
-% into 30-day non-overlapping windows. Each window is subject to LDA analysis. %Figure \ref{fig:commits}
+% To gather the data, we first extract the commit log comments from a  project's SCS repository.
+We filter out stop words and produce word distributions from these messages.
+These distributions are bucketed
+% into 30-day non-overlapping windows (periods). Each window is subject to LDA analysis. We arbitrarily create 20 topics, vectors of important words according to LDA.
+
+%Figure \ref{fig:commits}
 % depicts the general process for processing and analyzing the commit messages.
-\end{comment}
+
+Figure \ref{fig:process} gives an outline of the methodology followed.
+
+\begin{figure*}
+  \centering
+ \includegraphics[width=.55\textwidth]{process-model}
+ \caption{Research methodology process view.}
+  \label{fig:process}
+\end{figure*}
 
@@
 
-\shrink
 \subsection{Generating the data}
@@
 We extracted these messages and indexed them by creation time.
-We summarized each message as a word distribution but removed stop-words such as common English words like \emph{the} and \emph{at}.
+We summarized each message as a word distribution but removed stop-words such as \emph{the} and \emph{at}.
 
@@
 %A topic extracted in the first period Period 1, the most important %topic might have been ``security", with ``sql'' second, down on to %less relevant topics. 
-We arbitrarily set the number of topics to generate to $20$, because past experimentation showed that fewer topics might aggregate multiple unique topics while any more topics seemed to dilute the results and create indistinct topics.
-As well, more than $20$ topics quickly became infeasible for inspection and it was difficult to discern the difference in topics.
-
-\shrink
+We arbitrarily set the number of topics to generate to $20$, because past experimentation, e.g.
+\cite{Hindle09ICSM}, showed that fewer topics might aggregate multiple unique topics while any more topics seemed to dilute the results and create indistinct topics.
+Furthermore, more than $20$ topics quickly became infeasible for inspection.
+
 \subsection{Associating labels}
@@
 To construct these expanded word-lists, we used WordNet~\cite{Fellbaum1998}, an English-language ``lexical database'' that contains semantic relations between words, including meronymy and synonymy.
-We then added Boehm’s 1976 software quality model \cite{Boehm+:1976:ICSE}, and classified his eleven ‘ilities’ into their respective ISO9126 qualities.
+We then added Boehm's 1976 software quality model \cite{Boehm+:1976:ICSE}, and classified his eleven `ilities' into their respective ISO9126 qualities.
 We did the same for the quality model produced by McCall et al.
@@
 \end{tabular}
-	\caption{Qualities and associated signifiers – WordNet version (\textsf{exp2})}
+	\caption{Qualities and associated signifiers -- WordNet version (\textsf{exp2})}
 	\label{tbl:wnsig}
@@
 
-\shrink
 \subsection{Supervised learning}
@@
 
-\shrink
 \section{Observations and evaluation}
 
-\shrink
 \subsection{Word list similarity}
@@
 
-\shrink
 \subsection{Analysis of the unsupervised labelling}
@@
 For each quality we tried to assess whether the manual tag matched the unsupervised label assigned.
-Table \ref{tbl:maxdb-unsup-results} shows our results for MaxDB and MySQL.
+Figure \ref{fig:maxdb-unsup-results} shows our results for MaxDB and MySQL.
 In general results are poor.
-Using the F-Measure, the weighted average of precision and recall, where 1 is perfect, our best results are 0.6, a few at 0.4, and most around 0.2.
+Total ROC was slightly better than 0.5, and usability seems relatively accurate.
+
+\begin{figure*}
+  \centering
+ \includegraphics[width=.55\textwidth]{roc-unsup-pretty}
+ \caption{ROC values for each NFR, per project and wordlist.}
+  \label{fig:maxdb-unsup-results}
+\end{figure*}
+
+% Using the F-Measure, the weighted average of precision and recall, where 1 is perfect, our best results are 0.6, a few at 0.4, and most around 0.2.
 We achieved similar results using the Matthew's correlation coefficient (used to measure efficacy where classes are of different sizes) and ROC.
 
-\begin{table*}%[ht]
-	\centering
-
-\begin{tabular}{c|c|c|c|c|c|c}
-\toprule
-Experiment & Label & F1 & MCC & Precision & Recall & ROC  \\ 
-\midrule
-MaxDB exp2 & portability & 0.228 & 0.182 & 0.520 & 0.146 & 0.553 \\ 
- & efficiency & 0.217 & 0.125 & 0.237 & 0.200 & 0.558\\ 
- & reliability & 0.380 & 0.340 & 0.246 & 0.829 & 0.765 \\ 
- & functionality & 0.095 & 0.083 & 0.250 & 0.059 & 0.521 \\ 
- & maintainability & 0.092 & 0.123 & 0.571 & 0.050 & 0.520 \\ 
- & usability & 0.175 & 0.138 & 0.113 & 0.389 & 0.620 \\ 
- & total & 0.236 & 0.127 & 0.248 & 0.225 & 0.561 \\ 
-\midrule
-MySQL exp2 & portability & 0.138 & 0.211 & 1.000 & 0.074 & 0.537 \\ 
- & efficiency & 0.345 & 0.327 & 0.476 & 0.270 & 0.625 \\ 
- & reliability & 0.425 & 0.287 & 0.348 & 0.545 & 0.669 \\ 
- & functionality & 0.025 & 0.006 & 0.571 & 0.013 & 0.501 \\ 
- & maintainability & 0.000 & 0.000 & 0.000 & 0.000 & 0.500 \\ 
- & usability & 0.175 & 0.135 & 0.200 & 0.156 & 0.560 \\ 
- & total & 0.167 & 0.095 & 0.403 & 0.105 & 0.527 \\ 
-\midrule
-MaxDB exp3 & portability & 0.472 & 0.286 & 0.402 & 0.573 & 0.660 \\ 
- & efficiency & 0.223 & 0.068 & 0.130 & 0.778 & 0.549 \\ 
- & reliability & 0.257 & 0.196 & 0.149 & 0.927 & 0.652 \\ 
- & functionality & 0.236 & 0.187 & 0.138 & 0.824 & 0.665 \\ 
- & maintainability & 0.338 & 0.112 & 0.266 & 0.463 & 0.566 \\ 
- & usability & 0.108 & 0.094 & 0.057 & 0.944 & 0.595 \\ 
- & total & 0.258 & 0.093 & 0.160 & 0.671 & 0.568 \\ 
-\midrule
-MySQL exp3 & portability & 0.413 & 0.170 & 0.564 & 0.325 & 0.574 \\ 
- & efficiency & 0.158 & 0.105 & 0.089 & 0.703 & 0.608 \\ 
- & reliability & 0.388 & 0.240 & 0.260 & 0.758 & 0.660 \\ 
- & functionality & 0.652 & 0.240 & 0.655 & 0.649 & 0.620 \\ 
- & maintainability & 0.203 & 0.007 & 0.240 & 0.175 & 0.503 \\ 
- & usability & 0.105 & 0.013 & 0.057 & 0.688 & 0.513 \\ 
- & total & 0.362 & 0.076 & 0.284 & 0.499 & 0.544 \\
-\bottomrule
-\end{tabular}
-	\caption{Results for automatic topic labelling. F1 = f-measure, MCC = Matthew's correlation coeff., ROC = Area under ROC curve}
-	\label{tbl:maxdb-unsup-results}
-\end{table*}
-
+% \begin{table*}%[ht]
+% 	\centering
+
+% \begin{tabular}{c|c|c|c|c|c|c}
+% \toprule
+% Experiment & Label & F1 & MCC & Precision & Recall & ROC  \\ 
+% \midrule
+% MaxDB exp2 & portability & 0.228 & 0.182 & 0.520 & 0.146 & 0.553 \\ 
+%  & efficiency & 0.217 & 0.125 & 0.237 & 0.200 & 0.558\\ 
+%  & reliability & 0.380 & 0.340 & 0.246 & 0.829 & 0.765 \\ 
+%  & functionality & 0.095 & 0.083 & 0.250 & 0.059 & 0.521 \\ 
+%  & maintainability & 0.092 & 0.123 & 0.571 & 0.050 & 0.520 \\ 
+%  & usability & 0.175 & 0.138 & 0.113 & 0.389 & 0.620 \\ 
+%  & total & 0.236 & 0.127 & 0.248 & 0.225 & 0.561 \\ 
+% \midrule
+% MySQL exp2 & portability & 0.138 & 0.211 & 1.000 & 0.074 & 0.537 \\ 
+%  & efficiency & 0.345 & 0.327 & 0.476 & 0.270 & 0.625 \\ 
+%  & reliability & 0.425 & 0.287 & 0.348 & 0.545 & 0.669 \\ 
+%  & functionality & 0.025 & 0.006 & 0.571 & 0.013 & 0.501 \\ 
+%  & maintainability & 0.000 & 0.000 & 0.000 & 0.000 & 0.500 \\ 
+%  & usability & 0.175 & 0.135 & 0.200 & 0.156 & 0.560 \\ 
+%  & total & 0.167 & 0.095 & 0.403 & 0.105 & 0.527 \\ 
+% \midrule
+% MaxDB exp3 & portability & 0.472 & 0.286 & 0.402 & 0.573 & 0.660 \\ 
+%  & efficiency & 0.223 & 0.068 & 0.130 & 0.778 & 0.549 \\ 
+%  & reliability & 0.257 & 0.196 & 0.149 & 0.927 & 0.652 \\ 
+%  & functionality & 0.236 & 0.187 & 0.138 & 0.824 & 0.665 \\ 
+%  & maintainability & 0.338 & 0.112 & 0.266 & 0.463 & 0.566 \\ 
+%  & usability & 0.108 & 0.094 & 0.057 & 0.944 & 0.595 \\ 
+%  & total & 0.258 & 0.093 & 0.160 & 0.671 & 0.568 \\ 
+% \midrule
+% MySQL exp3 & portability & 0.413 & 0.170 & 0.564 & 0.325 & 0.574 \\ 
+%  & efficiency & 0.158 & 0.105 & 0.089 & 0.703 & 0.608 \\ 
+%  & reliability & 0.388 & 0.240 & 0.260 & 0.758 & 0.660 \\ 
+%  & functionality & 0.652 & 0.240 & 0.655 & 0.649 & 0.620 \\ 
+%  & maintainability & 0.203 & 0.007 & 0.240 & 0.175 & 0.503 \\ 
+%  & usability & 0.105 & 0.013 & 0.057 & 0.688 & 0.513 \\ 
+%  & total & 0.362 & 0.076 & 0.284 & 0.499 & 0.544 \\
+% \bottomrule
+% \end{tabular}
+% 	\caption{Results for automatic topic labelling. F1 = f-measure, MCC = Matthew's correlation coeff., ROC = Area under ROC curve}
+% 	\label{tbl:maxdb-unsup-results}
+% \end{table*}
 Based on these results we find that reliability and usability worked well for MaxDB in \textsf{exp2} and better in \textsf{exp3}.
@@
 %\subsection{Classifying topics}
-\shrink
+
 \subsection{Analysis of the supervised labelling}
@@
 
-% this was before I fixed an fmeasure bug
-% 
-% portability &  MySQL &  NaiveBayesMultinomial &  0.74 &  58.53 &  11.43 &  0.70  \\ 
-% efficiency &  MySQL &  NaiveBayes &  0.67 &  93.69 &  -7.68 &  0.88 \\ 
-% reliability &  MySQL &  NaiveBayes &  0.73 &  83.11 &  -12.80 &  0.74 \\ 
-% functionality &  MySQL &  DMNBtext &  0.81 &  54.44 &  21.67 &  0.76 \\ 
-% maintainability &  MySQL &  DMNBtext &  0.78 &  76.62 &  3.41 &  0.75 \\ 
-% usability &  MySQL &  NaiveBayes &  0.75 &  94.54 &  -5.80 &  0.90  \\ 
-% %total & MySQL & NaibeBayes & 0.75 & 
-%  \hline 
-% portability &  MaxDB &  NaiveBayes &  0.84 &  77.12 &  2.06 &  0.80 \\ 
-% efficiency &  MaxDB &  NaiveBayes &  0.62 &  88.43 &  -11.31 &  0.79 \\ 
-% reliability &  MaxDB &  DMNBtext &  0.84 &  89.46 &  3.86 &  0.92  \\ 
-% functionality &  MaxDB &  NaiveBayes &  0.67 &  91.26 &  -6.94 &  0.86  \\ 
-% maintainability &  MaxDB &  NaiveBayes &  0.70 &  79.43 &  -9.25 &  0.72  \\ 
-% usability &  MaxDB &  NaiveBayes &  0.56 &  95.37 &  -4.37 &  0.91  \\
-
 \end{tabular}
@@
 
-\shrink
 \subsection{Applying multiple labels to topics}
@@
 
-\shrink
 \subsection{Summary of techniques}
@@
 
-\shrink
 \subsection{Visualization}
@@
   % hi-res version
-  \includegraphics[height=.55\textheight]{maxdb-time-smear-plot-Equal}
+  %\includegraphics[height=.55\textheight]{maxdb-time-smear-plot-Equal}
+  \includegraphics[height=.55\textheight]{small-maxdb-time-smear-plot-Equal}
   % \includegraphics[height=.57\textheight]{small-maxdb-time-smear-plot-Equal}
@@
   % hi-res version
-  \includegraphics[width=.97\textwidth]{mysql-time-smear-plot-Equal}
+  %\includegraphics[width=.97\textwidth]{mysql-time-smear-plot-Equal}
+  \includegraphics[width=.97\textwidth]{small-mysql-time-smear-plot-Equal}
   %\includegraphics[width=.97\textwidth]{small-mysql-time-smear-plot-Equal}
@@
 
-\shrink
 \subsection{Comparing MaxDB and MySQL}
@@
 
-\shrink
 \subsection{Annotation observations}
@@
 
-\shrink
 \subsection{Effectiveness}
@@
 
-\shrink
 \subsection{Threats to validity}
@@
 
-\shrink
 \section{Related work}
@@
 
-\shrink
 \section{Conclusions and future work}

### 2010-08-13 icse/icse2011.tex
updated structure. There is no longer a separation between observations and results. More room for discussion now.
Signed-off-by: Neil Ernst <neil@neilernst.net>

@@
  \usepackage{booktabs}
-% \usepackage[colorinlistoftodos]{todonotes}
+\usepackage{subfigure}
+ \usepackage[colorinlistoftodos,backgroundcolor=yellow]{todonotes}
  \usepackage{color}
@@
 
-\title{What's in a name? Automated topic naming of software maintenance activities}
-
+%\title{What's in a name? Automated topic naming of software maintenance activities}
+\title{Automated topic naming to support analysis of software maintenance activities}
 \maketitle
@@
 \begin{abstract}
-%  Within the field of mining software repositories, many approaches,   such as topic modeling and concept location, rely on the automated   application of machine learning algorithms to corpora.  
-% Unfortunately the output of these tools is often difficult to distinguish and interpret as they are often so abstract.
-% Thus to have a meaningful discussion about the topics of software development, we must be able to devise appropriate labels for extracted topics. 
-% However, these approaches neither use domain-specific knowledge to improve results, nor contextualize those results for developers.
+  Within the field of mining software repositories, many approaches,   such as topic modeling and concept location, rely on the automated   application of machine learning algorithms to corpora.  
+ Unfortunately the output of these tools is often difficult to distinguish and interpret as they are often so abstract.
+ Thus to have a meaningful discussion about the topics of software development, we must be able to devise appropriate labels for extracted topics. 
+ However, these approaches neither use domain-specific knowledge to improve results, nor contextualize those results for developers.
 While too much specificity can produce non-generalizable results, too little produces broad learners that do not provide much immediately useful detail.
-% This paper implements \emph{labelled topic extraction}, in which topics are extracted from commit comments and given labels relating to a cross-project taxonomy.
+ This paper implements \emph{labelled topic extraction}, in which topics are extracted from commit comments and given labels relating to a cross-project taxonomy.
 We focus on non-functional requirements related to software quality as a potential generalization, since there is some shared belief that these qualities apply broadly across many software systems and their development artifacts.
-% We evaluated our approach with an experimental study on two large-scale database projects, MySQL and MaxDB.
+ We evaluated our approach with an experimental study on two large-scale database projects, MySQL and MaxDB.
 We extracted topics using Latent Dirichlet Allocation (LDA) from the commit log comments of their version control systems (CVS and BitKeeper).
 Our results were generalizable across the two projects, showing that non-functional requirements were commonly discussed, and we identified topic trends over time.
-% Our labelled topic extraction technique allowed us to devise appropriate, context-sensitive labels across these two projects, providing insight into software development activities.
+ Our labelled topic extraction technique allowed us to devise appropriate, context-sensitive labels across these two projects, providing insight into software development activities.
 \end{abstract}
@@
 
-This paper begins by introducing labelled topic extraction, demonstrating its usefulness compared to other approaches.
+This paper begins by introducing labelled topic extraction.
 We show that these labels with their topics can be learned and used to classify other data-sets.
@@
 
-% \section{Background}
+\section{Study design and execution}
+Figure \ref{fig:process} gives an outline of the methodology followed.
+We began by gathering source data; then applied unsupervised and supervised learning techniques to identify topics.
+We then use these topics to analyse the role of non-functional requirements in software maintenance.
 
-% We provide a brief overview of software repository mining and information retrieval.
-% This work is related to the mining software repositories (MSR)~\cite{msr} research community as it deals expressly with analyzing a project's source control repository, and the messages associated with revisions therein.
+\begin{figure}
+  \centering
+ \includegraphics[width=.45\textwidth]{process-model}
+ \caption{Research methodology process view.}
+  \label{fig:process}
+\end{figure}
 
-% \subsection{Definitions}
-% We will use the following terminology in the rest of the paper.
-% A \emph{message} is a block of text written by a developer. In this
-% paper, messages will be the CVS and BitKeeper commit log comments made
-% when a developer commits changes to files in a repository. A \emph{word
-%   distribution} is the summary of a message by word count. Each word
-% distribution will be over the words in all messages. However, most
-% words will not appear in each message. A word distribution is effectively
-% a word count divided by the message size. A \emph{topic} is a word
-% distribution, i.e., a set of words that form a word distribution that is
-% unique and independent within the set of documents in our total
-% corpus. One could think of a topic as a distribution of the centroid
-% of a group of messages. In this paper we often summarize topics by the
-% top ten most frequent words of their word distribution.  A \emph{trend}
-% is one or more similar topics that recur over time.  Trends are
-% particularly important, as they indicate long-lived and recurring
-% topics that may provide key insights into the development of the
-% project. A \emph{label} is part of a title we attach to a topic, whether manually or automatically.
+\subsection{Generating the data}
+To evaluate our approach, we sought candidate systems that were mature projects and had openly accessible source control repositories.
+We selected systems from the same application domain, so the functional requirements would be broadly similar.
+We selected MySQL and MaxDB as they were open-source, partially-commercial database systems.
+MaxDB started in the late 1970s as a research project, and was later acquired by SAP.
+As of version 7.500, released April 2007, the project has 940 thousand lines of C source code~\footnote{generated using David A.
+Wheeler's \emph{SLOCCount}.}.
+The MySQL project started in 1994 and MySQL 3.23 was released in early 2001.
+MySQL contains 320 thousand lines of C and C++ source code.
+\todo[inline]{why 3.23 and not newer?}
 
-% % area under the ROC 
-% % check
-% \emph{Area of ROC Curve} is the area under the Receiver Operating Characteristic (\emph{ROC}) curve, sometimes referred to as \emph{AUC}.
-ROC values reflect a score, similar to school letter-grades (A is 0.9, C is 0.6), reflecting how well a particular learner performed for the given data.
-A ROC result of 0.5 would be equivalent to a random learner (randomly classifying data).
-ROC maps to the more familiar concepts of precision/sensitivity and recall/specificity: it plots the true positive rate (sensitivity) versus the false positive rate (1 - specificity).
-A perfect learner has a ROC value of 1.0, reflecting perfect recall and precision.
+For each project, we used source control commit comments, the messages that programmers write when they commit revisions to a source control repository.
+This is the most readily accessible source of project interactions for outside researchers; bug reports and email messages were not available for both projects.
+An example of a typical commit message is: \textit{``history annotate diffs bug fixed (if mysql\_real\_connect() failed there were two pointers to malloc'ed strings, with memory corruption on free(), of course)''}.
+We extracted these messages and indexed them by creation time.
+We summarized each message as a word distribution but removed stop-words such as \emph{the} and \emph{at}.
+Stemming was performed in the later stages of our analysis.
 
-% \shrink
-% \subsection{Topic and Concept Extraction}
-% Topic extraction, sometimes called concept extraction, uses tools such as Latent Dirichlet Allocation (LDA)~\cite{Blei2003} and Latent Semantic Indexing (LSI) to extract
-% independent word distributions (topics) from
-% 	documents (commit log comments).
-Many researchers~\cite{marcus04wcre,Poshyvanyk2007,lukins08wcre,Linstead2007} have applied tools like LSI and LDA to mining software repositories, in particular analyzing source code, bugs or developer comments.
-% \begin{comment}
-% Ideally these extracted topics 
-% correlate with actual development topics that the developers discussed
-% during the development of the software project.
-For example, our technique might identify a collection of words: ``list next has iterator add'', which we then identify (manually) as concerning the topics of collections and the Iterator pattern.
-% \end{comment}
+From that data-set, we created an XML file which separated commits into 30 day periods.
+This size of period is smaller than the time between minor releases but large enough for there to be sufficient commits to analyze.
+For this set of periods, we created `topics' using Latent Dirichlet Allocation, a topic identification algorithm.
 
-% Typically a topic analysis tool like LDA will try to find $N$ independent word distributions found within the word distributions of all the messages.
+Typically a topic analysis tool like LDA will try to find $N$ independent word distributions within the word distributions of all the messages.
 Linear combinations of these $N$ word distributions are then meant to be able to recreate the word distributions of all of the underlying messages.
-These $N$ word distributions effectively form topics: cross cutting collections of words relevant to one or more documents.
-Our problem is that these topics are not easy to interpret, as the underlying pattern is not clear.
-We feel that automatic labelling or naming of these topics would be helpful with respect to interpreting the subject of a topic.
+These $N$ word distributions effectively form topics: cross-cutting collections of words relevant to one or more of our commit messages.
 LDA extracts topics in an unsupervised manner; the algorithm relies solely on the source data with no human intervention.
 
-% In topic analysis a single document, such as a commit message, can be related to multiple topics.
+In topic analysis a single document, such as a commit message, can be related to multiple topics.
 Representing documents as a mixture of topics maps well to source code repository commits, which often have more than one purpose~\cite{Hindle09ICSM}.
@@
 
-% One issue that arises with use of unsupervised techniques is how to label the topics.
+We applied Blei's LDA implementation~\cite{Blei2003} against the word distributions of these commits, and generated lists of topics per period. 
+We set the number of topics to generate to $20$, because past experimentation, e.g.
+\cite{Hindle09ICSM}, showed that fewer topics might aggregate multiple unique topics while any more topics seemed to dilute the results and create indistinct topics.
+Furthermore, more than $20$ topics quickly became infeasible for inspection.
+
+\subsection{Generating word lists}
+Topics are word distributions: lists of words ranked by frequency, which can be burdensome to interpret and hard to distinguish and understand.
 While the topic models themselves are generated automatically, what to make of them is less clear.
@@
 The labels are reasonable enough, but still require an expert in the field to determine them.
-Our technique is automated, because we match keywords from WordNet~\cite{Fellbaum1998} to words in the topic model.
+We sought to automate the process of naming the topics.
 
-% \shrink
-% \subsection{Supervised learning}
-% While unsupervised techniques (LSI and LDA are both unsupervised) are appealing in their lack of human intervention, and thus lower effort, supervised learners have the advantage of domain knowledge which typically means improved results.
-In supervised learning, the data is divided into slices.
-One slice is manually annotated by the domain expert, and the classifications he/she determines are applied to the remaining slices.
-In this paper, we employ the WEKA~\cite{weka09} and Mulan~\cite{mulan} machine learning frameworks in order to test supervised learning.
-
-\section{Methodology}
-
-% To gather the data, we first extract the commit log comments from a  project's SCS repository.
-We filter out stop words and produce word distributions from these messages.
-These distributions are bucketed
-% into 30-day non-overlapping windows (periods). Each window is subject to LDA analysis. We arbitrarily create 20 topics, vectors of important words according to LDA.
-
-%Figure \ref{fig:commits}
-% depicts the general process for processing and analyzing the commit messages.
-
-Figure \ref{fig:process} gives an outline of the methodology followed.
-
-\begin{figure*}
-  \centering
- \includegraphics[width=.55\textwidth]{process-model}
- \caption{Research methodology process view.}
-  \label{fig:process}
-\end{figure*}
-
-To evaluate our approach, we sought candidate systems that were
-mature projects and had openly accessible source control repositories.
-We also decided to select systems from the same application
-domain, as we felt the functional requirements would probably be
-broadly similar.
-%We wanted to compare two mature software projects, that had open %repositories, within the same application domain. 
-% This allows us to partially control for differences in functional %requirements. 
-We selected MySQL and MaxDB as they were open-source, partially-commercial database systems.
-MaxDB started in the late 1970s as a research project, and was later acquired by SAP.
-As of version 7.500, released April 2007, the project has 940 thousand lines of C source code~\footnote{generated using David A.
-Wheeler's \emph{SLOCCount}.}.
-%How many developers?
-%There are approximately ... developers working on it. 
-The MySQL project started in 1994 but MySQL 3.23 was released in early 2001. MySQL contains 320 thousand lines of C and C++ source code.
-
-\subsection{Generating the data}
-For each project, we used source control commit comments, the messages that programmers write when they commit revisions to a source control repository.
-We leveraged the data that we gathered in \cite{Hindle09ICSM} for this work.
-An example of a typical commit message is: \textit{``history annotate diffs bug fixed (if mysql\_real\_connect() failed there were two pointers to malloc'ed strings, with memory corruption on free(), of course)''}.
-We extracted these messages and indexed them by creation time.
-We summarized each message as a word distribution but removed stop-words such as \emph{the} and \emph{at}.
-
-From that data-set, we created an XML file which separated commits into
-monthly windows.
-This size of period is smaller than the time between minor releases but large enough for there to be sufficient commits to analyze.
-We applied Blei's LDA implementation~\cite{Blei2003} against the word distributions of these commits, and generated lists of topics per period.
-%A topic extracted in the first period Period 1, the most important %topic might have been ``security", with ``sql'' second, down on to %less relevant topics. 
-We arbitrarily set the number of topics to generate to $20$, because past experimentation, e.g.
-\cite{Hindle09ICSM}, showed that fewer topics might aggregate multiple unique topics while any more topics seemed to dilute the results and create indistinct topics.
-Furthermore, more than $20$ topics quickly became infeasible for inspection.
-
-\subsection{Associating labels}
-\begin{comment}
-% We would like to `name' our topic models for each period. In this way, we can see how certain words -- subjects -- occur more or less often with each period.
-\end{comment}
-
-Topics are word distributions: essentially lists of words ranked by frequency, which can be burdensome to interpret and hard to distinguish and understand.
-Once we had topics for each period, we tried to associate them with a label from a list of keywords and related terms.
-We performed simple string matching between these topics and our lists, `naming' a topic if it contained that word or words.
-We used several different word lists for comparison.
+Accordingly, we tried to associate each topic with a label from a list of keywords and related terms.
+We performed simple string matching between these topics and our lists, `naming' a topic if it contained that word or words (a topic might match several keywords).
+We used several different word lists for comparison, which we call \textsf{exp1,exp2,exp3} in the text which follows.
 
@@
 
-Our second word list set, \textsf{exp2}, relied on the ISO quality model (ISO9126)~\cite{iso9126}. ISO9126 describes six high-level quality requirements (listed in Table \ref{tbl:wnsig}). 
-%There is some debate about the significance and importance of the %terms in this model. 
-%However, it is ``an international standard and thus provides an %internationally accepted terminology for software quality \cite[p.
-%58]{Boegh2008},'' which is sufficient for the purposes of this %research.
-% NOTE I HAD TO CHANGE THE O HERE IN THE BIBTEX
-ISO9126 is ``an international standard and thus provides an internationally accepted terminology for software quality \cite[p. 58]{Boegh2008},'' that is sufficient for the purposes of this research.
-However, the terms extracted from ISO9126 may not capture all words associated with the labels.
+Our second word list set, \textsf{exp2}, relied on the ISO quality model (ISO9126)~\cite{iso9126}.
+ISO9126 describes six high-level non-functional requirements (listed in Table \ref{tbl:wnsig}).
+There is some debate about the significance and importance of the terms in this model.
+However, ISO9126 is ``an international standard and thus provides an internationally accepted terminology for software quality \cite[p.
+58]{Boegh2008},'' that is sufficient for the purposes of this research.
+The terms extracted from ISO9126 may not capture all words associated with the labels.
 For example, the term ``redundancy'' is one most would agree is relevant to discussion of reliability, but is not in the standard.
@@
 
-%Cite mockus et al
-
-To construct these expanded word-lists, we used WordNet~\cite{Fellbaum1998}, an English-language ``lexical database'' that contains semantic relations between words, including meronymy and synonymy.
+To construct these expanded word lists, we used WordNet~\cite{Fellbaum1998}, an English-language ``lexical database'' that contains semantic relations between words, including common related forms (similar to word stemming), meronymy and synonymy.
 We then added Boehm's 1976 software quality model \cite{Boehm+:1976:ICSE}, and classified his eleven `ilities' into their respective ISO9126 qualities.
@@
 \cite{mccall1977}.
-Finally, we analyzed two mailing lists from the KDE project to enhance the specificity of the sets.
-We selected KDE-Usability, which focuses on usability discussions for KDE as a whole; and KDE-Konqueror, a mailing list about a long-lived web browser project.
-For each high-level quality in ISO9126, we first searched for our existing labels; we then randomly sampled twenty-five mail messages that were relevant to that quality, and selected co-occurring terms relevant to that quality.
-For example, we add the term ``performance'' to the synonyms for \emph{efficiency}, since this term occurs in most KDE mail messages that discuss efficiency.
-%exp2
+Finally, we analyzed two mailing lists from another software project to expand our set with domain-specific terms.
+For example, we add the term ``performance'' to the synonyms for \emph{efficiency}, since this term occurs in most mail messages that discuss efficiency.
 
-For the third -- \textsf{exp3} -- list of quality labels, we extended the list from \textsf{exp2} using unfiltered WordNet similarity matches.
+For the third list of quality labels,  \textsf{exp3}, we extended the list from \textsf{exp2} using unfiltered WordNet similarity matches.
 Similarity in WordNet means siblings in a hypernym tree.
@@
 \end{tabular}
-	\caption{Qualities and associated signifiers -- WordNet version (\textsf{exp2})}
+	\caption{NFRs and associated signifiers -- \textsf{exp2}}
 	\label{tbl:wnsig}
@@
 
-\subsection{Supervised learning}
-
-\label{sec:suplearn}
-
-In order to validate how effective these word-bag approaches to topic labelling would be we had to make a data set to test against.
-For MySQL 3.23 and MaxDB 7.500, we manually annotated each extracted topic in each period with the same quality labels as \textsf{exp2} (software qualities).
-We looked at each period's topics, and assessed what the data -- consisting of the frequency-weighted word lists and messages -- suggested was the topic for that period.
-We were able to pinpoint the appropriate label using auxiliary information as well, such as the actual revisions and files that were related to this topic.
-For example, for the MaxDB topic consisting of a message ``exit() only used in non NPTL LINUX Versions'', we tagged that topic \emph{portability}.
-We compared against this data-set, but we also used it for our supervised learning based topic classification.
-%inter-rater reliability
-
-We first compared our previous analysis using label matching to our manual classifications to get an error rate for that process described below in Section \ref{sec:unsuplabelling}.
-
-For supervised learning, we used a suite of supervised classifiers from WEKA~\cite{weka09}.
-WEKA contains a suite of machine learning tools such as support vector machines and Bayes nets.
-We also used the multi-labelling add-on for WEKA, Mulan~\cite{mulan}\footnote{\url{http://mlkd.csd.auth.gr/multilabel.html}}.
-Traditional classifiers map our topics to a single class, whereas Mulan allows for a mixture of classes per topic, which maps to what we observed while manually labelling topics.
-
-To assess the performance of the supervised learners, we did a 10-fold cross-validation.
-This is when a set is partitioned into 10 partitions and then each partition is used once as a test set and 9 other times as part of the training set of 9 partitions.
-We have reported these results below in Section \ref{sec:suplabelling}.
-
-Finally, using this data, we evaluated two research questions (see
-Section \ref{sec:comparison}):
-\begin{enumerate}
-\item Do label frequencies change over time? That is, is a certain quality of more interest at one point in the life-cycle than some other? 
-\item  Do the different projects differ in their relative topic interest? That is, is a particular quality more important to one project than the other projects?  
-\end{enumerate}
-
-\section{Observations and evaluation}
-
-\subsection{Word list similarity}
-
+\subsection{Unsupervised labelling}
 % word list similarity
-
-In general, this approach did not work out well as common labels
-dominated the less common labels. The related words for
-\emph{correctness}, for example, tended to be too lengthy and
-non-specific. 
-Table \ref{tbl:wordlist} lists results. 
+Using our three word lists, we labeled our topics where there was a match between the list and the topic.
+In general, this approach did not work out well as common labels dominated the less common labels.
+The related words for \emph{correctness}, for example, tended to be too lengthy and non-specific.
+Table \ref{tbl:wordlist} lists broad labeling results for MaxDB as an example.
 A \emph{named topic} is a topic with a matching label.
@@
 
-%\begin{comment}
-%\XXX{this table should have mysql}
+\todo[inline]{this table should have mysql}
 \begin{table}
@@
 \midrule
-Topics &			500		 &			500  	 & 		500  \\
 Named topics   & 281 & 125 & 328  \\
@@
 \end{table}
-%\end{comment}
-%This is the counts
 
@@
 
-\subsection{Analysis of the unsupervised labelling}
+\subsubsection{Analysis of the unsupervised labelling}
 
@@
 Figure \ref{fig:maxdb-unsup-results} shows our results for MaxDB and MySQL.
-In general results are poor.
-Total ROC was slightly better than 0.5, and usability seems relatively accurate.
+Our error is measured using the ROC area-under-curve value.
+This value is the area under the Receiver Operating Characteristic (\emph{ROC}) curve, sometimes referred to as \emph{AUC}.
+ROC values reflect a score, similar to school letter-grades (A is 0.9, C is 0.6), reflecting how well a particular learner performed for the given data.
+A ROC result of 0.5 would be equivalent to a random learner (randomly classifying data).
+ROC maps to the more familiar concepts of precision/sensitivity and recall/specificity: it plots the true positive rate (sensitivity) versus the false positive rate (1 - specificity).
+A perfect learner has a ROC value of 1.0, reflecting perfect recall and precision.
 
@@
   \centering
- \includegraphics[width=.55\textwidth]{roc-unsup-pretty}
+ \includegraphics[width=.55\textwidth]{unsupervised-bar}
  \caption{ROC values for each NFR, per project and wordlist.}
@@
 
-% Using the F-Measure, the weighted average of precision and recall, where 1 is perfect, our best results are 0.6, a few at 0.4, and most around 0.2.
-We achieved similar results using the Matthew's correlation coefficient (used to measure efficacy where classes are of different sizes) and ROC.
-
-% \begin{table*}%[ht]
-% 	\centering
-
-% \begin{tabular}{c|c|c|c|c|c|c}
-% \toprule
-% Experiment & Label & F1 & MCC & Precision & Recall & ROC  \\ 
-% \midrule
-% MaxDB exp2 & portability & 0.228 & 0.182 & 0.520 & 0.146 & 0.553 \\ 
-%  & efficiency & 0.217 & 0.125 & 0.237 & 0.200 & 0.558\\ 
-%  & reliability & 0.380 & 0.340 & 0.246 & 0.829 & 0.765 \\ 
-%  & functionality & 0.095 & 0.083 & 0.250 & 0.059 & 0.521 \\ 
-%  & maintainability & 0.092 & 0.123 & 0.571 & 0.050 & 0.520 \\ 
-%  & usability & 0.175 & 0.138 & 0.113 & 0.389 & 0.620 \\ 
-%  & total & 0.236 & 0.127 & 0.248 & 0.225 & 0.561 \\ 
-% \midrule
-% MySQL exp2 & portability & 0.138 & 0.211 & 1.000 & 0.074 & 0.537 \\ 
-%  & efficiency & 0.345 & 0.327 & 0.476 & 0.270 & 0.625 \\ 
-%  & reliability & 0.425 & 0.287 & 0.348 & 0.545 & 0.669 \\ 
-%  & functionality & 0.025 & 0.006 & 0.571 & 0.013 & 0.501 \\ 
-%  & maintainability & 0.000 & 0.000 & 0.000 & 0.000 & 0.500 \\ 
-%  & usability & 0.175 & 0.135 & 0.200 & 0.156 & 0.560 \\ 
-%  & total & 0.167 & 0.095 & 0.403 & 0.105 & 0.527 \\ 
-% \midrule
-% MaxDB exp3 & portability & 0.472 & 0.286 & 0.402 & 0.573 & 0.660 \\ 
-%  & efficiency & 0.223 & 0.068 & 0.130 & 0.778 & 0.549 \\ 
-%  & reliability & 0.257 & 0.196 & 0.149 & 0.927 & 0.652 \\ 
-%  & functionality & 0.236 & 0.187 & 0.138 & 0.824 & 0.665 \\ 
-%  & maintainability & 0.338 & 0.112 & 0.266 & 0.463 & 0.566 \\ 
-%  & usability & 0.108 & 0.094 & 0.057 & 0.944 & 0.595 \\ 
-%  & total & 0.258 & 0.093 & 0.160 & 0.671 & 0.568 \\ 
-% \midrule
-% MySQL exp3 & portability & 0.413 & 0.170 & 0.564 & 0.325 & 0.574 \\ 
-%  & efficiency & 0.158 & 0.105 & 0.089 & 0.703 & 0.608 \\ 
-%  & reliability & 0.388 & 0.240 & 0.260 & 0.758 & 0.660 \\ 
-%  & functionality & 0.652 & 0.240 & 0.655 & 0.649 & 0.620 \\ 
-%  & maintainability & 0.203 & 0.007 & 0.240 & 0.175 & 0.503 \\ 
-%  & usability & 0.105 & 0.013 & 0.057 & 0.688 & 0.513 \\ 
-%  & total & 0.362 & 0.076 & 0.284 & 0.499 & 0.544 \\
-% \bottomrule
-% \end{tabular}
-% 	\caption{Results for automatic topic labelling. F1 = f-measure, MCC = Matthew's correlation coeff., ROC = Area under ROC curve}
-% 	\label{tbl:maxdb-unsup-results}
-% \end{table*}
 Based on these results we find that reliability and usability worked well for MaxDB in \textsf{exp2} and better in \textsf{exp3}.
@@
 
-%\subsection{Classifying topics}
+\subsection{Supervised labelling}
+\label{sec:suplearn}
+Supervised labelling requires expert analysis of the correct word to assign to a topic.
+%In order to validate how effective these word-bag approaches to topic labelling would be we created a validation data set.
+ For MySQL 3.23 and MaxDB 7.500, we manually annotated each extracted topic in each period with the same NFR labels as \textsf{exp2} (software qualities).
+We looked at each period's topics, and assessed what the data -- consisting of the frequency-weighted word lists and messages -- suggested was the topic for that period.
+We were able to pinpoint the appropriate label using auxiliary information as well, such as the actual revisions and files that were related to this topic.
+For example, for the MaxDB topic consisting of a message ``exit() only used in non NPTL LINUX Versions'', we tagged that topic \emph{portability}.
+We compared against this data-set, but we also used it for our supervised learning based topic classification.
+A topic may have more than one matching keyword.
+\todo[inline]{inter-rater reliability }
 
-\subsection{Analysis of the supervised labelling}
+We used a suite of supervised classifiers from WEKA~\cite{weka09}.
+WEKA is a suite of machine learning tools such as support vector machines and Bayes nets.
+We also used the multi-labelling add-on for WEKA, Mulan~\cite{mulan}\footnote{\url{http://mlkd.csd.auth.gr/multilabel.html}}.
+Traditional classifiers map our topics to a single class, whereas Mulan allows for a mixture of classes per topic, which maps to what we observed while manually labelling topics.
+
+To assess the performance of the supervised learners, we did a 10-fold cross-validation~\cite{Kohavi1995}, a common technique for evaluating machine learners.
+The original data is partitioned randomly into ten subsamples.
+One subsample is used as a training set and evaluated on the other nine subsamples.
+This is repeated nine more times, with each subsample used once as the training set.
+We have reported these results below in Section \ref{sec:suplabelling}.
+
+\subsubsection{Analysis of the supervised labelling}
 \label{sec:suplabelling}
+Because our data-set was of word counts we expected Bayesian techniques, often used in spam filtering, to perform well.
+We tried other learners that WEKA~\cite{weka09} provides: rule learners, decision tree learners, vector space learners, and support vector machines.
+Figure \ref{fig:best-learn-per-tag} shows the performance of the best performing learner per label.
+We considered the best learner for a label to be the one which had the highest ROC value for that label.
+Figure \ref{fig:best-learn-per-tag} uses the ZeroR learner as a baseline, a common practice in machine learning.
+ZeroR naively chooses the largest category all of the time.
+It occasionally outperforms non-naive learners; for labels which are not as common, this is to be expected because any miscategorization will hurt accuracy.
+This is why we use ROC values, as they can better represent performance on labels which are not applicable to the majority of samples.
 
-We took our annotated data-set and applied supervised learners to it.
-Because our data-set was of word counts we expected Bayesian techniques, often used in spam filtering, to perform well.
-We also tried other learners that WEKA~\cite{weka09} provides: rule learners, tree learners, vector space learners, and support vector machines.
-Table \ref{tab:best-learn-per-tag} shows the performance of the best performing learner per label.
-We considered the best learner for a label to be the one which had the highest ROC value for that label.
-Table \ref{tab:best-learn-per-tag} uses the ZeroR learner as a baseline, since it naively chooses the largest category all of the time.
-The ZeroR difference is often negative.
-For labels which are not as common, this can be expected because any miscategorization will hurt accuracy.
-This is why the F1 (F-measure) and the ROC values are useful, as they can better present performance on labels which are not applicable to the majority of samples.
-
-Table \ref{tab:best-learn-per-tag} shows that MaxDB and MySQL have quite different results, as the ROC values for reliability and functionality seem swapped between projects.
-It should be noted that for both projects Bayesian techniques did the best out of a wide variety of machine learners tested.
+Figure \ref{fig:best-learn-per-tag} shows that MaxDB and MySQL have quite different results, as the ROC values for reliability and functionality seem swapped between projects.
+For both projects Bayesian techniques did the best out of a wide variety of machine learners tested.
 Discriminative Multinomial Naive Bayes (DMNBtext), Naive Bayes (NaiveBayes) and Multinomial Naive Bayes (NaiveBayesMultinomal) are all based on Bayes's theorem and all assume, naively, that the features are independent.
@@
 
-The smaller the label the harder it is to get accurate results.
+The less-frequently occurring a label the harder it is to get accurate results, due to the high noise level.
 Nevertheless, these results are better than our previous word bag results of \textsf{exp2} and \textsf{exp3}, because the ROC values are sufficiently higher in most cases (other than MaxDB reliability and MySQL efficiency).
-
 The limitation of the approach we took here is that we assume labels are independent; however, labels could be correlated with each other.
@@
 
-% Cribbed from: 
-% src/validate/latex-out/best-learner-per-project-roc-table.txt
-\begin{table*}
+\begin{figure*}[ht]
 \centering
-\begin{tabular}{ccc|rrrr}
-\toprule
-Label & Project & Learner & ROC & ZeroR Acc. & ZeroR. Diff & F1\\
-\midrule
-
-portability &  MySQL &  NaiveBayesMultinomial &  0.74 &  58.53 &  11.43 &  0.62 \\
-efficiency &  MySQL &  NaiveBayes &  0.67 &  93.69 &  -7.68 &  0.23 \\ 
-reliability &  MySQL &  NaiveBayes &  0.73 &  83.11 &  -12.80 &  0.41 \\ 
-functionality &  MySQL &  DMNBtext &  0.81 &  54.44 &  21.67 &  0.77  \\ 
-maintainability &  MySQL &  DMNBtext &  0.78 &  76.62 &  3.41 &  0.32  \\ 
-usability &  MySQL &  NaiveBayes &  0.75 &  94.54 &  -5.80 &  0.21  \\ 
- \hline 
-portability &  MaxDB &  NaiveBayes &  0.84 &  77.12 &  2.06 &  0.61 \\ 
-efficiency &  MaxDB &  NaiveBayes &  0.62 &  88.43 &  -11.31 &  0.25  \\ 
-reliability &  MaxDB &  DMNBtext &  0.84 &  89.46 &  3.86 &  0.57  \\ 
-functionality &  MaxDB &  NaiveBayes &  0.67 &  91.26 &  -6.94 &  0.31  \\ 
-maintainability &  MaxDB &  NaiveBayes &  0.70 &  79.43 &  -9.25 &  0.42  \\ 
-usability &  MaxDB &  NaiveBayes &  0.56 &  95.37 &  -4.37 &  0.00  \\ 
- \hline
-
-\end{tabular}
-\caption{Per label, per project, the best learner for that label.
-ROC value rates learner performance, compared with the ZeroR learner ( a learner which just chooses the largest category all of the time).
-F1 is the F-measure for that particular learner.}
-\label{tab:best-learn-per-tag}
-\end{table*}
+\subfigure[MySQL]{
+\includegraphics[width=0.4\textwidth]{mysql-supervised}
+\label{fig:subfig1}
+}
+\subfigure[MaxDB]{
+\includegraphics[width=0.4\textwidth]{maxdb-supervised}
+\label{fig:subfig2}
+}
+\label{fig:best-learn-per-tag}
+\caption[]{ROC value (black) for the best learner per label, compared with ZeroR accuracy values (grey).}
+\end{figure*}
 
 \subsection{Applying multiple labels to topics}
-
-We applied the Mulan~\cite{mulan} library for Multi-Label learning to our data-set because intuitively, topics can have more than one label, much like how a particular source code revision can have more than one topic.
-Multi-label learning is more than just classifying entities with more than one label.
-It also includes methods for determining the performance of such techniques.
-The problem framed in the learners above has changed; instead of looking at the precision and recall of applying one label, we rank multiple labels at once.
+As noted in Section \ref{sec:wordlist}, each topic in our dataset can be composed of zero to all NFRs.
+For example, a commit message might address \textit{reliability} in the context of \textit{efficiency}, or make a \textit{maintainability} improvement in the source code that related to \textit{usability}.
+However, traditional machine learning techniques, such as Bayesian networks, can only map topics to a single class.
+The Mulan~\cite{mulan} library encapsulates several different machine learners which can handle this multi-label learning for our data-set.
+Mulan also includes methods for determining the performance of such techniques.
+% TOO DETAILED The problem framed in the learners above has changed; instead of looking at the precision and recall of applying one label, we rank multiple labels at once.
 We must check if the full subset of labels was applied, and then how much of that subset was applied.
 
-Another aspect of multi-label learning are micro versus macro measurements.
+In multi-label learning the technique can perform micro or macro measurements.
 Macro measurements are aggregated at a class or label level.
@@
 
-We have presented the results of Mulan's multi-label learners in Table \ref{tab:mulan}.
-Binary Relevance (BR), Calibrated Label Ranking (CLR) and Hierarchy Of Multi-label classifiERs (HOMER), performed the best.
-HOMER and BR act as a hierarchy of learners: BR is flat, while HOMER tries to build a deeper hierarchy to build a more accurate learner~\cite{mulan}.
+Fig. \ref{fig:mulan} presents the results of Mulan's best multi-label learners for our data. %Calibrated Label Ranking (CLR)
+The Hierarchy Of Multi-label classifiERs (HOMER) and Binary Relevance (BR) act as a hierarchy of learners: BR is flat, while HOMER tries to build a deeper hierarchy to build a more accurate learner~\cite{mulan}.
 These classifiers performed better than other multi-label classifiers.
-They have the best micro and macro ROC scores, although their results seem comparable to the naive Bayesian learners we used in Section \ref{sec:suplabelling}.
+They have the best micro and macro ROC scores, although their results are comparable to the naive Bayesian learners we used in Section \ref{sec:suplearn}.
+\todo[inline]{discuss this similiarity in more detail}
+\todo[inline]{what of CLR? Is it hierarchical?}
 
-\begin{comment}
+\begin{figure*}[ht]
+\centering
+\subfigure[MySQL]{
+\includegraphics[width=0.4\textwidth]{multi-label-results-mysql}
+\label{fig:subfig3}
+}
+\subfigure[MaxDB]{
+\includegraphics[width=0.4\textwidth]{multi-label-results-maxdb}
+\label{fig:subfig4}
+}
+\label{fig:mulan}
+\caption[]{MySQL and MaxDB macro (grey) and micro (black) ROC results per multi-label learner.}
+\end{figure*}
 
-\begin{table*}
-\centering
-\begin{tabular}{l|rrrrrrrr}
-
-  & BR & CLR & HOMER & IncludeLabels & LP & MLStacking & MLkNN & RAkEL \\
-
-\hline \\
-
-example Accuracy & 0.45 & 0.40 & 0.50 & 0.40 & 0.40 & 0.38 & 0.33 & 0.48 \\
-
-example F1 & 0.56 & 0.48 & 0.59 & 0.47 & 0.49 & 0.46 & 0.40 & 0.55 \\
-
-% example Hamming loss & 0.24 & 0.19 & 0.18 & 0.19 & 0.23 & 0.19 & 0.20 & 0.17 \\
-
-example Precision & 0.50 & 0.51 & 0.60 & 0.52 & 0.52 & 0.50 & 0.44 & 0.60 \\
-
-example Recall & 0.65 & 0.46 & 0.58 & 0.44 & 0.47 & 0.43 & 0.37 & 0.52 \\
-
-example Subset Accuracy & 0.19 & 0.24 & 0.31 & 0.24 & 0.23 & 0.23 & 0.20 & 0.31 \\
-
-label macro-ROC & 0.74 & 0.66 & 0.64 & 0.60 & 0.56 & 0.58 & 0.59 & 0.67 \\
-
-label macro-F1 & 0.46 & 0.30 & 0.43 & 0.28 & 0.32 & 0.27 & 0.22 & 0.39 \\
-
-label macro-Precision & 0.41 & 0.41 & 0.49 & 0.39 & 0.37 & 0.34 & 0.34 & 0.52 \\
-
-label macro-Recall & 0.56 & 0.26 & 0.40 & 0.24 & 0.31 & 0.23 & 0.19 & 0.33 \\
-
-label micro-ROC & 0.81 & 0.81 & 0.77 & 0.77 & 0.66 & 0.80 & 0.80 & 0.79 \\
-
-label micro-F1 & 0.59 & 0.53 & 0.61 & 0.51 & 0.50 & 0.51 & 0.45 & 0.59 \\
-
-label micro-Precision & 0.51 & 0.69 & 0.66 & 0.71 & 0.53 & 0.70 & 0.69 & 0.73 \\
-
-label micro-Recall & 0.68 & 0.43 & 0.56 & 0.41 & 0.48 & 0.41 & 0.34 & 0.49 \\
-
-% rank Avg. Precision & 0.76 & 0.77 & 0.69 & 0.72 & 0.53 & 0.74 & 0.76 & 0.73 \\
-
-% rank Coverage & 1.51 & 1.47 & 1.94 & 1.72 & 2.99 & 1.57 & 1.54 & 2.00 \\
-
-% rank One-error & 0.40 & 0.39 & 0.47 & 0.44 & 0.63 & 0.43 & 0.40 & 0.36 \\
-
-% rank Ranking Loss & 0.19 & 0.18 & 0.27 & 0.22 & 0.52 & 0.20 & 0.19 & 0.28 \\
-
-\end{tabular}
-\caption{Mysql Mulan results per Mulan learner}
-\end{table*}
-
-\begin{table*}
-% Cribbed from:
-% src/validate/output/maxdb.mulan.out.tex
-\centering
-\begin{tabular}{l|rrrrrrrr}
-
-  & BR & CLR & HOMER & IncludeLabels & LP & MLStacking & MLkNN & RAkEL \\
-
-\hline \\
-
-example-Accuracy & 0.49 & 0.47 & 0.51 & 0.44 & 0.42 & 0.47 & 0.43 & 0.48 \\
-
-example-Fmeasure & 0.55 & 0.50 & 0.53 & 0.45 & 0.44 & 0.48 & 0.43 & 0.50 \\
-
-example-HammingLoss & 0.19 & 0.14 & 0.14 & 0.14 & 0.17 & 0.13 & 0.13 & 0.13 \\
-
-example-Precision & 0.50 & 0.49 & 0.53 & 0.45 & 0.45 & 0.49 & 0.43 & 0.51 \\
-
-example-Recall & 0.61 & 0.50 & 0.53 & 0.45 & 0.43 & 0.47 & 0.43 & 0.50 \\
-
-example-SubsetAccuracy & 0.40 & 0.43 & 0.45 & 0.43 & 0.38 & 0.43 & 0.43 & 0.43 \\
-
-label-macro-AUC & NaN & NaN & NaN & NaN & NaN & NaN & NaN & NaN \\
-
-label-macro-F1 & 0.32 & 0.19 & 0.23 & 0.11 & 0.19 & 0.12 & 0.00 & 0.26 \\
-
-label-macro-Precision & 0.29 & 0.31 & 0.32 & 0.21 & 0.23 & 0.18 & 0.00 & 0.40 \\
-
-label-macro-Recall & 0.38 & 0.15 & 0.20 & 0.08 & 0.17 & 0.10 & 0.00 & 0.23 \\
-
-label-micro-AUC & 0.76 & 0.66 & 0.62 & 0.55 & 0.62 & 0.68 & 0.65 & 0.68 \\
-
-label-micro-F1 & 0.39 & 0.27 & 0.34 & 0.13 & 0.25 & 0.21 & 0.00 & 0.34 \\
-
-label-micro-Precision & 0.34 & 0.42 & 0.49 & 0.33 & 0.30 & 0.49 & 0.00 & 0.49 \\
-
-label-micro-Recall & 0.47 & 0.21 & 0.26 & 0.09 & 0.22 & 0.14 & 0.00 & 0.27 \\
-
-rank-AvgPrecision & 0.54 & 0.57 & 0.54 & 0.40 & 0.40 & 0.60 & 0.58 & 0.51 \\
-
-rank-Coverage & 1.40 & 1.23 & 1.43 & 1.84 & 1.93 & 1.16 & 1.20 & 1.54 \\
-
-rank-One-error & 0.81 & 0.79 & 0.81 & 0.90 & 0.89 & 0.76 & 0.79 & 0.79 \\
-
-rank-Ranking Loss & 0.41 & 0.35 & 0.41 & 0.58 & 0.60 & 0.32 & 0.33 & 0.46 \\
-
-\end{tabular}
-
-\caption{MaxDB Mulan results per Mulan learner}
-
-\end{table*}
-
-\end{comment}
-
-\begin{table*}%[h]
-\centering
-\begin{tabular}{l|ccc|ccc}
-\toprule
-  & MySQL & MySQL & MySQL &                         MaxDB & MaxDB & MaxDB  \\ 
-Performance Metric  & BR & CLR & HOMER &                              BR & CLR & HOMER \\
-
-\midrule
-%\hline \\
-
-% example Accuracy & 0.45 & 0.40 & 0.50 &           0.49  &  0.47  &  0.51 \\
-
-% example F1 & 0.56 & 0.48 & 0.59 &           0.55  &  0.50  &  0.53 \\
-
-% example Hamming Loss & 0.24 & 0.19 & 0.18 &        0.19  &  0.14  &  0.14 \\
-
-% example Precision & 0.50 & 0.51 & 0.60 &          0.50  &  0.49  &  0.53 \\
-
-% example Recall & 0.65 & 0.46 & 0.58 &             0.61  &  0.50  &  0.53 \\
-
-example Subset Accuracy & 0.19 & 0.24 & 0.31 &     0.40  &  0.43  &  0.45 \\ 
-      \hline                                                                      
-label macro-ROC & 0.74 & 0.66 & 0.64 &            NaN  &  NaN  &  NaN    \\
-
-label macro-F1 & 0.46 & 0.30 & 0.43 &             0.32  &  0.19  &  0.23 \\
-
-label macro-Precision & 0.41 & 0.41 & 0.49 &      0.29  &  0.31  &  0.32 \\
-
-label macro-Recall & 0.56 & 0.26 & 0.40 &         0.38  &  0.15  &  0.20 \\
-
-label micro-ROC & 0.81 & 0.81 & 0.77 &            0.76  &  0.66  &  0.62 \\
-
-label micro-F1 & 0.59 & 0.53 & 0.61 &             0.39  &  0.27  &  0.34 \\
-
-label micro-Precision & 0.51 & 0.69 & 0.66 &      0.34  &  0.42  &  0.49 \\
-
-label micro-Recall & 0.68 & 0.43 & 0.56 &         0.47  &  0.21  &  0.26 \\ 
-       \hline                                                                      
- rank Avg. Precision & 0.76 & 0.77 & 0.69 &          0.54  &  0.57  &  0.54 \\
-
- rank Coverage & 1.51 & 1.47 & 1.94 &              1.40  &  1.23  &  1.43 \\
-
- rank One-error & 0.40 & 0.39 & 0.47 &             0.81  &  0.79  &  0.81 \\
-
- rank Ranking Loss & 0.19 & 0.18 & 0.27 &          0.41  &  0.35  &  0.41 \\ 
-\hline
-\end{tabular}
-\caption{MySQL and MaxDB Mulan results per Multi-Label learner}
-\label{tab:mulan}
-\end{table*}
+\section{Discussion}
+Using the classification results from the previous sections, we evaluated two research questions:
+\begin{enumerate}
+\item Do label frequencies change over time? That is, is a certain quality of more interest at one point in the life-cycle than some other? 
+\item  Do the different projects differ in their relative topic interest? That is, is a particular quality more important to one project than the other projects?  
+\end{enumerate}
 
 \subsection{Summary of techniques}
+While unsupervised techniques (LSI and LDA are both unsupervised) are appealing in their lack of human intervention, and thus lower effort, supervised learners have the advantage of domain knowledge which typically means improved results.
 
-Very rarely did \textsf{exp2} and \textsf{exp3} (naive word matching) ever perform as well as the machine learners.
-For MaxDB, reliability was slightly better detected using the static word list of \textsf{exp2}.
+Very rarely did \textsf{exp2} and \textsf{exp3} (naive word matching) ever perform as well as the supervised machine learners.
+For MaxDB, \textit{reliability} was slightly better detected using the static word list of \textsf{exp2}.
 In general, the machine learners and \textsf{exp3} did better than \textsf{exp2} for both MaxDB and MySQL.
-For both MySQL and MaxDB usability was better served by \textsf{exp2}.
-Usability was a very infrequent label, however.
+For both MySQL and MaxDB \textit{usability} was better served by \textsf{exp2}.
+\textit{Usability} was a very infrequent label, however, which made it difficult to detect in any case.
 
-We found that the multi-label learners of BR, CLR and HOMER did not do as well for Macro-ROC and Micro-F1 as NaiveBayes and other NaiveBayes derived learners did.
+We found that the multi-label learners of BR, CLR and HOMER did not do as well for Macro-ROC as NaiveBayes and other NaiveBayes derived learners.
 This suggests that by sticking together multiple NaiveBayes learners we could probably label sets of topics effectively, but it would require a separate NaiveBayes learner per label.
 
-\begin{comment}
-MySQL Macro ROC 0.746 Micro F1 0.78 Mulan BR ROC 0.74 / 0.81 F1 0.46/0.59/0.56 
-MaxDB Macro ROC 0.7   Micro F1 0.83 Mulan BR ROC NaN  / 0.76 F1 0.32/0.39/0.55
-
-\end{comment}
-
-\begin{comment}
-           exp2 exp3 learn  B 2nd
-MaxDB port 0.55 0.66 0.74   L 3
-MaxDB eff  0.55 0.64 0.67   L 3
-MaxDB rel  0.76 0.65 0.73   2 2
-MaxDB fun  0.52 0.66 0.81   L 3
-MaxDB main 0.52 0.56 0.78   L 3
-MaxDB use  0.62 0.59 0.75   L 2
-
-mysql port 0.53 0.57 0.85   L 3
-mysql eff  0.62 0.60 0.62   2 2 
-mysql rel  0.66 0.66 0.84   L 3
-mysql fun  0.5  0.62 0.67   L 3
-mysql main 0.5  0.5  0.70   L 3
-mysql use  0.56 0.51 0.56   L 2
-\end{comment}
-
 \subsection{Visualization}
-
-We have created two visualizations of the manually labelled data.
+We created two visualizations of the manually labelled data.
 A problem we faced while visualizing was how to display tag overlaps.
@@
 
+\section{Limitations}
 \subsection{Effectiveness}
@@
 \subsection{Threats to validity}
-
-\begin{comment}
-\textbf{Construct validity} -- Focus on code commit messages, rather than mail or bug trackers. Choice of taxonomy
-
-\textbf{Internal validity} -- inter-rater reliability.
-
-\textbf{External validity} -- Our data originated from open-source projects in the database domain.
-Of these, the open-source nature of the project seems most problematic for external validity.
-However, both projects are developed by commercial entities.
-Our technique is dependent on development practises including meaningful commit messages, which might not be true in other domains.
-\end{comment}
-
-Our study suffers from multiple threats to validity.
 \emph{Construct validity} issues include that we used only commit messages rather than mail or bug tracker messages.
@@
 
-% Most of this related research examined project source code.
-Software mining of other project corpora is less common.
-Similarities exist with approaches that begin with structured taxonomies, as with the Hismo software ontology~\cite{girba06}.
-Rigby and Hassan~\cite{rigby07msr} used a general purpose taxonomy to classify developer email according to temperament.
-
 \section{Conclusions and future work}

### 2010-08-16 src/validate/quality_map.py
generating topic counts by NFR. Adjusted quality map to map Abram's names into 6 NFR types.
Signed-off-by: Neil Ernst <neil@neilernst.net>

@@ -3,33 +3,34 @@ import names as T
 t = T.Taxonomy()
 
 def quality_map():
-    quality_map = {'portability':[], 'efficiency':[], 'reliability':[], 'functionality':[], 'usability':[], 'maintainability':[]}
+    quality_map = {'none':[], 'portability':[], 'efficiency':[], 'reliability':[], 'functionality':[], 'usability':[], 'maintainability':[]}
     return quality_map
 
 
     
 #    ( "functionality" 0, "correctness" 1, "portability" 2, "maintainability" 3, "documentation" 4, "reliability" 5, "bug" 6, "testability" 7, "cleanup" 8, "unknown" 9, "optimization" 10, "efficiency" 11, "usability" 12, "ui" 13, "version control" 14, "feature" 15, "build" 16, "logging" 17, "modular" 18, "error" 19, "license" 20, "security" 21, "token replacement" 22, "version" 23, "integrity" 24, "concurrency" 25, "conformance" 26, "test" 27, "rename" 28, "distribution" 29, "compliance" 30, "standardized" 31, "standard" 32, "stability" 33, "specificiation" 34, "patch" 35, "interoperability" 36, "internationalization" 37, "interface" 38, "fault tolerance" 39, "external" 40, "accuracy" 41, "Understandability" 42 )
 
+""" Handle some of Abram's unorthodox annotations """ 
 def ext_annotation_map():
     am = {}
     am["cleanup"] = "maintainability"
-    am["unknown"] = ""
-    am["version control"] = ""
-    am["logging"] =""
-    am["license"] = ""
+    am["unknown"] = "none"
+    am["version control"] = "maintainability"
+    am["logging"] ="reliability"
+    am["license"] = "maintainability"
     am["token replacement"] = ""
-    am["version"] = ""
-    am["integrity"] = ""
-    am["concurrency"] = ""
-    am["test"] = ""
+    am["version"] = "maintainability"
+    am["integrity"] = "reliability"
+    am["concurrency"] = "efficiency"
+    am["test"] = "reliability"
     am["rename"] = "" #not sure here
-    am["distribution"] = ""
-    am["patch"] = ""
-    am["external"] = ""
-    am["accuracy"] = "",
+    am["distribution"] = "maintainability"
+    am["patch"] = "reliability"
+    am["external"] = "none"
+    am["accuracy"] = "functionality",
     am["Understandability"] = "understandability"
-    am["build"] = ""
-    am["none"] = ""
[…]

### 2010-08-16 src/validate/summary_labels.py

@@ -1,0 +1,42 @@
+""" Create a CSV of the project, period, topics, labels """
+import check_validity as cv
+import quality_map as Q
+import get_exps
+import lxml.etree
+import datetime
+
+""" Look something like this:
+Project, Period, Topic-#, Labels
+Mysql, 10343244, 2, Usability, Maintainability
+"""
+if __name__ == '__main__':
+    maxdb_file = 'label-summary.csv'
+    maxdb = open(maxdb_file,'w')
+    
+    exps = get_exps.get_exps() 
+    # test_period = '1098931753'
+    #     topics = load_period(test_period)
+    for project in exps: #['name', 'dir', [periods]]
+        str_write = project[0] + ',' # project name
+        periods_map = {}
+        for period in project[2]:
+            p_date = datetime.datetime.fromtimestamp(float(period)).strftime("%Y-%m-%d")
+            # maintain a list of qualities (counts) per period
+            quality_map = {'none':0,'portability':0, 'efficiency':0, 'reliability':0, 'functionality':0, 'usability':0, 'maintainability':0}
+            topics = cv.load_period(period, project[1])
+            i = 0
+            for topic in topics:
+                annotations = []
+                anno = topic.findall('Annotation') or topic.findall('annotation')
+                for a in anno:
+                    annotations.append(a.attrib['name'])
+                    annotations2 = Q.annotations_to_qualities( annotations ) # a converted list of annotations in that topic
+                    for q in annotations2:
+                        if q in quality_map.keys():
+                            quality_map[q] += 1
+            periods_map[p_date] = quality_map
+        for p in periods_map.keys():
+           # print  str(p) + ',' + ','.join([str(x) for x in periods_map[p].values()])
+            maxdb.write(project[0] + ',' + str(p) + ',' + ','.join([str(x) for x in periods_map[p].values()]) + '\n')
+            
+

### 2010-08-16 icse/icse2011.tex
updates to discussion section

@@
 For these projects, developers interact with abstractions of these artifacts, rather than the artifact itself.
-Such abstractions include project dashboards~\cite{}, tags (Peggy/Christoph), and complexity metrics (\cite{mccabe}).
+Such abstractions include project dashboards~\cite{kersten2005mylar}, tags~\cite{treude2010}, and complexity metrics \cite{mccabe1976complexity}.
 
-Developers do more than merely observe these abstractions, however \cite{}.
+Developers do more than merely observe these abstractions, however.
 They must also use them to assess progress and work with colleagues.
@@
 \subsection{Generating the data}
+\label{sec:wordlist}
 To evaluate our approach, we sought candidate systems that were mature projects and had openly accessible source control repositories.
@@
 
-\section{Discussion}
-Using the classification results from the previous sections, we evaluated two research questions:
-\begin{enumerate}
-\item Do label frequencies change over time? That is, is a certain quality of more interest at one point in the life-cycle than some other? 
-\item  Do the different projects differ in their relative topic interest? That is, is a particular quality more important to one project than the other projects?  
-\end{enumerate}
-
 \subsection{Summary of techniques}
@@
 This suggests that by sticking together multiple NaiveBayes learners we could probably label sets of topics effectively, but it would require a separate NaiveBayes learner per label.
+
+\section{NFRs and evolution} 
+\todo[inline]{talk about this topic in one section}
+\todo[inline]{reference project name Nomen more}
+Using the classification results from the previous sections, we evaluated two research questions:
+\begin{enumerate}
+\item Do label frequencies change over time? That is, is a certain quality of more interest at one point in the life-cycle than some other? 
+\item  Do the different projects differ in their relative topic interest? That is, is a particular quality more important to one project than the other projects?  
+\end{enumerate}
 
@@
 \end{figure*}
+
+\section{NFRs between projects}
 

### 2010-08-16 icse/icse2011.tex
updates for timeline view

@@
 MySQL contains 320 thousand lines of C and C++ source code.
-\todo[inline]{why 3.23 and not newer?}
+Choosing an older version of these projects allows us to focus on projects which have moved into the maintenance phase of the software lifecycle.
 
@@
   %\includegraphics[width=.97\textwidth]{mysql-time-smear-plot-Equal}
-  \includegraphics[width=.97\textwidth]{small-mysql-time-smear-plot-Equal}
   %\includegraphics[width=.97\textwidth]{small-mysql-time-smear-plot-Equal}
-  \caption{MySQL 3.23: Labelled topics related to non-functional software qualities plotted over time, continuous topics are trends that occur across time}
+  \includegraphics[width=.97\textwidth]{mysql-timeline}
+  \caption{MySQL 3.23:NFRs over time. Topic frequency is mapped to alpha (transparency) value.}
   \label{fig:mysql}

### 2010-08-16 src/validate/plot_timeline.py

@@ -1,0 +1,67 @@
+"""
+Plot timelines with no spaces. Input is a range of values per quality.
+"""
+import matplotlib.pyplot as plt
+from matplotlib.dates import YearLocator, MonthLocator, DateFormatter
+from matplotlib.patches import FancyBboxPatch, Rectangle
+
+import datetime
+
+def plot_timeline(proj, period_map):
+    """ Plot a timeline view a la ConcernLines to show trends in topic occurrence """
+    fig = plt.figure()
+    ax = fig.add_subplot(111)
+    # bar size
+    barwidth = 5
+    barheight = 5
+    ax.set_ylim(0,40)
+    ax.set_xlim(0,5*len(period_map.keys()))
+    ax.set_xlabel('Date')
+    ax.set_yticks([5,10,15,20,25,30,35])
+    ax.set_xticks([x*5 for x in range(len(period_map.keys())+1)])
+    ax.set_yticklabels(['None', 'Portability', 'Efficiency', 'Reliability', 'Functionality', 'Usability', 'Maintainability'])
+    ax.set_xticklabels([period_map.keys()])
+    ax.autoscale_view()
+    
+    # set intensity scale to 0-1 i.e. max-value is all out, the rest are in between
+    # this gives intensity relative to the highest value of all qualities. But we should probably normalize for a specific quality.
+    #max value is across all periods in that quality
+    #i am personally offended by this code
+    max_value_map = {'none':0,'portability':0, 'efficiency':0, 'reliability':0, 'functionality':0, 'usability':0, 'maintainability':0}
+    for qm in period_map.values():
+        #qm = period_map[p] # the quality map w/value
+        for q in qm.keys():
+            if qm[q] > max_value_map[q]: max_value_map[q] = qm[q]
+            
+    i = 0
+    for period in period_map.keys():
+        qm = period_map[period] # the quality map w/value
+        none_rect = Rectangle((i,2.5), barwidth,barheight,fill=True,lw=0, fc='red',alpha=(float(qm['none'])/float(max_value_map['none'])))
+        port_rect = Rectangle((i,7.5), barwidth,barheight,fill=True,lw=0, fc='red',alpha=(float(qm['portability'])/float(max_value_map['portability'])))
[…]

### 2010-08-16 src/validate/summary_labels.py

@@ -1,14 +1,13 @@
-""" Create a CSV of the project, period, topics, labels """
+""" Create a CSV of the project, period, topics, labels and plot a timeline"""
+
 import check_validity as cv
 import quality_map as Q
+import plot_timeline as pt
 import get_exps
+
 import lxml.etree
 import datetime
 
-""" Look something like this:
-Project, Period, Topic-#, Labels
-Mysql, 10343244, 2, Usability, Maintainability
-"""
 if __name__ == '__main__':
     maxdb_file = 'label-summary.csv'
     maxdb = open(maxdb_file,'w')
@@ -35,8 +34,11 @@ if __name__ == '__main__':
                         if q in quality_map.keys():
                             quality_map[q] += 1
             periods_map[p_date] = quality_map
-        for p in periods_map.keys():
-           # print  str(p) + ',' + ','.join([str(x) for x in periods_map[p].values()])
-            maxdb.write(project[0] + ',' + str(p) + ',' + ','.join([str(x) for x in periods_map[p].values()]) + '\n')
+        
+        pt.plot_timeline(project[0], periods_map)
+        
+        #for p in periods_map.keys():
+            # print  str(p) + ',' + ','.join([str(x) for x in periods_map[p].values()])
+            #maxdb.write(project[0] + ',' + str(p) + ',' + ','.join([str(x) for x in periods_map[p].values()]) + '\n')
             
 

### 2010-08-17 icse/icse2011.tex
Mild edits and todos inserted

@@
 \newcommand{\XXX}[1]{\textcolor{red}{{\it \textbf{[XXX: #1]}}}}
-\hyphenation{Max-DB}
+\hyphenation{Max-DB mysql\_-real\_-connect}
 
@@
 \begin{abstract}
+\todo[inline]{Topics and LDA sooner, Topics/concept location is a known
+  research area and we're saying they don't do enough}
   Within the field of mining software repositories, many approaches,   such as topic modeling and concept location, rely on the automated   application of machine learning algorithms to corpora.  
@@
 \section{Introduction}
+
 The act of developing software generates many artifacts, including source code, bug reports, commit messages, and test cases.
@@
 One way to identify topics is to use the artifacts themselves to derive them.
+\todo[inline]{it should be labels for topics not artifacts, I think the artifact part is almost secondary and the topic stuff is really what is up.
+Topics are used below and it isn't even clear what they are or what we're talking about.
+Naming is important but lets tell the reader what is being named first.}
 
@@
 To scale to meaningfully sized projects, we require automatic assistance to determine the topic label.
+\todo[inline]{Readers still don't know what topics are or what the point is yet..}
 
@@
   \item Topic mining of software has been limited to one project at a time.
-This is because traditional topic mining techniques are specific to a particular dataset.
+This is because traditional topic mining techniques are specific to a particular data-set.
 \textbf{Nomen} allows for comparisons \textit{between} projects.
@@
 
+\todo[inline]{Topics are poorly defined by this point and since we're building on research which didn't DO this, we should mention that research. Previous work probably should go here.}
+
 \section{Study design and execution}
@@
 We began by gathering source data; then applied unsupervised and supervised learning techniques to identify topics.
-We then use these topics to analyse the role of non-functional requirements in software maintenance.
+We then use these topics to analyze the role of non-functional requirements in software maintenance.
 
@@
 MySQL contains 320 thousand lines of C and C++ source code.
-Choosing an older version of these projects allows us to focus on projects which have moved into the maintenance phase of the software lifecycle.
+Choosing an older version of these projects allows us to focus on projects which have moved into the maintenance phase of the software life-cycle.
 
@@
 In this paper a topic is a set of tokens extracted from commit messages found within a project's source control system (SCS).
+\todo[inline]{These paragraphs should've come way earlier, in previous we should explain LDA so this doesn't need to be here.}
 
@@
 We performed simple string matching between these topics and our lists, `naming' a topic if it contained that word or words (a topic might match several keywords).
-We used several different word lists for comparison, which we call \textsf{exp1,exp2,exp3} in the text which follows.
+We used several different word lists for comparison, which we call \textsf{exp1, exp2, exp3} in the text which follows.
 
@@
 
+\todo[inline]{There should be a section here which is about manual annotation. It is really important and it is required by both analyses. Plus we can tell them how.}
+
 \subsection{Unsupervised labelling}
@@
  \caption{ROC values for each NFR, per project and wordlist.}
+\todo[inline]{this is too tiny in print}
+
   \label{fig:maxdb-unsup-results}
@@
 We also did not evaluate how well the learners performed together.
+\todo[inline]{Comment about correlation, we should link to the Mulan section, perfect segway since Mulan relies on correlation.}
 
@@
 \label{fig:best-learn-per-tag}
-\caption[]{ROC value (black) for the best learner per label, compared with ZeroR accuracy values (grey).}
+\caption[]{ROC value (black) for the best learner per label, compared with ZeroR accuracy values (grey).
+\todo[inline]{better discussion and graphical legend would be nice}
+}
 \end{figure*}
@@
 \subsection{Applying multiple labels to topics}
-As noted in Section \ref{sec:wordlist}, each topic in our dataset can be composed of zero to all NFRs.
+As noted in Section \ref{sec:wordlist}, each topic in our data-set can be composed of zero to all NFRs.
 For example, a commit message might address \textit{reliability} in the context of \textit{efficiency}, or make a \textit{maintainability} improvement in the source code that related to \textit{usability}.
@@
 They have the best micro and macro ROC scores, although their results are comparable to the naive Bayesian learners we used in Section \ref{sec:suplearn}.
-\todo[inline]{discuss this similiarity in more detail}
+\todo[inline]{discuss this similarity in more detail}
 \todo[inline]{what of CLR? Is it hierarchical?}
@@
 \label{fig:mulan}
-\caption[]{MySQL and MaxDB macro (grey) and micro (black) ROC results per multi-label learner.}
+\caption[]{MySQL and MaxDB macro (grey) and micro (black) ROC results per multi-label learner.
+\todo[inline]{better discussion and graphical legend would be nice}
+}
 \end{figure*}
@@
 Since this was MySQL 3.23, a stable branch, we expect that issues dealing with portability and maintenance would be the primary concerns of this branch: developers probably wanted it to work with current systems and thus had to update it.
+
+\todo[inline]{Update this}
 
@@
 
+%XXX can save space here!
 \section{Appendix}

### 2010-08-17 icse/icse2011.tex
fixed compile error on my part

@@
 \newcommand{\XXX}[1]{\textcolor{red}{{\it \textbf{[XXX: #1]}}}}
-\hyphenation{Max-DB mysql\_-real\_-connect}
+\hyphenation{Max-DB}
 
@@
 This is the most readily accessible source of project interactions for outside researchers; bug reports and email messages were not available for both projects.
-An example of a typical commit message is: \textit{``history annotate diffs bug fixed (if mysql\_real\_connect() failed there were two pointers to malloc'ed strings, with memory corruption on free(), of course)''}.
+An example of a typical commit message is: \textit{``history annotate diffs bug fixed (if mysql\-\_real\-\_connect() failed there were two pointers to malloc'ed strings, with memory corruption on free(), of course)''}.
 We extracted these messages and indexed them by creation time.

### 2010-08-17 icse/icse2011.tex
abram's remunging of the intro (which actually keeps most of the original content)

@@
 
-The act of developing software generates many artifacts, including source code, bug reports, commit messages, and test cases.
-It does not take long for a software project to grow to a point where keeping track of these artifacts exceeds easy comprehension.
-For these projects, developers interact with abstractions of these artifacts, rather than the artifact itself.
+When a developer addresses a topic of development, such as a requirement, maintenance, bug fixing, tests or other software engineering tasks, they often produce or modify multiple development artifacts.
+%
+These development artifacts include: source code, bug reports, revisions to source code, and even test cases. 
+%
+Topics that are addressed are sometimes not immediately resolved, some topics recur across time, yet others are only dealt with briefly.
+%
+The goal of concept location and topic analysis research is to semi-automatically recover some of these topics, and the related software artifacts.
+% get neil's data mining point in there
+The results of topic analysis and concept location are often very project specific, and not globally generalizable.
+% get the interpretation aspect
+As well much of the current research stops short of interpreting the purpose of the topic. That is, topics extracted are usually manually named and labelled by readers.
+% 
+Thus in this paper, we propose a method of interpreting and labelling some topics, and their related artifacts, that is generalizable and relevant to multiple projects.
+
+% % a rough sketch
+% The size of these artifacts is cumbersome for a developer or stakeholder to comprehend, thus they often have to be ``chunked'', grouped, modularized, organized and finally labelled.
+% Even more difficult is to pull together tangible trails or interests that pervade these artifacts.
+% These interests or topics are only really visible by analyzing multiple artifacts.
+% Other researchers have focused on extracting these topics from development artifacts, but often their approach leaves the organization or the labelling step to the user.
+% What we propose is to 
+% In this paper, we seek to automatically extract these interests, or topics that link artifacts such as source control commits and label them.
+
+% % Read for SE Data Mining
+% % Read for Topic Naming
+% % Read for previous research DID NOT
+% % thoughts, what does this do what are we introducing
+% % these abstractions ... how to link them
+
+% The act of developing software generates many artifacts, including source code, bug reports, commit messages, and test cases. 
+% It does not take long for a software project to grow to a point where keeping track of these artifacts exceeds easy comprehension. 
+% For these projects, developers interact with abstractions of these artifacts, rather than the artifact itself.
 Such abstractions include project dashboards~\cite{kersten2005mylar}, tags~\cite{treude2010}, and complexity metrics \cite{mccabe1976complexity}.
 
-Developers do more than merely observe these abstractions, however.
+% Developers do more than merely observe these abstractions, however.
 They must also use them to assess progress and work with colleagues.
@@
 One way to identify topics is to use the artifacts themselves to derive them.
-\todo[inline]{it should be labels for topics not artifacts, I think the artifact part is almost secondary and the topic stuff is really what is up.
+% \todo[inline]{it should be labels for topics not artifacts, I think the artifact part is almost secondary and the topic stuff is really what is up.
 Topics are used below and it isn't even clear what they are or what we're talking about.
@@
 
+% topics are numerous?
 Deriving such topics is difficult. 
+Raw topics from topic analysis are represented as unlabelled context-less word-lists or word distributions.
 Topics have to be identified by interpreting the prevalent words in a word distribution and by inspecting related documents. 
-This is impractical when one has to handle more than one hundred different topics.
+This manual approach is impractical when one has to handle more than one hundred different topics. 
 To scale to meaningfully sized projects, we require automatic assistance to determine the topic label.
-\todo[inline]{Readers still don't know what topics are or what the point is yet..}
+
+% why are topics useful
+Often developers interact with abstractions of these artifacts, rather than the artifact itself. 
+Topics are a valuable and automatically extractable abstraction of software artifacts.
+It abstracts these artifacts by their purpose or topic rather than their concrete representation.
+The topic of development is relevant because it shows anyone inspecting these topics what were the distinct issues or topics facing development at a certain time.
+Such abstractions include project dashboards~\cite{kersten2005mylar}, tags~\cite{treude2010}, and complexity metrics \cite{mccabe1976complexity}.
 
@@
 
+% harp on data mining more
+In general the data mining of software artifacts tends to be very project specific, yet NFRs are not. 
+NFRs are prevalent in almost all software systems.
+Thus our method of generalizing the data mining of these software artifacts is use the domain knowledge of software engineering, specifically NFRs. 
+This allows us to data mine software systems with the intention of comparing their NFR related development topics.
+% in order to achieve generalizable, cross project data mining results.
+
+%% RESEARCH DIDNT DO THIS AND WE ARE TOTALLY AWESOME AND WE DID
+% it is ok to repeat this here because it is clear and bullet pointed and the reviewers wont have any bloody excuse to not see it
 In this paper, we describe \textbf{Nomen}, which is a tool for \emph{labelled topic extraction}. \textbf{Nomen} addresses two gaps in the topic mining literature:
@@
 \textbf{Nomen} allows for comparisons \textit{between} projects.
-  \item Topic modeling creates bags of words that require interpretation by the user to assign meaning. Like (1), this means
+  \item Topic modeling creates word lists that require interpretation by the user to assign meaning. Like (1), this means
 that it is difficult to discuss results independent of the project context. Our technique automatically assigns labels agnostic of project.
@@
 
-\todo[inline]{Topics are poorly defined by this point and since we're building on research which didn't DO this, we should mention that research. Previous work probably should go here.}
+%\todo[inline]{Topics are poorly defined by this point and since we're building on research which didn't DO this, we should mention that research. Previous work probably should go here.}
 

### 2010-08-17 icse/icse2011.tex
Edit the intro again

@@
 
-When a developer addresses a topic of development, such as a requirement, maintenance, bug fixing, tests or other software engineering tasks, they often produce or modify multiple development artifacts.
+When a developer addresses a topic of development, such as a requirement, maintenance, bug fixes or other software engineering tasks, they often produce or modify multiple development artifacts.
 %
@@
 % get the interpretation aspect
-As well much of the current research stops short of interpreting the purpose of the topic. That is, topics extracted are usually manually named and labelled by readers.
+As well much of the current research stops short of interpreting the purpose of the topic. That is, extracted topics are usually manually named and labelled by those who read the topic words.
 % 
@@
 % why are topics useful
-Often developers interact with abstractions of these artifacts, rather than the artifact itself. 
+Often developers interact with abstractions of these artifacts that topics are derived from, rather than the artifact itself. 
 Topics are a valuable and automatically extractable abstraction of software artifacts.
 It abstracts these artifacts by their purpose or topic rather than their concrete representation.
-The topic of development is relevant because it shows anyone inspecting these topics what were the distinct issues or topics facing development at a certain time.
+The topic of development is relevant because it shows anyone inspecting these topics what are the distinct issues or topics facing development at a certain time.
 Such abstractions include project dashboards~\cite{kersten2005mylar}, tags~\cite{treude2010}, and complexity metrics \cite{mccabe1976complexity}.
@@
 NFRs are prevalent in almost all software systems.
-Thus our method of generalizing the data mining of these software artifacts is use the domain knowledge of software engineering, specifically NFRs. 
+Thus our method of generalizing the data mining of these software artifacts is to leverage  the domain knowledge of software engineering, specifically NFRs. 
 This allows us to data mine software systems with the intention of comparing their NFR related development topics.

### 2010-08-17 icse/icse2011.tex
Updates to figures. Created new figure directory for clarity and updated latex. Some figures made bigger, added legends.
Signed-off-by: Neil Ernst <neil@neilernst.net>

@@
   \centering
- \includegraphics[width=.45\textwidth]{process-model}
+ \includegraphics[width=.45\textwidth]{figures/process-model}
  \caption{Research methodology process view.}
@@
 \emph{Maintainability} &
-testability changeability analyzability stability maintain maintainable modularity modifiability understandability + interdependent dependency encapsulation decentralized modular\\ \hline
+testability changeability analyzability stability maintain maintainable modularity modifiability understandability interdependent dependency encapsulation decentralized modular\\ \hline
 \emph{Functionality} &
-security compliance accuracy interoperability suitability functional practicality functionality + compliant exploit certificate secured “buffer overflow” policy malicious trustworthy vulnerable vulnerability accurate secure vulnerability correctness accuracy\\ \hline
+security compliance accuracy interoperability suitability functional practicality functionality compliant exploit certificate secured ``buffer overflow'' policy malicious trustworthy vulnerable vulnerability accurate secure vulnerability correctness accuracy\\ \hline
 \emph{Portability} &
-conformance adaptability replaceability installability portable movableness movability portability + specification migration standardized l10n localization i18n internationalization documentation interoperability transferability\\ \hline
+conformance adaptability replaceability installability portable movableness movability portability specification migration standardized l10n localization i18n internationalization documentation interoperability transferability\\ \hline
 \emph{Efficiency} &
-“resource behaviour” “time behaviour” efficient efficiency + performance profiled optimize sluggish factor penalty slower faster slow fast optimization\\ \hline
+``resource behaviour'' ``time behaviour'' efficient efficiency performance profiled optimize sluggish factor penalty slower faster slow fast optimization\\ \hline
 \emph{Usability} &
-operability understandability learnability useable usable serviceable usefulness utility useableness usableness serviceableness serviceability usability + gui accessibility menu configure convention standard feature focus ui mouse icons ugly dialog guidelines click default human convention friendly user screen interface flexibility\\ \hline
+operability understandability learnability useable usable serviceable usefulness utility useableness usableness serviceableness serviceability usability gui accessibility menu configure convention standard feature focus ui mouse icons ugly dialog guidelines click default human convention friendly user screen interface flexibility\\ \hline
 \emph{Reliability} &
-“fault tolerance” recoverability maturity reliable dependable responsibleness responsibility reliableness reliability dependableness dependability + resilience integrity stability stable crash bug fails redundancy error failure\\
+``fault tolerance'' recoverability maturity reliable dependable responsibleness responsibility reliableness reliability dependableness dependability resilience integrity stability stable crash bug fails redundancy error failure\\
 \bottomrule
@@
 
-\todo[inline]{this table should have mysql}
+\todo[inline,color=green]{this table should have mysql}
 \begin{table}
@@
   \centering
- \includegraphics[width=.55\textwidth]{unsupervised-bar}
+ \includegraphics[width=.75\textwidth]{figures/unsupervised-bar}
  \caption{ROC values for each NFR, per project and wordlist.}
-\todo[inline]{this is too tiny in print}
+%\todo[inline]{this is too tiny in print}
 
@@
 \subfigure[MySQL]{
-\includegraphics[width=0.4\textwidth]{mysql-supervised}
+\includegraphics[width=0.4\textwidth]{figures/mysql-supervised}
 \label{fig:subfig1}
@@
 \subfigure[MaxDB]{
-\includegraphics[width=0.4\textwidth]{maxdb-supervised}
+\includegraphics[width=0.4\textwidth]{figures/maxdb-supervised}
 \label{fig:subfig2}
@@
 \subfigure[MySQL]{
-\includegraphics[width=0.4\textwidth]{multi-label-results-mysql}
+\includegraphics[width=0.4\textwidth]{figures/multi-label-results-mysql}
 \label{fig:subfig3}
@@
 \subfigure[MaxDB]{
-\includegraphics[width=0.4\textwidth]{multi-label-results-maxdb}
+\includegraphics[width=0.4\textwidth]{figures/multi-label-results-maxdb}
 \label{fig:subfig4}
@@
   %\includegraphics[height=.55\textheight]{maxdb-time-smear-plot-Equal}
-  \includegraphics[height=.55\textheight]{small-maxdb-time-smear-plot-Equal}
+  \includegraphics[height=.55\textheight]{figures/small-maxdb-time-smear-plot-Equal}
   % \includegraphics[height=.57\textheight]{small-maxdb-time-smear-plot-Equal}
@@
   %\includegraphics[width=.97\textwidth]{small-mysql-time-smear-plot-Equal}
-  \includegraphics[width=.97\textwidth]{mysql-timeline}
+  \includegraphics[width=.97\textwidth]{figures/mysql-timeline}
   \caption{MySQL 3.23:NFRs over time. Topic frequency is mapped to alpha (transparency) value.}

### 2010-08-17 src/validate/plot_timeline.py

@@ -57,7 +57,7 @@ def plot_timeline(proj, period_map):
     # ax.xaxis.set_major_locator(years)
     # ax.xaxis.set_major_formatter(yearsFmt)
     # ax.xaxis.set_minor_locator(months)
-    # ax.grid(True)
+    ax.grid(True)
     # ax.annotate('race interrupted', (61, 25),
     #             xytext=(0.8, 0.9), textcoords='axes fraction',
     #             arrowprops=dict(facecolor='black', shrink=0.05),

### 2010-08-17 icse/icse2011.tex
code and figures for properly plotting timelines. Matplotlib is overly configurable.
Signed-off-by: Neil Ernst <neil@neilernst.net>

@@
   \centering
- \includegraphics[width=.75\textwidth]{figures/unsupervised-bar}
+ \includegraphics[width=.85\textwidth]{figures/unsupervised-bar}
  \caption{ROC values for each NFR, per project and wordlist.}
@@
 \label{fig:best-learn-per-tag}
-\caption[]{ROC value (black) for the best learner per label, compared with ZeroR accuracy values (grey).
-\todo[inline]{better discussion and graphical legend would be nice}
+\caption[]{ROC value for the best learner per label, compared with ZeroR accuracy values.  Possible values range from 0--1.
+\todo[inline]{better discussion}
 }
@@
 \label{fig:mulan}
-\caption[]{MySQL and MaxDB macro (grey) and micro (black) ROC results per multi-label learner.
-\todo[inline]{better discussion and graphical legend would be nice}
+\caption[]{MySQL and MaxDB macro (grey) and micro (black) ROC results per multi-label learner. Possible values range from 0--1.
+\todo[inline]{better discussion}% and graphical legend would be nice}
 }
@@
   %\includegraphics[height=.55\textheight]{maxdb-time-smear-plot-Equal}
-  \includegraphics[height=.55\textheight]{figures/small-maxdb-time-smear-plot-Equal}
+  \includegraphics[width=\textwidth]{figures/maxdb-timeline}
   % \includegraphics[height=.57\textheight]{small-maxdb-time-smear-plot-Equal}
-  \caption{MaxDB 7.500: Labelled topics related to non-functional software qualities plotted over time, continuous topics are trends that occur across time}
+ % \caption{MaxDB 7.500: Labelled topics related to non-functional software qualities plotted over time, continuous topics are trends that occur across time}
+  \caption{MaxDB 7.500: NFRs over time. Topic frequency relative to that quality is mapped to alpha (transparency) value.}
  \label{fig:maxdb}
@@
   \centering
-  % hi-res version
  %\includegraphics[width=.97\textwidth]{mysql-time-smear-plot-Equal}
@@
   \includegraphics[width=.97\textwidth]{figures/mysql-timeline}
-  \caption{MySQL 3.23:NFRs over time. Topic frequency is mapped to alpha (transparency) value.}
+  \caption{MySQL 3.23: NFRs over time. Topic frequency relative to that quality is mapped to alpha (transparency) value.}
   \label{fig:mysql}

### 2010-08-17 src/validate/plot_timeline.py

@@ -1,26 +1,60 @@
 """
 Plot timelines with no spaces. Input is a range of values per quality.
 """
-import matplotlib.pyplot as plt
+# import matplotlib.pyplot as plt
+import matplotlib
+matplotlib.use('pdf')
+
 from matplotlib.dates import YearLocator, MonthLocator, DateFormatter
-from matplotlib.patches import FancyBboxPatch, Rectangle
+from matplotlib.patches import Rectangle
+from matplotlib.lines import Line2D     
+import matplotlib.pylab as pylab
+import matplotlib.font_manager as fm
+from pylab import rc                  
 
 import datetime
 
 def plot_timeline(proj, period_map):
     """ Plot a timeline view a la ConcernLines to show trends in topic occurrence """
-    fig = plt.figure()
+    
+    fig_dir = '/Users/nernst/Dropbox/research/abram/papers/naming-paper/icse/figures/'
+    #configure the figure
+    width = 17
+    height = 6
+    #rc("font", family="serif")
+    #rc("font", size=10)
+    # fractions of figure width
+    rc("figure.subplot", left=0.1)
+    rc("figure.subplot", right=0.98)
+    rc("figure.subplot", bottom=0.1) #no idea what the practical impact of this is ..
+    rc("figure.subplot", top=0.99)
+    fig = pylab.figure(figsize=(width, height))
+    #font = fm.FontProperties(fname = '/Library/Fonts/Helvetica.ttc')
     ax = fig.add_subplot(111)
+    years    = YearLocator()   # every year
+    months   = MonthLocator()  # every month
+    yearsFmt = DateFormatter('%Y')
+    months    = MonthLocator(range(1,13), bymonthday=1, interval=4)
+    monthsFmt = DateFormatter("%b %Y")
+    
     # bar size
-    barwidth = 5
+    barwidth = 30#datetime.date()
     barheight = 5
-    ax.set_ylim(0,40)
-    ax.set_xlim(0,5*len(period_map.keys()))
-    ax.set_xlabel('Date')
-    ax.set_yticks([5,10,15,20,25,30,35])
-    ax.set_xticks([x*5 for x in range(len(period_map.keys())+1)])
-    ax.set_yticklabels(['None', 'Portability', 'Efficiency', 'Reliability', 'Functionality', 'Usability', 'Maintainability'])
-    ax.set_xticklabels([period_map.keys()])
+    ax.set_ylim(0,35)
+    if proj == 'mysql':
[…]

### 2010-08-17 src/validate/summary_labels.py

@@ -34,9 +34,10 @@ if __name__ == '__main__':
                         if q in quality_map.keys():
                             quality_map[q] += 1
             periods_map[p_date] = quality_map
-        
+            #periods_map[perio] = quality_map
         pt.plot_timeline(project[0], periods_map)
         
+        #print data to csv
         #for p in periods_map.keys():
             # print  str(p) + ',' + ','.join([str(x) for x in periods_map[p].values()])
             #maxdb.write(project[0] + ',' + str(p) + ',' + ','.join([str(x) for x in periods_map[p].values()]) + '\n')

### 2010-08-17 src/validate/check_validity.py
Updated the check validity script to produce results.. some numbers have changed... not sure why

@@ -134,6 +134,9 @@ if __name__ == '__main__':
             rfile = open( report_file, "w" )
             print report
             rfile.write(report+'\n')
+            yes_no_report_file = wl + "." + report + ".yes_no_report.txt"
+            ynfile = open( yes_no_report_file, "w" )
+            ynfile.write(report+'\n')
             tp = make_count() #true positive  - matches but is
             fp = make_count() #false positive - matches but isn't
             fn = make_count() #false negative - doesn't match but does
@@ -182,7 +185,7 @@ if __name__ == '__main__':
                     if not res:
                         ress = "No"
                     #print res
-                    #rfile.write(ress+'\n')
+                    ynfile.write(ress+'\n')
                  
             tpc = sum(tp.values(),0)        
             fpc = sum(fp.values(),0)        
@@ -198,3 +201,4 @@ if __name__ == '__main__':
                 
             # 
             rfile.close()
+            ynfile.close()

### 2010-08-18 src/validate/check_validity.py
Trying to generate table 2

@@ -19,7 +19,8 @@ def load_period(period, data_dir = "../../data/maxdb-tagged/"):
     topics = doc.findall('Topic')
     #spme test establishing correct number
     return topics
-    
+ 
+
 def load_wordlists(expdir = '../exp3/' ): 
     """ load the datafile with the wordlists we want to use """
     quality_map = {'portability':[], 'efficiency':[], 'reliability':[], 'functionality':[], 'usability':[], 'maintainability':[]}
@@ -121,10 +122,9 @@ if __name__ == '__main__':
     #     topics = load_period(test_period)
     word_exps = get_exps.word_list_exps()
     for wl in word_exps:
-        wld = get_exps.word_list_dir( wl )
+        #wld = get_exps.word_list_dir( wl )
         print wl
-        quality_map = load_wordlists( wld )
-        
+        quality_map = get_exps.load_wordlists( wl )
         for exp in exps:
             #print exp
             ddir = exp[1]
@@ -141,11 +141,14 @@ if __name__ == '__main__':
             fp = make_count() #false positive - matches but isn't
             fn = make_count() #false negative - doesn't match but does
             tn = make_count() #true negative  - doesn't match but isn't
-            
+            yes = 0
+            no = 0
+            yncount = 0
             for period in periods:
                 pstr = "Period #" + period
                 #print pstr
                 #rfile.write(pstr+'\n')
+                ynfile.write(pstr+'\n')
                 topics = load_period(period, data_dir = ddir)
                 for t in topics:
                     annotations, thresh_el = create_element_list(t)
@@ -182,8 +185,13 @@ if __name__ == '__main__':
                             tn[a] += 1
 
                     ress = "Yes"
-                    if not res:
+                    #if not res: # return No if there were no automatches
+                    if len( matches) ==0  :
+                        no += 1
                         ress = "No"
+                    else:
+                        yes += 1
+                    yncount += 1
                     #print res
                     ynfile.write(ress+'\n')
                  
@@ -197,8 +205,15 @@ if __name__ == '__main__':
                 rfile.write(istr)
             istr = info_report_string("\ntotal", tpc, fpc, fnc, tnc)
             print istr
-            rfile.write(istr)
+            rfile.write(istr)        
                 
             # 
             rfile.close()
+            ynfile.write('################################\n');
+            ynfile.write('Named: ' + str(yes) + '\n')
+            ynfile.write('UnNamed: ' + str(no) + '\n')
[…]

### 2010-08-18 src/validate/get_exps.py

@@ -1,6 +1,6 @@
 
 def word_list_exps():
-    return [ 'exp'+str(x) for x in [2,3]]
+    return [ 'exp'+str(x) for x in [1,2,3]]
 
 def word_list_dir( wln ):
     return '../' + str( wln ) + '/'
@@ -26,3 +26,37 @@ def get_exps():
     mysql_dir = "../../data/mysql-tagged/"
     exprs = [ ["maxdb",maxdb_dir, maxdb_periods] , [ "mysql",mysql_dir, mysql_periods ] ]
     return exprs
+
+__exp1_wordlist__ = [
+"accuracy", "availability", "correctness", "efficiency", "flexibility", "integrity", "interoperability", "maintainability", "modifiability", "modularity", "performance", "portability", "reliability", "security", "testability", "traceability", "understandability", "usability", "verifiability"]
+
+"""NFRs for EXP2 and EXP3"""
+__exp2_wordlist__ = ['portability', 'efficiency', 'reliability', 'functionality', 'usability', 'maintainability']
+    
+"""Look up table of wordlists for exp1..exp3"""
+__word_lists__ = {
+    'exp1':__exp1_wordlist__,
+    'exp2':__exp2_wordlist__,
+    'exp3':__exp2_wordlist__,
+}
+
+def gen_exp_wordlist_quality_map( exp ):
+    """ Given an expirement return its wordlist's quality map  """
+    l = __word_lists__[exp]
+    h = {}
+    for x in l:
+        h[x] = []
+    return h
+
+ 
+def load_wordlists(exp = 'exp3' ): 
+    """ load the datafile with the wordlists we want to use """
+    quality_map = gen_exp_wordlist_quality_map( exp )
+    # {'portability':[], 'efficiency':[], 'reliability':[], 'functionality':[], 'usability':[], 'maintainability':[]}
+    expdir = word_list_dir( exp )
+    for q in quality_map.keys():        
+        q_file = open( expdir + 'wordlist.' + q )
+        for line in q_file:
+            quality_map[ q ].append( line.rstrip() )
+    return quality_map
+

### 2010-08-18 src/validate/info_theory.py

@@ -15,6 +15,7 @@ def false_positive_rate(fp,tn):
 def accuracy(tp,tn,fp,fn):
     return (tp + tn)/ (1.0 * tp + tn + fp + fn)
 def true_positive_rate(tp,fn):
+    if ((tp + fn) == 0): return 0
     return tp /(1.0 * ( tp + fn))
 def sensitivity(tp,fn):
     return true_positive_rate(tp,fn)

### 2010-08-18 icse/icse2011.tex
minor change to remove old figure
updated the table.. but unsure

@@
 
-\todo[inline,color=green]{this table should have mysql}
+%\todo[inline,color=green]{this table should have mysql}
 \begin{table}
+% This data came from running  
+%   python check_validity.py
+%  in src/validate
+% and then running
+% tail -n 5 *maxdb*yes_no*
+% tail -n 5 *mysql*yes_no*
 	\centering
@@
 \toprule
-Measure & \textsf{exp1} & \textsf{exp2} & \textsf{exp3} \\
+Project & Measure & \textsf{exp1} & \textsf{exp2} & \textsf{exp3} \\
 \midrule
-Named topics   & 281 & 125 & 328  \\
-Unnamed topics & 139 & 295 & 92   \\
+%Named topics   & 281 & 125 & 328  \\
+%Unnamed topics & 139 & 295 & 92   \\
+MaxDB 7.500 & Named Topics   & 305 & 183 & 330 \\
+MaxDB 7.500 & Unnamed Topics & 84  & 206 & 59   \\
+MaxDB 7.500 & Total  Topics  & 389 & 389 & 389 \\
+MySQL 3.23  & Named Topics   & 341 & 202 & 469 \\
+MySQL 3.23  & Unnamed Topics & 245 & 384 & 117 \\
+MySQL 3.23  & Total  Topics  & 586 & 586 & 586 \\
 \bottomrule
@@
   %\includegraphics[width=.97\textwidth]{small-mysql-time-smear-plot-Equal}
-  \includegraphics[width=.97\textwidth]{figures/mysql-timeline}
+  \includegraphics[width=\textwidth]{figures/mysql-timeline}
   \caption{MySQL 3.23: NFRs over time. Topic frequency relative to that quality is mapped to alpha (transparency) value.}

### 2010-08-18 src/exp2/quickreport.sh
more quick reports

@@ -1,1 +1,6 @@
-for file in wordlist.*; do echo $file; sh quickcount.sh $file; done | egrep -v ' 0$'
+for report in ../counts/*report
+do 
+	echo $report
+	cp $report ./lda.report
+	for file in wordlist.*; do echo $file; sh quickcount.sh $file; done | egrep -v ' 0$'
+done

### 2010-08-18 src/exp1/lda_relator_report.sh
Finally figured out how to produce those reports neil talked about Lda_relator! totally forgot!

@@ -1,0 +1,5 @@
+for proj in mysql maxdb
+do
+	cp ../counts/$proj.lda.report ./lda.report
+	~/projects/hiraldo_grok/hs/Lda_relator  > $proj.lda_relator.report
+done

### 2010-08-18 src/validate/lda_relator_counts.sh
lda relator stuff

@@ -1,0 +1,1 @@
+for exp in exp1 exp2 exp3; do echo $exp; for proj in mysql maxdb; do echo $proj;egrep '^Total'  ../$exp/$proj.lda_relator.report; done; done

### 2010-08-18 icse/icse2011.tex
Made that table finally work. So Sloppy.
ANyways check_validate.py now produces yes_no reports which count things for you
Updated the table again with new data from lda_relator

@@
 \begin{table}
-% This data came from running  
-%   python check_validity.py
-%  in src/validate
-% and then running
-% tail -n 5 *maxdb*yes_no*
-% tail -n 5 *mysql*yes_no*
+
+% % This data came from running  
+% %   python check_validity.py
+% %  in src/validate
+% % and then running
+% % tail -n 5 *maxdb*yes_no*
+% % tail -n 5 *mysql*yes_no*
+% ok the new way is to run lda_relator
+% src/validate/lda_relator_counts.sh
 	\centering
-\begin{tabular}{c|c|c|c}
+\begin{tabular}{l|c|c|c|c}
 \toprule
@@
 \midrule
+% NOTE THESE WERE THE OLD VALUES
+% WHY DID THEY CHANGE???
+% DUNNO
 %Named topics   & 281 & 125 & 328  \\
 %Unnamed topics & 139 & 295 & 92   \\
-MaxDB 7.500 & Named Topics   & 305 & 183 & 330 \\
-MaxDB 7.500 & Unnamed Topics & 84  & 206 & 59   \\
-MaxDB 7.500 & Total  Topics  & 389 & 389 & 389 \\
-MySQL 3.23  & Named Topics   & 341 & 202 & 469 \\
-MySQL 3.23  & Unnamed Topics & 245 & 384 & 117 \\
-MySQL 3.23  & Total  Topics  & 586 & 586 & 586 \\
+
+% these were check_validate
+% MaxDB 7.500 & Named Topics   & 305 & 183 & 330 \\
+% MaxDB 7.500 & Unnamed Topics & 84  & 206 & 59   \\
+% MaxDB 7.500 & Total  Topics  & 389 & 389 & 389 \\
+% MySQL 3.23  & Named Topics   & 341 & 202 & 469 \\
+% MySQL 3.23  & Unnamed Topics & 245 & 384 & 117 \\
+% MySQL 3.23  & Total  Topics  & 586 & 586 & 586 \\
+
+% these are lda relator
+
+MaxDB 7.500 & Named Topics   & 281 & 125 & 328 \\ % these numbers are
+                                % slightly different solely because
+                                % we're counting empty topics
+                                % maxdb has 80 empty topics
+MaxDB 7.500 & Unnamed Topics & 219  & 375 &  172  \\
+MaxDB 7.500 & Total  Topics  & 500 & 500 & 500 \\
+MySQL 3.23  & Named Topics   & 524 & 273 & 773 \\
+MySQL 3.23  & Unnamed Topics & 476 & 727 & 227 \\
+MySQL 3.23  & Total  Topics  & 1000 & 1000 & 1000 \\
+
 \bottomrule

### 2010-08-18 icse/icse2011.tex
CLR and Mulan segway

@@
 
-The less-frequently occurring a label the harder it is to get accurate results, due to the high noise level.
-Nevertheless, these results are better than our previous word bag results of \textsf{exp2} and \textsf{exp3}, because the ROC values are sufficiently higher in most cases (other than MaxDB reliability and MySQL efficiency).
-The limitation of the approach we took here is that we assume labels are independent; however, labels could be correlated with each other.
-We also did not evaluate how well the learners performed together.
-\todo[inline]{Comment about correlation, we should link to the Mulan section, perfect segway since Mulan relies on correlation.}
+The less-frequently occurring a label the harder it is to get accurate
+results, due to the high noise level. Nevertheless, these results are
+better than our previous word bag results of \textsf{exp2} and
+\textsf{exp3}, because the ROC values are sufficiently higher in most
+cases (other than MaxDB reliability and MySQL efficiency). The
+limitation of the approach we took here is that we assume labels are
+independent; however, labels could be correlated with each other. 
+The next section, Section \ref{sec:multilabel} on multi-labels
+addresses the issue of a lack of independence and correlation between labels.
+In the next section we will evaluate how well these learners perform
+together.
+%\todo[inline]{Comment about correlation, we should link to the Mulan section, perfect segway since Mulan relies on correlation.}
 
@@
 \subsection{Applying multiple labels to topics}
+\label{sec:multilabel}
+
 As noted in Section \ref{sec:wordlist}, each topic in our data-set can be composed of zero to all NFRs.
@@
 
-Fig. \ref{fig:mulan} presents the results of Mulan's best multi-label learners for our data. %Calibrated Label Ranking (CLR)
+Fig. \ref{fig:mulan} presents the results of Mulan's best multi-label
+learners for our data. Calibrated Label Ranking (CLR) is a learner
+that builds two layers, the first layer determines if an entity should
+be labelled, while the second layer determines what labels should be assigned.
 The Hierarchy Of Multi-label classifiERs (HOMER) and Binary Relevance (BR) act as a hierarchy of learners: BR is flat, while HOMER tries to build a deeper hierarchy to build a more accurate learner~\cite{mulan}.
@@
 \todo[inline]{discuss this similarity in more detail}
-\todo[inline]{what of CLR? Is it hierarchical?}
+%\todo[inline]{what of CLR? Is it hierarchical?}
 

### 2010-08-18 icse/icse2011.tex
minor intro edits

@@
   research area and we're saying they don't do enough}
+
   Within the field of mining software repositories, many approaches,   such as topic modeling and concept location, rely on the automated   application of machine learning algorithms to corpora.  
@@
 
-When a developer addresses a topic of development, such as a requirement, maintenance, bug fixes or other software engineering tasks, they often produce or modify multiple development artifacts.
+Topics in software engineering practice refer to specific `threads' between project participants.
+These topics can include particular requirements, maintenance requests, bug fixes or other software engineering tasks.
+%NEIL: define topic specifically as akin to email 'thread'.
+It sounded too vague.
+Also changed from 'developers' to 'participants' because developers sounds too coding oriented.
+Our work could apply to testers, users, etc.
+In addressing a topic of development participants often produce or modify multiple development artifacts.
 %
-These development artifacts include: source code, bug reports, revisions to source code, and even test cases. 
+These artifacts can include source code, bug reports, revisions to source code, and test cases. 
 %
-Topics that are addressed are sometimes not immediately resolved, some topics recur across time, yet others are only dealt with briefly.
+Topics can be in several potential states: they may not be immediately resolved, they may re-occur, yet others are only dealt with briefly  (e.g., a bug report that is closed).
+% semi-automatically sounds too uncertain. I think the introduction can be more muscular and bold. Don't give the reviewer any room to start doubting.
+The goal of concept location and topic analysis research is to recover these topics and the related software artifacts.
+% get neil's data mining point in there
+To date, the results of topic analysis and concept location are project specific, and not globally generalizable.
+% get the interpretation aspect
+Much of the current research stops short of interpreting the purpose of the topic. That is, extracted topics are usually manually named and labelled by those who read the topic words.
 % 
-The goal of concept location and topic analysis research is to semi-automatically recover some of these topics, and the related software artifacts.
-% get neil's data mining point in there
-The results of topic analysis and concept location are often very project specific, and not globally generalizable.
-% get the interpretation aspect
-As well much of the current research stops short of interpreting the purpose of the topic. That is, extracted topics are usually manually named and labelled by those who read the topic words.
-% 
-Thus in this paper, we propose a method of interpreting and labelling some topics, and their related artifacts, that is generalizable and relevant to multiple projects.
+In this paper we propose a method of interpreting and labelling topics, and their related artifacts, that is generalizable and relevant to multiple projects.
 
@@
 Deriving such topics is difficult. 
-Raw topics from topic analysis are represented as unlabelled context-less word-lists or word distributions.
+Raw topics from topic analysis are represented as unlabelled context-free word-lists or word distributions.
 Topics have to be identified by interpreting the prevalent words in a word distribution and by inspecting related documents. 
-This manual approach is impractical when one has to handle more than one hundred different topics. 
-To scale to meaningfully sized projects, we require automatic assistance to determine the topic label.
+This manual approach is impractical when one has to handle hundreds of different topics; to scale to meaningfully sized projects, we require automatic assistance to determine the topic label.
 
@@
 
-There are many avenues of further investigation.
+There are several avenues of further investigation.
 We want to investigate developer attitudes related to these labels: i.e., when we label a topic, was the developer expressing positive or negative qualities about that label?

### 2010-08-18 src/validate/plot_timeline.py
Figures with max(all) as denominator

@@ -71,13 +71,31 @@ def plot_timeline(proj, period_map):
         qm = period_map[period] # the quality map w/value
         y, m, d = period.split('-') # why I have to re create the datetime instance ... 
         date = datetime.date(int(y), int(m), int(d))
-        none_rect = Rectangle((date,0), barwidth,barheight,fill=True,lw=0, fc='red',alpha=(float(qm['none'])/float(max_value_map['none'])))
-        port_rect = Rectangle((date,5), barwidth,barheight,fill=True,lw=0, fc='red',alpha=(float(qm['portability'])/float(max_value_map['portability'])))
-        effc_rect = Rectangle((date,10),barwidth,barheight,fill=True,lw=0, fc='red',alpha=(float(qm['efficiency'])/float(max_value_map['efficiency'])))
-        reli_rect = Rectangle((date,15),barwidth,barheight,fill=True,lw=0, fc='red',alpha=(float(qm['reliability'])/float(max_value_map['reliability'])))
-        func_rect = Rectangle((date,20),barwidth,barheight,fill=True,lw=0, fc='red',alpha=(float(qm['functionality'])/float(max_value_map['functionality'])))
-        usab_rect = Rectangle((date,25),barwidth,barheight,fill=True,lw=0, fc='red',alpha=(float(qm['usability'])/float(max_value_map['usability'])))
-        main_rect = Rectangle((date,30),barwidth,barheight,fill=True,lw=0, fc='red',alpha=(float(qm['maintainability'])/float(max_value_map['maintainability'])))
+        
+        #now we draw a box with a transparent red fill. The transparency can be set based on several factors:
+        # - the ratio between that NFR's highest value and the current value (relative to self)
+        # - the ratio between the overall highest value and the current NFRs value (relative to all)
+        # - set the average topic number as a midpoint, and color based on distance from this average.
+        denominator = max(max_value_map.values())
+        
+        # relative to all
+        none_rect = Rectangle((date,0), barwidth,barheight,fill=True,lw=0, fc='red',alpha=(float(qm['none'])/float(denominator)))
+        port_rect = Rectangle((date,5), barwidth,barheight,fill=True,lw=0, fc='red',alpha=(float(qm['portability'])/float(denominator)))
[…]

### 2010-08-18 src/validate/plot_timeline.py
- Figures updated with black numbers of topic labels. Still using 'all' as denom. - Updated paper to show these timelines in a single figure with subfigures, more space that way. - Fixed reference titles to upper-case nouns.

@@ -14,6 +14,7 @@ from pylab import rc
 
 import datetime
 
+
 def plot_timeline(proj, period_map):
     """ Plot a timeline view a la ConcernLines to show trends in topic occurrence """
     
@@ -77,7 +78,6 @@ def plot_timeline(proj, period_map):
         # - the ratio between the overall highest value and the current NFRs value (relative to all)
         # - set the average topic number as a midpoint, and color based on distance from this average.
         denominator = max(max_value_map.values())
-        
         # relative to all
         none_rect = Rectangle((date,0), barwidth,barheight,fill=True,lw=0, fc='red',alpha=(float(qm['none'])/float(denominator)))
         port_rect = Rectangle((date,5), barwidth,barheight,fill=True,lw=0, fc='red',alpha=(float(qm['portability'])/float(denominator)))
@@ -106,13 +106,14 @@ def plot_timeline(proj, period_map):
         ax.add_patch(main_rect)
         
         #add text numbers to debug
-        ax.annotate(str(qm['none']), (date+num2date(5), 2.5) )
-        ax.annotate(str(qm['portability']), (date, 7.5) )
-        ax.annotate(str(qm['efficiency']), (date, 12.5) )
-        ax.annotate(str(qm['reliability']), (date, 17.5) )
-        ax.annotate(str(qm['functionality']), (date, 22.5) )
-        ax.annotate(str(qm['usability']), (date, 27.5) )
-        ax.annotate(str(qm['maintainability']), (date, 32.5) )
+        delta = datetime.timedelta(days=5)
+        ax.annotate(str(qm['none']),           (date+delta, 2.5) , color='black', size='small')
+        ax.annotate(str(qm['portability']),    (date+delta, 7.5),  color='black', size='small' )
+        ax.annotate(str(qm['efficiency']),     (date+delta, 12.5), color='black', size='small'  )
+        ax.annotate(str(qm['reliability']),    (date+delta, 17.5), color='black', size='small'  )
+        ax.annotate(str(qm['functionality']),  (date+delta, 22.5), color='black', size='small'  )
+        ax.annotate(str(qm['usability']),      (date+delta, 27.5), color='black', size='small'  )
+        ax.annotate(str(qm['maintainability']),(date+delta, 32.5), color='black', size='small'  )
         
         i += barwidth
         

### 2010-08-18 src/validate/plot_timeline.py
Changed the plots. Modified the absolute path (local relative output path now). Added a mini bar chart relative to the stream.

@@ -5,7 +5,7 @@ Plot timelines with no spaces. Input is a range of values per quality.
 import matplotlib
 matplotlib.use('pdf')
 
-from matplotlib.dates import YearLocator, MonthLocator, DateFormatter
+from matplotlib.dates import YearLocator, MonthLocator, DateFormatter, num2date
 from matplotlib.patches import Rectangle
 from matplotlib.lines import Line2D     
 import matplotlib.pylab as pylab
@@ -14,10 +14,13 @@ from pylab import rc
 
 import datetime
 
+
 def plot_timeline(proj, period_map):
     """ Plot a timeline view a la ConcernLines to show trends in topic occurrence """
     
-    fig_dir = '/Users/nernst/Dropbox/research/abram/papers/naming-paper/icse/figures/'
+    #fig_dir = '/Users/nernst/Dropbox/research/abram/papers/naming-paper/icse/figures/'
+    # sorry bro
+    fig_dir = './output/' # /Users/nernst/Dropbox/research/abram/papers/naming-paper/icse/figures/'
     #configure the figure
     width = 17
     height = 6
@@ -62,6 +65,16 @@ def plot_timeline(proj, period_map):
     #max value is across all periods in that quality
     #i am personally offended by this code
     max_value_map = {'none':0,'portability':0, 'efficiency':0, 'reliability':0, 'functionality':0, 'usability':0, 'maintainability':0}
+
+    # the order of display
+    value_order = [ 'none' , 'portability' , 'efficiency' , 'reliability' , 'functionality' , 'usability' , 'maintainability' ]
+    value_start = {}
+    # they y values of value
+    vo = 0
+    for value_name in value_order:
+        value_start[ value_name ] = vo
+        vo += 5
+
     for qm in period_map.values():
         for q in qm.keys():
             if qm[q] > max_value_map[q]: max_value_map[q] = qm[q]
@@ -79,34 +92,49 @@ def plot_timeline(proj, period_map):
         denominator = max(max_value_map.values())
         
         # relative to all
-        none_rect = Rectangle((date,0), barwidth,barheight,fill=True,lw=0, fc='red',alpha=(float(qm['none'])/float(denominator)))
-        port_rect = Rectangle((date,5), barwidth,barheight,fill=True,lw=0, fc='red',alpha=(float(qm['portability'])/float(denominator)))
[…]

### 2010-08-18 src/validate/test_timeline_plot.py

@@ -6,3 +6,4 @@ mysql_data = {'2004-04-11': {'none': 0, 'portability': 0, 'efficiency': 0, 'reli
 max_db_data = {'2004-08-28': {'none': 3, 'portability': 8, 'efficiency': 3, 'reliability': 1, 'functionality': 3, 'maintainability': 5, 'usability': 7}, '2004-11-26': {'none': 7, 'portability': 9, 'efficiency': 2, 'reliability': 0, 'functionality': 1, 'maintainability': 7, 'usability': 2}, '2006-05-20': {'none': 13, 'portability': 2, 'efficiency': 4, 'reliability': 1, 'functionality': 0, 'maintainability': 2, 'usability': 2}, '2005-11-21': {'none': 15, 'portability': 2, 'efficiency': 0, 'reliability': 1, 'functionality': 1, 'maintainability': 1, 'usability': 0}, '2004-09-27': {'none': 8, 'portability': 10, 'efficiency': 0, 'reliability': 2, 'functionality': 2, 'maintainability': 6, 'usability': 0}, '2006-01-20': {'none': 13, 'portability': 3, 'efficiency': 1, 'reliability': 1, 'functionality': 3, 'maintainability': 2, 'usability': 0}, '2005-12-21': {'none': 11, 'portability': 2, 'efficiency': 2, 'reliability': 1, 'functionality': 2, 'maintainability': 6, 'usability': 0}, '2005-08-23': {'none': 10, 'portability': 3, 'efficiency': 0, 'reliability': 5, 'functionality': 0, 'maintainability': 4, 'usability': 4}, '2006-04-20': {'none': 7, 'portability': 1, 'efficiency': 7, 'reliability': 1, 'functionality': 1, 'maintainability': 6, 'usability': 0}, '2004-07-29': {'none': 8, 'portability': 3, 'efficiency': 1, 'reliability': 2, 'functionality': 2, 'maintainability': 3, 'usability': 0}, '2004-10-27': {'none': 7, 'portability': 7, 'efficiency': 3, 'reliability': 2, 'functionality': 3, 'maintainability': 6, 'usability': 2}, '2005-01-25': {'none': 10, 'portability': 4, 'efficiency': 3, 'reliability': 0, 'functionality': 0, 'maintainability': 4, 'usability': 1}, '2005-07-24': {'none': 8, 'portability': 4, 'efficiency': 3, 'reliability': 1, 'functionality': 0, 'maintainability': 3, 'usa
[…]

### 2010-08-18 icse/icse2011.tex
participants -> stakeholders

@@
 
-Topics in software engineering practice refer to specific `threads' between project participants.
+Topics in software engineering practice refer to specific `threads' between project stakeholders.
 These topics can include particular requirements, maintenance requests, bug fixes or other software engineering tasks.
@@
 It sounded too vague.
-Also changed from 'developers' to 'participants' because developers sounds too coding oriented.
+Also changed from 'developers' to 'stakeholders' because developers sounds too coding oriented.
 Our work could apply to testers, users, etc.
-In addressing a topic of development participants often produce or modify multiple development artifacts.
+In addressing a topic of development stakeholders often produce or modify multiple development artifacts.
 %
@@
 
-\begin{figure*}
+\begin{figure*}[ht]
 \centering
-  % hi-res version
-  %\includegraphics[height=.55\textheight]{maxdb-time-smear-plot-Equal}
-  \includegraphics[width=\textwidth]{figures/maxdb-timeline}
-  % \includegraphics[height=.57\textheight]{small-maxdb-time-smear-plot-Equal}
- % \caption{MaxDB 7.500: Labelled topics related to non-functional software qualities plotted over time, continuous topics are trends that occur across time}
-  \caption{MaxDB 7.500: NFRs over time. Topic frequency relative to that quality is mapped to alpha (transparency) value.}
- \label{fig:maxdb}
+\subfigure[MaxDB 7.500]{
+\includegraphics[width=\textwidth]{figures/maxdb-timeline-all}
+\label{fig:mysql-timeline}
+}
+\subfigure[MySQL 3.23]{
+\includegraphics[width=\textwidth]{figures/mysql-timeline-all}
+\label{fig:maxdb-timeline}
+}
+\label{fig:timelines}
+\caption[]{NFRs over time. Numbers represent topics labelled with that quality. Box intensity (saturation) is mapped to label frequency relative to the largest label value.
+\todo[inline]{better discussion}
+}
 \end{figure*}
 
-\begin{figure*}%[h]
-  \centering
- %\includegraphics[width=.97\textwidth]{mysql-time-smear-plot-Equal}
-  %\includegraphics[width=.97\textwidth]{small-mysql-time-smear-plot-Equal}
-  \includegraphics[width=\textwidth]{figures/mysql-timeline}
-  \caption{MySQL 3.23: NFRs over time. Topic frequency relative to that quality is mapped to alpha (transparency) value.}
-  \label{fig:mysql}
-\end{figure*}
+% 
+% \begin{figure*}
+%   \centering
+%   % hi-res version
+%   %\includegraphics[height=.55\textheight]{maxdb-time-smear-plot-Equal}
+%   \includegraphics[width=\textwidth]{figures/maxdb-timeline-all}
+%   % \includegraphics[height=.57\textheight]{small-maxdb-time-smear-plot-Equal}
+%  % \caption{MaxDB 7.500: Labelled topics related to non-functional software qualities plotted over time, continuous topics are trends that occur across time}
+%   \caption{MaxDB 7.500: NFRs over time. Numbers represent topics labelled with that quality. Box intensity (saturation) is mapped to label frequency relative to the largest label value.}
+%  \label{fig:maxdb}
+% \end{figure*}
+% 
+% \begin{figure*}%[h]
+%   \centering
+%  %\includegraphics[width=.97\textwidth]{mysql-time-smear-plot-Equal}
+%   %\includegraphics[width=.97\textwidth]{small-mysql-time-smear-plot-Equal}
+%   \includegraphics[width=\textwidth]{figures/mysql-timeline-all}
+%   \caption{MySQL 3.23: NFRs over time. Topic frequency relative to that quality is mapped to alpha (transparency) value.}
+%   \label{fig:mysql}
+% \end{figure*}
 
@@
 
+% What did we do
+% What did we show
+% What did we do that was new?
+
+% use domain knowledge
 We demonstrated that static but domain-specific knowledge can improve unsupervised labelling of extracted topics. 

### 2010-08-18 icse/icse2011.tex
conclusion notes
reworked conclusions

@@
 
+\todo[inline]{More needs to be done here! More content!}
+
 \begin{comment}
@@
 
+%\todo[inline]{We need to summarize what we did}
+%\todo[inline]{We need to show our contribution}
+\todo[inline]{Need to juxtapose our work against previous work -- contribution}
+
 % What did we do
@@
 
-% use domain knowledge
-We demonstrated that static but domain-specific knowledge can improve unsupervised labelling of extracted topics. 
-Our \textsf{exp2} experiment used small accurate word bags to label topics but performed just as well as \textsf{exp3}, which used many more general terms from WordNet.
-We then showed that with some supervision, and by using efficient machine learners based on Naive Bayesian classifiers, we could improve the accuracy of automatic labelling topics even further.
+% para 1 we did this, we were successful
+In this paper we presented a cross-project data mining technique that was appropriate for mining and labelling non-functional requirements topics and artifacts from project histories.
+Our contributions include:
+\begin{itemize}
+\item We provide an unsupervised method of topic labelling (word-bags);
+%\item Our unsupervised method can be applied to projects without training;
+\item We demonstrate a supervised method of labelling topics by a single NFRs (machine learning);
+\item We demonstrate a supervised method of labelling topics by multiple NFRs (multi-label machine learning);
+\item We provide a method of cross-project analysis via topic labelling;
+\item We demonstrate a method of visualizing these topics across time.
+\end{itemize}
+We leveraged word-bags, machine learning classifiers and multi-label machine learning classifiers.  
+We evaluated these techniques using multiple experiments against two open source database systems and visualized the results.
+We found that our topic labelling techniques were effective.
 
+% para 2 we did this to validate and show we did it
+% Experiment description -- origin of wordbags
+We validated our topic labelling techniques using multiple experiments.
+Using unsupervised word-bag analysis we set up two NFR topic labelling experiments: 
+our \textsf{exp2} experiment used small but accurate word-bags, while \textsf{exp3} used large general word-bags derived from WordNet.
+The performance of these techniques differed per NFR but both \textsf{exp2} and \textsf{exp3} had similar performance.
+%Our \textsf{exp2} experiment used small accurate word bags to label topics but performed just as well as \textsf{exp3}, which used many more general terms from WordNet.
+% we could can improve auto labelling with ML
+Our third experiment showed that with some supervision, and by using efficient machine learners based on Naive Bayesian classifiers, we could improve the accuracy of automatic topic labelling even further.
+Our fourth experiment evaluated the effectiveness of multi-label learners, such as HOMER. 
+Both kinds of learners performed well with area under the ROC curve values between $.6$ and $.8$.
+
+% para 3 look while we did it we found this
+% total topics - maybe move above
+We showed that non-functional requirements are often trending topics, that non-functional requirements are quite common in developer topics, and that there are efficient methods of semi-automating and automating topic labelling.
+% We found NFRs
 Our manual inspection and annotation of the topics extracted from MySQL and MaxDB revealed that many of the extracted topics dealt with non-functional requirements, and these topics were spread across the entire history of a project.
+% Try to provide an example of NFRs but little context
 In the cases of MaxDB and MySQL, portability was a constant maintenance concern and was prevalent throughout the entire lifetime of the projects.
 
-We showed that non-functional requirements are often trending topics, that non-functional requirements are quite common in developer topics, and that there are efficient methods of semi-automating and automating topic labelling.
+% para 4 summarize and hammer contribution
+In conclusion,
+topic analysis and concept analysis research has produced topics that need to be manually labelled. 
+We found that topics without labels are difficult to interpret.
+Thus we leveraged software engineering knowledge to produce a method of domain specific topic labelling.
+The word-bag technique is not project specific, thus it has intrinsic value when analyzing multiple projects.
+We demonstrated multiple supervised and unsupervised methods of topic labelling, including multi-label learners.
+We provided a case study and visualization of NFR related topics in MaxDB 7.500 and MySQL 3.23. 
+During the case study we manually labelled 100s of topics by hand then used these annotations to validate the positive performance of topic labelling techniques.
+We showed that topics with labels are more useful than unlabelled topics, thus to furnish stakeholders with labelled topics we provided many methods of automatic topic labelling.
 
+\subsection{Future Work}
+
+% para 5 future work
+% Future work
 There are several avenues of further investigation.  
+% Talk to people
 We want to investigate developer attitudes related to these labels: i.e., when we label a topic, was the developer expressing positive or negative qualities about that label?
 It is difficult to map abstract qualities to particular messages.
-Is a ``quality'' discussion about more than just corrective maintenance?
-%In this study we take the position that any type of maintenance has to do with software quality.
+% This is disjointed and needs more context
+Is a ``quality'' discussion about more than just corrective maintenance? %In this study we take the position that any type of maintenance has to do with software quality.
+
+\todo[inline]{Future work is disjointed the question should be more clear}
 

### 2010-08-18 icse/icse2011.tex
some todo notes and future work

@@
 
-\todo[inline]{Update this}
+\todo[inline]{Update this, rewrite it to handle the new situation but be sure to keep the maintainability and portability}
 
@@
 \todo[inline]{More needs to be done here! More content!}
+\todo[inline]{Use the old diagrams to come up with interesting comparisons}
 
@@
 % This is disjointed and needs more context
-Is a ``quality'' discussion about more than just corrective maintenance? %In this study we take the position that any type of maintenance has to do with software quality.
+On question we seek to answer is if a ``quality'' discussion about more than just corrective maintenance?
+%In this study we take the position that any type of maintenance has to do with software quality.
 
-\todo[inline]{Future work is disjointed the question should be more clear}
+%\todo[inline]{Future work is disjointed the question should be more clear}
 

### 2010-08-19 src/validate/plot_timeline.py
minor change to figure code - cleanup

@@ -5,7 +5,7 @@ Plot timelines with no spaces. Input is a range of values per quality.
 import matplotlib
 matplotlib.use('pdf')
 
-from matplotlib.dates import YearLocator, MonthLocator, DateFormatter, num2date
+from matplotlib.dates import YearLocator, MonthLocator, DateFormatter
 from matplotlib.patches import Rectangle
 from matplotlib.lines import Line2D     
 import matplotlib.pylab as pylab
@@ -18,8 +18,6 @@ import datetime
 def plot_timeline(proj, period_map):
     """ Plot a timeline view a la ConcernLines to show trends in topic occurrence """
     
-    #fig_dir = '/Users/nernst/Dropbox/research/abram/papers/naming-paper/icse/figures/'
-    # sorry bro
     fig_dir = './output/' # /Users/nernst/Dropbox/research/abram/papers/naming-paper/icse/figures/'
     #configure the figure
     width = 17
@@ -69,7 +67,7 @@ def plot_timeline(proj, period_map):
     # the order of display
     value_order = [ 'none' , 'portability' , 'efficiency' , 'reliability' , 'functionality' , 'usability' , 'maintainability' ]
     value_start = {}
-    # they y values of value
+    # the y values of value
     vo = 0
     for value_name in value_order:
         value_start[ value_name ] = vo
@@ -79,7 +77,6 @@ def plot_timeline(proj, period_map):
         for q in qm.keys():
             if qm[q] > max_value_map[q]: max_value_map[q] = qm[q]
             
-    i = 0
     for period in period_map.keys():
         qm = period_map[period] # the quality map w/value
         y, m, d = period.split('-') # why I have to re create the datetime instance ... 
@@ -91,7 +88,8 @@ def plot_timeline(proj, period_map):
         # - set the average topic number as a midpoint, and color based on distance from this average.
         denominator = max(max_value_map.values())
         # relative to all
-        # loop over it instead of copy and paste
+        # loop over it instead of copy and paste 
+        #HAHAHA! Copy and paste rules. Code clones rule. Code smells rule.
         for value in value_order:
             #rect = Rectangle((date,value_start[value]), barwidth,barheight,fill=True,lw=0, fc='red',alpha=(float(qm[ value ])/float(max_value_map[ value ])))
[…]

### 2010-08-19 icse/icse2011.tex
icse crud
more icse'in it up
fixing viz section

--- 2010-08-18 icse/icse2011.tex
@@
 \usepackage{amsmath}
+\usepackage{url}
 \usepackage[pdfstartview=FitH,bookmarksopenlevel=3,bookmarks=true]{hyperref} %pdfpagescrop={92 112 523 778},
 
-\usepackage{url}
 \begin{document}
@@
 \maketitle
-\thispagestyle{empty}
+%\thispagestyle{empty}
 
@@
 
-Topics in software engineering practice refer to specific `threads' between project stakeholders.
+Topics in software engineering practice refer to specific `threads' between project participants.
 These topics can include particular requirements, maintenance requests, bug fixes or other software engineering tasks.
@@
 It sounded too vague.
-Also changed from 'developers' to 'stakeholders' because developers sounds too coding oriented.
+Also changed from 'developers' to 'participants' because developers sounds too coding oriented.
 Our work could apply to testers, users, etc.
-In addressing a topic of development stakeholders often produce or modify multiple development artifacts.
+In addressing a topic of development participants often produce or modify multiple development artifacts.
 %
@@
 Micro measurements are aggregated at a decision level.
+\todo[inline]{this is unclear, could do with a few more sentence explaining why we should care.}
  So a macro-ROC measurement is the average ROC over the ROC values for all labels, where a micro-ROC is the average ROC over all examples that were classified.
@@
 
-\subsection{Visualization}
-We created two visualizations of the manually labelled data.
-A problem we faced while visualizing was how to display tag overlaps.
-Our solution, while not optimal, was to assign a separate colour to each distinct set of annotations.
-Figures \ref{fig:maxdb} and \ref{fig:mysql} show extracted topics grouped by annotations.
-Annotations or labels that occur in adjacent windows are joined together as trends.
-For MaxDB (Figure \ref{fig:maxdb}) the largest trends were related to maintainability and portability.
+\subsection{Visualizations}
+Figures \ref{fig:maxdb-timeline} and \ref{fig:mysql-timeline} show the temporal patterns of label frequencies.
+There are two measures represented.
+One, the relative frequency, shown in the grey boxes, represents the number of labels with that NFR in that period, relative to the maximum number of topics assigned to that NFR.
+The second, absolute frequency, compares the number of topics labelled with that NFR per period relative to the maximum number of labelled topics overall.
+
+For example, in Fig. \ref{fig:mysql-timeline}, we see that the NFRs functionality, portability and maintainability contain more labeled topics, since these NFRs have more intense shaded boxes.
+
+% For MaxDB (Figure \ref{fig:maxdb}) the largest trends were related to maintainability and portability.
 MaxDB supports numerous platforms and portability was a constant issue facing the project's development.
@@
 \centering
+\subfigure[MySQL 3.23]{
+\includegraphics[width=\textwidth]{figures/mysql-timeline}
+\label{fig:maxdb-timeline}                               
+}                                                        
 \subfigure[MaxDB 7.500]{                                 
-\includegraphics[width=\textwidth]{figures/maxdb-timeline-all}
+\includegraphics[width=\textwidth]{figures/maxdb-timeline}
 \label{fig:mysql-timeline}
 }
-\subfigure[MySQL 3.23]{
-\includegraphics[width=\textwidth]{figures/mysql-timeline-all}
-\label{fig:maxdb-timeline}
-}
 \label{fig:timelines}
-\caption[]{NFRs over time. Numbers represent topics labelled with that quality. Box intensity (saturation) is mapped to label frequency relative to the largest label value.
+\caption[]{NFR label per period.
+Each box represents a 30-day period.
+Numbers represent number of topics in that period labelled with that NFR.
+Box intensity (saturation) is mapped to label frequency relative to the largest label value of \emph{all} NFRs.
+Grey bars indicate label frequency relative to \emph{that} NFR.
 \todo[inline]{better discussion}
@@
 
-% What did we do
-% What did we show
-% What did we do that was new?
-
-% use domain knowledge
 We demonstrated that static but domain-specific knowledge can improve unsupervised labelling of extracted topics.

### 2010-08-19 icse/icse2011.tex
abstract hacking

--- 2010-08-18 icse/icse2011.tex
@@
 \begin{document}
+
+\conferenceinfo{ICSE}{'11 Waikiki, Hawaii USA}
 \numberofauthors{5}
@@
 \maketitle
-\thispagestyle{empty}
+%\thispagestyle{empty}
 
 \begin{abstract}
-\todo[inline]{Topics and LDA sooner, Topics/concept location is a known
-  research area and we're saying they don't do enough}
+%\todo[inline]{Topics and LDA sooner, Topics/concept location is a known
+%  research area and we're saying they don't do enough}
 
+% our field
   Within the field of mining software repositories, many approaches,   such as topic modeling and concept location, rely on the automated   application of machine learning algorithms to corpora.  
- Unfortunately the output of these tools is often difficult to distinguish and interpret as they are often so abstract.
- Thus to have a meaningful discussion about the topics of software development, we must be able to devise appropriate labels for extracted topics. 
- However, these approaches neither use domain-specific knowledge to improve results, nor contextualize those results for developers.
+% the problem
+ Unfortunately the output of these tools is often difficult to distinguish and interpret as they are often so abstract, in particular topics that are automatically extracted are word-lists lacking semantic interpretation.
+% the solution 
+ Thus to have a meaningful discussion about the topics of software development, we must be able to interpret these topics and devise appropriate labels for extracted topics, yet current software engineering topic analysis and concept analysis research relies on manual naming and labelling of topics
+% the labelling research
+Current topic labelling approaches neither use domain-specific knowledge to improve results, nor contextualize those results for developers.   
+% a sub problem
 While too much specificity can produce non-generalizable results, too little produces broad learners that do not provide much immediately useful detail. 
+% what we do
 This paper implements \emph{labelled topic extraction}, in which topics are extracted from commit comments and given labels relating to a cross-project taxonomy. 
-We focus on non-functional requirements related to software quality as a potential generalization, since there is some shared belief that these qualities apply broadly across many software systems and their development artifacts.
+% data mining generalizable
+To ensure we have a generalizable project agnostic approach we label topics by non-functional requirements related to software quality, since there is some shared belief that these qualities apply broadly across many software systems and their development artifacts.
+% We used real world data
 We evaluated our approach with an experimental study on two large-scale database projects, MySQL and MaxDB. 
+% We used LDA, CVS, Bitkeeper
 We extracted topics using Latent Dirichlet Allocation (LDA) from the commit log comments of their version control systems (CVS and BitKeeper). 
+% We show NFRs exist
 Our results were generalizable across the two projects, showing that non-functional requirements were commonly discussed, and we identified topic trends over time. 	
- Our labelled topic extraction technique allowed us to devise appropriate, context-sensitive labels across these two projects, providing insight into software development activities.
+% we can label stuff
+Our labelled topic extraction technique allowed us to devise appropriate, context-sensitive labels across these two projects, providing insight into their software development activities.
 \end{abstract}
+
+\category{D.2.9}{Software Engineering}{Managment}[Lifecycle]
+\category{D.2.1}{Software Engineering}{Requirements/Specifications}[Tools]
+
+%\terms{Theory}
+
+\keywords{Topics, LDA, NFR}
 
@@
 NFRs are prevalent in almost all software systems.
+In this paper we will show that NFR related topics occur quite frequently.
 Thus our method of generalizing the data mining of these software artifacts is to leverage  the domain knowledge of software engineering, specifically NFRs. 
@@
 %XXX can save space here!
-\section{Appendix}
+%\section{Appendix}
+\appendix
+
 Our data and scripts are available at \url{http://softwareprocess.es/nomen/}

### 2010-08-19 icse/icse2011.tex

--- 2010-08-19 icse/icse2011.tex
@@
 \begin{abstract}
-\todo[inline]{Topics and LDA sooner, Topics/concept location is a known
-  research area and we're saying they don't do enough}
+%\todo[inline]{Topics and LDA sooner, Topics/concept location is a known
+%  research area and we're saying they don't do enough}
 
+% our field
   Within the field of mining software repositories, many approaches,   such as topic modeling and concept location, rely on the automated   application of machine learning algorithms to corpora.  
- Unfortunately the output of these tools is often difficult to distinguish and interpret as they are often so abstract.
- Thus to have a meaningful discussion about the topics of software development, we must be able to devise appropriate labels for extracted topics. 
- However, these approaches neither use domain-specific knowledge to improve results, nor contextualize those results for developers.
+% the problem
+ Unfortunately the output of these tools is often difficult to distinguish and interpret as they are often so abstract, in particular topics that are automatically extracted are word-lists lacking semantic interpretation.
+% the solution 
+ Thus to have a meaningful discussion about the topics of software development, we must be able to interpret these topics and devise appropriate labels for extracted topics, yet current software engineering topic analysis and concept analysis research relies on manual naming and labelling of topics
+% the labelling research
+Current topic labelling approaches neither use domain-specific knowledge to improve results, nor contextualize those results for developers.   
+% a sub problem
 While too much specificity can produce non-generalizable results, too little produces broad learners that do not provide much immediately useful detail. 
+% what we do
 This paper implements \emph{labelled topic extraction}, in which topics are extracted from commit comments and given labels relating to a cross-project taxonomy. 
-We focus on non-functional requirements related to software quality as a potential generalization, since there is some shared belief that these qualities apply broadly across many software systems and their development artifacts.
+% data mining generalizable
+To ensure we have a generalizable project agnostic approach we label topics by non-functional requirements related to software quality, since there is some shared belief that these qualities apply broadly across many software systems and their development artifacts.
+% We used real world data
 We evaluated our approach with an experimental study on two large-scale database projects, MySQL and MaxDB. 
+% We used LDA, CVS, Bitkeeper
 We extracted topics using Latent Dirichlet Allocation (LDA) from the commit log comments of their version control systems (CVS and BitKeeper). 
+% We show NFRs exist
 Our results were generalizable across the two projects, showing that non-functional requirements were commonly discussed, and we identified topic trends over time. 	
- Our labelled topic extraction technique allowed us to devise appropriate, context-sensitive labels across these two projects, providing insight into software development activities.
+% we can label stuff
+Our labelled topic extraction technique allowed us to devise appropriate, context-sensitive labels across these two projects, providing insight into their software development activities.
 \end{abstract}
+
+\category{D.2.9}{Software Engineering}{Managment}[Lifecycle]
+\category{D.2.1}{Software Engineering}{Requirements/Specifications}[Tools]
+
+%\terms{Theory}
+
+\keywords{Topics, LDA, NFR}
 

### 2010-08-19 icse/icse2011.tex
fixes for subfigures; fixed Figure references (!!!!)

@@
  \usepackage{booktabs}
-\usepackage{subfigure}
+\usepackage[caption=false]{subfig}
  \usepackage[colorinlistoftodos,backgroundcolor=yellow]{todonotes}
@@
 \usepackage[pdfstartview=FitH,bookmarksopenlevel=3,bookmarks=true]{hyperref} %pdfpagescrop={92 112 523 778},
+
+% % share affiliation and use less space
+% \def\sharedaffiliation{%
+% \end{tabular}
+% \begin{tabular}{c}}
+% %
 
@@
 
-\keywords{Topics, LDA, NFR}
+\keywords{Topic analysis, LDA, non-functional requirements}
 
@@
 MaxDB 7.500 & Total  Topics  & 500 & 500 & 500 \\
+\midrule
 MySQL 3.23  & Named Topics   & 524 & 273 & 773 \\
@@
 \end{tabular}
-	\caption{Automatic topic labelling for MaxDB 7.500}
+	\caption{Automatic topic labelling for MaxDB and MySQL}
 	\label{tbl:wordlist}
@@
   \centering
- \includegraphics[width=.85\textwidth]{figures/unsupervised-bar}
+ \includegraphics[width=\textwidth]{figures/unsupervised-bar}
  \caption{ROC values for each NFR, per project and wordlist.}
 %\todo[inline]{this is too tiny in print}
-
   \label{fig:maxdb-unsup-results}
@@
 
-Figure \ref{fig:best-learn-per-tag} shows that MaxDB and MySQL have quite different results, as the ROC values for reliability and functionality seem swapped between projects.
+Figure \ref{fig:best-learn-per-tag} shows that MaxDB and MySQL have quite different results, as the ROC values for \emph{reliability} and \emph{functionality} seem swapped between projects.
 For both projects Bayesian techniques did the best out of a wide variety of machine learners tested.
@@
 \centering
-\subfigure[MySQL]{
+\subfloat[MySQL]{
 \includegraphics[width=0.4\textwidth]{figures/mysql-supervised}
@@
 }
-\subfigure[MaxDB]{
+\subfloat[MaxDB]{
 \includegraphics[width=0.4\textwidth]{figures/maxdb-supervised}
@@
 }
-\label{fig:best-learn-per-tag}
 \caption[]{ROC value for the best learner per label, compared with ZeroR accuracy values.  Possible values range from 0--1.
@@
 }
+\label{fig:best-learn-per-tag}
 \end{figure*}
@@
 
-Fig. \ref{fig:mulan} presents the results of Mulan's best multi-label
+Figure~\ref{fig:mulan} presents the results of Mulan's best multi-label
 learners for our data. Calibrated Label Ranking (CLR) is a learner
@@
 \centering
-\subfigure[MySQL]{
+\subfloat[MySQL]{
 \includegraphics[width=0.4\textwidth]{figures/multi-label-results-mysql}
@@
 }
-\subfigure[MaxDB]{
+\subfloat[MaxDB]{
 \includegraphics[width=0.4\textwidth]{figures/multi-label-results-maxdb}
@@
 }
-\label{fig:mulan}
 \caption[]{MySQL and MaxDB macro (grey) and micro (black) ROC results per multi-label learner. Possible values range from 0--1.
@@
 }
+\label{fig:mulan}
 \end{figure*}
@@
 \centering
-\subfigure[MySQL 3.23]{
+\subfloat[MySQL 3.23]{
 \includegraphics[width=\textwidth]{figures/mysql-timeline}
@@
 }
-\subfigure[MaxDB 7.500]{                                 
+
+\subfloat[MaxDB 7.500]{                                 
 \includegraphics[width=\textwidth]{figures/maxdb-timeline}
@@
 }
-\label{fig:timelines}
 \caption[]{NFR label per period.
@@
 Grey bars indicate label frequency relative to \emph{that} NFR.
-\todo[inline]{better discussion}
+% \todo[inline]{better discussion}
 }
+\label{fig:timelines}
 \end{figure*}

### 2010-08-19 icse/icse2011.tex
more abstract?

--- 2010-08-19 icse/icse2011.tex
@@
 % the problem
- Unfortunately the output of these tools is often difficult to distinguish and interpret as they are often so abstract, in particular topics that are automatically extracted are word-lists lacking semantic interpretation.
+ Unfortunately the output of these tools is often difficult to distinguish and interpret as they are often so abstract, for instance topic modelling produces word-lists that lack semantic interpretation.
 % the solution 

### 2010-08-19 icse/icse2011.tex
changing intro, tightening language

--- 2010-08-19 icse/icse2011.tex
@@
 % our field
-  Within the field of mining software repositories, many approaches,   such as topic modeling and concept location, rely on the automated   application of machine learning algorithms to corpora.  
+%  Within the field of mining software repositories, a
+Approaches such as topic modeling and concept location rely on the automated application of machine learning algorithms to corpora in order to understand the latent subject matter of software development artifacts.
 % the problem
- Unfortunately the output of these tools is often difficult to distinguish and interpret as they are often so abstract, in particular topics that are automatically extracted are word-lists lacking semantic interpretation.
+ Unfortunately, the output of these tools is often difficult to distinguish and interpret.
+Results are usually abstract, as the topics that are automatically extracted are word-lists lacking semantic interpretation.
 % the solution 
- Thus to have a meaningful discussion about the topics of software development, we must be able to interpret these topics and devise appropriate labels for extracted topics, yet current software engineering topic analysis and concept analysis research relies on manual naming and labelling of topics
+ Therefore, in order to have a meaningful discussion about the topics of software development, we must be able to interpret and devise labels for these topics.
+However, current software engineering topic analysis and concept analysis research relies on manual naming and labelling of topics.
 % the labelling research
-Current topic labelling approaches neither use domain-specific knowledge to improve results, nor contextualize those results for developers.   
+Furthermore, these topic labelling approaches neither use domain-specific knowledge to improve results, nor contextualize those results for developers.   
 % a sub problem
@@
 % what we do
-This paper implements \emph{labelled topic extraction}, in which topics are extracted from commit comments and given labels relating to a cross-project taxonomy. 
+This paper introduces \emph{labelled topic extraction}, in which topics are extracted from commit comments and given labels relating to a cross-project taxonomy. 
 % data mining generalizable
-To ensure we have a generalizable project agnostic approach we label topics by non-functional requirements related to software quality, since there is some shared belief that these qualities apply broadly across many software systems and their development artifacts.
+To ensure we have a generalizable, project-agnostic approach we label topics using non-functional requirements,
+% these qualities 
+as these apply across many software systems.% and their development artifacts.
 % We used real world data
@@
 
-Topics in software engineering practice refer to specific `threads' between project participants.
+Topics in software engineering practice are specific shared discussions between project participants.
 These topics can include particular requirements, maintenance requests, bug fixes or other software engineering tasks.
@@
 %
-Topics can be in several potential states: they may not be immediately resolved, they may re-occur, yet others are only dealt with briefly  (e.g., a bug report that is closed).
+Topics are typically latent, i.e., not explicitly named, and can be in several potential states: they may not be immediately resolved, they may re-occur, yet others are only dealt with briefly  (e.g., a bug report that is closed).
 % semi-automatically sounds too uncertain. I think the introduction can be more muscular and bold. Don't give the reviewer any room to start doubting.
+
+% why are topics useful
+Often developers interact with abstractions of the artifacts from which topics are derived, rather than the artifact itself. 
+Such abstractions include project dashboards~\cite{kersten2005mylar}, tags~\cite{treude2010}, and complexity metrics \cite{mccabe1976complexity}. 
+Topics, too, are a valuable and automatically extractable abstraction of software artifacts.
+It abstracts these artifacts by their subject matter, rather than their concrete representation. 
+Much like Aspect-Oriented Programming, topic labeling can identify cross-cutting concerns in a software project~\cite{Baldi2008}.
+The topic of development is relevant because it identifies the distinct issues facing development at a certain time.
+
 The goal of concept location and topic analysis research is to recover these topics and the related software artifacts.
 % get neil's data mining point in there
-To date, the results of topic analysis and concept location are project specific, and not globally generalizable.
+To date, the results of topic analysis and concept location are project specific, and not generalizable.
 % get the interpretation aspect
-Much of the current research stops short of interpreting the purpose of the topic. That is, extracted topics are usually manually named and labelled by those who read the topic words.
-% 
-In this paper we propose a method of interpreting and labelling topics, and their related artifacts, that is generalizable and relevant to multiple projects.
+Much of the current research stops short of interpreting the purpose of the topic. 
+That is, extracted topics are usually manually named and labelled by those who read the topic words.
+We propose a method of interpreting and labelling topics, and their related artifacts, that is generalizable and relevant to multiple projects.
 
@@
 Topics have to be identified by interpreting the prevalent words in a word distribution and by inspecting related documents. 
-This manual approach is impractical when one has to handle hundreds of different topics; to scale to meaningfully sized projects, we require automatic assistance to determine the topic label.
-
-% why are topics useful
-Often developers interact with abstractions of these artifacts that topics are derived from, rather than the artifact itself. 
-Topics are a valuable and automatically extractable abstraction of software artifacts.
-It abstracts these artifacts by their purpose or topic rather than their concrete representation.
-The topic of development is relevant because it shows anyone inspecting these topics what are the distinct issues or topics facing development at a certain time.
-Such abstractions include project dashboards~\cite{kersten2005mylar}, tags~\cite{treude2010}, and complexity metrics \cite{mccabe1976complexity}.
+This manual approach is impractical when one has to handle hundreds of different topics. To scale to meaningfully sized projects, we need automatic assistance to determine the topic label.
 
@@
 In this paper we will show that NFR related topics occur quite frequently.
-Thus our method of generalizing the data mining of these software artifacts is to leverage  the domain knowledge of software engineering, specifically NFRs. 
-This allows us to data mine software systems with the intention of comparing their NFR related development topics.
+This informs our method of generalizing the data mining of these software artifacts to leverage the global domain knowledge of software engineering, specifically NFRs.
+In particular, we use software quality models to generalize topics across projects.
+This allows us to data mine software systems with the intention of comparing their NFR-related development topics.
 % in order to achieve generalizable, cross project data mining results.
@@
 
-This paper begins by introducing labelled topic extraction.
+This paper begins by introducing labelled topic extraction (Section \ref{sec:wordlist}).
 We show that these labels with their topics can be learned and used to classify other data-sets.
@@
 For \textsf{exp1}, our best performing labels (the labels matched with the most topics) were \emph{correctness} (182 topics) and \emph{testability} (121).
-We did not get good results for usability or accuracy, which were associated with fewer than ten topics.
-We also looked for correlation between our labels: Excluding double matches (self-correlation), our highest co-occurring terms were verifiability and traceability, and testability and correctness (76 and 62 matches, respectively).
+We did not get good results for \emph{usability} or \emph{accuracy}, which were associated with fewer than ten topics.
+We also looked for correlation between our labels: Excluding double matches (self-correlation), our highest co-occurring terms were \emph{verifiability} and \emph{traceability}, and \emph{testability} and \emph{correctness} (76 and 62 matches, respectively).
 
-For \textsf{exp2}, there are many more unnamed topics. Only reliability produces a lot of matches, mostly with the word `error'. Co-occurrence results were poor.
+For \textsf{exp2}, there are many more unnamed topics. Only \emph{reliability} produces a lot of matches, mostly with the word `error'. Co-occurrence results were poor.
 
@@
 See the following sections for our error analysis.
-We found a high of 265 topics for usability, with a low of 44 topics for maintainability.
-Common co-occurrences were reliability and usability, efficiency and reliability, and efficiency and usability (200, 190, and 150 topics in common, respectively).
+We found a high of 265 topics for \emph{usability}, with a low of 44 topics for \emph{maintainability}.
+Common co-occurrences were \emph{reliability} and \emph{usability}, \emph{efficiency} and \emph{reliability}, and \emph{efficiency} and \emph{usability} (200, 190, and 150 topics in common, respectively).
 
@@
 
-Based on these results we find that reliability and usability worked well for MaxDB in \textsf{exp2} and better in \textsf{exp3}.
-MySQL had reasonable results within \textsf{exp2} for reliability and efficiency.
-MySQL's results for efficiency did not improve in \textsf{exp3} but other qualities such as functionality did improve.
+Based on these results we find that \emph{reliability} and \emph{usability} worked well for MaxDB in \textsf{exp2} and better in \textsf{exp3}.
+MySQL had reasonable results within \textsf{exp2} for \emph{reliability} and \emph{efficiency}.
+MySQL's results for efficiency did not improve in \textsf{exp3} but other qualities such as \emph{functionality} did improve.
 If a \emph{C} grade performance has a ROC value of $0.6$ then most of these tests scored a grade of \emph{C} or less, but our results were still better than random chance.
@@
 \textsf{exp3}, because the ROC values are sufficiently higher in most
-cases (other than MaxDB reliability and MySQL efficiency). The
+cases (other than MaxDB \emph{reliability} and MySQL \emph{efficiency}). The
 limitation of the approach we took here is that we assume labels are
@@
 MySQL 3.23 also had more topics and a longer recorded history than MaxDB 7.500.
-We tagged more MySQL topics with annotations than MaxDB topics yet both shared similarities.
-In terms of non-functional requirements both projects had long running trends that focused on functionality, maintainability, and portability, yet MaxDB had more of a focus on efficiency and reliability.
+We tagged more MySQL topics with annotations than MaxDB topics, yet both shared similarities.
+In terms of non-functional requirements both projects had long running trends that focused on \emph{functionality}, \emph{maintainability}, and \emph{portability}, yet MaxDB had more of a focus on \emph{efficiency} and \emph{reliability}.
 MaxDB differed from MySQL since MaxDB was being actively developed, whereas halfway through our history of MySQL 3.23, other versions of MySQL were being actively developed: MySQL versions 4.0, 4.1, and eventually 5.0 and 5.1.
 In other words, MySQL 3.23 was being maintained rather than actively developed, whereas MaxDB 7.500 was being actively developed into MaxDB 7.6, and then maintained thereafter.
+\todo[inline,color=yellow]{expand on this with dates, and reference to visualizations}
 
@@
 \textbf{Do label frequencies change over time?} -- Yes, MySQL's label
-frequencies decreased as it got older. Usability and reliability
-labels became more and more infrequent as it matured. Maintainability topics
+frequencies decreased as it got older. \emph{Usability} and \emph{reliability}
+labels became more and more infrequent as it matured. \emph{Maintainability} topics
 became more prevalent as MaxDB matured.
@@
   interest?} -- Yes. MySQL 3.23 had proportionally more
-functionality labelled topics, while MaxDB had proportionally more
-efficiency and portability related topics.
+topics labelled \emph{functionality}, while MaxDB had proportionally more
+\emph{efficiency} and \emph{portability} related topics.
 
@@
 We found many topics that were not non-functional requirements (NFRs) but were often related to them.
-For instance, concurrency was mentioned often in the commit logs and was related to correctness and reliability, because concurrency was troublesome.
+For instance, concurrency was mentioned often in the commit logs and was related to \emph{correctness} and \emph{reliability}, because concurrency was troublesome.
 Configuration management and source control related changes appeared often and sometimes there were topics dedicated to configuration management.
-These kinds of changes are slightly related to maintainability.
+These kinds of changes are slightly related to \emph{maintainability}.
 A non-functional change that was not quality-related was licensing and copyright; many changes were simply to do with updating copyrights or ensuring copyright or license headers were applied to files.
@@
 In MySQL the optimizer is mentioned but often the change deals with correctness or another quality.
-Despite this difference, the name of the module could fool our learners into believing the change was always about efficiency.
-Perhaps a project specific word-bag is needed in order to avoid automated mistakes due to the names of entities and modules of a software project.
+Despite this difference, the name of the module could fool our learners into believing the change was always about \emph{efficiency}.
+In these cases the advantages of tailoring topic names to specific project terminologies are more clear: this would avoid automated mistakes due to the names of entities and modules of a software project.
 
@@
 Furthermore, our analysis techniques rely on a project's use of meaningful commit messages.
+\todo[inline,color=yellow]{update this with more detail}
 
 \section{Related work}
-
-\begin{comment}
-% Part of our effort with this project is to understand the qualitative and intentional aspects of requirements in software evolution, a notion we first discussed in \cite{ernst07icsm}.
-That idea is derived from, in part, work on narratives of software systems shown in academic work from Ant\'{o}n et al.~\cite{anton01}, or more general-purpose work from Waldo~\cite{waldo93}.
-
-% Part of our effort with this project is to understand the qualitative and intentional aspects of requirements in software evolution, a notion we first discussed in \cite{ernst07icsm}.
-That idea is derived from, in part, work on narratives of software systems shown in academic work like \cite{anton01}, or more general-purpose works like \cite{waldo93}.
-
-% Not sure how do handle this
-\end{comment}
 The idea of extracting higher-level `concerns' (also known as `concepts', `aspects' or `requirements') has been approached in two ways.
@@
 
-\begin{comment}
-% comment out or not
-% Massey~\cite{massey02icse} and Scacchi~\cite{scacchi02,scacchi05b} looked at the topic of requirements in open-source software.
+Massey~\cite{massey02icse} and Scacchi~\cite{scacchi02,scacchi05b} looked at the topic of requirements in open-source software.
 Their work discusses the source of the requirements and how they are used in the development process.

### 2010-08-19 icse/icse2011.tex
More edits

--- 2010-08-19 icse/icse2011.tex
@@
 
+% Maybe specify the hard work we did annotating the data
+
 % our field
@@
 Our work could apply to testers, users, etc.
-In addressing a topic of development stakeholders often produce or modify multiple development artifacts.
+While addressing a topic of development, stakeholders often produce or modify multiple development artifacts.
 %
@@
 %
-Topics can be in several potential states: they may not be immediately resolved, they may re-occur, yet others are only dealt with briefly  (e.g., a bug report that is closed).
+Topics can be in several potential states: they may not be immediately resolved, they may re-occur, yet other topics may be only dealt with briefly, such as a  bug report that is closed.
 % semi-automatically sounds too uncertain. I think the introduction can be more muscular and bold. Don't give the reviewer any room to start doubting.
-The goal of concept location and topic analysis research is to recover these topics and the related software artifacts.
+The goal of concept location and topic analysis research is to recover these topics and their related software artifacts.
 % get neil's data mining point in there
@@
 Deriving such topics is difficult. 
-Raw topics from topic analysis are represented as unlabelled context-free word-lists or word distributions.
+Topic analysis produces raw topics that are represented as unlabelled context-free word-lists or word distributions.
 Topics have to be identified by interpreting the prevalent words in a word distribution and by inspecting related documents. 
@@
 Topics are a valuable and automatically extractable abstraction of software artifacts.
-It abstracts these artifacts by their purpose or topic rather than their concrete representation.
+These artifacts are abstracted by their purpose or topic rather than their concrete representation.
 The topic of development is relevant because it shows anyone inspecting these topics what are the distinct issues or topics facing development at a certain time.
-Such abstractions include project dashboards~\cite{kersten2005mylar}, tags~\cite{treude2010}, and complexity metrics \cite{mccabe1976complexity}.
+Other such abstractions include project dashboards~\cite{kersten2005mylar}, tags~\cite{treude2010}, and complexity metrics \cite{mccabe1976complexity}.
 
@@
 
+\todo[inline]{We still haven't shown that topics are useful -- a reviewer comment}
+
 \section{Study design and execution}
-Figure \ref{fig:process} gives an outline of the methodology followed.
-We began by gathering source data; then applied unsupervised and supervised learning techniques to identify topics.
-We then use these topics to analyze the role of non-functional requirements in software maintenance.
+Figure \ref{fig:process} gives an outline of the methodology followed. We began by gathering source data;
+then we manually annotated the topics of these sources by their related NFRs;
+ then applied unsupervised and supervised learning techniques to identify topics. We then use these topics to analyze the role of non-functional requirements in software maintenance.
 

### 2010-08-19 icse/icse2011.tex

--- 2010-08-19 icse/icse2011.tex
@@
 
-Topics in software engineering practice are specific shared discussions between project participants.
+Topics in software engineering practice refer to specific shared threads of discussions between project stakeholders.
 These topics can include particular requirements, maintenance requests, bug fixes or other software engineering tasks.
@@
 It sounded too vague.
-Also changed from 'developers' to 'participants' because developers sounds too coding oriented.
+Also changed from 'developers' to 'stakeholders' because developers sounds too coding oriented.
 Our work could apply to testers, users, etc.
-In addressing a topic of development participants often produce or modify multiple development artifacts.
+While addressing a topic of development, stakeholders often produce or modify multiple development artifacts.
 %

### 2010-08-19 icse/icse2011.tex
Restore missing conclusions

@@
 \section{Conclusions and future work}
+\todo[inline]{Need to juxtapose our work against previous work -- contribution}
 
-We demonstrated that static but domain-specific knowledge can improve unsupervised labelling of extracted topics.
+% What did we do
+% What did we show
+% What did we do that was new?
+
+% para 1 we did this, we were successful
+In this paper we presented a cross-project data mining technique that was appropriate for mining and labelling non-functional requirements topics and artifacts from project histories.
+Our contributions include:
+\begin{itemize}
+\item We provide an unsupervised method of topic labelling (word-bags);
+%\item Our unsupervised method can be applied to projects without training;
+\item We demonstrate a supervised method of labelling topics by a single NFRs (machine learning);
+\item We demonstrate a supervised method of labelling topics by multiple NFRs (multi-label machine learning);
+\item We provide a method of cross-project analysis via topic labelling;
+\item We demonstrate a method of visualizing these topics across time.
+\end{itemize}
+We leveraged word-bags, machine learning classifiers and multi-label machine learning classifiers.  
+We evaluated these techniques using multiple experiments against two open source database systems and visualized the results.
+We found that our topic labelling techniques were effective.
+
+% para 2 we did this to validate and show we did it
+% Experiment description -- origin of wordbags
+We validated our topic labelling techniques using multiple experiments.
+Using unsupervised word-bag analysis we set up two NFR topic labelling experiments: 
+our \textsf{exp2} experiment used small but accurate word-bags, while \textsf{exp3} used large general word-bags derived from WordNet.
+The performance of these techniques differed per NFR but both \textsf{exp2} and \textsf{exp3} had similar performance.
+%Our \textsf{exp2} experiment used small accurate word bags to label topics but performed just as well as \textsf{exp3}, which used many more general terms from WordNet.
+% we could can improve auto labelling with ML
+Our third experiment showed that with some supervision, and by using efficient machine learners based on Naive Bayesian classifiers, we could improve the accuracy of automatic topic labelling even further.
+Our fourth experiment evaluated the effectiveness of multi-label learners, such as HOMER. 
+Both kinds of learners performed well with area under the ROC curve values between $.6$ and $.8$.
+
+% para 3 look while we did it we found this
+% total topics - maybe move above
+We showed that non-functional requirements are often trending topics, that non-functional requirements are quite common in developer topics, and that there are efficient methods of semi-automating and automating topic labelling.
+% We found NFRs
+Our manual inspection and annotation of the topics extracted from MySQL and MaxDB revealed that many of the extracted topics dealt with non-functional requirements, and these topics were spread across the entire history of a project.
+% Try to provide an example of NFRs but little context
+In the cases of MaxDB and MySQL, portability was a constant maintenance concern and was prevalent throughout the entire lifetime of the projects.
+
+% para 4 summarize and hammer contribution
+In conclusion,
+topic analysis and concept analysis research has produced topics that need to be manually labelled. 
+We found that topics without labels are difficult to interpret.
+Thus we leveraged software engineering knowledge to produce a method of domain specific topic labelling.
+The word-bag technique is not project specific, thus it has intrinsic value when analyzing multiple projects.
+We demonstrated multiple supervised and unsupervised methods of topic labelling, including multi-label learners.
+We provided a case study and visualization of NFR related topics in MaxDB 7.500 and MySQL 3.23. 
+During the case study we manually labelled 100s of topics by hand then used these annotations to validate the positive performance of topic labelling techniques.
+We showed that topics with labels are more useful than unlabelled topics, thus to furnish stakeholders with labelled topics we provided many methods of automatic topic labelling.
+
+\subsection{Future Work}
+
+ % para 5 future work
+ % Future work
+There are several avenues of further investigation.  
+ % Talk to people
+We want to investigate developer attitudes related to these labels: i.e., when we label a topic, was the developer expressing positive or negative qualities about that label?
+It is difficult to map abstract qualities to particular messages.
+ % This is disjointed and needs more context
+On question we seek to answer is if a ``quality'' discussion about more than just corrective maintenance?
+%In this study we take the position that any type of maintenance has to do with software quality.
+
+ %\todo[inline]{Future work is disjointed the question should be more clear}
+
+% Be careful how you merge!
+
+% We demonstrated that static but domain-specific knowledge can improve unsupervised labelling of extracted topics.
 Our \textsf{exp2} experiment used small accurate word bags to label topics but performed just as well as \textsf{exp3}, which used many more general terms from WordNet.
@@
 
-Our manual inspection and annotation of the topics extracted from MySQL and MaxDB revealed that many of the extracted topics dealt with non-functional requirements, and these topics were spread across the entire history of a project.
+% Our manual inspection and annotation of the topics extracted from MySQL and MaxDB revealed that many of the extracted topics dealt with non-functional requirements, and these topics were spread across the entire history of a project.
 In the cases of MaxDB and MySQL, portability was a constant maintenance concern and was prevalent throughout the entire lifetime of the projects.
 
-We showed that non-functional requirements are often trending topics, that non-functional requirements are quite common in developer topics, and that there are efficient methods of semi-automating and automating topic labelling.
+% We showed that non-functional requirements are often trending topics, that non-functional requirements are quite common in developer topics, and that there are efficient methods of semi-automating and automating topic labelling.
 
-There are several avenues of further investigation.
+% There are several avenues of further investigation.
 We want to investigate developer attitudes related to these labels: i.e., when we label a topic, was the developer expressing positive or negative qualities about that label?
@@
 Is a ``quality'' discussion about more than just corrective maintenance?
-%In this study we take the position that any type of maintenance has to do with software quality.
+%In th
+%is study we take the position that any type of maintenance has to do with software quality.
 

### 2010-08-19 icse/icse2011.tex
10 pages or less
generating the data - edit para 1
updated ROC definition/cite; updated unsupervised methodology; table placement

--- 2010-08-19 icse/icse2011.tex
@@
 \usepackage[caption=false]{subfig}
- \usepackage[colorinlistoftodos,backgroundcolor=yellow]{todonotes}
+ \usepackage[disable,colorinlistoftodos,backgroundcolor=yellow]{todonotes}
  \usepackage{color}
@@
 
-% Maybe specify the hard work we did annotating the data
-
 % our field
@@
 
-Topics in software engineering practice refer to specific shared threads of discussions between project stakeholders.
+Topics in software engineering practice are specific shared discussions between project participants.
 These topics can include particular requirements, maintenance requests, bug fixes or other software engineering tasks.
@@
 It sounded too vague.
-Also changed from 'developers' to 'stakeholders' because developers sounds too coding oriented.
+Also changed from 'developers' to 'participants' because developers sounds too coding oriented.
 Our work could apply to testers, users, etc.
-While addressing a topic of development, stakeholders often produce or modify multiple development artifacts.
+In addressing a topic of development participants often produce or modify multiple development artifacts.
 %
@@
 
-% % a rough sketch 
-% The size of these artifacts is cumbersome for a developer or stakeholder to comprehend, thus they often have to be ``chunked'', grouped, modularized, organized and finally labelled.
-% Even more difficult is to pull together tangible trails or interests that pervade these artifacts.
-% These interests or topics are only really visible by analyzing multiple artifacts.
-% Other researchers have focused on extracting these topics from development artifacts, but often their approach leaves the organization or the labelling step to the user.
-% What we propose is to 
-% In this paper, we seek to automatically extract these interests, or topics that link artifacts such as source control commits and label them.
-
-% % Read for SE Data Mining
-% % Read for Topic Naming
-% % Read for previous research DID NOT
-% % thoughts, what does this do what are we introducing
-% % these abstractions ... how to link them
-
-% The act of developing software generates many artifacts, including source code, bug reports, commit messages, and test cases. 
-% It does not take long for a software project to grow to a point where keeping track of these artifacts exceeds easy comprehension. 
-% For these projects, developers interact with abstractions of these artifacts, rather than the artifact itself.
-Such abstractions include project dashboards~\cite{kersten2005mylar}, tags~\cite{treude2010}, and complexity metrics \cite{mccabe1976complexity}.
-
-% Developers do more than merely observe these abstractions, however.
-They must also use them to assess progress and work with colleagues.
-To have meaningful discussions about how software development is progressing, we must be able to devise appropriate labels for our development artifacts.
-Such labels can be used to compare progress or understand how well a project is meeting requirements.
-These labels then become the topics for development conversations (such as the stream of commentary on a bug report).
-One way to identify topics is to use the artifacts themselves to derive them.
-% \todo[inline]{it should be labels for topics not artifacts, I think the artifact part is almost secondary and the topic stuff is really what is up.
-Topics are used below and it isn't even clear what they are or what we're talking about.
-Naming is important but lets tell the reader what is being named first.}
-
 % topics are numerous?
 Deriving such topics is difficult. 
-Topic analysis produces raw topics that are represented as unlabelled context-free word-lists or word distributions.
+Raw topics from topic analysis are represented as unlabelled context-free word-lists or word distributions.
 Topics have to be identified by interpreting the prevalent words in a word distribution and by inspecting related documents. 
@@
 NFRs are prevalent in almost all software systems.
-In this paper we will show that NFR related topics occur quite frequently.
+% In this paper we will show that NFR related topics occur quite frequently.
 This informs our method of generalizing the data mining of these software artifacts to leverage the global domain knowledge of software engineering, specifically NFRs.
@@
   \item Topic modeling creates word lists that require interpretation by the user to assign meaning. Like (1), this means
-that it is difficult to discuss results independent of the project context. Our technique automatically assigns labels agnostic of project.
+that it is difficult to discuss results independent of the project context. Our technique automatically assigns labels across projects.
 \end{enumerate}
@@
 This paper begins by introducing labelled topic extraction (Section \ref{sec:wordlist}).
-We show that these labels with their topics can be learned and used to classify other data-sets.
+We show that these labels with their topics can be learned and used to classify other data-sets, either without training (Section \ref{sec:unsuplabelling}), or with training (Section \ref{sec:suplearn}).
 We then present visualizations of named topics and their trends over time to aid communication and analysis.
-We use an exploratory case study of two open source database systems to show how named topics can be compared between projects.
+We use an exploratory case study of two open source database systems to show how named topics can be compared between projects  (Section \ref{sec:analysis}).
+The paper concludes with a discussion of limitations (Section \ref{sec:limit}) and related work (Section \ref{sec:related}).
 
@@
 
-\todo[inline]{We still haven't shown that topics are useful -- a reviewer comment}
-
 \section{Study design and execution}
-Figure \ref{fig:process} gives an outline of the methodology followed. We began by gathering source data;
-then we manually annotated the topics of these sources by their related NFRs;
- then applied unsupervised and supervised learning techniques to identify topics. We then use these topics to analyze the role of non-functional requirements in software maintenance.
+Figure \ref{fig:process} gives an outline of the methodology followed.
+We began by gathering source data.
+We then applied unsupervised and supervised learning techniques to identify topics.
+These topics were used to analyze the role of non-functional requirements in software maintenance.
 
@@
 MaxDB started in the late 1970s as a research project, and was later acquired by SAP.
-As of version 7.500, released April 2007, the project has 940 thousand lines of C source code~\footnote{generated using David A.
+As of version 7.500, released April 2007, the project has 940 thousand lines of C source code\footnote{generated using David A.
 Wheeler's \emph{SLOCCount}.}.
@@
 This size of period is smaller than the time between minor releases but large enough for there to be sufficient commits to analyze.
-For this set of periods, we created `topics' using Latent Dirichlet Allocation, a topic identification algorithm.
+For this set of periods, we created `topics' using Latent Dirichlet Allocation, a topic identification algorithm~\cite{Blei2003}.
 
 Typically a topic analysis tool like LDA will try to find $N$ independent word distributions within the word distributions of all the messages.
-Linear combinations of these $N$ word distributions are then meant to be able to recreate the word distributions of all of the underlying messages.
+Linear combinations of these $N$ word distributions are then used to recreate the word distributions of the underlying messages.
 These $N$ word distributions effectively form topics: cross-cutting collections of words relevant to one or more of our commit messages.
@@
 While the topic models themselves are generated automatically, what to make of them is less clear.
-For example, in our previous work~\cite{Hindle09ICSM}, as well as in Baldi et al.~\cite{Baldi2008}, topics are named manually: human experts read the highest-frequency members of a topic and assign a keyword accordingly.
+For example,
+% in our previous work~\cite{Hindle09ICSM}, as well as -- CITED a little too often
+in Baldi et al.~\cite{Baldi2008}, topics are named manually: human experts read the highest-frequency members of a topic and assign a keyword accordingly.
 E.g., for the word list \emph{``listener change remove add fire''}, Baldi et al.
@@
 
-\begin{table*}
+\begin{table*}[]
 	\centering
-\begin{tabular}{c|p{9cm}}
+\begin{tabular}{c|p{11cm}}
 \toprule
@@
 
-\todo[inline]{There should be a section here which is about manual annotation. It is really important and it is required by both analyses. Plus we can tell them how.}
+\noindent \textbf{Creating a validation corpus} -- For MySQL 3.23 and MaxDB 7.500, we manually annotated each extracted topic in each period with the same NFR labels as \textsf{exp2} (software qualities).
+We looked at each period's topics, and assessed what the data -- consisting of the frequency-weighted word lists and messages -- suggested was the topic for that period.
+We were able to pinpoint the appropriate label using auxiliary information as well, such as the actual revisions and files that were related to this topic.
+For example, for the MaxDB topic consisting of a message ``exit() only used in non NPTL LINUX Versions'', we tagged that topic \emph{portability}.
+We compared against this data-set, but we also used it for our supervised learning based topic classification.
+A topic may have more than one matching keyword.
+\todo[inline]{inter-rater reliability }
+% \todo[inline]{There should be a section here which is about manual annotation. It is really important and it is required by both analyses. Plus we can tell them how.}
 
 \subsection{Unsupervised labelling}
+\label{sec:unsuplabelling}
+
 % word list similarity
-Using our three word lists, we labeled our topics where there was a match between the list and the topic.
-In general, this approach did not work out well as common labels dominated the less common labels.
-The related words for \emph{correctness}, for example, tended to be too lengthy and non-specific.
+Using our three word lists, we labeled our topics where there was a match between a word in the list and the same word in the topic.
 Table \ref{tbl:wordlist} lists broad labeling results for MaxDB as an example.
@@
 %\todo[inline,color=green]{this table should have mysql}
-\begin{table}
+\begin{table}[h]
 
@@
 \end{tabular}
-	\caption{Automatic topic labelling for MaxDB and MySQL}
+	\caption{Automatic topic labelling for MaxDB and MySQL}% \todo[inline]{this stupid table is before table 1}}
 	\label{tbl:wordlist}
@@
 We also looked for correlation between our labels: Excluding double matches (self-correlation), our highest co-occurring terms were \emph{verifiability} and \emph{traceability}, and \emph{testability} and \emph{correctness} (76 and 62 matches, respectively).
+% See the following sections for our error analysis.
 
-For \textsf{exp2}, there are many more unnamed topics. Only \emph{reliability} produces a lot of matches, mostly with the word `error'. Co-occurrence results were poor.
+For \textsf{exp2}, there are more unnamed topics than \textsf{exp1}.
+Only \emph{reliability} produces a lot of matches, mostly with the word `error'.
+Co-occurrence results were poor.
+This suggests our word lists were overly restrictive.
 
-For \textsf{exp3}, we had many more named topics.
-As we mentioned, the word-lists are quite broad, so there are likely to be false-positives.
-See the following sections for our error analysis.
+For \textsf{exp3}, we generally labelled more topics.
+As we mentioned, the word-lists are quite broad, so there are likely to be false-positives (discussed below).
 We found a high of 265 topics for \emph{usability}, with a low of 44 topics for \emph{maintainability}.
@@
 \subsubsection{Analysis of the unsupervised labelling}
-
-\label{sec:unsuplabelling}
-
 Based on the labels, and our manual topic labelling, we compared the results of the unsupervised word matching approach.
-For each quality we tried to assess whether the manual tag matched the unsupervised label assigned.
+For each quality we assessed whether the manual tag matched the unsupervised label assigned.
 Figure \ref{fig:maxdb-unsup-results} shows our results for MaxDB and MySQL.
-Our error is measured using the ROC area-under-curve value.
-This value is the area under the Receiver Operating Characteristic (\emph{ROC}) curve, sometimes referred to as \emph{AUC}.
-ROC values reflect a score, similar to school letter-grades (A is 0.9, C is 0.6), reflecting how well a particular learner performed for the given data.
-A ROC result of 0.5 would be equivalent to a random learner (randomly classifying data).
+Our error is measured using the \emph{ROC} (Receiver Operating Characteristic) area-under-curve value~\cite{Fawcett2006861}.
+ROC values provide a score, similar to school letter-grades (A is 0.9, C is 0.6), reflecting how well a particular learner performed for the given data.
 ROC maps to the more familiar concepts of precision/sensitivity and recall/specificity: it plots the true positive rate (sensitivity) versus the false positive rate (1 - specificity).
 A perfect learner has a ROC value of 1.0, reflecting perfect recall and precision.
+A ROC result of 0.5 would be equivalent to a random learner (that is, issuing as many false positives as true positives).
+The AUC of a classifier is equivalent to the probability that the classifier will rank a randomly chosen positive instance higher than a randomly chosen negative instance.
 
+\todo[inline]{since ROC is new-sounding, I wonder if a more lengthy explanation would be merited. An obvious question is why not precision/recall?}
 \begin{figure*}
@@
 %In order to validate how effective these word-bag approaches to topic labelling would be we created a validation data set.
- For MySQL 3.23 and MaxDB 7.500, we manually annotated each extracted topic in each period with the same NFR labels as \textsf{exp2} (software qualities).
-We looked at each period's topics, and assessed what the data -- consisting of the frequency-weighted word lists and messages -- suggested was the topic for that period.
-We were able to pinpoint the appropriate label using auxiliary information as well, such as the actual revisions and files that were related to this topic.
-For example, for the MaxDB topic consisting of a message ``exit() only used in non NPTL LINUX Versions'', we tagged that topic \emph{portability}.
-We compared against this data-set, but we also used it for our supervised learning based topic classification.
-A topic may have more than one matching keyword.
-\todo[inline]{inter-rater reliability }
 
@@
 
-\section{NFRs and evolution} 
-\todo[inline]{talk about this topic in one section}
-\todo[inline]{reference project name Nomen more}
+\section{Using \textbf{Nomen} to understand software maintenance activities} 
+\label{sec:analysis}
+% \todo[inline]{talk about this topic in one section}
+% \todo[inline]{reference project name Nomen more}
 Using the classification results from the previous sections, we evaluated two research questions:
@@
 % \end{figure*}
-
-\section{NFRs between projects}
 
@@
 \section{Limitations}
+\label{sec:limit}
 \subsection{Effectiveness}
@@
 %{Discuss how these are CHEAP techniques and although somewhat %inaccurate they don't require a tonne of work}
+The unsupervised labelling had difficulty distinguishing between common labels and infrequent labels. The related words for \emph{correctness}, for example, tended to be too lengthy and non-specific. 
+\todo[inline,color=red]{Abram can you explain what you mean here?}
 
@@
 \section{Related work}
+\label{sec:related}
 The idea of extracting higher-level `concerns' (also known as `concepts', `aspects' or `requirements') has been approached in two ways.
@@
 \end{document}
+
+% % a rough sketch 
+% The size of these artifacts is cumbersome for a developer or stakeholder to comprehend, thus they often have to be ``chunked'', grouped, modularized, organized and finally labelled.
+% Even more difficult is to pull together tangible trails or interests that pervade these artifacts.
+% These interests or topics are only really visible by analyzing multiple artifacts.
+% Other researchers have focused on extracting these topics from development artifacts, but often their approach leaves the organization or the labelling step to the user.
+% What we propose is to 
+% In this paper, we seek to automatically extract these interests, or topics that link artifacts such as source control commits and label them.
+
+% % Read for SE Data Mining
+% % Read for Topic Naming
+% % Read for previous research DID NOT
+% % thoughts, what does this do what are we introducing
+% % these abstractions ... how to link them
+
+% The act of developing software generates many artifacts, including source code, bug reports, commit messages, and test cases. 
+% It does not take long for a software project to grow to a point where keeping track of these artifacts exceeds easy comprehension. 
+% For these projects, developers interact with abstractions of these artifacts, rather than the artifact itself.
+Such abstractions include project dashboards~\cite{kersten2005mylar}, tags~\cite{treude2010}, and complexity metrics \cite{mccabe1976complexity}.
+
+% Developers do more than merely observe these abstractions, however.
+They must also use them to assess progress and work with colleagues.
+To have meaningful discussions about how software development is progressing, we must be able to devise appropriate labels for our development artifacts.
+Such labels can be used to compare progress or understand how well a project is meeting requirements.
+These labels then become the topics for development conversations (such as the stream of commentary on a bug report).
+One way to identify topics is to use the artifacts themselves to derive them.
+% \todo[inline]{it should be labels for topics not artifacts, I think the artifact part is almost secondary and the topic stuff is really what is up.
+Topics are used below and it isn't even clear what they are or what we're talking about.
+Naming is important but lets tell the reader what is being named first.}

### 2010-08-19 icse/icse2011.tex
updated ROC definition/cite; updated unsupervised methodology; table placement

@@
 \usepackage[caption=false]{subfig}
- \usepackage[disable,colorinlistoftodos,backgroundcolor=yellow]{todonotes}
+ \usepackage[colorinlistoftodos,backgroundcolor=yellow]{todonotes} %disable
  \usepackage{color}
@@
 %\todo[inline]{Topics are poorly defined by this point and since we're building on research which didn't DO this, we should mention that research. Previous work probably should go here.}
+
+\todo[inline]{We still haven't shown that topics are useful -- a reviewer comment}
 
@@
 
-\todo[inline]{Update this}
+%\todo[inline]{Update this}
 
@@
 Furthermore, our analysis techniques rely on a project's use of meaningful commit messages.
-\todo[inline,color=yellow]{update this with more detail}
+\todo[inline]{update this with more detail}
 
@@
 \section{Conclusions and future work}
+%\todo[inline]{Need to juxtapose our work against previous work -- contribution}
 
-We demonstrated that static but domain-specific knowledge can improve unsupervised labelling of extracted topics.
+% What did we do
+% What did we show
+% What did we do that was new?
+
+% para 1 we did this, we were successful
+In this paper we presented a cross-project data mining technique that was appropriate for mining and labelling non-functional requirements topics and artifacts from project histories.
+Our contributions include:
+\begin{itemize}
+\item We provide an unsupervised method of topic labelling (word-bags);
+%\item Our unsupervised method can be applied to projects without training;
+\item We demonstrate a supervised method of labelling topics by a single NFRs (machine learning);
+\item We demonstrate a supervised method of labelling topics by multiple NFRs (multi-label machine learning);
+\item We provide a method of cross-project analysis via topic labelling;
+\item We demonstrate a method of visualizing these topics across time.
+\end{itemize}
+We leveraged word-bags, machine learning classifiers and multi-label machine learning classifiers.  
+We evaluated these techniques using multiple experiments against two open source database systems and visualized the results.
+We found that our topic labelling techniques were effective.
+
+% para 2 we did this to validate and show we did it
+% Experiment description -- origin of wordbags
+We validated our topic labelling techniques using multiple experiments.
+Using unsupervised word-bag analysis we set up two NFR topic labelling experiments: 
+our \textsf{exp2} experiment used small but accurate word-bags, while \textsf{exp3} used large general word-bags derived from WordNet.
+The performance of these techniques differed per NFR but both \textsf{exp2} and \textsf{exp3} had similar performance.
+%Our \textsf{exp2} experiment used small accurate word bags to label topics but performed just as well as \textsf{exp3}, which used many more general terms from WordNet.
+% we could can improve auto labelling with ML
+Our third experiment showed that with some supervision, and by using efficient machine learners based on Naive Bayesian classifiers, we could improve the accuracy of automatic topic labelling even further.
+Our fourth experiment evaluated the effectiveness of multi-label learners, such as HOMER. 
+Both kinds of learners performed well with area under the ROC curve values between $.6$ and $.8$.
+
+% para 3 look while we did it we found this
+% total topics - maybe move above
+We showed that non-functional requirements are often trending topics, that non-functional requirements are quite common in developer topics, and that there are efficient methods of semi-automating and automating topic labelling.
+% We found NFRs
+Our manual inspection and annotation of the topics extracted from MySQL and MaxDB revealed that many of the extracted topics dealt with non-functional requirements, and these topics were spread across the entire history of a project.
+% Try to provide an example of NFRs but little context
+In the cases of MaxDB and MySQL, portability was a constant maintenance concern and was prevalent throughout the entire lifetime of the projects.
+
+% para 4 summarize and hammer contribution
+In conclusion,
+topic analysis and concept analysis research has produced topics that need to be manually labelled. 
+We found that topics without labels are difficult to interpret.
+Thus we leveraged software engineering knowledge to produce a method of domain specific topic labelling.
+The word-bag technique is not project specific, thus it has intrinsic value when analyzing multiple projects.
+We demonstrated multiple supervised and unsupervised methods of topic labelling, including multi-label learners.
+We provided a case study and visualization of NFR related topics in MaxDB 7.500 and MySQL 3.23. 
+During the case study we manually labelled 100s of topics by hand then used these annotations to validate the positive performance of topic labelling techniques.
+We showed that topics with labels are more useful than unlabelled topics, thus to furnish stakeholders with labelled topics we provided many methods of automatic topic labelling.
+
+\subsection{Future Work}
+
+ % para 5 future work
+ % Future work
+There are several avenues of further investigation.  
+ % Talk to people
+We want to investigate developer attitudes related to these labels: i.e., when we label a topic, was the developer expressing positive or negative qualities about that label?
+It is difficult to map abstract qualities to particular messages.
+ % This is disjointed and needs more context
+On question we seek to answer is if a ``quality'' discussion about more than just corrective maintenance?
+%In this study we take the position that any type of maintenance has to do with software quality.
+
+ %\todo[inline]{Future work is disjointed the question should be more clear}
+
+% Be careful how you merge!
+
+% We demonstrated that static but domain-specific knowledge can improve unsupervised labelling of extracted topics.
 Our \textsf{exp2} experiment used small accurate word bags to label topics but performed just as well as \textsf{exp3}, which used many more general terms from WordNet.

### 2010-08-19 icse/icse2011.tex
generating the data edits -- note I replaced . with .\n

--- 2010-08-19 icse/icse2011.tex
@@
 We selected systems from the same application domain, so the functional requirements would be broadly similar. 
-We selected MySQL and MaxDB as they were open-source, partially-commercial database systems.
+We selected MySQL 3.23 and MaxDB 7.500 as they were open-source, partially-commercial database systems. 
 MaxDB started in the late 1970s as a research project, and was later acquired by SAP. 
-As of version 7.500, released April 2007, the project has 940 thousand lines of C source code~\footnote{generated using David A.
-Wheeler's \emph{SLOCCount}.}.
+As of version 7.500, released April 2007, the project has 940 thousand lines of C source code~\footnote{generated using David A. Wheeler's \emph{SLOCCount}.}. 
 The MySQL project started in 1994 and MySQL 3.23 was released in early 2001. 
@@
 For each project, we used source control commit comments, the messages that programmers write when they commit revisions to a source control repository. 
-This is the most readily accessible source of project interactions for outside researchers; bug reports and email messages were not available for both projects.
+This is the most readily accessible source of project interactions for outside researchers; bug trackers and email-list archives were not available for both projects. 
 An example of a typical commit message is: \textit{``history annotate diffs bug fixed (if mysql\-\_real\-\_connect() failed there were two pointers to malloc'ed strings, with memory corruption on free(), of course)''}.
@@
 We summarized each message as a word distribution but removed stop-words such as \emph{the} and \emph{at}. 
-Stemming was performed in the later stages of our analysis.
+Stemming was performed in the later stages of our analysis. % WAS IT?
 
-From that data-set, we created an XML file which separated commits into 30 day periods.
-This size of period is smaller than the time between minor releases but large enough for there to be sufficient commits to analyze.
-For this set of periods, we created `topics' using Latent Dirichlet Allocation, a topic identification algorithm.
+For the commit message data-sets of each project, we created an XML file which separated commits into 30 day periods. 
+This size of period, 30 days, is smaller than the time between minor releases but large enough for there to be sufficient commits to analyze. 
+For each period of each project, we input the messages of that period into Latent Dirichlet Allocation (LDA), a topic analysis algorithm, and recorded the topics of each period.
+%For each project and for each period of messages a project, we input these messages 
+%we created `topics' using Latent Dirichlet Allocation, a topic identification algorithm.
 
 Typically a topic analysis tool like LDA will try to find $N$ independent word distributions within the word distributions of all the messages. 
-Linear combinations of these $N$ word distributions are then meant to be able to recreate the word distributions of all of the underlying messages.
+Linear combinations of these $N$ word distributions are meant to represent and recreate the word distributions of any of the original messages. 
 These $N$ word distributions effectively form topics: cross-cutting collections of words relevant to one or more of our commit messages. 
-LDA extracts topics in an unsupervised manner; the algorithm relies solely on the source data with no human intervention.
+LDA extracts topics in an unsupervised manner; the algorithm relies solely on the source data, word distributions of messages, with no human intervention.
 
@@
 Representing documents as a mixture of topics maps well to source code repository commits, which often have more than one purpose~\cite{Hindle09ICSM}.  
-A topic represents both a word distribution and a group of commit log comments that are related to each other by their content.
+A topic, in this paper, represents both a word distribution and a group of commit log comments that are related to each other by their content.  
 In this paper a topic is a set of tokens extracted from commit messages found within a project's source control system (SCS).
-\todo[inline]{These paragraphs should've come way earlier, in previous we should explain LDA so this doesn't need to be here.}
+%\todo[inline]{These paragraphs should've come way earlier, in previous we should explain LDA so this doesn't need to be here.}
 
@@
 \section{Conclusions and future work}
-\todo[inline]{Need to juxtapose our work against previous work -- contribution}
+%\todo[inline]{Need to juxtapose our work against previous work -- contribution}
 

### 2010-08-19 icse/icse2011.tex

--- 2010-08-19 icse/icse2011.tex
@@
 \label{sec:wordlist}
+
 To evaluate our approach, we sought candidate systems that were mature projects and had openly accessible source control repositories. 
 We selected systems from the same application domain, so the functional requirements would be broadly similar. 
-We selected MySQL and MaxDB as they were open-source, partially-commercial database systems.
+We selected MySQL 3.23 and MaxDB 7.500 as they were open-source, partially-commercial database systems. 
 MaxDB started in the late 1970s as a research project, and was later acquired by SAP. 
-As of version 7.500, released April 2007, the project has 940 thousand lines of C source code\footnote{generated using David A.
-Wheeler's \emph{SLOCCount}.}.
+As of version 7.500, released April 2007, the project has 940 thousand lines of C source code~\footnote{generated using David A. Wheeler's \emph{SLOCCount}.}. 
 The MySQL project started in 1994 and MySQL 3.23 was released in early 2001. 
@@
 We summarized each message as a word distribution but removed stop-words such as \emph{the} and \emph{at}. 
-Stemming was performed in the later stages of our analysis.
+Stemming was performed in the later stages of our analysis. % WAS IT?
 
-From that data-set, we created an XML file which separated commits into 30 day periods.
-This size of period is smaller than the time between minor releases but large enough for there to be sufficient commits to analyze.
-For this set of periods, we created `topics' using Latent Dirichlet Allocation, a topic identification algorithm~\cite{Blei2003}.
+For the commit message data-sets of each project, we created an XML file which separated commits into 30 day periods. 
+This size of period, 30 days, is smaller than the time between minor releases but large enough for there to be sufficient commits to analyze. 
+For each period of each project, we input the messages of that period into Latent Dirichlet Allocation (LDA), a topic analysis algorithm~\cite{Blei2003}, and recorded the topics of each period.
+%For each project and for each period of messages a project, we input these messages 
+%we created `topics' using Latent Dirichlet Allocation, a topic identification algorithm.
 
 Typically a topic analysis tool like LDA will try to find $N$ independent word distributions within the word distributions of all the messages. 
-Linear combinations of these $N$ word distributions are then used to recreate the word distributions of the underlying messages.
+Linear combinations of these $N$ word distributions are meant to represent and recreate the word distributions of any of the original messages. 
 These $N$ word distributions effectively form topics: cross-cutting collections of words relevant to one or more of our commit messages. 
-LDA extracts topics in an unsupervised manner; the algorithm relies solely on the source data with no human intervention.
+LDA extracts topics in an unsupervised manner; the algorithm relies solely on the source data, word distributions of messages, with no human intervention.
 

### 2010-08-19 icse/icse2011.tex
generating word list edits

@@
 \subsection{Generating word lists}
-Topics are word distributions: lists of words ranked by frequency, which can be burdensome to interpret and hard to distinguish and understand.
-While the topic models themselves are generated automatically, what to make of them is less clear.
+\label{sec:genwordlist}
+
+Topics are word distributions, lists of words ranked by frequency, that are be burdensome to interpret and hard to distinguish and understand. 
+While the topic models themselves are generated automatically, how to interpret these topics is less clear.
 For example,
 % in our previous work~\cite{Hindle09ICSM}, as well as -- CITED a little too often
-in Baldi et al.~\cite{Baldi2008}, topics are named manually: human experts read the highest-frequency members of a topic and assign a keyword accordingly.
-E.g., for the word list \emph{``listener change remove add fire''}, Baldi et al.
-assign the keyword \emph{event-handling}.
+in Baldi et al.~\cite{Baldi2008}, topics are named manually: human experts read the highest-frequency members of a topic and assign a label accordingly. 
+Given the word list \emph{``listener change remove add fire''}, Baldi et al. would assign the label \emph{event-handling}. 
 The labels are reasonable enough, but still require an expert in the field to determine them. 
@@
 Accordingly, we tried to associate each topic with a label from a list of keywords and related terms. 
-We performed simple string matching between these topics and our lists, `naming' a topic if it contained that word or words (a topic might match several keywords).
-We used several different word lists for comparison, which we call \textsf{exp1, exp2, exp3} in the text which follows.
+We intersected the words of the topics and the words of our word-lists.
+We `named' a topic if any of its words matched any of the word-list's words.
+A topic could match more than one keyword.
+We used several different sets of word-lists for comparison, which we refer to as \textsf{exp1, exp2, exp3} in the text which follows.
 
-Our first word list set, \textsf{exp1}, was generated using the ontology described in Kayed et al.~\cite{5072519}.
+Our first word-list set, \textsf{exp1}, was generated using the ontology described in Kayed et al.~\cite{5072519}. 
 That paper constructs an ontology for software quality measurement using eighty source documents, including research papers and international standards. 
@@
 
-Our second word list set, \textsf{exp2}, relied on the ISO quality model (ISO9126)~\cite{iso9126}.
+Our second word list set, \textsf{exp2}, relied on the ISO quality model, ISO9126~\cite{iso9126}. 
 ISO9126 describes six high-level non-functional requirements (listed in Table \ref{tbl:wnsig}). 
@@
 We then added Boehm's 1976 software quality model \cite{Boehm+:1976:ICSE}, and classified his eleven `ilities' into their respective ISO9126 qualities. 
-We did the same for the quality model produced by McCall et al.
-\cite{mccall1977}.
-Finally, we analyzed two mailing lists from another software project to expand our set with domain-specific terms.
+We did the same for the quality model produced by McCall et al. \cite{mccall1977}. 
+Finally, we analyzed two mailing lists from another software project, KDE, to expand our set with domain-specific terms. 
 For example, we add the term ``performance'' to the synonyms for \emph{efficiency}, since this term occurs in most mail messages that discuss efficiency.
 
-For the third list of quality labels,  \textsf{exp3}, we extended the list from \textsf{exp2} using unfiltered WordNet similarity matches.
+For the third set of word-lists,  \textsf{exp3}, we extended the word-lists from \textsf{exp2} using unfiltered WordNet similarity matches. 
 Similarity in WordNet means siblings in a hypernym tree. 
 We do not include these words here for space considerations (but see the Appendix for our data repository). 
-It is not clear the words associated with our labels are specific enough, however: for example, the label \emph{maintainability} is associated with words \emph{ease} and \emph{ownership}.
+It is not clear the words associated with our labels are specific enough. For example, the label \emph{maintainability} is associated with words \emph{ease} and \emph{ownership}.
 
@@
 
-\noindent \textbf{Creating a validation corpus} -- For MySQL 3.23 and MaxDB 7.500, we manually annotated each extracted topic in each period with the same NFR labels as \textsf{exp2} (software qualities).
-We looked at each period's topics, and assessed what the data -- consisting of the frequency-weighted word lists and messages -- suggested was the topic for that period.
-We were able to pinpoint the appropriate label using auxiliary information as well, such as the actual revisions and files that were related to this topic.
+\noindent \textbf{Creating a validation corpus} --- For MySQL 3.23 and MaxDB 7.500, we manually annotated each extracted topic in each period with the same NFR labels as \textsf{exp2} (software qualities).
+We looked at each period's topics, and assessed what the data --- consisting of the frequency-weighted word lists and messages --- suggested was the topic for that period. 
+We were able to pinpoint the appropriate label using auxiliary information as well, such as the actual revisions and files that were related to the topic being annotated.
 For example, for the MaxDB topic consisting of a message ``exit() only used in non NPTL LINUX Versions'', we tagged that topic \emph{portability}. 
-We compared against this data-set, but we also used it for our supervised learning based topic classification.
-A topic may have more than one matching keyword.
+We compared against this data-set, but we also used it for our supervised machine learning based topic classification. 
+% A topic may have more than one matching keyword. % ok what?
 \todo[inline]{inter-rater reliability }

### 2010-08-19 icse/icse2011.tex
moving comments on methodology up in the outline, closer to section it came from

--- 2010-08-19 icse/icse2011.tex
@@
 \label{sec:wordlist}
-
 To evaluate our approach, we sought candidate systems that were mature projects and had openly accessible source control repositories.
 We selected systems from the same application domain, so the functional requirements would be broadly similar.
-We selected MySQL 3.23 and MaxDB 7.500 as they were open-source, partially-commercial database systems. 
+We selected MySQL and MaxDB as they were open-source, partially-commercial database systems.
 MaxDB started in the late 1970s as a research project, and was later acquired by SAP.
-As of version 7.500, released April 2007, the project has 940 thousand lines of C source code~\footnote{generated using David A. Wheeler's \emph{SLOCCount}.}. 
+As of version 7.500, released April 2007, the project has 940 thousand lines of C source code\footnote{generated using David A.
+Wheeler's \emph{SLOCCount}.}.
 The MySQL project started in 1994 and MySQL 3.23 was released in early 2001.
@@
 For each project, we used source control commit comments, the messages that programmers write when they commit revisions to a source control repository.
-This is the most readily accessible source of project interactions for outside researchers; bug trackers and email-list archives were not available for both projects. 
+This is the most readily accessible source of project interactions for outside researchers; bug reports and email messages were not available for both projects.
 An example of a typical commit message is: \textit{``history annotate diffs bug fixed (if mysql\-\_real\-\_connect() failed there were two pointers to malloc'ed strings, with memory corruption on free(), of course)''}.
@@
 We summarized each message as a word distribution but removed stop-words such as \emph{the} and \emph{at}.
-Stemming was performed in the later stages of our analysis. % WAS IT?
+Stemming was performed in the later stages of our analysis.
 
-For the commit message data-sets of each project, we created an XML file which separated commits into 30 day periods. 
-This size of period, 30 days, is smaller than the time between minor releases but large enough for there to be sufficient commits to analyze. 
-For each period of each project, we input the messages of that period into Latent Dirichlet Allocation (LDA), a topic analysis algorithm~\cite{Blei2003}, and recorded the topics of each period.
-%For each project and for each period of messages a project, we input these messages 
-%we created `topics' using Latent Dirichlet Allocation, a topic identification algorithm.
+From that data-set, we created an XML file which separated commits into 30 day periods.
+This size of period is smaller than the time between minor releases but large enough for there to be sufficient commits to analyze.
+For this set of periods, we created `topics' using Latent Dirichlet Allocation, a topic identification algorithm~\cite{Blei2003}.
 
 Typically a topic analysis tool like LDA will try to find $N$ independent word distributions within the word distributions of all the messages.
-Linear combinations of these $N$ word distributions are meant to represent and recreate the word distributions of any of the original messages. 
+Linear combinations of these $N$ word distributions are then used to recreate the word distributions of the underlying messages.
 These $N$ word distributions effectively form topics: cross-cutting collections of words relevant to one or more of our commit messages.
-LDA extracts topics in an unsupervised manner; the algorithm relies solely on the source data, word distributions of messages, with no human intervention.
+LDA extracts topics in an unsupervised manner; the algorithm relies solely on the source data with no human intervention.
 
@@
 Representing documents as a mixture of topics maps well to source code repository commits, which often have more than one purpose~\cite{Hindle09ICSM}.
-A topic, in this paper, represents both a word distribution and a group of commit log comments that are related to each other by their content.  
+A topic represents both a word distribution and a group of commit log comments that are related to each other by their content.
 In this paper a topic is a set of tokens extracted from commit messages found within a project's source control system (SCS).
-%\todo[inline]{These paragraphs should've come way earlier, in previous we should explain LDA so this doesn't need to be here.}
+\todo[inline]{These paragraphs should've come way earlier, in previous we should explain LDA so this doesn't need to be here.}
 
@@
 MySQL's results for efficiency did not improve in \textsf{exp3} but other qualities such as \emph{functionality} did improve.
-If a \emph{C} grade performance has a ROC value of $0.6$ then most of these tests scored a grade of \emph{C} or less, but our results were still better than random chance.
+If a \emph{C} grade performance has a ROC value of $0.6$ then most of these tests scored a grade of \emph{C} or less, but our classifier nonetheless performed substantially better than random.
 
@@
 
+\subsection{Annotation observations}
+We found many topics that were not non-functional requirements (NFRs) but were often related to them.
+For instance, concurrency was mentioned often in the commit logs and was related to \emph{correctness} and \emph{reliability}, because concurrency was troublesome.
+Configuration management and source control related changes appeared often and sometimes there were topics dedicated to configuration management.
+These kinds of changes are slightly related to \emph{maintainability}.
+A non-functional change that was not quality-related was licensing and copyright; many changes were simply to do with updating copyrights or ensuring copyright or license headers were applied to files.
+
+We noticed that occasionally the names of modules would conflict with words related to other non-functional requirements.
+For instance, optimizers are very common modules in database systems: both MySQL and MaxDB have optimizer modules.
+In MySQL the optimizer is mentioned but often the change deals with correctness or another quality.
+Despite this difference, the name of the module could fool our learners into believing the change was always about \emph{efficiency}.
+In these cases the advantages of tailoring topic names to specific project terminologies are more clear: this would avoid automated mistakes due to the names of entities and modules of a software project.
+
+\begin{comment}
+A quick mind dump before I forget:
+* portability was very popular
+* there were many instances of concurrency
+* There were often topics of crap commits with no comment
+* might need more than the top words to compare
+* Things like tests were noticable more by file type
+* Configuration management is not part of the software quality ontology?
+* I didn't tag none I tagged unknown, so that's some inconsistency between us
+* There were a lot of copyright changes in mysql
+* mysql had optimizer module which did not mean optimization
+* I tried to assign reliability/bug for reliability bugs
+  and Functionality/Correctness for fixes
+* Some security changes came later
+* Mysql 3.23 transitioned from a project being developed to
+  a project being maintained, more frequent back-port bug-fixes
+\end{comment}
+
 \section{Using \textbf{Nomen} to understand software maintenance activities} 
@@
 
-For example, in Fig. \ref{fig:mysql-timeline}, we see that the NFRs functionality, portability and maintainability contain more labeled topics, since these NFRs have more intense shaded boxes.
+For example, in Fig.
+\ref{fig:mysql-timeline}, we see that the NFRs \emph{functionality}, \emph{portability} and \emph{maintainability} contain more labeled topics, since these NFRs have more intense shaded boxes.
 
@@
 Box intensity (saturation) is mapped to label frequency relative to the largest label value of \emph{all} NFRs.
-Grey bars indicate label frequency relative to \emph{that} NFR.
+Grey bars indicate label frequency relative to \emph{a particular} NFR.
 % \todo[inline]{better discussion}
@@
 
-\end{comment}
-
-\subsection{Annotation observations}
-
-We found many topics that were not non-functional requirements (NFRs) but were often related to them.
-For instance, concurrency was mentioned often in the commit logs and was related to \emph{correctness} and \emph{reliability}, because concurrency was troublesome.
-Configuration management and source control related changes appeared often and sometimes there were topics dedicated to configuration management.
-These kinds of changes are slightly related to \emph{maintainability}.
-A non-functional change that was not quality-related was licensing and copyright; many changes were simply to do with updating copyrights or ensuring copyright or license headers were applied to files.
-
-We noticed that occasionally the names of modules would conflict with words related to other non-functional requirements.
-For instance, optimizers are very common modules in database systems: both MySQL and MaxDB have optimizer modules.
-In MySQL the optimizer is mentioned but often the change deals with correctness or another quality.
-Despite this difference, the name of the module could fool our learners into believing the change was always about \emph{efficiency}.
-In these cases the advantages of tailoring topic names to specific project terminologies are more clear: this would avoid automated mistakes due to the names of entities and modules of a software project.
-
-\begin{comment}
-A quick mind dump before I forget:
-* portability was very popular
-* there were many instances of concurrency
-* There were often topics of crap commits with no comment
-* might need more than the top words to compare
-* Things like tests were noticable more by file type
-* Configuration management is not part of the software quality ontology?
-* I didn't tag none I tagged unknown, so that's some inconsistency between us
-* There were a lot of copyright changes in mysql
-* mysql had optimizer module which did not mean optimization
-* I tried to assign reliability/bug for reliability bugs
-  and Functionality/Correctness for fixes
-* Some security changes came later
-* Mysql 3.23 transitioned from a project being developed to
-  a project being maintained, more frequent back-port bug-fixes
 \end{comment}
@@
 % para 1 we did this, we were successful
-In this paper we presented a cross-project data mining technique that was appropriate for mining and labelling non-functional requirements topics and artifacts from project histories.
+In this paper we presented a cross-project data mining technique that was appropriate for mining and labelling latent software development topics using a non-functional requirements taxonomy.
+%topics and artifacts from project histories.
 Our contributions include:

### 2010-08-19 icse/icse2011.tex
version for Jorge

@@
 \usepackage[caption=false]{subfig}
- \usepackage[colorinlistoftodos,backgroundcolor=yellow]{todonotes} %disable
+ \usepackage[colorinlistoftodos,backgroundcolor=yellow,disable]{todonotes} %disable
  \usepackage{color}
@@
 There are two measures represented. 
-One, the relative frequency, shown in the grey boxes, represents the number of labels with that NFR in that period, relative to the maximum number of topics assigned to that NFR.
-The second, absolute frequency, compares the number of topics labelled with that NFR per period relative to the maximum number of labelled topics overall.
+One, the relative frequency, shown in the grey boxes, represents the number of labels with that NFR in that period, 
+relative to the maximum number of topics assigned to that NFR. 
+The second, absolute frequency, compares the number of topics labelled with that NFR per period 
+relative to the maximum number of labelled topics overall.
 
 For example, in Fig.
-\ref{fig:mysql-timeline}, we see that the NFRs \emph{functionality}, \emph{portability} and \emph{maintainability} contain more labeled topics, since these NFRs have more intense shaded boxes.
+\ref{fig:mysql-timeline}, we see that the NFRs \emph{functionality}, \emph{portability} and \emph{maintainability} contain more labeled topics, since these NFRs have more intensely shaded boxes.
+We note that the gap in the MaxDB data is due to an interruption in the source data, having to do with the infrastructure for releasing data, rather than specific occurrences in the project.
+\todo[inline]{can I pin this down more ?}
+
+\todo[inline]{try to pin down these reasons below}%  MySQL 3.23 was put on maintenance.
+% 
+% MySQL forks repositories per every version and 3.23 was supported for a
+% long time. there was 4.0 4.1 5.0 5.1
+% 
+% This is the MySQL 3.23 branch
+% 
+% Same thing with MaxDB 7.5
+% 7.6 is in a different repository
 
@@
 %\todo[inline]{Need to juxtapose our work against previous work -- contribution}
+\todo[inline]{tighten... typically conclusions are much shorter.}
 

### 2010-08-19 icse/icse2011.tex
changes to visualization discussion. Added notes on releases of both dbs.

@@
 \usepackage[caption=false]{subfig}
- \usepackage[colorinlistoftodos,backgroundcolor=yellow,disable]{todonotes} %disable
+ \usepackage[colorinlistoftodos,backgroundcolor=yellow]{todonotes} %disable
  \usepackage{color}
@@
 
-\conferenceinfo{ICSE}{'11 Waikiki, Hawaii USA}
+\conferenceinfo{ICSE}{'11 Waikiki, Hawaii, USA}
 \numberofauthors{5}
@@
 
-\subsection{Summary of techniques}
-While unsupervised techniques (LSI and LDA are both unsupervised) are appealing in their lack of human intervention, and thus lower effort, supervised learners have the advantage of domain knowledge which typically means improved results.
-
-Very rarely did \textsf{exp2} and \textsf{exp3} (naive word matching) ever perform as well as the supervised machine learners.
-For MaxDB, \textit{reliability} was slightly better detected using the static word list of \textsf{exp2}.
-In general, the machine learners and \textsf{exp3} did better than \textsf{exp2} for both MaxDB and MySQL.
-For both MySQL and MaxDB \textit{usability} was better served by \textsf{exp2}.
-\textit{Usability} was a very infrequent label, however, which made it difficult to detect in any case.
-
-We found that the multi-label learners of BR, CLR and HOMER did not do as well for Macro-ROC as NaiveBayes and other NaiveBayes derived learners.
-This suggests that by sticking together multiple NaiveBayes learners we could probably label sets of topics effectively, but it would require a separate NaiveBayes learner per label.
-
 \subsection{Annotation observations}
@@
 
+\subsection{Summary of techniques}
+While unsupervised techniques (LSI and LDA are both unsupervised) are appealing in their lack of human intervention, and thus lower effort, supervised learners have the advantage of domain knowledge which typically means improved results.
+Our manual annotations were fairly quick to do, taking on average five minutes per 30 day period.
+
+Very rarely did \textsf{exp2} and \textsf{exp3} (naïve word matching) ever perform as well as the supervised machine learners.
+For MaxDB, \textit{reliability} was slightly better detected using the static word list of \textsf{exp2}.
+In general, the machine learners and \textsf{exp3} did better than \textsf{exp2} for both MaxDB and MySQL. 
+For both MySQL and MaxDB \textit{usability} was better served by \textsf{exp2}. 
+\textit{Usability} was a very infrequent label, however, which made it difficult to detect in any case.
+
+We found that the multi-label learners of BR, CLR and HOMER did not do as well for Macro-ROC as NaiveBayes and other NaiveBayes derived learners. 
+This suggests that by sticking together multiple NaiveBayes learners we could probably label sets of topics effectively, but it would require a separate NaiveBayes learner per label.
+
 \section{Using \textbf{Nomen} to understand software maintenance activities} 
@@
 % \todo[inline]{reference project name Nomen more}
+
 Using the classification results from the previous sections, we evaluated two research questions:
@@
 
-For example, in Fig.
+\textbf{Understanding the figures} -- For example, in Fig.
 \ref{fig:mysql-timeline}, we see that the NFRs \emph{functionality}, \emph{portability} and \emph{maintainability} contain more labeled topics, since these NFRs have more intensely shaded boxes.
-We note that the gap in the MaxDB data is due to an interruption in the source data, having to do with the infrastructure for releasing data, rather than specific occurrences in the project.
+We note that the gap in the MaxDB data is due to an interruption in the source data, having to do with the infrastructure for releasing developer artifacts, rather than specific occurrences in the project.
 \todo[inline]{can I pin this down more ?}
+There are more \textsf{None} labels in MaxDB because a number of topics were to do with code cleanup or automated checking using tools like \texttt{sutcheck}, a tool-specific to the development process at MaxDB.
 
-\todo[inline]{try to pin down these reasons below}%  MySQL 3.23 was put on maintenance.
+\textbf{Timeline of key events} -- Figs.
+\ref{} show key events for each of our projects.
+Several external events (external to `normal' development activity, that is) affected our projects.
+For example, both projects show a high number of NFRs recognized at the first period of analysis.
+This is due to our window choice: we deliberately targeted our analysis to
+MySQL 3.23 First announced Major update \& End-of life
+MaxDB 7.500 First announced
+
+Note that MySQL's release process was haphazard at this point: they would release a patch version whenever the need arose.
+By version 5.1, the company moved to a milestone release, with new releases targeted for every six months.
+
+\todo[inline]{Does our visualization capture the entire timeline?}
+
+\todo[inline]{try to pin down these reasons below}
+%  MySQL 3.23 was put on maintenance.
 % 

### 2010-08-19 icse/icse2011.tex
Unsupervised labelling edit part I

--- 2010-08-19 icse/icse2011.tex
@@
 Using our three word lists, we labeled our topics where there was a match between a word in the list and the same word in the topic.
-Table \ref{tbl:wordlist} lists broad labeling results for MaxDB as an example.
 A \emph{named topic} is a topic with a matching label. 
-There are \{periods X 20\} topics per project as we told LDA to extract 20 topics per period.
+There are roughly 20 times the number of topics as there are periods per project. 
+This is because we told LDA to extract 20 topics per period. 
 All experiments were run on MaxDB 7.500 and MySQL 3.23 data.
+Table \ref{tbl:wordlist} shows how many topics were labelled for MaxDB
+and MySQL.
 
@@
 
-For \textsf{exp1}, our best performing labels (the labels matched with the most topics) were \emph{correctness} (182 topics) and \emph{testability} (121).
+For \textsf{exp1}, our best performing labels, those with the most topics, were \emph{correctness} (182 topics) and \emph{testability} (121). 
 We did not get good results for \emph{usability} or \emph{accuracy}, which were associated with fewer than ten topics. 
-We also looked for correlation between our labels: Excluding double matches (self-correlation), our highest co-occurring terms were \emph{verifiability} and \emph{traceability}, and \emph{testability} and \emph{correctness} (76 and 62 matches, respectively).
+We also looked for correlation between our labels: Excluding double matches (self-correlation), our highest co-occurring terms were \emph{verifiability} with \emph{traceability}, and \emph{testability} with \emph{correctness} (76 and 62 matches, respectively).
 % See the following sections for our error analysis.
@@
 Only \emph{reliability} produces a lot of matches, mostly with the word `error'. 
-Co-occurrence results were poor.
-This suggests our word lists were overly restrictive.
+Co-occurrence results were poor. This suggests our word lists were overly restrictive.
 
@@
 We found a high of 265 topics for \emph{usability}, with a low of 44 topics for \emph{maintainability}. 
-Common co-occurrences were \emph{reliability} and \emph{usability}, \emph{efficiency} and \emph{reliability}, and \emph{efficiency} and \emph{usability} (200, 190, and 150 topics in common, respectively).
+Common co-occurrences were \emph{reliability} with \emph{usability}, \emph{efficiency} with \emph{reliability}, and \emph{efficiency} with \emph{usability} (200, 190, and 150 topics in common, respectively).
 

### 2010-08-19 icse/icse2011.tex
unsupervised edits II

@@
 \subsubsection{Analysis of the unsupervised labelling}
+
 Based on the labels, and our manual topic labelling, we compared the results of the unsupervised word matching approach. 
-For each quality we assessed whether the manual tag matched the unsupervised label assigned.
+For each quality we assessed whether unsupervised labels matches the manual annotations.
 Figure \ref{fig:maxdb-unsup-results} shows our results for MaxDB and MySQL. 
-Our error is measured using the \emph{ROC} (Receiver Operating Characteristic) area-under-curve value~\cite{Fawcett2006861}.
+Our performance is measured using the \emph{ROC} (Receiver Operating Characteristic) area-under-curve value~\cite{Fawcett2006861}. 
 ROC values provide a score, similar to school letter-grades (A is 0.9, C is 0.6), reflecting how well a particular learner performed for the given data. 
@@
 
-\todo[inline]{since ROC is new-sounding, I wonder if a more lengthy explanation would be merited. An obvious question is why not precision/recall?}
+\todo[inline]{since ROC is new-sounding, I wonder if a more lengthy explanation would be merited. An obvious question is why not precision/recall? -- ROC is hot right now and it combines both}
 \begin{figure*}

### 2010-08-20 icse/icse2011.tex
effectiveness

@@
 %{Discuss how these are CHEAP techniques and although somewhat %inaccurate they don't require a tonne of work}
-The unsupervised labelling had difficulty distinguishing between common labels and infrequent labels. The related words for \emph{correctness}, for example, tended to be too lengthy and non-specific. 
-\todo[inline,color=red]{Abram can you explain what you mean here?}
+The unsupervised labelling had difficulty distinguishing between common labels and infrequent labels, often the learners would mislabel a topic deserving of an infrequent label with a more common label.
+The word-lists for \emph{correctness}, tended to be too lengthy, non-specific and broad, especially if WordNet words were used as it is a loosely defined concept in common parlance.
 
-With ROC values ranging from $0.6$ to $0.8$ we can see there is promise in these attempts.
+With ROC values ranging from $0.6$ to $0.8$ we can see there is promise in these methods.
 \textsf{exp2} and \textsf{exp3} both indicate that static information can be used to help label topics without any training whatsoever. 
-If the techniques used in \textsf{exp2} and \textsf{exp3} were combined with the supervised techniques we could probably reduce the training effort.
+If the techniques used in \textsf{exp2} and \textsf{exp3} were combined with the supervised techniques we could probably reduce the training effort by boosting training sets with topics classified with the unsupervised techniques.
 Both Naive Bayesian learners and the word-list approaches were computationally efficient.  
-These results are promising, because the results are accurate enough to be useful, while still cheap enough to execute to be feasible as an automated or semi-automated method of labelling topics by their software qualities.
+%These results are promising, because the accuracy of the techniques is good enough to be useful, while the run-time performance is still cheap enough to execute to be feasible as an automated or semi-automated method of labelling topics by their software qualities.
+These results are promising because they indicate that these techniques are accurate enough to be useful while still maintaining acceptable run-time performance.
 
@@
 We want to investigate developer attitudes related to these labels: i.e., when we label a topic, was the developer expressing positive or negative qualities about that label?  
-It is difficult to map abstract qualities to particular messages.
+It is difficult to map abstract qualities to particular messages. % what?
  % This is disjointed and needs more context

### 2010-08-20 icse/icse2011.tex
threats to validity

@@
 \subsection{Threats to validity}
+
 \emph{Construct validity} issues include that we used only commit messages rather than mail or bug tracker messages. 
-We also chose our taxonomy and the data to study.
-\emph{Internal validity} issues are to do with inter-rater reliability.
-\emph{External validity} issues are that our data originated from OSS database projects and thus might not be applicable to commercially developed software.
+We also chose our taxonomy and the data to study. Also, we did not show these results to the stakeholders.
+
+\emph{Internal validity} issues are include inter-rater reliability and our lack of rival explanation building. 
+We improved internal validity by trying to correlate and explain the behaviours observed in the analysis with the historical records of the projects.
+
+\emph{External validity} was addressed by our use of multiple case studies, but there are still issues. 
+Our data originated from OSS database projects and thus might not be applicable to commercially developed software. 
 Furthermore, our analysis techniques rely on a project's use of meaningful commit messages.
-\todo[inline]{update this with more detail}
+
+% in Case Study Research: Design Case Studies
+\emph{Reliability} was addressed by each annotator following the same protocol and using the same annotations. 
+Unfortunately only two annotators were used, their annotating could be biased and we did not analyze for inter-rater reliability.
+
+%\todo[inline]{update this with more detail}
 

### 2010-08-20 icse/icse2011.tex
ROC rebuttal

@@
 
-\todo[inline]{since ROC is new-sounding, I wonder if a more lengthy explanation would be merited. An obvious question is why not precision/recall? -- ROC is hot right now and it combines both}
+%\todo[inline]{since ROC is new-sounding, I wonder if a more lengthy explanation would be merited. An obvious question is why not precision/recall? -- ROC is hot right now and it combines both}
+% Rebuttal: ROC is not new and it is what the machine learning community, search based software engineering and now the MSR community are using +  it has natural meaningful quality mappings.
 \begin{figure*}
@@
 
+% can always be cut down
+
 \emph{Construct validity} issues include that we used only commit messages rather than mail or bug tracker messages. 
+Conversely construct validity was addressed in our validation as we used the files, topics and revisions to annotate the commit messages.
 We also chose our taxonomy and the data to study. Also, we did not show these results to the stakeholders.
@@
 We improved internal validity by trying to correlate and explain the behaviours observed in the analysis with the historical records of the projects.
+We did not attempt to match our results to any particular model or theory.
 

### 2010-08-20 icse/icse2011.tex
comparing edits

@@
 
-MySQL and MaxDB's machine learners did make decisions based off some shared words.
-Words that were used to classify topics that were shared between MySQL and MaxDB included: \textsf{bug, code, compiler, database, HP UX, delete, memory, missing, problems, removed, add, added, changed, problem, and test}.
+MySQL and MaxDB's machine learners made some decisions based off some shared words: \textsf{bug, code, compiler, database, HP UX, delete, memory, missing, problems, removed, add, added, changed, problem, and test}.
 Adding these words to the word bags of \textsf{exp2} and \textsf{exp3} could improve performance while ensuring they were only domain specific.

### 2010-08-20 icse/icse2011.tex
explain the gap

@@
 The second, absolute frequency, compares the number of topics labelled with that NFR per period 
-relative to the maximum number of labelled topics overall.
+relative to the maximum number of labelled topics overall, for that project.
 
 \textbf{Understanding the figures} -- For example, in Fig.
-\ref{fig:mysql-timeline}, we see that the NFRs \emph{functionality}, \emph{portability} and \emph{maintainability} contain more labeled topics, since these NFRs have more intensely shaded boxes.
-We note that the gap in the MaxDB data is due to an interruption in the source data, having to do with the infrastructure for releasing developer artifacts, rather than specific occurrences in the project.
-\todo[inline]{can I pin this down more ?}
+\ref{fig:mysql-timeline}, we see that the NFRs \emph{functionality}, \emph{portability} and \emph{maintainability} contain more labeled topics, since these NFRs have been more intensely shaded.
+% boxes.
+not sure boxes is important..
+%We note that the gap in the MaxDB data is due to an interruption in the source data, having to do with the infrastructure for releasing developer artifacts, rather than specific occurrences in the project.
+% I grepped query1.mrs2 http://softwareprocess.es/y/y/maxdb-76.tar.gz
+% and                   http://softwareprocess.es/y/maxdb-7500.tar.gz
+The gap in the MaxDB is due to the fact that other versions of MaxDB exist on different repositories and during this period from February 2005 to Jun 2005 the focus of development shifted to MaxDB 7.6.
+Immediately after this period of inactivity, there was a large deluge of commits added to the MaxDB 7.500 repositories.
+%\todo[inline]{can I pin this down more ?}
 There are more \textsf{None} labels in MaxDB because a number of topics were to do with code cleanup or automated checking using tools like \texttt{sutcheck}, a tool-specific to the development process at MaxDB.

### 2010-08-20 icse/icse2011.tex
editing viz section

--- 2010-08-20 icse/icse2011.tex
@@
 
-%\todo[inline]{since ROC is new-sounding, I wonder if a more lengthy explanation would be merited. An obvious question is why not precision/recall? -- ROC is hot right now and it combines both}
-% Rebuttal: ROC is not new and it is what the machine learning community, search based software engineering and now the MSR community are using +  it has natural meaningful quality mappings.
+\todo[inline]{since ROC is new-sounding, I wonder if a more lengthy explanation would be merited. An obvious question is why not precision/recall? -- ROC is hot right now and it combines both}
 \begin{figure*}
@@
 \caption[]{MySQL and MaxDB macro (grey) and micro (black) ROC results per multi-label learner. Possible values range from 0--1.
-\todo[inline]{better discussion}% and graphical legend would be nice}
+%\todo[inline]{better discussion}% and graphical legend would be nice}
 }
@@
 We found many topics that were not non-functional requirements (NFRs) but were often related to them.
-For instance, concurrency was mentioned often in the commit logs and was related to \emph{correctness} and \emph{reliability}, because concurrency was troublesome.
+For instance, concurrency was mentioned often in the commit logs and was related to \emph{correctness} and \emph{reliability}, because dealing with concurrency was troublesome.
 Configuration management and source control related changes appeared often and sometimes there were topics dedicated to configuration management.
@@
 
-\begin{comment}
-A quick mind dump before I forget:
-* portability was very popular
-* there were many instances of concurrency
-* There were often topics of crap commits with no comment
-* might need more than the top words to compare
-* Things like tests were noticable more by file type
-* Configuration management is not part of the software quality ontology?
-* I didn't tag none I tagged unknown, so that's some inconsistency between us
-* There were a lot of copyright changes in mysql
-* mysql had optimizer module which did not mean optimization
-* I tried to assign reliability/bug for reliability bugs
-  and Functionality/Correctness for fixes
-* Some security changes came later
-* Mysql 3.23 transitioned from a project being developed to
-  a project being maintained, more frequent back-port bug-fixes
-\end{comment}
-
 \subsection{Summary of techniques}
-While unsupervised techniques (LSI and LDA are both unsupervised) are appealing in their lack of human intervention, and thus lower effort, supervised learners have the advantage of domain knowledge which typically means improved results.
+While unsupervised techniques such as LDA is appealing in its lack of human intervention, and thus lower effort, supervised learners have the advantage of domain knowledge which typically means improved results.
 Our manual annotations were fairly quick to do, taking on average five minutes per 30 day period.
@@
 % \todo[inline]{reference project name Nomen more}
-
-Using the classification results from the previous sections, we evaluated two research questions:
+Since our ROC values produced acceptable results, we now turn to \emph{applying} those techniques to understand software maintenance activities. We evaluated two research questions:
 \begin{enumerate}
@@
 \end{enumerate}
+
+There are several possible applications of \textsf{Nomen}'s techniques. 
+For outside users of the software, \textsf{Nomen} can provide insight into that company's development practices. 
+Are they focused on usability, for example?
+Within the organization \textsf{Nomen} acts as a historical sanity check. 
+Confirmation bias is the difference between what people think happened and what actually happened. 
+Some software releases, for instance, are poorly received since they are quite buggy. 
+\textsf{Nomen} would help explain why this was the case, since it can identify NFRs which are dominating developer attention.
 
@@
 
-% 
-% \begin{figure*}
-%   \centering
-%   % hi-res version
-%   %\includegraphics[height=.55\textheight]{maxdb-time-smear-plot-Equal}
-%   \includegraphics[width=\textwidth]{figures/maxdb-timeline-all}
-%   % \includegraphics[height=.57\textheight]{small-maxdb-time-smear-plot-Equal}
-%  % \caption{MaxDB 7.500: Labelled topics related to non-functional software qualities plotted over time, continuous topics are trends that occur across time}
-%   \caption{MaxDB 7.500: NFRs over time. Numbers represent topics labelled with that quality. Box intensity (saturation) is mapped to label frequency relative to the largest label value.}
-%  \label{fig:maxdb}
-% \end{figure*}
-% 
-% \begin{figure*}%[h]
-%   \centering
-%  %\includegraphics[width=.97\textwidth]{mysql-time-smear-plot-Equal}
-%   %\includegraphics[width=.97\textwidth]{small-mysql-time-smear-plot-Equal}
-%   \includegraphics[width=\textwidth]{figures/mysql-timeline-all}
-%   \caption{MySQL 3.23: NFRs over time. Topic frequency relative to that quality is mapped to alpha (transparency) value.}
-%   \label{fig:mysql}
-% \end{figure*}
-
 \subsection{Comparing MaxDB and MySQL}
@@
 %{Discuss how these are CHEAP techniques and although somewhat %inaccurate they don't require a tonne of work}
-The unsupervised labelling had difficulty distinguishing between common labels and infrequent labels, often the learners would mislabel a topic deserving of an infrequent label with a more common label.
-The word-lists for \emph{correctness}, tended to be too lengthy, non-specific and broad, especially if WordNet words were used as it is a loosely defined concept in common parlance.
+The unsupervised labelling had difficulty distinguishing between common labels and infrequent labels. The related words for \emph{correctness}, for example, tended to be too lengthy and non-specific. 
+\todo[inline,color=red]{Abram can you explain what you mean here?}
 
-With ROC values ranging from $0.6$ to $0.8$ we can see there is promise in these methods.
+With ROC values ranging from $0.6$ to $0.8$ we can see there is promise in these attempts.
 \textsf{exp2} and \textsf{exp3} both indicate that static information can be used to help label topics without any training whatsoever.
-If the techniques used in \textsf{exp2} and \textsf{exp3} were combined with the supervised techniques we could probably reduce the training effort by boosting training sets with topics classified with the unsupervised techniques.
+If the techniques used in \textsf{exp2} and \textsf{exp3} were combined with the supervised techniques we could probably reduce the training effort.
 Both Naive Bayesian learners and the word-list approaches were computationally efficient.
-%These results are promising, because the accuracy of the techniques is good enough to be useful, while the run-time performance is still cheap enough to execute to be feasible as an automated or semi-automated method of labelling topics by their software qualities.
-These results are promising because they indicate that these techniques are accurate enough to be useful while still maintaining acceptable run-time performance.
+These results are promising, because the results are accurate enough to be useful, while still cheap enough to execute to be feasible as an automated or semi-automated method of labelling topics by their software qualities.
 
 \subsection{Threats to validity}
-
-% can always be cut down
-
 \emph{Construct validity} issues include that we used only commit messages rather than mail or bug tracker messages.
-Conversely construct validity was addressed in our validation as we used the files, topics and revisions to annotate the commit messages.
-We also chose our taxonomy and the data to study. Also, we did not show these results to the stakeholders.
-
-\emph{Internal validity} issues are include inter-rater reliability and our lack of rival explanation building. 
-We improved internal validity by trying to correlate and explain the behaviours observed in the analysis with the historical records of the projects.
-We did not attempt to match our results to any particular model or theory.
-
-\emph{External validity} was addressed by our use of multiple case studies, but there are still issues. 
-Our data originated from OSS database projects and thus might not be applicable to commercially developed software. 
+We also chose our taxonomy and the data to study.
+\emph{Internal validity} issues are to do with inter-rater reliability.
+\emph{External validity} issues are that our data originated from OSS database projects and thus might not be applicable to commercially developed software.
 Furthermore, our analysis techniques rely on a project's use of meaningful commit messages.
-
-% in Case Study Research: Design Case Studies
-\emph{Reliability} was addressed by each annotator following the same protocol and using the same annotations. 
-Unfortunately only two annotators were used, their annotating could be biased and we did not analyze for inter-rater reliability.
-
-%\todo[inline]{update this with more detail}
+\todo[inline]{update this with more detail}
 
@@
 We want to investigate developer attitudes related to these labels: i.e., when we label a topic, was the developer expressing positive or negative qualities about that label?
-It is difficult to map abstract qualities to particular messages. % what?
+It is difficult to map abstract qualities to particular messages.
  % This is disjointed and needs more context
@@
 \end{document}
-
+% \begin{comment}
+% A quick mind dump before I forget:
+% * portability was very popular
+% * there were many instances of concurrency
+% * There were often topics of crap commits with no comment
+% * might need more than the top words to compare
+% * Things like tests were noticable more by file type
+% * Configuration management is not part of the software quality ontology?
+% * I didn't tag none I tagged unknown, so that's some inconsistency between us
+% * There were a lot of copyright changes in mysql
+% * mysql had optimizer module which did not mean optimization
+% * I tried to assign reliability/bug for reliability bugs
+%   and Functionality/Correctness for fixes
+% * Some security changes came later
+% * Mysql 3.23 transitioned from a project being developed to
+%   a project being maintained, more frequent back-port bug-fixes
+% \end{comment}
 % % a rough sketch 

### 2010-08-20 icse/icse2011.tex

--- 2010-08-20 icse/icse2011.tex
@@
 \end{figure*}
-
-% 
-% \begin{figure*}
-%   \centering
-%   % hi-res version
-%   %\includegraphics[height=.55\textheight]{maxdb-time-smear-plot-Equal}
-%   \includegraphics[width=\textwidth]{figures/maxdb-timeline-all}
-%   % \includegraphics[height=.57\textheight]{small-maxdb-time-smear-plot-Equal}
-%  % \caption{MaxDB 7.500: Labelled topics related to non-functional software qualities plotted over time, continuous topics are trends that occur across time}
-%   \caption{MaxDB 7.500: NFRs over time. Numbers represent topics labelled with that quality. Box intensity (saturation) is mapped to label frequency relative to the largest label value.}
-%  \label{fig:maxdb}
-% \end{figure*}
-% 
-% \begin{figure*}%[h]
-%   \centering
-%  %\includegraphics[width=.97\textwidth]{mysql-time-smear-plot-Equal}
-%   %\includegraphics[width=.97\textwidth]{small-mysql-time-smear-plot-Equal}
-%   \includegraphics[width=\textwidth]{figures/mysql-timeline-all}
-%   \caption{MySQL 3.23: NFRs over time. Topic frequency relative to that quality is mapped to alpha (transparency) value.}
-%   \label{fig:mysql}
-% \end{figure*}
 

### 2010-08-20 icse/icse2011.tex
edit summary of techniques

--- 2010-08-20 icse/icse2011.tex
@@
 \subsection{Summary of techniques}
-While unsupervised techniques (LSI and LDA are both unsupervised) are appealing in their lack of human intervention, and thus lower effort, supervised learners have the advantage of domain knowledge which typically means improved results.
+While unsupervised techniques (LSI and LDA are both unsupervised) are appealing in their lack of human intervention, and thus lower effort, supervised learners have the advantage of domain knowledge (training sets) which typically means improved results.
 Our manual annotations were fairly quick to do, taking on average five minutes per 30 day period.
@@
 
-We found that the multi-label learners of BR, CLR and HOMER did not do as well for Macro-ROC as NaiveBayes and other NaiveBayes derived learners. 
-This suggests that by sticking together multiple NaiveBayes learners we could probably label sets of topics effectively, but it would require a separate NaiveBayes learner per label.
+We found that the multi-label learners of BR, CLR and HOMER did not do as well for Macro-ROC as NaiveBayes and other naive Bayes derived learners. 
+This suggests that by combining together multiple NaiveBayes learners we could probably label sets of topics effectively, but it would require a separate NaiveBayes learner per label.
 
@@
 
-\textbf{Understanding the figures} -- For example, in Fig.
-\ref{fig:mysql-timeline}, we see that the NFRs \emph{functionality}, \emph{portability} and \emph{maintainability} contain more labeled topics, since these NFRs have been more intensely shaded.
+\textbf{Understanding the figures} -- For example, in Figure \ref{fig:mysql-timeline}, we see that the NFRs \emph{functionality}, \emph{portability} and \emph{maintainability} contain more labeled topics, since these NFRs have been more intensely shaded.
 % boxes.
@@
 
-\textbf{Timeline of key events} -- Figs.
-\ref{} show key events for each of our projects.
+%XXX 
+\textbf{Timeline of key events} -- Figs. \ref{} show key events for each of our projects.  % where's the figure?
 Several external events (external to `normal' development activity, that is) affected our projects. 
@@
 This is due to our window choice: we deliberately targeted our analysis to 
-MySQL 3.23 First announced Major update \& End-of life
-MaxDB 7.500 First announced
+when MySQL 3.23 was first announced %Major update \& End-of life
+and when MaxDB 7.500 was first announced.
 
@@
 
-\todo[inline]{try to pin down these reasons below}
+%\todo[inline]{try to pin down these reasons below}
 %  MySQL 3.23 was put on maintenance.

### 2010-08-20 icse/icse2011.tex
annotation observations edits

@@
 We found many topics that were not non-functional requirements (NFRs) but were often related to them. 
-For instance, concurrency was mentioned often in the commit logs and was related to \emph{correctness} and \emph{reliability}, because concurrency was troublesome.
-Configuration management and source control related changes appeared often and sometimes there were topics dedicated to configuration management.
+For instance, concurrency was mentioned often in the commit logs and was related to \emph{correctness} and \emph{reliability}, likely because concurrent code is prone to bugs such as race conditions.%are
+% often  was troublesome. 
+\emph{Configuration management} and \emph{source control} related changes appeared often and sometimes there were topics dedicated to configuration management. 
 These kinds of changes are slightly related to \emph{maintainability}. 
@@
 For instance, optimizers are very common modules in database systems: both MySQL and MaxDB have optimizer modules. 
-In MySQL the optimizer is mentioned but often the change deals with correctness or another quality.
+In MySQL the optimizer is mentioned but often the change addresses  correctness or another quality. 
 Despite this difference, the name of the module could fool our learners into believing the change was always about \emph{efficiency}. 
-In these cases the advantages of tailoring topic names to specific project terminologies are more clear: this would avoid automated mistakes due to the names of entities and modules of a software project.
+In these cases the advantages of tailoring topic names to specific project terminologies are more clear. 
+Project specific word-lists would avoid automated mistakes due to the names of entities and modules of a software project.
 

### 2010-08-20 icse/icse2011.tex
applying multiple labels to topics edits

@@
 
-As noted in Section \ref{sec:wordlist}, each topic in our data-set can be composed of zero to all NFRs.
+As noted in Section \ref{sec:wordlist}, each topic in our data-set can be composed of zero, one, or more NFRs. 
 For example, a commit message might address \textit{reliability} in the context of \textit{efficiency}, or make a \textit{maintainability} improvement in the source code that related to \textit{usability}.
-However, traditional machine learning techniques, such as Bayesian networks, can only map topics to a single class.
-The Mulan~\cite{mulan} library encapsulates several different machine learners which can handle this multi-label learning for our data-set.
+However, traditional machine learning techniques, such as naive Bayes, can only map topics to a single class. 
+The Mulan~\cite{mulan} library encapsulates several different multi-label machine learners which can label elements with multiple labels.
 Mulan also includes methods for determining the performance of such techniques.
@@
 
-In multi-label learning the technique can perform micro or macro measurements.
-Macro measurements are aggregated at a class or label level.
-Micro measurements are aggregated at a decision level.
-\todo[inline]{this is unclear, could do with a few more sentence explaining why we should care.}
- So a macro-ROC measurement is the average ROC over the ROC values for all labels, where a micro-ROC is the average ROC over all examples that were classified.
-Unfortunately for MaxDB, the macro-ROC values are undefined because of poor performance of one of the labels.
+Two perspectives to evaluate multi-label learners are with micro or macro measurements (used in Figure \ref{fig:mulan}).
+Macro measurements are aggregated at a class or label level (per class) while micro measurements are at the element level (per element).
+A macro-ROC measurement is the average ROC over the ROC values for all labels, where a micro-ROC is the average ROC over all examples that were classified. 
+Unfortunately for MaxDB, the macro-ROC values are undefined because of poor performance of one of the labels, see Figure \ref{fig:mulan}.
 
-Figure~\ref{fig:mulan} presents the results of Mulan's best multi-label
-learners for our data. Calibrated Label Ranking (CLR) is a learner
-that builds two layers, the first layer determines if an entity should
-be labelled, while the second layer determines what labels should be assigned.
+Figure~\ref{fig:mulan} presents the results of Mulan's best multi-label learners for our data. 
+Calibrated Label Ranking (CLR) is a learner that builds two layers, the first layer determines if an entity should be labelled, while the second layer determines what labels should be assigned.
 The Hierarchy Of Multi-label classifiERs (HOMER) and Binary Relevance (BR) act as a hierarchy of learners: BR is flat, while HOMER tries to build a deeper hierarchy to build a more accurate learner~\cite{mulan}.
@@
 \todo[inline]{discuss this similarity in more detail}
-%\todo[inline]{what of CLR? Is it hierarchical?}
 

### 2010-08-20 icse/icse2011.tex
write up for analysis of timelines
Signed-off-by: Neil Ernst <neil@neilernst.net>

--- 2010-08-20 icse/icse2011.tex
@@
 
-\begin{comment}
-A quick mind dump before I forget:
-* portability was very popular
-* there were many instances of concurrency
-* There were often topics of crap commits with no comment
-* might need more than the top words to compare
-* Things like tests were noticable more by file type
-* Configuration management is not part of the software quality ontology?
-* I didn't tag none I tagged unknown, so that's some inconsistency between us
-* There were a lot of copyright changes in mysql
-* mysql had optimizer module which did not mean optimization
-* I tried to assign reliability/bug for reliability bugs
-  and Functionality/Correctness for fixes
-* Some security changes came later
-* Mysql 3.23 transitioned from a project being developed to
-  a project being maintained, more frequent back-port bug-fixes
-\end{comment}
-
 \subsection{Summary of techniques}
-While unsupervised techniques (LSI and LDA are both unsupervised) are appealing in their lack of human intervention, and thus lower effort, supervised learners have the advantage of domain knowledge (training sets) which typically means improved results.
+While unsupervised techniques such as LDA is appealing in its lack of human intervention, and thus lower effort, supervised learners have the advantage of domain knowledge which typically means improved results.
 Our manual annotations were fairly quick to do, taking on average five minutes per 30 day period.
@@
 
-We found that the multi-label learners of BR, CLR and HOMER did not do as well for Macro-ROC as NaiveBayes and other naive Bayes derived learners. 
-This suggests that by combining together multiple NaiveBayes learners we could probably label sets of topics effectively, but it would require a separate NaiveBayes learner per label.
+We found that the multi-label learners of BR, CLR and HOMER did not do as well for Macro-ROC as NaiveBayes and other NaiveBayes derived learners. 
+This suggests that by sticking together multiple NaiveBayes learners we could probably label sets of topics effectively, but it would require a separate NaiveBayes learner per label.
 
@@
 
-\subsection{Visualizations}
+There are several possible applications of \textsf{Nomen}'s techniques. 
+For outside users of the software, \textsf{Nomen} can provide insight into that company's development practices. 
+Are they focused on usability, for example?
+Within the organization \textsf{Nomen} acts as a historical sanity check. 
+Confirmation bias is the difference between what people think happened and what actually happened. 
+Some software releases, for instance, are poorly received since they are quite buggy. 
+\textsf{Nomen} would help explain why this was the case, since it can identify NFRs which are dominating developer attention.
+
+\subsection{Timeline visualization}
 Figures \ref{fig:maxdb-timeline} and \ref{fig:mysql-timeline} show the temporal patterns of label frequencies. 
@@
 relative to the maximum number of labelled topics overall, for that project.
+The topmost row in the diagrams is reserved for historical events of note.
+We refer to these occurrences in the discussion which follows.
 
-\textbf{Understanding the figures} -- For example, in Figure \ref{fig:mysql-timeline}, we see that the NFRs \emph{functionality}, \emph{portability} and \emph{maintainability} contain more labeled topics, since these NFRs have been more intensely shaded.
+\textbf{Understanding the figures} -- In Fig.
+\ref{fig:mysql-timeline}, we see that the NFRs \emph{functionality}, \emph{portability} and \emph{maintainability} contain more labeled topics, since these NFRs have been more intensely shaded.
 % boxes.
@@
 % and                   http://softwareprocess.es/y/maxdb-7500.tar.gz
-The gap in the MaxDB is due to the fact that other versions of MaxDB exist on different repositories and during this period from February 2005 to Jun 2005 the focus of development shifted to MaxDB 7.6.
-Immediately after this period of inactivity, there was a large deluge of commits added to the MaxDB 7.500 repositories.
-%\todo[inline]{can I pin this down more ?}
 There are more \textsf{None} labels in MaxDB because a number of topics were to do with code cleanup or automated checking using tools like \texttt{sutcheck}, a tool-specific to the development process at MaxDB.
 
-%XXX 
-\textbf{Timeline of key events} -- Figs. \ref{} show key events for each of our projects.  % where's the figure?
-Several external events (external to `normal' development activity, that is) affected our projects. 
+\textbf{Timeline of key events} -- As discussed earlier, the top row in each figure shows key events for each project.
+\textsf{Nomen} can pick out the underlying NFR activity behind these events.
 For example, both projects show a high number of NFRs recognized at the first period of analysis.
-This is due to our window choice: we deliberately targeted our analysis to 
-when MySQL 3.23 was first announced %Major update \& End-of life
-and when MaxDB 7.500 was first announced.
+This is due to our window choice: we deliberately targeted our analysis to coincide with the development lifecycle of a particular major release of the software.
+For MaxDB, version 7.5.00  was released in December of 2003.
+From analyzing the mailing list for external validation, we know that release 7.5.00.23 saw the development of PHP interfaces, possibly accounting for the simultaneous increase in the \emph{portability} NFR at the same time.
+The gap in MaxDB (Fig. \ref{fig:maxdb-timeline}) is due to a shift in development focus (from February 2005 to Jun 2005) to MaxDB 7.6, which is released in June 2005.
 
-Note that MySQL's release process was haphazard at this point: they would release a patch version whenever the need arose. 
+For MySQL (Fig.
+\ref{fig:mysql-timeline}), we similarly validated our NFR patterns with external mailing list data.
+This release of MySQL was the first to be licenced under the GPL.
+Version 3.23.31 (January, 2001) was the production release (non-beta), and we see a flurry of topics labeled with \emph{functionality} and \emph{maintainability}, as one might expect.
+After this point, we can consider this version to enter the maintenance phase of the lifecycle.
+
+%Note that MySQL's release process was haphazard at this point: they would release a patch version whenever the need arose.
 By version 5.1, the company moved to a milestone release, with new releases targeted for every six months.
 
-\todo[inline]{Does our visualization capture the entire timeline?}
+% \todo[inline]{Does our visualization capture the entire timeline?}
 
 %\todo[inline]{try to pin down these reasons below}
-%  MySQL 3.23 was put on maintenance.
-% 
-% MySQL forks repositories per every version and 3.23 was supported for a
-% long time. there was 4.0 4.1 5.0 5.1
-% 
-% This is the MySQL 3.23 branch
-% 
-% Same thing with MaxDB 7.5
-% 7.6 is in a different repository
-
-% For MaxDB (Figure \ref{fig:maxdb}) the largest trends were related to maintainability and portability.
-MaxDB supports numerous platforms and portability was a constant issue facing the project's development.
-MySQL was different, with many topics overlapping, so there were many different subsets.
-Functionality trends were prevalent throughout its history (the largest reddish streak across the top).
-Two combination  tags of ``functionality portability'' (orange) and ``maintainability portability'' (teal) streaked across most of the history of MySQL.
-Since this was MySQL 3.23, a stable branch, we expect that issues dealing with portability and maintenance would be the primary concerns of this branch: developers probably wanted it to work with current systems and thus had to update it.
-
-%\todo[inline]{Update this}
 
@@
 \includegraphics[width=\textwidth]{figures/mysql-timeline}
-\label{fig:maxdb-timeline}                               
+\label{fig:mysql-timeline}                               
 }
@@
 \includegraphics[width=\textwidth]{figures/maxdb-timeline}
-\label{fig:mysql-timeline}
+\label{fig:maxdb-timeline}
 }

### 2010-08-20 icse/icse2011.tex
analysis of supervised labelling edits

--- 2010-08-20 icse/icse2011.tex
@@
 Figure \ref{fig:best-learn-per-tag} shows the performance of the best performing learner per label. 
-We considered the best learner for a label to be the one which had the highest ROC value for that label.
-Figure \ref{fig:best-learn-per-tag} uses the ZeroR learner as a baseline, a common practice in machine learning.
-ZeroR naively chooses the largest category all of the time.
-It occasionally outperforms non-naive learners; for labels which are not as common, this is to be expected because any miscategorization will hurt accuracy.
-This is why we use ROC values, as they can better represent performance on labels which are not applicable to the majority of samples.
+ We considered the best learner for a label to be the learner which had the highest ROC value for that label. 
+Figure \ref{fig:best-learn-per-tag} uses the ZeroR learner as a baseline, a common practice in machine learning,
+as ZeroR naively chooses the largest category all of the time. 
+ZeroR occasionally outperforms other learners; for labels which are not as common, this is to be expected because any miscategorization will hurt accuracy. 
+This is why we use ROC values, instead of accuracy, as they can better represent performance on labels which are not applicable to the majority of samples.
 
@@
 
-The less-frequently occurring a label the harder it is to get accurate
+The less-frequently occurring a label, the harder it is to get accurate
 results, due to the high noise level. Nevertheless, these results are

### 2010-08-20 icse/icse2011.tex
supervised labelling edits

@@
 
-We used a suite of supervised classifiers from WEKA~\cite{weka09}.
-WEKA is a suite of machine learning tools such as support vector machines and Bayes nets.
+We used a suite of supervised classifiers, WEKA~\cite{weka09}. 
+WEKA includes machine learning tools such as support vector machines and Bayes-nets. 
 We also used the multi-labelling add-on for WEKA, Mulan~\cite{mulan}\footnote{\url{http://mlkd.csd.auth.gr/multilabel.html}}. 
@@
 To assess the performance of the supervised learners, we did a 10-fold cross-validation~\cite{Kohavi1995}, a common technique for evaluating machine learners. 
-The original data is partitioned randomly into ten subsamples.
-One subsample is used as a training set and evaluated on the other nine subsamples.
-This is repeated nine more times, with each subsample used once as the training set.
+The original data is partitioned randomly into ten sub-samples. Each sample is used to test against a train set composed of the nine other samples.
+%One subsample is used as a training set and evaluated on the other nine sub-samples. 
+%This is repeated nine more times, with each subsample used once as the training set. 
 We have reported these results below in Section \ref{sec:suplabelling}.

### 2010-08-20 icse/icse2011.tex

--- 2010-08-20 icse/icse2011.tex
@@
 We found many topics that were not non-functional requirements (NFRs) but were often related to them. 
-For instance, concurrency was mentioned often in the commit logs and was related to \emph{correctness} and \emph{reliability}, because dealing with concurrency was troublesome.
-Configuration management and source control related changes appeared often and sometimes there were topics dedicated to configuration management.
+For instance, concurrency was mentioned often in the commit logs and was related to \emph{correctness} and \emph{reliability}, likely because concurrent code is prone to bugs such as race conditions.%are
+% often  was troublesome. 
+\emph{Configuration management} and \emph{source control} related changes appeared often and sometimes there were topics dedicated to configuration management. 
 These kinds of changes are slightly related to \emph{maintainability}. 
@@
 For instance, optimizers are very common modules in database systems: both MySQL and MaxDB have optimizer modules. 
-In MySQL the optimizer is mentioned but often the change deals with correctness or another quality.
+In MySQL the optimizer is mentioned but often the change addresses  correctness or another quality. 
 Despite this difference, the name of the module could fool our learners into believing the change was always about \emph{efficiency}. 
-In these cases the advantages of tailoring topic names to specific project terminologies are more clear: this would avoid automated mistakes due to the names of entities and modules of a software project.
+In these cases the advantages of tailoring topic names to specific project terminologies are more clear. 
+Project specific word-lists would avoid automated mistakes due to the names of entities and modules of a software project.
 
 \subsection{Summary of techniques}
-While unsupervised techniques such as LDA is appealing in its lack of human intervention, and thus lower effort, supervised learners have the advantage of domain knowledge which typically means improved results.
+While unsupervised techniques such as LDA is appealing in its lack of human intervention, and thus lower effort, 
+supervised learners have the advantage of domain knowledge which typically means improved results. 
 Our manual annotations were fairly quick to do, taking on average five minutes per 30 day period.
@@
 
-\textbf{Understanding the figures} -- In Fig.
-\ref{fig:mysql-timeline}, we see that the NFRs \emph{functionality}, \emph{portability} and \emph{maintainability} contain more labeled topics, since these NFRs have been more intensely shaded.
+\textbf{Understanding the figures} -- For example, in Figure \ref{fig:mysql-timeline}, we see that the NFRs \emph{functionality}, \emph{portability} and \emph{maintainability} contain more labeled topics, since these NFRs have been more intensely shaded.
 % boxes.
 not sure boxes is important..
-%We note that the gap in the MaxDB data is due to an interruption in the source data, having to do with the infrastructure for releasing developer artifacts, rather than specific occurrences in the project.
 % I grepped query1.mrs2 http://softwareprocess.es/y/y/maxdb-76.tar.gz
@@
 For example, both projects show a high number of NFRs recognized at the first period of analysis.
-This is due to our window choice: we deliberately targeted our analysis to coincide with the development lifecycle of a particular major release of the software.
-For MaxDB, version 7.5.00  was released in December of 2003.
+This is due to our window choice: we deliberately targeted our analysis to when MySQL 3.23 was first announced %Major update \& End-of life
+and when MaxDB 7.500 was first announced. For MaxDB, version 7.5.00  was released in December of 2003. 
 From analyzing the mailing list for external validation, we know that release 7.5.00.23 saw the development of PHP interfaces, possibly accounting for the simultaneous increase in the \emph{portability} NFR at the same time.
@@
 Version 3.23.31 (January, 2001) was the production release (non-beta), and we see a flurry of topics labeled with \emph{functionality} and \emph{maintainability}, as one might expect.
-After this point, we can consider this version to enter the maintenance phase of the lifecycle.
+After this point, we can consider this version to have entered the maintenance phase of the lifecycle.
 

### 2010-08-20 icse/icse2011.tex
finished visualization analysis

@@
 Version 3.23.31 (January, 2001) was the production release (non-beta), and we see a flurry of topics labeled with \emph{functionality} and \emph{maintainability}, as one might expect.
-After this point, we can consider this version to have entered the maintenance phase of the lifecycle.
-
-%Note that MySQL's release process was haphazard at this point: they would release a patch version whenever the need arose.
-By version 5.1, the company moved to a milestone release, with new releases targeted for every six months.
-
-% \todo[inline]{Does our visualization capture the entire timeline?}
-
-%\todo[inline]{try to pin down these reasons below}
+After this point, this version enters the maintenance phase of the lifecycle.
+In May 2001, there is an increase in the number of topics labelled with \emph{portability}.
+This might be related to release 3.23.38, which focused on Windows compatibility.
+Similarly, in Augsut, 2002, both \emph{functionality} and \emph{portability} are high in our tool, and mailing list data suggests this is related to the release of version 3.23.52, a general bug fix with focus on security (encoded as the \emph{functionality} NFR in the ISO9126 model).
+After this point, efforts shift to the newer releases (4.0, 4.1, 5.0).
 

### 2010-08-20 icse/icse2011.tex
minor corrections

@@
 In other words, MySQL 3.23 was being maintained rather than actively developed, whereas MaxDB 7.500 was being actively developed into MaxDB 7.6, and then maintained thereafter.
-\todo[inline,color=yellow]{expand on this with dates, and reference to visualizations}
+%\todo[inline,color=yellow]{expand on this with dates, and reference to visualizations}
 
-MySQL and MaxDB's machine learners made some decisions based off some shared words: \textsf{bug, code, compiler, database, HP UX, delete, memory, missing, problems, removed, add, added, changed, problem, and test}.
+MySQL and MaxDB's machine learners made some decisions based off a few shared words: \textsf{bug, code, compiler, database, HP UX, delete, memory, missing, problems, removed, add, added, changed, problem, and test}.
 Adding these words to the word bags of \textsf{exp2} and \textsf{exp3} could improve performance while ensuring they were only domain specific.

### 2010-08-20 icse/icse2011.tex
abstract tighten so it fits in one column

@@
 % a sub problem
-While too much specificity can produce non-generalizable results, too little produces broad learners that do not provide much immediately useful detail. 
+% While too much specificity can produce non-generalizable results, too little produces broad learners that do not provide much immediately useful detail. 
 % what we do
@@
 % data mining generalizable
-To ensure we have a generalizable, project-agnostic approach we label topics using non-functional requirements,
+Our approach is 
+%To ensure we have a 
+generalizable and project-agnostic as we label topics using non-functional requirements,
 % these qualities 
@@
 % We used LDA, CVS, Bitkeeper
-We extracted topics using Latent Dirichlet Allocation (LDA) from the commit log comments of their version control systems (CVS and BitKeeper). 
+We extracted topics using Latent Dirichlet Allocation (LDA) from the commit-log comments of their version control systems (CVS and BitKeeper). 
 % We show NFRs exist
@@
 % we can label stuff
-Our labelled topic extraction technique allowed us to devise appropriate, context-sensitive labels across these two projects, providing insight into their software development activities.
+Labelled topic extraction allowed us to devise appropriate, context-sensitive labels across these two projects, providing insight into their software development activities.
 \end{abstract}
@@
 Topics, too, are a valuable and automatically extractable abstraction of software artifacts.
-It abstracts these artifacts by their subject matter, rather than their concrete representation. 
+Topics abstracts these artifacts by their subject matter, rather than their concrete representation. 
 Much like Aspect-Oriented Programming, topic labeling can identify cross-cutting concerns in a software project~\cite{Baldi2008}.

### 2010-08-20 icse/icse2011.tex
PARAGRAPH ARGUES ABOUT THE CONCRETE APPLICATION OF TOPIC MODELLING AND LABELLED TOPIC EXTRACTION IT NEEDS TO BE TIGHTER BUT IT ALSO NEEDS TO BE HERE

@@
 
+%%%% THIS PARAGRAPH ARGUES ABOUT THE CONCRETE APPLICATION OF 
+%%%% TOPIC MODELLING AND LABELLED TOPIC EXTRACTION
+%%%% IT NEEDS TO BE TIGHTER BUT IT ALSO NEEDS TO BE HERE
+There are many practical applications of topic modelling and \emph{labelled  topic extraction}.
+These methods are useful for providing overviews of development.
+We can improve project dashboards, used by managers, by showing both the current topics of development as well those that relate to NFRs.
+Some quality assurance questions can be addressed by seeking evidence within repositories that NFR related topics are being addressed.
+Topic modelling can be used to check if development is on track and focused.
+% This is a very long example it'd be nice if it was tighter
+Perhaps development has been sidetracked by cascading issues that appeared during implementation.
+For instance, a developer adding an optimization to a RDBMS might have to change multiple modules in order to provide that optimization code with enough information to operate. 
+This issue might trigger a series of cascading changes that might be difficult to discover when trying to explain why the project's release date has been missed. 
+Labelled Topic extraction could be used to detect this distinct focus of development.
+
 %% RESEARCH DIDNT DO THIS AND WE ARE TOTALLY AWESOME AND WE DID
@@
 
-\todo[inline]{We still haven't shown that topics are useful -- a reviewer comment}
+%\todo[inline]{We still haven't shown that topics are useful -- a reviewer comment}
 

### 2010-08-20 icse/icse2011.tex
updated process model, changed subsubsec to simpler heading for space

--- 2010-08-20 icse/icse2011.tex
@@
 \section{Introduction}
-
 Topics in software engineering practice refer to specific shared threads of discussions between project stakeholders. 
@@
 This paper begins by introducing labelled topic extraction (Section \ref{sec:wordlist}).
-We show that these labels with their topics can be learned and used to classify other data-sets, either without training (Section \ref{sec:unsuplabelling}), or with training (Section \ref{sec:suplearn}).
+We show that these labels with their topics can be learned and used to classify other data-sets, either without training (Section \ref{sec:unsuplabelling}), or with training (Section \ref{sec:suplabelling}).
 We then present visualizations of named topics and their trends over time to aid communication and analysis.
@@
 
-\subsubsection{Analysis of the unsupervised labelling}
-
-Based on the labels, and our manual topic labelling, we compared the results of the unsupervised word matching approach. 
+\textbf{Analysis of the unsupervised labelling} -- Based on the labels, and our manual topic labelling, 
+we compared the results of the unsupervised word matching approach. 
 For each quality we assessed whether unsupervised labels matches the manual annotations.
@@
 \subsection{Supervised labelling}
-\label{sec:suplearn}
+\label{sec:suplabelling}
 Supervised labelling requires expert analysis of the correct word to assign to a topic.
@@
 
-\subsubsection{Analysis of the supervised labelling}
-\label{sec:suplabelling}
+\textbf{Analysis of the supervised labelling} -- 
+% \label{sec:suplabelling}
 Because our data-set was of word counts we expected Bayesian techniques, often used in spam filtering, to perform well. 
@@
 These classifiers performed better than other multi-label classifiers. 
-They have the best micro and macro ROC scores, although their results are comparable to the naive Bayesian learners we used in Section \ref{sec:suplearn}.
+They have the best micro and macro ROC scores, although their results are comparable to the naive Bayesian learners we used in Section \ref{sec:suplabelling}.
 \todo[inline]{discuss this similarity in more detail}
@@
 
-With respect to the questions raised in section \ref{sec:suplearn}:
+With respect to the questions raised in section \ref{sec:suplabelling}:
 
@@
 \emph{efficiency} and \emph{portability} related topics.
-
-\begin{comment}
- shared-report.sh ; cat output/shared-j48-words
-
-      2 bug
-      2 code
-      2 compiler
-      2 database
-      2 Delete_
-      2 HP_UX
-      2 memory
-      2 missing
-      2 monty_mysql_com
-      2 monty_mysql_fi
-      2 problems
-      2 Removed
-      2 work__home_bk_mysql
-      4 4_0
-      4 Add
-      4 added
-      4 Changed
-      4 problem
-      6 Added
-      6 test
-
-\end{comment}
 
@@
 The presence of a well-maintained set of tags is obviously essential to the success of this technique.
-
-\begin{comment}
-%We maintain that these results require more substantive manipulation before being reported to, for example, management.
-Like any large-scale data analysis project, it requires skill to derive the implications.
-Hooking the backend results directly to a reporting tool is unlikely to produce useful results.
-
-% There has been an explosion of interest in mining OSS repositories.
-Godfrey and Tu \cite{godfrey00} were among the first to use this data to assess software evolution.
-They discussed how well the growth of Linux was predicted by laws that Lehman and colleagues \cite{lehman85} proposed.
-More recently, Herraiz et al.
-\cite{herraiz07icsm} predicted OSS growth using time series analysis.
-Like us, they refer to release windows as a useful unit of analysis.
-\end{comment}
 
@@
 
- %\todo[inline]{Future work is disjointed the question should be more clear}
-
-% Be careful how you merge!
-
-% We demonstrated that static but domain-specific knowledge can improve unsupervised labelling of extracted topics.
-Our \textsf{exp2} experiment used small accurate word bags to label topics but performed just as well as \textsf{exp3}, which used many more general terms from WordNet.
-We then showed that with some supervision, and by using efficient machine learners based on Naive Bayesian classifiers, we could improve the accuracy of automatic labelling topics even further.
-
-% Our manual inspection and annotation of the topics extracted from MySQL and MaxDB revealed that many of the extracted topics dealt with non-functional requirements, and these topics were spread across the entire history of a project.
-In the cases of MaxDB and MySQL, portability was a constant maintenance concern and was prevalent throughout the entire lifetime of the projects.
-
-% We showed that non-functional requirements are often trending topics, that non-functional requirements are quite common in developer topics, and that there are efficient methods of semi-automating and automating topic labelling.
-
-% There are several avenues of further investigation.
-We want to investigate developer attitudes related to these labels: i.e., when we label a topic, was the developer expressing positive or negative qualities about that label?
-It is difficult to map abstract qualities to particular messages.
-Is a ``quality'' discussion about more than just corrective maintenance?
-%In th
-%is study we take the position that any type of maintenance has to do with software quality.
-
-%XXX can save space here!
-%\section{Appendix}
 \appendix
@@
 \end{document}
+
+% \begin{comment}
+%  shared-report.sh ; cat output/shared-j48-words 
+% 
+%       2 bug
+%       2 code
+%       2 compiler
+%       2 database
+%       2 Delete_
+%       2 HP_UX
+%       2 memory
+%       2 missing
+%       2 monty_mysql_com
+%       2 monty_mysql_fi
+%       2 problems
+%       2 Removed
+%       2 work__home_bk_mysql
+%       4 4_0
+%       4 Add
+%       4 added
+%       4 Changed
+%       4 problem
+%       6 Added
+%       6 test
+% 
+% \end{comment}
+
 % \begin{comment}
@@
 Naming is important but lets tell the reader what is being named first.}
+
+ %\todo[inline]{Future work is disjointed the question should be more clear}
+
+% Be careful how you merge!
+
+% We demonstrated that static but domain-specific knowledge can improve unsupervised labelling of extracted topics.
+Our \textsf{exp2} experiment used small accurate word bags to label topics but performed just as well as \textsf{exp3}, which used many more general terms from WordNet.
+We then showed that with some supervision, and by using efficient machine learners based on Naive Bayesian classifiers, we could improve the accuracy of automatic labelling topics even further.
+
+% Our manual inspection and annotation of the topics extracted from MySQL and MaxDB revealed that many of the extracted topics dealt with non-functional requirements, and these topics were spread across the entire history of a project.
+In the cases of MaxDB and MySQL, portability was a constant maintenance concern and was prevalent throughout the entire lifetime of the projects.
+
+% We showed that non-functional requirements are often trending topics, that non-functional requirements are quite common in developer topics, and that there are efficient methods of semi-automating and automating topic labelling.
+
+% There are several avenues of further investigation.
+We want to investigate developer attitudes related to these labels: i.e., when we label a topic, was the developer expressing positive or negative qualities about that label?
+It is difficult to map abstract qualities to particular messages.
+Is a ``quality'' discussion about more than just corrective maintenance?
+%In th
+%is study we take the position that any type of maintenance has to do with software quality.
+
+%XXX can save space here!
+%\section{Appendix}

### 2010-08-20 icse/icse2011.tex
added todos back
Explain 1 diagram

@@
  \includegraphics[width=\textwidth]{figures/unsupervised-bar}
- \caption{ROC values for each NFR, per project and wordlist.}
+ \caption{Performance, ROC values, of Unsupervised topic labelling for
+   each NFR and per word-list. This graph shows how well the
+   unsupervised topic labelling matched our manual annotations.}
+
 %\todo[inline]{this is too tiny in print}

### 2010-08-20 icse/icse2011.tex
shortened abstract again

--- 2010-08-20 icse/icse2011.tex
@@
 %  Within the field of mining software repositories, a
-Approaches such as topic modeling and concept location rely on the automated application of machine learning algorithms to corpora in order to understand the latent subject matter of software development artifacts.
+To understand the latent subject matter of software development artifacts, researchers have introduced topic modeling and concept location.
+These techniques rely on the automated application of machine learning algorithms to corpora.
 % the problem
- Unfortunately, the output of these tools is often difficult to distinguish and interpret.
-Results are usually abstract, as the topics that are automatically extracted are word-lists lacking semantic interpretation.
+Their outputs are word-lists lacking semantic interpretation, which are difficult to distinguish and interpret. 
+%Results are usually abstract, as the topics that are automatically extracted 
 % the solution
- Therefore, in order to have a meaningful discussion about the topics of software development, we must be able to interpret and devise labels for these topics.
-However, current software engineering topic analysis and concept analysis research relies on manual naming and labelling of topics.
+
+In order to have a meaningful discussion about the topics of software development, we must be able to interpret and devise labels for these topics. 
+%However, current software engineering topic analysis and concept analysis research relies on manual naming and labelling of topics.
 % the labelling research
-Furthermore, these topic labelling approaches neither use domain-specific knowledge to improve results, nor contextualize those results for developers.   
+Current topic labelling approaches are manual, and do not use domain-specific knowledge to improve results. Nor to they contextualize those results for developers.   
 % a sub problem
@@
 % what we do
-This paper introduces \emph{labelled topic extraction}, in which topics are extracted from commit comments and given labels relating to a cross-project taxonomy. 
+
+To solve this problem, this paper introduces \emph{labelled topic extraction}, 
+in which topics are extracted from commit comments and given labels relating to a cross-project taxonomy. 
 % data mining generalizable
@@
 %To ensure we have a 
-generalizable and project-agnostic as we label topics using non-functional requirements,
+generalizable and project-agnostic, as we label topics using non-functional requirements,
 % these qualities 
@@
 % We show NFRs exist
-Our results were generalizable across the two projects, showing that non-functional requirements were commonly discussed, and we identified topic trends over time. 	
+%Our results were generalizable across the two projects, showing that non-functional requirements were commonly discussed, 
+%and we identified topic trends over time. 	
 % we can label stuff
-Labelled topic extraction allowed us to devise appropriate, context-sensitive labels across these two projects, providing insight into their software development activities.
+Labelled topic extraction allowed us to devise appropriate, context-sensitive labels across these two projects, providing insight into their software development activities over time.
 \end{abstract}

### 2010-08-20 icse/icse2011.tex
15:25 <abez> we can't really address inter-rater reliability with what we have 15:25 <abez> the mention in validity threats should be enough 15:25 <neilernst> yes, i agree
Removed ZeroR part, updated figure
Signed-off-by: Neil Ernst <neil@neilernst.net>

@@
  We considered the best learner for a label to be the learner which had the highest ROC value for that label. 
-Figure \ref{fig:best-learn-per-tag} uses the ZeroR learner as a baseline, a common practice in machine learning,
-as ZeroR naively chooses the largest category all of the time. 
-ZeroR occasionally outperforms other learners; for labels which are not as common, this is to be expected because any miscategorization will hurt accuracy. 
-This is why we use ROC values, instead of accuracy, as they can better represent performance on labels which are not applicable to the majority of samples.
+%Figure \ref{fig:best-learn-per-tag} uses the ZeroR learner as a baseline, a common practice in machine learning,
+%as ZeroR naively chooses the largest category all of the time. 
+%ZeroR occasionally outperforms other learners; for labels which are not as common, this is to be expected because any miscategorization will hurt accuracy. 
+%This is why we use ROC values, instead of accuracy, as they can better represent performance on labels which are not applicable to the majority of samples.
 
@@
 
-\begin{figure*}[ht]
+\begin{figure}[ht]
 \centering
-\subfloat[MySQL]{
-\includegraphics[width=0.4\textwidth]{figures/mysql-supervised}
-\label{fig:subfig1}
-}
-\subfloat[MaxDB]{
-\includegraphics[width=0.4\textwidth]{figures/maxdb-supervised}
-\label{fig:subfig2}
-}
-\caption[]{ROC value for the best learner per label, compared with ZeroR accuracy values.  Possible values range from 0--1.
-\todo[inline]{better discussion}
+% \subfloat[MySQL]{
+\includegraphics[width=0.45\textwidth]{figures/both-supervised}
+% \label{fig:subfig1}
+% }
+% \subfloat[MaxDB]{
+% \includegraphics[width=0.4\textwidth]{figures/maxdb-supervised}
+% \label{fig:subfig2}
+% }
+\caption[]{ROC value for the best learner per label. Possible values range from 0--1.
+%\todo[inline]{better discussion}
 }
 \label{fig:best-learn-per-tag}
-\end{figure*}
+\end{figure}
 

### 2010-08-20 icse/icse2011.tex
Updated intro with reference to NFRs and ISO

@@
 % why are topics useful
-Often developers interact with abstractions of the artifacts from which topics are derived, rather than the artifact itself. 
+Developers often interact with abstractions of the artifacts from which topics are derived, rather than the artifact itself. 
 Such abstractions include project dashboards~\cite{kersten2005mylar}, tags~\cite{treude2010}, and complexity metrics \cite{mccabe1976complexity}. 
@@
 Topics abstracts these artifacts by their subject matter, rather than their concrete representation. 
-Much like Aspect-Oriented Programming, topic labeling can identify cross-cutting concerns in a software project~\cite{Baldi2008}.
-The topic of development is relevant because it identifies the distinct issues facing development at a certain time.
+Much like aspect-oriented programming, topic labeling can identify cross-cutting concerns in a software project (as in~\cite{Baldi2008}).
+The topic of development is relevant because it identifies the distinct issues facing development at a particular time.
 
@@
 That is, extracted topics are usually manually named and labelled by those who read the topic words.
-We propose a method of interpreting and labelling topics, and their related artifacts, that is generalizable and relevant to multiple projects.
 
@@
 Deriving such topics is difficult. 
-Topic analysis produces raw topics that are represented as unlabelled context-free word-lists or word distributions.
+Topic analysis produces% raw topics that are represented as
+ unlabelled context-free word-lists or word distributions.
 Topics have to be identified by interpreting the prevalent words in a word distribution and by inspecting related documents. 
 This manual approach is impractical when one has to handle hundreds of different topics. To scale to meaningfully sized projects, we need automatic assistance to determine the topic label.
+We propose a method of interpreting and labelling topics, and their related artifacts, that is generalizable and relevant to multiple projects.
 
@@
 % harp on data mining more
-In general the data mining of software artifacts tends to be very project specific, yet NFRs are not. 
-NFRs are prevalent in almost all software systems.
+In general, the mining of software artifacts tends to be very project specific, yet NFRs are not.
+There is a series of standards on NFRs (such as \cite{iso9126}) specifically intended to apply to projects of varying types (indicating this is possible, if perhaps not universally accepted).
+% NFRs are prevalent in almost all software systems.
 % In this paper we will show that NFR related topics occur quite frequently.
@@
 In particular, we use software quality models to generalize topics across projects.
-This allows us to data mine software systems with the intention of comparing their NFR-related development topics.
+This allows us to mine software systems with the intention of comparing their NFR-related development topics.
 % in order to achieve generalizable, cross project data mining results.

### 2010-08-20 icse/icse2011.tex
classifier similarity

@@
 % A topic may have more than one matching keyword. % ok what?
-\todo[inline]{inter-rater reliability }
+%\todo[inline]{inter-rater reliability }
 % \todo[inline]{There should be a section here which is about manual annotation. It is really important and it is required by both analyses. Plus we can tell them how.}
@@
 The Hierarchy Of Multi-label classifiERs (HOMER) and Binary Relevance (BR) act as a hierarchy of learners: BR is flat, while HOMER tries to build a deeper hierarchy to build a more accurate learner~\cite{mulan}.
-These classifiers performed better than other multi-label classifiers. 
-They have the best micro and macro ROC scores, although their results are comparable to the naive Bayesian learners we used in Section \ref{sec:suplabelling}.
-\todo[inline]{discuss this similarity in more detail}
+These classifiers performed better than other multi-label classifiers as they have the best micro and macro ROC scores. 
+The multi-label and single-label learners had similar performance: for MySQL, BR and NaiveBayes had similar macro-ROC scores of $0.74$.
+% although their results are comparable to the naive Bayesian learners we used in Section \ref{sec:suplabelling}.
+%\todo[inline]{discuss this similarity in more detail}
 

### 2010-08-20 icse/icse2011.tex
tighten conclusions and word-bags

@@
 results, due to the high noise level. Nevertheless, these results are
-better than our previous word bag results of \textsf{exp2} and
+better than our previous word-list results of \textsf{exp2} and
 \textsf{exp3}, because the ROC values are sufficiently higher in most
@@
 MySQL and MaxDB's machine learners made some decisions based off a few shared words: \textsf{bug, code, compiler, database, HP UX, delete, memory, missing, problems, removed, add, added, changed, problem, and test}.
-Adding these words to the word bags of \textsf{exp2} and \textsf{exp3} could improve performance while ensuring they were only domain specific.
+Adding these words to the word-lists of \textsf{exp2} and \textsf{exp3} could improve performance while ensuring they were only domain specific.
 
@@
 % para 1 we did this, we were successful
-In this paper we presented a cross-project data mining technique that was appropriate for mining and labelling latent software development topics using a non-functional requirements taxonomy.
+In this paper we presented a cross-project data mining technique that was appropriate for mining and labelling latent software development topics using a non-functional requirements (NFRs) taxonomy.
 %topics and artifacts from project histories.
@@
 \begin{itemize}
-\item We provide an unsupervised method of topic labelling (word-bags);
+\item We provide an unsupervised method of topic labelling (word-lists);
 %\item Our unsupervised method can be applied to projects without training;
@@
 \end{itemize}
-We leveraged word-bags, machine learning classifiers and multi-label machine learning classifiers.  
-We evaluated these techniques using multiple experiments against two open source database systems and visualized the results.
-We found that our topic labelling techniques were effective.
+
+%We leveraged word-bags, machine learning classifiers and multi-label machine learning classifiers.  
+%We evaluated these techniques using multiple experiments against two open source database systems and visualized the results.
+%We found that our topic labelling techniques were effective.
 
@@
 We validated our topic labelling techniques using multiple experiments.
-Using unsupervised word-bag analysis we set up two NFR topic labelling experiments: 
-our \textsf{exp2} experiment used small but accurate word-bags, while \textsf{exp3} used large general word-bags derived from WordNet.
+Using unsupervised word-list analysis we set up two NFR topic labelling experiments: 
+our \textsf{exp2} experiment used small but accurate word-lists, while \textsf{exp3} used large general word-lists derived from WordNet.
 The performance of these techniques differed per NFR but both \textsf{exp2} and \textsf{exp3} had similar performance.
@@
 % total topics - maybe move above
-We showed that non-functional requirements are often trending topics, that non-functional requirements are quite common in developer topics, and that there are efficient methods of semi-automating and automating topic labelling.
+We showed that NFRs are often trending topics, that NFRs are quite common in developer topics, and that there are efficient methods of semi-automating and automating topic labelling.
 % We found NFRs
-Our manual inspection and annotation of the topics extracted from MySQL and MaxDB revealed that many of the extracted topics dealt with non-functional requirements, and these topics were spread across the entire history of a project.
+Our manual inspection and annotation of the topics extracted from MySQL and MaxDB revealed that many of the extracted topics dealt with NFRs, and these topics were spread across the entire history of a project.
 % Try to provide an example of NFRs but little context
@@
 Thus we leveraged software engineering knowledge to produce a method of domain specific topic labelling.
-The word-bag technique is not project specific, thus it has intrinsic value when analyzing multiple projects.
+The word-list technique is not project specific, thus it has intrinsic value when analyzing multiple projects.
 We demonstrated multiple supervised and unsupervised methods of topic labelling, including multi-label learners.

### 2010-08-20 icse/icse2011.tex
removed 2 references for space considerations (authors were ref'd twice)

@@
 
-Massey~\cite{massey02icse} and Scacchi~\cite{scacchi02,scacchi05b} looked at the topic of requirements in open-source software.
+%Massey~\cite{massey02icse} and Scacchi~\cite{scacchi02,scacchi05b} looked at the topic of requirements in open-source software. 
+% this reduces the line length and drops a ref even though we still have a scacchi kickin around
+Massey~\cite{massey02icse} and Scacchi~\cite{scacchi05b} looked at the topic of requirements in open-source software. 
 Their work discusses the source of the requirements and how they are used in the development process.
@@
 
-Hindle et al.
+% XXX vanity or space.. 
+% Hindle et al.
 \cite{Hindle2007} examined release patterns in OSS.
@@
 %\todo[inline]{Need to juxtapose our work against previous work -- contribution}
-\todo[inline]{tighten... typically conclusions are much shorter.}
+%\todo[inline]{tighten... typically conclusions are much shorter.}
 

### 2010-08-20 icse/icse2011.tex
hacking the abstract a little more

@@
 \newcommand{\XXX}[1]{\textcolor{red}{{\it \textbf{[XXX: #1]}}}}
-\hyphenation{Max-DB}
+\hyphenation{Max-DB data-base}
 
@@
 %  Within the field of mining software repositories, a
-To understand the latent subject matter of software development artifacts, researchers have introduced topic modeling and concept location.
-These techniques rely on the automated application of machine learning algorithms to corpora.
+To understand the latent subject matter of software development artifacts, researchers have proposed topic modeling and concept location.
+%These techniques rely on the automated application of machine learning algorithms to corpora.
+These techniques use unsupervised machine learning algorithms find topics or concepts.
 % the problem
-Their outputs are word-lists lacking semantic interpretation, which are difficult to distinguish and interpret. 
+These topics are word-lists lacking semantic interpretation, which are difficult to distinguish and interpret. 
 %Results are usually abstract, as the topics that are automatically extracted 
 % the solution 
-
 In order to have a meaningful discussion about the topics of software development, we must be able to interpret and devise labels for these topics. 
@@
 % the labelling research
-Current topic labelling approaches are manual, and do not use domain-specific knowledge to improve results. Nor to they contextualize those results for developers.   
+Current topic labelling approaches are manual, and do not use domain-specific knowledge to improve, contextualize, or describe results for the developers.
 % a sub problem
@@
 % We used LDA, CVS, Bitkeeper
-We extracted topics using Latent Dirichlet Allocation (LDA) from the commit-log comments of their version control systems (CVS and BitKeeper). 
+We extracted topics using Latent Dirichlet Allocation (LDA) from the commit-log comments of their source control systems (CVS and BitKeeper). 
 % We show NFRs exist
@@
 % we can label stuff
-Labelled topic extraction allowed us to devise appropriate, context-sensitive labels across these two projects, providing insight into their software development activities over time.
+Labelled topic extraction produces appropriate, context-sensitive labels relevant to these projects, providing insight into their evolving software development activities.
 \end{abstract}

### 2010-08-20 icse/icse2011.tex
Added a comment
loosen layout rules
Changed a fig caption

@@
 
-\begin{table*}[]
+\begin{table*}
 	\centering
@@
 %\todo[inline,color=green]{this table should have mysql}
-\begin{table}[h]
+\begin{table}
 
@@
 
-\begin{figure}[ht]
+\begin{figure}
 \centering
@@
 % }
-\caption[]{ROC value for the best learner per label. Possible values range from 0--1.
+\caption[]{ROC value for the best learner per label for MaxDB and MySQL. ROC values range from 0--1.
 %\todo[inline]{better discussion}
@@
 A macro-ROC measurement is the average ROC over the ROC values for all labels, where a micro-ROC is the average ROC over all examples that were classified. 
-Unfortunately for MaxDB, the macro-ROC values are undefined because of poor performance of one of the labels, see Figure \ref{fig:mulan}.
+Unfortunately for MaxDB, the macro-ROC values are undefined because of poor performance of one of the labels.%, see Figure \ref{fig:mulan}.
 
@@
 
-\begin{figure*}[ht]
+\begin{figure*}
 \centering
@@
 \label{sec:limit}
+
+%XXX can just remove this
 \subsection{Effectiveness}

### 2010-08-20 icse/icse2011.tex
related work start

@@
 \label{sec:related}
+%XXX really not sure about what the two-ways are
 The idea of extracting higher-level `concerns' (also known as `concepts', `aspects' or `requirements') has been approached in two ways.
@@
 Cleland-Huang and her colleagues published work on mining requirements documents for non-functional requirements (NFR) (quality requirements)~\cite{Cleland-Huang2006}. 
-One approach they tried was similar to this one, with keywords mined from NFR catalogues found in their previous work~\cite{chung99}.
+Their approach was similar, as they mined keywords from their own NFR catalogues~\cite{chung99}.
+%One approach they tried was similar to this one, with keywords mined from NFR catalogues found in their previous work~\cite{chung99}. 
 They managed recall of 80\% with precision of 57\% for the Security NFR, but could not find a reliable source of keywords for other NFRs. 
@@
 
-Mens et al.~\cite{mens08icsm} conducted an empirical study of Eclipse, the open source software (OSS) source code editor, to verify the claims of Lehman~\cite{lehman97sms}.
-They concerned themselves with source code only, and found Law Seven, ``Declining Quality'', to be too difficult to assess: ``[we lacked an] appropriate measurement of the evolving quality of the system as perceived by the users \cite[p.
+% Not sure we need this
+% Mens et al.~\cite{mens08icsm} conducted an empirical study of Eclipse, the open source software (OSS) source code editor, to verify the claims of Lehman~\cite{lehman97sms}.
+%  They concerned themselves with source code only, and found Law Seven, ``Declining Quality'', to be too difficult to assess: ``[we lacked an] appropriate measurement of the evolving quality of the system as perceived by the users \cite[p.
 388]{mens08icsm}''.
-This paper examines the notions of quality in terms of a consistent ontology, as Mens et al.
-call for in their conclusions.
+% This paper examines the notions of quality in terms of a consistent ontology, as Mens et al. call for in their conclusions.
 

### 2010-08-20 icse/icse2011.tex
related work start

--- 2010-08-20 icse/icse2011.tex
@@
 
+% Not sure we need this
 Mens et al.~\cite{mens08icsm} conducted an empirical study of Eclipse, the open source software (OSS) source code editor, to verify the claims of Lehman~\cite{lehman97sms}.
@@
 388]{mens08icsm}''.
-This paper examines the notions of quality in terms of a consistent ontology, as Mens et al.
-call for in their conclusions.
+This paper examines the notions of quality in terms of a consistent ontology, as Mens et al. call for in their conclusions.
 

### 2010-08-20 icse/icse2011.tex
related work edits

--- 2010-08-20 icse/icse2011.tex
@@
 \label{sec:related}
-%XXX really not sure about what the two-ways are
-The idea of extracting higher-level `concerns' (also known as `concepts', `aspects' or `requirements') has been approached in two ways.
+
+The idea of extracting higher-level `concerns', also known as `concepts', `aspects' or `requirements', has been approached from documentation-based and repositories-based perspectives.
 
 Cleland-Huang and her colleagues published work on mining requirements documents for non-functional requirements (NFR) (quality requirements)~\cite{Cleland-Huang2006}. 
-Their approach was similar, as they mined keywords from their own NFR catalogues~\cite{chung99}.
+Their approach was similar to ours, as they mined keywords from their own NFR catalogues~\cite{chung99}.
 %One approach they tried was similar to this one, with keywords mined from NFR catalogues found in their previous work~\cite{chung99}. 
@@
 Instead, they developed a supervised classifier by using human experts to identify an NFR training set. 
-There are several reasons we did not follow this route. 
-One, we believe we have a more comprehensive set of terms due to the taxonomy we chose. 
-Secondly, we wanted to compare across projects. 
-Their technique was not compared across different projects and the applicability of the training set to different corpora is unclear. 
-A common taxonomy allows us to make inter-project comparison (subject to the assumption that all projects conceive of these terms in the same way). 
-Thirdly, while the objective of Cleland-Huang's study was to identify new NFRs (for system development) our study assumes these NFRs are latent in the textual documents of the project. 
-Finally, the source text we use is less structured than their requirements documents.
+Our worked differed because we had a more comprehensive set of terms based on the taxonomy we chose.
+We also wanted a common taxonomy that allows us to make cross-project comparison, which was not one of their objectives.
+The objective of Cleland-Huang's study was to identify new NFRs for system development, yet our objective was to recover those latent NFRs from commit-log messages of the project. 
+Another difference was that our data was far less structured that the requirements documents used in their study.
+%There are several reasons we did not follow this route. 
+%One, we believe we have a more comprehensive set of terms due to the taxonomy we chose. 
+%Secondly, we wanted to compare across projects. 
+%Their technique was not compared across different projects and the applicability of the training set to different corpora is unclear. 
+%A common taxonomy allows us to make cross-project comparison. 
+%Although subject to the assumption that all projects conceive of these terms in the same way.
+%Thirdly, while the objective of Cleland-Huang's study was to identify new NFRs (for system development) our study assumes these NFRs are latent in the commit-log messages of the project.
 
@@
 
-The other approach is to start with code repositories, and try to extract concerns from there. 
+The other approach is to extract concerns from software repositories.
 Marcus et al.~\cite{marcus04wcre} describe their use of Latent Semantic Indexing to identify commonly occurring concerns for software maintenance. 
-Some results were interesting, but their precision was quite low. 
+%Some results were interesting, but their precision was quite low. 
 ConcernLines~\cite{treude09cl} shows tag occurrence using colour intensity. 
-They mined change request tags (such as `milestone 3') and used these to make evolutionary analyses of a single product. 
+They mined 
+developer created tags 
+%change request tags (such as `milestone 3') and used these 
+in order to analyze the evolution of a single product.
+% make evolutionary analyses of a single product. 
 The presence of a well-maintained set of tags is obviously essential to the success of this technique.
@@
 % Not sure we need this
-% Mens et al.~\cite{mens08icsm} conducted an empirical study of Eclipse, the open source software (OSS) source code editor, to verify the claims of Lehman~\cite{lehman97sms}.
+Mens et al.~\cite{mens08icsm} conducted an empirical study of Eclipse, a popular Java IDE, to verify the claims of Lehman~\cite{lehman97sms}.
 % They concerned themselves with source code only, and found Law Seven, ``Declining Quality'', to be too difficult to assess: ``[we lacked an] appropriate measurement of the evolving quality of the system as perceived by the users \cite[p.
 388]{mens08icsm}''.
-% This paper examines the notions of quality in terms of a consistent ontology, as Mens et al. call for in their conclusions.
+This paper examines the notions of quality in terms of a consistent ontology, as Mens et al. call for in their conclusions.
 
@@
 German~\cite{german03gnome} looked at GNOME specifically, and listed several sources for requirements: leader interest, mimicry, brainstorming, and prototypes.
-None of this work  addressed quality requirements in OSS, nor did it examine requirements trends.
+%None of this work  addressed quality requirements in OSS, nor did it examine requirements trends.
 

### 2010-08-20 icse/icse2011.tex
fixed intro example, removed it from discussion section.

@@
 %%%% IT NEEDS TO BE TIGHTER BUT IT ALSO NEEDS TO BE HERE
-There are many practical applications of topic modelling and \emph{labelled  topic extraction}.
-These methods are useful for providing overviews of development.
-We can improve project dashboards, used by managers, by showing both the current topics of development as well those that relate to NFRs.
-Some quality assurance questions can be addressed by seeking evidence within repositories that NFR related topics are being addressed.
+There are many practical applications of \emph{labelled topic extraction}.
+These methods are useful for providing abstractions of development activities and artifacts.
+Particularly in large projects, summarization of developer activities is essential to maintenance.
+Rather than looking for specific concepts in a project~\cite{marcus04wcre}, we are interested in understanding a software project relative to common properties of all software.
+We can improve project dashboards, used by managers, by showing both the current topics of development as well those that relate to domain-independent taxonomies, such as NFRs.
+Some quality assurance questions can be addressed by seeking evidence within repositories that NFR-related topics are being addressed.
 Topic modelling can be used to check if development is on track and focused.
+For outside users of the software, our technique can provide insight into that company's development practices. 
+Are they focused on usability, for example?
+Within the organization \emph{labelled topic extraction} acts as a historical sanity check. 
+\emph{Confirmation bias} is the difference between what people think happened and what actually happened. 
+Some software releases, for instance, are poorly received since they are quite buggy. 
+Our technique would help explain why this was the case, since it can identify NFRs which are dominating developer attention.
 % This is a very long example it'd be nice if it was tighter
-Perhaps development has been sidetracked by cascading issues that appeared during implementation.
-For instance, a developer adding an optimization to a RDBMS might have to change multiple modules in order to provide that optimization code with enough information to operate. 
-This issue might trigger a series of cascading changes that might be difficult to discover when trying to explain why the project's release date has been missed. 
-Labelled Topic extraction could be used to detect this distinct focus of development.
+% Perhaps development has been sidetracked by cascading issues that appeared during implementation.
+% For instance, a developer adding an optimization to a RDBMS might have to change multiple modules in order to provide that optimization code with enough information to operate. 
+% This issue might trigger a series of cascading changes that might be difficult to discover when trying to explain why the project's release date has been missed. 
+% Labelled Topic extraction could be used to detect this distinct focus of development.
 
@@
 % it is ok to repeat this here because it is clear and bullet pointed and the reviewers wont have any bloody excuse to not see it
-In this paper, we describe \textbf{Nomen}, which is a tool for \emph{labelled topic extraction}. \textbf{Nomen} addresses two gaps in the topic mining literature:
+In this paper, we describe \emph{labelled topic extraction}. It addresses two gaps in the topic mining literature:
 \begin{enumerate}
@@
 This is because traditional topic mining techniques are specific to a particular data-set.
-\textbf{Nomen} allows for comparisons \textit{between} projects.
+\textit{Labelled topic extraction} allows for comparisons \textit{between} projects.
   \item Topic modeling creates word lists that require interpretation by the user to assign meaning. Like (1), this means
@@
 
-\section{Using \textbf{Nomen} to understand software maintenance activities} 
+\section{Understanding software maintenance activities} 
 \label{sec:analysis}
@@
 
-There are several possible applications of \textsf{Nomen}'s techniques. 
-For outside users of the software, \textsf{Nomen} can provide insight into that company's development practices. 
-Are they focused on usability, for example?
-Within the organization \textsf{Nomen} acts as a historical sanity check. 
-Confirmation bias is the difference between what people think happened and what actually happened. 
-Some software releases, for instance, are poorly received since they are quite buggy. 
-\textsf{Nomen} would help explain why this was the case, since it can identify NFRs which are dominating developer attention.
+% There are several possible applications of \textsf{Nomen}'s techniques. 
+% For outside users of the software, \textsf{Nomen} can provide insight into that company's development practices. 
+% Are they focused on usability, for example?
+% Within the organization \textsf{Nomen} acts as a historical sanity check. 
+% Confirmation bias is the difference between what people think happened and what actually happened. 
+% Some software releases, for instance, are poorly received since they are quite buggy. 
+% \textsf{Nomen} would help explain why this was the case, since it can identify NFRs which are dominating developer attention.
 

### 2010-08-20 icse/icse2011.tex
removed name "Nomen" - it wasn't used in enough places and was just confusing. Replaced with 'labelled topic extraction'.

@@
 
-Typically a topic analysis tool like LDA will try to find $N$ independent word distributions within the word distributions of all the messages. 
+A topic analysis tool like LDA will try to find $N$ independent word distributions within the word distributions of all the messages. 
 Linear combinations of these $N$ word distributions are meant to represent and recreate the word distributions of any of the original messages. 
@@
 Figure \ref{fig:maxdb-unsup-results} shows our results for MaxDB and MySQL. 
-Our performance is measured using the \emph{ROC} (Receiver Operating Characteristic) area-under-curve value~\cite{Fawcett2006861}. 
+Our performance is measured using the Receiver Operating Characteristic area-under-curve value~\cite{Fawcett2006861}, abbreviated \emph{ROC}. 
 ROC values provide a score, similar to school letter-grades (A is 0.9, C is 0.6), reflecting how well a particular learner performed for the given data. 
@@
  \includegraphics[width=\textwidth]{figures/unsupervised-bar}
- \caption{Performance, ROC values, of Unsupervised topic labelling for
+ \caption{Performance, ROC values, of unsupervised topic labelling for
    each NFR and per word-list. This graph shows how well the
@@
 \textbf{Timeline of key events} -- As discussed earlier, the top row in each figure shows key events for each project.
-\textsf{Nomen} can pick out the underlying NFR activity behind these events.
+\textit{Labelled topic extraction} can pick out the underlying NFR activity behind these events.
 For example, both projects show a high number of NFRs recognized at the first period of analysis.
@@
 Grey bars indicate label frequency relative to \emph{a particular} NFR.
+Dashed vertical lines relate a project milestone (``*Key Event*'') to our topic windows.
 % \todo[inline]{better discussion}

### 2010-08-20 icse/icse2011.tex
improved future work

@@
 % the solution 
-In order to have a meaningful discussion about the topics of software development, we must be able to interpret and devise labels for these topics. 
+In order to have a meaningful discussion about the topics of software development, we must be able to attach meaningful labels to these topics. 
 %However, current software engineering topic analysis and concept analysis research relies on manual naming and labelling of topics.
@@
 % these qualities 
-as these apply across many software systems.% and their development artifacts.
+which apply across software systems.% and their development artifacts.
 % We used real world data
@@
 % we can label stuff
-Labelled topic extraction produces appropriate, context-sensitive labels relevant to these projects, providing insight into their evolving software development activities.
+Our labelled topic extraction technique produces appropriate, context-sensitive labels relevant to these projects, providing fresh insight into their evolving software development activities.
 \end{abstract}
@@
 Such abstractions include project dashboards~\cite{kersten2005mylar}, tags~\cite{treude2010}, and complexity metrics \cite{mccabe1976complexity}. 
-Topics, too, are a valuable and automatically extractable abstraction of software artifacts.
-Topics abstracts these artifacts by their subject matter, rather than their concrete representation. 
+Without the structured information which dashboards typically rely on, topics become a valuable and automatically extractable abstraction of software artifacts.
+Topics abstract software artifacts by their subject matter, rather than their concrete representation. 
 Much like aspect-oriented programming, topic labeling can identify cross-cutting concerns in a software project (as in~\cite{Baldi2008}).
@@
 
-This paper begins by introducing labelled topic extraction (Section \ref{sec:wordlist}).
-We show that these labels with their topics can be learned and used to classify other data-sets, either without training (Section \ref{sec:unsuplabelling}), or with training (Section \ref{sec:suplabelling}).
+This paper first introduces labelled topic extraction, and then uses that technique to analyze development activity in two large-scale examples.
+We begin by describing how we generated our data (Section \ref{sec:wordlist}).
+We show that labels with their topics can be learned and used to classify other data-sets, either without training (Section \ref{sec:unsuplabelling}), or with training (Section \ref{sec:suplabelling}).
 We then present visualizations of named topics and their trends over time to aid communication and analysis.
@@
 MaxDB started in the late 1970s as a research project, and was later acquired by SAP. 
-As of version 7.500, released April 2007, the project has 940 thousand lines of C source code~\footnote{generated using David A. Wheeler's \emph{SLOCCount}.}. 
+As of version 7.500, released April 2007, the project has 940 thousand lines of C source code\footnote{generated using David A. Wheeler's \emph{SLOCCount}.}. 
 The MySQL project started in 1994 and MySQL 3.23 was released in early 2001. 
@@
 58]{Boegh2008},'' that is sufficient for the purposes of this research.
+We use these qualities later on as classes in supervised labelling.
 The terms extracted from ISO9126 may not capture all words associated with the labels.  
@@
 For example, for the MaxDB topic consisting of a message ``exit() only used in non NPTL LINUX Versions'', we tagged that topic \emph{portability}. 
-We compared against this data-set, but we also used it for our supervised machine learning based topic classification. 
+We compared against this data-set, but we also used the data-set for our supervised machine learning based topic classification. 
 % A topic may have more than one matching keyword. % ok what?
@@
 \label{sec:suplabelling}
-Supervised labelling requires expert analysis of the correct word to assign to a topic.
-%In order to validate how effective these word-bag approaches to topic labelling would be we created a validation data set.
+Supervised labelling requires expert analysis of the correct class/label to assign to a topic.
+In our approach, we use the ISO9126 standard~\cite{iso9126} for our classes, but other taxonomies are also applicable.%In order to validate how effective these word-bag approaches to topic labelling would be we created a validation data set.
 
@@
 
-\begin{figure*}
+\begin{figure*}[]
 \centering
 \subfloat[MySQL]{
-\includegraphics[width=0.4\textwidth]{figures/multi-label-results-mysql}
+\includegraphics[width=0.3\textwidth]{figures/multi-label-results-mysql}
 \label{fig:subfig3}
 }
+
 \subfloat[MaxDB]{
-\includegraphics[width=0.4\textwidth]{figures/multi-label-results-maxdb}
+\includegraphics[width=0.3\textwidth]{figures/multi-label-results-maxdb}
 \label{fig:subfig4}
@@
 \subsection{Summary of techniques}
-While unsupervised techniques such as LDA is appealing in its lack of human intervention, and thus lower effort, 
-supervised learners have the advantage of domain knowledge which typically means improved results. 
+While an unsupervised technique such as LDA is appealing in its lack of human intervention, and thus lower effort, 
+supervised learners have the advantage of domain knowledge, which typically means improved results. 
 Our manual annotations were fairly quick to do, taking on average five minutes per 30 day period.
@@
 In terms of non-functional requirements both projects had long running trends that focused on \emph{functionality}, \emph{maintainability}, and \emph{portability}, yet MaxDB had more of a focus on \emph{efficiency} and \emph{reliability}.
-MaxDB differed from MySQL since MaxDB was being actively developed, whereas halfway through our history of MySQL 3.23, other versions of MySQL were being actively developed: MySQL versions 4.0, 4.1, and eventually 5.0 and 5.1.
-In other words, MySQL 3.23 was being maintained rather than actively developed, whereas MaxDB 7.500 was being actively developed into MaxDB 7.6, and then maintained thereafter.
+\todo[inline]{expand}
 %\todo[inline,color=yellow]{expand on this with dates, and reference to visualizations}
@@
 \textbf{Do label frequencies change over time?} -- Yes, MySQL's label
-frequencies decreased as it got older. \emph{Usability} and \emph{reliability}
+frequencies decreased as that release aged. \emph{Usability} and \emph{reliability}
 labels became more and more infrequent as it matured. \emph{Maintainability} topics
@@
 %XXX can just remove this
-\subsection{Effectiveness}
+%\subsection{Effectiveness}
 
 %{Discuss how these are CHEAP techniques and although somewhat %inaccurate they don't require a tonne of work}
-The unsupervised labelling had difficulty distinguishing between common labels and infrequent labels, often the learners would mislabel a topic deserving of an infrequent label with a more common label.
-The word-lists for \emph{correctness}, tended to be too lengthy, non-specific and broad, especially if WordNet words were used as it is a loosely defined concept in common parlance.
+The unsupervised labelling had difficulty distinguishing between common labels and infrequent labels.
+The learners would occasionally mislabel a topic deserving of an infrequent label with a more common label.
+The word-lists for \emph{correctness} tended to be too lengthy, non-specific and broad, especially if WordNet words were used, as correctness is a loosely defined concept in common parlance.
 
@@
 
-\emph{Construct validity} issues include that we used only commit messages rather than mail or bug tracker messages. 
-Conversely construct validity was addressed in our validation as we used the files, topics and revisions to annotate the commit messages.
-We also chose our taxonomy and the data to study. Also, we did not show these results to the stakeholders.
+\emph{Construct validity} -- we used only commit messages rather than mail or bug tracker messages. 
+Conversely, construct validity was addressed in our validation as we used the files, topics and revisions to annotate the commit messages.
+We also chose our taxonomy and the data to study. We did not show these results to the stakeholders.
 
-\emph{Internal validity} issues are include inter-rater reliability and our lack of rival explanation building. 
+\emph{Internal validity} -- include inter-rater reliability and our lack of rival explanation building. 
 We improved internal validity by trying to correlate and explain the behaviours observed in the analysis with the historical records of the projects.
@@
 \emph{Reliability} was addressed by each annotator following the same protocol and using the same annotations. 
-Unfortunately only two annotators were used, their annotating could be biased and we did not analyze for inter-rater reliability.
+Unfortunately only two annotators were used; their annotations could be biased as we did not analyze for inter-rater reliability.
 
@@
 They do not use external domain-specific information as we do.
-In \cite{ernst10refsq}, we describe our earlier project, similar to this, to identify change in quality requirements in GNOME software projects; our approach is solely text-matching, however, it does not leverage machine learning strategies.
+In \cite{ernst10refsq}, we describe our earlier project, similar to this, to identify change in quality requirements in GNOME software projects; our approach is solely text-matching, however, and does not leverage machine learning strategies.
 
@@
 % para 1 we did this, we were successful
-In this paper we presented a cross-project data mining technique that was appropriate for mining and labelling latent software development topics using a non-functional requirements (NFRs) taxonomy.
+In this paper we presented a cross-project data mining technique, \emph{labelled topic extraction}.
+The technique leveraged  a non-functional requirements (NFRs) taxonomy to mine and label latent software development topics.
 %topics and artifacts from project histories.
@@
 \begin{itemize}
-\item We provide an unsupervised method of topic labelling (word-lists);
+\item We provided an unsupervised method of topic labelling (word-lists);
 %\item Our unsupervised method can be applied to projects without training;
-\item We demonstrate a supervised method of labelling topics by a single NFRs (machine learning);
-\item We demonstrate a supervised method of labelling topics by multiple NFRs (multi-label machine learning);
-\item We provide a method of cross-project analysis via topic labelling;
-\item We demonstrate a method of visualizing these topics across time.
+\item We demonstrated a supervised method of labelling topics by a single NFRs (machine learning);
+\item We demonstrated a supervised method of labelling topics by multiple NFRs (multi-label machine learning);
+\item We provided a method of cross-project analysis via topic labelling;
+\item We demonstrated a method of visualizing these topics across time.
 \end{itemize}
@@
 Our fourth experiment evaluated the effectiveness of multi-label learners, such as HOMER. 
-Both kinds of learners performed well with area under the ROC curve values between $.6$ and $.8$.
+Both kinds of learners performed well with area under the ROC curve values between $0.6$ and $0.8$.
 
@@
 % total topics - maybe move above
-We showed that NFRs are often trending topics, that NFRs are quite common in developer topics, and that there are efficient methods of semi-automating and automating topic labelling.
+We showed that NFRs are often trending topics; that NFRs are quite common in developer topics; and that there are efficient methods of semi-automating and automating topic labelling.
 % We found NFRs
@@
 
-\subsection{Future Work}
+%\subsection{Future Work}
 
@@
 There are several avenues of further investigation.  
+More external validation would be useful. 
+Although we validated our comparisons using a mailing list for each project, interviews with developers would provide more detail. 
+We also think multi-label learning techniques, although in their infancy, are crucial in understanding cross-cutting concerns such as NFRs. 
+Finally, we would like to extend this analysis to other domains, to see what patterns might occur in, for example, an end-user facing software product.
  % Talk to people
-We want to investigate developer attitudes related to these labels: i.e., when we label a topic, was the developer expressing positive or negative qualities about that label?  
-It is difficult to map abstract qualities to particular messages. % what?
- % This is disjointed and needs more context
-On question we seek to answer is if a ``quality'' discussion about more than just corrective maintenance?
+% We want to investigate developer attitudes related to these labels: i.e., when we label a topic, was the developer expressing positive or negative qualities about that label?  
+% It is difficult to map abstract qualities to particular messages. % what?
+%  % This is disjointed and needs more context
+% On question we seek to answer is if a ``quality'' discussion about more than just corrective maintenance?
 %In this study we take the position that any type of maintenance has to do with software quality.

### 2010-08-20 icse/icse2011.tex
Laid out that one figure properly.
Note: if you comment the mens et al. paragraph the paper fits just fine.
Messing with conclusions.

@@
 
-\begin{figure*}[]
+\begin{figure}[]
 \centering
@@
 \label{fig:mulan}
-\end{figure*}
+\end{figure}
 
@@
 
-\textbf{Do the different projects differ in their relative topic
-  interest?} -- Yes. MySQL 3.23 had proportionally more
+\textbf{Do the different projects differ in their relative topic interest?} -- 
+Yes. MySQL 3.23 had proportionally more
 topics labelled \emph{functionality}, while MaxDB had proportionally more
 \emph{efficiency} and \emph{portability} related topics.
+ In both cases portability was a constant maintenance concern and was prevalent throughout the entire lifetime of the projects.
+\todo[inline]{fix}
 
@@
 
-% Not sure we need this
-Mens et al.~\cite{mens08icsm} conducted an empirical study of Eclipse, a popular Java IDE, to verify the claims of Lehman~\cite{lehman97sms}.
-% They concerned themselves with source code only, and found Law Seven, ``Declining Quality'', to be too difficult to assess: ``[we lacked an] appropriate measurement of the evolving quality of the system as perceived by the users \cite[p.
+% % Not sure we need this
+% Mens et al.~\cite{mens08icsm} conducted an empirical study of Eclipse, a popular Java IDE, to verify the claims of Lehman~\cite{lehman97sms}.
+% % They concerned themselves with source code only, and found Law Seven, ``Declining Quality'', to be too difficult to assess: ``[we lacked an] appropriate measurement of the evolving quality of the system as perceived by the users \cite[p.
 388]{mens08icsm}''.
-This paper examines the notions of quality in terms of a consistent ontology, as Mens et al. call for in their conclusions.
+% This paper examines the notions of quality in terms of a consistent ontology, as Mens
+et al. call for in their conclusions.
 
@@
 \item We provided a method of cross-project analysis via topic labelling;
-\item We demonstrated a method of visualizing these topics across time.
+\item We demonstrated a method of visualizing these topics across time, which we used to analyze maintenance activities.
 \end{itemize}
@@
 % Experiment description -- origin of wordbags
+% NEIL - shortening this 
 We validated our topic labelling techniques using multiple experiments.
-Using unsupervised word-list analysis we set up two NFR topic labelling experiments: 
-our \textsf{exp2} experiment used small but accurate word-lists, while \textsf{exp3} used large general word-lists derived from WordNet.
-The performance of these techniques differed per NFR but both \textsf{exp2} and \textsf{exp3} had similar performance.
+We conducted unsupervised labelling using word-lists.
+Our next approach was supervised, using single-label and multi-label learners.
+% Using unsupervised word-list analysis we set up two NFR topic labelling experiments: 
+% our \textsf{exp2} experiment used small but accurate word-lists, while \textsf{exp3} used large general word-lists derived from WordNet.
+% The performance of these techniques differed per NFR but both \textsf{exp2} and \textsf{exp3} had similar performance.
 %Our \textsf{exp2} experiment used small accurate word bags to label topics but performed just as well as \textsf{exp3}, which used many more general terms from WordNet.
 % we could can improve auto labelling with ML
-Our third experiment showed that with some supervision, and by using efficient machine learners based on Naive Bayesian classifiers, we could improve the accuracy of automatic topic labelling even further.
-Our fourth experiment evaluated the effectiveness of multi-label learners, such as HOMER. 
+% Our third experiment showed that with some supervision, and by using efficient machine learners based on Naive Bayesian classifiers, we could improve the accuracy of automatic topic labelling even further.
+% Our fourth experiment evaluated the effectiveness of multi-label learners, such as HOMER. 
 Both kinds of learners performed well with area under the ROC curve values between $0.6$ and $0.8$.
@@
 % total topics - maybe move above
-We showed that NFRs are often trending topics; that NFRs are quite common in developer topics; and that there are efficient methods of semi-automating and automating topic labelling.
+These labelling techniques allowed us to investigate the occurrence of non-functional requirements in our projects.
+We showed that NFRs are often trending topics; that NFRs are quite common in developer topics; and that there are efficient methods of automating topic labelling.
+Although it may seem like common knowledge that maintenance is a concern as a project matures, our technique was able to show this independently.
+
 % We found NFRs
-Our manual inspection and annotation of the topics extracted from MySQL and MaxDB revealed that many of the extracted topics dealt with NFRs, and these topics were spread across the entire history of a project.
+% Our manual inspection and annotation of the topics extracted from MySQL and MaxDB revealed that many of the extracted topics dealt with NFRs, and these topics were spread across the entire history of a project.
 % Try to provide an example of NFRs but little context
-In the cases of MaxDB and MySQL, portability was a constant maintenance concern and was prevalent throughout the entire lifetime of the projects.
+%too specific in this sectionIn the cases of MaxDB and MySQL, portability was a constant maintenance concern and was prevalent throughout the entire lifetime of the projects.
 
 % para 4 summarize and hammer contribution
-In conclusion,
-topic analysis and concept analysis research has produced topics that need to be manually labelled. 
-We found that topics without labels are difficult to interpret.
-Thus we leveraged software engineering knowledge to produce a method of domain specific topic labelling.
-The word-list technique is not project specific, thus it has intrinsic value when analyzing multiple projects.
+% NE not a fan of this phrasing -> % In conclusion, 
+Previous topic analysis research produced topics that needed to be manually labelled, but topics without labels are difficult to interpret.
+To improve this, we leveraged software engineering standards to produce a method of domain specific topic labelling.
+Since the word-list technique is not project specific, it can be used to compare multiple projects.
 We demonstrated multiple supervised and unsupervised methods of topic labelling, including multi-label learners.
 We provided a case study and visualization of NFR related topics in MaxDB 7.500 and MySQL 3.23. 
-During the case study we manually labelled 100s of topics by hand then used these annotations to validate the positive performance of topic labelling techniques.
-We showed that topics with labels are more useful than unlabelled topics, thus to furnish stakeholders with labelled topics we provided many methods of automatic topic labelling.
+During the case study we manually labelled hundreds of topics by hand then used these annotations to validate the positive performance of topic labelling techniques.
+We showed that topics with labels are more useful than unlabelled topics. To furnish stakeholders with labelled topics, our technique provides many methods of automatic topic labelling.
 

### 2010-08-20 icse/icse2011.tex
updated abstract

--- 2010-08-20 icse/icse2011.tex
@@
 \begin{abstract}
-%\todo[inline]{Topics and LDA sooner, Topics/concept location is a known
-%  research area and we're saying they don't do enough}
+% %\todo[inline]{Topics and LDA sooner, Topics/concept location is a known
+% %  research area and we're saying they don't do enough}
 
-% Maybe specify the hard work we did annotating the data
+% % Maybe specify the hard work we did annotating the data
 
-% our field
-%  Within the field of mining software repositories, a
-To understand the latent subject matter of software development artifacts, researchers have proposed topic modeling and concept location.
-%These techniques rely on the automated application of machine learning algorithms to corpora.
-These techniques use unsupervised machine learning algorithms find topics or concepts.
-% the problem
-These topics are word-lists lacking semantic interpretation, which are difficult to distinguish and interpret. 
-%Results are usually abstract, as the topics that are automatically extracted 
-% the solution 
-In order to have a meaningful discussion about the topics of software development, we must be able to attach meaningful labels to these topics. 
-%However, current software engineering topic analysis and concept analysis research relies on manual naming and labelling of topics.
-% the labelling research
+% % our field
+% %  Within the field of mining software repositories, a
+% To understand the latent subject matter of software development artifacts, researchers have proposed topic modeling and concept location.
+% %These techniques rely on the automated application of machine learning algorithms to corpora.
+% These techniques use unsupervised machine learning algorithms find topics or concepts.
+% % the problem
+% These topics are word-lists lacking semantic interpretation, which are difficult to distinguish and interpret. 
+% %Results are usually abstract, as the topics that are automatically extracted 
+% % the solution 
+% In order to have a meaningful discussion about the topics of software development, we must be able to attach meaningful labels to these topics. 
+% %However, current software engineering topic analysis and concept analysis research relies on manual naming and labelling of topics.
+% % the labelling research
+% Current topic labelling approaches are manual, and do not use domain-specific knowledge to improve, contextualize, or describe results for the developers.
+% % a sub problem
+% % While too much specificity can produce non-generalizable results, too little produces broad learners that do not provide much immediately useful detail. 
+% % what we do
+
+% To solve this problem, this paper introduces \emph{labelled topic extraction}, 
+% in which topics are extracted from commit comments and given labels relating to a cross-project taxonomy. 
+% % data mining generalizable
+% Our approach is 
+% %To ensure we have a 
+% generalizable and project-agnostic, as we label topics using non-functional requirements,
+% % these qualities 
+% which apply across software systems.% and their development artifacts.
+% % We used real world data
+% We evaluated our approach with an experimental study on two large-scale database projects, MySQL and MaxDB. 
+% % We used LDA, CVS, Bitkeeper
+% We extracted topics using Latent Dirichlet Allocation (LDA) from the commit-log comments of their source control systems (CVS and BitKeeper). 
+% % We show NFRs exist
+% %Our results were generalizable across the two projects, showing that non-functional requirements were commonly discussed, 
+% %and we identified topic trends over time. 	
+% % we can label stuff
+% Our labelled topic extraction technique produces appropriate, context-sensitive labels relevant to these projects, providing fresh insight into their evolving software development activities.
+
+Researchers have used topic modeling and concept location to understand the latent topics of software development artifacts.
+These techniques use unsupervised machine-learning algorithms recover topics.
+These topics are word-lists and are difficult to distinguish and interpret. 
+Topics are not meaningful until they have been named or interpreted.
 Current topic labelling approaches are manual, and do not use domain-specific knowledge to improve, contextualize, or describe results for the developers.
-% a sub problem
-% While too much specificity can produce non-generalizable results, too little produces broad learners that do not provide much immediately useful detail. 
-% what we do
+We propose a solution: \emph{labelled topic extraction}.
+Topics are extracted using Latent Dirichlet Allocation (LDA) from commit-log comments extracted from source control systems such as CVS and BitKeeper. 
+Then these topics given labels relating to a generalizable cross-project taxonomy consisting of non-functional requirements (NFRs).
+Our approach was evaluated  with experiments and case studies on two large-scale RDBMS projects: MySQl and MaxDB.
+Labelled topic extraction produces appropriate, context-sensitive labels relevant to these projects, which provides fresh insight into their evolving software development activities.
 
-To solve this problem, this paper introduces \emph{labelled topic extraction}, 
-in which topics are extracted from commit comments and given labels relating to a cross-project taxonomy. 
-% data mining generalizable
-Our approach is 
-%To ensure we have a 
-generalizable and project-agnostic, as we label topics using non-functional requirements,
-% these qualities 
-which apply across software systems.% and their development artifacts.
-% We used real world data
-We evaluated our approach with an experimental study on two large-scale database projects, MySQL and MaxDB. 
-% We used LDA, CVS, Bitkeeper
-We extracted topics using Latent Dirichlet Allocation (LDA) from the commit-log comments of their source control systems (CVS and BitKeeper). 
-% We show NFRs exist
-%Our results were generalizable across the two projects, showing that non-functional requirements were commonly discussed, 
-%and we identified topic trends over time. 	
-% we can label stuff
-Our labelled topic extraction technique produces appropriate, context-sensitive labels relevant to these projects, providing fresh insight into their evolving software development activities.
 \end{abstract}

### 2010-08-20 icse/icse2011.tex
more abstract

@@
 Researchers have used topic modeling and concept location to understand the latent topics of software development artifacts.
-These techniques use unsupervised machine-learning algorithms recover topics.
+These techniques use unsupervised machine-learning algorithms to recover topics.
 These topics are word-lists and are difficult to distinguish and interpret. 
@@
 Topics are extracted using Latent Dirichlet Allocation (LDA) from commit-log comments extracted from source control systems such as CVS and BitKeeper. 
-Then these topics given labels relating to a generalizable cross-project taxonomy consisting of non-functional requirements (NFRs).
+These topics are given labels relating to a generalizable cross-project taxonomy consisting of non-functional requirements.
 Our approach was evaluated  with experiments and case studies on two large-scale RDBMS projects: MySQl and MaxDB.

### 2010-08-20 icse/icse2011.tex
Fix the figure
viz discussion

@@
 
-\begin{figure}[]
+\begin{figure*}[]
 \centering
@@
 }
-
 \subfloat[MaxDB]{
@@
 \label{fig:mulan}
-\end{figure}
+\end{figure*}
 
@@
 
-\subsection{Timeline visualization}
+\subsection{Timeline analysis}
 Figures \ref{fig:maxdb-timeline} and \ref{fig:mysql-timeline} show the temporal patterns of label frequencies. 
@@
 
-\textbf{Understanding the figures} -- For example, in Figure \ref{fig:mysql-timeline}, we see that the NFRs \emph{functionality}, \emph{portability} and \emph{maintainability} contain more labeled topics, since these NFRs have been more intensely shaded.
+%\noindent \textbf{Understanding the figures} -- 
+In Fig.
+\ref{fig:mysql-timeline}, we see that the NFRs \emph{functionality}, \emph{portability} and \emph{maintainability} contain more labeled topics, since these NFRs have been more intensely shaded.
 % boxes.
@@
 
-\textbf{Timeline of key events} -- As discussed earlier, the top row in each figure shows key events for each project.
+%\noindent \textbf{Timeline of key events} -- 
+As discussed earlier, the top row in each figure shows key events for each project.
 \textit{Labelled topic extraction} can pick out the underlying NFR activity behind these events.
@@
 
-\subsection{Comparing MaxDB and MySQL}
+% \subsection{Comparing MaxDB and MySQL}
 
-\label{sec:comparison}
+% \label{sec:comparison}
 %\XXX{Not sure there's enough here to do that!}
-
-We observed that MySQL had more topic subsets than MaxDB. 
-MySQL 3.23 also had more topics and a longer recorded history than MaxDB 7.500. 
-We tagged more MySQL topics with annotations than MaxDB topics, yet both shared similarities. 
-In terms of non-functional requirements both projects had long running trends that focused on \emph{functionality}, \emph{maintainability}, and \emph{portability}, yet MaxDB had more of a focus on \emph{efficiency} and \emph{reliability}.
-\todo[inline]{expand}
+% We observed that MySQL had more topic subsets than MaxDB. 
+% MySQL 3.23 also had more topics and a longer recorded history than MaxDB 7.500. 
+% We tagged more MySQL topics with annotations than MaxDB topics, yet both shared similarities. 
+% In terms of non-functional requirements both projects had long running trends that focused on \emph{functionality}, \emph{maintainability}, and \emph{portability}, yet MaxDB had more of a focus on \emph{efficiency} and \emph{reliability}.
+% \todo[inline]{expand}
 %\todo[inline,color=yellow]{expand on this with dates, and reference to visualizations}
 
-MySQL and MaxDB's machine learners made some decisions based off a few shared words: \textsf{bug, code, compiler, database, HP UX, delete, memory, missing, problems, removed, add, added, changed, problem, and test}.
-Adding these words to the word-lists of \textsf{exp2} and \textsf{exp3} could improve performance while ensuring they were only domain specific.
+This analysis allows to address the questions raised in section \ref{sec:suplabelling}:
 
-With respect to the questions raised in section \ref{sec:suplabelling}:
-
-\textbf{Do label frequencies change over time?} -- Yes, MySQL's label
-frequencies decreased as that release aged. \emph{Usability} and \emph{reliability}
-labels became more and more infrequent as it matured. \emph{Maintainability} topics
-became more prevalent as MaxDB matured.
+\textbf{Do label frequencies change over time?} -- In both projects the frequencies generally decreased with age. 
+However, there are variations within our NFR labels. In MySQL, \emph{usability} and \emph{efficiency} do not appear very often in topics. 
+A proportionately smaller number of commits addressed these NFRs.
+As discussed in the section previous, certain peaks in topic numbers coincide with particular emphases in the development team, such as new releases or bug fixes.
+This suggests that maintenance activity is not necessarily strictly decreasing with time, but rather episodic and responsive to outside stimuli. 
+In MaxDB, we can observe that \emph{Maintainability} topics became more prevalent as MaxDB matures. This is likely due to our analysis time horizon being shorter than the MySQL product.
 
@@
 
-\section{Limitations}
+\section{Discussion}
 \label{sec:limit}
@@
 \textsf{exp2} and \textsf{exp3} both indicate that static information can be used to help label topics without any training whatsoever. 
+MySQL and MaxDB's machine learners made some decisions based off a few shared words: \textsf{bug, code, compiler, database, HP UX, delete, memory, missing, problems, removed, add, added, changed, problem, and test}.
+Adding these words to the word-lists of \textsf{exp2} and \textsf{exp3} could improve performance while ensuring they were only domain specific.
+
 If the techniques used in \textsf{exp2} and \textsf{exp3} were combined with the supervised techniques we could probably reduce the training effort by boosting training sets with topics classified with the unsupervised techniques.
@@
 During the case study we manually labelled hundreds of topics by hand then used these annotations to validate the positive performance of topic labelling techniques.
-We showed that topics with labels are more useful than unlabelled topics. To furnish stakeholders with labelled topics, our technique provides many methods of automatic topic labelling.
+The implication of our paper is that topics with labels are more useful than unlabelled topics.
+To furnish stakeholders with labelled topics, our technique provides many methods of automatic topic labelling.
 

### 2010-08-20 icse/icse2011.tex
more on comparison

@@
 This suggests that maintenance activity is not necessarily strictly decreasing with time, but rather episodic and responsive to outside stimuli. 
-In MaxDB, we can observe that \emph{Maintainability} topics became more prevalent as MaxDB matures. This is likely due to our analysis time horizon being shorter than the MySQL product.
+In MaxDB, we can observe that \emph{Maintainability} topics became more prevalent as MaxDB matures.
+This is likely due to our analysis time horizon being shorter than equivalent span for the MySQL product.
 
@@
 topics labelled \emph{functionality}, while MaxDB had proportionally more
-\emph{efficiency} and \emph{portability} related topics.
- In both cases portability was a constant maintenance concern and was prevalent throughout the entire lifetime of the projects.
-\todo[inline]{fix}
+\emph{efficiency} related topics. MaxDB was a very mature release `donated' to the open-source community. 
+MySQL, on the other hand, was in its relative infancy. 
+Security problems were more common (security is a component of `functionality' in the ISO 9126 model). 
+In both cases \emph{portability} was a constant maintenance concern and was prevalent throughout the entire lifetime of the projects. 
+% \todo[inline]{fix}
 

### 2010-08-20 icse/icse2011.tex
more edits of comparison

--- 2010-08-20 icse/icse2011.tex
@@
 Version 3.23.31 (January, 2001) was the production release (non-beta), and we see a flurry of topics labeled with \emph{functionality} and \emph{maintainability}, as one might expect.
-After this point, this version enters the maintenance phase of the lifecycle.
+After this point, this version enters the maintenance phase of its lifecycle.
 In May 2001, there is an increase in the number of topics labelled with \emph{portability}.
 This might be related to release 3.23.38, which focused on Windows compatibility.
-Similarly, in Augsut, 2002, both \emph{functionality} and \emph{portability} are high in our tool, and mailing list data suggests this is related to the release of version 3.23.52, a general bug fix with focus on security (encoded as the \emph{functionality} NFR in the ISO9126 model).
+Similarly, in Augsut, 2002, both \emph{functionality} and \emph{portability} are frequent, and mailing list data suggests this is related to the release of version 3.23.52, a general bug fix with a focus on security (encoded as the \emph{functionality} NFR in the ISO9126 model).
 After this point, efforts shift to the newer releases (4.0, 4.1, 5.0).
@@
 
-This analysis allows to address the questions raised in section \ref{sec:suplabelling}:
+This analysis allows to address the questions raised in Section \ref{sec:suplabelling}:
 
@@
 A proportionately smaller number of commits addressed these NFRs.
-As discussed in the section previous, certain peaks in topic numbers coincide with particular emphases in the development team, such as new releases or bug fixes.
+As discussed in the previous section, certain peaks in topic numbers coincide with a particular emphasis from the development team on issues such as new releases or bug fixes.
 This suggests that maintenance activity is not necessarily strictly decreasing with time, but rather episodic and responsive to outside stimuli. 
-In MaxDB, we can observe that \emph{Maintainability} topics became more prevalent as MaxDB matures. This is likely due to our analysis time horizon being shorter than the MySQL product.
+In MaxDB, we can observe that \emph{Maintainability} topics became more prevalent as MaxDB matures. 
+This is likely due to our analysis time horizon being shorter than the MySQL product.
 

### 2010-08-21 icse/icse2011.tex
threats to validity read and edit

@@
  In both cases portability was a constant maintenance concern and was prevalent throughout the entire lifetime of the projects.
-\todo[inline]{fix}
+% XXX fix?
+%\todo[inline]{fix}
 
@@
 
-If the techniques used in \textsf{exp2} and \textsf{exp3} were combined with the supervised techniques we could probably reduce the training effort by boosting training sets with topics classified with the unsupervised techniques.
+If the techniques used in \textsf{exp2} and \textsf{exp3} were combined with the supervised techniques we could reduce the training effort by boosting training sets with topics classified with the unsupervised techniques.
 Both Naive Bayesian learners and the word-list approaches were computationally efficient.  
@@
 % can always be cut down
+Our work faced multiple threats to validity and we attempted to address many of them.
 
@@
 Conversely, construct validity was addressed in our validation as we used the files, topics and revisions to annotate the commit messages. 
+We used mailing-lists to verify the purpose behind some events.
 We also chose our taxonomy and the data to study. We did not show these results to the stakeholders.
 
-\emph{Internal validity} -- include inter-rater reliability and our lack of rival explanation building. 
+\emph{Internal validity} -- includes inter-rater reliability and our lack of rival explanation building. 
 We improved internal validity by trying to correlate and explain the behaviours observed in the analysis with the historical records of the projects.
@@
 
-%\todo[inline]{update this with more detail}
-
 \section{Related work}
@@
 
-The idea of extracting higher-level `concerns', also known as `concepts', `aspects' or `requirements', has been approached from documentation-based and repositories-based perspectives.
+The idea of extracting higher-level `concerns', also known as `concepts', `aspects' or `requirements', has been approached from documentation-based and repository-based perspectives.
 
@@
 %One approach they tried was similar to this one, with keywords mined from NFR catalogues found in their previous work~\cite{chung99}. 
-They managed recall of 80\% with precision of 57\% for the Security NFR, but could not find a reliable source of keywords for other NFRs. 
+They managed a recall of 80\% with precision of 57\% for the Security NFR, but could not find a reliable source of keywords for other NFRs. 
 Instead, they developed a supervised classifier by using human experts to identify an NFR training set. 

### 2010-08-21 icse/icse2011.tex
conclusions and future work editted

@@
 Our worked differed because we had a more comprehensive set of terms based on the taxonomy we chose.
-We also wanted a common taxonomy that allows us to make cross-project comparison, which was not one of their objectives.
+We also wanted a common taxonomy that allows us to make cross-project comparison; this was not one of their objectives.
 The objective of Cleland-Huang's study was to identify new NFRs for system development, yet our objective was to recover those latent NFRs from commit-log messages of the project. 
-Another difference was that our data was far less structured that the requirements documents used in their study.
+Another difference was that our data was far less structured than the requirements documents used in their study.
 %There are several reasons we did not follow this route. 
@@
 In the same vein, Mockus and Votta~\cite{Mockus00} studied a large-scale industrial change-tracking system. 
-They also leveraged WordNet, but only for word roots. 
-They felt the synonyms would be non-specific and cause errors. 
-A nice contribution was access to system developers, with whom they could validate their labels. 
-Since we try to bridge different organizations, these interviews are infeasible (particularly in the distributed world of open-source software).
+They also leveraged WordNet, but only for word roots as they felt the synonyms would be non-specific and cause errors. 
+Mockus et al. had access to system developers, with whom they could validate their labels. 
+Since we try to bridge different organizations, these kind of interviews are infeasible (particularly in the distributed world of open-source software).
 
@@
 
-% % Not sure we need this
-% Mens et al.~\cite{mens08icsm} conducted an empirical study of Eclipse, a popular Java IDE, to verify the claims of Lehman~\cite{lehman97sms}.
-% % They concerned themselves with source code only, and found Law Seven, ``Declining Quality'', to be too difficult to assess: ``[we lacked an] appropriate measurement of the evolving quality of the system as perceived by the users \cite[p.
+%XXXXXXXXXXXX OPTIONAL
+% Not sure we need this
+Mens et al.~\cite{mens08icsm} conducted an empirical study of Eclipse, a popular Java IDE, to verify the claims of Lehman~\cite{lehman97sms}.
+% They concerned themselves with source code only, and found Law Seven, ``Declining Quality'', to be too difficult to assess: ``[we lacked an] appropriate measurement of the evolving quality of the system as perceived by the users \cite[p.
 388]{mens08icsm}''.
-% This paper examines the notions of quality in terms of a consistent ontology, as Mens
+This paper examines the notions of quality in terms of a consistent ontology, as Mens
 et al. call for in their conclusions.
@@
 % XXX vanity or space.. 
-% Hindle et al.
+%XXXXXXXXXXXX OPTIONAL
+ Hindle et al.
 \cite{Hindle2007} examined release patterns in OSS.
@@
 
+%XXXXXXXXXXXXXXXX I think we need to suggest a couple of uses.
+
 % para 4 summarize and hammer contribution
@@
 We provided a case study and visualization of NFR related topics in MaxDB 7.500 and MySQL 3.23. 
-During the case study we manually labelled hundreds of topics by hand then used these annotations to validate the positive performance of topic labelling techniques.
+During the case study we manually labelled hundreds of topics by hand and then used these annotations to validate the positive performance of topic labelling techniques.
 The implication of our paper is that topics with labels are more useful than unlabelled topics.
@@
 
-%\subsection{Future Work}
+%XXXXXXXX Optional  
+\subsection{Future Work}
 
@@
 We also think multi-label learning techniques, although in their infancy, are crucial in understanding cross-cutting concerns such as NFRs. 
+We also want to leverage different kinds of artifacts to discover threads of NFR-related discussions that occur between multiple kinds of artifacts.
 Finally, we would like to extend this analysis to other domains, to see what patterns might occur in, for example, an end-user facing software product.

### 2010-08-21 icse/icse2011.tex

@@
 In MaxDB, we can observe that \emph{Maintainability} topics became more prevalent as MaxDB matures. 
-This is likely due to our analysis time horizon being shorter than the MySQL product.
+This is likely due to our analysis time horizon being shorter than equivalent span for the MySQL product.
 
@@
 topics labelled \emph{functionality}, while MaxDB had proportionally more
-\emph{efficiency} and \emph{portability} related topics.
- In both cases portability was a constant maintenance concern and was prevalent throughout the entire lifetime of the projects.
-% XXX fix?
-%\todo[inline]{fix}
+\emph{efficiency} related topics. MaxDB was a very mature release `donated' to the open-source community. 
+MySQL, on the other hand, was in its relative infancy. 
+Security problems were more common (security is a component of `functionality' in the ISO 9126 model). 
+In both cases \emph{portability} was a constant maintenance concern and was prevalent throughout the entire lifetime of the projects.
 

### 2010-08-21 icse/icse2011.tex
more edits...
heh 10 pages
more abstract

@@
 We propose a solution: \emph{labelled topic extraction}.
-Topics are extracted using Latent Dirichlet Allocation (LDA) from commit-log comments extracted from source control systems such as CVS and BitKeeper. 
+Topics are extracted using Latent Dirichlet Allocation (LDA) from commit-log comments recovered from source control systems such as CVS and BitKeeper. 
 These topics are given labels relating to a generalizable cross-project taxonomy consisting of non-functional requirements.
@@
 In MaxDB, we can observe that \emph{Maintainability} topics became more prevalent as MaxDB matures. 
-This is likely due to our analysis time horizon being shorter than equivalent span for the MySQL product.
+This is likely due to our analysis time horizon being shorter than an equivalent span for the MySQL product. %XXXXXXXXXXXXXXXXX what
 
@@
 %XXXXXXXX Optional  
-\subsection{Future Work}
+%\subsection{Future Work}
 

### 2010-08-21 icse/icse2011.tex
intro edits hard

@@
 % why are topics useful
-Developers often interact with abstractions of the artifacts from which topics are derived, rather than the artifact itself. 
+Developers often interact with abstractions of the artifacts from which topics are derived, rather than the artifacts themselves.
 Such abstractions include project dashboards~\cite{kersten2005mylar}, tags~\cite{treude2010}, and complexity metrics \cite{mccabe1976complexity}. 
@@
 Much like aspect-oriented programming, topic labeling can identify cross-cutting concerns in a software project (as in~\cite{Baldi2008}).
-The topic of development is relevant because it identifies the distinct issues facing development at a particular time.
+The topics of development are relevant because they identify the distinct issues facing development at a particular time.
 
@@
 % get the interpretation aspect
-Much of the current research stops short of interpreting the purpose of the topic. 
+Much of the current research stops short of automatically interpreting the purpose of the topic. 
 That is, extracted topics are usually manually named and labelled by those who read the topic words.
@@
 Deriving such topics is difficult. 
-Topic analysis produces% raw topics that are represented as
+Topic analysis produces % raw topics that are represented as
  unlabelled context-free word-lists or word distributions.
@@
 In general, the mining of software artifacts tends to be very project specific, yet NFRs are not. 
-There is a series of standards on NFRs (such as \cite{iso9126}) specifically intended to apply to projects of varying types (indicating this is possible, if perhaps not universally accepted).
+% I was told if you're going to use a paren, you don't need to say it.
+There is a series of standards on NFRs, such as \cite{iso9126}, specifically intended to apply to projects of varying types.
+These standards indicate our approach is possible.%, if perhaps not universally accepted).
 % NFRs are prevalent in almost all software systems.
 % In this paper we will show that NFR related topics occur quite frequently.
+%XXXXXX this sentence is confusing
 This informs our method of generalizing the data mining of these software artifacts to leverage the global domain knowledge of software engineering, specifically NFRs. 
@@
 Topic modelling can be used to check if development is on track and focused.
-For outside users of the software, our technique can provide insight into that company's development practices. 
-Are they focused on usability, for example?
-Within the organization \emph{labelled topic extraction} acts as a historical sanity check. 
-\emph{Confirmation bias} is the difference between what people think happened and what actually happened. 
+For outside users of the software, our technique can provide insight into that company's development practices in order to answer questions such as, ``Are they focused on usability?''
+Within the organization \emph{labelled topic extraction} can act as a historical sanity check, used to avoid \emph{confirmation bias}, which is the difference between what people think happened and what actually happened.
 Some software releases, for instance, are poorly received since they are quite buggy. 
@@
 \textit{Labelled topic extraction} allows for comparisons \textit{between} projects. 
-  \item Topic modeling creates word lists that require interpretation by the user to assign meaning. Like (1), this means
-that it is difficult to discuss results independent of the project context. Our technique automatically assigns labels across projects.
+  \item Topic modeling creates word lists that require interpretation by the user to assign meaning. 
+Like (1), this means
+that it is difficult to discuss results independent of the project context. 
+Our technique automatically assigns labels across projects.
 \end{enumerate}

### 2010-08-21 icse/icse2011.tex
generating the data
minor edits

@@
 Much like aspect-oriented programming, topic labeling can identify cross-cutting concerns in a software project (as in~\cite{Baldi2008}).
-The topics of development are relevant because they identify the distinct issues facing development at a particular time.
+The topics of development are relevant because they identify the distinct issues facing stakeholders at a particular time.
 
@@
 % get the interpretation aspect
-Much of the current research stops short of automatically interpreting the purpose of the topic. 
+Much of the current research stops short of automatically interpreting the purpose of the topic, e.g.~\cite{Baldi2008,Kuhn2007,marcus04wcre}. 
 That is, extracted topics are usually manually named and labelled by those who read the topic words.
@@
 % topics are numerous?
-Deriving such topics is difficult. 
+Deriving and labelling topics is a challenging machine learning problem. 
 Topic analysis produces % raw topics that are represented as
@@
 Topics have to be identified by interpreting the prevalent words in a word distribution and by inspecting related documents. 
-This manual approach is impractical when one has to handle hundreds of different topics. To scale to meaningfully sized projects, we need automatic assistance to determine the topic label.
+This manual approach is impractical when handling hundreds of different topics.
+To scale to large projects (thousands of lines of code and multiple releases), we need automatic assistance to determine the topic label.
 We propose a method of interpreting and labelling topics, and their related artifacts, that is generalizable and relevant to multiple projects.
@@
 %XXXXXX this sentence is confusing
-This informs our method of generalizing the data mining of these software artifacts to leverage the global domain knowledge of software engineering, specifically NFRs. 
+%This informs our method of generalizing the data mining of these software artifacts to leverage the global domain knowledge of software engineering, specifically NFRs. 
 In particular, we use software quality models to generalize topics across projects. 
@@
 MySQL contains 320 thousand lines of C and C++ source code.  
-Choosing an older version of these projects allows us to focus on projects which have moved into the maintenance phase of the software life-cycle.
+Choosing an older version of these projects allowed us to focus on projects which have moved into the maintenance phase of the software life-cycle.
 
@@
 For the commit message data-sets of each project, we created an XML file which separated commits into 30 day periods. 
-This size of period, 30 days, is smaller than the time between minor releases but large enough for there to be sufficient commits to analyze. 
-For each period of each project, we input the messages of that period into Latent Dirichlet Allocation (LDA), a topic analysis algorithm~\cite{Blei2003}, and recorded the topics of each period.
+This size of period, 30 days, is smaller than the time between minor releases but large enough for there to be sufficient commits to analyze~\cite{Hindle09ICSM}. 
+For each 30 day period of each project, we input the messages of that period into Latent Dirichlet Allocation (LDA), a topic analysis algorithm~\cite{Blei2003}, and recorded the topics the algorithm returned.
 %For each project and for each period of messages a project, we input these messages 
@@
 We applied Blei's LDA implementation~\cite{Blei2003} against the word distributions of these commits, and generated lists of topics per period. 
-We set the number of topics to generate to $20$, because past experimentation, e.g.
-\cite{Hindle09ICSM}, showed that fewer topics might aggregate multiple unique topics while any more topics seemed to dilute the results and create indistinct topics.
-Furthermore, more than $20$ topics quickly became infeasible for inspection.
+We set the number of topics to generate to $20$, because past experimentation showed that fewer topics might aggregate multiple unique topics while any more topics seemed to dilute the results and create indistinct topics~\cite{Hindle09ICSM}.
+% Furthermore, more than $20$ topics quickly became infeasible for inspection.
 
@@
 In MaxDB, we can observe that \emph{Maintainability} topics became more prevalent as MaxDB matures. 
-This is likely due to our analysis time horizon being shorter than an equivalent span for the MySQL product. %XXXXXXXXXXXXXXXXX what
+This is likely due to our analysis time horizon being shorter than an equivalent span for the MySQL product. 
+%%XXXXXXXXXXXXXXXXX what
 
@@
 During the case study we manually labelled hundreds of topics by hand and then used these annotations to validate the positive performance of topic labelling techniques.
-The implication of our paper is that topics with labels are more useful than unlabelled topics.
+The implication of our paper is that topics with labels are more useful for analyzing software development activities than are unlabelled topics.
 To furnish stakeholders with labelled topics, our technique provides many methods of automatic topic labelling.
@@
 % \todo[inline]{it should be labels for topics not artifacts, I think the artifact part is almost secondary and the topic stuff is really what is up.
-Topics are used below and it isn't even clear what they are or what we're talking about.
+Topics are used below and it isn't even clear what they are or what we're talking aout.
 Naming is important but lets tell the reader what is being named first.}

### 2010-08-21 icse/icse2011.tex
unsupervised
supervised labelling edits

--- 2010-08-21 icse/icse2011.tex
@@
 In this paper a topic is a set of tokens extracted from commit messages found within a project's source control system (SCS).
-%\todo[inline]{These paragraphs should've come way earlier, in previous we should explain LDA so this doesn't need to be here.}
 
@@
 MySQL had reasonable results within \textsf{exp2} for \emph{reliability} and \emph{efficiency}. 
-MySQL's results for efficiency did not improve in \textsf{exp3} but other qualities such as \emph{functionality} did improve. 
+MySQL's results for \emph{efficiency} did not improve in \textsf{exp3} but other qualities such as \emph{functionality} did improve. 
 If a \emph{C} grade performance has a ROC value of $0.6$ then most of these tests scored a grade of \emph{C} or less, but our classifier nonetheless performed substantially better than random.
@@
 %This is repeated nine more times, with each subsample used once as the training set. 
-We have reported these results below in Section \ref{sec:suplabelling}.
+We have reported these results below.% in Section \ref{sec:suplabelling}.
 
@@
 For both projects Bayesian techniques did the best out of a wide variety of machine learners tested. 
-Discriminative Multinomial Naive Bayes (DMNBtext), Naive Bayes (NaiveBayes) and Multinomial Naive Bayes (NaiveBayesMultinomal) are all based on Bayes's theorem and all assume, naively, that the features are independent.
+Our best learners, Discriminative Multinomial Naive Bayes, Naive Bayes  and Multinomial Naive Bayes  are all based on Bayes's theorem and all assume, naively, that the features are independent. 
 The features we used are word counts per message. 

### 2010-08-21 icse/icse2011.tex
annotation observations
summary of techniques
understanding software ...

@@
 For example, a commit message might address \textit{reliability} in the context of \textit{efficiency}, or make a \textit{maintainability} improvement in the source code that related to \textit{usability}.
-However, traditional machine learning techniques, such as naive Bayes, can only map topics to a single class. 
+However, traditional machine learning techniques, such as Naive Bayes, can only map topics to a single class. 
 The Mulan~\cite{mulan} library encapsulates several different multi-label machine learners which can label elements with multiple labels.
@@
 % although their results are comparable to the naive Bayesian learners we used in Section \ref{sec:suplabelling}.
-%\todo[inline]{discuss this similarity in more detail}
+% \todo[inline]{discuss this similarity in more detail}
 
@@
 \begin{enumerate}
-\item Do label frequencies change over time? That is, is a certain quality of more interest at one point in the life-cycle than some other? 
-\item  Do the different projects differ in their relative topic interest? That is, is a particular quality more important to one project than the other projects?  
+\item Do label frequencies change over time? Is a certain quality of more interest at one point in the life-cycle than some other? 
+\item  Do the different projects differ in their relative topic interest? Is a particular quality more important to one project than the other projects?  
 \end{enumerate}

### 2010-08-21 icse/icse2011.tex
time-line analysis

@@
 relative to the maximum number of labelled topics overall, for that project. 
-The topmost row in the diagrams is reserved for historical events of note.
+The topmost row in each the diagram is reserved for historical events of note. 
 We refer to these occurrences in the discussion which follows.
@@
 %\noindent \textbf{Understanding the figures} -- 
-In Fig.
-\ref{fig:mysql-timeline}, we see that the NFRs \emph{functionality}, \emph{portability} and \emph{maintainability} contain more labeled topics, since these NFRs have been more intensely shaded.
+In Figure \ref{fig:mysql-timeline}, we see that the NFRs \emph{functionality}, \emph{portability} and \emph{maintainability} contain more labeled topics, since these NFRs have been more intensely shaded.
 % boxes.
@@
 From analyzing the mailing list for external validation, we know that release 7.5.00.23 saw the development of PHP interfaces, possibly accounting for the simultaneous increase in the \emph{portability} NFR at the same time.
-The gap in MaxDB (Fig. \ref{fig:maxdb-timeline}) is due to a shift in development focus (from February 2005 to Jun 2005) to MaxDB 7.6, which is released in June 2005.
-
-For MySQL (Fig.
-\ref{fig:mysql-timeline}), we similarly validated our NFR patterns with external mailing list data.
+The gap in MaxDB (Figure \ref{fig:maxdb-timeline}) is due to a shift in development focus (from February 2005 to Jun 2005) to MaxDB 7.6, which is released in June 2005.
+
+For MySQL (Figure \ref{fig:mysql-timeline}), we similarly validated our NFR patterns with external mailing list data. 
 This release of MySQL was the first to be licenced under the GPL. 
-Version 3.23.31 (January, 2001) was the production release (non-beta), and we see a flurry of topics labeled with \emph{functionality} and \emph{maintainability}, as one might expect.
-After this point, this version enters the maintenance phase of its lifecycle.
+Version 3.23.31 (January, 2001) was the production release (non-beta), and we see a flurry of topics labelled with \emph{functionality} and \emph{maintainability}, as one might expect. 
+After this point, this version enters the maintenance phase of its life-cycle. 
 In May 2001, there is an increase in the number of topics labelled with \emph{portability}. 
 This might be related to release 3.23.38, which focused on Windows compatibility. 
-Similarly, in Augsut, 2002, both \emph{functionality} and \emph{portability} are frequent, and mailing list data suggests this is related to the release of version 3.23.52, a general bug fix with a focus on security (encoded as the \emph{functionality} NFR in the ISO9126 model).
+Similarly, in August, 2002, both \emph{functionality} and \emph{portability} are frequent, and mailing list data suggests this is related to the release of version 3.23.52, a general bug fix with a focus on security (encoded as the \emph{functionality} NFR in the ISO9126 model).
 After this point, efforts shift to the newer releases (4.0, 4.1, 5.0).

### 2010-08-21 icse/icse2011.tex
appendix
Do label frequencies ...

@@
 In MaxDB, we can observe that \emph{Maintainability} topics became more prevalent as MaxDB matures. 
-This is likely due to our analysis time horizon being shorter than an equivalent span for the MySQL product. %XXXXXXXXXXXXXXXXX what
+This is likely due to our analysis time horizon being shorter than an equivalent span for the MySQL project. %XXXXXXXXXXXXXXXXX what
 
@@
 
-Our data and scripts are available at \url{http://softwareprocess.es/nomen/}
+Our data and scripts are available at:
+
+ \url{http://softwareprocess.es/nomen/}
 

### 2010-08-21 icse/icse2011.tex

@@
 In MaxDB, we can observe that \emph{Maintainability} topics became more prevalent as MaxDB matures. 
-This is likely due to our analysis time horizon being shorter than an equivalent span for the MySQL project. %XXXXXXXXXXXXXXXXX what
+This is likely due to our analysis timeframe for MaxDB being shorter than the timeframe for the MySQL product.
 

### 2010-08-21 src/validate/plot_timeline.py
removed edge colors from bars; removed numbers.

@@ -91,21 +91,24 @@ def plot_timeline(proj, period_map):
         # relative to all
         # loop over it instead of copy and paste 
         #HAHAHA! Copy and paste rules. Code clones rule. Code smells rule.
+        delta = datetime.timedelta(days=5)
+        i = 0
         for value in value_order:
             #rect = Rectangle((date,value_start[value]), barwidth,barheight,fill=True,lw=0, fc='red',alpha=(float(qm[ value ])/float(max_value_map[ value ])))
-            rect = Rectangle((date,value_start[value]), barwidth,barheight,zorder=2,fill=True,lw=0, fc='red',alpha=(float(qm[ value ])/float(denominator)))
+            intensity = (float(qm[ value ])/float(denominator))
+            # if intensity > 0:
+            #     pass#intensity = intensity + 0.2 # the lowest alpha possible
+            # if intensity >= 1: intensity = 1 
+            rect = Rectangle((date,value_start[value]), barwidth,barheight,zorder=2,fill=True,lw=0, fc='red',alpha=intensity)
             ax.add_patch( rect )
             # now draw the boxes relative the stream itself
             # 0.49 is for half height
             h = 0.49 * barheight * float(qm[ value ]) / float( max_value_map[ value ])
-            vrect = Rectangle((date,value_start[value]), barwidth, h,fill=True,zorder=3,lw=1, fc='gray',)#alpha=(float(qm[ value ])/float(max_value_map[ value ])))
+            vrect = Rectangle((date,value_start[value]), barwidth, h,fill=True,zorder=3,lw=1, ec='gray',fc='gray',)#alpha=(float(qm[ value ])/float(max_value_map[ value ])))
             ax.add_patch( vrect )
         
-        #add text numbers to debug
-        delta = datetime.timedelta(days=5)
-        i = 0
-        for value in value_order:
-            ax.annotate(qm[value],  (date+delta, 2.5+5*i) , color='black', size='small')
+            #add text numbers to debug
+            #ax.annotate("%.2f" % intensity,  (date+delta, 2.5+5*i) , color='black', size='small') #qm[value]
             i += 1
         
     for i in range(len(period_map.keys()[0])):

### 2010-08-21 icse/icse2011.tex
minor for figure
final edits

@@
 There is a series of standards on NFRs, such as \cite{iso9126}, specifically intended to apply to projects of varying types.
-These standards indicate our approach is possible.%, if perhaps not universally accepted).
+These standards indicate our approach is possible. %, if perhaps not universally accepted).
 % NFRs are prevalent in almost all software systems.
@@
 This paper first introduces labelled topic extraction, and then uses that technique to analyze development activity in two large-scale examples. 
-We begin by describing how we generated our data (Section \ref{sec:wordlist}). 
+We begin by describing how we generated our data (Sections  \ref{sec:wordlist} and \ref{sec:genwordlist}). 
 We show that labels with their topics can be learned and used to classify other data-sets, either without training (Section \ref{sec:unsuplabelling}), or with training (Section \ref{sec:suplabelling}).
@@
 
-Topics are word distributions, lists of words ranked by frequency, that are be burdensome to interpret and hard to distinguish and understand. 
+Topics are word distributions, lists of words ranked by frequency, that are burdensome to interpret and hard to distinguish and understand. 
 While the topic models themselves are generated automatically, how to interpret these topics is less clear.
@@
 We did the same for the quality model produced by McCall et al. \cite{mccall1977}. 
-Finally, we analyzed two mailing lists from another software project, KDE, to expand our set with domain-specific terms. 
+Finally, we analyzed two mailing lists from another software project to expand our set with domain-specific terms. 
 For example, we add the term ``performance'' to the synonyms for \emph{efficiency}, since this term occurs in most mail messages that discuss efficiency.
@@
 
-\textbf{Analysis of the unsupervised labelling} -- Based on the labels, and our manual topic labelling, 
+\noindent \textbf{Analysis of the unsupervised labelling} -- Based on the labels, and our manual topic labelling, 
 we compared the results of the unsupervised word matching approach. 
-For each quality we assessed whether unsupervised labels matches the manual annotations.
+For each quality we assessed whether unsupervised labels matched the manual annotations.
 Figure \ref{fig:maxdb-unsup-results} shows our results for MaxDB and MySQL. 
@@
 A ROC result of 0.5 would be equivalent to a random learner (that is, issuing as many false positives as true positives). 
-The AUC of a classifier is equivalent to the probability that the classifier will rank a randomly chosen positive instance higher than a randomly chosen negative instance.
+The area under the ROC of a classifier is equivalent to the probability that the classifier will rank a randomly chosen positive instance higher than a randomly chosen negative instance.
 
@@
  \includegraphics[width=\textwidth]{figures/unsupervised-bar}
- \caption{Performance, ROC values, of unsupervised topic labelling for
+ \caption{Performance, ROC values (range: 0-1), of unsupervised topic labelling for
    each NFR and per word-list. This graph shows how well the
@@
 Supervised labelling requires expert analysis of the correct class/label to assign to a topic.
-In our approach, we use the ISO9126 standard~\cite{iso9126} for our classes, but other taxonomies are also applicable.%In order to validate how effective these word-bag approaches to topic labelling would be we created a validation data set.
+In our approach, we use the top-level NFRs in the ISO 9126 standard~\cite{iso9126} for our classes, but other taxonomies are also applicable.%In order to validate how effective these word-bag approaches to topic labelling would be we created a validation data set.
 
@@
 WEKA includes machine learning tools such as support vector machines and Bayes-nets. 
-We also used the multi-labelling add-on for WEKA, Mulan~\cite{mulan}\footnote{\url{http://mlkd.csd.auth.gr/multilabel.html}}. 
+We also used the multi-labelling add-on for WEKA, Mulan~\cite{mulan}. %\footnote{\url{http://mlkd.csd.auth.gr/multilabel.html}}. 
 Traditional classifiers map our topics to a single class, whereas Mulan allows for a mixture of classes per topic, which maps to what we observed while manually labelling topics.
@@
 To assess the performance of the supervised learners, we did a 10-fold cross-validation~\cite{Kohavi1995}, a common technique for evaluating machine learners. 
-The original data is partitioned randomly into ten sub-samples. Each sample is used to test against a train set composed of the nine other samples.
+The original data is partitioned randomly into ten sub-samples. Each sample is used to test against a training set composed of the nine other samples.
 %One subsample is used as a training set and evaluated on the other nine sub-samples. 
@@
 independent; however, labels could be correlated with each other. 
-The next section, Section \ref{sec:multilabel} on multi-labels
+The next section (\ref{sec:multilabel})
 addresses the issue of a lack of independence and correlation between labels.
@@
 % }
-\caption[]{ROC value for the best learner per label for MaxDB and MySQL. ROC values range from 0--1.
+\caption[]{ROC value for the best learner per label for MaxDB and MySQL. Values range from 0--1.
 %\todo[inline]{better discussion}
@@
 A macro-ROC measurement is the average ROC over the ROC values for all labels, where a micro-ROC is the average ROC over all examples that were classified. 
-Unfortunately for MaxDB, the macro-ROC values are undefined because of poor performance of one of the labels.%, see Figure \ref{fig:mulan}.
+For MaxDB, the macro-ROC values are undefined because of poor performance of one of the labels.%, see Figure \ref{fig:mulan}.
 
@@
 We found many topics that were not non-functional requirements (NFRs) but were often related to them. 
-For instance, concurrency was mentioned often in the commit logs and was related to \emph{correctness} and \emph{reliability}, likely because concurrent code is prone to bugs such as race conditions.%are
+For instance, concurrency was mentioned often in the commit logs and was related to \emph{correctness} and \emph{reliability}, likely because concurrent code is prone to bugs such as race conditions.
+%are
 % often  was troublesome. 
-\emph{Configuration management} and \emph{source control} related changes appeared often and sometimes there were topics dedicated to configuration management. 
-These kinds of changes are slightly related to \emph{maintainability}. 
+\emph{Configuration management} and \emph{source control} related changes appeared often; % and sometimes there were topics dedicated to configuration management. 
+these kinds of changes are slightly related to \emph{maintainability}. 
 A non-functional change that was not quality-related was licensing and copyright; many changes were simply to do with updating copyrights or ensuring copyright or license headers were applied to files.
+In these cases we assigned the \emph{None} class.
 
@@
 supervised learners have the advantage of domain knowledge, which typically means improved results. 
-Our manual annotations were fairly quick to do, taking on average five minutes per 30 day period.
-
-Very rarely did \textsf{exp2} and \textsf{exp3} (naïve word matching) ever perform as well as the supervised machine learners. 
+Our manual annotations were fairly quick to do, taking only a few minutes per 30 day period.
+
+Very rarely did \textsf{exp2} and \textsf{exp3} (naive word matching) ever perform as well as the supervised machine learners. 
 For MaxDB, \textit{reliability} was slightly better detected using the static word list of \textsf{exp2}. 
@@
 \subsection{Timeline analysis}
-Figures \ref{fig:maxdb-timeline} and \ref{fig:mysql-timeline} show the temporal patterns of label frequencies. 
+Figures \ref{fig:mysql-timeline} and \ref{fig:maxdb-timeline} show the temporal patterns of label frequencies. 
 There are two measures represented. 
@@
 relative to the maximum number of labelled topics overall, for that project. 
-The topmost row in each the diagram is reserved for historical events of note. 
+The topmost row in each diagram is reserved for historical events of note. 
 We refer to these occurrences in the discussion which follows.
@@
 This release of MySQL was the first to be licenced under the GPL. 
-Version 3.23.31 (January, 2001) was the production release (non-beta), and we see a flurry of topics labelled with \emph{functionality} and \emph{maintainability}, as one might expect. 
+Version 3.23.31 (January, 2001) was the production release (non-beta), and we see a flurry of topics labelled with \emph{functionality} and \emph{maintainability}. 
 After this point, this version enters the maintenance phase of its life-cycle. 
@@
 This might be related to release 3.23.38, which focused on Windows compatibility. 
-Similarly, in August, 2002, both \emph{functionality} and \emph{portability} are frequent, and mailing list data suggests this is related to the release of version 3.23.52, a general bug fix with a focus on security (encoded as the \emph{functionality} NFR in the ISO9126 model).
+Similarly, in August, 2002, both \emph{functionality} and \emph{portability} are frequent, and mailing list data suggests this is related to the release of version 3.23.52, a general bug fix with a focus on security (a component of the \emph{functionality} NFR in the ISO9126 model).
 After this point, efforts shift to the newer releases (4.0, 4.1, 5.0).
@@
 }
-	\caption[]{NFR label per period.
-Each box represents a 30-day period.
-Numbers represent number of topics in that period labelled with that NFR.
-Box intensity (saturation) is mapped to label frequency relative to the largest label value of \emph{all} NFRs.
+	\caption[]{NFR label per period. Each box represents a 30-day period. %Numbers represent number of topics in that period labelled with that NFR. 
+	Grid cell intensity (saturation) is mapped to label frequency relative to the largest label value of \emph{all} NFRs.
 Grey bars indicate label frequency relative to \emph{a particular} NFR.
-Dashed vertical lines relate a project milestone (``*Key Event*'') to our topic windows.
+Dashed vertical lines relate a project milestone (*Key Event*) to our topic windows.
 % \todo[inline]{better discussion}
@@
 We validated our topic labelling techniques using multiple experiments. 
-We conducted unsupervised labelling using word-lists. 
+We first conducted unsupervised labelling using word-lists. 
 Our next approach was supervised, using single-label and multi-label learners. 
@@
 We also think multi-label learning techniques, although in their infancy, are crucial in understanding cross-cutting concerns such as NFRs. 
-We also want to leverage different kinds of artifacts to discover threads of NFR-related discussions that occur between multiple kinds of artifacts.
-Finally, we would like to extend this analysis to other domains, to see what patterns might occur in, for example, an end-user facing software product.
+We want to leverage different kinds of artifacts to discover threads of NFR-related discussions that occur between multiple kinds of artifacts.
+Finally, we would like to extend this analysis to other domains, to see what patterns might occur in, for example, a consumer-facing software product.
  % Talk to people

### 2010-12-15 msr/msr2011.tex
icse -> msr 2011
Added inline todos

--- 2010-08-21 icse/icse2011.tex
@@
 
-\conferenceinfo{ICSE}{'11 Waikiki, Hawaii, USA}
+\conferenceinfo{MSR}{'11 Waikiki, Hawaii, USA}
+
 \numberofauthors{5}
+
 \author{
@@
 \section{Introduction}
+
+\todo[inline]{Strength story: Neil says
+  And ultimately, I think the story needs more strengthening. Maybe we
+  need a scenario: you work at Bank X. You are curious about how much
+  recent activity has been done on software maintainability
+  vs. security. Our technique will allow you to compare different IT
+  projects using the same keywords, and ultimately, to generate lists
+  showing relative frequencies.}
+
+\todo[inline]{mention project dashboards}
+
+\todo[inline]{Usefulness needs to be immediate and apparent}
+
+\todo[inline]{ It needs to be clear there is value in labelling.
+  We need to make it clear that other researchers weren't labelling them
+ automatically and presented them as labelled and thus it is obvious from
+  their presentation that labels matters + need a reading argument. The
+ 2nd reviewer picked on this a bit but still didn't get it. I thought
+ some of this had been clear.}
+
+\todo[inline]{Topic should be defined clearly (italics and hilighted)}
+
+\todo[inline]{Move previous work up}
+
 Topics in software engineering practice refer to specific shared threads of discussions between project stakeholders. 
@@
 
+\todo[inline]{Make it clear where the data comes from! Neil says: Also, I
+  didn't get the impression they understood that we weren't generating
+  labels from the data, but rather from independent taxonomies. Of
+  course if you use labels from the data it isn't comparable. But the
+  point of ISO 9126 etc. was to avoid exactly this problem.}
+
+\todo[inline]{Need to hammer it in that the word bags are not project specific and
+ are inherited so it isn't a supervised method}
+
 Topics are word distributions, lists of words ranked by frequency, that are burdensome to interpret and hard to distinguish and understand. 
@@
 
+\todo[inline]{* clarify the mailinglist stuff we did}
 To construct these expanded word lists, we used WordNet~\cite{Fellbaum1998}, an English-language ``lexical database'' that contains semantic relations between words, including common related forms (similar to word stemming), meronymy and synonymy.
@@
 
+\todo[inline]{* the hierarchy of tags was not clear}
+
 % word list similarity
@@
 
+\todo[inline]{Should we use F-1 as well to satisfy everyone?}
+
 \noindent \textbf{Analysis of the unsupervised labelling} -- Based on the labels, and our manual topic labelling, 
@@
 Furthermore, our analysis techniques rely on a project's use of meaningful commit messages. 
+\todo[inline]{If we compare with the same domain we don't get to say it generalizes a lot?}
 

### 2011-01-05 msr/msr2011.tex
names
Moving some paragraphs around in introduction. Attempting to define topics. Moved the previous work section. Talked about generalizability since we only did databases.

@@
 
-\numberofauthors{5}
+\numberofauthors{4}
 
@@
 Abram Hindle\\
-   \affaddr{David Cheriton School of Computer Science}\\
-   \affaddr{University of Waterloo}\\
-   \affaddr{Waterloo, Ontario, CANADA}\\
-   \email{ahindle@uwaterloo.ca}
+   \affaddr{Department of Computer Science}\\
+   \affaddr{University of California, Davis} \\
+   \affaddr{Davis, CA, USA}\\
+   \email{abram.hindle@softwareprocess.es}
 \alignauthor  Neil A. Ernst\\
@@
    \email{migod@uwaterloo.ca}
-\alignauthor Richard C. Holt \\
-   \affaddr{David Cheriton School of Computer Science}\\
-   \affaddr{University of Waterloo}\\
-   \affaddr{Waterloo, Ontario, CANADA}\\
-   \email{holt@uwaterloo.ca}
 \alignauthor John Mylopoulos \\
@@
 
-\todo[inline]{Move previous work up}
+% \todo[inline]{Move previous work up}
+
+% An attempt Motivate topics
+%XXX is this the correct terminology for commits/revisions
+The development of software does not occur in a vacuum, revisions to
+source code are often not independent of the past, the specification or
+the task at hand.
+Often the purpose of revisions to source code relate to one or more topics, whether it be a maintenance activity, a quality issue, a requirement or feature or a response to a bug report.
+Many of these purposes and tasks exhibit themselves as latent topics which cross-cut and compose many of development artifacts: version control revisions, bug reports, and mailing-list discussions.
 
@@
 %
-Topics are typically latent, i.e., not explicitly named, and can be in several potential states: they may not be immediately resolved, they may re-occur, yet others are only dealt with briefly  (e.g., a bug report that is closed).
-% semi-automatically sounds too uncertain. I think the introduction can be more muscular and bold. Don't give the reviewer any room to start doubting.
-
-% why are topics useful
-Developers often interact with abstractions of the artifacts from which topics are derived, rather than the artifacts themselves.
-Such abstractions include project dashboards~\cite{kersten2005mylar}, tags~\cite{treude2010}, and complexity metrics \cite{mccabe1976complexity}. 
-Without the structured information which dashboards typically rely on, topics become a valuable and automatically extractable abstraction of software artifacts.
-Topics abstract software artifacts by their subject matter, rather than their concrete representation. 
-Much like aspect-oriented programming, topic labeling can identify cross-cutting concerns in a software project (as in~\cite{Baldi2008}).
-The topics of development are relevant because they identify the distinct issues facing stakeholders at a particular time.
-
-The goal of concept location and topic analysis research is to recover these topics and the related software artifacts.
-% get neil's data mining point in there
-To date, the results of topic analysis and concept location are project specific, and not generalizable.
-% get the interpretation aspect
-Much of the current research stops short of automatically interpreting the purpose of the topic, e.g.~\cite{Baldi2008,Kuhn2007,marcus04wcre}. 
-That is, extracted topics are usually manually named and labelled by those who read the topic words.
-
-% topics are numerous?
-Deriving and labelling topics is a challenging machine learning problem. 
-Topic analysis produces % raw topics that are represented as
- unlabelled context-free word-lists or word distributions.
-Topics have to be identified by interpreting the prevalent words in a word distribution and by inspecting related documents. 
-This manual approach is impractical when handling hundreds of different topics.
-To scale to large projects (thousands of lines of code and multiple releases), we need automatic assistance to determine the topic label.
-We propose a method of interpreting and labelling topics, and their related artifacts, that is generalizable and relevant to multiple projects.
-
-In our previous work we dealt with topic trends, which are topics that recur over time~\cite{Hindle09ICSM}. 
-We observed that topic trends were often non-functional requirements (NFRs). 
-NFRs have the property of being cross-domain and widely applicable. 
-In this sense, they are useful abstractions for developer conversations about different software projects.
-
-% harp on data mining more
-In general, the mining of software artifacts tends to be very project specific, yet NFRs are not. 
-% I was told if you're going to use a paren, you don't need to say it.
-There is a series of standards on NFRs, such as \cite{iso9126}, specifically intended to apply to projects of varying types.
-These standards indicate our approach is possible. %, if perhaps not universally accepted).
-% NFRs are prevalent in almost all software systems.
-% In this paper we will show that NFR related topics occur quite frequently.
-%XXXXXX this sentence is confusing
-%This informs our method of generalizing the data mining of these software artifacts to leverage the global domain knowledge of software engineering, specifically NFRs. 
-In particular, we use software quality models to generalize topics across projects. 
-This allows us to mine software systems with the intention of comparing their NFR-related development topics.
-% in order to achieve generalizable, cross project data mining results.
-
-%%%% THIS PARAGRAPH ARGUES ABOUT THE CONCRETE APPLICATION OF 
-%%%% TOPIC MODELLING AND LABELLED TOPIC EXTRACTION
-%%%% IT NEEDS TO BE TIGHTER BUT IT ALSO NEEDS TO BE HERE
-There are many practical applications of \emph{labelled topic extraction}.
+Topics are typically latent, i.e., not explicitly named, and can be in several potential states: they may not be immediately resolved, they may re-occur, yet others are only dealt with briefly  (e.g., a bug report that is closed)
+Within this paper we define \emph{topics} as latent word distributions that describe and compose the documents being analyzed. We then defined \emph{labelled topics} as \emph{topics} that have been
+
+There are many practical applications of \emph{labelled topic
+  extraction} 
+ranging from summarization, project dashboards, and software process recovery.
 These methods are useful for providing abstractions of development activities and artifacts.
@@
 
+% semi-automatically sounds too uncertain. I think the introduction can be more muscular and bold. Don't give the reviewer any room to start doubting.
+
+% why are topics useful
+Developers often interact with abstractions of the artifacts from which topics are derived, rather than the artifacts themselves.
+Such abstractions include project dashboards~\cite{kersten2005mylar}, tags~\cite{treude2010}, and complexity metrics \cite{mccabe1976complexity}. 
+Without the structured information which dashboards typically rely on, topics become a valuable and automatically extractable abstraction of software artifacts.
+Topics abstract software artifacts by their subject matter, rather than their concrete representation. 
+Much like aspect-oriented programming, topic labeling can identify cross-cutting concerns in a software project (as in~\cite{Baldi2008}).
+The topics of development are relevant because they identify the distinct issues facing stakeholders at a particular time.
+
+The goal of concept location and topic analysis research is to recover these topics and the related software artifacts.
+% get neil's data mining point in there
+To date, the results of topic analysis and concept location are project specific, and not generalizable.
+% get the interpretation aspect
+Much of the current research stops short of automatically interpreting the purpose of the topic, e.g.~\cite{Baldi2008,Kuhn2007,marcus04wcre}. 
+That is, extracted topics are usually manually named and labelled by those who read the topic words.
+
+% topics are numerous?
+Deriving and labelling topics is a challenging machine learning problem. 
+Topic analysis produces % raw topics that are represented as
+ unlabelled context-free word-lists or word distributions.
+Topics have to be identified by interpreting the prevalent words in a word distribution and by inspecting related documents. 
+This manual approach is impractical when handling hundreds of different topics.
+To scale to large projects (thousands of lines of code and multiple releases), we need automatic assistance to determine the topic label.
+We propose a method of interpreting and labelling topics, and their related artifacts, that is generalizable and relevant to multiple projects.
+
+In our previous work we dealt with topic trends, which are topics that recur over time~\cite{Hindle09ICSM}. 
+We observed that topic trends were often non-functional requirements (NFRs). 
+NFRs have the property of being cross-domain and widely applicable. 
+In this sense, they are useful abstractions for developer conversations about different software projects.
+
+% harp on data mining more
+In general, the mining of software artifacts tends to be very project specific, yet NFRs are not. 
+% I was told if you're going to use a paren, you don't need to say it.
+There is a series of standards on NFRs, such as \cite{iso9126}, specifically intended to apply to projects of varying types.
+These standards indicate our approach is possible. %, if perhaps not universally accepted).
+% NFRs are prevalent in almost all software systems.
+% In this paper we will show that NFR related topics occur quite frequently.
+%XXXXXX this sentence is confusing
+%This informs our method of generalizing the data mining of these software artifacts to leverage the global domain knowledge of software engineering, specifically NFRs. 
+In particular, we use software quality models to generalize topics across projects. 
+This allows us to mine software systems with the intention of comparing their NFR-related development topics.
+% in order to achieve generalizable, cross project data mining results.
+
+%%%% THIS PARAGRAPH ARGUES ABOUT THE CONCRETE APPLICATION OF 
+%%%% TOPIC MODELLING AND LABELLED TOPIC EXTRACTION
+%%%% IT NEEDS TO BE TIGHTER BUT IT ALSO NEEDS TO BE HERE
+
 %% RESEARCH DIDNT DO THIS AND WE ARE TOTALLY AWESOME AND WE DID
@@
 This paper first introduces labelled topic extraction, and then uses that technique to analyze development activity in two large-scale examples.
-We begin by describing how we generated our data (Sections  \ref{sec:wordlist} and \ref{sec:genwordlist}). 
+Next we cover much of the previous work that we leverage (Section \ref{sec:related}).
+Then we follow by describing how we generated our data (Sections  \ref{sec:wordlist} and \ref{sec:genwordlist}). 
 We show that labels with their topics can be learned and used to classify other data-sets, either without training (Section \ref{sec:unsuplabelling}), or with training (Section \ref{sec:suplabelling}).
@@
 We use an exploratory case study of two open source database systems to show how named topics can be compared between projects  (Section \ref{sec:analysis}). 
-The paper concludes with a discussion of limitations (Section \ref{sec:limit}) and related work (Section \ref{sec:related}).
+The paper concludes with a discussion of limitations (Section \ref{sec:limit})
 
@@
 %\todo[inline]{We still haven't shown that topics are useful -- a reviewer comment}
+
+\section{Previous Work}
+\label{sec:related}
+
+The idea of extracting higher-level `concerns', also known as `concepts', `aspects' or `requirements', has been approached from documentation-based and repository-based perspectives.
+
+Cleland-Huang and her colleagues published work on mining requirements documents for non-functional requirements (NFR) (quality requirements)~\cite{Cleland-Huang2006}. 
+Their approach was similar to ours, as they mined keywords from their own NFR catalogues~\cite{chung99}.
+%One approach they tried was similar to this one, with keywords mined from NFR catalogues found in their previous work~\cite{chung99}. 
+They managed a recall of 80\% with precision of 57\% for the Security NFR, but could not find a reliable source of keywords for other NFRs. 
+Instead, they developed a supervised classifier by using human experts to identify an NFR training set. 
+Our worked differed because we had a more comprehensive set of terms based on the taxonomy we chose.
+We also wanted a common taxonomy that allows us to make cross-project comparison; this was not one of their objectives.
+The objective of Cleland-Huang's study was to identify new NFRs for system development, yet our objective was to recover those latent NFRs from commit-log messages of the project. 
+Another difference was that our data was far less structured than the requirements documents used in their study.
+%There are several reasons we did not follow this route. 
+%One, we believe we have a more comprehensive set of terms due to the taxonomy we chose. 
+%Secondly, we wanted to compare across projects. 
+%Their technique was not compared across different projects and the applicability of the training set to different corpora is unclear. 
+%A common taxonomy allows us to make cross-project comparison. 
+%Although subject to the assumption that all projects conceive of these terms in the same way.
+%Thirdly, while the objective of Cleland-Huang's study was to identify new NFRs (for system development) our study assumes these NFRs are latent in the commit-log messages of the project.
+
+In the same vein, Mockus and Votta~\cite{Mockus00} studied a large-scale industrial change-tracking system. 
+They also leveraged WordNet, but only for word roots as they felt the synonyms would be non-specific and cause errors. 
+Mockus et al. had access to system developers, with whom they could validate their labels. 
+Since we try to bridge different organizations, these kind of interviews are infeasible (particularly in the distributed world of open-source software).
+
+The other approach is to extract concerns from software repositories.
+Marcus et al.~\cite{marcus04wcre} describe their use of Latent Semantic Indexing to identify commonly occurring concerns for software maintenance. 
+%Some results were interesting, but their precision was quite low. 
+ConcernLines~\cite{treude09cl} shows tag occurrence using colour intensity. 
+They mined 
+developer created tags 
+%change request tags (such as `milestone 3') and used these 
+in order to analyze the evolution of a single product.
+% make evolutionary analyses of a single product. 
+The presence of a well-maintained set of tags is obviously essential to the success of this technique.
+
+%XXXXXXXXXXXX OPTIONAL
+% Not sure we need this
+Mens et al.~\cite{mens08icsm} conducted an empirical study of Eclipse, a popular Java IDE, to verify the claims of Lehman~\cite{lehman97sms}.
+% They concerned themselves with source code only, and found Law Seven, ``Declining Quality'', to be too difficult to assess: ``[we lacked an] appropriate measurement of the evolving quality of the system as perceived by the users \cite[p.
+388]{mens08icsm}''.
+This paper examines the notions of quality in terms of a consistent ontology, as Mens
+et al. call for in their conclusions.
+
+Mei et al.~\cite{Mei2007} use context information to automatically name topics. 
+They describe probabilistic labelling, using the frequency distribution of words in a topic to create a meaningful phrase. 
+They do not use external domain-specific information as we do.
+In \cite{ernst10refsq}, we describe our earlier project, similar to this, to identify change in quality requirements in GNOME software projects; our approach is solely text-matching, however, and does not leverage machine learning strategies.
+
+%Massey~\cite{massey02icse} and Scacchi~\cite{scacchi02,scacchi05b} looked at the topic of requirements in open-source software. 
+% this reduces the line length and drops a ref even though we still have a scacchi kickin around
+Massey~\cite{massey02icse} and Scacchi~\cite{scacchi05b} looked at the topic of requirements in open-source software. 
+Their work discusses the source of the requirements and how they are used in the development process.
+German~\cite{german03gnome} looked at GNOME specifically, and listed several sources for requirements: leader interest, mimicry, brainstorming, and prototypes.
+%None of this work  addressed quality requirements in OSS, nor did it examine requirements trends.
+
+% XXX vanity or space.. 
+%XXXXXXXXXXXX OPTIONAL
+ Hindle et al.
+\cite{Hindle2007} examined release patterns in OSS.
+They showed that there is a difference between projects regarding maintenance techniques.
+This supports our result that software qualities are not discussed with the same frequency across projects.
 
@@
 We also chose our taxonomy and the data to study. We did not show these results to the stakeholders.
+We only studied database systems.
 
@@
 Our data originated from OSS database projects and thus might not be applicable to commercially developed software. 
-Furthermore, our analysis techniques rely on a project's use of meaningful commit messages. 
-\todo[inline]{If we compare with the same domain we don't get to say it generalizes a lot?}
+Furthermore, our analysis techniques rely on a project's use of meaningful commit messages. The OSS projects investigated were database systems, thus 
+our results might not generalize to other domains.
+% \todo[inline]{If we compare with the same domain we don't get to say it generalizes a lot?}
 
@@
 Unfortunately only two annotators were used; their annotations could be biased as we did not analyze for inter-rater reliability.
-
-\section{Related work}
-\label{sec:related}
-
-The idea of extracting higher-level `concerns', also known as `concepts', `aspects' or `requirements', has been approached from documentation-based and repository-based perspectives.
-
-Cleland-Huang and her colleagues published work on mining requirements documents for non-functional requirements (NFR) (quality requirements)~\cite{Cleland-Huang2006}. 
-Their approach was similar to ours, as they mined keywords from their own NFR catalogues~\cite{chung99}.
-%One approach they tried was similar to this one, with keywords mined from NFR catalogues found in their previous work~\cite{chung99}. 
-They managed a recall of 80\% with precision of 57\% for the Security NFR, but could not find a reliable source of keywords for other NFRs. 
-Instead, they developed a supervised classifier by using human experts to identify an NFR training set. 
-Our worked differed because we had a more comprehensive set of terms based on the taxonomy we chose.
-We also wanted a common taxonomy that allows us to make cross-project comparison; this was not one of their objectives.
-The objective of Cleland-Huang's study was to identify new NFRs for system development, yet our objective was to recover those latent NFRs from commit-log messages of the project. 
-Another difference was that our data was far less structured than the requirements documents used in their study.
-%There are several reasons we did not follow this route. 
-%One, we believe we have a more comprehensive set of terms due to the taxonomy we chose. 
-%Secondly, we wanted to compare across projects. 
-%Their technique was not compared across different projects and the applicability of the training set to different corpora is unclear. 
-%A common taxonomy allows us to make cross-project comparison. 
-%Although subject to the assumption that all projects conceive of these terms in the same way.
-%Thirdly, while the objective of Cleland-Huang's study was to identify new NFRs (for system development) our study assumes these NFRs are latent in the commit-log messages of the project.
-
-In the same vein, Mockus and Votta~\cite{Mockus00} studied a large-scale industrial change-tracking system. 
-They also leveraged WordNet, but only for word roots as they felt the synonyms would be non-specific and cause errors. 
-Mockus et al. had access to system developers, with whom they could validate their labels. 
-Since we try to bridge different organizations, these kind of interviews are infeasible (particularly in the distributed world of open-source software).
-
-The other approach is to extract concerns from software repositories.
-Marcus et al.~\cite{marcus04wcre} describe their use of Latent Semantic Indexing to identify commonly occurring concerns for software maintenance. 
-%Some results were interesting, but their precision was quite low. 
-ConcernLines~\cite{treude09cl} shows tag occurrence using colour intensity. 
-They mined 
-developer created tags 
-%change request tags (such as `milestone 3') and used these 
-in order to analyze the evolution of a single product.
-% make evolutionary analyses of a single product. 
-The presence of a well-maintained set of tags is obviously essential to the success of this technique.
-
-%XXXXXXXXXXXX OPTIONAL
-% Not sure we need this
-Mens et al.~\cite{mens08icsm} conducted an empirical study of Eclipse, a popular Java IDE, to verify the claims of Lehman~\cite{lehman97sms}.
-% They concerned themselves with source code only, and found Law Seven, ``Declining Quality'', to be too difficult to assess: ``[we lacked an] appropriate measurement of the evolving quality of the system as perceived by the users \cite[p.
-388]{mens08icsm}''.
-This paper examines the notions of quality in terms of a consistent ontology, as Mens
-et al. call for in their conclusions.
-
-Mei et al.~\cite{Mei2007} use context information to automatically name topics. 
-They describe probabilistic labelling, using the frequency distribution of words in a topic to create a meaningful phrase. 
-They do not use external domain-specific information as we do.
-In \cite{ernst10refsq}, we describe our earlier project, similar to this, to identify change in quality requirements in GNOME software projects; our approach is solely text-matching, however, and does not leverage machine learning strategies.
-
-%Massey~\cite{massey02icse} and Scacchi~\cite{scacchi02,scacchi05b} looked at the topic of requirements in open-source software. 
-% this reduces the line length and drops a ref even though we still have a scacchi kickin around
-Massey~\cite{massey02icse} and Scacchi~\cite{scacchi05b} looked at the topic of requirements in open-source software. 
-Their work discusses the source of the requirements and how they are used in the development process.
-German~\cite{german03gnome} looked at GNOME specifically, and listed several sources for requirements: leader interest, mimicry, brainstorming, and prototypes.
-%None of this work  addressed quality requirements in OSS, nor did it examine requirements trends.
-
-% XXX vanity or space.. 
-%XXXXXXXXXXXX OPTIONAL
- Hindle et al.
-\cite{Hindle2007} examined release patterns in OSS.
-They showed that there is a difference between projects regarding maintenance techniques.
-This supports our result that software qualities are not discussed with the same frequency across projects.
 

### 2011-01-05 msr/msr2011.tex
more intro edits

@@
 
-\todo[inline]{mention project dashboards}
+% \todo[inline]{mention project dashboards}
 
@@
 Topics are typically latent, i.e., not explicitly named, and can be in several potential states: they may not be immediately resolved, they may re-occur, yet others are only dealt with briefly  (e.g., a bug report that is closed)
-Within this paper we define \emph{topics} as latent word distributions that describe and compose the documents being analyzed. We then defined \emph{labelled topics} as \emph{topics} that have been
+Within this paper we define \emph{topics} as latent word distributions that describe and compose the documents (version control commits) being analyzed. 
+We then defined \emph{labelled topics} as \emph{topics} that have been interpreted and labelled. 
+The distinction is important because word distributions can consistent of many words, thus if one has many topics, reading and interpreting these words is prohibitive.
+For example, if you read only 10 words per topic from 500 topics you have to read and interpret 5000 words, this is 5 or more pages of text!
+Thus \emph{labelled topics} save interpretation time.
 

### 2011-01-10 msr/msr2011.tex
More intro fiddling
migod's abstract

@@
    \affaddr{Davis, CA, USA}\\
-   \email{abram.hindle@softwareprocess.es}
+   \email{abram@softwareprocess.es}
 \alignauthor  Neil A. Ernst\\
@@
 
-Researchers have used topic modeling and concept location to understand the latent topics of software development artifacts.
-These techniques use unsupervised machine-learning algorithms to recover topics.
-These topics are word-lists and are difficult to distinguish and interpret. 
-Topics are not meaningful until they have been named or interpreted.
-Current topic labelling approaches are manual, and do not use domain-specific knowledge to improve, contextualize, or describe results for the developers.
-We propose a solution: \emph{labelled topic extraction}.
-Topics are extracted using Latent Dirichlet Allocation (LDA) from commit-log comments recovered from source control systems such as CVS and BitKeeper. 
-These topics are given labels relating to a generalizable cross-project taxonomy consisting of non-functional requirements.
-Our approach was evaluated  with experiments and case studies on two large-scale RDBMS projects: MySQl and MaxDB.
-Labelled topic extraction produces appropriate, context-sensitive labels relevant to these projects, which provides fresh insight into their evolving software development activities.
+Researchers have employed a variety of techniques to extract underlying
+topics that relate to software development artifacts.  Typically, these
+techniques use unsupervised machine-learning algorithms to suggest
+candidate word-lists that recur together often.  However, word-lists are
+difficult to 
+interpret in the absence of meaningful summary labels.  Current topic modeling 
+techniques assume manual labelling and do not use domain-specific knowledge
+to improve, contextualize, or describe results for the developers.  
+We propose a solution: \emph{automated labelled topic extraction}.
+  Topics are extracted 
+using  Latent Dirichlet Allocation (LDA) from commit-log comments recovered
+from source control systems such as CVS and BitKeeper.  
+These topics are given labels relating to a generalizable
+cross-project taxonomy, consisting of non-functional
+requirements. Our approach was evaluated with experiments and case studies
+on two large-scale RDBMS projects: MySQL and MaxDB.  The case studies
+show that labelled topic extraction can produce appropriate,
+context-sensitive labels relevant to these projects, which provides fresh
+insight into their evolving software development activities.
+
+% Researchers have used topic modeling and concept location to understand the latent topics of software development artifacts.
+% These techniques use unsupervised machine-learning algorithms to recover topics.
+% These topics are word-lists and are difficult to distinguish and interpret. 
+% Topics are not meaningful until they have been named or interpreted.
+% Current topic labelling approaches are manual, and do not use domain-specific knowledge to improve, contextualize, or describe results for the developers.
+% We propose a solution: \emph{labelled topic extraction}.
+% Topics are extracted using Latent Dirichlet Allocation (LDA) from commit-log comments recovered from source control systems such as CVS and BitKeeper. 
+% These topics are given labels relating to a generalizable cross-project taxonomy consisting of non-functional requirements.
+% Our approach was evaluated  with experiments and case studies on two large-scale RDBMS projects: MySQl and MaxDB.
+% Labelled topic extraction produces appropriate, context-sensitive labels relevant to these projects, which provides fresh insight into their evolving software development activities.
 
@@
 Our technique would help explain why this was the case, since it can identify NFRs which are dominating developer attention.
+Thus developers can use these techniques to get an overview of the current and past development topics.
+For awareness, managers can use these techniques to gain an overview of topics of development.
+
 % This is a very long example it'd be nice if it was tighter

### 2011-01-10 msr/msr2011.tex
f-measures added

@@
 
-\todo[inline]{Topic should be defined clearly (italics and hilighted)}
+%\todo[inline]{Topic should be defined clearly (italics and hilighted)}
 
@@
 Within this paper we define \emph{topics} as latent word distributions that describe and compose the documents (version control commits) being analyzed. 
-We then defined \emph{labelled topics} as \emph{topics} that have been interpreted and labelled. 
+We define \emph{labelled topics} as \emph{topics} that have been interpreted and labelled. 
 The distinction is important because word distributions can consistent of many words, thus if one has many topics, reading and interpreting these words is prohibitive.
@@
 
-Our first word-list set, \textsf{exp1}, was generated using the ontology described in Kayed et al.~\cite{5072519}. 
+Our first word-list set, \textsf{exp1}, was generated using the ontology described in Kayed et al.~\cite{5072519} .
 That paper constructs an ontology for software quality measurement using eighty source documents, including research papers and international standards. 
@@
 
-\todo[inline]{Should we use F-1 as well to satisfy everyone?}
+%\todo[inline]{Should we use F-1 as well to satisfy everyone?}
 
@@
 
+% EXP2
+For \textsf{exp2}, the average F-measure (macro-F1) for MaxDB was $0.24$ with a range $0.091$ to
+$0.37$, and $0.16$ for MySQL with a range of $0$ to $0.41$.
+MaxDB had an average precision and recall of $0.25$ and $0.22$
+while MySQL had $0.41$ and $0.10$ respectively.
+
+% EXP3
+For \textsf{exp3}, the average F-measure (macro-F1) for MaxDB was $0.26$ with a range $0.11$ to
+$0.47$, and $0.36$ for MySQL with a range of $0.10$ to $0.65$.
+MaxDB had an average precision and recall of $0.16$ and $0.67$
+while MySQL had $0.3$ and $0.48$ respectively.
+
 Based on these results we find that \emph{reliability} and \emph{usability} worked well for MaxDB in \textsf{exp2} and better in \textsf{exp3}. 
@@
 If a \emph{C} grade performance has a ROC value of $0.6$ then most of these tests scored a grade of \emph{C} or less, but our classifier nonetheless performed substantially better than random.
+
+\begin{comment}
+maxdb
+
+portability,0.773778920308,0.228070175439,76,12,0.181684100119,0.52,0.14606741573,0.553033707865,288,0.96,13
+efficiency,0.832904884319,0.21686746988,36,29,0.124642098177,0.236842105263,0.2,0.557848837209,315,0.915697674419,9
+reliability,0.714652956298,0.379888268156,7,104,0.340426350577,0.246376811594,0.829268292683,0.765208858985,244,0.701149425287,34
+functionality,0.902313624679,0.0952380952381,32,6,0.0834236909204,0.25,0.0588235294118,0.520961060481,349,0.983098591549,2
+maintainability,0.796915167095,0.0919540229885,76,3,0.122505166175,0.571428571429,0.05,0.520145631068,306,0.990291262136,4
+usability,0.830334190231,0.175,11,55,0.138109307712,0.112903225806,0.388888888889,0.620320455226,316,0.851752021563,7
+
+total,0.808483290488,0.235897435897,238,209,0.126930463023,0.248201438849,0.224755700326,0.560823829443,1818,0.896891958559,69mysql
+
+portability,0.616040955631,0.137931034483,225,0,0.211497874858,1.0,0.0740740740741,0.537037037037,343,1.0,18
+efficiency,0.924914675768,0.3125,33,11,0.297817974114,0.47619047619,0.232558139535,0.606150156324,532,0.979742173112,10
+reliability,0.726962457338,0.411764705882,61,99,0.242477699079,0.361290322581,0.478632478632,0.633772529295,370,0.788912579957,56
+functionality,0.457337883959,0.0245398773006,315,3,0.00597425429824,0.571428571429,0.012539184953,0.500651614948,264,0.988764044944,4
+maintainability,0.686006825939,0,184,0,0,0,0.0,0.5,402,1.0,0
+usability,0.919795221843,0.175438596491,27,20,0.13508225345,0.2,0.15625,0.560074458484,534,0.963898916968,5
+
+total,0.721843003413,0.159793814433,845,133,0.0857635592199,0.411504424779,0.0991471215352,0.523778370698,2445,0.9484ahindle@wish:~/projects/lda-paper/src/validate$ cat NOTES 
+columns of maxdb.report.text
+
+    out['accuracy'] = acc
+    out['f1'] = f1
+    out['fn'] = fn
+    out['fp'] = fp
+    out['mcc'] = mcc
+    out['precision'] = pre
+    out['recall'] = rec
+    out['roca'] = roca
+    out['tnr'] = tnr
+    out['tn'] = tn
+    out['tp'] = tp
+
+cat exp3.*.report.txt 
+maxdb
+
+portability,0.706940874036,0.472222222222,38,76,0.286382363909,0.40157480315,0.573033707865,0.659850187266,224,0.746666666667,51
+efficiency,0.372750642674,0.222929936306,10,234,0.067549899325,0.130111524164,0.777777777778,0.548772609819,110,0.31976744186,35
+reliability,0.434447300771,0.256756756757,3,217,0.195966959965,0.149019607843,0.926829268293,0.651633024951,131,0.376436781609,38
+functionality,0.53470437018,0.236286919831,6,175,0.186902353269,0.137931034483,0.823529411765,0.665285832643,180,0.507042253521,28
+maintainability,0.627249357326,0.337899543379,43,102,0.111672018872,0.26618705036,0.4625,0.566201456311,207,0.669902912621,37
+usability,0.27763496144,0.107936507937,1,280,0.0937953415493,0.0572390572391,0.944444444444,0.594863731656,91,0.245283018868,17
+
+total,0.492287917738,0.257983719474,101,1084,0.0926022435775,0.159689922481,0.671009771987,0.568114654124,943,0.46521953626,206
+mysql
+
+portability,0.616040955631,0.412532637076,164,61,0.170139234681,0.564285714286,0.325102880658,0.57363015753,282,0.822157434402,79
+efficiency,0.532423208191,0.184523809524,12,262,0.124342460899,0.105802047782,0.720930232558,0.619212814253,281,0.517495395948,31
+reliability,0.588737201365,0.404938271605,35,206,0.209194239572,0.284722222222,0.700854700855,0.630811145737,263,0.560767590618,82
+functionality,0.622866894198,0.651968503937,112,109,0.240455709256,0.655063291139,0.648902821317,0.620331560471,158,0.591760299625,207
+maintainability,0.621160409556,0.218309859155,153,69,-0.00390279540262,0.31,0.16847826087,0.498418234912,333,0.828358208955,31
+usability,0.358361774744,0.104761904762,10,366,0.0128981856383,0.0567010309278,0.6875,0.513425090253,188,0.339350180505,22
+
+total,0.556598407281,0.367032074706,486,1073,0.058598680978,0.296393442623,0.481876332623,0.532831106575,1505,0.583785880528,452
+\end{comment}
 

### 2011-01-10 msr/msr2011.tex
ML F-Measures

--- 2011-01-10 msr/msr2011.tex
@@
 If a \emph{C} grade performance has a ROC value of $0.6$ then most of these tests scored a grade of \emph{C} or less, but our classifier nonetheless performed substantially better than random.
+
+% \begin{comment}
+% maxdb
+
+% portability,0.773778920308,0.228070175439,76,12,0.181684100119,0.52,0.14606741573,0.553033707865,288,0.96,13
+% efficiency,0.832904884319,0.21686746988,36,29,0.124642098177,0.236842105263,0.2,0.557848837209,315,0.915697674419,9
+% reliability,0.714652956298,0.379888268156,7,104,0.340426350577,0.246376811594,0.829268292683,0.765208858985,244,0.701149425287,34
+% functionality,0.902313624679,0.0952380952381,32,6,0.0834236909204,0.25,0.0588235294118,0.520961060481,349,0.983098591549,2
+% maintainability,0.796915167095,0.0919540229885,76,3,0.122505166175,0.571428571429,0.05,0.520145631068,306,0.990291262136,4
+% usability,0.830334190231,0.175,11,55,0.138109307712,0.112903225806,0.388888888889,0.620320455226,316,0.851752021563,7
+
+% total,0.808483290488,0.235897435897,238,209,0.126930463023,0.248201438849,0.224755700326,0.560823829443,1818,0.896891958559,69mysql
+
+% portability,0.616040955631,0.137931034483,225,0,0.211497874858,1.0,0.0740740740741,0.537037037037,343,1.0,18
+% efficiency,0.924914675768,0.3125,33,11,0.297817974114,0.47619047619,0.232558139535,0.606150156324,532,0.979742173112,10
+% reliability,0.726962457338,0.411764705882,61,99,0.242477699079,0.361290322581,0.478632478632,0.633772529295,370,0.788912579957,56
+% functionality,0.457337883959,0.0245398773006,315,3,0.00597425429824,0.571428571429,0.012539184953,0.500651614948,264,0.988764044944,4
+% maintainability,0.686006825939,0,184,0,0,0,0.0,0.5,402,1.0,0
+% usability,0.919795221843,0.175438596491,27,20,0.13508225345,0.2,0.15625,0.560074458484,534,0.963898916968,5
+
+% total,0.721843003413,0.159793814433,845,133,0.0857635592199,0.411504424779,0.0991471215352,0.523778370698,2445,0.9484ahindle@wish:~/projects/lda-paper/src/validate$ cat NOTES 
+% columns of maxdb.report.text
+
+%     out['accuracy'] = acc
+%     out['f1'] = f1
+%     out['fn'] = fn
+%     out['fp'] = fp
+%     out['mcc'] = mcc
+%     out['precision'] = pre
+%     out['recall'] = rec
+%     out['roca'] = roca
+%     out['tnr'] = tnr
+%     out['tn'] = tn
+%     out['tp'] = tp
+
+% cat exp3.*.report.txt 
+% maxdb
+
+% portability,0.706940874036,0.472222222222,38,76,0.286382363909,0.40157480315,0.573033707865,0.659850187266,224,0.746666666667,51
+% efficiency,0.372750642674,0.222929936306,10,234,0.067549899325,0.130111524164,0.777777777778,0.548772609819,110,0.31976744186,35
+% reliability,0.434447300771,0.256756756757,3,217,0.195966959965,0.149019607843,0.926829268293,0.651633024951,131,0.376436781609,38
+% functionality,0.53470437018,0.236286919831,6,175,0.186902353269,0.137931034483,0.823529411765,0.665285832643,180,0.507042253521,28
+% maintainability,0.627249357326,0.337899543379,43,102,0.111672018872,0.26618705036,0.4625,0.566201456311,207,0.669902912621,37
+% usability,0.27763496144,0.107936507937,1,280,0.0937953415493,0.0572390572391,0.944444444444,0.594863731656,91,0.245283018868,17
+
+% total,0.492287917738,0.257983719474,101,1084,0.0926022435775,0.159689922481,0.671009771987,0.568114654124,943,0.46521953626,206
+% mysql
+
+% portability,0.616040955631,0.412532637076,164,61,0.170139234681,0.564285714286,0.325102880658,0.57363015753,282,0.822157434402,79
+% efficiency,0.532423208191,0.184523809524,12,262,0.124342460899,0.105802047782,0.720930232558,0.619212814253,281,0.517495395948,31
+% reliability,0.588737201365,0.404938271605,35,206,0.209194239572,0.284722222222,0.700854700855,0.630811145737,263,0.560767590618,82
+% functionality,0.622866894198,0.651968503937,112,109,0.240455709256,0.655063291139,0.648902821317,0.620331560471,158,0.591760299625,207
+% maintainability,0.621160409556,0.218309859155,153,69,-0.00390279540262,0.31,0.16847826087,0.498418234912,333,0.828358208955,31
+% usability,0.358361774744,0.104761904762,10,366,0.0128981856383,0.0567010309278,0.6875,0.513425090253,188,0.339350180505,22
+
+% total,0.556598407281,0.367032074706,486,1073,0.058598680978,0.296393442623,0.481876332623,0.532831106575,1505,0.583785880528,452
+% \end{comment}
 
@@
 Figure \ref{fig:best-learn-per-tag} shows the performance of the best performing learner per label. 
- We considered the best learner for a label to be the learner which had the highest ROC value for that label. 
+ We considered the best learner for a label to be the learner which
+ had the highest ROC value for that label. 
+The best learner is important because one uses a single learner per label.
 %Figure \ref{fig:best-learn-per-tag} uses the ZeroR learner as a baseline, a common practice in machine learning,
@@
 %http://nlp.stanford.edu/IR-book/html/htmledition/naive-bayes-text-classification-1.html
+
+% # F measure
+% mysql
+% > v <- c( 0.64 , 0.23  ,0.41 , 0.77  ,0.62 , 0.21 ) 
+% > summary(v)
+%    Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
+%   0.210   0.275   0.515   0.480   0.635   0.770 
+% maxdb
+% > v <- c(0.61, 0.25, 0.59, 0.32, 0.42, 0.17)
+% > summary(v)
+%    Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
+%  0.1700  0.2675  0.3700  0.3933  0.5475  0.6100
+
+The range of F-measures for MySQL was $0.21$ to $0.77$ with an average
+of $0.48$, while MaxDB had a range of $0.17$ to $0.61$ with an average
+of $0.39$.
 

### 2011-01-11 msr/msr2011.tex
Latest changes
some notes

@@
 Abram Hindle\\
-   \affaddr{Department of Computer Science}\\
+   \affaddr{Dept.of Computer Science}\\
    \affaddr{University of California, Davis} \\
@@
 \alignauthor  Neil A. Ernst\\
-   \affaddr{Department of Computer Science}\\
+   \affaddr{Dept. of Computer Science}\\
    \affaddr{University of Toronto}\\
@@
 \alignauthor John Mylopoulos \\
-   \affaddr{Department of Computer Science}\\
+   \affaddr{Dept. of Computer Science}\\
    \affaddr{University of Toronto}\\
@@
 
-Accordingly, we tried to associate each topic with a label from a list of keywords and related terms. 
+Accordingly, we tried to associate each topic with a label from a list
+of keywords and related terms. These keywords were determined before
+hand and were not extracted from the project themselves.
 We intersected the words of the topics and the words of our word-lists.
@@
 \todo[inline]{* clarify the mailinglist stuff we did}
+\todo[inline]{This part doesn't make a lot of sense}
 To construct these expanded word lists, we used WordNet~\cite{Fellbaum1998}, an English-language ``lexical database'' that contains semantic relations between words, including common related forms (similar to word stemming), meronymy and synonymy.
@@
 % word list similarity
-Using our three word lists, we labeled our topics where there was a match between a word in the list and the same word in the topic.
+Using our three word lists (\textsf{exp1}, \textsf{exp2}, \textsf{exp3}), we labeled our topics where there was a match between a word in the list and the same word in the topic.
 A \emph{named topic} is a topic with a matching label. 

### 2011-01-13 icsm09/dist.sh
removing ICSM files for space
Signed-off-by: Neil Ernst <neil@neilernst.net>

@@ -1,9 +1,0 @@
-#make clean paper1.pdf
-#FILE=indentation.paper1.`date +"%Y%m%d.%H%M%S.%h.%d.%Y"`.pdf
-#cp paper1.pdf $FILE
-#vex $FILE
-FILEO=lda-paper.pdf
-make clean $FILEO 
-FILE=lda-paper.`date +"%Y%m%d.%H%M%S.%h.%d.%Y"`.pdf
-cp $FILEO $FILE
-vex $FILE

### 2011-01-13 icsm09/lda-c-dist/lda-data.c

@@ -1,64 +1,0 @@
-// (C) Copyright 2004, David M. Blei (blei [at] cs [dot] cmu [dot] edu)
-
-// This file is part of LDA-C.
-
-// LDA-C is free software; you can redistribute it and/or modify it under
-// the terms of the GNU General Public License as published by the Free
-// Software Foundation; either version 2 of the License, or (at your
-// option) any later version.
-
-// LDA-C is distributed in the hope that it will be useful, but WITHOUT
-// ANY WARRANTY; without even the implied warranty of MERCHANTABILITY or
-// FITNESS FOR A PARTICULAR PURPOSE.  See the GNU General Public License
-// for more details.
-
-// You should have received a copy of the GNU General Public License
-// along with this program; if not, write to the Free Software
-// Foundation, Inc., 59 Temple Place, Suite 330, Boston, MA 02111-1307
-// USA
-
-#include "lda-data.h"
-
-corpus* read_data(char* data_filename)
-{
-    FILE *fileptr;
-    int length, count, word, n, nd, nw;
-    corpus* c;
-
-    printf("reading data from %s\n", data_filename);
-    c = malloc(sizeof(corpus));
-    fileptr = fopen(data_filename, "r");
-    nd = 0; nw = 0;
-    while ((fscanf(fileptr, "%10d", &length) != EOF))
-    {
-	c->docs = (document*) realloc(c->docs, sizeof(document)*(nd+1));
-	c->docs[nd].length = length;
-	c->docs[nd].total = 0;
-	c->docs[nd].words = malloc(sizeof(int)*length);
-	c->docs[nd].counts = malloc(sizeof(int)*length);
-	for (n = 0; n < length; n++)
-	{
-	    fscanf(fileptr, "%10d:%10d", &word, &count);
-	    word = word - OFFSET;
-	    c->docs[nd].words[n] = word;
-	    c->docs[nd].counts[n] = count;
-	    c->docs[nd].total += count;
-	    if (word >= nw) { nw = word + 1; }
-	}
-	nd++;
-    }
-    fclose(fileptr);
-    c->num_docs = nd;
-    c->num_terms = nw;
-    printf("number of docs    : %d\n", nd);
-    printf("number of terms   : %d\n", nw);
-    return(c);
-}
-
-int max_corpus_length(corpus* c)
-{
-    int n, max = 0;
-    for (n = 0; n < c->num_docs; n++)
-	if (c->docs[n].length > max) max = c->docs[n].length;
-    return(max);
-}

### 2011-01-13 icsm09/lda-c-dist/lda-estimate.c

@@ -1,337 +1,0 @@
-// (C) Copyright 2004, David M. Blei (blei [at] cs [dot] cmu [dot] edu)
-
-// This file is part of LDA-C.
-
-// LDA-C is free software; you can redistribute it and/or modify it under
-// the terms of the GNU General Public License as published by the Free
-// Software Foundation; either version 2 of the License, or (at your
-// option) any later version.
-
-// LDA-C is distributed in the hope that it will be useful, but WITHOUT
-// ANY WARRANTY; without even the implied warranty of MERCHANTABILITY or
-// FITNESS FOR A PARTICULAR PURPOSE.  See the GNU General Public License
-// for more details.
-
-// You should have received a copy of the GNU General Public License
-// along with this program; if not, write to the Free Software
-// Foundation, Inc., 59 Temple Place, Suite 330, Boston, MA 02111-1307
-// USA
-
-#include "lda-estimate.h"
-
-/*
- * perform inference on a document and update sufficient statistics
- *
- */
-
-double doc_e_step(document* doc, double* gamma, double** phi,
-                  lda_model* model, lda_suffstats* ss)
-{
-    double likelihood;
-    int n, k;
-
-    // posterior inference
-
-    likelihood = lda_inference(doc, model, gamma, phi);
-
-    // update sufficient statistics
-
-    double gamma_sum = 0;
-    for (k = 0; k < model->num_topics; k++)
-    {
-        gamma_sum += gamma[k];
-        ss->alpha_suffstats += digamma(gamma[k]);
-    }
-    ss->alpha_suffstats -= model->num_topics * digamma(gamma_sum);
-
-    for (n = 0; n < doc->length; n++)
-    {
-        for (k = 0; k < model->num_topics; k++)
-        {
-            ss->class_word[k][doc->words[n]] += doc->counts[n]*phi[n][k];
-            ss->class_total[k] += doc->counts[n]*phi[n][k];
-        }
-    }
-
-    ss->num_docs = ss->num_docs + 1;
-
-    return(likelihood);
-}
-
-
-/*
- * writes the word assignments line for a document to a file
- *
- */
-
-void write_word_assignment(FILE* f, document* doc, double** phi, lda_model* model)
-{
-    int n;
-
-    fprintf(f, "%03d", doc->length);
[…]

### 2011-01-13 icsm09/lda-c-dist/lda-estimate.h

@@ -1,49 +1,0 @@
-#ifndef LDA_ESTIMATE_H
-#define LDA_ESTIMATE_H
-
-#include <stdlib.h>
-#include <stdio.h>
-#include <math.h>
-#include <float.h>
-#include <string.h>
-
-#include "lda.h"
-#include "lda-data.h"
-#include "lda-inference.h"
-#include "lda-model.h"
-#include "lda-alpha.h"
-#include "utils.h"
-#include "cokus.h"
-
-int LAG = 5;
-
-float EM_CONVERGED;
-int EM_MAX_ITER;
-int ESTIMATE_ALPHA;
-double INITIAL_ALPHA;
-int NTOPICS;
-
-double doc_e_step(document* doc,
-                  double* gamma,
-                  double** phi,
-                  lda_model* model,
-                  lda_suffstats* ss);
-
-void save_gamma(char* filename,
-                double** gamma,
-                int num_docs,
-                int num_topics);
-
-void run_em(char* start,
-            char* directory,
-            corpus* corpus);
-
-void read_settings(char* filename);
-
-void infer(char* model_root,
-           char* save,
-           corpus* corpus);
-
-#endif
-
-

### 2011-01-13 icsm09/lda-c-dist/lda-inference.c

@@ -1,127 +1,0 @@
-// (C) Copyright 2004, David M. Blei (blei [at] cs [dot] cmu [dot] edu)
-
-// This file is part of LDA-C.
-
-// LDA-C is free software; you can redistribute it and/or modify it under
-// the terms of the GNU General Public License as published by the Free
-// Software Foundation; either version 2 of the License, or (at your
-// option) any later version.
-
-// LDA-C is distributed in the hope that it will be useful, but WITHOUT
-// ANY WARRANTY; without even the implied warranty of MERCHANTABILITY or
-// FITNESS FOR A PARTICULAR PURPOSE.  See the GNU General Public License
-// for more details.
-
-// You should have received a copy of the GNU General Public License
-// along with this program; if not, write to the Free Software
-// Foundation, Inc., 59 Temple Place, Suite 330, Boston, MA 02111-1307
-// USA
-
-#include "lda-inference.h"
-
-/*
- * variational inference
- *
- */
-
-double lda_inference(document* doc, lda_model* model, double* var_gamma, double** phi)
-{
-    double converged = 1;
-    double phisum = 0, likelihood = 0;
-    double likelihood_old = 0, oldphi[model->num_topics];
-    int k, n, var_iter;
-    double digamma_gam[model->num_topics];
-
-    // compute posterior dirichlet
-
-    for (k = 0; k < model->num_topics; k++)
-    {
-        var_gamma[k] = model->alpha + (doc->total/((double) model->num_topics));
-        digamma_gam[k] = digamma(var_gamma[k]);
-        for (n = 0; n < doc->length; n++)
-            phi[n][k] = 1.0/model->num_topics;
-    }
-    var_iter = 0;
-
-    while ((converged > VAR_CONVERGED) && ((var_iter < VAR_MAX_ITER) || (VAR_MAX_ITER == -1)))
-    {
-	var_iter++;
-	for (n = 0; n < doc->length; n++)
-	{
-            phisum = 0;
-            for (k = 0; k < model->num_topics; k++)
-            {
-                oldphi[k] = phi[n][k];
-                phi[n][k] =
-                    digamma_gam[k] +
-                    model->log_prob_w[k][doc->words[n]];
-
-                if (k > 0)
-                    phisum = log_sum(phisum, phi[n][k]);
-                else
-                    phisum = phi[n][k]; // note, phi is in log space
-            }
-
[…]

### 2011-01-13 icsm09/lda-c-dist/runit/data/split.sh

@@ -1,1 +1,0 @@
-sed -re 's/^([0-9\-]*[\t ])*//' ev1 | split -d -a 3 - ev1.

### 2011-01-13 icsm09/presentation/parser.pl

@@ -1,68 +1,0 @@
-use strict;
-my $first = 1;
-my $depth = 0;
-my $lastitem = 0;
-while(<>) {
-	chomp;
-	s/  /\t/g;
-        $_ = filter_stars($_);
-	if (/^\=$/) {
-		$depth = processDepth($depth,"");
-		print '\end{itemize}',$/;
-		$lastitem = 0;
-	} elsif (/^\=(.+)$/) {
-		$depth = processDepth($depth,"");
-		if ($first) {
-			$first = 0;
-		} else {
-			print '\end{itemize}',$/;
-		}
-		print "\\sslide{$1}$/";
-		print "\\begin{itemize}$/";
-		$lastitem = 0;
-	} elsif (/^(\t*)\s*\*\s*(.*)$/) {
-		print STDERR "=====> $_$/";
-		$depth = processDepth($depth,$1);
-		my $i = $2;
-		print ("\t"x$depth);
-		print "\\item $i$/";
-		$lastitem = 1;
-	} else {
-		$lastitem = 0;
-		print "$_$/";
-	}
-}
-sub processDepth {
-	my ($depth,$one) = @_;
-	my $length = length($one);
-	if ($depth!=$length) {
-		if ($depth > $length) {
-			my $d = $depth - $length;
-			foreach (1..($depth-$length)) {
-				print ("\t"x($d - $_));
-				print "\\end{itemize}$/";
-			}
-		} else {
-			foreach (1..($length-$depth)) {
-				print ("\t"x($depth+$_));
-				print "\\item " if ($_ >= 1 && !$lastitem);
-				print "\\begin{itemize}$/";
-			}
-		}
-		$depth = $length;
-	}
-	print STDERR "$depth $length\n";
-	return $depth;
-}
-sub filter_stars {
-    my $line = $_[0];
-    if ($line =~ /^(\*+)\*/) {
-        warn "FILTER $line";
-        my $replace = "\t" x (length($1));
-        $replace .= "*";
-        $line =~ s/^\*+//;
-        $line = $replace . $line
-    }
-    return $line;
-}
-

### 2011-01-13 msr/msr2011.tex
Neil's modifications to the exp2 word lists (and minor other changes).

@@
 \alignauthor John Mylopoulos \\
-   \affaddr{Dept. of Computer Science}\\
-   \affaddr{University of Toronto}\\
-   \affaddr{Toronto, Ontario, CANADA}\\
-   \email{jm@cs.toronto.edu}
+   \affaddr{Dept. Information Eng. and Computer Science}\\
+   \affaddr{University of Trento}\\
+   \affaddr{Trento, Italy}\\
+   \email{jm@disi.unitn.it}
 }
@@
 \newcommand{\shrink}{}
-
 \section{Introduction}
@@
 %XXXXXXXXXXXX OPTIONAL
- Hindle et al.
+% NE 0113 this seems too early to mention 'result'.
+% Hindle et al.
 \cite{Hindle2007} examined release patterns in OSS.
@@
 MySQL contains 320 thousand lines of C and C++ source code.  
-Choosing an older version of these projects allowed us to focus on projects which have moved into the maintenance phase of the software life-cycle.
+Choosing an older version of these projects allowed us to focus on projects which have moved further into the maintenance phase of the software life-cycle.
 
@@
 For example,
-% in our previous work~\cite{Hindle09ICSM}, as well as -- CITED a little too often
+% in our previous work~\cite{Hindle09ICSM}, as well as -- CITED a little too often (vanity cites!)
 in Baldi et al.~\cite{Baldi2008}, topics are named manually: human experts read the highest-frequency members of a topic and assign a label accordingly. 
@@
 The labels are reasonable enough, but still require an expert in the field to determine them. 
-We sought to automate the process of naming the topics.
-
-Accordingly, we tried to associate each topic with a label from a list
-of keywords and related terms. These keywords were determined before
-hand and were not extracted from the project themselves.
+Furthermore, these labels are project-specific, because they are generated from the data of that project. E.g., we might have a label called `Oracle' in the MySQL case, since Oracle owns MySQL.
+
+In this paper, we had two objectives.
+One, we sought to automate the process of naming the topics.
+Two, we sought to label topics with independent terms, in order to facilitate comparison between projects.
+
+Accordingly, we associated each topic with a label from a list
+of keywords (`wordlists', in our parlance) and related terms. These wordlists were determined a priori and were not extracted from the projects themselves.
 We intersected the words of the topics and the words of our word-lists.
-We `named' a topic if any of its words matched any of the word-list's words.
+We `labeled' a topic if any of its words matched any of the word-list's words.
 A topic could match more than one keyword.
@@
 
-Our second word list set, \textsf{exp2}, relied on the ISO quality model, ISO9126~\cite{iso9126}. 
-ISO9126 describes six high-level non-functional requirements (listed in Table \ref{tbl:wnsig}). 
-There is some debate about the significance and importance of the terms in this model. 
+Our second word-list, \textsf{exp2}, is derived from the ISO quality model, ISO9126~\cite{iso9126}. 
+ISO9126 describes six high-level non-functional requirements (NFRs), maintainability, functionality, portability, efficiency, usability, and reliability. 
+There is some debate about the terms in this model. 
 However, ISO9126 is ``an international standard and thus provides an internationally accepted terminology for software quality \cite[p.
 58]{Boegh2008},'' that is sufficient for the purposes of this research.
-We use these qualities later on as classes in supervised labelling.
-The terms extracted from ISO9126 may not capture all words associated with the labels.  
+%We use these qualities later on as classes in supervised labelling. NE confusing the issue here, I think.
+The terms from ISO9126 may not capture all words occuring in the topics that are nonetheless associated with one of the NFRs. 
 For example, the term ``redundancy'' is one most would agree is relevant to discussion of reliability, but is not in the standard. 
-We therefore took the words from the taxonomy and expanded them.
-
-\todo[inline]{* clarify the mailinglist stuff we did}
-\todo[inline]{This part doesn't make a lot of sense}
-To construct these expanded word lists, we used WordNet~\cite{Fellbaum1998}, an English-language ``lexical database'' that contains semantic relations between words, including common related forms (similar to word stemming), meronymy and synonymy.
-We then added Boehm's 1976 software quality model \cite{Boehm+:1976:ICSE}, and classified his eleven `ilities' into their respective ISO9126 qualities. 
+We therefore took the NFRs from the ISO9126 and added to them.
+
+%\todo[inline]{* clarify the mailinglist stuff we did}
+%\todo[inline]{This part doesn't make a lot of sense}
+To construct these expanded word-lists, we used WordNet~\cite{Fellbaum1998}, an English-language ``lexical database'' that contains semantic relations between words, including common related forms (similar to word stemming), meronymy and synonymy.
+We then added Boehm's software quality model \cite{Boehm+:1976:ICSE}, and classified his eleven `ilities' into their respective ISO9126 NFRs. 
 We did the same for the quality model produced by McCall et al. \cite{mccall1977}. 
-Finally, we analyzed two mailing lists from another software project to expand our set with domain-specific terms. 
-For example, we add the term ``performance'' to the synonyms for \emph{efficiency}, since this term occurs in most mail messages that discuss efficiency.
+We then did a simple random analysis of mailing list messages from an open source project, KDE.
+If we judged a given message to contain terms that were related to one of the NFRs in ISO9126, we added it to our word-list.
+This allowed us to expand our word-lists with more software-specific terms.
+%Finally, we randomly analyzed two mailing lists from another software project to expand our set with domain-specific terms. 
+% Less info is probably better here...
+%For example, we add the term ``performance'' to the synonyms for \emph{efficiency}, since this term occurs in most mail messages that discuss efficiency.
+Table \ref{tbl:wnsig} shows the labels (NFRs) and word-lists we used for matching.
 
@@
 We do not include these words here for space considerations (but see the Appendix for our data repository). 
-It is not clear the words associated with our labels are specific enough. For example, the label \emph{maintainability} is associated with words \emph{ease} and \emph{ownership}.
+It is not clear the words associated with our labels in \textsf{exp3} are specific enough.
+For example, the label \emph{maintainability} is associated with words \emph{ease} and \emph{ownership}.
+In general, as we proceed from word-list in \textsf{exp1} to that in \textsf{exp3}, our lists get more generic.
 

### 2011-01-14 msr/msr2011.tex
minor wording changes, added draft intro.

@@
    \affaddr{University of Trento}\\
-   \affaddr{Trento, Italy}\\
-   \email{jm@disi.unitn.it}
+   \affaddr{Trento, ITALY}\\
+   \email{\normalsize jm@disi.unitn.it}
 }
@@
 techniques use unsupervised machine-learning algorithms to suggest
-candidate word-lists that recur together often.  However, word-lists are
+candidate word-lists.  However, word-lists are
 difficult to 
@@
 
+\todo[inline]{neil: possible reword: Topic modeling is a machine learning technique which creates multinomial distributions over words found in a text corpus.
+Topic models are useful in software engineering because they can identify, from text corpora such as source code, change logs, email, and others, co-occuring words.
+For example, a topic extracted from a set of changelogs might be ``listener change remove add fire''.
+A topic model gives developers an overview of where activity is occurring and gives managers or maintainers a sense of project history.
+A major challenge is to accurately identify the label for a topic (in the previous case, perhaps ``event-handling'').
+Current approaches are manual: they rely on human expertise to recognize the appropriate label.
+Furthermore, to date the results of topic modeling are project specific, and not generalizable.
+}
+
 % An attempt Motivate topics
 %XXX is this the correct terminology for commits/revisions
-The development of software does not occur in a vacuum, revisions to
-source code are often not independent of the past, the specification or
+%The development of software does not occur in a vacuum, 
+Revisions to
+source code are not independent of past revisions, the specification or
 the task at hand.
@@
 Many of these purposes and tasks exhibit themselves as latent topics which cross-cut and compose many of development artifacts: version control revisions, bug reports, and mailing-list discussions.
+\todo[inline]{NeiL: I still find this clumsily worded}
 
@@
 Developers often interact with abstractions of the artifacts from which topics are derived, rather than the artifacts themselves.
-Such abstractions include project dashboards~\cite{kersten2005mylar}, tags~\cite{treude2010}, and complexity metrics \cite{mccabe1976complexity}. 
+Such abstractions include project dashboards~\cite{kersten2005mylar} and tags~\cite{treude2010}%, and complexity metrics \cite{mccabe1976complexity}. 
 Without the structured information which dashboards typically rely on, topics become a valuable and automatically extractable abstraction of software artifacts.
@@
 They do not use external domain-specific information as we do.
-In \cite{ernst10refsq}, we describe our earlier project, similar to this, to identify change in quality requirements in GNOME software projects; our approach is solely text-matching, however, and does not leverage machine learning strategies.
+%In \cite{ernst10refsq}, we describe our earlier project, similar to this, to identify change in quality requirements in GNOME software projects; our approach is solely text-matching, however, and does not leverage machine learning strategies.
 
@@
 
-\todo[inline]{Make it clear where the data comes from! Neil says: Also, I
-  didn't get the impression they understood that we weren't generating
-  labels from the data, but rather from independent taxonomies. Of
-  course if you use labels from the data it isn't comparable. But the
-  point of ISO 9126 etc. was to avoid exactly this problem.}
-
-\todo[inline]{Need to hammer it in that the word bags are not project specific and
- are inherited so it isn't a supervised method}
+% %\todo[inline]{Make it clear where the data comes from! Neil says: Also, I
+%   didn't get the impression they understood that we weren't generating
+%   labels from the data, but rather from independent taxonomies. Of
+%   course if you use labels from the data it isn't comparable. But the
+%   point of ISO 9126 etc. was to avoid exactly this problem.}
+
+% \todo[inline]{Need to hammer it in that the word bags are not project specific and
+%  are inherited so it isn't a supervised method}
 
 Topics are word distributions, lists of words ranked by frequency, that are burdensome to interpret and hard to distinguish and understand. 
-While the topic models themselves are generated automatically, how to interpret these topics is less clear.
+While the topics themselves are generated automatically, how to interpret these topics is less clear.
 For example,
@@
 in Baldi et al.~\cite{Baldi2008}, topics are named manually: human experts read the highest-frequency members of a topic and assign a label accordingly. 
-Given the word list \emph{``listener change remove add fire''}, Baldi et al. would assign the label \emph{event-handling}. 
+Given the topic \emph{``listener change remove add fire''}, Baldi et al. would assign the label \emph{event-handling}. 
 The labels are reasonable enough, but still require an expert in the field to determine them. 
@@
 
-In this paper, we had two objectives.
-One, we sought to automate the process of naming the topics.
-Two, we sought to label topics with independent terms, in order to facilitate comparison between projects.
+This paper does this differently. First of all, we automated the process of naming the topics. Secondly, we labelled topics with project-independent terms, in order to permit cross-project comparison.
 
@@
 There are roughly 20 times the number of topics as there are periods per project. 
-This is because we told LDA to extract 20 topics per period. 
+This is because LDA extracted 20 topics per period. 
 All experiments were run on MaxDB 7.500 and MySQL 3.23 data.
@@
 \toprule
-Project & Measure & \textsf{exp1} & \textsf{exp2} & \textsf{exp3} \\
+\textbf{Project} & \textbf{Measure} & \textsf{exp1} & \textsf{exp2} & \textsf{exp3} \\
 \midrule
@@
                                 % maxdb has 80 empty topics
-MaxDB 7.500 & Unnamed Topics & 219  & 375 &  172  \\
-MaxDB 7.500 & Total  Topics  & 500 & 500 & 500 \\
+ & Unnamed Topics & 219  & 375 &  172  \\
+ & Total  Topics  & 500 & 500 & 500 \\
 \midrule
 MySQL 3.23  & Named Topics   & 524 & 273 & 773 \\
-MySQL 3.23  & Unnamed Topics & 476 & 727 & 227 \\
-MySQL 3.23  & Total  Topics  & 1000 & 1000 & 1000 \\
+  & Unnamed Topics & 476 & 727 & 227 \\
+  & Total  Topics  & 1000 & 1000 & 1000 \\
 
@@
   \centering
- \includegraphics[width=\textwidth]{figures/unsupervised-bar}
+ \includegraphics[width=0.6\textwidth]{figures/unsupervised-bar}
  \caption{Performance, ROC values (range: 0-1), of unsupervised topic labelling for

### 2011-01-14 msr/msr2011.tex
prototype with neil's intro. It seems much simpler to me. Maybe expand a little on the benefits.

@@
 
-\todo[inline]{neil: possible reword: Topic modeling is a machine learning technique which creates multinomial distributions over words found in a text corpus.
+Topic modeling is a machine learning technique which creates multinomial distributions over words found in a text corpus.
 Topic models are useful in software engineering because they can identify, from text corpora such as source code, change logs, email, and others, co-occuring words.
@@
 Furthermore, to date the results of topic modeling are project specific, and not generalizable.
-}
-
-% An attempt Motivate topics
-%XXX is this the correct terminology for commits/revisions
-%The development of software does not occur in a vacuum, 
-Revisions to
-source code are not independent of past revisions, the specification or
-the task at hand.
-Often the purpose of revisions to source code relate to one or more topics, whether it be a maintenance activity, a quality issue, a requirement or feature or a response to a bug report.
-Many of these purposes and tasks exhibit themselves as latent topics which cross-cut and compose many of development artifacts: version control revisions, bug reports, and mailing-list discussions.
-\todo[inline]{NeiL: I still find this clumsily worded}
-
-Topics in software engineering practice refer to specific shared threads of discussions between project stakeholders. 
-These topics can include particular requirements, maintenance requests, bug fixes or other software engineering tasks.
-%NEIL: define topic specifically as akin to email 'thread'.
-It sounded too vague.
-Also changed from 'developers' to 'stakeholders' because developers sounds too coding oriented.
-Our work could apply to testers, users, etc.
-While addressing a topic of development, stakeholders often produce or modify multiple development artifacts.
-%
-These artifacts can include source code, bug reports, revisions to source code, and test cases. 
-%
-Topics are typically latent, i.e., not explicitly named, and can be in several potential states: they may not be immediately resolved, they may re-occur, yet others are only dealt with briefly  (e.g., a bug report that is closed)
-Within this paper we define \emph{topics} as latent word distributions that describe and compose the documents (version control commits) being analyzed. 
-We define \emph{labelled topics} as \emph{topics} that have been interpreted and labelled. 
-The distinction is important because word distributions can consistent of many words, thus if one has many topics, reading and interpreting these words is prohibitive.
-For example, if you read only 10 words per topic from 500 topics you have to read and interpret 5000 words, this is 5 or more pages of text!
-Thus \emph{labelled topics} save interpretation time.
-
-There are many practical applications of \emph{labelled topic
-  extraction} 
-ranging from summarization, project dashboards, and software process recovery.
-These methods are useful for providing abstractions of development activities and artifacts.
-Particularly in large projects, summarization of developer activities is essential to maintenance. 
-Rather than looking for specific concepts in a project~\cite{marcus04wcre}, we are interested in understanding a software project relative to common properties of all software.
-We can improve project dashboards, used by managers, by showing both the current topics of development as well those that relate to domain-independent taxonomies, such as NFRs.
-Some quality assurance questions can be addressed by seeking evidence within repositories that NFR-related topics are being addressed.
-Topic modelling can be used to check if development is on track and focused.
-For outside users of the software, our technique can provide insight into that company's development practices in order to answer questions such as, ``Are they focused on usability?''
-Within the organization \emph{labelled topic extraction} can act as a historical sanity check, used to avoid \emph{confirmation bias}, which is the difference between what people think happened and what actually happened.
-Some software releases, for instance, are poorly received since they are quite buggy. 
-Our technique would help explain why this was the case, since it can identify NFRs which are dominating developer attention.
-Thus developers can use these techniques to get an overview of the current and past development topics.
-For awareness, managers can use these techniques to gain an overview of topics of development.
-
-% This is a very long example it'd be nice if it was tighter
-% Perhaps development has been sidetracked by cascading issues that appeared during implementation.
-% For instance, a developer adding an optimization to a RDBMS might have to change multiple modules in order to provide that optimization code with enough information to operate. 
-% This issue might trigger a series of cascading changes that might be difficult to discover when trying to explain why the project's release date has been missed. 
-% Labelled Topic extraction could be used to detect this distinct focus of development.
-
-% semi-automatically sounds too uncertain. I think the introduction can be more muscular and bold. Don't give the reviewer any room to start doubting.
-
-% why are topics useful
-Developers often interact with abstractions of the artifacts from which topics are derived, rather than the artifacts themselves.
-Such abstractions include project dashboards~\cite{kersten2005mylar} and tags~\cite{treude2010}%, and complexity metrics \cite{mccabe1976complexity}. 
-Without the structured information which dashboards typically rely on, topics become a valuable and automatically extractable abstraction of software artifacts.
-Topics abstract software artifacts by their subject matter, rather than their concrete representation. 
-Much like aspect-oriented programming, topic labeling can identify cross-cutting concerns in a software project (as in~\cite{Baldi2008}).
-The topics of development are relevant because they identify the distinct issues facing stakeholders at a particular time.
-
-The goal of concept location and topic analysis research is to recover these topics and the related software artifacts.
-% get neil's data mining point in there
-To date, the results of topic analysis and concept location are project specific, and not generalizable.
-% get the interpretation aspect
-Much of the current research stops short of automatically interpreting the purpose of the topic, e.g.~\cite{Baldi2008,Kuhn2007,marcus04wcre}. 
-That is, extracted topics are usually manually named and labelled by those who read the topic words.
-
-% topics are numerous?
-Deriving and labelling topics is a challenging machine learning problem. 
-Topic analysis produces % raw topics that are represented as
- unlabelled context-free word-lists or word distributions.
-Topics have to be identified by interpreting the prevalent words in a word distribution and by inspecting related documents. 
-This manual approach is impractical when handling hundreds of different topics.
-To scale to large projects (thousands of lines of code and multiple releases), we need automatic assistance to determine the topic label.
-We propose a method of interpreting and labelling topics, and their related artifacts, that is generalizable and relevant to multiple projects.
+
+Changes to software systems, revisions, do not occur without
+reason. 
+The reason or purpose is often related to requirements 
+Often one revision might have multiple purposes ranging from
+maintenance issues, quality issues, requirements, features or a
+response to a bug report.
+These purposes often are exhibited as latent topics that relate to
+many development artifacts such as:
+ version control revisions, bug reports, and mailing-list discussions.
 
@@
 Instead, they developed a supervised classifier by using human experts to identify an NFR training set. 
-Our worked differed because we had a more comprehensive set of terms based on the taxonomy we chose.
+Our work differed because we had a more comprehensive set of terms based on the taxonomy we chose.
 We also wanted a common taxonomy that allows us to make cross-project comparison; this was not one of their objectives.
@@
 MaxDB started in the late 1970s as a research project, and was later acquired by SAP. 
-As of version 7.500, released April 2007, the project has 940 thousand lines of C source code\footnote{generated using David A. Wheeler's \emph{SLOCCount}.}. 
+As of version 7.500, released April 2007, the project has 940 thousand lines of C source code\footnote{generated using David A. Wheeler's \emph{SLOCCount}, dwheeler.com/sloccount.}. 
 The MySQL project started in 1994 and MySQL 3.23 was released in early 2001. 
@@
 For example, for the MaxDB topic consisting of a message ``exit() only used in non NPTL LINUX Versions'', we tagged that topic \emph{portability}. 
+%\todo[inline]{this confused people. Is it truly unsupervised? Yes, but it needs to be clearer.}
 We compared against this data-set, but we also used the data-set for our supervised machine learning based topic classification. 
@@
 A \emph{named topic} is a topic with a matching label. 
-There are roughly 20 times the number of topics as there are periods per project. 
-This is because LDA extracted 20 topics per period. 
-All experiments were run on MaxDB 7.500 and MySQL 3.23 data.
+All experiments were run on MaxDB 7.500 and MySQL 3.23 data. LDA
+extracted 20 topics per period for each.
+% Is this good enough
+This labelling is unsupervised because the corpus is not derived from 
+the project being labelled and we did not label the project's topics
+ourselves for a training set. The theory behind this technique is that
+because most software often addresses similar issues, we can use the
+domain knowledge of software to label relevant topics.
+\todo[inline]{Neil does this explain the unsupervised aspect well enough?}
+
 Table \ref{tbl:wordlist} shows how many topics were labelled for MaxDB
@@
 Figure \ref{fig:maxdb-unsup-results} shows our results for MaxDB and MySQL. 
-Our performance is measured using the Receiver Operating Characteristic area-under-curve value~\cite{Fawcett2006861}, abbreviated \emph{ROC}. 
+\todo[inline]{NEIL: this section should be updated to explain F-measure}
+Our performance is measured using the Receiver Operating
+Characteristic area-under-curve value~\cite{Fawcett2006861},
+abbreviated \emph{ROC}, and the F-measure. 
 ROC values provide a score, similar to school letter-grades (A is 0.9, C is 0.6), reflecting how well a particular learner performed for the given data. 
@@
 The area under the ROC of a classifier is equivalent to the probability that the classifier will rank a randomly chosen positive instance higher than a randomly chosen negative instance.
+F-measure is the harmonic mean of precision and recall.
 
@@
 % Rebuttal: ROC is not new and it is what the machine learning community, search based software engineering and now the MSR community are using +  it has natural meaningful quality mappings.
-\begin{figure*}
+\begin{figure}
   \centering
- \includegraphics[width=0.6\textwidth]{figures/unsupervised-bar}
+ \includegraphics[width=0.45\textwidth]{figures/unsupervised-bar}
  \caption{Performance, ROC values (range: 0-1), of unsupervised topic labelling for
-   each NFR and per word-list. This graph shows how well the
+   each NFR and per word-list. Dashed line indicates the performance of a random classifier. This graph shows how well the
    unsupervised topic labelling matched our manual annotations.}
@@
   \label{fig:maxdb-unsup-results}
-\end{figure*}
-
-% EXP2
+\end{figure}
+
+\begin{figure}
+  \centering
+ \includegraphics[width=0.45\textwidth]{figures/unsupervised-bar-f1}
+ \caption{Performance, F1 values (range: 0-1), of unsupervised topic labelling for
+   each NFR and per word-list.}
+  \label{fig:maxdb-unsup-results-f1}
+\end{figure}
+
+%\todo[inline]{where is exp1 report?}
+
+%EXP1 could not be evaluated properly!
+Because our ground truth annotations were only relevant to ISO-9126,
+we estimate that \textsf{exp1} had poor
+performance via the overlap between ISO-9126 and the Kayed ontology.
+For \textsf{exp1} the F-measures for MaxDB were from $0$ to $0.18$ with an average
+of $0.03$, and for MySQL were from $0$ to $0.16$ with an average of
+$0.05$.
+% 0.177215189873 / 6
+% .02953586497883333333
+% (0.0298507462687 + 0.0661157024793 + 0.0634920634921 + 0 + 0.115384615385) / 6
+% .04580718793751666666
+
+% EXP2 source is naming-paper/src/validate/exp2.maxdb.report.txt
+% (change as needed). Second column (after the name) is F1.
 For \textsf{exp2}, the average F-measure (macro-F1) for MaxDB was $0.24$ with a range $0.091$ to
@@
 
-Based on these results we find that \emph{reliability} and \emph{usability} worked well for MaxDB in \textsf{exp2} and better in \textsf{exp3}. 
+Based on these results we find that \emph{reliability} and
+\emph{usability} worked well for MaxDB in \textsf{exp2} and better in
+\textsf{exp3}. 
+\textsf{exp1} performed poorly.
 MySQL had reasonable results within \textsf{exp2} for \emph{reliability} and \emph{efficiency}. 
@@
 If a \emph{C} grade performance has a ROC value of $0.6$ then most of these tests scored a grade of \emph{C} or less, but our classifier nonetheless performed substantially better than random.
-
-% \begin{comment}
-% maxdb
-
-% portability,0.773778920308,0.228070175439,76,12,0.181684100119,0.52,0.14606741573,0.553033707865,288,0.96,13
-% efficiency,0.832904884319,0.21686746988,36,29,0.124642098177,0.236842105263,0.2,0.557848837209,315,0.915697674419,9
-% reliability,0.714652956298,0.379888268156,7,104,0.340426350577,0.246376811594,0.829268292683,0.765208858985,244,0.701149425287,34
-% functionality,0.902313624679,0.0952380952381,32,6,0.0834236909204,0.25,0.0588235294118,0.520961060481,349,0.983098591549,2
-% maintainability,0.796915167095,0.0919540229885,76,3,0.122505166175,0.571428571429,0.05,0.520145631068,306,0.990291262136,4
-% usability,0.830334190231,0.175,11,55,0.138109307712,0.112903225806,0.388888888889,0.620320455226,316,0.851752021563,7
-
-% total,0.808483290488,0.235897435897,238,209,0.126930463023,0.248201438849,0.224755700326,0.560823829443,1818,0.896891958559,69mysql
-
-% portability,0.616040955631,0.137931034483,225,0,0.211497874858,1.0,0.0740740740741,0.537037037037,343,1.0,18
-% efficiency,0.924914675768,0.3125,33,11,0.297817974114,0.47619047619,0.232558139535,0.606150156324,532,0.979742173112,10
-% reliability,0.726962457338,0.411764705882,61,99,0.242477699079,0.361290322581,0.478632478632,0.633772529295,370,0.788912579957,56
-% functionality,0.457337883959,0.0245398773006,315,3,0.00597425429824,0.571428571429,0.012539184953,0.500651614948,264,0.988764044944,4
-% maintainability,0.686006825939,0,184,0,0,0,0.0,0.5,402,1.0,0
-% usability,0.919795221843,0.175438596491,27,20,0.13508225345,0.2,0.15625,0.560074458484,534,0.963898916968,5
-
-% total,0.721843003413,0.159793814433,845,133,0.0857635592199,0.411504424779,0.0991471215352,0.523778370698,2445,0.9484ahindle@wish:~/projects/lda-paper/src/validate$ cat NOTES 
-% columns of maxdb.report.text
-
-%     out['accuracy'] = acc
-%     out['f1'] = f1
-%     out['fn'] = fn
-%     out['fp'] = fp
-%     out['mcc'] = mcc
-%     out['precision'] = pre
-%     out['recall'] = rec
-%     out['roca'] = roca
-%     out['tnr'] = tnr
-%     out['tn'] = tn
-%     out['tp'] = tp
-
-% cat exp3.*.report.txt 
-% maxdb
-
-% portability,0.706940874036,0.472222222222,38,76,0.286382363909,0.40157480315,0.573033707865,0.659850187266,224,0.746666666667,51
-% efficiency,0.372750642674,0.222929936306,10,234,0.067549899325,0.130111524164,0.777777777778,0.548772609819,110,0.31976744186,35
-% reliability,0.434447300771,0.256756756757,3,217,0.195966959965,0.149019607843,0.926829268293,0.651633024951,131,0.376436781609,38
-% functionality,0.53470437018,0.236286919831,6,175,0.186902353269,0.137931034483,0.823529411765,0.665285832643,180,0.507042253521,28
-% maintainability,0.627249357326,0.337899543379,43,102,0.111672018872,0.26618705036,0.4625,0.566201456311,207,0.669902912621,37
-% usability,0.27763496144,0.107936507937,1,280,0.0937953415493,0.0572390572391,0.944444444444,0.594863731656,91,0.245283018868,17
-
-% total,0.492287917738,0.257983719474,101,1084,0.0926022435775,0.159689922481,0.671009771987,0.568114654124,943,0.46521953626,206
-% mysql
-
-% portability,0.616040955631,0.412532637076,164,61,0.170139234681,0.564285714286,0.325102880658,0.57363015753,282,0.822157434402,79
-% efficiency,0.532423208191,0.184523809524,12,262,0.124342460899,0.105802047782,0.720930232558,0.619212814253,281,0.517495395948,31
-% reliability,0.588737201365,0.404938271605,35,206,0.209194239572,0.284722222222,0.700854700855,0.630811145737,263,0.560767590618,82
-% functionality,0.622866894198,0.651968503937,112,109,0.240455709256,0.655063291139,0.648902821317,0.620331560471,158,0.591760299625,207
-% maintainability,0.621160409556,0.218309859155,153,69,-0.00390279540262,0.31,0.16847826087,0.498418234912,333,0.828358208955,31
-% usability,0.358361774744,0.104761904762,10,366,0.0128981856383,0.0567010309278,0.6875,0.513425090253,188,0.339350180505,22
-
-% total,0.556598407281,0.367032074706,486,1073,0.058598680978,0.296393442623,0.481876332623,0.532831106575,1505,0.583785880528,452
-% \end{comment}
 

### 2011-01-14 msr/msr2011.tex
prototype with neil's intro. It seems much simpler to me. Maybe expand a little on the benefits.

--- 2011-01-14 msr/msr2011.tex
@@
 Furthermore, to date the results of topic modeling are project specific, and not generalizable.
-}
-
-% An attempt Motivate topics
-%XXX is this the correct terminology for commits/revisions
-%The development of software does not occur in a vacuum, 
-Revisions to
-source code are not independent of past revisions, the specification or
-the task at hand.
-Often the purpose of revisions to source code relate to one or more topics, whether it be a maintenance activity, a quality issue, a requirement or feature or a response to a bug report.
-Many of these purposes and tasks exhibit themselves as latent topics which cross-cut and compose many of development artifacts: version control revisions, bug reports, and mailing-list discussions.
-\todo[inline]{NeiL: I still find this clumsily worded}
-
-Topics in software engineering practice refer to specific shared threads of discussions between project stakeholders. 
-These topics can include particular requirements, maintenance requests, bug fixes or other software engineering tasks.
+
+% % An attempt Motivate topics
+% %XXX is this the correct terminology for commits/revisions
+% %The development of software does not occur in a vacuum, 
+% Revisions to
+% source code are not independent of past revisions, the specification or
+% the task at hand.
+% Often the purpose of revisions to source code relate to one or more topics, whether it be a maintenance activity, a quality issue, a requirement or feature or a response to a bug report.
+% Many of these purposes and tasks exhibit themselves as latent topics which cross-cut and compose many of development artifacts: version control revisions, bug reports, and mailing-list discussions.
+% \todo[inline]{NeiL: I still find this clumsily worded}
+
+% Topics in software engineering practice refer to specific shared threads of discussions between project stakeholders. 
+% These topics can include particular requirements, maintenance requests, bug fixes or other software engineering tasks.
 %NEIL: define topic specifically as akin to email 'thread'.
@@
 Our work could apply to testers, users, etc.
-While addressing a topic of development, stakeholders often produce or modify multiple development artifacts.
-%
-These artifacts can include source code, bug reports, revisions to source code, and test cases. 
-%
-Topics are typically latent, i.e., not explicitly named, and can be in several potential states: they may not be immediately resolved, they may re-occur, yet others are only dealt with briefly  (e.g., a bug report that is closed)
-Within this paper we define \emph{topics} as latent word distributions that describe and compose the documents (version control commits) being analyzed. 
-We define \emph{labelled topics} as \emph{topics} that have been interpreted and labelled. 
-The distinction is important because word distributions can consistent of many words, thus if one has many topics, reading and interpreting these words is prohibitive.
-For example, if you read only 10 words per topic from 500 topics you have to read and interpret 5000 words, this is 5 or more pages of text!
-Thus \emph{labelled topics} save interpretation time.
-
-There are many practical applications of \emph{labelled topic
-  extraction} 
-ranging from summarization, project dashboards, and software process recovery.
-These methods are useful for providing abstractions of development activities and artifacts.
-Particularly in large projects, summarization of developer activities is essential to maintenance. 
-Rather than looking for specific concepts in a project~\cite{marcus04wcre}, we are interested in understanding a software project relative to common properties of all software.
-We can improve project dashboards, used by managers, by showing both the current topics of development as well those that relate to domain-independent taxonomies, such as NFRs.
-Some quality assurance questions can be addressed by seeking evidence within repositories that NFR-related topics are being addressed.
-Topic modelling can be used to check if development is on track and focused.
-For outside users of the software, our technique can provide insight into that company's development practices in order to answer questions such as, ``Are they focused on usability?''
-Within the organization \emph{labelled topic extraction} can act as a historical sanity check, used to avoid \emph{confirmation bias}, which is the difference between what people think happened and what actually happened.
-Some software releases, for instance, are poorly received since they are quite buggy. 
-Our technique would help explain why this was the case, since it can identify NFRs which are dominating developer attention.
-Thus developers can use these techniques to get an overview of the current and past development topics.
-For awareness, managers can use these techniques to gain an overview of topics of development.
+% While addressing a topic of development, stakeholders often produce or modify multiple development artifacts.
+% %
+% These artifacts can include source code, bug reports, revisions to source code, and test cases. 
+% %
+% Topics are typically latent, i.e., not explicitly named, and can be in several potential states: they may not be immediately resolved, they may re-occur, yet others are only dealt with briefly  (e.g., a bug report that is closed)
+% Within this paper we define \emph{topics} as latent word distributions that describe and compose the documents (version control commits) being analyzed. 
+% We define \emph{labelled topics} as \emph{topics} that have been interpreted and labelled. 
+% The distinction is important because word distributions can consistent of many words, thus if one has many topics, reading and interpreting these words is prohibitive.
+% For example, if you read only 10 words per topic from 500 topics you have to read and interpret 5000 words, this is 5 or more pages of text!
+% Thus \emph{labelled topics} save interpretation time.
+
+% There are many practical applications of \emph{labelled topic
+%   extraction} 
+% ranging from summarization, project dashboards, and software process recovery.
+% These methods are useful for providing abstractions of development activities and artifacts.
+% Particularly in large projects, summarization of developer activities is essential to maintenance. 
+% Rather than looking for specific concepts in a project~\cite{marcus04wcre}, we are interested in understanding a software project relative to common properties of all software.
+% We can improve project dashboards, used by managers, by showing both the current topics of development as well those that relate to domain-independent taxonomies, such as NFRs.
+% Some quality assurance questions can be addressed by seeking evidence within repositories that NFR-related topics are being addressed.
+% Topic modelling can be used to check if development is on track and focused.
+% For outside users of the software, our technique can provide insight into that company's development practices in order to answer questions such as, ``Are they focused on usability?''
+% Within the organization \emph{labelled topic extraction} can act as a historical sanity check, used to avoid \emph{confirmation bias}, which is the difference between what people think happened and what actually happened.
+% Some software releases, for instance, are poorly received since they are quite buggy. 
+% Our technique would help explain why this was the case, since it can identify NFRs which are dominating developer attention.
+% Thus developers can use these techniques to get an overview of the current and past development topics.
+% For awareness, managers can use these techniques to gain an overview of topics of development.
 
@@
 % why are topics useful
-Developers often interact with abstractions of the artifacts from which topics are derived, rather than the artifacts themselves.
-Such abstractions include project dashboards~\cite{kersten2005mylar} and tags~\cite{treude2010}%, and complexity metrics \cite{mccabe1976complexity}. 
-Without the structured information which dashboards typically rely on, topics become a valuable and automatically extractable abstraction of software artifacts.
-Topics abstract software artifacts by their subject matter, rather than their concrete representation. 
-Much like aspect-oriented programming, topic labeling can identify cross-cutting concerns in a software project (as in~\cite{Baldi2008}).
-The topics of development are relevant because they identify the distinct issues facing stakeholders at a particular time.
-
-The goal of concept location and topic analysis research is to recover these topics and the related software artifacts.
-% get neil's data mining point in there
-To date, the results of topic analysis and concept location are project specific, and not generalizable.
-% get the interpretation aspect
-Much of the current research stops short of automatically interpreting the purpose of the topic, e.g.~\cite{Baldi2008,Kuhn2007,marcus04wcre}. 
-That is, extracted topics are usually manually named and labelled by those who read the topic words.
-
-% topics are numerous?
-Deriving and labelling topics is a challenging machine learning problem. 
-Topic analysis produces % raw topics that are represented as
- unlabelled context-free word-lists or word distributions.
-Topics have to be identified by interpreting the prevalent words in a word distribution and by inspecting related documents. 
-This manual approach is impractical when handling hundreds of different topics.
+% Developers often interact with abstractions of the artifacts from which topics are derived, rather than the artifacts themselves.
+% Such abstractions include project dashboards~\cite{kersten2005mylar} and tags~\cite{treude2010}%, and complexity metrics \cite{mccabe1976complexity}. 
+% Without the structured information which dashboards typically rely on, topics become a valuable and automatically extractable abstraction of software artifacts.
+% Topics abstract software artifacts by their subject matter, rather than their concrete representation. 
+% Much like aspect-oriented programming, topic labeling can identify cross-cutting concerns in a software project (as in~\cite{Baldi2008}).
+% The topics of development are relevant because they identify the distinct issues facing stakeholders at a particular time.
+
+% The goal of concept location and topic analysis research is to recover these topics and the related software artifacts.
+% % get neil's data mining point in there
+% To date, the results of topic analysis and concept location are project specific, and not generalizable.
+% % get the interpretation aspect
+% Much of the current research stops short of automatically interpreting the purpose of the topic, e.g.~\cite{Baldi2008,Kuhn2007,marcus04wcre}. 
+% That is, extracted topics are usually manually named and labelled by those who read the topic words.
+
+% % topics are numerous?
+% Deriving and labelling topics is a challenging machine learning problem. 
+% Topic analysis produces % raw topics that are represented as
+%  unlabelled context-free word-lists or word distributions.
+% Topics have to be identified by interpreting the prevalent words in a word distribution and by inspecting related documents. 
+% This manual approach is impractical when handling hundreds of different topics.
 To scale to large projects (thousands of lines of code and multiple releases), we need automatic assistance to determine the topic label.
-We propose a method of interpreting and labelling topics, and their related artifacts, that is generalizable and relevant to multiple projects.
+% We propose a method of interpreting and labelling topics, and their related artifacts, that is generalizable and relevant to multiple projects.
 

### 2011-01-14 msr/msr2011.tex
add todo on unsup/sup distinctin
New F1 plot. Some text changes, added a few todos.

--- 2011-01-14 msr/msr2011.tex
@@
 For example, for the MaxDB topic consisting of a message ``exit() only used in non NPTL LINUX Versions'', we tagged that topic \emph{portability}. 
+\todo[inline]{this confused people. Is it truly unsupervised? Yes, but it needs to be clearer.}
 We compared against this data-set, but we also used the data-set for our supervised machine learning based topic classification. 
@@
 A \emph{named topic} is a topic with a matching label. 
-There are roughly 20 times the number of topics as there are periods per project. 
-This is because LDA extracted 20 topics per period. 
-All experiments were run on MaxDB 7.500 and MySQL 3.23 data.
+All experiments were run on MaxDB 7.500 and MySQL 3.23 data. LDA extracted 20 topics per period for each.
+
 Table \ref{tbl:wordlist} shows how many topics were labelled for MaxDB
@@
 Figure \ref{fig:maxdb-unsup-results} shows our results for MaxDB and MySQL. 
+\todo[inline]{NEIL: this section should be updated to explain F-measure}
 Our performance is measured using the Receiver Operating Characteristic area-under-curve value~\cite{Fawcett2006861}, abbreviated \emph{ROC}. 
@@
   \label{fig:maxdb-unsup-results}
-\end{figure*}
-
-% EXP2
+\end{figure}
+
+\begin{figure}
+  \centering
+ \includegraphics[width=0.45\textwidth]{figures/unsupervised-bar-f1}
+ \caption{Performance, F1 values (range: 0-1), of unsupervised topic labelling for
+   each NFR and per word-list.}
+  \label{fig:maxdb-unsup-results-f1}
+\end{figure}
+
+\todo[inline]{where is exp1 report?}
+
+% EXP2 source is naming-paper/src/validate/exp2.maxdb.report.txt (change as needed). Second column is F1.
 For \textsf{exp2}, the average F-measure (macro-F1) for MaxDB was $0.24$ with a range $0.091$ to

### 2011-01-21 msr/msr2011.tex
minor todo added
minor edits for ICSE reviewer

@@
 
+\todo[inline]{more on the case study early on as a way of showing our technique is useful for insight into dev activities}
+
 % An attempt Motivate topics

### 2011-01-21 msr/msr2011.tex

--- 2011-01-14 msr/msr2011.tex
@@
 %   and Functionality/Correctness for fixes
-% * Some security changes came later
+% * some security changes came later
 % * Mysql 3.23 transitioned from a project being developed to

### 2011-01-21 msr/msr2011.tex
exp1 added, an intro todo handled about clarity, unsupervised explained

--- 2011-01-21 msr/msr2011.tex
@@
 %The development of software does not occur in a vacuum, 
-Revisions to
-source code are not independent of past revisions, the specification or
-the task at hand.
-Often the purpose of revisions to source code relate to one or more topics, whether it be a maintenance activity, a quality issue, a requirement or feature or a response to a bug report.
-Many of these purposes and tasks exhibit themselves as latent topics which cross-cut and compose many of development artifacts: version control revisions, bug reports, and mailing-list discussions.
-\todo[inline]{NeiL: I still find this clumsily worded}
+% Revisions to
+% source code are not independent of past revisions, the specification or
+% the task at hand.
+% Often the purpose of revisions to source code relate to one or more topics, whether it be a maintenance activity, a quality issue, a requirement or feature or a response to a bug report.
+% Many of these purposes and tasks exhibit themselves as latent topics which cross-cut and compose many of development artifacts: version control revisions, bug reports, and mailing-list discussions.
+%
+Changes to software systems, revisions, do not occur without
+reason. 
+The reason or purpose is often related to requirements 
+Often one revision might have multiple purposes ranging from
+maintenance issues, quality issues, requirements, features or a
+response to a bug report.
+These purposes often are exhibited as latent topics that relate to
+many development artifacts such as:
+ version control revisions, bug reports, and mailing-list discussions.
+%\todo[inline]{NeiL: I still find this clumsily worded}
 
@@
 For example, for the MaxDB topic consisting of a message ``exit() only used in non NPTL LINUX Versions'', we tagged that topic \emph{portability}. 
-\todo[inline]{this confused people. Is it truly unsupervised? Yes, but it needs to be clearer.}
+%\todo[inline]{this confused people. Is it truly unsupervised? Yes, but it needs to be clearer.}
 We compared against this data-set, but we also used the data-set for our supervised machine learning based topic classification. 
@@
 A \emph{named topic} is a topic with a matching label. 
-All experiments were run on MaxDB 7.500 and MySQL 3.23 data. LDA extracted 20 topics per period for each.
+All experiments were run on MaxDB 7.500 and MySQL 3.23 data. LDA
+extracted 20 topics per period for each.
+% Is this good enough
+This labelling is unsupervised because the corpus is not derived from 
+the project being labelled and we did not label the project's topics
+ourselves for a training set. The theory behind this technique is that
+because most software often addresses similar issues, we can use the
+domain knowledge of software to label relevant topics.
+\todo[inline]{Neil does this explain the unsupervised aspect well enough?}
 
@@
 
-\todo[inline]{where is exp1 report?}
-
-% EXP2 source is naming-paper/src/validate/exp2.maxdb.report.txt (change as needed). Second column is F1.
+%\todo[inline]{where is exp1 report?}
+
+%EXP1 could not be evaluated properly!
+Because our ground truth annotations were only relevant to ISO-9126,
+we estimate that \textsf{exp1} had poor
+performance via the overlap between ISO-9126 and the Kayed ontology.
+For \textsf{exp1} the F-measures for MaxDB were from $0$ to $0.18$ with an average
+of $0.03$, and for MySQL were from $0$ to $0.16$ with an average of
+$0.05$.
+% 0.177215189873 / 6
+% .02953586497883333333
+% (0.0298507462687 + 0.0661157024793 + 0.0634920634921 + 0 + 0.115384615385) / 6
+% .04580718793751666666
+
+% EXP2 source is naming-paper/src/validate/exp2.maxdb.report.txt
+% (change as needed). Second column (after the name) is F1.
 For \textsf{exp2}, the average F-measure (macro-F1) for MaxDB was $0.24$ with a range $0.091$ to

### 2011-01-21 src/validate/check_validity.py

@@ -21,14 +21,14 @@ def load_period(period, data_dir = "../../data/maxdb-tagged/"):
     return topics
  
 
-def load_wordlists(expdir = '../exp3/' ): 
-    """ load the datafile with the wordlists we want to use """
-    quality_map = {'portability':[], 'efficiency':[], 'reliability':[], 'functionality':[], 'usability':[], 'maintainability':[]}
-    for q in quality_map.keys():
-        q_file = open( expdir + 'wordlist.' + q )
-        for line in q_file:
-            quality_map[ q ].append( line.rstrip() )
-    return quality_map
+#def load_wordlists(expdir = '../exp3/' ): 
+#    """ load the datafile with the wordlists we want to use """
+#    quality_map = {'portability':[], 'efficiency':[], 'reliability':[], 'functionality':[], 'usability':[], 'maintainability':[]}
+#    for q in quality_map.keys():
+#        q_file = open( expdir + 'wordlist.' + q )
+#        for line in q_file:
+#            quality_map[ q ].append( line.rstrip() )
+#    return quality_map
     
 def create_element_list(topic):
     """Parse each period and create a list of elements and annotations in that topic, 

### 2011-01-21 msr/msr2011.tex
moved discussion of findings to discussion section ensmallened the author emails revised title updated sources of error

--- 2011-01-21 msr/msr2011.tex
@@
 Abram Hindle\\
-   \affaddr{Dept.of Computer Science}\\
+   \affaddr{Dept. of Computer Science}\\
    \affaddr{University of California, Davis} \\
    \affaddr{Davis, CA, USA}\\
-   \email{abram@softwareprocess.es}
+   \email{\large abram@softwareprocess.es}
 \alignauthor  Neil A. Ernst\\
@@
    \affaddr{Toronto, Ontario, CANADA}\\
-   \email{nernst@cs.toronto.edu}
+   \email{\large nernst@cs.toronto.edu}
 \and
@@
    \affaddr{Waterloo, Ontario, CANADA}\\
-   \email{migod@uwaterloo.ca}
+   \email{\large migod@uwaterloo.ca}
 \alignauthor John Mylopoulos \\
@@
    \affaddr{Trento, ITALY}\\
-   \email{\normalsize jm@disi.unitn.it}
+   \email{\large jm@disi.unitn.it}
 }
@@
 %\title{What's in a name? Automated topic naming of software maintenance activities}
-\title{Automated topic naming to support analysis of software maintenance activities}
+\title{Automated topic naming to support cross-project analysis of software maintenance activities}
 \maketitle
@@
 
-Topic modeling is a machine learning technique which creates multinomial distributions over words found in a text corpus.
-Topic models are useful in software engineering because they can identify, from text corpora such as source code, change logs, email, and others, co-occuring words.
-For example, a topic extracted from a set of changelogs might be ``listener change remove add fire''.
-A topic model gives developers an overview of where activity is occurring and gives managers or maintainers a sense of project history.
-A major challenge is to accurately identify the label for a topic (in the previous case, perhaps ``event-handling'').
+Topic modeling is a machine learning technique which creates multinomial distributions over words from a text corpus.
+Topic models are useful in software engineering because they can identify co-occuring words from text corpora such as source code, commit comments, email, and others.
+For example, a topic extracted from a set of commit logs might be\emph{ ``listener change remove add fire''}.
+A topic model is a summary of the key concepts in the corpus.
+Among other uses, it can quickly give developers an overview of where activity is occurring and gives managers or maintainers a sense of project history.
+A major challenge is to accurately identify the summary label for a topic (in the previous case, perhaps \emph{``event-handling''}).
 Current approaches are manual: they rely on human expertise to recognize the appropriate label.
 Furthermore, to date the results of topic modeling are project specific, and not generalizable.
+This paper presents an approach, labelled topic extraction, which addresses these two problems.
 
@@
 
-\subsection{Annotation observations}
-We found many topics that were not non-functional requirements (NFRs) but were often related to them. 
-For instance, concurrency was mentioned often in the commit logs and was related to \emph{correctness} and \emph{reliability}, likely because concurrent code is prone to bugs such as race conditions.
-%are
-% often  was troublesome. 
-\emph{Configuration management} and \emph{source control} related changes appeared often; % and sometimes there were topics dedicated to configuration management. 
-these kinds of changes are slightly related to \emph{maintainability}. 
-A non-functional change that was not quality-related was licensing and copyright; many changes were simply to do with updating copyrights or ensuring copyright or license headers were applied to files.
-In these cases we assigned the \emph{None} class.
-
-We noticed that occasionally the names of modules would conflict with words related to other non-functional requirements. 
-For instance, optimizers are very common modules in database systems: both MySQL and MaxDB have optimizer modules. 
-In MySQL the optimizer is mentioned but often the change addresses  correctness or another quality. 
-Despite this difference, the name of the module could fool our learners into believing the change was always about \emph{efficiency}. 
-In these cases the advantages of tailoring topic names to specific project terminologies are more clear. 
-Project specific word-lists would avoid automated mistakes due to the names of entities and modules of a software project.
-
-\subsection{Summary of techniques}
-While an unsupervised technique such as LDA is appealing in its lack of human intervention, and thus lower effort, 
-supervised learners have the advantage of domain knowledge, which typically means improved results. 
-Our manual annotations were fairly quick to do, taking only a few minutes per 30 day period.
-
-Very rarely did \textsf{exp2} and \textsf{exp3} (naive word matching) ever perform as well as the supervised machine learners. 
-For MaxDB, \textit{reliability} was slightly better detected using the static word list of \textsf{exp2}. 
-In general, the machine learners and \textsf{exp3} did better than \textsf{exp2} for both MaxDB and MySQL. 
-For both MySQL and MaxDB \textit{usability} was better served by \textsf{exp2}. 
-\textit{Usability} was a very infrequent label, however, which made it difficult to detect in any case.
-
-We found that the multi-label learners of BR, CLR and HOMER did not do as well for Macro-ROC as NaiveBayes and other naive Bayes derived learners. 
-This suggests that by combining together multiple NaiveBayes learners we could probably label sets of topics effectively, but it would require a separate NaiveBayes learner per label.
-
 \section{Understanding software maintenance activities} 
@@
 %\subsection{Effectiveness}
+\subsection{Annotation observations}
+We found many topics that were not non-functional requirements (NFRs) but were often related to them. 
+For instance, concurrency was mentioned often in the commit logs and was related to \emph{correctness} and \emph{reliability}, likely because concurrent code is prone to bugs such as race conditions.
+%are
+% often  was troublesome. 
+\emph{Configuration management} and \emph{source control} related changes appeared often; % and sometimes there were topics dedicated to configuration management. 
+these kinds of changes are slightly related to \emph{maintainability}. 
+A non-functional change that was not quality-related was licensing and copyright; many changes were simply to do with updating copyrights or ensuring copyright or license headers were applied to files.
+In these cases we assigned the \emph{None} class.
+
+We noticed that occasionally the names of modules would conflict with words related to other non-functional requirements. 
+For instance, optimizers are very common modules in database systems: both MySQL and MaxDB have optimizer modules. 
+In MySQL the optimizer is mentioned but often the change addresses  correctness or another quality. 
+Despite this difference, the name of the module could fool our learners into believing the change was always about \emph{efficiency}. 
+In these cases the advantages of tailoring topic names to specific project terminologies are more clear. 
+Project specific word-lists would avoid automated mistakes due to the names of entities and modules of a software project.
+
+\subsection{Summary of techniques}
+While an unsupervised technique such as LDA is appealing in its lack of human intervention, and thus lower effort, 
+supervised learners have the advantage of domain knowledge, which typically means improved results. 
+Our manual annotations were fairly quick to do, taking only a few minutes per 30 day period.
+
+Very rarely did \textsf{exp2} and \textsf{exp3} (naive word matching) ever perform as well as the supervised machine learners. 
+For MaxDB, \textit{reliability} was slightly better detected using the static word list of \textsf{exp2}. 
+In general, the machine learners and \textsf{exp3} did better than \textsf{exp2} for both MaxDB and MySQL. 
+For both MySQL and MaxDB \textit{usability} was better served by \textsf{exp2}. 
+\textit{Usability} was a very infrequent label, however, which made it difficult to detect in any case.
+
+We found that the multi-label learners of BR, CLR and HOMER did not do as well for Macro-ROC as NaiveBayes and other naive Bayes derived learners. 
+This suggests that by combining together multiple NaiveBayes learners we could probably label sets of topics effectively, but it would require a separate NaiveBayes learner per label.
 
@@
 % can always be cut down
-Our work faced multiple threats to validity and we attempted to address many of them.
+%Our work faced multiple threats to validity and we attempted to address many of them.
 
 \emph{Construct validity} -- we used only commit messages rather than mail or bug tracker messages.
-Conversely, construct validity was addressed in our validation as we used the files, topics and revisions to annotate the commit messages. 
-We used mailing-lists to verify the purpose behind some events.
-We also chose our taxonomy and the data to study. We did not show these results to the stakeholders.
-We only studied database systems.
-
-\emph{Internal validity} -- includes inter-rater reliability and our lack of rival explanation building. 
+Since we required a seamless record, it was not possible to obtain other data sources.
+Possibly they would have influenced our results, but there would be a degree of correlation between the corpora.
+%Conversely, construct validity was addressed in our validation as we used the files, topics and revisions to annotate the commit messages. 
+%We used mailing-lists to verify the purpose behind some events.
+Our taxonomy for software NFRs is subject to dispute, but seems generally accepted.
+Finally, there are exogenous sources, such as in-person discussions, which we did not access, an omission as noted in \cite{aranda09icse}.
+%We did not show these results to the stakeholders.
+
+\emph{Internal validity} -- %includes inter-rater reliability and our lack of rival explanation building. 
 We improved internal validity by trying to correlate and explain the behaviours observed in the analysis with the historical records of the projects.
-We did not attempt to match our results to any particular model or theory.
-
-\emph{External validity} was addressed by our use of multiple case studies, but there are still issues. 
+We did not attempt to match our results to any particular model or theory, as such theories are sparse in
+
+\emph{External validity} -- %was addressed by our use of two case studies, but there are still issues. 
 Our data originated from OSS database projects and thus might not be applicable to commercially developed software. 
 Furthermore, our analysis techniques rely on a project's use of meaningful commit messages. The OSS projects investigated were database systems, thus 
-our results might not generalize to other domains.
+our results might not generalize to other domains. %We only studied database systems.
 % \todo[inline]{If we compare with the same domain we don't get to say it generalizes a lot?}
@@
 % in Case Study Research: Design Case Studies
-\emph{Reliability} was addressed by each annotator following the same protocol and using the same annotations. 
-Unfortunately only two annotators were used; their annotations could be biased as we did not analyze for inter-rater reliability.
+\emph{Reliability} -- each annotator followed the same protocol and used the same annotations. 
+However, only two annotators were used; their annotations could be biased as we did not analyze for inter-rater reliability.
 

### 2011-01-21 msr/msr2011.tex
moved discussion of findings to discussion section ensmallened the author emails revised title updated sources of error

@@
 
-% % An attempt Motivate topics
-% %XXX is this the correct terminology for commits/revisions
-% %The development of software does not occur in a vacuum, 
-% Revisions to
-% source code are not independent of past revisions, the specification or
-% the task at hand.
-% Often the purpose of revisions to source code relate to one or more topics, whether it be a maintenance activity, a quality issue, a requirement or feature or a response to a bug report.
-% Many of these purposes and tasks exhibit themselves as latent topics which cross-cut and compose many of development artifacts: version control revisions, bug reports, and mailing-list discussions.
-% \todo[inline]{NeiL: I still find this clumsily worded}
-
-% Topics in software engineering practice refer to specific shared threads of discussions between project stakeholders. 
-% These topics can include particular requirements, maintenance requests, bug fixes or other software engineering tasks.
-%NEIL: define topic specifically as akin to email 'thread'.
-It sounded too vague.
-Also changed from 'developers' to 'stakeholders' because developers sounds too coding oriented.
-Our work could apply to testers, users, etc.
-% While addressing a topic of development, stakeholders often produce or modify multiple development artifacts.
-% %
-% These artifacts can include source code, bug reports, revisions to source code, and test cases. 
-% %
-% Topics are typically latent, i.e., not explicitly named, and can be in several potential states: they may not be immediately resolved, they may re-occur, yet others are only dealt with briefly  (e.g., a bug report that is closed)
-% Within this paper we define \emph{topics} as latent word distributions that describe and compose the documents (version control commits) being analyzed. 
-% We define \emph{labelled topics} as \emph{topics} that have been interpreted and labelled. 
-% The distinction is important because word distributions can consistent of many words, thus if one has many topics, reading and interpreting these words is prohibitive.
-% For example, if you read only 10 words per topic from 500 topics you have to read and interpret 5000 words, this is 5 or more pages of text!
-% Thus \emph{labelled topics} save interpretation time.
-
-% There are many practical applications of \emph{labelled topic
-%   extraction} 
-% ranging from summarization, project dashboards, and software process recovery.
-% These methods are useful for providing abstractions of development activities and artifacts.
-% Particularly in large projects, summarization of developer activities is essential to maintenance. 
-% Rather than looking for specific concepts in a project~\cite{marcus04wcre}, we are interested in understanding a software project relative to common properties of all software.
-% We can improve project dashboards, used by managers, by showing both the current topics of development as well those that relate to domain-independent taxonomies, such as NFRs.
-% Some quality assurance questions can be addressed by seeking evidence within repositories that NFR-related topics are being addressed.
-% Topic modelling can be used to check if development is on track and focused.
-% For outside users of the software, our technique can provide insight into that company's development practices in order to answer questions such as, ``Are they focused on usability?''
-% Within the organization \emph{labelled topic extraction} can act as a historical sanity check, used to avoid \emph{confirmation bias}, which is the difference between what people think happened and what actually happened.
-% Some software releases, for instance, are poorly received since they are quite buggy. 
-% Our technique would help explain why this was the case, since it can identify NFRs which are dominating developer attention.
-% Thus developers can use these techniques to get an overview of the current and past development topics.
-% For awareness, managers can use these techniques to gain an overview of topics of development.
-
-% This is a very long example it'd be nice if it was tighter
-% Perhaps development has been sidetracked by cascading issues that appeared during implementation.
-% For instance, a developer adding an optimization to a RDBMS might have to change multiple modules in order to provide that optimization code with enough information to operate. 
-% This issue might trigger a series of cascading changes that might be difficult to discover when trying to explain why the project's release date has been missed. 
-% Labelled Topic extraction could be used to detect this distinct focus of development.
-
-% semi-automatically sounds too uncertain. I think the introduction can be more muscular and bold. Don't give the reviewer any room to start doubting.
-
-% why are topics useful
-% Developers often interact with abstractions of the artifacts from which topics are derived, rather than the artifacts themselves.
-% Such abstractions include project dashboards~\cite{kersten2005mylar} and tags~\cite{treude2010}%, and complexity metrics \cite{mccabe1976complexity}. 
-% Without the structured information which dashboards typically rely on, topics become a valuable and automatically extractable abstraction of software artifacts.
-% Topics abstract software artifacts by their subject matter, rather than their concrete representation. 
-% Much like aspect-oriented programming, topic labeling can identify cross-cutting concerns in a software project (as in~\cite{Baldi2008}).
-% The topics of development are relevant because they identify the distinct issues facing stakeholders at a particular time.
-
-% The goal of concept location and topic analysis research is to recover these topics and the related software artifacts.
-% % get neil's data mining point in there
-% To date, the results of topic analysis and concept location are project specific, and not generalizable.
-% % get the interpretation aspect
-% Much of the current research stops short of automatically interpreting the purpose of the topic, e.g.~\cite{Baldi2008,Kuhn2007,marcus04wcre}. 
-% That is, extracted topics are usually manually named and labelled by those who read the topic words.
-
-% % topics are numerous?
-% Deriving and labelling topics is a challenging machine learning problem. 
-% Topic analysis produces % raw topics that are represented as
-%  unlabelled context-free word-lists or word distributions.
-% Topics have to be identified by interpreting the prevalent words in a word distribution and by inspecting related documents. 
-% This manual approach is impractical when handling hundreds of different topics.
-To scale to large projects (thousands of lines of code and multiple releases), we need automatic assistance to determine the topic label.
-% We propose a method of interpreting and labelling topics, and their related artifacts, that is generalizable and relevant to multiple projects.
+Changes to software systems, revisions, do not occur without
+reason. 
+The reason or purpose is often related to requirements 
+Often one revision might have multiple purposes ranging from
+maintenance issues, quality issues, requirements, features or a
+response to a bug report.
+These purposes often are exhibited as latent topics that relate to
+many development artifacts such as:
+ version control revisions, bug reports, and mailing-list discussions.
 
@@
 \todo[inline]{NEIL: this section should be updated to explain F-measure}
-Our performance is measured using the Receiver Operating Characteristic area-under-curve value~\cite{Fawcett2006861}, abbreviated \emph{ROC}. 
+Our performance is measured using the Receiver Operating
+Characteristic area-under-curve value~\cite{Fawcett2006861},
+abbreviated \emph{ROC}, and the F-measure. 
 ROC values provide a score, similar to school letter-grades (A is 0.9, C is 0.6), reflecting how well a particular learner performed for the given data. 
@@
 %   and Functionality/Correctness for fixes
-% * some security changes came later
+% * Some security changes came later
 % * Mysql 3.23 transitioned from a project being developed to

### 2011-01-21 msr/msr2011.tex
minor fixes, removed commented intro for clarity, fixing wording

@@
 % %\todo[inline]{Topics and LDA sooner, Topics/concept location is a known
-% %  research area and we're saying they don't do enough}
-
-% % Maybe specify the hard work we did annotating the data
-
-% % our field
-% %  Within the field of mining software repositories, a
-% To understand the latent subject matter of software development artifacts, researchers have proposed topic modeling and concept location.
-% %These techniques rely on the automated application of machine learning algorithms to corpora.
-% These techniques use unsupervised machine learning algorithms find topics or concepts.
-% % the problem
-% These topics are word-lists lacking semantic interpretation, which are difficult to distinguish and interpret. 
-% %Results are usually abstract, as the topics that are automatically extracted 
-% % the solution 
-% In order to have a meaningful discussion about the topics of software development, we must be able to attach meaningful labels to these topics. 
-% %However, current software engineering topic analysis and concept analysis research relies on manual naming and labelling of topics.
-% % the labelling research
-% Current topic labelling approaches are manual, and do not use domain-specific knowledge to improve, contextualize, or describe results for the developers.
-% % a sub problem
-% % While too much specificity can produce non-generalizable results, too little produces broad learners that do not provide much immediately useful detail. 
-% % what we do
-
-% To solve this problem, this paper introduces \emph{labelled topic extraction}, 
-% in which topics are extracted from commit comments and given labels relating to a cross-project taxonomy. 
-% % data mining generalizable
-% Our approach is 
-% %To ensure we have a 
-% generalizable and project-agnostic, as we label topics using non-functional requirements,
-% % these qualities 
-% which apply across software systems.% and their development artifacts.
-% % We used real world data
-% We evaluated our approach with an experimental study on two large-scale database projects, MySQL and MaxDB. 
-% % We used LDA, CVS, Bitkeeper
-% We extracted topics using Latent Dirichlet Allocation (LDA) from the commit-log comments of their source control systems (CVS and BitKeeper). 
-% % We show NFRs exist
-% %Our results were generalizable across the two projects, showing that non-functional requirements were commonly discussed, 
-% %and we identified topic trends over time. 	
-% % we can label stuff
-% Our labelled topic extraction technique produces appropriate, context-sensitive labels relevant to these projects, providing fresh insight into their evolving software development activities.
-
 Researchers have employed a variety of techniques to extract underlying
@@
 insight into their evolving software development activities.
-
-% Researchers have used topic modeling and concept location to understand the latent topics of software development artifacts.
-% These techniques use unsupervised machine-learning algorithms to recover topics.
-% These topics are word-lists and are difficult to distinguish and interpret. 
-% Topics are not meaningful until they have been named or interpreted.
-% Current topic labelling approaches are manual, and do not use domain-specific knowledge to improve, contextualize, or describe results for the developers.
-% We propose a solution: \emph{labelled topic extraction}.
-% Topics are extracted using Latent Dirichlet Allocation (LDA) from commit-log comments recovered from source control systems such as CVS and BitKeeper. 
-% These topics are given labels relating to a generalizable cross-project taxonomy consisting of non-functional requirements.
-% Our approach was evaluated  with experiments and case studies on two large-scale RDBMS projects: MySQl and MaxDB.
-% Labelled topic extraction produces appropriate, context-sensitive labels relevant to these projects, which provides fresh insight into their evolving software development activities.
-
 \end{abstract}
@@
 This paper presents an approach, labelled topic extraction, which addresses these two problems.
-
-Changes to software systems, revisions, do not occur without
-reason. 
-The reason or purpose is often related to requirements 
-Often one revision might have multiple purposes ranging from
-maintenance issues, quality issues, requirements, features or a
-response to a bug report.
-These purposes often are exhibited as latent topics that relate to
-many development artifacts such as:
- version control revisions, bug reports, and mailing-list discussions.
 
@@
 Cleland-Huang and her colleagues published work on mining requirements documents for non-functional requirements (NFR) (quality requirements)~\cite{Cleland-Huang2006}. 
-Their approach was similar to ours, as they mined keywords from their own NFR catalogues~\cite{chung99}.
+Their approach was similar to ours, as they mined keywords from NFR catalogues.
 %One approach they tried was similar to this one, with keywords mined from NFR catalogues found in their previous work~\cite{chung99}. 
@@
 Our work differed because we had a more comprehensive set of terms based on the taxonomy we chose.
-We also wanted a common taxonomy that allows us to make cross-project comparison; this was not one of their objectives.
-The objective of Cleland-Huang's study was to identify new NFRs for system development, yet our objective was to recover those latent NFRs from commit-log messages of the project. 
+We make cross-project comparisons, while they focused on a single project.
+%The objective of Cleland-Huang's study was to identify new NFRs for system development, yet our objective was to recover those latent NFRs from commit-log messages of the project. 
 Another difference was that our data was far less structured than the requirements documents used in their study.
@@
 
-In the same vein, Mockus and Votta~\cite{Mockus00} studied a large-scale industrial change-tracking system. 
-They also leveraged WordNet, but only for word roots as they felt the synonyms would be non-specific and cause errors. 
+Similarly, Mockus and Votta~\cite{Mockus00} studied a large-scale industrial change-tracking system. 
+They also leveraged WordNet, but only for word roots as they felt the synonyms would be non-specific and cause errors, problems we also encountered. 
 Mockus et al. had access to system developers, with whom they could validate their labels. 
-Since we try to bridge different organizations, these kind of interviews are infeasible (particularly in the distributed world of open-source software).
+Since we try to bridge different organizations, instead of a single project, these kind of interviews are infeasible (particularly in the distributed world of open-source software).
 
 The other approach is to extract concerns from software repositories.
-Marcus et al.~\cite{marcus04wcre} describe their use of Latent Semantic Indexing to identify commonly occurring concerns for software maintenance. 
+Marcus et al.~\cite{marcus04wcre} describe their use of Latent Semantic Indexing (LSI) to identify commonly occurring concerns for software maintenance.
+The concerns are given by the user, and LSI used to retrieve them from a corpus.
+Topic modelling generates topics that are independent of a user query, and relate only to word frequencies in the corpus.
 %Some results were interesting, but their precision was quite low. 
@@
 They do not use external domain-specific information as we do.
-%In \cite{ernst10refsq}, we describe our earlier project, similar to this, to identify change in quality requirements in GNOME software projects; our approach is solely text-matching, however, and does not leverage machine learning strategies.
+
+Finally, in \cite{ernst10refsq}, we describe our earlier project, similar to this, to identify change in quality requirements in GNOME software projects; our approach is solely text-matching, however, and does not leverage machine learning strategies.
 
@@
 % this reduces the line length and drops a ref even though we still have a scacchi kickin around
-Massey~\cite{massey02icse} and Scacchi~\cite{scacchi05b} looked at the topic of requirements in open-source software. 
-Their work discusses the source of the requirements and how they are used in the development process.
+%Massey~\cite{massey02icse} and Scacchi~\cite{scacchi05b} looked at the topic of requirements in open-source software. 
+%Their work discusses the source of the requirements and how they are used in the development process.
 German~\cite{german03gnome} looked at GNOME specifically, and listed several sources for requirements: leader interest, mimicry, brainstorming, and prototypes.
@@
 Figure \ref{fig:process} gives an outline of the methodology followed. 
-We began by gathering source data. 
-We then applied unsupervised and supervised learning techniques to identify topics. 
+We began by gathering source data and creating topic models. We created labels using a well-known NFR taxonomy. 
+We then applied unsupervised and supervised learning techniques to try to match our labels to the topics. 
 These topics were used to analyze the role of non-functional requirements in software maintenance.
@@
 To evaluate our approach, we sought candidate systems that were mature projects and had openly accessible source control repositories. 
-We selected systems from the same application domain, so the functional requirements would be broadly similar. 
+We selected systems from the same application domain, to control for differences in functional, rather than non-functional, requirements. 
 We selected MySQL 3.23 and MaxDB 7.500 as they were open-source, partially-commercial database systems. 
@@
 For example, for the MaxDB topic consisting of a message ``exit() only used in non NPTL LINUX Versions'', we tagged that topic \emph{portability}. 
-%\todo[inline]{this confused people. Is it truly unsupervised? Yes, but it needs to be clearer.}
+\todo[inline]{this confused people. Is it truly unsupervised? Yes, but it needs to be clearer.}
 We compared against this data-set, but we also used the data-set for our supervised machine learning based topic classification. 
@@
 A \emph{named topic} is a topic with a matching label. 
-All experiments were run on MaxDB 7.500 and MySQL 3.23 data. LDA
-extracted 20 topics per period for each.
-% Is this good enough
-This labelling is unsupervised because the corpus is not derived from 
-the project being labelled and we did not label the project's topics
-ourselves for a training set. The theory behind this technique is that
-because most software often addresses similar issues, we can use the
-domain knowledge of software to label relevant topics.
-\todo[inline]{Neil does this explain the unsupervised aspect well enough?}
+All experiments were run on MaxDB 7.500 and MySQL 3.23 data. LDA extracted 20 topics per period for each.
 
@@
 
-\noindent \textbf{Analysis of the unsupervised labelling} -- Based on the labels, and our manual topic labelling, 
-we compared the results of the unsupervised word matching approach. 
-For each quality we assessed whether unsupervised labels matched the manual annotations.
+\noindent \textbf{Analysis of the unsupervised labelling} -- %Based on the labels, and our manual topic labelling, 
+%we compared the results of the unsupervised word matching approach. 
+For each quality we assessed whether unsupervised labels matched the manual annotations. To be clear, the manual annotations were not used to train the labeling process.
 Figure \ref{fig:maxdb-unsup-results} shows our results for MaxDB and MySQL. 
 \todo[inline]{NEIL: this section should be updated to explain F-measure}
-Our performance is measured using the Receiver Operating
-Characteristic area-under-curve value~\cite{Fawcett2006861},
-abbreviated \emph{ROC}, and the F-measure. 
+Our performance is measured using the Receiver Operating Characteristic area-under-curve value~\cite{Fawcett2006861}, abbreviated \emph{ROC}. 
 ROC values provide a score, similar to school letter-grades (A is 0.9, C is 0.6), reflecting how well a particular learner performed for the given data. 
@@
 A ROC result of 0.5 would be equivalent to a random learner (that is, issuing as many false positives as true positives). 
-The area under the ROC of a classifier is equivalent to the probability that the classifier will rank a randomly chosen positive instance higher than a randomly chosen negative instance.
-F-measure is the harmonic mean of precision and recall.
+The %area under the -- NE confusing to have area again ... and not importantana
+ROC of a classifier is equivalent to the probability that the classifier will rank a randomly chosen positive instance higher than a randomly chosen negative instance.
 
@@
 
-%\todo[inline]{where is exp1 report?}
-
-%EXP1 could not be evaluated properly!
-Because our ground truth annotations were only relevant to ISO-9126,
-we estimate that \textsf{exp1} had poor
-performance via the overlap between ISO-9126 and the Kayed ontology.
-For \textsf{exp1} the F-measures for MaxDB were from $0$ to $0.18$ with an average
-of $0.03$, and for MySQL were from $0$ to $0.16$ with an average of
-$0.05$.
-% 0.177215189873 / 6
-% .02953586497883333333
-% (0.0298507462687 + 0.0661157024793 + 0.0634920634921 + 0 + 0.115384615385) / 6
-% .04580718793751666666
-
-% EXP2 source is naming-paper/src/validate/exp2.maxdb.report.txt
-% (change as needed). Second column (after the name) is F1.
+\todo[inline]{where is exp1 report?}
+
+% EXP2 source is naming-paper/src/validate/exp2.maxdb.report.txt (change as needed). Second column is F1.
 For \textsf{exp2}, the average F-measure (macro-F1) for MaxDB was $0.24$ with a range $0.091$ to
@@
 
-Based on these results we find that \emph{reliability} and
-\emph{usability} worked well for MaxDB in \textsf{exp2} and better in
-\textsf{exp3}. 
-\textsf{exp1} performed poorly.
+Based on these results we find that \emph{reliability} and \emph{usability} worked well for MaxDB in \textsf{exp2} and better in \textsf{exp3}. 
 MySQL had reasonable results within \textsf{exp2} for \emph{reliability} and \emph{efficiency}. 

### 2011-01-21 msr/msr2011.tex
minor fixes, removed commented intro for clarity, fixing wording

@@
 
+Changes to software systems, revisions, do not occur without
+reason. 
+The reason or purpose is often related to requirements 
+Often one revision might have multiple purposes ranging from
+maintenance issues, quality issues, requirements, features or a
+response to a bug report.
+These purposes often are exhibited as latent topics that relate to
+many development artifacts such as:
+ version control revisions, bug reports, and mailing-list discussions.
+
 In our previous work we dealt with topic trends, which are topics that recur over time~\cite{Hindle09ICSM}. 
@@
 A ROC result of 0.5 would be equivalent to a random learner (that is, issuing as many false positives as true positives).
+
 The %area under the -- NE confusing to have area again ... and not importantana
 ROC of a classifier is equivalent to the probability that the classifier will rank a randomly chosen positive instance higher than a randomly chosen negative instance.
+F-measure is the harmonic mean of precision and recall.
 
@@
 % Our fourth experiment evaluated the effectiveness of multi-label learners, such as HOMER. 
-Both kinds of learners performed well with area under the ROC curve values between $0.6$ and $0.8$.
+Both kinds of learners performed well with 
+%area under the 
+ROC values between $0.6$ and $0.8$.
 

### 2011-01-21 msr/msr2011.tex
More about F-measure

--- 2011-01-21 msr/msr2011.tex
@@
 Abram Hindle\\
-   \affaddr{Dept. of Computer Science}\\
+   \affaddr{Dept.of Computer Science}\\
    \affaddr{University of California, Davis} \\
    \affaddr{Davis, CA, USA}\\
-   \email{\large abram@softwareprocess.es}
+   \email{abram@softwareprocess.es}
 \alignauthor  Neil A. Ernst\\
@@
    \affaddr{Toronto, Ontario, CANADA}\\
-   \email{\large nernst@cs.toronto.edu}
+   \email{nernst@cs.toronto.edu}
 \and
@@
    \affaddr{Waterloo, Ontario, CANADA}\\
-   \email{\large migod@uwaterloo.ca}
+   \email{migod@uwaterloo.ca}
 \alignauthor John Mylopoulos \\
@@
    \affaddr{Trento, ITALY}\\
-   \email{\large jm@disi.unitn.it}
+   \email{\normalsize jm@disi.unitn.it}
 }
@@
 %\title{What's in a name? Automated topic naming of software maintenance activities}
-\title{Automated topic naming to support cross-project analysis of software maintenance activities}
+\title{Automated topic naming to support analysis of software maintenance activities}
 \maketitle
@@
 
-Topic modeling is a machine learning technique which creates multinomial distributions over words from a text corpus.
-Topic models are useful in software engineering because they can identify co-occuring words from text corpora such as source code, commit comments, email, and others.
-For example, a topic extracted from a set of commit logs might be\emph{ ``listener change remove add fire''}.
-A topic model is a summary of the key concepts in the corpus.
-Among other uses, it can quickly give developers an overview of where activity is occurring and gives managers or maintainers a sense of project history.
-A major challenge is to accurately identify the summary label for a topic (in the previous case, perhaps \emph{``event-handling''}).
+\todo[inline]{neil: possible reword: Topic modeling is a machine learning technique which creates multinomial distributions over words found in a text corpus.
+Topic models are useful in software engineering because they can identify, from text corpora such as source code, change logs, email, and others, co-occuring words.
+For example, a topic extracted from a set of changelogs might be ``listener change remove add fire''.
+A topic model gives developers an overview of where activity is occurring and gives managers or maintainers a sense of project history.
+A major challenge is to accurately identify the label for a topic (in the previous case, perhaps ``event-handling'').
 Current approaches are manual: they rely on human expertise to recognize the appropriate label.
 Furthermore, to date the results of topic modeling are project specific, and not generalizable.
-This paper presents an approach, labelled topic extraction, which addresses these two problems.
-
+}
+
+\todo[inline]{more on the case study early on as a way of showing our technique is useful for insight into dev activities}
+
+% An attempt Motivate topics
+%XXX is this the correct terminology for commits/revisions
+%The development of software does not occur in a vacuum, 
+% Revisions to
+% source code are not independent of past revisions, the specification or
+% the task at hand.
+% Often the purpose of revisions to source code relate to one or more topics, whether it be a maintenance activity, a quality issue, a requirement or feature or a response to a bug report.
+% Many of these purposes and tasks exhibit themselves as latent topics which cross-cut and compose many of development artifacts: version control revisions, bug reports, and mailing-list discussions.
+%
 Changes to software systems, revisions, do not occur without
@@
  version control revisions, bug reports, and mailing-list discussions.
+%\todo[inline]{NeiL: I still find this clumsily worded}
+
+Topics in software engineering practice refer to specific shared threads of discussions between project stakeholders. 
+These topics can include particular requirements, maintenance requests, bug fixes or other software engineering tasks.
+%NEIL: define topic specifically as akin to email 'thread'.
+It sounded too vague.
+Also changed from 'developers' to 'stakeholders' because developers sounds too coding oriented.
+Our work could apply to testers, users, etc.
+While addressing a topic of development, stakeholders often produce or modify multiple development artifacts.
+%
+These artifacts can include source code, bug reports, revisions to source code, and test cases. 
+%
+Topics are typically latent, i.e., not explicitly named, and can be in several potential states: they may not be immediately resolved, they may re-occur, yet others are only dealt with briefly  (e.g., a bug report that is closed)
+Within this paper we define \emph{topics} as latent word distributions that describe and compose the documents (version control commits) being analyzed. 
+We define \emph{labelled topics} as \emph{topics} that have been interpreted and labelled. 
+The distinction is important because word distributions can consistent of many words, thus if one has many topics, reading and interpreting these words is prohibitive.
+For example, if you read only 10 words per topic from 500 topics you have to read and interpret 5000 words, this is 5 or more pages of text!
+Thus \emph{labelled topics} save interpretation time.
+
+There are many practical applications of \emph{labelled topic
+  extraction} 
+ranging from summarization, project dashboards, and software process recovery.
+These methods are useful for providing abstractions of development activities and artifacts.
+Particularly in large projects, summarization of developer activities is essential to maintenance. 
+Rather than looking for specific concepts in a project~\cite{marcus04wcre}, we are interested in understanding a software project relative to common properties of all software.
+We can improve project dashboards, used by managers, by showing both the current topics of development as well those that relate to domain-independent taxonomies, such as NFRs.
+Some quality assurance questions can be addressed by seeking evidence within repositories that NFR-related topics are being addressed.
+Topic modelling can be used to check if development is on track and focused.
+For outside users of the software, our technique can provide insight into that company's development practices in order to answer questions such as, ``Are they focused on usability?''
+Within the organization \emph{labelled topic extraction} can act as a historical sanity check, used to avoid \emph{confirmation bias}, which is the difference between what people think happened and what actually happened.
+Some software releases, for instance, are poorly received since they are quite buggy. 
+Our technique would help explain why this was the case, since it can identify NFRs which are dominating developer attention.
+Thus developers can use these techniques to get an overview of the current and past development topics.
+For awareness, managers can use these techniques to gain an overview of topics of development.
+
+% This is a very long example it'd be nice if it was tighter
+% Perhaps development has been sidetracked by cascading issues that appeared during implementation.
+% For instance, a developer adding an optimization to a RDBMS might have to change multiple modules in order to provide that optimization code with enough information to operate. 
+% This issue might trigger a series of cascading changes that might be difficult to discover when trying to explain why the project's release date has been missed. 
+% Labelled Topic extraction could be used to detect this distinct focus of development.
+
+% semi-automatically sounds too uncertain. I think the introduction can be more muscular and bold. Don't give the reviewer any room to start doubting.
+
+% why are topics useful
+Developers often interact with abstractions of the artifacts from which topics are derived, rather than the artifacts themselves.
+Such abstractions include project dashboards~\cite{kersten2005mylar} and tags~\cite{treude2010}%, and complexity metrics \cite{mccabe1976complexity}. 
+Without the structured information which dashboards typically rely on, topics become a valuable and automatically extractable abstraction of software artifacts.
+Topics abstract software artifacts by their subject matter, rather than their concrete representation. 
+Much like aspect-oriented programming, topic labeling can identify cross-cutting concerns in a software project (as in~\cite{Baldi2008}).
+The topics of development are relevant because they identify the distinct issues facing stakeholders at a particular time.
+
+The goal of concept location and topic analysis research is to recover these topics and the related software artifacts.
+% get neil's data mining point in there
+To date, the results of topic analysis and concept location are project specific, and not generalizable.
+% get the interpretation aspect
+Much of the current research stops short of automatically interpreting the purpose of the topic, e.g.~\cite{Baldi2008,Kuhn2007,marcus04wcre}. 
+That is, extracted topics are usually manually named and labelled by those who read the topic words.
+
+% topics are numerous?
+Deriving and labelling topics is a challenging machine learning problem. 
+Topic analysis produces % raw topics that are represented as
+ unlabelled context-free word-lists or word distributions.
+Topics have to be identified by interpreting the prevalent words in a word distribution and by inspecting related documents. 
+This manual approach is impractical when handling hundreds of different topics.
+To scale to large projects (thousands of lines of code and multiple releases), we need automatic assistance to determine the topic label.
+We propose a method of interpreting and labelling topics, and their related artifacts, that is generalizable and relevant to multiple projects.
 
@@
 The area under the ROC of a classifier is equivalent to the probability that the classifier will rank a randomly chosen positive instance higher than a randomly chosen negative instance.
-F-measure is the harmonic mean of precision and recall.
+F-measure is harmonic mean of precision and recall.
 
@@
 
+\subsection{Annotation observations}
+We found many topics that were not non-functional requirements (NFRs) but were often related to them. 
+For instance, concurrency was mentioned often in the commit logs and was related to \emph{correctness} and \emph{reliability}, likely because concurrent code is prone to bugs such as race conditions.
+%are
+% often  was troublesome. 
+\emph{Configuration management} and \emph{source control} related changes appeared often; % and sometimes there were topics dedicated to configuration management. 
+these kinds of changes are slightly related to \emph{maintainability}. 
+A non-functional change that was not quality-related was licensing and copyright; many changes were simply to do with updating copyrights or ensuring copyright or license headers were applied to files.
+In these cases we assigned the \emph{None} class.
+
+We noticed that occasionally the names of modules would conflict with words related to other non-functional requirements. 
+For instance, optimizers are very common modules in database systems: both MySQL and MaxDB have optimizer modules. 
+In MySQL the optimizer is mentioned but often the change addresses  correctness or another quality. 
+Despite this difference, the name of the module could fool our learners into believing the change was always about \emph{efficiency}. 
+In these cases the advantages of tailoring topic names to specific project terminologies are more clear. 
+Project specific word-lists would avoid automated mistakes due to the names of entities and modules of a software project.
+
+\subsection{Summary of techniques}
+While an unsupervised technique such as LDA is appealing in its lack of human intervention, and thus lower effort, 
+supervised learners have the advantage of domain knowledge, which typically means improved results. 
+Our manual annotations were fairly quick to do, taking only a few minutes per 30 day period.
+
+Very rarely did \textsf{exp2} and \textsf{exp3} (naive word matching) ever perform as well as the supervised machine learners. 
+For MaxDB, \textit{reliability} was slightly better detected using the static word list of \textsf{exp2}. 
+In general, the machine learners and \textsf{exp3} did better than \textsf{exp2} for both MaxDB and MySQL. 
+For both MySQL and MaxDB \textit{usability} was better served by \textsf{exp2}. 
+\textit{Usability} was a very infrequent label, however, which made it difficult to detect in any case.
+
+We found that the multi-label learners of BR, CLR and HOMER did not do as well for Macro-ROC as NaiveBayes and other naive Bayes derived learners. 
+This suggests that by combining together multiple NaiveBayes learners we could probably label sets of topics effectively, but it would require a separate NaiveBayes learner per label.
+
 \section{Understanding software maintenance activities} 
@@
 %\subsection{Effectiveness}
-\subsection{Annotation observations}
-We found many topics that were not non-functional requirements (NFRs) but were often related to them. 
-For instance, concurrency was mentioned often in the commit logs and was related to \emph{correctness} and \emph{reliability}, likely because concurrent code is prone to bugs such as race conditions.
-%are
-% often  was troublesome. 
-\emph{Configuration management} and \emph{source control} related changes appeared often; % and sometimes there were topics dedicated to configuration management. 
-these kinds of changes are slightly related to \emph{maintainability}. 
-A non-functional change that was not quality-related was licensing and copyright; many changes were simply to do with updating copyrights or ensuring copyright or license headers were applied to files.
-In these cases we assigned the \emph{None} class.
-
-We noticed that occasionally the names of modules would conflict with words related to other non-functional requirements. 
-For instance, optimizers are very common modules in database systems: both MySQL and MaxDB have optimizer modules. 
-In MySQL the optimizer is mentioned but often the change addresses  correctness or another quality. 
-Despite this difference, the name of the module could fool our learners into believing the change was always about \emph{efficiency}. 
-In these cases the advantages of tailoring topic names to specific project terminologies are more clear. 
-Project specific word-lists would avoid automated mistakes due to the names of entities and modules of a software project.
-
-\subsection{Summary of techniques}
-While an unsupervised technique such as LDA is appealing in its lack of human intervention, and thus lower effort, 
-supervised learners have the advantage of domain knowledge, which typically means improved results. 
-Our manual annotations were fairly quick to do, taking only a few minutes per 30 day period.
-
-Very rarely did \textsf{exp2} and \textsf{exp3} (naive word matching) ever perform as well as the supervised machine learners. 
-For MaxDB, \textit{reliability} was slightly better detected using the static word list of \textsf{exp2}. 
-In general, the machine learners and \textsf{exp3} did better than \textsf{exp2} for both MaxDB and MySQL. 
-For both MySQL and MaxDB \textit{usability} was better served by \textsf{exp2}. 
-\textit{Usability} was a very infrequent label, however, which made it difficult to detect in any case.
-
-We found that the multi-label learners of BR, CLR and HOMER did not do as well for Macro-ROC as NaiveBayes and other naive Bayes derived learners. 
-This suggests that by combining together multiple NaiveBayes learners we could probably label sets of topics effectively, but it would require a separate NaiveBayes learner per label.
 
@@
 % can always be cut down
-%Our work faced multiple threats to validity and we attempted to address many of them.
+Our work faced multiple threats to validity and we attempted to address many of them.
 
 \emph{Construct validity} -- we used only commit messages rather than mail or bug tracker messages. 
-Since we required a seamless record, it was not possible to obtain other data sources.
-Possibly they would have influenced our results, but there would be a degree of correlation between the corpora.
-%Conversely, construct validity was addressed in our validation as we used the files, topics and revisions to annotate the commit messages. 
-%We used mailing-lists to verify the purpose behind some events.
-Our taxonomy for software NFRs is subject to dispute, but seems generally accepted.
-Finally, there are exogenous sources, such as in-person discussions, which we did not access, an omission as noted in \cite{aranda09icse}.
-%We did not show these results to the stakeholders.
-
-\emph{Internal validity} -- %includes inter-rater reliability and our lack of rival explanation building. 
+Conversely, construct validity was addressed in our validation as we used the files, topics and revisions to annotate the commit messages. 
+We used mailing-lists to verify the purpose behind some events.
+We also chose our taxonomy and the data to study. We did not show these results to the stakeholders.
+We only studied database systems.
+
+\emph{Internal validity} -- includes inter-rater reliability and our lack of rival explanation building. 
 We improved internal validity by trying to correlate and explain the behaviours observed in the analysis with the historical records of the projects.
-We did not attempt to match our results to any particular model or theory, as such theories are sparse in
-
-\emph{External validity} -- %was addressed by our use of two case studies, but there are still issues. 
+We did not attempt to match our results to any particular model or theory.
+
+\emph{External validity} was addressed by our use of multiple case studies, but there are still issues. 
 Our data originated from OSS database projects and thus might not be applicable to commercially developed software. 
 Furthermore, our analysis techniques rely on a project's use of meaningful commit messages. The OSS projects investigated were database systems, thus 
-our results might not generalize to other domains. %We only studied database systems.
+our results might not generalize to other domains.
 % \todo[inline]{If we compare with the same domain we don't get to say it generalizes a lot?}
@@
 % in Case Study Research: Design Case Studies
-\emph{Reliability} -- each annotator followed the same protocol and used the same annotations. 
-However, only two annotators were used; their annotations could be biased as we did not analyze for inter-rater reliability.
+\emph{Reliability} was addressed by each annotator following the same protocol and using the same annotations. 
+Unfortunately only two annotators were used; their annotations could be biased as we did not analyze for inter-rater reliability.
 

### 2011-01-22 msr/msr2011.tex
um, yeah. i don't understand rebase i guess.
removed Mens, added Ernst ref
removed F-measure figure, updated supervised figure with random bar.

--- 2011-01-21 msr/msr2011.tex
@@
 % Not sure we need this
-Mens et al.~\cite{mens08icsm} conducted an empirical study of Eclipse, a popular Java IDE, to verify the claims of Lehman~\cite{lehman97sms}.
+%Mens et al.~\cite{mens08icsm} conducted an empirical study of Eclipse, a popular Java IDE, to verify the claims of Lehman~\cite{lehman97sms}.
 % They concerned themselves with source code only, and found Law Seven, ``Declining Quality'', to be too difficult to assess: ``[we lacked an] appropriate measurement of the evolving quality of the system as perceived by the users \cite[p.
 388]{mens08icsm}''.
-This paper examines the notions of quality in terms of a consistent ontology, as Mens
-et al. call for in their conclusions.
+%This paper examines the notions of quality in terms of a consistent ontology, as Menset al. call for in their conclusions.
 
@@
 
-\begin{figure}
-  \centering
- \includegraphics[width=0.45\textwidth]{figures/unsupervised-bar-f1}
- \caption{Performance, F1 values (range: 0-1), of unsupervised topic labelling for
-   each NFR and per word-list.}
-  \label{fig:maxdb-unsup-results-f1}
-\end{figure}
+% \begin{figure}
+%   \centering
+%  \includegraphics[width=0.45\textwidth]{figures/unsupervised-bar-f1}
+%  \caption{Performance, F1 values (range: 0-1), of unsupervised topic labelling for
+%    each NFR and per word-list.}
+%   \label{fig:maxdb-unsup-results-f1}
+% \end{figure}
 
@@
 % \textsf{Nomen} would help explain why this was the case, since it can identify NFRs which are dominating developer attention.
+\todo[inline]{does this section use supervised learning, and if so what ones?}
 

### 2011-01-22 msr/msr2011.tex
commented out todos in intro

@@
 
-\todo[inline]{Strength story: Neil says
-  And ultimately, I think the story needs more strengthening. Maybe we
-  need a scenario: you work at Bank X. You are curious about how much
-  recent activity has been done on software maintainability
-  vs. security. Our technique will allow you to compare different IT
-  projects using the same keywords, and ultimately, to generate lists
-  showing relative frequencies.}
+%\todo[inline]{Strength story: Neil says
+  % And ultimately, I think the story needs more strengthening. Maybe we
+  % need a scenario: you work at Bank X. You are curious about how much
+  % recent activity has been done on software maintainability
+  % vs. security. Our technique will allow you to compare different IT
+  % projects using the same keywords, and ultimately, to generate lists
+  % showing relative frequencies.}
 
@@
 
-\todo[inline]{Usefulness needs to be immediate and apparent}
-
-\todo[inline]{ It needs to be clear there is value in labelling.
-  We need to make it clear that other researchers weren't labelling them
- automatically and presented them as labelled and thus it is obvious from
-  their presentation that labels matters + need a reading argument. The
- 2nd reviewer picked on this a bit but still didn't get it. I thought
- some of this had been clear.}
+% % \todo[inline]{Usefulness needs to be immediate and apparent}
+
+% \todo[inline]{ It needs to be clear there is value in labelling.
+ %  We need to make it clear that other researchers weren't labelling them
+ % automatically and presented them as labelled and thus it is obvious from
+ %  their presentation that labels matters + need a reading argument. The
+ % 2nd reviewer picked on this a bit but still didn't get it. I thought
+ % some of this had been clear.}
 
@@
 
+One challenge for software maintenance is to identify the reasons behind software development activities.
+Identifying these reasons provides insight into \emph{why} certain changes were made.
 Topic modeling is a machine learning technique which creates multinomial distributions over words from a text corpus.
-Topic models are useful in software engineering because they can identify co-occuring words from text corpora such as source code, commit comments, email, and others.
-For example, a topic extracted from a set of commit logs might be\emph{ ``listener change remove add fire''}.
-A topic model is a summary of the key concepts in the corpus.
-Among other uses, it can quickly give developers an overview of where activity is occurring and gives managers or maintainers a sense of project history.
-A major challenge is to accurately identify the summary label for a topic (in the previous case, perhaps \emph{``event-handling''}).
+Topic models are useful in software maintenance because they summarize the key concepts in a corpus, such as source code, commit comments, or email, by identifying statistically co-occuring words.
+Among other uses, it can quickly give developers an overview of where activity occurred, and gives managers or maintainers a sense of project history.
+For example, a topic extracted from a set of commit logs might be \emph{ ``listener change remove add fire''}.
+A major challenge for topic modeling is to accurately identify the summary label for a topic (in the previous case, perhaps \emph{``event-handling''}).
 Current approaches are manual: they rely on human expertise to recognize the appropriate label.
 Furthermore, to date the results of topic modeling are project specific, and not generalizable.
-This paper presents an approach, labelled topic extraction, which addresses these two problems.
-
-Changes to software systems, revisions, do not occur without
-reason. 
-The reason or purpose is often related to requirements 
-Often one revision might have multiple purposes ranging from
-maintenance issues, quality issues, requirements, features or a
-response to a bug report.
-These purposes often are exhibited as latent topics that relate to
-many development artifacts such as:
- version control revisions, bug reports, and mailing-list discussions.
+This paper presents an approach, \emph{labelled topic extraction}, which addresses these two problems.
+
+% Changes to software systems, revisions, do not occur without
+% reason. 
+% The reason or purpose is often related to requirements 
+% Often one revision might have multiple purposes ranging from
+% maintenance issues, quality issues, requirements, features or a
+% response to a bug report.
+% These purposes often are exhibited as latent topics that relate to
+% many development artifacts such as:
+%  version control revisions, bug reports, and mailing-list discussions.
 

### 2011-01-22 msr/msr2011.tex
added contribs to intro.

@@
 
-This paper first introduces labelled topic extraction, and then uses that technique to analyze development activity in two large-scale examples.
-Next we cover much of the previous work that we leverage (Section \ref{sec:related}).
-Then we follow by describing how we generated our data (Sections  \ref{sec:wordlist} and \ref{sec:genwordlist}). 
-We show that labels with their topics can be learned and used to classify other data-sets, either without training (Section \ref{sec:unsuplabelling}), or with training (Section \ref{sec:suplabelling}).
-We then present visualizations of named topics and their trends over time to aid communication and analysis. 
+This paper makes the following contributions: 
+\begin{itemize}
+\item We introduce the concept of labelled topic extraction, using a non-functional requirements (NFR) taxonomy for our labels. 
+\item We evaluate an unsupervised method of automated topic labelling (word-lists);
+%\item Our unsupervised method can be applied to projects without training;
+\item We evaluate a supervised method of automatically labelling topics with a single NFR (machine learning);
+\item We evaluate a supervised method of labelling topics with multiple NFRs (multi-label machine learning);
+\item We provide a method of cross-project analysis via topic labelling;
+\item We apply these techniques to visualizing NFRs over time, to analyze maintenance activities.
+\end{itemize}
+
+%We first introduces labelled topic extraction, and then uses that technique to analyze development activity in two large-scale examples.
+We begin with related work in Section \ref{sec:related}.
+Next, we describe how we generated our data (Section \ref{sec:wordlist}) and how we created the word-lists used in unsupervised classification (Section \ref{sec:genwordlist}). 
+%We show that labels with their topics can be learned and used to classify other data-sets, either without training 
+%Our goal is to label a given topic with either an NFR from our taxonomy,
+ Section \ref{sec:unsuplabelling} presents our technique for using word-lists to do unsupervised classification, and describes the results.
+In Section \ref{sec:suplabelling}, we manually annotate the topics, and use those annotations as training data for supervised classification.
+%We then present visualizations of named topics and their trends over time to aid communication and analysis. 
 We use an exploratory case study of two open source database systems to show how named topics can be compared between projects  (Section \ref{sec:analysis}). 
-The paper concludes with a discussion of limitations (Section \ref{sec:limit})
+The paper concludes with a discussion of limitations (Section \ref{sec:limit}), and some conclusions.
 

### 2011-01-22 msr/msr2011.tex
abstract: make the labels from NFRs point clearer.
intro changes/clarifications on claims

@@
 from source control systems such as CVS and BitKeeper.  
-These topics are given labels relating to a generalizable
+These topics are given labels from to a generalizable
 cross-project taxonomy, consisting of non-functional
@@
 that it is difficult to discuss results independent of the project context. 
-Our technique automatically assigns labels across projects.
+Our technique automatically, or with some initial training, assigns labels across projects.
 \end{enumerate}
@@
 %We then present visualizations of named topics and their trends over time to aid communication and analysis. 
-We use an exploratory case study of two open source database systems to show how named topics can be compared between projects  (Section \ref{sec:analysis}). 
+To demonstrate an application of labelled topic extraction, we use an exploratory case study of two open source database systems to show how named topics can be compared between projects  (Section \ref{sec:analysis}).
 The paper concludes with a discussion of limitations (Section \ref{sec:limit}), and some conclusions.

### 2011-01-22 msr/msr2011.tex
made subfig for unsupervised.

@@
 % Rebuttal: ROC is not new and it is what the machine learning community, search based software engineering and now the MSR community are using +  it has natural meaningful quality mappings.
-\begin{figure}
+\begin{figure*}[]
 \centering
+\subfloat[ROC]{
  \includegraphics[width=0.45\textwidth]{figures/unsupervised-bar}
- \caption{Performance, ROC values (range: 0-1), of unsupervised topic labelling for
-   each NFR and per word-list. Dashed line indicates the performance of a random classifier. This graph shows how well the
-   unsupervised topic labelling matched our manual annotations.}
-
-%\todo[inline]{this is too tiny in print}
+\label{fig:maxdb-unsup-results-roc}
+}
+\subfloat[F1]{
+\includegraphics[width=0.45\textwidth]{figures/unsupervised-bar-f1}
+\label{fig:maxdb-unsup-results-f1}
+}
+\caption[]{Performance of unsupervised topic labelling for each NFR per word-list.
+A) ROC values -- dashed line indicates the performance of a random classifier.
+B) F1 measure.
+Possible values range from 0--1.
+%\todo[inline]{better discussion}% and graphical legend would be nice}
+}
 \label{fig:maxdb-unsup-results}
-\end{figure}
+\end{figure*}
+
+% \begin{figure}
+%   \centering
+%  \includegraphics[width=0.45\textwidth]{figures/unsupervised-bar}
+%  \caption{Performance, ROC values (range: 0-1), of unsupervised topic labelling for
+%    each NFR and per word-list. Dashed line indicates the performance of a random classifier. This graph shows how well the
+%    unsupervised topic labelling matched our manual annotations.}
+
+% %\todo[inline]{this is too tiny in print}
+%   \label{fig:maxdb-unsup-results}
+% \end{figure}
 

### 2011-01-22 msr/msr2011.tex
moved word-lists into "unsupervised class" section as this is more related.

@@
 We begin with related work in Section \ref{sec:related}.
-Next, we describe how we generated our data (Section \ref{sec:wordlist}) and how we created the word-lists used in unsupervised classification (Section \ref{sec:genwordlist}). 
+Next, we describe how we generated our data (Section \ref{sec:wordlist}).
+For unsupervised classification (Section \ref{sec:unsuplabelling}), we begin by creating word-lists to signifiy when a topic matches an NFR label.
+We then apply our classifier and analyze the results.
+%and how we created the word-lists used in unsupervised classification.
 %We show that labels with their topics can be learned and used to classify other data-sets, either without training 
 %Our goal is to label a given topic with either an NFR from our taxonomy,
- Section \ref{sec:unsuplabelling} presents our technique for using word-lists to do unsupervised classification, and describes the results.
+ %Section \ref{sec:unsuplabelling} presents our technique for using word-lists to do unsupervised classification, and describes the results.
 In Section \ref{sec:suplabelling}, we manually annotate the topics, and use those annotations as training data for supervised classification.  
@@
 To demonstrate an application of labelled topic extraction, we use an exploratory case study of two open source database systems to show how named topics can be compared between projects  (Section \ref{sec:analysis}).
-The paper concludes with a discussion of limitations (Section \ref{sec:limit}), and some conclusions.
+The paper concludes with a discussion of limitations (Section \ref{sec:limit}), and future work.
 
@@
 Figure \ref{fig:process} gives an outline of the methodology followed. 
-We began by gathering source data and creating topic models. We created labels using a well-known NFR taxonomy. 
-We then applied unsupervised and supervised learning techniques to try to match our labels to the topics. 
-These topics were used to analyze the role of non-functional requirements in software maintenance.
+We began by gathering source data and creating topic models. For unsupervised labelling, we generate three sets of wordlists as signifiers for NFRs.
+%e created labels using a well-known NFR taxonomy. 
+In supervised learning, we trained our data with manual annotations in order to match topics with NFRs.
+Finally, these topics were used to analyze the role of non-functional requirements in software maintenance.
 
@@
 
-\subsection{Generating word lists}
-\label{sec:genwordlist}
-
+\subsection{Unsupervised labelling}
+\label{sec:unsuplabelling}
+
+%\subsection{Generating word lists}
+%\label{sec:genwordlist}
+\textbf{Generating word lists}
 % %\todo[inline]{Make it clear where the data comes from! Neil says: Also, I
@@
 
-This paper does this differently. First of all, we automated the process of naming the topics. Secondly, we labelled topics with project-independent terms, in order to permit cross-project comparison.
+%% NEIL: this should be in the intro or conclusion, not this section.
+%This paper does this differently.
+First of all, we automated the process of naming the topics.
+Secondly, we labelled topics with project-independent terms, in order to permit cross-project comparison.
 
@@
 
-\subsection{Unsupervised labelling}
-\label{sec:unsuplabelling}
-
-\todo[inline]{* the hierarchy of tags was not clear}
+\textbf{Generating labels automatically}
+\todo[inline]{NEIL: this could be a subsubsection}
+%\todo[inline]{* the hierarchy of tags was not clear}
 

### 2011-01-22 msr/msr2011.tex
minor fixes to wording, some clarification for reviews (why we didn't use email)

@@
 For each project, we used source control commit comments, the messages that programmers write when they commit revisions to a source control repository. 
-This is the most readily accessible source of project interactions for outside researchers; bug trackers and email-list archives were not available for both projects. 
+This is the most readily accessible source of project interactions for outside researchers, since developers often or always (e.g., Git) write such message.
+Our approach is more generalizable if it does not assume the presence of other corpora.
+%bug trackers and email-list archives were not available for both projects.
 An example of a typical commit message is: \textit{``history annotate diffs bug fixed (if mysql\-\_real\-\_connect() failed there were two pointers to malloc'ed strings, with memory corruption on free(), of course)''}.
@@
 
-A topic analysis tool like LDA will try to find $N$ independent word distributions within the word distributions of all the messages. 
+A topic analysis tool such as LDA will try to find $N$ independent word distributions within the word distributions of all the messages. 
 Linear combinations of these $N$ word distributions are meant to represent and recreate the word distributions of any of the original messages. 
@@
 Representing documents as a mixture of topics maps well to source code repository commits, which often have more than one purpose~\cite{Hindle09ICSM}.  
-A topic, in this paper, represents both a word distribution and a group of commit log comments that are related to each other by their content.  
-In this paper a topic is a set of tokens extracted from commit messages found within a project's source control system (SCS).
+For this paper, a topic represents a word distribution generated from a group of commit log comments which are related by their content.  
+%In this paper a topic is a set of tokens extracted from commit messages found within a project's source control system (SCS).
 

### 2011-01-22 msr/msr2011.tex
moved the ISO discussion into the data generation section. Having it first makes it very clear that our compariosn are based on this.

@@
 
+\noindent \textbf{The high-level labels} --- To facilitate cross-project comparison, we used a taxonomy of NFRs. This taxonomy is based on the ISO quality model, ISO9126~\cite{iso9126}. 
+ISO9126 describes six high-level non-functional requirements (NFRs), maintainability, functionality, portability, efficiency, usability, and reliability. 
+There is some debate about the terms in this model. 
+However, ISO9126 is ``an international standard and thus provides an internationally accepted terminology for software quality \cite[p.
+58]{Boegh2008},'' that is sufficient for the purposes of this research.
+We claim that these NFRs are maintenance concerns (to varying degrees) in all software projects, and are therefore well-suited to comparison between projects.
+
+\noindent \textbf{Creating a validation corpus} --- To evaluate both unsupervised and supervised classification, we created a validation set of manually labelled topics.
+For MySQL 3.23 and MaxDB 7.500, we annotated each extracted topic in each period with the same NFR labels as described in the preceding paragraph.
+We looked at each period's topics, and assessed what the data --- consisting of the frequency-weighted word lists and messages --- suggested was the topic for that period. 
+We were able to pinpoint the appropriate label using auxiliary information as well, such as the actual revisions and files that were related to the topic being annotated.
+For example, for the MaxDB topic consisting of a message ``exit() only used in non NPTL LINUX Versions'', we tagged that topic \emph{portability}. 
+%\todo[inline]{this confused people. Is it truly unsupervised? Yes, but it needs to be clearer.}
+%We compared against this data-set, but we also used the data-set for our supervised machine learning based topic classification. 
+% A topic may have more than one matching keyword. % ok what?
+%\todo[inline]{inter-rater reliability }
+% \todo[inline]{There should be a section here which is about manual annotation. It is really important and it is required by both analyses. Plus we can tell them how.}
+
 \subsection{Unsupervised labelling}
@@
 
-Our second word-list, \textsf{exp2}, is derived from the ISO quality model, ISO9126~\cite{iso9126}. 
-ISO9126 describes six high-level non-functional requirements (NFRs), maintainability, functionality, portability, efficiency, usability, and reliability. 
-There is some debate about the terms in this model. 
-However, ISO9126 is ``an international standard and thus provides an internationally accepted terminology for software quality \cite[p.
-58]{Boegh2008},'' that is sufficient for the purposes of this research.
+Our second word-list, \textsf{exp2}, uses the ISO9126 taxonomy described above (Section \ref{sec:wordlist}) to seed the wordlists.
 %We use these qualities later on as classes in supervised labelling. NE confusing the issue here, I think.
@@
 \end{table*}
-
-\noindent \textbf{Creating a validation corpus} --- For MySQL 3.23 and MaxDB 7.500, we manually annotated each extracted topic in each period with the same NFR labels as \textsf{exp2} (software qualities).
-We looked at each period's topics, and assessed what the data --- consisting of the frequency-weighted word lists and messages --- suggested was the topic for that period. 
-We were able to pinpoint the appropriate label using auxiliary information as well, such as the actual revisions and files that were related to the topic being annotated.
-For example, for the MaxDB topic consisting of a message ``exit() only used in non NPTL LINUX Versions'', we tagged that topic \emph{portability}. 
-%\todo[inline]{this confused people. Is it truly unsupervised? Yes, but it needs to be clearer.}
-We compared against this data-set, but we also used the data-set for our supervised machine learning based topic classification. 
-% A topic may have more than one matching keyword. % ok what?
-%\todo[inline]{inter-rater reliability }
-% \todo[inline]{There should be a section here which is about manual annotation. It is really important and it is required by both analyses. Plus we can tell them how.}
 

### 2011-01-22 msr/msr2011.tex
updated related work by moving it from unsupervised section.

@@
 %Some results were interesting, but their precision was quite low. 
-ConcernLines~\cite{treude09cl} shows tag occurrence using colour intensity. 
+With ConcernLines, \cite{treude09cl} shows tag occurrence using colour intensity. 
 They mined 
@@
 %This paper examines the notions of quality in terms of a consistent ontology, as Menset al. call for in their conclusions.
+
+% in our previous work~\cite{Hindle09ICSM}, as well as -- CITED a little too often (vanity cites!)
+In Baldi et al.~\cite{Baldi2008}, topics are named manually: human experts read the highest-frequency members of a topic and assign a label accordingly. 
+As discussed earlier, given the topic \emph{``listener change remove add fire''}, Baldi et al. would assign the label \emph{event-handling}. 
+The labels are reasonable enough, but still require an expert in the field to determine them. 
+Furthermore, these labels are project-specific, because they are generated from the data of that project. E.g., we might have a label called `Oracle' in the MySQL case, since Oracle owns MySQL.
+
+This paper does this differently. First of all, we automated the process of naming the topics. Secondly, we labelled topics with project-independent terms, in order to permit cross-project comparison.
 

### 2011-01-22 msr/msr2011.tex
use word-list rather than wordlist as spelling clear up what a label is

@@
 Figure \ref{fig:process} gives an outline of the methodology followed. 
-We began by gathering source data and creating topic models. For unsupervised labelling, we generate three sets of wordlists as signifiers for NFRs.
+We began by gathering source data and creating topic models. For unsupervised labelling, we generate three sets of word-lists as signifiers for NFRs.
 %e created labels using a well-known NFR taxonomy. 
@@
 %\label{sec:genwordlist}
-\textbf{Generating word lists}
 % %\todo[inline]{Make it clear where the data comes from! Neil says: Also, I
@@
 
-Topics are word distributions, lists of words ranked by frequency, that are burdensome to interpret and hard to distinguish and understand. 
-While the topics themselves are generated automatically, how to interpret these topics is less clear.
-For example,
-% in our previous work~\cite{Hindle09ICSM}, as well as -- CITED a little too often (vanity cites!)
-in Baldi et al.~\cite{Baldi2008}, topics are named manually: human experts read the highest-frequency members of a topic and assign a label accordingly. 
-Given the topic \emph{``listener change remove add fire''}, Baldi et al. would assign the label \emph{event-handling}. 
-The labels are reasonable enough, but still require an expert in the field to determine them. 
-Furthermore, these labels are project-specific, because they are generated from the data of that project. E.g., we might have a label called `Oracle' in the MySQL case, since Oracle owns MySQL.
-
-%% NEIL: this should be in the intro or conclusion, not this section.
-%This paper does this differently.
-First of all, we automated the process of naming the topics.
-Secondly, we labelled topics with project-independent terms, in order to permit cross-project comparison.
-
-Accordingly, we associated each topic with a label from a list
-of keywords (`wordlists', in our parlance) and related terms. These wordlists were determined a priori and were not extracted from the projects themselves.
+%Topics are word distributions, lists of words ranked by frequency, that are burdensome to interpret and hard to distinguish and understand. 
+%While the topics themselves are generated automatically, how to interpret these topics is less clear.
+%For example,
+
+\noindent\textbf{Generating word lists} --- In order to automatically label each topic with an NFR from ISO9126, we associate each NFR with a list
+of keywords (`word-lists', in our parlance). These word-lists were determined a priori and were not extracted from the projects themselves.
 We intersected the words of the topics and the words of our word-lists.
 We `labeled' a topic if any of its words matched any of the word-list's words.
-A topic could match more than one keyword.
+A topic could match more than one NFR.
 We used several different sets of word-lists for comparison, which we refer to as \textsf{exp1, exp2, exp3} in the text which follows.
@@
 
-Our second word-list, \textsf{exp2}, uses the ISO9126 taxonomy described above (Section \ref{sec:wordlist}) to seed the wordlists.
+Our second word-list, \textsf{exp2}, uses the ISO9126 taxonomy described above (Section \ref{sec:wordlist}) to seed the word-lists.
 %We use these qualities later on as classes in supervised labelling. NE confusing the issue here, I think.
@@
 
-\textbf{Generating labels automatically}
 \todo[inline]{NEIL: this could be a subsubsection}
 %\todo[inline]{* the hierarchy of tags was not clear}
-
-% word list similarity
-Using our three word lists (\textsf{exp1}, \textsf{exp2}, \textsf{exp3}), we labeled our topics where there was a match between a word in the list and the same word in the topic.
-A \emph{named topic} is a topic with a matching label. 
+\noindent \textbf{Generating labels automatically} -- Using our three word-lists (\textsf{exp1}, \textsf{exp2}, \textsf{exp3}), we labeled our topics with an NFR where there was a match between a word in the list and the same word somewhere in the distribution of words that constitute the topic.
+A \emph{named topic} is a topic with a matching NFR label.
+Unnamed topics occur where this no such match, which may indicate either a lack of precision, or simply that this topic is not associated with non-functional requirements.
 All experiments were run on MaxDB 7.500 and MySQL 3.23 data. LDA
@@
 domain knowledge of software to label relevant topics.
-\todo[inline]{Neil does this explain the unsupervised aspect well enough?}
+%\todo[inline]{Neil does this explain the unsupervised aspect well enough?}
 

### 2011-01-22 msr/msr2011.tex
moved ROC discussion into section 3.1, clearer there I think. Also changed ROC unsupervised figure.

@@
 
+%\todo[inline]{NEIL: this section should be updated to explain F-measure}
+We validate classification performance using the Receiver Operating
+Characteristic area-under-curve value~\cite{Fawcett2006861},
+abbreviated \emph{ROC}, and the F-measure.  F-measure is the harmonic mean of precision and recall.
+
+\todo[inline]{NEIL: abram, i got rid of the letter-grade thing because I'm worried people latch on to the idea we are C- students. I feel anything better than random is our criteria here}
+ROC values provide a score %, similar to school letter-grades (A is 0.9, C is 0.6),
+ reflecting how well a particular learner performed for the given data. 
+ROC maps to the more familiar concepts of precision/sensitivity and recall/specificity: it plots the true positive rate (sensitivity) versus the false positive rate (1 - specificity). 
+A perfect learner has a ROC value of 1.0, reflecting perfect recall and precision. 
+A ROC result of 0.5 would be equivalent to a random learner (that is, issuing as many false positives as true positives). 
+The %area under the -- NE confusing to have area again ... and not importantana
+ROC of a classifier is equivalent to the probability that the classifier will rank a randomly chosen positive instance higher than a randomly chosen negative instance.
+
+We consider our labelling classifiers acceptable if they outperform a random classifier (0.5). 
 \subsection{Unsupervised labelling}
@@
 %\todo[inline]{* the hierarchy of tags was not clear}
-\noindent \textbf{Generating labels automatically} -- Using our three word-lists (\textsf{exp1}, \textsf{exp2}, \textsf{exp3}), we labeled our topics with an NFR where there was a match between a word in the list and the same word somewhere in the distribution of words that constitute the topic.
+\noindent \textbf{Generating labels automatically} --- Using our three word-lists (\textsf{exp1}, \textsf{exp2}, \textsf{exp3}), we labeled our topics with an NFR where there was a match between a word in the list and the same word somewhere in the distribution of words that constitute the topic.
 A \emph{named topic} is a topic with a matching NFR label.
@@
 This labelling is unsupervised because the corpus is not derived from 
-the project being labelled and we did not label the project's topics
+the project being labelled, and we did not label the project's topics
 ourselves for a training set. The theory behind this technique is that
@@
 
-\noindent \textbf{Analysis of the unsupervised labelling} -- %Based on the labels, and our manual topic labelling, 
+\noindent \textbf{Analysis of the unsupervised labelling} --- %Based on the labels, and our manual topic labelling, 
 %we compared the results of the unsupervised word matching approach. 
-For each quality we assessed whether unsupervised labels matched the manual annotations. To be clear, the manual annotations were not used to train the labeling process.
-Figure \ref{fig:maxdb-unsup-results} shows our results for MaxDB and MySQL. 
-\todo[inline]{NEIL: this section should be updated to explain F-measure}
-Our performance is measured using the Receiver Operating
-Characteristic area-under-curve value~\cite{Fawcett2006861},
-abbreviated \emph{ROC}, and the F-measure. 
-ROC values provide a score, similar to school letter-grades (A is 0.9, C is 0.6), reflecting how well a particular learner performed for the given data. 
-ROC maps to the more familiar concepts of precision/sensitivity and recall/specificity: it plots the true positive rate (sensitivity) versus the false positive rate (1 - specificity). 
-A perfect learner has a ROC value of 1.0, reflecting perfect recall and precision. 
-A ROC result of 0.5 would be equivalent to a random learner (that is, issuing as many false positives as true positives).
-
-The %area under the -- NE confusing to have area again ... and not importantana
-ROC of a classifier is equivalent to the probability that the classifier will rank a randomly chosen positive instance higher than a randomly chosen negative instance.
-F-measure is the harmonic mean of precision and recall.
+For each quality we assessed whether unsupervised labels matched the manual annotations. %To be clear, the manual annotations were not used to train the labeling process.
+As described in \ref{sec:wordlist} we used both ROC and F-1 measures to evaluate the performance of the classification.
+Figure \ref{fig:maxdb-unsup-results} shows our ROC results for MaxDB and MySQL. We describe F-1 results in the text below.
 
@@
 % Rebuttal: ROC is not new and it is what the machine learning community, search based software engineering and now the MSR community are using +  it has natural meaningful quality mappings.
-\begin{figure*}[]
-\centering
-\subfloat[ROC]{
- \includegraphics[width=0.45\textwidth]{figures/unsupervised-bar}
-\label{fig:maxdb-unsup-results-roc}
-}
-\subfloat[F1]{
-\includegraphics[width=0.45\textwidth]{figures/unsupervised-bar-f1}
-\label{fig:maxdb-unsup-results-f1}
-}
-\caption[]{Performance of unsupervised topic labelling for each NFR per word-list.
+% \begin{figure}[t]
+% \centering
+% \subfloat[ROC]{
+%  \includegraphics[width=0.45\textwidth]{figures/unsupervised-bar}
+% \label{fig:maxdb-unsup-results-roc}
+% }
+% %\subfloat[F1]{
+% %\includegraphics[width=0.45\textwidth]{figures/unsupervised-bar-f1}
+% %\label{fig:maxdb-unsup-results-f1}
+% %}
+% \caption[]{Performance of unsupervised topic labelling for each NFR per word-list.
 A) ROC values -- dashed line indicates the performance of a random classifier.
@@
 Possible values range from 0--1.
-%\todo[inline]{better discussion}% and graphical legend would be nice}
-}
+% %\todo[inline]{better discussion}% and graphical legend would be nice}
+% }
+% \label{fig:maxdb-unsup-results}
+% \end{figure}
+
+\begin{figure*}
+  \centering
+ \includegraphics[width=\textwidth]{figures/unsupervised-bar}
+ \caption{Performance, ROC values (range: 0-1), of unsupervised topic labelling for
+   each NFR and per word-list. Dashed line indicates the performance of a random classifier. This graph shows how well the
+   unsupervised topic labelling matched our manual annotations.}
+
+%\todo[inline]{this is too tiny in print}
   \label{fig:maxdb-unsup-results}
 \end{figure*}
-
-% \begin{figure}
-%   \centering
-%  \includegraphics[width=0.45\textwidth]{figures/unsupervised-bar}
-%  \caption{Performance, ROC values (range: 0-1), of unsupervised topic labelling for
-%    each NFR and per word-list. Dashed line indicates the performance of a random classifier. This graph shows how well the
-%    unsupervised topic labelling matched our manual annotations.}
-
-% %\todo[inline]{this is too tiny in print}
-%   \label{fig:maxdb-unsup-results}
-% \end{figure}
 

### 2011-01-23 msr/msr2011.tex
made more subsubsections

@@
 
+% AH: Topics take a long to read
+% AH: Project Dashboards, summarization, software process recovery
+
+Concrete applications of topics and \emph{labelled topic extraction}
+range from project dashboards to annotating software artifacts with
+NFR-related tags such as revisions and bug reports. 
+Project dashboard are typically employed by managers and are used to
+provide quick summaries of a the efforts put into a software project,
+labelled topic extraction would allow managers to track efforts
+related to topics, specifically NFRs.
+These techniques also allow the annotation of bug
+reports and other software artifacts with the NFRs, such as
+portability that they relate to, this would enable querying of bug
+reports and artifacts by relevant NFRs. 
+For instance one could look
+for portability relevant revisions and bug reports.
+
 %% RESEARCH DIDNT DO THIS AND WE ARE TOTALLY AWESOME AND WE DID
@@
 
-\noindent \textbf{The high-level labels} --- To facilitate cross-project comparison, we used a taxonomy of NFRs. This taxonomy is based on the ISO quality model, ISO9126~\cite{iso9126}. 
+\subsubsection{The high-level labels}
+
+%\noindent \textbf{The high-level labels} ---
+
+To facilitate cross-project comparison, we used a taxonomy of NFRs. This taxonomy is based on the ISO quality model, ISO9126~\cite{iso9126}. 
 ISO9126 describes six high-level non-functional requirements (NFRs), maintainability, functionality, portability, efficiency, usability, and reliability. 
@@
 
-\noindent \textbf{Creating a validation corpus} --- To evaluate both unsupervised and supervised classification, we created a validation set of manually labelled topics.
+%\noindent \textbf{Creating a validation corpus} --- 
+\subsubsection{Creating a validation corpus}
+To evaluate both unsupervised and supervised classification, we created a validation set of manually labelled topics.
 For MySQL 3.23 and MaxDB 7.500, we annotated each extracted topic in each period with the same NFR labels as described in the preceding paragraph.
@@
 
-\todo[inline]{NEIL: abram, i got rid of the letter-grade thing because I'm worried people latch on to the idea we are C- students. I feel anything better than random is our criteria here}
+%\todo[inline]{NEIL: abram, i got rid of the letter-grade thing because I'm worried people latch on to the idea we are C- students. I feel anything better than random is our criteria here}
 ROC values provide a score %, similar to school letter-grades (A is 0.9, C is 0.6),
@@
 We consider our labelling classifiers acceptable if they outperform a random classifier (0.5).
+
 \subsection{Unsupervised labelling}
 \label{sec:unsuplabelling}
+
+In this section we describe how to label topics based on dictionaries
+mined from sources external to the projects themselves.
 
@@
 
-\noindent\textbf{Generating word lists} --- In order to automatically label each topic with an NFR from ISO9126, we associate each NFR with a list
+\subsubsection{Generating word lists}
+
+%\noindent\textbf{Generating word lists} ---
+
+In order to automatically label each topic with an NFR from ISO9126, we associate each NFR with a list
 of keywords (`word-lists', in our parlance). These word-lists were determined a priori and were not extracted from the projects themselves.
@@
 
-\todo[inline]{NEIL: this could be a subsubsection}
+\subsubsection{Automatic Labelled Topic Extraction}
+
+%\todo[inline]{NEIL: this could be a subsubsection}
 %\todo[inline]{* the hierarchy of tags was not clear}
-\noindent \textbf{Generating labels automatically} --- Using our three word-lists (\textsf{exp1}, \textsf{exp2}, \textsf{exp3}), we labeled our topics with an NFR where there was a match between a word in the list and the same word somewhere in the distribution of words that constitute the topic.
+%\noindent \textbf{Generating labels automatically} ---
+
+Using our three word-lists (\textsf{exp1}, \textsf{exp2}, \textsf{exp3}), we labeled our topics with an NFR where there was a match between a word in the list and the same word somewhere in the distribution of words that constitute the topic.
 A \emph{named topic} is a topic with a matching NFR label.
@@
 
-\noindent \textbf{Analysis of the unsupervised labelling} --- %Based on the labels, and our manual topic labelling, 
+\subsubsection{Analysis of the unsupervised labelling} %--- %Based
+                                %on the labels, and our manual topic
+                                %labelling, 
+%\noindent \textbf{Analysis of the unsupervised labelling} --- %Based on the labels, and our manual topic labelling, 
 %we compared the results of the unsupervised word matching approach. 
@@
 
-\textbf{Analysis of the supervised labelling} -- 
+\subsubsection{Analysis of the supervised labelling}
 % \label{sec:suplabelling}
@@
 % \todo[inline]{reference project name Nomen more}
-Since our ROC values produced acceptable results, we now turn to \emph{applying} those techniques to understand software maintenance activities. We evaluated two research questions:
+%In order to test  our ROC values produced acceptable results, we now turn to
+%\emph{applying} those techniques to understand software maintenance
+%activities. 
+We evaluated two research questions using the ground truth data that we
+annotated by hand:
 \begin{enumerate}
@@
 % \textsf{Nomen} would help explain why this was the case, since it can identify NFRs which are dominating developer attention.
-\todo[inline]{does this section use supervised learning, and if so what ones?}
+%\todo[inline]{does this section use supervised learning, and if so what ones?}
 
 \subsection{Timeline analysis}
-Figures \ref{fig:mysql-timeline} and \ref{fig:maxdb-timeline} show the temporal patterns of label frequencies. 
+Figures \ref{fig:mysql-timeline} and \ref{fig:maxdb-timeline} show the
+temporal patterns of label frequencies based on the manually annotated
+topics, although this visualization can be generated from the
+results of labelled topic extraction.
 There are two measures represented. 
@@
 
-This analysis allows to address the questions raised in Section \ref{sec:suplabelling}:
-
-\textbf{Do label frequencies change over time?} -- In both projects the frequencies generally decreased with age. 
+%This analysis allows to address the questions raised in Section \ref{sec:suplabelling}:
+This analysis allows to address the questions raised in Section \ref{sec:analysis}:
+
+%\textbf{Do label frequencies change over time?} -- 
+\subsubsection{Do label frequencies change over time?}
+In both projects the frequencies generally decreased with age. 
 However, there are variations within our NFR labels. In MySQL, \emph{usability} and \emph{efficiency} do not appear very often in topics. 
@@
 
-\textbf{Do the different projects differ in their relative topic interest?} -- 
+%\textbf{Do the different projects differ in their relative topic
+%interest?} --
+
+\subsubsection{Do the different projects differ in their relative
+  topic interest?}
+
 Yes. MySQL 3.23 had proportionally more

### 2011-01-24 msr/msr2011.tex
expanded topic modeling defn by a line

@@
 Topic modeling is a machine learning technique which creates multinomial distributions over words from a text corpus.
+The technique infers the hidden structure of a corpus using posterior inference (the probability of the hidden structure given the data).
 Topic models are useful in software maintenance because they summarize the key concepts in a corpus, such as source code, commit comments, or email, by identifying statistically co-occuring words.
@@
 Concrete applications of topics and \emph{labelled topic extraction}
-range from project dashboards to annotating software artifacts with
-NFR-related tags such as revisions and bug reports. 
+range from project dashboards to annotating software artifacts such as revisions and bug reports with
+NFR-related tags.
 Project dashboard are typically employed by managers and are used to
-provide quick summaries of a the efforts put into a software project,
-labelled topic extraction would allow managers to track efforts
-related to topics, specifically NFRs.
-These techniques also allow the annotation of bug
-reports and other software artifacts with the NFRs, such as
-portability that they relate to, this would enable querying of bug
+provide quick summaries of the effort put into a software project. In this case, 
+labelled topic extraction would allow managers to track effort
+related to topics, in this paper, NFRs such as usability.
+These techniques also allow the annotation of commit comments
+ and other software artifacts with NFRs. This would enable querying of bug
 reports and artifacts by relevant NFRs. 
 For instance one could look
-for portability relevant revisions and bug reports.
+for usability-relevant revisions and bug reports.
 

### 2011-01-24 msr/msr2011.tex
moved future work to discussion. shortened conclusions to be more succinct and keep claims reasonable.

@@
 
-\category{D.2.9}{Software Engineering}{Managment}[Lifecycle]
+\category{D.2.9}{Software Engineering}{Management}[Lifecycle]
 \category{D.2.1}{Software Engineering}{Requirements/Specifications}[Tools]
@@
 The technique infers the hidden structure of a corpus using posterior inference (the probability of the hidden structure given the data).
-Topic models are useful in software maintenance because they summarize the key concepts in a corpus, such as source code, commit comments, or email, by identifying statistically co-occuring words.
+Topic models are useful in software maintenance because they summarize the key concepts in a corpus, such as source code, commit comments, or email, by identifying statistically co-occurring words.
 Among other uses, it can quickly give developers an overview of where activity occurred, and gives managers or maintainers a sense of project history.
@@
 Next, we describe how we generated our data (Section \ref{sec:wordlist}).
-For unsupervised classification (Section \ref{sec:unsuplabelling}), we begin by creating word-lists to signifiy when a topic matches an NFR label.
+For unsupervised classification (Section \ref{sec:unsuplabelling}), we begin by creating word-lists to signify when a topic matches an NFR label.
 We then apply our classifier and analyze the results.
@@
 Characteristic area-under-curve value~\cite{Fawcett2006861},
-abbreviated \emph{ROC}, and the F-measure.  F-measure is the harmonic mean of precision and recall.
+abbreviated \emph{ROC}, and the F-measure.  F-measure is the harmonic mean of precision and recall, i.e., $2 * (P * R) / (P + R)$.
 
@@
 %We use these qualities later on as classes in supervised labelling. NE confusing the issue here, I think.
-The terms from ISO9126 may not capture all words occuring in the topics that are nonetheless associated with one of the NFRs. 
+The terms from ISO9126 may not capture all words occurring in the topics that are nonetheless associated with one of the NFRs. 
 For example, the term ``redundancy'' is one most would agree is relevant to discussion of reliability, but is not in the standard. 
@@
   \centering
- \includegraphics[width=\textwidth]{figures/unsupervised-bar}
+ \includegraphics[width=0.8\textwidth]{figures/unsupervised-bar}
  \caption{Performance, ROC values (range: 0-1), of unsupervised topic labelling for
@@
 % \subfloat[MySQL]{
-\includegraphics[width=0.45\textwidth]{figures/both-supervised}
+\includegraphics[width=0.4\textwidth]{figures/both-supervised}
 % \label{fig:subfig1}
@@
 % }
-\caption[]{ROC value for the best learner per label for MaxDB and MySQL. Values range from 0--1.
+\caption[]{ROC value for the best learner per label for MaxDB and MySQL. Values range from 0--1.  Dashed line indicates the performance of a random classifier.
 %\todo[inline]{better discussion}
@@
 }
-\caption[]{MySQL and MaxDB macro (grey) and micro (black) ROC results per multi-label learner. Possible values range from 0--1.
+\caption[]{MySQL and MaxDB macro (grey) and micro (black) ROC results per multi-label learner. Possible values range from 0--1.  Dashed line indicates the performance of a random classifier.
 %\todo[inline]{better discussion}% and graphical legend would be nice}
@@
 supervised learners have the advantage of domain knowledge, which typically means improved results. 
-Our manual annotations were fairly quick to do, taking only a few minutes per 30 day period.
-
-Very rarely did \textsf{exp2} and \textsf{exp3} (naive word matching) ever perform as well as the supervised machine learners. 
+% NEIL - we got criticized for making it sound like this is the obvious thing to do... and then why do you need a learner?
+%Our manual annotations were fairly quick to do, taking only a few minutes per 30 day period. 
+Creating annotated topics (i.e., manual labels) for training is painstaking, but with a suitably representative set of topics, the effort is acceptable.
+To annotate \emph{all} topics took us approximately 20 hours per project, but we estimate only 10\% of the topics need annotation to produce useful results.
+
+Very rarely did \textsf{exp2} and \textsf{exp3} (unsupervised word matching) ever perform as well as the supervised machine learners. 
 For MaxDB, \textit{reliability} was slightly better detected using the static word list of \textsf{exp2}. 
@@
 The learners would occasionally mislabel a topic deserving of an infrequent label with a more common label.
-The word-lists for \emph{correctness} tended to be too lengthy, non-specific and broad, especially if WordNet words were used, as correctness is a loosely defined concept in common parlance.
+The word-lists for \emph{correctness} tended to be too lengthy, non-specific and broad, especially if WordNet words were used, since the NFRs are typically loosely defined concepts in common parlance.
 
@@
 
-If the techniques used in \textsf{exp2} and \textsf{exp3} were combined with the supervised techniques we could reduce the training effort by boosting training sets with topics classified with the unsupervised techniques.
+If the techniques used in \textsf{exp2} and \textsf{exp3} were combined with the supervised techniques, we could reduce the training effort by boosting training sets with topics classified with the unsupervised techniques.
 Both Naive Bayesian learners and the word-list approaches were computationally efficient.  
@@
 \subsection{Threats to validity}
-
 % can always be cut down
 %Our work faced multiple threats to validity and we attempted to address many of them.
-
 \emph{Construct validity} -- we used only commit messages rather than mail or bug tracker messages.
@@
 
-\section{Conclusions and future work}
+\subsection{Future work}
+There are several avenues of further investigation.  
+More external validation would be useful. 
+Although we validated our comparisons using a mailing list for each project, interviews with developers would provide more detail. 
+We also think multi-label learning techniques, although in their infancy, are crucial in understanding cross-cutting concerns such as NFRs. 
+We want to leverage different kinds of artifacts to discover threads of NFR-related discussions that occur between multiple kinds of artifacts.
+Finally, we would like to extend this analysis to other domains, to see what patterns might occur in, for example, a consumer-facing software product.
+
+\section{Conclusions}
+This paper presented a cross-project data mining technique, \emph{labelled topic extraction}.
+The technique leveraged a non-functional requirements (NFRs) taxonomy to mine and label latent software development topics.
+Previous topic analysis research produced topics that needed to be manually labelled, but topics without labels are difficult to interpret, and manually labelling them is costly.
+Furthermore, previous work was project-specific and incomparable.
+To improve this, we leveraged software engineering standards, specifically the ISO quality taxonomy, to produce a method of supervised and unsupervised topic labelling.
+Since the word-list technique is not project specific, it can be used to compare multiple projects.%topics and artifacts from project histories.
+% Our contributions include:
+% \begin{itemize}
+% \item We provided an unsupervised method of topic labelling (word-lists);
+% %\item Our unsupervised method can be applied to projects without training;
+% \item We demonstrated a supervised method of labelling topics by a single NFRs (machine learning);
+% \item We demonstrated a supervised method of labelling topics by multiple NFRs (multi-label machine learning);
+% \item We provided a method of cross-project analysis via topic labelling;
+% \item We demonstrated a method of visualizing these topics across time, which we used to analyze maintenance activities.
+% \end{itemize}
+
+We validated our topic labelling techniques using multiple experiments. 
+We first conducted unsupervised labelling using word-lists. 
+Our next approach was supervised, using single-label and multi-label learners. 
+Both kinds of learners performed well with average
+%area under the 
+ROC values between $0.6$ and $0.8$. These results, along with the efficient performance of our learners, demonstrates that labelled topic extraction is a promising approach for understanding
+
+These labelling techniques allowed us to investigate the occurrence of non-functional requirements in our projects. 
+%Although it may seem like common knowledge that maintenance is a concern as a project matures, our technique was able to show this independently.
+We showed that NFRs are often trending topics in commit comments
+%, and that NFRs are quite common as developer topics.
+%; and that there are efficient methods of automating topic labelling.
+
+%%%HMMM
+
+% We demonstrated multiple supervised and unsupervised methods of topic labelling, including multi-label learners.
+% We provided a case study and visualization of NFR related topics in MaxDB 7.500 and MySQL 3.23. 
+% During the case study we manually labelled hundreds of topics by hand and then used these annotations to validate the positive performance of topic labelling techniques.
+% The implication of our paper is that topics with labels are more useful for analyzing software development activities than are unlabelled topics.
+To furnish stakeholders with labelled topics, our technique provides many methods of automatic topic labelling.
+
+\appendix
+
+Our data and scripts are available at:
+
+ \url{http://softwareprocess.es/nomen/}
+
+\bibliographystyle{abbrv}
+%\bibliographystyle{IEEEtran}
+\bibliography{icpc}
+
+\end{document}
+
+%%% Unused stuff from conclusions
 %\todo[inline]{Need to juxtapose our work against previous work -- contribution}
@@
 % para 1 we did this, we were successful
-In this paper we presented a cross-project data mining technique, \emph{labelled topic extraction}.
-The technique leveraged  a non-functional requirements (NFRs) taxonomy to mine and label latent software development topics.
-%topics and artifacts from project histories.
-Our contributions include:
-\begin{itemize}
-\item We provided an unsupervised method of topic labelling (word-lists);
-%\item Our unsupervised method can be applied to projects without training;
-\item We demonstrated a supervised method of labelling topics by a single NFRs (machine learning);
-\item We demonstrated a supervised method of labelling topics by multiple NFRs (multi-label machine learning);
-\item We provided a method of cross-project analysis via topic labelling;
-\item We demonstrated a method of visualizing these topics across time, which we used to analyze maintenance activities.
-\end{itemize}
+% We found NFRs
+% Our manual inspection and annotation of the topics extracted from MySQL and MaxDB revealed that many of the extracted topics dealt with NFRs, and these topics were spread across the entire history of a project.
+% Try to provide an example of NFRs but little context
+%too specific in this sectionIn the cases of MaxDB and MySQL, portability was a constant maintenance concern and was prevalent throughout the entire lifetime of the projects.
+
+%XXXXXXXXXXXXXXXX I think we need to suggest a couple of uses.
+
+% para 4 summarize and hammer contribution
+% NE not a fan of this phrasing -> % In conclusion,
 
@@
 % NEIL - shortening this
-We validated our topic labelling techniques using multiple experiments. 
-We first conducted unsupervised labelling using word-lists. 
-Our next approach was supervised, using single-label and multi-label learners. 
+
+ % Talk to people
+% We want to investigate developer attitudes related to these labels: i.e., when we label a topic, was the developer expressing positive or negative qualities about that label?  
+% It is difficult to map abstract qualities to particular messages. % what?
+%  % This is disjointed and needs more context
+% On question we seek to answer is if a ``quality'' discussion about more than just corrective maintenance?
+%In this study we take the position that any type of maintenance has to do with software quality.
 % Using unsupervised word-list analysis we set up two NFR topic labelling experiments: 
@@
 % Our fourth experiment evaluated the effectiveness of multi-label learners, such as HOMER.
-Both kinds of learners performed well with 
-%area under the 
-ROC values between $0.6$ and $0.8$.
-
-% para 3 look while we did it we found this
-% total topics - maybe move above
-These labelling techniques allowed us to investigate the occurrence of non-functional requirements in our projects.
-We showed that NFRs are often trending topics; that NFRs are quite common in developer topics; and that there are efficient methods of automating topic labelling.
-Although it may seem like common knowledge that maintenance is a concern as a project matures, our technique was able to show this independently.
-
-% We found NFRs
-% Our manual inspection and annotation of the topics extracted from MySQL and MaxDB revealed that many of the extracted topics dealt with NFRs, and these topics were spread across the entire history of a project.
-% Try to provide an example of NFRs but little context
-%too specific in this sectionIn the cases of MaxDB and MySQL, portability was a constant maintenance concern and was prevalent throughout the entire lifetime of the projects.
-
-%XXXXXXXXXXXXXXXX I think we need to suggest a couple of uses.
-
-% para 4 summarize and hammer contribution
-% NE not a fan of this phrasing -> % In conclusion, 
-Previous topic analysis research produced topics that needed to be manually labelled, but topics without labels are difficult to interpret.
-To improve this, we leveraged software engineering standards to produce a method of domain specific topic labelling.
-Since the word-list technique is not project specific, it can be used to compare multiple projects.
-We demonstrated multiple supervised and unsupervised methods of topic labelling, including multi-label learners.
-We provided a case study and visualization of NFR related topics in MaxDB 7.500 and MySQL 3.23. 
-During the case study we manually labelled hundreds of topics by hand and then used these annotations to validate the positive performance of topic labelling techniques.
-The implication of our paper is that topics with labels are more useful for analyzing software development activities than are unlabelled topics.
-To furnish stakeholders with labelled topics, our technique provides many methods of automatic topic labelling.
-
-%XXXXXXXX Optional  
-%\subsection{Future Work}
-
- % para 5 future work
- % Future work
-There are several avenues of further investigation.  
-More external validation would be useful. 
-Although we validated our comparisons using a mailing list for each project, interviews with developers would provide more detail. 
-We also think multi-label learning techniques, although in their infancy, are crucial in understanding cross-cutting concerns such as NFRs. 
-We want to leverage different kinds of artifacts to discover threads of NFR-related discussions that occur between multiple kinds of artifacts.
-Finally, we would like to extend this analysis to other domains, to see what patterns might occur in, for example, a consumer-facing software product.
- % Talk to people
-% We want to investigate developer attitudes related to these labels: i.e., when we label a topic, was the developer expressing positive or negative qualities about that label?  
-% It is difficult to map abstract qualities to particular messages. % what?
-%  % This is disjointed and needs more context
-% On question we seek to answer is if a ``quality'' discussion about more than just corrective maintenance?
-%In this study we take the position that any type of maintenance has to do with software quality.
-
-\appendix
-
-Our data and scripts are available at:
-
- \url{http://softwareprocess.es/nomen/}
-
-\bibliographystyle{abbrv}
-%\bibliographystyle{IEEEtran}
-\bibliography{icpc}
-
-\end{document}
 

### 2011-01-25 msr/msr2011.tex
reformat a paragraph so I can edit it..

@@
 from source control systems such as CVS and BitKeeper.  
-These topics are given labels from to a generalizable
+These topics are given labels from a generalizable
 cross-project taxonomy, consisting of non-functional
@@
 
-One challenge for software maintenance is to identify the reasons behind software development activities.
-Identifying these reasons provides insight into \emph{why} certain changes were made.
-Topic modeling is a machine learning technique which creates multinomial distributions over words from a text corpus.
-The technique infers the hidden structure of a corpus using posterior inference (the probability of the hidden structure given the data).
-Topic models are useful in software maintenance because they summarize the key concepts in a corpus, such as source code, commit comments, or email, by identifying statistically co-occurring words.
-Among other uses, it can quickly give developers an overview of where activity occurred, and gives managers or maintainers a sense of project history.
-For example, a topic extracted from a set of commit logs might be \emph{ ``listener change remove add fire''}.
+One challenge for software maintenance is to identify the reasons
+behind software development activities. 
+Identifying these reasons provides insight into \emph{why} certain
+changes were made. 
+Topic modeling is a machine learning technique which creates
+multinomial distributions over words from a text corpus. 
+The technique infers the hidden structure of a corpus using posterior
+inference (the probability of the hidden structure given the data). 
+Topic models are useful in software maintenance because they summarize
+the key concepts in a corpus, such as source code, commit comments, or
+email, by identifying statistically co-occurring words. 
+Among other uses, it can quickly give developers an overview of where
+activity occurred, and gives managers or maintainers a sense of
+project history. 
+For example, a topic extracted from a set of commit logs might be
+\emph{ ``listener change remove add fire''}. 
 A major challenge for topic modeling is to accurately identify the summary label for a topic (in the previous case, perhaps \emph{``event-handling''}). 
-Current approaches are manual: they rely on human expertise to recognize the appropriate label.
-Furthermore, to date the results of topic modeling are project specific, and not generalizable.
+Current approaches are manual: they rely on human expertise to
+recognize the appropriate label. 
+Furthermore, to date the results of topic modeling are project
+specific, and not generalizable. 
 This paper presents an approach, \emph{labelled topic extraction}, which addresses these two problems.

### 2011-01-25 msr/msr2011.tex
more intro edits

@@
 The technique infers the hidden structure of a corpus using posterior
-inference (the probability of the hidden structure given the data). 
+inference: the probability of the hidden structure given the data. 
 Topic models are useful in software maintenance because they summarize
 the key concepts in a corpus, such as source code, commit comments, or
-email, by identifying statistically co-occurring words. 
+mailing-list messages, by identifying statistically co-occurring words. 
 Among other uses, it can quickly give developers an overview of where
@@
 \emph{ ``listener change remove add fire''}. 
-A major challenge for topic modeling is to accurately identify the summary label for a topic (in the previous case, perhaps \emph{``event-handling''}). 
+A major challenge for topic modeling is to accurately identify the
+summary label for a topic; in the previous case, perhaps
+\emph{``event-handling''} would be appropriate. 
 Current approaches are manual: they rely on human expertise to

### 2011-01-25 msr/msr2011.tex
another tiny intro edit

@@
 % I was told if you're going to use a paren, you don't need to say it.
-There is a series of standards on NFRs, such as \cite{iso9126}, specifically intended to apply to projects of varying types.
+There is a series of standards on NFRs, such as ISO9126~\cite{iso9126}, specifically intended to apply to projects of varying types.
 These standards indicate our approach is possible. %, if perhaps not universally accepted).
@@
 These techniques also allow the annotation of commit comments
- and other software artifacts with NFRs. This would enable querying of bug
-reports and artifacts by relevant NFRs. 
+ and other software artifacts with NFRs. 
+This would enable querying of bug reports and artifacts by relevant NFRs. 
 For instance one could look
-for usability-relevant revisions and bug reports.
+for \emph{usability}-relevant revisions and bug reports.
 

### 2011-01-25 msr/msr2011.tex
edit contributions
previous work edits

@@
 \item We provide a method of cross-project analysis via topic labelling;
-\item We apply these techniques to visualizing NFRs over time, to analyze maintenance activities.
+\item We apply these techniques to visualize NFRs over time, and to analyze maintenance activities.
 \end{itemize}
@@
 
-The idea of extracting higher-level `concerns', also known as `concepts', `aspects' or `requirements', has been approached from documentation-based and repository-based perspectives.
+The idea of extracting higher-level \emph{concerns}, also known as \emph{concepts}, \emph{aspects} or \emph{requirements}, has been approached from documentation-based and repository-based perspectives.
 
@@
 Furthermore, these labels are project-specific, because they are generated from the data of that project. E.g., we might have a label called `Oracle' in the MySQL case, since Oracle owns MySQL. 
-
-This paper does this differently. First of all, we automated the process of naming the topics. Secondly, we labelled topics with project-independent terms, in order to permit cross-project comparison.
+We do this differently.
+%This paper does this differently. 
+First of all, we automated the process of naming the topics. Secondly, we labelled topics with project-independent terms, in order to permit cross-project comparison.
 

### 2011-01-25 msr/msr2011.tex
study design and execution edits
latest edits

@@
 Figure \ref{fig:process} gives an outline of the methodology followed. 
-We began by gathering source data and creating topic models. For unsupervised labelling, we generate three sets of word-lists as signifiers for NFRs.
+We began by gathering source data and creating topic models. For unsupervised labelling, we generated three sets of word-lists as signifiers for NFRs.
 %e created labels using a well-known NFR taxonomy. 
@@
 The MySQL project started in 1994 and MySQL 3.23 was released in early 2001. 
-MySQL contains 320 thousand lines of C and C++ source code.  
+MySQL contains $320 000$ lines of C and C++ source code.  
 Choosing an older version of these projects allowed us to focus on projects which have moved further into the maintenance phase of the software life-cycle.
@@
 We summarized each message as a word distribution but removed stop-words such as \emph{the} and \emph{at}. 
-Stemming was performed in the later stages of our analysis. % WAS IT?
+%Stemming was performed in the later stages of our analysis. % WAS IT?
 
@@
 This size of period, 30 days, is smaller than the time between minor releases but large enough for there to be sufficient commits to analyze~\cite{Hindle09ICSM}. 
-For each 30 day period of each project, we input the messages of that period into Latent Dirichlet Allocation (LDA), a topic analysis algorithm~\cite{Blei2003}, and recorded the topics the algorithm returned.
+For each 30 day period of each project, we input the messages of that period into Latent Dirichlet Allocation (LDA), a topic analysis algorithm~\cite{Blei2003}, and recorded the topics the algorithm extracted.
 %For each project and for each period of messages a project, we input these messages 
@@
 
-A topic analysis tool such as LDA will try to find $N$ independent word distributions within the word distributions of all the messages. 
+A topic analysis tool such as LDA will try to find $N$ independent
+word distributions within the word distributions of all input messages. 
 Linear combinations of these $N$ word distributions are meant to represent and recreate the word distributions of any of the original messages. 

### 2011-01-25 msr/msr2011.tex
1/2 editted

@@
 For example, the label \emph{maintainability} is associated with words \emph{ease} and \emph{ownership}.
-In general, as we proceed from word-list in \textsf{exp1} to that in \textsf{exp3}, our lists get more generic.
+In general, as we proceed from word-list in \textsf{exp1} to that in \textsf{exp3}, our lists become more generic.
 
@@
 Using our three word-lists (\textsf{exp1}, \textsf{exp2}, \textsf{exp3}), we labeled our topics with an NFR where there was a match between a word in the list and the same word somewhere in the distribution of words that constitute the topic.
-A \emph{named topic} is a topic with a matching NFR label.
-Unnamed topics occur where this no such match, which may indicate either a lack of precision, or simply that this topic is not associated with non-functional requirements.
+A \emph{named topic} is a topic with a matching NFR label. Unnamed
+topics occur where there is no such match, which may indicate either a lack of precision, or simply that this topic is not associated with non-functional requirements.
 All experiments were run on MaxDB 7.500 and MySQL 3.23 data. LDA
-extracted 20 topics per period for each.
+extracted 20 topics per period for each project.
 % Is this good enough

### 2011-01-25 msr/msr2011.tex
todos for viz

--- 2011-01-25 msr/msr2011.tex
@@
 from source control systems such as CVS and BitKeeper.  
-These topics are given labels from a generalizable
+These topics are given labels from to a generalizable
 cross-project taxonomy, consisting of non-functional
@@
 
-One challenge for software maintenance is to identify the reasons
-behind software development activities. 
-Identifying these reasons provides insight into \emph{why} certain
-changes were made. 
-Topic modeling is a machine learning technique which creates
-multinomial distributions over words from a text corpus. 
-The technique infers the hidden structure of a corpus using posterior
-inference: the probability of the hidden structure given the data. 
-Topic models are useful in software maintenance because they summarize
-the key concepts in a corpus, such as source code, commit comments, or
-mailing-list messages, by identifying statistically co-occurring words. 
-Among other uses, it can quickly give developers an overview of where
-activity occurred, and gives managers or maintainers a sense of
-project history. 
-For example, a topic extracted from a set of commit logs might be
-\emph{ ``listener change remove add fire''}. 
-A major challenge for topic modeling is to accurately identify the
-summary label for a topic; in the previous case, perhaps
-\emph{``event-handling''} would be appropriate. 
-Current approaches are manual: they rely on human expertise to
-recognize the appropriate label. 
-Furthermore, to date the results of topic modeling are project
-specific, and not generalizable. 
+One challenge for software maintenance is to identify the reasons behind software development activities.
+Identifying these reasons provides insight into \emph{why} certain changes were made.
+Topic modeling is a machine learning technique which creates multinomial distributions over words from a text corpus.
+The technique infers the hidden structure of a corpus using posterior inference (the probability of the hidden structure given the data).
+Topic models are useful in software maintenance because they summarize the key concepts in a corpus, such as source code, commit comments, or email, by identifying statistically co-occurring words.
+Among other uses, it can quickly give developers an overview of where activity occurred, and gives managers or maintainers a sense of project history.
+For example, a topic extracted from a set of commit logs might be \emph{ ``listener change remove add fire''}.
+A major challenge for topic modeling is to accurately identify the summary label for a topic (in the previous case, perhaps \emph{``event-handling''}).
+Current approaches are manual: they rely on human expertise to recognize the appropriate label.
+Furthermore, to date the results of topic modeling are project specific, and not generalizable.
 This paper presents an approach, \emph{labelled topic extraction}, which addresses these two problems.
@@
 % I was told if you're going to use a paren, you don't need to say it.
-There is a series of standards on NFRs, such as ISO9126~\cite{iso9126}, specifically intended to apply to projects of varying types.
+There is a series of standards on NFRs, such as \cite{iso9126}, specifically intended to apply to projects of varying types.
 These standards indicate our approach is possible. %, if perhaps not universally accepted).
@@
 These techniques also allow the annotation of commit comments
- and other software artifacts with NFRs. 
-This would enable querying of bug reports and artifacts by relevant NFRs. 
+ and other software artifacts with NFRs. This would enable querying of bug
+reports and artifacts by relevant NFRs. 
 For instance one could look
-for \emph{usability}-relevant revisions and bug reports.
+for usability-relevant revisions and bug reports.
 
@@
 \item We provide a method of cross-project analysis via topic labelling;
-\item We apply these techniques to visualize NFRs over time, and to analyze maintenance activities.
+\item We apply these techniques to visualizing NFRs over time, to analyze maintenance activities.
 \end{itemize}
@@
 
-The idea of extracting higher-level \emph{concerns}, also known as \emph{concepts}, \emph{aspects} or \emph{requirements}, has been approached from documentation-based and repository-based perspectives.
+The idea of extracting higher-level `concerns', also known as `concepts', `aspects' or `requirements', has been approached from documentation-based and repository-based perspectives.
 
@@
 Furthermore, these labels are project-specific, because they are generated from the data of that project. E.g., we might have a label called `Oracle' in the MySQL case, since Oracle owns MySQL.
-We do this differently.
-%This paper does this differently. 
-First of all, we automated the process of naming the topics. Secondly, we labelled topics with project-independent terms, in order to permit cross-project comparison.
+
+This paper does this differently. First of all, we automated the process of naming the topics. Secondly, we labelled topics with project-independent terms, in order to permit cross-project comparison.
 
@@
 %activities. 
+\todo[inline]{more on intro ... why should the reader care about this. It needs to show how the stuff in the intro (applications) is possible}
+\todo[inline]{fix section headings}
+
 We evaluated two research questions using the ground truth data that we
@@
 %This analysis allows to address the questions raised in Section \ref{sec:suplabelling}:
-This analysis allows to address the questions raised in Section \ref{sec:analysis}:
+This analysis addresses the questions raised in Section \ref{sec:analysis}:
 
@@
 
+\todo[inline]{link the actual data, which is interesting, to somethign actionable for a developer or manager}
 \section{Discussion}

### 2011-01-25 msr/msr2011.tex
Mike's changes to intro

@@
 
-One challenge for software maintenance is to identify the reasons
-behind software development activities. 
-Identifying these reasons provides insight into \emph{why} certain
-changes were made. 
+%One challenge for software maintenance is to identify the reasons
+%behind software development activities. 
+%Identifying these reasons provides insight into \emph{why} certain
+%changes were made. 
+%migod
+A key problem for practising software maintainers is gaining an understanding of \emph{why} a system has evolved the way it has.
+Looking back on streams of artifacts scattered across different repositories, inferring what activities were performed, when, and for what reasons, is hard to do without expert advice from the developers involved.
+In this work we seek to provide automated help in labelling development topics that have been extracted from commit logs.
+
 Topic modeling is a machine learning technique which creates
@@
 mailing-list messages, by identifying statistically co-occurring words. 
-Among other uses, it can quickly give developers an overview of where
-activity occurred, and gives managers or maintainers a sense of
+Among other uses, it can quickly give developers an overview of where significant
+activity has occurred, and gives managers or maintainers a sense of
 project history. 
-For example, a topic extracted from a set of commit logs might be
-\emph{ ``listener change remove add fire''}. 
-A major challenge for topic modeling is to accurately identify the
-summary label for a topic; in the previous case, perhaps
-\emph{``event-handling''} would be appropriate. 
-Current approaches are manual: they rely on human expertise to
-recognize the appropriate label. 
-Furthermore, to date the results of topic modeling are project
-specific, and not generalizable. 
-This paper presents an approach, \emph{labelled topic extraction}, which addresses these two problems.
+For example, a topic extracted from a set of commit logs might consist of the following terms: \emph{ ``listener change remove add fire''}. 
+%A major challenge for topic modeling is to accurately identify the
+%summary label for a topic; in the previous case, perhaps
+%\emph{``event-handling''} would be appropriate. 
+%Current approaches are manual: they rely on human expertise to
+%recognize the appropriate label. 
+%Furthermore, to date the results of topic modeling are project
+%specific, and not generalizable. 
+%This paper presents an approach, \emph{labelled topic extraction}, which addresses these two problems.
+
+%migod
+While machine learning techniques can automatically identify clumps of commonly recurring terms, devising an appropriate summary label for each clump/topic is harder.
+For example, the word list above might reasonably be labelled as \emph{"event handling"} by a human who understands the domain well, despite the fact that label does not appear in the word list itself.
+Current approaches to topic labelling rely on manual intervention by human experts, and also are limited to project-specific topic labels.
+In this paper, we introduce \emph{labelled topic extraction}, an approach to topic labelling that is both automatic and creates labels that are project independent.
 
@@
 
-In our previous work we dealt with topic trends, which are topics that recur over time~\cite{Hindle09ICSM}. 
-We observed that topic trends were often non-functional requirements (NFRs). 
-NFRs have the property of being cross-domain and widely applicable. 
-In this sense, they are useful abstractions for developer conversations about different software projects.
+%In our previous work we dealt with topic trends, which are topics that recur over time~\cite{Hindle09ICSM}. 
+%We observed that topic trends were often non-functional requirements (NFRs). 
+%NFRs have the property of being cross-domain and widely applicable. 
+%In this sense, they are useful abstractions for developer conversations about different software projects.
 
 % harp on data mining more
-In general, the mining of software artifacts tends to be very project specific, yet NFRs are not. 
+%In general, the mining of software artifacts tends to be very project specific, yet NFRs are not. 
 % I was told if you're going to use a paren, you don't need to say it.
-There is a series of standards on NFRs, such as ISO9126~\cite{iso9126}, specifically intended to apply to projects of varying types.
-These standards indicate our approach is possible. %, if perhaps not universally accepted).
+%There is a series of standards on NFRs, such as ISO9126~\cite{iso9126}, specifically intended to apply to projects of varying types.
+%These standards indicate our approach is possible. %, if perhaps not universally accepted).
 % NFRs are prevalent in almost all software systems.
@@
 %This informs our method of generalizing the data mining of these software artifacts to leverage the global domain knowledge of software engineering, specifically NFRs. 
-In particular, we use software quality models to generalize topics across projects. 
-This allows us to mine software systems with the intention of comparing their NFR-related development topics.
+%In particular, we use software quality models to generalize topics across projects. 
+%This allows us to mine software systems with the intention of comparing their NFR-related development topics.
 % in order to achieve generalizable, cross project data mining results.
@@
 % AH: Project Dashboards, summarization, software process recovery
+
+%migod
+In general, the fruits of mining software artifacts are often project specific and hard to generalize.
+However, in our previous work we investigated \emph{topic trends} --- that is, topics that recur over time --- we observed that topic trends often corresponded to non-functional requirements (NFRs) \cite{Hindle09ICSM}.
+This is encouraging, as NFRs have the property of being cross-domain and widely applicable.
+In this sense, they are useful abstractions for developer conversations about different software projects.
+Furthermore, there is a series of standards on NFRs, such as ISO9126 \cite{iso9126}, that are specifically intended to apply to projects of varying types; this suggests that our goal of trying to extract NFR-related development topics, such as those related to software quality models, holds promise.
 
@@
 labelled topic extraction would allow managers to track effort
-related to topics, in this paper, NFRs such as usability.
-These techniques also allow the annotation of commit comments
+related to topics, such as usability or other NFRs.
+These techniques also allow for the annotation of commit comments
  and other software artifacts with NFRs. 
@@
 \begin{itemize}
-\item We introduce the concept of labelled topic extraction, using a non-functional requirements (NFR) taxonomy for our labels. 
-\item We evaluate an unsupervised method of automated topic labelling (word-lists);
+\item introduces the concept of labelled topic extraction, using a non-functional requirements (NFR) taxonomy for our labels. 
+\item evaluates an unsupervised method of automated topic labelling (word-lists);
 %\item Our unsupervised method can be applied to projects without training;
-\item We evaluate a supervised method of automatically labelling topics with a single NFR (machine learning);
-\item We evaluate a supervised method of labelling topics with multiple NFRs (multi-label machine learning);
-\item We provide a method of cross-project analysis via topic labelling;
-\item We apply these techniques to visualize NFRs over time, and to analyze maintenance activities.
+\item evaluates a supervised method of automatically labelling topics with a single NFR (machine learning);
+\item evaluates a supervised method of labelling topics with multiple NFRs (multi-label machine learning);
+\item provides a method of cross-project analysis via topic labelling; and
+\item applies these techniques to visualize NFRs over time, and to analyze maintenance activities.
 \end{itemize}

### 2011-01-26 msr/msr2011.tex
page 3 of migod edits

@@
 %We first introduces labelled topic extraction, and then uses that technique to analyze development activity in two large-scale examples.
-We begin with related work in Section \ref{sec:related}.
+We begin by discussing related work in Section \ref{sec:related}.
 Next, we describe how we generated our data (Section \ref{sec:wordlist}).
@@
 \section{Study design and execution}
-Figure \ref{fig:process} gives an outline of the methodology followed. 
+Figure \ref{fig:process} gives an outline of methodology.
 We began by gathering source data and creating topic models. For unsupervised labelling, we generated three sets of word-lists as signifiers for NFRs.
 %e created labels using a well-known NFR taxonomy. 
-In supervised learning, we trained our data with manual annotations in order to match topics with NFRs.
-Finally, these topics were used to analyze the role of non-functional requirements in software maintenance.
+In supervised learning, we trained our data with manual annotations in order to match topics with NFRs. Finally, these topics were used to analyze the role of NFRs in software maintenance.
 
@@
 MaxDB started in the late 1970s as a research project, and was later acquired by SAP. 
-As of version 7.500, released April 2007, the project has 940 thousand lines of C source code\footnote{generated using David A. Wheeler's \emph{SLOCCount}, dwheeler.com/sloccount.}. 
+As of version 7.500, released April 2007, the project has over $940,000$ lines of C source code\footnote{generated using David A. Wheeler's \emph{SLOCCount}, dwheeler.com/sloccount.}. 
 The MySQL project started in 1994 and MySQL 3.23 was released in early 2001. 
-MySQL contains $320 000$ lines of C and C++ source code.  
-Choosing an older version of these projects allowed us to focus on projects which have moved further into the maintenance phase of the software life-cycle.
-
-For each project, we used source control commit comments, the messages that programmers write when they commit revisions to a source control repository. 
-This is the most readily accessible source of project interactions for outside researchers, since developers often or always (e.g., Git) write such message.
+MySQL contains $320,000$ lines of C and C++ source code.  
+%Choosing an older version of these projects allowed us to focus on projects which have moved further into the maintenance phase of the software life-cycle.
+For our study, we explicitly chose older versions of mature projects from a stable problem domain to increase the likelihood that we would encounter primarily maintenance activities in our studies.
+
+For each project, we used source control commit comments, the messages
+that programmers write when they commit revisions to a source control
+repository. 
+Commit comments are often studied by researchers, as they are the most readily accessible source of project interactions, and developers are often required to create them by the repository mechanism (e.g., Git).
+Additionally, relying only on commit comments makes our approach more generalizable, as we do not assume the presence of other artifact corpora.
+%This is the most readily accessible source of project interactions for outside researchers, since developers often or always (e.g., Git) write such message.
 Our approach is more generalizable if it does not assume the presence of other corpora.
 %bug trackers and email-list archives were not available for both projects.
-An example of a typical commit message is: \textit{``history annotate diffs bug fixed (if mysql\-\_real\-\_connect() failed there were two pointers to malloc'ed strings, with memory corruption on free(), of course)''}.
+An example of a typical commit message, from MySQL, is: \textit{``history annotate diffs bug fixed (if mysql\-\_real\-\_connect() failed there were two pointers to malloc'ed strings, with memory corruption on free(), of course)''}.
 We extracted these messages and indexed them by creation time. 
-We summarized each message as a word distribution but removed stop-words such as \emph{the} and \emph{at}. 
+We summarized each message as a word distribution but removed stop-words such as ``\emph{the}'' and ``\emph{at}''. 
 %Stemming was performed in the later stages of our analysis. % WAS IT?
 
-For the commit message data-sets of each project, we created an XML file which separated commits into 30 day periods. 
-This size of period, 30 days, is smaller than the time between minor releases but large enough for there to be sufficient commits to analyze~\cite{Hindle09ICSM}. 
+For the commit message data-sets of each project, we created an XML file that separated commits into 30 day periods. 
+%This size of period, 30 days, is smaller than the time between minor
+%releases but large enough for there to be sufficient commits to
+%analyze
+We chose a period size of 30 days as it is smaller than the time between minor releases but large enough for there to be sufficient commits to analyze~\cite{Hindle09ICSM}. 
 For each 30 day period of each project, we input the messages of that period into Latent Dirichlet Allocation (LDA), a topic analysis algorithm~\cite{Blei2003}, and recorded the topics the algorithm extracted.
@@
 These $N$ word distributions effectively form topics: cross-cutting collections of words relevant to one or more of our commit messages. 
-LDA extracts topics in an unsupervised manner; the algorithm relies solely on the source data, word distributions of messages, with no human intervention.
+LDA extracts topics in an unsupervised manner; the algorithm relies
+solely on the source data and word distributions of messages, with no human intervention.
 
@@
 To facilitate cross-project comparison, we used a taxonomy of NFRs. This taxonomy is based on the ISO quality model, ISO9126~\cite{iso9126}. 
-ISO9126 describes six high-level non-functional requirements (NFRs), maintainability, functionality, portability, efficiency, usability, and reliability. 
-There is some debate about the terms in this model. 
-However, ISO9126 is ``an international standard and thus provides an internationally accepted terminology for software quality \cite[p.
+ISO9126 describes six high-level NFRs: maintainability, functionality,
+portability, efficiency, usability, and reliability~\footnote{While there may be lingering debate in some circles about these terms, an ISO standard seems like a reasonable starting point for our work.}
+%There is some debate about the terms in this model. 
+%However, ISO9126 is ``an international standard and thus provides an internationally accepted terminology for software quality \cite[p.
 58]{Boegh2008},'' that is sufficient for the purposes of this research.
@@
 \subsubsection{Creating a validation corpus}
-To evaluate both unsupervised and supervised classification, we created a validation set of manually labelled topics.
-For MySQL 3.23 and MaxDB 7.500, we annotated each extracted topic in each period with the same NFR labels as described in the preceding paragraph.
+To evaluate both unsupervised and supervised classification, we
+created a validation set of manually labelled topics. For MySQL 3.23
+and MaxDB 7.500, we annotated each extracted topic in each period with
+the six NFR labels listed above.
 We looked at each period's topics, and assessed what the data --- consisting of the frequency-weighted word lists and messages --- suggested was the topic for that period. 
@@
 Characteristic area-under-curve value~\cite{Fawcett2006861},
-abbreviated \emph{ROC}, and the F-measure.  F-measure is the harmonic mean of precision and recall, i.e., $2 * (P * R) / (P + R)$.
+abbreviated \emph{ROC}, and the F-measure, which is the harmonic mean of precision and recall, i.e., $2 * (P * R) / (P + R)$.
 

### 2011-01-26 msr/msr2011.tex
migod edits II

@@
 
-In order to automatically label each topic with an NFR from ISO9126, we associate each NFR with a list
+In order to automatically label each topic with 
+one of the six high-level NFRs,
+%an NFR from ISO9126, 
+we associate each NFR with a list
 of keywords (`word-lists', in our parlance). These word-lists were determined a priori and were not extracted from the projects themselves.
 We intersected the words of the topics and the words of our word-lists.
-We `labeled' a topic if any of its words matched any of the word-list's words.
+We ``labeled'' a topic if any of its words matched any of the word-list's words.
 A topic could match more than one NFR.
-We used several different sets of word-lists for comparison, which we refer to as \textsf{exp1, exp2, exp3} in the text which follows.
+We used several different sets of word-lists for comparison, which we
+refer to as \textsf{exp1}, \textsf{exp2}, and \textsf{exp3} in the text which follows.
 
@@
 The terms from ISO9126 may not capture all words occurring in the topics that are nonetheless associated with one of the NFRs. 
-For example, the term ``redundancy'' is one most would agree is relevant to discussion of reliability, but is not in the standard. 
+For example, the term ``redundancy'' is one 
+we considered to be
+relevant to discussion of reliability, but is not in the standard. 
 We therefore took the NFRs from the ISO9126 and added to them.
@@
 the project being labelled, and we did not label the project's topics
-ourselves for a training set. The theory behind this technique is that
+ourselves for a training set. The motivation behind this technique is that
 because most software often addresses similar issues, we can use the
@@
 
-For \textsf{exp1}, our best performing labels, those with the most topics, were \emph{correctness} (182 topics) and \emph{testability} (121). 
+For \textsf{exp1}, our best performing 
+labels --- that is, those with the most topics --- were
+%labels, those with the most topics, 
+\emph{correctness} (182 topics) and \emph{testability} (121). 
 We did not get good results for \emph{usability} or \emph{accuracy}, which were associated with fewer than ten topics. 
@@
 For \textsf{exp2}, there are more unnamed topics than \textsf{exp1}. 
-Only \emph{reliability} produces a lot of matches, mostly with the word `error'. 
+Only \emph{reliability} produces a lot of matches, mostly with the word ``error''. 
 Co-occurrence results were poor. This suggests our word lists were overly restrictive.
@@
 For \textsf{exp3}, we generally labelled more topics. 
-As we mentioned, the word-lists are quite broad, so there are likely to be false-positives (discussed below). 
+As we mentioned, the word-lists are broad, so there are likely to be false-positives (discussed below). 
 We found a high of 265 topics for \emph{usability}, with a low of 44 topics for \emph{maintainability}. 
@@
 %EXP1 could not be evaluated properly!
-Because our ground truth annotations were only relevant to ISO-9126,
+Because our ground truth annotations were relevant only to ISO9126,
 we estimate that \textsf{exp1} had poor
@@
 topics labelled \emph{functionality}, while MaxDB had proportionally more
-\emph{efficiency} related topics. MaxDB was a very mature release `donated' to the open-source community. 
+\emph{efficiency} related topics. MaxDB was a very mature release ``donated'' to the open-source community. 
 MySQL, on the other hand, was in its relative infancy. 
@@
 We found many topics that were not non-functional requirements (NFRs) but were often related to them. 
-For instance, concurrency was mentioned often in the commit logs and was related to \emph{correctness} and \emph{reliability}, likely because concurrent code is prone to bugs such as race conditions.
-%are
+For instance, concurrency was mentioned often in the commit logs and
+was related to \emph{correctness} and \emph{reliability}, possibly because concurrent code is prone to bugs such as race conditions. %are 
 % often  was troublesome. 
@@
 \section{Conclusions}
-This paper presented a cross-project data mining technique, \emph{labelled topic extraction}.
-The technique leveraged a non-functional requirements (NFRs) taxonomy to mine and label latent software development topics.
-Previous topic analysis research produced topics that needed to be manually labelled, but topics without labels are difficult to interpret, and manually labelling them is costly.
-Furthermore, previous work was project-specific and incomparable.
+This paper presented a cross-project data mining technique,
+\emph{labelled topic extraction}. The technique leveraged a
+non-functional requirements (NFRs) taxonomy to mine and label latent
+software development topics. Previous topic analysis research produced
+topics that needed to be manually labelled, but topics without labels
+are difficult to interpret, and manually labelling them is
+costly. Furthermore, previous work was project specific and incomparable. 
 To improve this, we leveraged software engineering standards, specifically the ISO quality taxonomy, to produce a method of supervised and unsupervised topic labelling.

### 2011-01-26 msr/msr2011.tex
latest edits

@@
 MySQL had reasonable results within \textsf{exp2} for \emph{reliability} and \emph{efficiency}. 
-MySQL's results for \emph{efficiency} did not improve in \textsf{exp3} but other qualities such as \emph{functionality} did improve. 
-If a \emph{C} grade performance has a ROC value of $0.6$ then most of these tests scored a grade of \emph{C} or less, but our classifier nonetheless performed substantially better than random.
+MySQL's results for \emph{efficiency} did not improve in \textsf{exp3}
+but other qualities such as \emph{functionality} did improve. 
+Many of ROC scores were $0.6$ or less, but our classifier nonetheless performed substantially better than random.
 

### 2011-01-26 msr/msr2011.tex
more abram editting

@@
 %activities. 
-\todo[inline]{more on intro ... why should the reader care about this. It needs to show how the stuff in the intro (applications) is possible}
-\todo[inline]{fix section headings}
-
+\todo[inline]{more on intro ... why should the reader care about
+  this. It needs to show how the stuff in the intro (applications) is
+  possible}
+%%% XXX NEIL I don't get this.. VVVVV
+%\todo[inline]{fix section headings}
+
+In this section we investigate the history of the two projects that we
+studied.
+We demonstrate the value of topic analysis as it is capable of showing
+the topic of development efforts over time in each project.
 We evaluated two research questions using the ground truth data that we
@@
 % and                   http://softwareprocess.es/y/maxdb-7500.tar.gz
-There are more \textsf{None} labels in MaxDB because a number of topics were to do with code cleanup or automated checking using tools like \texttt{sutcheck}, a tool-specific to the development process at MaxDB.
+There are more \textsf{None} labels in MaxDB because a number of topics were to do with code cleanup or automated checking using tools like \texttt{sutcheck}, a tool-specific to the development process of MaxDB.
 
@@
 MySQL, on the other hand, was in its relative infancy. 
-Security problems were more common (security is a component of `functionality' in the ISO 9126 model). 
+Security problems were more common (security is a component of \emph{functionality} in the ISO 9126 model). 
 In both cases \emph{portability} was a constant maintenance concern and was prevalent throughout the entire lifetime of the projects.
 
-\todo[inline]{link the actual data, which is interesting, to somethign actionable for a developer or manager}
+\subsection{Summary}
+
+\todo[inline]{link the actual data, which is interesting, to somethign
+  actionable for a developer or manager}
+
 \section{Discussion}
@@
 We improved internal validity by trying to correlate and explain the behaviours observed in the analysis with the historical records of the projects.
-We did not attempt to match our results to any particular model or theory, as such theories are sparse in
+We did not attempt to match our results to any particular model or theory.
 

### 2011-01-26 msr/msr2011.tex
updates to layout, and rewrite of analysis section

@@
 \begin{itemize}
-\item introduces the concept of labelled topic extraction, using a non-functional requirements (NFR) taxonomy for our labels. 
+\item introduces the concept of labelled topic extraction, using a non-functional requirements (NFR) taxonomy for our labels; 
 \item evaluates an unsupervised method of automated topic labelling (word-lists);
@@
   \centering
- \includegraphics[width=0.8\textwidth]{figures/unsupervised-bar}
+ \includegraphics[width=0.7\textwidth]{figures/unsupervised-bar}
  \caption{Performance, ROC values (range: 0-1), of unsupervised topic labelling for
@@
  had the highest ROC value for that label. 
-The best learner is important because one uses a single learner per label.
+The best learner is important because one uses a single learner per label. In applying our technique, we expect the user to select, for each NFR, the best learner possible.
 %Figure \ref{fig:best-learn-per-tag} uses the ZeroR learner as a baseline, a common practice in machine learning,
@@
 Figure \ref{fig:best-learn-per-tag} shows that MaxDB and MySQL have quite different results, as the ROC values for \emph{reliability} and \emph{functionality} seem swapped between projects.
-For both projects Bayesian techniques did the best out of a wide variety of machine learners tested. 
-Our best learners, Discriminative Multinomial Naive Bayes, Naive Bayes  and Multinomial Naive Bayes  are all based on Bayes's theorem and all assume, naively, that the features are independent. 
-The features we used are word counts per message. 
-One beneficial aspect of this result is that it suggests we can have very fast training and classifying  since Naive Bayes can be calculated in $O(N)$ for $N$ features.
-%http://nlp.stanford.edu/IR-book/html/htmledition/naive-bayes-text-classification-1.html
-
-% # F measure
-% mysql
-% > v <- c( 0.64 , 0.23  ,0.41 , 0.77  ,0.62 , 0.21 ) 
-% > summary(v)
-%    Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
-%   0.210   0.275   0.515   0.480   0.635   0.770 
-% maxdb
-% > v <- c(0.61, 0.25, 0.59, 0.32, 0.42, 0.17)
-% > summary(v)
-%    Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
-%  0.1700  0.2675  0.3700  0.3933  0.5475  0.6100
-
-The range of F-measures for MySQL was $0.21$ to $0.77$ with an average
-of $0.48$, while MaxDB had a range of $0.17$ to $0.61$ with an average
-of $0.39$.
-
-The less-frequently occurring a label, the harder it is to get accurate
-results, due to the high noise level. Nevertheless, these results are
-better than our previous word-list results of \textsf{exp2} and
-\textsf{exp3}, because the ROC values are sufficiently higher in most
-cases (other than MaxDB \emph{reliability} and MySQL \emph{efficiency}). The
-limitation of the approach we took here is that we assume labels are
-independent; however, labels could be correlated with each other. 
-The next section (\ref{sec:multilabel})
-addresses the issue of a lack of independence and correlation between labels.
-In the next section we will evaluate how well these learners perform
-together.
-%\todo[inline]{Comment about correlation, we should link to the Mulan section, perfect segway since Mulan relies on correlation.}
 
@@
 
+For both projects Bayesian techniques did the best out of a wide variety of machine learners tested. 
+Our best learners, Discriminative Multinomial Naive Bayes, Naive Bayes  and Multinomial Naive Bayes  are all based on Bayes's theorem and all assume, naively, that the features are independent. 
+The features we used are word counts per message. 
+One beneficial aspect of this result is that it suggests we can have very fast training and classifying  since Naive Bayes can be calculated in $O(N)$ for $N$ features.
+%http://nlp.stanford.edu/IR-book/html/htmledition/naive-bayes-text-classification-1.html
+
+% # F measure
+% mysql
+% > v <- c( 0.64 , 0.23  ,0.41 , 0.77  ,0.62 , 0.21 ) 
+% > summary(v)
+%    Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
+%   0.210   0.275   0.515   0.480   0.635   0.770 
+% maxdb
+% > v <- c(0.61, 0.25, 0.59, 0.32, 0.42, 0.17)
+% > summary(v)
+%    Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
+%  0.1700  0.2675  0.3700  0.3933  0.5475  0.6100
+
+The range of F-measures for MySQL was $0.21$ to $0.77$ with an average
+of $0.48$, while MaxDB had a range of $0.17$ to $0.61$ with an average
+of $0.39$.
+
+The less-frequently occurring a label, the harder it is to get accurate
+results, due to the high noise level. Nevertheless, these results are
+better than our previous word-list results of \textsf{exp2} and
+\textsf{exp3}, because the ROC values are sufficiently higher in most
+cases (other than MaxDB \emph{reliability} and MySQL \emph{efficiency}). The
+limitation of the approach we took here is that we assume labels are
+independent; however, labels could be correlated with each other. 
+The next section (\ref{sec:multilabel})
+addresses the issue of a lack of independence and correlation between labels.
+In the next section we will evaluate how well these learners perform
+together.
+%\todo[inline]{Comment about correlation, we should link to the Mulan section, perfect segway since Mulan relies on correlation.}
+
 \subsection{Applying multiple labels to topics}
@@
 We must check if the full subset of labels was applied, and then how much of that subset was applied.
-
-Two perspectives to evaluate multi-label learners are with micro or macro measurements (used in Figure \ref{fig:mulan}).
-Macro measurements are aggregated at a class or label level (per class) while micro measurements are at the element level (per element).
-A macro-ROC measurement is the average ROC over the ROC values for all labels, where a micro-ROC is the average ROC over all examples that were classified. 
-For MaxDB, the macro-ROC values are undefined because of poor performance of one of the labels.%, see Figure \ref{fig:mulan}.
-
-Figure~\ref{fig:mulan} presents the results of Mulan's best multi-label learners for our data. 
-Calibrated Label Ranking (CLR) is a learner that builds two layers, the first layer determines if an entity should be labelled, while the second layer determines what labels should be assigned.
-The Hierarchy Of Multi-label classifiERs (HOMER) and Binary Relevance (BR) act as a hierarchy of learners: BR is flat, while HOMER tries to build a deeper hierarchy to build a more accurate learner~\cite{mulan}.
-These classifiers performed better than other multi-label classifiers as they have the best micro and macro ROC scores. 
-The multi-label and single-label learners had similar performance: for MySQL, BR and NaiveBayes had similar macro-ROC scores of $0.74$.
-% although their results are comparable to the naive Bayesian learners we used in Section \ref{sec:suplabelling}.
-% \todo[inline]{discuss this similarity in more detail}
 
@@
 
+Two perspectives to evaluate multi-label learners are with micro or macro measurements (used in Figure \ref{fig:mulan}).
+Macro measurements are aggregated at a class or label level (per class) while micro measurements are at the element level (per element).
+A macro-ROC measurement is the average ROC over the ROC values for all labels, where a micro-ROC is the average ROC over all examples that were classified. 
+For MaxDB, the macro-ROC values are undefined because of poor performance of one of the labels.%, see Figure \ref{fig:mulan}.
+
+Figure~\ref{fig:mulan} presents the results of Mulan's best multi-label learners for our data. 
+Calibrated Label Ranking (CLR) is a learner that builds two layers. The first layer determines if an entity should be labelled, while the second layer determines what labels should be assigned.
+The Hierarchy Of Multi-label classifiERs (HOMER) and Binary Relevance (BR) act as a hierarchy of learners: BR is flat, while HOMER tries to build a deeper hierarchy for a more accurate learner~\cite{mulan}.
+These classifiers performed better than other multi-label classifiers as they have the best micro and macro ROC scores. 
+The multi-label and single-label learners had similar performance: for MySQL, BR and NaiveBayes had similar macro-ROC scores of $0.74$.
+% although their results are comparable to the naive Bayesian learners we used in Section \ref{sec:suplabelling}.
+% \todo[inline]{discuss this similarity in more detail}
+
 \section{Understanding software maintenance activities} 
@@
 %activities. 
-\todo[inline]{more on intro ... why should the reader care about
-  this. It needs to show how the stuff in the intro (applications) is
-  possible}
+%\todo[inline]{more on intro ... why should the reader care about
+ % this. It needs to show how the stuff in the intro (applications) is
+ % possible}
 %%% XXX NEIL I don't get this.. VVVVV
 %\todo[inline]{fix section headings}
-
-In this section we investigate the history of the two projects that we
-studied.
-We demonstrate the value of topic analysis as it is capable of showing
-the topic of development efforts over time in each project.
-We evaluated two research questions using the ground truth data that we
-annotated by hand:
+As we mentioned in the introduction, a key issue in software maintenance is understanding \emph{why} a system has evolved the way it has. 
+In this section we demonstrate the value of labelled topic extraction in addressing this issue. 
+We investigate the history of the two large-scale database systems that we studied. 
+We use our technique to show the topic of development efforts over time in each project.
+%We evaluated two research questions using the ground truth data that we annotated by hand:
+We motivated our investigation with two research questions:
 \begin{enumerate}
-\item Do label frequencies change over time? Is a certain quality of more interest at one point in the life-cycle than some other? 
-\item  Do the different projects differ in their relative topic interest? Is a particular quality more important to one project than the other projects?  
+\item \emph{Do NFR frequencies change over time?} If a particular NFR was of more interest at one point in the life-cycle than another, this suggests that development activity shifted focus.
+For example, if a developer expected to see a recent focus on \emph{reliability}, but instead \emph{usability} dominated, he or she might re-prioritize up-coming work items.
+\item \emph{Do projects differ in their relative interest in NFRs?} A program manager would be interested in knowing whether a particular NFR, such as \emph{reliability}, was more important for one project than another.
+This could be to confirm the initial design goals, or to track the progress on that quarter's objectives.
 \end{enumerate}
-
-% There are several possible applications of \textsf{Nomen}'s techniques. 
-% For outside users of the software, \textsf{Nomen} can provide insight into that company's development practices. 
-% Are they focused on usability, for example?
-% Within the organization \textsf{Nomen} acts as a historical sanity check. 
-% Confirmation bias is the difference between what people think happened and what actually happened. 
-% Some software releases, for instance, are poorly received since they are quite buggy. 
-% \textsf{Nomen} would help explain why this was the case, since it can identify NFRs which are dominating developer attention.
-%\todo[inline]{does this section use supervised learning, and if so what ones?}
 
@@
 Figures \ref{fig:mysql-timeline} and \ref{fig:maxdb-timeline} show the
-temporal patterns of label frequencies based on the manually annotated
-topics, although this visualization can be generated from the
-results of labelled topic extraction.
+temporal patterns of NFR frequencies.
+%NEIL- I think this (below) is just confusing, although maybe important to note.
+% based on the manually annotated
+%topics, although this visualization can be generated from the
+%results of labelled topic extraction.
 There are two measures represented. 
-One, the relative frequency, shown in the grey boxes, represents the number of labels with that NFR in that period, 
+One, the relative frequency, shown in the grey histogram boxes, represents the number of topics with that NFR in that period, 
 relative to the maximum number of topics assigned to that NFR.
-The second, absolute frequency, compares the number of topics labelled with that NFR per period 
+For example, in Figure \ref{fig:mysql-timeline} we see a spike in \emph{portability} and \emph{functionality} frequency in September 2002.
+The second, absolute frequency, is shown using cell intensity compares the number of topics labelled with that NFR per period 
 relative to the maximum number of labelled topics overall, for that project.
-The topmost row in each diagram is reserved for historical events of note. 
-We refer to these occurrences in the discussion which follows.
+For instance, Figure \ref{fig:mysql-timeline} shows that the NFRs \emph{functionality}, \emph{portability} and \emph{maintainability} contain more labelled topics, since these NFRs have been more intensely shaded.
+The topmost row in each diagram lists historical events for that project (such as a release). 
+%We refer to these occurrences in the discussion which follows.
+
+\begin{figure*}[]
+\centering
+\subfloat[MySQL 3.23]{
+\includegraphics[width=\textwidth]{figures/mysql-timeline}
+\label{fig:mysql-timeline}                               
+}
+
+\subfloat[MaxDB 7.500]{                                 
+\includegraphics[width=\textwidth]{figures/maxdb-timeline}
+\label{fig:maxdb-timeline}
+}
+	\caption[]{NFR label per period. Each cell represents a 30-day period. %Numbers represent number of topics in that period labelled with that NFR. 
+	Grid cell intensity (saturation) is mapped to label frequency relative to the largest label value of \emph{all} NFRs.
+Grey histogram bars indicate label frequency relative to that particular NFR's largest label value.
+Dashed vertical lines relate a project milestone (*Key Event*) to our topic windows.
+% \todo[inline]{better discussion}
+}
+\label{fig:timelines}
+\end{figure*}
 
 %\noindent \textbf{Understanding the figures} -- 
-In Figure \ref{fig:mysql-timeline}, we see that the NFRs \emph{functionality}, \emph{portability} and \emph{maintainability} contain more labeled topics, since these NFRs have been more intensely shaded.
-% boxes.
-not sure boxes is important..
+ % boxes. not sure boxes is important..
 % I grepped query1.mrs2 http://softwareprocess.es/y/y/maxdb-76.tar.gz
 % and                   http://softwareprocess.es/y/maxdb-7500.tar.gz
-There are more \textsf{None} labels in MaxDB because a number of topics were to do with code cleanup or automated checking using tools like \texttt{sutcheck}, a tool-specific to the development process of MaxDB.
+%There are more \textsf{None} labels in MaxDB because a number of topics were to do with code cleanup or automated checking using tools like \texttt{sutcheck}, a tool specific to the development process of MaxDB.
 
 %\noindent \textbf{Timeline of key events} -- 
+We analyzed each project's developer mailing list for external validation. 
 As discussed earlier, the top row in each figure shows key events for each project. 
@@
 and when MaxDB 7.500 was first announced. For MaxDB, version 7.5.00  was released in December of 2003. 
-From analyzing the mailing list for external validation, we know that release 7.5.00.23 saw the development of PHP interfaces, possibly accounting for the simultaneous increase in the \emph{portability} NFR at the same time.
+We know that release 7.5.00.23 saw the development of PHP interfaces, possibly accounting for the simultaneous increase in the \emph{portability} NFR at the same time.
 The gap in MaxDB (Figure \ref{fig:maxdb-timeline}) is due to a shift in development focus (from February 2005 to Jun 2005) to MaxDB 7.6, which is released in June 2005.
@@
 
-\begin{figure*}[ht]
-\centering
-\subfloat[MySQL 3.23]{
-\includegraphics[width=\textwidth]{figures/mysql-timeline}
-\label{fig:mysql-timeline}                               
-}
-
-\subfloat[MaxDB 7.500]{                                 
-\includegraphics[width=\textwidth]{figures/maxdb-timeline}
-\label{fig:maxdb-timeline}
-}
-	\caption[]{NFR label per period. Each box represents a 30-day period. %Numbers represent number of topics in that period labelled with that NFR. 
-	Grid cell intensity (saturation) is mapped to label frequency relative to the largest label value of \emph{all} NFRs.
-Grey bars indicate label frequency relative to \emph{a particular} NFR.
-Dashed vertical lines relate a project milestone (*Key Event*) to our topic windows.
-% \todo[inline]{better discussion}
-}
-\label{fig:timelines}
-\end{figure*}
-
 % \subsection{Comparing MaxDB and MySQL}
@@
 %\todo[inline,color=yellow]{expand on this with dates, and reference to visualizations}
-
+\subsection{Research questions}
 %This analysis allows to address the questions raised in Section \ref{sec:suplabelling}:
-This analysis addresses the questions raised in Section \ref{sec:analysis}:
+This analysis addresses our research questions: %raised in Section \ref{sec:analysis}:
 
@@
 
-%\textbf{Do the different projects differ in their relative topic
-%interest?} --
-
-\subsubsection{Do the different projects differ in their relative
-  topic interest?}
-
+\subsubsection{Do  projects differ in their relative  topic interest?}
 Yes. MySQL 3.23 had proportionally more
@@
 
-\subsection{Summary}
-
-\todo[inline]{link the actual data, which is interesting, to somethign
-  actionable for a developer or manager}
+%\subsection{Summary}
+
+%\todo[inline]{link the actual data, which is interesting, to somethign
+ % actionable for a developer or manager}
 
@@
 
-We found that the multi-label learners of BR, CLR and HOMER did not do as well for Macro-ROC as NaiveBayes and other naive Bayes derived learners. 
-This suggests that by combining together multiple NaiveBayes learners we could probably label sets of topics effectively, but it would require a separate NaiveBayes learner per label.
-
-%{Discuss how these are CHEAP techniques and although somewhat %inaccurate they don't require a tonne of work}
 The unsupervised labelling had difficulty distinguishing between common labels and infrequent labels. 
@@
 The word-lists for \emph{correctness} tended to be too lengthy, non-specific and broad, especially if WordNet words were used, since the NFRs are typically loosely defined concepts in common parlance.
+
+We found that the multi-label learners of BR, CLR and HOMER only did as well or worse for Macro-ROC as the single-label NaiveBayes and other naive Bayes-derived learners. 
+This suggests that by combining together multiple NaiveBayes learners we could probably label sets of topics effectively, but it would require a separate NaiveBayes learner per label.
+
+%{Discuss how these are CHEAP techniques and although somewhat %inaccurate they don't require a tonne of work}
 

### 2011-01-26 msr/msr2011.tex
updates to conclusion, various resizing of figures for space reasons.

@@
 \end{tabular}
-	\caption{NFRs and associated signifiers -- \textsf{exp2}}
+	\caption{NFRs and associated word-lists -- \textsf{exp2}}
 	\label{tbl:wnsig}
@@
   \centering
- \includegraphics[width=0.7\textwidth]{figures/unsupervised-bar}
+ \includegraphics[width=0.8\textwidth]{figures/unsupervised-bar}
  \caption{Performance, ROC values (range: 0-1), of unsupervised topic labelling for
@@
 % \subfloat[MySQL]{
-\includegraphics[width=0.4\textwidth]{figures/both-supervised}
+\includegraphics[width=0.45\textwidth]{figures/both-supervised}
 % \label{fig:subfig1}
@@
 
-\subsection{Timeline analysis}
+%\subsection{Timeline analysis}
 Figures \ref{fig:mysql-timeline} and \ref{fig:maxdb-timeline} show the
@@
 For example, in Figure \ref{fig:mysql-timeline} we see a spike in \emph{portability} and \emph{functionality} frequency in September 2002.
-The second, absolute frequency, is shown using cell intensity compares the number of topics labelled with that NFR per period 
-relative to the maximum number of labelled topics overall, for that project.
+The second, absolute frequency, is shown using cell intensity, and compares the number of topics labelled with that NFR per period 
+relative to the maximum number of labelled topics overall.
 For instance, Figure \ref{fig:mysql-timeline} shows that the NFRs \emph{functionality}, \emph{portability} and \emph{maintainability} contain more labelled topics, since these NFRs have been more intensely shaded.
@@
 We analyzed each project's developer mailing list for external validation. 
-As discussed earlier, the top row in each figure shows key events for each project. 
-\textit{Labelled topic extraction} can pick out the underlying NFR activity behind these events. 
+%As discussed earlier, the top row in each figure shows key events for each project. 
+We use \textit{labelled topic extraction} to pick out the underlying NFR activity behind these events. 
 For example, both projects show a high number of NFRs recognized at the first period of analysis. 
@@
 
-For MySQL (Figure \ref{fig:mysql-timeline}), we similarly validated our NFR patterns with external mailing list data. 
-This release of MySQL was the first to be licenced under the GPL. 
+%For MySQL (Figure \ref{fig:mysql-timeline}), we similarly validated our NFR patterns with external mailing list data. 
+The release of MySQL we study  (Figure \ref{fig:mysql-timeline}) was the first to be licenced under the GPL. 
 Version 3.23.31 (January, 2001) was the production release (non-beta), and we see a flurry of topics labelled with \emph{functionality} and \emph{maintainability}. 
@@
 Similarly, in August, 2002, both \emph{functionality} and \emph{portability} are frequent, and mailing list data suggests this is related to the release of version 3.23.52, a general bug fix with a focus on security (a component of the \emph{functionality} NFR in the ISO9126 model).
-After this point, efforts shift to the newer releases (4.0, 4.1, 5.0).
+After this point, efforts shift to the newer releases (4.0, 4.1, 5.0). With respect to our research questions:
 
@@
 %\todo[inline,color=yellow]{expand on this with dates, and reference to visualizations}
-\subsection{Research questions}
+%\subsection{Research questions}
 %This analysis allows to address the questions raised in Section \ref{sec:suplabelling}:
-This analysis addresses our research questions: %raised in Section \ref{sec:analysis}:
-
-%\textbf{Do label frequencies change over time?} -- 
-\subsubsection{Do label frequencies change over time?}
+%This analysis addresses our research questions: %raised in Section \ref{sec:analysis}:
+
+\noindent\textbf{1. Do NFR frequencies change over time?} --- 
+%\subsubsection{Do label frequencies change over time?}
 In both projects the frequencies generally decreased with age. 
@@
 A proportionately smaller number of commits addressed these NFRs.
-As discussed in the previous section, certain peaks in topic numbers coincide with a particular emphasis from the development team on issues such as new releases or bug fixes.
+%As discussed in the previous section, 
+Certain peaks in topic numbers coincide with a particular emphasis from the development team on issues such as new releases or bug fixes.
 This suggests that maintenance activity is not necessarily strictly decreasing with time, but rather episodic and responsive to outside stimuli. 
@@
 
-\subsubsection{Do  projects differ in their relative  topic interest?}
+%\subsubsection{Do  projects differ in their relative  topic interest?}
+\noindent\textbf{2. Do  projects differ in their relative  topic interest?} ---
 Yes. MySQL 3.23 had proportionally more
 topics labelled \emph{functionality}, while MaxDB had proportionally more
-\emph{efficiency} related topics. MaxDB was a very mature release ``donated'' to the open-source community. 
-MySQL, on the other hand, was in its relative infancy. 
-Security problems were more common (security is a component of \emph{functionality} in the ISO 9126 model). 
-In both cases \emph{portability} was a constant maintenance concern and was prevalent throughout the entire lifetime of the projects.
-
+\emph{efficiency} related topics. MaxDB was a very mature release ``donated'' to the open-source community, 
+whereas MySQL was in its relative infancy, and  
+security problems were more common (security is a component of \emph{functionality} in the ISO 9126 model). 
+In both cases \emph{portability} was a constant maintenance concern and was prevalent throughout the lifetime of the projects. It may surprise developers how often portability arises as a concern.
+%On one hand, this analysis is unsurprising; on the other hand, it provides confirmation that our approach identifies actual patterns.
 %\subsection{Summary}
@@
 \emph{External validity} -- %was addressed by our use of two case studies, but there are still issues. 
-Our data originated from OSS database projects and thus might not be applicable to commercially developed software. 
-Furthermore, our analysis techniques rely on a project's use of meaningful commit messages. The OSS projects investigated were database systems, thus 
-our results might not generalize to other domains. %We only studied database systems.
+Our data originated from OSS database projects and thus might not be applicable to commercially developed software or other domains. 
+Furthermore, our analysis techniques rely on a project's use of meaningful commit messages, although we feel this is the most frequently occurring form of developer comments. 
+%The OSS projects investigated were database systems, thus 
+%our results might not generalize to other domains. %We only studied database systems.
 % \todo[inline]{If we compare with the same domain we don't get to say it generalizes a lot?}
@@
 This paper presented a cross-project data mining technique,
-\emph{labelled topic extraction}. The technique leveraged a
-non-functional requirements (NFRs) taxonomy to mine and label latent
-software development topics. Previous topic analysis research produced
-topics that needed to be manually labelled, but topics without labels
-are difficult to interpret, and manually labelling them is
-costly. Furthermore, previous work was project specific and incomparable. 
-To improve this, we leveraged software engineering standards, specifically the ISO quality taxonomy, to produce a method of supervised and unsupervised topic labelling.
-Since the word-list technique is not project specific, it can be used to compare multiple projects.%topics and artifacts from project histories.
+\emph{labelled topic extraction}. %The technique leveraged a
+%non-functional requirements (NFRs) taxonomy to mine and label latent
+%software development topics. 
+Previous topic analysis research produced project-specific
+topics that needed to be manually labelled.
+%, but topics without labels
+%are difficult to interpret, and manually labelling them is
+%costly. Furthermore, previous work was project-specific and thus incomparable. 
+To improve on this, we leveraged software engineering standards, specifically the ISO quality taxonomy, to produce a method of partially-automated (supervised) and fully-automated (unsupervised) topic labelling.
+Since the word-list technique is not project-specific, we used it to compare two distinct projects, where we showed our technique produced interesting insight into maintenance activity.
+%topics and artifacts from project histories.
 % Our contributions include:
@@
 
-We validated our topic labelling techniques using multiple experiments. 
+We validated our topic labelling techniques using multiple experiments. % on unsupervised learners, supervised and multi-label learners. 
 We first conducted unsupervised labelling using word-lists. 
@@
 %area under the 
-ROC values between $0.6$ and $0.8$. These results, along with the efficient performance of our learners, demonstrates that labelled topic extraction is a promising approach for understanding
-
-These labelling techniques allowed us to investigate the occurrence of non-functional requirements in our projects. 
+ROC values between $0.6$ and $0.8$.
+These results, along with the efficient performance of our learners, demonstrate that labelled topic extraction is a promising approach for understanding the occurrence of non-functional requirements in software projects.
 %Although it may seem like common knowledge that maintenance is a concern as a project matures, our technique was able to show this independently.
-We showed that NFRs are often trending topics in commit comments
+%We showed that NFRs are often trending topics in commit comments
 %, and that NFRs are quite common as developer topics.

### 2011-01-26 msr/msr2011.tex
Tiny edits

--- 2011-01-26 msr/msr2011.tex
@@
 \begin{itemize}
-\item introduces the concept of labelled topic extraction, using a non-functional requirements (NFR) taxonomy for our labels; 
-\item evaluates an unsupervised method of automated topic labelling (word-lists);
+\item we introduce the concept of labelled topic extraction, using a non-functional requirements (NFR) taxonomy for our labels; 
+\item we evaluate an unsupervised method of automated topic labelling (word-lists);
 %\item Our unsupervised method can be applied to projects without training;
-\item evaluates a supervised method of automatically labelling topics with a single NFR (machine learning);
-\item evaluates a supervised method of labelling topics with multiple NFRs (multi-label machine learning);
-\item provides a method of cross-project analysis via topic labelling; and
-\item applies these techniques to visualize NFRs over time, and to analyze maintenance activities.
+\item we evaluate a supervised method of automatically labelling topics with a single NFR (machine learning);
+\item we evaluate a supervised method of labelling topics with multiple NFRs (multi-label machine learning);
+\item we provide a method of cross-project analysis via topic labelling; and
+\item we apply these techniques to visualize NFRs over time, and to analyze maintenance activities.
 \end{itemize}
@@
 
-\begin{figure*}[]
+\begin{figure*}[t]
 \centering

### 2011-03-24 msr/msr2011.tex
finalish conclusion edit
Can we apply to bug reports? Why versus how?     - [X] The question that needs to be better answered is what           application this technique has to commit messages?           - If it is only workable on commit messages, is it limited           to providing coarse comparisons of the past history of           projects?           - response: answered in discussion section. No it isn't             limited but we didn't test it.     - [X] - Are the commit messages close enough to bug report titles           that the approach might work on bug reports as well?

--- 2011-01-26 msr/msr2011.tex
@@
 %migod
-A key problem for practising software maintainers is gaining an understanding of \emph{why} a system has evolved the way it has.
-Looking back on streams of artifacts scattered across different repositories, inferring what activities were performed, when, and for what reasons, is hard to do without expert advice from the developers involved.
+A key problem for practising software maintainers is gaining an
+understanding of \emph{why} a system has evolved the way it has. 
+Looking back on streams of artifacts scattered across different
+repositories, inferring what activities were performed, when, and for
+what reasons, is hard to do without expert advice from the developers
+involved. 
+%new stuff
+% edit this. awkward
+This is different than \emph{how}, because the behaviour of change itself is the
+how, not the \emph{why}
 In this work we seek to provide automated help in labelling development topics that have been extracted from commit logs.
@@
 As we mentioned in the introduction, a key issue in software maintenance is understanding \emph{why} a system has evolved the way it has. 
-In this section we demonstrate the value of labelled topic extraction in addressing this issue. 
+In this section we demonstrate the value of labelled topic extraction
+in addressing this issue. 
+% more why
+Labelled topics address \emph{why}  because they show why changes
+occurred rather than \emph{how}. 
+The \emph{how} of a change is the change itself, the purpose or
+\emph{why} is what we are after.
 We investigate the history of the two large-scale database systems that we studied. 
@@
 %These results are promising, because the accuracy of the techniques is good enough to be useful, while the run-time performance is still cheap enough to execute to be feasible as an automated or semi-automated method of labelling topics by their software qualities.
-These results are promising because they indicate that these techniques are accurate enough to be useful while still maintaining acceptable run-time performance.
+These results are promising because they indicate that these
+techniques are accurate enough to be useful while still maintaining
+acceptable run-time performance.
+
+Since this work focuses on labelling natural language commit log
+comments we feel it can adapted to other natural language artifacts
+such as mailing-list discussions and bug reports, even though that was
+not evaluated in this paper. Bug reports might not exhibit the same
+behaviour as commits in terms of dominant topics.
 

### 2011-03-24 msr/msr2011.tex
re-laid-out paragraph (no content change)

@@
 %migod
-While machine learning techniques can automatically identify clumps of commonly recurring terms, devising an appropriate summary label for each clump/topic is harder.
-For example, the word list above might reasonably be labelled as \emph{"event handling"} by a human who understands the domain well, despite the fact that label does not appear in the word list itself.
-Current approaches to topic labelling rely on manual intervention by human experts, and also are limited to project-specific topic labels.
+While machine learning techniques can automatically identify clumps of
+commonly recurring terms, devising an appropriate summary label for
+each clump/topic is harder.  
+For example, the word list above might reasonably be labelled as
+\emph{"event handling"} by a human who understands the domain well,
+despite the fact that label does not appear in the word list itself.  
+Current approaches to topic labelling rely on manual intervention by
+human experts, and also are limited to project-specific topic labels.  
 In this paper, we introduce \emph{labelled topic extraction}, an approach to topic labelling that is both automatic and creates labels that are project independent.

### 2011-03-24 msr/msr2011.tex
Moved example and addressed log messages     - [X] The last sentence of the second paragraph of the second           column of the first page: Seems the example does not belong           here. Based on the sentence before the reader would expect           an example of how topic models help developers to get a           sense of project history.           - response: example has been moved to avoid confusion     - [X] Commit messages were used to extract the topics. It has not           discussed whether all commits have a log message (at least           for the two system which were studied).

@@
 project history. 
-For example, a topic extracted from a set of commit logs might consist of the following terms: \emph{ ``listener change remove add fire''}. 
 %A major challenge for topic modeling is to accurately identify the
@@
 each clump/topic is harder.  
-For example, the word list above might reasonably be labelled as
-\emph{"event handling"} by a human who understands the domain well,
+A given topic extracted from a set of commit logs might consist of the following terms: \emph{ ``listener change remove add fire''}. 
+This topic might reasonably be labelled as
+``\emph{event handling}'' by a human who understands the domain well,
 despite the fact that label does not appear in the word list itself.  
@@
 repository. 
+Most commits we observed had commit comments.
 Commit comments are often studied by researchers, as they are the most readily accessible source of project interactions, and developers are often required to create them by the repository mechanism (e.g., Git).

### 2011-03-24 msr/msr2011.tex
Inter-rater and annotation details - fixed some threats to validity
    - [X] It would also be useful to provide inter-rater reliability numbers for the
          commit messages annotated. What did you use as the final labels for the supervised
          learning? The union of the annotators? The intersection?
          - response: added in the reliability section of threats to validity
          - response: added more detail to "creating a validation
            corpus"
    - [X] In Threats to Validity, you state you required a "seamless record". What do you
          mean by this phrase?

@@
 created a validation set of manually labelled topics. For MySQL 3.23
-and MaxDB 7.500, we annotated each extracted topic in each period with
+and MaxDB 7.500, we (the first two authors) annotated each extracted topic in each period with
 the six NFR labels listed above.
+Annotators did not annotate each other's annotations but some brief
+inspection of annotations was used to confirm that the annotators were
+acting similarly.
 We looked at each period's topics, and assessed what the data --- consisting of the frequency-weighted word lists and messages --- suggested was the topic for that period. 
 We were able to pinpoint the appropriate label using auxiliary information as well, such as the actual revisions and files that were related to the topic being annotated.
-For example, for the MaxDB topic consisting of a message ``exit() only used in non NPTL LINUX Versions'', we tagged that topic \emph{portability}. 
+For example, for the MaxDB topic consisting of a message ``exit() only
+used in non NPTL LINUX Versions'', we tagged that topic
+\emph{portability}. 
+Given the top-level annotations of \emph{none}, \emph{portability},
+\emph{efficiency}, \emph{reliability}, \emph{functionality},
+\emph{usability}, and \emph{maintainability}, the annotators annotated each topic
+with the relevant label and sometimes we used finer-grained
+annotations that would be aggregated up to one of these higher-level labels.
 %\todo[inline]{this confused people. Is it truly unsupervised? Yes, but it needs to be clearer.}
@@
 %Our work faced multiple threats to validity and we attempted to address many of them.
-\emph{Construct validity} -- we used only commit messages rather than mail or bug tracker messages.
-Since we required a seamless record, it was not possible to obtain other data sources.
+\emph{Construct validity} -- we used only commit messages rather than
+mail or bug tracker messages. 
+To extend further we would need matching repositories for each project.
+%Each case study was BecauseSince we required a seamless record, it was not possible to obtain
+%other data sources. 
 Possibly they would have influenced our results, but there would be a degree of correlation between the corpora.
@@
 % in Case Study Research: Design Case Studies
-\emph{Reliability} -- each annotator followed the same protocol and used the same annotations. 
-However, only two annotators were used; their annotations could be biased as we did not analyze for inter-rater reliability.
+\emph{Reliability} -- each annotator, the first two authors, followed the same protocol and used the same annotations. 
+However, only two annotators were used; their annotations could be
+biased as we did not analyze for inter-rater reliability because they
+did not rate the same documents.
 

### 2011-03-24 msr/msr2011.tex
layout a paragraph

@@
 %migod
-In general, the fruits of mining software artifacts are often project specific and hard to generalize.
-However, in our previous work we investigated \emph{topic trends} --- that is, topics that recur over time --- we observed that topic trends often corresponded to non-functional requirements (NFRs) \cite{Hindle09ICSM}.
-This is encouraging, as NFRs have the property of being cross-domain and widely applicable.
-In this sense, they are useful abstractions for developer conversations about different software projects.
+In general, the fruits of mining software artifacts are often project
+specific and hard to generalize.  
+However, in our previous work we
+investigated \emph{topic trends} --- that is, topics that recur over
+time --- we observed that topic trends often corresponded to
+non-functional requirements (NFRs)~\cite{Hindle09ICSM}.  
+This is encouraging, as NFRs have the property of being cross-domain
+and widely applicable. 
+In this sense, they are useful abstractions for developer
+conversations about different software projects.  
 Furthermore, there is a series of standards on NFRs, such as ISO9126 \cite{iso9126}, that are specifically intended to apply to projects of varying types; this suggests that our goal of trying to extract NFR-related development topics, such as those related to software quality models, holds promise.

### 2011-03-24 msr/msr2011.tex
- [X] 6: "trending topics within commits are often related to NFRs"      seems unsupported -- define "often". There is no comparison with      other possible labels, and little discussion of the percentage of      commits that are assigned topics versus how many should be      assigned to topics.      - addressed in introduction

@@
 time --- we observed that topic trends often corresponded to
-non-functional requirements (NFRs)~\cite{Hindle09ICSM}.  
+non-functional requirements (NFRs)~\cite{Hindle09ICSM}, which is
+further emphasized in this paper due to the large numbers of NFR
+labelled topics.  
 This is encouraging, as NFRs have the property of being cross-domain

### 2011-03-24 msr/msr2011.tex
- [X] The authors need to further demonstrate the utility of       labelled topics in practice. Is understanding why a system       evolved a key problem? If so, it should be justified with a       citation or numbers.       - citing the secret life of bugs in this case         where Dr. Aranda found that finding out why was really hard :)
Discuss features, fix wordnet stuff, respond to example criticism
    - [X] In supervised labeling (3.3) it has not been discussed what
          features of commit messages has been used in
          training/testing learners.
          - response: in super

@@
 Similarly, Mockus and Votta~\cite{Mockus00} studied a large-scale industrial change-tracking system. 
-They also leveraged WordNet, but only for word roots as they felt the synonyms would be non-specific and cause errors, problems we also encountered. 
+Mockus and Votta leveraged WordNet~\cite{Fellbaum1998}, an English-language ``lexical database'' that contains semantic relations between words, including common related forms (similar to word stemming), meronymy and synonymy.
+They only used WordNet for word roots as they felt the synonyms would be non-specific and cause errors, problems we also encountered. 
 Mockus et al. had access to system developers, with whom they could validate their labels. 
@@
 % in our previous work~\cite{Hindle09ICSM}, as well as -- CITED a little too often (vanity cites!)
-In Baldi et al.~\cite{Baldi2008}, topics are named manually: human experts read the highest-frequency members of a topic and assign a label accordingly. 
+In Baldi et al.~\cite{Baldi2008}, topics are named manually: human
+experts read the highest-frequency members of a topic and assign a
+label accordingly. 
+%XXX Possible to cut this
 As discussed earlier, given the topic \emph{``listener change remove add fire''}, Baldi et al. would assign the label \emph{event-handling}. 
@@
 %\todo[inline]{This part doesn't make a lot of sense}
-To construct these expanded word-lists, we used WordNet~\cite{Fellbaum1998}, an English-language ``lexical database'' that contains semantic relations between words, including common related forms (similar to word stemming), meronymy and synonymy.
+To construct these expanded word-lists, we used
+WordNet~\cite{Fellbaum1998}.
+%, an English-language ``lexical database'' that contains semantic relations between words, including common related forms (similar to word stemming), meronymy and synonymy. 
 We then added Boehm's software quality model \cite{Boehm+:1976:ICSE}, and classified his eleven `ilities' into their respective ISO9126 NFRs. 
@@
 Traditional classifiers map our topics to a single class, whereas Mulan allows for a mixture of classes per topic, which maps to what we observed while manually labelling topics.
+The \emph{features} we used are word counts/occurrence  per topic, if
+a word occurs frequently enough in a topic we consider it a feature of
+the topic.
 
@@
 Our best learners, Discriminative Multinomial Naive Bayes, Naive Bayes  and Multinomial Naive Bayes  are all based on Bayes's theorem and all assume, naively, that the features are independent. 
-The features we used are word counts per message. 
+%The features we used are word counts per message. 
 One beneficial aspect of this result is that it suggests we can have very fast training and classifying  since Naive Bayes can be calculated in $O(N)$ for $N$ features.
@@
 %\todo[inline]{fix section headings}
-As we mentioned in the introduction, a key issue in software maintenance is understanding \emph{why} a system has evolved the way it has. 
+As we mentioned in the introduction, a key issue in software maintenance is understanding \emph{why} a system has evolved the way it has~\cite{aranda09icse}. 
 In this section we demonstrate the value of labelled topic extraction

### 2011-03-24 msr/msr2011.tex
Discuss features, fix wordnet stuff, respond to example criticism
    - [X] In supervised labeling (3.3) it has not been discussed what
          features of commit messages has been used in
          training/testing learners.
          - response: in supervised labelling we described the word count
            approach used.
    - [X] The third paragraph of the second page: "They also leveraged
          WordNet ..." This is the first time "WordNet" is mentioned
          without having been introduced before.
          - reponse: we moved the refs around in previous work
    - [X] The parag

@@
 \item introduces the concept of labelled topic extraction, using a non-functional requirements (NFR) taxonomy for our labels;
-\item evaluates an unsupervised method of automated topic labelling (word-lists);
+
+%\item evaluates an unsupervised method of automated topic labelling (word-lists);
+%%\item Our unsupervised method can be applied to projects without training;
+%%\item 
+%evaluates a supervised method of automatically labelling topics with a single NFR (machine learning);
+%%\item 
+%evaluates a supervised method of labelling topics with multiple NFRs
+%(multi-label machine learning);
+
+\item evaluates three kinds of automatic topic labelling methods:
+  semi-unsupervised labelling (word-lists) of topics, supervised labelling of
+  topics with a single NFR (machine learning), 
+  and supervised labelling of topics with multiple NFRs (multi-label
+  machine learning)
+
+%  of automated topic labelling (word-lists);
 %\item Our unsupervised method can be applied to projects without training;
-\item evaluates a supervised method of automatically labelling topics with a single NFR (machine learning);
-\item evaluates a supervised method of labelling topics with multiple NFRs (multi-label machine learning);
+%\item 
+%evaluates a supervised method of automatically labelling topics with a single NFR (machine learning);
+%\item 
+%evaluates a supervised method of labelling topics with multiple NFRs (multi-label machine learning);
+
 \item provides a method of cross-project analysis via topic labelling; and
-\item applies these techniques to visualize NFRs over time, and to analyze maintenance activities.
+ applies these techniques to visualize NFRs over time, and to analyze maintenance activities.
 \end{itemize}

### 2011-03-24 msr/msr2011.tex
"our work" & mei Different & our own work difference     - [X] "our work" -- this work or [10]? past tense confusing           - response: dealt with it     - [X] It seems like the authors are building off of Mei et al's work       by adding domain specific information. If this is not the case,       authors should clarify.           - response: clarification added     - [X] Authors need to better differentiate the current work from           previous research ([6], [10])           - response: addressed in previous work

@@
 
-Cleland-Huang and her colleagues published work on mining requirements documents for non-functional requirements (NFR) (quality requirements)~\cite{Cleland-Huang2006}. 
-Their approach was similar to ours, as they mined keywords from NFR catalogues.
+Cleland-Huang and her colleagues have research on mining requirements documents for non-functional requirements (NFR) (quality requirements)~\cite{Cleland-Huang2006}. 
+Their approach is similar to ours, as they mined keywords from NFR catalogues.
 %One approach they tried was similar to this one, with keywords mined from NFR catalogues found in their previous work~\cite{chung99}. 
-They managed a recall of 80\% with precision of 57\% for the Security NFR, but could not find a reliable source of keywords for other NFRs. 
+They demonstrated a recall of $80\%$ with precision of $57\%$ for the \emph{security} NFR, but could not find a reliable source of keywords for other NFRs. 
 Instead, they developed a supervised classifier by using human experts to identify an NFR training set. 
-Our work differed because we had a more comprehensive set of terms based on the taxonomy we chose.
+Our research differs because we use a more comprehensive set of terms based on the taxonomy we chose.
 We make cross-project comparisons, while they focused on a single project.
 %The objective of Cleland-Huang's study was to identify new NFRs for system development, yet our objective was to recover those latent NFRs from commit-log messages of the project. 
-Another difference was that our data was far less structured than the requirements documents used in their study.
+Another difference is that our data was far less structured than the requirements documents used in their study.
 %There are several reasons we did not follow this route. 
@@
 They only used WordNet for word roots as they felt the synonyms would be non-specific and cause errors, problems we also encountered. 
-Mockus et al. had access to system developers, with whom they could validate their labels. 
-Since we try to bridge different organizations, instead of a single project, these kind of interviews are infeasible (particularly in the distributed world of open-source software).
-
-The other approach is to extract concerns from software repositories.
-Marcus et al.~\cite{marcus04wcre} describe their use of Latent Semantic Indexing (LSI) to identify commonly occurring concerns for software maintenance.
+Mockus et al. validated their labels with system developers.
+Since we study multiple projects, instead of a single project, these
+kind of interviews are somewhat infeasible (particularly in the distributed world of open-source software).
+
+Another approach is to extract concerns from software repositories.
+Marcus et al.~\cite{marcus04wcre} use Latent Semantic Indexing (LSI) to identify commonly occurring concerns for software maintenance.
 The concerns are given by the user, and LSI used to retrieve them from a corpus.
@@
 %Some results were interesting, but their precision was quite low. 
-With ConcernLines, \cite{treude09cl} shows tag occurrence using colour intensity. 
-They mined 
+With ConcernLines, Treude et al.~\cite{treude09cl} show tag occurrence using colour intensity. 
+They mine
 developer created tags 
@@
 The labels are reasonable enough, but still require an expert in the field to determine them. 
-Furthermore, these labels are project-specific, because they are generated from the data of that project. E.g., we might have a label called `Oracle' in the MySQL case, since Oracle owns MySQL. 
+Furthermore, these labels are project-specific, because they are generated from the data of that project.
+E.g., we might have a label called \emph{`Oracle'} in the MySQL case, since Oracle owns MySQL.
 We do this differently.
@@
 They describe probabilistic labelling, using the frequency distribution of words in a topic to create a meaningful phrase. 
-They do not use external domain-specific information as we do.
-
-Finally, in \cite{ernst10refsq}, we describe our earlier project, similar to this, to identify change in quality requirements in GNOME software projects; our approach is solely text-matching, however, and does not leverage machine learning strategies.
+They do not use external domain-specific information as we do and we
+do not generate phrases from the topics.
+
+Finally, in Ernst et al.~\cite{ernst10refsq}, we describe our earlier
+project, similar to this, that identifies changes in quality requirements
+in GNOME software projects; this approach was more exploratory, had
+less validation, uses different word-lists, solely uses text-matching,
+and does not leverage machine learning strategies. 
+This research extends Hindle et al.~\cite{Hindle09ICSM} with labelled
+topics, NFRs and new visualizations.
 

### 2011-03-24 msr/msr2011.tex
justification for validity of eval data     - [X] the authors need to provide more justification for the      validity of the evaluation data, preferably include annotators      outside the authors, and evaluate agreement. What justification      do the authors have that the data is accurate?           - response: the annotators have much experience in software development             and did inspect some of the labels that the other             annotators were producing. Unfortunately we lack the data             to run inter-rater reliability but this was already             pointed out in t

@@
 topics that relate to software development artifacts.  Typically, these
-techniques use unsupervised machine-learning algorithms to suggest
+techniques use semi-unsupervised machine-learning algorithms to suggest
 candidate word-lists.  However, word-lists are
@@
 Next, we describe how we generated our data (Section \ref{sec:wordlist}).
-For unsupervised classification (Section \ref{sec:unsuplabelling}), we begin by creating word-lists to signify when a topic matches an NFR label.
+For semi-unsupervised classification (Section \ref{sec:unsuplabelling}), we begin by creating word-lists to signify when a topic matches an NFR label.
 We then apply our classifier and analyze the results.
@@
 Figure \ref{fig:process} gives an outline of methodology.
-We began by gathering source data and creating topic models. For unsupervised labelling, we generated three sets of word-lists as signifiers for NFRs.
+We began by gathering source data and creating topic models. For semi-unsupervised labelling, we generated three sets of word-lists as signifiers for NFRs.
 %e created labels using a well-known NFR taxonomy. 
@@
 \subsubsection{Creating a validation corpus}
-To evaluate both unsupervised and supervised classification, we
+To evaluate both semi-unsupervised and supervised classification, we
 created a validation set of manually labelled topics. For MySQL 3.23
@@
 
-\subsection{Unsupervised labelling}
+\subsection{Semi-unsupervised labelling}
 \label{sec:unsuplabelling}
@@
 % Is this good enough
-This labelling is unsupervised because the corpus is not derived from 
+This labelling is \emph{semi-unsupervised} because the corpus is not derived from 
 the project being labelled, and we did not label the project's topics
@@
 
-\subsubsection{Analysis of the unsupervised labelling} %--- %Based
+\subsubsection{Analysis of the semi-unsupervised labelling} %--- %Based
                                 %on the labels, and our manual topic
@@
 %we compared the results of the unsupervised word matching approach. 
-For each quality we assessed whether unsupervised labels matched the manual annotations. %To be clear, the manual annotations were not used to train the labeling process.
+For each quality we assessed whether semi-unsupervised labels matched the manual annotations. %To be clear, the manual annotations were not used to train the labeling process.
 As described in \ref{sec:wordlist} we used both ROC and F-1 measures to evaluate the performance of the classification.
@@
  \includegraphics[width=0.8\textwidth]{figures/unsupervised-bar}
- \caption{Performance, ROC values (range: 0-1), of unsupervised topic labelling for
+ \caption{Performance, ROC values (range: 0-1), of semi-unsupervised topic labelling for
    each NFR and per word-list. Dashed line indicates the performance of a random classifier. This graph shows how well the
-   unsupervised topic labelling matched our manual annotations.}
+   semi-unsupervised topic labelling matched our manual annotations.}
 
@@
 
-Very rarely did \textsf{exp2} and \textsf{exp3} (unsupervised word matching) ever perform as well as the supervised machine learners. 
+Very rarely did \textsf{exp2} and \textsf{exp3} (semi-unsupervised word matching) ever perform as well as the supervised machine learners. 
 For MaxDB, \textit{reliability} was slightly better detected using the static word list of \textsf{exp2}. 
@@
 
-The unsupervised labelling had difficulty distinguishing between common labels and infrequent labels. 
+The semi-unsupervised labelling had difficulty distinguishing between common labels and infrequent labels. 
 The learners would occasionally mislabel a topic deserving of an infrequent label with a more common label.
@@
 
-If the techniques used in \textsf{exp2} and \textsf{exp3} were combined with the supervised techniques, we could reduce the training effort by boosting training sets with topics classified with the unsupervised techniques.
+If the techniques used in \textsf{exp2} and \textsf{exp3} were combined with the supervised techniques, we could reduce the training effort by boosting training sets with topics classified with the semi-unsupervised techniques.
 Both Naive Bayesian learners and the word-list approaches were computationally efficient.  
@@
 We improved internal validity by trying to correlate and explain the behaviours observed in the analysis with the historical records of the projects.
-We did not attempt to match our results to any particular model or theory.
+We did not attempt to match our results to any particular model or
+theory.
+Also we did not have enough manpower to assess inter-rater reliability.
 
@@
 %costly. Furthermore, previous work was project-specific and thus incomparable. 
-To improve on this, we leveraged software engineering standards, specifically the ISO quality taxonomy, to produce a method of partially-automated (supervised) and fully-automated (unsupervised) topic labelling.
+To improve on this, we leveraged software engineering standards, specifically the ISO quality taxonomy, to produce a method of partially-automated (supervised) and fully-automated (semi-unsupervised) topic labelling.
 Since the word-list technique is not project-specific, we used it to compare two distinct projects, where we showed our technique produced interesting insight into maintenance activity.
@@
 We validated our topic labelling techniques using multiple experiments. % on unsupervised learners, supervised and multi-label learners. 
-We first conducted unsupervised labelling using word-lists. 
+We first conducted semi-unsupervised labelling using word-lists. 
 Our next approach was supervised, using single-label and multi-label learners. 

### 2011-03-24 msr/msr2011.tex
- [X] "This would enable querying of bug reports and software       artifacts" how would developers/managers benefit from this?           - response: we cleared up the text a bit but the gist is             that programmers and developers can validate behaviour and             effort relevant to NFRs.
- [X] Why does "best performing" = "most topics"? Will bigger      topics help a programmer/manager more in practice? Isn't accuracy      of the topics a concern here, rather than size?           - response: we agree. It has been dealt with in the text

@@
 This would enable querying of bug reports and artifacts by relevant NFRs. 
-For instance one could look
-for \emph{usability}-relevant revisions and bug reports.
+For instance a manager could confirm if their developers were
+focusing on \emph{usability} by looking for
+ \emph{usability}-relevant revisions and bug reports.
 
@@
 
-For \textsf{exp1}, our best performing 
-labels --- that is, those with the most topics --- were
+% was best performing
+For \textsf{exp1} the labels with the most topics
+were
 %labels, those with the most topics, 
@@
 We did not get good results for \emph{usability} or \emph{accuracy}, which were associated with fewer than ten topics. 
-We also looked for correlation between our labels: Excluding double matches (self-correlation), our highest co-occurring terms were \emph{verifiability} with \emph{traceability}, and \emph{testability} with \emph{correctness} (76 and 62 matches, respectively).
+We also looked for correlation between our labels: excluding double matches (self-correlation), our highest co-occurring terms were \emph{verifiability} with \emph{traceability}, and \emph{testability} with \emph{correctness} (76 and 62 matches, respectively).
 % See the following sections for our error analysis.

### 2011-03-24 msr/msr2011.tex
lay out paragraph

@@
 relative to the maximum number of labelled topics overall. 
-For instance, Figure \ref{fig:mysql-timeline} shows that the NFRs \emph{functionality}, \emph{portability} and \emph{maintainability} contain more labelled topics, since these NFRs have been more intensely shaded.
+For instance, Figure \ref{fig:mysql-timeline} shows that the NFRs \emph{functionality}, \emph{portability} and \emph{maintainability} contain more labelled topics, since these NFRs have been more intensely shaded
 The topmost row in each diagram lists historical events for that project (such as a release). 

### 2011-03-24 msr/msr2011.tex
Neil audit this one please:     - [X] 4: "Fig 5a shows func. port. and maint. contain more       labelled topics" -- what are the practical implications?           - reponse: we suggested the practical implications of the opposite,             why is effeciency periodic?             addressed in text.

--- 2011-03-24 msr/msr2011.tex
@@
 relative to the maximum number of labelled topics overall. 
-For instance, Figure \ref{fig:mysql-timeline} shows that the NFRs \emph{functionality}, \emph{portability} and \emph{maintainability} contain more labelled topics, since these NFRs have been more intensely shaded.
+For instance, Figure \ref{fig:mysql-timeline} shows that the NFRs
+\emph{functionality}, \emph{portability} and \emph{maintainability}
+contain more labelled topics, since these NFRs have been more
+intensely shaded; 
+one interesting stream is \emph{efficiency} which shows periodic
+activity, does this suggest that once maybe efficiency related changes
+have longer lasting effects?
 The topmost row in each diagram lists historical events for that project (such as a release). 

### 2011-03-24 msr/msr2011.tex
- [X] please provide a cite for project dashboards     - [X] Submit copyright

@@
 NFR-related tags.
-Project dashboard are typically employed by managers and are used to
+Project dashboard~\cite{dashboard} are typically employed by managers and are used to
 provide quick summaries of the effort put into a software project. In this case, 
@@
 For example, if a developer expected to see a recent focus on \emph{reliability}, but instead \emph{usability} dominated, he or she might re-prioritize up-coming work items.
-\item \emph{Do projects differ in their relative interest in NFRs?} A program manager would be interested in knowing whether a particular NFR, such as \emph{reliability}, was more important for one project than another.
+\item \emph{Do projects differ in their relative interest in NFRs?} A
+  project manager, especially a systems-manager, would be interested in knowing whether a particular NFR, such as \emph{reliability}, was more important for one project than another.
 This could be to confirm the initial design goals, or to track the progress on that quarter's objectives.

### 2011-03-24 msr/msr2011.tex
- [X] 4RQ2: What purpose does knowing how projects differ in       topics serve? Might be of use for a high-level view of the       system for managers, but authors need to justify           - response: we have addressed this in the text
Previous work, present term and terms     - [X] Previous work is written as if it should come at the end of       the paper. Too much terminology is used that has not yet been       introduced.           - response: we have addressed this issue and added             definitions to the undefined and clarified what our work             was.     - [X] should use

@@
 
-The idea of extracting higher-level \emph{concerns}, also known as \emph{concepts}, \emph{aspects} or \emph{requirements}, has been approached from documentation-based and repository-based perspectives.
-
-Cleland-Huang and her colleagues have research on mining requirements documents for non-functional requirements (NFR) (quality requirements)~\cite{Cleland-Huang2006}. 
+The idea of extracting higher-level \emph{concerns}, also known as
+\emph{topics}, \emph{concepts}, \emph{aspects} or \emph{requirements},
+has been approached from documentation-based and repository-based
+perspectives.
+
+Cleland-Huang and her colleagues have investigated on mining requirements documents for non-functional requirements (NFR) (quality requirements)~\cite{Cleland-Huang2006}. 
 Their approach is similar to ours, as they mined keywords from NFR catalogues.
@@
 Our research differs because we use a more comprehensive set of terms based on the taxonomy we chose.
-We make cross-project comparisons, while they focused on a single project.
+In this paper we make cross-project comparisons instead of focusing
+on a single project.
+They relied on requirements documents instead of version control
+histories that we use.
 %The objective of Cleland-Huang's study was to identify new NFRs for system development, yet our objective was to recover those latent NFRs from commit-log messages of the project.
-Another difference is that our data was far less structured than the requirements documents used in their study.
+
 %There are several reasons we did not follow this route. 
@@
 Mockus and Votta leveraged WordNet~\cite{Fellbaum1998}, an English-language ``lexical database'' that contains semantic relations between words, including common related forms (similar to word stemming), meronymy and synonymy.
-They only used WordNet for word roots as they felt the synonyms would be non-specific and cause errors, problems we also encountered. 
+They used WordNet for word roots as they felt the synonyms would be non-specific and cause errors, problems we also encountered. 
 Mockus et al. validated their labels with system developers.
@@
 and does not leverage machine learning strategies. 
-This research extends Hindle et al.~\cite{Hindle09ICSM} with labelled
+Hindle et al.~\cite{Hindle09ICSM} propose a windowed method of topic
+analysis, we extend this with labelled
 topics, NFRs and new visualizations.
@@
 \item \emph{Do projects differ in their relative interest in NFRs?} A
-  project manager, especially a systems-manager, would be interested in knowing whether a particular NFR, such as \emph{reliability}, was more important for one project than another.
-This could be to confirm the initial design goals, or to track the progress on that quarter's objectives.
+  project manager, especially a systems-manager, would be interested
+  in knowing whether a particular NFR, such as \emph{reliability}, was
+  more important for one project than another. This could be to
+  confirm the initial design goals, or to track the progress on that
+  quarter's objectives.  The difference in NFR proportion is
+  interesting because it implies a difference in focus between two projects.
 \end{enumerate}

### 2011-03-24 msr/msr2011.tex
hahah 10 pages :O booyah! now just need to make it reads well!

@@
 This is different than \emph{how}, because the behaviour of change itself is the
-how, not the \emph{why}
-In this work we seek to provide automated help in labelling development topics that have been extracted from commit logs.
+\emph{how}.
+In this work we seek to provide automated in labelling development topics extracted from commit logs.
 
@@
 human experts, and also are limited to project-specific topic labels.  
-In this paper, we introduce \emph{labelled topic extraction}, an approach to topic labelling that is both automatic and creates labels that are project independent.
+In this paper, we introduce \emph{labelled topic extraction}, an
+approach to topic labelling that creates labels automatically that are
+project independent.
+%is both automatic and creates labels that are project independent.
 
@@
 labelled topic extraction would allow managers to track effort
-related to topics, such as usability or other NFRs.
+related to NFR topics, such as usability.
 These techniques also allow for the annotation of commit comments
@@
 This would enable querying of bug reports and artifacts by relevant NFRs. 
-For instance a manager could confirm if their developers were
-focusing on \emph{usability} by looking for
+For instance a manager can confirm if their developers were
+focused on \emph{usability} by looking for
  \emph{usability}-relevant revisions and bug reports.
@@
 
-The idea of extracting higher-level \emph{concerns}, also known as
-\emph{topics}, \emph{concepts}, \emph{aspects} or \emph{requirements},
+The idea of extracting higher-level \emph{concerns} and \emph{topics}, also known as
+ \emph{concepts}, \emph{aspects} or \emph{requirements},
 has been approached from documentation-based and repository-based
@@
 Cleland-Huang and her colleagues have investigated on mining requirements documents for non-functional requirements (NFR) (quality requirements)~\cite{Cleland-Huang2006}. 
-Their approach is similar to ours, as they mined keywords from NFR catalogues.
+Their approach is similar to ours, as they mined keywords from NFR
+catalogues.
+They differ because they mine requirements documents where as we mine revisions.
 %One approach they tried was similar to this one, with keywords mined from NFR catalogues found in their previous work~\cite{chung99}. 
@@
 on a single project.
-They relied on requirements documents instead of version control
-histories that we use.
+%They relied on requirements documents instead of version control
+%histories that we use.
 %The objective of Cleland-Huang's study was to identify new NFRs for system development, yet our objective was to recover those latent NFRs from commit-log messages of the project.
@@
 %Choosing an older version of these projects allowed us to focus on projects which have moved further into the maintenance phase of the software life-cycle.
-For our study, we explicitly chose older versions of mature projects from a stable problem domain to increase the likelihood that we would encounter primarily maintenance activities in our studies.
+We explicitly chose older versions of mature projects from a stable problem domain to increase the likelihood that we would encounter primarily maintenance activities in our studies.
 
@@
 We extracted these messages and indexed them by creation time. 
-We summarized each message as a word distribution but removed stop-words such as ``\emph{the}'' and ``\emph{at}''. 
+We summarized each message as a word distribution minus stop-words such as ``\emph{the}'' and ``\emph{at}''. 
 %Stemming was performed in the later stages of our analysis. % WAS IT?
@@
 We applied Blei's LDA implementation~\cite{Blei2003} against the word distributions of these commits, and generated lists of topics per period. 
-We set the number of topics to generate to $20$, because past experimentation showed that fewer topics might aggregate multiple unique topics while any more topics seemed to dilute the results and create indistinct topics~\cite{Hindle09ICSM}.
+We set the number of topics to generate to $20$, because past experimentation showed that fewer topics might aggregate multiple unique topics while any more topics diluted the results and create indistinct topics~\cite{Hindle09ICSM}.
 % Furthermore, more than $20$ topics quickly became infeasible for inspection.
@@
 
-\begin{figure}
+\begin{figure}[t]
 \centering

### 2011-03-25 msr/msr2011.tex
intro and previous work

@@
 \newcommand{\shrink}{}
+
+\vspace*{1em}
 \section{Introduction}
@@
 understanding of \emph{why} a system has evolved the way it has. 
+This is different than \emph{how} a system has evolved, because the behaviour of change itself is the
+\emph{how}.
 Looking back on streams of artifacts scattered across different
@@
 % edit this. awkward
-This is different than \emph{how}, because the behaviour of change itself is the
-\emph{how}.
-In this work we seek to provide automated in labelling development topics extracted from commit logs.
+In this work we seek to provide a method of automatically labelling  development topics extracted from commit logs.
 
 Topic modeling is a machine learning technique which creates
-multinomial distributions over words from a text corpus. 
+multinomial distributions of words extracted from a text corpus. 
 The technique infers the hidden structure of a corpus using posterior
@@
 This topic might reasonably be labelled as
-``\emph{event handling}'' by a human who understands the domain well,
-despite the fact that label does not appear in the word list itself.  
+``\emph{event handling}'' by a developer who understands the domain well,
+despite the fact that this label does not appear in the word list itself.  
 Current approaches to topic labelling rely on manual intervention by
@@
 labelled topic extraction would allow managers to track effort
-related to NFR topics, such as usability.
+related to NFR topics, such as portability.
 These techniques also allow for the annotation of commit comments
@@
 \item evaluates three kinds of automatic topic labelling methods:
-  semi-unsupervised labelling (word-lists) of topics, supervised labelling of
+  semi-unsupervised labelling  of topics (word-lists), supervised labelling of
   topics with a single NFR (machine learning), 
   and supervised labelling of topics with multiple NFRs (multi-label
-  machine learning)
+  machine learning);
 
@@
 
-Cleland-Huang and her colleagues have investigated on mining requirements documents for non-functional requirements (NFR) (quality requirements)~\cite{Cleland-Huang2006}. 
+Cleland-Huang and her colleagues have investigated mining requirements documents for non-functional requirements (NFR) (quality requirements)~\cite{Cleland-Huang2006}. 
 Their approach is similar to ours, as they mined keywords from NFR
@@
 Our research differs because we use a more comprehensive set of terms based on the taxonomy we chose.
-In this paper we make cross-project comparisons instead of focusing
+Another difference is that we make cross-project comparisons instead of focusing
 on a single project.
@@
 Mockus and Votta leveraged WordNet~\cite{Fellbaum1998}, an English-language ``lexical database'' that contains semantic relations between words, including common related forms (similar to word stemming), meronymy and synonymy.
-They used WordNet for word roots as they felt the synonyms would be non-specific and cause errors, problems we also encountered. 
+They used WordNet for word roots as they felt the synonyms would be
+non-specific and cause errors.
+%, problems we also encountered. 
 Mockus et al. validated their labels with system developers.
@@
 Another approach is to extract concerns from software repositories.
-Marcus et al.~\cite{marcus04wcre} use Latent Semantic Indexing (LSI) to identify commonly occurring concerns for software maintenance.
-The concerns are given by the user, and LSI used to retrieve them from a corpus.
+Marcus et al.~\cite{marcus04wcre} use Latent Semantic Indexing (LSI)
+to identify commonly occurring concerns for software maintenance. 
+The
+concerns are given by the user, and LSI is used to retrieve them from
+a corpus. 
 Topic modelling generates topics that are independent of a user query, and relate only to word frequencies in the corpus.
 %Some results were interesting, but their precision was quite low. 
-With ConcernLines, Treude et al.~\cite{treude09cl} show tag occurrence using colour intensity. 
+With ConcernLines, Treude et al.~\cite{treude09cl} show tag occurrence
+using colour and intensity. 
 They mine
@@
 %This paper does this differently. 
-First of all, we automated the process of naming the topics. Secondly, we labelled topics with project-independent terms, in order to permit cross-project comparison.
+First of all, we automate the process of naming the topics. Secondly, we label topics with project-independent terms, in order to permit cross-project comparison.
 
@@
 %we compared the results of the unsupervised word matching approach. 
-For each quality we assessed whether semi-unsupervised labels matched the manual annotations. %To be clear, the manual annotations were not used to train the labeling process.
+For each quality we assessed whether semi-unsupervised labels matched the manual annotations. %To be clear, the manual annotations were not used to train the labelling process.
 As described in \ref{sec:wordlist} we used both ROC and F-1 measures to evaluate the performance of the classification.

### 2011-03-25 msr/msr2011.tex
up to automatic labelled topic extraction

@@
 \section{Study design and execution}
-Figure \ref{fig:process} gives an outline of methodology.
+Figure \ref{fig:process} gives an outline of our methodology.
 We began by gathering source data and creating topic models. For semi-unsupervised labelling, we generated three sets of word-lists as signifiers for NFRs.
@@
 MaxDB started in the late 1970s as a research project, and was later acquired by SAP. 
-As of version 7.500, released April 2007, the project has over $940,000$ lines of C source code\footnote{generated using David A. Wheeler's \emph{SLOCCount}, dwheeler.com/sloccount.}. 
+As of version 7.500, released April 2007, the project has over $940,000$ lines of C source code\footnote{generated using David A. Wheeler's \emph{SLOCCount}, http://dwheeler.com/sloccount.}. 
 The MySQL project started in 1994 and MySQL 3.23 was released in early 2001. 
@@
 Linear combinations of these $N$ word distributions are meant to represent and recreate the word distributions of any of the original messages. 
-These $N$ word distributions effectively form topics: cross-cutting collections of words relevant to one or more of our commit messages. 
+These $N$ word distributions effectively form \emph{topics}: cross-cutting collections of words relevant to one or more of our commit messages. 
 LDA extracts topics in an unsupervised manner; the algorithm relies
@@
 We applied Blei's LDA implementation~\cite{Blei2003} against the word distributions of these commits, and generated lists of topics per period. 
-We set the number of topics to generate to $20$, because past experimentation showed that fewer topics might aggregate multiple unique topics while any more topics diluted the results and create indistinct topics~\cite{Hindle09ICSM}.
+We set the number of topics to generate to $20$, because past experimentation showed that fewer topics might aggregate multiple unique topics while any more topics dilutes the results and creates indistinct topics~\cite{Hindle09ICSM}.
 % Furthermore, more than $20$ topics quickly became infeasible for inspection.
@@
 58]{Boegh2008},'' that is sufficient for the purposes of this research.
-We claim that these NFRs are maintenance concerns (to varying degrees) in all software projects, and are therefore well-suited to comparison between projects.
+We claim that these NFRs are maintenance concerns (to varying degrees) in all software projects, and are therefore well-suited for comparisons between projects.
 
@@
 created a validation set of manually labelled topics. For MySQL 3.23
-and MaxDB 7.500, we (the first two authors) annotated each extracted topic in each period with
+and MaxDB 7.500, the annotators (the first two authors) annotated each extracted topic in each period with
 the six NFR labels listed above.
-Annotators did not annotate each other's annotations but some brief
+Annotators did not annotate each other's annotations, but some brief
 inspection of annotations was used to confirm that the annotators were
 acting similarly.
-We looked at each period's topics, and assessed what the data --- consisting of the frequency-weighted word lists and messages --- suggested was the topic for that period. 
+We looked at each period's topics, and assessed what the data ---
+consisting of the frequency-weighted word lists and messages ---
+suggested was the label for that topic. 
 We were able to pinpoint the appropriate label using auxiliary information as well, such as the actual revisions and files that were related to the topic being annotated.
@@
 \emph{usability}, and \emph{maintainability}, the annotators annotated each topic
-with the relevant label and sometimes we used finer-grained
+with the relevant label and sometimes they used finer-grained
 annotations that would be aggregated up to one of these higher-level labels.
@@
 %, an English-language ``lexical database'' that contains semantic relations between words, including common related forms (similar to word stemming), meronymy and synonymy. 
-We then added Boehm's software quality model \cite{Boehm+:1976:ICSE}, and classified his eleven `ilities' into their respective ISO9126 NFRs. 
-We did the same for the quality model produced by McCall et al. \cite{mccall1977}. 
+We then added Boehm's software quality model~\cite{Boehm+:1976:ICSE}, and classified his eleven `ilities' into their respective ISO9126 NFRs. 
+We did the same for the quality model produced by McCall et al.~\cite{mccall1977}. 
 We then did a simple random analysis of mailing list messages from an open source project, KDE.

### 2011-03-26 msr/msr2011.tex
latest edits

@@
 All experiments were run on MaxDB 7.500 and MySQL 3.23 data. LDA
-extracted 20 topics per period for each project.
+extracted $20$ topics per period for each project.
 % Is this good enough
 This labelling is \emph{semi-unsupervised} because the corpus is not derived from 
-the project being labelled, and we did not label the project's topics
+the project being analyzed, and we did not label the project's topics
 ourselves for a training set. The motivation behind this technique is that
-because most software often addresses similar issues, we can use the
+because most software often addresses similar issues, thus we can use the
 domain knowledge of software to label relevant topics.
@@
 \emph{correctness} (182 topics) and \emph{testability} (121). 
-We did not get good results for \emph{usability} or \emph{accuracy}, which were associated with fewer than ten topics. 
-We also looked for correlation between our labels: excluding double matches (self-correlation), our highest co-occurring terms were \emph{verifiability} with \emph{traceability}, and \emph{testability} with \emph{correctness} (76 and 62 matches, respectively).
+We did not see many results for \emph{usability} or \emph{accuracy}, which were associated with fewer than ten topics. 
+We also looked for correlations between our labels: excluding double matches (self-correlation), our highest co-occurring terms were \emph{verifiability} with \emph{traceability}, and \emph{testability} with \emph{correctness} (76 and 62 matches, respectively).
 % See the following sections for our error analysis.
@@
 As we mentioned, the word-lists are broad, so there are likely to be false-positives (discussed below). 
-We found a high of 265 topics for \emph{usability}, with a low of 44 topics for \emph{maintainability}. 
+The most frequent label, 265 topics, was \emph{usability}, and the
+least frequent label, 44 topics, was \emph{maintainability}. 
 Common co-occurrences were \emph{reliability} with \emph{usability}, \emph{efficiency} with \emph{reliability}, and \emph{efficiency} with \emph{usability} (200, 190, and 150 topics in common, respectively).
@@
  \includegraphics[width=0.8\textwidth]{figures/unsupervised-bar}
- \caption{Performance, ROC values (range: 0-1), of semi-unsupervised topic labelling for
+ \caption{Performance, ROC values (range: $0$--$1$), of semi-unsupervised topic labelling for
    each NFR and per word-list. Dashed line indicates the performance of a random classifier. This graph shows how well the
@@
 \label{sec:suplabelling}
-Supervised labelling requires expert analysis of the correct class/label to assign to a topic.
+Supervised labelling requires expert analysis of the correct
+class/label to assign a label to a topic.
 In our approach, we use the top-level NFRs in the ISO 9126 standard~\cite{iso9126} for our classes, but other taxonomies are also applicable.%In order to validate how effective these word-bag approaches to topic labelling would be we created a validation data set.
@@
 % }
-\caption[]{ROC value for the best learner per label for MaxDB and MySQL. Values range from 0--1.  Dashed line indicates the performance of a random classifier.
+\caption[]{ROC value for the best learner per label for MaxDB and MySQL. Values range from $0$--$1$.  Dashed line indicates the performance of a random classifier.
 %\todo[inline]{better discussion}

### 2011-03-26 msr/msr2011.tex
more edits

@@
  \caption{Performance, ROC values (range: $0$--$1$), of semi-unsupervised topic labelling for
-   each NFR and per word-list. Dashed line indicates the performance of a random classifier. This graph shows how well the
+   each NFR and per word-list. The dashed line indicates the performance of a random classifier. This graph shows how well the
    semi-unsupervised topic labelling matched our manual annotations.}
@@
 For both projects Bayesian techniques did the best out of a wide variety of machine learners tested. 
-Our best learners, Discriminative Multinomial Naive Bayes, Naive Bayes  and Multinomial Naive Bayes  are all based on Bayes's theorem and all assume, naively, that the features are independent. 
+Our best learners, Discriminative Multinomial Naive Bayes, Naive Bayes
+and Multinomial Naive Bayes  are all based on Bayes's theorem and all
+assume, naively, that the features supplied are independent. 
 %The features we used are word counts per message. 
@@
 The next section (\ref{sec:multilabel})
-addresses the issue of a lack of independence and correlation between labels.
-In the next section we will evaluate how well these learners perform
-together.
+addresses the issue of a lack of independence and correlation between
+labels
+using multi-label learners.
+%In the next section we will evaluate how well these learners perform
+%together.
 %\todo[inline]{Comment about correlation, we should link to the Mulan section, perfect segway since Mulan relies on correlation.}
@@
 % more why
-Labelled topics address \emph{why}  because they show why changes
+Labelled topics address \emph{why}  because they show reasons why changes
 occurred rather than \emph{how}. 
@@
 \item \emph{Do NFR frequencies change over time?} If a particular NFR was of more interest at one point in the life-cycle than another, this suggests that development activity shifted focus.
-For example, if a developer expected to see a recent focus on \emph{reliability}, but instead \emph{usability} dominated, he or she might re-prioritize up-coming work items.
+For example, if a developer expected to see a recent focus on \emph{reliability}, but instead \emph{usability} dominated, they might re-prioritize up-coming work items.
 \item \emph{Do projects differ in their relative interest in NFRs?} A
@@
 	\caption[]{NFR label per period. Each cell represents a 30-day period. %Numbers represent number of topics in that period labelled with that NFR. 
-	Grid cell intensity (saturation) is mapped to label frequency relative to the largest label value of \emph{all} NFRs.
-Grey histogram bars indicate label frequency relative to that particular NFR's largest label value.
-Dashed vertical lines relate a project milestone (*Key Event*) to our topic windows.
+	Grid cell intensity (saturation) is mapped to label frequency
+        relative to the largest label count of \emph{all} NFRs. Grey
+        histogram bars indicate label frequency relative to that
+        particular NFR's largest label count. Dashed vertical lines
+        relate a project milestone (a \emph{key event}) to our topic windows. 
 % \todo[inline]{better discussion}

### 2011-03-26 msr/msr2011.tex
round 1 done!

@@
 we estimate that \textsf{exp1} had poor
-performance via the overlap between ISO-9126 and the Kayed ontology.
+performance via the overlap between ISO9126 and the Kayed ontology.
 For \textsf{exp1} the F-measures for MaxDB were from $0$ to $0.18$ with an average
@@
 class/label to assign a label to a topic.
-In our approach, we use the top-level NFRs in the ISO 9126 standard~\cite{iso9126} for our classes, but other taxonomies are also applicable.%In order to validate how effective these word-bag approaches to topic labelling would be we created a validation data set.
+In our approach, we use the top-level NFRs in the ISO9126 standard~\cite{iso9126} for our classes, but other taxonomies are also applicable.%In order to validate how effective these word-bag approaches to topic labelling would be we created a validation data set.
 
@@
 For example, both projects show a high number of NFRs recognized at the first period of analysis. 
-This is due to our window choice: we deliberately targeted our analysis to when MySQL 3.23 was first announced %Major update \& End-of life
-and when MaxDB 7.500 was first announced. For MaxDB, version 7.5.00  was released in December of 2003. 
+This is due to our window choice: we deliberately targeted our
+analysis to when both MySQL 3.23
+and MaxDB 7.500 were first announced. For MaxDB, version 7.5.00  was released in December of 2003. 
 We know that release 7.5.00.23 saw the development of PHP interfaces, possibly accounting for the simultaneous increase in the \emph{portability} NFR at the same time.
-The gap in MaxDB (Figure \ref{fig:maxdb-timeline}) is due to a shift in development focus (from February 2005 to Jun 2005) to MaxDB 7.6, which is released in June 2005.
+The gap in MaxDB (Figure \ref{fig:maxdb-timeline}) is due to a shift in development focus (from February 2005 to June 2005) to MaxDB 7.6, which is released in June 2005.
 
@@
 whereas MySQL was in its relative infancy, and  
-security problems were more common (security is a component of \emph{functionality} in the ISO 9126 model). 
+security problems were more common (security is a component of \emph{functionality} in the ISO9126 model). 
 In both cases \emph{portability} was a constant maintenance concern and was prevalent throughout the lifetime of the projects. It may surprise developers how often portability arises as a concern.
@@
 \subsection{Annotation observations}
-We found many topics that were not non-functional requirements (NFRs) but were often related to them. 
+We found many topics that were not actually non-functional requirements (NFRs) but were often related to them. 
 For instance, concurrency was mentioned often in the commit logs and
@@
 %We used mailing-lists to verify the purpose behind some events.
-Our taxonomy for software NFRs is subject to dispute, but seems generally accepted.
-Finally, there are exogenous sources, such as in-person discussions, which we did not access, an omission as noted in \cite{aranda09icse}.
-%We did not show these results to the stakeholders.
+Our taxonomy for software NFRs is subject to dispute, but seems
+generally accepted. Finally, there are exogenous sources, such as
+in-person discussions, which we did not access, an omission as noted
+in Aranda et al.~\cite{aranda09icse} %We did not show these results to the stakeholders.
 
@@
 %costly. Furthermore, previous work was project-specific and thus incomparable. 
-To improve on this, we leveraged software engineering standards, specifically the ISO quality taxonomy, to produce a method of partially-automated (supervised) and fully-automated (semi-unsupervised) topic labelling.
+To improve on this, we leveraged software engineering standards,
+specifically the ISO9126 quality taxonomy, to produce a method of partially-automated (supervised) and fully-automated (semi-unsupervised) topic labelling.
 Since the word-list technique is not project-specific, we used it to compare two distinct projects, where we showed our technique produced interesting insight into maintenance activity.

### 2011-03-26 msr/msr2011.tex
10 pages

@@
  \usepackage{color}
+
+\usepackage{balance}
 
@@
 % %
+
+% Only for camera ready -- prevent TeX from rendering widows
+\widowpenalty=9999
+\clubpenalty=9999
 
@@
 
-\vspace*{1em}
+\newpage
+
 \section{Introduction}
@@
 MaxDB started in the late 1970s as a research project, and was later acquired by SAP. 
-As of version 7.500, released April 2007, the project has over $940,000$ lines of C source code\footnote{generated using David A. Wheeler's \emph{SLOCCount}, http://dwheeler.com/sloccount.}. 
+As of version 7.500, released April 2007, the project has over $940,000$ lines of C source code\footnote{generated using David A. Wheeler's \emph{SLOCCount}, \tiny{http://dwheeler.com/sloccount}.}. 
 The MySQL project started in 1994 and MySQL 3.23 was released in early 2001. 
@@
 %\noindent \textbf{Creating a validation corpus} ---
+
+%FAT
+\newpage
+
 \subsubsection{Creating a validation corpus}
@@
 ROC of a classifier is equivalent to the probability that the classifier will rank a randomly chosen positive instance higher than a randomly chosen negative instance.
-
 We consider our labelling classifiers acceptable if they outperform a random classifier (0.5).
@@
 In this section we describe how to label topics based on dictionaries
-mined from sources external to the projects themselves.
+mined from sources external to the projects.
 
@@
 The labels we used:
+
+%FAT we can remove this
+\newpage
 \begin{quotation}
@@
 
+%FAT
+\balance
+\newpage
+
 For \textsf{exp2}, there are more unnamed topics than \textsf{exp1}. 
@@
 
-Based on these results we find that \emph{reliability} and
+Thus we find that \emph{reliability} and
 \emph{usability} worked well for MaxDB in \textsf{exp2} and better in
@@
 but other qualities such as \emph{functionality} did improve. 
-Many of ROC scores were $0.6$ or less, but our classifier nonetheless performed substantially better than random.
-
+Many of ROC scores were $0.6$ or less, but our classifier still performed substantially better than random.
+
+%\vspace*{-1em}
 \subsection{Supervised labelling}
@@
 
-We used a suite of supervised classifiers, WEKA~\cite{weka09}. 
-WEKA includes machine learning tools such as support vector machines and Bayes-nets. 
+We used a suite of supervised classifiers, WEKA~\cite{weka09},
+that includes machine learning tools such as support vector machines and Bayes-nets. 
 We also used the multi-labelling add-on for WEKA, Mulan~\cite{mulan}. %\footnote{\url{http://mlkd.csd.auth.gr/multilabel.html}}. 
-Traditional classifiers map our topics to a single class, whereas Mulan allows for a mixture of classes per topic, which maps to what we observed while manually labelling topics.
+Traditional classifiers label topics with a single class, whereas Mulan allows for a mixture of classes per topic, which is what we observed while manually labelling topics.
 The \emph{features} we used are word counts/occurrence  per topic, if
@@
 We tried other learners that WEKA~\cite{weka09} provides: rule learners, decision tree learners, vector space learners, and support vector machines.  
-Figure \ref{fig:best-learn-per-tag} shows the performance of the best performing learner per label. 
- We considered the best learner for a label to be the learner which
- had the highest ROC value for that label. 
-The best learner is important because one uses a single learner per label. In applying our technique, we expect the user to select, for each NFR, the best learner possible.
+Figure \ref{fig:best-learn-per-tag} shows the performance of the best
+performing learner per label: 
+% We considered the best learner for a label to be 
+the learner that had the highest ROC value for that label. 
+The best learner is important because one uses a single learner per
+label. When applying our technique, for each NFR one should select the best learner possible.
 %Figure \ref{fig:best-learn-per-tag} uses the ZeroR learner as a baseline, a common practice in machine learning,
@@
 
+\balance
+
 \subsection{Applying multiple labels to topics}
@@
 \end{figure*}
+
+%FAT 
+%\newpage
 
@@
 \begin{enumerate}
-\item \emph{Do NFR frequencies change over time?} If a particular NFR was of more interest at one point in the life-cycle than another, this suggests that development activity shifted focus.
-For example, if a developer expected to see a recent focus on \emph{reliability}, but instead \emph{usability} dominated, they might re-prioritize up-coming work items.
+\item \emph{Do NFR frequencies change over time?} If a particular NFR
+  was of more interest at one point in the life-cycle than another,
+  this suggests that development activity shifted focus. For example,
+  if a developer expected to see a recent focus on \emph{reliability},
+  but instead \emph{usability} dominated, they might re-prioritize
+  upcoming work items.
+
 \item \emph{Do projects differ in their relative interest in NFRs?} A
@@
 \end{enumerate}
+
+%FAT
+
+%\newpage
 
@@
 
-\noindent\textbf{1. Do NFR frequencies change over time?} --- 
+\subsection{Do NFR frequencies change over time?}
+%\noindent\textbf{1. Do NFR frequencies change over time?} --- 
 %\subsubsection{Do label frequencies change over time?}
@@
 In MaxDB, we can observe that \emph{Maintainability} topics became more prevalent as MaxDB matures. 
-This is likely due to our analysis timeframe for MaxDB being shorter than the timeframe for the MySQL product.
-
+This is likely due to our analysis time-frame for MaxDB being shorter than the time-frame for the MySQL product.
+
+%FAT
+\newpage
+
+\subsection{Do  projects differ in their relative  topic interest?}
 %\subsubsection{Do  projects differ in their relative  topic interest?}
-\noindent\textbf{2. Do  projects differ in their relative  topic interest?} ---
+%\noindent\textbf{2. Do  projects differ in their relative  topic interest?} ---
 Yes. MySQL 3.23 had proportionally more
@@
 
+%FAT
+%\newpage
 \subsection{Summary of techniques}
@@
 We improved internal validity by trying to correlate and explain the behaviours observed in the analysis with the historical records of the projects.
-We did not attempt to match our results to any particular model or
-theory.
-Also we did not have enough manpower to assess inter-rater reliability.
+We did not attempt to match our results to any particular model.
+We unable to assess inter-rater reliability.
 
@@
 did not rate the same documents.
+
+\balance
+\newpage
 
@@
 
+\balance
+
 \bibliographystyle{abbrv}
@@
 \bibliography{icpc}
+
+\balance
 

### 2011-03-26 msr/msr2011.tex
added General Terms
back to where I started?
updated the fonts

@@
 % Only for camera ready -- prevent TeX from rendering widows
-\widowpenalty=9999
-\clubpenalty=9999
+%\widowpenalty=9999
+%\clubpenalty=9999
+\clubpenalty = 10000
+\widowpenalty = 10000
+\displaywidowpenalty = 10000
 
@@
 
-\conferenceinfo{MSR}{'11 Waikiki, Hawaii, USA}
+\conferenceinfo{MSR}{'11, May 21-22, 2011, Waikiki, Honolulu, HI, USA}
+\CopyrightYear{2011}
+\crdata{978-1-4503-0574-7/11/05}
 
@@
 
+\vspace*{-0.4em}
+
 \category{D.2.9}{Software Engineering}{Management}[Lifecycle]
@@
 
-%\terms{Theory}
+\vspace*{-1em}
+
+\terms{Human Factors, Management, Measurement}
 
@@
 \newcommand{\shrink}{}
-
-\newpage
 
@@
 MaxDB started in the late 1970s as a research project, and was later acquired by SAP. 
-As of version 7.500, released April 2007, the project has over $940,000$ lines of C source code\footnote{generated using David A. Wheeler's \emph{SLOCCount}, \tiny{http://dwheeler.com/sloccount}.}. 
+As of version 7.500, released April 2007, the project has over $940,000$ lines of C source code\footnote{generated using David A. Wheeler's \emph{SLOCCount}, {http://dwheeler.com/sloccount}.}. 
 The MySQL project started in 1994 and MySQL 3.23 was released in early 2001. 
@@
 %FAT
-\newpage
+%\newpage
 
@@
 %FAT we can remove this
+%\balance
+%LFAT
 \newpage
@@
 
-%FAT
+%LFAT
 \balance
@@
 
-\balance
+%\balance
 
@@
         particular NFR's largest label count. Dashed vertical lines
-        relate a project milestone (a \emph{key event}) to our topic windows. 
+        relate a project milestone (\emph{*Key events*}) to our topic windows. 
 % \todo[inline]{better discussion}
@@
 
-%FAT
+%LFAT
+\balance
 \newpage
-
 \subsection{Do projects differ in their relative  topic interest?}
@@
 
+%LFAT
 \balance

### 2011-03-26 msr/msr2011.tex
latest greatest stuff

@@
 multinomial distributions of words extracted from a text corpus. 
-The technique infers the hidden structure of a corpus using posterior
+This technique infers the hidden structure of a corpus using posterior
 inference: the probability of the hidden structure given the data. 
@@
 % it is ok to repeat this here because it is clear and bullet pointed and the reviewers wont have any bloody excuse to not see it
-In this paper, we describe \emph{labelled topic extraction}. It addresses two gaps in the topic mining literature:
+In this paper, we describe \emph{automated labelled topic extraction}. It addresses two gaps in the topic mining literature:
 \begin{enumerate}
@@
 This is because traditional topic mining techniques are specific to a particular data-set. 
-\textit{Labelled topic extraction} allows for comparisons \textit{between} projects. 
+\textit{Automated labelled topic extraction} allows for comparisons \textit{between} projects. 
   \item Topic modeling creates word lists that require interpretation by the user to assign meaning. 
@@
 E.g., we might have a label called \emph{`Oracle'} in the MySQL case, since Oracle owns MySQL.
-We do this differently.
+%We do this differently.
 %This paper does this differently. 
-First of all, we automate the process of naming the topics. Secondly, we label topics with project-independent terms, in order to permit cross-project comparison.
+Our approach differs: first of all, we automate the process of naming the topics; secondly, we label topics with project-independent terms, in order to permit cross-project comparison.
 
@@
 They describe probabilistic labelling, using the frequency distribution of words in a topic to create a meaningful phrase. 
-They do not use external domain-specific information as we do and we
+They do not use external domain-specific information as we do, but we
 do not generate phrases from the topics.
@@
 Hindle et al.~\cite{Hindle09ICSM} propose a windowed method of topic
-analysis, we extend this with labelled
+analysis that we extend with labelled
 topics, NFRs and new visualizations.
@@
 ISO9126 describes six high-level NFRs: maintainability, functionality,
-portability, efficiency, usability, and reliability~\footnote{While there may be lingering debate in some circles about these terms, an ISO standard seems like a reasonable starting point for our work.}
+portability, efficiency, usability, and reliability~\footnote{While there may be lingering debate in some circles about these terms, an ISO standard seems like a reasonable starting point for our work.}.
 %There is some debate about the terms in this model. 
@@
 \emph{usability}, and \emph{maintainability}, the annotators annotated each topic
-with the relevant label and sometimes they used finer-grained
+with the relevant label. Sometimes they used finer-grained
 annotations that would be aggregated up to one of these higher-level labels.
@@
 we associate each NFR with a list
-of keywords (`word-lists', in our parlance). These word-lists were determined a priori and were not extracted from the projects themselves.
+of keywords (\emph{word-lists}, in our parlance). These word-lists were determined a priori and were not extracted from the projects themselves.
 We intersected the words of the topics and the words of our word-lists.
-We ``labeled'' a topic if any of its words matched any of the word-list's words.
+We ``labelled'' a topic if any of its words matched any of the word-list's words.
 A topic could match more than one NFR.
@@
 we considered to be
-relevant to discussion of reliability, but is not in the standard. 
+relevant to discussion of \emph{reliability}, but is not in the standard. 
 We therefore took the NFRs from the ISO9126 and added to them.
@@
 
-Using our three word-lists (\textsf{exp1}, \textsf{exp2}, \textsf{exp3}), we labeled our topics with an NFR where there was a match between a word in the list and the same word somewhere in the distribution of words that constitute the topic.
-A \emph{named topic} is a topic with a matching NFR label. Unnamed
-topics occur where there is no such match, which may indicate either a lack of precision, or simply that this topic is not associated with non-functional requirements.
+Using our three word-lists (\textsf{exp1}, \textsf{exp2}, \textsf{exp3}), we labelled our topics with an NFR where there was a match between a word in the list and the same word somewhere in the distribution of words that constitute the topic.
+A \emph{named topic} is a topic with a matching NFR label. \emph{Unnamed
+topics} occur where there is no such match, which may indicate either a lack of precision, or simply that this topic is not associated with non-functional requirements.
 All experiments were run on MaxDB 7.500 and MySQL 3.23 data. LDA
@@
 For each quality we assessed whether semi-unsupervised labels matched the manual annotations. %To be clear, the manual annotations were not used to train the labelling process.
-As described in \ref{sec:wordlist} we used both ROC and F-1 measures to evaluate the performance of the classification.
+As described in Section \ref{sec:wordlist} we used both ROC and F-1 measures to evaluate the performance of the classification.
 Figure \ref{fig:maxdb-unsup-results} shows our ROC results for MaxDB and MySQL. We describe F-1 results in the text below.
@@
 but other qualities such as \emph{functionality} did improve. 
-Many of ROC scores were $0.6$ or less, but our classifier still performed substantially better than random.
+Many ROC scores were $0.6$ or less, but our classifier still performed substantially better than random.
 
@@
 }
-\caption[]{MySQL and MaxDB macro (grey) and micro (black) ROC results per multi-label learner. Possible values range from 0--1.  Dashed line indicates the performance of a random classifier.
+\caption[]{MySQL and MaxDB macro (grey) and micro (black) ROC results per multi-label learner. Possible values range from $0$--$1$.  Dashed line indicates the performance of a random classifier.
 %\todo[inline]{better discussion}% and graphical legend would be nice}
@@
 These classifiers performed better than other multi-label classifiers as they have the best micro and macro ROC scores. 
-The multi-label and single-label learners had similar performance: for MySQL, BR and NaiveBayes had similar macro-ROC scores of $0.74$.
+The multi-label and single-label learners had similar performance: for
+MySQL, BR and Naive Bayes had similar macro-ROC scores of $0.74$.
 % although their results are comparable to the naive Bayesian learners we used in Section \ref{sec:suplabelling}.
@@
 Similarly, in August, 2002, both \emph{functionality} and \emph{portability} are frequent, and mailing list data suggests this is related to the release of version 3.23.52, a general bug fix with a focus on security (a component of the \emph{functionality} NFR in the ISO9126 model).
-After this point, efforts shift to the newer releases (4.0, 4.1, 5.0). With respect to our research questions:
+After this point, efforts shift to the newer releases (4.0, 4.1,
+5.0). We now address our research questions:
 
@@
 these kinds of changes are slightly related to \emph{maintainability}. 
-A non-functional change that was not quality-related was licensing and copyright; many changes were simply to do with updating copyrights or ensuring copyright or license headers were applied to files.
-In these cases we assigned the \emph{None} class.
+A non-functional change that was not quality-related was licensing and
+copyright; many changes were simply to do with updating copyrights or
+ensuring copyright or license headers were applied to files. In these
+cases we assigned the \emph{None} label to the topic.
 
@@
 
-We found that the multi-label learners of BR, CLR and HOMER only did as well or worse for Macro-ROC as the single-label NaiveBayes and other naive Bayes-derived learners. 
-This suggests that by combining together multiple NaiveBayes learners we could probably label sets of topics effectively, but it would require a separate NaiveBayes learner per label.
+We found that the multi-label learners of BR, CLR and HOMER only did
+as well or worse for Macro-ROC as the single-label Naive Bayes and other naive Bayes-derived learners. 
+This suggests that by combining together multiple Naive Bayes learners
+we could probably label sets of topics effectively, but it would
+require a separate Naive Bayes learner per label.
 
@@
 Since this work focuses on labelling natural language commit log
-comments we feel it can adapted to other natural language artifacts
+comments we feel it can be adapted to other natural language artifacts
 such as mailing-list discussions and bug reports, even though that was
@@
 We did not attempt to match our results to any particular model.
-We unable to assess inter-rater reliability.
+We were unable to assess inter-rater reliability.
 

### 2011-03-28 msr/msr2011.tex
Updated requirements

@@
    \affaddr{Davis, CA, USA}\\
-   \email{\large abram@softwareprocess.es}
+   \email{abram@softwareprocess.es}
 \alignauthor  Neil A. Ernst\\
@@
    \affaddr{Trento, ITALY}\\
-   \email{\large jm@disi.unitn.it}
+   \email{jm@disi.unitn.it}
 }
@@
 %\title{What's in a name? Automated topic naming of software maintenance activities}
-\title{Automated topic naming to support cross-project analysis of software maintenance activities}
+\title{Automated Topic Naming to Support Cross-project Analysis of Software Maintenance Activities}
 \maketitle
@@
 
-\section{Study design and execution}
+\section{Study Design and Execution}
 Figure \ref{fig:process} gives an outline of our methodology.
@@
 
-\subsection{Generating the data}
+\subsection{Generating the Data}
 \label{sec:wordlist}
@@
 
-\subsubsection{The high-level labels}
+\subsubsection{The High-level Labels}
 
@@
 
-\subsubsection{Creating a validation corpus}
+\subsubsection{Creating a Validation Corpus}
 To evaluate both semi-unsupervised and supervised classification, we
@@
 
-\subsection{Semi-unsupervised labelling}
+\subsection{Semi-unsupervised Labelling}
 \label{sec:unsuplabelling}
@@
 
-\subsubsection{Generating word lists}
+\subsubsection{Generating Word Lists}
 
@@
 
-\subsubsection{Analysis of the semi-unsupervised labelling} %--- %Based
+\subsubsection{Analysis of the Semi-unsupervised Labelling} %--- %Based
                                 %on the labels, and our manual topic
@@
 %\vspace*{-1em}
-\subsection{Supervised labelling}
+\subsection{Supervised Labelling}
 \label{sec:suplabelling}
@@
 
-\subsubsection{Analysis of the supervised labelling}
+\subsubsection{Analysis of the Supervised Labelling}
 % \label{sec:suplabelling}
@@
 
-\subsection{Applying multiple labels to topics}
+\subsection{Applying Multiple Labels to Topics}
 \label{sec:multilabel}
@@
 
-\section{Understanding software maintenance activities} 
+\section{Understanding Software Maintenance Activities} 
 \label{sec:analysis}
@@
 
-\subsection{Do NFR frequencies change over time?}
+\subsection{Do NFR Frequencies Change over Time?}
 %\noindent\textbf{1. Do NFR frequencies change over time?} --- 
@@
 \newpage
-\subsection{Do projects differ in their relative  topic interest?}
+\subsection{Do Projects Differ in their Relative  Topic Interest?}
 %\subsubsection{Do  projects differ in their relative  topic interest?}
@@
 %\subsection{Effectiveness}
-\subsection{Annotation observations}
+\subsection{Annotation Observations}
 We found many topics that were not actually non-functional requirements (NFRs) but were often related to them. 
@@
 %\newpage
-\subsection{Summary of techniques}
+\subsection{Summary of Techniques}
 While an unsupervised technique such as LDA is appealing in its lack of human intervention, and thus lower effort, 
@@
 
-\subsection{Threats to validity}
+\subsection{Threats to Validity}
 % can always be cut down
@@
 
-\subsection{Future work}
+\subsection{Future Work}
 There are several avenues of further investigation.  
@@
 
-\appendix
-
-Our data and scripts are available at:
-
- \url{http://softwareprocess.es/nomen/}
-
 \balance
@@
 \balance
+
+\appendix
+
+Our data and scripts are available at:
+
+\url{http://softwareprocess.es/nomen/}
 

### 2011-08-09 ese/ese.tex
initial journal template

%
\RequirePackage{fix-cm}
%
%\documentclass{svjour3}                     % onecolumn (standard format)
%\documentclass[smallcondensed]{svjour3}     % onecolumn (ditto)
\documentclass[smallextended]{svjour3}       % onecolumn (second format)
%\documentclass[twocolumn]{svjour3}          % twocolumn
%
\smartqed  % flush right qed marks, e.g. at end of proof
%
\usepackage{graphicx}
%
% \usepackage{mathptmx}      % use Times fonts if available on your TeX system
%
% insert here the call for the packages your document requires
%\usepackage{latexsym}
% etc.
%
% please place your own definitions here and don't use \def but
% \newcommand{}{}
%
% Insert the name of "your journal" with
% \journalname{Empirical Software Engineering}
%
\begin{document}

\title{Insert your title here
%\thanks{Grants or other notes
%about the article that should go on the front page should be
%placed here. General acknowledgments should be placed at the end of the article.}
}

%\titlerunning{Short form of title}        % if too long for running head

\author{First Author         \and
        Second Author %etc.
}

%\authorrunning{Short form of author list} % if too long for running head

\institute{F. Author \at
              first address \\
              Tel.: +123-45-678910\\
              Fax: +123-45-678910\\
              \email{fauthor@example.com}           %  \\
%             \emph{Present address:} of F. Author  %  if needed
           \and
           S. Author \at
              second address
}

\date{Received: date / Accepted: date}
% The correct dates will be entered by the editor

\maketitle

\begin{abstract}

\keywords{First keyword \and Second keyword \and More}

\end{abstract}

\section{Introduction}

%\begin{acknowledgements}
%\end{acknowledgements}

% BibTeX users please use one of
%\bibliographystyle{spbasic}      % basic style, author-year citations
\bibliographystyle{spmpsci}      % mathematics and physical sciences
%\bibliographystyle{spphys}       % APS-like style for physics
%\bibliography{}   % name your BibTeX data base

\end{document}
% end of file template.tex

### 2011-08-10 ese/ese.tex
added todos, some sections on PG background. ALso added third research question about devs

@@
 \usepackage{graphicx}
-%
-% \usepackage{mathptmx}      % use Times fonts if available on your TeX system
-%
-% insert here the call for the packages your document requires
-%\usepackage{latexsym}
-% etc.
-%
-% please place your own definitions here and don't use \def but
-% \newcommand{}{}
-%
-% Insert the name of "your journal" with
-% \journalname{Empirical Software Engineering}
-%
+ \usepackage{booktabs}
+\usepackage[caption=false]{subfig}
+ \usepackage[colorinlistoftodos,backgroundcolor=yellow]{todonotes} %disable
+ \usepackage{color}
+
+\usepackage{balance}
+
+\newcommand{\XXX}[1]{\textcolor{red}{{\it \textbf{[XXX: #1]}}}}
+\hyphenation{Max-DB data-base}
+
+\usepackage{comment}
+\newtheorem{mydef}{definition}
+\newcommand{\slbl}[1]{\textbf{\textsf{#1}}}
+\usepackage{multirow}%,multicolumn}
+\usepackage{slashbox}
+\usepackage{wrapfig}
+\usepackage{booktabs, colortbl}
+\usepackage{amsmath}
+\usepackage{url}
+\usepackage[pdfstartview=FitH,bookmarksopenlevel=3,bookmarks=true]{hyperref} %pdfpagescrop={92 112 523 778},
+\journalname{Empirical Software Engineering}
+
 \begin{document}
 
-\title{Insert your title here
+\title{Automated Topic Naming
 %\thanks{Grants or other notes
@@
 }
-
+\subtitle{Supporting Cross-project Analysis of Software Maintenance Activities}
 %\titlerunning{Short form of title}        % if too long for running head
 
-\author{First Author         \and
-        Second Author %etc.
+%\numberofauthors{4}
+
+\author{
+Abram Hindle \and
+Neil A. Ernst \and
+Michael W. Godfrey \and
+John Mylopoulos
 }
-
 %\authorrunning{Short form of author list} % if too long for running head
 
-\institute{F. Author \at
-              first address \\
-              Tel.: +123-45-678910\\
-              Fax: +123-45-678910\\
-              \email{fauthor@example.com}           %  \\
-%             \emph{Present address:} of F. Author  %  if needed
+\institute{
+  Abram Hindle \at 
+  Dept. of Computer Science\\
+  University of California, Davis \\
+  Davis, CA, USA\\
+  \email{abram@softwareprocess.es}
 \and
-           S. Author \at
-              second address
+  Neil A. Ernst \at
+  Dept. of Computer Science\\
+  University of Toronto\\
+  Toronto, Ontario, CANADA\\
+  \email{nernst@cs.toronto.edu}
+\and
+  Michael W. Godfrey \at
+  David Cheriton School of Computer Science\\
+  University of Waterloo\\
+  Waterloo, Ontario, CANADA\\
+  \email{migod@uwaterloo.ca}
+\and 
+  John Mylopoulos \at
+  Dept. Information Eng. and Computer Science\\
+  University of Trento\\
+  Trento, ITALY\\
+  \email{jm@disi.unitn.it}
 }
@@
 \begin{abstract}
-
-\keywords{First keyword \and Second keyword \and More}
+% %\todo[inline]{Topics and LDA sooner, Topics/concept location is a known
+Researchers have employed a variety of techniques to extract underlying
+topics that relate to software development artifacts.  Typically, these
+techniques use semi-unsupervised machine-learning algorithms to suggest
+candidate word-lists.  However, word-lists are
+difficult to 
+interpret in the absence of meaningful summary labels.  Current topic modeling 
+techniques assume manual labelling and do not use domain-specific knowledge
+to improve, contextualize, or describe results for the developers.  
+We propose a solution: \emph{automated labelled topic extraction}.
+  Topics are extracted 
+using  Latent Dirichlet Allocation (LDA) from commit-log comments recovered
+from source control systems such as CVS and BitKeeper.  
+These topics are given labels from a generalizable
+cross-project taxonomy, consisting of non-functional
+requirements. Our approach was evaluated with experiments and case studies
+on two large-scale RDBMS projects: MySQL and MaxDB.  The case studies
+show that labelled topic extraction can produce appropriate,
+context-sensitive labels relevant to these projects, which provides fresh
+insight into their evolving software development activities.
+%160 words
+\keywords{ Software maintenance \and Repository mining \and Latent dirichlet allocation \and Topic models}
 
@@
 \section{Introduction}
+A key problem for practising software maintainers is gaining an
+understanding of \emph{why} a system has evolved the way it has. 
+This is different than \emph{how} a system has evolved, because the behaviour of change itself is the
+\emph{how}.
+Looking back on streams of artifacts scattered across different
+repositories, inferring what activities were performed, when, and for
+what reasons, is hard to do without expert advice from the developers
+involved. 
+%new stuff
+% edit this. awkward
+In this work we seek to provide a method of automatically labelling  development topics extracted from commit logs.
+
+Topic modeling is a machine learning technique which creates
+multinomial distributions of words extracted from a text corpus. 
+This technique infers the hidden structure of a corpus using posterior
+inference: the probability of the hidden structure given the data. 
+Topic models are useful in software maintenance because they summarize
+the key concepts in a corpus, such as source code, commit comments, or
+mailing-list messages, by identifying statistically co-occurring words. 
+Among other uses, it can quickly give developers an overview of where significant
+activity has occurred, and gives managers or maintainers a sense of
+project history. 
+%A major challenge for topic modeling is to accurately identify the
+%summary label for a topic; in the previous case, perhaps
+%\emph{``event-handling''} would be appropriate. 
+%Current approaches are manual: they rely on human expertise to
+%recognize the appropriate label. 
+%Furthermore, to date the results of topic modeling are project
+%specific, and not generalizable. 
+%This paper presents an approach, \emph{labelled topic extraction}, which addresses these two problems.
+
+%migod
+While machine learning techniques can automatically identify clumps of
+commonly recurring terms, devising an appropriate summary label for
+each clump/topic is harder.  
+A given topic extracted from a set of commit logs might consist of the following terms: \emph{ ``listener change remove add fire''}. 
+This topic might reasonably be labelled as
+``\emph{event handling}'' by a developer who understands the domain well,
+despite the fact that this label does not appear in the word list itself.  
+Current approaches to topic labelling rely on manual intervention by
+human experts, and also are limited to project-specific topic labels.  
+In this paper, we introduce \emph{labelled topic extraction}, an
+approach to topic labelling that creates labels automatically that are
+project independent.
+%is both automatic and creates labels that are project independent.
+
+% Changes to software systems, revisions, do not occur without
+% reason. 
+% The reason or purpose is often related to requirements 
+% Often one revision might have multiple purposes ranging from
+% maintenance issues, quality issues, requirements, features or a
+% response to a bug report.
+% These purposes often are exhibited as latent topics that relate to
+% many development artifacts such as:
+%  version control revisions, bug reports, and mailing-list discussions.
+
+%In our previous work we dealt with topic trends, which are topics that recur over time~\cite{Hindle09ICSM}. 
+%We observed that topic trends were often non-functional requirements (NFRs). 
+%NFRs have the property of being cross-domain and widely applicable. 
+%In this sense, they are useful abstractions for developer conversations about different software projects.
+
+% harp on data mining more
+%In general, the mining of software artifacts tends to be very project specific, yet NFRs are not. 
+% I was told if you're going to use a paren, you don't need to say it.
+%There is a series of standards on NFRs, such as ISO9126~\cite{iso9126}, specifically intended to apply to projects of varying types.
+%These standards indicate our approach is possible. %, if perhaps not universally accepted).
+% NFRs are prevalent in almost all software systems.
+% In this paper we will show that NFR related topics occur quite frequently.
+%XXXXXX this sentence is confusing
+%This informs our method of generalizing the data mining of these software artifacts to leverage the global domain knowledge of software engineering, specifically NFRs. 
+%In particular, we use software quality models to generalize topics across projects. 
+%This allows us to mine software systems with the intention of comparing their NFR-related development topics.
+% in order to achieve generalizable, cross project data mining results.
+
+%%%% THIS PARAGRAPH ARGUES ABOUT THE CONCRETE APPLICATION OF 
+%%%% TOPIC MODELLING AND LABELLED TOPIC EXTRACTION
+%%%% IT NEEDS TO BE TIGHTER BUT IT ALSO NEEDS TO BE HERE
+
+% AH: Topics take a long to read
+% AH: Project Dashboards, summarization, software process recovery
+
+%migod
+In general, the fruits of mining software artifacts are often project
+specific and hard to generalize.  
+However, in our previous work we
+investigated \emph{topic trends} --- that is, topics that recur over
+time --- we observed that topic trends often corresponded to
+non-functional requirements (NFRs)~\cite{Hindle09ICSM}, which is
+further emphasized in this paper due to the large numbers of NFR
+labelled topics.  
+This is encouraging, as NFRs have the property of being cross-domain
+and widely applicable. 
+In this sense, they are useful abstractions for developer
+conversations about different software projects.  
+Furthermore, there is a series of standards on NFRs, such as ISO9126 \cite{iso9126}, that are specifically intended to apply to projects of varying types; this suggests that our goal of trying to extract NFR-related development topics, such as those related to software quality models, holds promise.
+
+Concrete applications of topics and \emph{labelled topic extraction}
+range from project dashboards to annotating software artifacts such as revisions and bug reports with
+NFR-related tags.
+Project dashboard~\cite{dashboard} are typically employed by managers and are used to
+provide quick summaries of the effort put into a software project. In this case, 
+labelled topic extraction would allow managers to track effort
+related to NFR topics, such as portability.
+These techniques also allow for the annotation of commit comments
+ and other software artifacts with NFRs. 
+This would enable querying of bug reports and artifacts by relevant NFRs. 
+For instance a manager can confirm if their developers were
+focused on \emph{usability} by looking for
+ \emph{usability}-relevant revisions and bug reports.
+
+%% RESEARCH DIDNT DO THIS AND WE ARE TOTALLY AWESOME AND WE DID
+% it is ok to repeat this here because it is clear and bullet pointed and the reviewers wont have any bloody excuse to not see it
+In this paper, we describe \emph{automated labelled topic extraction}. It addresses two gaps in the topic mining literature:
+\begin{enumerate}
+  \item Topic mining of software has been limited to one project at a time. 
+This is because traditional topic mining techniques are specific to a particular data-set. 
+\textit{Automated labelled topic extraction} allows for comparisons \textit{between} projects. 
+  \item Topic modeling creates word lists that require interpretation by the user to assign meaning. 
+Like (1), this means
+that it is difficult to discuss results independent of the project context. 
+Our technique automatically, or with some initial training, assigns labels across projects.
+\end{enumerate}
+
+This paper makes the following contributions: 
+\begin{itemize}
+\item introduces the concept of labelled topic extraction, using a non-functional requirements (NFR) taxonomy for our labels;
+
+%\item evaluates an unsupervised method of automated topic labelling (word-lists);
+%%\item Our unsupervised method can be applied to projects without training;
+%%\item 
+%evaluates a supervised method of automatically labelling topics with a single NFR (machine learning);
+%%\item 
+%evaluates a supervised method of labelling topics with multiple NFRs
+%(multi-label machine learning);
+
+\item evaluates three kinds of automatic topic labelling methods:
+  semi-unsupervised labelling  of topics (word-lists), supervised labelling of
+  topics with a single NFR (machine learning), 
+  and supervised labelling of topics with multiple NFRs (multi-label
+  machine learning);
+
+%  of automated topic labelling (word-lists);
+%\item Our unsupervised method can be applied to projects without training;
+%\item 
+%evaluates a supervised method of automatically labelling topics with a single NFR (machine learning);
+%\item 
+%evaluates a supervised method of labelling topics with multiple NFRs (multi-label machine learning);
+
+\item provides a method of cross-project analysis via topic labelling; and
+ applies these techniques to visualize NFRs over time, and to analyze maintenance activities.
+\end{itemize}
+
+%We first introduces labelled topic extraction, and then uses that technique to analyze development activity in two large-scale examples.
+We begin by discussing related work in Section \ref{sec:related}.
+Next, we describe how we generated our data (Section \ref{sec:wordlist}).
+For semi-unsupervised classification (Section \ref{sec:unsuplabelling}), we begin by creating word-lists to signify when a topic matches an NFR label.
+We then apply our classifier and analyze the results.
+%and how we created the word-lists used in unsupervised classification.
+%We show that labels with their topics can be learned and used to classify other data-sets, either without training 
+%Our goal is to label a given topic with either an NFR from our taxonomy,
+ %Section \ref{sec:unsuplabelling} presents our technique for using word-lists to do unsupervised classification, and describes the results.
+In Section \ref{sec:suplabelling}, we manually annotate the topics, and use those annotations as training data for supervised classification.  
+%We then present visualizations of named topics and their trends over time to aid communication and analysis. 
+To demonstrate an application of labelled topic extraction, we use an exploratory case study of two open source database systems to show how named topics can be compared between projects  (Section \ref{sec:analysis}).
+The paper concludes with a discussion of limitations (Section \ref{sec:limit}), and future work.
+
+%\todo[inline]{Topics are poorly defined by this point and since we're building on research which didn't DO this, we should mention that research. Previous work probably should go here.}
+
+%\todo[inline]{We still haven't shown that topics are useful -- a reviewer comment}
+
+\section{Previous Work}
+\label{sec:related}
+
+The idea of extracting higher-level \emph{concerns} and \emph{topics}, also known as
+ \emph{concepts}, \emph{aspects} or \emph{requirements},
+has been approached from documentation-based and repository-based
+perspectives.
+
+Cleland-Huang and her colleagues have investigated mining requirements documents for non-functional requirements (NFR) (quality requirements)~\cite{Cleland-Huang2006}. 
+Their approach is similar to ours, as they mined keywords from NFR
+catalogues.
+They differ because they mine requirements documents where as we mine revisions.
+%One approach they tried was similar to this one, with keywords mined from NFR catalogues found in their previous work~\cite{chung99}. 
+They demonstrated a recall of $80\%$ with precision of $57\%$ for the \emph{security} NFR, but could not find a reliable source of keywords for other NFRs. 
+Instead, they developed a supervised classifier by using human experts to identify an NFR training set. 
+Our research differs because we use a more comprehensive set of terms based on the taxonomy we chose.
+Another difference is that we make cross-project comparisons instead of focusing
+on a single project.
+%They relied on requirements documents instead of version control
+%histories that we use.
+%The objective of Cleland-Huang's study was to identify new NFRs for system development, yet our objective was to recover those latent NFRs from commit-log messages of the project.
+
+%There are several reasons we did not follow this route. 
+%One, we believe we have a more comprehensive set of terms due to the taxonomy we chose. 
+%Secondly, we wanted to compare across projects. 
+%Their technique was not compared across different projects and the applicability of the training set to different corpora is unclear. 
+%A common taxonomy allows us to make cross-project comparison. 
+%Although subject to the assumption that all projects conceive of these terms in the same way.
+%Thirdly, while the objective of Cleland-Huang's study was to identify new NFRs (for system development) our study assumes these NFRs are latent in the commit-log messages of the project.
+
+Similarly, Mockus and Votta~\cite{Mockus00} studied a large-scale industrial change-tracking system. 
+Mockus and Votta leveraged WordNet~\cite{Fellbaum1998}, an English-language ``lexical database'' that contains semantic relations between words, including common related forms (similar to word stemming), meronymy and synonymy.
+They used WordNet for word roots as they felt the synonyms would be
+non-specific and cause errors.
+%, problems we also encountered. 
+Mockus et al. validated their labels with system developers.
+Since we study multiple projects, instead of a single project, these
+kind of interviews are somewhat infeasible (particularly in the distributed world of open-source software).
+
+Another approach is to extract concerns from software repositories.
+Marcus et al.~\cite{marcus04wcre} use Latent Semantic Indexing (LSI)
+to identify commonly occurring concerns for software maintenance. 
+The
+concerns are given by the user, and LSI is used to retrieve them from
+a corpus. 
+Topic modelling generates topics that are independent of a user query, and relate only to word frequencies in the corpus.
+%Some results were interesting, but their precision was quite low. 
+With ConcernLines, Treude et al.~\cite{treude09cl} show tag occurrence
+using colour and intensity. 
+They mine
+developer created tags 
+%change request tags (such as `milestone 3') and used these 
+in order to analyze the evolution of a single product.
+% make evolutionary analyses of a single product. 
+The presence of a well-maintained set of tags is obviously essential to the success of this technique.
+
+%XXXXXXXXXXXX OPTIONAL
+% Not sure we need this
+%Mens et al.~\cite{mens08icsm} conducted an empirical study of Eclipse, a popular Java IDE, to verify the claims of Lehman~\cite{lehman97sms}.
+% They concerned themselves with source code only, and found Law Seven, ``Declining Quality'', to be too difficult to assess: ``[we lacked an] appropriate measurement of the evolving quality of the system as perceived by the users \cite[p.
+388]{mens08icsm}''.
+%This paper examines the notions of quality in terms of a consistent ontology, as Menset al. call for in their conclusions.
+
+% in our previous work~\cite{Hindle09ICSM}, as well as -- CITED a little too often (vanity cites!)
+In Baldi et al.~\cite{Baldi2008}, topics are named manually: human
+experts read the highest-frequency members of a topic and assign a
+label accordingly. 
+%XXX Possible to cut this
+As discussed earlier, given the topic \emph{``listener change remove add fire''}, Baldi et al. would assign the label \emph{event-handling}. 
+The labels are reasonable enough, but still require an expert in the field to determine them. 
+Furthermore, these labels are project-specific, because they are generated from the data of that project.
+E.g., we might have a label called \emph{`Oracle'} in the MySQL case, since Oracle owns MySQL.
+%We do this differently.
+%This paper does this differently. 
+Our approach differs: first of all, we automate the process of naming the topics; secondly, we label topics with project-independent terms, in order to permit cross-project comparison.
+
+Mei et al.~\cite{Mei2007} use context information to automatically name topics. 
+They describe probabilistic labelling, using the frequency distribution of words in a topic to create a meaningful phrase. 
+They do not use external domain-specific information as we do, but we
+do not generate phrases from the topics.
+
+Finally, in Ernst et al.~\cite{ernst10refsq}, we describe our earlier
+project, similar to this, that identifies changes in quality requirements
+in GNOME software projects; this approach was more exploratory, had
+less validation, uses different word-lists, solely uses text-matching,
+and does not leverage machine learning strategies. 
+Hindle et al.~\cite{Hindle09ICSM} propose a windowed method of topic
+analysis that we extend with labelled
+topics, NFRs and new visualizations.
+
+%Massey~\cite{massey02icse} and Scacchi~\cite{scacchi02,scacchi05b} looked at the topic of requirements in open-source software. 
+% this reduces the line length and drops a ref even though we still have a scacchi kickin around
+%Massey~\cite{massey02icse} and Scacchi~\cite{scacchi05b} looked at the topic of requirements in open-source software. 
+%Their work discusses the source of the requirements and how they are used in the development process.
+German~\cite{german03gnome} looked at GNOME specifically, and listed several sources for requirements: leader interest, mimicry, brainstorming, and prototypes.
+%None of this work  addressed quality requirements in OSS, nor did it examine requirements trends.
+
+% XXX vanity or space.. 
+%XXXXXXXXXXXX OPTIONAL
+% NE 0113 this seems too early to mention 'result'.
+% Hindle et al.
+\cite{Hindle2007} examined release patterns in OSS.
+They showed that there is a difference between projects regarding maintenance techniques.
+This supports our result that software qualities are not discussed with the same frequency across projects.
+
+\section{Study Design and Execution}
+Figure \ref{fig:process} gives an outline of our methodology.
+We began by gathering source data and creating topic models. For semi-unsupervised labelling, we generated three sets of word-lists as signifiers for NFRs.
+%e created labels using a well-known NFR taxonomy. 
+In supervised learning, we trained our data with manual annotations in order to match topics with NFRs. Finally, these topics were used to analyze the role of NFRs in software maintenance.
+
+\begin{figure}
+  \centering
+ \includegraphics[width=.45\textwidth]{figures/process-model}
+ \caption{Research methodology process view.}
+  \label{fig:process}
+\end{figure}
+
+\subsection{Generating the Data}
+\label{sec:wordlist}
+
+To evaluate our approach, we sought candidate systems that were mature projects and had openly accessible source control repositories. 
+We selected systems from the same application domain, to control for differences in functional, rather than non-functional, requirements. 
+We used three different open-source, partially-commercial database systems:
+\begin{itemize}
+\item  MySQL 3.23.
+Started in 1994 and MySQL 3.23 was released in early 2001.
+MySQL contains $320,000$ lines of C and C++ source code\footnote{generated using David A.
+Wheeler's \emph{SLOCCount}, {http://dwheeler.com/sloccount}.}.
+\item MaxDB 7.500. Started in the late 1970s as a research project, and was later acquired by SAP. As of version 7.500, released April 2007, the project has over $940,000$ lines of C source code. 
+\item PostgreSQL 7.3. Started in the 1980s as a Berkeley research project. PostgreSQL 7.3 contains $306,000$ lines of C code.
+\end{itemize}
+
+%Choosing an older version of these projects allowed us to focus on projects which have moved further into the maintenance phase of the software life-cycle.
+We explicitly chose older versions of mature projects from a stable problem domain to increase the likelihood that we would encounter primarily maintenance activities in our studies.
+
+For each project, we used source control commit comments, the messages
+that programmers write when they commit revisions to a source control
+repository. 
+Most commits we observed had commit comments.
+Commit comments are often studied by researchers, as they are the most readily accessible source of project interactions, and developers are often required to create them by the repository mechanism (e.g., Git).
+Additionally, relying only on commit comments makes our approach more generalizable, as we do not assume the presence of other artifact corpora.
+%This is the most readily accessible source of project interactions for outside researchers, since developers often or always (e.g., Git) write such message.
+Our approach is more generalizable if it does not assume the presence of other corpora.
+%bug trackers and email-list archives were not available for both projects.
+An example of a typical commit message, from MySQL, is: \textit{``history annotate diffs bug fixed (if mysql\-\_real\-\_connect() failed there were two pointers to malloc'ed strings, with memory corruption on free(), of course)''}.
+We extracted these messages and indexed them by creation time. 
+We summarized each message as a word distribution minus stop-words such as ``\emph{the}'' and ``\emph{at}''. 
+%Stemming was performed in the later stages of our analysis. % WAS IT?
+
+For the commit message data-sets of each project, we created an XML file that separated commits into 30 day periods. 
+%This size of period, 30 days, is smaller than the time between minor
+%releases but large enough for there to be sufficient commits to
+%analyze
+We chose a period size of 30 days as it is smaller than the time between minor releases but large enough for there to be sufficient commits to analyze~\cite{Hindle09ICSM}. 
+For each 30 day period of each project, we input the messages of that period into Latent Dirichlet Allocation (LDA), a topic analysis algorithm~\cite{Blei2003}, and recorded the topics the algorithm extracted.
+%For each project and for each period of messages a project, we input these messages 
+%we created `topics' using Latent Dirichlet Allocation, a topic identification algorithm.
+
+A topic analysis tool such as LDA will try to find $N$ independent
+word distributions within the word distributions of all input messages. 
+Linear combinations of these $N$ word distributions are meant to represent and recreate the word distributions of any of the original messages. 
+These $N$ word distributions effectively form \emph{topics}: cross-cutting collections of words relevant to one or more of our commit messages. 
+LDA extracts topics in an unsupervised manner; the algorithm relies
+solely on the source data and word distributions of messages, with no human intervention.
+
+In topic analysis a single document, such as a commit message, can be related to multiple topics. 
+Representing documents as a mixture of topics maps well to source code repository commits, which often have more than one purpose~\cite{Hindle09ICSM}.  
+For this paper, a topic represents a word distribution generated from a group of commit log comments which are related by their content.  
+%In this paper a topic is a set of tokens extracted from commit messages found within a project's source control system (SCS).
+
+We applied Blei's LDA implementation~\cite{Blei2003} against the word distributions of these commits, and generated lists of topics per period. 
+We set the number of topics to generate to $20$, because past experimentation showed that fewer topics might aggregate multiple unique topics while any more topics dilutes the results and creates indistinct topics~\cite{Hindle09ICSM}.
+% Furthermore, more than $20$ topics quickly became infeasible for inspection.
+
+\subsubsection{The High-level Labels}
+
+%\noindent \textbf{The high-level labels} ---
+
+To facilitate cross-project comparison, we used a taxonomy of NFRs. This taxonomy is based on the ISO quality model, ISO9126~\cite{iso9126}. 
+ISO9126 describes six high-level NFRs: maintainability, functionality,
+portability, efficiency, usability, and reliability~\footnote{While there may be lingering debate in some circles about these terms, an ISO standard seems like a reasonable starting point for our work.}.
+%There is some debate about the terms in this model. 
+%However, ISO9126 is ``an international standard and thus provides an internationally accepted terminology for software quality \cite[p.
+58]{Boegh2008},'' that is sufficient for the purposes of this research.
+We claim that these NFRs are maintenance concerns (to varying degrees) in all software projects, and are therefore well-suited for comparisons between projects.
+
+%\noindent \textbf{Creating a validation corpus} ---
+
+%FAT
+%\newpage
+
+\subsubsection{Creating a Validation Corpus}
+To evaluate both semi-unsupervised and supervised classification, we
+created a validation set of manually labelled topics. For MySQL 3.23
+and MaxDB 7.500, the annotators (the first two authors) annotated each extracted topic in each period with
+the six NFR labels listed above.
+Annotators did not annotate each other's annotations, but some brief
+inspection of annotations was used to confirm that the annotators were
+acting similarly.
+We looked at each period's topics, and assessed what the data ---
+consisting of the frequency-weighted word lists and messages ---
+suggested was the label for that topic. 
+We were able to pinpoint the appropriate label using auxiliary information as well, such as the actual revisions and files that were related to the topic being annotated.
+For example, for the MaxDB topic consisting of a message ``exit() only
+used in non NPTL LINUX Versions'', we tagged that topic
+\emph{portability}. 
+Given the top-level annotations of \emph{none}, \emph{portability},
+\emph{efficiency}, \emph{reliability}, \emph{functionality},
+\emph{usability}, and \emph{maintainability}, the annotators annotated each topic
+with the relevant label. Sometimes they used finer-grained
+annotations that would be aggregated up to one of these higher-level labels.
+%\todo[inline]{this confused people. Is it truly unsupervised? Yes, but it needs to be clearer.}
+%We compared against this data-set, but we also used the data-set for our supervised machine learning based topic classification. 
+% A topic may have more than one matching keyword. % ok what?
+%\todo[inline]{inter-rater reliability }
+% \todo[inline]{There should be a section here which is about manual annotation. It is really important and it is required by both analyses. Plus we can tell them how.}
+
+%\todo[inline]{NEIL: this section should be updated to explain F-measure}
+We validate classification performance using the Receiver Operating
+Characteristic area-under-curve value~\cite{Fawcett2006861},
+abbreviated \emph{ROC}, and the F-measure, which is the harmonic mean of precision and recall, i.e., $2 * (P * R) / (P + R)$.
+
+%\todo[inline]{NEIL: abram, i got rid of the letter-grade thing because I'm worried people latch on to the idea we are C- students. I feel anything better than random is our criteria here}
+ROC values provide a score %, similar to school letter-grades (A is 0.9, C is 0.6),
+ reflecting how well a particular learner performed for the given data. 
+ROC maps to the more familiar concepts of precision/sensitivity and recall/specificity: it plots the true positive rate (sensitivity) versus the false positive rate (1 - specificity). 
+A perfect learner has a ROC value of 1.0, reflecting perfect recall and precision. 
+A ROC result of 0.5 would be equivalent to a random learner (that is, issuing as many false positives as true positives). 
+The %area under the -- NE confusing to have area again ... and not importantana
+ROC of a classifier is equivalent to the probability that the classifier will rank a randomly chosen positive instance higher than a randomly chosen negative instance.
+We consider our labelling classifiers acceptable if they outperform a random classifier (0.5).
+
+\subsection{Semi-unsupervised Labelling}
+\label{sec:unsuplabelling}
+
+In this section we describe how to label topics based on dictionaries
+mined from sources external to the projects.
+
+%\subsection{Generating word lists}
+%\label{sec:genwordlist}
+% %\todo[inline]{Make it clear where the data comes from! Neil says: Also, I
+%   didn't get the impression they understood that we weren't generating
+%   labels from the data, but rather from independent taxonomies. Of
+%   course if you use labels from the data it isn't comparable. But the
+%   point of ISO 9126 etc. was to avoid exactly this problem.}
+
+% \todo[inline]{Need to hammer it in that the word bags are not project specific and
+%  are inherited so it isn't a supervised method}
+
+%Topics are word distributions, lists of words ranked by frequency, that are burdensome to interpret and hard to distinguish and understand. 
+%While the topics themselves are generated automatically, how to interpret these topics is less clear.
+%For example,
+
+\subsubsection{Generating Word Lists}
+
+%\noindent\textbf{Generating word lists} ---
+
+In order to automatically label each topic with 
+one of the six high-level NFRs,
+%an NFR from ISO9126, 
+we associate each NFR with a list
+of keywords (\emph{word-lists}, in our parlance). These word-lists were determined a priori and were not extracted from the projects themselves.
+We intersected the words of the topics and the words of our word-lists.
+We ``labelled'' a topic if any of its words matched any of the word-list's words.
+A topic could match more than one NFR.
+We used several different sets of word-lists for comparison, which we
+refer to as \textsf{exp1}, \textsf{exp2}, and \textsf{exp3} in the text which follows.
+
+Our first word-list set, \textsf{exp1}, was generated using the ontology described in Kayed et al.~\cite{5072519} .
+That paper constructs an ontology for software quality measurement using eighty source documents, including research papers and international standards. 
+The labels we used:
+
+%FAT we can remove this
+%\balance
+%LFAT
+\newpage
+\begin{quotation}
+\small \noindent \textsf{
+integrity, security,
+interoperability, testability, maintainability, traceability,
+accuracy, modifiability, understandability, availability, modularity,
+usability, correctness, performance, verifiability, efficiency,
+portability, flexibility, reliability.
+}
+\end{quotation}
+
+Our second word-list, \textsf{exp2}, uses the ISO9126 taxonomy described above (Section \ref{sec:wordlist}) to seed the word-lists.
+%We use these qualities later on as classes in supervised labelling. NE confusing the issue here, I think.
+The terms from ISO9126 may not capture all words occurring in the topics that are nonetheless associated with one of the NFRs. 
+For example, the term ``redundancy'' is one 
+we considered to be
+relevant to discussion of \emph{reliability}, but is not in the standard. 
+We therefore took the NFRs from the ISO9126 and added to them.
+
+%\todo[inline]{* clarify the mailinglist stuff we did}
+%\todo[inline]{This part doesn't make a lot of sense}
+To construct these expanded word-lists, we used
+WordNet~\cite{Fellbaum1998}.
+%, an English-language ``lexical database'' that contains semantic relations between words, including common related forms (similar to word stemming), meronymy and synonymy. 
+We then added Boehm's software quality model~\cite{Boehm+:1976:ICSE}, and classified his eleven `ilities' into their respective ISO9126 NFRs. 
+We did the same for the quality model produced by McCall et al.~\cite{mccall1977}. 
+We then did a simple random analysis of mailing list messages from an open source project, KDE.
+If we judged a given message to contain terms that were related to one of the NFRs in ISO9126, we added it to our word-list.
+This allowed us to expand our word-lists with more software-specific terms.
+%Finally, we randomly analyzed two mailing lists from another software project to expand our set with domain-specific terms. 
+% Less info is probably better here...
+%For example, we add the term ``performance'' to the synonyms for \emph{efficiency}, since this term occurs in most mail messages that discuss efficiency.
+Table \ref{tbl:wnsig} shows the labels (NFRs) and word-lists we used for matching.
+
+For the third set of word-lists, \textsf{exp3}, we extended the word-lists from \textsf{exp2} using unfiltered WordNet similarity matches. 
+Similarity in WordNet means siblings in a hypernym tree. 
+We do not include these words here for space considerations (but see the Appendix for our data repository). 
+It is not clear the words associated with our labels in \textsf{exp3} are specific enough.
+For example, the label \emph{maintainability} is associated with words \emph{ease} and \emph{ownership}.
+In general, as we proceed from word-list in \textsf{exp1} to that in \textsf{exp3}, our lists become more generic.
+
+\begin{table*}
+	\centering
+\begin{tabular}{c|p{11cm}}
+\toprule
+\textbf{Label} & \textbf{Related terms} \\
+\midrule
+\emph{Maintainability} &
+testability changeability analyzability stability maintain maintainable modularity modifiability understandability interdependent dependency encapsulation decentralized modular\\ \hline
+\emph{Functionality} &
+security compliance accuracy interoperability suitability functional practicality functionality compliant exploit certificate secured ``buffer overflow'' policy malicious trustworthy vulnerable vulnerability accurate secure vulnerability correctness accuracy\\ \hline
+\emph{Portability} &
+conformance adaptability replaceability installability portable movableness movability portability specification migration standardized l10n localization i18n internationalization documentation interoperability transferability\\ \hline
+\emph{Efficiency} &
+``resource behaviour'' ``time behaviour'' efficient efficiency performance profiled optimize sluggish factor penalty slower faster slow fast optimization\\ \hline
+\emph{Usability} &
+operability understandability learnability useable usable serviceable usefulness utility useableness usableness serviceableness serviceability usability gui accessibility menu configure convention standard feature focus ui mouse icons ugly dialog guidelines click default human convention friendly user screen interface flexibility\\ \hline
+\emph{Reliability} &
+``fault tolerance'' recoverability maturity reliable dependable responsibleness responsibility reliableness reliability dependableness dependability resilience integrity stability stable crash bug fails redundancy error failure\\
+\bottomrule
+\end{tabular}
+	\caption{NFRs and associated word-lists -- \textsf{exp2}}
+	\label{tbl:wnsig}
+
+\end{table*}
+
+\subsubsection{Automatic Labelled Topic Extraction}
+
+%\todo[inline]{NEIL: this could be a subsubsection}
+%\todo[inline]{* the hierarchy of tags was not clear}
+%\noindent \textbf{Generating labels automatically} ---
+
+Using our three word-lists (\textsf{exp1}, \textsf{exp2}, \textsf{exp3}), we labelled our topics with an NFR where there was a match between a word in the list and the same word somewhere in the distribution of words that constitute the topic.
+A \emph{named topic} is a topic with a matching NFR label. \emph{Unnamed
+topics} occur where there is no such match, which may indicate either a lack of precision, or simply that this topic is not associated with non-functional requirements.
+All experiments were run on MaxDB 7.500 and MySQL 3.23 data. LDA
+extracted $20$ topics per period for each project.
+% Is this good enough
+This labelling is \emph{semi-unsupervised} because the corpus is not derived from 
+the project being analyzed, and we did not label the project's topics
+ourselves for a training set. The motivation behind this technique is that
+because most software often addresses similar issues, thus we can use the
+domain knowledge of software to label relevant topics.
+%\todo[inline]{Neil does this explain the unsupervised aspect well enough?}
+
+Table \ref{tbl:wordlist} shows how many topics were labelled for MaxDB
+and MySQL.
+
+%\todo[inline,color=green]{this table should have mysql}
+\begin{table}
+
+% % This data came from running  
+% %   python check_validity.py
+% %  in src/validate
+% % and then running
+% % tail -n 5 *maxdb*yes_no*
+% % tail -n 5 *mysql*yes_no*
+% ok the new way is to run lda_relator
+% src/validate/lda_relator_counts.sh
+	\centering
+\begin{tabular}{l|c|c|c|c}
+\toprule
+\textbf{Project} & \textbf{Measure} & \textsf{exp1} & \textsf{exp2} & \textsf{exp3} \\
+\midrule
+% NOTE THESE WERE THE OLD VALUES
+% WHY DID THEY CHANGE???
+% DUNNO
+%Named topics   & 281 & 125 & 328  \\
+%Unnamed topics & 139 & 295 & 92   \\
+
+% these were check_validate
+% MaxDB 7.500 & Named Topics   & 305 & 183 & 330 \\
+% MaxDB 7.500 & Unnamed Topics & 84  & 206 & 59   \\
+% MaxDB 7.500 & Total  Topics  & 389 & 389 & 389 \\
+% MySQL 3.23  & Named Topics   & 341 & 202 & 469 \\
+% MySQL 3.23  & Unnamed Topics & 245 & 384 & 117 \\
+% MySQL 3.23  & Total  Topics  & 586 & 586 & 586 \\
+
+% these are lda relator
+
+MaxDB 7.500 & Named Topics   & 281 & 125 & 328 \\ % these numbers are
+                                % slightly different solely because
+                                % we're counting empty topics
+                                % maxdb has 80 empty topics
+ & Unnamed Topics & 219  & 375 &  172  \\
+ & Total  Topics  & 500 & 500 & 500 \\
+\midrule
+MySQL 3.23  & Named Topics   & 524 & 273 & 773 \\
+  & Unnamed Topics & 476 & 727 & 227 \\
+  & Total  Topics  & 1000 & 1000 & 1000 \\
+
+\bottomrule
+\end{tabular}
+	\caption{Automatic topic labelling for MaxDB and MySQL}% \todo[inline]{this stupid table is before table 1}}
+	\label{tbl:wordlist}
+
+\end{table}
+
+% was best performing
+For \textsf{exp1} the labels with the most topics
+were
+%labels, those with the most topics, 
+\emph{correctness} (182 topics) and \emph{testability} (121). 
+We did not see many results for \emph{usability} or \emph{accuracy}, which were associated with fewer than ten topics. 
+We also looked for correlations between our labels: excluding double matches (self-correlation), our highest co-occurring terms were \emph{verifiability} with \emph{traceability}, and \emph{testability} with \emph{correctness} (76 and 62 matches, respectively).
+% See the following sections for our error analysis.
+
+%LFAT
+\balance
+\newpage
+
+For \textsf{exp2}, there are more unnamed topics than \textsf{exp1}. 
+Only \emph{reliability} produces a lot of matches, mostly with the word ``error''. 
+Co-occurrence results were poor. This suggests our word lists were overly restrictive.
+
+For \textsf{exp3}, we generally labelled more topics. 
+As we mentioned, the word-lists are broad, so there are likely to be false-positives (discussed below). 
+The most frequent label, 265 topics, was \emph{usability}, and the
+least frequent label, 44 topics, was \emph{maintainability}. 
+Common co-occurrences were \emph{reliability} with \emph{usability}, \emph{efficiency} with \emph{reliability}, and \emph{efficiency} with \emph{usability} (200, 190, and 150 topics in common, respectively).
+
+%\todo[inline]{Should we use F-1 as well to satisfy everyone?}
+
+\subsubsection{Analysis of the Semi-unsupervised Labelling} %--- %Based
+                                %on the labels, and our manual topic
+                                %labelling, 
+%\noindent \textbf{Analysis of the unsupervised labelling} --- %Based on the labels, and our manual topic labelling, 
+%we compared the results of the unsupervised word matching approach. 
+For each quality we assessed whether semi-unsupervised labels matched the manual annotations. %To be clear, the manual annotations were not used to train the labelling process.
+As described in Section \ref{sec:wordlist} we used both ROC and F-1 measures to evaluate the performance of the classification.
+Figure \ref{fig:maxdb-unsup-results} shows our ROC results for MaxDB and MySQL. We describe F-1 results in the text below.
+
+%\todo[inline]{since ROC is new-sounding, I wonder if a more lengthy explanation would be merited. An obvious question is why not precision/recall? -- ROC is hot right now and it combines both}
+% Rebuttal: ROC is not new and it is what the machine learning community, search based software engineering and now the MSR community are using +  it has natural meaningful quality mappings.
+% \begin{figure}[t]
+% \centering
+% \subfloat[ROC]{
+%  \includegraphics[width=0.45\textwidth]{figures/unsupervised-bar}
+% \label{fig:maxdb-unsup-results-roc}
+% }
+% %\subfloat[F1]{
+% %\includegraphics[width=0.45\textwidth]{figures/unsupervised-bar-f1}
+% %\label{fig:maxdb-unsup-results-f1}
+% %}
+% \caption[]{Performance of unsupervised topic labelling for each NFR per word-list.
+A) ROC values -- dashed line indicates the performance of a random classifier.
+B) F1 measure.
+Possible values range from 0--1.
+% %\todo[inline]{better discussion}% and graphical legend would be nice}
+% }
+% \label{fig:maxdb-unsup-results}
+% \end{figure}
+
+\begin{figure*}[t]
+  \centering
+ \includegraphics[width=0.8\textwidth]{figures/unsupervised-bar}
+ \caption{Performance, ROC values (range: $0$--$1$), of semi-unsupervised topic labelling for
+   each NFR and per word-list. The dashed line indicates the performance of a random classifier. This graph shows how well the
+   semi-unsupervised topic labelling matched our manual annotations.}
+
+%\todo[inline]{this is too tiny in print}
+  \label{fig:maxdb-unsup-results}
+\end{figure*}
+
+% \begin{figure}
+%   \centering
+%  \includegraphics[width=0.45\textwidth]{figures/unsupervised-bar-f1}
+%  \caption{Performance, F1 values (range: 0-1), of unsupervised topic labelling for
+%    each NFR and per word-list.}
+%   \label{fig:maxdb-unsup-results-f1}
+% \end{figure}
+
+%\todo[inline]{where is exp1 report?}
+
+%EXP1 could not be evaluated properly!
+Because our ground truth annotations were relevant only to ISO9126,
+we estimate that \textsf{exp1} had poor
+performance via the overlap between ISO9126 and the Kayed ontology.
+For \textsf{exp1} the F-measures for MaxDB were from $0$ to $0.18$ with an average
+of $0.03$, and for MySQL were from $0$ to $0.16$ with an average of
+$0.05$.
+% 0.177215189873 / 6
+% .02953586497883333333
+% (0.0298507462687 + 0.0661157024793 + 0.0634920634921 + 0 + 0.115384615385) / 6
+% .04580718793751666666
+
+% EXP2 source is naming-paper/src/validate/exp2.maxdb.report.txt
+% (change as needed). Second column (after the name) is F1.
+For \textsf{exp2}, the average F-measure (macro-F1) for MaxDB was $0.24$ with a range $0.091$ to
+$0.37$, and $0.16$ for MySQL with a range of $0$ to $0.41$.
+MaxDB had an average precision and recall of $0.25$ and $0.22$
+while MySQL had $0.41$ and $0.10$ respectively.
+
+% EXP3
+For \textsf{exp3}, the average F-measure (macro-F1) for MaxDB was $0.26$ with a range $0.11$ to
+$0.47$, and $0.36$ for MySQL with a range of $0.10$ to $0.65$.
+MaxDB had an average precision and recall of $0.16$ and $0.67$
+while MySQL had $0.3$ and $0.48$ respectively.
+
+Thus we find that \emph{reliability} and
+\emph{usability} worked well for MaxDB in \textsf{exp2} and better in
+\textsf{exp3}. 
+\textsf{exp1} performed poorly.
+MySQL had reasonable results within \textsf{exp2} for \emph{reliability} and \emph{efficiency}. 
+MySQL's results for \emph{efficiency} did not improve in \textsf{exp3}
+but other qualities such as \emph{functionality} did improve. 
+Many ROC scores were $0.6$ or less, but our classifier still performed substantially better than random.
+
+%\vspace*{-1em}
+\subsection{Supervised Labelling}
+\label{sec:suplabelling}
+Supervised labelling requires expert analysis of the correct
+class/label to assign a label to a topic.
+In our approach, we use the top-level NFRs in the ISO9126 standard~\cite{iso9126} for our classes, but other taxonomies are also applicable.%In order to validate how effective these word-bag approaches to topic labelling would be we created a validation data set.
+
+We used a suite of supervised classifiers, WEKA~\cite{weka09},
+that includes machine learning tools such as support vector machines and Bayes-nets. 
+We also used the multi-labelling add-on for WEKA, Mulan~\cite{mulan}. %\footnote{\url{http://mlkd.csd.auth.gr/multilabel.html}}. 
+Traditional classifiers label topics with a single class, whereas Mulan allows for a mixture of classes per topic, which is what we observed while manually labelling topics.
+The \emph{features} we used are word counts/occurrence  per topic, if
+a word occurs frequently enough in a topic we consider it a feature of
+the topic.
+
+To assess the performance of the supervised learners, we did a 10-fold cross-validation~\cite{Kohavi1995}, a common technique for evaluating machine learners. 
+The original data is partitioned randomly into ten sub-samples. Each sample is used to test against a training set composed of the nine other samples.
+%One subsample is used as a training set and evaluated on the other nine sub-samples. 
+%This is repeated nine more times, with each subsample used once as the training set. 
+We have reported these results below.% in Section \ref{sec:suplabelling}.
+
+\subsubsection{Analysis of the Supervised Labelling}
+% \label{sec:suplabelling}
+Because our data-set was of word counts we expected Bayesian techniques, often used in spam filtering, to perform well. 
+We tried other learners that WEKA~\cite{weka09} provides: rule learners, decision tree learners, vector space learners, and support vector machines.  
+Figure \ref{fig:best-learn-per-tag} shows the performance of the best
+performing learner per label: 
+% We considered the best learner for a label to be 
+the learner that had the highest ROC value for that label. 
+The best learner is important because one uses a single learner per
+label. When applying our technique, for each NFR one should select the best learner possible.
+%Figure \ref{fig:best-learn-per-tag} uses the ZeroR learner as a baseline, a common practice in machine learning,
+%as ZeroR naively chooses the largest category all of the time. 
+%ZeroR occasionally outperforms other learners; for labels which are not as common, this is to be expected because any miscategorization will hurt accuracy. 
+%This is why we use ROC values, instead of accuracy, as they can better represent performance on labels which are not applicable to the majority of samples.
+
+Figure \ref{fig:best-learn-per-tag} shows that MaxDB and MySQL have quite different results, as the ROC values for \emph{reliability} and \emph{functionality} seem swapped between projects.
+
+\begin{figure}[t]
+\centering
+% \subfloat[MySQL]{
+\includegraphics[width=0.45\textwidth]{figures/both-supervised}
+% \label{fig:subfig1}
+% }
+% \subfloat[MaxDB]{
+% \includegraphics[width=0.4\textwidth]{figures/maxdb-supervised}
+% \label{fig:subfig2}
+% }
+\caption[]{ROC value for the best learner per label for MaxDB and MySQL. Values range from $0$--$1$.  Dashed line indicates the performance of a random classifier.
+%\todo[inline]{better discussion}
+}
+\label{fig:best-learn-per-tag}
+\end{figure}
+
+For both projects Bayesian techniques did the best out of a wide variety of machine learners tested. 
+Our best learners, Discriminative Multinomial Naive Bayes, Naive Bayes
+and Multinomial Naive Bayes  are all based on Bayes's theorem and all
+assume, naively, that the features supplied are independent. 
+%The features we used are word counts per message. 
+One beneficial aspect of this result is that it suggests we can have very fast training and classifying  since Naive Bayes can be calculated in $O(N)$ for $N$ features.
+%http://nlp.stanford.edu/IR-book/html/htmledition/naive-bayes-text-classification-1.html
+
+% # F measure
+% mysql
+% > v <- c( 0.64 , 0.23  ,0.41 , 0.77  ,0.62 , 0.21 ) 
+% > summary(v)
+%    Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
+%   0.210   0.275   0.515   0.480   0.635   0.770 
+% maxdb
+% > v <- c(0.61, 0.25, 0.59, 0.32, 0.42, 0.17)
+% > summary(v)
+%    Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
+%  0.1700  0.2675  0.3700  0.3933  0.5475  0.6100
+
+The range of F-measures for MySQL was $0.21$ to $0.77$ with an average
+of $0.48$, while MaxDB had a range of $0.17$ to $0.61$ with an average
+of $0.39$.
+
+The less-frequently occurring a label, the harder it is to get accurate
+results, due to the high noise level. Nevertheless, these results are
+better than our previous word-list results of \textsf{exp2} and
+\textsf{exp3}, because the ROC values are sufficiently higher in most
+cases (other than MaxDB \emph{reliability} and MySQL \emph{efficiency}). The
+limitation of the approach we took here is that we assume labels are
+independent; however, labels could be correlated with each other. 
+The next section (\ref{sec:multilabel})
+addresses the issue of a lack of independence and correlation between
+labels
+using multi-label learners.
+%In the next section we will evaluate how well these learners perform
+%together.
+%\todo[inline]{Comment about correlation, we should link to the Mulan section, perfect segway since Mulan relies on correlation.}
+
+%\balance
+
+\subsection{Applying Multiple Labels to Topics}
+\label{sec:multilabel}
+
+As noted in Section \ref{sec:wordlist}, each topic in our data-set can be composed of zero, one, or more NFRs. 
+For example, a commit message might address \textit{reliability} in the context of \textit{efficiency}, or make a \textit{maintainability} improvement in the source code that related to \textit{usability}.
+However, traditional machine learning techniques, such as Naive Bayes, can only map topics to a single class. 
+The Mulan~\cite{mulan} library encapsulates several different multi-label machine learners which can label elements with multiple labels.
+Mulan also includes methods for determining the performance of such techniques.
+% TOO DETAILED The problem framed in the learners above has changed; instead of looking at the precision and recall of applying one label, we rank multiple labels at once.
+We must check if the full subset of labels was applied, and then how much of that subset was applied.
+
+\begin{figure*}[t]
+\centering
+\subfloat[MySQL]{
+\includegraphics[width=0.3\textwidth]{figures/multi-label-results-mysql}
+\label{fig:subfig3}
+}
+\subfloat[MaxDB]{
+\includegraphics[width=0.3\textwidth]{figures/multi-label-results-maxdb}
+\label{fig:subfig4}
+}
+\caption[]{MySQL and MaxDB macro (grey) and micro (black) ROC results per multi-label learner. Possible values range from $0$--$1$.  Dashed line indicates the performance of a random classifier.
+%\todo[inline]{better discussion}% and graphical legend would be nice}
+}
+\label{fig:mulan}
+\end{figure*}
+
+%FAT 
+%\newpage
+
+Two perspectives to evaluate multi-label learners are with micro or macro measurements (used in Figure \ref{fig:mulan}).
+Macro measurements are aggregated at a class or label level (per class) while micro measurements are at the element level (per element).
+A macro-ROC measurement is the average ROC over the ROC values for all labels, where a micro-ROC is the average ROC over all examples that were classified. 
+For MaxDB, the macro-ROC values are undefined because of poor performance of one of the labels.%, see Figure \ref{fig:mulan}.
+
+Figure~\ref{fig:mulan} presents the results of Mulan's best multi-label learners for our data. 
+Calibrated Label Ranking (CLR) is a learner that builds two layers. The first layer determines if an entity should be labelled, while the second layer determines what labels should be assigned.
+The Hierarchy Of Multi-label classifiERs (HOMER) and Binary Relevance (BR) act as a hierarchy of learners: BR is flat, while HOMER tries to build a deeper hierarchy for a more accurate learner~\cite{mulan}.
+These classifiers performed better than other multi-label classifiers as they have the best micro and macro ROC scores. 
+The multi-label and single-label learners had similar performance: for
+MySQL, BR and Naive Bayes had similar macro-ROC scores of $0.74$.
+% although their results are comparable to the naive Bayesian learners we used in Section \ref{sec:suplabelling}.
+% \todo[inline]{discuss this similarity in more detail}
+
+\section{Understanding Software Maintenance Activities} 
+\label{sec:analysis}
+% \todo[inline]{talk about this topic in one section}
+% \todo[inline]{reference project name Nomen more}
+%In order to test  our ROC values produced acceptable results, we now turn to
+%\emph{applying} those techniques to understand software maintenance
+%activities. 
+%\todo[inline]{more on intro ... why should the reader care about
+ % this. It needs to show how the stuff in the intro (applications) is
+ % possible}
+%%% XXX NEIL I don't get this.. VVVVV
+%\todo[inline]{fix section headings}
+As we mentioned in the introduction, a key issue in software maintenance is understanding \emph{why} a system has evolved the way it has~\cite{aranda09icse}. 
+In this section we demonstrate the value of labelled topic extraction
+in addressing this issue. 
+% more why
+Labelled topics address \emph{why}  because they show reasons why changes
+occurred rather than \emph{how}. 
+The \emph{how} of a change is the change itself, the purpose or
+\emph{why} is what we are after.
+We investigate the history of the two large-scale database systems that we studied. 
+We use our technique to show the topic of development efforts over time in each project.
+%We evaluated two research questions using the ground truth data that we annotated by hand:
+We motivated our investigation with three research questions:
+\begin{enumerate}
+\item \emph{Do NFR frequencies change over time?} If a particular NFR
+  was of more interest at one point in the life-cycle than another,
+  this suggests that development activity shifted focus. For example,
+  if a developer expected to see a recent focus on \emph{reliability},
+  but instead \emph{usability} dominated, they might re-prioritize
+  upcoming work items.
+
+\item \emph{Do projects differ in their relative interest in NFRs?} A
+  project manager, especially a systems-manager, would be interested
+  in knowing whether a particular NFR, such as \emph{reliability}, was
+  more important for one project than another. This could be to
+  confirm the initial design goals, or to track the progress on that
+  quarter's objectives.  The difference in NFR proportion is
+  interesting because it implies a difference in focus between two projects.
+
+\item \emph{Do different developers work on different NFRs?} For a given project, it is reasonable to think that developers are either assigned (in commercial organizations) or choose (in open-source organizations) to a particular NFR.
+For example, one developer might be more junior, and take responsibility for the low-impact reliability fixes.
+Another, more senior developer might assume responsibility for major improvements such as efficiency improvements.
+
+\end{enumerate}
+
+%FAT
+
+%\newpage
+
+%\subsection{Timeline analysis}
+Figures \ref{fig:mysql-timeline} and \ref{fig:maxdb-timeline} show the
+temporal patterns of NFR frequencies.
+%NEIL- I think this (below) is just confusing, although maybe important to note.
+% based on the manually annotated
+%topics, although this visualization can be generated from the
+%results of labelled topic extraction.
+There are two measures represented. 
+One, the relative frequency, shown in the grey histogram boxes, represents the number of topics with that NFR in that period, 
+relative to the maximum number of topics assigned to that NFR. 
+For example, in Figure \ref{fig:mysql-timeline} we see a spike in \emph{portability} and \emph{functionality} frequency in September 2002.
+The second, absolute frequency, is shown using cell intensity, and compares the number of topics labelled with that NFR per period 
+relative to the maximum number of labelled topics overall. 
+For instance, Figure \ref{fig:mysql-timeline} shows that the NFRs
+\emph{functionality}, \emph{portability} and \emph{maintainability}
+contain more labelled topics, since these NFRs have been more
+intensely shaded; 
+one interesting stream is \emph{efficiency} which shows periodic
+activity, does this suggest that once maybe efficiency related changes
+have longer lasting effects?
+The topmost row in each diagram lists historical events for that project (such as a release). 
+%We refer to these occurrences in the discussion which follows.
+
+\begin{figure*}[t]
+\centering
+\subfloat[MySQL 3.23]{
+\includegraphics[width=\textwidth]{figures/mysql-timeline.png}
+\label{fig:mysql-timeline}                               
+}
+
+\subfloat[MaxDB 7.500]{                                 
+\includegraphics[width=\textwidth]{figures/maxdb-timeline.png}
+\label{fig:maxdb-timeline}
+}
+	\caption[]{NFR label per period. Each cell represents a 30-day period. %Numbers represent number of topics in that period labelled with that NFR. 
+	Grid cell intensity (saturation) is mapped to label frequency
+        relative to the largest label count of \emph{all} NFRs. Grey
+        histogram bars indicate label frequency relative to that
+        particular NFR's largest label count. Dashed vertical lines
+        relate a project milestone (\emph{*Key events*}) to our topic windows. 
+% \todo[inline]{better discussion}
+}
+\label{fig:timelines}
+\end{figure*}
+
+%\noindent \textbf{Understanding the figures} -- 
+ % boxes. not sure boxes is important..
+% I grepped query1.mrs2 http://softwareprocess.es/y/y/maxdb-76.tar.gz
+% and                   http://softwareprocess.es/y/maxdb-7500.tar.gz
+%There are more \textsf{None} labels in MaxDB because a number of topics were to do with code cleanup or automated checking using tools like \texttt{sutcheck}, a tool specific to the development process of MaxDB.
+
+%\noindent \textbf{Timeline of key events} -- 
+We analyzed each project's developer mailing list for external validation. 
+%As discussed earlier, the top row in each figure shows key events for each project. 
+We use \textit{labelled topic extraction} to pick out the underlying NFR activity behind these events. 
+For example, both projects show a high number of NFRs recognized at the first period of analysis. 
+This is due to our window choice: we deliberately targeted our
+analysis to when both MySQL 3.23
+and MaxDB 7.500 were first announced. For MaxDB, version 7.5.00  was released in December of 2003. 
+We know that release 7.5.00.23 saw the development of PHP interfaces, possibly accounting for the simultaneous increase in the \emph{portability} NFR at the same time.
+The gap in MaxDB (Figure \ref{fig:maxdb-timeline}) is due to a shift in development focus (from February 2005 to June 2005) to MaxDB 7.6, which is released in June 2005.
+
+%For MySQL (Figure \ref{fig:mysql-timeline}), we similarly validated our NFR patterns with external mailing list data. 
+The release of MySQL we study  (Figure \ref{fig:mysql-timeline}) was the first to be licenced under the GPL. 
+Version 3.23.31 (January, 2001) was the production release (non-beta), and we see a flurry of topics labelled with \emph{functionality} and \emph{maintainability}. 
+After this point, this version enters the maintenance phase of its life-cycle. 
+In May 2001, there is an increase in the number of topics labelled with \emph{portability}. 
+This might be related to release 3.23.38, which focused on Windows compatibility. 
+Similarly, in August, 2002, both \emph{functionality} and \emph{portability} are frequent, and mailing list data suggests this is related to the release of version 3.23.52, a general bug fix with a focus on security (a component of the \emph{functionality} NFR in the ISO9126 model).
+After this point, efforts shift to the newer releases (4.0, 4.1,
+5.0). 
+\todo{add PGSQL disucssion here. Schema additions? Lots of functionality?}
+
+We now address our research questions:
+
+% \subsection{Comparing MaxDB and MySQL}
+
+% \label{sec:comparison}
+%\XXX{Not sure there's enough here to do that!}
+% We observed that MySQL had more topic subsets than MaxDB. 
+% MySQL 3.23 also had more topics and a longer recorded history than MaxDB 7.500. 
+% We tagged more MySQL topics with annotations than MaxDB topics, yet both shared similarities. 
+% In terms of non-functional requirements both projects had long running trends that focused on \emph{functionality}, \emph{maintainability}, and \emph{portability}, yet MaxDB had more of a focus on \emph{efficiency} and \emph{reliability}.
+% \todo[inline]{expand}
+%\todo[inline,color=yellow]{expand on this with dates, and reference to visualizations}
+%\subsection{Research questions}
+%This analysis allows to address the questions raised in Section \ref{sec:suplabelling}:
+%This analysis addresses our research questions: %raised in Section \ref{sec:analysis}:
+
+\subsection{Do NFR Frequencies Change over Time?}
+%\noindent\textbf{1. Do NFR frequencies change over time?} --- 
+%\subsubsection{Do label frequencies change over time?}
+\todo{add more on PGSQL}
+In both projects the frequencies generally decreased with age. 
+However, there are variations within our NFR labels. In MySQL, \emph{usability} and \emph{efficiency} do not appear very often in topics. 
+A proportionately smaller number of commits addressed these NFRs.
+%As discussed in the previous section, 
+Certain peaks in topic numbers coincide with a particular emphasis from the development team on issues such as new releases or bug fixes.
+This suggests that maintenance activity is not necessarily strictly decreasing with time, but rather episodic and responsive to outside stimuli. 
+In MaxDB, we can observe that \emph{Maintainability} topics became more prevalent as MaxDB matures. 
+This is likely due to our analysis time-frame for MaxDB being shorter than the time-frame for the MySQL product.
+
+%LFAT
+\balance
+\newpage
+\subsection{Do Projects Differ in their Relative  Topic Interest?}
+%\subsubsection{Do  projects differ in their relative  topic interest?}
+%\noindent\textbf{2. Do  projects differ in their relative  topic interest?} ---
+Yes. MySQL 3.23 had proportionally more
+topics labelled \emph{functionality}, while MaxDB had proportionally more
+\emph{efficiency} related topics. MaxDB was a very mature release ``donated'' to the open-source community, 
+whereas MySQL was in its relative infancy, and  
+security problems were more common (security is a component of \emph{functionality} in the ISO9126 model). 
+In both cases \emph{portability} was a constant maintenance concern and was prevalent throughout the lifetime of the projects. It may surprise developers how often portability arises as a concern.
+%On one hand, this analysis is unsurprising; on the other hand, it provides confirmation that our approach identifies actual patterns.
+%\subsection{Summary}
+
+%\todo[inline]{link the actual data, which is interesting, to somethign
+ % actionable for a developer or manager}
+
+\subsection{Do different developers work on different NFRs?}
+\todo{figure this out!}
+
+\section{Discussion}
+\label{sec:limit}
+
+%XXX can just remove this
+%\subsection{Effectiveness}
+\subsection{Annotation Observations}
+We found many topics that were not actually non-functional requirements (NFRs) but were often related to them. 
+For instance, concurrency was mentioned often in the commit logs and
+was related to \emph{correctness} and \emph{reliability}, possibly because concurrent code is prone to bugs such as race conditions. %are 
+% often  was troublesome. 
+\emph{Configuration management} and \emph{source control} related changes appeared often; % and sometimes there were topics dedicated to configuration management. 
+these kinds of changes are slightly related to \emph{maintainability}. 
+A non-functional change that was not quality-related was licensing and
+copyright; many changes were simply to do with updating copyrights or
+ensuring copyright or license headers were applied to files. In these
+cases we assigned the \emph{None} label to the topic.
+
+We noticed that occasionally the names of modules would conflict with words related to other non-functional requirements. 
+For instance, optimizers are very common modules in database systems: both MySQL and MaxDB have optimizer modules. 
+In MySQL the optimizer is mentioned but often the change addresses  correctness or another quality. 
+Despite this difference, the name of the module could fool our learners into believing the change was always about \emph{efficiency}. 
+In these cases the advantages of tailoring topic names to specific project terminologies are more clear. 
+Project specific word-lists would avoid automated mistakes due to the names of entities and modules of a software project.
+
+%FAT
+%\newpage
+\subsection{Summary of Techniques}
+While an unsupervised technique such as LDA is appealing in its lack of human intervention, and thus lower effort, 
+supervised learners have the advantage of domain knowledge, which typically means improved results. 
+% NEIL - we got criticized for making it sound like this is the obvious thing to do... and then why do you need a learner?
+%Our manual annotations were fairly quick to do, taking only a few minutes per 30 day period. 
+Creating annotated topics (i.e., manual labels) for training is painstaking, but with a suitably representative set of topics, the effort is acceptable.
+To annotate \emph{all} topics took us approximately 20 hours per project, but we estimate only 10\% of the topics need annotation to produce useful results.
+
+Very rarely did \textsf{exp2} and \textsf{exp3} (semi-unsupervised word matching) ever perform as well as the supervised machine learners. 
+For MaxDB, \textit{reliability} was slightly better detected using the static word list of \textsf{exp2}. 
+In general, the machine learners and \textsf{exp3} did better than \textsf{exp2} for both MaxDB and MySQL. 
+For both MySQL and MaxDB \textit{usability} was better served by \textsf{exp2}. 
+\textit{Usability} was a very infrequent label, however, which made it difficult to detect in any case.
+
+The semi-unsupervised labelling had difficulty distinguishing between common labels and infrequent labels. 
+The learners would occasionally mislabel a topic deserving of an infrequent label with a more common label.
+The word-lists for \emph{correctness} tended to be too lengthy, non-specific and broad, especially if WordNet words were used, since the NFRs are typically loosely defined concepts in common parlance.
+
+We found that the multi-label learners of BR, CLR and HOMER only did
+as well or worse for Macro-ROC as the single-label Naive Bayes and other naive Bayes-derived learners. 
+This suggests that by combining together multiple Naive Bayes learners
+we could probably label sets of topics effectively, but it would
+require a separate Naive Bayes learner per label.
+
+%{Discuss how these are CHEAP techniques and although somewhat %inaccurate they don't require a tonne of work}
+
+With ROC values ranging from $0.6$ to $0.8$ we can see there is promise in these methods.
+\textsf{exp2} and \textsf{exp3} both indicate that static information can be used to help label topics without any training whatsoever. 
+MySQL and MaxDB's machine learners made some decisions based off a few shared words: \textsf{bug, code, compiler, database, HP UX, delete, memory, missing, problems, removed, add, added, changed, problem, and test}.
+Adding these words to the word-lists of \textsf{exp2} and \textsf{exp3} could improve performance while ensuring they were only domain specific.
+
+If the techniques used in \textsf{exp2} and \textsf{exp3} were combined with the supervised techniques, we could reduce the training effort by boosting training sets with topics classified with the semi-unsupervised techniques.
+Both Naive Bayesian learners and the word-list approaches were computationally efficient.  
+%These results are promising, because the accuracy of the techniques is good enough to be useful, while the run-time performance is still cheap enough to execute to be feasible as an automated or semi-automated method of labelling topics by their software qualities.
+These results are promising because they indicate that these
+techniques are accurate enough to be useful while still maintaining
+acceptable run-time performance.
+
+Since this work focuses on labelling natural language commit log
+comments we feel it can be adapted to other natural language artifacts
+such as mailing-list discussions and bug reports, even though that was
+not evaluated in this paper. Bug reports might not exhibit the same
+behaviour as commits in terms of dominant topics.
+
+\subsection{Threats to Validity}
+% can always be cut down
+%Our work faced multiple threats to validity and we attempted to address many of them.
+\emph{Construct validity} -- we used only commit messages rather than
+mail or bug tracker messages. 
+To extend further we would need matching repositories for each project.
+%Each case study was BecauseSince we required a seamless record, it was not possible to obtain
+%other data sources. 
+Possibly they would have influenced our results, but there would be a degree of correlation between the corpora.
+%Conversely, construct validity was addressed in our validation as we used the files, topics and revisions to annotate the commit messages. 
+%We used mailing-lists to verify the purpose behind some events.
+Our taxonomy for software NFRs is subject to dispute, but seems
+generally accepted. Finally, there are exogenous sources, such as
+in-person discussions, which we did not access, an omission as noted
+in Aranda et al.~\cite{aranda09icse} %We did not show these results to the stakeholders.
+
+\emph{Internal validity} -- %includes inter-rater reliability and our lack of rival explanation building. 
+We improved internal validity by trying to correlate and explain the behaviours observed in the analysis with the historical records of the projects.
+We did not attempt to match our results to any particular model.
+We were unable to assess inter-rater reliability.
+
+\emph{External validity} -- %was addressed by our use of two case studies, but there are still issues. 
+Our data originated from OSS database projects and thus might not be applicable to commercially developed software or other domains. 
+Furthermore, our analysis techniques rely on a project's use of meaningful commit messages, although we feel this is the most frequently occurring form of developer comments. 
+%The OSS projects investigated were database systems, thus 
+%our results might not generalize to other domains. %We only studied database systems.
+% \todo[inline]{If we compare with the same domain we don't get to say it generalizes a lot?}
+
+% in Case Study Research: Design Case Studies
+\emph{Reliability} -- each annotator, the first two authors, followed the same protocol and used the same annotations. 
+However, only two annotators were used; their annotations could be
+biased as we did not analyze for inter-rater reliability because they
+did not rate the same documents.
+
+%LFAT
+\balance
+\newpage
+
+\subsection{Future Work}
+There are several avenues of further investigation.  
+More external validation would be useful. 
+Although we validated our comparisons using a mailing list for each project, interviews with developers would provide more detail. 
+We also think multi-label learning techniques, although in their infancy, are crucial in understanding cross-cutting concerns such as NFRs. 
+We want to leverage different kinds of artifacts to discover threads of NFR-related discussions that occur between multiple kinds of artifacts.
+Finally, we would like to extend this analysis to other domains, to see what patterns might occur in, for example, a consumer-facing software product.
+
+\section{Conclusions}
+This paper presented a cross-project data mining technique,
+\emph{labelled topic extraction}. %The technique leveraged a
+%non-functional requirements (NFRs) taxonomy to mine and label latent
+%software development topics. 
+Previous topic analysis research produced project-specific
+topics that needed to be manually labelled.
+%, but topics without labels
+%are difficult to interpret, and manually labelling them is
+%costly. Furthermore, previous work was project-specific and thus incomparable. 
+To improve on this, we leveraged software engineering standards,
+specifically the ISO9126 quality taxonomy, to produce a method of partially-automated (supervised) and fully-automated (semi-unsupervised) topic labelling.
+Since the word-list technique is not project-specific, we used it to compare two distinct projects, where we showed our technique produced interesting insight into maintenance activity.
+%topics and artifacts from project histories.
+% Our contributions include:
+% \begin{itemize}
+% \item We provided an unsupervised method of topic labelling (word-lists);
+% %\item Our unsupervised method can be applied to projects without training;
+% \item We demonstrated a supervised method of labelling topics by a single NFRs (machine learning);
+% \item We demonstrated a supervised method of labelling topics by multiple NFRs (multi-label machine learning);
+% \item We provided a method of cross-project analysis via topic labelling;
+% \item We demonstrated a method of visualizing these topics across time, which we used to analyze maintenance activities.
+% \end{itemize}
+
+We validated our topic labelling techniques using multiple experiments. % on unsupervised learners, supervised and multi-label learners. 
+We first conducted semi-unsupervised labelling using word-lists. 
+Our next approach was supervised, using single-label and multi-label learners. 
+Both kinds of learners performed well with average
+%area under the 
+ROC values between $0.6$ and $0.8$.
+These results, along with the efficient performance of our learners, demonstrate that labelled topic extraction is a promising approach for understanding the occurrence of non-functional requirements in software projects.
+%Although it may seem like common knowledge that maintenance is a concern as a project matures, our technique was able to show this independently.
+%We showed that NFRs are often trending topics in commit comments
+%, and that NFRs are quite common as developer topics.
+%; and that there are efficient methods of automating topic labelling.
+
+%%%HMMM
+
+Our data and scripts are available at:
+
+\url{http://softwareprocess.es/nomen/}
+
+% We demonstrated multiple supervised and unsupervised methods of topic labelling, including multi-label learners.
+% We provided a case study and visualization of NFR related topics in MaxDB 7.500 and MySQL 3.23. 
+% During the case study we manually labelled hundreds of topics by hand and then used these annotations to validate the positive performance of topic labelling techniques.
+% The implication of our paper is that topics with labels are more useful for analyzing software development activities than are unlabelled topics.
+To furnish stakeholders with labelled topics, our technique provides many methods of automatic topic labelling.
 
@@
 %\bibliographystyle{spphys}       % APS-like style for physics
-%\bibliography{}   % name your BibTeX data base
+\bibliography{ese}   % name your BibTeX data base
 

### 2011-08-11 ese/ese.tex

@@
 insight into their evolving software development activities.
-%160 words
+
 \keywords{ Software maintenance \and Repository mining \and Latent dirichlet allocation \and Topic models}
@@
 We selected systems from the same application domain, to control for differences in functional, rather than non-functional, requirements. 
-We used three different open-source, partially-commercial database systems:
-\begin{itemize}
-\item  MySQL 3.23.
-Started in 1994 and MySQL 3.23 was released in early 2001.
-MySQL contains $320,000$ lines of C and C++ source code\footnote{generated using David A.
-Wheeler's \emph{SLOCCount}, {http://dwheeler.com/sloccount}.}.
-\item MaxDB 7.500. Started in the late 1970s as a research project, and was later acquired by SAP. As of version 7.500, released April 2007, the project has over $940,000$ lines of C source code. 
-\item PostgreSQL 7.3. Started in the 1980s as a Berkeley research project. PostgreSQL 7.3 contains $306,000$ lines of C code.
-\end{itemize}
-
+We selected MySQL 3.23 and MaxDB 7.500 as they were open-source, partially-commercial database systems. 
+MaxDB started in the late 1970s as a research project, and was later acquired by SAP. 
+As of version 7.500, released April 2007, the project has over $940,000$ lines of C source code\footnote{generated using David A. Wheeler's \emph{SLOCCount}, {http://dwheeler.com/sloccount}.}. 
+The MySQL project started in 1994 and MySQL 3.23 was released in early 2001. 
+MySQL contains $320,000$ lines of C and C++ source code.  
 %Choosing an older version of these projects allowed us to focus on projects which have moved further into the maintenance phase of the software life-cycle.
@@
 %We evaluated two research questions using the ground truth data that we annotated by hand:
-We motivated our investigation with three research questions:
+We motivated our investigation with two research questions:
 \begin{enumerate}
@@
   interesting because it implies a difference in focus between two projects.
-
-\item \emph{Do different developers work on different NFRs?} For a given project, it is reasonable to think that developers are either assigned (in commercial organizations) or choose (in open-source organizations) to a particular NFR.
-For example, one developer might be more junior, and take responsibility for the low-impact reliability fixes.
-Another, more senior developer might assume responsibility for major improvements such as efficiency improvements.
-
 \end{enumerate}
@@
 After this point, efforts shift to the newer releases (4.0, 4.1,
-5.0). 
-\todo{add PGSQL disucssion here. Schema additions? Lots of functionality?}
-
-We now address our research questions:
+5.0). We now address our research questions:
 
@@
 %\subsubsection{Do label frequencies change over time?}
-\todo{add more on PGSQL}
 In both projects the frequencies generally decreased with age. 
@@
  % actionable for a developer or manager}
-
-\subsection{Do different developers work on different NFRs?}
-\todo{figure this out!}
 

### 2011-08-23 ese/ese.tex
before line-shortening

--- 2011-08-10 ese/ese.tex
@@
 requirements. Our approach was evaluated with experiments and case studies
-on two large-scale RDBMS projects: MySQL and MaxDB.  The case studies
+on three large-scale RDBMS projects: MySQL, PostgreSQL and MaxDB.  The case studies
 show that labelled topic extraction can produce appropriate,
@@
 
+\item examines how NFRs correlate with the work of individual developers;
 \item provides a method of cross-project analysis via topic labelling; and
  applies these techniques to visualize NFRs over time, and to analyze maintenance activities.
+
 \end{itemize}

### 2011-08-23 ese/ese.tex
some more todos and line wrapping for easier diff

@@
 %XXXXXX this sentence is confusing
-%This informs our method of generalizing the data mining of these software artifacts to leverage the global domain knowledge of software engineering, specifically NFRs. 
+%This informs our method of generalizing the data mining of these software artifacts to leverage the global domain knowledge of software engineering,
+%specifically NFRs. 
 %In particular, we use software quality models to generalize topics across projects. 
@@
 conversations about different software projects.  
-Furthermore, there is a series of standards on NFRs, such as ISO9126 \cite{iso9126}, that are specifically intended to apply to projects of varying types; this suggests that our goal of trying to extract NFR-related development topics, such as those related to software quality models, holds promise.
+Furthermore, there is a series of standards on NFRs, such as ISO9126 \cite{iso9126}, that are specifically intended to apply to projects of varying
+types; this suggests that our goal of trying to extract NFR-related development topics, such as those related to software quality models, holds promise.
 
@@
 
-%% RESEARCH DIDNT DO THIS AND WE ARE TOTALLY AWESOME AND WE DID
-% it is ok to repeat this here because it is clear and bullet pointed and the reviewers wont have any bloody excuse to not see it
 In this paper, we describe \emph{automated labelled topic extraction}. It addresses two gaps in the topic mining literature:
@@
 We begin by discussing related work in Section \ref{sec:related}.
-Next, we describe how we generated our data (Section \ref{sec:wordlist}).
-For semi-unsupervised classification (Section \ref{sec:unsuplabelling}), we begin by creating word-lists to signify when a topic matches an NFR label.
-We then apply our classifier and analyze the results.
+Next, we describe how we generated our data (Section \ref{sec:wordlist}). For semi-unsupervised classification (Section \ref{sec:unsuplabelling}), we
+begin by creating word-lists to signify when a topic matches an NFR label. We then apply our classifier and analyze the results. 
 %and how we created the word-lists used in unsupervised classification. 
@@
 %We then present visualizations of named topics and their trends over time to aid communication and analysis. 
-To demonstrate an application of labelled topic extraction, we use an exploratory case study of two open source database systems to show how named topics can be compared between projects  (Section \ref{sec:analysis}).
+To demonstrate an application of labelled topic extraction, we use an exploratory case study of two open source database systems to show how named
+topics can be compared between projects  (Section \ref{sec:analysis}). 
 The paper concludes with a discussion of limitations (Section \ref{sec:limit}), and future work.
-
-%\todo[inline]{Topics are poorly defined by this point and since we're building on research which didn't DO this, we should mention that research. Previous work probably should go here.}
-
-%\todo[inline]{We still haven't shown that topics are useful -- a reviewer comment}
 
@@
 
-Cleland-Huang and her colleagues have investigated mining requirements documents for non-functional requirements (NFR) (quality requirements)~\cite{Cleland-Huang2006}. 
+Cleland-Huang and her colleagues have investigated mining requirements documents for non-functional requirements (NFR) (quality
+requirements)~\cite{Cleland-Huang2006}. 
 Their approach is similar to ours, as they mined keywords from NFR
@@
 %histories that we use.
-%The objective of Cleland-Huang's study was to identify new NFRs for system development, yet our objective was to recover those latent NFRs from commit-log messages of the project.
+%The objective of Cleland-Huang's study was to identify new NFRs for system development, 
+%yet our objective was to recover those latent NFRs from commit-log messages of the project.
 
@@
 %Although subject to the assumption that all projects conceive of these terms in the same way.
-%Thirdly, while the objective of Cleland-Huang's study was to identify new NFRs (for system development) our study assumes these NFRs are latent in the commit-log messages of the project.
+%Thirdly, while the objective of Cleland-Huang's study was to identify new NFRs (for system development) 
+%our study assumes these NFRs are latent in the commit-log messages of the project.
 
 Similarly, Mockus and Votta~\cite{Mockus00} studied a large-scale industrial change-tracking system. 
-Mockus and Votta leveraged WordNet~\cite{Fellbaum1998}, an English-language ``lexical database'' that contains semantic relations between words, including common related forms (similar to word stemming), meronymy and synonymy.
+Mockus and Votta leveraged WordNet~\cite{Fellbaum1998}, an English-language ``lexical database'' that contains semantic relations between words,
+including common related forms (similar to word stemming), meronymy and synonymy.
 They used WordNet for word roots as they felt the synonyms would be
@@
 %Mens et al.~\cite{mens08icsm} conducted an empirical study of Eclipse, a popular Java IDE, to verify the claims of Lehman~\cite{lehman97sms}.
-% They concerned themselves with source code only, and found Law Seven, ``Declining Quality'', to be too difficult to assess: ``[we lacked an] appropriate measurement of the evolving quality of the system as perceived by the users \cite[p.
-388]{mens08icsm}''.
+% They concerned themselves with source code only, and found Law Seven, ``Declining Quality'', to be too difficult to assess:
+% ``[we lacked an] appropriate measurement of the evolving quality of the system as perceived by the users \cite[p. 388]{mens08icsm}''. 
 %This paper examines the notions of quality in terms of a consistent ontology, as Menset al. call for in their conclusions.
@@
 The labels are reasonable enough, but still require an expert in the field to determine them. 
-Furthermore, these labels are project-specific, because they are generated from the data of that project.
-E.g., we might have a label called \emph{`Oracle'} in the MySQL case, since Oracle owns MySQL.
+Furthermore, these labels are project-specific, because they are generated from the data of that project. E.g., we might have a label called \emph{`Oracle'}
+in the MySQL case, since Oracle owns MySQL. 
 %We do this differently.
 %This paper does this differently. 
-Our approach differs: first of all, we automate the process of naming the topics; secondly, we label topics with project-independent terms, in order to permit cross-project comparison.
+Our approach differs: first of all, we automate the process of naming the topics; secondly, we label topics with project-independent terms, in order
+to permit cross-project comparison.
 
@@
 %Their work discusses the source of the requirements and how they are used in the development process. 
-German~\cite{german03gnome} looked at GNOME specifically, and listed several sources for requirements: leader interest, mimicry, brainstorming, and prototypes.
+% German~\cite{german03gnome} looked at GNOME specifically, and listed several sources for requirements: leader interest, mimicry, brainstorming, and prototypes. 
 %None of this work  addressed quality requirements in OSS, nor did it examine requirements trends.
-
+\todo{re-add this related work?}
 % XXX vanity or space.. 
@@
 % NE 0113 this seems too early to mention 'result'.
-% Hindle et al.
-\cite{Hindle2007} examined release patterns in OSS.
-They showed that there is a difference between projects regarding maintenance techniques.
-This supports our result that software qualities are not discussed with the same frequency across projects.
+% Hindle et al. \cite{Hindle2007} examined release patterns in OSS. They showed that there is a difference between projects regarding maintenance
+% techniques. This supports our result that software qualities are not discussed with the same frequency across projects.
 
@@
 %e created labels using a well-known NFR taxonomy. 
-In supervised learning, we trained our data with manual annotations in order to match topics with NFRs. Finally, these topics were used to analyze the role of NFRs in software maintenance.
+In supervised learning, we trained our data with manual annotations in order to match topics with NFRs. Finally, these topics were used to analyze the
+role of NFRs in software maintenance.
 
@@
 \begin{itemize}
-\item  MySQL 3.23.
-Started in 1994 and MySQL 3.23 was released in early 2001.
-MySQL contains $320,000$ lines of C and C++ source code\footnote{generated using David A.
-Wheeler's \emph{SLOCCount}, {http://dwheeler.com/sloccount}.}.
-\item MaxDB 7.500. Started in the late 1970s as a research project, and was later acquired by SAP. As of version 7.500, released April 2007, the project has over $940,000$ lines of C source code. 
+\item  MySQL 3.23. Started in 1994 and MySQL 3.23 was released in early 2001. MySQL contains $320,000$ lines of C and C++ source code\footnote{generated
+using David A. Wheeler's \emph{SLOCCount}, {http://dwheeler.com/sloccount}.}.
+\item MaxDB 7.500. Started in the late 1970s as a research project, and was later acquired by SAP. As of version 7.500, released April 2007, the project
+has over $940,000$ lines of C source code. 
 \item PostgreSQL 7.3. Started in the 1980s as a Berkeley research project. PostgreSQL 7.3 contains $306,000$ lines of C code.
@@
 %Choosing an older version of these projects allowed us to focus on projects which have moved further into the maintenance phase of the software life-cycle.
-We explicitly chose older versions of mature projects from a stable problem domain to increase the likelihood that we would encounter primarily maintenance activities in our studies.
+We explicitly chose older versions of mature projects from a stable problem domain to increase the likelihood that we would encounter primarily
+maintenance activities in our studies.
 
@@
 Most commits we observed had commit comments.
-Commit comments are often studied by researchers, as they are the most readily accessible source of project interactions, and developers are often required to create them by the repository mechanism (e.g., Git).
-Additionally, relying only on commit comments makes our approach more generalizable, as we do not assume the presence of other artifact corpora.
-%This is the most readily accessible source of project interactions for outside researchers, since developers often or always (e.g., Git) write such message.
-Our approach is more generalizable if it does not assume the presence of other corpora.
+Commit comments are often studied by researchers, as they are the most readily accessible source of project interactions, and developers are often
+required to create them by the repository mechanism (e.g., Git).  Additionally, relying only on commit comments makes our approach more generalizable,
+as we do not assume the presence of other artifact corpora.
+%This is the most readily accessible source of project interactions for outside researchers, since developers often or always (e.g., Git) write such
+%message. Our approach is more generalizable if it does not assume the presence of other corpora. 
 %bug trackers and email-list archives were not available for both projects. 
-An example of a typical commit message, from MySQL, is: \textit{``history annotate diffs bug fixed (if mysql\-\_real\-\_connect() failed there were two pointers to malloc'ed strings, with memory corruption on free(), of course)''}.
+An example of a typical commit message, from MySQL, is: \textit{``history annotate diffs bug fixed (if mysql\-\_real\-\_connect() failed there were
+two pointers to malloc'ed strings, with memory corruption on free(), of course)''}. 
 We extracted these messages and indexed them by creation time. 
@@
 %analyze
-We chose a period size of 30 days as it is smaller than the time between minor releases but large enough for there to be sufficient commits to analyze~\cite{Hindle09ICSM}. 
-For each 30 day period of each project, we input the messages of that period into Latent Dirichlet Allocation (LDA), a topic analysis algorithm~\cite{Blei2003}, and recorded the topics the algorithm extracted.
+We chose a period size of 30 days as it is smaller than the time between minor releases but large enough for there to be sufficient commits to
+analyze~\cite{Hindle09ICSM}. 
+For each 30 day period of each project, we input the messages of that period into Latent Dirichlet Allocation (LDA), a topic analysis
+algorithm~\cite{Blei2003}, and recorded the topics the algorithm extracted.
 %For each project and for each period of messages a project, we input these messages 
@@
 We applied Blei's LDA implementation~\cite{Blei2003} against the word distributions of these commits, and generated lists of topics per period. 
-We set the number of topics to generate to $20$, because past experimentation showed that fewer topics might aggregate multiple unique topics while any more topics dilutes the results and creates indistinct topics~\cite{Hindle09ICSM}.
+We set the number of topics to generate to $20$, because past experimentation showed that fewer topics might aggregate multiple unique topics while
+any more topics dilutes the results and creates indistinct topics~\cite{Hindle09ICSM}. 
 % Furthermore, more than $20$ topics quickly became infeasible for inspection.
@@
 ISO9126 describes six high-level NFRs: maintainability, functionality,
-portability, efficiency, usability, and reliability~\footnote{While there may be lingering debate in some circles about these terms, an ISO standard seems like a reasonable starting point for our work.}.
+portability, efficiency, usability, and reliability~\footnote{While there may be lingering debate in some circles about these terms, an ISO standard
+seems like a reasonable starting point for our work.}.
 %There is some debate about the terms in this model. 
-%However, ISO9126 is ``an international standard and thus provides an internationally accepted terminology for software quality \cite[p.
-58]{Boegh2008},'' that is sufficient for the purposes of this research.
-We claim that these NFRs are maintenance concerns (to varying degrees) in all software projects, and are therefore well-suited for comparisons between projects.
+%However, ISO9126 is ``an international standard and thus provides an 
+%internationally accepted terminology for software quality \cite[p. 58]{Boegh2008},'' that is sufficient for the purposes of this research.  
+We claim that these NFRs are maintenance concerns (to varying degrees) in all software projects, and are therefore well-suited for comparisons between
+projects.
 
@@
 suggested was the label for that topic. 
-We were able to pinpoint the appropriate label using auxiliary information as well, such as the actual revisions and files that were related to the topic being annotated.
+We were able to pinpoint the appropriate label using auxiliary information as well, such as the actual revisions and files that were related to the
+topic being annotated.
 For example, for the MaxDB topic consisting of a message ``exit() only
@@
 %\todo[inline]{inter-rater reliability }
-% \todo[inline]{There should be a section here which is about manual annotation. It is really important and it is required by both analyses. Plus we can tell them how.}
+% \todo[inline]{There should be a section here which is about manual annotation. 
+%It is really important and it is required by both analyses. Plus we can tell them how.}
 
@@
 
-%\todo[inline]{NEIL: abram, i got rid of the letter-grade thing because I'm worried people latch on to the idea we are C- students. I feel anything better than random is our criteria here}
 ROC values provide a score %, similar to school letter-grades (A is 0.9, C is 0.6),
  reflecting how well a particular learner performed for the given data. 
-ROC maps to the more familiar concepts of precision/sensitivity and recall/specificity: it plots the true positive rate (sensitivity) versus the false positive rate (1 - specificity). 
+ROC maps to the more familiar concepts of precision/sensitivity and recall/specificity: it plots the true positive rate (sensitivity) versus the false
+positive rate (1 - specificity). 
 A perfect learner has a ROC value of 1.0, reflecting perfect recall and precision. 
@@
 The %area under the -- NE confusing to have area again ... and not importantana
-ROC of a classifier is equivalent to the probability that the classifier will rank a randomly chosen positive instance higher than a randomly chosen negative instance.
+ROC of a classifier is equivalent to the probability that the classifier will rank a randomly chosen positive instance higher than a randomly chosen
+negative instance.
 We consider our labelling classifiers acceptable if they outperform a random classifier (0.5).
@@
 WordNet~\cite{Fellbaum1998}.
-%, an English-language ``lexical database'' that contains semantic relations between words, including common related forms (similar to word stemming), meronymy and synonymy. 
+%, an English-language ``lexical database'' that contains semantic relations between words,
+% including common related forms (similar to word stemming), meronymy and synonymy. 
 We then added Boehm's software quality model~\cite{Boehm+:1976:ICSE}, and classified his eleven `ilities' into their respective ISO9126 NFRs. 
 We did the same for the quality model produced by McCall et al.~\cite{mccall1977}. 
-We then did a simple random analysis of mailing list messages from an open source project, KDE.
-If we judged a given message to contain terms that were related to one of the NFRs in ISO9126, we added it to our word-list.
-This allowed us to expand our word-lists with more software-specific terms.
+We then did a simple random analysis of mailing list messages from an open source project, KDE. If we judged a given message to contain terms that were
+related to one of the NFRs in ISO9126, we added it to our word-list. This allowed us to expand our word-lists with more software-specific terms.
 %Finally, we randomly analyzed two mailing lists from another software project to expand our set with domain-specific terms. 
@@
 We do not include these words here for space considerations (but see the Appendix for our data repository). 
-It is not clear the words associated with our labels in \textsf{exp3} are specific enough.
-For example, the label \emph{maintainability} is associated with words \emph{ease} and \emph{ownership}.
-In general, as we proceed from word-list in \textsf{exp1} to that in \textsf{exp3}, our lists become more generic.
+It is not clear the words associated with our labels in \textsf{exp3} are specific enough. For example, the label \emph{maintainability} is associated with
+words \emph{ease} and \emph{ownership}. In general, as we proceed from word-list in \textsf{exp1} to that in \textsf{exp3}, our lists become more generic.
 
@@
 \emph{Maintainability} &
-testability changeability analyzability stability maintain maintainable modularity modifiability understandability interdependent dependency encapsulation decentralized modular\\ \hline
+testability changeability analyzability stability maintain maintainable modularity modifiability understandability interdependent dependency encapsulation
+decentralized modular\\ \hline
 \emph{Functionality} &
-security compliance accuracy interoperability suitability functional practicality functionality compliant exploit certificate secured ``buffer overflow'' policy malicious trustworthy vulnerable vulnerability accurate secure vulnerability correctness accuracy\\ \hline
+security compliance accuracy interoperability suitability functional practicality functionality compliant exploit certificate secured ``buffer overflow''
+policy malicious trustworthy vulnerable vulnerability accurate secure vulnerability correctness accuracy\\ \hline
 \emph{Portability} &
-conformance adaptability replaceability installability portable movableness movability portability specification migration standardized l10n localization i18n internationalization documentation interoperability transferability\\ \hline
+conformance adaptability replaceability installability portable movableness movability portability specification migration standardized l10n localization
+i18n internationalization documentation interoperability transferability\\ \hline
 \emph{Efficiency} &
-``resource behaviour'' ``time behaviour'' efficient efficiency performance profiled optimize sluggish factor penalty slower faster slow fast optimization\\ \hline
+``resource behaviour'' ``time behaviour'' efficient efficiency performance profiled optimize sluggish factor penalty slower faster slow fast optimization\\
+\hline
 \emph{Usability} &
-operability understandability learnability useable usable serviceable usefulness utility useableness usableness serviceableness serviceability usability gui accessibility menu configure convention standard feature focus ui mouse icons ugly dialog guidelines click default human convention friendly user screen interface flexibility\\ \hline
+operability understandability learnability useable usable serviceable usefulness utility useableness usableness serviceableness serviceability usability
+gui accessibility menu configure convention standard feature focus ui mouse icons ugly dialog guidelines click default human convention friendly user
+screen interface flexibility\\ \hline
 \emph{Reliability} &
-``fault tolerance'' recoverability maturity reliable dependable responsibleness responsibility reliableness reliability dependableness dependability resilience integrity stability stable crash bug fails redundancy error failure\\
+``fault tolerance'' recoverability maturity reliable dependable responsibleness responsibility reliableness reliability dependableness dependability
+resilience integrity stability stable crash bug fails redundancy error failure\\ 
 \bottomrule
@@
 
-Using our three word-lists (\textsf{exp1}, \textsf{exp2}, \textsf{exp3}), we labelled our topics with an NFR where there was a match between a word in the list and the same word somewhere in the distribution of words that constitute the topic.
+Using our three word-lists (\textsf{exp1}, \textsf{exp2}, \textsf{exp3}), we labelled our topics with an NFR where there was a match between a word in
+the list and the same word somewhere in the distribution of words that constitute the topic.
 A \emph{named topic} is a topic with a matching NFR label. \emph{Unnamed
-topics} occur where there is no such match, which may indicate either a lack of precision, or simply that this topic is not associated with non-functional requirements.
+topics} occur where there is no such match, which may indicate either a lack of precision, or simply that this topic is not associated with non-functional
+requirements.
 All experiments were run on MaxDB 7.500 and MySQL 3.23 data. LDA
@@
 
+\todo{update with PGSQL}
 \bottomrule
@@
 We did not see many results for \emph{usability} or \emph{accuracy}, which were associated with fewer than ten topics. 
-We also looked for correlations between our labels: excluding double matches (self-correlation), our highest co-occurring terms were \emph{verifiability} with \emph{traceability}, and \emph{testability} with \emph{correctness} (76 and 62 matches, respectively).
+We also looked for correlations between our labels: excluding double matches (self-correlation), our highest co-occurring terms were \emph{verifiability}
+with \emph{traceability}, and \emph{testability} with \emph{correctness} (76 and 62 matches, respectively).
 % See the following sections for our error analysis.
-
-%LFAT
-\balance
-\newpage
 
@@
 least frequent label, 44 topics, was \emph{maintainability}. 
-Common co-occurrences were \emph{reliability} with \emph{usability}, \emph{efficiency} with \emph{reliability}, and \emph{efficiency} with \emph{usability} (200, 190, and 150 topics in common, respectively).
+Common co-occurrences were \emph{reliability} with \emph{usability}, \emph{efficiency} with \emph{reliability}, and \emph{efficiency} with \emph{usability}
+(200, 190, and 150 topics in common, respectively).
 
@@
 %we compared the results of the unsupervised word matching approach. 
-For each quality we assessed whether semi-unsupervised labels matched the manual annotations. %To be clear, the manual annotations were not used to train the labelling process.
+For each quality we assessed whether semi-unsupervised labels matched the manual annotations. %To be clear, the manual annotations were not used to
+train the labelling process.
 As described in Section \ref{sec:wordlist} we used both ROC and F-1 measures to evaluate the performance of the classification.
@@
 
-%\todo[inline]{since ROC is new-sounding, I wonder if a more lengthy explanation would be merited. An obvious question is why not precision/recall? -- ROC is hot right now and it combines both}
-% Rebuttal: ROC is not new and it is what the machine learning community, search based software engineering and now the MSR community are using +  it has natural meaningful quality mappings.
+%\todo[inline]{since ROC is new-sounding, I wonder if a more lengthy explanation would be merited. An obvious question is why not precision/recall? --
+% ROC is hot right now and it combines both}
+% Rebuttal: ROC is not new and it is what the machine learning community, search based software engineering and now the MSR community are using +
+%it has natural meaningful quality mappings.
 % \begin{figure}[t]
@@
 % %}
-% \caption[]{Performance of unsupervised topic labelling for each NFR per word-list.
-A) ROC values -- dashed line indicates the performance of a random classifier.
-B) F1 measure.
-Possible values range from 0--1.
+% \caption[]{Performance of unsupervised topic labelling for each NFR per word-list. A) ROC values -- 
+% dashed line indicates the performance of a random classifier. B) F1 measure. Possible values range from 0--1.
 % %\todo[inline]{better discussion}% and graphical legend would be nice}
@@
 Supervised labelling requires expert analysis of the correct
-class/label to assign a label to a topic.
-In our approach, we use the top-level NFRs in the ISO9126 standard~\cite{iso9126} for our classes, but other taxonomies are also applicable.%In order to validate how effective these word-bag approaches to topic labelling would be we created a validation data set.
+class/label to assign a label to a topic. In our approach, we use the top-level NFRs in the ISO9126 standard~\cite{iso9126} for our classes, but other
+taxonomies are also applicable.%In order to validate how effective these word-bag approaches to topic labelling would be we created a validation data set.
 
@@
 We also used the multi-labelling add-on for WEKA, Mulan~\cite{mulan}. %\footnote{\url{http://mlkd.csd.auth.gr/multilabel.html}}. 
-Traditional classifiers label topics with a single class, whereas Mulan allows for a mixture of classes per topic, which is what we observed while manually labelling topics.
+Traditional classifiers label topics with a single class, whereas Mulan allows for a mixture of classes per topic, which is what we observed while
+manually labelling topics.
 The \emph{features} we used are word counts/occurrence	per topic, if
@@
 
-To assess the performance of the supervised learners, we did a 10-fold cross-validation~\cite{Kohavi1995}, a common technique for evaluating machine learners. 
+To assess the performance of the supervised learners, we did a 10-fold cross-validation~\cite{Kohavi1995}, a common technique for evaluating machine
+learners. 
 The original data is partitioned randomly into ten sub-samples. Each sample is used to test against a training set composed of the nine other samples.
@@
 
-Figure \ref{fig:best-learn-per-tag} shows that MaxDB and MySQL have quite different results, as the ROC values for \emph{reliability} and \emph{functionality} seem swapped between projects.
+Figure \ref{fig:best-learn-per-tag} shows that MaxDB and MySQL have quite different results, as the ROC values for \emph{reliability} and \emph{functionality}
+seem swapped between projects.
 
@@
 % }
-\caption[]{ROC value for the best learner per label for MaxDB and MySQL. Values range from $0$--$1$.  Dashed line indicates the performance of a random classifier.
+\caption[]{ROC value for the best learner per label for MaxDB and MySQL. Values range from $0$--$1$.  Dashed line indicates the performance of a random
+classifier.
 %\todo[inline]{better discussion}
@@
 %The features we used are word counts per message. 
-One beneficial aspect of this result is that it suggests we can have very fast training and classifying  since Naive Bayes can be calculated in $O(N)$ for $N$ features.
+One beneficial aspect of this result is that it suggests we can have very fast training and classifying  since Naive Bayes can be calculated in $O(N)$
+for $N$ features.
 %http://nlp.stanford.edu/IR-book/html/htmledition/naive-bayes-text-classification-1.html
@@
 As noted in Section \ref{sec:wordlist}, each topic in our data-set can be composed of zero, one, or more NFRs. 
-For example, a commit message might address \textit{reliability} in the context of \textit{efficiency}, or make a \textit{maintainability} improvement in the source code that related to \textit{usability}.
+For example, a commit message might address \textit{reliability} in the context of \textit{efficiency}, or make a \textit{maintainability} improvement
+in the source code that related to \textit{usability}. 
 However, traditional machine learning techniques, such as Naive Bayes, can only map topics to a single class. 
@@
 Mulan also includes methods for determining the performance of such techniques.
-% TOO DETAILED The problem framed in the learners above has changed; instead of looking at the precision and recall of applying one label, we rank multiple labels at once.
-We must check if the full subset of labels was applied, and then how much of that subset was applied.
+% TOO DETAILED The problem framed in the learners above has changed; instead of looking at the precision and recall of applying one label, we rank
+% multiple labels at once. We must check if the full subset of labels was applied, and then how much of that subset was applied.
 
@@
 }
-\caption[]{MySQL and MaxDB macro (grey) and micro (black) ROC results per multi-label learner. Possible values range from $0$--$1$.  Dashed line indicates the performance of a random classifier.
+\caption[]{MySQL and MaxDB macro (grey) and micro (black) ROC results per multi-label learner. Possible values range from $0$--$1$.  Dashed line
+indicates the performance of a random classifier.
 %\todo[inline]{better discussion}% and graphical legend would be nice}
@@
 Figure~\ref{fig:mulan} presents the results of Mulan's best multi-label learners for our data. 
-Calibrated Label Ranking (CLR) is a learner that builds two layers. The first layer determines if an entity should be labelled, while the second layer determines what labels should be assigned.
-The Hierarchy Of Multi-label classifiERs (HOMER) and Binary Relevance (BR) act as a hierarchy of learners: BR is flat, while HOMER tries to build a deeper hierarchy for a more accurate learner~\cite{mulan}.
+Calibrated Label Ranking (CLR) is a learner that builds two layers. The first layer determines if an entity should be labelled, while the second layer
+determines what labels should be assigned.
+The Hierarchy Of Multi-label classifiERs (HOMER) and Binary Relevance (BR) act as a hierarchy of learners: BR is flat, while HOMER tries to build a
+deeper hierarchy for a more accurate learner~\cite{mulan}. 
 These classifiers performed better than other multi-label classifiers as they have the best micro and macro ROC scores. 
@@
 
-\item \emph{Do different developers work on different NFRs?} For a given project, it is reasonable to think that developers are either assigned (in commercial organizations) or choose (in open-source organizations) to a particular NFR.
-For example, one developer might be more junior, and take responsibility for the low-impact reliability fixes.
-Another, more senior developer might assume responsibility for major improvements such as efficiency improvements.
+\item \emph{Do different developers work on different NFRs?} For a given project, it is reasonable to think that developers are either assigned
+(in commercial organizations) or choose (in open-source organizations) to a particular NFR. For example, one developer might be more junior, and take
+responsibility for the low-impact reliability fixes. Another, more senior developer might assume responsibility for major improvements such as efficiency
+improvements.
 
@@
 }
+\subfloat[PostgreSQL 7.2]{					
+\includegraphics[width=\textwidth]{figures/pgsql-timeline}
+\label{fig:pgsql-timeline}
+}
 	\caption[]{NFR label per period. Each cell represents a 30-day period. %Numbers represent number of topics in that period labelled with that NFR. 
@@
 % and			http://softwareprocess.es/y/maxdb-7500.tar.gz
-%There are more \textsf{None} labels in MaxDB because a number of topics were to do with code cleanup or automated checking using tools like \texttt{sutcheck}, a tool specific to the development process of MaxDB.
+%There are more \textsf{None} labels in MaxDB because a number of topics were to do with code cleanup or automated checking using tools like
+%\texttt{sutcheck}, a tool specific to the development process of MaxDB.
 
@@
 and MaxDB 7.500 were first announced. For MaxDB, version 7.5.00  was released in December of 2003. 
-We know that release 7.5.00.23 saw the development of PHP interfaces, possibly accounting for the simultaneous increase in the \emph{portability} NFR at the same time.
-The gap in MaxDB (Figure \ref{fig:maxdb-timeline}) is due to a shift in development focus (from February 2005 to June 2005) to MaxDB 7.6, which is released in June 2005.
+We know that release 7.5.00.23 saw the development of PHP interfaces, possibly accounting for the simultaneous increase in the \emph{portability}
+NFR at the same time.
+The gap in MaxDB (Figure \ref{fig:maxdb-timeline}) is due to a shift in development focus (from February 2005 to June 2005) to MaxDB 7.6, which is
+released in June 2005.
 
@@
 The release of MySQL we study  (Figure \ref{fig:mysql-timeline}) was the first to be licenced under the GPL. 
-Version 3.23.31 (January, 2001) was the production release (non-beta), and we see a flurry of topics labelled with \emph{functionality} and \emph{maintainability}. 
+Version 3.23.31 (January, 2001) was the production release (non-beta), and we see a flurry of topics labelled with \emph{functionality} and
+\emph{maintainability}. 
 After this point, this version enters the maintenance phase of its life-cycle. 
@@
 This might be related to release 3.23.38, which focused on Windows compatibility. 
-Similarly, in August, 2002, both \emph{functionality} and \emph{portability} are frequent, and mailing list data suggests this is related to the release of version 3.23.52, a general bug fix with a focus on security (a component of the \emph{functionality} NFR in the ISO9126 model).
+Similarly, in August, 2002, both \emph{functionality} and \emph{portability} are frequent, and mailing list data suggests this is related to the release
+of version 3.23.52, a general bug fix with a focus on security (a component of the \emph{functionality} NFR in the ISO9126 model). 
 After this point, efforts shift to the newer releases (4.0, 4.1,
@@
 % We tagged more MySQL topics with annotations than MaxDB topics, yet both shared similarities. 
-% In terms of non-functional requirements both projects had long running trends that focused on \emph{functionality}, \emph{maintainability}, and \emph{portability}, yet MaxDB had more of a focus on \emph{efficiency} and \emph{reliability}.
+% In terms of non-functional requirements both projects had long running trends that focused on \emph{functionality}, \emph{maintainability}, and
+%\emph{portability}, yet MaxDB had more of a focus on \emph{efficiency} and \emph{reliability}. 
 % \todo[inline]{expand}
@@
 security problems were more common (security is a component of \emph{functionality} in the ISO9126 model). 
-In both cases \emph{portability} was a constant maintenance concern and was prevalent throughout the lifetime of the projects. It may surprise developers how often portability arises as a concern.
+In both cases \emph{portability} was a constant maintenance concern and was prevalent throughout the lifetime of the projects. It may surprise developers
+how often portability arises as a concern.
 %On one hand, this analysis is unsurprising; on the other hand, it provides confirmation that our approach identifies actual patterns.
@@
 % often  was troublesome. 
-\emph{Configuration management} and \emph{source control} related changes appeared often; % and sometimes there were topics dedicated to configuration management. 
+\emph{Configuration management} and \emph{source control} related changes appeared often; % and sometimes there were topics dedicated to configuration
+management. 
 these kinds of changes are slightly related to \emph{maintainability}. 
@@
 %Our manual annotations were fairly quick to do, taking only a few minutes per 30 day period. 
-Creating annotated topics (i.e., manual labels) for training is painstaking, but with a suitably representative set of topics, the effort is acceptable.
-To annotate \emph{all} topics took us approximately 20 hours per project, but we estimate only 10\% of the topics need annotation to produce useful results.
+Creating annotated topics (i.e., manual labels) for training is painstaking, but with a suitably representative set of topics, the effort is acceptable. To
+annotate \emph{all} topics took us approximately 20 hours per project, but we estimate only 10\% of the topics need annotation to produce useful results.
 
@@
 The learners would occasionally mislabel a topic deserving of an infrequent label with a more common label.
-The word-lists for \emph{correctness} tended to be too lengthy, non-specific and broad, especially if WordNet words were used, since the NFRs are typically loosely defined concepts in common parlance.
+The word-lists for \emph{correctness} tended to be too lengthy, non-specific and broad, especially if WordNet words were used, since the NFRs are
+typically loosely defined concepts in common parlance.
 
@@
 \textsf{exp2} and \textsf{exp3} both indicate that static information can be used to help label topics without any training whatsoever. 
-MySQL and MaxDB's machine learners made some decisions based off a few shared words: \textsf{bug, code, compiler, database, HP UX, delete, memory, missing, problems, removed, add, added, changed, problem, and test}.
+MySQL and MaxDB's machine learners made some decisions based off a few shared words: \textsf{bug, code, compiler, database, HP UX, delete, memory,
+missing, problems, removed, add, added, changed, problem, and test}. 
 Adding these words to the word-lists of \textsf{exp2} and \textsf{exp3} could improve performance while ensuring they were only domain specific.
 
-If the techniques used in \textsf{exp2} and \textsf{exp3} were combined with the supervised techniques, we could reduce the training effort by boosting training sets with topics classified with the semi-unsupervised techniques.
+If the techniques used in \textsf{exp2} and \textsf{exp3} were combined with the supervised techniques, we could reduce the training effort by boosting
+training sets with topics classified with the semi-unsupervised techniques.
 Both Naive Bayesian learners and the word-list approaches were computationally efficient.  
-%These results are promising, because the accuracy of the techniques is good enough to be useful, while the run-time performance is still cheap enough to execute to be feasible as an automated or semi-automated method of labelling topics by their software qualities.
+%These results are promising, because the accuracy of the techniques is good enough to be useful, while the run-time performance is still cheap enough
+%to execute to be feasible as an automated or semi-automated method of labelling topics by their software qualities.
 These results are promising because they indicate that these
@@
 We did not attempt to match our results to any particular model.
-We were unable to assess inter-rater reliability.
+We assessed inter-rater reliability (IRR), a measure of the correlation between the manual annotations the first two authors created. Perfect IRR (1.0) indicates complete agreement 
+on the appropriate annotation for that topic. We verified IRR using manually annotated PostgreSQL topics, and achieved a result of \todo[inline]{insert here}. 
+Some of the reasons for this are a disagreement over the appropriate classification for documentation topics (should this be considered
+usability or portability?). A short education session on what a given NFR from the ISO standard encompasses would reduce this error rate.
 
@@
 Our data originated from OSS database projects and thus might not be applicable to commercially developed software or other domains. 
-Furthermore, our analysis techniques rely on a project's use of meaningful commit messages, although we feel this is the most frequently occurring form of developer comments. 
+Furthermore, our analysis techniques rely on a project's use of meaningful commit messages, although we feel this is the most frequently occurring form
+of developer comments. 
 %The OSS projects investigated were database systems, thus 
@@
 specifically the ISO9126 quality taxonomy, to produce a method of partially-automated (supervised) and fully-automated (semi-unsupervised) topic labelling.
-Since the word-list technique is not project-specific, we used it to compare two distinct projects, where we showed our technique produced interesting insight into maintenance activity.
-%topics and artifacts from project histories.
+Since the word-list technique is not project-specific, we used it to compare two distinct projects, where we showed our technique produced interesting
+insight into maintenance activity. %topics and artifacts from project histories.
 % Our contributions include:
@@
 %area under the 
-ROC values between $0.6$ and $0.8$.
-These results, along with the efficient performance of our learners, demonstrate that labelled topic extraction is a promising approach for understanding the occurrence of non-functional requirements in software projects.
+ROC values between $0.6$ and $0.8$. These results, along with the efficient performance of our learners, demonstrate that labelled topic extraction is
+a promising approach for understanding the occurrence of non-functional requirements in software projects. 
 %Although it may seem like common knowledge that maintenance is a concern as a project matures, our technique was able to show this independently.
@@
 % We provided a case study and visualization of NFR related topics in MaxDB 7.500 and MySQL 3.23. 
-% During the case study we manually labelled hundreds of topics by hand and then used these annotations to validate the positive performance of topic labelling techniques.
-% The implication of our paper is that topics with labels are more useful for analyzing software development activities than are unlabelled topics.
-To furnish stakeholders with labelled topics, our technique provides many methods of automatic topic labelling.
+% During the case study we manually labelled hundreds of topics by hand and then used these annotations to validate the positive performance of topic
+%labelling techniques.
+% The implication of our paper is that topics with labels are more useful for analyzing software development activities than are unlabelled topics. To
+%furnish stakeholders with labelled topics, our technique provides many methods of automatic topic labelling.
 

### 2011-08-23 ese/ese.tex
updating pg timeline plot

@@
 %\noindent \textbf{Creating a validation corpus} ---
-
-%FAT
-%\newpage
 
@@
 
-%FAT we can remove this
-%\balance
-%LFAT
-\newpage
 \begin{quotation}
@@
   & Total  Topics  & 1000 & 1000 & 1000 \\
-
-\todo{update with PGSQL}
+\midrule
+PgSQL 3.23  & Named Topics   & - & - & - \\
+  & Unnamed Topics & - & - & - \\
+  & Total  Topics  & - & - & - \\
+
 \bottomrule
@@
 \end{table}
-
+\todo{update with PGSQL}
 % was best performing
@@
 \end{figure*}
-
-%FAT 
-%\newpage
 
@@
 
-%FAT
-
-%\newpage
-
 %\subsection{Timeline analysis}
-Figures \ref{fig:mysql-timeline} and \ref{fig:maxdb-timeline} show the
+Figures \ref{fig:mysql-timeline} and \ref{fig:maxdb-timeline} and \ref{fig:pgsql-timeline} show the
 temporal patterns of NFR frequencies.
@@
 
-\begin{figure*}[t]
+\begin{figure}[t]
 \centering
@@
 }
+
 \subfloat[PostgreSQL 7.2]{					
@@
 \label{fig:timelines}
-\end{figure*}
+% Timelines created using src/validate/summary_labels.py
+\end{figure}
 
@@
 
-%LFAT
-\balance
-\newpage
 \subsection{Do Projects Differ in their Relative  Topic Interest?}
@@
 
-%FAT
-%\newpage
 \subsection{Summary of Techniques}
@@
 did not rate the same documents.
-
-%LFAT
-\balance
-\newpage
 

### 2011-08-23 src/validate/plot_timeline.py

@@ -48,7 +48,7 @@ def plot_timeline(proj, period_map):
     if proj == 'maxdb':
         ax.set_xlim(datetime.date(2004,06,29),datetime.date(2006,6,19))#5*len(period_map.keys()
     if proj == 'pgsql':
-        ax.set_xlim(datetime.date(2002,02,7),datetime.date(2004,6,28))
+        ax.set_xlim(datetime.date(2002,02,1),datetime.date(2004,6,28))
 
     #ax.set_xlabel('Date')
     ax.set_yticks([2.5,7.5,12.5,17.5,22.5,27.5,32.5,37.5])
@@ -120,7 +120,7 @@ def plot_timeline(proj, period_map):
         if proj == 'maxdb':                                                                            
             l = Line2D([datetime.date(2004,06,29),datetime.date(2006,6,19)],[5*i,5*i]  ,lw = 0.5,  color='black')#5*len(period_map.keys()
         if proj == 'pgsql':
-            l = Line2D([datetime.date(2002,02,7),datetime.date(2004,6,28)],[5*i,5*i]  ,lw = 0.5,  color='black')#5*len(period_map.keys()
+            l = Line2D([datetime.date(2002,02,1),datetime.date(2004,6,28)],[5*i,5*i]  ,lw = 0.5,  color='black')#5*len(period_map.keys()
         ax.add_line(l)                              
 
     # label the model with key events
@@ -135,7 +135,7 @@ def plot_timeline(proj, period_map):
         events_rel = ["7.5.00.15", "7.5.00.23", "7.5.00.24 PHP","7.6.00 prod.","7.5.00.34"]
     if proj == 'pgsql':
          events_dates = [datetime.date(2002,02,04), datetime.date(2002,11,27), datetime.date(2003,11,17), datetime.date(2004,6,1) ]
-         events_rel = ["7.2", "7.3", "7.4", "7.5/8.0 alpha"]
+         events_rel = ["7.2", "7.3", "7.4", "7.5/8"]
         
     l = Line2D(events_dates,[37.5],ls=' ',marker='o')
     ax.add_line(l)

### 2011-08-23 src/validate/summary_labels.py

@@ -9,8 +9,8 @@ import lxml.etree
 import datetime
 
 if __name__ == '__main__':
-    maxdb_file = 'pg-label-summary.csv'
-    maxdb = open(maxdb_file,'w')
+    #maxdb_file = 'pg-label-summary.csv'
+    #maxdb = open(maxdb_file,'w')
     
     exps = get_exps.get_exps() 
     # test_period = '1098931753'

### 2011-08-23 ese/ese.tex
added some text on pg numbers and timeline

@@
 We use \textit{labelled topic extraction} to pick out the underlying NFR activity behind these events. 
-For example, both projects show a high number of NFRs recognized at the first period of analysis. 
+For example, both MaxDB and MySQL show a high number of NFRs recognized at the first period of analysis. 
 This is due to our window choice: we deliberately targeted our
@@
 5.0).
-\todo{add PGSQL disucssion here. Schema additions? Lots of functionality?}
+
+By contrast, the PostgreSQL timeline  (Figure \ref{fig:pgsql-timeline}) is extracted from a central trunk and is not version-specific. Therefore development 
+tends to be focused on releases 7.3, 7.4 and 7.5/8.0 alpha. For example, a priority for the 8.0 candidate was a Windows-native port of the source code, which seems to 
+correlate with the \emph{portability} NFR increasing in frequency in mid-2004. Before the 7.4 release, \emph{usability,functionality} and \emph{portability} all 
+increase in frequency, possibly reflecting the interest in adding features and documentation for the release.
 
@@
 %\subsubsection{Do label frequencies change over time?}
-\todo{add more on PGSQL}
-In both projects the frequencies generally decreased with age. 
+%\todo{add more on PGSQL}
+In both MaxDB and MySQL the frequencies generally decreased with age. 
 However, there are variations within our NFR labels. In MySQL, \emph{usability} and \emph{efficiency} do not appear very often in topics. 
@@
 
+In PostgreSQL, by comparison, the frequencies seem to by cyclic, since we are not studying a maintenance-phase for the product, but rather ongoing
+feature addition and usability improvements.
+
 \subsection{Do Projects Differ in their Relative  Topic Interest?}
-%\subsubsection{Do  projects differ in their relative  topic interest?}
-%\noindent\textbf{2. Do  projects differ in their relative  topic interest?} ---
 Yes. MySQL 3.23 had proportionally more
@@
 how often portability arises as a concern.
-%On one hand, this analysis is unsurprising; on the other hand, it provides confirmation that our approach identifies actual patterns.
-%\subsection{Summary}
-
-%\todo[inline]{link the actual data, which is interesting, to somethign
- % actionable for a developer or manager}
+\todo{add the pgsql data}
 
 \subsection{Do different developers work on different NFRs?}
-\todo{figure this out!}
+We did a t-test to assess whether there was statistically significant differences in the types of NFRs developers chose to work on, for the PostgreSQL project only. 
+\todo{add more, figure}
 
@@
 
-%XXX can just remove this
-%\subsection{Effectiveness}
 \subsection{Annotation Observations}

### 2011-08-24 src/validate/get_exps.py
trying to combine both pgsqla and pgsqln

@@ -33,8 +33,11 @@ def get_exps():
     maxdb_dir = "../../data/maxdb-tagged/"
     mysql_dir = "../../data/mysql-tagged/"
     pg_dir = "../../data/pgsql-tagged/"
+    pga_dir = "../../data/pgsqla-tagged/"
+    pgn_dir = "../../data/pgsqln-tagged/"
 
-    exprs = [ ["maxdb",maxdb_dir, maxdb_periods] , [ "mysql",mysql_dir, mysql_periods ], ["pgsql",pg_dir, pgsql_periods]]
+    #exprs = [ ["maxdb",maxdb_dir, maxdb_periods] , [ "mysql",mysql_dir, mysql_periods ], ["pgsql",pg_dir, pgsql_periods]]
+    exprs = [ ["maxdb",maxdb_dir, maxdb_periods] , [ "mysql",mysql_dir, mysql_periods ], ["pgsqla",pg_dir, pgsql_periods] , ["pgsqln",pg_dir, pgsql_periods]]]
     return exprs
 
 __exp1_wordlist__ = [

### 2011-08-24 src/validate/run-weka-classifier-single-trials.sh

@@ -1,8 +1,10 @@
 #!/bin/bash
 # We make experimental make files
 # We run each annotation against each classifier
-CPUS=4
-perl make-single-class-arff-files.pl pgsql && make -f Makefile.pgsql -j $CPUS all
+CPUS=3
+perl make-single-class-arff-files.pl pgsqla && make -f Makefile.pgsql -j $CPUS all
+perl make-single-class-arff-files.pl pgsqln && make -f Makefile.pgsql -j $CPUS all
+#perl make-single-class-arff-files.pl pgsql && make -f Makefile.pgsql -j $CPUS all
 perl make-single-class-arff-files.pl mysql && make -f Makefile.mysql -j $CPUS all
 perl make-single-class-arff-files.pl maxdb && make -f Makefile.maxdb -j $CPUS all
-perl make-single-class-arff-files.pl maxdb__mysql && make -f Makefile.maxdb__mysql -j $CPUS all
+#perl make-single-class-arff-files.pl maxdb__mysql && make -f Makefile.maxdb__mysql -j $CPUS all

### 2011-08-24 ese/ese.tex
added IRR discussion to paper

@@
 We did not attempt to match our results to any particular model.
-We assessed inter-rater reliability (IRR), a measure of the correlation between the manual annotations the first two authors created. Perfect IRR (1.0) indicates complete agreement 
-on the appropriate annotation for that topic. We verified IRR using manually annotated PostgreSQL topics, and achieved a result of \todo[inline]{insert here}. 
+We assessed inter-rater reliability (IRR) using Cohen's Kappa to measure the level of agreement between the manual 
+annotations the first two authors created, under some model of random agreement. Perfect IRR (1.0) indicates 
+complete agreement on the appropriate annotation for that topic. We verified IRR using manually annotated PostgreSQL topics, 
+and achieved a result of 0.108 (slight agreement), with a correlation coefficient of 0.108. 
 Some of the reasons for this are a disagreement over the appropriate classification for documentation topics (should this be considered
-usability or portability?). A short education session on what a given NFR from the ISO standard encompasses would reduce this error rate.
+usability or portability?), and a disagreement on functionality as a software quality. A short education session on the meaning of a given NFR from 
+the ISO standard would reduce this error rate.
+In general there is disagreement over what constitutes a non-functional requirement \cite{Glinz_2007}.
 

### 2011-08-25 ese/ese.tex
some interrater reliability

--- 2011-08-23 ese/ese.tex
@@
   Dept. of Computer Science\\
-  University of California, Davis \\
-  Davis, CA, USA\\
+  University of Alberta\\
+  Edmonton, AB, CANADA\\
   \email{abram@softwareprocess.es}
@@
 
+*** We also clustered the authors by similarity and we found that
+among some of the most prolific postgresql authors, their NFR focus
+was distinctly different *** Some confounds are that postgresql
+authors commit the patches of others.
+
 \section{Discussion}
@@
 Project specific word-lists would avoid automated mistakes due to the names of entities and modules of a software project.
+
+\subsection{Inter-rater reliability}
+
+Authors Neil and Abram each annotated the PostgreSQL topics in order
+to determine inter-rater reliability. While when they read each others
+annotations they generally agreed, the inter-rater reliability did not
+bear this out.
+
+%>    A_portability   A_functionality     A_reliability A_maintainability
+%>      0.153548387      -0.013588517       0.004573794       0.082001031
+%>     A_efficiency       A_usability            A_none
+%>      0.231496539       0.009287926       0.062133646
+
+% > Spearman says
+% >   A_portability   A_functionality     A_reliability A_maintainability
+% >      0.253037469      -0.014415214       0.004579664       0.082215613
+% >     A_efficiency       A_usability            A_none
+% >      0.258037216       0.014436230       0.080667468
+
+\begin{table}
+\begin{tabular}{l|c|c}
+\hline
+            & Cohen's Kappa & Spearman Correlation \\ \hline
+Portability & 0.154 & 0.253  \\
+Functionality & -0.014 & -0.014 \\
+Reliability & 0.005 & 0.005 \\
+Maintainability & 0.082 & 0.082 \\
+Efficiency & 0.231 & 0.258 \\
+Usability & 0.009 & 0.014 \\
+None &      0.062 & 0.081 \\ \hline 
+Everything & 0.107 & 0.108 \\ \hline
+\end{tabular}
+\caption{Inter-rater Reliability on PostgreSQL}
+\label{tab:interr}
+\end{table}
+
+We feel these results are poor. The aggregate view of with a Kappa of
+$0.1$ indicates there is some agreement but not a lot. We found that
+there was a lot of agreement in terms of lack of a tag, but often more
+disagreement when a tag should be applied.
+
+After some discussion we concluded that usability was a source of
+disagreement. Do user manuals improve usability? Is a command-line
+option a usability issue? These kinds of questions illustrate some of
+the agreement, disagreement and ambiguity about these labels. Further
+more, both annotators found that the annotation was time-consuming and
+difficult.
+
+% I think this is largely due to lack of training.
+So I will write it up as such - as an interesting point to say that agreeing on the meaning of "software quality" is difficult, etc.
+Ideally we would now do a sample training session where we go over say one period and determine the appropriate annotations, then recalculate Kappa to see what changes.
+
+% The other tack is to ignore the IRR data completely as we did in the workshop. But I don't think it hurts our results that much ¦ just confirms the difficulty of annotation.
+
+% Neil
 
@@
 We assessed inter-rater reliability (IRR), a measure of the correlation between the manual annotations the first two authors created. Perfect IRR (1.0) indicates complete agreement 
-on the appropriate annotation for that topic. We verified IRR using manually annotated PostgreSQL topics, and achieved a result of \todo[inline]{insert here}. 
+on the appropriate annotation for that topic. We verified IRR using
+manually annotated PostgreSQL topics, and achieved a result of $0.1$.
 Some of the reasons for this are a disagreement over the appropriate classification for documentation topics (should this be considered

### 2011-08-25 ese/ese.tex
neil's changes to ese.tex

--- 2011-08-24 ese/ese.tex
@@
 seems like a reasonable starting point for our work.}.
-%There is some debate about the terms in this model. 
-%However, ISO9126 is ``an international standard and thus provides an 
-%internationally accepted terminology for software quality \cite[p. 58]{Boegh2008},'' that is sufficient for the purposes of this research.  
+There is some debate about the terms in this model. 
+However, ISO9126 is ``an international standard and thus provides an 
+internationally accepted terminology for software quality \cite[p. 58]{Boegh2008},'' that is sufficient for the purposes of this research.  
 We claim that these NFRs are maintenance concerns (to varying degrees) in all software projects, and are therefore well-suited for comparisons between
@@
 
-%\noindent \textbf{Creating a validation corpus} ---
-
 \subsubsection{Creating a Validation Corpus}
 To evaluate both semi-unsupervised and supervised classification, we
-created a validation set of manually labelled topics. For MySQL 3.23
-and MaxDB 7.500, the annotators (the first two authors) annotated each extracted topic in each period with
+created a validation set of manually labelled topics. For each project, 
+the annotators (the first two authors) annotated each extracted topic in each period with
 the six NFR labels listed above.
@@
 
-%\subsection{Generating word lists}
-%\label{sec:genwordlist}
-% %\todo[inline]{Make it clear where the data comes from! Neil says: Also, I
-%   didn't get the impression they understood that we weren't generating
-%   labels from the data, but rather from independent taxonomies. Of
-%   course if you use labels from the data it isn't comparable. But the
-%   point of ISO 9126 etc. was to avoid exactly this problem.}
-
-% \todo[inline]{Need to hammer it in that the word bags are not project specific and
-%  are inherited so it isn't a supervised method}
-
-%Topics are word distributions, lists of words ranked by frequency, that are burdensome to interpret and hard to distinguish and understand. 
-%While the topics themselves are generated automatically, how to interpret these topics is less clear.
-%For example,
-
 \subsubsection{Generating Word Lists}
-
-%\noindent\textbf{Generating word lists} ---
 
@@
 
-%\todo[inline]{* clarify the mailinglist stuff we did}
-%\todo[inline]{This part doesn't make a lot of sense}
 To construct these expanded word-lists, we used
@@
 
-%\todo[inline]{NEIL: this could be a subsubsection}
-%\todo[inline]{* the hierarchy of tags was not clear}
-%\noindent \textbf{Generating labels automatically} ---
-
 Using our three word-lists (\textsf{exp1}, \textsf{exp2}, \textsf{exp3}), we labelled our topics with an NFR where there was a match between a word in
@@
 requirements.
-All experiments were run on MaxDB 7.500 and MySQL 3.23 data. LDA
+All experiments were run on the datasets for each project (e.g., PostgreSQL, MySQL, MaxDB). LDA
 extracted $20$ topics per period for each project.
-% Is this good enough
 This labelling is \emph{semi-unsupervised} because the corpus is not derived from 
@@
 ourselves for a training set. The motivation behind this technique is that
-because most software often addresses similar issues, thus we can use the
+because most software often addresses similar issues, we can use the
 domain knowledge of software to label relevant topics.
-%\todo[inline]{Neil does this explain the unsupervised aspect well enough?}
 

### 2011-08-25 ese/ese.tex

@@
 Project specific word-lists would avoid automated mistakes due to the names of entities and modules of a software project.
+
+\subsection{Inter-rater reliability}
+
+Authors Neil and Abram each annotated the PostgreSQL topics in order
+to determine inter-rater reliability. While when they read each others
+annotations they generally agreed, the inter-rater reliability did not
+bear this out.
+
+%>    A_portability   A_functionality     A_reliability A_maintainability
+%>      0.153548387      -0.013588517       0.004573794       0.082001031
+%>     A_efficiency       A_usability            A_none
+%>      0.231496539       0.009287926       0.062133646
+
+% > Spearman says
+% >   A_portability   A_functionality     A_reliability A_maintainability
+% >      0.253037469      -0.014415214       0.004579664       0.082215613
+% >     A_efficiency       A_usability            A_none
+% >      0.258037216       0.014436230       0.080667468
+
+\begin{table}
+\centering
+\begin{tabular}{l|c|c}
+\hline
+            & Cohen's Kappa & Spearman Correlation \\ \hline
+Portability & 0.154 & 0.253  \\
+Functionality & -0.014 & -0.014 \\
+Reliability & 0.005 & 0.005 \\
+Maintainability & 0.082 & 0.082 \\
+Efficiency & 0.231 & 0.258 \\
+Usability & 0.009 & 0.014 \\
+None &      0.062 & 0.081 \\ \hline 
+Everything & 0.107 & 0.108 \\ \hline
+\end{tabular}
+\caption{Inter-rater Reliability on PostgreSQL}
+\label{tab:interr}
+\end{table}
+
+Table \ref{tab:iterr} describes the Cohen Kappa and the Spearman
+correlation of our per topic annotations for each tag. We had to
+evaluate inter-rater reliability this way because 1 topic could be
+tagged with more than 1 annotation.
+
+We feel these results are poor. The aggregate view of with a Kappa of
+$0.1$ indicates there is some agreement but not a lot. We found that
+there was a lot of agreement in terms of lack of a tag, but often more
+disagreement when a tag should be applied.
+
+After some discussion we concluded that usability was a source of
+disagreement. Do user manuals improve usability? Is a command-line
+option a usability issue? These kinds of questions illustrate some of
+the agreement, disagreement and ambiguity about these labels. Further
+more, both annotators found that the annotation was time-consuming and
+difficult.
+
+% I think this is largely due to lack of training.
+So I will write it up as such - as an interesting point to say that agreeing on the meaning of "software quality" is difficult, etc.
+Ideally we would now do a sample training session where we go over say one period and determine the appropriate annotations, then recalculate Kappa to see what changes.
+
+% The other tack is to ignore the IRR data completely as we did in the workshop. But I don't think it hurts our results that much ¦ just confirms the difficulty of annotation.
+
+% Neil
 
@@
 and achieved a result of 0.108 (slight agreement), with a correlation coefficient of 0.108.
+
 Some of the reasons for this are a disagreement over the appropriate classification for documentation topics (should this be considered

### 2011-08-25 ese/ese.tex
reverted to abram's IRR work
fixed some wording to add pgsql

--- 2011-08-25 ese/ese.tex
@@
 To evaluate both semi-unsupervised and supervised classification, we
-created a validation set of manually labelled topics. For MySQL 3.23
-and MaxDB 7.500, the annotators (the first two authors) annotated each extracted topic in each period with
+created a validation set of manually labelled topics. For each project,
+ the annotators (the first two authors) annotated each extracted topic in each period with
 the six NFR labels listed above.
@@
 requirements.
-All experiments were run on MaxDB 7.500 and MySQL 3.23 data. LDA
+All experiments were run on the datasets for each project (e.g., PostgreSQL, MySQL, MaxDB). LDA
 extracted $20$ topics per period for each project.
@@
 ourselves for a training set. The motivation behind this technique is that
-because most software often addresses similar issues, thus we can use the
+because most software often addresses similar issues, we can use the 
 domain knowledge of software to label relevant topics.
@@
 \begin{table}
+\centering
 \begin{tabular}{l|c|c}
@@
 \end{table}
+
+Table \ref{tab:iterr} describes the Cohen Kappa and the Spearman
+correlation of our per topic annotations for each tag. We had to
+evaluate inter-rater reliability this way because 1 topic could be
+tagged with more than 1 annotation.
 

### 2011-08-26 ese/ese.tex
updated Named/Unnamed table

@@
 %
+
 \RequirePackage{fix-cm}
+
 %
+
 %\documentclass{svjour3}                     % onecolumn (standard format)
+
 %\documentclass[smallcondensed]{svjour3}     % onecolumn (ditto)
+
 \documentclass[smallextended]{svjour3}       % onecolumn (second format)
+
 %\documentclass[twocolumn]{svjour3}          % twocolumn
+
 %
+
 \smartqed  % flush right qed marks, e.g. at end of proof
+
 %
+
 \usepackage{graphicx}
+
  \usepackage{booktabs}
+
 \usepackage[caption=false]{subfig}
+
  \usepackage[colorinlistoftodos,backgroundcolor=yellow]{todonotes} %disable
+
  \usepackage{color}
@@
 \newcommand{\XXX}[1]{\textcolor{red}{{\it \textbf{[XXX: #1]}}}}
+
 \hyphenation{Max-DB data-base}
@@
 \usepackage{comment}
+
 \newtheorem{mydef}{definition}
+
 \newcommand{\slbl}[1]{\textbf{\textsf{#1}}}
+
 \usepackage{multirow}%,multicolumn}
+
 \usepackage{slashbox}
+
 \usepackage{wrapfig}
+
 \usepackage{booktabs, colortbl}
+
 \usepackage{amsmath}
+
 \usepackage{url}
+
 \usepackage[pdfstartview=FitH,bookmarksopenlevel=3,bookmarks=true]{hyperref} %pdfpagescrop={92 112 523 778},
+
 \journalname{Empirical Software Engineering}
@@
 \title{Automated Topic Naming
+
 %\thanks{Grants or other notes
+
 %about the article that should go on the front page should be
+
 %placed here. General acknowledgments should be placed at the end of the article.}
+
 }
+
 \subtitle{Supporting Cross-project Analysis of Software Maintenance Activities}
+
 %\titlerunning{Short form of title}        % if too long for running head
@@
 \author{
+
 Abram Hindle \and
+
 Neil A. Ernst \and
+
 Michael W. Godfrey \and
+
 John Mylopoulos
+
 }
+
 %\authorrunning{Short form of author list} % if too long for running head
@@
 \institute{
+
   Abram Hindle \at
+
   Dept. of Computer Science\\
+
   University of Alberta\\
+
   Edmonton, AB, CANADA\\
+
   \email{abram@softwareprocess.es}
+
 \and
+
   Neil A. Ernst \at
+
   Dept. of Computer Science\\
+
   University of Toronto\\
+
   Toronto, Ontario, CANADA\\
+
   \email{nernst@cs.toronto.edu}
+
 \and
+
   Michael W. Godfrey \at
+
   David Cheriton School of Computer Science\\
+
   University of Waterloo\\
+
   Waterloo, Ontario, CANADA\\
+
   \email{migod@uwaterloo.ca}
+
 \and
+
   John Mylopoulos \at
+
   Dept. Information Eng. and Computer Science\\
+
   University of Trento\\
+
   Trento, ITALY\\
+
   \email{jm@disi.unitn.it}
+
 }
@@
 \date{Received: date / Accepted: date}
+
 % The correct dates will be entered by the editor
@@
 \begin{abstract}
+
 % %\todo[inline]{Topics and LDA sooner, Topics/concept location is a known
+
 Researchers have employed a variety of techniques to extract underlying
+
 topics that relate to software development artifacts.  Typically, these
+
 techniques use semi-unsupervised machine-learning algorithms to suggest
+
 candidate word-lists.  However, word-lists are
+
 difficult to
+
 interpret in the absence of meaningful summary labels.  Current topic modeling
+
 techniques assume manual labelling and do not use domain-specific knowledge
+
 to improve, contextualize, or describe results for the developers.
+
 We propose a solution: \emph{automated labelled topic extraction}.
+
   Topics are extracted
+
 using  Latent Dirichlet Allocation (LDA) from commit-log comments recovered
+
 from source control systems such as CVS and BitKeeper.
+
 These topics are given labels from a generalizable
+
 cross-project taxonomy, consisting of non-functional
+
 requirements. Our approach was evaluated with experiments and case studies
+
 on three large-scale RDBMS projects: MySQL, PostgreSQL and MaxDB.  The case studies
+
 show that labelled topic extraction can produce appropriate,
+
 context-sensitive labels relevant to these projects, which provides fresh
+
 insight into their evolving software development activities.
+
 %160 words
+
 \keywords{ Software maintenance \and Repository mining \and Latent dirichlet allocation \and Topic models}
@@
 \section{Introduction}
+
 A key problem for practising software maintainers is gaining an
+
 understanding of \emph{why} a system has evolved the way it has.
+
 This is different than \emph{how} a system has evolved, because the behaviour of change itself is the
+
 \emph{how}.
+
 Looking back on streams of artifacts scattered across different
+
 repositories, inferring what activities were performed, when, and for
+
 what reasons, is hard to do without expert advice from the developers
+
 involved.
+
 %new stuff
+
 % edit this. awkward
+
 In this work we seek to provide a method of automatically labelling  development topics extracted from commit logs.
@@
 Topic modeling is a machine learning technique which creates
+
 multinomial distributions of words extracted from a text corpus.
+
 This technique infers the hidden structure of a corpus using posterior
+
 inference: the probability of the hidden structure given the data.
+
 Topic models are useful in software maintenance because they summarize
+
 the key concepts in a corpus, such as source code, commit comments, or
+
 mailing-list messages, by identifying statistically co-occurring words.
+
 Among other uses, it can quickly give developers an overview of where significant
+
 activity has occurred, and gives managers or maintainers a sense of
+
 project history.
+
 %A major challenge for topic modeling is to accurately identify the
+
 %summary label for a topic; in the previous case, perhaps
+
 %\emph{``event-handling''} would be appropriate.
+
 %Current approaches are manual: they rely on human expertise to
+
 %recognize the appropriate label.
+
 %Furthermore, to date the results of topic modeling are project
+
 %specific, and not generalizable.
+
 %This paper presents an approach, \emph{labelled topic extraction}, which addresses these two problems.
@@
 %migod
+
 While machine learning techniques can automatically identify clumps of
+
 commonly recurring terms, devising an appropriate summary label for
+
 each clump/topic is harder.
+
 A given topic extracted from a set of commit logs might consist of the following terms: \emph{ ``listener change remove add fire''}.
+
 This topic might reasonably be labelled as
+
 ``\emph{event handling}'' by a developer who understands the domain well,
+
 despite the fact that this label does not appear in the word list itself.
+
 Current approaches to topic labelling rely on manual intervention by
+
 human experts, and also are limited to project-specific topic labels.
+
 In this paper, we introduce \emph{labelled topic extraction}, an
+
 approach to topic labelling that creates labels automatically that are
+
 project independent.
+
 %is both automatic and creates labels that are project independent.
@@
 % Changes to software systems, revisions, do not occur without
+
 % reason.
+
 % The reason or purpose is often related to requirements
+
 % Often one revision might have multiple purposes ranging from
+
 % maintenance issues, quality issues, requirements, features or a
+
 % response to a bug report.
+
 % These purposes often are exhibited as latent topics that relate to
+
 % many development artifacts such as:
+
 %  version control revisions, bug reports, and mailing-list discussions.
@@
 %In our previous work we dealt with topic trends, which are topics that recur over time~\cite{Hindle09ICSM}.
+
 %We observed that topic trends were often non-functional requirements (NFRs).
+
 %NFRs have the property of being cross-domain and widely applicable.
+
 %In this sense, they are useful abstractions for developer conversations about different software projects.
@@
 % harp on data mining more
+
 %In general, the mining of software artifacts tends to be very project specific, yet NFRs are not.
+
 % I was told if you're going to use a paren, you don't need to say it.
+
 %There is a series of standards on NFRs, such as ISO9126~\cite{iso9126}, specifically intended to apply to projects of varying types.
+
 %These standards indicate our approach is possible. %, if perhaps not universally accepted).
+
 % NFRs are prevalent in almost all software systems.
+
 % In this paper we will show that NFR related topics occur quite frequently.
+
 %XXXXXX this sentence is confusing
+
 %This informs our method of generalizing the data mining of these software artifacts to leverage the global domain knowledge of software engineering,
+
 %specifically NFRs.
+
 %In particular, we use software quality models to generalize topics across projects.
+
 %This allows us to mine software systems with the intention of comparing their NFR-related development topics.
+
 % in order to achieve generalizable, cross project data mining results.
@@
 %%%% THIS PARAGRAPH ARGUES ABOUT THE CONCRETE APPLICATION OF
+
 %%%% TOPIC MODELLING AND LABELLED TOPIC EXTRACTION
+
 %%%% IT NEEDS TO BE TIGHTER BUT IT ALSO NEEDS TO BE HERE
@@
 % AH: Topics take a long to read
+
 % AH: Project Dashboards, summarization, software process recovery
@@
 %migod
+
 In general, the fruits of mining software artifacts are often project
+
 specific and hard to generalize.
+
 However, in our previous work we
+
 investigated \emph{topic trends} --- that is, topics that recur over
+
 time --- we observed that topic trends often corresponded to
+
 non-functional requirements (NFRs)~\cite{Hindle09ICSM}, which is
+
 further emphasized in this paper due to the large numbers of NFR
+
 labelled topics.
+
 This is encouraging, as NFRs have the property of being cross-domain
+
 and widely applicable.
+
 In this sense, they are useful abstractions for developer
+
 conversations about different software projects.
+
 Furthermore, there is a series of standards on NFRs, such as ISO9126 \cite{iso9126}, that are specifically intended to apply to projects of varying
+
 types; this suggests that our goal of trying to extract NFR-related development topics, such as those related to software quality models, holds promise.
@@
 Concrete applications of topics and \emph{labelled topic extraction}
+
 range from project dashboards to annotating software artifacts such as revisions and bug reports with
+
 NFR-related tags.
+
 Project dashboard~\cite{dashboard} are typically employed by managers and are used to
+
 provide quick summaries of the effort put into a software project. In this case,
+
 labelled topic extraction would allow managers to track effort
+
 related to NFR topics, such as portability.
+
 These techniques also allow for the annotation of commit comments
+
  and other software artifacts with NFRs.
+
 This would enable querying of bug reports and artifacts by relevant NFRs.
+
 For instance a manager can confirm if their developers were
+
 focused on \emph{usability} by looking for
+
  \emph{usability}-relevant revisions and bug reports.
@@
 In this paper, we describe \emph{automated labelled topic extraction}. It addresses two gaps in the topic mining literature:
+
 \begin{enumerate}
+
   \item Topic mining of software has been limited to one project at a time.
+
 This is because traditional topic mining techniques are specific to a particular data-set.
+
 \textit{Automated labelled topic extraction} allows for comparisons \textit{between} projects.
+
   \item Topic modeling creates word lists that require interpretation by the user to assign meaning.
+
 Like (1), this means
+
 that it is difficult to discuss results independent of the project context.
+
 Our technique automatically, or with some initial training, assigns labels across projects.
+
 \end{enumerate}
@@
 This paper makes the following contributions:
+
 \begin{itemize}
+
 \item introduces the concept of labelled topic extraction, using a non-functional requirements (NFR) taxonomy for our labels;
@@
 %\item evaluates an unsupervised method of automated topic labelling (word-lists);
+
 %%\item Our unsupervised method can be applied to projects without training;
+
 %%\item
+
 %evaluates a supervised method of automatically labelling topics with a single NFR (machine learning);
+
 %%\item
+
 %evaluates a supervised method of labelling topics with multiple NFRs
+
 %(multi-label machine learning);
@@
 \item evaluates three kinds of automatic topic labelling methods:
+
   semi-unsupervised labelling  of topics (word-lists), supervised labelling of
+
   topics with a single NFR (machine learning),
+
   and supervised labelling of topics with multiple NFRs (multi-label
+
   machine learning);
@@
 %  of automated topic labelling (word-lists);
+
 %\item Our unsupervised method can be applied to projects without training;
+
 %\item
+
 %evaluates a supervised method of automatically labelling topics with a single NFR (machine learning);
+
 %\item
+
 %evaluates a supervised method of labelling topics with multiple NFRs (multi-label machine learning);
@@
 \item examines how NFRs correlate with the work of individual developers;
+
 \item provides a method of cross-project analysis via topic labelling; and
+
  applies these techniques to visualize NFRs over time, and to analyze maintenance activities.
@@
 %We first introduces labelled topic extraction, and then uses that technique to analyze development activity in two large-scale examples.
+
 We begin by discussing related work in Section \ref{sec:related}.
+
 Next, we describe how we generated our data (Section \ref{sec:wordlist}). For semi-unsupervised classification (Section \ref{sec:unsuplabelling}), we
+
 begin by creating word-lists to signify when a topic matches an NFR label. We then apply our classifier and analyze the results.
+
 %and how we created the word-lists used in unsupervised classification.
+
 %We show that labels with their topics can be learned and used to classify other data-sets, either without training
+
 %Our goal is to label a given topic with either an NFR from our taxonomy,
+
  %Section \ref{sec:unsuplabelling} presents our technique for using word-lists to do unsupervised classification, and describes the results.
+
 In Section \ref{sec:suplabelling}, we manually annotate the topics, and use those annotations as training data for supervised classification.
+
 %We then present visualizations of named topics and their trends over time to aid communication and analysis.
+
 To demonstrate an application of labelled topic extraction, we use an exploratory case study of two open source database systems to show how named
+
 topics can be compared between projects  (Section \ref{sec:analysis}).
+
 The paper concludes with a discussion of limitations (Section \ref{sec:limit}), and future work.
@@
 \section{Previous Work}
+
 \label{sec:related}
@@
 The idea of extracting higher-level \emph{concerns} and \emph{topics}, also known as
+
  \emph{concepts}, \emph{aspects} or \emph{requirements},
+
 has been approached from documentation-based and repository-based
+
 perspectives.
@@
 Cleland-Huang and her colleagues have investigated mining requirements documents for non-functional requirements (NFR) (quality
+
 requirements)~\cite{Cleland-Huang2006}.
+
 Their approach is similar to ours, as they mined keywords from NFR
+
 catalogues.
+
 They differ because they mine requirements documents where as we mine revisions.
+
 %One approach they tried was similar to this one, with keywords mined from NFR catalogues found in their previous work~\cite{chung99}.
+
 They demonstrated a recall of $80\%$ with precision of $57\%$ for the \emph{security} NFR, but could not find a reliable source of keywords for other NFRs.
+
 Instead, they developed a supervised classifier by using human experts to identify an NFR training set.
+
 Our research differs because we use a more comprehensive set of terms based on the taxonomy we chose.
+
 Another difference is that we make cross-project comparisons instead of focusing
+
 on a single project.
+
 %They relied on requirements documents instead of version control
+
 %histories that we use.
+
 %The objective of Cleland-Huang's study was to identify new NFRs for system development,
+
 %yet our objective was to recover those latent NFRs from commit-log messages of the project.
@@
 %There are several reasons we did not follow this route.
+
 %One, we believe we have a more comprehensive set of terms due to the taxonomy we chose.
+
 %Secondly, we wanted to compare across projects.
+
 %Their technique was not compared across different projects and the applicability of the training set to different corpora is unclear.
+
 %A common taxonomy allows us to make cross-project comparison.
+
 %Although subject to the assumption that all projects conceive of these terms in the same way.
+
 %Thirdly, while the objective of Cleland-Huang's study was to identify new NFRs (for system development)
+
 %our study assumes these NFRs are latent in the commit-log messages of the project.
@@
 Similarly, Mockus and Votta~\cite{Mockus00} studied a large-scale industrial change-tracking system.
+
 Mockus and Votta leveraged WordNet~\cite{Fellbaum1998}, an English-language ``lexical database'' that contains semantic relations between words,
+
 including common related forms (similar to word stemming), meronymy and synonymy.
+
 They used WordNet for word roots as they felt the synonyms would be
+
 non-specific and cause errors.
+
 %, problems we also encountered.
+
 Mockus et al. validated their labels with system developers.
+
 Since we study multiple projects, instead of a single project, these
+
 kind of interviews are somewhat infeasible (particularly in the distributed world of open-source software).
@@
 Another approach is to extract concerns from software repositories.
+
 Marcus et al.~\cite{marcus04wcre} use Latent Semantic Indexing (LSI)
+
 to identify commonly occurring concerns for software maintenance.
+
 The
+
 concerns are given by the user, and LSI is used to retrieve them from
+
 a corpus.
+
 Topic modelling generates topics that are independent of a user query, and relate only to word frequencies in the corpus.
+
 %Some results were interesting, but their precision was quite low.
+
 With ConcernLines, Treude et al.~\cite{treude09cl} show tag occurrence
+
 using colour and intensity.
+
 They mine
+
 developer created tags
+
 %change request tags (such as `milestone 3') and used these
+
 in order to analyze the evolution of a single product.
+
 % make evolutionary analyses of a single product.
+
 The presence of a well-maintained set of tags is obviously essential to the success of this technique.
@@
 %XXXXXXXXXXXX OPTIONAL
+
 % Not sure we need this
+
 %Mens et al.~\cite{mens08icsm} conducted an empirical study of Eclipse, a popular Java IDE, to verify the claims of Lehman~\cite{lehman97sms}.
+
 % They concerned themselves with source code only, and found Law Seven, ``Declining Quality'', to be too difficult to assess:
+
 % ``[we lacked an] appropriate measurement of the evolving quality of the system as perceived by the users \cite[p. 388]{mens08icsm}''.
+
 %This paper examines the notions of quality in terms of a consistent ontology, as Menset al. call for in their conclusions.
@@
 % in our previous work~\cite{Hindle09ICSM}, as well as -- CITED a little too often (vanity cites!)
+
 In Baldi et al.~\cite{Baldi2008}, topics are named manually: human
+
 experts read the highest-frequency members of a topic and assign a
+
 label accordingly.
+
 %XXX Possible to cut this
+
 As discussed earlier, given the topic \emph{``listener change remove add fire''}, Baldi et al. would assign the label \emph{event-handling}.
+
 The labels are reasonable enough, but still require an expert in the field to determine them.
+
 Furthermore, these labels are project-specific, because they are generated from the data of that project. E.g., we might have a label called \emph{`Oracle'}
+
 in the MySQL case, since Oracle owns MySQL.
+
 %We do this differently.
+
 %This paper does this differently.
+
 Our approach differs: first of all, we automate the process of naming the topics; secondly, we label topics with project-independent terms, in order
+
 to permit cross-project comparison.
@@
 Mei et al.~\cite{Mei2007} use context information to automatically name topics.
+
 They describe probabilistic labelling, using the frequency distribution of words in a topic to create a meaningful phrase.
+
 They do not use external domain-specific information as we do, but we
+
 do not generate phrases from the topics.
@@
 Finally, in Ernst et al.~\cite{ernst10refsq}, we describe our earlier
+
 project, similar to this, that identifies changes in quality requirements
+
 in GNOME software projects; this approach was more exploratory, had
+
 less validation, uses different word-lists, solely uses text-matching,
+
 and does not leverage machine learning strategies.
+
 Hindle et al.~\cite{Hindle09ICSM} propose a windowed method of topic
+
 analysis that we extend with labelled
+
 topics, NFRs and new visualizations.
@@
 %Massey~\cite{massey02icse} and Scacchi~\cite{scacchi02,scacchi05b} looked at the topic of requirements in open-source software.
+
 % this reduces the line length and drops a ref even though we still have a scacchi kickin around
+
 %Massey~\cite{massey02icse} and Scacchi~\cite{scacchi05b} looked at the topic of requirements in open-source software.
+
 %Their work discusses the source of the requirements and how they are used in the development process.
+
 % German~\cite{german03gnome} looked at GNOME specifically, and listed several sources for requirements: leader interest, mimicry, brainstorming, and prototypes.
+
 %None of this work  addressed quality requirements in OSS, nor did it examine requirements trends.
+
 \todo{re-add this related work?}
+
 % XXX vanity or space..
+
 %XXXXXXXXXXXX OPTIONAL
+
 % NE 0113 this seems too early to mention 'result'.
+
 % Hindle et al. \cite{Hindle2007} examined release patterns in OSS. They showed that there is a difference between projects regarding maintenance
+
 % techniques. This supports our result that software qualities are not discussed with the same frequency across projects.
@@
 \section{Study Design and Execution}
+
 Figure \ref{fig:process} gives an outline of our methodology.
+
 We began by gathering source data and creating topic models. For semi-unsupervised labelling, we generated three sets of word-lists as signifiers for NFRs.
+
 %e created labels using a well-known NFR taxonomy.
+
 In supervised learning, we trained our data with manual annotations in order to match topics with NFRs. Finally, these topics were used to analyze the
+
 role of NFRs in software maintenance.
@@
 \begin{figure}
+
   \centering
+
  \includegraphics[width=.45\textwidth]{figures/process-model}
+
  \caption{Research methodology process view.}
+
   \label{fig:process}
+
 \end{figure}
@@
 \subsection{Generating the Data}
+
 \label{sec:wordlist}
@@
 To evaluate our approach, we sought candidate systems that were mature projects and had openly accessible source control repositories.
+
 We selected systems from the same application domain, to control for differences in functional, rather than non-functional, requirements.
+
 We used three different open-source, partially-commercial database systems:
+
 \begin{itemize}
+
 \item  MySQL 3.23. Started in 1994 and MySQL 3.23 was released in early 2001. MySQL contains $320,000$ lines of C and C++ source code\footnote{generated
+
 using David A. Wheeler's \emph{SLOCCount}, {http://dwheeler.com/sloccount}.}.
+
 \item MaxDB 7.500. Started in the late 1970s as a research project, and was later acquired by SAP. As of version 7.500, released April 2007, the project
+
 has over $940,000$ lines of C source code.
+
 \item PostgreSQL 7.3. Started in the 1980s as a Berkeley research project. PostgreSQL 7.3 contains $306,000$ lines of C code.
+
 \end{itemize}
@@
 %Choosing an older version of these projects allowed us to focus on projects which have moved further into the maintenance phase of the software life-cycle.
+
 We explicitly chose older versions of mature projects from a stable problem domain to increase the likelihood that we would encounter primarily
+
 maintenance activities in our studies.
@@
 For each project, we used source control commit comments, the messages
+
 that programmers write when they commit revisions to a source control
+
 repository.
+
 Most commits we observed had commit comments.
+
 Commit comments are often studied by researchers, as they are the most readily accessible source of project interactions, and developers are often
+
 required to create them by the repository mechanism (e.g., Git).  Additionally, relying only on commit comments makes our approach more generalizable,
+
 as we do not assume the presence of other artifact corpora.
+
 %This is the most readily accessible source of project interactions for outside researchers, since developers often or always (e.g., Git) write such
+
 %message. Our approach is more generalizable if it does not assume the presence of other corpora.
+
 %bug trackers and email-list archives were not available for both projects.
+
 An example of a typical commit message, from MySQL, is: \textit{``history annotate diffs bug fixed (if mysql\-\_real\-\_connect() failed there were
+
 two pointers to malloc'ed strings, with memory corruption on free(), of course)''}.
+
 We extracted these messages and indexed them by creation time.
+
 We summarized each message as a word distribution minus stop-words such as ``\emph{the}'' and ``\emph{at}''.
+
 %Stemming was performed in the later stages of our analysis. % WAS IT?
@@
 For the commit message data-sets of each project, we created an XML file that separated commits into 30 day periods.
+
 %This size of period, 30 days, is smaller than the time between minor
+
 %releases but large enough for there to be sufficient commits to
+
 %analyze
+
 We chose a period size of 30 days as it is smaller than the time between minor releases but large enough for there to be sufficient commits to
+
 analyze~\cite{Hindle09ICSM}.
+
 For each 30 day period of each project, we input the messages of that period into Latent Dirichlet Allocation (LDA), a topic analysis
+
 algorithm~\cite{Blei2003}, and recorded the topics the algorithm extracted.
+
 %For each project and for each period of messages a project, we input these messages
+
 %we created `topics' using Latent Dirichlet Allocation, a topic identification algorithm.
@@
 A topic analysis tool such as LDA will try to find $N$ independent
+
 word distributions within the word distributions of all input messages.
+
 Linear combinations of these $N$ word distributions are meant to represent and recreate the word distributions of any of the original messages.
+
 These $N$ word distributions effectively form \emph{topics}: cross-cutting collections of words relevant to one or more of our commit messages.
+
 LDA extracts topics in an unsupervised manner; the algorithm relies
+
 solely on the source data and word distributions of messages, with no human intervention.
@@
 In topic analysis a single document, such as a commit message, can be related to multiple topics.
+
 Representing documents as a mixture of topics maps well to source code repository commits, which often have more than one purpose~\cite{Hindle09ICSM}.
+
 For this paper, a topic represents a word distribution generated from a group of commit log comments which are related by their content.
+
 %In this paper a topic is a set of tokens extracted from commit messages found within a project's source control system (SCS).
@@
 We applied Blei's LDA implementation~\cite{Blei2003} against the word distributions of these commits, and generated lists of topics per period.
+
 We set the number of topics to generate to $20$, because past experimentation showed that fewer topics might aggregate multiple unique topics while
+
 any more topics dilutes the results and creates indistinct topics~\cite{Hindle09ICSM}.
+
 % Furthermore, more than $20$ topics quickly became infeasible for inspection.
@@
 To facilitate cross-project comparison, we used a taxonomy of NFRs. This taxonomy is based on the ISO quality model, ISO9126~\cite{iso9126}.
+
 ISO9126 describes six high-level NFRs: maintainability, functionality,
+
 portability, efficiency, usability, and reliability~\footnote{While there may be lingering debate in some circles about these terms, an ISO standard
+
 seems like a reasonable starting point for our work.}.
+
 There is some debate about the terms in this model.
+
 However, ISO9126 is ``an international standard and thus provides an
+
 internationally accepted terminology for software quality \cite[p. 58]{Boegh2008},'' that is sufficient for the purposes of this research.
+
 We claim that these NFRs are maintenance concerns (to varying degrees) in all software projects, and are therefore well-suited for comparisons between
+
 projects.
@@
 \subsubsection{Creating a Validation Corpus}
+
 To evaluate both semi-unsupervised and supervised classification, we
+
 created a validation set of manually labelled topics. For each project,
  the annotators (the first two authors) annotated each extracted topic in each period with
+
 the six NFR labels listed above.
+
 Annotators did not annotate each other's annotations, but some brief
+
 inspection of annotations was used to confirm that the annotators were
+
 acting similarly.
+
 We looked at each period's topics, and assessed what the data ---
+
 consisting of the frequency-weighted word lists and messages ---
+
 suggested was the label for that topic.
+
 We were able to pinpoint the appropriate label using auxiliary information as well, such as the actual revisions and files that were related to the
+
 topic being annotated.
+
 For example, for the MaxDB topic consisting of a message ``exit() only
+
 used in non NPTL LINUX Versions'', we tagged that topic
+
 \emph{portability}.
+
 Given the top-level annotations of \emph{none}, \emph{portability},
+
 \emph{efficiency}, \emph{reliability}, \emph{functionality},
+
 \emph{usability}, and \emph{maintainability}, the annotators annotated each topic
+
 with the relevant label. Sometimes they used finer-grained
+
 annotations that would be aggregated up to one of these higher-level labels.
+
 %\todo[inline]{this confused people. Is it truly unsupervised? Yes, but it needs to be clearer.}
+
 %We compared against this data-set, but we also used the data-set for our supervised machine learning based topic classification.
+
 % A topic may have more than one matching keyword. % ok what?
+
 %\todo[inline]{inter-rater reliability }
+
 % \todo[inline]{There should be a section here which is about manual annotation.
+
 %It is really important and it is required by both analyses. Plus we can tell them how.}
@@
 %\todo[inline]{NEIL: this section should be updated to explain F-measure}
+
 We validate classification performance using the Receiver Operating
+
 Characteristic area-under-curve value~\cite{Fawcett2006861},
+
 abbreviated \emph{ROC}, and the F-measure, which is the harmonic mean of precision and recall, i.e., $2 * (P * R) / (P + R)$.
@@
 ROC values provide a score %, similar to school letter-grades (A is 0.9, C is 0.6),
+
  reflecting how well a particular learner performed for the given data.
+
 ROC maps to the more familiar concepts of precision/sensitivity and recall/specificity: it plots the true positive rate (sensitivity) versus the false
+
 positive rate (1 - specificity).
+
 A perfect learner has a ROC value of 1.0, reflecting perfect recall and precision.
+
 A ROC result of 0.5 would be equivalent to a random learner (that is, issuing as many false positives as true positives).
+
 The %area under the -- NE confusing to have area again ... and not importantana
+
 ROC of a classifier is equivalent to the probability that the classifier will rank a randomly chosen positive instance higher than a randomly chosen
+
 negative instance.
+
 We consider our labelling classifiers acceptable if they outperform a random classifier (0.5).
@@
 \subsection{Semi-unsupervised Labelling}
+
 \label{sec:unsuplabelling}
@@
 In this section we describe how to label topics based on dictionaries
+
 mined from sources external to the projects.
@@
 %\subsection{Generating word lists}
+
 %\label{sec:genwordlist}
+
 % %\todo[inline]{Make it clear where the data comes from! Neil says: Also, I
+
 %   didn't get the impression they understood that we weren't generating
+
 %   labels from the data, but rather from independent taxonomies. Of
+
 %   course if you use labels from the data it isn't comparable. But the
+
 %   point of ISO 9126 etc. was to avoid exactly this problem.}
@@
 % \todo[inline]{Need to hammer it in that the word bags are not project specific and
+
 %  are inherited so it isn't a supervised method}
@@
 %Topics are word distributions, lists of words ranked by frequency, that are burdensome to interpret and hard to distinguish and understand.
+
 %While the topics themselves are generated automatically, how to interpret these topics is less clear.
+
 %For example,
@@
 In order to automatically label each topic with
+
 one of the six high-level NFRs,
+
 %an NFR from ISO9126,
+
 we associate each NFR with a list
+
 of keywords (\emph{word-lists}, in our parlance). These word-lists were determined a priori and were not extracted from the projects themselves.
+
 We intersected the words of the topics and the words of our word-lists.
+
 We ``labelled'' a topic if any of its words matched any of the word-list's words.
+
 A topic could match more than one NFR.
+
 We used several different sets of word-lists for comparison, which we
+
 refer to as \textsf{exp1}, \textsf{exp2}, and \textsf{exp3} in the text which follows.
@@
 Our first word-list set, \textsf{exp1}, was generated using the ontology described in Kayed et al.~\cite{5072519} .
+
 That paper constructs an ontology for software quality measurement using eighty source documents, including research papers and international standards.
+
 The labels we used:
@@
 \begin{quotation}
+
 \small \noindent \textsf{
+
 integrity, security,
+
 interoperability, testability, maintainability, traceability,
+
 accuracy, modifiability, understandability, availability, modularity,
+
 usability, correctness, performance, verifiability, efficiency,
+
 portability, flexibility, reliability.
+
 }
+
 \end{quotation}
@@
 Our second word-list, \textsf{exp2}, uses the ISO9126 taxonomy described above (Section \ref{sec:wordlist}) to seed the word-lists.
+
 %We use these qualities later on as classes in supervised labelling. NE confusing the issue here, I think.
+
 The terms from ISO9126 may not capture all words occurring in the topics that are nonetheless associated with one of the NFRs.
+
 For example, the term ``redundancy'' is one
+
 we considered to be
+
 relevant to discussion of \emph{reliability}, but is not in the standard.
+
 We therefore took the NFRs from the ISO9126 and added to them.
@@
 %\todo[inline]{* clarify the mailinglist stuff we did}
+
 %\todo[inline]{This part doesn't make a lot of sense}
+
 To construct these expanded word-lists, we used
+
 WordNet~\cite{Fellbaum1998}.
+
 %, an English-language ``lexical database'' that contains semantic relations between words,
+
 % including common related forms (similar to word stemming), meronymy and synonymy.
+
 We then added Boehm's software quality model~\cite{Boehm+:1976:ICSE}, and classified his eleven `ilities' into their respective ISO9126 NFRs.
+
 We did the same for the quality model produced by McCall et al.~\cite{mccall1977}.
+
 We then did a simple random analysis of mailing list messages from an open source project, KDE. If we judged a given message to contain terms that were
+
 related to one of the NFRs in ISO9126, we added it to our word-list. This allowed us to expand our word-lists with more software-specific terms.
+
 %Finally, we randomly analyzed two mailing lists from another software project to expand our set with domain-specific terms.
+
 % Less info is probably better here...
+
 %For example, we add the term ``performance'' to the synonyms for \emph{efficiency}, since this term occurs in most mail messages that discuss efficiency.
+
 Table \ref{tbl:wnsig} shows the labels (NFRs) and word-lists we used for matching.
@@
 For the third set of word-lists, \textsf{exp3}, we extended the word-lists from \textsf{exp2} using unfiltered WordNet similarity matches.
+
 Similarity in WordNet means siblings in a hypernym tree.
+
 We do not include these words here for space considerations (but see the Appendix for our data repository).
+
 It is not clear the words associated with our labels in \textsf{exp3} are specific enough. For example, the label \emph{maintainability} is associated with
+
 words \emph{ease} and \emph{ownership}. In general, as we proceed from word-list in \textsf{exp1} to that in \textsf{exp3}, our lists become more generic.
@@
 \begin{table*}
+
 	\centering
+
 \begin{tabular}{c|p{11cm}}
+
 \toprule
+
 \textbf{Label} & \textbf{Related terms} \\
+
 \midrule
+
 \emph{Maintainability} &
+
 testability changeability analyzability stability maintain maintainable modularity modifiability understandability interdependent dependency encapsulation
+
 decentralized modular\\ \hline
+
 \emph{Functionality} &
+
 security compliance accuracy interoperability suitability functional practicality functionality compliant exploit certificate secured ``buffer overflow''
+
 policy malicious trustworthy vulnerable vulnerability accurate secure vulnerability correctness accuracy\\ \hline
+
 \emph{Portability} &
+
 conformance adaptability replaceability installability portable movableness movability portability specification migration standardized l10n localization
+
 i18n internationalization documentation interoperability transferability\\ \hline
+
 \emph{Efficiency} &
+
 ``resource behaviour'' ``time behaviour'' efficient efficiency performance profiled optimize sluggish factor penalty slower faster slow fast optimization\\
+
 \hline
+
 \emph{Usability} &
+
 operability understandability learnability useable usable serviceable usefulness utility useableness usableness serviceableness serviceability usability
+
 gui accessibility menu configure convention standard feature focus ui mouse icons ugly dialog guidelines click default human convention friendly user
+
 screen interface flexibility\\ \hline
+
 \emph{Reliability} &
+
 ``fault tolerance'' recoverability maturity reliable dependable responsibleness responsibility reliableness reliability dependableness dependability
+
 resilience integrity stability stable crash bug fails redundancy error failure\\
+
 \bottomrule
+
 \end{tabular}
+
 	\caption{NFRs and associated word-lists -- \textsf{exp2}}
+
 	\label{tbl:wnsig}
@@
 %\todo[inline]{NEIL: this could be a subsubsection}
+
 %\todo[inline]{* the hierarchy of tags was not clear}
+
 %\noindent \textbf{Generating labels automatically} ---
@@
 Using our three word-lists (\textsf{exp1}, \textsf{exp2}, \textsf{exp3}), we labelled our topics with an NFR where there was a match between a word in
+
 the list and the same word somewhere in the distribution of words that constitute the topic.
+
 A \emph{named topic} is a topic with a matching NFR label. \emph{Unnamed
+
 topics} occur where there is no such match, which may indicate either a lack of precision, or simply that this topic is not associated with non-functional
+
 requirements.
+
 All experiments were run on the datasets for each project (e.g., PostgreSQL, MySQL, MaxDB). LDA
 extracted $20$ topics per period for each project.
+
 % Is this good enough
+
 This labelling is \emph{semi-unsupervised} because the corpus is not derived from
+
 the project being analyzed, and we did not label the project's topics
+
 ourselves for a training set. The motivation behind this technique is that
+
 because most software often addresses similar issues, we can use the 
 domain knowledge of software to label relevant topics.
+
 %\todo[inline]{Neil does this explain the unsupervised aspect well enough?}
@@
 Table \ref{tbl:wordlist} shows how many topics were labelled for MaxDB
+
 and MySQL.
@@
 %\todo[inline,color=green]{this table should have mysql}
+
 \begin{table}
@@
 % % This data came from running
+
 % %   python check_validity.py
+
 % %  in src/validate
+
 % % and then running
+
 % % tail -n 5 *maxdb*yes_no*
+
 % % tail -n 5 *mysql*yes_no*
+
 % ok the new way is to run lda_relator
+
 % src/validate/lda_relator_counts.sh
-	\centering
-\begin{tabular}{l|c|c|c|c}
-\toprule
-\textbf{Project} & \textbf{Measure} & \textsf{exp1} & \textsf{exp2} & \textsf{exp3} \\
-\midrule
+
 % NOTE THESE WERE THE OLD VALUES
@@
 
-MaxDB 7.500 & Named Topics   & 281 & 125 & 328 \\ % these numbers are
-				% slightly different solely because
-				% we're counting empty topics
-				% maxdb has 80 empty topics
- & Unnamed Topics & 219  & 375 &  172  \\
- & Total  Topics  & 500 & 500 & 500 \\
+% MaxDB 7.500 & Named Topics   & 281 & 125 & 328 \\ % these numbers are
+% 				% slightly different solely because
+% 				% we're counting empty topics
+% 				% maxdb has 80 empty topics
+%  & Unnamed Topics & 219  & 375 &  172  \\
+%  & Total  Topics  & 500 & 500 & 500 \\
+% \midrule
+% MySQL 3.23  & Named Topics   & 524 & 273 & 773 \\
+%   & Unnamed Topics & 476 & 727 & 227 \\
+%   & Total  Topics  & 1000 & 1000 & 1000 \\
+% \midrule
+% PgSQL 3.23  & Named Topics   & - & - & - \\
+%   & Unnamed Topics & - & - & - \\
+%   & Total  Topics  & - & - & - \\
+	\centering
+
+\begin{tabular}{l|c|c|c|c}
+
+\toprule
+
+\textbf{Project} & \textbf{Measure} & \textsf{exp1} & \textsf{exp2} & \textsf{exp3} \\
+
 \midrule
-MySQL 3.23  & Named Topics   & 524 & 273 & 773 \\
-  & Unnamed Topics & 476 & 727 & 227 \\
+
+% updated for ESE: taken from latest-reports.tar.gz, last five lines of 
+% yes_no_report.txt, using pgsqln instead of pgsqla
+ MaxDB 7.500 & Named Topics   & 305 & 183 & 330 \\
+             & Unnamed Topics & 84  & 206 & 59  \\
+             & Total Topics   & 400 & 400 & 400 \\
+\midrule
+ MySQL 3.23  & Named Topics   & 341 & 202 & 469 \\
+             & Unnamed Topics & 245 & 384 & 117  \\
              & Total Topics   & 1000 & 1000 & 1000 \\
 \midrule
-PgSQL 3.23  & Named Topics   & - & - & - \\
-  & Unnamed Topics & - & - & - \\
-  & Total  Topics  & - & - & - \\
-
+ PgSQL 7.3   & Named Topics   & 639 & 543 & 640 \\
+             & Unnamed Topics & 1   & 97  & 0  \\
+             & Total Topics   & 640 & 640 & 640 \\
+\midrule
 \bottomrule
+
 \end{tabular}
-	\caption{Automatic topic labelling for MaxDB and MySQL}% \todo[inline]{this stupid table is before table 1}}
+
+	\caption{Automatic topic labelling for MaxDB, MySQL and PostgreSQL}
 	\label{tbl:wordlist}
@@
 \end{table}
-\todo{update with PGSQL}
-% was best performing
+
 For \textsf{exp1} the labels with the most topics
+
 were
-%labels, those with the most topics, 
+
 \emph{correctness} (182 topics) and \emph{testability} (121).
+
 We did not see many results for \emph{usability} or \emph{accuracy}, which were associated with fewer than ten topics.
+
 We also looked for correlations between our labels: excluding double matches (self-correlation), our highest co-occurring terms were \emph{verifiability}
+
 with \emph{traceability}, and \emph{testability} with \emph{correctness} (76 and 62 matches, respectively).
+
 % See the following sections for our error analysis.
@@
 For \textsf{exp2}, there are more unnamed topics than \textsf{exp1}.
+
 Only \emph{reliability} produces a lot of matches, mostly with the word ``error''.
+
 Co-occurrence results were poor. This suggests our word lists were overly restrictive.
@@
 For \textsf{exp3}, we generally labelled more topics.
+
 As we mentioned, the word-lists are broad, so there are likely to be false-positives (discussed below).
+
 The most frequent label, 265 topics, was \emph{usability}, and the
+
 least frequent label, 44 topics, was \emph{maintainability}.
+
 Common co-occurrences were \emph{reliability} with \emph{usability}, \emph{efficiency} with \emph{reliability}, and \emph{efficiency} with \emph{usability}
+
 (200, 190, and 150 topics in common, respectively).
@@
 \subsubsection{Analysis of the Semi-unsupervised Labelling} %--- %Based
+
 				%on the labels, and our manual topic
+
 				%labelling,
+
 %\noindent \textbf{Analysis of the unsupervised labelling} --- %Based on the labels, and our manual topic labelling,
+
 %we compared the results of the unsupervised word matching approach.
+
 For each quality we assessed whether semi-unsupervised labels matched the manual annotations. %To be clear, the manual annotations were not used to
+
 train the labelling process.
+
 As described in Section \ref{sec:wordlist} we used both ROC and F-1 measures to evaluate the performance of the classification.
+
 Figure \ref{fig:maxdb-unsup-results} shows our ROC results for MaxDB and MySQL. We describe F-1 results in the text below.
@@
 %\todo[inline]{since ROC is new-sounding, I wonder if a more lengthy explanation would be merited. An obvious question is why not precision/recall? --
+
 % ROC is hot right now and it combines both}
+
 % Rebuttal: ROC is not new and it is what the machine learning community, search based software engineering and now the MSR community are using +
+
 %it has natural meaningful quality mappings.
+
 % \begin{figure}[t]
+
 % \centering
+
 % \subfloat[ROC]{
+
 %  \includegraphics[width=0.45\textwidth]{figures/unsupervised-bar}
+
 % \label{fig:maxdb-unsup-results-roc}
+
 % }
+
 % %\subfloat[F1]{
+
 % %\includegraphics[width=0.45\textwidth]{figures/unsupervised-bar-f1}
+
 % %\label{fig:maxdb-unsup-results-f1}
+
 % %}
+
 % \caption[]{Performance of unsupervised topic labelling for each NFR per word-list. A) ROC values --
+
 % dashed line indicates the performance of a random classifier. B) F1 measure. Possible values range from 0--1.
+
 % %\todo[inline]{better discussion}% and graphical legend would be nice}
+
 % }
+
 % \label{fig:maxdb-unsup-results}
+
 % \end{figure}
@@
 \begin{figure*}[t]
+
   \centering
+
  \includegraphics[width=0.8\textwidth]{figures/unsupervised-bar}
+
  \caption{Performance, ROC values (range: $0$--$1$), of semi-unsupervised topic labelling for
+
    each NFR and per word-list. The dashed line indicates the performance of a random classifier. This graph shows how well the
+
    semi-unsupervised topic labelling matched our manual annotations.}
@@
 %\todo[inline]{this is too tiny in print}
+
   \label{fig:maxdb-unsup-results}
+
 \end{figure*}
@@
 % \begin{figure}
+
 %   \centering
+
 %  \includegraphics[width=0.45\textwidth]{figures/unsupervised-bar-f1}
+
 %  \caption{Performance, F1 values (range: 0-1), of unsupervised topic labelling for
+
 %    each NFR and per word-list.}
+
 %   \label{fig:maxdb-unsup-results-f1}
+
 % \end{figure}
@@
 %EXP1 could not be evaluated properly!
+
 Because our ground truth annotations were relevant only to ISO9126,
+
 we estimate that \textsf{exp1} had poor
+
 performance via the overlap between ISO9126 and the Kayed ontology.
+
 For \textsf{exp1} the F-measures for MaxDB were from $0$ to $0.18$ with an average
+
 of $0.03$, and for MySQL were from $0$ to $0.16$ with an average of
+
 $0.05$.
+
 % 0.177215189873 / 6
+
 % .02953586497883333333
+
 % (0.0298507462687 + 0.0661157024793 + 0.0634920634921 + 0 + 0.115384615385) / 6
+
 % .04580718793751666666
@@
 % EXP2 source is naming-paper/src/validate/exp2.maxdb.report.txt
+
 % (change as needed). Second column (after the name) is F1.
+
 For \textsf{exp2}, the average F-measure (macro-F1) for MaxDB was $0.24$ with a range $0.091$ to
+
 $0.37$, and $0.16$ for MySQL with a range of $0$ to $0.41$.
+
 MaxDB had an average precision and recall of $0.25$ and $0.22$
+
 while MySQL had $0.41$ and $0.10$ respectively.
@@
 % EXP3
+
 For \textsf{exp3}, the average F-measure (macro-F1) for MaxDB was $0.26$ with a range $0.11$ to
+
 $0.47$, and $0.36$ for MySQL with a range of $0.10$ to $0.65$.
+
 MaxDB had an average precision and recall of $0.16$ and $0.67$
+
 while MySQL had $0.3$ and $0.48$ respectively.
@@
 Thus we find that \emph{reliability} and
+
 \emph{usability} worked well for MaxDB in \textsf{exp2} and better in
+
 \textsf{exp3}.
+
 \textsf{exp1} performed poorly.
+
 MySQL had reasonable results within \textsf{exp2} for \emph{reliability} and \emph{efficiency}.
+
 MySQL's results for \emph{efficiency} did not improve in \textsf{exp3}
+
 but other qualities such as \emph{functionality} did improve.
+
 Many ROC scores were $0.6$ or less, but our classifier still performed substantially better than random.
@@
 %\vspace*{-1em}
+
 \subsection{Supervised Labelling}
+
 \label{sec:suplabelling}
+
 Supervised labelling requires expert analysis of the correct
+
 class/label to assign a label to a topic. In our approach, we use the top-level NFRs in the ISO9126 standard~\cite{iso9126} for our classes, but other
+
 taxonomies are also applicable.%In order to validate how effective these word-bag approaches to topic labelling would be we created a validation data set.
@@
 We used a suite of supervised classifiers, WEKA~\cite{weka09},
+
 that includes machine learning tools such as support vector machines and Bayes-nets.
+
 We also used the multi-labelling add-on for WEKA, Mulan~\cite{mulan}. %\footnote{\url{http://mlkd.csd.auth.gr/multilabel.html}}.
+
 Traditional classifiers label topics with a single class, whereas Mulan allows for a mixture of classes per topic, which is what we observed while
+
 manually labelling topics.
+
 The \emph{features} we used are word counts/occurrence	per topic, if
+
 a word occurs frequently enough in a topic we consider it a feature of
+
 the topic.
@@
 To assess the performance of the supervised learners, we did a 10-fold cross-validation~\cite{Kohavi1995}, a common technique for evaluating machine
+
 learners.
+
 The original data is partitioned randomly into ten sub-samples. Each sample is used to test against a training set composed of the nine other samples.
+
 %One subsample is used as a training set and evaluated on the other nine sub-samples.
+
 %This is repeated nine more times, with each subsample used once as the training set.
+
 We have reported these results below.% in Section \ref{sec:suplabelling}.
@@
 \subsubsection{Analysis of the Supervised Labelling}
+
 % \label{sec:suplabelling}
+
 Because our data-set was of word counts we expected Bayesian techniques, often used in spam filtering, to perform well.
+
 We tried other learners that WEKA~\cite{weka09} provides: rule learners, decision tree learners, vector space learners, and support vector machines.
+
 Figure \ref{fig:best-learn-per-tag} shows the performance of the best
+
 performing learner per label:
+
 % We considered the best learner for a label to be
+
 the learner that had the highest ROC value for that label.
+
 The best learner is important because one uses a single learner per
+
 label. When applying our technique, for each NFR one should select the best learner possible.
+
 %Figure \ref{fig:best-learn-per-tag} uses the ZeroR learner as a baseline, a common practice in machine learning,
+
 %as ZeroR naively chooses the largest category all of the time.
+
 %ZeroR occasionally outperforms other learners; for labels which are not as common, this is to be expected because any miscategorization will hurt accuracy.
+
 %This is why we use ROC values, instead of accuracy, as they can better represent performance on labels which are not applicable to the majority of samples.
@@
 Figure \ref{fig:best-learn-per-tag} shows that MaxDB and MySQL have quite different results, as the ROC values for \emph{reliability} and \emph{functionality}
+
 seem swapped between projects.
@@
 \begin{figure}[t]
+
 \centering
+
 % \subfloat[MySQL]{
+
 \includegraphics[width=0.45\textwidth]{figures/both-supervised}
+
 % \label{fig:subfig1}
+
 % }
+
 % \subfloat[MaxDB]{
+
 % \includegraphics[width=0.4\textwidth]{figures/maxdb-supervised}
+
 % \label{fig:subfig2}
+
 % }
+
 \caption[]{ROC value for the best learner per label for MaxDB and MySQL. Values range from $0$--$1$.  Dashed line indicates the performance of a random
+
 classifier.
+
 %\todo[inline]{better discussion}
+
 }
+
 \label{fig:best-learn-per-tag}
+
 \end{figure}
@@
 For both projects Bayesian techniques did the best out of a wide variety of machine learners tested.
+
 Our best learners, Discriminative Multinomial Naive Bayes, Naive Bayes
+
 and Multinomial Naive Bayes  are all based on Bayes's theorem and all
+
 assume, naively, that the features supplied are independent.
+
 %The features we used are word counts per message.
+
 One beneficial aspect of this result is that it suggests we can have very fast training and classifying  since Naive Bayes can be calculated in $O(N)$
+
 for $N$ features.
+
 %http://nlp.stanford.edu/IR-book/html/htmledition/naive-bayes-text-classification-1.html
@@
 % # F measure
+
 % mysql
+
 % > v <- c( 0.64 , 0.23  ,0.41 , 0.77  ,0.62 , 0.21 )
+
 % > summary(v)
+
 %    Min. 1st Qu.  Median    Mean 3rd Qu.    Max.
+
 %   0.210   0.275   0.515   0.480   0.635   0.770
+
 % maxdb
+
 % > v <- c(0.61, 0.25, 0.59, 0.32, 0.42, 0.17)
+
 % > summary(v)
+
 %    Min. 1st Qu.  Median    Mean 3rd Qu.    Max.
+
 %  0.1700  0.2675  0.3700  0.3933  0.5475  0.6100
@@
 The range of F-measures for MySQL was $0.21$ to $0.77$ with an average
+
 of $0.48$, while MaxDB had a range of $0.17$ to $0.61$ with an average
+
 of $0.39$.
@@
 The less-frequently occurring a label, the harder it is to get accurate
+
 results, due to the high noise level. Nevertheless, these results are
+
 better than our previous word-list results of \textsf{exp2} and
+
 \textsf{exp3}, because the ROC values are sufficiently higher in most
+
 cases (other than MaxDB \emph{reliability} and MySQL \emph{efficiency}). The
+
 limitation of the approach we took here is that we assume labels are
+
 independent; however, labels could be correlated with each other.
+
 The next section (\ref{sec:multilabel})
+
 addresses the issue of a lack of independence and correlation between
+
 labels
+
 using multi-label learners.
+
 %In the next section we will evaluate how well these learners perform
+
 %together.
+
 %\todo[inline]{Comment about correlation, we should link to the Mulan section, perfect segway since Mulan relies on correlation.}
@@
 \subsection{Applying Multiple Labels to Topics}
+
 \label{sec:multilabel}
@@
 As noted in Section \ref{sec:wordlist}, each topic in our data-set can be composed of zero, one, or more NFRs.
+
 For example, a commit message might address \textit{reliability} in the context of \textit{efficiency}, or make a \textit{maintainability} improvement
+
 in the source code that related to \textit{usability}.
+
 However, traditional machine learning techniques, such as Naive Bayes, can only map topics to a single class.
+
 The Mulan~\cite{mulan} library encapsulates several different multi-label machine learners which can label elements with multiple labels.
+
 Mulan also includes methods for determining the performance of such techniques.
+
 % TOO DETAILED The problem framed in the learners above has changed; instead of looking at the precision and recall of applying one label, we rank
+
 % multiple labels at once. We must check if the full subset of labels was applied, and then how much of that subset was applied.
@@
 \begin{figure*}[t]
+
 \centering
+
 \subfloat[MySQL]{
+
 \includegraphics[width=0.3\textwidth]{figures/multi-label-results-mysql}
+
 \label{fig:subfig3}
+
 }
+
 \subfloat[MaxDB]{
+
 \includegraphics[width=0.3\textwidth]{figures/multi-label-results-maxdb}
+
 \label{fig:subfig4}
+
 }
+
 \caption[]{MySQL and MaxDB macro (grey) and micro (black) ROC results per multi-label learner. Possible values range from $0$--$1$.  Dashed line
+
 indicates the performance of a random classifier.
+
 %\todo[inline]{better discussion}% and graphical legend would be nice}
+
 }
+
 \label{fig:mulan}
+
 \end{figure*}
@@
 Two perspectives to evaluate multi-label learners are with micro or macro measurements (used in Figure \ref{fig:mulan}).
+
 Macro measurements are aggregated at a class or label level (per class) while micro measurements are at the element level (per element).
+
 A macro-ROC measurement is the average ROC over the ROC values for all labels, where a micro-ROC is the average ROC over all examples that were classified.
+
 For MaxDB, the macro-ROC values are undefined because of poor performance of one of the labels.%, see Figure \ref{fig:mulan}.
@@
 Figure~\ref{fig:mulan} presents the results of Mulan's best multi-label learners for our data.
+
 Calibrated Label Ranking (CLR) is a learner that builds two layers. The first layer determines if an entity should be labelled, while the second layer
+
 determines what labels should be assigned.
+
 The Hierarchy Of Multi-label classifiERs (HOMER) and Binary Relevance (BR) act as a hierarchy of learners: BR is flat, while HOMER tries to build a
+
 deeper hierarchy for a more accurate learner~\cite{mulan}.
+
 These classifiers performed better than other multi-label classifiers as they have the best micro and macro ROC scores.
+
 The multi-label and single-label learners had similar performance: for
+
 MySQL, BR and Naive Bayes had similar macro-ROC scores of $0.74$.
+
 % although their results are comparable to the naive Bayesian learners we used in Section \ref{sec:suplabelling}.
+
 % \todo[inline]{discuss this similarity in more detail}
@@
 \section{Understanding Software Maintenance Activities}
+
 \label{sec:analysis}
+
 % \todo[inline]{talk about this topic in one section}
+
 % \todo[inline]{reference project name Nomen more}
+
 %In order to test  our ROC values produced acceptable results, we now turn to
+
 %\emph{applying} those techniques to understand software maintenance
+
 %activities.
+
 %\todo[inline]{more on intro ... why should the reader care about
+
  % this. It needs to show how the stuff in the intro (applications) is
+
  % possible}
+
 %%% XXX NEIL I don't get this.. VVVVV
+
 %\todo[inline]{fix section headings}
+
 As we mentioned in the introduction, a key issue in software maintenance is understanding \emph{why} a system has evolved the way it has~\cite{aranda09icse}.
+
 In this section we demonstrate the value of labelled topic extraction
+
 in addressing this issue.
+
 % more why
+
 Labelled topics address \emph{why}  because they show reasons why changes
+
 occurred rather than \emph{how}.
+
 The \emph{how} of a change is the change itself, the purpose or
+
 \emph{why} is what we are after.
+
 We investigate the history of the two large-scale database systems that we studied.
+
 We use our technique to show the topic of development efforts over time in each project.
+
 %We evaluated two research questions using the ground truth data that we annotated by hand:
+
 We motivated our investigation with three research questions:
+
 \begin{enumerate}
+
 \item \emph{Do NFR frequencies change over time?} If a particular NFR
+
   was of more interest at one point in the life-cycle than another,
+
   this suggests that development activity shifted focus. For example,
+
   if a developer expected to see a recent focus on \emph{reliability},
+
   but instead \emph{usability} dominated, they might re-prioritize
+
   upcoming work items.
@@
 \item \emph{Do projects differ in their relative interest in NFRs?} A
+
   project manager, especially a systems-manager, would be interested
+
   in knowing whether a particular NFR, such as \emph{reliability}, was
+
   more important for one project than another. This could be to
+
   confirm the initial design goals, or to track the progress on that
+
   quarter's objectives.  The difference in NFR proportion is
+
   interesting because it implies a difference in focus between two projects.
@@
 \item \emph{Do different developers work on different NFRs?} For a given project, it is reasonable to think that developers are either assigned
+
 (in commercial organizations) or choose (in open-source organizations) to a particular NFR. For example, one developer might be more junior, and take
+
 responsibility for the low-impact reliability fixes. Another, more senior developer might assume responsibility for major improvements such as efficiency
+
 improvements.
@@
 %\subsection{Timeline analysis}
+
 Figures \ref{fig:mysql-timeline} and \ref{fig:maxdb-timeline} and \ref{fig:pgsql-timeline} show the
+
 temporal patterns of NFR frequencies.
+
 %NEIL- I think this (below) is just confusing, although maybe important to note.
+
 % based on the manually annotated
+
 %topics, although this visualization can be generated from the
+
 %results of labelled topic extraction.
+
 There are two measures represented.
+
 One, the relative frequency, shown in the grey histogram boxes, represents the number of topics with that NFR in that period,
+
 relative to the maximum number of topics assigned to that NFR.
+
 For example, in Figure \ref{fig:mysql-timeline} we see a spike in \emph{portability} and \emph{functionality} frequency in September 2002.
+
 The second, absolute frequency, is shown using cell intensity, and compares the number of topics labelled with that NFR per period
+
 relative to the maximum number of labelled topics overall.
+
 For instance, Figure \ref{fig:mysql-timeline} shows that the NFRs
+
 \emph{functionality}, \emph{portability} and \emph{maintainability}
+
 contain more labelled topics, since these NFRs have been more
+
 intensely shaded;
+
 one interesting stream is \emph{efficiency} which shows periodic
+
 activity, does this suggest that once maybe efficiency related changes
+
 have longer lasting effects?
+
 The topmost row in each diagram lists historical events for that project (such as a release).
+
 %We refer to these occurrences in the discussion which follows.
@@
 \begin{figure}[t]
+
 \centering
+
 \subfloat[MySQL 3.23]{
+
 \includegraphics[width=\textwidth]{figures/mysql-timeline.png}
+
 \label{fig:mysql-timeline}
+
 }
@@
 \subfloat[MaxDB 7.500]{
+
 \includegraphics[width=\textwidth]{figures/maxdb-timeline.png}
+
 \label{fig:maxdb-timeline}
+
 }
@@
 \subfloat[PostgreSQL 7.2]{
+
 \includegraphics[width=\textwidth]{figures/pgsql-timeline}
+
 \label{fig:pgsql-timeline}
+
 }
+
 	\caption[]{NFR label per period. Each cell represents a 30-day period. %Numbers represent number of topics in that period labelled with that NFR.
+
 	Grid cell intensity (saturation) is mapped to label frequency
+
 	relative to the largest label count of \emph{all} NFRs. Grey
+
 	histogram bars indicate label frequency relative to that
+
 	particular NFR's largest label count. Dashed vertical lines
+
 	relate a project milestone (\emph{*Key events*}) to our topic windows.
+
 % \todo[inline]{better discussion}
+
 }
+
 \label{fig:timelines}
+
 % Timelines created using src/validate/summary_labels.py
+
 \end{figure}
@@
 %\noindent \textbf{Understanding the figures} --
+
  % boxes. not sure boxes is important..
+
 % I grepped query1.mrs2 http://softwareprocess.es/y/y/maxdb-76.tar.gz
+
 % and			http://softwareprocess.es/y/maxdb-7500.tar.gz
+
 %There are more \textsf{None} labels in MaxDB because a number of topics were to do with code cleanup or automated checking using tools like
+
 %\texttt{sutcheck}, a tool specific to the development process of MaxDB.
@@
 %\noindent \textbf{Timeline of key events} --
+
 We analyzed each project's developer mailing list for external validation.
+
 %As discussed earlier, the top row in each figure shows key events for each project.
+
 We use \textit{labelled topic extraction} to pick out the underlying NFR activity behind these events.
+
 For example, both MaxDB and MySQL show a high number of NFRs recognized at the first period of analysis.
+
 This is due to our window choice: we deliberately targeted our
+
 analysis to when both MySQL 3.23
+
 and MaxDB 7.500 were first announced. For MaxDB, version 7.5.00  was released in December of 2003.
+
 We know that release 7.5.00.23 saw the development of PHP interfaces, possibly accounting for the simultaneous increase in the \emph{portability}
+
 NFR at the same time.
+
 The gap in MaxDB (Figure \ref{fig:maxdb-timeline}) is due to a shift in development focus (from February 2005 to June 2005) to MaxDB 7.6, which is
+
 released in June 2005.
@@
 %For MySQL (Figure \ref{fig:mysql-timeline}), we similarly validated our NFR patterns with external mailing list data.
+
 The release of MySQL we study  (Figure \ref{fig:mysql-timeline}) was the first to be licenced under the GPL.
+
 Version 3.23.31 (January, 2001) was the production release (non-beta), and we see a flurry of topics labelled with \emph{functionality} and
+
 \emph{maintainability}.
+
 After this point, this version enters the maintenance phase of its life-cycle.
+
 In May 2001, there is an increase in the number of topics labelled with \emph{portability}.
+
 This might be related to release 3.23.38, which focused on Windows compatibility.
+
 Similarly, in August, 2002, both \emph{functionality} and \emph{portability} are frequent, and mailing list data suggests this is related to the release
+
 of version 3.23.52, a general bug fix with a focus on security (a component of the \emph{functionality} NFR in the ISO9126 model).
+
 After this point, efforts shift to the newer releases (4.0, 4.1,
+
 5.0).
@@
 By contrast, the PostgreSQL timeline  (Figure \ref{fig:pgsql-timeline}) is extracted from a central trunk and is not version-specific. Therefore development
+
 tends to be focused on releases 7.3, 7.4 and 7.5/8.0 alpha. For example, a priority for the 8.0 candidate was a Windows-native port of the source code, which seems to
+
 correlate with the \emph{portability} NFR increasing in frequency in mid-2004. Before the 7.4 release, \emph{usability,functionality} and \emph{portability} all
+
 increase in frequency, possibly reflecting the interest in adding features and documentation for the release.
@@
 % \label{sec:comparison}
+
 %\XXX{Not sure there's enough here to do that!}
+
 % We observed that MySQL had more topic subsets than MaxDB.
+
 % MySQL 3.23 also had more topics and a longer recorded history than MaxDB 7.500.
+
 % We tagged more MySQL topics with annotations than MaxDB topics, yet both shared similarities.
+
 % In terms of non-functional requirements both projects had long running trends that focused on \emph{functionality}, \emph{maintainability}, and
+
 %\emph{portability}, yet MaxDB had more of a focus on \emph{efficiency} and \emph{reliability}.
+
 % \todo[inline]{expand}
+
 %\todo[inline,color=yellow]{expand on this with dates, and reference to visualizations}
+
 %\subsection{Research questions}
+
 %This analysis allows to address the questions raised in Section \ref{sec:suplabelling}:
+
 %This analysis addresses our research questions: %raised in Section \ref{sec:analysis}:
@@
 \subsection{Do NFR Frequencies Change over Time?}
+
 %\noindent\textbf{1. Do NFR frequencies change over time?} ---
+
 %\subsubsection{Do label frequencies change over time?}
+
 %\todo{add more on PGSQL}
+
 In both MaxDB and MySQL the frequencies generally decreased with age.
+
 However, there are variations within our NFR labels. In MySQL, \emph{usability} and \emph{efficiency} do not appear very often in topics.
+
 A proportionately smaller number of commits addressed these NFRs.
+
 %As discussed in the previous section,
+
 Certain peaks in topic numbers coincide with a particular emphasis from the development team on issues such as new releases or bug fixes.
+
 This suggests that maintenance activity is not necessarily strictly decreasing with time, but rather episodic and responsive to outside stimuli.
+
 In MaxDB, we can observe that \emph{Maintainability} topics became more prevalent as MaxDB matures.
+
 This is likely due to our analysis time-frame for MaxDB being shorter than the time-frame for the MySQL product.
@@
 In PostgreSQL, by comparison, the frequencies seem to by cyclic, since we are not studying a maintenance-phase for the product, but rather ongoing
+
 feature addition and usability improvements.
@@
 \subsection{Do Projects Differ in their Relative  Topic Interest?}
+
 Yes. MySQL 3.23 had proportionally more
+
 topics labelled \emph{functionality}, while MaxDB had proportionally more
+
 \emph{efficiency} related topics. MaxDB was a very mature release ``donated'' to the open-source community,
+
 whereas MySQL was in its relative infancy, and
+
 security problems were more common (security is a component of \emph{functionality} in the ISO9126 model).
+
 In both cases \emph{portability} was a constant maintenance concern and was prevalent throughout the lifetime of the projects. It may surprise developers
+
 how often portability arises as a concern.
+
 \todo{add the pgsql data}
@@
 \subsection{Do different developers work on different NFRs?}
+
 We did a t-test to assess whether there was statistically significant differences in the types of NFRs developers chose to work on, for the PostgreSQL project only.
+
 \todo{add more, figure}
@@
 *** We also clustered the authors by similarity and we found that
+
 among some of the most prolific postgresql authors, their NFR focus
+
 was distinctly different *** Some confounds are that postgresql
+
 authors commit the patches of others.
@@
 \section{Discussion}
+
 \label{sec:limit}
@@
 \subsection{Annotation Observations}
+
 We found many topics that were not actually non-functional requirements (NFRs) but were often related to them.
+
 For instance, concurrency was mentioned often in the commit logs and
+
 was related to \emph{correctness} and \emph{reliability}, possibly because concurrent code is prone to bugs such as race conditions. %are
+
 % often  was troublesome.
+
 \emph{Configuration management} and \emph{source control} related changes appeared often; % and sometimes there were topics dedicated to configuration
+
 management.
+
 these kinds of changes are slightly related to \emph{maintainability}.
+
 A non-functional change that was not quality-related was licensing and
+
 copyright; many changes were simply to do with updating copyrights or
+
 ensuring copyright or license headers were applied to files. In these
+
 cases we assigned the \emph{None} label to the topic.
@@
 We noticed that occasionally the names of modules would conflict with words related to other non-functional requirements.
+
 For instance, optimizers are very common modules in database systems: both MySQL and MaxDB have optimizer modules.
+
 In MySQL the optimizer is mentioned but often the change addresses  correctness or another quality.
+
 Despite this difference, the name of the module could fool our learners into believing the change was always about \emph{efficiency}.
+
 In these cases the advantages of tailoring topic names to specific project terminologies are more clear.
+
 Project specific word-lists would avoid automated mistakes due to the names of entities and modules of a software project.
@@
 Authors Neil and Abram each annotated the PostgreSQL topics in order
+
 to determine inter-rater reliability. While when they read each others
+
 annotations they generally agreed, the inter-rater reliability did not
+
 bear this out.
@@
 %>    A_portability   A_functionality     A_reliability A_maintainability
+
 %>      0.153548387      -0.013588517       0.004573794       0.082001031
+
 %>     A_efficiency       A_usability            A_none
+
 %>      0.231496539       0.009287926       0.062133646
@@
 % > Spearman says
+
 % >   A_portability   A_functionality     A_reliability A_maintainability
+
 % >      0.253037469      -0.014415214       0.004579664       0.082215613
+
 % >     A_efficiency       A_usability            A_none
+
 % >      0.258037216       0.014436230       0.080667468
@@
 \begin{table}
+
 \centering
+
 \begin{tabular}{l|c|c}
+
 \hline
+
             & Cohen's Kappa & Spearman Correlation \\ \hline
+
 Portability & 0.154 & 0.253  \\
+
 Functionality & -0.014 & -0.014 \\
+
 Reliability & 0.005 & 0.005 \\
+
 Maintainability & 0.082 & 0.082 \\
+
 Efficiency & 0.231 & 0.258 \\
+
 Usability & 0.009 & 0.014 \\
+
 None &      0.062 & 0.081 \\ \hline
+
 Everything & 0.107 & 0.108 \\ \hline
+
 \end{tabular}
+
 \caption{Inter-rater Reliability on PostgreSQL}
+
 \label{tab:interr}
+
 \end{table}
@@
 Table \ref{tab:iterr} describes the Cohen Kappa and the Spearman
+
 correlation of our per topic annotations for each tag. We had to
+
 evaluate inter-rater reliability this way because 1 topic could be
+
 tagged with more than 1 annotation.
@@
 We feel these results are poor. The aggregate view of with a Kappa of
+
 $0.1$ indicates there is some agreement but not a lot. We found that
+
 there was a lot of agreement in terms of lack of a tag, but often more
+
 disagreement when a tag should be applied.
@@
 After some discussion we concluded that usability was a source of
+
 disagreement. Do user manuals improve usability? Is a command-line
+
 option a usability issue? These kinds of questions illustrate some of
+
 the agreement, disagreement and ambiguity about these labels. Further
+
 more, both annotators found that the annotation was time-consuming and
+
 difficult.
@@
 \subsection{Summary of Techniques}
+
 While an unsupervised technique such as LDA is appealing in its lack of human intervention, and thus lower effort,
+
 supervised learners have the advantage of domain knowledge, which typically means improved results.
+
 % NEIL - we got criticized for making it sound like this is the obvious thing to do... and then why do you need a learner?
+
 %Our manual annotations were fairly quick to do, taking only a few minutes per 30 day period.
+
 Creating annotated topics (i.e., manual labels) for training is painstaking, but with a suitably representative set of topics, the effort is acceptable. To
+
 annotate \emph{all} topics took us approximately 20 hours per project, but we estimate only 10\% of the topics need annotation to produce useful results.
@@
 Very rarely did \textsf{exp2} and \textsf{exp3} (semi-unsupervised word matching) ever perform as well as the supervised machine learners.
+
 For MaxDB, \textit{reliability} was slightly better detected using the static word list of \textsf{exp2}.
+
 In general, the machine learners and \textsf{exp3} did better than \textsf{exp2} for both MaxDB and MySQL.
+
 For both MySQL and MaxDB \textit{usability} was better served by \textsf{exp2}.
+
 \textit{Usability} was a very infrequent label, however, which made it difficult to detect in any case.
@@
 The semi-unsupervised labelling had difficulty distinguishing between common labels and infrequent labels.
+
 The learners would occasionally mislabel a topic deserving of an infrequent label with a more common label.
+
 The word-lists for \emph{correctness} tended to be too lengthy, non-specific and broad, especially if WordNet words were used, since the NFRs are
+
 typically loosely defined concepts in common parlance.
@@
 We found that the multi-label learners of BR, CLR and HOMER only did
+
 as well or worse for Macro-ROC as the single-label Naive Bayes and other naive Bayes-derived learners.
+
 This suggests that by combining together multiple Naive Bayes learners
+
 we could probably label sets of topics effectively, but it would
+
 require a separate Naive Bayes learner per label.
@@
 With ROC values ranging from $0.6$ to $0.8$ we can see there is promise in these methods.
+
 \textsf{exp2} and \textsf{exp3} both indicate that static information can be used to help label topics without any training whatsoever.
+
 MySQL and MaxDB's machine learners made some decisions based off a few shared words: \textsf{bug, code, compiler, database, HP UX, delete, memory,
+
 missing, problems, removed, add, added, changed, problem, and test}.
+
 Adding these words to the word-lists of \textsf{exp2} and \textsf{exp3} could improve performance while ensuring they were only domain specific.
@@
 If the techniques used in \textsf{exp2} and \textsf{exp3} were combined with the supervised techniques, we could reduce the training effort by boosting
+
 training sets with topics classified with the semi-unsupervised techniques.
+
 Both Naive Bayesian learners and the word-list approaches were computationally efficient.
+
 %These results are promising, because the accuracy of the techniques is good enough to be useful, while the run-time performance is still cheap enough
+
 %to execute to be feasible as an automated or semi-automated method of labelling topics by their software qualities.
+
 These results are promising because they indicate that these
+
 techniques are accurate enough to be useful while still maintaining
+
 acceptable run-time performance.
@@
 Since this work focuses on labelling natural language commit log
+
 comments we feel it can be adapted to other natural language artifacts
+
 such as mailing-list discussions and bug reports, even though that was
+
 not evaluated in this paper. Bug reports might not exhibit the same
+
 behaviour as commits in terms of dominant topics.
@@
 \subsection{Threats to Validity}
+
 % can always be cut down
+
 %Our work faced multiple threats to validity and we attempted to address many of them.
+
 \emph{Construct validity} -- we used only commit messages rather than
+
 mail or bug tracker messages.
+
 To extend further we would need matching repositories for each project.
+
 %Each case study was BecauseSince we required a seamless record, it was not possible to obtain
+
 %other data sources.
+
 Possibly they would have influenced our results, but there would be a degree of correlation between the corpora.
+
 %Conversely, construct validity was addressed in our validation as we used the files, topics and revisions to annotate the commit messages.
+
 %We used mailing-lists to verify the purpose behind some events.
+
 Our taxonomy for software NFRs is subject to dispute, but seems
+
 generally accepted. Finally, there are exogenous sources, such as
+
 in-person discussions, which we did not access, an omission as noted
+
 in Aranda et al.~\cite{aranda09icse} %We did not show these results to the stakeholders.
@@
 \emph{Internal validity} -- %includes inter-rater reliability and our lack of rival explanation building.
+
 We improved internal validity by trying to correlate and explain the behaviours observed in the analysis with the historical records of the projects.
+
 We did not attempt to match our results to any particular model.
+
 We assessed inter-rater reliability (IRR), a measure of the correlation between the manual annotations the first two authors created. Perfect IRR (1.0) indicates complete agreement
+
 on the appropriate annotation for that topic. We verified IRR using
+
 manually annotated PostgreSQL topics, and achieved a result of $0.1$.
+
 Some of the reasons for this are a disagreement over the appropriate classification for documentation topics (should this be considered
+
 usability or portability?). A short education session on what a given NFR from the ISO standard encompasses would reduce this error rate.
@@
 \emph{External validity} -- %was addressed by our use of two case studies, but there are still issues.
+
 Our data originated from OSS database projects and thus might not be applicable to commercially developed software or other domains.
+
 Furthermore, our analysis techniques rely on a project's use of meaningful commit messages, although we feel this is the most frequently occurring form
+
 of developer comments.
+
 %The OSS projects investigated were database systems, thus
+
 %our results might not generalize to other domains. %We only studied database systems.
+
 % \todo[inline]{If we compare with the same domain we don't get to say it generalizes a lot?}
@@
 % in Case Study Research: Design Case Studies
+
 \emph{Reliability} -- each annotator, the first two authors, followed the same protocol and used the same annotations.
+
 However, only two annotators were used; their annotations could be
+
 biased as we did not analyze for inter-rater reliability because they
+
 did not rate the same documents.
@@
 \subsection{Future Work}
+
 There are several avenues of further investigation.
+
 More external validation would be useful.
+
 Although we validated our comparisons using a mailing list for each project, interviews with developers would provide more detail.
+
 We also think multi-label learning techniques, although in their infancy, are crucial in understanding cross-cutting concerns such as NFRs.
+
 We want to leverage different kinds of artifacts to discover threads of NFR-related discussions that occur between multiple kinds of artifacts.
+
 Finally, we would like to extend this analysis to other domains, to see what patterns might occur in, for example, a consumer-facing software product.
@@
 \section{Conclusions}
+
 This paper presented a cross-project data mining technique,
+
 \emph{labelled topic extraction}. %The technique leveraged a
+
 %non-functional requirements (NFRs) taxonomy to mine and label latent
+
 %software development topics.
+
 Previous topic analysis research produced project-specific
+
 topics that needed to be manually labelled.
+
 %, but topics without labels
+
 %are difficult to interpret, and manually labelling them is
+
 %costly. Furthermore, previous work was project-specific and thus incomparable.
+
 To improve on this, we leveraged software engineering standards,
+
 specifically the ISO9126 quality taxonomy, to produce a method of partially-automated (supervised) and fully-automated (semi-unsupervised) topic labelling.
+
 Since the word-list technique is not project-specific, we used it to compare two distinct projects, where we showed our technique produced interesting
+
 insight into maintenance activity. %topics and artifacts from project histories.
+
 % Our contributions include:
+
 % \begin{itemize}
+
 % \item We provided an unsupervised method of topic labelling (word-lists);
+
 % %\item Our unsupervised method can be applied to projects without training;
+
 % \item We demonstrated a supervised method of labelling topics by a single NFRs (machine learning);
+
 % \item We demonstrated a supervised method of labelling topics by multiple NFRs (multi-label machine learning);
+
 % \item We provided a method of cross-project analysis via topic labelling;
+
 % \item We demonstrated a method of visualizing these topics across time, which we used to analyze maintenance activities.
+
 % \end{itemize}
@@
 We validated our topic labelling techniques using multiple experiments. % on unsupervised learners, supervised and multi-label learners.
+
 We first conducted semi-unsupervised labelling using word-lists.
+
 Our next approach was supervised, using single-label and multi-label learners.
+
 Both kinds of learners performed well with average
+
 %area under the
+
 ROC values between $0.6$ and $0.8$. These results, along with the efficient performance of our learners, demonstrate that labelled topic extraction is
+
 a promising approach for understanding the occurrence of non-functional requirements in software projects.
+
 %Although it may seem like common knowledge that maintenance is a concern as a project matures, our technique was able to show this independently.
+
 %We showed that NFRs are often trending topics in commit comments
+
 %, and that NFRs are quite common as developer topics.
+
 %; and that there are efficient methods of automating topic labelling.
@@
 % We demonstrated multiple supervised and unsupervised methods of topic labelling, including multi-label learners.
+
 % We provided a case study and visualization of NFR related topics in MaxDB 7.500 and MySQL 3.23.
+
 % During the case study we manually labelled hundreds of topics by hand and then used these annotations to validate the positive performance of topic
+
 %labelling techniques.
+
 % The implication of our paper is that topics with labels are more useful for analyzing software development activities than are unlabelled topics. To
+
 %furnish stakeholders with labelled topics, our technique provides many methods of automatic topic labelling.
@@
 %\begin{acknowledgements}
+
 %\end{acknowledgements}
@@
 % BibTeX users please use one of
+
 %\bibliographystyle{spbasic}	  % basic style, author-year citations
+
 \bibliographystyle{spmpsci}	 % mathematics and physical sciences
+
 %\bibliographystyle{spphys}	  % APS-like style for physics
+
 \bibliography{ese}   % name your BibTeX data base
@@
 \end{document}
+
 % end of file template.tex

### 2011-08-28 ese/ese.tex
removed 'total' number for topics

@@
 
-A \emph{named topic} is a topic with a matching NFR label. \emph{Unnamed
-
-topics} occur where there is no such match, which may indicate either a lack of precision, or simply that this topic is not associated with non-functional
+A \emph{named topic} is a topic with a match. 
+\emph{Unnamed topics} occur where there is no such match. 
+This may indicate either a lack of precision in the word-lists, or simply that this topic is not associated with non-functional
 
@@
 %   & Total  Topics  & 1000 & 1000 & 1000 \\
-% \midrule
-% PgSQL 3.23  & Named Topics   & - & - & - \\
-%   & Unnamed Topics & - & - & - \\
-%   & Total  Topics  & - & - & - \\
 	\centering
@@
              & Unnamed Topics & 84  & 206 & 59  \\
-             & Total Topics   & 400 & 400 & 400 \\
+             %& Total Topics   & 400 & 400 & 400 \\
 \midrule
@@
              & Unnamed Topics & 245 & 384 & 117  \\
-             & Total Topics   & 1000 & 1000 & 1000 \\
+             %& Total Topics   & 1000 & 1000 & 1000 \\
 \midrule
@@
              & Unnamed Topics & 1   & 97  & 0  \\
-             & Total Topics   & 640 & 640 & 640 \\
+             %& Total Topics   & 640 & 640 & 640 \\
 \midrule
@@
 
+\todo{this discussion is about which DB?}
 For \textsf{exp1} the labels with the most topics
@@
 
-For each quality we assessed whether semi-unsupervised labels matched the manual annotations. %To be clear, the manual annotations were not used to
+For each quality we assessed whether semi-unsupervised labels matched the manual annotations. 
+To be clear, the manual annotations were not used to
 

### 2011-08-29 ese/ese.tex
updated unsup numbers for PGSQL based on exp2/3

@@
 
-\todo{this discussion is about which DB?}
+\todo{this discussion is about which DB? where do the #s come from?}
 For \textsf{exp1} the labels with the most topics
@@
 
-%\todo[inline]{where is exp1 report?}
-
-%EXP1 could not be evaluated properly!
-
 Because our ground truth annotations were relevant only to ISO9126,
@@
 
+\todo{update these numbers but from where? }
 % 0.177215189873 / 6
@@
 
-% EXP2 source is naming-paper/src/validate/exp2.maxdb.report.txt
+% EXP2 source is latest-reports/exp2-unsup-results.xls. Numbers reflect latest runs as of 08/2011.
 
@@
 
+% column format is determined by info-theory.py : accuracy	f1	fn	fp	mcc	precisions	recall	roca	tn	tnr	tp
 For \textsf{exp2}, the average F-measure (macro-F1) for MaxDB was  $0.24$ with a range $0.091$ to
 
-$0.37$, and $0.16$ for MySQL with a range of $0$ to $0.41$.
-
+$0.37$, and $0.16$ for MySQL with a range of $0$ to $0.41$. PostgreSQL had an average F-measure of $0.30$ %using pgsqln
+ with a range of $0.09$ to $0.38$.
 MaxDB had an average precision and recall of $0.25$ and $0.22$
 
-while MySQL had $0.41$ and $0.10$ respectively.
+while MySQL had $0.41$ and $0.10$ and PostgreSQL $0.31$ and $0.29$, respectively.
 
@@
 $0.47$, and $0.36$ for MySQL with a range of $0.10$ to $0.65$. 
-
+PostgreSQL had an average of $0.42$ with a range of $0.31$ to $0.54$.
 MaxDB had an average precision and recall of $0.16$ and $0.67$
 
-while MySQL had $0.3$ and $0.48$ respectively.
-
-Thus we find that \emph{reliability} and
+while MySQL had $0.3$ and $0.48$. PostgreSQL had precision and recall of $0.27$ and $0.95$, respectively.
+
+Based on this we found that \emph{reliability} and
 
@@
 but other qualities such as \emph{functionality} did improve. 
-
+For PostgreSQL in \textsf{exp2}, \emph{reliability} and \emph{efficiency} were likewise the most accurate,
+while \emph{functionality} remained poor. \emph{Functionality} improved dramatically by \textsf{exp3}.
 Many ROC scores were $0.6$ or less, but our classifier still performed substantially better than random.

### 2011-08-29 ese/ese.tex
removed ctrl-m line endings from abram - I think we can tell the server to ignore these. Also fixed small bug where hash sign was used

@@
+\RequirePackage{fix-cm}
 %
-
-\RequirePackage{fix-cm}
-
+%\documentclass{svjour3}                     % onecolumn (standard format)
+%\documentclass[smallcondensed]{svjour3}     % onecolumn (ditto)
+\documentclass[smallextended]{svjour3}       % onecolumn (second format)
+%\documentclass[twocolumn]{svjour3}          % twocolumn
 %
-
-%\documentclass{svjour3}                     % onecolumn (standard format)
-
-%\documentclass[smallcondensed]{svjour3}     % onecolumn (ditto)
-
-\documentclass[smallextended]{svjour3}       % onecolumn (second format)
-
-%\documentclass[twocolumn]{svjour3}          % twocolumn
-
+\smartqed  % flush right qed marks, e.g. at end of proof
 %
-
-\smartqed  % flush right qed marks, e.g. at end of proof
-
-%
-
 \usepackage{graphicx}
-
  \usepackage{booktabs}
-
 \usepackage[caption=false]{subfig}
-
  \usepackage[colorinlistoftodos,backgroundcolor=yellow]{todonotes} %disable
-
  \usepackage{color}
@@
 \newcommand{\XXX}[1]{\textcolor{red}{{\it \textbf{[XXX: #1]}}}}
-
 \hyphenation{Max-DB data-base}
@@
 \usepackage{comment}
-
 \newtheorem{mydef}{definition}
-
 \newcommand{\slbl}[1]{\textbf{\textsf{#1}}}
-
 \usepackage{multirow}%,multicolumn}
-
 \usepackage{slashbox}
-
 \usepackage{wrapfig}
-
 \usepackage{booktabs, colortbl}
-
 \usepackage{amsmath}
-
 \usepackage{url}
-
 \usepackage[pdfstartview=FitH,bookmarksopenlevel=3,bookmarks=true]{hyperref} %pdfpagescrop={92 112 523 778},
-
 \journalname{Empirical Software Engineering}
@@
 
-\title{Automated Topic Naming
-
+\title{Automated Topic Naming}
 %\thanks{Grants or other notes
-
 %about the article that should go on the front page should be
-
 %placed here. General acknowledgments should be placed at the end of the article.}
 
+\subtitle{Supporting Cross-project Analysis of Software Maintenance Activities}
+%\titlerunning{Short form of title}        % if too long for running head
+
+%\numberofauthors{4}
+
+\author{
+Abram Hindle \and
+Neil A. Ernst \and
+Michael W. Godfrey \and
+John Mylopoulos
 }
-
-\subtitle{Supporting Cross-project Analysis of Software Maintenance Activities}
-
-%\titlerunning{Short form of title}        % if too long for running head
-
-%\numberofauthors{4}
-
-\author{
-
-Abram Hindle \and
-
-Neil A. Ernst \and
-
-Michael W. Godfrey \and
-
-John Mylopoulos
-
+%\authorrunning{Short form of author list} % if too long for running head
+
+\institute{
+  Abram Hindle \at 
+  Dept. of Computer Science\\
+  University of Alberta\\
+  Edmonton, AB, CANADA\\
+  \email{abram@softwareprocess.es}
+\and
+  Neil A. Ernst \at
+  Dept. of Computer Science\\
+  University of British Columbia\\
+  Vancouver, BC, CANADA\\
+  \email{nernst@cs.ubc.ca}
+\and
+  Michael W. Godfrey \at
+  David Cheriton School of Computer Science\\
+  University of Waterloo\\
+  Waterloo, Ontario, CANADA\\
+  \email{migod@uwaterloo.ca}
+\and 
+  John Mylopoulos \at
+  Dept. Information Eng. and Computer Science\\
+  University of Trento\\
+  Trento, ITALY\\
+  \email{jm@disi.unitn.it}
 }
 
-%\authorrunning{Short form of author list} % if too long for running head
-
-\institute{
-
-  Abram Hindle \at
-
-  Dept. of Computer Science\\
-
-  University of Alberta\\
-
-  Edmonton, AB, CANADA\\
-
-  \email{abram@softwareprocess.es}
-
-\and
-
-  Neil A. Ernst \at
-
-  Dept. of Computer Science\\
-
-  University of Toronto\\
-
-  Toronto, Ontario, CANADA\\
-
-  \email{nernst@cs.toronto.edu}
-
-\and
-
-  Michael W. Godfrey \at
-
-  David Cheriton School of Computer Science\\
-
-  University of Waterloo\\
-
-  Waterloo, Ontario, CANADA\\
-
-  \email{migod@uwaterloo.ca}
-
-\and
-
-  John Mylopoulos \at
-
-  Dept. Information Eng. and Computer Science\\
-
-  University of Trento\\
-
-  Trento, ITALY\\
-
-  \email{jm@disi.unitn.it}
-
-}
-
 \date{Received: date / Accepted: date}
-
 % The correct dates will be entered by the editor
@@
 \begin{abstract}
-
 % %\todo[inline]{Topics and LDA sooner, Topics/concept location is a known
-
 Researchers have employed a variety of techniques to extract underlying
-
 topics that relate to software development artifacts.  Typically, these
-
 techniques use semi-unsupervised machine-learning algorithms to suggest
-
 candidate word-lists.  However, word-lists are
-
 difficult to 
-
 interpret in the absence of meaningful summary labels.  Current topic modeling 
-
 techniques assume manual labelling and do not use domain-specific knowledge
-
 to improve, contextualize, or describe results for the developers.  
-
 We propose a solution: \emph{automated labelled topic extraction}.
-
   Topics are extracted 
-
 using  Latent Dirichlet Allocation (LDA) from commit-log comments recovered
-
 from source control systems such as CVS and BitKeeper.  
-
 These topics are given labels from a generalizable
-
 cross-project taxonomy, consisting of non-functional
-
 requirements. Our approach was evaluated with experiments and case studies
-
 on three large-scale RDBMS projects: MySQL, PostgreSQL and MaxDB.  The case studies
-
 show that labelled topic extraction can produce appropriate,
-
 context-sensitive labels relevant to these projects, which provides fresh
-
 insight into their evolving software development activities.
-
 %160 words
-
 \keywords{ Software maintenance \and Repository mining \and Latent dirichlet allocation \and Topic models}
@@
 \section{Introduction}
-
 A key problem for practising software maintainers is gaining an
-
 understanding of \emph{why} a system has evolved the way it has. 
-
 This is different than \emph{how} a system has evolved, because the behaviour of change itself is the
-
 \emph{how}.
-
 Looking back on streams of artifacts scattered across different
-
 repositories, inferring what activities were performed, when, and for
-
 what reasons, is hard to do without expert advice from the developers
-
 involved. 
-
 %new stuff
-
 % edit this. awkward
-
 In this work we seek to provide a method of automatically labelling  development topics extracted from commit logs.
@@
 Topic modeling is a machine learning technique which creates
-
 multinomial distributions of words extracted from a text corpus. 
-
 This technique infers the hidden structure of a corpus using posterior
-
 inference: the probability of the hidden structure given the data. 
-
 Topic models are useful in software maintenance because they summarize
-
 the key concepts in a corpus, such as source code, commit comments, or
-
 mailing-list messages, by identifying statistically co-occurring words. 
-
 Among other uses, it can quickly give developers an overview of where significant
-
 activity has occurred, and gives managers or maintainers a sense of
-
 project history. 
-
 %A major challenge for topic modeling is to accurately identify the
-
 %summary label for a topic; in the previous case, perhaps
-
 %\emph{``event-handling''} would be appropriate. 
-
 %Current approaches are manual: they rely on human expertise to
-
 %recognize the appropriate label. 
-
 %Furthermore, to date the results of topic modeling are project
-
 %specific, and not generalizable. 
-
 %This paper presents an approach, \emph{labelled topic extraction}, which addresses these two problems.
@@
 %migod
-
 While machine learning techniques can automatically identify clumps of
-
 commonly recurring terms, devising an appropriate summary label for
-
 each clump/topic is harder.  
-
 A given topic extracted from a set of commit logs might consist of the following terms: \emph{ ``listener change remove add fire''}. 
-
 This topic might reasonably be labelled as
-
 ``\emph{event handling}'' by a developer who understands the domain well,
-
 despite the fact that this label does not appear in the word list itself.  
-
 Current approaches to topic labelling rely on manual intervention by
-
 human experts, and also are limited to project-specific topic labels.  
-
 In this paper, we introduce \emph{labelled topic extraction}, an
-
 approach to topic labelling that creates labels automatically that are
-
 project independent.
-
 %is both automatic and creates labels that are project independent.
@@
 % Changes to software systems, revisions, do not occur without
-
 % reason. 
-
 % The reason or purpose is often related to requirements 
-
 % Often one revision might have multiple purposes ranging from
-
 % maintenance issues, quality issues, requirements, features or a
-
 % response to a bug report.
-
 % These purposes often are exhibited as latent topics that relate to
-
 % many development artifacts such as:
-
 %  version control revisions, bug reports, and mailing-list discussions.
@@
 %In our previous work we dealt with topic trends, which are topics that recur over time~\cite{Hindle09ICSM}. 
-
 %We observed that topic trends were often non-functional requirements (NFRs). 
-
 %NFRs have the property of being cross-domain and widely applicable. 
-
 %In this sense, they are useful abstractions for developer conversations about different software projects.
@@
 % harp on data mining more
-
 %In general, the mining of software artifacts tends to be very project specific, yet NFRs are not. 
-
 % I was told if you're going to use a paren, you don't need to say it.
-
 %There is a series of standards on NFRs, such as ISO9126~\cite{iso9126}, specifically intended to apply to projects of varying types.
-
 %These standards indicate our approach is possible. %, if perhaps not universally accepted).
-
 % NFRs are prevalent in almost all software systems.
-
 % In this paper we will show that NFR related topics occur quite frequently.
-
 %XXXXXX this sentence is confusing
-
 %This informs our method of generalizing the data mining of these software artifacts to leverage the global domain knowledge of software engineering,
-
 %specifically NFRs. 
-
 %In particular, we use software quality models to generalize topics across projects. 
-
 %This allows us to mine software systems with the intention of comparing their NFR-related development topics.
-
 % in order to achieve generalizable, cross project data mining results.
@@
 %%%% THIS PARAGRAPH ARGUES ABOUT THE CONCRETE APPLICATION OF 
-
 %%%% TOPIC MODELLING AND LABELLED TOPIC EXTRACTION
-
 %%%% IT NEEDS TO BE TIGHTER BUT IT ALSO NEEDS TO BE HERE
@@
 % AH: Topics take a long to read
-
 % AH: Project Dashboards, summarization, software process recovery
@@
 %migod
-
 In general, the fruits of mining software artifacts are often project
-
 specific and hard to generalize.  
-
 However, in our previous work we
-
 investigated \emph{topic trends} --- that is, topics that recur over
-
 time --- we observed that topic trends often corresponded to
-
 non-functional requirements (NFRs)~\cite{Hindle09ICSM}, which is
-
 further emphasized in this paper due to the large numbers of NFR
-
 labelled topics.  
-
 This is encouraging, as NFRs have the property of being cross-domain
-
 and widely applicable. 
-
 In this sense, they are useful abstractions for developer
-
 conversations about different software projects.  
-
 Furthermore, there is a series of standards on NFRs, such as ISO9126 \cite{iso9126}, that are specifically intended to apply to projects of varying
-
 types; this suggests that our goal of trying to extract NFR-related development topics, such as those related to software quality models, holds promise.
@@
 Concrete applications of topics and \emph{labelled topic extraction}
-
 range from project dashboards to annotating software artifacts such as revisions and bug reports with
-
 NFR-related tags.
-
 Project dashboard~\cite{dashboard} are typically employed by managers and are used to
-
 provide quick summaries of the effort put into a software project. In this case, 
-
 labelled topic extraction would allow managers to track effort
-
 related to NFR topics, such as portability.
-
 These techniques also allow for the annotation of commit comments
-
  and other software artifacts with NFRs. 
-
 This would enable querying of bug reports and artifacts by relevant NFRs. 
-
 For instance a manager can confirm if their developers were
-
 focused on \emph{usability} by looking for
-
  \emph{usability}-relevant revisions and bug reports.
@@
 In this paper, we describe \emph{automated labelled topic extraction}. It addresses two gaps in the topic mining literature:
-
 \begin{enumerate}
-
   \item Topic mining of software has been limited to one project at a time. 
-
 This is because traditional topic mining techniques are specific to a particular data-set. 
-
 \textit{Automated labelled topic extraction} allows for comparisons \textit{between} projects. 
-
   \item Topic modeling creates word lists that require interpretation by the user to assign meaning. 
-
 Like (1), this means
-
 that it is difficult to discuss results independent of the project context. 
-
 Our technique automatically, or with some initial training, assigns labels across projects.
-
 \end{enumerate}
@@
 This paper makes the following contributions: 
-
 \begin{itemize}
-
 \item introduces the concept of labelled topic extraction, using a non-functional requirements (NFR) taxonomy for our labels;
@@
 %\item evaluates an unsupervised method of automated topic labelling (word-lists);
-
 %%\item Our unsupervised method can be applied to projects without training;
-
 %%\item 
-
 %evaluates a supervised method of automatically labelling topics with a single NFR (machine learning);
-
 %%\item 
-
 %evaluates a supervised method of labelling topics with multiple NFRs
-
 %(multi-label machine learning);
@@
 \item evaluates three kinds of automatic topic labelling methods:
-
   semi-unsupervised labelling  of topics (word-lists), supervised labelling of
-
   topics with a single NFR (machine learning), 
-
   and supervised labelling of topics with multiple NFRs (multi-label
-
   machine learning);
@@
 %  of automated topic labelling (word-lists);
-
 %\item Our unsupervised method can be applied to projects without training;
-
 %\item 
-
 %evaluates a supervised method of automatically labelling topics with a single NFR (machine learning);
-
 %\item 
-
 %evaluates a supervised method of labelling topics with multiple NFRs (multi-label machine learning);
@@
 \item examines how NFRs correlate with the work of individual developers;
-
 \item provides a method of cross-project analysis via topic labelling; and
-
  applies these techniques to visualize NFRs over time, and to analyze maintenance activities.
@@
 %We first introduces labelled topic extraction, and then uses that technique to analyze development activity in two large-scale examples.
-
 We begin by discussing related work in Section \ref{sec:related}.
-
 Next, we describe how we generated our data (Section \ref{sec:wordlist}). For semi-unsupervised classification (Section \ref{sec:unsuplabelling}), we
-
 begin by creating word-lists to signify when a topic matches an NFR label. We then apply our classifier and analyze the results. 
-
 %and how we created the word-lists used in unsupervised classification. 
-
 %We show that labels with their topics can be learned and used to classify other data-sets, either without training 
-
 %Our goal is to label a given topic with either an NFR from our taxonomy,
-
  %Section \ref{sec:unsuplabelling} presents our technique for using word-lists to do unsupervised classification, and describes the results.
-
 In Section \ref{sec:suplabelling}, we manually annotate the topics, and use those annotations as training data for supervised classification.  
-
 %We then present visualizations of named topics and their trends over time to aid communication and analysis. 
-
 To demonstrate an application of labelled topic extraction, we use an exploratory case study of two open source database systems to show how named
-
 topics can be compared between projects  (Section \ref{sec:analysis}). 
-
 The paper concludes with a discussion of limitations (Section \ref{sec:limit}), and future work.
@@
 \section{Previous Work}
-
 \label{sec:related}
@@
 The idea of extracting higher-level \emph{concerns} and \emph{topics}, also known as
-
  \emph{concepts}, \emph{aspects} or \emph{requirements},
-
 has been approached from documentation-based and repository-based
-
 perspectives.
@@
 Cleland-Huang and her colleagues have investigated mining requirements documents for non-functional requirements (NFR) (quality
-
 requirements)~\cite{Cleland-Huang2006}. 
-
 Their approach is similar to ours, as they mined keywords from NFR
-
 catalogues.
-
 They differ because they mine requirements documents where as we mine revisions.
-
 %One approach they tried was similar to this one, with keywords mined from NFR catalogues found in their previous work~\cite{chung99}. 
-
 They demonstrated a recall of $80\%$ with precision of $57\%$ for the \emph{security} NFR, but could not find a reliable source of keywords for other NFRs. 
-
 Instead, they developed a supervised classifier by using human experts to identify an NFR training set. 
-
 Our research differs because we use a more comprehensive set of terms based on the taxonomy we chose.
-
 Another difference is that we make cross-project comparisons instead of focusing
-
 on a single project.
-
 %They relied on requirements documents instead of version control
-
 %histories that we use.
-
 %The objective of Cleland-Huang's study was to identify new NFRs for system development, 
-
 %yet our objective was to recover those latent NFRs from commit-log messages of the project.
@@
 %There are several reasons we did not follow this route. 
-
 %One, we believe we have a more comprehensive set of terms due to the taxonomy we chose. 
-
 %Secondly, we wanted to compare across projects. 
-
 %Their technique was not compared across different projects and the applicability of the training set to different corpora is unclear. 
-
 %A common taxonomy allows us to make cross-project comparison. 
-
 %Although subject to the assumption that all projects conceive of these terms in the same way.
-
 %Thirdly, while the objective of Cleland-Huang's study was to identify new NFRs (for system development) 
-
 %our study assumes these NFRs are latent in the commit-log messages of the project.
@@
 Similarly, Mockus and Votta~\cite{Mockus00} studied a large-scale industrial change-tracking system. 
-
 Mockus and Votta leveraged WordNet~\cite{Fellbaum1998}, an English-language ``lexical database'' that contains semantic relations between words,
-
 including common related forms (similar to word stemming), meronymy and synonymy.
-
 They used WordNet for word roots as they felt the synonyms would be
-
 non-specific and cause errors.
-
 %, problems we also encountered. 
-
 Mockus et al. validated their labels with system developers.
-
 Since we study multiple projects, instead of a single project, these
-
 kind of interviews are somewhat infeasible (particularly in the distributed world of open-source software).
@@
 Another approach is to extract concerns from software repositories.
-
 Marcus et al.~\cite{marcus04wcre} use Latent Semantic Indexing (LSI)
-
 to identify commonly occurring concerns for software maintenance. 
-
 The
-
 concerns are given by the user, and LSI is used to retrieve them from
-
 a corpus. 
-
 Topic modelling generates topics that are independent of a user query, and relate only to word frequencies in the corpus.
-
 %Some results were interesting, but their precision was quite low. 
-
 With ConcernLines, Treude et al.~\cite{treude09cl} show tag occurrence
-
 using colour and intensity. 
-
 They mine
-
 developer created tags 
-
 %change request tags (such as `milestone 3') and used these 
-
 in order to analyze the evolution of a single product.
-
 % make evolutionary analyses of a single product. 
-
 The presence of a well-maintained set of tags is obviously essential to the success of this technique.
@@
 %XXXXXXXXXXXX OPTIONAL
-
 % Not sure we need this
-
 %Mens et al.~\cite{mens08icsm} conducted an empirical study of Eclipse, a popular Java IDE, to verify the claims of Lehman~\cite{lehman97sms}.
-
 % They concerned themselves with source code only, and found Law Seven, ``Declining Quality'', to be too difficult to assess:
-
 % ``[we lacked an] appropriate measurement of the evolving quality of the system as perceived by the users \cite[p. 388]{mens08icsm}''. 
-
 %This paper examines the notions of quality in terms of a consistent ontology, as Menset al. call for in their conclusions.
@@
 % in our previous work~\cite{Hindle09ICSM}, as well as -- CITED a little too often (vanity cites!)
-
 In Baldi et al.~\cite{Baldi2008}, topics are named manually: human
-
 experts read the highest-frequency members of a topic and assign a
-
 label accordingly. 
-
 %XXX Possible to cut this
-
 As discussed earlier, given the topic \emph{``listener change remove add fire''}, Baldi et al. would assign the label \emph{event-handling}. 
-
 The labels are reasonable enough, but still require an expert in the field to determine them. 
-
 Furthermore, these labels are project-specific, because they are generated from the data of that project. E.g., we might have a label called \emph{`Oracle'}
-
 in the MySQL case, since Oracle owns MySQL. 
-
 %We do this differently.
-
 %This paper does this differently. 
-
 Our approach differs: first of all, we automate the process of naming the topics; secondly, we label topics with project-independent terms, in order
-
 to permit cross-project comparison.
@@
 Mei et al.~\cite{Mei2007} use context information to automatically name topics. 
-
 They describe probabilistic labelling, using the frequency distribution of words in a topic to create a meaningful phrase. 
-
 They do not use external domain-specific information as we do, but we
-
 do not generate phrases from the topics.
@@
 Finally, in Ernst et al.~\cite{ernst10refsq}, we describe our earlier
-
 project, similar to this, that identifies changes in quality requirements
-
 in GNOME software projects; this approach was more exploratory, had
-
 less validation, uses different word-lists, solely uses text-matching,
-
 and does not leverage machine learning strategies. 
-
 Hindle et al.~\cite{Hindle09ICSM} propose a windowed method of topic
-
 analysis that we extend with labelled
-
 topics, NFRs and new visualizations.
@@
 %Massey~\cite{massey02icse} and Scacchi~\cite{scacchi02,scacchi05b} looked at the topic of requirements in open-source software. 
-
 % this reduces the line length and drops a ref even though we still have a scacchi kickin around
-
 %Massey~\cite{massey02icse} and Scacchi~\cite{scacchi05b} looked at the topic of requirements in open-source software. 
-
 %Their work discusses the source of the requirements and how they are used in the development process. 
-
 % German~\cite{german03gnome} looked at GNOME specifically, and listed several sources for requirements: leader interest, mimicry, brainstorming, and prototypes. 
-
 %None of this work  addressed quality requirements in OSS, nor did it examine requirements trends.
-
 \todo{re-add this related work?}
-
 % XXX vanity or space.. 
-
 %XXXXXXXXXXXX OPTIONAL
-
 % NE 0113 this seems too early to mention 'result'.
-
 % Hindle et al. \cite{Hindle2007} examined release patterns in OSS. They showed that there is a difference between projects regarding maintenance
-
 % techniques. This supports our result that software qualities are not discussed with the same frequency across projects.
@@
 \section{Study Design and Execution}
-
 Figure \ref{fig:process} gives an outline of our methodology.
-
 We began by gathering source data and creating topic models. For semi-unsupervised labelling, we generated three sets of word-lists as signifiers for NFRs.
-
 %e created labels using a well-known NFR taxonomy. 
-
 In supervised learning, we trained our data with manual annotations in order to match topics with NFRs. Finally, these topics were used to analyze the
-
 role of NFRs in software maintenance.
@@
 \begin{figure}
-
   \centering
-
  \includegraphics[width=.45\textwidth]{figures/process-model}
-
  \caption{Research methodology process view.}
-
   \label{fig:process}
-
 \end{figure}
@@
 \subsection{Generating the Data}
-
 \label{sec:wordlist}
@@
 To evaluate our approach, we sought candidate systems that were mature projects and had openly accessible source control repositories. 
-
 We selected systems from the same application domain, to control for differences in functional, rather than non-functional, requirements. 
-
 We used three different open-source, partially-commercial database systems:
-
 \begin{itemize}
-
 \item  MySQL 3.23. Started in 1994 and MySQL 3.23 was released in early 2001. MySQL contains $320,000$ lines of C and C++ source code\footnote{generated
-
 using David A. Wheeler's \emph{SLOCCount}, {http://dwheeler.com/sloccount}.}.
-
 \item MaxDB 7.500. Started in the late 1970s as a research project, and was later acquired by SAP. As of version 7.500, released April 2007, the project
-
 has over $940,000$ lines of C source code. 
-
 \item PostgreSQL 7.3. Started in the 1980s as a Berkeley research project. PostgreSQL 7.3 contains $306,000$ lines of C code.
-
 \end{itemize}
@@
 %Choosing an older version of these projects allowed us to focus on projects which have moved further into the maintenance phase of the software life-cycle.
-
 We explicitly chose older versions of mature projects from a stable problem domain to increase the likelihood that we would encounter primarily
-
 maintenance activities in our studies.
@@
 For each project, we used source control commit comments, the messages
-
 that programmers write when they commit revisions to a source control
-
 repository. 
-
 Most commits we observed had commit comments.
-
 Commit comments are often studied by researchers, as they are the most readily accessible source of project interactions, and developers are often
-
 required to create them by the repository mechanism (e.g., Git).  Additionally, relying only on commit comments makes our approach more generalizable,
-
 as we do not assume the presence of other artifact corpora.
-
 %This is the most readily accessible source of project interactions for outside researchers, since developers often or always (e.g., Git) write such
-
 %message. Our approach is more generalizable if it does not assume the presence of other corpora. 
-
 %bug trackers and email-list archives were not available for both projects. 
-
 An example of a typical commit message, from MySQL, is: \textit{``history annotate diffs bug fixed (if mysql\-\_real\-\_connect() failed there were
-
 two pointers to malloc'ed strings, with memory corruption on free(), of course)''}. 
-
 We extracted these messages and indexed them by creation time. 
-
 We summarized each message as a word distribution minus stop-words such as ``\emph{the}'' and ``\emph{at}''. 
-
 %Stemming was performed in the later stages of our analysis. % WAS IT?
@@
 For the commit message data-sets of each project, we created an XML file that separated commits into 30 day periods. 
-
 %This size of period, 30 days, is smaller than the time between minor
-
 %releases but large enough for there to be sufficient commits to
-
 %analyze
-
 We chose a period size of 30 days as it is smaller than the time between minor releases but large enough for there to be sufficient commits to
-
 analyze~\cite{Hindle09ICSM}. 
-
 For each 30 day period of each project, we input the messages of that period into Latent Dirichlet Allocation (LDA), a topic analysis
-
 algorithm~\cite{Blei2003}, and recorded the topics the algorithm extracted.
-
 %For each project and for each period of messages a project, we input these messages 
-
 %we created `topics' using Latent Dirichlet Allocation, a topic identification algorithm.
@@
 A topic analysis tool such as LDA will try to find $N$ independent
-
 word distributions within the word distributions of all input messages. 
-
 Linear combinations of these $N$ word distributions are meant to represent and recreate the word distributions of any of the original messages. 
-
 These $N$ word distributions effectively form \emph{topics}: cross-cutting collections of words relevant to one or more of our commit messages. 
-
 LDA extracts topics in an unsupervised manner; the algorithm relies
-
 solely on the source data and word distributions of messages, with no human intervention.
@@
 In topic analysis a single document, such as a commit message, can be related to multiple topics. 
-
 Representing documents as a mixture of topics maps well to source code repository commits, which often have more than one purpose~\cite{Hindle09ICSM}.  
-
 For this paper, a topic represents a word distribution generated from a group of commit log comments which are related by their content.  
-
 %In this paper a topic is a set of tokens extracted from commit messages found within a project's source control system (SCS).
@@
 We applied Blei's LDA implementation~\cite{Blei2003} against the word distributions of these commits, and generated lists of topics per period. 
-
 We set the number of topics to generate to $20$, because past experimentation showed that fewer topics might aggregate multiple unique topics while
-
 any more topics dilutes the results and creates indistinct topics~\cite{Hindle09ICSM}. 
-
 % Furthermore, more than $20$ topics quickly became infeasible for inspection.
@@
 To facilitate cross-project comparison, we used a taxonomy of NFRs. This taxonomy is based on the ISO quality model, ISO9126~\cite{iso9126}. 
-
 ISO9126 describes six high-level NFRs: maintainability, functionality,
-
 portability, efficiency, usability, and reliability~\footnote{While there may be lingering debate in some circles about these terms, an ISO standard
-
 seems like a reasonable starting point for our work.}.
-
 There is some debate about the terms in this model. 
-
 However, ISO9126 is ``an international standard and thus provides an 
-
 internationally accepted terminology for software quality \cite[p. 58]{Boegh2008},'' that is sufficient for the purposes of this research.  
-
 We claim that these NFRs are maintenance concerns (to varying degrees) in all software projects, and are therefore well-suited for comparisons between
-
 projects.
@@
 \subsubsection{Creating a Validation Corpus}
-
 To evaluate both semi-unsupervised and supervised classification, we
-
 created a validation set of manually labelled topics. For each project,
  the annotators (the first two authors) annotated each extracted topic in each period with
-
 the six NFR labels listed above.
-
 Annotators did not annotate each other's annotations, but some brief
-
 inspection of annotations was used to confirm that the annotators were
-
 acting similarly.
-
 We looked at each period's topics, and assessed what the data ---
-
 consisting of the frequency-weighted word lists and messages ---
-
 suggested was the label for that topic. 
-
 We were able to pinpoint the appropriate label using auxiliary information as well, such as the actual revisions and files that were related to the
-
 topic being annotated.
-
 For example, for the MaxDB topic consisting of a message ``exit() only
-
 used in non NPTL LINUX Versions'', we tagged that topic
-
 \emph{portability}. 
-
 Given the top-level annotations of \emph{none}, \emph{portability},
-
 \emph{efficiency}, \emph{reliability}, \emph{functionality},
-
 \emph{usability}, and \emph{maintainability}, the annotators annotated each topic
-
 with the relevant label. Sometimes they used finer-grained
-
 annotations that would be aggregated up to one of these higher-level labels.
-
 %\todo[inline]{this confused people. Is it truly unsupervised? Yes, but it needs to be clearer.}
-
 %We compared against this data-set, but we also used the data-set for our supervised machine learning based topic classification. 
-
 % A topic may have more than one matching keyword. % ok what?
-
 %\todo[inline]{inter-rater reliability }
-
 % \todo[inline]{There should be a section here which is about manual annotation. 
-
 %It is really important and it is required by both analyses. Plus we can tell them how.}
@@
 %\todo[inline]{NEIL: this section should be updated to explain F-measure}
-
 We validate classification performance using the Receiver Operating
-
 Characteristic area-under-curve value~\cite{Fawcett2006861},
-
 abbreviated \emph{ROC}, and the F-measure, which is the harmonic mean of precision and recall, i.e., $2 * (P * R) / (P + R)$.
@@
 ROC values provide a score %, similar to school letter-grades (A is 0.9, C is 0.6),
-
  reflecting how well a particular learner performed for the given data. 
-
 ROC maps to the more familiar concepts of precision/sensitivity and recall/specificity: it plots the true positive rate (sensitivity) versus the false
-
 positive rate (1 - specificity). 
-
 A perfect learner has a ROC value of 1.0, reflecting perfect recall and precision. 
-
 A ROC result of 0.5 would be equivalent to a random learner (that is, issuing as many false positives as true positives). 
-
 The %area under the -- NE confusing to have area again ... and not importantana
-
 ROC of a classifier is equivalent to the probability that the classifier will rank a randomly chosen positive instance higher than a randomly chosen
-
 negative instance.
-
 We consider our labelling classifiers acceptable if they outperform a random classifier (0.5).
@@
 \subsection{Semi-unsupervised Labelling}
-
 \label{sec:unsuplabelling}
@@
 In this section we describe how to label topics based on dictionaries
-
 mined from sources external to the projects.
@@
 %\subsection{Generating word lists}
-
 %\label{sec:genwordlist}
-
 % %\todo[inline]{Make it clear where the data comes from! Neil says: Also, I
-
 %   didn't get the impression they understood that we weren't generating
-
 %   labels from the data, but rather from independent taxonomies. Of
-
 %   course if you use labels from the data it isn't comparable. But the
-
 %   point of ISO 9126 etc. was to avoid exactly this problem.}
@@
 % \todo[inline]{Need to hammer it in that the word bags are not project specific and
-
 %  are inherited so it isn't a supervised method}
@@
 %Topics are word distributions, lists of words ranked by frequency, that are burdensome to interpret and hard to distinguish and understand. 
-
 %While the topics themselves are generated automatically, how to interpret these topics is less clear.
-
 %For example,
@@
 In order to automatically label each topic with 
-
 one of the six high-level NFRs,
-
 %an NFR from ISO9126, 
-
 we associate each NFR with a list
-
 of keywords (\emph{word-lists}, in our parlance). These word-lists were determined a priori and were not extracted from the projects themselves.
-
 We intersected the words of the topics and the words of our word-lists.
-
 We ``labelled'' a topic if any of its words matched any of the word-list's words.
-
 A topic could match more than one NFR.
-
 We used several different sets of word-lists for comparison, which we
-
 refer to as \textsf{exp1}, \textsf{exp2}, and \textsf{exp3} in the text which follows.
@@
 Our first word-list set, \textsf{exp1}, was generated using the ontology described in Kayed et al.~\cite{5072519} .
-
 That paper constructs an ontology for software quality measurement using eighty source documents, including research papers and international standards. 
-
 The labels we used:
@@
 \begin{quotation}
-
 \small \noindent \textsf{
-
 integrity, security,
-
 interoperability, testability, maintainability, traceability,
-
 accuracy, modifiability, understandability, availability, modularity,
-
 usability, correctness, performance, verifiability, efficiency,
-
 portability, flexibility, reliability.
-
 }
-
 \end{quotation}
@@
 Our second word-list, \textsf{exp2}, uses the ISO9126 taxonomy described above (Section \ref{sec:wordlist}) to seed the word-lists.
-
 %We use these qualities later on as classes in supervised labelling. NE confusing the issue here, I think.
-
 The terms from ISO9126 may not capture all words occurring in the topics that are nonetheless associated with one of the NFRs. 
-
 For example, the term ``redundancy'' is one 
-
 we considered to be
-
 relevant to discussion of \emph{reliability}, but is not in the standard. 
-
 We therefore took the NFRs from the ISO9126 and added to them.
@@
 %\todo[inline]{* clarify the mailinglist stuff we did}
-
 %\todo[inline]{This part doesn't make a lot of sense}
-
 To construct these expanded word-lists, we used
-
 WordNet~\cite{Fellbaum1998}.
-
 %, an English-language ``lexical database'' that contains semantic relations between words,
-
 % including common related forms (similar to word stemming), meronymy and synonymy. 
-
 We then added Boehm's software quality model~\cite{Boehm+:1976:ICSE}, and classified his eleven `ilities' into their respective ISO9126 NFRs. 
-
 We did the same for the quality model produced by McCall et al.~\cite{mccall1977}. 
-
 We then did a simple random analysis of mailing list messages from an open source project, KDE. If we judged a given message to contain terms that were
-
 related to one of the NFRs in ISO9126, we added it to our word-list. This allowed us to expand our word-lists with more software-specific terms.
-
 %Finally, we randomly analyzed two mailing lists from another software project to expand our set with domain-specific terms. 
-
 % Less info is probably better here...
-
 %For example, we add the term ``performance'' to the synonyms for \emph{efficiency}, since this term occurs in most mail messages that discuss efficiency.
-
 Table \ref{tbl:wnsig} shows the labels (NFRs) and word-lists we used for matching.
@@
 For the third set of word-lists, \textsf{exp3}, we extended the word-lists from \textsf{exp2} using unfiltered WordNet similarity matches. 
-
 Similarity in WordNet means siblings in a hypernym tree. 
-
 We do not include these words here for space considerations (but see the Appendix for our data repository). 
-
 It is not clear the words associated with our labels in \textsf{exp3} are specific enough. For example, the label \emph{maintainability} is associated with
-
 words \emph{ease} and \emph{ownership}. In general, as we proceed from word-list in \textsf{exp1} to that in \textsf{exp3}, our lists become more generic.
@@
 \begin{table*}
-
 	\centering
-
 \begin{tabular}{c|p{11cm}}
-
 \toprule
-
 \textbf{Label} & \textbf{Related terms} \\
-
 \midrule
-
 \emph{Maintainability} &
-
 testability changeability analyzability stability maintain maintainable modularity modifiability understandability interdependent dependency encapsulation
-
 decentralized modular\\ \hline
-
 \emph{Functionality} &
-
 security compliance accuracy interoperability suitability functional practicality functionality compliant exploit certificate secured ``buffer overflow''
-
 policy malicious trustworthy vulnerable vulnerability accurate secure vulnerability correctness accuracy\\ \hline
-
 \emph{Portability} &
-
 conformance adaptability replaceability installability portable movableness movability portability specification migration standardized l10n localization
-
 i18n internationalization documentation interoperability transferability\\ \hline
-
 \emph{Efficiency} &
-
 ``resource behaviour'' ``time behaviour'' efficient efficiency performance profiled optimize sluggish factor penalty slower faster slow fast optimization\\
-
 \hline
-
 \emph{Usability} &
-
 operability understandability learnability useable usable serviceable usefulness utility useableness usableness serviceableness serviceability usability
-
 gui accessibility menu configure convention standard feature focus ui mouse icons ugly dialog guidelines click default human convention friendly user
-
 screen interface flexibility\\ \hline
-
 \emph{Reliability} &
-
 ``fault tolerance'' recoverability maturity reliable dependable responsibleness responsibility reliableness reliability dependableness dependability
-
 resilience integrity stability stable crash bug fails redundancy error failure\\ 
-
 \bottomrule
-
 \end{tabular}
-
 	\caption{NFRs and associated word-lists -- \textsf{exp2}}
-
 	\label{tbl:wnsig}
@@
 %\todo[inline]{NEIL: this could be a subsubsection}
-
 %\todo[inline]{* the hierarchy of tags was not clear}
-
 %\noindent \textbf{Generating labels automatically} ---
@@
 Using our three word-lists (\textsf{exp1}, \textsf{exp2}, \textsf{exp3}), we labelled our topics with an NFR where there was a match between a word in
-
 the list and the same word somewhere in the distribution of words that constitute the topic.
-
 A \emph{named topic} is a topic with a match. 
@@
 This may indicate either a lack of precision in the word-lists, or simply that this topic is not associated with non-functional
-
 requirements.
-
 All experiments were run on the datasets for each project (e.g., PostgreSQL, MySQL, MaxDB). LDA
 extracted $20$ topics per period for each project.
-
 % Is this good enough
-
 This labelling is \emph{semi-unsupervised} because the corpus is not derived from 
-
 the project being analyzed, and we did not label the project's topics
-
 ourselves for a training set. The motivation behind this technique is that
-
 because most software often addresses similar issues, we can use the 
 domain knowledge of software to label relevant topics.
-
 %\todo[inline]{Neil does this explain the unsupervised aspect well enough?}
@@
 Table \ref{tbl:wordlist} shows how many topics were labelled for MaxDB
-
 and MySQL.
@@
 %\todo[inline,color=green]{this table should have mysql}
-
 \begin{table}
-
+\centering
 % % This data came from running  
-
 % %   python check_validity.py
-
 % %  in src/validate
-
 % % and then running
-
 % % tail -n 5 *maxdb*yes_no*
-
 % % tail -n 5 *mysql*yes_no*
-
 % ok the new way is to run lda_relator
-
 % src/validate/lda_relator_counts.sh
-
 % NOTE THESE WERE THE OLD VALUES
@@
 %   & Total  Topics  & 1000 & 1000 & 1000 \\
-	\centering
 
 \begin{tabular}{l|c|c|c|c}
-
 \toprule
-
 \textbf{Project} & \textbf{Measure} & \textsf{exp1} & \textsf{exp2} & \textsf{exp3} \\
-
 \midrule
-
 % updated for ESE: taken from latest-reports.tar.gz, last five lines of 
@@
              %& Total Topics   & 640 & 640 & 640 \\
-\midrule
 \bottomrule
-
 \end{tabular}
-
 	\caption{Automatic topic labelling for MaxDB, MySQL and PostgreSQL}
@@
 
-\todo{this discussion is about which DB? where do the #s come from?}
+\todo{this discussion is about which DB? where do the numbers come from?}
 For \textsf{exp1} the labels with the most topics
-
 were
-
 \emph{correctness} (182 topics) and \emph{testability} (121). 
-
 We did not see many results for \emph{usability} or \emph{accuracy}, which were associated with fewer than ten topics. 
-
 We also looked for correlations between our labels: excluding double matches (self-correlation), our highest co-occurring terms were \emph{verifiability}
-
 with \emph{traceability}, and \emph{testability} with \emph{correctness} (76 and 62 matches, respectively).
-
 % See the following sections for our error analysis.
@@
 For \textsf{exp2}, there are more unnamed topics than \textsf{exp1}. 
-
 Only \emph{reliability} produces a lot of matches, mostly with the word ``error''. 
-
 Co-occurrence results were poor. This suggests our word lists were overly restrictive.
@@
 For \textsf{exp3}, we generally labelled more topics. 
-
 As we mentioned, the word-lists are broad, so there are likely to be false-positives (discussed below). 
-
 The most frequent label, 265 topics, was \emph{usability}, and the
-
 least frequent label, 44 topics, was \emph{maintainability}. 
-
 Common co-occurrences were \emph{reliability} with \emph{usability}, \emph{efficiency} with \emph{reliability}, and \emph{efficiency} with \emph{usability}
-
 (200, 190, and 150 topics in common, respectively).
@@
 \subsubsection{Analysis of the Semi-unsupervised Labelling} %--- %Based
-
 				%on the labels, and our manual topic
-
 				%labelling, 
-
 %\noindent \textbf{Analysis of the unsupervised labelling} --- %Based on the labels, and our manual topic labelling, 
-
 %we compared the results of the unsupervised word matching approach. 
-
 For each quality we assessed whether semi-unsupervised labels matched the manual annotations. 
 To be clear, the manual annotations were not used to
-
 train the labelling process.
-
 As described in Section \ref{sec:wordlist} we used both ROC and F-1 measures to evaluate the performance of the classification.
-
 Figure \ref{fig:maxdb-unsup-results} shows our ROC results for MaxDB and MySQL. We describe F-1 results in the text below.
@@
 %\todo[inline]{since ROC is new-sounding, I wonder if a more lengthy explanation would be merited. An obvious question is why not precision/recall? --
-
 % ROC is hot right now and it combines both}
-
 % Rebuttal: ROC is not new and it is what the machine learning community, search based software engineering and now the MSR community are using +
-
 %it has natural meaningful quality mappings.
-
 % \begin{figure}[t]
-
 % \centering
-
 % \subfloat[ROC]{
-
 %  \includegraphics[width=0.45\textwidth]{figures/unsupervised-bar}
-
 % \label{fig:maxdb-unsup-results-roc}
-
 % }
-
 % %\subfloat[F1]{
-
 % %\includegraphics[width=0.45\textwidth]{figures/unsupervised-bar-f1}
-
 % %\label{fig:maxdb-unsup-results-f1}
-
 % %}
-
 % \caption[]{Performance of unsupervised topic labelling for each NFR per word-list. A) ROC values -- 
-
 % dashed line indicates the performance of a random classifier. B) F1 measure. Possible values range from 0--1.
-
 % %\todo[inline]{better discussion}% and graphical legend would be nice}
-
 % }
-
 % \label{fig:maxdb-unsup-results}
-
 % \end{figure}
@@
 \begin{figure*}[t]
-
   \centering
-
  \includegraphics[width=0.8\textwidth]{figures/unsupervised-bar}
-
  \caption{Performance, ROC values (range: $0$--$1$), of semi-unsupervised topic labelling for
-
    each NFR and per word-list. The dashed line indicates the performance of a random classifier. This graph shows how well the
-
    semi-unsupervised topic labelling matched our manual annotations.}
@@
 %\todo[inline]{this is too tiny in print}
-
   \label{fig:maxdb-unsup-results}
-
 \end{figure*}
@@
 % \begin{figure}
-
 %   \centering
-
 %  \includegraphics[width=0.45\textwidth]{figures/unsupervised-bar-f1}
-
 %  \caption{Performance, F1 values (range: 0-1), of unsupervised topic labelling for
-
 %    each NFR and per word-list.}
-
 %   \label{fig:maxdb-unsup-results-f1}
-
 % \end{figure}
@@
 Because our ground truth annotations were relevant only to ISO9126,
-
 we estimate that \textsf{exp1} had poor
-
 performance via the overlap between ISO9126 and the Kayed ontology.
-
 For \textsf{exp1} the F-measures for MaxDB were from $0$ to $0.18$ with an average
-
 of $0.03$, and for MySQL were from $0$ to $0.16$ with an average of
-
 $0.05$.
-
 \todo{update these numbers but from where? }
 % 0.177215189873 / 6
-
 % .02953586497883333333
-
 % (0.0298507462687 + 0.0661157024793 + 0.0634920634921 + 0 + 0.115384615385) / 6
-
 % .04580718793751666666
@@
 % EXP2 source is latest-reports/exp2-unsup-results.xls. Numbers reflect latest runs as of 08/2011.
-
 % (change as needed). Second column (after the name) is F1.
-
 % column format is determined by info-theory.py : accuracy	f1	fn	fp	mcc	precisions	recall	roca	tn	tnr	tp
 For \textsf{exp2}, the average F-measure (macro-F1) for MaxDB was  $0.24$ with a range $0.091$ to
-
 $0.37$, and $0.16$ for MySQL with a range of $0$ to $0.41$. PostgreSQL had an average F-measure of $0.30$ %using pgsqln
@@
 MaxDB had an average precision and recall of $0.25$ and $0.22$
-
 while MySQL had $0.41$ and $0.10$ and PostgreSQL $0.31$ and $0.29$, respectively.
@@
 % EXP3
-
 For \textsf{exp3}, the average F-measure (macro-F1) for MaxDB was $0.26$ with a range $0.11$ to
-
 $0.47$, and $0.36$ for MySQL with a range of $0.10$ to $0.65$. 
@@
 MaxDB had an average precision and recall of $0.16$ and $0.67$
-
 while MySQL had $0.3$ and $0.48$. PostgreSQL had precision and recall of $0.27$ and $0.95$, respectively.
@@
 Based on this we found that \emph{reliability} and
-
 \emph{usability} worked well for MaxDB in \textsf{exp2} and better in
-
 \textsf{exp3}. 
-
 \textsf{exp1} performed poorly.
-
 MySQL had reasonable results within \textsf{exp2} for \emph{reliability} and \emph{efficiency}. 
-
 MySQL's results for \emph{efficiency} did not improve in \textsf{exp3}
-
 but other qualities such as \emph{functionality} did improve. 
@@
 %\vspace*{-1em}
-
 \subsection{Supervised Labelling}
-
 \label{sec:suplabelling}
-
 Supervised labelling requires expert analysis of the correct
-
 class/label to assign a label to a topic. In our approach, we use the top-level NFRs in the ISO9126 standard~\cite{iso9126} for our classes, but other
-
 taxonomies are also applicable.%In order to validate how effective these word-bag approaches to topic labelling would be we created a validation data set.
@@
 We used a suite of supervised classifiers, WEKA~\cite{weka09},
-
 that includes machine learning tools such as support vector machines and Bayes-nets. 
-
 We also used the multi-labelling add-on for WEKA, Mulan~\cite{mulan}. %\footnote{\url{http://mlkd.csd.auth.gr/multilabel.html}}. 
-
 Traditional classifiers label topics with a single class, whereas Mulan allows for a mixture of classes per topic, which is what we observed while
-
 manually labelling topics.
-
 The \emph{features} we used are word counts/occurrence	per topic, if
-
 a word occurs frequently enough in a topic we consider it a feature of
-
 the topic.
@@
 To assess the performance of the supervised learners, we did a 10-fold cross-validation~\cite{Kohavi1995}, a common technique for evaluating machine
-
 learners. 
-
 The original data is partitioned randomly into ten sub-samples. Each sample is used to test against a training set composed of the nine other samples.
-
 %One subsample is used as a training set and evaluated on the other nine sub-samples. 
-
 %This is repeated nine more times, with each subsample used once as the training set. 
-
 We have reported these results below.% in Section \ref{sec:suplabelling}.
@@
 \subsubsection{Analysis of the Supervised Labelling}
-
 % \label{sec:suplabelling}
-
 Because our data-set was of word counts we expected Bayesian techniques, often used in spam filtering, to perform well. 
-
 We tried other learners that WEKA~\cite{weka09} provides: rule learners, decision tree learners, vector space learners, and support vector machines.  
-
 Figure \ref{fig:best-learn-per-tag} shows the performance of the best
-
 performing learner per label: 
-
 % We considered the best learner for a label to be 
-
 the learner that had the highest ROC value for that label. 
-
 The best learner is important because one uses a single learner per
-
 label. When applying our technique, for each NFR one should select the best learner possible.
-
 %Figure \ref{fig:best-learn-per-tag} uses the ZeroR learner as a baseline, a common practice in machine learning,
-
 %as ZeroR naively chooses the largest category all of the time. 
-
 %ZeroR occasionally outperforms other learners; for labels which are not as common, this is to be expected because any miscategorization will hurt accuracy. 
-
 %This is why we use ROC values, instead of accuracy, as they can better represent performance on labels which are not applicable to the majority of samples.
@@
 Figure \ref{fig:best-learn-per-tag} shows that MaxDB and MySQL have quite different results, as the ROC values for \emph{reliability} and \emph{functionality}
-
 seem swapped between projects.
@@
 \begin{figure}[t]
-
 \centering
-
 % \subfloat[MySQL]{
-
 \includegraphics[width=0.45\textwidth]{figures/both-supervised}
-
 % \label{fig:subfig1}
-
 % }
-
 % \subfloat[MaxDB]{
-
 % \includegraphics[width=0.4\textwidth]{figures/maxdb-supervised}
-
 % \label{fig:subfig2}
-
 % }
-
 \caption[]{ROC value for the best learner per label for MaxDB and MySQL. Values range from $0$--$1$.  Dashed line indicates the performance of a random
-
 classifier.
-
 %\todo[inline]{better discussion}
-
 }
-
 \label{fig:best-learn-per-tag}
-
 \end{figure}
@@
 For both projects Bayesian techniques did the best out of a wide variety of machine learners tested. 
-
 Our best learners, Discriminative Multinomial Naive Bayes, Naive Bayes
-
 and Multinomial Naive Bayes  are all based on Bayes's theorem and all
-
 assume, naively, that the features supplied are independent. 
-
 %The features we used are word counts per message. 
-
 One beneficial aspect of this result is that it suggests we can have very fast training and classifying  since Naive Bayes can be calculated in $O(N)$
-
 for $N$ features.
-
 %http://nlp.stanford.edu/IR-book/html/htmledition/naive-bayes-text-classification-1.html
@@
 % # F measure
-
 % mysql
-
 % > v <- c( 0.64 , 0.23  ,0.41 , 0.77  ,0.62 , 0.21 ) 
-
 % > summary(v)
-
 %    Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
-
 %   0.210   0.275   0.515   0.480   0.635   0.770 
-
 % maxdb
-
 % > v <- c(0.61, 0.25, 0.59, 0.32, 0.42, 0.17)
-
 % > summary(v)
-
 %    Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
-
 %  0.1700  0.2675  0.3700  0.3933  0.5475  0.6100
@@
 The range of F-measures for MySQL was $0.21$ to $0.77$ with an average
-
 of $0.48$, while MaxDB had a range of $0.17$ to $0.61$ with an average
-
 of $0.39$.
@@
 The less-frequently occurring a label, the harder it is to get accurate
-
 results, due to the high noise level. Nevertheless, these results are
-
 better than our previous word-list results of \textsf{exp2} and
-
 \textsf{exp3}, because the ROC values are sufficiently higher in most
-
 cases (other than MaxDB \emph{reliability} and MySQL \emph{efficiency}). The
-
 limitation of the approach we took here is that we assume labels are
-
 independent; however, labels could be correlated with each other. 
-
 The next section (\ref{sec:multilabel})
-
 addresses the issue of a lack of independence and correlation between
-
 labels
-
 using multi-label learners.
-
 %In the next section we will evaluate how well these learners perform
-
 %together.
-
 %\todo[inline]{Comment about correlation, we should link to the Mulan section, perfect segway since Mulan relies on correlation.}
@@
 \subsection{Applying Multiple Labels to Topics}
-
 \label{sec:multilabel}
@@
 As noted in Section \ref{sec:wordlist}, each topic in our data-set can be composed of zero, one, or more NFRs. 
-
 For example, a commit message might address \textit{reliability} in the context of \textit{efficiency}, or make a \textit{maintainability} improvement
-
 in the source code that related to \textit{usability}. 
-
 However, traditional machine learning techniques, such as Naive Bayes, can only map topics to a single class. 
-
 The Mulan~\cite{mulan} library encapsulates several different multi-label machine learners which can label elements with multiple labels.
-
 Mulan also includes methods for determining the performance of such techniques.
-
 % TOO DETAILED The problem framed in the learners above has changed; instead of looking at the precision and recall of applying one label, we rank
-
 % multiple labels at once. We must check if the full subset of labels was applied, and then how much of that subset was applied.
@@
 \begin{figure*}[t]
-
 \centering
-
 \subfloat[MySQL]{
-
 \includegraphics[width=0.3\textwidth]{figures/multi-label-results-mysql}
-
 \label{fig:subfig3}
-
 }
-
 \subfloat[MaxDB]{
-
 \includegraphics[width=0.3\textwidth]{figures/multi-label-results-maxdb}
-
 \label{fig:subfig4}
-
 }
-
 \caption[]{MySQL and MaxDB macro (grey) and micro (black) ROC results per multi-label learner. Possible values range from $0$--$1$.  Dashed line
-
 indicates the performance of a random classifier.
-
 %\todo[inline]{better discussion}% and graphical legend would be nice}
-
 }
-
 \label{fig:mulan}
-
 \end{figure*}
@@
 Two perspectives to evaluate multi-label learners are with micro or macro measurements (used in Figure \ref{fig:mulan}).
-
 Macro measurements are aggregated at a class or label level (per class) while micro measurements are at the element level (per element).
-
 A macro-ROC measurement is the average ROC over the ROC values for all labels, where a micro-ROC is the average ROC over all examples that were classified. 
-
 For MaxDB, the macro-ROC values are undefined because of poor performance of one of the labels.%, see Figure \ref{fig:mulan}.
@@
 Figure~\ref{fig:mulan} presents the results of Mulan's best multi-label learners for our data. 
-
 Calibrated Label Ranking (CLR) is a learner that builds two layers. The first layer determines if an entity should be labelled, while the second layer
-
 determines what labels should be assigned.
-
 The Hierarchy Of Multi-label classifiERs (HOMER) and Binary Relevance (BR) act as a hierarchy of learners: BR is flat, while HOMER tries to build a
-
 deeper hierarchy for a more accurate learner~\cite{mulan}. 
-
 These classifiers performed better than other multi-label classifiers as they have the best micro and macro ROC scores. 
-
 The multi-label and single-label learners had similar performance: for
-
 MySQL, BR and Naive Bayes had similar macro-ROC scores of $0.74$.
-
 % although their results are comparable to the naive Bayesian learners we used in Section \ref{sec:suplabelling}.
-
 % \todo[inline]{discuss this similarity in more detail}
@@
 \section{Understanding Software Maintenance Activities} 
-
 \label{sec:analysis}
-
 % \todo[inline]{talk about this topic in one section}
-
 % \todo[inline]{reference project name Nomen more}
-
 %In order to test  our ROC values produced acceptable results, we now turn to
-
 %\emph{applying} those techniques to understand software maintenance
-
 %activities. 
-
 %\todo[inline]{more on intro ... why should the reader care about
-
  % this. It needs to show how the stuff in the intro (applications) is
-
  % possible}
-
 %%% XXX NEIL I don't get this.. VVVVV
-
 %\todo[inline]{fix section headings}
-
 As we mentioned in the introduction, a key issue in software maintenance is understanding \emph{why} a system has evolved the way it has~\cite{aranda09icse}. 
-
 In this section we demonstrate the value of labelled topic extraction
-
 in addressing this issue. 
-
 % more why
-
 Labelled topics address \emph{why}  because they show reasons why changes
-
 occurred rather than \emph{how}. 
-
 The \emph{how} of a change is the change itself, the purpose or
-
 \emph{why} is what we are after.
-
 We investigate the history of the two large-scale database systems that we studied. 
-
 We use our technique to show the topic of development efforts over time in each project.
-
 %We evaluated two research questions using the ground truth data that we annotated by hand:
-
 We motivated our investigation with three research questions:
-
 \begin{enumerate}
-
 \item \emph{Do NFR frequencies change over time?} If a particular NFR
-
   was of more interest at one point in the life-cycle than another,
-
   this suggests that development activity shifted focus. For example,
-
   if a developer expected to see a recent focus on \emph{reliability},
-
   but instead \emph{usability} dominated, they might re-prioritize
-
   upcoming work items.
@@
 \item \emph{Do projects differ in their relative interest in NFRs?} A
-
   project manager, especially a systems-manager, would be interested
-
   in knowing whether a particular NFR, such as \emph{reliability}, was
-
   more important for one project than another. This could be to
-
   confirm the initial design goals, or to track the progress on that
-
   quarter's objectives.  The difference in NFR proportion is
-
   interesting because it implies a difference in focus between two projects.
@@
 \item \emph{Do different developers work on different NFRs?} For a given project, it is reasonable to think that developers are either assigned
-
 (in commercial organizations) or choose (in open-source organizations) to a particular NFR. For example, one developer might be more junior, and take
-
 responsibility for the low-impact reliability fixes. Another, more senior developer might assume responsibility for major improvements such as efficiency
-
 improvements.
@@
 %\subsection{Timeline analysis}
-
 Figures \ref{fig:mysql-timeline} and \ref{fig:maxdb-timeline} and \ref{fig:pgsql-timeline} show the
-
 temporal patterns of NFR frequencies.
-
 %NEIL- I think this (below) is just confusing, although maybe important to note.
-
 % based on the manually annotated
-
 %topics, although this visualization can be generated from the
-
 %results of labelled topic extraction.
-
 There are two measures represented. 
-
 One, the relative frequency, shown in the grey histogram boxes, represents the number of topics with that NFR in that period, 
-
 relative to the maximum number of topics assigned to that NFR. 
-
 For example, in Figure \ref{fig:mysql-timeline} we see a spike in \emph{portability} and \emph{functionality} frequency in September 2002.
-
 The second, absolute frequency, is shown using cell intensity, and compares the number of topics labelled with that NFR per period 
-
 relative to the maximum number of labelled topics overall. 
-
 For instance, Figure \ref{fig:mysql-timeline} shows that the NFRs
-
 \emph{functionality}, \emph{portability} and \emph{maintainability}
-
 contain more labelled topics, since these NFRs have been more
-
 intensely shaded; 
-
 one interesting stream is \emph{efficiency} which shows periodic
-
 activity, does this suggest that once maybe efficiency related changes
-
 have longer lasting effects?
-
 The topmost row in each diagram lists historical events for that project (such as a release). 
-
 %We refer to these occurrences in the discussion which follows.
@@
 \begin{figure}[t]
-
 \centering
-
 \subfloat[MySQL 3.23]{
-
 \includegraphics[width=\textwidth]{figures/mysql-timeline.png}
-
 \label{fig:mysql-timeline}				 
-
 }
@@
 \subfloat[MaxDB 7.500]{					
-
 \includegraphics[width=\textwidth]{figures/maxdb-timeline.png}
-
 \label{fig:maxdb-timeline}
-
 }
@@
 \subfloat[PostgreSQL 7.2]{					
-
 \includegraphics[width=\textwidth]{figures/pgsql-timeline}
-
 \label{fig:pgsql-timeline}
-
 }
-
 	\caption[]{NFR label per period. Each cell represents a 30-day period. %Numbers represent number of topics in that period labelled with that NFR. 
-
 	Grid cell intensity (saturation) is mapped to label frequency
-
 	relative to the largest label count of \emph{all} NFRs. Grey
-
 	histogram bars indicate label frequency relative to that
-
 	particular NFR's largest label count. Dashed vertical lines
-
 	relate a project milestone (\emph{*Key events*}) to our topic windows. 
-
 % \todo[inline]{better discussion}
-
 }
-
 \label{fig:timelines}
-
 % Timelines created using src/validate/summary_labels.py
-
 \end{figure}
@@
 %\noindent \textbf{Understanding the figures} -- 
-
  % boxes. not sure boxes is important..
-
 % I grepped query1.mrs2 http://softwareprocess.es/y/y/maxdb-76.tar.gz
-
 % and			http://softwareprocess.es/y/maxdb-7500.tar.gz
-
 %There are more \textsf{None} labels in MaxDB because a number of topics were to do with code cleanup or automated checking using tools like
-
 %\texttt{sutcheck}, a tool specific to the development process of MaxDB.
@@
 %\noindent \textbf{Timeline of key events} -- 
-
 We analyzed each project's developer mailing list for external validation. 
-
 %As discussed earlier, the top row in each figure shows key events for each project. 
-
 We use \textit{labelled topic extraction} to pick out the underlying NFR activity behind these events. 
-
 For example, both MaxDB and MySQL show a high number of NFRs recognized at the first period of analysis. 
-
 This is due to our window choice: we deliberately targeted our
-
 analysis to when both MySQL 3.23
-
 and MaxDB 7.500 were first announced. For MaxDB, version 7.5.00  was released in December of 2003. 
-
 We know that release 7.5.00.23 saw the development of PHP interfaces, possibly accounting for the simultaneous increase in the \emph{portability}
-
 NFR at the same time.
-
 The gap in MaxDB (Figure \ref{fig:maxdb-timeline}) is due to a shift in development focus (from February 2005 to June 2005) to MaxDB 7.6, which is
-
 released in June 2005.
@@
 %For MySQL (Figure \ref{fig:mysql-timeline}), we similarly validated our NFR patterns with external mailing list data. 
-
 The release of MySQL we study  (Figure \ref{fig:mysql-timeline}) was the first to be licenced under the GPL. 
-
 Version 3.23.31 (January, 2001) was the production release (non-beta), and we see a flurry of topics labelled with \emph{functionality} and
-
 \emph{maintainability}. 
-
 After this point, this version enters the maintenance phase of its life-cycle. 
-
 In May 2001, there is an increase in the number of topics labelled with \emph{portability}. 
-
 This might be related to release 3.23.38, which focused on Windows compatibility. 
-
 Similarly, in August, 2002, both \emph{functionality} and \emph{portability} are frequent, and mailing list data suggests this is related to the release
-
 of version 3.23.52, a general bug fix with a focus on security (a component of the \emph{functionality} NFR in the ISO9126 model). 
-
 After this point, efforts shift to the newer releases (4.0, 4.1,
-
 5.0).
@@
 By contrast, the PostgreSQL timeline  (Figure \ref{fig:pgsql-timeline}) is extracted from a central trunk and is not version-specific. Therefore development 
-
 tends to be focused on releases 7.3, 7.4 and 7.5/8.0 alpha. For example, a priority for the 8.0 candidate was a Windows-native port of the source code, which seems to 
-
 correlate with the \emph{portability} NFR increasing in frequency in mid-2004. Before the 7.4 release, \emph{usability,functionality} and \emph{portability} all 
-
 increase in frequency, possibly reflecting the interest in adding features and documentation for the release.
@@
 % \label{sec:comparison}
-
 %\XXX{Not sure there's enough here to do that!}
-
 % We observed that MySQL had more topic subsets than MaxDB. 
-
 % MySQL 3.23 also had more topics and a longer recorded history than MaxDB 7.500. 
-
 % We tagged more MySQL topics with annotations than MaxDB topics, yet both shared similarities. 
-
 % In terms of non-functional requirements both projects had long running trends that focused on \emph{functionality}, \emph{maintainability}, and
-
 %\emph{portability}, yet MaxDB had more of a focus on \emph{efficiency} and \emph{reliability}. 
-
 % \todo[inline]{expand}
-
 %\todo[inline,color=yellow]{expand on this with dates, and reference to visualizations}
-
 %\subsection{Research questions}
-
 %This analysis allows to address the questions raised in Section \ref{sec:suplabelling}:
-
 %This analysis addresses our research questions: %raised in Section \ref{sec:analysis}:
@@
 \subsection{Do NFR Frequencies Change over Time?}
-
 %\noindent\textbf{1. Do NFR frequencies change over time?} --- 
-
 %\subsubsection{Do label frequencies change over time?}
-
 %\todo{add more on PGSQL}
-
 In both MaxDB and MySQL the frequencies generally decreased with age. 
-
 However, there are variations within our NFR labels. In MySQL, \emph{usability} and \emph{efficiency} do not appear very often in topics. 
-
 A proportionately smaller number of commits addressed these NFRs.
-
 %As discussed in the previous section, 
-
 Certain peaks in topic numbers coincide with a particular emphasis from the development team on issues such as new releases or bug fixes.
-
 This suggests that maintenance activity is not necessarily strictly decreasing with time, but rather episodic and responsive to outside stimuli. 
-
 In MaxDB, we can observe that \emph{Maintainability} topics became more prevalent as MaxDB matures. 
-
 This is likely due to our analysis time-frame for MaxDB being shorter than the time-frame for the MySQL product.
@@
 In PostgreSQL, by comparison, the frequencies seem to by cyclic, since we are not studying a maintenance-phase for the product, but rather ongoing
-
 feature addition and usability improvements.
@@
 \subsection{Do Projects Differ in their Relative  Topic Interest?}
-
 Yes. MySQL 3.23 had proportionally more
-
 topics labelled \emph{functionality}, while MaxDB had proportionally more
-
 \emph{efficiency} related topics. MaxDB was a very mature release ``donated'' to the open-source community, 
-
 whereas MySQL was in its relative infancy, and	
-
 security problems were more common (security is a component of \emph{functionality} in the ISO9126 model). 
-
 In both cases \emph{portability} was a constant maintenance concern and was prevalent throughout the lifetime of the projects. It may surprise developers
-
 how often portability arises as a concern.
-
 \todo{add the pgsql data}
@@
 \subsection{Do different developers work on different NFRs?}
-
 We did a t-test to assess whether there was statistically significant differences in the types of NFRs developers chose to work on, for the PostgreSQL project only. 
-
 \todo{add more, figure}
@@
 *** We also clustered the authors by similarity and we found that
-
 among some of the most prolific postgresql authors, their NFR focus
-
 was distinctly different *** Some confounds are that postgresql
-
 authors commit the patches of others.
@@
 \section{Discussion}
-
 \label{sec:limit}
@@
 \subsection{Annotation Observations}
-
 We found many topics that were not actually non-functional requirements (NFRs) but were often related to them. 
-
 For instance, concurrency was mentioned often in the commit logs and
-
 was related to \emph{correctness} and \emph{reliability}, possibly because concurrent code is prone to bugs such as race conditions. %are 
-
 % often  was troublesome. 
-
 \emph{Configuration management} and \emph{source control} related changes appeared often; % and sometimes there were topics dedicated to configuration
-
 management. 
-
 these kinds of changes are slightly related to \emph{maintainability}. 
-
 A non-functional change that was not quality-related was licensing and
-
 copyright; many changes were simply to do with updating copyrights or
-
 ensuring copyright or license headers were applied to files. In these
-
 cases we assigned the \emph{None} label to the topic.
@@
 We noticed that occasionally the names of modules would conflict with words related to other non-functional requirements. 
-
 For instance, optimizers are very common modules in database systems: both MySQL and MaxDB have optimizer modules. 
-
 In MySQL the optimizer is mentioned but often the change addresses  correctness or another quality. 
-
 Despite this difference, the name of the module could fool our learners into believing the change was always about \emph{efficiency}. 
-
 In these cases the advantages of tailoring topic names to specific project terminologies are more clear. 
-
 Project specific word-lists would avoid automated mistakes due to the names of entities and modules of a software project.
@@
 Authors Neil and Abram each annotated the PostgreSQL topics in order
-
 to determine inter-rater reliability. While when they read each others
-
 annotations they generally agreed, the inter-rater reliability did not
-
 bear this out.
@@
 %>    A_portability   A_functionality     A_reliability A_maintainability
-
 %>      0.153548387      -0.013588517       0.004573794       0.082001031
-
 %>     A_efficiency       A_usability            A_none
-
 %>      0.231496539       0.009287926       0.062133646
@@
 % > Spearman says
-
 % >   A_portability   A_functionality     A_reliability A_maintainability
-
 % >      0.253037469      -0.014415214       0.004579664       0.082215613
-
 % >     A_efficiency       A_usability            A_none
-
 % >      0.258037216       0.014436230       0.080667468
@@
 \begin{table}
-
 \centering
-
 \begin{tabular}{l|c|c}
-
 \hline
-
             & Cohen's Kappa & Spearman Correlation \\ \hline
-
 Portability & 0.154 & 0.253  \\
-
 Functionality & -0.014 & -0.014 \\
-
 Reliability & 0.005 & 0.005 \\
-
 Maintainability & 0.082 & 0.082 \\
-
 Efficiency & 0.231 & 0.258 \\
-
 Usability & 0.009 & 0.014 \\
-
 None &      0.062 & 0.081 \\ \hline 
-
 Everything & 0.107 & 0.108 \\ \hline
-
 \end{tabular}
-
 \caption{Inter-rater Reliability on PostgreSQL}
-
 \label{tab:interr}
-
 \end{table}
@@
 Table \ref{tab:iterr} describes the Cohen Kappa and the Spearman
-
 correlation of our per topic annotations for each tag. We had to
-
 evaluate inter-rater reliability this way because 1 topic could be
-
 tagged with more than 1 annotation.
@@
 We feel these results are poor. The aggregate view of with a Kappa of
-
 $0.1$ indicates there is some agreement but not a lot. We found that
-
 there was a lot of agreement in terms of lack of a tag, but often more
-
 disagreement when a tag should be applied.
@@
 After some discussion we concluded that usability was a source of
-
 disagreement. Do user manuals improve usability? Is a command-line
-
 option a usability issue? These kinds of questions illustrate some of
-
 the agreement, disagreement and ambiguity about these labels. Further
-
 more, both annotators found that the annotation was time-consuming and
-
 difficult.
@@
 \subsection{Summary of Techniques}
-
 While an unsupervised technique such as LDA is appealing in its lack of human intervention, and thus lower effort, 
-
 supervised learners have the advantage of domain knowledge, which typically means improved results. 
-
 % NEIL - we got criticized for making it sound like this is the obvious thing to do... and then why do you need a learner?
-
 %Our manual annotations were fairly quick to do, taking only a few minutes per 30 day period. 
-
 Creating annotated topics (i.e., manual labels) for training is painstaking, but with a suitably representative set of topics, the effort is acceptable. To
-
 annotate \emph{all} topics took us approximately 20 hours per project, but we estimate only 10\% of the topics need annotation to produce useful results.
@@
 Very rarely did \textsf{exp2} and \textsf{exp3} (semi-unsupervised word matching) ever perform as well as the supervised machine learners. 
-
 For MaxDB, \textit{reliability} was slightly better detected using the static word list of \textsf{exp2}. 
-
 In general, the machine learners and \textsf{exp3} did better than \textsf{exp2} for both MaxDB and MySQL. 
-
 For both MySQL and MaxDB \textit{usability} was better served by \textsf{exp2}. 
-
 \textit{Usability} was a very infrequent label, however, which made it difficult to detect in any case.
@@
 The semi-unsupervised labelling had difficulty distinguishing between common labels and infrequent labels. 
-
 The learners would occasionally mislabel a topic deserving of an infrequent label with a more common label.
-
 The word-lists for \emph{correctness} tended to be too lengthy, non-specific and broad, especially if WordNet words were used, since the NFRs are
-
 typically loosely defined concepts in common parlance.
@@
 We found that the multi-label learners of BR, CLR and HOMER only did
-
 as well or worse for Macro-ROC as the single-label Naive Bayes and other naive Bayes-derived learners. 
-
 This suggests that by combining together multiple Naive Bayes learners
-
 we could probably label sets of topics effectively, but it would
-
 require a separate Naive Bayes learner per label.
@@
 With ROC values ranging from $0.6$ to $0.8$ we can see there is promise in these methods.
-
 \textsf{exp2} and \textsf{exp3} both indicate that static information can be used to help label topics without any training whatsoever. 
-
 MySQL and MaxDB's machine learners made some decisions based off a few shared words: \textsf{bug, code, compiler, database, HP UX, delete, memory,
-
 missing, problems, removed, add, added, changed, problem, and test}. 
-
 Adding these words to the word-lists of \textsf{exp2} and \textsf{exp3} could improve performance while ensuring they were only domain specific.
@@
 If the techniques used in \textsf{exp2} and \textsf{exp3} were combined with the supervised techniques, we could reduce the training effort by boosting
-
 training sets with topics classified with the semi-unsupervised techniques.
-
 Both Naive Bayesian learners and the word-list approaches were computationally efficient.  
-
 %These results are promising, because the accuracy of the techniques is good enough to be useful, while the run-time performance is still cheap enough
-
 %to execute to be feasible as an automated or semi-automated method of labelling topics by their software qualities.
-
 These results are promising because they indicate that these
-
 techniques are accurate enough to be useful while still maintaining
-
 acceptable run-time performance.
@@
 Since this work focuses on labelling natural language commit log
-
 comments we feel it can be adapted to other natural language artifacts
-
 such as mailing-list discussions and bug reports, even though that was
-
 not evaluated in this paper. Bug reports might not exhibit the same
-
 behaviour as commits in terms of dominant topics.
@@
 \subsection{Threats to Validity}
-
 % can always be cut down
-
 %Our work faced multiple threats to validity and we attempted to address many of them.
-
 \emph{Construct validity} -- we used only commit messages rather than
-
 mail or bug tracker messages. 
-
 To extend further we would need matching repositories for each project.
-
 %Each case study was BecauseSince we required a seamless record, it was not possible to obtain
-
 %other data sources. 
-
 Possibly they would have influenced our results, but there would be a degree of correlation between the corpora.
-
 %Conversely, construct validity was addressed in our validation as we used the files, topics and revisions to annotate the commit messages. 
-
 %We used mailing-lists to verify the purpose behind some events.
-
 Our taxonomy for software NFRs is subject to dispute, but seems
-
 generally accepted. Finally, there are exogenous sources, such as
-
 in-person discussions, which we did not access, an omission as noted
-
 in Aranda et al.~\cite{aranda09icse} %We did not show these results to the stakeholders.
@@
 \emph{Internal validity} -- %includes inter-rater reliability and our lack of rival explanation building. 
-
 We improved internal validity by trying to correlate and explain the behaviours observed in the analysis with the historical records of the projects.
-
 We did not attempt to match our results to any particular model.
-
 We assessed inter-rater reliability (IRR), a measure of the correlation between the manual annotations the first two authors created. Perfect IRR (1.0) indicates complete agreement 
-
 on the appropriate annotation for that topic. We verified IRR using
-
 manually annotated PostgreSQL topics, and achieved a result of $0.1$.
-
 Some of the reasons for this are a disagreement over the appropriate classification for documentation topics (should this be considered
-
 usability or portability?). A short education session on what a given NFR from the ISO standard encompasses would reduce this error rate.
@@
 \emph{External validity} -- %was addressed by our use of two case studies, but there are still issues. 
-
 Our data originated from OSS database projects and thus might not be applicable to commercially developed software or other domains. 
-
 Furthermore, our analysis techniques rely on a project's use of meaningful commit messages, although we feel this is the most frequently occurring form
-
 of developer comments. 
-
 %The OSS projects investigated were database systems, thus 
-
 %our results might not generalize to other domains. %We only studied database systems.
-
 % \todo[inline]{If we compare with the same domain we don't get to say it generalizes a lot?}
@@
 % in Case Study Research: Design Case Studies
-
 \emph{Reliability} -- each annotator, the first two authors, followed the same protocol and used the same annotations. 
-
 However, only two annotators were used; their annotations could be
-
 biased as we did not analyze for inter-rater reliability because they
-
 did not rate the same documents.
@@
 \subsection{Future Work}
-
 There are several avenues of further investigation.  
-
 More external validation would be useful. 
-
 Although we validated our comparisons using a mailing list for each project, interviews with developers would provide more detail. 
-
 We also think multi-label learning techniques, although in their infancy, are crucial in understanding cross-cutting concerns such as NFRs. 
-
 We want to leverage different kinds of artifacts to discover threads of NFR-related discussions that occur between multiple kinds of artifacts.
-
 Finally, we would like to extend this analysis to other domains, to see what patterns might occur in, for example, a consumer-facing software product.
@@
 \section{Conclusions}
-
 This paper presented a cross-project data mining technique,
-
 \emph{labelled topic extraction}. %The technique leveraged a
-
 %non-functional requirements (NFRs) taxonomy to mine and label latent
-
 %software development topics. 
-
 Previous topic analysis research produced project-specific
-
 topics that needed to be manually labelled.
-
 %, but topics without labels
-
 %are difficult to interpret, and manually labelling them is
-
 %costly. Furthermore, previous work was project-specific and thus incomparable. 
-
 To improve on this, we leveraged software engineering standards,
-
 specifically the ISO9126 quality taxonomy, to produce a method of partially-automated (supervised) and fully-automated (semi-unsupervised) topic labelling.
-
 Since the word-list technique is not project-specific, we used it to compare two distinct projects, where we showed our technique produced interesting
-
 insight into maintenance activity. %topics and artifacts from project histories.
-
 % Our contributions include:
-
 % \begin{itemize}
-
 % \item We provided an unsupervised method of topic labelling (word-lists);
-
 % %\item Our unsupervised method can be applied to projects without training;
-
 % \item We demonstrated a supervised method of labelling topics by a single NFRs (machine learning);
-
 % \item We demonstrated a supervised method of labelling topics by multiple NFRs (multi-label machine learning);
-
 % \item We provided a method of cross-project analysis via topic labelling;
-
 % \item We demonstrated a method of visualizing these topics across time, which we used to analyze maintenance activities.
-
 % \end{itemize}
@@
 We validated our topic labelling techniques using multiple experiments. % on unsupervised learners, supervised and multi-label learners. 
-
 We first conducted semi-unsupervised labelling using word-lists. 
-
 Our next approach was supervised, using single-label and multi-label learners. 
-
 Both kinds of learners performed well with average
-
 %area under the 
-
 ROC values between $0.6$ and $0.8$. These results, along with the efficient performance of our learners, demonstrate that labelled topic extraction is
-
 a promising approach for understanding the occurrence of non-functional requirements in software projects. 
-
 %Although it may seem like common knowledge that maintenance is a concern as a project matures, our technique was able to show this independently.
-
 %We showed that NFRs are often trending topics in commit comments
-
 %, and that NFRs are quite common as developer topics.
-
 %; and that there are efficient methods of automating topic labelling.
@@
 % We demonstrated multiple supervised and unsupervised methods of topic labelling, including multi-label learners.
-
 % We provided a case study and visualization of NFR related topics in MaxDB 7.500 and MySQL 3.23. 
-
 % During the case study we manually labelled hundreds of topics by hand and then used these annotations to validate the positive performance of topic
-
 %labelling techniques.
-
 % The implication of our paper is that topics with labels are more useful for analyzing software development activities than are unlabelled topics. To
-
 %furnish stakeholders with labelled topics, our technique provides many methods of automatic topic labelling.
@@
 %\begin{acknowledgements}
-
 %\end{acknowledgements}
@@
 % BibTeX users please use one of
-
 %\bibliographystyle{spbasic}	  % basic style, author-year citations
-
 \bibliographystyle{spmpsci}	 % mathematics and physical sciences
-
 %\bibliographystyle{spphys}	  % APS-like style for physics
-
 \bibliography{ese}   % name your BibTeX data base
@@
 \end{document}
-
 % end of file template.tex

### 2011-08-30 ese/ese.tex
updated figures for Weka results and changed wording in paper

@@
 
-% \begin{figure}
-%   \centering
-%  \includegraphics[width=0.45\textwidth]{figures/unsupervised-bar-f1}
-%  \caption{Performance, F1 values (range: 0-1), of unsupervised topic labelling for
-%    each NFR and per word-list.}
-%   \label{fig:maxdb-unsup-results-f1}
-% \end{figure}
-
 Because our ground truth annotations were relevant only to ISO9126,
@@
 \centering
-% \subfloat[MySQL]{
 \includegraphics[width=0.45\textwidth]{figures/both-supervised}
-% \label{fig:subfig1}
-% }
-% \subfloat[MaxDB]{
-% \includegraphics[width=0.4\textwidth]{figures/maxdb-supervised}
-% \label{fig:subfig2}
-% }
-\caption[]{ROC value for the best learner per label for MaxDB and MySQL. Values range from $0$--$1$.  Dashed line indicates the performance of a random
+\caption[]{ROC value for the best learner per label for MaxDB, MySQL and PgSQL. Values range from $0$--$1$.  Dashed line indicates the performance of a random
 classifier.
-%\todo[inline]{better discussion}
 }
@@
 The range of F-measures for MySQL was $0.21$ to $0.77$ with an average
-of $0.48$, while MaxDB had a range of $0.17$ to $0.61$ with an average
-of $0.39$.
+of $0.48$. MaxDB had a range of $0.17$ to $0.61$ with an average
+of $0.39$. Finally, PostgreSQL had a range of $0.04$ to $0.9$ and an average of $0.43$.
 
@@
 \textsf{exp3}, because the ROC values are sufficiently higher in most
-cases (other than MaxDB \emph{reliability} and MySQL \emph{efficiency}). The
+cases (other than MaxDB \emph{reliability}, MySQL \emph{efficiency}, and PgSQL \emph{usability}). The
 limitation of the approach we took here is that we assume labels are
@@
 \label{sec:analysis}
-% \todo[inline]{talk about this topic in one section}
-% \todo[inline]{reference project name Nomen more}
-%In order to test  our ROC values produced acceptable results, we now turn to
-%\emph{applying} those techniques to understand software maintenance
-%activities. 
-%\todo[inline]{more on intro ... why should the reader care about
- % this. It needs to show how the stuff in the intro (applications) is
- % possible}
-%%% XXX NEIL I don't get this.. VVVVV
-%\todo[inline]{fix section headings}
 As we mentioned in the introduction, a key issue in software maintenance is understanding \emph{why} a system has evolved the way it has~\cite{aranda09icse}. 
@@
 
-We now address our research questions:
+In the following sections we turn to our research questions:
 

### 2011-09-01 src/validate/CrossValidationExperiment.java
Updating to a newer version of mulan

@@ -19,6 +19,7 @@ import mulan.classifier.transformation.MultiClassLearner;
 import mulan.classifier.transformation.MultiLabelStacking;
 import mulan.data.MultiLabelInstances;
 import mulan.evaluation.Evaluation;
+import mulan.evaluation.MultipleEvaluation;
 import mulan.evaluation.Evaluator;
 import mulan.transformations.multiclass.Copy;
 import mulan.transformations.multiclass.Ignore;
@@ -39,7 +40,8 @@ public class CrossValidationExperiment {
 
 	public static void main(String[] args) {
 		String[] methodsToCompare = { "HOMER", "BR", "CLR", "MLkNN", "MC-Copy",
-				"IncludeLabels", "MC-Ignore", "RAkEL", "LP", "MLStacking" };
+				//"IncludeLabels", "MC-Ignore", "RAkEL", "LP", "MLStacking" };
+				"IncludeLabels", "MC-Ignore", "RAkEL", "LP" };
 
 		try {
 			String path = Utils.getOption("path", args);
@@ -49,7 +51,7 @@ public class CrossValidationExperiment {
 					+ filestem + ".arff", path + filestem + ".xml");
 
 			Evaluator eval = new Evaluator();
-			Evaluation results = null;
+			MultipleEvaluation results = null;
 			boolean debugOn = true;
 			int numFolds = 10;
 
@@ -136,17 +138,17 @@ public class CrossValidationExperiment {
 					results = eval.crossValidate(homer, dataSet, numFolds);
 					
 				}
-				if (methodsToCompare[i].equals("MLStacking")) {
-					J48 baseClassifier = new J48();
-					J48 metaClassifier = new J48();
-					baseClassifier.setUseLaplace(true);
-					metaClassifier.setUseLaplace(true);
-					MultiLabelStacking mls = new MultiLabelStacking(
-							baseClassifier, metaClassifier, 10);
-					mls.setDebug( debugOn );
-					mls.setPhival(0.06);
-					results = eval.crossValidate(mls, dataSet, numFolds);
-				}
+				//if (methodsToCompare[i].equals("MLStacking")) {
+				//	J48 baseClassifier = new J48();
+				//	J48 metaClassifier = new J48();
+				//	baseClassifier.setUseLaplace(true);
+				//	metaClassifier.setUseLaplace(true);
+				//	MultiLabelStacking mls = new MultiLabelStacking(
+				//			baseClassifier, metaClassifier, 10);
[…]

### 2011-09-01 src/validate/count_annotations.py

@@ -140,7 +140,7 @@ def period_file( ndir, period ):
 
 # returns the etree 
 def load_period_file( ndir, period ):
-    warnings.warn( str(period) )
+    #warnings.warn( str(period) )
     tree = lxml.etree.parse( period_file( ndir, period ) )
     return tree
 

### 2011-09-01 src/validate/run-mulan.sh

@@ -1,5 +1,7 @@
 #java -cp mulan-1.0.1.jar;weka-3.7.0.jar mulan.examples.TrainTestExperiment -path
 #java -cp mulan-1.0.1.jar:weka-36.jar mulan.examples.CrossValidationExperiment -path output/ -filestem $1 | tee output/$1.mulan.out
 #java -cp mulan-1.0.1.jar:weka-3-6-2.jar mulan.examples.CrossValidationExperiment -path output/ -filestem $1 | tee output/$1.mulan.out
-javac -cp mulan-1.0.1.jar:weka-3-7-0.jar:. CrossValidationExperiment.java 
-java -cp mulan-1.0.1.jar:weka-3-7-0.jar:. CrossValidationExperiment -path output/ -filestem $1 | tee output/$1.mulan.out
+#CP=mulan-1.0.1.jar:weka-3-7-0.jar:. 
+CP=mulan-1.3/mulan-1.3.0/mulan.jar:mulan-1.3/mulan-1.3.0/weka.jar:.
+make -f Makefile.mulan CrossValidationExperiment.class
+java -cp $CP CrossValidationExperiment -path output/ -filestem $1 | tee output/$1.mulan.out

### 2011-09-01 src/validate/topics2arff.py

@@ -29,6 +29,7 @@ def annotations_of_topic( topic ):
     annotations, thresh_el = V.create_element_list( topic )
     return annotations
 
+
 def arff_char( c ):
     if (c.isalnum()):
         return c
@@ -47,6 +48,33 @@ def arff_header( name, columns ):
            [ "@DATA" ]
     return [ x + "\n" for x in strs ]
 
+def arff_escape_words(words):    
+    word_columns = [ i_hate_python_and_unicode(arff_escape(w)) for w in words]
+    wcdict = {}
+    out_word_columns = []
+    for word in word_columns:
+       wword = word
+       if (word in wcdict):
+         wword += "__DUPLICATE__"+str(wcdict[word])
+	 wcdict[word] += 1
+	 wcdict[wword] = 1
+       else:
+         wcdict[word] = 1
+       #if ("Bj" in wword):
+       #    print("Found a Bjorklund "+word+" -- "+wword)           
+       out_word_columns.append( wword )
+    # a check
+    #for word in out_word_columns:
+    #    count = 0
+    #    for word2 in out_word_columns:
+    #        if word == word2:
+    #            count += 1
+    #    if (count > 1):
+    #        print(word + " is a duplicate with count " + str(count))
+    return out_word_columns
+
+
+
 def to_arff( name, xperiods ): #name, inputdir, periods ):
     topics = A.flatten([ A.get_topics(x) for x in xperiods ])
     # summarize the words and order by most common
@@ -63,9 +91,15 @@ def to_arff( name, xperiods ): #name, inputdir, periods ):
     # total # of columns
     ncolumns = nannos + nwords
     allcolumns = words + annos
+
+    # words to ARFF Escapes
+    out_word_columns = arff_escape_words(words)
+
+    
     # header is a list of strings
     # too much arff right here.. oh well
-    header = arff_header( name , ["W_"+arff_escape(w)+" INTEGER" for w in words] + ["A_"+arff_escape(a)+" {0,1}" for a in annos] )
+
+    header = arff_header( name , ["W_"+w+" INTEGER" for w in out_word_columns] + ["A_"+arff_escape(a)+" {0,1}" for a in annos] )
     data = []
     for t in topics:
         tannotations, twords = V.create_element_list( t )
@@ -147,7 +181,7 @@ if __name__ == '__main__':
         for line in arff:
             if (type("") != type(line)):
                 warnings.warn("A line is not a string!"+str(type(line)) + str(line))
[…]

### 2011-09-02 ese/ese.tex
revised paper for new ML learner figures
updated weka dataset for PG in graph

@@
 ISO9126 describes six high-level NFRs: maintainability, functionality,
-portability, efficiency, usability, and reliability~\footnote{While there may be lingering debate in some circles about these terms, an ISO standard
-seems like a reasonable starting point for our work.}.
+portability, efficiency, usability, and reliability.\footnote{While there may be lingering debate in some circles about these terms, an ISO standard
+seems like a reasonable starting point for our work.} 
 There is some debate about the terms in this model. 
@@
 	\centering
-\begin{tabular}{c|p{11cm}}
+\begin{tabular}{c|p{9cm}}
 \toprule
@@
 Figure \ref{fig:best-learn-per-tag} shows that MaxDB and MySQL have quite different results, as the ROC values for \emph{reliability} and \emph{functionality}
-seem swapped between projects.
-
+ swap between projects. For PostgreSQL, the performance is nearly always poorer than the other two systems. 
+\todo{expand on why?}
+
+% PG data from abram's email of 09/01/2011 using first column and pgsqln.
 \begin{figure}[t]
@@
 
-The range of F-measures for MySQL was $0.21$ to $0.77$ with an average
-of $0.48$. MaxDB had a range of $0.17$ to $0.61$ with an average
-of $0.39$. Finally, PostgreSQL had a range of $0.04$ to $0.9$ and an average of $0.43$.
+The range of F-measures for MySQL was $0.21$ to $0.77$ with a mean
+of $0.48$. MaxDB had a range of $0.17$ to $0.61$ with a mean
+\todo{NE: I believe these numbers are wrong}
+of $0.39$. Finally, PostgreSQL had a range of $0.58$ to $0.69$ and a mean of $0.64$.
 
@@
 \textsf{exp3}, because the ROC values are sufficiently higher in most
-cases (other than MaxDB \emph{reliability}, MySQL \emph{efficiency}, and PgSQL \emph{usability}). The
+cases (other than MaxDB \emph{reliability}, MySQL \emph{efficiency}, and PgSQL \emph{maintainability}). The
 limitation of the approach we took here is that we assume labels are
@@
 % multiple labels at once. We must check if the full subset of labels was applied, and then how much of that subset was applied.
+
+% numbers from mulan.tar.gz/maxdb.mulan.out.tex. I picked the two best learners for each db.
+% The numbers are the micro and macro AUC numbers.
+% MaxDB macro: BR NaN CLR NaN Homer NaN mlkNN Nan
+% MaxDB micro: BR 0.77 CLR 0.67 Homer 0.61 mlkNN 0.67
+% mysql macro: BR 0.74 CLR 0.65 mlKNN 0.61
+% mysql micro: BR 0.81 CLR 0.8 mlkNN 0.78
+% PgSQLn macro: BR 0.64 MlkNN 0.52
+% PGSQLn micro: BR 0.65 MlkNN 0.59
 
@@
 }
-\caption[]{MySQL and MaxDB macro (grey) and micro (black) ROC results per multi-label learner. Possible values range from $0$--$1$.  Dashed line
+\subfloat[PgSQL]{
+\includegraphics[width=0.3\textwidth]{figures/multi-label-results-pgsql}
+\label{fig:subfig5}
+}
+\caption[]{MySQL, PgSQL and MaxDB macro (grey) and micro (black) ROC results per multi-label learner. Possible values range from $0$--$1$.  Dashed line
 indicates the performance of a random classifier.

### 2011-09-02 ese/ese.tex
supervised numbers and plot now accurate

@@
 
+%src/validate/latex-out/best-learner-per-project-roc
 The range of F-measures for MySQL was $0.21$ to $0.77$ with a mean
 of $0.48$. MaxDB had a range of $0.17$ to $0.61$ with a mean
-\todo{NE: I believe these numbers are wrong}
-of $0.39$. Finally, PostgreSQL had a range of $0.58$ to $0.69$ and a mean of $0.64$.
+of $0.39$. Finally, PostgreSQL had a range of $0.04$ to $0.9$ and a mean of $0.43$.
 

### 2011-09-02 src/validate/output/maxdb.mulan.out.tex
Mulan parser and new run of data

@@
-\begin{tabular}{l|rrrrrrrr}
+\begin{tabular}{l|rrrrrrrrr}
 
-  & BR & CLR & HOMER & IncludeLabels & LP & MLStacking & MLkNN & RAkEL \\
+  & BR & CLR & HOMER & IncludeLabels & LP & MC-Copy & MC-Ignore & MLkNN & RAkEL \\
 
@@
 
-example-Accuracy & 0.49 & 0.47 & 0.51 & 0.44 & 0.42 & 0.47 & 0.43 & 0.48 \\
+Average Precision & 0.53 & 0.59 & 0.52 & 0.56 & 0.47 & 0.52 & 0.55 & 0.58 & 0.54 \\
 
-example-Fmeasure & 0.55 & 0.50 & 0.53 & 0.45 & 0.44 & 0.48 & 0.43 & 0.50 \\
+Coverage & 1.41 & 1.19 & 1.44 & 1.33 & 1.82 & 1.52 & 1.54 & 1.21 & 1.49 \\
 
-example-HammingLoss & 0.19 & 0.14 & 0.14 & 0.14 & 0.17 & 0.13 & 0.13 & 0.13 \\
+ErrorSetSize & 1.42 & 1.13 & 1.38 & 1.29 & 1.87 & 1.49 & 1.54 & 1.14 & 1.43 \\
 
-example-Precision & 0.50 & 0.49 & 0.53 & 0.45 & 0.45 & 0.49 & 0.43 & 0.51 \\
+Example-Based Accuracy & NaN & NaN & NaN & NaN & NaN &  &  & NaN & NaN \\
 
-example-Recall & 0.61 & 0.50 & 0.53 & 0.45 & 0.43 & 0.47 & 0.43 & 0.50 \\
+Example-Based F Measure & NaN & NaN & NaN & NaN & NaN &  &  & NaN & NaN \\
 
-example-SubsetAccuracy & 0.40 & 0.43 & 0.45 & 0.43 & 0.38 & 0.43 & 0.43 & 0.43 \\
+Example-Based Precision & NaN & NaN & NaN & NaN & NaN &  &  & NaN & NaN \\
 
-label-macro-AUC & NaN & NaN & NaN & NaN & NaN & NaN & NaN & NaN \\
+Example-Based Recall & NaN & NaN & NaN & NaN & NaN &  &  & NaN & NaN \\
 
-label-macro-F1 & 0.32 & 0.19 & 0.23 & 0.11 & 0.19 & 0.12 & 0.00 & 0.26 \\
+Hamming Loss & 0.19 & 0.14 & 0.13 & 0.14 & 0.17 &  &  & 0.13 & 0.14 \\
 
-label-macro-Precision & 0.29 & 0.31 & 0.32 & 0.21 & 0.23 & 0.18 & 0.00 & 0.40 \\
+IsError & 0.44 & 0.41 & 0.45 & 0.42 & 0.46 & 0.44 & 0.41 & 0.42 & 0.42 \\
 
-label-macro-Recall & 0.38 & 0.15 & 0.20 & 0.08 & 0.17 & 0.10 & 0.00 & 0.23 \\
+Macro-averaged AUC & NaN & NaN & NaN & NaN & NaN &  &  & NaN & NaN \\
 
-label-micro-AUC & 0.76 & 0.66 & 0.62 & 0.55 & 0.62 & 0.68 & 0.65 & 0.68 \\
+Macro-averaged F-Measure & NaN & NaN & NaN & NaN & NaN &  &  & NaN & NaN \\
 
-label-micro-F1 & 0.39 & 0.27 & 0.34 & 0.13 & 0.25 & 0.21 & 0.00 & 0.34 \\
+Macro-averaged Precision & NaN & NaN & NaN & NaN & NaN &  &  & NaN & NaN \\
 
-label-micro-Precision & 0.34 & 0.42 & 0.49 & 0.33 & 0.30 & 0.49 & 0.00 & 0.49 \\
+Macro-averaged Recall & NaN & NaN & NaN & NaN & NaN &  &  & NaN & NaN \\
 
-label-micro-Recall & 0.47 & 0.21 & 0.26 & 0.09 & 0.22 & 0.14 & 0.00 & 0.27 \\
+Mean Average Precision & NaN & NaN & NaN & NaN & NaN &  &  & NaN & NaN \\
 
-rank-AvgPrecision & 0.54 & 0.57 & 0.54 & 0.40 & 0.40 & 0.60 & 0.58 & 0.51 \\
+Micro-averaged AUC & 0.77 & 0.67 & 0.61 & 0.77 & 0.74 &  &  & 0.69 & 0.72 \\
 
-rank-Coverage & 1.40 & 1.23 & 1.43 & 1.84 & 1.93 & 1.16 & 1.20 & 1.54 \\
+Micro-averaged F-Measure & 0.39 & 0.36 & 0.37 & 0.34 & 0.33 &  &  & NaN & 0.34 \\
 
-rank-One-error & 0.81 & 0.79 & 0.81 & 0.90 & 0.89 & 0.76 & 0.79 & 0.79 \\
+Micro-averaged Precision & 0.34 & 0.44 & 0.52 & 0.44 & 0.35 &  &  & NaN & 0.44 \\
 
-rank-Ranking Loss & 0.41 & 0.35 & 0.41 & 0.58 & 0.60 & 0.32 & 0.33 & 0.46 \\
+Micro-averaged Recall & 0.47 & 0.31 & 0.30 & 0.29 & 0.32 &  &  & 0.00 & 0.28 \\
+
+OneError & 0.83 & 0.78 & 0.83 & 0.79 & 0.83 & 0.81 & 0.78 & 0.79 & 0.79 \\
+
+Ranking Loss & 0.23 & 0.19 & 0.24 & 0.22 & 0.31 &  &  & 0.19 & 0.24 \\
+
+Subset Accuracy & 0.41 & 0.48 & 0.50 & 0.50 & 0.53 &  &  & 0.43 & 0.51 \\
 

### 2011-09-02 src/validate/output/maxdb.mulan.out.transpose.tex
Mulan parser and new run of data

@@
-\begin{tabular}{l|rrrrrrrrrrrrrrrrrr}
+\begin{tabular}{l|rrrrrrrrrrrrrrrrrrrrr}
 
-  & example-Accuracy & example-Fmeasure & example-HammingLoss & example-Precision & example-Recall & example-SubsetAccuracy & label-macro-AUC & label-macro-F1 & label-macro-Precision & label-macro-Recall & label-micro-AUC & label-micro-F1 & label-micro-Precision & label-micro-Recall & rank-AvgPrecision & rank-Coverage & rank-One-error & rank-Ranking Loss \\
+  & Average Precision & Coverage & ErrorSetSize & Example-Based Accuracy & Example-Based F Measure & Example-Based Precision & Example-Based Recall & Hamming Loss & IsError & Macro-averaged AUC & Macro-averaged F-Measure & Macro-averaged Precision & Macro-averaged Recall & Mean Average Precision & Micro-averaged AUC & Micro-averaged F-Measure & Micro-averaged Precision & Micro-averaged Recall & OneError & Ranking Loss & Subset Accuracy \\
 
@@
 
-BR & 0.49 & 0.55 & 0.19 & 0.50 & 0.61 & 0.40 & NaN & 0.32 & 0.29 & 0.38 & 0.76 & 0.39 & 0.34 & 0.47 & 0.54 & 1.40 & 0.81 & 0.41 \\
+BR & 0.53 & 1.41 & 1.42 & NaN & NaN & NaN & NaN & 0.19 & 0.44 & NaN & NaN & NaN & NaN & NaN & 0.77 & 0.39 & 0.34 & 0.47 & 0.83 & 0.23 & 0.41 \\
 
-CLR & 0.47 & 0.50 & 0.14 & 0.49 & 0.50 & 0.43 & NaN & 0.19 & 0.31 & 0.15 & 0.66 & 0.27 & 0.42 & 0.21 & 0.57 & 1.23 & 0.79 & 0.35 \\
+CLR & 0.59 & 1.19 & 1.13 & NaN & NaN & NaN & NaN & 0.14 & 0.41 & NaN & NaN & NaN & NaN & NaN & 0.67 & 0.36 & 0.44 & 0.31 & 0.78 & 0.19 & 0.48 \\
 
-HOMER & 0.51 & 0.53 & 0.14 & 0.53 & 0.53 & 0.45 & NaN & 0.23 & 0.32 & 0.20 & 0.62 & 0.34 & 0.49 & 0.26 & 0.54 & 1.43 & 0.81 & 0.41 \\
+HOMER & 0.52 & 1.44 & 1.38 & NaN & NaN & NaN & NaN & 0.13 & 0.45 & NaN & NaN & NaN & NaN & NaN & 0.61 & 0.37 & 0.52 & 0.30 & 0.83 & 0.24 & 0.50 \\
 
-IncludeLabels & 0.44 & 0.45 & 0.14 & 0.45 & 0.45 & 0.43 & NaN & 0.11 & 0.21 & 0.08 & 0.55 & 0.13 & 0.33 & 0.09 & 0.40 & 1.84 & 0.90 & 0.58 \\
+IncludeLabels & 0.56 & 1.33 & 1.29 & NaN & NaN & NaN & NaN & 0.14 & 0.42 & NaN & NaN & NaN & NaN & NaN & 0.77 & 0.34 & 0.44 & 0.29 & 0.79 & 0.22 & 0.50 \\
 
-LP & 0.42 & 0.44 & 0.17 & 0.45 & 0.43 & 0.38 & NaN & 0.19 & 0.23 & 0.17 & 0.62 & 0.25 & 0.30 & 0.22 & 0.40 & 1.93 & 0.89 & 0.60 \\
+LP & 0.47 & 1.82 & 1.87 & NaN & NaN & NaN & NaN & 0.17 & 0.46 & NaN & NaN & NaN & NaN & NaN & 0.74 & 0.33 & 0.35 & 0.32 & 0.83 & 0.31 & 0.53 \\
 
-MLStacking & 0.47 & 0.48 & 0.13 & 0.49 & 0.47 & 0.43 & NaN & 0.12 & 0.18 & 0.10 & 0.68 & 0.21 & 0.49 & 0.14 & 0.60 & 1.16 & 0.76 & 0.32 \\
+MC-Copy & 0.52 & 1.52 & 1.49 &  &  &  &  &  & 0.44 &  &  &  &  &  &  &  &  &  & 0.81 &  &  \\
 
-MLkNN & 0.43 & 0.43 & 0.13 & 0.43 & 0.43 & 0.43 & NaN & 0.00 & 0.00 & 0.00 & 0.65 & 0.00 & 0.00 & 0.00 & 0.58 & 1.20 & 0.79 & 0.33 \\
+MC-Ignore & 0.55 & 1.54 & 1.54 &  &  &  &  &  & 0.41 &  &  &  &  &  &  &  &  &  & 0.78 &  &  \\
 
-RAkEL & 0.48 & 0.50 & 0.13 & 0.51 & 0.50 & 0.43 & NaN & 0.26 & 0.40 & 0.23 & 0.68 & 0.34 & 0.49 & 0.27 & 0.51 & 1.54 & 0.79 & 0.46 \\
+MLkNN & 0.58 & 1.21 & 1.14 & NaN & NaN & NaN & NaN & 0.13 & 0.42 & NaN & NaN & NaN & NaN & NaN & 0.69 & NaN & NaN & 0.00 & 0.79 & 0.19 & 0.43 \\
+
+RAkEL & 0.54 & 1.49 & 1.43 & NaN & NaN & NaN & NaN & 0.14 & 0.42 & NaN & NaN & NaN & NaN & NaN & 0.72 & 0.34 & 0.44 & 0.28 & 0.79 & 0.24 & 0.51 \\
 

### 2011-09-02 src/validate/output/mysql.mulan.out.tex
Updating to a newer version of mulan
Mulan parser and new run of data

@@
 
-example-Accuracy & 0.45 & 0.40 & 0.50 & 0.40 & 0.40 & 0.38 & 0.33 & 0.48 \\
+Average Precision & 0.76 & 0.76 & 0.69 & 0.73 & 0.53 & 0.72 & 0.68 & 0.75 & 0.74 \\
 
-example-Fmeasure & 0.56 & 0.48 & 0.59 & 0.47 & 0.49 & 0.46 & 0.40 & 0.55 \\
+Coverage & 1.67 & 1.64 & 2.24 & 1.87 & 3.14 & 1.88 & 2.35 & 1.69 & 2.01 \\
 
-example-HammingLoss & 0.24 & 0.19 & 0.18 & 0.19 & 0.23 & 0.19 & 0.20 & 0.17 \\
+ErrorSetSize & 1.23 & 1.18 & 1.90 & 1.48 & 3.49 & 1.49 & 2.10 & 1.25 & 1.59 \\
 
-example-Precision & 0.50 & 0.51 & 0.60 & 0.52 & 0.52 & 0.50 & 0.44 & 0.60 \\
+Example-Based Accuracy & NaN & NaN & NaN & NaN & NaN &  &  & NaN & NaN \\
 
-example-Recall & 0.65 & 0.46 & 0.58 & 0.44 & 0.47 & 0.43 & 0.37 & 0.52 \\
+Example-Based F Measure & NaN & NaN & NaN & NaN & NaN &  &  & NaN & NaN \\
 
-example-SubsetAccuracy & 0.19 & 0.24 & 0.31 & 0.24 & 0.23 & 0.23 & 0.20 & 0.31 \\
+Example-Based Precision & NaN & NaN & NaN & NaN & NaN &  &  & NaN & NaN \\
 
-label-macro-AUC & 0.74 & 0.66 & 0.64 & 0.60 & 0.56 & 0.58 & 0.59 & 0.67 \\
+Example-Based Recall & NaN & NaN & NaN & NaN & NaN &  &  & NaN & NaN \\
 
-label-macro-F1 & 0.46 & 0.30 & 0.43 & 0.28 & 0.32 & 0.27 & 0.22 & 0.39 \\
+Hamming Loss & 0.25 & 0.21 & 0.20 & 0.21 & 0.25 &  &  & 0.22 & 0.20 \\
 
-label-macro-Precision & 0.41 & 0.41 & 0.49 & 0.39 & 0.37 & 0.34 & 0.34 & 0.52 \\
+IsError & 0.55 & 0.52 & 0.62 & 0.57 & 0.74 & 0.57 & 0.64 & 0.54 & 0.55 \\
 
-label-macro-Recall & 0.56 & 0.26 & 0.40 & 0.24 & 0.31 & 0.23 & 0.19 & 0.33 \\
+Macro-averaged AUC & 0.74 & 0.65 & 0.64 & 0.61 & 0.59 &  &  & 0.62 & 0.68 \\
 
-label-micro-AUC & 0.81 & 0.81 & 0.77 & 0.77 & 0.66 & 0.80 & 0.80 & 0.79 \\
+Macro-averaged F-Measure & NaN & NaN & NaN & NaN & NaN &  &  & NaN & NaN \\
 
-label-micro-F1 & 0.59 & 0.53 & 0.61 & 0.51 & 0.50 & 0.51 & 0.45 & 0.59 \\
+Macro-averaged Precision & 0.42 & NaN & NaN & NaN & NaN &  &  & NaN & NaN \\
 
-label-micro-Precision & 0.51 & 0.69 & 0.66 & 0.71 & 0.53 & 0.70 & 0.69 & 0.73 \\
+Macro-averaged Recall & 0.57 & 0.30 & 0.45 & 0.27 & 0.33 &  &  & 0.22 & 0.34 \\
 
-label-micro-Recall & 0.68 & 0.43 & 0.56 & 0.41 & 0.48 & 0.41 & 0.34 & 0.49 \\
+Mean Average Precision & 0.53 & 0.44 & 0.45 & 0.42 & 0.39 &  &  & 0.41 & 0.50 \\
 
-rank-AvgPrecision & 0.76 & 0.77 & 0.69 & 0.72 & 0.53 & 0.74 & 0.76 & 0.73 \\
+Micro-averaged AUC & 0.81 & 0.80 & 0.74 & 0.77 & 0.66 &  &  & 0.79 & 0.80 \\
 
-rank-Coverage & 1.51 & 1.47 & 1.94 & 1.72 & 2.99 & 1.57 & 1.54 & 2.00 \\
+Micro-averaged F-Measure & 0.59 & 0.54 & 0.61 & 0.52 & 0.50 &  &  & 0.48 & 0.56 \\
 
-rank-One-error & 0.40 & 0.39 & 0.47 & 0.44 & 0.63 & 0.43 & 0.40 & 0.36 \\
+Micro-averaged Precision & 0.52 & 0.65 & 0.63 & 0.67 & 0.53 &  &  & 0.67 & 0.66 \\
 
-rank-Ranking Loss & 0.19 & 0.18 & 0.27 & 0.22 & 0.52 & 0.20 & 0.19 & 0.28 \\
+Micro-averaged Recall & 0.69 & 0.47 & 0.59 & 0.43 & 0.46 &  &  & 0.38 & 0.49 \\
+
+OneError & 0.39 & 0.38 & 0.45 & 0.41 & 0.62 & 0.42 & 0.44 & 0.39 & 0.37 \\
+
+Ranking Loss & 0.18 & 0.18 & 0.28 & 0.22 & 0.50 &  &  & 0.19 & 0.24 \\
+
+Subset Accuracy & 0.18 & 0.22 & 0.30 & 0.23 & 0.25 &  &  & 0.22 & 0.29 \\
 

### 2011-09-02 src/validate/output/mysql.mulan.out.transpose.tex
Mulan parser and new run of data

@@
 
-BR & 0.45 & 0.56 & 0.24 & 0.50 & 0.65 & 0.19 & 0.74 & 0.46 & 0.41 & 0.56 & 0.81 & 0.59 & 0.51 & 0.68 & 0.76 & 1.51 & 0.40 & 0.19 \\
+BR & 0.76 & 1.67 & 1.23 & NaN & NaN & NaN & NaN & 0.25 & 0.55 & 0.74 & NaN & 0.42 & 0.57 & 0.53 & 0.81 & 0.59 & 0.52 & 0.69 & 0.39 & 0.18 & 0.18 \\
 
-CLR & 0.40 & 0.48 & 0.19 & 0.51 & 0.46 & 0.24 & 0.66 & 0.30 & 0.41 & 0.26 & 0.81 & 0.53 & 0.69 & 0.43 & 0.77 & 1.47 & 0.39 & 0.18 \\
+CLR & 0.76 & 1.64 & 1.18 & NaN & NaN & NaN & NaN & 0.21 & 0.52 & 0.65 & NaN & NaN & 0.30 & 0.44 & 0.80 & 0.54 & 0.65 & 0.47 & 0.38 & 0.18 & 0.22 \\
 
-HOMER & 0.50 & 0.59 & 0.18 & 0.60 & 0.58 & 0.31 & 0.64 & 0.43 & 0.49 & 0.40 & 0.77 & 0.61 & 0.66 & 0.56 & 0.69 & 1.94 & 0.47 & 0.27 \\
+HOMER & 0.69 & 2.24 & 1.90 & NaN & NaN & NaN & NaN & 0.20 & 0.62 & 0.64 & NaN & NaN & 0.45 & 0.45 & 0.74 & 0.61 & 0.63 & 0.59 & 0.45 & 0.28 & 0.30 \\
 
-IncludeLabels & 0.40 & 0.47 & 0.19 & 0.52 & 0.44 & 0.24 & 0.60 & 0.28 & 0.39 & 0.24 & 0.77 & 0.51 & 0.71 & 0.41 & 0.72 & 1.72 & 0.44 & 0.22 \\
+IncludeLabels & 0.73 & 1.87 & 1.48 & NaN & NaN & NaN & NaN & 0.21 & 0.57 & 0.61 & NaN & NaN & 0.27 & 0.42 & 0.77 & 0.52 & 0.67 & 0.43 & 0.41 & 0.22 & 0.23 \\
 
-LP & 0.40 & 0.49 & 0.23 & 0.52 & 0.47 & 0.23 & 0.56 & 0.32 & 0.37 & 0.31 & 0.66 & 0.50 & 0.53 & 0.48 & 0.53 & 2.99 & 0.63 & 0.52 \\
+LP & 0.53 & 3.14 & 3.49 & NaN & NaN & NaN & NaN & 0.25 & 0.74 & 0.59 & NaN & NaN & 0.33 & 0.39 & 0.66 & 0.50 & 0.53 & 0.46 & 0.62 & 0.50 & 0.25 \\
 
-MLStacking & 0.38 & 0.46 & 0.19 & 0.50 & 0.43 & 0.23 & 0.58 & 0.27 & 0.34 & 0.23 & 0.80 & 0.51 & 0.70 & 0.41 & 0.74 & 1.57 & 0.43 & 0.20 \\
+MC-Copy & 0.72 & 1.88 & 1.49 &  &  &  &  &  & 0.57 &  &  &  &  &  &  &  &  &  & 0.42 &  &  \\
 
-MLkNN & 0.33 & 0.40 & 0.20 & 0.44 & 0.37 & 0.20 & 0.59 & 0.22 & 0.34 & 0.19 & 0.80 & 0.45 & 0.69 & 0.34 & 0.76 & 1.54 & 0.40 & 0.19 \\
+MC-Ignore & 0.68 & 2.35 & 2.10 &  &  &  &  &  & 0.64 &  &  &  &  &  &  &  &  &  & 0.44 &  &  \\
 
-RAkEL & 0.48 & 0.55 & 0.17 & 0.60 & 0.52 & 0.31 & 0.67 & 0.39 & 0.52 & 0.33 & 0.79 & 0.59 & 0.73 & 0.49 & 0.73 & 2.00 & 0.36 & 0.28 \\
+MLkNN & 0.75 & 1.69 & 1.25 & NaN & NaN & NaN & NaN & 0.22 & 0.54 & 0.62 & NaN & NaN & 0.22 & 0.41 & 0.79 & 0.48 & 0.67 & 0.38 & 0.39 & 0.19 & 0.22 \\
+
+RAkEL & 0.74 & 2.01 & 1.59 & NaN & NaN & NaN & NaN & 0.20 & 0.55 & 0.68 & NaN & NaN & 0.34 & 0.50 & 0.80 & 0.56 & 0.66 & 0.49 & 0.37 & 0.24 & 0.29 \\
 

### 2011-09-02 src/validate/output/pgsqla.mulan.out.tex
Mulan parser and new run of data

\begin{tabular}{l|rrrrrrrrr}

  & BR & CLR & HOMER & IncludeLabels & LP & MC-Copy & MC-Ignore & MLkNN & RAkEL \\

\hline \\

Average Precision & 0.78 & 0.78 & 0.74 & 0.69 & 0.59 & 0.83 & 0.68 & 0.85 & 0.80 \\

Coverage & 2.06 & 2.01 & 2.35 & 2.66 & 3.30 & 1.83 & 3.00 & 1.74 & 2.15 \\

ErrorSetSize & 1.48 & 1.40 & 1.81 & 2.29 & 3.49 & 1.16 & 2.72 & 1.02 & 1.52 \\

Example-Based Accuracy & 0.51 & NaN & 0.58 & NaN & NaN &  &  & 0.56 & NaN \\

Example-Based F Measure & NaN & NaN & 0.75 & NaN & NaN &  &  & 0.70 & 0.70 \\

Example-Based Precision & NaN & NaN & 0.71 & NaN & NaN &  &  & NaN & NaN \\

Example-Based Recall & NaN & NaN & NaN & NaN & NaN &  &  & NaN & NaN \\

Hamming Loss & 0.25 & 0.25 & 0.20 & 0.26 & 0.25 &  &  & 0.20 & 0.21 \\

IsError & 0.58 & 0.60 & 0.62 & 0.69 & 0.76 & 0.47 & 0.67 & 0.45 & 0.53 \\

Macro-averaged AUC & 0.63 & 0.56 & 0.54 & 0.51 & 0.54 &  &  & 0.54 & 0.53 \\

Macro-averaged F-Measure & NaN & NaN & NaN & NaN & NaN &  &  & NaN & NaN \\

Macro-averaged Precision & NaN & NaN & NaN & NaN & NaN &  &  & NaN & NaN \\

Macro-averaged Recall & 0.46 & 0.37 & 0.38 & 0.32 & 0.37 &  &  & 0.29 & 0.33 \\

Mean Average Precision & 0.45 & 0.40 & 0.39 & 0.39 & 0.39 &  &  & 0.38 & 0.39 \\

Micro-averaged AUC & 0.81 & 0.81 & 0.79 & 0.70 & 0.73 &  &  & 0.84 & 0.80 \\

Micro-averaged F-Measure & 0.64 & 0.62 & 0.69 & 0.58 & 0.60 &  &  & 0.67 & 0.66 \\

Micro-averaged Precision & 0.60 & 0.60 & 0.68 & 0.60 & 0.60 &  &  & 0.72 & 0.67 \\

Micro-averaged Recall & 0.69 & 0.64 & 0.70 & 0.57 & 0.61 &  &  & 0.63 & 0.64 \\

OneError & 0.31 & 0.28 & 0.38 & 0.42 & 0.57 & 0.21 & 0.36 & 0.19 & 0.25 \\

Ranking Loss & 0.20 & 0.19 & 0.24 & 0.31 & 0.46 &  &  & 0.14 & 0.20 \\

Subset Accuracy & 0.19 & 0.17 & 0.27 & 0.18 & 0.20 &  &  & 0.28 & 0.26 \\

\end{tabular}

### 2011-09-02 src/validate/output/pgsqla.mulan.out.transpose.tex
Mulan parser and new run of data

\begin{tabular}{l|rrrrrrrrrrrrrrrrrrrrr}

  & Average Precision & Coverage & ErrorSetSize & Example-Based Accuracy & Example-Based F Measure & Example-Based Precision & Example-Based Recall & Hamming Loss & IsError & Macro-averaged AUC & Macro-averaged F-Measure & Macro-averaged Precision & Macro-averaged Recall & Mean Average Precision & Micro-averaged AUC & Micro-averaged F-Measure & Micro-averaged Precision & Micro-averaged Recall & OneError & Ranking Loss & Subset Accuracy \\

\hline \\

BR & 0.78 & 2.06 & 1.48 & 0.51 & NaN & NaN & NaN & 0.25 & 0.58 & 0.63 & NaN & NaN & 0.46 & 0.45 & 0.81 & 0.64 & 0.60 & 0.69 & 0.31 & 0.20 & 0.19 \\

CLR & 0.78 & 2.01 & 1.40 & NaN & NaN & NaN & NaN & 0.25 & 0.60 & 0.56 & NaN & NaN & 0.37 & 0.40 & 0.81 & 0.62 & 0.60 & 0.64 & 0.28 & 0.19 & 0.17 \\

HOMER & 0.74 & 2.35 & 1.81 & 0.58 & 0.75 & 0.71 & NaN & 0.20 & 0.62 & 0.54 & NaN & NaN & 0.38 & 0.39 & 0.79 & 0.69 & 0.68 & 0.70 & 0.38 & 0.24 & 0.27 \\

IncludeLabels & 0.69 & 2.66 & 2.29 & NaN & NaN & NaN & NaN & 0.26 & 0.69 & 0.51 & NaN & NaN & 0.32 & 0.39 & 0.70 & 0.58 & 0.60 & 0.57 & 0.42 & 0.31 & 0.18 \\

LP & 0.59 & 3.30 & 3.49 & NaN & NaN & NaN & NaN & 0.25 & 0.76 & 0.54 & NaN & NaN & 0.37 & 0.39 & 0.73 & 0.60 & 0.60 & 0.61 & 0.57 & 0.46 & 0.20 \\

MC-Copy & 0.83 & 1.83 & 1.16 &  &  &  &  &  & 0.47 &  &  &  &  &  &  &  &  &  & 0.21 &  &  \\

MC-Ignore & 0.68 & 3.00 & 2.72 &  &  &  &  &  & 0.67 &  &  &  &  &  &  &  &  &  & 0.36 &  &  \\

MLkNN & 0.85 & 1.74 & 1.02 & 0.56 & 0.70 & NaN & NaN & 0.20 & 0.45 & 0.54 & NaN & NaN & 0.29 & 0.38 & 0.84 & 0.67 & 0.72 & 0.63 & 0.19 & 0.14 & 0.28 \\

RAkEL & 0.80 & 2.15 & 1.52 & NaN & 0.70 & NaN & NaN & 0.21 & 0.53 & 0.53 & NaN & NaN & 0.33 & 0.39 & 0.80 & 0.66 & 0.67 & 0.64 & 0.25 & 0.20 & 0.26 \\

\end{tabular}

### 2011-09-02 src/validate/output/pgsqln.mulan.out.tex
Mulan parser and new run of data

\begin{tabular}{l|rrrrrrrrr}

  & BR & CLR & HOMER & IncludeLabels & LP & MC-Copy & MC-Ignore & MLkNN & RAkEL \\

\hline \\

Average Precision & 0.60 & 0.56 & 0.56 & 0.53 & 0.47 & 0.53 & 0.52 & 0.57 & 0.55 \\

Coverage & 2.49 & 2.67 & 2.78 & 2.94 & 3.26 & 2.99 & 3.02 & 2.71 & 2.86 \\

ErrorSetSize & 2.41 & 2.67 & 2.74 & 3.01 & 3.57 & 3.07 & 3.18 & 2.68 & 2.86 \\

Example-Based Accuracy & NaN & NaN & NaN & NaN & NaN &  &  & NaN & NaN \\

Example-Based F Measure & NaN & NaN & NaN & NaN & NaN &  &  & NaN & NaN \\

Example-Based Precision & NaN & NaN & NaN & NaN & NaN &  &  & NaN & NaN \\

Example-Based Recall & NaN & NaN & NaN & NaN & NaN &  &  & NaN & NaN \\

Hamming Loss & 0.34 & 0.33 & 0.32 & 0.34 & 0.36 &  &  & 0.27 & 0.30 \\

IsError & 0.74 & 0.77 & 0.80 & 0.82 & 0.87 & 0.80 & 0.81 & 0.77 & 0.81 \\

Macro-averaged AUC & 0.64 & 0.58 & 0.58 & 0.55 & 0.53 &  &  & 0.53 & 0.56 \\

Macro-averaged F-Measure & 0.41 & NaN & 0.37 & NaN & 0.28 &  &  & NaN & NaN \\

Macro-averaged Precision & 0.38 & 0.37 & 0.39 & 0.33 & 0.30 &  &  & NaN & 0.30 \\

Macro-averaged Recall & 0.48 & 0.34 & 0.39 & 0.31 & 0.29 &  &  & 0.00 & 0.22 \\

Mean Average Precision & 0.43 & 0.38 & 0.39 & 0.35 & 0.33 &  &  & 0.34 & 0.36 \\

Micro-averaged AUC & 0.65 & 0.60 & 0.60 & 0.56 & 0.55 &  &  & 0.59 & 0.58 \\

Micro-averaged F-Measure & 0.43 & 0.36 & 0.41 & 0.34 & 0.31 &  &  & NaN & 0.30 \\

Micro-averaged Precision & 0.39 & 0.38 & 0.40 & 0.35 & 0.31 &  &  & NaN & 0.39 \\

Micro-averaged Recall & 0.49 & 0.35 & 0.42 & 0.33 & 0.30 &  &  & 0.01 & 0.24 \\

OneError & 0.59 & 0.63 & 0.64 & 0.68 & 0.75 & 0.66 & 0.66 & 0.62 & 0.64 \\

Ranking Loss & 0.35 & 0.39 & 0.40 & 0.44 & 0.52 &  &  & 0.40 & 0.42 \\

Subset Accuracy & 0.08 & 0.10 & 0.10 & 0.09 & 0.11 &  &  & 0.04 & 0.10 \\

\end{tabular}

### 2011-09-02 src/validate/output/pgsqln.mulan.out.transpose.tex
Mulan parser and new run of data

\begin{tabular}{l|rrrrrrrrrrrrrrrrrrrrr}

  & Average Precision & Coverage & ErrorSetSize & Example-Based Accuracy & Example-Based F Measure & Example-Based Precision & Example-Based Recall & Hamming Loss & IsError & Macro-averaged AUC & Macro-averaged F-Measure & Macro-averaged Precision & Macro-averaged Recall & Mean Average Precision & Micro-averaged AUC & Micro-averaged F-Measure & Micro-averaged Precision & Micro-averaged Recall & OneError & Ranking Loss & Subset Accuracy \\

\hline \\

BR & 0.60 & 2.49 & 2.41 & NaN & NaN & NaN & NaN & 0.34 & 0.74 & 0.64 & 0.41 & 0.38 & 0.48 & 0.43 & 0.65 & 0.43 & 0.39 & 0.49 & 0.59 & 0.35 & 0.08 \\

CLR & 0.56 & 2.67 & 2.67 & NaN & NaN & NaN & NaN & 0.33 & 0.77 & 0.58 & NaN & 0.37 & 0.34 & 0.38 & 0.60 & 0.36 & 0.38 & 0.35 & 0.63 & 0.39 & 0.10 \\

HOMER & 0.56 & 2.78 & 2.74 & NaN & NaN & NaN & NaN & 0.32 & 0.80 & 0.58 & 0.37 & 0.39 & 0.39 & 0.39 & 0.60 & 0.41 & 0.40 & 0.42 & 0.64 & 0.40 & 0.10 \\

IncludeLabels & 0.53 & 2.94 & 3.01 & NaN & NaN & NaN & NaN & 0.34 & 0.82 & 0.55 & NaN & 0.33 & 0.31 & 0.35 & 0.56 & 0.34 & 0.35 & 0.33 & 0.68 & 0.44 & 0.09 \\

LP & 0.47 & 3.26 & 3.57 & NaN & NaN & NaN & NaN & 0.36 & 0.87 & 0.53 & 0.28 & 0.30 & 0.29 & 0.33 & 0.55 & 0.31 & 0.31 & 0.30 & 0.75 & 0.52 & 0.11 \\

MC-Copy & 0.53 & 2.99 & 3.07 &  &  &  &  &  & 0.80 &  &  &  &  &  &  &  &  &  & 0.66 &  &  \\

MC-Ignore & 0.52 & 3.02 & 3.18 &  &  &  &  &  & 0.81 &  &  &  &  &  &  &  &  &  & 0.66 &  &  \\

MLkNN & 0.57 & 2.71 & 2.68 & NaN & NaN & NaN & NaN & 0.27 & 0.77 & 0.53 & NaN & NaN & 0.00 & 0.34 & 0.59 & NaN & NaN & 0.01 & 0.62 & 0.40 & 0.04 \\

RAkEL & 0.55 & 2.86 & 2.86 & NaN & NaN & NaN & NaN & 0.30 & 0.81 & 0.56 & NaN & 0.30 & 0.22 & 0.36 & 0.58 & 0.30 & 0.39 & 0.24 & 0.64 & 0.42 & 0.10 \\

\end{tabular}

### 2011-09-02 ese/ese.tex
updated figures for PG comparison

@@
 
-% numbers from mulan.tar.gz/maxdb.mulan.out.tex. I picked the two best learners for each db.
-% The numbers are the micro and macro AUC numbers.
-% MaxDB macro: BR NaN CLR NaN Homer NaN mlkNN Nan
-% MaxDB micro: BR 0.77 CLR 0.67 Homer 0.61 mlkNN 0.67
-% mysql macro: BR 0.74 CLR 0.65 mlKNN 0.61
-% mysql micro: BR 0.81 CLR 0.8 mlkNN 0.78
-% PgSQLn macro: BR 0.64 MlkNN 0.52
-% PGSQLn micro: BR 0.65 MlkNN 0.59
+% Data source is src/validate/output/<db>.mulan.out.tex
+% pgsqln & br & clr & homer
+% macro-auc .64 .58 .58
+% micro .65 .6 .6
+%pgsqla
+% macro .63 .56 .54
+% micro .81 .81 .79
+% maxdb (no macro)
+% micro .77 .67 .61
+% mysql
+% macro .74 .65 .64
+% micro .81 .8 .74
 
@@
 \subfloat[MySQL]{
-\includegraphics[width=0.3\textwidth]{figures/multi-label-results-mysql}
+\includegraphics[width=0.4\textwidth]{figures/multi-label-results-mysql}
 \label{fig:subfig3}
@@
 \subfloat[MaxDB]{
-\includegraphics[width=0.3\textwidth]{figures/multi-label-results-maxdb}
+\includegraphics[width=0.4\textwidth]{figures/multi-label-results-maxdb}
 \label{fig:subfig4}
 }
-\subfloat[PgSQL]{
-\includegraphics[width=0.3\textwidth]{figures/multi-label-results-pgsql}
-\label{fig:subfig5}
-}
-\caption[]{MySQL, PgSQL and MaxDB macro (grey) and micro (black) ROC results per multi-label learner. Possible values range from $0$--$1$.  Dashed line
-indicates the performance of a random classifier.
-%\todo[inline]{better discussion}% and graphical legend would be nice}
+\caption[]{MySQL and MaxDB macro and micro-ROC results per multi-label learner. Possible values range from $0$--$1$.  Dashed line indicates the performance of a random classifier.
 }
@@
 \end{figure*}
+
+\begin{figure}
+\centering
+\subfloat[Macro-ROC]{
+\includegraphics[width=0.4\textwidth]{figures/pg-mulan-macro-compare}
+\label{fig:pgcomp-a}
+}
+\subfloat[Micro-ROC]{
+\includegraphics[width=0.4\textwidth]{figures/pg-mulan-micro-compare}
+\label{fig:pgcomp-b}
+}
+\caption{Author \#1 and author \#2 ROC results for PostgreSQL.} 
+\label{fig:pgcomp}
+\end{figure}
 
@@
 deeper hierarchy for a more accurate learner~\cite{mulan}.
+
+Figure~\ref{fig:pgcomp} shows the ROC results for the PostgreSQL product.
+In this figure, we show the relative differences when we use different training data sets.
+In \ref{fig:pgcomp-a} we see that ROC results are very similar.
+In the other figure, however, author \#1 has dramatically better performance.
+We speculate that this is due to the particular annotation decisions made by author \#1; in some sense he performed better.
+\todo{check this writeup of comparison.
+T-test useful}
+
 These classifiers performed better than other multi-label classifiers as they have the best micro and macro ROC scores. 
@@
 MySQL, BR and Naive Bayes had similar macro-ROC scores of $0.74$.
-% although their results are comparable to the naive Bayesian learners we used in Section \ref{sec:suplabelling}.
-% \todo[inline]{discuss this similarity in more detail}
 

### 2011-09-02 ese/ese.tex
fixed missing ref error

@@
 However, traditional machine learning techniques, such as Naive Bayes, can only map topics to a single class. 
-The Mulan~\cite{mulan} library encapsulates several different multi-label machine learners which can label elements with multiple labels.
+tab:iterrThe Mulan~\cite{mulan} library encapsulates several different multi-label machine learners which can label elements with multiple labels.
 Mulan also includes methods for determining the performance of such techniques.
@@
 
-Table \ref{tab:iterr} describes the Cohen Kappa and the Spearman
+Table \ref{tab:interr} describes the Cohen Kappa and the Spearman
 correlation of our per topic annotations for each tag. We had to

### 2011-09-03 ese/ese.tex
writing

--- 2011-09-02 ese/ese.tex
@@
 \subsection{Do different developers work on different NFRs?}
-We did a t-test to assess whether there was statistically significant differences in the types of NFRs developers chose to work on, for the PostgreSQL project only. 
-\todo{add more, figure}
-
-*** We also clustered the authors by similarity and we found that
-among some of the most prolific postgresql authors, their NFR focus
-was distinctly different *** Some confounds are that postgresql
-authors commit the patches of others.
+
+%data/clusters.R
+
+We wanted to see how similar developers were to each other in terms of
+the NFRs that worked on. First we did pairwise $X^2$ (chi-squared)
+tests and Kolmogorov-Smirnov tests for each developer pair and we
+found that $3/10$ of these tests the developers were significantly
+different, that is their distributions were not similar. This implies
+that there are authors who focus on different NFRs or different
+proportion of NFRs.
+
+In Figure \ref{fig:authorcluster} we clustered developers using the
+Ward method of Hierarchical clsutering. We've clustered the authors
+into 2 clusters and 5 clusters.
+
+\begin{figure}
+  \centering
+  \includegraphics[width=.9\textwidth]{figures/postgresql-author-cluster}
+  \caption{PostgreSQL Author Clusters: Authors clustered by cosine
+    distance of their NFR contributions}
+\label{fig:authorcluster}
+\end{figure}
 

### 2011-09-03 ese/ese.tex
why did this take so much time?

@@
 In Figure \ref{fig:authorcluster} we clustered developers using the
-Ward method of Hierarchical clsutering. We've clustered the authors
-into 2 clusters and 5 clusters.
+Ward method of Hierarchical clustering.  This diagram shows that
+\texttt{petere}, \texttt{tgl} and \texttt{momjian} form their own
+cluster (they are also the most frequent committers). This is
+interesting because it means that the important developers are not
+committing code that fits the same NFR profile. In terms of the larger
+cluster we can see that \texttt{momjian} and \texttt{tgl} are in the
+same cluster, but \texttt{petere} is not. The most frequent
+committers do not share the same clusters. This implies that
+developers do work on different sets of NFRs and have different
+software quality focus.
+
+Figure \ref{fig:authorcluster} clustered the authors into $2$ clusters
+and $6$ clusters. We chose $6$ because the difference in width between
+entities and centroids was minimal at $6$ and the derivative of was
+the highest in that region as well.
+
+It should be noted that much of PostgreSQL development originates from
+the mailinglist and the developers are often pushing integrations.
 

### 2011-09-03 data/clusters.R

@@ -1,0 +1,116 @@
+v <- read.csv("Author_NFRs.csv")
+vv <- v[1:18,]
+for (i in 2:17) {
+  vv[,i] <- as.numeric(vv[,i])
+}
+tv <- t(vv[,2:8])
+authors <- matrix(tv,nrow=7,dimnames=list(labels(tv)[[1]],vv[,1]))
+#hc <- hclust(as.dist(cor(authors)),method="ward")
+#hc <- hclust(as.dist(cor(authors)),method="centroid")
+#
+hc <- hclust(as.dist(cor(authors)),method="ward")
+pdf("postgresql-author-cluster.pdf")
+plot(hc,sub="Organized into 2 and 6 clusters",xlab="PostgreSQL Authors")
+rect.hclust(hc,k=2,border="blue")
+#rect.hclust(hc,k=3,border="red")
+#rect.hclust(hc,k=4,border="green")
+rect.hclust(hc,k=6,border="purple")
+dev.off()
+
+nauthors <- authors
+normalize <- function(x) { x / sum(x) }
+for (author in labels(nauthors)[[2]]) { nauthors[,author] <-
+                                          normalize(authors[,author])
+                                      }
+nauthor <- length(labels(authors)[[2]])
+authorNames = labels(authors)[[2]]
+chisqm <- matrix(c(1:(nauthor*nauthor))*0,ncol=nauthor,dimnames=list(authorNames,authorNames))
+ksm <- matrix(c(1:(nauthor*nauthor))*0,ncol=nauthor,dimnames=list(authorNames,authorNames))
+
+for (author1 in labels(authors)[[2]]) {
+  for (author2 in labels(authors)[[2]]) {
+    print(paste(author1,author2))
+    pv   <- chisq.test(authors[,c(author1,author2)])$p.value
+    kspv <- ks.test(as.numeric(makeAuthorData(authors[,author1])),
+                    as.numeric(makeAuthorData(authors[,author2])))$p.value
+    print(paste(pv,kspv))
+    chisqm[author1,author2] <- pv
+    ksm[author1,author2] <- pv
+  }
+}
[…]

### 2011-09-04 ese/ese.tex
more author analysis

@@
 \begin{abstract}
-% %\todo[inline]{Topics and LDA sooner, Topics/concept location is a known
+
 Researchers have employed a variety of techniques to extract underlying
@@
 % this reduces the line length and drops a ref even though we still have a scacchi kickin around
-%Massey~\cite{massey02icse} and Scacchi~\cite{scacchi05b} looked at the topic of requirements in open-source software. 
-%Their work discusses the source of the requirements and how they are used in the development process. 
-% German~\cite{german03gnome} looked at GNOME specifically, and listed several sources for requirements: leader interest, mimicry, brainstorming, and prototypes. 
-%None of this work  addressed quality requirements in OSS, nor did it examine requirements trends.
-\todo{re-add this related work?}
-% XXX vanity or space.. 
-%XXXXXXXXXXXX OPTIONAL
+Massey~\cite{massey02icse} and Scacchi~\cite{scacchi05b} looked at the topic of requirements in open-source software. 
+Their work discusses the source of the requirements and how they are used in the development process. 
+ German~\cite{german03gnome} looked at GNOME specifically, and listed several sources for requirements: leader interest, mimicry, brainstorming, and prototypes. 
+None of this work  addressed quality requirements in OSS, nor did it examine requirements trends.
 % NE 0113 this seems too early to mention 'result'.
@@
 annotations that would be aggregated up to one of these higher-level labels.
-%\todo[inline]{this confused people. Is it truly unsupervised? Yes, but it needs to be clearer.}
 %We compared against this data-set, but we also used the data-set for our supervised machine learning based topic classification. 
 % A topic may have more than one matching keyword. % ok what?
-%\todo[inline]{inter-rater reliability }
-% \todo[inline]{There should be a section here which is about manual annotation. 
-%It is really important and it is required by both analyses. Plus we can tell them how.}
-
-%\todo[inline]{NEIL: this section should be updated to explain F-measure}
+
 We validate classification performance using the Receiver Operating
@@
 
-%\subsection{Generating word lists}
-%\label{sec:genwordlist}
-% %\todo[inline]{Make it clear where the data comes from! Neil says: Also, I
-%   didn't get the impression they understood that we weren't generating
-%   labels from the data, but rather from independent taxonomies. Of
-%   course if you use labels from the data it isn't comparable. But the
-%   point of ISO 9126 etc. was to avoid exactly this problem.}
-
-% \todo[inline]{Need to hammer it in that the word bags are not project specific and
-%  are inherited so it isn't a supervised method}
-
-%Topics are word distributions, lists of words ranked by frequency, that are burdensome to interpret and hard to distinguish and understand. 
-%While the topics themselves are generated automatically, how to interpret these topics is less clear.
-%For example,
-
 \subsubsection{Generating Word Lists}
@@
 
-%\todo[inline]{* clarify the mailinglist stuff we did}
-%\todo[inline]{This part doesn't make a lot of sense}
 To construct these expanded word-lists, we used
 WordNet~\cite{Fellbaum1998}.
-%, an English-language ``lexical database'' that contains semantic relations between words,
-% including common related forms (similar to word stemming), meronymy and synonymy. 
 We then added Boehm's software quality model~\cite{Boehm+:1976:ICSE}, and classified his eleven `ilities' into their respective ISO9126 NFRs. 
@@
 
-%\todo[inline]{NEIL: this could be a subsubsection}
-%\todo[inline]{* the hierarchy of tags was not clear}
 %\noindent \textbf{Generating labels automatically} ---
@@
 domain knowledge of software to label relevant topics.
-%\todo[inline]{Neil does this explain the unsupervised aspect well enough?}
 
@@
 
-%\todo[inline,color=green]{this table should have mysql}
 \begin{table}
@@
 
-%\todo[inline]{Should we use F-1 as well to satisfy everyone?}
-
 \subsubsection{Analysis of the Semi-unsupervised Labelling} %--- %Based
@@
 
-%\todo[inline]{since ROC is new-sounding, I wonder if a more lengthy explanation would be merited. An obvious question is why not precision/recall? --
-% ROC is hot right now and it combines both}
-% Rebuttal: ROC is not new and it is what the machine learning community, search based software engineering and now the MSR community are using +
-%it has natural meaningful quality mappings.
 % \begin{figure}[t]
@@
 % dashed line indicates the performance of a random classifier. B) F1 measure. Possible values range from 0--1.
-% %\todo[inline]{better discussion}% and graphical legend would be nice}
+
 % }
@@
 
-%\todo[inline]{this is too tiny in print}
   \label{fig:maxdb-unsup-results}
@@
 	relate a project milestone (\emph{*Key events*}) to our topic windows.
-% \todo[inline]{better discussion}
+
 }
@@
 %\emph{portability}, yet MaxDB had more of a focus on \emph{efficiency} and \emph{reliability}. 
-% \todo[inline]{expand}
-%\todo[inline,color=yellow]{expand on this with dates, and reference to visualizations}
 %\subsection{Research questions}
@@
 %\subsubsection{Do label frequencies change over time?}
-%\todo{add more on PGSQL}
+
 In both MaxDB and MySQL the frequencies generally decreased with age. 
@@
 It should be noted that much of PostgreSQL development originates from
-the mailinglist and the developers are often pushing integrations.
+the mailinglist and the developers are often pushing integrations. If
+we compare the total to each author we find that 1 quarter of the
+authors have a similarity of $0.47$ or less. This is depicted in
+Figure \ref{fig:authorsim} where we can see that not all authors are
+as similar to the global trend as others. Of
+course, we found that number of commits had a strong correlation
+($0.59$ Pearson) between the similarity to the aggregate and the
+number of topics associated with that author.
+
+Our theory is that the less frequent committers perhaps are more
+focused and less general, thus their distributions of topics are
+different than the main developers who commit code in many different
+contexts. Perhaps the lead developers ``wear more hats'' or have more
+time to ``wear more hats''.
+
+Given the top 15 PostgreSQL devs we found that proportionally many
+devs had a focus on NFR topics. Authors with top proporitions per NFR
+were:
+\texttt{dennis} for usability; 
+\texttt{scrappy} and \texttt{meskes} for portability;
+\texttt{inoue} and \texttt{neilc} for efficiency;
+\texttt{jurka} and \texttt{joe} for reliability;
+\texttt{thomas} and \texttt{weick} for functionality;
+and \texttt{ishii} and \texttt{scrappy} for maintainability.
+Given these results we found that yes indeed there was often a focus
+of $1/3$ to $1/10$ of commits dedicated to a particular NFR.
+
+% Dennis is into usability
+% scrappy and meskes portability
+% Inoue and neilc for effiency
+% jurka joe and barry for Reliabilityb
+% thomas weick teodor for Functionality
+% ishii, scrappy, darcy for maintainability
+
+%\todo{Now look at the authors, what are they into?}
+
+% tgl == momjian
+% > nauthors
+%                     momjian     thomas      davec      neilc      ishii darcy
+% None            0.003       0.150       0.048       0.020       0.081       0.15
+% Portability     0.211       0.083       0.177       0.173       0.122       0.20
+% Efficiency      0.114       0.100       0.080       0.152       0.132       0.00
+% Reliability     0.149       0.100       0.137       0.159       0.102       0.15
+% Functionality   0.220       0.300       0.266       0.145       0.214       0.20
+% Maintainability 0.132       0.150       0.145       0.104       0.224       0.20
+% Usability       0.167       0.116       0.145       0.243       0.122       0.10
+%                      barry      wieck      inoue      pgsql teodor jurka
+% None            0.03488372 0.04444444 0.04166667 0.06521739 0.0250  0.04
+% Portability     0.13372093 0.17777778 0.09722222 0.19565217 0.2250  0.18
+% Efficiency      0.13372093 0.08888889 0.26388889 0.15217391 0.1250  0.12
+% Reliability     0.19767442 0.11111111 0.12500000 0.06521739 0.0875  0.24
+% Functionality   0.20930233 0.28888889 0.23611111 0.19565217 0.2375  0.16
+% Maintainability 0.13372093 0.06666667 0.13888889 0.17391304 0.1250  0.14
+% Usability       0.15697674 0.22222222 0.09722222 0.15217391 0.1750  0.12
+%                         tgl    scrappy     petere     dennis     meskes
+% None            0.006704981 0.08333333 0.02296451 0.06451613 0.02298851
+% Portability     0.207854406 0.25000000 0.20250522 0.22580645 0.23754789
+% Efficiency      0.119731801 0.08333333 0.09812109 0.12903226 0.08045977
+% Reliability     0.150383142 0.16666667 0.12734864 0.09677419 0.14559387
+% Functionality   0.217432950 0.12500000 0.20668058 0.12903226 0.22222222
+% Maintainability 0.129310345 0.20833333 0.14196242 0.00000000 0.10727969
+% Usability       0.168582375 0.08333333 0.20041754 0.35483871 0.18390805
+%                        joe
+% None            0.06122449
+% Portability     0.22448980
+% Efficiency      0.08163265
+% Reliability     0.22448980
+% Functionality   0.12244898
+% Maintainability 0.10204082
+% Usability       0.18367347
 
@@
 \label{fig:authorcluster}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=.9\textwidth]{figures/author-distance-from-aggregate}
+  \caption{PostgreSQL Author Similarity to the Global Distribution of NFR topics}
+\label{fig:authorsim}
 \end{figure}
@@
 %our results might not generalize to other domains. %We only studied database systems.
-% \todo[inline]{If we compare with the same domain we don't get to say it generalizes a lot?}
 

### 2011-09-04 data/clusters.R

@@ -1,3 +1,5 @@
+
+
 v <- read.csv("Author_NFRs.csv")
 vv <- v[1:18,]
 for (i in 2:17) {
@@ -19,6 +21,13 @@ dev.off()
 
 nauthors <- authors
 normalize <- function(x) { x / sum(x) }
+
+makeAuthorData <- function(x) {
+  l <- as.factor(labels(nauthors)[[1]])
+  unlist(sapply(c(1:length(l)), function(i) { rep(l[i], x[i]) }))
+}
+
+
 for (author in labels(nauthors)[[2]]) { nauthors[,author] <-
                                           normalize(authors[,author])
                                       }
@@ -40,14 +49,32 @@ for (author1 in labels(authors)[[2]]) {
 }
 chisqm[!is.finite(chisqm)] <- 0
 ksm[!is.finite(ksm)] <- 0
-
-makeAuthorData <- function(x) {
-  l <- as.factor(labels(nauthors)[[1]])
-  unlist(sapply(c(1:length(l)), function(i) { rep(l[i], x[i]) }))
-}
+general<-sapply(labels(authors)[[1]],function(x){sum(authors[x,])})
+ngeneral<-sapply(labels(nauthors)[[1]],function(x){sum(nauthors[x,])})
 
 sapply( authorNames, function(x) { median(chisqm[,x]) })
-#
+
+pdf("author-distance-from-aggregate.pdf")
+d <- sapply(labels(authors)[[2]],function(x){cor(general,authors[,x])})
+asize <- sapply(labels(authors)[[2]],function(x){sum(authors[,x])})
+print(cor(d,asize))
+print(cor(d,asize,method="spearman"))
+plot(d,asize,main="Author Correlation with NFR distribution versus # of Author Topics",xlab="Similarity of Author NFR topics to global distribution",ylab="# of Topics associated with Author")
+text(d,asize,labels(authors)[[2]],pos=3)
+dev.off()
+
+
+pdf("author-distance-from-aggregate-normalized.pdf")
+nd <- sapply(labels(nauthors)[[2]],function(x){cor(general,nauthors[,x])})
+nasize <- sapply(labels(nauthors)[[2]],function(x){sum(nauthors[,x])})
+print(cor(nd,asize))
+print(cor(nd,asize,method="spearman"))
[…]

### 2011-09-05 ese/ese.tex
46% through and not a lot of problems

@@
 NFR-related tags.
-Project dashboard~\cite{dashboard} are typically employed by managers and are used to
+Project dashboards~\cite{dashboard} are typically employed by managers and are used to
 provide quick summaries of the effort put into a software project. In this case, 
 labelled topic extraction would allow managers to track effort
-related to NFR topics, such as portability.
+related to NFR topics, such as \emph{portability}.
 These techniques also allow for the annotation of commit comments
@@
 %e created labels using a well-known NFR taxonomy. 
-In supervised learning, we trained our data with manual annotations in order to match topics with NFRs. Finally, these topics were used to analyze the
+With supervised learning, we trained our data with manual annotations in order to match topics with NFRs. Finally, these topics were used to analyze the
 role of NFRs in software maintenance.
@@
 positive rate (1 - specificity). 
-A perfect learner has a ROC value of 1.0, reflecting perfect recall and precision. 
-A ROC result of 0.5 would be equivalent to a random learner (that is, issuing as many false positives as true positives). 
+A perfect learner has a ROC value of $1.0$, reflecting perfect recall and precision. 
+A ROC result of $0.5$ would be equivalent to a random learner (that is, issuing as many false positives as true positives). 
 The %area under the -- NE confusing to have area again ... and not importantana
@@
 relevant to discussion of \emph{reliability}, but is not in the standard. 
-We therefore took the NFRs from the ISO9126 and added to them.
+We therefore took the NFRs from the ISO9126 and added terms to them.
 

### 2011-09-05 ese/ese.tex
63%

@@
 As described in Section \ref{sec:wordlist} we used both ROC and F-1 measures to evaluate the performance of the classification.
-Figure \ref{fig:maxdb-unsup-results} shows our ROC results for MaxDB and MySQL. We describe F-1 results in the text below.
+Figure \ref{fig:maxdb-unsup-results} shows our ROC results for PostgreSQL, MaxDB and MySQL. We describe F-1 results in the text below.
 
@@
 
-For both projects Bayesian techniques did the best out of a wide variety of machine learners tested. 
+For all projects Bayesian techniques did the best out of a wide variety of machine learners tested. 
 Our best learners, Discriminative Multinomial Naive Bayes, Naive Bayes
-and Multinomial Naive Bayes  are all based on Bayes's theorem and all
+and Multinomial Naive Bayes  are all based on Bayes' theorem and all
 assume, naively, that the features supplied are independent. 
 %The features we used are word counts per message. 
-One beneficial aspect of this result is that it suggests we can have very fast training and classifying  since Naive Bayes can be calculated in $O(N)$
+One beneficial aspect of this result is that it suggests we can have
+very fast training and classifying  since training or classifying one
+instance of Naive Bayes can be calculated in $O(N)$
 for $N$ features.

### 2011-09-05 ese/ese.tex
migod's changes

@@
 
-Topic modeling is a machine learning technique which creates
+Topic modeling is a machine learning technique that creates
 multinomial distributions of words extracted from a text corpus. 
@@
 %We then present visualizations of named topics and their trends over time to aid communication and analysis. 
-To demonstrate an application of labelled topic extraction, we use an exploratory case study of two open source database systems to show how named
+To demonstrate an application of labelled topic extraction, we use an exploratory case study of three open source database systems to show how named
 topics can be compared between projects  (Section \ref{sec:analysis}). 
@@
 
-Cleland-Huang and her colleagues have investigated mining requirements documents for non-functional requirements (NFR) (quality
-requirements)~\cite{Cleland-Huang2006}. 
-Their approach is similar to ours, as they mined keywords from NFR
-catalogues.
-They differ because they mine requirements documents where as we mine revisions.
+Cleland-Huang and her colleagues have investigated mining requirements
+documents for non-functional requirements (NFR) (software
+qualities)~\cite{Cleland-Huang2006}.  Their approach is similar to
+ours, as they mined keywords from NFR catalogues.  They differ because
+they mine requirements documents where as we mine revisions.
 %One approach they tried was similar to this one, with keywords mined from NFR catalogues found in their previous work~\cite{chung99}. 
@@
 However, traditional machine learning techniques, such as Naive Bayes, can only map topics to a single class. 
-tab:iterrThe Mulan~\cite{mulan} library encapsulates several different multi-label machine learners which can label elements with multiple labels.
+The Mulan~\cite{mulan} library encapsulates several different multi-label machine learners which can label elements with multiple labels.
 Mulan also includes methods for determining the performance of such techniques.
@@
 \emph{why} is what we are after.
-We investigate the history of the two large-scale database systems that we studied. 
+We investigate the history of the three large-scale database systems that we studied. 
 We use our technique to show the topic of development efforts over time in each project.
@@
   quarter's objectives.  The difference in NFR proportion is
-  interesting because it implies a difference in focus between two projects.
+  interesting because it implies a difference in focus between projects.
 
 \item \emph{Do different developers work on different NFRs?} For a given project, it is reasonable to think that developers are either assigned
-(in commercial organizations) or choose (in open-source organizations) to a particular NFR. For example, one developer might be more junior, and take
+(in commercial organizations) or choose (in open-source organizations) to work on a particular NFR. For example, one developer might be more junior, and take
 responsibility for the low-impact reliability fixes. Another, more senior developer might assume responsibility for major improvements such as efficiency
@@
 
-Our theory is that the less frequent committers perhaps are more
+Our theory is that the less frequent committers are more
 focused and less general, thus their distributions of topics are
 different than the main developers who commit code in many different
-contexts. Perhaps the lead developers ``wear more hats'' or have more
+contexts.  The lead developers either ``wear more hats'' or have more
 time to ``wear more hats''.
@@
 
-Authors Neil and Abram each annotated the PostgreSQL topics in order
-to determine inter-rater reliability. While when they read each others
-annotations they generally agreed, the inter-rater reliability did not
-bear this out.
+To determine inter-rater reliability two of the authors---Ernst and Hindle---each annotated the PostgreSQL topics,
+and then evaluated each other's annotations.
 
@@
 option a usability issue? These kinds of questions illustrate some of
-the agreement, disagreement and ambiguity about these labels. Further
-more, both annotators found that the annotation was time-consuming and
+the agreement, disagreement and ambiguity about these labels. Furthermore, 
+both annotators found that the annotation was time-consuming and
 difficult.
@@
 specifically the ISO9126 quality taxonomy, to produce a method of partially-automated (supervised) and fully-automated (semi-unsupervised) topic labelling.
-Since the word-list technique is not project-specific, we used it to compare two distinct projects, where we showed our technique produced interesting
+Since the word-list technique is not project-specific, we used it to compare three distinct projects, where we showed our technique produced interesting
 insight into maintenance activity. %topics and artifacts from project histories.

### 2011-09-05 ese/ese.tex
john m. changes

@@
 requirements. Our approach was evaluated with experiments and case studies
-on three large-scale RDBMS projects: MySQL, PostgreSQL and MaxDB.  The case studies
+on three large-scale Relational Database Management Systems (RDBMS) projects: MySQL, PostgreSQL and MaxDB.  
+The case studies
 show that labelled topic extraction can produce appropriate,
@@
 understanding of \emph{why} a system has evolved the way it has. 
-This is different than \emph{how} a system has evolved, because the behaviour of change itself is the
-\emph{how}.
+This is different from \emph{how} a system has evolved. 
+% NE this phrase confuses me: "because the change of behaviour itself is the \emph{how}.
 Looking back on streams of artifacts scattered across different
 repositories, inferring what activities were performed, when, and for
-what reasons, is hard to do without expert advice from the developers
+what reasons, is hard without expert advice from the developers
 involved. 
@@
 qualities)~\cite{Cleland-Huang2006}.  Their approach is similar to
-ours, as they mined keywords from NFR catalogues.  They differ because
+ours, as they mine keywords from NFR catalogues.  They differ from our approach because
 they mine requirements documents where as we mine revisions.
 %One approach they tried was similar to this one, with keywords mined from NFR catalogues found in their previous work~\cite{chung99}. 
-They demonstrated a recall of $80\%$ with precision of $57\%$ for the \emph{security} NFR, but could not find a reliable source of keywords for other NFRs. 
+Their approach resulted in a recall of $80\%$ with precision of $57\%$ for the \emph{security} NFR, but could not find a reliable source of keywords for other NFRs. 
 Instead, they developed a supervised classifier by using human experts to identify an NFR training set. 
-Our research differs because we use a more comprehensive set of terms based on the taxonomy we chose.
+Our research is different because we use a more comprehensive set of terms based on a taxonomy that is an integral part of our framework.
 Another difference is that we make cross-project comparisons instead of focusing
@@
 
-Similarly, Mockus and Votta~\cite{Mockus00} studied a large-scale industrial change-tracking system. 
-Mockus and Votta leveraged WordNet~\cite{Fellbaum1998}, an English-language ``lexical database'' that contains semantic relations between words,
+Similarly, Mockus and Votta~\cite{Mockus00} study a large-scale industrial change-tracking system. 
+Mockus and Votta leverage WordNet~\cite{Fellbaum1998}, an English-language ``lexical database'' that contains semantic relations between words,
 including common related forms (similar to word stemming), meronymy and synonymy.
-They used WordNet for word roots as they felt the synonyms would be
+They use WordNet for word roots as they felt the synonyms would be
 non-specific and cause errors.
 %, problems we also encountered. 
-Mockus et al. validated their labels with system developers.
+Mockus et al. validate their labels with system developers.
 Since we study multiple projects, instead of a single project, these
-kind of interviews are somewhat infeasible (particularly in the distributed world of open-source software).
+kind of interviews are not feasible (particularly in the distributed world of open-source software).
 
@@
 Figure \ref{fig:process} gives an outline of our methodology.
-We began by gathering source data and creating topic models. For semi-unsupervised labelling, we generated three sets of word-lists as signifiers for NFRs.
+We begin by gathering source data and creating topic models. For semi-unsupervised labelling, we generate three sets of word-lists as signifiers for NFRs.
 %e created labels using a well-known NFR taxonomy. 
-With supervised learning, we trained our data with manual annotations in order to match topics with NFRs. Finally, these topics were used to analyze the
+With supervised learning, we train our data with manual annotations in order to match topics with NFRs. Finally, these topics are used to analyze the
 role of NFRs in software maintenance.

### 2011-09-05 src/exp1/lda_relator_report.sh
exp1 pgsql resultsd

@@ -1,4 +1,5 @@
-for proj in mysql maxdb
+#for proj in mysql maxdb pgsql
+for proj in pgsql
 do
 	cp ../counts/$proj.lda.report ./lda.report
 	~/projects/hiraldo_grok/hs/Lda_relator  > $proj.lda_relator.report

### 2011-09-05 ese/ese.tex
exp1 pgsql resultsd
exp1 to exp3 postgresql updates

@@
 
-\todo{this discussion is about which DB? where do the numbers come from?}
+%\DONE{this discussion is about which DB? where do the numbers come from?}
 For \textsf{exp1} the labels with the most topics
 were
-\emph{correctness} (182 topics) and \emph{testability} (121). 
-We did not see many results for \emph{usability} or \emph{accuracy}, which were associated with fewer than ten topics. 
-We also looked for correlations between our labels: excluding double matches (self-correlation), our highest co-occurring terms were \emph{verifiability}
+\emph{correctness} (182/305/748 topics) and \emph{testability} (121/238/625). 
+We did not see many results for \emph{usability} (4/0/138) or
+\emph{accuracy} (3/0/27), which were associated with fewer than ten topics. 
+We also looked for correlations between our labels: excluding double
+matches (self-correlation), our highest co-occurring terms were
+\emph{verifiability} or \emph{correctness}
 with \emph{traceability}, and \emph{testability} with \emph{correctness} (76 and 62 matches, respectively).
@@
 For \textsf{exp2}, there are more unnamed topics than \textsf{exp1}. 
-Only \emph{reliability} produces a lot of matches, mostly with the word ``error''. 
+Only \emph{reliability} produces the most matches, mostly due to the word ``error''. 
 Co-occurrence results were poor. This suggests our word lists were overly restrictive.
+For PostgreSQL \emph{reliability} and  \emph{usability} co-occurred
+with \emph{portability} and \emph{efficiency}.
 
@@
 As we mentioned, the word-lists are broad, so there are likely to be false-positives (discussed below). 
-The most frequent label, 265 topics, was \emph{usability}, and the
-least frequent label, 44 topics, was \emph{maintainability}. 
-Common co-occurrences were \emph{reliability} with \emph{usability}, \emph{efficiency} with \emph{reliability}, and \emph{efficiency} with \emph{usability}
-(200, 190, and 150 topics in common, respectively).
+The most frequent label across all projects was \emph{usability}, and the
+least frequent label was \emph{maintainability}. 
+Common co-occurrences were \emph{reliability} with \emph{usability}, \emph{efficiency} with \emph{reliability}, and \emph{efficiency} with \emph{usability}.
+%(200, 190, and 150 topics in common, respectively).
 
@@
 For \textsf{exp1} the F-measures for MaxDB were from $0$ to $0.18$ with an average
-of $0.03$, and for MySQL were from $0$ to $0.16$ with an average of
-$0.05$.
-\todo{update these numbers but from where? }
+of $0.03$, for MySQL were from $0$ to $0.16$ with an average of
+$0.05$, and for PostgreSQL $0$ to $0.15$ with an average of $0.07$.
+
+%\todo{update these numbers but from where? }
 % 0.177215189873 / 6
@@
 % .04580718793751666666
+
+% pgsql
+% portability 
+% 0.0910746812386
+% efficiency 
+% 0.154639175258
+% reliability
+% 0.0470588235294
+% maintainability
+% 0
+% usability
+% 0.0720720720721
+%> v <- c(0.0910746812386 , 0.154639175258 , 0.0470588235294 , 0 , 0.0720720720721)
+%> summary(v)
+%   Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
+%0.00000 0.04706 0.07207 0.07297 0.09107 0.15460
 

### 2011-09-05 ese/ese.tex
expand on why postgresql performs soo poorly?

@@
 
-%\todo{update these numbers but from where? }
+%\DONE{update these numbers but from where? }
 % 0.177215189873 / 6
@@
 
-Figure \ref{fig:best-learn-per-tag} shows that MaxDB and MySQL have quite different results, as the ROC values for \emph{reliability} and \emph{functionality}
- swap between projects. For PostgreSQL, the performance is nearly always poorer than the other two systems. 
-\todo{expand on why?}
+Figure \ref{fig:best-learn-per-tag} shows that MaxDB and MySQL have
+quite different results, as the ROC values for \emph{reliability} and
+\emph{functionality} swap between projects. For PostgreSQL, the
+performance is nearly always more poor than the other two systems. The
+reason for this lack of performance could be that the number of
+topics, $N$ could be non-optimal for PostgreSQL. Perhaps topics were
+getting too mixed and it was becoming hard to distinguish one topic
+from the next. PostgreSQL's dataset was the largest of the 3 projects,
+the XML data-dump that described the PostgreSQL topics that we
+annotated was 8X larger than MySQL and 1.7X larger than MaxDB. These
+size differences arise from the number of commits, the number of files
+and the verbosity of the commit descriptions.
+
+%\DONE{expand on why?}
 

### 2011-09-06 ese/ese.tex
macro-micro multilabel learner comparison

@@
 %together.
-%\todo[inline]{Comment about correlation, we should link to the Mulan section, perfect segway since Mulan relies on correlation.}
+%\DONE[inline]{Comment about correlation, we should link to the Mulan section, perfect segway since Mulan relies on correlation.}
 
@@
 
-Figure~\ref{fig:pgcomp} shows the ROC results for the PostgreSQL product.
-In this figure, we show the relative differences when we use different training data sets.
-In \ref{fig:pgcomp-a} we see that ROC results are very similar.
-In the other figure, however, author \#1 has dramatically better performance.
-We speculate that this is due to the particular annotation decisions made by author \#1; in some sense he performed better.
-\todo{check this writeup of comparison.
-T-test useful}
+Figure~\ref{fig:pgcomp} shows the ROC results for the PostgreSQL
+product. In this figure, we show the relative differences when we use
+different training data sets. In \ref{fig:pgcomp-a} we see that ROC
+results are very similar. In the other figure, however, author \#1 has
+dramatically better performance. We speculate that this is due to the
+particular annotation decisions made by author \#1; in some sense he
+performed better. 
+The difference in Macro-ROC was not signifcant, but the difference in
+Micro-ROC was as the p-value of Student's T-test was $< 0.001$. 
+%\DONE{check this writeup of comparison. T-test useful}
 
@@
 security problems were more common (security is a component of \emph{functionality} in the ISO9126 model). 
-In both cases \emph{portability} was a constant maintenance concern and was prevalent throughout the lifetime of the projects. It may surprise developers
+PostgreSQL had more NFR topics in general, but \emph{portability},
+\emph{functionality} and \emph{usability} were more prevalent in
+PostgreSQL. One notable difference between PostgreSQL and the other
+two projects is that many of the commits for PostgreSQL were UI
+changes and documentation changes. PostgreSQL also seemed to focus
+more on reliability later in the case than earlier like MySQL.
+In all cases \emph{portability} was a constant maintenance concern and was prevalent throughout the lifetime of the projects. It may surprise developers
 how often portability arises as a concern.
-\todo{add the pgsql data}
+
+%\DONE{add the pgsql data}
 
@@
 
-%\todo{Now look at the authors, what are they into?}
+%\DONE{Now look at the authors, what are they into?}
 

### 2011-09-06 data/ttest-for-multilabel.R

@@ -1,0 +1,17 @@
+Abram.microAUC <- c(0.81 , 0.81 , 0.79 , 0.70 , 0.73  , 0.84 , 0.80)
+Abram.macroAUC <- c(0.63 , 0.56 , 0.54 , 0.51 , 0.54  , 0.54 , 0.53)
+Neil.macroAUC  <- c( 0.64 , 0.58 , 0.58 , 0.55 , 0.53 , 0.53 , 0.56 )
+Neil.microAUC  <- c( 0.65 , 0.60 , 0.60 , 0.56 , 0.55 , 0.59 , 0.58 )
+
+
+pvalues <- c(
+  t.test(Abram.microAUC,Neil.microAUC)$p.value,
+  t.test(Abram.macroAUC,Neil.macroAUC)$p.value,
+  wilcox.test(Abram.microAUC,Neil.microAUC)$p.value,
+  wilcox.test(Abram.macroAUC,Neil.macroAUC)$p.value
+             )
+tests <- as.factor(c("t.test","t.test","wilcox.test","wilcox.test"))
+x <- c()
+x$tests <- tests
+x$pvalues <- pvalues
+             

### 2011-09-06 ese/ese.tex
global both issues

@@
 We noticed that occasionally the names of modules would conflict with words related to other non-functional requirements. 
-For instance, optimizers are very common modules in database systems: both MySQL and MaxDB have optimizer modules. 
+For instance, optimizers are very common modules in database systems:
+all three projects, MySQL, MaxDB, and PostgreSQL have optimizer modules. 
 In MySQL the optimizer is mentioned but often the change addresses  correctness or another quality. 
@@
 For MaxDB, \textit{reliability} was slightly better detected using the static word list of \textsf{exp2}. 
-In general, the machine learners and \textsf{exp3} did better than \textsf{exp2} for both MaxDB and MySQL. 
+In general, the machine learners and \textsf{exp3} did better than
+\textsf{exp2} for MySQL and MaxDB, yet for PostgreSQL the
+\textsf{exp2} wordlists performed better.
 For both MySQL and MaxDB \textit{usability} was better served by \textsf{exp2}. 
@@
 
-With ROC values ranging from $0.6$ to $0.8$ we can see there is promise in these methods.
+With ROC values ranging from $0.6$ to $0.8$ for MySQL and MaxDB and
+$0.47$ to $0.6$ for PostgreSQL ,we can see there is promise in these methods.
 \textsf{exp2} and \textsf{exp3} both indicate that static information can be used to help label topics without any training whatsoever. 

### 2011-09-06 ese/ese.tex
added dates

@@
 requirements. Our approach was evaluated with experiments and case studies
-on three large-scale Relational Database Management Systems (RDBMS) projects: MySQL, PostgreSQL and MaxDB.  
+on three large-scale Relational Database Management System (RDBMS) projects: MySQL, PostgreSQL and MaxDB.  
 The case studies
@@
 This topic might reasonably be labelled as
-``\emph{event handling}'' by a developer who understands the domain well,
+\emph{``event handling''} by a developer who understands the domain well,
 despite the fact that this label does not appear in the word list itself.  
@@
 In this paper, we introduce \emph{labelled topic extraction}, an
-approach to topic labelling that creates labels automatically that are
-project independent.
+approach to topic labelling that labels topics automatically with
+labels that are project independent.
 %is both automatic and creates labels that are project independent.
@@
 
-Similarly, Mockus and Votta~\cite{Mockus00} study a large-scale industrial change-tracking system. 
-Mockus and Votta leverage WordNet~\cite{Fellbaum1998}, an English-language ``lexical database'' that contains semantic relations between words,
+Similarly, Mockus and Votta~\cite{Mockus00} studied a large-scale industrial change-tracking system. 
+Mockus and Votta leveraged WordNet~\cite{Fellbaum1998}, an English-language ``lexical database'' that contains semantic relations between words,
 including common related forms (similar to word stemming), meronymy and synonymy.
-They use WordNet for word roots as they felt the synonyms would be
+They used WordNet for word roots as they felt the synonyms would be
 non-specific and cause errors.
 %, problems we also encountered. 
-Mockus et al. validate their labels with system developers.
+Mockus et al. validated their labels with system developers.
 Since we study multiple projects, instead of a single project, these
-kind of interviews are not feasible (particularly in the distributed world of open-source software).
+kind of interviews were not feasible (particularly in the distributed world of open-source software).
 
 Another approach is to extract concerns from software repositories.
-Marcus et al.~\cite{marcus04wcre} use Latent Semantic Indexing (LSI)
+Marcus et al.~\cite{marcus04wcre} used Latent Semantic Indexing (LSI)
 to identify commonly occurring concerns for software maintenance. 
@@
 %Some results were interesting, but their precision was quite low. 
-With ConcernLines, Treude et al.~\cite{treude09cl} show tag occurrence
+With ConcernLines, Treude et al.~\cite{treude09cl} showed tag occurrence
 using colour and intensity. 
-They mine
+They mined
 developer created tags 
@@
 The labels are reasonable enough, but still require an expert in the field to determine them. 
-Furthermore, these labels are project-specific, because they are generated from the data of that project. E.g., we might have a label called \emph{`Oracle'}
+Furthermore, these labels are project-specific, because they are
+generated from the data of that project. For example we might have a label called \emph{`Oracle'}
 in the MySQL case, since Oracle owns MySQL. 
@@
 \begin{itemize}
-\item  MySQL 3.23. Started in 1994 and MySQL 3.23 was released in early 2001. MySQL contains $320,000$ lines of C and C++ source code\footnote{generated
-using David A. Wheeler's \emph{SLOCCount}, {http://dwheeler.com/sloccount}.}.
+\item  MySQL 3.23. Started in 1994 and MySQL 3.23 was released in early 2001. MySQL contains $320,000$ lines of C and C++ source code~\footnote{generated
+using David A. Wheeler's \emph{SLOCCount},
+{http://dwheeler.com/sloccount}.}. We used MySQL 3.23 history from July
+31st, 2000 to August 9th 2004.
 \item MaxDB 7.500. Started in the late 1970s as a research project, and was later acquired by SAP. As of version 7.500, released April 2007, the project
-has over $940,000$ lines of C source code. 
-\item PostgreSQL 7.3. Started in the 1980s as a Berkeley research project. PostgreSQL 7.3 contains $306,000$ lines of C code.
+has over $940,000$ lines of C source
+code~\footnote{{http://www.sdn.sap.com/irj/sdn/maxdb}}. 
+ We used MaxDB 7.500 history from June
+29th, 2004 to June 19th 2006.
+\item PostgreSQL 7.3. Started in the 1980s as a Berkeley research
+  project~\footnote{{http://www.postgresql.org/docs/7.3/static/}}. PostgreSQL
+  7.3 contains $306,000$ lines of C code.
+ We used PostgreSQL 7.3 history from May 9th, 2002 to 
+ August 26th 2004.
+
 \end{itemize}

### 2011-09-06 ese/ese.tex
dates

@@
 using David A. Wheeler's \emph{SLOCCount},
-{http://dwheeler.com/sloccount}.}. We used MySQL 3.23 history from July
-31st, 2000 to August 9th 2004.
+{http://dwheeler.com/sloccount}.}. We used the MySQL 3.23 version control  history from July
+31st, 2000 to August 9th, 2004.
 \item MaxDB 7.500. Started in the late 1970s as a research project, and was later acquired by SAP. As of version 7.500, released April 2007, the project
@@
 code~\footnote{{http://www.sdn.sap.com/irj/sdn/maxdb}}. 
- We used MaxDB 7.500 history from June
-29th, 2004 to June 19th 2006.
+ We used the MaxDB 7.500 version control history from June
+29th, 2004 to June 19th, 2006.
 \item PostgreSQL 7.3. Started in the 1980s as a Berkeley research
@@
   7.3 contains $306,000$ lines of C code.
- We used PostgreSQL 7.3 history from May 9th, 2002 to 
- August 26th 2004.
+ We used the PostgreSQL 7.3 version control  history from May 9th, 2002 to 
+ August 26th, 2004.
 

### 2011-09-06 ese/ese.tex
edits

--- 2011-09-06 ese/ese.tex
@@
 using David A. Wheeler's \emph{SLOCCount},
-{http://dwheeler.com/sloccount}.}. We used MySQL 3.23 history from July
-31st, 2000 to August 9th 2004.
+{http://dwheeler.com/sloccount}.}. We used the MySQL 3.23 source control  history from July
+31st, 2000 to August 9th, 2004.
 \item MaxDB 7.500. Started in the late 1970s as a research project, and was later acquired by SAP. As of version 7.500, released April 2007, the project
@@
 code~\footnote{{http://www.sdn.sap.com/irj/sdn/maxdb}}. 
- We used MaxDB 7.500 history from June
-29th, 2004 to June 19th 2006.
+ We used the MaxDB 7.500 source control history from June
+29th, 2004 to June 19th, 2006.
 \item PostgreSQL 7.3. Started in the 1980s as a Berkeley research
@@
   7.3 contains $306,000$ lines of C code.
- We used PostgreSQL 7.3 history from May 9th, 2002 to 
- August 26th 2004.
+ We used the PostgreSQL 7.3 source control  history from May 9th, 2002 to 
+ August 26th, 2004.
 
@@
 To facilitate cross-project comparison, we used a taxonomy of NFRs. This taxonomy is based on the ISO quality model, ISO9126~\cite{iso9126}. 
-ISO9126 describes six high-level NFRs: maintainability, functionality,
-portability, efficiency, usability, and reliability.\footnote{While there may be lingering debate in some circles about these terms, an ISO standard
-seems like a reasonable starting point for our work.} 
+ISO9126 describes six high-level NFRs: \emph{maintainability, functionality,
+portability, efficiency, usability,} and \emph{reliability}~\footnote{While there may be lingering debate in some circles about these terms, an ISO standard
+seems like a reasonable starting point for our work.}.
 There is some debate about the terms in this model. 
@@
 inspection of annotations was used to confirm that the annotators were
-acting similarly.
+acting similarly. PostgreSQL was annotated by both annotators in order
+to evaluate inter-rater reliability.
 We looked at each period's topics, and assessed what the data ---
@@
 
-We validate classification performance using the Receiver Operating
-Characteristic area-under-curve value~\cite{Fawcett2006861},
+We validate classification performance using the area under the curve
+of Receiver Operating
+Characteristic~\cite{Fawcett2006861},
 abbreviated \emph{ROC}, and the F-measure, which is the harmonic mean of precision and recall, i.e., $2 * (P * R) / (P + R)$.

### 2011-09-06 ese/ese.tex
66% editted

@@
 inspection of annotations was used to confirm that the annotators were
-acting similarly. PostgreSQL was annotated by both annotators in order
+acting similarly. PostgreSQL (PgSQL) was annotated by both annotators in order
 to evaluate inter-rater reliability.
@@
 WordNet~\cite{Fellbaum1998}.
-We then added Boehm's software quality model~\cite{Boehm+:1976:ICSE}, and classified his eleven `ilities' into their respective ISO9126 NFRs. 
+We then added Boehm's software quality model~\cite{Boehm+:1976:ICSE}, and classified his eleven `\emph{ilities}' into their respective ISO9126 NFRs. 
 We did the same for the quality model produced by McCall et al.~\cite{mccall1977}. 
@@
 It is not clear the words associated with our labels in \textsf{exp3} are specific enough. For example, the label \emph{maintainability} is associated with
-words \emph{ease} and \emph{ownership}. In general, as we proceed from word-list in \textsf{exp1} to that in \textsf{exp3}, our lists become more generic.
+words \emph{ease} and \emph{ownership}. In general, as we proceed from word-lists in \textsf{exp1} to that in \textsf{exp3}, our lists become more generic.
 
@@
 were
-\emph{correctness} (182/305/748 topics) and \emph{testability} (121/238/625). 
+\emph{correctness} (182/305/748 -- MySQL, MaxDB and Postgresql topics) and \emph{testability} (121/238/625). 
 We did not see many results for \emph{usability} (4/0/138) or
@@
 \emph{verifiability} or \emph{correctness}
-with \emph{traceability}, and \emph{testability} with \emph{correctness} (76 and 62 matches, respectively).
+with \emph{traceability}, and \emph{testability} with \emph{correctness}.% (76 and 62 matches, respectively).
 % See the following sections for our error analysis.
@@
 MaxDB had an average precision and recall of $0.16$ and $0.67$
-while MySQL had $0.3$ and $0.48$. PostgreSQL had precision and recall of $0.27$ and $0.95$, respectively.
+while MySQL had $0.3$ and $0.48$. PostgreSQL had average precision and recall of $0.27$ and $0.95$, respectively.
 
@@
 while \emph{functionality} remained poor. \emph{Functionality} improved dramatically by \textsf{exp3}.
-Many ROC scores were $0.6$ or less, but our classifier still performed substantially better than random.
+Many ROC scores were $0.6$ or less, but our classifier, in most cases, still performed substantially better than random.
 
@@
 The best learner is important because one uses a single learner per
-label. When applying our technique, for each NFR one should select the best learner possible.
+label. When applying our technique, for each NFR one should select the
+best performing learner possible.
 %Figure \ref{fig:best-learn-per-tag} uses the ZeroR learner as a baseline, a common practice in machine learning,
@@
 getting too mixed and it was becoming hard to distinguish one topic
-from the next. PostgreSQL's dataset was the largest of the 3 projects,
+from the next. PostgreSQL's dataset was the largest of the 3 projects:
 the XML data-dump that described the PostgreSQL topics that we
-annotated was 8X larger than MySQL and 1.7X larger than MaxDB. These
+annotated was 8 times larger than MySQL and 1.7 times larger than MaxDB. These
 size differences arise from the number of commits, the number of files
@@
 One beneficial aspect of this result is that it suggests we can have
-very fast training and classifying  since training or classifying one
-instance of Naive Bayes can be calculated in $O(N)$
+very fast training and classifying  since training on or classifying one
+instance with Naive Bayes can be calculated in $O(N)$
 for $N$ features.
@@
 \textsf{exp3}, because the ROC values are sufficiently higher in most
-cases (other than MaxDB \emph{reliability}, MySQL \emph{efficiency}, and PgSQL \emph{maintainability}). The
+cases (other than MaxDB \emph{reliability}, MySQL \emph{efficiency}, and PostgreSQL \emph{maintainability}). The
 limitation of the approach we took here is that we assume labels are
@@
 
-Two perspectives to evaluate multi-label learners are with micro or macro measurements (used in Figure \ref{fig:mulan}).
+Two perspectives used to evaluate multi-label learners are with micro or macro measurements (used in Figure \ref{fig:mulan}).
 Macro measurements are aggregated at a class or label level (per class) while micro measurements are at the element level (per element).
@@
 product. In this figure, we show the relative differences when we use
-different training data sets. In \ref{fig:pgcomp-a} we see that ROC
+different training data sets from different annotators. In \ref{fig:pgcomp-a} we see that ROC
 results are very similar. In the other figure, however, author \#1 has
@@
 performed better. 
-The difference in Macro-ROC was not signifcant, but the difference in
-Micro-ROC was as the p-value of Student's T-test was $< 0.001$. 
+The difference in Macro-ROC was not significant, but the difference in
+Micro-ROC was, as the p-value of Student's T-test was $< 0.001$. 
 %\DONE{check this writeup of comparison. T-test useful}

### 2011-09-06 ese/ese.tex
some editting

@@
 
-In PostgreSQL, by comparison, the frequencies seem to by cyclic, since we are not studying a maintenance-phase for the product, but rather ongoing
+In PostgreSQL, by comparison, the frequencies seem to be cyclic, since we are not studying a maintenance-phase for the product, but rather ongoing
 feature addition and usability improvements.
@@
 \subsection{Do Projects Differ in their Relative  Topic Interest?}
-Yes. MySQL 3.23 had proportionally more
+\emph{Yes}. MySQL 3.23 had proportionally more
 topics labelled \emph{functionality}, while MaxDB had proportionally more
@@
 changes and documentation changes. PostgreSQL also seemed to focus
-more on reliability later in the case than earlier like MySQL.
+more on \emph{reliability} later in the case study than earlier like MySQL.
 In all cases \emph{portability} was a constant maintenance concern and was prevalent throughout the lifetime of the projects. It may surprise developers
-how often portability arises as a concern.
+how often \emph{portability} arises as a concern.
 
@@
 We wanted to see how similar developers were to each other in terms of
-the NFRs that worked on. First we did pairwise $X^2$ (chi-squared)
+the NFRs that worked on. 
+Within this section we focused on PostgreSQL.
+First we did pairwise $X^2$ (chi-squared)
 tests and Kolmogorov-Smirnov tests for each developer pair and we
@@
 
-In Figure \ref{fig:authorcluster} we clustered developers using the
+In Figure \ref{fig:authorcluster} we clustered the top developers from
+PostgreSQL using the
 Ward method of Hierarchical clustering.  This diagram shows that
@@
 Figure \ref{fig:authorcluster} clustered the authors into $2$ clusters
-and $6$ clusters. We chose $6$ because the difference in width between
-entities and centroids was minimal at $6$ and the derivative of was
-the highest in that region as well.
+and $6$ clusters. We chose $6$ because the difference in distance
+between entities and centroids was minimal at $6$ clusters, and the
+derivative of distance was the highest in that region as well.
 
@@
 the mailinglist and the developers are often pushing integrations. If
-we compare the total to each author we find that 1 quarter of the
-authors have a similarity of $0.47$ or less. This is depicted in
-Figure \ref{fig:authorsim} where we can see that not all authors are
-as similar to the global trend as others. Of
-course, we found that number of commits had a strong correlation
-($0.59$ Pearson) between the similarity to the aggregate and the
-number of topics associated with that author.
+we compare the global NFR distribution to each author we find that 1
+quarter of the authors have a similarity of $0.47$ or less. This is
+depicted in Figure \ref{fig:authorsim} where we can see that not all
+authors are as similar to the global trend as others. Of course, we
+found that number of commits correlated with an author's similarity to the
+global NFR distribution ($0.59$ Pearson).
 
@@
 
-Given the top 15 PostgreSQL devs we found that proportionally many
-devs had a focus on NFR topics. Authors with top proporitions per NFR
+Given the top 15 PostgreSQL developers we found that proportionally many
+developers had a focus on NFR topics. Authors with top proporitions per NFR
 were:

### 2011-09-06 ese/ese.tex
figure size

@@
   \centering
- \includegraphics[width=.45\textwidth]{figures/process-model}
+ \includegraphics[height=.6\textheight]{figures/process-model}
  \caption{Research methodology process view.}
@@
   \centering
- \includegraphics[width=0.8\textwidth]{figures/unsupervised-bar}
+ \includegraphics[width=0.9\textwidth]{figures/unsupervised-bar}
  \caption{Performance, ROC values (range: $0$--$1$), of semi-unsupervised topic labelling for
@@
 \centering
-\includegraphics[width=0.45\textwidth]{figures/both-supervised}
+\includegraphics[width=0.7\textwidth]{figures/both-supervised}
 \caption[]{ROC value for the best learner per label for MaxDB, MySQL and PgSQL. Values range from $0$--$1$.  Dashed line indicates the performance of a random
@@
 \emph{Configuration management} and \emph{source control} related changes appeared often; % and sometimes there were topics dedicated to configuration
-management. 
+% management. 
 these kinds of changes are slightly related to \emph{maintainability}. 
@@
 Table \ref{tab:interr} describes the Cohen Kappa and the Spearman
-correlation of our per topic annotations for each tag. We had to
+correlation of our per topic annotations for each NFR. We had to
 evaluate inter-rater reliability this way because 1 topic could be
-tagged with more than 1 annotation.
+tagged with more than 1 NFR.
 

### 2011-09-06 ese/ese.tex
PS stuff
ensure PS compilation
finally!

@@
 \subfloat[MySQL 3.23]{
-\includegraphics[width=\textwidth]{figures/mysql-timeline.png}
+%\includegraphics[width=\textwidth]{figures/mysql-timeline.png}
+\includegraphics[width=\textwidth]{figures/mysql-timeline}
 \label{fig:mysql-timeline}				 
@@
 \subfloat[MaxDB 7.500]{					
-\includegraphics[width=\textwidth]{figures/maxdb-timeline.png}
+%\includegraphics[width=\textwidth]{figures/maxdb-timeline.png}
+\includegraphics[width=\textwidth]{figures/maxdb-timeline}
 \label{fig:maxdb-timeline}
@@
 \textsf{exp2} for MySQL and MaxDB, yet for PostgreSQL the
-\textsf{exp2} wordlists performed better.
+\textsf{exp2} word-lists performed better.
 For both MySQL and MaxDB \textit{usability} was better served by \textsf{exp2}. 
@@
 With ROC values ranging from $0.6$ to $0.8$ for MySQL and MaxDB and
-$0.47$ to $0.6$ for PostgreSQL ,we can see there is promise in these methods.
+$0.47$ to $0.6$ for PostgreSQL, we can see there is promise in these methods.
 \textsf{exp2} and \textsf{exp3} both indicate that static information can be used to help label topics without any training whatsoever. 
@@
 % can always be cut down
-%Our work faced multiple threats to validity and we attempted to address many of them.
+
+Our work faced multiple threats to validity and we attempted to address many of them:
+
 \emph{Construct validity} -- we used only commit messages rather than
@@
 %We used mailing-lists to verify the purpose behind some events.
-Our taxonomy for software NFRs is subject to dispute, but seems
+Our taxonomy for software NFRs is subject to dispute, but seems to be
 generally accepted. Finally, there are exogenous sources, such as
@@
 on the appropriate annotation for that topic. We verified IRR using
-manually annotated PostgreSQL topics, and achieved a result of $0.1$.
+manually annotated PostgreSQL topics, and achieved a weak result of $0.1$.
 Some of the reasons for this are a disagreement over the appropriate classification for documentation topics (should this be considered
@@
 \emph{Reliability} -- each annotator, the first two authors, followed the same protocol and used the same annotations. 
-However, only two annotators were used; their annotations could be
-biased as we did not analyze for inter-rater reliability because they
-did not rate the same documents.
+However, only two annotators were used; their annotations exhibit some
+bias as suggested by the weak inter-rater reliability for
+PostgreSQL. Inter-rater reliability could not be checked for MySQL and
+MaxDB because annotators did not rate the same documents.
 
@@
 We want to leverage different kinds of artifacts to discover threads of NFR-related discussions that occur between multiple kinds of artifacts.
-Finally, we would like to extend this analysis to other domains, to see what patterns might occur in, for example, a consumer-facing software product.
+Finally, we would like to extend this analysis to other domains, to
+see what patterns might occur within those domains, such as consumer-facing software products.
 

### 2011-09-06 ese/ese.tex
spell checking

@@
 %160 words
-\keywords{ Software maintenance \and Repository mining \and Latent dirichlet allocation \and Topic models}
+\keywords{ Software maintenance \and Repository mining \and Latent Dirichlet allocation \and Topic models}
 
@@
 requirements.
-All experiments were run on the datasets for each project (e.g., PostgreSQL, MySQL, MaxDB). LDA
+All experiments were run on the data-sets for each project (e.g., PostgreSQL, MySQL, MaxDB). LDA
 extracted $20$ topics per period for each project.
@@
 were
-\emph{correctness} (182/305/748 -- MySQL, MaxDB and Postgresql topics) and \emph{testability} (121/238/625). 
+\emph{correctness} (182/305/748 -- MySQL, MaxDB and PostgreSQL topics) and \emph{testability} (121/238/625). 
 We did not see many results for \emph{usability} (4/0/138) or
@@
 getting too mixed and it was becoming hard to distinguish one topic
-from the next. PostgreSQL's dataset was the largest of the 3 projects:
+from the next. PostgreSQL's data-set was the largest of the 3 projects:
 the XML data-dump that described the PostgreSQL topics that we
@@
 
-By contrast, the PostgreSQL timeline  (Figure \ref{fig:pgsql-timeline}) is extracted from a central trunk and is not version-specific. Therefore development 
+By contrast, the PostgreSQL time-line  (Figure \ref{fig:pgsql-timeline}) is extracted from a central trunk and is not version-specific. Therefore development 
 tends to be focused on releases 7.3, 7.4 and 7.5/8.0 alpha. For example, a priority for the 8.0 candidate was a Windows-native port of the source code, which seems to 
@@
 Given the top 15 PostgreSQL developers we found that proportionally many
-developers had a focus on NFR topics. Authors with top proporitions per NFR
+developers had a focus on NFR topics. Authors with top proportions per NFR
 were:

### 2011-09-06 src/validate/arffdedup.pl
a missing file

@@ -1,0 +1,8 @@
+#!/usr/bin/perl
+use strict;
+use ARFF;
+my $filename = $ARGV[0];
+warn "$filename -> $filename";
+my $arff = ARFF->new( filename => $filename );
+$arff->drop_duplicate_columns();
+$arff->writeout( $filename );

### 2012-01-15 ese/ese.tex
- [X] Talk about Treude's concernlines      In the related work part you mention ConcernLines by Truede. Could you      be more precise and specify where the tags come from (source code, cvs      logs, ...)

@@
 They mined
-developer created tags 
-%change request tags (such as `milestone 3') and used these 
+developer created 
+change request tags from IBM Jazz
+repositories and used these 
 in order to analyze the evolution of a single product.
+Change requests in Jazz allow for users to annotate each change
+request and check-in with tags such as ``ui'', ``usability'' or
+``milestone 3''. 
 % make evolutionary analyses of a single product. 

### 2012-01-15 ese/ese.tex
- [X] Address single domain      - On page 5 you mention that you chose the same application domain to        control for differences in functional requirements. While I do indeed        see that the functional requirements of each of the 3 database systems        would be similar, it might very well be that they are not        identical. Say for example that there is an import functionality in        MySQL, which is absent from the other 2. That would create an extra        functional requirement. Could you discuss this further in the paper        what the benefits and consequences are of you

@@
 With ConcernLines, Treude et al.~\cite{treude09cl} showed tag occurrence
-using colour and intensity. 
+using colour and intensity, and our plots that rely on color and intensity
+have a similar look and feel to Treude et al.'s plots.
 They mined
@@
 We explicitly chose older versions of mature projects from a stable problem domain to increase the likelihood that we would encounter primarily
-maintenance activities in our studies.
+maintenance activities in our studies. We felt that a single domain
+would allow for cross-project comparison. At the same time we
+recognize that a problem-domain alone does guarantee functional and
+non-functional similarity.
+For instance each database system has a different focus, PostgreSQL
+tends to focus on fulfilling much of the SQL92 specification and adding
+more features while MySQL has been slow to adopt much of the spec.
+ As a consequence our choice of looking at a
+single domain limits our generalizability. What we show in this paper
+will be somewhat biased towards database software.
 
@@
 of developer comments. 
-%The OSS projects investigated were database systems, thus 
-%our results might not generalize to other domains. %We only studied database systems.
+While we tried to control for application domain variability, OSS
+projects investigated were database systems, thus our results might
+not generalize well to other domains.
 

### 2012-01-15 ese/ese.tex
- [X] Most commits we observed had commit comments. Can you be more precise      here and quantify?          - MaxDB 7.500          -    Min. 1st Qu.  Median    Mean 3rd Qu.    Max.               5.0    83.0    94.0   111.5   138.0  1149.0          - summary(v$V1)            Min. 1st Qu.  Median    Mean 3rd Qu.    Max.            5.0    83.0    94.0   111.5   138.0  1149.0          - summary(v$V2)            Min. 1st Qu.  Median    Mean 3rd Qu.    Max.            1.000   6.000   8.000   7.872   8.000 110.000          - summary(mysql$V1)            Min. 1st Qu.  Median    Mean 3rd Qu.    Max.

@@
 repository. 
-Most commits we observed had commit comments.
+Most commits we observed had commit comments (90\% in MySQL 3.23,
+98.5\% in PostgreSQL and 99.99\% in MaxDB 7.500).
 Commit comments are often studied by researchers, as they are the most readily accessible source of project interactions, and developers are often
@@
 A topic analysis tool such as LDA will try to find $N$ independent
-word distributions within the word distributions of all input messages. 
+word distributions within the word distributions of all input
+messages. 
+If there are not $N$ independent word distributions, the topics
+produce tend to be duplicates of each other, that is they share the
+top terms. During this study we found that when $N$ was 20 we did not
+witness duplicate topics very often.
 Linear combinations of these $N$ word distributions are meant to represent and recreate the word distributions of any of the original messages. 

### 2012-01-16 ese/ese.tex
- [X] On page 9 you mention that you put the term "redundancy" under the          flag of reliability. This is quite possible in the context of RDBMS,          but... it could also be used to indicate code cloning (code          redundancy). How did you cope with this double meaning?          Threat to validity and IRR issues

@@
 we considered to be
-relevant to discussion of \emph{reliability}, but is not in the standard. 
+relevant to discussion of \emph{reliability}, but is not in the
+standard. We recognize though that terms like this might be used in a
+different context with a different meaning, like code-cloning.
 We therefore took the NFRs from the ISO9126 and added terms to them.
@@
 
+Our word-lists were built up of words that were assumed to be relevant
+to those topics, our automated analysis was ignorant of the multi-uses
+of words and thus topics could be flagged inappropriately. For
+instance, if the word \emph{redundancy} is used would it reference
+reliability or cloned code? This issue is why we checked the
+performance of the techniques, although we did not explicitly check
+for these cases.
+
 \emph{Internal validity} -- %includes inter-rater reliability and our lack of rival explanation building. 

### 2012-01-16 ese/ese.tex
- [X] Defend ROC (class sizes)

@@
 However, ISO9126 is ``an international standard and thus provides an 
-internationally accepted terminology for software quality \cite[p. 58]{Boegh2008},'' that is sufficient for the purposes of this research.  
+internationally accepted terminology for software quality
+\cite[p. 58]{Boegh2008},'' that is sufficient for the purposes of this
+research.  
+\emph{Performance} is an example of RDBMS word related to the \emph{efficiency} NFR.
 We claim that these NFRs are maintenance concerns (to varying degrees) in all software projects, and are therefore well-suited for comparisons between
@@
 Characteristic~\cite{Fawcett2006861},
-abbreviated \emph{ROC}, and the F-measure, which is the harmonic mean of precision and recall, i.e., $2 * (P * R) / (P + R)$.
+abbreviated \emph{ROC}, sometimes called \emph{AUC}, and the F-measure, which is the harmonic mean of precision and recall, i.e., $2 * (P * R) / (P + R)$.
 
@@
 negative instance.
+
+We argue for using ROC over F-Measure because ROC suffers less from
+bias than F-Measure and as often F-Measure skews towards the positive
+class~\cite{flach-icml03,Forman:2010:ACS:1882471.1882479}, especially
+in the case of class imbalance. Forman et
+al.~\cite{Forman:2010:ACS:1882471.1882479} demonstrate that cross-fold
+validation will often produce low F-measures given high class
+imbalance and the presence of false negatives.
+Although
+recent work~\cite{Forman:2010:ACS:1882471.1882479} has suggested that while ROC suffers less bias than
+the average F-measure ($F_{avg}$) for cross-folds, F-measures ($F_{tp,fp}$) that are calculated from
+the sums of true-positives,
+false-positive, and false-negatives across folds tend to exhibit no
+bias. Unfortunately our experimental framework lacks a robust way to
+calculate $F_{tp,fp}$, thus we still provide ROC and $F_{avg}$.
 We consider our labelling classifiers acceptable if they outperform a random classifier (0.5).
@@
 Similarity in WordNet means siblings in a hypernym tree. 
-We do not include these words here for space considerations (but see the Appendix for our data repository). 
+We do not include these words here for space
+considerations~\footenote{For our data repository visit \url{http://softwareprocess.es/nomen/}}. 
 It is not clear the words associated with our labels in \textsf{exp3} are specific enough. For example, the label \emph{maintainability} is associated with
@@
 Using our three word-lists (\textsf{exp1}, \textsf{exp2}, \textsf{exp3}), we labelled our topics with an NFR where there was a match between a word in
-the list and the same word somewhere in the distribution of words that constitute the topic.
+the list and the same word somewhere in the frequency distribution of words that constitute the topic.
 A \emph{named topic} is a topic with a match. 

### 2012-01-17 ese/ese.tex
2 minor fixes

@@
 
-For the third set of word-lists, \textsf{exp3}, we extended the word-lists from \textsf{exp2} using unfiltered WordNet similarity matches. 
+For the third set of word-lists, \textsf{exp3}, we extended the word-lists from \textsf{exp2} using WordNet similarity matches. 
 Similarity in WordNet means siblings in a hypernym tree. 
 We do not include these words here for space
-considerations~\footenote{For our data repository visit \url{http://softwareprocess.es/nomen/}}. 
-It is not clear the words associated with our labels in \textsf{exp3} are specific enough. For example, the label \emph{maintainability} is associated with
-words \emph{ease} and \emph{ownership}. In general, as we proceed from word-lists in \textsf{exp1} to that in \textsf{exp3}, our lists become more generic.
+considerations~\footnote{For our data repository visit \url{http://softwareprocess.es/nomen/}}. 
+Wordnet similarity is a very broad match. For example, the label \emph{maintainability} is associated with
+words \emph{ease} and \emph{ownership}, and the word \emph{performance} has a `sense' which refers to musical performances -- which is not related to software development.
+In general, as we proceed from word-lists in \textsf{exp1} to that in \textsf{exp3}, our lists become more generic.
 
@@
 
-Table \ref{tbl:wordlist} shows how many topics were labelled for MaxDB
-and MySQL.
+Table \ref{tbl:wordlist} shows how many topics were labelled for MaxDB, MySQL and PostgreSQL.
 

### 2012-01-17 ese/ese.tex
affiliation update
**  5. What about developer style    5. Reviewer 4 raises the some observed results may simply be due to    the differences among developer styles and not necessarily    reflecting different types of work.    - re: We added developer style threats to the end of the section:      Do different developers work on different NFRs?    - re: We added this concern to threats to validity.    - [X] Add to threats to validity.    - [X] Warn up front about developer style    - [X] Cite large changes paper that style can be an issue

@@
   Abram Hindle \at 
-  Dept. of Computer Science\\
+  Dept. of Computing Science\\
   University of Alberta\\
@@
 of $1/3$ to $1/10$ of commits dedicated to a particular NFR.
+
+One potential confound for this analysis is that we could be
+describing developer style instead of developer focus. Developer style
+would be a developers likelihood to use or not use terms found within
+our dictionaries and training corpus. Further more a top developer, in
+terms of frequency of commits, will have more samples and more likely
+take on a more general role. We found in a previous
+paper~\cite{largechanges} that often the words used to describe a commit
+and the actual author of the commit were often correlated. Thus based
+on that observation developer style might be a confounding concern.
 
@@
 
+Developer style is another confounding issue, if we are searching for
+these words we are relying on developers to use these words. This
+study might be exploiting the behaviour or style of a few
+developers. If one developer did not describe their commits well or
+used fewer terms it is likely they would be associated with a NFR
+topic regardless of the actual purpose of their commits.
+
 \emph{Internal validity} -- %includes inter-rater reliability and our lack of rival explanation building. 

### 2012-01-17 ese/ese.tex
- [X] Defend low ROC in threats to validity          - Re: we explained some reasons for low ROC scores.    - [X] Defend ROC (class sizes)          - Re: in the "Creating a Validation Corpus" we addressed why we            used ROC over F-Measure: class imbalance leads to bias in            F-Measure. And we cited relevant work that discussed this            particular issue.    - [X] Skewed classes in threats to validity          - Re: in threats to validity, internal validity, we discussed skewed classes

@@
 We did not attempt to match our results to any particular model.
+
+We primarily relied on ROC (the area under the receiver operating
+characteristic curve), as our measurement to compare the effectiveness
+of semi-supervised and supervised learners. ROC is less biased than
+F-Measure in cross-folds
+validation~\cite{flach-icml03,Forman:2010:ACS:1882471.1882479} in the
+case of class imbalance. We
+feel that the fixed point of $0.5$ for random results allows us to
+better describe the predictability achieved by some of these
+techniques against random and ZeroR learners. 
+We also include
+F-Measure results to allow for the reader to validate against
+F-Measure if they are more comfortable with it. Thus because of
+F-Measure biased handling of class-imbalance and the inherent
+class-imbalance that our data suffers from we chose ROC.
+
+%XXX If we have more time try to elaborate more on this
+One potential issue with some of our results is that we achieved very
+low ROC and F-Measure scores.  Performance was not uniformly low,
+often the more common classes exhibited better performance than the
+more rare classes.  Often this was amplified by class imbalance. In
+future work we plan to investigate techniques such as sub-sampling and
+boot-strapping in order to improve performance. Some of the ROC
+performance could be due to a lack of coherence in tagging which was
+shown by our low inter-rater reliability score.
+
 We assessed inter-rater reliability (IRR), a measure of the correlation between the manual annotations the first two authors created. Perfect IRR (1.0) indicates complete agreement 

### 2012-01-17 ese/ese.tex
unsure
- [X] In intro provide more details    - [X] In Previous work and intro add more about the comparison          between the two papers          I would actually think that it is appropriate that you mention your          previous (MSR) paper somewhere in the introduction and briefly explain          in 1 sentence how this paper extends the previous work.

@@
  %Section \ref{sec:unsuplabelling} presents our technique for using word-lists to do unsupervised classification, and describes the results.
+
+This work extends our previous work~\cite{msr2011}, published at The
+Working Conference on Mining Software Repositories 2011
+(MSR'2011). 
+Within this new paper that builds up upon the previous
+work, we added more machine learning results and another case-study: PostgreSQL. 
+We used two authors to
+each annotate the same PostgreSQL topics in order to compare
+them. 
+Thus we used this new case-study to test inter-rater
+reliability, described in Section \ref{sec:inter-rater}. 
+In Section
+\ref{sec:developers}, we further extended the MSR'2011 paper with an
+analysis of PostgreSQL authors and their association with NFRs and
+topics.
+
 In Section \ref{sec:suplabelling}, we manually annotate the topics, and use those annotations as training data for supervised classification.  
@@
 \subsection{Do different developers work on different NFRs?}
+
+\label{sec:developers}
 

### 2012-01-17 ese/ese.tex
moved a paragraph
1. The reviewers would like to see discussion of the relationship of    the current draft and the previous MSR 2011 paper and more detailed    discussion of previous papers related to the current draft.    - [X] In intro provide more details    - [ ] In Previous work provide MORE details

@@
 
-Finally, in Ernst et al.~\cite{ernst10refsq}, we describe our earlier
-project, similar to this, that identifies changes in quality requirements
-in GNOME software projects; this approach was more exploratory, had
-less validation, uses different word-lists, solely uses text-matching,
-and does not leverage machine learning strategies. 
-Hindle et al.~\cite{Hindle09ICSM} propose a windowed method of topic
-analysis that we extend with labelled
-topics, NFRs and new visualizations.
-
 %Massey~\cite{massey02icse} and Scacchi~\cite{scacchi02,scacchi05b} looked at the topic of requirements in open-source software. 
@@
 % techniques. This supports our result that software qualities are not discussed with the same frequency across projects.
+
+Finally, 
+this paper and our MSR 2011 paper~\cite{msr2011}, are based off of the
+work of Ernst
+et al.~\cite{ernst10refsq} and the work of Hindle et
+al.~\cite{Hindle09ICSM}.
+In Ernst et al.~\cite{ernst10refsq}, we describe our earlier
+project, similar to this, that identifies changes in quality requirements
+in GNOME software projects; this approach was more exploratory, had
+less validation, uses different word-lists, solely uses text-matching,
+and does not leverage machine learning strategies. 
+This work, and our
+MSR'2011 paper~\cite{msr2011}, extend this work by using 
+word-lists to label extracted topics.
+ Hindle et
+al.~\cite{Hindle09ICSM}
+ propose a windowed method of topic
+analysis that we extend with labelled
+topics, NFRs and new visualizations. This windowed method was
+to bucket documents  by time windows (such as months), that could overlap if
+necessary, and then use LDA to extract documents from these
+bucket. These topics were then related to topics in other adjacent
+buckets, and joined if deemed sufficiently similar. This work, and our
+MSR'2011 paper~\cite{msr2011}, extend this method by labelling topics
+and providing an alternative visualization, the topic time-line
+described in Section \ref{sec:analysis}.
 
@@
 %\subsection{Timeline analysis}
-Figures \ref{fig:mysql-timeline} and \ref{fig:maxdb-timeline} and \ref{fig:pgsql-timeline} show the
+Topic time-lines are depicted in Figures \ref{fig:mysql-timeline} and
+\ref{fig:maxdb-timeline} and \ref{fig:pgsql-timeline}. These topic time-lines show the
 temporal patterns of NFR frequencies.

### 2012-01-17 ese/ese.tex
4. Reviewer 3 raises that the extremely low IRR (~0.1) threatens the    validity of the results of these experiments, and its threats to    validity should be made more prominent.    - [X] Threats to validity updated for low IRR    - [X] Suggestions on improvement

@@
 option a usability issue? These kinds of questions illustrate some of
-the agreement, disagreement and ambiguity about these labels. Furthermore, 
+the agreement, disagreement and ambiguity about these labels. 
+Thus we would recommend that future annotators make a concerted effort
+to train and discuss when one applies an annotation.
+Furthermore, 
 both annotators found that the annotation was time-consuming and
 difficult.
+
+We feel that this resulting IRR score, although lackluster, indicates
+the underlying difficulty of annotating such a data-set.
+Consistent training of the annotators (the people) might avoid low IRR scores, or
+using more annotators.
+Other improvements could be gained by approaching some of the original developers
+and using their expertise to annotate the data according to their intent.
+Thus improvements
+could be gained by producing more robust training samples that use
+multiple raters, or experts, who are similarly trained and willing to discuss and
+negotiate disagreements.
 
@@
 Some of the reasons for this are a disagreement over the appropriate classification for documentation topics (should this be considered
-usability or portability?). A short education session on what a given NFR from the ISO standard encompasses would reduce this error rate.
+usability or portability?). A short education session on what a given
+NFR from the ISO standard encompasses would reduce this error rate.
+Weak IRR results could also be responsible for some the low ROC
+performance, if each rater is picking up on different information it
+might make it more difficult to train a machine learner to do the
+same.
+Furthermore weak IRR results threatens the validity of the
+experiments, if raters cannot agree then consistently labelling by
+ISO-9126 related NFRs could be quite difficult.
 
@@
 %area under the 
-ROC values between $0.6$ and $0.8$. These results, along with the efficient performance of our learners, demonstrate that labelled topic extraction is
+ROC values between $0.6$ and $0.8$. 
+These results were confounded by our low inter-rater reliability score
+on the PostgreSQL data-set, suggesting that annotators need careful
+and thoughtful training.
+With appropriate rigor in annotation, these results, along with the performance of our learners,
+demonstrate that labelled topic extraction can be
 a promising approach for understanding the occurrence of non-functional requirements in software projects. 

### 2012-01-17 ese/ese.tex
4. Reviewer 3 raises that the extremely low IRR (~0.1) threatens the    validity of the results of these experiments, and its threats to    validity should be made more prominent.    - [X] Threats to validity updated for low IRR    - [X] Suggestions on improvement
related work in progress

--- 2012-01-17 ese/ese.tex
@@
  %Section \ref{sec:unsuplabelling} presents our technique for using word-lists to do unsupervised classification, and describes the results.
-
-This work extends our previous work~\cite{msr2011}, published at The
-Working Conference on Mining Software Repositories 2011
-(MSR'2011). 
-Within this new paper that builds up upon the previous
-work, we added more machine learning results and another case-study: PostgreSQL. 
-We used two authors to
-each annotate the same PostgreSQL topics in order to compare
-them. 
-Thus we used this new case-study to test inter-rater
-reliability, described in Section \ref{sec:inter-rater}. 
-In Section
-\ref{sec:developers}, we further extended the MSR'2011 paper with an
-analysis of PostgreSQL authors and their association with NFRs and
-topics.
-
 In Section \ref{sec:suplabelling}, we manually annotate the topics, and use those annotations as training data for supervised classification.  
@@
 ours, as they mine keywords from NFR catalogues.  They differ from our approach because
-they mine requirements documents where as we mine revisions.
-%One approach they tried was similar to this one, with keywords mined from NFR catalogues found in their previous work~\cite{chung99}. 
+they mine requirements documents whereas we mine revisions.
+One approach they tried was similar to this one, with keywords mined from NFR catalogues found in their previous work~\cite{chung99}. 
 Their approach resulted in a recall of $80\%$ with precision of $57\%$ for the \emph{security} NFR, but could not find a reliable source of keywords for other NFRs. 
@@
 on a single project.
-%They relied on requirements documents instead of version control
-%histories that we use.
-%The objective of Cleland-Huang's study was to identify new NFRs for system development, 
-%yet our objective was to recover those latent NFRs from commit-log messages of the project.
+They relied on requirements documents instead of version control histories that we use.
+The objective of Cleland-Huang's study was to identify new NFRs for system development, 
+yet our objective was to recover those latent NFRs from commit-log messages of the project.
 
@@
 \subsection{Do different developers work on different NFRs?}
-
-\label{sec:developers}
 

### 2012-01-17 ese/ese.tex
Read through previous work section, some changes.

@@
 documents for non-functional requirements (NFR) (software
-qualities)~\cite{Cleland-Huang2006}.  Their approach is similar to
-ours, as they mine keywords from NFR catalogues.  They differ from our approach because
-they mine requirements documents whereas we mine revisions.
-One approach they tried was similar to this one, with keywords mined from NFR catalogues found in their previous work~\cite{chung99}. 
+qualities)~\cite{Cleland-Huang2006}.  
+%Their approach is similar to
+%ours, as they mine keywords from NFR catalogues.  They differ from our approach because
+%they mine requirements documents whereas we mine revisions.
+One approach they tried was similar to this one, with keywords mined from NFR catalogues found in Chung et al.~\cite{corollary99}. 
 Their approach resulted in a recall of $80\%$ with precision of $57\%$ for the \emph{security} NFR, but could not find a reliable source of keywords for other NFRs. 
@@
 on a single project.
-They relied on requirements documents instead of version control histories that we use.
+They relied on relatively well-structured requirements documents instead of version control histories that we use.
 The objective of Cleland-Huang's study was to identify new NFRs for system development, 
@@
 %Some results were interesting, but their precision was quite low.
+
 With ConcernLines, Treude et al.~\cite{treude09cl} showed tag occurrence
@@
 % NE 0113 this seems too early to mention 'result'.
-% Hindle et al. \cite{Hindle2007} examined release patterns in OSS. They showed that there is a difference between projects regarding maintenance
-% techniques. This supports our result that software qualities are not discussed with the same frequency across projects.
+In Hindle et al. \cite{Hindle2007} we examined release patterns in OSS. That work showed that there is a difference between projects regarding maintenance
+techniques. This supports the result described in this paper, that software qualities are not discussed with the same frequency across projects.
 
 Finally, 
-this paper and our MSR 2011 paper~\cite{msr2011}, are based off of the
-work of Ernst
-et al.~\cite{ernst10refsq} and the work of Hindle et
+this paper and our MSR 2011 paper~\cite{msr2011} are based off of the
+work of Ernst and Mylopoulos~\cite{ernst10refsq} and the work of Hindle et
 al.~\cite{Hindle09ICSM}.
-In Ernst et al.~\cite{ernst10refsq}, we describe our earlier
-project, similar to this, that identifies changes in quality requirements
-in GNOME software projects; this approach was more exploratory, had
+In Ernst and Mylopoulos~\cite{ernst10refsq}, we describe an earlier
+project that identifies changes in quality requirements
+in GNOME software projects (GNOME is a Linux desktop environment). This earlier approach was more exploratory, had
 less validation, uses different word-lists, solely uses text-matching,
 and does not leverage machine learning strategies.  
-This work, and our
-MSR'2011 paper~\cite{msr2011}, extend this work by using 
-word-lists to label extracted topics.
+Our approach in this paper, and our
+MSR 2011 paper~\cite{msr2011}, extends Ernst and Mylopoulos~\cite{ernst10refsq} by using 
+word-lists to label topics, which are completely absent in the earlier work.
  Hindle et
@@
 bucket. These topics were then related to topics in other adjacent
-buckets, and joined if deemed sufficiently similar. This work, and our
-MSR'2011 paper~\cite{msr2011}, extend this method by labelling topics
+buckets, and joined if deemed sufficiently similar. This paper, and our
+MSR'2011 paper~\cite{msr2011}, extend the method of Hindle et
+al.~\cite{Hindle09ICSM} by labelling topics
 and providing an alternative visualization, the topic time-line
@@
 ISO9126 describes six high-level NFRs: \emph{maintainability, functionality,
-portability, efficiency, usability,} and \emph{reliability}~\footnote{While there may be lingering debate in some circles about these terms, an ISO standard
-seems like a reasonable starting point for our work.}.
-There is some debate about the terms in this model. 
+portability, efficiency, usability,} and \emph{reliability}
+%\footnote{While there may be lingering debate in some circles about these terms, an ISO standard
+%seems like a reasonable starting point for our work.}.
+There is some debate about the terms in this model, and whether they are a) the correct terms and b) correctly organized. 
 However, ISO9126 is ``an international standard and thus provides an 

### 2012-01-17 ese/ese.tex
KDE reffed

@@
 We did the same for the quality model produced by McCall et al.~\cite{mccall1977}. 
-We then did a simple random analysis of mailing list messages from an open source project, KDE. If we judged a given message to contain terms that were
+We then did a simple random analysis of mailing list messages from an open source ecosystem, KDE.
+Like the GNOME study we conducted in Ernst and Mylopoulos \cite{ernst10refsq}, KDE contains a suite of different products covering a variety of software categories.
+If we judged a given message to contain terms that were
 related to one of the NFRs in ISO9126, we added it to our word-list. This allowed us to expand our word-lists with more software-specific terms.

### 2012-01-17 ese/ese.tex
Updated exp1-3 discussion on usability/accuracy

@@
 
-Our first word-list set, \textsf{exp1}, was generated using the ontology described in Kayed et al.~\cite{5072519} .
+Our first word-list set, \textsf{exp1}, was generated using the ontology described in Kayed et al.~\cite{5072519}
 That paper constructs an ontology for software quality measurement using eighty source documents, including research papers and international standards. 
@@
 We did not see many results for \emph{usability} (4/0/138) or
-\emph{accuracy} (3/0/27), which were associated with fewer than ten topics. 
+\emph{accuracy} (3/0/27), which were infrequently matched.
+Note the significantly higher result for \emph{usability} in PostgreSQL -- this suggests a difference in how this project is discussing usability, at least with respect to our analysis.
 We also looked for correlations between our labels: excluding double
@@
 As we mentioned, the word-lists are broad, so there are likely to be false-positives (discussed below). 
-The most frequent label across all projects was \emph{usability}, and the
-least frequent label was \emph{maintainability}. 
+The most frequent label across all projects (for this wordlist, and unlike \textsf{exp1}) was \emph{usability}, and the
+least frequent label was \emph{maintainability}. This implies that our signifiers for usability in this experiment were fairly broad.
 Common co-occurrences were \emph{reliability} with \emph{usability}, \emph{efficiency} with \emph{reliability}, and \emph{efficiency} with \emph{usability}.

### 2012-01-17 src/validate/interrater.R
Trying to do interrater

@@ -18,11 +18,13 @@ kappa2(matrix(c( 1*(runif(600)>.9), 1*(runif(600)>.9)), ncol=2))
 kappa2(matrix(c( 1*(runif(600)>.99), 1*(runif(600)>.99)), ncol=2))
 kappa2(matrix(c( (runif(600)>.5), (runif(600)>.5)), ncol=2))
 
+normit <- function(x) { x*4 + 1 } 
+
 weight <- "unweighted"
 #for (weight in c("unweighted","equal","squared")) {
   for (i in c(1:7)) {
     print(paste(names(pa)[[i]],weight))
-    mat <- matrix(c( pa[,i], pn[,i] ), ncol=2)
+    mat <- matrix(c( normit(pa[,i]), normit(pn[,i]) ), ncol=2)
     print(kappa2(mat, weight=weight))
     print(kappam.fleiss(mat))
     #print(cor(pa[,i],pn[,i]))
@@ -61,3 +63,70 @@ kappa2(mat)$value
     print(paste("Pos Agree: ",length(pa[pa[,i]+pn[,i]==2,i])))
     print(paste("Neg Agree: ",length(pa[pa[,i]+pn[,i]==0,i])))
   }
+
+
+
+sapply(names(pa),function(i) {
+  mat <- matrix(c( normit(pa[,i]), normit(pn[,i]) ), ncol=2)
+  kappa2(mat)$value
+})
+
+kappaof <- function(sai,pai) {
+    mat <- matrix(c( sai, pai ), ncol=2)
+    kappa2(mat)$value
+}
+
+# repeat N times
+# sample from the distribution itself
+# and see what the average Kappa score is per class
+paunion <- pa
+for (i in names(pa)) {
+  paunion[,i] <- (pa[,i] & pn[,i]) * 1
+}
+kappaofself <- function( pan, N) {
+  sapply(names(pan),
+         function(x){
+           mean(sapply(c(1:N),
+                       function(x){
+                         sapply(names(pan),function(i){
+                           kappaof(sample(pan[,i],630,replace=T),pan[,i])})})[x,])})
+}
+# Abram annotations 1000 times
+                                        #    A_portability   A_functionality     A_reliability A_maintainability 
+#     0.0084965406      0.0062652051     -0.0055896254     -0.0008711916 
+#     A_efficiency       A_usability            A_none 
+#    -0.0021828228      0.0031867318     -0.0008958907 
+
+# kappaofself( pn, 1000)
+#  A_functionality     A_portability       A_usability     A_reliability 
[…]

### 2012-01-17 src/validate/interrater.R
more interrater

@@ -126,7 +126,7 @@ kappaofunif <- function( pan, N, threshold) {
 #     -0.001374283      -0.001031894      -0.001149971
 cc <- c(0,0.01,0.05,0.1,0.2,0.3,0.4,0.5,0.6,0.7,0.8,0.9,0.95,0.99,1.0)
 for (i in cc) {
-  kappaofunit(pa,1000,i)
-  kappaofunit(pn,1000,i)
-  kappaofunit(paunion,1000,i)
+  kappaofunif(pa,1000,i)
+  kappaofunif(pn,1000,i)
+  kappaofunif(paunion,1000,i)
 }

### 2012-01-17 ese/ese.tex
minor fixes for supervised

--- 2012-01-17 ese/ese.tex
@@
 %we compared the results of the unsupervised word matching approach. 
-For each quality we assessed whether semi-unsupervised labels matched the manual annotations. 
+For each quality in the high-level ISO 9126 taxonomy (namely, \emph{Maintainability, Usability, Reliability, Efficiency, Portability, Functionality}) we assessed whether the semi-unsupervised labels for a topic matched the manual annotations we created for the validation corpus.
 To be clear, the manual annotations were not used to
@@
 performance via the overlap between ISO9126 and the Kayed ontology.
-For \textsf{exp1} the F-measures for MaxDB were from $0$ to $0.18$ with an average
+For \textsf{exp1} the F-measures for the NFRs for MaxDB were from $0$ to $0.18$ with an average (of all NFRs)
 of $0.03$, for MySQL were from $0$ to $0.16$ with an average of
@@
 Traditional classifiers label topics with a single class, whereas Mulan allows for a mixture of classes per topic, which is what we observed while
-manually labelling topics.
+manually labelling topics. For example, a given topic (word distribution) may be `about' both usability and maintainability, if this topic was a product of a discussion on design tradeoffs.
 The \emph{features} we used are word counts/occurrence	per topic, if

### 2012-01-17 ese/ese.tex
R2 fixes in section 4.3

@@
 correlate with the \emph{portability} NFR increasing in frequency in mid-2004. Before the 7.4 release, \emph{usability,functionality} and \emph{portability} all 
-increase in frequency, possibly reflecting the interest in adding features and documentation for the release.
-
-In the following sections we turn to our research questions:
+increase in frequency, possibly reflecting the interest in adding features and documentation for the release. In the following sections we turn to our research questions:
 
@@
 We wanted to see how similar developers were to each other in terms of
-the NFRs that worked on. 
+the NFRs that they worked on. 
 Within this section we focused on PostgreSQL.
-First we did pairwise $X^2$ (chi-squared)
-tests and Kolmogorov-Smirnov tests for each developer pair and we
-found that $3/10$ of these tests the developers were significantly
-different, that is their distributions were not similar. This implies
+We first conducted pairwise $X^2$ (chi-squared)
+tests and Kolmogorov-Smirnov tests for each developer pair. We
+found that in $3/10$ of these tests, the developers were significantly
+different, i.e., their distributions were not similar. This implies
 that there are authors who focus on different NFRs or different
@@
 the mailinglist and the developers are often pushing integrations. If
-we compare the global NFR distribution to each author we find that 1
-quarter of the authors have a similarity of $0.47$ or less. This is
-depicted in Figure \ref{fig:authorsim} where we can see that not all
+we compare the global NFR distribution to each author we find that 25\%
+ of the authors have a similarity of $0.47$ or less. This is
+depicted in Figure \ref{fig:authorsim}, where we can see that not all
 authors are as similar to the global trend as others. Of course, we
@@
 and \texttt{ishii} and \texttt{scrappy} for maintainability.
-Given these results we found that yes indeed there was often a focus
+Given these results we found that there indeed was a focus
 of $1/3$ to $1/10$ of commits dedicated to a particular NFR.

### 2012-01-17 ese/ese.tex
Changes to intro per comments. Removed commented out text to enhance readability.

@@
 using  Latent Dirichlet Allocation (LDA) from commit-log comments recovered
-from source control systems such as CVS and BitKeeper.  
+from source control systems. % such as CVS and BitKeeper.  
 These topics are given labels from a generalizable
@@
 \section{Introduction}
-A key problem for practising software maintainers is gaining an
-understanding of \emph{why} a system has evolved the way it has. 
+A key problem for practicing software maintainers is gaining an
+understanding of \emph{why} a system has evolved the way it has \cite{Mockus00}. 
 This is different from \emph{how} a system has evolved. 
@@
 In this work we seek to provide a method of automatically labelling  development topics extracted from commit logs.
-
-Topic modeling is a machine learning technique that creates
+Concrete applications of topics and \emph{labelled topic extraction}, which this paper introduces,
+range from project dashboards to annotating software artifacts such as revisions and bug reports with
+NFR-related tags.
+Project dashboards~\cite{dashboard} are typically employed by managers and are used to
+provide quick summaries of the effort put into a software project. In this case, 
+labelled topic extraction would allow managers to track effort
+related to NFR topics, such as \emph{portability}.
+These techniques also allow for the annotation of commit comments
+ and other software artifacts with NFRs. 
+This would enable querying of bug reports and artifacts by relevant NFRs. 
+For instance a manager can confirm if their developers were
+focused on \emph{usability} by looking for
+ \emph{usability}-relevant revisions and bug reports.
+
+Topic modeling (such as Latent Dirichlet Allocation ~\cite{Blei2003}) is a machine learning technique that creates
 multinomial distributions of words extracted from a text corpus. 
@@
 project history.
-%A major challenge for topic modeling is to accurately identify the
-%summary label for a topic; in the previous case, perhaps
-%\emph{``event-handling''} would be appropriate. 
-%Current approaches are manual: they rely on human expertise to
-%recognize the appropriate label. 
-%Furthermore, to date the results of topic modeling are project
-%specific, and not generalizable. 
-%This paper presents an approach, \emph{labelled topic extraction}, which addresses these two problems.
 
@@
 labels that are project independent.
-%is both automatic and creates labels that are project independent.
-
-% Changes to software systems, revisions, do not occur without
-% reason. 
-% The reason or purpose is often related to requirements 
-% Often one revision might have multiple purposes ranging from
-% maintenance issues, quality issues, requirements, features or a
-% response to a bug report.
-% These purposes often are exhibited as latent topics that relate to
-% many development artifacts such as:
-%  version control revisions, bug reports, and mailing-list discussions.
-
-%In our previous work we dealt with topic trends, which are topics that recur over time~\cite{Hindle09ICSM}. 
-%We observed that topic trends were often non-functional requirements (NFRs). 
-%NFRs have the property of being cross-domain and widely applicable. 
-%In this sense, they are useful abstractions for developer conversations about different software projects.
-
-% harp on data mining more
-%In general, the mining of software artifacts tends to be very project specific, yet NFRs are not. 
-% I was told if you're going to use a paren, you don't need to say it.
-%There is a series of standards on NFRs, such as ISO9126~\cite{iso9126}, specifically intended to apply to projects of varying types.
-%These standards indicate our approach is possible. %, if perhaps not universally accepted).
-% NFRs are prevalent in almost all software systems.
-% In this paper we will show that NFR related topics occur quite frequently.
-%XXXXXX this sentence is confusing
-%This informs our method of generalizing the data mining of these software artifacts to leverage the global domain knowledge of software engineering,
-%specifically NFRs. 
-%In particular, we use software quality models to generalize topics across projects. 
-%This allows us to mine software systems with the intention of comparing their NFR-related development topics.
-% in order to achieve generalizable, cross project data mining results.
-
-%%%% THIS PARAGRAPH ARGUES ABOUT THE CONCRETE APPLICATION OF 
-%%%% TOPIC MODELLING AND LABELLED TOPIC EXTRACTION
-%%%% IT NEEDS TO BE TIGHTER BUT IT ALSO NEEDS TO BE HERE
-
-% AH: Topics take a long to read
-% AH: Project Dashboards, summarization, software process recovery
 
@@
 
-Concrete applications of topics and \emph{labelled topic extraction}
-range from project dashboards to annotating software artifacts such as revisions and bug reports with
-NFR-related tags.
-Project dashboards~\cite{dashboard} are typically employed by managers and are used to
-provide quick summaries of the effort put into a software project. In this case, 
-labelled topic extraction would allow managers to track effort
-related to NFR topics, such as \emph{portability}.
-These techniques also allow for the annotation of commit comments
- and other software artifacts with NFRs. 
-This would enable querying of bug reports and artifacts by relevant NFRs. 
-For instance a manager can confirm if their developers were
-focused on \emph{usability} by looking for
- \emph{usability}-relevant revisions and bug reports.
-
 In this paper, we describe \emph{automated labelled topic extraction}. It addresses two gaps in the topic mining literature:
@@
 
-This work extends our previous work~\cite{msr2011}, published at The
+This work extends our previous work~\cite{msr2011}, published at the
 Working Conference on Mining Software Repositories 2011
-(MSR'2011). 
-Within this new paper that builds up upon the previous
-work, we added more machine learning results and another case-study: PostgreSQL. 
-We used two authors to
-each annotate the same PostgreSQL topics in order to compare
+(MSR'2011). The major extensions in this paper are as follows:
+\begin{itemize}
+	\item We added more machine learning results and another case-study: PostgreSQL. 
+	\item We used two authors to each annotate the same PostgreSQL topics in order to compare
 them. 
-Thus we used this new case-study to test inter-rater
-reliability, described in Section \ref{sec:inter-rater}. 
-In Section
-\ref{sec:developers}, we further extended the MSR'2011 paper with an
-analysis of PostgreSQL authors and their association with NFRs and
-topics.
+	\item We used this new case-study to test inter-rater reliability, described in Section \ref{sec:inter-rater}. 
+	\item We conducted an analysis of PostgreSQL authors and their association with NFRs and
+topics (Section \ref{sec:developers}).
+\end{itemize}
 
@@
 
-Finally, 
-this paper and our MSR 2011 paper~\cite{msr2011} are based off of the
+This paper and our MSR 2011 paper~\cite{msr2011} are based off of the
 work of Ernst and Mylopoulos~\cite{ernst10refsq} and the work of Hindle et
@@
 project that identifies changes in quality requirements
-in GNOME software projects (GNOME is a Linux desktop environment). This earlier approach was more exploratory, had
-less validation, uses different word-lists, solely uses text-matching,
+in GNOME software projects (GNOME is a Linux desktop environment). Unlike this paper, this earlier approach was more exploratory and had
+less validation. In particular, it  uses different word-lists, solely uses text-matching,
 and does not leverage machine learning strategies.  

### 2012-01-17 ese/ese.tex
fixes for review comments, including new process model diagram.

@@
 % edit this. awkward
-In this work we seek to provide a method of automatically labelling  development topics extracted from commit logs.
+In this work we provide a method of automatically labelling  development topics extracted from commit logs.
+
 Concrete applications of topics and \emph{labelled topic extraction}, which this paper introduces,
@@
 In this section we describe how to label topics based on dictionaries
-mined from sources external to the projects.
+mined from sources external to the projects. We call this `semi-unsupervised' because while there is no ``training set'', we do seed the wordlists manually.
 
@@
 we associate each NFR with a list
-of keywords (\emph{word-lists}, in our parlance). These word-lists were determined a priori and were not extracted from the projects themselves.
+of keywords (\emph{word-lists}, in our parlance).
+These word-lists were determined a priori and were not extracted from the projects themselves, using the methodology explained below.
+In general, these lists are project and domain independent.
 We intersected the words of the topics and the words of our word-lists.
@@
 That paper constructs an ontology for software quality measurement using eighty source documents, including research papers and international standards. 
-The labels we used:
+The labels we used are:
 
@@
 were
-\emph{correctness} (182/305/748 -- MySQL, MaxDB and PostgreSQL topics) and \emph{testability} (121/238/625). 
+\emph{correctness} (182/305/748, which represent MySQL, MaxDB and PostgreSQL topic counts, respectively) and \emph{testability} (121/238/625). 
 We did not see many results for \emph{usability} (4/0/138) or
@@
 For each quality in the high-level ISO 9126 taxonomy (namely, \emph{Maintainability, Usability, Reliability, Efficiency, Portability, Functionality}) we assessed whether the semi-unsupervised labels for a topic matched the manual annotations we created for the validation corpus.
-To be clear, the manual annotations were not used to
+Recall that the manual annotations were not used to
 train the labelling process.
 As described in Section \ref{sec:wordlist} we used both ROC and F-1 measures to evaluate the performance of the classification.
-Figure \ref{fig:maxdb-unsup-results} shows our ROC results for PostgreSQL, MaxDB and MySQL. We describe F-1 results in the text below.
+Figure \ref{fig:maxdb-unsup-results} shows our ROC results for PostgreSQL, MaxDB and MySQL. We omit \textsf{exp1} due to poor results. We describe F-1 results in the text below.
 
@@
 Because our ground truth annotations were relevant only to ISO9126,
-we estimate that \textsf{exp1} had poor
-performance via the overlap between ISO9126 and the Kayed ontology.
+ \textsf{exp1} had poor
+performance due to the overlap between ISO9126 and the Kayed ontology (i.e., we annotated topics with labels which did not appear in the validation corpus). 
 For \textsf{exp1} the F-measures for the NFRs for MaxDB were from $0$ to $0.18$ with an average (of all NFRs)
@@
 label. When applying our technique, for each NFR one should select the
-best performing learner possible.
+best performing learner possible. Given the tractability of these techniques, a tool which applied all learners and presented the best result should be feasible. 
 %Figure \ref{fig:best-learn-per-tag} uses the ZeroR learner as a baseline, a common practice in machine learning,
@@
 \emph{functionality} swap between projects. For PostgreSQL, the
-performance is nearly always more poor than the other two systems. The
+performance is nearly always poorer than the other two systems. The
 reason for this lack of performance could be that the number of
-topics, $N$ could be non-optimal for PostgreSQL. Perhaps topics were
+topics, $N$, could be non-optimal for PostgreSQL. Perhaps topics were
 getting too mixed and it was becoming hard to distinguish one topic
@@
 
-As noted in Section \ref{sec:wordlist}, each topic in our data-set can be composed of zero, one, or more NFRs. 
+As noted in Section \ref{sec:wordlist}, each topic in our data-set can be composed of zero or more NFRs. 
 For example, a commit message might address \textit{reliability} in the context of \textit{efficiency}, or make a \textit{maintainability} improvement
@@
 Possibly they would have influenced our results, but there would be a degree of correlation between the corpora.
+ It is possible for a given label to occur across the arbitrary 30-day boundary we set. We suspect but have not proved that this is insignificant. 
 %Conversely, construct validity was addressed in our validation as we used the files, topics and revisions to annotate the commit messages. 
@@
 generally accepted. Finally, there are exogenous sources, such as
-in-person discussions, which we did not access, an omission as noted
-in Aranda et al.~\cite{aranda09icse} %We did not show these results to the stakeholders.
+in-person discussions, which we did not access.  %We did not show these results to the stakeholders.
 
@@
 
-Our data and scripts are available at:
-
-\url{http://softwareprocess.es/nomen/}
+Our data and scripts are available at \url{http://softwareprocess.es/nomen/}
 
@@
 
-%\begin{acknowledgements}
-%\end{acknowledgements}
+% 
+% \begin{acknowledgements}
+% 	Thanks to the helpful comments from the external reviewers.
+% \end{acknowledgements}
 

### 2012-01-17 ese/ese.tex
fixed most of the minor todos

@@
 The Mulan~\cite{mulan} library encapsulates several different multi-label machine learners which can label elements with multiple labels.
-Mulan also includes methods for determining the performance of such techniques.
+Mulan also includes methods for determining the performance of these learners.
 % TOO DETAILED The problem framed in the learners above has changed; instead of looking at the precision and recall of applying one label, we rank
@@
 
-Two perspectives used to evaluate multi-label learners are with micro or macro measurements (used in Figure \ref{fig:mulan}).
+Two perspectives used to evaluate multi-label learners are with micro or macro measurements (shown in Figure \ref{fig:mulan}a).
 Macro measurements are aggregated at a class or label level (per class) while micro measurements are at the element level (per element).
@@
 
-Figure~\ref{fig:mulan} presents the results of Mulan's best multi-label learners for our data. 
+Figure~\ref{fig:mulan} presents the results of Mulan's best multi-label learners for the MaxDB and MySQL projects, and  Figure~\ref{fig:pgcomp} for PostgreSQL. 
 Calibrated Label Ranking (CLR) is a learner that builds two layers. The first layer determines if an entity should be labelled, while the second layer
@@
   in knowing whether a particular NFR, such as \emph{reliability}, was
-  more important for one project than another. This could be to
+  more important for one project than another. This question could be used to
   confirm the initial design goals, or to track the progress on that
@@
 %\subsection{Timeline analysis}
-Topic time-lines are depicted in Figures \ref{fig:mysql-timeline} and
+Topic time-lines are depicted in Figures \ref{fig:mysql-timeline}, 
 \ref{fig:maxdb-timeline} and \ref{fig:pgsql-timeline}. These topic time-lines show the
@@
 One, the relative frequency, shown in the grey histogram boxes, represents the number of topics with that NFR in that period, 
-relative to the maximum number of topics assigned to that NFR. 
+relative to the maximum number of topics assigned to the NFR. 
 For example, in Figure \ref{fig:mysql-timeline} we see a spike in \emph{portability} and \emph{functionality} frequency in September 2002.
-The second, absolute frequency, is shown using cell intensity, and compares the number of topics labelled with that NFR per period 
+The second, absolute frequency, is shown using cell intensity, and compares the number of topics labelled with the NFR per period 
 relative to the maximum number of labelled topics overall. 
@@
 contain more labelled topics, since these NFRs have been more
-intensely shaded; 
-one interesting stream is \emph{efficiency} which shows periodic
-activity, does this suggest that once maybe efficiency related changes
-have longer lasting effects?
+intensely shaded. One interesting stream is \emph{efficiency}, which shows periodic
+activity, and suggests that efficiency-related changes
+have longer lasting effects.
 The topmost row in each diagram lists historical events for that project (such as a release). 
@@
 %\noindent \textbf{Timeline of key events} -- 
-We analyzed each project's developer mailing list for external validation. 
+We analyzed each project's developer mailing list for external validation (the body of the email).
 %As discussed earlier, the top row in each figure shows key events for each project. 
@@
 %For MySQL (Figure \ref{fig:mysql-timeline}), we similarly validated our NFR patterns with external mailing list data. 
-The release of MySQL we study  (Figure \ref{fig:mysql-timeline}) was the first to be licenced under the GPL. 
-Version 3.23.31 (January, 2001) was the production release (non-beta), and we see a flurry of topics labelled with \emph{functionality} and
+The development period of MySQL we studied  (Figure \ref{fig:mysql-timeline}) saw the first releases to be licensed under the GPL. 
+Version 3.23.31 (January, 2001) was the production release (non-beta), and the timeline view shows a flurry of topics labelled with \emph{functionality} and
 \emph{maintainability}. 
@@
 
-By contrast, the PostgreSQL time-line  (Figure \ref{fig:pgsql-timeline}) is extracted from a central trunk and is not version-specific. Therefore development 
+By contrast, the PostgreSQL time-line  (Figure \ref{fig:pgsql-timeline}) is extracted from a central trunk and is not version-specific (due to differences in how the projects manage branches).
+Therefore development
 tends to be focused on releases 7.3, 7.4 and 7.5/8.0 alpha. For example, a priority for the 8.0 candidate was a Windows-native port of the source code, which seems to 
-correlate with the \emph{portability} NFR increasing in frequency in mid-2004. Before the 7.4 release, \emph{usability,functionality} and \emph{portability} all 
+correlate with the \emph{portability} NFR increasing in frequency in mid-2004. Before the 7.4 release, \emph{usability, functionality} and \emph{portability} all 
 increase in frequency, possibly reflecting the interest in adding features and documentation for the release. In the following sections we turn to our research questions:
@@
 
-In PostgreSQL, by comparison, the frequencies seem to be cyclic, since we are not studying a maintenance-phase for the product, but rather ongoing
+In PostgreSQL, by comparison, the frequencies seem to become somewhat cyclic, since we are not studying a maintenance-phase for the product, but rather ongoing
 feature addition and usability improvements.
@@
 A non-functional change that was not quality-related was licensing and
-copyright; many changes were simply to do with updating copyrights or
+copyright; many changes concerned updating copyrights or
 ensuring copyright or license headers were applied to files. In these
@@
 correlation of our per topic annotations for each NFR. We had to
-evaluate inter-rater reliability this way because 1 topic could be
-tagged with more than 1 NFR.
-
-We feel these results are poor. The aggregate view of with a Kappa of
-$0.1$ indicates there is some agreement but not a lot. We found that
+evaluate inter-rater reliability this way because a single topic could be
+tagged with more than one NFR.
+
+We feel these results are poor. The aggregate view of a Kappa of
+$0.1$ indicates there is some weak agreement. We found that
 there was a lot of agreement in terms of lack of a tag, but often more
-disagreement when a tag should be applied.
-
-After some discussion we concluded that usability was a source of
+disagreement when a tag should be applied. After some discussion we concluded that usability was a source of
 disagreement. Do user manuals improve usability? Is a command-line
@@
 to train and discuss when one applies an annotation.
-Furthermore, 
-both annotators found that the annotation was time-consuming and
-difficult.
+%Furthermore, 
+%both annotators found that the annotation was time-consuming and
+%difficult.
 
@@
 For both MySQL and MaxDB \textit{usability} was better served by \textsf{exp2}. 
-\textit{Usability} was a very infrequent label, however, which made it difficult to detect in any case.
+\textit{Usability} was a very infrequent label, however, which made it difficult to detect for both approaches.
 
@@
 With ROC values ranging from $0.6$ to $0.8$ for MySQL and MaxDB and
-$0.47$ to $0.6$ for PostgreSQL, we can see there is promise in these methods.
+$0.47$ to $0.6$ for PostgreSQL, we can see there is promise in supervised methods.
 \textsf{exp2} and \textsf{exp3} both indicate that static information can be used to help label topics without any training whatsoever. 
@@
 
-Since this work focuses on labelling natural language commit log
-comments we feel it can be adapted to other natural language artifacts
-such as mailing-list discussions and bug reports, even though that was
-not evaluated in this paper. Bug reports might not exhibit the same
+While this work focuses on labelling natural language commit log
+comments, we feel it can be adapted to other natural language software artifacts, 
+such as mailing-list discussions and bug reports. Bug reports might not exhibit the same
 behaviour as commits in terms of dominant topics.
@@
 Our taxonomy for software NFRs is subject to dispute, but seems to be
-generally accepted. Finally, there are exogenous sources, such as
+generally accepted.  A future approach should consider a different taxonomy, such as one created by surveying developers on what ``types" of tasks they work on.
+Finally, there are exogenous sources, such as
 in-person discussions, which we did not access.  %We did not show these results to the stakeholders.

### 2012-01-17 ese/ese.tex
interrater reliability

--- 2012-01-17 ese/ese.tex
@@
 
+Empirically how poor were our results? We decided to compare our
+results against simulated or random results. In the first simulation
+we sampled with replacement from our own NFR label distributions to
+produce random ratings that looked like our own. We then applied the
+Kappa-statistic on this sampling versus our labelling, and repeated
+this 10000 times, we found that
+for 4 out of 7 NFR labels, our labels IRR were greater than the random
+IRR measures 95\% of the time. For portability and efficiency, our IRR
+was greater 100\% of the time both for Neil's ratings, Abram's ratings
+and the union of both Neil and Abram. Figure \ref{fig:irr} depicts the IRR of
+these sampled ratings versus our IRR using the Kappa statistic.
+
+We also achieved similar results if the ratings were pulled from a
+uniform distributions. What this does indicate is that our labels and
+ratings were definitely better than random except for functionality,
+reliability and usability.
+
+\begin{figure}
+  \centering
+  \includegraphics[width=0.9\textwidth]{figures/self-sample-Both}
+  \caption{Measured IRR versus IRR of labeling random simulations.}
+  \label{fig:irr}
+\end{figure}
+
 After some discussion we concluded that usability was a source of

### 2012-01-17 src/validate/interrater.R

@@ -1,4 +1,5 @@
 #neil and abram annotations!
+cores <- 8
 pgsqln <- read.csv("output/pgsqln.arff.csv",header=TRUE)
 pgsqla <- read.csv("output/pgsqla.arff.csv",header=TRUE)
 nn <- length(pgsqln)
@@ -57,6 +58,7 @@ mat <- matrix(c( paall, pnall ), ncol=2)
 kappam.fleiss(mat)$value
 kappa2(mat)$value
 
+
   for (i in c(1:7)) {
     print(paste(names(pa)[[i]]))
     print(paste("Pos Disagree: ",length(pa[pa[,i]+pn[,i]==1,i])))
@@ -72,9 +74,11 @@ sapply(names(pa),function(i) {
 })
 
 kappaof <- function(sai,pai) {
-    mat <- matrix(c( sai, pai ), ncol=2)
-    kappa2(mat)$value
+  #mat <- matrix(c( sai, pai ), ncol=2)
+  mat <- matrix(c( as.factor(sai), as.factor(pai) ), ncol=2)
+  kappa2(mat)$value
 }
+
 
 # repeat N times
 # sample from the distribution itself
@@ -91,7 +95,119 @@ kappaofself <- function( pan, N) {
                          sapply(names(pan),function(i){
                            kappaof(sample(pan[,i],630,replace=T),pan[,i])})})[x,])})
 }
-# Abram annotations 1000 times
+
+library(irr)
+library(multicore)
+
+paunion <- pa
+for (i in names(pa)) {
+  paunion[,i] <- (pa[,i] & pn[,i]) * 1
+}
+
+ratings <- list()
+ratings[["Abram"]] <- pa
+ratings[["Neil"]] <- pn
+ratings[["Both"]] <- paunion
+panames <- c("Portability","Functionality","Reliability","Maintainability","Effeciency", "Usability", "None")
+
+for (person in c("Abram","Neil","Both")) {
+pan <- ratings[[person]]
+i <- names(pan)[1]
+#
+#           mean(
+#                mclapply(c(1:N),
+#                sapply(c(1:N),
+#                       function(x){
+#                         sapply(names(pan),function(i){
+#                           kappaof(sample(pan[,i],630,replace=T),pan[,i])})})[x,])})
+#}
+
+
+#versus a sample
+
+N <- 1000
+kappas <- lapply(names(pa),function(acolumn) {
+  unlist(
+         mclapply(
+                  c(1:N),
[…]

### 2012-01-17 ese/ese.tex

--- 2012-01-17 ese/ese.tex
@@
 to train and discuss when one applies an annotation.
-%Furthermore, 
-%both annotators found that the annotation was time-consuming and
-%difficult.
+
+Empirically how poor were our results? We decided to compare our
+results against simulated or random results. In the first simulation
+we sampled with replacement from our own NFR label distributions to
+produce random ratings that looked like our own. We then applied the
+Kappa-statistic on this sampling versus our labelling, and repeated
+this 10000 times, we found that
+for 4 out of 7 NFR labels, our labels IRR were greater than the random
+IRR measures 95\% of the time. For portability and efficiency, our IRR
+was greater 100\% of the time both for Neil's ratings, Abram's ratings
+and the union of both Neil and Abram. Figure \ref{fig:irr} depicts the IRR of
+these sampled ratings versus our IRR using the Kappa statistic.
+
+We also achieved similar results if the ratings were pulled from a
+uniform distributions. What this does indicate is that our labels and
+ratings were definitely better than random except for functionality,
+reliability and usability.
+
+\begin{figure}
+  \centering
+  \includegraphics[width=0.9\textwidth]{figures/self-sample-Both}
+  \caption{Measured IRR versus IRR of labeling random simulations.}
+  \label{fig:irr}
+\end{figure}
 

### 2012-01-18 ese/ese.tex
- [X] Using ROC instead of F-Measure as your primary measure (for      graphs) was a surprise.  Why did you choose this? A cynical      reader would suggest because ROC values are higher (not my      point-of-view but you should be aware of this point).
- [X] 3.2.3 These f-measures are very low, potentially making the      approach not usable, consider discussing why you think that an      approach with such a low f-measure is usable.

@@
 A perfect learner has a ROC value of $1.0$, reflecting perfect recall and precision. 
-A ROC result of $0.5$ would be equivalent to a random learner (that is, issuing as many false positives as true positives). 
+A ROC result of $0.5$ would be equivalent to a random learner (that
+is, issuing as many false positives as true positives). While we
+recognize that using $0.5$ as the base-line means our ROC scores will
+look much larger than our F-Measure scores, we feel that the knowledge
+that the random selection is $0.5$ or worse is helpful for interpreting our results.
 The %area under the -- NE confusing to have area again ... and not importantana
@@
 were
+%XXX Neil                    V
 \emph{correctness} (182/305/748, which represent MySQL, MaxDB and PostgreSQL topic counts, respectively) and \emph{testability} (121/238/625). 
@@
 while \emph{functionality} remained poor. \emph{Functionality} improved dramatically by \textsf{exp3}.
-Many ROC scores were $0.6$ or less, but our classifier, in most cases, still performed substantially better than random.
+Our F-measure scores were low and many ROC scores were $0.6$ or less, but our classifier, in most cases,
+still performed substantially better than random ($0.5$), even in the
+face of heavy class-imbalance for qualities such as \emph{usability} 
+and \emph{efficiency}. While there is much room for improvement, we are seeing some
+correlation between our quality word lists and relevant topics.
 
@@
 %These results are promising, because the accuracy of the techniques is good enough to be useful, while the run-time performance is still cheap enough
-%to execute to be feasible as an automated or semi-automated method of labelling topics by their software qualities.
+%to execute to be feasible as an automated or semi-automated method of
+%labelling topics by their software qualities.
+Low F-measures and ROC scores are a concern for some of these
+techniques, perhaps the word lists need to be re-enforced or made
+robust in the face of heavy class imbalance.
 These results are promising because they indicate that these

### 2012-01-18 ese/ese.tex
Lots of edits, did we solve the 748 problem? I thought we did.

@@
 Characteristic~\cite{Fawcett2006861},
-abbreviated \emph{ROC}, sometimes called \emph{AUC}, and the F-measure, which is the harmonic mean of precision and recall, i.e., $2 * (P * R) / (P + R)$.
+abbreviated \emph{ROC}, sometimes called \emph{AUC}, and the
+F-measure, which is the harmonic mean of precision and recall, i.e.,
+$2 * (P * R) / (P + R)$. 
+Throughout the paper we will provide F-measure scores so that readers
+who are more familiar with F-measure than ROC can intuitively
+interpret the results.
 
@@
 
-Table \ref{tbl:wordlist} shows how many topics were labelled for MaxDB, MySQL and PostgreSQL.
+Table \ref{tbl:wordlist} shows how many topics were labelled for
+MaxDB, MySQL and PostgreSQL.  Notice how PostgreSQL had far fewer
+unlabelled topics, PostgreSQL also was under going far more
+development as we evaluated 20330 PostgreSQL commits versus 8664 MaxDB
+commits and 12056 commits. We suspect that PostgreSQL topics were
+flooded with many terms and perhaps we needed more 20 topics per month
+for PostgreSQL as this result indicates there is a lot of overlap in
+the topics.
 
@@
 were
-%XXX Neil                    V
-\emph{correctness} (182/305/748, which represent MySQL, MaxDB and PostgreSQL topic counts, respectively) and \emph{testability} (121/238/625). 
+%XXX Neil                    V changed from 748 is that right?
+\emph{correctness} (182/305/640, which represent MySQL, MaxDB and PostgreSQL topic counts, respectively) and \emph{testability} (121/238/625). 
 We did not see many results for \emph{usability} (4/0/138) or
@@
 
+PostgreSQL's size in terms of commits was equal to MySQL and MaxDB
+combined. This imbalance in size, combined with the choice of 20
+topics per month produced topics that represented too many issues or
+threads within PostgreSQL's development. Choice of number of topics
+should probably be tuned to the project.
+
 We primarily relied on ROC (the area under the receiver operating

### 2012-01-18 ese/ese.tex
I hope this isn't some phantom commit where I wrecked something

@@
 Kappa-statistic on this sampling versus our labelling, and repeated
-this 10000 times, we found that
+this 100,000 times, we found that
 for 4 out of 7 NFR labels, our labels IRR were greater than the random
-IRR measures 95\% of the time. For portability and efficiency, our IRR
+IRR measures 96\% of the time. For portability and efficiency, our IRR
 was greater 100\% of the time both for Neil's ratings, Abram's ratings
@@
 negotiate disagreements.
+
+% [1] "Person  Abram"
+% [1]  3.093543e-06  1.269107e-04 -5.132487e-05 -1.449561e-05  7.953201e-05
+% [6] -2.791593e-05  6.246161e-05
+% [1] "Person ECDF measures Abram"
+% [1] 0.99992 0.36418 0.55153 0.97717 1.00000 0.64988 0.96132
+% [1] "Person  Neil"
+% [1] -1.426950e-04 -8.394072e-06 -1.108948e-04  8.568950e-06 -1.531356e-06
+% [6]  3.387974e-05 -9.010485e-05
+% [1] "Person ECDF measures Neil"
+% [1] 0.99994 0.36777 0.55371 0.97775 1.00000 0.59585 0.92487
+% [1] "Person  Both"
+% [1] -1.026767e-04  2.685101e-05  2.550572e-06  5.996015e-05 -4.787163e-05
+% [6] -1.982510e-04 -8.891684e-05
+% [1] "Person ECDF measures Both"
+% [1] 0.99995 0.37230 0.57847 0.96622 0.99998 0.90559 0.99859
 

### 2012-01-18 ese/ese.tex
more fixes on e.g timeline critiques

@@
 We argue for using ROC over F-Measure because ROC suffers less from
-bias than F-Measure and as often F-Measure skews towards the positive
+bias than F-Measure and F-Measure often skews towards the positive
 class~\cite{flach-icml03,Forman:2010:ACS:1882471.1882479}, especially
@@
 performance is nearly always poorer than the other two systems. The
-reason for this lack of performance could be that the number of
-topics, $N$, could be non-optimal for PostgreSQL. Perhaps topics were
-getting too mixed and it was becoming hard to distinguish one topic
+reason for this lack of performance could be that parameter we chose for number of
+topics, $N$, could be non-optimal for PostgreSQL. One explanation is that given the size of the PostgreSQL dataset, it was becoming hard to distinguish one topic
 from the next. PostgreSQL's data-set was the largest of the 3 projects:
-the XML data-dump that described the PostgreSQL topics that we
+the XML datafile that described the PostgreSQL topics that we
 annotated was 8 times larger than MySQL and 1.7 times larger than MaxDB. These
@@
 
-For all projects Bayesian techniques did the best out of a wide variety of machine learners tested. 
+For all projects Bayesian techniques did the best out of a wide variety of machine learners tested, due to the large number of features they can handle.
 Our best learners, Discriminative Multinomial Naive Bayes, Naive Bayes
@@
 %NEIL- I think this (below) is just confusing, although maybe important to note.
-% based on the manually annotated
-%topics, although this visualization can be generated from the
-%results of labelled topic extraction.
+This is generated from the manually annotated
+topics, although this visualization can be generated from the
+results of labelled topic extraction. Note that there are no unlabelled topics in this dataset.
+
 There are two measures represented. 
@@
 contain more labelled topics, since these NFRs have been more
-intensely shaded. One interesting stream is \emph{efficiency}, which shows periodic
+intensely shaded. One interesting stream is \emph{efficiency} in PostgreSQL, which shows periodic
 activity, and suggests that efficiency-related changes
-have longer lasting effects.
+have longer lasting effects. A more systematic analysis of periodicity is necessary to properly conclude this.
 The topmost row in each diagram lists historical events for that project (such as a release). 
@@
 of version 3.23.52, a general bug fix with a focus on security (a component of the \emph{functionality} NFR in the ISO9126 model). 
-After this point, efforts shift to the newer releases (4.0, 4.1,
-5.0).
+After this point, efforts shift to the newer releases (4.0, 4.1, 5.0) and subsequently becomes more functionality oriented.
 

### 2012-01-18 ese/ese.tex
Major rewrite of developer clustering section.

@@
 
-% \subsection{Comparing MaxDB and MySQL}
-
-% \label{sec:comparison}
-%\XXX{Not sure there's enough here to do that!}
-% We observed that MySQL had more topic subsets than MaxDB. 
-% MySQL 3.23 also had more topics and a longer recorded history than MaxDB 7.500. 
-% We tagged more MySQL topics with annotations than MaxDB topics, yet both shared similarities. 
-% In terms of non-functional requirements both projects had long running trends that focused on \emph{functionality}, \emph{maintainability}, and
-%\emph{portability}, yet MaxDB had more of a focus on \emph{efficiency} and \emph{reliability}. 
-%\subsection{Research questions}
-%This analysis allows to address the questions raised in Section \ref{sec:suplabelling}:
-%This analysis addresses our research questions: %raised in Section \ref{sec:analysis}:
-
 \subsection{Do NFR Frequencies Change over Time?}
-%\noindent\textbf{1. Do NFR frequencies change over time?} --- 
-%\subsubsection{Do label frequencies change over time?}
-
 In both MaxDB and MySQL the frequencies generally decreased with age. 
@@
 
-We wanted to see how similar developers were to each other in terms of
-the NFRs that they worked on. 
-Within this section we focused on PostgreSQL.
-We first conducted pairwise $X^2$ (chi-squared)
+We wanted to see whether developers in the PostgreSQL project\footnote{Since the MySQL and MaxDB data had poor records for developer ids, we focused on PostgreSQL.} worked on similar NFRs.
+For example, one developer might be a usability expert, while
+another developer is focused on security.
+We assume that the NFR labels associated with a commit implicitly reflect work on those NFRs.
+Being able to report on where effort is being expended is useful for tracking productivity (among other things).
+%Within this section we focused on PostgreSQL.
+We explored this issue by extracting, for each topic in the PostgreSQL dataset, the developer names responsible for the commits used to generate that topic, and 
+the NFRs that are associated with the topic.
+This produced a map which associated developer name with NFR labels.
+The six NFR labels (\emph{maintainability, usability, efficiency, functionality, portability, reliability}) and the \emph{None} label describe a 7-dimensional space, into which we can position developers, based on the frequency with which each label occurs in the map.
+
+Using this space, we first conducted pairwise $X^2$ (chi-squared)
 tests and Kolmogorov-Smirnov tests for each developer pair. We
-found that in $3/10$ of these tests, the developers were significantly
-different, i.e., their distributions were not similar. This implies
-that there are authors who focus on different NFRs or different
-proportion of NFRs.
-
-In Figure \ref{fig:authorcluster} we clustered the top developers from
+found that in $3/10$ of these tests, the two developers were significantly
+different (i.e. distant in the 7-d space). This implies
+that there are developers who focus on different NFRs or different
+proportions of NFRs.
+
+In Figure \ref{fig:authorcluster} we clustered the developers from
 PostgreSQL using the
-Ward method of Hierarchical clustering.  This diagram shows that
-\texttt{petere}, \texttt{tgl} and \texttt{momjian} form their own
-cluster (they are also the most frequent committers). This is
+Ward method of Hierarchical clustering on the 7-d space.  
+This will group developers according to a distance metric. 
+We clustered the authors into both $2$-clusters
+and $6$-clusters. We chose $6$-clusters because the difference in distance
+between entities and centroids was minimal at $6$-clusters, and the
+derivative of distance was the highest in that region as well.
+
+\begin{figure}
+  \centering
+  \includegraphics[width=.9\textwidth]{figures/postgresql-author-cluster}
+  \caption{PostgreSQL Author Clusters: Authors clustered by cosine
+    distance of their NFR contributions. The outer rectangles (blue) reflect the 2-clusters; the inner rectangles (purple) reflect 6-clusters.}
+\label{fig:authorcluster}
+\end{figure}
+
+Figure \ref{fig:authorcluster} shows that when we use $6$-clusters, 
+\texttt{petere}, \texttt{tgl} and \texttt{momjian} form their own individual
+clusters (i.e., are distant from one another in the 7-d space). 
+%NEIL - not sure the relevance of this ... needs more explaining. --> (they are also the most frequent committers). 
+This is
 interesting because it means that the important developers are not
-committing code that fits the same NFR profile. In terms of the larger
-cluster we can see that \texttt{momjian} and \texttt{tgl} are in the
+committing code that fits the same NFR profile. In terms of the 2-clusters,
+ we can see that \texttt{momjian} and \texttt{tgl} are in the
 same cluster, but \texttt{petere} is not. The most frequent
-committers do not share the same clusters. This implies that
-developers do work on different sets of NFRs and have different
-software quality focus.
-
-Figure \ref{fig:authorcluster} clustered the authors into $2$ clusters
-and $6$ clusters. We chose $6$ because the difference in distance
-between entities and centroids was minimal at $6$ clusters, and the
-derivative of distance was the highest in that region as well.
-
-It should be noted that much of PostgreSQL development originates from
-the mailinglist and the developers are often pushing integrations. If
-we compare the global NFR distribution to each author we find that 25\%
- of the authors have a similarity of $0.47$ or less. This is
-depicted in Figure \ref{fig:authorsim}, where we can see that not all
-authors are as similar to the global trend as others. Of course, we
-found that number of commits correlated with an author's similarity to the
-global NFR distribution ($0.59$ Pearson).
-
-Our theory is that the less frequent committers are more
-focused and less general, thus their distributions of topics are
-different than the main developers who commit code in many different
-contexts.  The lead developers either ``wear more hats'' or have more
-time to ``wear more hats''.
-
-Given the top 15 PostgreSQL developers we found that proportionally many
-developers had a focus on NFR topics. Authors with top proportions per NFR
-were:
-\texttt{dennis} for usability; 
-\texttt{scrappy} and \texttt{meskes} for portability;
-\texttt{inoue} and \texttt{neilc} for efficiency;
-\texttt{jurka} and \texttt{joe} for reliability;
-\texttt{thomas} and \texttt{weick} for functionality;
-and \texttt{ishii} and \texttt{scrappy} for maintainability.
-Given these results we found that there indeed was a focus
-of $1/3$ to $1/10$ of commits dedicated to a particular NFR.
+committers do not share the same clusters, even at the coarsest level. 
+This implies that
+developers in this sample do work on different sets of NFRs and have different
+software quality focuses.
+
+% NEIL --> Explain the reason why this sentence is included: 
+%It should be noted that much of PostgreSQL development originates from
+%the project mailing list, and the developers are often pushing integrations. 
+If we compare the global NFR distribution (that is, the relative global frequency of each NFR label)
+to each author we find that 25\%
+ of the authors have a similarity of $0.47$ or less. In other words, for these authors (gathered to the lower left of Figure  \ref{fig:authorsim}),
+ their NFR distribution does not match well with the global distribution.
+We found that number of commits correlated with an author's similarity to the
+global NFR distribution ($0.59$ Pearson), i.e., the variables ``number of topics associated with a developer'' and ``number of commits'' are not discriminative.
+
+\begin{figure}
+  \centering
+  \includegraphics[width=.9\textwidth]{figures/author-distance-from-aggregate}
+  \caption{PostgreSQL Author Similarity to the Global Distribution of NFR topics}
+\label{fig:authorsim}
+\end{figure}
+
+Our working theory is that less frequent committers are more
+specialized, e.g., interested in a specific NFR, while the main developers (i.e., frequent committers) either ``wear more hats'' or have more
+time to ``wear more hats''. An interesting extension would be to compare this data with 
+the PostgreSQL source code files, to see which developer touched which file.
+
+We say that an author (read developer) is ``proportionately interested'' in an NFR if, for all NFRs with which he or she is associated, a given NFR receives the plurality of his or her commits.
+This is a measure of relative interest and is independent of number of commits (subject to the caveat about frequent committers, above).
+If we look at which of the top 15 developers were proportionately interested in a given NFR, we find the associations identified in Table \ref{tbl:devinterest}.
+Our data also showed that for these top developers, between $1/10$ and $1/3$ of their commits were labelled with a single NFR.
+
+\begin{table}
+\begin{tabular}{l|c}
+\toprule
+\textbf{NFR} & \textbf{Developer} \\
+\midrule
+\emph{Usability} & \texttt{dennis, neilc} \\
+\emph{Portability} & \texttt{scrappy, meskes} \\
+\emph{Efficiency} & \texttt{inoue, neilc} \\
+\emph{Reliability} & \texttt{jurka, joe} \\
+\emph{Functionality} & \texttt{thomas, weick} \\
+\emph{Maintainability} & \texttt{scrappy, ishii}\\
+\bottomrule
+\end{tabular}
+	\caption{Developer interest in NFRs}
+	\label{tbl:devinterest}
+\end{table}
 
@@
 describing developer style instead of developer focus. Developer style
-would be a developers likelihood to use or not use terms found within
-our dictionaries and training corpus. Further more a top developer, in
-terms of frequency of commits, will have more samples and more likely
-take on a more general role. We found in a previous
-paper~\cite{largechanges} that often the words used to describe a commit
-and the actual author of the commit were often correlated. Thus based
-on that observation developer style might be a confounding concern.
+would be a developer's likelihood to use or not use terms found within
+our dictionaries and training corpus in their commit messages. Furthermore, a top developer, in
+terms of number of commits, will have more samples, and thus be more likely
+take on a more general role. For example, we found in a previous
+project~\cite{largechanges} that there was a good correlation between the words used to describe a commit
+and the actual author of the commit. 
+%were often correlated. Thus based
+%on that observation developer style might be a confounding concern.
 
@@
 
-\begin{figure}
-  \centering
-  \includegraphics[width=.9\textwidth]{figures/postgresql-author-cluster}
-  \caption{PostgreSQL Author Clusters: Authors clustered by cosine
-    distance of their NFR contributions}
-\label{fig:authorcluster}
-\end{figure}
-
-\begin{figure}
-  \centering
-  \includegraphics[width=.9\textwidth]{figures/author-distance-from-aggregate}
-  \caption{PostgreSQL Author Similarity to the Global Distribution of NFR topics}
-\label{fig:authorsim}
-\end{figure}
-
 \section{Discussion}

### 2012-01-18 ese/ese.tex
abstract attempt

@@
 
-Researchers have employed a variety of techniques to extract underlying
-topics that relate to software development artifacts.  Typically, these
-techniques use semi-unsupervised machine-learning algorithms to suggest
-candidate word-lists.  However, word-lists are
+%Researchers have employed a variety of techniques to extract underlying
+%topics that relate to software development artifacts.  Typically, these
+%techniques use semi-unsupervised machine-learning algorithms to suggest
+%candidate word-lists, topics.  
+Software repositories provide a deluge of software artifacts to
+analyze. Researchers have attempted to summarize, categorize, and relate these
+artifacts by using semi-unsupervised machine-learning
+algorithms, such as Latent Dirichlet Allocation (LDA), used for concept and topic analysis to suggest candidate
+word-lists or topics that describe and relate software artifacts.
+However, these word-lists and topics are
 difficult to 
@@
   Topics are extracted 
-using  Latent Dirichlet Allocation (LDA) from commit-log comments recovered
+using  LDA from commit-log comments recovered
 from source control systems. % such as CVS and BitKeeper.  

### 2012-01-18 ese/ese.tex
intro comments from migod

--- 2012-01-18 ese/ese.tex
@@
 show that labelled topic extraction can produce appropriate,
-context-sensitive labels relevant to these projects, which provides fresh
+context-sensitive labels that are relevant to these projects, and provide fresh
 insight into their evolving software development activities.
@@
 % edit this. awkward
-In this work we provide a method of automatically labelling  development topics extracted from commit logs.
-
-Concrete applications of topics and \emph{labelled topic extraction}, which this paper introduces,
-range from project dashboards to annotating software artifacts such as revisions and bug reports with
-NFR-related tags.
-Project dashboards~\cite{dashboard} are typically employed by managers and are used to
-provide quick summaries of the effort put into a software project. In this case, 
-labelled topic extraction would allow managers to track effort
-related to NFR topics, such as \emph{portability}.
-These techniques also allow for the annotation of commit comments
- and other software artifacts with NFRs. 
-This would enable querying of bug reports and artifacts by relevant NFRs. 
-For instance a manager can confirm if their developers were
-focused on \emph{usability} by looking for
- \emph{usability}-relevant revisions and bug reports.
+In this work we provide a method of automatically labelling  development topics extracted from commit logs called \emph{labelled topic extraction}.
+
+Concrete applications of \emph{labelled topic extraction} include the annotation of development artifacts with NFR-related tags and the creation of project dashboards.
+Annotating software development artifacts with NFR-related tags would allow developers to create detailed directed queries of different artifact kinds that concern the same NFR; for example, a developer could browse the recent history of performance-related bug reports and code check-ins.
+Project dashboards distill detailed information about a software system into a simpler and more abstract view that summarizes key aspects of the development effort \cite{dashboard}; labelled topic extraction would allow managers to track effort related to specific NFR topics, such as \emph{usability} or \emph{portability}.
 
@@
 Topic models are useful in software maintenance because they summarize
-the key concepts in a corpus, such as source code, commit comments, or
-mailing-list messages, by identifying statistically co-occurring words. 
-Among other uses, it can quickly give developers an overview of where significant
-activity has occurred, and gives managers or maintainers a sense of
-project history.
+the key concepts in a corpus -- such as source code, commit comments, or
+mailing-list messages -- by identifying words that commonly occur together. 
+Among other uses, topic modelling can quickly give developers an overview of where significant
+activity has occurred, and provide managers or maintainers an enhanced understanding of the project's 
+history.
 
@@
 In this paper, we introduce \emph{labelled topic extraction}, an
-approach to topic labelling that labels topics automatically with
-labels that are project independent.
+approach that automatically suggests project-independent labels for topics.
 
@@
 \begin{itemize}
-\item introduces the concept of labelled topic extraction, using a non-functional requirements (NFR) taxonomy for our labels;
-
-%\item evaluates an unsupervised method of automated topic labelling (word-lists);
-%%\item Our unsupervised method can be applied to projects without training;
-%%\item 
-%evaluates a supervised method of automatically labelling topics with a single NFR (machine learning);
-%%\item 
-%evaluates a supervised method of labelling topics with multiple NFRs
-%(multi-label machine learning);
-
-\item evaluates three kinds of automatic topic labelling methods:
+\item We introduce the concept of labelled topic extraction, using a taxonomy of non-functional requirements (NFR) for our labels; 
+\item We evaluate three kinds of automatic topic labelling methods:
   semi-unsupervised labelling  of topics (word-lists), supervised labelling of
@@
   machine learning);
-
-%  of automated topic labelling (word-lists);
-%\item Our unsupervised method can be applied to projects without training;
-%\item 
-%evaluates a supervised method of automatically labelling topics with a single NFR (machine learning);
-%\item 
-%evaluates a supervised method of labelling topics with multiple NFRs (multi-label machine learning);
-
-\item examines how NFRs correlate with the work of individual developers;
-\item provides a method of cross-project analysis via topic labelling; and
- applies these techniques to visualize NFRs over time, and to analyze maintenance activities.
+\item We examine how NFRs correlate with the work of individual developers;
+\item We provide a method of cross-project analysis via topic labelling, and
+we apply these techniques to visualize NFRs over time, and to analyze maintenance activities.
 
@@
 
-This work extends our previous work~\cite{msr2011}, published at the
-Working Conference on Mining Software Repositories 2011
-(MSR'2011). The major extensions in this paper are as follows:
+This work extends our previous work~\cite{msr2011}. The major extensions in this paper are as follows:
 \begin{itemize}
-	\item We added more machine learning results and another case-study: PostgreSQL. 
+	%\item We added new results on developer activities.
+	\item We conducted another case-study on PostgreSQL. 
 	\item We used two authors to each annotate the same PostgreSQL topics in order to compare
-them. 
+these annotations. 
 	\item We used this new case-study to test inter-rater reliability, described in Section \ref{sec:inter-rater}. 
@@
 The objective of Cleland-Huang's study was to identify new NFRs for system development, 
-yet our objective was to recover those latent NFRs from commit-log messages of the project.
+while our objective was to recover those latent NFRs from commit-log messages of the project.
 
@@
 \subsection{Inter-rater reliability}
-
+\label{sec:inter-rater}
 To determine inter-rater reliability two of the authors---Ernst and Hindle---each annotated the PostgreSQL topics,

### 2012-01-18 ese/ese.tex
- [X] 3.2.2: what preprocessing steps were taken before applying LDA to the commit messages? For example, were the terms stemmed? Were any identifiers split? Or were the words in the commits just delimited using non-alphanumeric characters?

--- 2012-01-18 ese/ese.tex
@@
 We extracted these messages and indexed them by creation time. 
-We summarized each message as a word distribution minus stop-words such as ``\emph{the}'' and ``\emph{at}''. 
+Each word in the message was
+stripped of punctuation and converted to lowercase.
+We summarized each message as a word distribution minus stop-words
+such as ``\emph{the}'' and ``\emph{at}''. We did not stem or apply any
+other transforms to the messages.
+Our stop words are derived from the Natural Language Toolkit (NLTK) stop-word English stop-word
+list~\footenote{NLTK: \url{http://www.nltk.org/}.
+
 %Stemming was performed in the later stages of our analysis. % WAS IT?
@@
 % Is this good enough
+Each change-log message was lightly processed before applying LDA:
+words were converted to lowercase with punctuation removed and then
+stop words were removed.
 This labelling is \emph{semi-unsupervised} because the corpus is not derived from 

### 2012-01-18 ese/ese.tex
I guess I should've compiled before commit!
hurrrf compile before commit again
section 2-3 comments from migod

--- 2012-01-18 ese/ese.tex
@@
 %they mine requirements documents whereas we mine revisions.
-One approach they tried was similar to this one, with keywords mined from NFR catalogues found in Chung et al.~\cite{corollary99}. 
+One approach they tried was similar to this one, with keywords mined from NFR catalogues found in Chung et al.~\cite{chung99}. 
 Their approach resulted in a recall of $80\%$ with precision of $57\%$ for the \emph{security} NFR, but could not find a reliable source of keywords for other NFRs. 
@@
 The labels are reasonable enough, but still require an expert in the field to determine them. 
-Furthermore, these labels are project-specific, because they are
+Furthermore, these labels are project specific, because they are
 generated from the data of that project. For example we might have a label called \emph{`Oracle'}
@@
 %This paper does this differently. 
-Our approach differs: first of all, we automate the process of naming the topics; secondly, we label topics with project-independent terms, in order
+Our approach differs in two important ways: we automate the process of naming the topics, and we label topics with project-independent terms, in order
 to permit cross-project comparison.
@@
 We used three different open-source, partially-commercial database systems:
-\begin{itemize}
-\item  MySQL 3.23. Started in 1994 and MySQL 3.23 was released in early 2001. MySQL contains $320,000$ lines of C and C++ source code~\footnote{generated
+\begin{description}
+\item [ MySQL 3.23] --- Started in 1994 and MySQL 3.23 was released in early 2001. MySQL contains $320,000$ lines of C and C++ source code~\footnote{generated
 using David A. Wheeler's \emph{SLOCCount},
@@
 31st, 2000 to August 9th, 2004.
-\item MaxDB 7.500. Started in the late 1970s as a research project, and was later acquired by SAP. As of version 7.500, released April 2007, the project
+\item [MaxDB 7.500] --- Started in the late 1970s as a research project, and was later acquired by SAP. As of version 7.500, released April 2007, the project
 has over $940,000$ lines of C source
@@
 29th, 2004 to June 19th, 2006.
-\item PostgreSQL 7.3. Started in the 1980s as a Berkeley research
+\item [PostgreSQL 7.3] --- Started in the 1980s as a Berkeley research
   project~\footnote{{http://www.postgresql.org/docs/7.3/static/}}. PostgreSQL
@@
 
-\end{itemize}
+\end{description}
 
@@
 would allow for cross-project comparison. At the same time we
-recognize that a problem-domain alone does guarantee functional and
+recognize that problem domain alone does not guarantee functional and
 non-functional similarity.
-For instance each database system has a different focus, PostgreSQL
+For instance, each database system has a different focus. PostgreSQL
 tends to focus on fulfilling much of the SQL92 specification and adding
-more features while MySQL has been slow to adopt much of the spec.
- As a consequence our choice of looking at a
-single domain limits our generalizability. What we show in this paper
+more features, while MySQL has been slow to adopt much of that specification in favour of community-requested features.
+The consequence of our choice to look at a
+single domain is to limit generalizability. What we show in this paper
 will be somewhat biased towards database software.
 
-For each project, we used source control commit comments, the messages
+For each project, we studied source control commit comments, the messages
 that programmers write when they commit revisions to a source control
 repository. 
-Most commits we observed had commit comments (90\% in MySQL 3.23,
-98.5\% in PostgreSQL and 99.99\% in MaxDB 7.500).
+Most commits we observed had comments: 90\% in MySQL 3.23,
+98.5\% in PostgreSQL and 99.99\% in MaxDB 7.500.
 Commit comments are often studied by researchers, as they are the most readily accessible source of project interactions, and developers are often
@@
 
-For the commit message data-sets of each project, we created an XML file that separated commits into 30 day periods. 
+For the commit message data-sets of each project, we created an XML file that partitioned commits into 30 day periods. 
 %This size of period, 30 days, is smaller than the time between minor
@@
 If there are not $N$ independent word distributions, the topics
-produce tend to be duplicates of each other, that is they share the
-top terms. During this study we found that when $N$ was 20 we did not
-witness duplicate topics very often.
+produced tend to be duplicates of each other, that is, they share the
+top terms. During this study we found that if we used a value of 20 for $N$, duplicate topics were infrequent.
 Linear combinations of these $N$ word distributions are meant to represent and recreate the word distributions of any of the original messages.
-These $N$ word distributions effectively form \emph{topics}: cross-cutting collections of words relevant to one or more of our commit messages. 
+In other words, these topics are cross-cutting collections of words relevant to one or more of our commit messages.
 LDA extracts topics in an unsupervised manner; the algorithm relies
 solely on the source data and word distributions of messages, with no human intervention.
-
-In topic analysis a single document, such as a commit message, can be related to multiple topics. 
-Representing documents as a mixture of topics maps well to source code repository commits, which often have more than one purpose~\cite{Hindle09ICSM}.  
-For this paper, a topic represents a word distribution generated from a group of commit log comments which are related by their content.  
+%We applied Blei's LDA implementation~\cite{Blei2003} against the word distributions of these commits, and generated lists of topics per period.
+
+% In topic analysis a single document, such as a commit message, can be related to multiple topics. 
+% Representing documents as a mixture of topics maps well to source code repository commits, which often have more than one purpose~\cite{Hindle09ICSM}.  
+% For this paper, a topic represents a word distribution generated from a group of commit log comments which are related by their content.  
 %In this paper a topic is a set of tokens extracted from commit messages found within a project's source control system (SCS).
 
-We applied Blei's LDA implementation~\cite{Blei2003} against the word distributions of these commits, and generated lists of topics per period. 
-We set the number of topics to generate to $20$, because past experimentation showed that fewer topics might aggregate multiple unique topics while
-any more topics dilutes the results and creates indistinct topics~\cite{Hindle09ICSM}. 
+%We set the number of topics to generate to $20$, because past experimentation showed that fewer topics might aggregate multiple unique topics while
+%any more topics dilutes the results and creates indistinct topics~\cite{Hindle09ICSM}. 
 % Furthermore, more than $20$ topics quickly became infeasible for inspection.
@@
 
-To facilitate cross-project comparison, we used a taxonomy of NFRs. This taxonomy is based on the ISO quality model, ISO9126~\cite{iso9126}. 
-ISO9126 describes six high-level NFRs: \emph{maintainability, functionality,
-portability, efficiency, usability,} and \emph{reliability}
+To facilitate cross-project comparison, we used a taxonomy of NFRs.
+This taxonomy is based on the ISO quality model, ISO9126~\cite{iso9126}, which describes six high-level NFRs: \emph{maintainability, functionality,
+portability, efficiency, usability,} and \emph{reliability}. 
 %\footnote{While there may be lingering debate in some circles about these terms, an ISO standard
 %seems like a reasonable starting point for our work.}.
-There is some debate about the terms in this model, and whether they are a) the correct terms and b) correctly organized. 
+There is some debate about the terms in this model \cite{Boegh2008}, and whether they are a) the correct terms and b) correctly organized. 
 However, ISO9126 is ``an international standard and thus provides an 
 internationally accepted terminology for software quality
-\cite[p. 58]{Boegh2008},'' that is sufficient for the purposes of this
+\cite[p. 58]{Boegh2008},'' and so we consider that it is sufficient for the purposes of this
 research.  
-\emph{Performance} is an example of RDBMS word related to the \emph{efficiency} NFR.
-We claim that these NFRs are maintenance concerns (to varying degrees) in all software projects, and are therefore well-suited for comparisons between
+\emph{Performance} is an example of an RDBMS word related to the \emph{efficiency} NFR.
+We claim that these NFRs are maintenance concerns (to varying degrees) in all software projects, and are therefore well suited for comparisons between
 projects.
@@
 the six NFR labels listed above.
-Annotators did not annotate each other's annotations, but some brief
-inspection of annotations was used to confirm that the annotators were
-acting similarly. PostgreSQL (PgSQL) was annotated by both annotators in order
+Annotators did not annotate each other's annotations
+%NEIL --> not really! also makes IRR confusing., but some brief
+%inspection of annotations was used to confirm that the annotators were acting similarly. 
+PostgreSQL (PgSQL) was annotated by both annotators in order
 to evaluate inter-rater reliability.
-We looked at each period's topics, and assessed what the data ---
-consisting of the frequency-weighted word lists and messages ---
-suggested was the label for that topic. 
-We were able to pinpoint the appropriate label using auxiliary information as well, such as the actual revisions and files that were related to the
+We looked at each period's topics, and assessed what the data---consisting of the frequency-weighted word lists and messages---suggested was the most appropriate labels for that topic. 
+We selected the appropriate labels using auxiliary information as well, such as the actual revisions and files that were related to the
 topic being annotated.
@@
 \emph{portability}. 
-Given the top-level annotations of \emph{none}, \emph{portability},
+Given the top-level annotations of \emph{portability},
 \emph{efficiency}, \emph{reliability}, \emph{functionality},
 \emph{usability}, and \emph{maintainability}, the annotators annotated each topic
-with the relevant label. Sometimes they used finer-grained
+with the relevant label. We added a catch-all label, \emph{none}, which we used when none of the six NFRs was suitable. In some instances, we used finer-grained
 annotations that would be aggregated up to one of these higher-level labels.
@@
 Characteristic~\cite{Fawcett2006861},
-abbreviated \emph{ROC}, sometimes called \emph{AUC}, and the
+abbreviated \emph{ROC}, sometimes called Area Under the Curve or \emph{AUC}, and the
 F-measure, which is the harmonic mean of precision and recall, i.e.,
@@
 We argue for using ROC over F-Measure because ROC suffers less from
-bias than F-Measure and F-Measure often skews towards the positive
+bias, as F-Measure often skews towards the positive
 class~\cite{flach-icml03,Forman:2010:ACS:1882471.1882479}, especially
@@
 we associate each NFR with a list
-of keywords (\emph{word-lists}, in our parlance).
+of keywords or \emph{word-lists}, in our parlance.
 These word-lists were determined a priori and were not extracted from the projects themselves, using the methodology explained below.
@@
 
-Our first word-list set, \textsf{exp1}, was generated using the ontology described in Kayed et al.~\cite{5072519}
-That paper constructs an ontology for software quality measurement using eighty source documents, including research papers and international standards. 
-The labels we used are:
-
-\begin{quotation}
-\small \noindent \textsf{
+Our first word-list set, \textsf{exp1}, was generated using the ontology for software quality measurement described in Kayed et al.~\cite{5072519}, which was constructed using 80 source documents, including research papers and international standards.
+The labels we used were:
+
+\vspace{2mm}
+\parbox{0.7\textwidth}{
+\small \textsf{
 integrity, security,
@@
 portability, flexibility, reliability.
-}
-\end{quotation}
+}}
+\vspace{2mm}
 
@@
 relevant to discussion of \emph{reliability}, but is not in the
-standard. We recognize though that terms like this might be used in a
+standard. We recognize that terms like this might be used in a
 different context with a different meaning, like code-cloning.
@@
 We did the same for the quality model produced by McCall et al.~\cite{mccall1977}. 
-We then did a simple random analysis of mailing list messages from an open source ecosystem, KDE.
+We then conducted a simple random analysis of mailing list messages from an open source ecosystem, KDE.
 Like the GNOME study we conducted in Ernst and Mylopoulos \cite{ernst10refsq}, KDE contains a suite of different products covering a variety of software categories.
@@
 Wordnet similarity is a very broad match. For example, the label \emph{maintainability} is associated with
-words \emph{ease} and \emph{ownership}, and the word \emph{performance} has a `sense' which refers to musical performances -- which is not related to software development.
+words \emph{ease} and \emph{ownership}, and the word \emph{performance} has a `sense' that refers to musical performances, which is obviously unrelated to software development.
 In general, as we proceed from word-lists in \textsf{exp1} to that in \textsf{exp3}, our lists become more generic.

### 2012-01-19 ese/ese.tex
migod changes in section 4/5

@@
 
-%\noindent \textbf{Generating labels automatically} ---
-
 Using our three word-lists (\textsf{exp1}, \textsf{exp2}, \textsf{exp3}), we labelled our topics with an NFR where there was a match between a word in
@@
 \emph{accuracy} (3/0/27), which were infrequently matched.
-Note the significantly higher result for \emph{usability} in PostgreSQL -- this suggests a difference in how this project is discussing usability, at least with respect to our analysis.
+Note the significantly higher result for \emph{usability} in PostgreSQL---this suggests a difference in how project developers are discussing usability, at least with respect to our analysis.
 We also looked for correlations between our labels: excluding double
@@
 For \textsf{exp3}, we generally labelled more topics. 
-As we mentioned, the word-lists are broad, so there are likely to be false-positives (discussed below). 
+As mentioned above, the word-lists are broad, so there are likely to be false-positives (discussed below). 
 The most frequent label across all projects (for this wordlist, and unlike \textsf{exp1}) was \emph{usability}, and the
@@
 To assess the performance of the supervised learners, we did a 10-fold cross-validation~\cite{Kohavi1995}, a common technique for evaluating machine
-learners. 
-The original data is partitioned randomly into ten sub-samples. Each sample is used to test against a training set composed of the nine other samples.
+learners, where the original data is partitioned randomly into ten sub-samples, and each sample is used to test against a training set composed of the nine other samples.
 %One subsample is used as a training set and evaluated on the other nine sub-samples. 
 %This is repeated nine more times, with each subsample used once as the training set. 
-We have reported these results below.% in Section \ref{sec:suplabelling}.
+We discuss these results below.% in Section \ref{sec:suplabelling}.
 
@@
 % \label{sec:suplabelling}
-Because our data-set was of word counts we expected Bayesian techniques, often used in spam filtering, to perform well. 
-We tried other learners that WEKA~\cite{weka09} provides: rule learners, decision tree learners, vector space learners, and support vector machines.  
+Because our data-set consisted of word counts we expected Bayesian techniques to perform well.
+Bayesian techniques are often used in spam filtering, which is similarly interested in word distributions.
+We tried other learners that WEKA~\cite{weka09} provides, including rule learners, decision tree learners, vector space learners, and support vector machines.  
 Figure \ref{fig:best-learn-per-tag} shows the performance of the best
@@
 
-For all projects Bayesian techniques did the best out of a wide variety of machine learners tested, due to the large number of features they can handle.
-Our best learners, Discriminative Multinomial Naive Bayes, Naive Bayes
-and Multinomial Naive Bayes  are all based on Bayes' theorem and all
-assume, naively, that the features supplied are independent. 
+Although we used a variety of machine learners, we found that Bayesian techniques performed the best on all projects. We believe this is due to the large number of features they can handle.
+Our best-performing learners---Discriminative Multinomial Naive Bayes, Naive Bayes
+and Multinomial Naive Bayes---are all based on Bayes' theorem and all make the naive assumption
+that the features supplied are independent. 
 %The features we used are word counts per message. 
@@
 For example, a commit message might address \textit{reliability} in the context of \textit{efficiency}, or make a \textit{maintainability} improvement
-in the source code that related to \textit{usability}. 
-However, traditional machine learning techniques, such as Naive Bayes, can only map topics to a single class. 
+in the source code that relates to \textit{usability}. 
+However, traditional machine learning techniques, such as Naive Bayes, can map topics to only a single class. 
 The Mulan~\cite{mulan} library encapsulates several different multi-label machine learners which can label elements with multiple labels.
@@
 % more why
-Labelled topics address \emph{why}  because they show reasons why changes
-occurred rather than \emph{how}. 
-The \emph{how} of a change is the change itself, the purpose or
+Labelled topics address \emph{why} questions by associating a commit with the underlying software qualities which motivated that change.
+The \emph{how} of a change is the change itself, the software quality behind it, the
 \emph{why} is what we are after.
@@
 We motivated our investigation with three research questions:
+\renewcommand{\labelenumi}{\textbf{RQ\arabic{enumi}. }}
 \begin{enumerate}

### 2012-01-19 ese/ese.tex

@@
 We extracted these messages and indexed them by creation time. 
-We summarized each message as a word distribution minus stop-words such as ``\emph{the}'' and ``\emph{at}''. 
+Each word in the message was
+stripped of punctuation and converted to lowercase.
+We summarized each message as a word distribution minus stop-words
+such as ``\emph{the}'' and ``\emph{at}''. We did not stem or apply any
+other transforms to the messages.
+Our stop words are derived from the Natural Language Toolkit (NLTK) stop-word English stop-word
+list~\footnote{NLTK: \url{http://www.nltk.org/}}.
+
 %Stemming was performed in the later stages of our analysis. % WAS IT?

### 2012-01-19 ese/ese.tex
some rewording in IRR section, migod's comment there

@@
 
-\subsection{Do NFR Frequencies Change over Time?}
+\subsection{RQ1: Do NFR Frequencies Change over Time?}
 In both MaxDB and MySQL the frequencies generally decreased with age. 
@@
 
-\subsection{Do Projects Differ in their Relative  Topic Interest?}
-\emph{Yes}. MySQL 3.23 had proportionally more
+\subsection{RQ2: Do Projects Differ in Their Relative  Topic Interest?}
+We found significant overall variation in the projects. MySQL 3.23 had proportionally more
 topics labelled \emph{functionality}, while MaxDB had proportionally more
@@
 
-\subsection{Do different developers work on different NFRs?}
+\subsection{RQ3: Do Different Developers Work on Different NFRs?}
 
@@
 Our working theory is that less frequent committers are more
-specialized, e.g., interested in a specific NFR, while the main developers (i.e., frequent committers) either ``wear more hats'' or have more
-time to ``wear more hats''. An interesting extension would be to compare this data with 
+specialized, e.g., interested in a specific NFR, while the main developers (i.e., frequent committers) either have wider responsibility or have more
+time to be widely involved. An interesting extension would be to compare this data with 
 the PostgreSQL source code files, to see which developer touched which file.
@@
 describing developer style instead of developer focus. Developer style
-would be a developer's likelihood to use or not use terms found within
+would be a developer's likelihood to use terms found within
 our dictionaries and training corpus in their commit messages. Furthermore, a top developer, in
@@
 project~\cite{largechanges} that there was a good correlation between the words used to describe a commit
-and the actual author of the commit. 
+and the author of the commit. 
 %were often correlated. Thus based
@@
 \subsection{Annotation Observations}
-We found many topics that were not actually non-functional requirements (NFRs) but were often related to them. 
+We found many topics that were not non-functional requirements (NFRs) but were often related to them. 
 For instance, concurrency was mentioned often in the commit logs and
@@
 % often  was troublesome. 
-\emph{Configuration management} and \emph{source control} related changes appeared often; % and sometimes there were topics dedicated to configuration
+Topics related to \emph{Configuration management} and \emph{source control} appeared often; % and sometimes there were topics dedicated to configuration
 % management. 
@@
 and then evaluated each other's annotations. 
+Table \ref{tab:interr} describes the Cohen Kappa and the Spearman
+correlation of our per-topic annotations for each NFR. We 
+evaluated inter-rater reliability using each NFR, because a single topic could be
+tagged with more than one NFR.
 
@@
 
-Table \ref{tab:interr} describes the Cohen Kappa and the Spearman
-correlation of our per topic annotations for each NFR. We had to
-evaluate inter-rater reliability this way because a single topic could be
-tagged with more than one NFR.
-
 We feel these results are poor. The aggregate view of a Kappa of
@@
 there was a lot of agreement in terms of lack of a tag, but often more
-disagreement when a tag should be applied. After some discussion we concluded that usability was a source of
+disagreement when a tag should be applied. After some discussion we concluded that 
+usability was a primary source of
 disagreement. Do user manuals improve usability? Is a command-line
 option a usability issue? These kinds of questions illustrate some of
-the agreement, disagreement and ambiguity about these labels. 
-Thus we would recommend that future annotators make a concerted effort
-to train and discuss when one applies an annotation.
-
-Empirically how poor were our results? We decided to compare our
-results against simulated or random results. In the first simulation
+the agreement, disagreement, and ambiguity about these labels. 
+We therefore recommend that future annotators train and discuss how and when an annotation is appropriate.
+
+To evaluate our results empirically, we compared them the annotations that would result if the seven labels
+ were applied randomly. In the first simulation
 we sampled with replacement from our own NFR label distributions to
 produce random ratings that looked like our own. We then applied the
-Kappa-statistic on this sampling versus our labelling, and repeated
-this 100,000 times, we found that
-for 4 out of 7 NFR labels, our labels IRR were greater than the random
-IRR measures 96\% of the time. For portability and efficiency, our IRR
+Kappa-statistic on this sampling, compared it with our labelling, and repeated
+this 100,000 times. We found that
+for 4 out of 7 NFR labels, the IRR values for our labels were greater than the random
+IRR measures 96\% of the time. For \emph{portability} and \emph{efficiency}, our IRR
 was greater 100\% of the time both for Neil's ratings, Abram's ratings
 and the union of both Neil and Abram. Figure \ref{fig:irr} depicts the IRR of
-these sampled ratings versus our IRR using the Kappa statistic.
-
-We also achieved similar results if the ratings were pulled from a
-uniform distributions. What this does indicate is that our labels and
-ratings were definitely better than random except for functionality,
-reliability and usability.
+these sampled ratings versus our IRR using the Kappa statistic. We also achieved similar results if the ratings were pulled from a
+uniform distributions. These experiments indicate that our labels and
+ratings were an improvement on random annotations in all cases, with the exceptions of \emph{functionality},
+\emph{reliability} and \emph{usability}.
 
@@
 
-We feel that this resulting IRR score, although lackluster, indicates
-the underlying difficulty of annotating such a data-set.
-Consistent training of the annotators (the people) might avoid low IRR scores, or
-using more annotators.
+We feel that this resulting IRR score indicates
+the underlying difficulty of annotating such a data-set. To improve IRR,
+consistent training of the annotators or the use of more annotators might help.
 Other improvements could be gained by approaching some of the original developers
-and using their expertise to annotate the data according to their intent.
-Thus improvements
-could be gained by producing more robust training samples that use
+and using their expertise to annotate the data according to their original, recollected intent.
+Improvements
+could be therefore be gained by producing more robust training samples that use
 multiple raters, or experts, who are similarly trained and willing to discuss and

### 2012-01-19 ese/ese.tex
finished migod's comments\nremoved a redundant section on IRR

@@
 %Our manual annotations were fairly quick to do, taking only a few minutes per 30 day period. 
-Creating annotated topics (i.e., manual labels) for training is painstaking, but with a suitably representative set of topics, the effort is acceptable. To
+Creating annotated topics (i.e., manual labels) for training is painstaking, but with a suitably representative set of topics, we feel that the effort is acceptable for non-academic use. To
 annotate \emph{all} topics took us approximately 20 hours per project, but we estimate only 10\% of the topics need annotation to produce useful results.
@@
 
-We found that the multi-label learners of BR, CLR and HOMER only did
+We found that the multi-label learners of BR, CLR and HOMER performed only 
 as well or worse for Macro-ROC as the single-label Naive Bayes and other naive Bayes-derived learners. 
@@
 
-Our work faced multiple threats to validity and we attempted to address many of them:
-
-\emph{Construct validity} -- we used only commit messages rather than
+Our work faced multiple threats to validity and we have attempted to address them:
+
+\paragraph{Construct validity} -- we used only commit messages rather than
 mail or bug tracker messages. 
@@
 
-Developer style is another confounding issue, if we are searching for
-these words we are relying on developers to use these words. This
+Developer style is another confounding issue; if we are searching for
+words in a word-list, we are relying on developers to use these words. This
 study might be exploiting the behaviour or style of a few
@@
 
-\emph{Internal validity} -- %includes inter-rater reliability and our lack of rival explanation building. 
+\paragraph{Internal validity} -- %includes inter-rater reliability and our lack of rival explanation building. 
 We improved internal validity by trying to correlate and explain the behaviours observed in the analysis with the historical records of the projects.
@@
 low ROC and F-Measure scores.  Performance was not uniformly low,
-often the more common classes exhibited better performance than the
+as the more common classes often exhibited better performance than the
 more rare classes.  Often this was amplified by class imbalance. In
@@
 performance could be due to a lack of coherence in tagging which was
-shown by our low inter-rater reliability score.
-
-We assessed inter-rater reliability (IRR), a measure of the correlation between the manual annotations the first two authors created. Perfect IRR (1.0) indicates complete agreement 
-on the appropriate annotation for that topic. We verified IRR using
-manually annotated PostgreSQL topics, and achieved a weak result of $0.1$.
-Some of the reasons for this are a disagreement over the appropriate classification for documentation topics (should this be considered
-usability or portability?). A short education session on what a given
-NFR from the ISO standard encompasses would reduce this error rate.
-Weak IRR results could also be responsible for some the low ROC
-performance, if each rater is picking up on different information it
-might make it more difficult to train a machine learner to do the
-same.
-Furthermore weak IRR results threatens the validity of the
-experiments, if raters cannot agree then consistently labelling by
-ISO-9126 related NFRs could be quite difficult.
-
-\emph{External validity} -- %was addressed by our use of two case studies, but there are still issues. 
+shown by our low inter-rater reliability score. Inter-rater reliability is a threat, and we discussed it in Section \ref{sec:inter-rater}. 
+% We assessed inter-rater reliability (IRR), a measure of the correlation between the manual annotations the first two authors created. Perfect IRR (1.0) indicates complete agreement 
+% on the appropriate annotation for that topic. We verified IRR using
+% manually annotated PostgreSQL topics, and achieved a weak result of $0.1$.
+% Some of the reasons for this are a disagreement over the appropriate classification for documentation topics (should this be considered
+% usability or portability?). A short education session on what a given
+% NFR from the ISO standard encompasses would reduce this error rate.
+% Weak IRR results could also be responsible for some the low ROC
+% performance, if each rater is picking up on different information it
+% might make it more difficult to train a machine learner to do the
+% same.
+% Furthermore weak IRR results threatens the validity of the
+% experiments, if raters cannot agree then consistently labelling by
+% ISO-9126 related NFRs could be quite difficult.
+
+\paragraph{External validity} -- %was addressed by our use of two case studies, but there are still issues. 
 Our data originated from OSS database projects and thus might not be applicable to commercially developed software or other domains. 
@@
 % in Case Study Research: Design Case Studies
-\emph{Reliability} -- each annotator, the first two authors, followed the same protocol and used the same annotations. 
+\paragraph{Reliability} -- each annotator, the first two authors, followed the same protocol and used the same annotations. 
 However, only two annotators were used; their annotations exhibit some
@@
 PostgreSQL. Inter-rater reliability could not be checked for MySQL and
-MaxDB because annotators did not rate the same documents.
+MaxDB because annotators did not rate the same documents. This is discussed at length in Section \ref{sec:inter-rater}
 

### 2012-01-19 ese/ese.tex
minor tweak in intro
Updated clustering

--- 2012-01-19 ese/ese.tex
@@
 %they mine requirements documents whereas we mine revisions.
-One approach they tried was similar to this one, with keywords mined from NFR catalogues found in Chung et al.~\cite{chung99}. 
+One approach they tried was similar to this one, with keywords mined from NFR catalogues found in Chung et al.~\cite{corollary99}. 
 Their approach resulted in a recall of $80\%$ with precision of $57\%$ for the \emph{security} NFR, but could not find a reliable source of keywords for other NFRs. 
@@
 The labels are reasonable enough, but still require an expert in the field to determine them. 
-Furthermore, these labels are project specific, because they are
+Furthermore, these labels are project-specific, because they are
 generated from the data of that project. For example we might have a label called \emph{`Oracle'}
@@
 %This paper does this differently. 
-Our approach differs in two important ways: we automate the process of naming the topics, and we label topics with project-independent terms, in order
+Our approach differs: first of all, we automate the process of naming the topics; secondly, we label topics with project-independent terms, in order
 to permit cross-project comparison.
@@
 We used three different open-source, partially-commercial database systems:
-\begin{description}
-\item [ MySQL 3.23] --- Started in 1994 and MySQL 3.23 was released in early 2001. MySQL contains $320,000$ lines of C and C++ source code~\footnote{generated
+\begin{itemize}
+\item  MySQL 3.23. Started in 1994 and MySQL 3.23 was released in early 2001. MySQL contains $320,000$ lines of C and C++ source code~\footnote{generated
 using David A. Wheeler's \emph{SLOCCount},
@@
 31st, 2000 to August 9th, 2004.
-\item [MaxDB 7.500] --- Started in the late 1970s as a research project, and was later acquired by SAP. As of version 7.500, released April 2007, the project
+\item MaxDB 7.500. Started in the late 1970s as a research project, and was later acquired by SAP. As of version 7.500, released April 2007, the project
 has over $940,000$ lines of C source
@@
 29th, 2004 to June 19th, 2006.
-\item [PostgreSQL 7.3] --- Started in the 1980s as a Berkeley research
+\item PostgreSQL 7.3. Started in the 1980s as a Berkeley research
   project~\footnote{{http://www.postgresql.org/docs/7.3/static/}}. PostgreSQL
@@
 
-\end{description}
+\end{itemize}
 
@@
 would allow for cross-project comparison. At the same time we
-recognize that problem domain alone does not guarantee functional and
+recognize that a problem-domain alone does guarantee functional and
 non-functional similarity.
-For instance, each database system has a different focus. PostgreSQL
+For instance each database system has a different focus, PostgreSQL
 tends to focus on fulfilling much of the SQL92 specification and adding
-more features, while MySQL has been slow to adopt much of that specification in favour of community-requested features.
-The consequence of our choice to look at a
-single domain is to limit generalizability. What we show in this paper
+more features while MySQL has been slow to adopt much of the spec.
+ As a consequence our choice of looking at a
+single domain limits our generalizability. What we show in this paper
 will be somewhat biased towards database software.
 
-For each project, we studied source control commit comments, the messages
+For each project, we used source control commit comments, the messages
 that programmers write when they commit revisions to a source control
 repository. 
-Most commits we observed had comments: 90\% in MySQL 3.23,
-98.5\% in PostgreSQL and 99.99\% in MaxDB 7.500.
+Most commits we observed had commit comments (90\% in MySQL 3.23,
+98.5\% in PostgreSQL and 99.99\% in MaxDB 7.500).
 Commit comments are often studied by researchers, as they are the most readily accessible source of project interactions, and developers are often
@@
 
-For the commit message data-sets of each project, we created an XML file that partitioned commits into 30 day periods. 
+For the commit message data-sets of each project, we created an XML file that separated commits into 30 day periods. 
 %This size of period, 30 days, is smaller than the time between minor
@@
 If there are not $N$ independent word distributions, the topics
-produced tend to be duplicates of each other, that is, they share the
-top terms. During this study we found that if we used a value of 20 for $N$, duplicate topics were infrequent.
+produce tend to be duplicates of each other, that is they share the
+top terms. During this study we found that when $N$ was 20 we did not
+witness duplicate topics very often.
 Linear combinations of these $N$ word distributions are meant to represent and recreate the word distributions of any of the original messages. 
-In other words, these topics are cross-cutting collections of words relevant to one or more of our commit messages.
+These $N$ word distributions effectively form \emph{topics}: cross-cutting collections of words relevant to one or more of our commit messages. 
 LDA extracts topics in an unsupervised manner; the algorithm relies
 solely on the source data and word distributions of messages, with no human intervention.
-%We applied Blei's LDA implementation~\cite{Blei2003} against the word distributions of these commits, and generated lists of topics per period.
-
-% In topic analysis a single document, such as a commit message, can be related to multiple topics. 
-% Representing documents as a mixture of topics maps well to source code repository commits, which often have more than one purpose~\cite{Hindle09ICSM}.  
-% For this paper, a topic represents a word distribution generated from a group of commit log comments which are related by their content.  
+
+In topic analysis a single document, such as a commit message, can be related to multiple topics. 
+Representing documents as a mixture of topics maps well to source code repository commits, which often have more than one purpose~\cite{Hindle09ICSM}.  
+For this paper, a topic represents a word distribution generated from a group of commit log comments which are related by their content.  
 %In this paper a topic is a set of tokens extracted from commit messages found within a project's source control system (SCS).
 
-%We set the number of topics to generate to $20$, because past experimentation showed that fewer topics might aggregate multiple unique topics while
-%any more topics dilutes the results and creates indistinct topics~\cite{Hindle09ICSM}. 
+We applied Blei's LDA implementation~\cite{Blei2003} against the word distributions of these commits, and generated lists of topics per period. 
+We set the number of topics to generate to $20$, because past experimentation showed that fewer topics might aggregate multiple unique topics while
+any more topics dilutes the results and creates indistinct topics~\cite{Hindle09ICSM}. 
 % Furthermore, more than $20$ topics quickly became infeasible for inspection.
@@
 
-To facilitate cross-project comparison, we used a taxonomy of NFRs.
-This taxonomy is based on the ISO quality model, ISO9126~\cite{iso9126}, which describes six high-level NFRs: \emph{maintainability, functionality,
-portability, efficiency, usability,} and \emph{reliability}. 
+To facilitate cross-project comparison, we used a taxonomy of NFRs. This taxonomy is based on the ISO quality model, ISO9126~\cite{iso9126}. 
+ISO9126 describes six high-level NFRs: \emph{maintainability, functionality,
+portability, efficiency, usability,} and \emph{reliability}
 %\footnote{While there may be lingering debate in some circles about these terms, an ISO standard
 %seems like a reasonable starting point for our work.}.
-There is some debate about the terms in this model \cite{Boegh2008}, and whether they are a) the correct terms and b) correctly organized. 
+There is some debate about the terms in this model, and whether they are a) the correct terms and b) correctly organized. 
 However, ISO9126 is ``an international standard and thus provides an 
 internationally accepted terminology for software quality
-\cite[p. 58]{Boegh2008},'' and so we consider that it is sufficient for the purposes of this
+\cite[p. 58]{Boegh2008},'' that is sufficient for the purposes of this
 research.  
-\emph{Performance} is an example of an RDBMS word related to the \emph{efficiency} NFR.
-We claim that these NFRs are maintenance concerns (to varying degrees) in all software projects, and are therefore well suited for comparisons between
+\emph{Performance} is an example of RDBMS word related to the \emph{efficiency} NFR.
+We claim that these NFRs are maintenance concerns (to varying degrees) in all software projects, and are therefore well-suited for comparisons between
 projects.
@@
 the six NFR labels listed above.
-Annotators did not annotate each other's annotations
-%NEIL --> not really! also makes IRR confusing., but some brief
-%inspection of annotations was used to confirm that the annotators were acting similarly. 
-PostgreSQL (PgSQL) was annotated by both annotators in order
+Annotators did not annotate each other's annotations, but some brief
+inspection of annotations was used to confirm that the annotators were
+acting similarly. PostgreSQL (PgSQL) was annotated by both annotators in order
 to evaluate inter-rater reliability.
-We looked at each period's topics, and assessed what the data---consisting of the frequency-weighted word lists and messages---suggested was the most appropriate labels for that topic. 
-We selected the appropriate labels using auxiliary information as well, such as the actual revisions and files that were related to the
+We looked at each period's topics, and assessed what the data ---
+consisting of the frequency-weighted word lists and messages ---
+suggested was the label for that topic. 
+We were able to pinpoint the appropriate label using auxiliary information as well, such as the actual revisions and files that were related to the
 topic being annotated.
@@
 \emph{portability}. 
-Given the top-level annotations of \emph{portability},
+Given the top-level annotations of \emph{none}, \emph{portability},
 \emph{efficiency}, \emph{reliability}, \emph{functionality},
 \emph{usability}, and \emph{maintainability}, the annotators annotated each topic
-with the relevant label. We added a catch-all label, \emph{none}, which we used when none of the six NFRs was suitable. In some instances, we used finer-grained
+with the relevant label. Sometimes they used finer-grained
 annotations that would be aggregated up to one of these higher-level labels.
@@
 Characteristic~\cite{Fawcett2006861},
-abbreviated \emph{ROC}, sometimes called Area Under the Curve or \emph{AUC}, and the
+abbreviated \emph{ROC}, sometimes called \emph{AUC}, and the
 F-measure, which is the harmonic mean of precision and recall, i.e.,
@@
 We argue for using ROC over F-Measure because ROC suffers less from
-bias, as F-Measure often skews towards the positive
+bias than F-Measure and F-Measure often skews towards the positive
 class~\cite{flach-icml03,Forman:2010:ACS:1882471.1882479}, especially
@@
 we associate each NFR with a list
-of keywords or \emph{word-lists}, in our parlance.
+of keywords (\emph{word-lists}, in our parlance).
 These word-lists were determined a priori and were not extracted from the projects themselves, using the methodology explained below.
@@
 
-Our first word-list set, \textsf{exp1}, was generated using the ontology for software quality measurement described in Kayed et al.~\cite{5072519}, which was constructed using 80 source documents, including research papers and international standards.
-The labels we used were:
-
-\vspace{2mm}
-\parbox{0.7\textwidth}{
-\small \textsf{
+Our first word-list set, \textsf{exp1}, was generated using the ontology described in Kayed et al.~\cite{5072519}
+That paper constructs an ontology for software quality measurement using eighty source documents, including research papers and international standards. 
+The labels we used are:
+
+\begin{quotation}
+\small \noindent \textsf{
 integrity, security,
@@
 portability, flexibility, reliability.
-}}
-\vspace{2mm}
+}
+\end{quotation}
 
@@
 relevant to discussion of \emph{reliability}, but is not in the
-standard. We recognize that terms like this might be used in a
+standard. We recognize though that terms like this might be used in a
 different context with a different meaning, like code-cloning.
@@
 We did the same for the quality model produced by McCall et al.~\cite{mccall1977}. 
-We then conducted a simple random analysis of mailing list messages from an open source ecosystem, KDE.
+We then did a simple random analysis of mailing list messages from an open source ecosystem, KDE.
 Like the GNOME study we conducted in Ernst and Mylopoulos \cite{ernst10refsq}, KDE contains a suite of different products covering a variety of software categories.
@@
 Wordnet similarity is a very broad match. For example, the label \emph{maintainability} is associated with
-words \emph{ease} and \emph{ownership}, and the word \emph{performance} has a `sense' that refers to musical performances, which is obviously unrelated to software development.
+words \emph{ease} and \emph{ownership}, and the word \emph{performance} has a `sense' which refers to musical performances -- which is not related to software development.
 In general, as we proceed from word-lists in \textsf{exp1} to that in \textsf{exp3}, our lists become more generic.
@@
 
+%\noindent \textbf{Generating labels automatically} ---
+
 Using our three word-lists (\textsf{exp1}, \textsf{exp2}, \textsf{exp3}), we labelled our topics with an NFR where there was a match between a word in
@@
 \emph{accuracy} (3/0/27), which were infrequently matched.
-Note the significantly higher result for \emph{usability} in PostgreSQL---this suggests a difference in how project developers are discussing usability, at least with respect to our analysis.
+Note the significantly higher result for \emph{usability} in PostgreSQL -- this suggests a difference in how this project is discussing usability, at least with respect to our analysis.
 We also looked for correlations between our labels: excluding double
@@
 For \textsf{exp3}, we generally labelled more topics. 
-As mentioned above, the word-lists are broad, so there are likely to be false-positives (discussed below). 
+As we mentioned, the word-lists are broad, so there are likely to be false-positives (discussed below). 
 The most frequent label across all projects (for this wordlist, and unlike \textsf{exp1}) was \emph{usability}, and the
@@
 To assess the performance of the supervised learners, we did a 10-fold cross-validation~\cite{Kohavi1995}, a common technique for evaluating machine
-learners, where the original data is partitioned randomly into ten sub-samples, and each sample is used to test against a training set composed of the nine other samples.
+learners. 
+The original data is partitioned randomly into ten sub-samples. Each sample is used to test against a training set composed of the nine other samples.
 %One subsample is used as a training set and evaluated on the other nine sub-samples. 
 %This is repeated nine more times, with each subsample used once as the training set. 
-We discuss these results below.% in Section \ref{sec:suplabelling}.
+We have reported these results below.% in Section \ref{sec:suplabelling}.
 
@@
 % \label{sec:suplabelling}
-Because our data-set consisted of word counts we expected Bayesian techniques to perform well.
-Bayesian techniques are often used in spam filtering, which is similarly interested in word distributions.
-We tried other learners that WEKA~\cite{weka09} provides, including rule learners, decision tree learners, vector space learners, and support vector machines.  
+Because our data-set was of word counts we expected Bayesian techniques, often used in spam filtering, to perform well. 
+We tried other learners that WEKA~\cite{weka09} provides: rule learners, decision tree learners, vector space learners, and support vector machines.  
 Figure \ref{fig:best-learn-per-tag} shows the performance of the best
@@
 
-Although we used a variety of machine learners, we found that Bayesian techniques performed the best on all projects. We believe this is due to the large number of features they can handle.
-Our best-performing learners---Discriminative Multinomial Naive Bayes, Naive Bayes
-and Multinomial Naive Bayes---are all based on Bayes' theorem and all make the naive assumption
-that the features supplied are independent. 
+For all projects Bayesian techniques did the best out of a wide variety of machine learners tested, due to the large number of features they can handle.
+Our best learners, Discriminative Multinomial Naive Bayes, Naive Bayes
+and Multinomial Naive Bayes  are all based on Bayes' theorem and all
+assume, naively, that the features supplied are independent. 
 %The features we used are word counts per message. 
@@
 For example, a commit message might address \textit{reliability} in the context of \textit{efficiency}, or make a \textit{maintainability} improvement
-in the source code that relates to \textit{usability}. 
-However, traditional machine learning techniques, such as Naive Bayes, can map topics to only a single class. 
+in the source code that related to \textit{usability}. 
+However, traditional machine learning techniques, such as Naive Bayes, can only map topics to a single class. 
 The Mulan~\cite{mulan} library encapsulates several different multi-label machine learners which can label elements with multiple labels.
@@
 % more why
-Labelled topics address \emph{why} questions by associating a commit with the underlying software qualities which motivated that change.
-The \emph{how} of a change is the change itself, the software quality behind it, the
+Labelled topics address \emph{why}  because they show reasons why changes
+occurred rather than \emph{how}. 
+The \emph{how} of a change is the change itself, the purpose or
 \emph{why} is what we are after.
@@
 We motivated our investigation with three research questions:
-\renewcommand{\labelenumi}{\textbf{RQ\arabic{enumi}. }}
 \begin{enumerate}
@@
 PostgreSQL using the
-Ward method of Hierarchical clustering on the 7-d space.  
-This will group developers according to a distance metric. 
+Ward method of Hierarchical clustering on the 7-dimensional space.  
+This will group developers according to the Euclidean distance metric.
+%We chose cosine distance because it is often used in IR research to
+%compare documents, in this case we are comparing the labels of topics
+%associated with each author's commits.
+The height in Figure \ref{fig:authorcluster} describes the euclidean
+distance between elements, in terms of the path 
 We clustered the authors into both $2$-clusters
@@
   \includegraphics[width=.9\textwidth]{figures/postgresql-author-cluster}
-  \caption{PostgreSQL Author Clusters: Authors clustered by cosine
-    distance of their NFR contributions. The outer rectangles (blue) reflect the 2-clusters; the inner rectangles (purple) reflect 6-clusters.}
+  \caption{PostgreSQL Author Clusters: Authors clustered by euclidean
+    distance of their NFR contributions. The outer rectangles (blue)
+    reflect the clustering of authors into 2 clusters; while the inner rectangles
+    (purple) reflect clustering of authors into 6 clusters. The height is a
+    measure of cluster similarity, Euclidean distance, where smaller
+    values are more similar. Individuals that share
+    branches are more similar to each other individuals who do
+    not. The longer the path, in terms of height, between two
+    individuals, the more different their NFR contributions are.}
 \label{fig:authorcluster}
@@
 Figure \ref{fig:authorcluster} shows that when we use $6$-clusters, 
-\texttt{petere}, \texttt{tgl} and \texttt{momjian} form their own individual
-clusters (i.e., are distant from one another in the 7-d space). 
+\texttt{petere}, one of the major contributors, is in his own cluster
+while two other major contributors, \texttt{tgl} and \texttt{momjian}
+form their own cluster (i.e., they are distant from one another in the 7-dimensional space). 
 %NEIL - not sure the relevance of this ... needs more explaining. --> (they are also the most frequent committers). 
-This is
-interesting because it means that the important developers are not
-committing code that fits the same NFR profile. In terms of the 2-clusters,
- we can see that \texttt{momjian} and \texttt{tgl} are in the
-same cluster, but \texttt{petere} is not. The most frequent
-committers do not share the same clusters, even at the coarsest level. 
-This implies that
-developers in this sample do work on different sets of NFRs and have different
+This is interesting because it means that the important developers are
+have different proportions of NFR-relevant contributions; they have different
+focuses during development. In terms of the 2-clusters, we can see that \texttt{momjian}
+and \texttt{tgl} are in the same cluster, but \texttt{petere} is
+not. The most frequent committers do not share the same clusters, even
+at the coarsest level of clustering, 2 clusters.  This implies that developers in
+this sample do work on different sets of NFRs and have different
 software quality focuses.

### 2012-01-19 data/clusters.R

@@ -10,14 +10,71 @@ authors <- matrix(tv,nrow=7,dimnames=list(labels(tv)[[1]],vv[,1]))
 #hc <- hclust(as.dist(cor(authors)),method="ward")
 #hc <- hclust(as.dist(cor(authors)),method="centroid")
 #
-hc <- hclust(as.dist(cor(authors)),method="ward")
-pdf("postgresql-author-cluster.pdf")
+
+pdf("oldpostgresql-author-cluster.pdf")
+
+#hc <- hclust(as.dist(cor(authors)),method="ward")
+#plot(hc,sub="Organized into 2 and 6 clusters",xlab="PostgreSQL Authors")
+#rect.hclust(hc,k=2,border="blue")
+##rect.hclust(hc,k=3,border="red")
+##rect.hclust(hc,k=4,border="green")
+#rect.hclust(hc,k=6,border="purple")
+
+
+hc <- hclust(as.dist(1-cor(authors)),method="ward")
+plot(hc,sub="Organized into 2 and 6 clusters",xlab="PostgreSQL Authors")
+rect.hclust(hc,k=2,border="blue")
+#rect.hclust(hc,k=3,border="red")
+#rect.hclust(hc,k=4,border="green")
+rect.hclust(hc,k=6,border="purple")
+
+
+dev.off()
+
+asqr <- cor(authors)
+asqr <- asqr^2
+hc <- hclust(as.dist(asqr),method="ward")
+plot(hc,sub="Organized into 2 and 6 clusters",xlab="PostgreSQL Authors")
+rect.hclust(hc,k=2,border="blue")
+#rect.hclust(hc,k=3,border="red")
+#rect.hclust(hc,k=4,border="green")
+rect.hclust(hc,k=6,border="purple")
+
+
+
+library(lsa)
+pdf("postgresql-author-cluster-cosine.pdf")
+hc <- hclust(as.dist(1-cosine(authors)),method="ward")
 plot(hc,sub="Organized into 2 and 6 clusters",xlab="PostgreSQL Authors")
 rect.hclust(hc,k=2,border="blue")
 #rect.hclust(hc,k=3,border="red")
 #rect.hclust(hc,k=4,border="green")
 rect.hclust(hc,k=6,border="purple")
 dev.off()
+
+# grr fixed it to this. The easiest to explain clustering
+pdf("postgresql-author-cluster.pdf")
+hc <- hclust(dist(t(authors)),method="ward")
[…]

### 2012-01-19 ese/ese.tex

--- 2012-01-19 ese/ese.tex
@@
  %Section \ref{sec:unsuplabelling} presents our technique for using word-lists to do unsupervised classification, and describes the results.
+In Section \ref{sec:suplabelling}, we manually annotate the topics, and use those annotations as training data for supervised classification.  
+%We then present visualizations of named topics and their trends over time to aid communication and analysis. 
+To demonstrate an application of labelled topic extraction, we use an exploratory case study of three open source database systems to show how named
+topics can be compared between projects  (Section \ref{sec:analysis}). 
+The paper concludes with a discussion of limitations (Section \ref{sec:limit}), and future work.
 
@@
 \end{itemize}
-
-In Section \ref{sec:suplabelling}, we manually annotate the topics, and use those annotations as training data for supervised classification.  
-%We then present visualizations of named topics and their trends over time to aid communication and analysis. 
-To demonstrate an application of labelled topic extraction, we use an exploratory case study of three open source database systems to show how named
-topics can be compared between projects  (Section \ref{sec:analysis}). 
-The paper concludes with a discussion of limitations (Section \ref{sec:limit}), and future work.
 
@@
 PostgreSQL. Inter-rater reliability could not be checked for MySQL and
-MaxDB because annotators did not rate the same documents. This is discussed at length in Section \ref{sec:inter-rater}
+MaxDB because annotators did not rate the same documents. This is discussed at length in Section \ref{sec:inter-rater}.
 

### 2012-01-19 ese/ese.tex
- [X] Fig 8 needs more explanation of how it should be read. For       example, what do values in each quadrant mean? For instance,       morjan in the top right means that s/he commits on many topics       and matches the global distribution. This implies that..

@@
 to each author we find that 25\%
- of the authors have a similarity of $0.47$ or less. In other words, for these authors (gathered to the lower left of Figure  \ref{fig:authorsim}),
- their NFR distribution does not match well with the global distribution.
+ of the authors have a similarity (Pearson correlation) of $0.47$ or less. In other words, for these authors (gathered to the lower left of Figure  \ref{fig:authorsim}),
+ their NFR distribution does not match well with the global
+ distribution.
+Authors \texttt{momjian} and \texttt{tgl} are in the top right of
+Figure \ref{fig:authorsim}
+    top right, dominate in topic count and commit count, thus they are
+    very similar to the global distribution of NFR topics.
+Figure \ref{fig:authorsim} demonstrates that although a  few authors dominate in
+the number of commits many authors exhibit different behaviour in
+terms of the NFR-relevant topics of their commits. Furthermore there
+are no authors in the top left corner, indicating that the authors who
+commit less, do not contribute with a similar NFR-topic distribution
+as \texttt{momjian} or \texttt{tgl}.
 We found that number of commits correlated with an author's similarity to the
@@
   \includegraphics[width=.9\textwidth]{figures/author-distance-from-aggregate}
-  \caption{PostgreSQL Author Similarity to the Global Distribution of NFR topics}
+  % \caption{PostgreSQL Author Similarity to the Global Distribution of
+  %   NFR topics versus Commits}
+  \caption{Author commit count versus the similarity of the NFR
+    distribution of an Author to the Global Distribution of NFR
+    topics. Similarity is Pearson correlation between the counts of
+    NFR relevant topics associated with an Author and the total counts
+    of NFR relevant topics. }
 \label{fig:authorsim}

### 2012-01-19 ese/ese.tex
- [X] Different clustering algorithms result in different trees:      how sensitive the results were to algorithm/distance measure      choices?

@@
 A macro-ROC measurement is the average ROC over the ROC values for all labels, where a micro-ROC is the average ROC over all examples that were classified. 
-For MaxDB, the macro-ROC values are undefined because of poor performance of one of the labels.%, see Figure \ref{fig:mulan}.
+For MaxDB, the macro-ROC values are undefined because of poor
+performance of one of the labels.%, see Figure \ref{fig:mulan}. 
+During cross-folds validation if the class imbalance or poor learner
+performance causes a division by zero when evaluating one of the NFR
+labels (in this case \emph{usability}), creating an undefined
+value. This undefined 
+value is propagated through to the mean of ROC per NFR label, causing
+the end calculation of macro-ROC to be undefined.
 
@@
 Using this space, we first conducted pairwise $X^2$ (chi-squared)
-tests and Kolmogorov-Smirnov tests for each developer pair. We
-found that in $3/10$ of these tests, the two developers were significantly
-different (i.e. distant in the 7-d space). This implies
+tests and Kolmogorov-Smirnov tests for each developer 
+ pair on their
+NFR distributions
+(18 developers, 306 pair-wise tests).
+We
+found that in $27\%$ of these tests, the two developers were significantly
+different (i.e. distant in the 7-d space and have significant p-values less than $0.05$). This implies
 that there are developers who focus on different NFRs or different
-proportions of NFRs.
+proportions of NFRs, but the majority of developers have similar NFR distributions.
 
@@
 while two other major contributors, \texttt{tgl} and \texttt{momjian}
-form their own cluster (i.e., they are distant from one another in the 7-dimensional space). 
+form their own cluster (i.e., they are distant from one another in the
+7-dimensional space). 
 %NEIL - not sure the relevance of this ... needs more explaining. --> (they are also the most frequent committers). 
@@
 \begin{table}
+\centering
 \begin{tabular}{l|c}
@@
 \end{table}
+
+Based on Figure \label{fig:authorsim} we wanted to see which clusters
+\texttt{momjian}, \texttt{tgl}, and \texttt{petere} would be in if we
+evaluated the clusters proportionately, that is if we use distance
+measures that normalize the data and ignore magnitude (e.g. turning an
+author NFR contributions into a histogram or unit-vector) such as
+cosine distance, euclidean distance of unit vectors, or Pearson
+correlation. Euclidean distance uses both angle and magnitude of a
+vector. For all 3 non-magnitude distances measures we tend to see that
+the 3 largest contributors, \texttt{momjian}, \texttt{tgl}, and
+\texttt{petere} inhabit the same clusters when we use 2 clusters or 6
+clusters. This confirms the observations in
+Figure \label{fig:authorsim}, which uses Pearson correlation, that the larger contributors tend to be
+proportionally similar.
 

### 2012-01-19 ese/ese.tex
Alright I think we got that covered..
Almost done

@@
 \label{sec:analysis}
-As we mentioned in the introduction, a key issue in software maintenance is understanding \emph{why} a system has evolved the way it has~\cite{aranda09icse}. 
+%XXX Neil should we remove this cite?
+As we mentioned in the introduction, a key issue in software
+maintenance is understanding \emph{why} a system has evolved the way
+it has: Aranda et al.~\cite{aranda09icse} explored \emph{why}, in
+terms of how a bug actually occurred and was documented.
 In this section we demonstrate the value of labelled topic extraction
@@
 the NFRs that are associated with the topic. 
+The analysis in this section used Neil's annotations.
 This produced a map which associated developer name with NFR labels.
@@
 projects investigated were database systems, thus our results might
-not generalize well to other domains.
+not generalize well to other domains. Generaizability is enhanced by
+the fact that all three projects
+were maintained by different developers and different teams using
+different development processes.
 

### 2012-01-19 ese/ese.tex
- [X] MiGOD: So you had difficulty agreeing on labels.  Did it       actually make a difference to the results?  (ie not the labels       themselves but the rest of the work)

@@
 
+Were our results harmed by low IRR scores? For MaxDB and MySQL since
+each annotator (Neil or Abram) acted as the sole Oracle, the results
+for either system still stand, but construct validity is harmed in
+terms of did if an NFR actually meant the NFR that is being
+evaluated. The measured empirical performance of each system still
+stands, but the comparison between systems in terms of NFRs might be
+harmed by low IRR scores.
+
 We feel that this resulting IRR score indicates

### 2012-01-19 ese/ese-major-response.tex
ESE Major Response and TODOS update

\documentclass{article}

\usepackage{comment,fullpage}

\begin{document}
Dear EMSE Reviewers,

We would like to thank you for your thorough reviews. We addressed all
5 major revision points and we then we tried to address all of the
smaller reviewer comments, edits. We have a giant TODO file that was
compiled from the very thorough four reviews our paper received, and
we achieved, met, and integrated all of those comments, edits, and
requirements into this revision of the paper.

\begin{quotation}
 The reviewers would like to see discussion of the relationship of
 the current draft and the previous MSR 2011 paper and more detailed
 discussion of previous papers related to the current draft. 
\end{quotation}

In this new revision, within the introduction and the previous work we
addressed the provenance of this paper and clearly indicated what we
added to this paper to make it different and more thorough. We
addressed our better coverage of machine learning, our addition of the
PostgreSQL case study, our inter-rater reliability study, our author
analysis study, etc.

We also fleshed out much of the previous work that both authors had
worked on in the past relevant to this paper, in order to provide more
context.

\begin{quotation}
     The reviewers have several requests for more details of the study
   design choices (e.g., choosing projects from the same application
   domain, using ROC instead of F-Measure as your primary measure) and
   some related discussion on the threats to validity.
\end{quotation}

Throughout the paper we added more details, especially in terms of
methodology and what statistical methods we were using and how.

In the section ``Creating a Validation Corpus'' we added why we chose
ROC over F-Measure and described some of the weakness of F-Measure
based evaluation. We also motivated why we did provide both
measures. We also discussed some of these issues in threats to
validity.

Furthermore we addressed issues of low ROC and F-measure scores in
this section, in threats to validity, and in the conclusions.

We also addressed reviewer concerns about the use of a single domain.

Examples of further details include: we discussed the pre-processing
of changelog comments for LDA in better detail; we described our
clustering methodology clearer and discussed significance of many of
the comparisons; we provided some more explanation of LDA and the
topics it extracts.

\begin{comment}
         - Re: in the "Creating a Validation Corpus" we addressed why we
           used ROC over F-Measure: class imbalance leads to bias in
           F-Measure. And we cited relevant work that discussed this
           particular issue.

         - Re: we explained some reasons for low ROC scores.

         - Re: in threats to validity, internal validity, we discussed skewed classes

     - Re: we responded on page 6 and in the threats to validity about
       the issues of a single domain

    - Re: discussed LDA word distributions

    - in the clustering section we described our methodology clearer
    - we better described LDA preprocessing
\end{comment}

\begin{quotation}
  Fix writing: The reviewers point out quite some inconsistent and
  unclear writing in various places of the paper.
\end{quotation}

In an effort to clean up the paper and appease the reviewers, we
collated a big TODO list from the reviewer comments and ensured all of
the comments were addressed and all edits were executed. We feel that
cleaned up the paper and edited it very careful, with the help of the
reviewer comments of course.  As per reviewer requests we softened
some of the claims in light of low ROC and IRR scores when applicable.
We recognize there were a lot of edits suggested by the reviewers and
we did our due diligence and attended to all that we could.

\begin{quotation}
 Reviewer 3 raises that the extremely low IRR (~0.1) threatens the
   validity of the results of these experiments, and its threats to
   validity should be made more prominent.

\end{quotation}

Within this new revisions, we addressed this issue in the inter-rater
reliability section and referenced this discussion in the threats to
validity and the conclusion.

Furthermore we provided more evidence that our IRR results were mostly
better than noise, we did extensive simulations sampling our ratings
and then comparing those ratings against our ratings to find that the
majority of our labels received higher IRR scores than 96\% to 99.9\%
of the random simulations. This allows us to confidently show that our
labels were primarily better than random. But we recognize that the
average of the values is low and we address the threats this causes to
this work and what it might mean for practitioners and other
researchers when replicating this study or methodology.

We also addressed the reviewers concerns that the IRR scores would
affect our ROC and F-Measure scores. We addressed this in the IRR
section and the threats to validity.

\begin{quotation}
 Reviewer 4 raises the some observed results may simply be due to
 the differences among developer styles and not necessarily
 reflecting different types of work.  
\end{quotation}

In this new version, we addressed these issues ``Do different
developers work on different NFRs?''  and the threats to validity. We
also better explained the methodology behind comparing these authors
and the topics they are associated with.

In conclusion we addressed the reviews to fullest extent possible,
both in terms of major criticisms and minor criticisms. We feel that
this new revision is far more robust and show ease the concerns of the
reviewers.

Thanks you,

Abram, Neil, Michael and John

\end{document}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: t
%%% End:

### 2012-01-19 ese/ese.tex
spell checking

@@
 In this section we describe how to label topics based on dictionaries
-mined from sources external to the projects. We call this `semi-unsupervised' because while there is no ``training set'', we do seed the wordlists manually.
+mined from sources external to the projects. We call this `semi-unsupervised' because while there is no ``training set'', we do seed the word-lists manually.
 
@@
 As mentioned above, the word-lists are broad, so there are likely to be false-positives (discussed below). 
-The most frequent label across all projects (for this wordlist, and unlike \textsf{exp1}) was \emph{usability}, and the
+The most frequent label across all projects (for this worldliest, and unlike \textsf{exp1}) was \emph{usability}, and the
 least frequent label was \emph{maintainability}. This implies that our signifiers for usability in this experiment were fairly broad.
@@
 Traditional classifiers label topics with a single class, whereas Mulan allows for a mixture of classes per topic, which is what we observed while
-manually labelling topics. For example, a given topic (word distribution) may be `about' both usability and maintainability, if this topic was a product of a discussion on design tradeoffs.
+manually labelling topics. For example, a given topic (word distribution) may be `about' both usability and maintainability, if this topic was a product of a discussion on design trade offs.
 The \emph{features} we used are word counts/occurrence	per topic, if
@@
 reason for this lack of performance could be that parameter we chose for number of
-topics, $N$, could be non-optimal for PostgreSQL. One explanation is that given the size of the PostgreSQL dataset, it was becoming hard to distinguish one topic
+topics, $N$, could be non-optimal for PostgreSQL. One explanation is that given the size of the PostgreSQL datasets, it was becoming hard to distinguish one topic
 from the next. PostgreSQL's data-set was the largest of the 3 projects:
-the XML datafile that described the PostgreSQL topics that we
+the XML datable that described the PostgreSQL topics that we
 annotated was 8 times larger than MySQL and 1.7 times larger than MaxDB. These
@@
 topics, although this visualization can be generated from the
-results of labelled topic extraction. Note that there are no unlabelled topics in this dataset.
+results of labelled topic extraction. Note that there are no unlabelled topics in this data-set.
 
@@
 The development period of MySQL we studied  (Figure \ref{fig:mysql-timeline}) saw the first releases to be licensed under the GPL. 
-Version 3.23.31 (January, 2001) was the production release (non-beta), and the timeline view shows a flurry of topics labelled with \emph{functionality} and
+Version 3.23.31 (January, 2001) was the production release (non-beta), and the timberline view shows a flurry of topics labelled with \emph{functionality} and
 \emph{maintainability}. 
@@
 %Within this section we focused on PostgreSQL.
-We explored this issue by extracting, for each topic in the PostgreSQL dataset, the developer names responsible for the commits used to generate that topic, and 
+We explored this issue by extracting, for each topic in the PostgreSQL data-set, the developer names responsible for the commits used to generate that topic, and 
 the NFRs that are associated with the topic. 
@@
 projects investigated were database systems, thus our results might
-not generalize well to other domains. Generaizability is enhanced by
+not generalize well to other domains. Generalizability is enhanced by
 the fact that all three projects

### 2012-01-19 ese/ese.tex
minor edit
latest greatest edits?
minor fixes and typos

@@
 insight into their evolving software development activities.
-%160 words
+%173 words
 \keywords{ Software maintenance \and Repository mining \and Latent Dirichlet allocation \and Topic models}
@@
 
-Topic modeling (such as Latent Dirichlet Allocation ~\cite{Blei2003}) is a machine learning technique that creates
+Topic modeling (such as Latent Dirichlet Allocation~\cite{Blei2003}) is a machine learning technique that creates
 multinomial distributions of words extracted from a text corpus. 
@@
 product. In this figure, we show the relative differences when we use
-different training data sets from different annotators. In \ref{fig:pgcomp-a} we see that ROC
+different training data-sets from different annotators. In \ref{fig:pgcomp-a} we see that ROC
 results are very similar. In the other figure, however, author \#1 has
@@
 
-\subsection{RQ1: Do NFR Frequencies Change over Time?}
+\subsection{RQ1: Do NFR Frequencies Change Over Time?}
 In both MaxDB and MySQL the frequencies generally decreased with age. 
@@
 
-Based on Figure \label{fig:authorsim} we wanted to see which clusters
+Based on Figure \ref{fig:authorsim} we wanted to see which clusters
 \texttt{momjian}, \texttt{tgl}, and \texttt{petere} would be in if we
@@
 measures that normalize the data and ignore magnitude (e.g. turning an
-author NFR contributions into a histogram or unit-vector) such as
-cosine distance, euclidean distance of unit vectors, or Pearson
+author's NFR contributions into a histogram or unit-vector) such as
+cosine distance, Euclidean distance of unit vectors, or Pearson
 correlation. Euclidean distance uses both angle and magnitude of a
-vector. For all 3 non-magnitude distances measures we tend to see that
-the 3 largest contributors, \texttt{momjian}, \texttt{tgl}, and
-\texttt{petere} inhabit the same clusters when we use 2 clusters or 6
-clusters. This confirms the observations in
-Figure \label{fig:authorsim}, which uses Pearson correlation, that the larger contributors tend to be
+vector. For all three non-magnitude distances measures we tend to see that
+the three largest contributors, \texttt{momjian}, \texttt{tgl}, and
+\texttt{petere} inhabit the same clusters when we use 2-clusters or 6-clusters. This confirms the observations in
+Figure \ref{fig:authorsim}, which uses Pearson correlation, that the larger contributors tend to be
 proportionally similar.
@@
 % often  was troublesome. 
-Topics related to \emph{Configuration management} and \emph{source control} appeared often; % and sometimes there were topics dedicated to configuration
+Topics related to \emph{configuration management} and \emph{source control} appeared often; % and sometimes there were topics dedicated to configuration
 % management. 
@@
 
-We feel these results are poor. The aggregate view of a Kappa of
+These results are fairly poor. The aggregate view of a Kappa of
 $0.1$ indicates there is some weak agreement. We found that
-there was a lot of agreement in terms of lack of a tag, but often more
-disagreement when a tag should be applied. After some discussion we concluded that 
-usability was a primary source of
-disagreement. Do user manuals improve usability? Is a command-line
+there was good agreement in terms of lack of an annotation, but
+disagreement regarding which annotation to apply. After some discussion we concluded that 
+\emph{usability} was a primary source of
+disagreement. For instance, should we annotate a commit which updates the user manual as a usability-related change? Is adding a command-line
 option a usability issue? These kinds of questions illustrate some of
 the agreement, disagreement, and ambiguity about these labels. 
-We therefore recommend that future annotators train and discuss how and when an annotation is appropriate.
-
-To evaluate our results empirically, we compared them the annotations that would result if the seven labels
+We therefore recommend that future annotators train and discuss how and when an annotation is appropriate. This ought to be
+easier to do if an internal team is using our tool, since (presumably) there is a greater shared understanding of project activities, and therefore,
+potentially greater rater agreement.
+
+To evaluate our results empirically, we compared them to the annotations that would result if the seven labels
  were applied randomly. In the first simulation
@@
 was greater 100\% of the time both for Neil's ratings, Abram's ratings
-and the union of both Neil and Abram. Figure \ref{fig:irr} depicts the IRR of
-these sampled ratings versus our IRR using the Kappa statistic. We also achieved similar results if the ratings were pulled from a
+and the union of both Neil and Abram. One explanation of the positive result for \emph{portability} is that such changes are often 
+accompanied by clear indicator terms such as \emph{Windows, Linux, SunOS}, etc.
+
+Figure \ref{fig:irr} depicts the IRR of
+these sampled ratings versus our IRR using the Kappa statistic. We also achieved similar results if the ratings were pulled from
 uniform distributions. These experiments indicate that our labels and
@@
   \includegraphics[width=0.9\textwidth]{figures/self-sample-Both}
-  \caption{Measured IRR versus IRR of labeling random simulations.}
+  \caption{Measured IRR versus IRR of labeling random simulations. Perfect IRR is 1.0. }
   \label{fig:irr}
@@
 
-Were our results harmed by low IRR scores? For MaxDB and MySQL since
-each annotator (Neil or Abram) acted as the sole Oracle, the results
-for either system still stand, but construct validity is harmed in
-terms of did if an NFR actually meant the NFR that is being
-evaluated. The measured empirical performance of each system still
+Were our results harmed by low IRR scores? For MaxDB and MySQL, since
+each annotator (Neil or Abram) acted as the sole oracle, the results
+for either system still stand, but construct validity is harmed since 
+it is not clear if the annotated NFR actually represented the NFR relevant to that commit.
+ The measured empirical performance of each system still
 stands, but the comparison between systems in terms of NFRs might be
@@
 
-We feel that this resulting IRR score indicates
+The low IRR scores indicate
 the underlying difficulty of annotating such a data-set. To improve IRR,

### 2012-01-19 ese/ese.tex
abram hates commas

@@
 Authors \texttt{momjian} and \texttt{tgl} are in the top right of
-Figure \ref{fig:authorsim}
-    top right, dominate in topic count and commit count, thus they are
+Figure \ref{fig:authorsim}, dominate in topic count and commit count, and so are
     very similar to the global distribution of NFR topics.
 Figure \ref{fig:authorsim} demonstrates that although a  few authors dominate in
-the number of commits many authors exhibit different behaviour in
-terms of the NFR-relevant topics of their commits. Furthermore there
+the number of commits, many authors exhibit different behaviour in
+terms of the NFR-relevant topics of their commits. Furthermore, there
 are no authors in the top left corner, indicating that the authors who

### 2012-01-19 ese/ese.tex
explained IRR a little better

@@
 To evaluate our results empirically, we compared them to the annotations that would result if the seven labels
- were applied randomly. In the first simulation
+ were applied randomly. In the first simulation, as shown by Figure \ref{fig:irr},
 we sampled with replacement from our own NFR label distributions to
 produce random ratings that looked like our own. We then applied the
-Kappa-statistic on this sampling, compared it with our labelling, and repeated
-this 100,000 times. We found that
+Kappa-statistic on this sampling and our labelling, and repeated
+this 100,000 times. We then compared this distribution of simulated
+IRR ratings against our our IRR scores. We found that
 for 4 out of 7 NFR labels, the IRR values for our labels were greater than the random
@@
 was greater 100\% of the time both for Neil's ratings, Abram's ratings
-and the union of both Neil and Abram. One explanation of the positive result for \emph{portability} is that such changes are often 
+and the union of both Neil and Abram. 
+We know this by evaluating our IRR ratings against the empirical
+cumulative distribution function (ECDF) of the IRR ratings of the 100,000
+simulation runs.
+One explanation of the positive result for \emph{portability} is that such changes are often 
 accompanied by clear indicator terms such as \emph{Windows, Linux, SunOS}, etc.
@@
 Figure \ref{fig:irr} depicts the IRR of
-these sampled ratings versus our IRR using the Kappa statistic. We also achieved similar results if the ratings were pulled from
+these simulated sampled ratings versus our IRR using the Kappa statistic. We also achieved similar results if the ratings were pulled from
 uniform distributions. These experiments indicate that our labels and
@@
   \includegraphics[width=0.9\textwidth]{figures/self-sample-Both}
-  \caption{Measured IRR versus IRR of labeling random simulations. Perfect IRR is 1.0. }
+  \caption{Measured IRR versus IRR of random labeling
+    simulations. Perfect IRR is 1.0.
+    The red line and points highlight the inter-rater reliability of
+    NFR topics labels
+    measured on PostgreSQL. 
+    The box-plots are the distribution of 100,000 IRR ratings between
+    random simulation, drawn from our distribution of ratings, and our
+    labelling of NFRs.
+    Note how for 4 of the 7 NFRs it is
+    distinctly higher than the median of the simulations.
+  }
   \label{fig:irr}

### 2012-01-19 src/validate/interrater.R

@@ -108,9 +108,9 @@ ratings <- list()
 ratings[["Abram"]] <- pa
 ratings[["Neil"]] <- pn
 ratings[["Both"]] <- paunion
-panames <- c("Portability","Functionality","Reliability","Maintainability","Effeciency", "Usability", "None")
+panames <- c("Portability","Functionality","Reliability","Maintainability","Efficiency", "Usability", "None")
 
-for (person in c("Abram","Neil","Both")) {
+for (person in c("Both","Abram","Neil")) {
 pan <- ratings[[person]]
 i <- names(pan)[1]
 #
@@ -126,6 +126,7 @@ i <- names(pan)[1]
 #versus a sample
 
 N <- 100000
+#N <- 1000
 kappas <- lapply(names(pa),function(acolumn) {
   unlist(
          mclapply(
@@ -147,6 +148,7 @@ pankappas <- sapply(names(pa),function(i) {
   boxplot(kappas, names=panames,ylim=c(miny,maxy))
   title(paste("Human Ratings (",person,") versus Sampled Random Ratings"))
   lines(pankappas, type="o",col="red",lw=3)
+  legend("topleft", c("Interrater Reliability per NFR Topic Label"), bty="o",lwd=3, col=c("red"),  lty=c(1));
   dev.off()
   print(paste("Person ", person))
   print(sapply(kappas,mean))

### 2012-01-19 ese/ese.tex
edits before class

@@
 % edit this. awkward
-In this work we provide a method of automatically labelling  development topics extracted from commit logs called \emph{labelled topic extraction}.
+In this work we provide a method of automatically labelling
+development topics extracted from commit logs, this method is called \emph{labelled topic extraction}.
 
@@
 time --- we observed that topic trends often corresponded to
-non-functional requirements (NFRs)~\cite{Hindle09ICSM}, which is
+\emph{non-functional requirements} (NFRs)~\cite{Hindle09ICSM}, which is
 further emphasized in this paper due to the large numbers of NFR
@@
 these annotations. 
-	\item We used this new case-study to test inter-rater reliability, described in Section \ref{sec:inter-rater}. 
+	\item We used this new case-study to test \emph{inter-rater
+          reliability} (IRR), described in Section \ref{sec:inter-rater}. 
 	\item We conducted an analysis of PostgreSQL authors and their association with NFRs and
@@
 % NE 0113 this seems too early to mention 'result'.
-In Hindle et al. \cite{Hindle2007} we examined release patterns in OSS. That work showed that there is a difference between projects regarding maintenance
+In Hindle et al.~\cite{Hindle2007} we examined release patterns in OSS. That work showed that there is a difference between projects regarding maintenance
 techniques. This supports the result described in this paper, that software qualities are not discussed with the same frequency across projects.

### 2012-01-20 ese/ese.tex
edits up to supervised labelling

@@
 repositories and used these 
-in order to analyze the evolution of a single product.
+to analyze the evolution of a single product.
 Change requests in Jazz allow for users to annotate each change
@@
 
+% Did a little validation here
 Table \ref{tbl:wordlist} shows how many topics were labelled for
@@
 unlabelled topics, PostgreSQL also was under going far more
-development as we evaluated 20330 PostgreSQL commits versus 8664 MaxDB
-commits and 12056 commits. We suspect that PostgreSQL topics were
-flooded with many terms and perhaps we needed more 20 topics per month
+development as we evaluated 7744 PostgreSQL commits versus 8664 MaxDB
+commits and 6223 MySQL commits. 
+In terms of change-log comment words PostgreSQL had 164724, while
+MaxDB had 68203, and MySQL had 101081. This implies there were
+more PostgreSQL
+terms for the topic analyzer to produce topics with.
+And thus we suspect that PostgreSQL topics were
+flooded with many terms, indicating that perhaps we needed more 20 topics per month
 for PostgreSQL as this result indicates there is a lot of overlap in
 the topics.
+
+%abez@burn:~/projects/query1-data/postgresql$ sort -u query1.mrs2 | egrep "\|200[234]\-[0-9]" | awk -F\| '{print $5}' | wc
+%   7744  164724 1340914
+%abez@burn:~/projects/query1-data/maxdb-7500/dist$ sort -u query1.mrs2 | awk -F\| '{print $5}' | wc
+%   8664   68203  974487
+%abez@burn:~/projects/query1-data/mysql-3.23$ sort -u query1.mrs2 | awk -F\| '{print $5}' | wc
+%   6223  101081  628971
 
@@
 As mentioned above, the word-lists are broad, so there are likely to be false-positives (discussed below). 
-The most frequent label across all projects (for this worldliest, and unlike \textsf{exp1}) was \emph{usability}, and the
+The most frequent label across all projects (for this word-list, and unlike \textsf{exp1}) was \emph{usability}, and the
 least frequent label was \emph{maintainability}. This implies that our signifiers for usability in this experiment were fairly broad.
@@
 As described in Section \ref{sec:wordlist} we used both ROC and F-1 measures to evaluate the performance of the classification.
-Figure \ref{fig:maxdb-unsup-results} shows our ROC results for PostgreSQL, MaxDB and MySQL. We omit \textsf{exp1} due to poor results. We describe F-1 results in the text below.
+Figure \ref{fig:maxdb-unsup-results} shows our ROC results for
+PostgreSQL, MaxDB and MySQL. We omit plots of \textsf{exp1} due to poor results. We describe F-1 results in the text below.
 

### 2012-01-20 ese/ese.tex
latest edits

@@
 The analysis in this section used Neil's annotations.
-This produced a map which associated developer name with NFR labels.
-The six NFR labels (\emph{maintainability, usability, efficiency, functionality, portability, reliability}) and the \emph{None} label describe a 7-dimensional space, into which we can position developers, based on the frequency with which each label occurs in the map.
+This produced a map which associated developer name with NFR
+labels. The six NFR labels (\emph{maintainability, usability,
+  efficiency, functionality, portability, reliability}) and the
+\emph{None} label describe a 7-dimensional space (essentially a distribution), into which we can position developers, based on the frequency with which each label occurs in the map.
 
@@
 cosine distance, Euclidean distance of unit vectors, or Pearson
-correlation. Euclidean distance uses both angle and magnitude of a
+correlation distance ($1 - r$). Whereas, Euclidean distance uses both angle and magnitude of a
 vector. For all three non-magnitude distances measures we tend to see that
@@
 this 100,000 times. We then compared this distribution of simulated
-IRR ratings against our our IRR scores. We found that
+IRR ratings against our own IRR ratings. We found that
 for 4 out of 7 NFR labels, the IRR values for our labels were greater than the random
@@
 
-PostgreSQL's size in terms of commits was equal to MySQL and MaxDB
-combined. This imbalance in size, combined with the choice of 20
-topics per month produced topics that represented too many issues or
-threads within PostgreSQL's development. Choice of number of topics
-should probably be tuned to the project.
+PostgreSQL was larger than MySQL and MaxDB. This imbalance in size,
+combined with the choice of 20 topics per month produced topics that
+represented too many issues or threads within PostgreSQL's
+development. Choice of number of topics should probably be tuned to
+the project.
 
@@
 %XXX If we have more time try to elaborate more on this
-One potential issue with some of our results is that we achieved very
+One potential issue with some of our results is that we often achieved very
 low ROC and F-Measure scores.  Performance was not uniformly low,
@@
 more rare classes.  Often this was amplified by class imbalance. In
-future work we plan to investigate techniques such as sub-sampling and
+our future work we plan to investigate techniques such as sub-sampling and
 boot-strapping in order to improve performance. Some of the ROC

### 2012-01-20 ese/ese.tex
minor edit

@@
 characteristic curve), as our measurement to compare the effectiveness
-of semi-supervised and supervised learners. ROC is less biased than
+of semi-unsupervised and supervised learners. ROC is less biased than
 F-Measure in cross-folds

### 2012-01-20 ese/ese.tex
33% audio edit

@@
 
-%migod
+% migod
 In general, the fruits of mining software artifacts are often project
@@
 Furthermore, these labels are project specific, because they are
-generated from the data of that project. For example we might have a label called \emph{`Oracle'}
+generated from the data of that project. For example we might have a label called \emph{``Oracle''}
 in the MySQL case, since Oracle owns MySQL. 
@@
 Commit comments are often studied by researchers, as they are the most readily accessible source of project interactions, and developers are often
-required to create them by the repository mechanism (e.g., Git).  Additionally, relying only on commit comments makes our approach more generalizable,
+required to create them by the repository mechanism (e.g., CVS).  Additionally, relying only on commit comments makes our approach more generalizable,
 as we do not assume the presence of other artifact corpora.
@@
 other transforms to the messages.
-Our stop words are derived from the Natural Language Toolkit (NLTK) stop-word English stop-word
+Our stop words are derived from the Natural Language Toolkit (NLTK) English stop-word
 list~\footnote{NLTK: \url{http://www.nltk.org/}}.
@@
 To evaluate both semi-unsupervised and supervised classification, we
-created a validation set of manually labelled topics. For each project,
+created a validation set of manually labelled topics. Per each project,
  the annotators (the first two authors) annotated each extracted topic in each period with
 the six NFR labels listed above.
-Annotators did not annotate each other's annotations
+Annotators did not annotate each other's annotations, except that
 %NEIL --> not really! also makes IRR confusing., but some brief
@@
 
-We validate classification performance using the area under the curve
+We validate classification performance using the \emph{area under the curve
 of Receiver Operating
-Characteristic~\cite{Fawcett2006861},
+Characteristic}~\cite{Fawcett2006861},
 abbreviated \emph{ROC}, sometimes called Area Under the Curve or \emph{AUC}, and the
@@
 look much larger than our F-Measure scores, we feel that the knowledge
-that the random selection is $0.5$ or worse is helpful for interpreting our results.
+that random selection is $0.5$ or worse is helpful for interpreting our results.
 The %area under the -- NE confusing to have area again ... and not importantana
@@
 calculate $F_{tp,fp}$, thus we still provide ROC and $F_{avg}$.
-We consider our labelling classifiers acceptable if they outperform a random classifier (0.5).
+We consider our labelling classifiers acceptable if they outperform a random classifier ($0.5$).
 
@@
 In this section we describe how to label topics based on dictionaries
-mined from sources external to the projects. We call this `semi-unsupervised' because while there is no ``training set'', we do seed the word-lists manually.
+mined from sources external to the projects. We call this ``semi-unsupervised'' because while there is no ``training set'', we do seed the word-lists manually.
 
@@
 We then conducted a simple random analysis of mailing list messages from an open source ecosystem, KDE.
-Like the GNOME study we conducted in Ernst and Mylopoulos \cite{ernst10refsq}, KDE contains a suite of different products covering a variety of software categories.
+Like the GNOME study we conducted in Ernst and Mylopoulos~\cite{ernst10refsq}, KDE contains a suite of different products covering a variety of software categories.
 If we judged a given message to contain terms that were

### 2012-01-20 ese/ese-major-response.tex
formatting response

@@
 
-We would like to thank you for your thorough reviews. We addressed all
-5 major revision points and we then we tried to address all of the
-smaller reviewer comments, edits. We have a giant TODO file that was
-compiled from the very thorough four reviews our paper received, and
-we achieved, met, and integrated all of those comments, edits, and
-requirements into this revision of the paper.
+We would like to thank you for your thorough reviews. In our revised paper,
+we have addressed all five  major revision points. In addition, we incorporated nearly all of the
+minor reviewer comments and corrections. 
+% one exception: the timeline figure is still fairly small.
+%We have a giant TODO file that was 
+% compiled from the very thorough four reviews our paper received, and
+% we achieved, met, and integrated all of those comments, edits, and
+% requirements into this revision of the paper.
 
-\begin{quotation}
- The reviewers would like to see discussion of the relationship of
+We now explain how we have amended the paper to incorporate the major revisions requested.
+
+\subsubsection*{1. The reviewers would like to see discussion of the relationship of
  the current draft and the previous MSR 2011 paper and more detailed
- discussion of previous papers related to the current draft. 
-\end{quotation}
+ discussion of previous papers related to the current draft.}
 
@@
 
-\begin{quotation}
-     The reviewers have several requests for more details of the study
+\subsubsection*{2. The reviewers have several requests for more details of the study
    design choices (e.g., choosing projects from the same application
    domain, using ROC instead of F-Measure as your primary measure) and
-   some related discussion on the threats to validity.
-\end{quotation}
+   some related discussion on the threats to validity.}
 
@@
 
-\begin{quotation}
-  Fix writing: The reviewers point out quite some inconsistent and
-  unclear writing in various places of the paper.
-\end{quotation}
+\subsubsection*{
+  3. Fix writing: The reviewers point out quite some inconsistent and
+  unclear writing in various places of the paper.}
 
@@
 
-\begin{quotation}
- Reviewer 3 raises that the extremely low IRR (~0.1) threatens the
+\subsubsection*{4.  Reviewer 3 raises that the extremely low IRR (~0.1) threatens the
    validity of the results of these experiments, and its threats to
-   validity should be made more prominent.
-
-\end{quotation}
+   validity should be made more prominent.}
 
@@
 
-\begin{quotation}
- Reviewer 4 raises the some observed results may simply be due to
+\subsubsection*{5. Reviewer 4 raises the some observed results may simply be due to
  the differences among developer styles and not necessarily
- reflecting different types of work.  
-\end{quotation}
+ reflecting different types of work. }
 

### 2012-01-20 ese/ese.tex
formatting response

@@
 maintenance is understanding \emph{why} a system has evolved the way
-it has: Aranda et al.~\cite{aranda09icse} explored \emph{why}, in
-terms of how a bug actually occurred and was documented.
+it has. 
+% Aranda et al.~\cite{aranda09icse} explored \emph{why}, in
+%terms of how a bug actually occurred and was documented.
 In this section we demonstrate the value of labelled topic extraction

### 2012-01-20 ese/ese.tex
100% audio edit -- I switched West Indies mode at 270 words per minute -- pretty cool :O

--- 2012-01-20 ese/ese.tex
@@
 We do not include these words here for space
-considerations~\footnote{For our data repository visit \url{http://softwareprocess.es/nomen/}}. 
+considerations~\footnote{For our word lists visit \url{http://softwareprocess.es/nomen/}}. 
 Wordnet similarity is a very broad match. For example, the label \emph{maintainability} is associated with
@@
 And thus we suspect that PostgreSQL topics were
-flooded with many terms, indicating that perhaps we needed more 20 topics per month
-for PostgreSQL as this result indicates there is a lot of overlap in
+flooded with many terms, indicating that perhaps we needed more than 20 topics per month
+for PostgreSQL, as this result indicates there is a lot of overlap in
 the topics.
@@
 We also looked for correlations between our labels: excluding double
-matches (self-correlation), our highest co-occurring terms were
+matches (self-correlation), our highest co-occurring labels were
 \emph{verifiability} or \emph{correctness}
@@
 train the labelling process.
-As described in Section \ref{sec:wordlist} we used both ROC and F-1 measures to evaluate the performance of the classification.
+As described in Section \ref{sec:wordlist} we used both ROC and F-measure measures to evaluate the performance of the classification.
 Figure \ref{fig:maxdb-unsup-results} shows our ROC results for
-PostgreSQL, MaxDB and MySQL. We omit plots of \textsf{exp1} due to poor results. We describe F-1 results in the text below.
+PostgreSQL, MaxDB and MySQL. We omit plots of \textsf{exp1} due to poor results. We describe F-measure results in the text below.
 
@@
 $0.47$, and $0.36$ for MySQL with a range of $0.10$ to $0.65$. 
-PostgreSQL had an average of $0.42$ with a range of $0.31$ to $0.54$.
+PostgreSQL had an average F-measure of $0.42$ with a range of $0.31$ to $0.54$.
 MaxDB had an average precision and recall of $0.16$ and $0.67$
@@
 The best learner is important because one uses a single learner per
-label. When applying our technique, for each NFR one should select the
+label. When applying our technique, per each NFR one should select the
 best performing learner possible. Given the tractability of these techniques, a tool which applied all learners and presented the best result should be feasible. 
@@
 The development period of MySQL we studied  (Figure \ref{fig:mysql-timeline}) saw the first releases to be licensed under the GPL. 
-Version 3.23.31 (January, 2001) was the production release (non-beta), and the timberline view shows a flurry of topics labelled with \emph{functionality} and
+Version 3.23.31 (January, 2001) was the production release (non-beta), and the time-line view shows a flurry of topics labelled with \emph{functionality} and
 \emph{maintainability}. 
@@
   \includegraphics[width=.9\textwidth]{figures/postgresql-author-cluster}
-  \caption{PostgreSQL Author Clusters: Authors clustered by euclidean
+  \caption{PostgreSQL Author Clusters: Authors clustered by Euclidean
     distance of their NFR contributions. The outer rectangles (blue)
@@
     simulations. Perfect IRR is 1.0.
-    The red line and points highlight the inter-rater reliability of
-    NFR topics labels
-    measured on PostgreSQL. 
+    The red line and points highlight the measured inter-rater reliability of
+    NFR topics labels between annotators of PostgreSQL.
     The box-plots are the distribution of 100,000 IRR ratings between
@@
     labelling of NFRs.
-    Note how for 4 of the 7 NFRs it is
+    Note how for 4 of the 7 NFRs the measure IRR is
     distinctly higher than the median of the simulations.
@@
 Our taxonomy for software NFRs is subject to dispute, but seems to be
-generally accepted.  A future approach should consider a different taxonomy, such as one created by surveying developers on what ``types" of tasks they work on.
+generally accepted.  A future approach should consider a different taxonomy, such as one created by surveying developers on what ``types'' of tasks they work on.
 Finally, there are exogenous sources, such as
@@
 F-Measure if they are more comfortable with it. Thus because of
-F-Measure biased handling of class-imbalance and the inherent
+F-Measure's biased handling of class-imbalance and the inherent
 class-imbalance that our data suffers from we chose ROC.

### 2012-01-20 ese/ese-major-response.tex
minor edits

@@
 
-Within this new revisions, we addressed this issue in the inter-rater
+Within this new revision, we addressed this issue in the inter-rater
 reliability section and referenced this discussion in the threats to

### 2012-01-20 ese/ese.tex
black and white compatible

@@
   \caption{PostgreSQL Author Clusters: Authors clustered by Euclidean
-    distance of their NFR contributions. The outer rectangles (blue)
+    distance of their NFR contributions. The outer rectangles (black)
     reflect the clustering of authors into 2 clusters; while the inner rectangles
-    (purple) reflect clustering of authors into 6 clusters. The height is a
+    (dim grey) reflect clustering of authors into 6 clusters. The height is a
     measure of cluster similarity, Euclidean distance, where smaller

### 2012-01-20 data/clusters.R

@@ -56,10 +56,10 @@ dev.off()
 pdf("postgresql-author-cluster.pdf")
 hc <- hclust(dist(t(authors)),method="ward")
 plot(hc,sub="Organized into 2 and 6 clusters",xlab="PostgreSQL Authors")
-rect.hclust(hc,k=2,border="blue")
+rect.hclust(hc,k=2,border="black")
 #rect.hclust(hc,k=3,border="red")
 #rect.hclust(hc,k=4,border="green")
-rect.hclust(hc,k=6,border="purple")
+rect.hclust(hc,k=6,border="dimgrey")
 dev.off()
 
 

### 2012-01-20 ese/ese-major-response.tex
response cleaned up

@@
 Dear EMSE Reviewers,
+
+\vspace{2mm}
 
@@
 % requirements into this revision of the paper.
-
 We now explain how we have amended the paper to incorporate the major revisions requested.
@@
 
-In this new revision, within the introduction and the previous work we
-addressed the provenance of this paper and clearly indicated what we
-added to this paper to make it different and more thorough. We
+In this new revision, we have expanded the introduction and the previous work sections to address
+the provenance of this paper, and included a detailed description of the work added to 
+this paper to extend the conference publication. In particular, we highlighted how we 
 addressed our better coverage of machine learning, our addition of the
-PostgreSQL case study, our inter-rater reliability study, our author
-analysis study, etc.
+PostgreSQL case study, our inter-rater reliability study, and our author
+analysis study.
 
-We also fleshed out much of the previous work that both authors had
-worked on in the past relevant to this paper, in order to provide more
+We also expanded on the previous work that the authors had conducted 
+ which was relevant to this paper, in order to provide more
 context.
@@
 
-Throughout the paper we added more details, especially in terms of
-methodology and what statistical methods we were using and how.
+Throughout the paper we added explanations of 
+methodology choices and choice of statistical methods.
+In Section 3.1.2, ``Creating a Validation Corpus'', we added an explanation of our choice of 
+ROC over F-Measure, referencing other researchers in describing some of the weakness of F-Measure
+based evaluation. We now motivate why both measures were provided.
+We also discussed some of these issues in Section 5.4, ``Threats to
+Validity''.
 
-In the section ``Creating a Validation Corpus'' we added why we chose
-ROC over F-Measure and described some of the weakness of F-Measure
-based evaluation. We also motivated why we did provide both
-measures. We also discussed some of these issues in threats to
-validity.
+Furthermore, we now address issues of low ROC and F-measure scores in
+both 3.1.2 and 5.4, as well as in the conclusions. We added a discussion regarding concerns about the use of a single domain in section 3.1.
 
-Furthermore we addressed issues of low ROC and F-measure scores in
-this section, in threats to validity, and in the conclusions.
-
-We also addressed reviewer concerns about the use of a single domain.
-
-Examples of further details include: we discussed the pre-processing
+Examples of further amendments: we discussed the pre-processing
 of changelog comments for LDA in better detail; we described our
-clustering methodology clearer and discussed significance of many of
-the comparisons; we provided some more explanation of LDA and the
+clustering methodology more clearly, and discussed the statistical significance of
+the comparisons; and we provided some more explanation of LDA and the
 topics it extracts.
-
-\begin{comment}
-         - Re: in the "Creating a Validation Corpus" we addressed why we
-           used ROC over F-Measure: class imbalance leads to bias in
-           F-Measure. And we cited relevant work that discussed this
-           particular issue.
-
-         - Re: we explained some reasons for low ROC scores.
-
-         - Re: in threats to validity, internal validity, we discussed skewed classes
-
-     - Re: we responded on page 6 and in the threats to validity about
-       the issues of a single domain
-
-    - Re: discussed LDA word distributions
-
-    - in the clustering section we described our methodology clearer
-    - we better described LDA preprocessing
-\end{comment}
 
@@
 
-In an effort to clean up the paper and appease the reviewers, we
-collated a big TODO list from the reviewer comments and ensured all of
-the comments were addressed and all edits were executed. We feel that
-cleaned up the paper and edited it very careful, with the help of the
-reviewer comments of course.  As per reviewer requests we softened
-some of the claims in light of low ROC and IRR scores when applicable.
+We collated a big TODO list from the reviewer comments and ensured all of
+the comments were addressed and all edits were executed. 
+%We feel that cleaned up the paper and edited it very careful, with the help of the
+%reviewer comments of course.  
+As per reviewer requests we softened some of the claims in light of low ROC and IRR scores when applicable.
 We recognize there were a lot of edits suggested by the reviewers and
-we did our due diligence and attended to all that we could.
+we did our due diligence and attended to all that we could. We regret that making the figure showing timelines 
+larger would have required three pages of figures. 
+%We note that this figure is vector-based, and so zooms appropriately
+%when read on a device.
 
@@
 
-Within this new revision, we addressed this issue in the inter-rater
-reliability section and referenced this discussion in the threats to
-validity and the conclusion.
+Within this new revision, we addressed this issue in a new section on inter-rater
+reliability (5.3) and referenced this discussion in the threats to
+validity (5.4) and the conclusion.
 
-Furthermore we provided more evidence that our IRR results were mostly
-better than noise, we did extensive simulations sampling our ratings
-and then comparing those ratings against our ratings to find that the
+Furthermore, we expanded the discussion with statistical evidence that our IRR results were mostly
+better than noise. We did extensive simulations sampling our ratings,
+and then comparing those ratings against our ratings. We found that the
 majority of our labels received higher IRR scores than 96\% to 99.9\%
-of the random simulations. This allows us to confidently show that our
+of the random simulations. This allows us to confidently state that our
 labels were primarily better than random. But we recognize that the
-average of the values is low and we address the threats this causes to
-this work and what it might mean for practitioners and other
-researchers when replicating this study or methodology.
+average of the values is low, and Section 5.3 address the threats this causes to
+this paper, and what it might mean for practitioners and other
+researchers when replicating this study. Section 5.3 also addresses the reviewers concerns that the IRR scores would
+affect our ROC and F-Measure scores.
 
-We also addressed the reviewers concerns that the IRR scores would
-affect our ROC and F-Measure scores. We addressed this in the IRR
-section and the threats to validity.
-
-\subsubsection*{5. Reviewer 4 raises the some observed results may simply be due to
+\subsubsection*{5. Reviewer 4 raises the [issue that] some observed results may simply be due to
  the differences among developer styles and not necessarily
@@
 
-In this new version, we addressed these issues ``Do different
-developers work on different NFRs?''  and the threats to validity. We
+We discuss this possibility in Section 4.3, ``Do different
+developers work on different NFRs?'',  and in Section 5.3, the threats to validity. We
 also better explained the methodology behind comparing these authors
@@
 
-In conclusion we addressed the reviews to fullest extent possible,
+We believe we have addressed the reviews to the fullest extent possible,
 both in terms of major criticisms and minor criticisms. We feel that
-this new revision is far more robust and show ease the concerns of the
+this new revision is far more robust and should ease the concerns of the
 reviewers.
 
-Thanks you,
+\vspace{2mm}
+
+Thank you,
+
+\vspace{3mm}
 

### 2012-01-20 ese/ese-major-response.tex

@@
 
-Within this new revision, we addressed this issue in a new section on inter-rater
-reliability (5.3) and referenced this discussion in the threats to
-validity (5.4) and the conclusion.
-
-Furthermore, we expanded the discussion with statistical evidence that our IRR results were mostly
+We addressed this issue in a new section on inter-rater
+reliability (5.3), and referenced this discussion in the threats to
+validity (5.4) and the conclusion.  Furthermore, we expanded the discussion with statistical evidence that our IRR results were mostly
 better than noise. We did extensive simulations sampling our ratings,

### 2012-03-15 ese/ese.tex
updated todos for 'minor changes'
minor fixes

@@
 However, in our previous work we
-investigated \emph{topic trends} --- that is, topics that recur over
-time --- we observed that topic trends often corresponded to
+investigated \emph{topic trends}---that is, topics that recur over
+time---where we observed that topic trends often corresponded to
 \emph{non-functional requirements} (NFRs)~\cite{Hindle09ICSM}, which is
@@
 %We first introduces labelled topic extraction, and then uses that technique to analyze development activity in two large-scale examples.
-We begin by discussing related work in Section \ref{sec:related}.
-Next, we describe how we generated our data (Section \ref{sec:wordlist}). For semi-unsupervised classification (Section \ref{sec:unsuplabelling}), we
+We begin by discussing related work in Section~\ref{sec:related}.
+Next, we describe how we generated our data (Section~\ref{sec:wordlist}). For semi-unsupervised classification (Section~\ref{sec:unsuplabelling}), we
 begin by creating word-lists to signify when a topic matches an NFR label. We then apply our classifier and analyze the results. 
@@
  %Section \ref{sec:unsuplabelling} presents our technique for using word-lists to do unsupervised classification, and describes the results.
-In Section \ref{sec:suplabelling}, we manually annotate the topics, and use those annotations as training data for supervised classification.  
+In Section~\ref{sec:suplabelling}, we manually annotate the topics, and use those annotations as training data for supervised classification.  
 %We then present visualizations of named topics and their trends over time to aid communication and analysis. 
 To demonstrate an application of labelled topic extraction, we use an exploratory case study of three open source database systems to show how named
-topics can be compared between projects  (Section \ref{sec:analysis}). 
-The paper concludes with a discussion of limitations (Section \ref{sec:limit}), and future work.
+topics can be compared between projects  (Section~\ref{sec:analysis}). 
+The paper concludes with a discussion of limitations (Section~\ref{sec:limit}), and future work.
 
@@
 request and check-in with tags such as ``ui'', ``usability'' or
-``milestone 3''. 
+``milestone~3''. 
 % make evolutionary analyses of a single product. 
@@
 necessary, and then use LDA to extract documents from these
-bucket. These topics were then related to topics in other adjacent
+buckets. These topics were then related to topics in other adjacent
 buckets, and joined if deemed sufficiently similar. This paper, and our
@@
 Each word in the message was
-stripped of punctuation and converted to lowercase.
+stripped of punctuation and converted to lower-case.
 We summarized each message as a word distribution minus stop-words
@@
 This might be related to release 3.23.38, which focused on Windows compatibility. 
-Similarly, in August, 2002, both \emph{functionality} and \emph{portability} are frequent, and mailing list data suggests this is related to the release
+Similarly, in August 2002, both \emph{functionality} and \emph{portability} are frequent, and mailing list data suggests this is related to the release
 of version 3.23.52, a general bug fix with a focus on security (a component of the \emph{functionality} NFR in the ISO9126 model). 
@@
 Ward method of Hierarchical clustering on the 7-dimensional space.  
-This will group developers according to the Euclidean distance metric.
+This method groups developers according to the Euclidean distance metric.
 %We chose cosine distance because it is often used in IR research to
@@
 The height in Figure \ref{fig:authorcluster} describes the euclidean
-distance between elements, in terms of the path 
+distance between elements, in terms of the path. 
 We clustered the authors into both $2$-clusters
@@
 %NEIL - not sure the relevance of this ... needs more explaining. --> (they are also the most frequent committers). 
-This is interesting because it means that the important developers are
+This is interesting because it means that the important developers
 have different proportions of NFR-relevant contributions; they have different
@@
 \texttt{momjian}, \texttt{tgl}, and \texttt{petere} would be in if we
-evaluated the clusters proportionately, that is if we use distance
+evaluated the clusters proportionately, that is, if we use distance
 measures that normalize the data and ignore magnitude (e.g. turning an
@@
 cosine distance, Euclidean distance of unit vectors, or Pearson
-correlation distance ($1 - r$). Whereas, Euclidean distance uses both angle and magnitude of a
-vector. For all three non-magnitude distances measures we tend to see that
+correlation distance ($1 - r$). 
+%Whereas Euclidean distance uses both angle and magnitude of a vector. 
+For all three non-magnitude distances measures we tend to see that
 the three largest contributors, \texttt{momjian}, \texttt{tgl}, and
@@
 our dictionaries and training corpus in their commit messages. Furthermore, a top developer, in
-terms of number of commits, will have more samples, and thus be more likely
+terms of number of commits, will have more samples, and thus be more likely to 
 take on a more general role. For example, we found in a previous

### 2012-03-16 ese/ese.tex
author reference

@@
 the NFRs that are associated with the topic. 
-The analysis in this section used Neil's annotations.
+The analysis in this section used author 2's annotations.
 This produced a map which associated developer name with NFR
@@
 \label{sec:inter-rater}
-To determine inter-rater reliability two of the authors---Ernst and Hindle---each annotated the PostgreSQL topics,
+To determine inter-rater reliability two of the authors---author 1 and author 2---each annotated the PostgreSQL topics,
 and then evaluated each other's annotations. 
@@
 IRR measures 96\% of the time. For \emph{portability} and \emph{efficiency}, our IRR
-was greater 100\% of the time both for Neil's ratings, Abram's ratings
-and the union of both Neil and Abram. 
+was greater 100\% of the time both for author 2's ratings, author 1's ratings
+and the union of both authors 1 and 2. 
 We know this by evaluating our IRR ratings against the empirical
@@
 Were our results harmed by low IRR scores? For MaxDB and MySQL, since
-each annotator (Neil or Abram) acted as the sole oracle, the results
+each annotator (author 2 or author 1) acted as the sole oracle, the results
 for either system still stand, but construct validity is harmed since 

### 2012-03-16 ese/ese.tex
minor fixes and wording

@@
 artifacts by using semi-unsupervised machine-learning
-algorithms, such as Latent Dirichlet Allocation (LDA), used for concept and topic analysis to suggest candidate
+algorithms, such as Latent Dirichlet Allocation (LDA). LDA is used for concept and topic analysis to suggest candidate
 word-lists or topics that describe and relate software artifacts.
@@
 difficult to  
-interpret in the absence of meaningful summary labels.  Current topic modeling 
-techniques assume manual labelling and do not use domain-specific knowledge
+interpret in the absence of meaningful summary labels.  Current attempts to interpret topics
+%topic modeling 
+%techniques 
+assume manual labelling and do not use domain-specific knowledge
 to improve, contextualize, or describe results for the developers.  
@@
 
-Concrete applications of \emph{labelled topic extraction} include the annotation of development artifacts with NFR-related tags and the creation of project dashboards.
+Concrete applications of \emph{labelled topic extraction} include the annotation of development artifacts with non-functional requirements (NFR) keywords, and the creation of project dashboards, which can summarize project activity in a single location.
 Annotating software development artifacts with NFR-related tags would allow developers to create detailed directed queries of different artifact kinds that concern the same NFR; for example, a developer could browse the recent history of performance-related bug reports and code check-ins.
@@
 time---where we observed that topic trends often corresponded to
-\emph{non-functional requirements} (NFRs)~\cite{Hindle09ICSM}, which is
-further emphasized in this paper due to the large numbers of NFR
-labelled topics.  
+\emph{non-functional requirements} (NFRs)~\cite{Hindle09ICSM}.
+%which is further emphasized in this paper due to the large numbers of NFR labelled topics.  
 This is encouraging, as NFRs have the property of being cross-domain
@@
 	%\item We added new results on developer activities.
-	\item We conducted another case-study on PostgreSQL. 
+	\item We added a third case study using PostgreSQL, in order to consider another set of data while remaining in the database domain.
 	\item We used two authors to each annotate the same PostgreSQL topics in order to compare
@@
 and \emph{efficiency}. While there is much room for improvement, we are seeing some
-correlation between our quality word lists and relevant topics.
+correlation between our NFR word lists (\textsf{exp2} and \textsf{exp3}) and relevant topics.
 
@@
 \paragraph{External validity} -- %was addressed by our use of two case studies, but there are still issues. 
-Our data originated from OSS database projects and thus might not be applicable to commercially developed software or other domains. 
+Our data originated from OSS database projects and thus might not be applicable to commercially developed software or software projects in other domains. 
 Furthermore, our analysis techniques rely on a project's use of meaningful commit messages, although we feel this is the most frequently occurring form
@@
 Our next approach was supervised, using single-label and multi-label learners. 
-Both kinds of learners performed well with average
+Both kinds of learners performed noticeably better than random with average
 %area under the 

### 2012-03-16 src/validate/plot_timeline.py
Fixed figure text label sizes for 2 and 6

@@ -21,10 +21,12 @@ def plot_timeline(proj, period_map):
     
     fig_dir = './output/' # /Users/nernst/Dropbox/research/abram/papers/naming-paper/icse/figures/'
     #configure the figure
-    width = 17
+    width = 22
     height = 6
     #rc("font", family="serif")
     #rc("font", size=10)
+    #set default font size
+    matplotlib.rcParams['font.size'] = 24
     # fractions of figure width
     rc("figure.subplot", left=0.1)
     rc("figure.subplot", right=0.98)
@@ -47,12 +49,13 @@ def plot_timeline(proj, period_map):
         ax.set_xlim(datetime.date(2000,07,31),datetime.date(2004,8,9))#5*len(period_map.keys()
     if proj == 'maxdb':
         ax.set_xlim(datetime.date(2004,06,29),datetime.date(2006,6,19))#5*len(period_map.keys()
-    if proj == 'pgsql':
+    if proj == 'pgsqln' or proj == 'pgsqla':
         ax.set_xlim(datetime.date(2002,02,1),datetime.date(2004,6,28))
 
     #ax.set_xlabel('Date')
+    ax.tick_params(axis='both',direction='in',length=8,width=3,top=False,right=False,zorder=15)
     ax.set_yticks([2.5,7.5,12.5,17.5,22.5,27.5,32.5,37.5])
-    ax.set_yticklabels(['None', 'Portability', 'Efficiency', 'Reliability', 'Functionality', 'Usability', 'Maintainability','*Key Events*'])#,fontproperties = font)
+    ax.set_yticklabels(['None', 'Portability', 'Efficiency', 'Reliability', 'Functionality', 'Usability', 'Maintainability','*Key Events*'], rotation=30,rotation_mode='anchor')#,fontproperties = font)
     ax.xaxis_date()
     ax.xaxis.set_major_locator(months)
     ax.xaxis.set_major_formatter(monthsFmt)
@@ -119,7 +122,7 @@ def plot_timeline(proj, period_map):
             l = Line2D([datetime.date(2000,07,31),datetime.date(2004,8,9)],[5*i,5*i]  , lw = 0.5, color='black')#5*len(period_map.keys()
         if proj == 'maxdb':                                                                            
[…]

### 2012-03-19 ese/ese.tex
Fixed figure text label sizes for 2 and 6
fixed labels fig 8
refactored section 3. I moved the results into a separate section per reviewer #3. I think he/she is probably right. Also added a blurb on F-measure per #2.

@@
 
-\section{Study Design and Execution}
+\section{Study Design}
 Figure \ref{fig:process} gives an outline of our methodology.
@@
 With supervised learning, we train our data with manual annotations in order to match topics with NFRs. Finally, these topics are used to analyze the
-role of NFRs in software maintenance.
+role of NFRs in software maintenance. We discuss the results in Section \ref{sec:obs}.
 
@@
   \centering
- \includegraphics[height=.6\textheight]{figures/process-model}
+ \includegraphics[height=.5\textheight]{figures/process-model}
  \caption{Research methodology process view.}
@@
 % A topic may have more than one matching keyword. % ok what?
-
-We validate classification performance using the \emph{area under the curve
-of Receiver Operating
-Characteristic}~\cite{Fawcett2006861},
-abbreviated \emph{ROC}, sometimes called Area Under the Curve or \emph{AUC}, and the
-F-measure, which is the harmonic mean of precision and recall, i.e.,
-$2 * (P * R) / (P + R)$. 
-Throughout the paper we will provide F-measure scores so that readers
-who are more familiar with F-measure than ROC can intuitively
-interpret the results.
-
-ROC values provide a score %, similar to school letter-grades (A is 0.9, C is 0.6),
- reflecting how well a particular learner performed for the given data. 
-ROC maps to the more familiar concepts of precision/sensitivity and recall/specificity: it plots the true positive rate (sensitivity) versus the false
-positive rate (1 - specificity). 
-A perfect learner has a ROC value of $1.0$, reflecting perfect recall and precision. 
-A ROC result of $0.5$ would be equivalent to a random learner (that
-is, issuing as many false positives as true positives). While we
-recognize that using $0.5$ as the base-line means our ROC scores will
-look much larger than our F-Measure scores, we feel that the knowledge
-that random selection is $0.5$ or worse is helpful for interpreting our results.
-The %area under the -- NE confusing to have area again ... and not importantana
-ROC of a classifier is equivalent to the probability that the classifier will rank a randomly chosen positive instance higher than a randomly chosen
-negative instance.
-
-We argue for using ROC over F-Measure because ROC suffers less from
-bias, as F-Measure often skews towards the positive
-class~\cite{flach-icml03,Forman:2010:ACS:1882471.1882479}, especially
-in the case of class imbalance. Forman et
-al.~\cite{Forman:2010:ACS:1882471.1882479} demonstrate that cross-fold
-validation will often produce low F-measures given high class
-imbalance and the presence of false negatives.
-Although
-recent work~\cite{Forman:2010:ACS:1882471.1882479} has suggested that while ROC suffers less bias than
-the average F-measure ($F_{avg}$) for cross-folds, F-measures ($F_{tp,fp}$) that are calculated from
-the sums of true-positives,
-false-positive, and false-negatives across folds tend to exhibit no
-bias. Unfortunately our experimental framework lacks a robust way to
-calculate $F_{tp,fp}$, thus we still provide ROC and $F_{avg}$.
-We consider our labelling classifiers acceptable if they outperform a random classifier ($0.5$).
 
@@
 
+\subsection{Supervised Labelling}
+\label{sec:suplabelling}
+Supervised labelling requires expert analysis of the correct
+class/label to assign a label to a topic. In our approach, we use the top-level NFRs in the ISO9126 standard~\cite{iso9126} for our classes, but other
+taxonomies are also applicable.%In order to validate how effective these word-bag approaches to topic labelling would be we created a validation data set.
+
+We used a suite of supervised classifiers, WEKA~\cite{weka09},
+that includes machine learning tools such as support vector machines and Bayes-nets. 
+We also used the multi-labelling add-on for WEKA, Mulan~\cite{mulan}. %\footnote{\url{http://mlkd.csd.auth.gr/multilabel.html}}. 
+Traditional classifiers label topics with a single class, whereas Mulan allows for a mixture of classes per topic, which is what we observed while
+manually labelling topics. For example, a given topic (word distribution) may be `about' both usability and maintainability, if this topic was a product of a discussion on design trade offs.
+The \emph{features} we used are word counts/occurrence	per topic, if
+a word occurs frequently enough in a topic we consider it a feature of
+the topic.
+
+To assess the performance of the supervised learners, we did a 10-fold cross-validation~\cite{Kohavi1995}, a common technique for evaluating machine
+learners, where the original data is partitioned randomly into ten sub-samples, and each sample is used to test against a training set composed of the nine other samples.
+%One subsample is used as a training set and evaluated on the other nine sub-samples. 
+%This is repeated nine more times, with each subsample used once as the training set. 
+We discuss these results below.% in Section \ref{sec:suplabelling}.
+
+\subsection{Applying Multiple Labels to Topics}
+\label{sec:multilabel}
+As noted in Section \ref{sec:wordlist}, each topic in our data-set can be composed of zero or more NFRs. 
+For example, a commit message might address \textit{reliability} in the context of \textit{efficiency}, or make a \textit{maintainability} improvement
+in the source code that relates to \textit{usability}. 
+However, traditional machine learning techniques, such as Naive Bayes, can map topics to only a single class. 
+The Mulan~\cite{mulan} library encapsulates several different multi-label machine learners which can label elements with multiple labels.
+Mulan also includes methods for determining the performance of these learners.
+
+Two perspectives used to evaluate multi-label learners are with micro or macro measurements (shown in Figure \ref{fig:mulan}a).
+Macro measurements are aggregated at a class or label level (per class) while micro measurements are at the element level (per element).
+A macro-ROC measurement is the average ROC over the ROC values for all labels, where a micro-ROC is the average ROC over all examples that were classified. 
+For MaxDB, the macro-ROC values are undefined because of poor
+performance of one of the labels.%, see Figure \ref{fig:mulan}. 
+During cross-folds validation if the class imbalance or poor learner
+performance causes a division by zero when evaluating one of the NFR
+labels (in this case \emph{usability}), creating an undefined
+value. This undefined 
+value is propagated through to the mean of ROC per NFR label, causing
+the end calculation of macro-ROC to be undefined.
+
+\section{Observations and Analysis}
+\label{sec:obs}
+\subsection{Measuring Classification Performance}
+We validate classification performance using the \emph{area under the curve
+of Receiver Operating
+Characteristic}~\cite{Fawcett2006861},
+abbreviated \emph{ROC}, sometimes called Area Under the Curve or \emph{AUC}, and the
+F-measure, which is the harmonic mean of precision and recall, i.e.,
+$2 * (P * R) / (P + R)$. 
+In this domain, F-measure values reflect the performance of our
+approach with respect to false-positive/false-negative (and the duals) results.
+A higher F-measure indicates that more topics were correctly labelled.
+Throughout the paper we will provide F-measure scores so that readers
+who are more familiar with F-measure than ROC can intuitively
+interpret the results.
+
+ROC values provide a score %, similar to school letter-grades (A is 0.9, C is 0.6),
+ reflecting how well a particular learner performed for the given data. 
+ROC maps to the more familiar concepts of precision/sensitivity and recall/specificity: it plots the true positive rate (sensitivity) versus the false
+positive rate (1 - specificity). 
+A perfect learner has a ROC value of $1.0$, reflecting perfect recall and precision. 
+A ROC result of $0.5$ would be equivalent to a random learner (that
+is, issuing as many false positives as true positives). While we
+recognize that using $0.5$ as the base-line means our ROC scores will
+look much larger than our F-Measure scores, we feel that the knowledge
+that random selection is $0.5$ or worse is helpful for interpreting our results.
+The %area under the -- NE confusing to have area again ... and not importantana
+ROC of a classifier is equivalent to the probability that the classifier will rank a randomly chosen positive instance higher than a randomly chosen
+negative instance.
+
+We argue for using ROC over F-Measure because ROC suffers less from
+bias, as F-Measure often skews towards the positive
+class~\cite{flach-icml03,Forman:2010:ACS:1882471.1882479}, especially
+in the case of class imbalance. Forman et
+al.~\cite{Forman:2010:ACS:1882471.1882479} demonstrate that cross-fold
+validation will often produce low F-measures given high class
+imbalance and the presence of false negatives.
+Although
+recent work~\cite{Forman:2010:ACS:1882471.1882479} has suggested that while ROC suffers less bias than
+the average F-measure ($F_{avg}$) for cross-folds, F-measures ($F_{tp,fp}$) that are calculated from
+the sums of true-positives,
+false-positive, and false-negatives across folds tend to exhibit no
+bias. Unfortunately our experimental framework lacks a robust way to
+calculate $F_{tp,fp}$, thus we still provide ROC and $F_{avg}$.
+We consider our labelling classifiers acceptable if they outperform a random classifier ($0.5$).
+
+\subsection{Analysis of the Semi-Unsupervised Labelling}
 % Did a little validation here
@@
 
-\subsubsection{Analysis of the Semi-unsupervised Labelling} %--- %Based
+%\subsubsection{Analysis of the Semi-unsupervised Labelling} %--- %Based
 				%on the labels, and our manual topic
@@
 %\vspace*{-1em}
-\subsection{Supervised Labelling}
-\label{sec:suplabelling}
-Supervised labelling requires expert analysis of the correct
-class/label to assign a label to a topic. In our approach, we use the top-level NFRs in the ISO9126 standard~\cite{iso9126} for our classes, but other
-taxonomies are also applicable.%In order to validate how effective these word-bag approaches to topic labelling would be we created a validation data set.
-
-We used a suite of supervised classifiers, WEKA~\cite{weka09},
-that includes machine learning tools such as support vector machines and Bayes-nets. 
-We also used the multi-labelling add-on for WEKA, Mulan~\cite{mulan}. %\footnote{\url{http://mlkd.csd.auth.gr/multilabel.html}}. 
-Traditional classifiers label topics with a single class, whereas Mulan allows for a mixture of classes per topic, which is what we observed while
-manually labelling topics. For example, a given topic (word distribution) may be `about' both usability and maintainability, if this topic was a product of a discussion on design trade offs.
-The \emph{features} we used are word counts/occurrence	per topic, if
-a word occurs frequently enough in a topic we consider it a feature of
-the topic.
-
-To assess the performance of the supervised learners, we did a 10-fold cross-validation~\cite{Kohavi1995}, a common technique for evaluating machine
-learners, where the original data is partitioned randomly into ten sub-samples, and each sample is used to test against a training set composed of the nine other samples.
-%One subsample is used as a training set and evaluated on the other nine sub-samples. 
-%This is repeated nine more times, with each subsample used once as the training set. 
-We discuss these results below.% in Section \ref{sec:suplabelling}.
-
-\subsubsection{Analysis of the Supervised Labelling}
+
+\subsection{Analysis of the Supervised Labelling}
 % \label{sec:suplabelling}
@@
 
-\subsection{Applying Multiple Labels to Topics}
-\label{sec:multilabel}
-
-As noted in Section \ref{sec:wordlist}, each topic in our data-set can be composed of zero or more NFRs. 
-For example, a commit message might address \textit{reliability} in the context of \textit{efficiency}, or make a \textit{maintainability} improvement
-in the source code that relates to \textit{usability}. 
-However, traditional machine learning techniques, such as Naive Bayes, can map topics to only a single class. 
-The Mulan~\cite{mulan} library encapsulates several different multi-label machine learners which can label elements with multiple labels.
-Mulan also includes methods for determining the performance of these learners.
+\subsection{Analysis of the Multiple Label Approach}
+
 % TOO DETAILED The problem framed in the learners above has changed; instead of looking at the precision and recall of applying one label, we rank
@@
 \end{figure}
-
-Two perspectives used to evaluate multi-label learners are with micro or macro measurements (shown in Figure \ref{fig:mulan}a).
-Macro measurements are aggregated at a class or label level (per class) while micro measurements are at the element level (per element).
-A macro-ROC measurement is the average ROC over the ROC values for all labels, where a micro-ROC is the average ROC over all examples that were classified. 
-For MaxDB, the macro-ROC values are undefined because of poor
-performance of one of the labels.%, see Figure \ref{fig:mulan}. 
-During cross-folds validation if the class imbalance or poor learner
-performance causes a division by zero when evaluating one of the NFR
-labels (in this case \emph{usability}), creating an undefined
-value. This undefined 
-value is propagated through to the mean of ROC per NFR label, causing
-the end calculation of macro-ROC to be undefined.
 
@@
 \subfloat[PostgreSQL 7.2]{					
-\includegraphics[width=\textwidth]{figures/pgsql-timeline}
+\includegraphics[width=\textwidth]{figures/pgsqln-timeline}
 \label{fig:pgsql-timeline}
@@
   \centering
-  \includegraphics[width=.9\textwidth]{figures/author-distance-from-aggregate}
+  \includegraphics[width=.9\textwidth]{figures/author-distance-from-aggregate-manlab}
   % \caption{PostgreSQL Author Similarity to the Global Distribution of

### 2012-03-19 ese/ese.tex
updated figure 7 with dashed lines

@@
 Figure \ref{fig:maxdb-unsup-results} shows our ROC results for
-PostgreSQL, MaxDB and MySQL. We omit plots of \textsf{exp1} due to poor results. We describe F-measure results in the text below.
+PostgreSQL, MaxDB and MySQL.
+We did not calculate the ROC of \textsf{exp1} because the labels used were completely different than the manual annotation labels.
+We describe F-measure results in the text below.
 
@@
   \caption{PostgreSQL Author Clusters: Authors clustered by Euclidean
-    distance of their NFR contributions. The outer rectangles (black)
+    distance of their NFR contributions. The two outer rectangles (thicker black)
     reflect the clustering of authors into 2 clusters; while the inner rectangles
-    (dim grey) reflect clustering of authors into 6 clusters. The height is a
+    (dashed) reflect clustering of authors into 6 clusters. The height is a
     measure of cluster similarity, Euclidean distance, where smaller

### 2012-04-02 src/validate/interrater.R
inter-rater for figure 9

@@ -1,5 +1,6 @@
 #neil and abram annotations!
 cores <- 8
+cores <- 4
 pgsqln <- read.csv("output/pgsqln.arff.csv",header=TRUE)
 pgsqla <- read.csv("output/pgsqla.arff.csv",header=TRUE)
 nn <- length(pgsqln)
@@ -110,7 +111,8 @@ ratings[["Neil"]] <- pn
 ratings[["Both"]] <- paunion
 panames <- c("Portability","Functionality","Reliability","Maintainability","Efficiency", "Usability", "None")
 
-for (person in c("Both","Abram","Neil")) {
+#for (person in c("Both","Abram","Neil")) {
+for (person in c("Both")) {
 pan <- ratings[[person]]
 i <- names(pan)[1]
 #
@@ -142,21 +144,27 @@ pankappas <- sapply(names(pa),function(i) {
    kappaof(pa[,i],pn[,i])
 })
 
-  pdf(paste("self-sample",person,"pdf",sep="."),width=12,height=8)
+  heights=c(3,4,5,6,8,9,10)
+
   miny <- min(c(sapply(kappas,min), min(pankappas)))
   maxy <- max(c(sapply(kappas,max), max(pankappas)))
-  boxplot(kappas, names=panames,ylim=c(miny,maxy))
-  title(paste("Human Ratings (",person,") versus Sampled Random Ratings"))
-  lines(pankappas, type="o",col="red",lw=3)
-  legend("topleft", c("Interrater Reliability per NFR Topic Label"), bty="o",lwd=3, col=c("red"),  lty=c(1));
-  dev.off()
+
+  for (height in heights) {
+    width = 1.5*height
+    pdf(paste("self-sample",person,"pdf",sep="."),width=7.5,height=5)
+    boxplot(kappas, names=panames,ylim=c(miny,maxy))
+    title(paste("Human Ratings (",person,") versus Sampled Random Ratings"))
+    lines(pankappas, type="o",col="red",lw=3)
+    legend("topleft", c("Interrater Reliability per NFR Topic Label"), bty="o",lwd=3, col=c("red"),  lty=c(1));
+    dev.off()
+  }
   print(paste("Person ", person))
   print(sapply(kappas,mean))
   print(paste("Person ECDF measures", person))
   print(sapply(c(1:7),function(i){ecdf(kappas[[i]])(pankappas[i])}))
 }
 
-
+exit(0);
[…]

### 2012-04-02 ese/ese.tex
fix 1 little figure bug
- [X]  Discussion of multiple labels        I did not find the discussion arguing that multiple labels in        Pg were caused by too many terms in each 30-day window to be        convincing. It seems that the overall increase in change-log        words is less than double of that in MySQL (165K words in Pg vs        101K in MySQL), yet the differences in topic overlap appear to        be quite notable. The argument also appears to conflict with an        earlier statement that N=20 was chosen because the overlap in        topics (two topics sharing most frequent terms

@@
 produced tend to be duplicates of each other, that is, they share the
-top terms. During this study we found that if we used a value of 20 for $N$, duplicate topics were infrequent.
+top terms. During this study we found that if we used a value of 20
+for $N$, duplicate topics were infrequent in MaxDB and MySQL. The
+PostgreSQL study was added later and we followed the same methodology
+to extract topics fro PostgreSQL, the choice of $N$ = 20 was not
+necessarily the optimal choice for PostgreSQL.
 Linear combinations of these $N$ word distributions are meant to represent and recreate the word distributions of any of the original messages.
@@
 size differences arise from the number of commits, the number of files
-and the verbosity of the commit descriptions.
-
+and the verbosity of the commit descriptions. The density of
+information with in the PostgreSQL is a concern as well. Not only are
+PostgreSQL commits larger, when raw commit text are compressed with \texttt{gzip} per
+each project, PostgreSQL (402kb) is 5.1 times the size of MaxDB (80kb) and 11.8 the size
+of MySQL (34kb). PostgreSQL commit messages often include email messages from
+the from the mailing list where the patch originated.
+
+% 
+\begin{comment}
+for file in ~/projects/query1-data/postgresql/query1.mrs2 \
+            ~/projects/query1-data/maxdb-7500/dist/query1.mrs2 \
+            ~/projects/query1-data/mysql-3.23/query1.mrs2 
+do
+sort -u $file | egrep "\|200[234]\-[0-9]" | awk -F\| '{print $5}' | sed -e 's/NEWLINE/\n/g' | gzip -c - | wc -c
+done
+\end{comment}
 %\DONE{expand on why?}
@@
 
-\subfloat[PostgreSQL 7.2]{					
-\includegraphics[width=\textwidth]{figures/pgsqln-timeline}
+\subfloat[PostgreSQL 7.3]{					
+\includegraphics[width=\textwidth]{figures/pgsql-timeline}
+%\includegraphics[width=\textwidth]{figures/pgsqln-timeline}
 \label{fig:pgsql-timeline}

### 2012-04-06 ese/ese.tex
- [ ] There appear to be some discrepancies in numbers.          According to Table 2, there were 640 topics in Pg. 20 topics          per 30-day windo 32 30-day windows for Pg, yet the time          period appears to contain fewer such windows.  For the other          two projects the numbers of topics are not even divisible          by 20. Some clarification would help. It appears that Pg has          625 of 640 topics with at least two labels: testability and          correctness based on exp1. It may be worth giving he reader          some idea about how many topics have exactly one label

@@
 top terms. During this study we found that if we used a value of 20
-for $N$, duplicate topics were infrequent in MaxDB and MySQL. The
+for $N$, duplicate topics were infrequent in MaxDB and MySQL. 
+Sometimes LDA produces topics that have a
+very low association to documents, in this paper we removed topics
+that
+exhibit very low document association. Our threshold for
+topic inclusion was that a topic needed at least 1
+document with a gamma (the variational
+posterior Dirichlet~\cite{Blei2003} of a topic to a document, similar
+to an association weight) of $1.0$ or greater.
+The
 PostgreSQL study was added later and we followed the same methodology
@@
 Table \ref{tbl:wordlist} shows how many topics were labelled for
-MaxDB, MySQL and PostgreSQL.  Notice how PostgreSQL had far fewer
+MaxDB, MySQL and PostgreSQL.  
+The total topics are often less than 20 times the number of months
+studied because in some months there was not enough data for LDA to
+extract 20 actual topics as some topics were not actually strongly
+associated with any documents. 
+Notice how PostgreSQL had far fewer
 unlabelled topics, PostgreSQL also was under going far more

### 2012-04-08 ese/authors.R
authors

@@ -1,0 +1,9 @@
+v <- read.table("authors.csv")
+x <- c()
+x$counts <- v$V1
+x$author <- v$V2
+tm <- sum(x$counts[x$author %in% c("tgl","momjian")])
+tmp <- sum(x$counts[x$author %in% c("tgl","momjian","petere")])
+
+ts <- sum(x$counts)
+tm/ts

### 2012-04-08 ese/ese.tex
- [X] Validity on authors         Although briefly noted in the validity section, the two most         frequent committers may be defining the global distribution of         NFR topics in Pg, if, for example they have produced the bulk         of commits for the topic extraction. This, perhaps, could be         mentioned earlier with a fraction of commits these two authors         produced.      - In section {RQ3: Do Different Developers Work on Different        NFRs?} we added text that showed the proportions of the top 3        authors and later mentioned this domination of the the global

--- 2012-04-02 ese/ese.tex
@@
 top terms. During this study we found that if we used a value of 20
-for $N$, duplicate topics were infrequent in MaxDB and MySQL. The
+for $N$, duplicate topics were infrequent in MaxDB and MySQL. 
+Sometimes LDA produces topics that have a
+very low association to documents, and in this paper we removed topics
+that
+exhibit very low document association. Our threshold for
+topic inclusion was that a topic needed to be related to at least 1
+document with a gamma (the variational
+posterior Dirichlet~\cite{Blei2003} of a topic to a document, similar
+to an association weight) of $1.0$ or greater.
+The
 PostgreSQL study was added later and we followed the same methodology
-to extract topics fro PostgreSQL, the choice of $N$ = 20 was not
+to extract topics fro PostgreSQL, the choice of $N = 20$ was not
 necessarily the optimal choice for PostgreSQL.
@@
 with the relevant label. We added a catch-all label, \emph{none}, which we used when none of the six NFRs was suitable. In some instances, we used finer-grained
-annotations that would be aggregated up to one of these higher-level labels. 
+annotations that would be aggregated up to one of these higher-level
+labels. 
 %We compared against this data-set, but we also used the data-set for our supervised machine learning based topic classification. 
@@
 topics, although this visualization can be generated from the
-results of labelled topic extraction. Note that there are no unlabelled topics in this data-set.
+results of labelled topic extraction. Note that there are no
+unlabelled topics in this data-set. Furthermore these plots are
+display data about the interpretation of NFRs relevant to topics;
+different annotators might produce different results.
 
@@
 %data/clusters.R
+%hindle1@burn:~/oldhome/projects/query1-data/postgresql$ sort -u query1.mrs2 | egrep "\|200[234]\-[0-9]" | awk -F: '{print $1}' | sort | uniq -c > ~/projects/lda-paper/ese/authors.csv
 
@@
 
+% > tmp/ts
+% [1] 0.8601498
+% > tm/ts
+% [1] 0.7855114
+
 Figure \ref{fig:authorcluster} shows that when we use $6$-clusters, 
@@
 7-dimensional space). 
+\texttt{tgl} and \texttt{momjian} make up $79\%$ of the commits, and
+together \texttt{tgl}, \texttt{momjian} and \texttt{petere} make up
+$86\%$ of all commits.
 %NEIL - not sure the relevance of this ... needs more explaining. --> (they are also the most frequent committers). 
@@
  distribution.
+As stated before, the  3 most frequent authors produced $86\%$ of the
+commits and thus they dominate the global NFR distribution.
 Authors \texttt{momjian} and \texttt{tgl} are in the top right of

### 2012-04-08 ese/ese.tex
latest attempt to address minor revisions

@@
 \usepackage{url}
-\usepackage[pdfstartview=FitH,bookmarksopenlevel=3,bookmarks=true]{hyperref} %pdfpagescrop={92 112 523 778},
+\usepackage[pdfstartview=FFitH,bookmarksopenlevel=3,bookmarks=true]{hyperref} %pdfpagescrop={92 112 523 778},
 \journalname{Empirical Software Engineering}
@@
 approach with respect to false-positive/false-negative (and the duals) results.
-A higher F-measure indicates that more topics were correctly labelled.
+A higher F-measure indicates that more topics were correctly labelled,
+$0$ is the worst, $1$ is the best, $0.5$ indicates potentially good performance, 
+and above $0.10$ gives an indication of perhaps better than random
+performance but it is hard to tell as F-measure doesn't account for
+true negatives.
 Throughout the paper we will provide F-measure scores so that readers
@@
 Figure \ref{fig:maxdb-unsup-results} shows our ROC results for
-PostgreSQL, MaxDB and MySQL.
-We did not calculate the ROC of \textsf{exp1} because the labels used were completely different than the manual annotation labels.
-We describe F-measure results in the text below.
+PostgreSQL, MaxDB and MySQL. We did not calculate the ROC of
+\textsf{exp1} because the labels used were completely different than
+the manual annotation labels, but we provide an F-measure estimate
+based on the possible overlap (we could not apply the same tools to
+calculate ROC in this case). We describe F-measure results in the text below.
 
@@
 In all cases \emph{portability} was a constant maintenance concern and was prevalent throughout the lifetime of the projects. It may surprise developers
-how often \emph{portability} arises as a concern.
+how often \emph{portability} arises as a concern, but that might
+depend on the quality of the annotations or the accuracy of the
+automatic labeller.
 

### 2012-04-08 ese/ese.tex
minor detailed response

@@
 We want to leverage different kinds of artifacts to discover threads of NFR-related discussions that occur between multiple kinds of artifacts.
-Finally, we would like to extend this analysis to other domains, to
-see what patterns might occur within those domains, such as consumer-facing software products.
+Finally, we would like to extend this analysis to many other domains of
+software projects, such as smart-phone apps, web-browsers, office-productivity tools, and web-services,
+to see what patterns might occur within those domains. These software
+project domains might share different architectural patterns and
+domain concepts relevant to NFRs that could be missing from our word-lists.
 

### 2012-04-09 ese/ese.tex
minor spelling problem

@@
 PostgreSQL study was added later and we followed the same methodology
-to extract topics fro PostgreSQL, the choice of $N = 20$ was not
+to extract topics from PostgreSQL, the choice of $N = 20$ was not
 necessarily the optimal choice for PostgreSQL.
